### 2013-11-10 distrproj.tex
12.03.2010 (Vassilis) Initial commit.

\documentclass[a4paper,10pt]{article}
\usepackage[utf-8]{inputenc}

\usepackage{paralist}

\usepackage{graphicx}

%opening
\title{Alleviating the Topology Mismatch Problem with Location-Aware Overlays: A Survey}
\author{
Vassilis Moustakas\\
Department of Informatics and Telecommunications,\\
National and Kapodistrian University of Athens\\
\mbox{\texttt{b.moystakas@di.uoa.gr}}
}

\begin{document}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Peer-to-peer systems is a rapidly growing application architecture. The overlay network abstraction on top of a best-effort infrastructure provides a powerfull tool to realize ever-wanted exotic features like anonymity, robustness, load balancing, QoS at the application layer and many more. Unfortunately weaknesses of early effectuations constrained applications from unleashing the full potential of the paradigm. One of them is the topology mismatch problem between the overlay network and the physical underlying one. This survey is on the recent research developments in alleviating the problem, in unstructured schemes, using location-aware overlay construction algorithms.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\emph{Peer-to-peer} systems gained significant attention due to their plethora of unique features in supporting file sharing among huge numbers of networked users, where each such \emph{peer} acts both as a resource provider and a resource consumer. Such serverless systems, proved to be able to achieve outstanding aggregate resource capacities as participants join the system\footnote{In peer-to-peer file-sharing systems there has been observed mitigation of this self-scaling property by the \emph{free-rider} problem\cite{adar_gnutellafreeriders_2000}}, without requiring additional expenditure for infrastructure.

As these systems evolved, three main architectures have been emerged:
\begin{itemize}
  \item \emph{centralized}
  \item \emph{decentralized structured}
  \item \emph{decentralized unstructured}
\end{itemize}

\emph{Centralized} architectures were the first to recognize that requests for popular content need not be sent to a central server but instead could be handled by the many hosts that already posses the content. \emph{Napster}, for example, maintained a \emph{central index server} based on file lists provided by participating peers. The central index server was queried by the users and it returned pointers to the actual content. Thus, by centralizing search while distributing downloads, Napster achieved a high functional design that was widely acknowledged, at the time, as ``the fastest growing Internet application ever''. Of course, the architecture is not without flaws. The centralized search facility is its ``Achilles' heel'' as it can be prooven a bottleneck for the system's scalability as well as a central point of failure and vulnerable to \emph{Denial-of-Service (DoS)} attacks.

This centralized approach, has been eventually, replaced by \emph{decentralized}, that distributed both the search and the download capabilities. In the \emph{decentralized unstructured} architectures, file placement is random, which means there is no correlation with the network topology whatsoever \cite{yang_improvep2psearch_2002}. The most important properties of such systems are that they:
\begin{itemize}
  \item support inherent heterogeneity of peers,
  \item are highly resilient to peer failures, and
  \item incur low maintenance overhead at peer arrivals and departures.
\end{itemize}
In the literature, they can be also found under the term \emph{broadcast-based} systems, because they use \emph{message flooding} among peers (Gnutella\cite{gnutella}) or among superpeers (KaZaA\cite{kazaa}) to propagate queries but with no specific destination. Each (super)peer receiving a query
\begin{inparaenum}[\itshape i\upshape)]
  \item decreases the incoming query message's lifetime which is measured in \emph{time-to-live (TTL)} hops,
  \item back-propagates its results, and
  \item if the lifetime of the incoming message did not expired (TTL did not reach zero), it forwards the message to all its neighbours except from the incoming link.
\end{inparaenum}
Early incarnations of the Gnutella protocol, though were considered far from scalable \cite{ritter_gnucantscale_2001} because:
\begin{itemize}
  \item randomly choosing logical neighbours without knowledge of the underlying network causes mismatch between physical and application-level topologies, and/or
  \item queries may be flooded to multiple paths merging in one peer while in such case one of the paths would have been enough, and/or
  \item peers may forward the same message to each other before they receive the query messages from each other
\end{itemize}
More generally, studies upon peer-to-peer traffic \cite{seroiu_analysiscds_2002, sen_analyzep2ptraffic_2004} with measurements on some popular systems, such as FastTrack, Gnutella and DirectConnect have shown that it contributes the largest portion of the overall Internet traffic. \cite{matei_mapgnutella_2002} showed that, even given that 95 percent of any two nodes are less than 7 hops away from each other, a flooding-based algorithm, like the one used by Gnutella, can generate 330TB/month in a 50.000 node network!

Recently \emph{decentralized structured} schemes have been proposed \cite{ratnasamy_can_2001, stoica_chord_2001, zhao_tapestry_2001, antony_pastry_2001}. These, form an application-level topology that provides a hash table look-up and routing service, called \emph{Distributed Hash Table (DHT)} that requires only $O \left ( log n \right )$ steps for a hash-table-like \texttt{look-up()} search that also guarantees perfect recall even if only a single copy of the searched data exists in the system\footnote{In comparison to Gnutella, that requires $O \left ( n \right )$ to reliably locate a specific file}.

Studying the behaviour of the various peer-to-peer schemes \cite{matei_mapgnutella_2002, lv_randomwalks_2002, merugu_str2unstr_2003} showed that the ad-hoc network topology of unstructured overlay networks that preserve \emph{Power Law} and \emph{Small World} characteristics\footnote{Power Law describes the node degree while the Small World describes characteristics of path length and clustering coefficient. The clustering coefficient for a node $\upsilon$ in a graph $G = \left( V, E \right)$ is defined as the ratio of the existing connections between $\upsilon$'s neighbouring nodes to $\gamma \times \left( \gamma - 1 \right)$, where $\gamma$ is the number of neighbouring nodes of $\upsilon$. High cluster coefficient means that neighbouring nodes of any node $\upsilon$ likely connect one another.} \cite{faloutsos_powerlaw_1999, saroiu_measurefileshare_2002} offer a more promising approach. Particularly:
\begin{itemize}
  \item Peer-to-peer clients are extremely \emph{transient}. Unstructured systems can have high maintenance traffic in delivering messages, updating the mapping, discovering failures and replicating lost data or pointers, making them insufficient on highly volatile networks.
  \item \emph{Keyword searches} versus \emph{exact-match queries}. In DHTs there is a tight control between the data placement and the topology of the network. For this reason it is hard to efficiently support partially matched queries while Gnutella and other similar systems effortlessly support keyword searches and other complex queries since the mechanism is realized locally, on a node-by-node basis.
  \item Popular content is located at multiple peers and thus it is more likely for a flooding-based search to return results. DHTs, on the other hand, fit better in the systems which require ability to reliably locate content, even in the extreme case that only a single-copy exists in the network.
\end{itemize}
For this reasons, efforts have been placed for optimizing the efficiency of decentralized unstructured peer-to-peer networks. Research mainly focuses on
\begin{inparaenum}[\itshape i\upshape)]
  \item reducing unnecessary, redundant communication traffic, and
  \item exploiting physical locality to reduce communication response.
\end{inparaenum}
The goal can be achieved at, both, the application-level network as well as the underlying physical one. In the first case by refining the message relay techniques, while in the second one, by adaptively reconstructing the application network to map as well as possible to the the physical network.

This survey begins with a quick review of the conception of the overlay network. It continues with the phenomenon of the topology mismatch between the overlay and the underlying networks and ultimately end up presenting a series of distinguished research work that has been published in the recent literature concerning the alleviation of such a problematic state, focusing on decentralized unstructured peer-to-peer architectures.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overlay Networks}
An \emph{overlay} network is an abstract, logical interconnection formed on top of another network. It consists of a set of nodes, the computing elements, that are being connected by virtual (logical) point-to-point links. For example, the Asymmetric Digital Subscriber Line (ADSL) is, actually, an overlay network on top of the Public Switched Telephone Network (PSTN).

The abstraction of overlay networks has been proposed as a way to implement efficient, fully distributed, application layer, services on top of the best-effort IP layer forming a widely ranged family of protocols collectively known as \emph{peer-to-peer} or \emph{p2p} for short.

Initial efforts targeted on the implementation of one-to-many addressing schemes that could replace \emph{IP multicast} in providing higher quality of streaming media through \emph{quality of service (QoS)} guarantees. IP multicast as well as other proposals such as \emph{IntServ} and \emph{DiffServ}\cite{cisco_diffserv_2005} have not seen wide acceptance (yet?), largely because they require changes to all routers of the network. On the other hand, an overlay network can be incrementally deployed on end-systems running the overlay protocol software, without cooperation from the ISPs. Academic research includes \cite{chu_esm_2000}, \cite{jannotti_overcast_2000} and \cite{kwon_tag_2002} among others.

Overlay networks were also implemented in order to back the routing of messages to destinations whose address is not known in advance. This resulted in the appearance of well kwown peer-to-peer protocols such as \cite{gnutella} and \cite{maymounkov_kademlia_2002} widely used for digital content sharing among their network nodes.

Other special perpose overlay networks include \cite{anderson_ron_2001} for resilient routing; \cite{subramanian_overqos_2004} for quality of service guarantees; and \cite{clarke_freenet_2001} for anonymity; to name just a few.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Topology Mismatch Problem}
An overlay network is able to control the sequence of peers, a message traverses, before reaching its destination but it is completely unaware of how the actual packets are switched at the underlying infrastructure along the overlay path. For example, a single logical point-to-point link on the overlay, most of the time corresponds to multiple physical links, in the underlying layer.

Additionally a link in the underlying network often serves the mapping of several overlay paths causing increase of the physical link's \emph{stress}\cite{chu_esm_2002}.  Furthermore, the stochastic behaviour of peers randomly joining and leaving the peer-to-peer network all cause the overlay to physical mapping to create such unnecessary, redundant traffic that impacts the efficiency of the network in terms of average response time. The phenomenon is reffered to as the \emph{topology mismatch problem}.

For example, assume nodes $A$, $B$, $C$ and $D$ are connected through the physical network shown in Figure~\ref{figure:phys} and peer $A$ sends a message to peer $C$ (through $B$ of course). If these peers participate in an overlay network according to one of the setups of Figures~\ref{figure:over1} and~\ref{figure:over2} then users will yield different performances. In the fist overlay, the message will traverse the following sequence of links in the physical layer: $3 \rightarrow 7 \rightarrow 1 \rightarrow 2 \rightarrow 2 \rightarrow 1 \rightarrow 7 \rightarrow 8 \rightarrow 5 \rightarrow 5 \rightarrow 4$. In the second overlay: $3 \rightarrow 8 \rightarrow 4$. For the first case the cost is $45 ms$ while in the second $15 ms$. Therefore the second overlay is more congruent with the underlying physical network than the other one and thus more efficient.

\begin{figure}
\centering
  \includegraphics[scale=0.4]{img/phys.jpeg}
\caption{Interconnection of nodes in the physical level.}
\label{figure:phys}
\end{figure}

\begin{figure}
\centering
  \includegraphics[scale=0.4]{img/over1.jpeg}
\caption{An overlay setup}
\label{figure:over1}
\end{figure}

\begin{figure}
\centering
  \includegraphics[scale=0.4]{img/over2.jpeg}
\caption{An (other) overlay setup}
\label{figure:over2}
\end{figure}

The problem of constructing an optimal overlay is known to be \emph{NP-hard}\cite{chawathe_scattercast_2000}. Moreover, The Internet is an interesting environment where end-to-end latencies demonstrate triangle inequality violations (TIVs). These delay space TIV is a consequence of the Internet's structure and routing policies and thus will remain a property of the Internet for the foreseeable future \cite{zheng_irprtt_2005}. TIVs affect network coordinate \cite{cox_vivaldi_2004, wong_meridian_2005} and positioning \cite{ng_gnp_2001} and makes difficult the construction of delay aware overlays. A number of heuristic approaches have been proposed in an effort to finding an algorithm that performes reasonably and consistently well.

Several protocols for dealing with the problems derived from the topology mismatch problem in decentralised unstructured peer-to-peer systems have been proposed in the literature. Each implementation has different goals but all share some common ground on the aspect of the network functionality they focus on, in order to achieve them. Thus, algorithms can be categorized into three basic types:
\begin{itemize}
  \item \emph{forwarding-based}. A peer selects only a subset of its neighbours to re-broadcast query messages. The selection is made using one or a combination of various statistical metrics. Examples of such metrics, could be the number of responses received by a neighbour or the connection latency of the link between the nodes e.t.c. These approaches enhance search efficiency but have also several drawbacks. First they drastically reduce the search scope. Expanding the search scope, on the other hand, is no easy task because the overhead of forming multicast trees is proportional to the multicast group size. Also such schemes do not consider dynamic joining and leaving of peers so they do not scale well on dynamic environments.
  \item \emph{cache based}. These approaches can be further divided into:
  \begin{itemize}
    \item \emph{data index caching}, and
    \item \emph{content index caching}.
  \end{itemize}
  Additionally caching policies can be divided into:
  \begin{itemize}
    \item \emph{centralized}, and
    \item \emph{local}.
  \end{itemize}
  Centralized peer-to-peer systems provide centralized index servers to keep indices of shared files of all peer (e.g. Napster). In the local caching policy, each or some peers maintain information about other peers (like the superpeer scheme of Kazaa). Caching can significantly reduce traffic costs and response time but duplication of messages still exist, limiting scalability of such approaches.
  \item \emph{overlay optimization}. This approach can be subdivided furtherlly into:
  \begin{itemize}
    \item \emph{Spanning tree based}. Construction of a rich graph and the build its minimum spanning tree. These approaches incur large traffic overhead to the system \cite{chu_esm_2000, chu_esm_2002}.
    \item \emph{Cluster based}. Selection to link physically closer nodes with each other. These approaches can shrink the search scope significantly while mapping accuracy is not always guaranteed.
    \item \emph{Minimum latency first}. Use of latency as a metric to calculate distance among peers. They require global latency information ``landmarks''\footnote{Measuring latency between peers and stable Internet servers.}.
  \end{itemize}
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Algorithms}
Having understood the nature of the problem that peer-to-peer architecture face, this section starts the discussion of the academic work that has been conducted in the field the last few years. What follows this introduction, does not claim to be a thorough citation of all known protocols that are available out there, but instead, a carefull selection of those that left a distinctive fingerprint contribution in the efforts of the research community to alleviate the topology mismatch problem.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Narada}
\emph{End System Multicast} \cite{chu_esm_2000, chu_esm_2001, chu_esm_2002} as well as other  proposes an architecture, where end-systems implement all multicast related functionality (including membership management and packet replication). In \emph{Narada}, the protocol developed in the paper's context, end-systems
\begin{inparaenum}[\itshape i\upshape)]
  \item self-organize, in a fully distributed manner,
  \item into an efficient overlay structure, both from the network and the application perspective,
  \item with self improving properties,
  \item adapting to network dynamics\footnote{Long-term variations in Internet path characteristics (i.e. bandwidth, latency e.t.c.)}.
\end{inparaenum}

What Narada really does is to construct overlay spanning trees in a two-step process:
\begin{enumerate}
  \item It builds a richer connected graph \footnote{ENarada \cite{li_enarada_2008} used Gossip protocol for the construction}, termed a \emph{mesh}, while trying to ensure a couple of desired performance properties:
    \begin{itemize}
      \item path quality between any pair of members should be comparable to the quality of the unicast path between them, and
      \item each member should have a limited number of neighbours on the mesh.
    \end{itemize}

  \item It computes minimum spanning trees out of the mesh, each rooted at the corresponding source member.
\end{enumerate}

Narada incrementally improves the quality of the mesh by adding and dropping overlay links. Members probe each other at random and new links are added depending on what is the gain in doing so. On the other hand, participants, continuously  monitor the utility of existing links, and drop them if they are found to be not usefull.

Narada protocol achieves relatively good performance for small and medium sized groups involving tens to hundreds of members. This is not, by no means, what a contemporary peer-to-peer system needs to be characterized as scalable, but this work is considered pioneering\footnote{Another similar approach is \emph{Scattercast}\cite{chawathe_scattercast_2000}}, in that it was the first to have conducted a detailed Internet evaluation to analyse and demonstrate the feasibility of an overlay-based, application layer, multicast architecture that can dynamically adapt to bandwidth and latency properties of the physical underlying infrastructure.

\subsection{Gia}
\emph{Gia} \cite{chawathe_gia_2003} is a decentralized unstructured peer-to-peer file sharing system with the following characteristics:
\begin{inparaenum}[\itshape i\upshape)]
  \item replaces Gnutella's flooding with random walks \cite{lv_randomwalks_2002},
  \item recognizes the importance of overlay network's topology, while using the random walks, and exploits it by incorporating a topology adaptation algorithm,
  \item introduces a token-based flow control mechanism, and
  \item detects the significant heterogeneity in peer bandwidth, processing power, disk speed e.t.c. and takes advantage of it.
\end{inparaenum}

More specifically, there are four key components in the design of \emph{Gia} which are summarized bellow:
\begin{enumerate}
  \item A \emph{dynamic topology adaptation} protocol that puts participating nodes within short reach of high capacity nodes so that these \emph{high-degree} nodes, which due to their high connectivity receive most of the queries, actually have the capacity to handle them.
  \item An \emph{active flow control} scheme is used to avoid overloaded hot-spots. Heterogeneity is detected and flow control tokens are given to nodes based on the available capacity.
  \item Every node maintains pointers to to the content that is offered by their immediate neighbours, creating a \emph{one-hop replication} of pointers, scheme.
  \item A \emph{search protocol} based on random walks, that is biased towards directing queries to high-capacity nodes that are typically best able to answer these queries.
\end{enumerate}

\subsection{Adaptive Overlay Topology Optimization}
\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{liu_auto_2003} is an algorithm for building an overlay multicast tree among each source node and its direct logical neighbours in order to alleviate the mismatching problem while providing a larger query coverage range. It includes the steps of
\begin{inparaenum}[\itshape i\upshape)]
  \item \emph{Selective Flooding (SF)}, and
  \item \emph{Active Topology (AT)}
\end{inparaenum}
which are furtherlly discussed in the following paragraphs.

\subsubsection*{Selective Flooding (SF)}
Instead of flooding to all neighbours, SF builds an overlay multicast tree among each peer and its immediate logical neighbours. This reduces overall message burden to the network by making some neighbours non-flooding. The tree is formed using a minimum spanning tree algorithm. For this reason a peer needs to know the costs to all its logical neighbours (e.g. the network delay) as well as between any pair of neighbours. Additional, cost probing, messages need to be added to the original messages defined by the Gnutella protocol.

LTM's SF effectiveness has be proven to be detached from the different physical or overlay topologies. On the other hand, SF is more effective with large number of logical neighbours. It can reach an average optimization rate of 87.4 percent on a logical topology with an average of 30 logical neighbours.

\subsubsection*{Active Topology (AT)}
At this step the overlay topology is reorganized. Each peer, independently, makes optimizations in the overlay network to alleviate topology mismatching by replacing non-flooding neighbours, with closer nodes. The approach incorporated by the paper is called \emph{Randomized AT} algorithm which picks up a candidate peer at random among the non-flooding neighbour's neighbours. Whenever a new neighbour cost table is received or there is a change of neighbours, the source peer has to re-calculate the multicast tree and apply the randomized AT algorithm.

Different numbers of average logical neighbours has little to do with the effectiveness of AT. If the source has $n$ non-flooding peers, there are $n$ potential neighbour replacements. The overhead to exhaust all $n$ possible replacements can be high, so in practice, after each replacement the source peer can decide whether it needs to find another candidate peer. This is done by computing the cost improvement ratio greater than some predefined termination threshold. The larger the threshold, the slower, in the number of optimization steps, the reduction of the normalized average distance. As a whole the average response time is significantly reduced when more optimization steps taken.

\subsection{Location-aware Topology Matching}
\emph{Location-aware topology matching} or \emph{LTM} \cite{liu_ltm_2004}, is an algorithm of building an efficient overlay by disconnecting low productive connections and choosing physically closer nodes as logical neighbours while still retaining the search scope and reducing response time of queries. This is done by issuing detector messages by peers, to a small region so that receivers of the detector can record relative delay information and cut inefficient or redundant logical links as well as add closer nodes as direct neighbours.

Three operations are defined in LTMband are furtherly discussed in the following paragraphs.

\subsubsection*{TTL2-detector flooding}
\emph{TTL2-detector} messages are designed based on the specification of the Gnutella protocol. Suppose that $d(i, S, v)$ denotes the TTL2-detector which has the message ID of $i$ with TTL value of $v$ and is initiated by $S$. In addition to Gnutella's unified 23-byte header, such a message has a body of two types:
\begin{itemize}
  \item short format $d(i, S, 1)$ if $S$ is a direct neighbour of the receiver
  \item long format $d(i, S, 0)$ if $S$ is within a two-hop distance from the receiver
\end{itemize}
This way a peer can calculate link costs from within a one- or two-hop area around it and thus reach peers in $N(S)$ and $N^2(S)$ sets, respectively, in the paper's context. Each peer floods a TTL-detector periodically while it calculates link cost to another peer by receiving other peer's detector messages and checking:
\begin{itemize}
  \item the Source Timestamp field and its own clock upon time of receipt for one-hop messages
  \item the Source Timestamp, TTL1 Timestamp and its own clock upon receipt for two-hop messages
\end{itemize}
Of course peer clocks must be synchronized in order to do such checking.

\subsubsection*{Low productive connection cutting}
There are three cases for any peer $P$ who receives $d(i, S, v)$ multiple times:
\begin{inparaenum}[\itshape i\upshape)]
  \item $P$ receives both $d(i, S, 1)$ and $d(i, S, 0)$
  \item $P$ receives multiple $d(i, S, 0)$s from different paths, and P randomly chooses to process one
  \item $P$ receives one $d(i, S, 1)$ and multiple $d(i, S, 0)$s, and $P$ processes $d(i, S, 1)$ and one randomly selected $d(i, S, 0)$
\end{inparaenum}
If the link with the largest cost is found and is a direct neighbour then the connection is put in a will-cut list and stays there for a certain period of time. If it is not, then it is handled by other peers. After that period, connections are cut and recorded to $P$'s cut-list.

\subsubsection*{Source probing}
For a peer $P\in(N^2(S) - N(S))$ who receives only one $d(i, S, 0)$, the cost of $PS$ is obtained (with list look-up or probing). Then $P$ compares it with the cost form each hop and if $PS$ has the largest cost, $P$ will not keep this connection, while otherwise the connection will be created.

\paragraph{}
Supposing $n$ is the number of peers, $c_n$ is the average number of neighbours and $c_e$ is the average cost of logical links, then in the flooding-based search the traffic incurred by one query from an arbitrary peer in a peer-to-peer network is $O(n)$. As observed in the Gnutella network \cite{sripanidkulchai_gnutella_2001}, each peer issues $0.3$ queries per minute in average, thus the per minute traffic incurred by the network with $n$ peers is $O(n^2)$. Because each $d(i, S, v)$ has a TTL of $2$ in each source peer, the traffic for one time LTM optimization in all peers is at most $2nc_n^2c_e$. If each peer uses LTM $k$ times per minute, the total traffic incurred is $2knc_n^2c_e$. Simulation shows the best value for $k$ is $2$ or $3$. So, the traffic overhead caused by LTM to the network is $O(n)$.

TTLj-detectors, with $j > 2$, would detect and break cycles with more than 4 links. LTM though, does not use such detectors because detector-flood traffic would increase significantly, and cut links between two end-peers, could cause queries initiated by them to traverse a path much more expensive than the cost on the the cut link.

LTM disadvantages are
\begin{inparaenum}[\itshape i\upshape)]
  \item disagreement of measured delay due to unsynchronized clocks causes problems when deciding the cut positions, which can influence the network connectivity, and
  \item the network delay metric mainly focuses on disabling the connections between peers physically far away without considering the shortcuts created by powerful peers.
\end{inparaenum}

\begin{figure}
\centering
  \includegraphics[scale=0.4]{img/ltm_multid.jpeg}
\caption{Peer $P$ receives $d(i, S, v)$ multiple times}
\label{figure:ltm_multid}
\end{figure}

\begin{figure}
\centering
  \includegraphics[scale=0.4]{img/ltm_example.jpeg}
\caption{A full example of LTM}
\label{figure:ltm_example}
\end{figure}

\subsection{Scalable Bipartite Overlay}
\emph{Scalable Bipartite Overlay (SBO)} \cite{liu_bipartite_2007} employs an efficient strategy to select query forwarding paths and logical neighbours. The topology construction and optimization of SBO consist of four phases:
\begin{enumerate}[\itshape i\upshape)]
  \item When a new peer is joining the network, it will randomly take an initial colour; say red or white. This colour is not to be changed for the time the peer stays connected. If it leaves and then rejoins it will, again, have to pick a colour, randomly. Thus all peers are separated into two groups, red or white. Then the bootstrap host will provide, the joining peer, with a list of active peers, including their colour information, in order for the later to establish connections to different colour peers. This way, all peers form a bipartite overlay.
  \item White peers probe distances with their immediate (red) neighbours, form a cost table and send this table to their corresponding red neighbours.
  \item Each red peer builds a minimum spanning tree using the obtained neighbour cost tables. These are two-hop diameter trees so white peers do not need to build one. The links that are part of a minimum spanning tree are called \emph{forwarding connections (FC)} while the rest \emph{non-forwarding (NFC)}.
  \item Having a minimum spanning tree with two hops a red peer is able to send its queries within that range. Some white peers, though, have become non-forwarding neighbours. In this phase, such a (white) neighbour will try to find another red peer being two hops away from its current red neighbour to replace the later as its new neighbour.
\end{enumerate}

\begin{figure}
\centering
  \includegraphics[scale=0.4]{img/sbo_efficient_forward.jpeg}
\caption{A red peer $P$ has a small overlay topology of $N(P)$ and $N^2(P)$ and computes the efficient forwarding paths}
\label{figure:sbo_efficient_forward}
\end{figure}

\begin{figure}
\centering
  \includegraphics[scale=0.4]{img/sbo_neighbour_replace.jpeg}
\caption{Neighbor replacement}
\label{figure:sbo_neighbour_replace}
\end{figure}

In a static enviromnent LTM may reduce traffic cost by around 80 to 85 percent while SBO reduces traffic cost between 85 and 90 percent. However, LTM  is proved to converge in around 2-3 steps while SBO needs 4-5 steps. Moreover LTM reduces response time by more than 60 percent in 3 steps while SBO needs 8. In a dynamic environment (10 minute average peer lifetime, 0.3 queries/sec by each peer) SBO and LTM reduce the average traffic cost per query (including the overhead due to the optimization steps) by 85 and 80 percent, respectively. Moreover LTM reduces the response time per query to 30 percent while SBO to 35 percent.

\subsection{Adaptive Connection Establishment}
\emph{Adaptive Connection Establishment (ACE)} \cite{liu_ace_2004} builds an overlay multicast tree among each source node and the peers within a certain diameter from the source peer and optimizes the neighbour connections that are not in that tree.

ACE indicates three phases for its algorithm:
\begin{enumerate}[\itshape i\upshape)]
  \item Calculate cost between nodes using network delay as a metric. Each peer probes the costs with its immediate logical neighbours and forms a \emph{neighbour cost table (NCT)} using a special routing message type. Two neighbouring peers exchange their NCTs in order for every peer to obtain the cost between any pair of its local neighbours forming a small overlay topology.
  \item Based on obtained NCTs a minimum spanning tree among each peer and its immediate neighbours is built (Figure~\ref{figure:ace_phase2}).
  \item Physically far away neighbours are replaced by physically close neighbours. In ACE a peer, say $S$, probes the distance between one of its non-flooding neighbour's neighbour, say $G$ and $H$ respectively. If the link to neighbour's neighbour is smaller than that to the neighbour, the later is cut. If this is not the case but $S$ finds that cost of $GH$ is even larger than that of the $SH$, $S$ will keep $H$ as a new neighbour. Obviously, if $SH$ is larger than $SG$ and $GH$, the connection will not be established and $S$ will continue probing another neighbour's neighbour. The algorithm described is conducted within $1$-neighbour closure (among its source peer and all its direct neighbours) but the optimization scope can be enlarged. The larger the scope, the better topology matching improvement but also the greater the computational overhead (Figure~\ref{figure:ace_phase3}).
\end{enumerate}

Simulations in \cite{liu_acesims_2004} show that the average scope of each query to cover the same scope of nodes is reduced by about 65 percent without losing any autonomy feature, while the average response time can be reduced by 35 percent. Larger diameter topologies lead to better topology optimization rate but also to higher communication and computation overhead. It was also found that it is more effective in higher connectivity dense topologies. Compared to LTM, it comes short of convergence speed. In \cite{ni_mismatch_2004} shows reduction of both total traffic (90 percent) and response time (80 percent) to message queries without shrinking the search scope. SBO, on the other hand, achieves approximately 85 percent reduction on traffic cost and about 60 percent reduction on query response time. Last but not least, it is concluded that work must be done on incorporating a more sophisticated selection policy for candidate non-flooding peers.

\begin{figure}
\centering
  \includegraphics[scale=0.4]{img/ace_phase2.jpeg}
\caption{Second phase in ACE}
\label{figure:ace_phase2}
\end{figure}

\begin{figure}
\centering
  \includegraphics[scale=0.4]{img/ace_phase3.jpeg}
\caption{Third phase in ACE}
\label{figure:ace_phase3}
\end{figure}

\subsection{Hops Adaptive Neighbour Discovery}
Paper \cite{chen_hand_2006} proposes a new algorithm named \emph{Hops Adaptive Neighbour Discovery (HAND)} which uses a fully distributed triple hop adjustment strategy to address the topology mismatch problem.

The advantages of the algorithm compared to other approaches are that
\begin{inparaenum}[\itshape i\upshape)]
  \item it does not need any clock synchronization,
  \item it is a fully distributed algorithm making it robust and reliable in decentralized systems,
  \item the traffic overhead of the triple hop adjustment is very low,
  \item it is applicable to dynamic peer-to-peer environments, and
  \item maintains lower query response time.
\end{inparaenum}

The algorithm's ultimate goal is the optimal overlay, the \emph{Logical Communication Network (LCN)} in the paper's context. The key concept of the algorithm is that a graph $G^{*}$ that describes an LCN and a graph $G$ that describes the current overlay are matched only if all peer hop sequences $(v_1, v_2, \ldots, v_k)$ in $G$ exist in $G^{*}$ and in the same order. In practice triple sequences $(v_1, v_2, v_3)$ are used.

The mismatching detection is done in the following way. Suppose we want to verify peer sequence $v_2-v_1-v_3$ (see Figure~\ref{figure:hand_com_overlay}). A pair of probing messages are sent from $v_1$ to $v_2$ and $v_3$. Suppose delays of $(v_1,v_2)$ and $(v_1,v_3)$ are are denoted as $x$ and $z$, respectively. When the probing message arrives to $v_2$ it forwards it directly to $v_3$. Similarly, when the probing message arrives to $v_3$, it forwards it directly to $v_2$. These last steps are performed in order to obtain delays of $(v_2,v_3)$ and $(v_3,v_2)$ physical paths, respectively, denoted by $y$.
\begin{itemize}
  \item If $y=z-x\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched and should be adjusted to $v_1-v_2-v_3$ by deleting edge $(v_1,v_3)$ and adding a new $(v_2,v_3)$ (see Figure~\ref{figure:hand_matchtriple_lemma3}).
  \item if $y=x-z\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched and should be adjusted to $v_1-v_3-v_2$ by deleting edge $(v_1,v_2)$ and adding a new $(v_3,v_2)$ (see Figure~\ref{figure:hand_matchtriple_lemma4}).
\end{itemize}
$\varepsilon$ is a small positive real number denoting additional delays caused by possible forwarding and jitter delays.

\begin{figure}
\centering
  \includegraphics[scale=0.4]{img/hand_com_overlay.jpeg}
\caption{The communication in overlay}
\label{figure:hand_com_overlay}
\end{figure}

\begin{figure}
\centering
  \includegraphics[scale=0.4]{img/hand_matchtriple_lemma3.jpeg}
\caption{The matching triple in paper's Lemma 3}
\label{figure:hand_matchtriple_lemma3}
\end{figure}

\begin{figure}
\centering
  \includegraphics[scale=0.4]{img/hand_matchtriple_lemma4.jpeg}
\caption{The matching triple in paper's Lemma 4}
\label{figure:hand_matchtriple_lemma4}
\end{figure}

Measurements conducted for evaluation perposes showed that in a static environment the algorithm can effectively decrease traffic cost by about 77 percent and shorten the query response time by about 49 percent in less than two minutes. In a dynamic environment it shows similar behaviour and with the size of the overlay network having little impact on the effectiveness of the algorithm. Compared to LTM both algorithms have almost the same traffic reduction rate, however on the response time reduction rate HAND has a higher one by about 4 percent. The traffic overhead of HAND is much less than that of LTM by an average of 55 percent.

\subsection{Distributed Cycle Minimization Protocol}
\emph{Distributed Cycle Minimization Protocol (DCMP)} which is introduced in \cite{zhu_dcmp_2008} is a dynamic, fully decentralized protocol that promises significant reduction of duplicate messages. To achieve that, it uses the slightly different approach of eliminating unnecessary cycles, while retaining the connectivity of the network and preserving fault resilience and load balancing properties of unstructured peer-to-peer schemas by avoiding the creation of a hierarchical organization.

DCMP aims at cutting the cycle paths at strategic locations. Any peer that detects a duplicate message can initiate the cutting process which consists of the following steps:
\begin{enumerate}
  \item Peers in the cycle elect a leader called \emph{GatePeer}\footnote{GatePeers are important for maintaining the connectivity and optimal structure of the network while peers join and leave randomly.}.
  \item The cycle is cut at a well-defind point with respect to the GatePeer
\end{enumerate}

The first step after detecting a duplicate message by some peer is to gather information from all peers in the cycle using a new type of control message called \emph{Information Collecting Message} or \emph{ICM}. ICM contains:
\begin{inparaenum}[\itshape i\upshape)]
  \item a \emph{GUID}\footnote{Globally Unique IDentifier assigned to every query message generated by any node.} field same as the one of the duplicate message,
  \item \emph{DetectionID} field which represents the direction of the connection where the duplicate was identified\footnote{This ensures the uniqueness of the ICM messages because as it travels through many cyclic paths, multiple peers will detect the duplicates and initiate an ICM message.}, and
  \item \emph{Node Information Vector (NIV)} which contains information (bandwidth, CPU power, etc) about peers that propagated the ICM.
\end{inparaenum}

Suppose $A$ detected the duplicate as depicted in Figure~\ref{figure:dcmp}. It then emits an ICM to $B$ and $F$, that initially contains information only about itself. Each peer that receives the ICM, appends its information and propagates it along the reverse path of the original message. Since two copies of ICM are sent, at some point, a peer, say peer $D$, will receive a duplicate ICM. Using the information in the NIVs of the ICMs, $D$, decides to cut (for example) the EF connection. To inform the other peers about its decision, it emits a \emph{Cut Message (CM)} which contains the GUID and DetectionID of the corresponding ICM and an additional field that identifies the connection to be cut. $D$, then, forwards the CM in the reverse directions from where the ICM arrived. Similarly CMs received by any peer are propagated toward the reverse path of the corresponding ICM. Eventually, either peer $E$ or peer $F$ will receive the CM and cut the connection, thus eliminating the cycle.

Receiving a duplicate ICM denotes the existence of a cycle. The opposite is not true though. For example, if the cycle contains $2 \times TTL$ edges, it will not be detected, because the ICM messages will be discarded before they locate it. There is a tradeoff between preserving the connectivity of the network and minimizing the duplicates that makes such a possibility to be safely ignored.

\begin{figure}
\centering
  \includegraphics[scale=0.4]{img/dcmp.jpeg}
\caption{Cycle elimination methods in DCMP}
\label{figure:dcmp}
\end{figure}

Experiments in \cite{zhu_dcmp_2008} showed that DCMP incurs a lower delay, returns more results and decreases the number of duplicate messages by 22\%, compared to LTM. Additionally DCMP has one to two orders of magnitude less overhead, because it adopts a more ``LAZY'' approach than broadcasting control messages periodically like LTM does.

\subsection{Two-Hop-Away Neighbour Comparison and Selection}
Work in \cite{liu_thancs_2005, liu_thancs_2008} proposes a distributed heuristic called \emph{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} that
\begin{inparaenum}[\itshape i\upshape)]
  \item is completely distributed and needs no global knowledge,
  \item presents trivial overhead compared to the query cost savings
  \item its convergent speed of the algorithm is fast enough (faster than minimum spanning tree approaches) so that is effective to dynamic environments, and
  \item does not shrink the search scope.
\end{inparaenum}

THANCS is considered a \emph{local search method}, in the sense that it targets in finding a locally optimum solution, by exploiting knowledge within a 2-hop radius. The algorithm consists of two main components: \emph{piggybacking neighbour distance on queries} and \emph{neighbour comparison and selection} which are furtherly discussed bellow.

\paragraph*{Piggybacking neighbour distance on queries.}
Using network delay as a metric for measuring the distance, each peer probes distances with its immediate naighbours and stores information locally. For this reason a special query message type, \emph{Piggy Message (PM)}, is introduced. It is 6 bytes long and includes two fields: Neighbour IP Address and Neighbour Distance. A peer $P$ constructs a PM for its neighbour $Q$, which contains $Q$'s IP address and $Q$'s distance from $P$. When $P$ receives a query from $Q$, this PM will be piggybacked by the query that goes to all other neighbours of peer $P$. Upon receiving such a query message, each of the other neighbours will detach the PM, record the $PQ$ distance and process the query. This PM will not be further forwarded. In selecting which incoming queries  should piggyback a PM the paper proposes the \emph{pure propability-based (PPB)} and the \emph{new neighbour triggered (NNT)} policies.

\paragraph*{Neighbour comparison and selection.}
Figure~\ref{figure:thancs} illustrates this component of the THANCS algorithm. A peer $S$ probes the distance to all known unprobed $N^2(S)$\footnote{$N^2(S)$ denotes the set of peers being two hops away from $S$, while $N(S)$ denotes the set of direct logical neighbours of $S$.}. The distance of $SP$ is known to $S$. Upon receiving a PM from node $P$ with the distance of $PQ$, $S$ follows one of the following:
\begin{itemize}
  \item $Q \in \left( N(S) \cap N^2(S) \right)$, i.e. $Q$ is direct neighbour of $S$. In this case $S$ will compare cost of $SQ$, $SP$ and $PQ$. If the most costly connection is one of $SQ$ or $SP$ the corresponding link will be put into a \emph{will-cut list}\footnote{Links are not immediately disconnected when put in the will-cut list. There are useless for forwarding but are kept active, for some time, in order to serve query responses that are traveling to the source peer along the inverse search path.}. If the most costly connection is $PQ$ then $S$ will do nothing, as the fully distributed nature of the algorithm will give a chance to either $P$ or $Q$ to cut the connection.
  \item $Q \in \left( N^2(S) - N(S) \right)$, i.e. $Q$ is a two-hop-away neighbour of $S$. If $S$ hasn't probed $Q$ before\footnote{A distance cache is maintained and looked up in such a case.}, it probes and stores the result in the \emph{distance cache}. Having the distance of $SQ$, $S$ compares costs of $SQ$, $SP$ and $PQ$. If $SQ$ is the most costly, $S$ will not establish the connection. If $SP$ is the most costly, $S$ will establish connection $SQ$ and put $SP$ in the will-cut list. If $PQ$ is the longest, $S$ will keep  the connection with both $P$ and $Q$, expecting that $P$ or $Q$ will eventually disconnect link $PQ$, later.
\end{itemize}

\begin{figure}
\centering
  \includegraphics[scale=0.4]{img/thancs.jpeg}
\caption{Probing two-hop-away neighbours}
\label{figure:thancs}
\end{figure}

In a static environment THANCS has been proven to be effective; optimizing 45 percent out of the 60 percent of mismatched paths, constructing a nearly optimal overlay. This leads to a 60 percent reduction in traffic cost as well as a 40 percent decrease in query response time. In dynamic environments (Gnutella 0.6/Limewire super-peer-like and Ion flat-like), THANCS saves up to 70 percent of the traffic cost in the super-peer topology and 55 percent for the flat one. Average response time is also decreased by 60 and 45 percent, respectively. Generally, THANCS has similar performance to LTM, without needing synchronization. SBO, incurring half the  overhead of AOTO, reduces the traffic cost the most, while THANCS has lower response time and converges faster than SBO. THANCS is, thus, more suitable for a more dynamic environment. In addition, THANCS is easy to implement and its operation overhead is trivial, compared with the other three approaches. This design, however, has the limitation of not being easily extend to also support non-flooding-based systems.

\subsection{mOverlay}
Zhang et al. in \cite{zhang_moverlay_2004} focus on two aspects of the overlay construction. First, to achieve \emph{efficiency} of communication, a protocol should minimize unecessary long-distance hops and redundant traffic. Second, to be \emph{scalable} the overlay should be constructed in a distributed fashion and maintenance cost (data and locality management) should be minimized. Both of the above observations, are positively affected by the exploitation of node proximity at the underlying network. To accomplish it, the paper introduces the \emph{group} concept, according to which non-static (dynamic) landmarks are used to compute proximity, resulting in a highly robust and scalable overlay construction with reduced maintenance cost as well. The dynamic nature of landmarks help towards load-balance, meaning that hot-spots are avoided.

The proposed architecture, called \emph{mOverlay}, is a \emph{clustering approach} which creates a two-level hierarchical network, where on the top level we have connections between groups while on the bottom we have connections between hosts inside groups, thus attempting to recreate \emph{Small-World}-like properties for the constructed overlay network. The notion of a \emph{group}, in the paper's context, is a set of hosts that are close to each other with respect to any position $P$ in the underlying network. The distance between hosts can be 
\begin{inparaenum}[\itshape i\upshape)]
  \item network latency,
  \item round-trip time,
  \item minimum bandwidth on the links along the path connecting the two nodes, or
  \item some other user-defined cost metric between the two nodes.
\end{inparaenum}
These nodes are refered to as neighbouring. Similarly groups can exchange messages with their neighbouring ones\footnote{Groups nearby, in the underlying physical network.}.

According to the \emph{grouping criterion} set by the authors, a new host $Q$ belongs to some group $A$ if the distance between $Q$ and group $A$'s neighboring groups is the same as the distance between group $A$ and it's neighbouring groups. The neighbour groups thus, play the role of the dynamic landmarks and for accuracy and performance reasons the nearest of them are chosen to play that role.

\paragraph{Locating process.} A new coming host, $Q$, first connects to a globally known host cache called the \emph{rendezvous point (RP)} in order to retrieve the starting point in the overlay, say $A$ in group $1$. Host $Q$ then, measures its distance to host $A$. At the same time, the later, sends information about the neighbour groups of group $1$ back to host $Q$. This list is called \emph{candidate group list}, and the newcoming host sequnentially measures its distance to each of them in seek for the closest one. If the \emph{grouping criterion} is met, host $Q$ belongs to group $1$. If not, a boot host from the closest group is found and the algorithm is re-run until the criterion is met or after a predefined number of repetitions. In the later case, $Q$ creates a new group comprising itself only. The above protocol does not favour hotspots as it spreads the probability of visiting a group across the whole overlay and limits the overhead in the level of $O \left ( log N \right )$.

\paragraph{General overlay operations.} A set of additional protocols, are also introduced, similar to those found in traditional unstructred networks, but modified focusing on scalability and robustness. For example a protocol for \emph{group formation} is introduced that exploits the inherent characteristic of proximity, in the overlay, in order to efficiently detect the neighbouring groups of a newly formated group from the set of adjasent groups of its closest neighbour. Additionaly, during \emph{group joining} the coresponding protocol denotes the exchange of important information for group maintenance. This can be furtherly improved by \emph{information sharing} between nodes of the same group, functionality handled by a dedicated flood-like protocol\footnote{Since nodes that belong to the same group are physically close this can be achieved at a minimum price.}. Moreover, another set of distributed protocols handle the \emph{information update}. The information that needs update, in the proposed architecture, is
\begin{inparaenum}[\itshape i\upshape)]
  \item the host cache, when a new node joins, and
  \item the neighbours of groups, when a close-by group is generated.
\end{inparaenum}
Finally, in case of \emph{host failure} or \emph{host departure} the system is able to maintain its stability since there are defined operations for periodical host cache update and group leader selection if one leaves or dies.

\subsection{Distributed Domain Name Order}
\cite{zeinalipour-yazti_ddno_2005} proposes the \emph{Distributed Domain Name Order (DDNO)} technique which makes unstructured overlay networks, topologically aware. It results in a flat overlay topology but with some changes it can be utilized in \emph{superpeer} environments. In DDNO, a node of degree $d$, tries to connect to $\frac{d}{2}$ nodes that belong to the same domain (\emph{sibling} connections) and another $\frac{d}{2}$ of random nodes (\emph{random} connections). Connecting to \emph{sibling} nodes ensures the reduction of long distance message travelling and thus resulting to better performance, while \emph{random} nodes keep the structure connected.

\subsubsection*{Discovering \emph{random} neighbours}
Initially, when a new coming node wants to join a network $N$, uses a \emph{hostcache} mechanism to provide it with a list of nodes with which to establish its $\frac{d}{2}$ \emph{random} connections.

\subsubsection*{Discovering \emph{sibling} neighbours}
In order to discover the rest $\frac{d}{2}$, \emph{sibling}, nodes, the \emph{lookupDN} procedure is initiated. In this phase, a special $l$ walker message is multicasted, by the newcomer, in search of a node that knows\footnote{A special structure is used called ZoneCache that contains information about which nodes are reachable in an $r$-hop radius.} how to guide $n$ to a \emph{sibling} node.

Figure~\ref{figure:ddno_lookupdn} shows an $l$ message emitted by node $n$ which travels along the path [$a$, $b$, $c$, $e$, $b$, $d$]. At node $d$, $l$ finds information to make a decision on which neighbour to follow next\footnote{Contrary to the random choices, made at the previous hops.}. Ultimately $l$ arrives at $m$, which is a \emph{sibling} node to $n$. $m$ then issues a broadcast message to all its own \emph{siblings}. Each of the receiving nodes, $m$ included, will respond to $n$ with a designated $l^{'}$ message if it is willing to accept new connections and out of these answers, node $n$, will attempt to establish its, remaining, $\frac{d}{2}$ \emph{sibling} connections.

Unfortunately, use of DDNO results in networks which have a uniform distribution of node degrees, losing the valuable properties of Power Law and Small World networks, as mentioned previously in this survey.

\begin{figure}
\centering
  \includegraphics[scale=0.375]{img/ddno_lookupdn.jpeg}
\caption{Domain name look-up in a \emph{DDNO} topology}
\label{figure:ddno_lookupdn}
\end{figure}

\subsection{Critical Topology-Aware Grouping}
\emph{Critical Topology-Aware Grouping (CTAG)} presented in \cite{zhao_ctag_2006}, is a grouping algorithm that tries to exploit low-cost and low-delay communication of physically close-by peers. The grouping strategy is based on the \emph{IANA}\footnote{Internet Assigned Numbers Authority} and the respective \emph{Regionanal Internet Registry (RIR)}'s IP assignment strategies, according to which nodes within the same organization are always addressed from the same block. The paper proposes the \emph{Adjacency Measurement (AM)} technique which uses the longest matching IP segment criterion to calculate node proximity.

Observations made in various studies (e.g. \cite{matei_mapgnutella_2002}) concerning the distribution of nodes among \emph{Internet Service Providers (ISPs)} and \emph{Autonomous Systems (ASs)} have shown that only $2$ to $5$ percent of Gnutella connections link peers within a single AS while more than $40$ percent of all Gnutella peers are located within the top 10 ASs. Similarly, measurements in \cite{zeinalipour-yazti_gnudc_2002} used a $244,000$ IPs test-bed and results have shown that $45$ percent of the nodes belonged to only $10$ large ISPs and $58$ percent belong to only $20$. Such results mean that most overlay generated traffic crosses AS borders increasing topology mismatch cost.

\emph{CTAG} focuses on both the construction of the overlay as well as dynamically revising it during node interaction, phases called \emph{bootstrapping grouping} and \emph{dynamic revision}, respectively in the paper's context.

\subsubsection*{Bootstrapping grouping}
The \emph{Gnutella Web Caching} mechanism has been modified in order for a new coming node to choose the closest \emph{GWC} in order to retrieve the node list for bootstrapping.

\subsubsection*{Dynamic revision}
Similarly to the bootstrapping phase, \emph{AM} metric is used to store hosts' addresses. read from \emph{X-Try} headers during handshake or from \emph{QueryHit} messages. Additionally, when a node reaches the max neighbour connections, node disconnects the neighbours with the lowest \emph{AM}.

\subsection{Peer-exchange Routing Optimization Protocols}
Work presented in \cite{qiu_prop_2007} introduces \emph{peer-exchange}\footnote{A series of exchanges of neighbours between two peers. One exchange can be viewed as a pair of cut-add operations.} as a basic operation to adaptively adjust the connections of the overlay network, and efficiently reduce the average logical link latency of the whole system. There are two points that differentiate this scheme from others:
\begin{inparaenum}[\itshape i\upshape)]
  \item it utilizes the collaboration of two peers to optimize their neighbourhood environment, than simply letting each node to ``selfishly'' choose its own strategy, and
  \item can be deployed effortlessly on both unstructured and structured peer-to-peer systems.
\end{inparaenum}

\emph{PROP-G}\footnote{\emph{G} stands for \emph{generic}.} is the simple and direct variant of \emph{peer-exchange} in which two neighbours swap all their neighbours. As depicted in figure~\ref{figure:prop-g} the mechanism can be viewed as, nodes, exchanging overlay ``positions''. Thus, the overall overlay topology is not affected by the this operation.

\begin{figure}
\centering
  \includegraphics[scale=0.4]{img/prop-g.jpeg}
\caption{PROP-G in which a neighbours are exchanged.}
\label{figure:prop-g}
\end{figure}

The alternative protocol proposed, called \emph{PROP-O}\footnote{\emph{O} stands for optimized}, ensures that the degree of each participating node remains the same by selectively choose the same number of neighbours for the exchange. With this mechanism, 2 nodes exchange the same number of neighbours in order not to change the degree

\begin{figure}
\centering
  \includegraphics[scale=0.4]{img/prop-o.jpeg}
\caption{PROP-O exchanging \emph{m} neighbours.}
\label{figure:prop-o}
\end{figure}

\subsection{T2MC (unstrucured)}
\emph{T2MC}\cite{shi_t2mc_2008} exploits some properties of the Internet paradigm and clusters nodes belonging to the same ISP without any centralised control or predefined system parameterization. The algorithm considers the dynamic nature of peers and exploits the stable nature of routers in order to build a topology location relationship among end peers.
Some special routers split the physical network into autonomous system domains. Using a Traceroute mechanism T2MC searches for latency leaps among the path to a host in order to form ``near'' and ``remote'' router clusters. This is achieved by a 2-Means Classification, defined by the following steps:
1.the peer chooses the minimum and maximum Latency results from the Traceroute for initializing the cntroinds of two sets ``first'' and ``second''.
2.The peer calculates, for all hops allong its tracerouted path, the absolute distance to the centroids of both sets and assigns the routers to that centroid with which it has the smaller absolute distance.
3.The peer calculates the latncy mean and variance value of two sets
4.If the variance is larger than a predefined threshold then the algorithm takes a loop from step 2 picking the two latency mean values as new centroids of sets ``first'' and ``second''. Ultimately peer will end up with two sets having the minimum intra-set variance.

Finally the peer chooses the router from ``second'' set with the minimum hops attribute and sets it as a threshold. The selected router and all others whose hops attribute is larger than the threshold are classified as ``remote'' router cluster. The remaining are classified as ``near''. From the ``near'' class, the peer chooses the one with maximum hops attribute as its edge router, and registers it along with the all the ``near'' cluster into the DHT of the p2p overlay. As new peers join the network, those that share the same edge router or any of the members of the ``near'' router clustersm ther would gather to form a ``close'' peer cluster. As Edge routers can provide more valuable information than other members of the ``near'' set, T2MC was designed to prioritize interaction of peers and edge gateways.
TODO: figure t2mc.jpeg (a) stars represent edge routers (b) hop leaps
The use of Traceroute as a tool for implementing the distance measuring infrastructure raise concearns about its efficiency and scalability. Being, primarilly, a network diagnostic utility, it is concearned too heavy weighted and intrucive for use in a larger scale\cite{ratnasamy_binning_2002}. Additionally, disabling ICMP is a common administrative policy for edge sites to enforce security, while dumping BGP routing tables\cite{krishnamurthy_bgpclust_2000} is not directly available to the application layer.

\subsection{Unnamed Unstructured!!}
(TODO: to be reviewed)
In \cite{hsiao_redblue_2009}, Hsiao et al, claim to construct topology-aware unstructured overlays that \emph{guarantee} performance qualities in terms of
\begin{inparaenum}[\itshape i\upshape)]
  \item the expected communication latency among any two overlay peers regardless of the network size, and
  \item the broadcasting scope of each participating peer.
\end{inparaenum}

The algorithm constructs an undirected graph $G = \left( V, E \right)$ comprised by two subgraphs. The first, namely $G^{\left( red \right)} = \left( V^{\left( red \right)}, E^{\left( red \right)} \right)$ in the paper's context, includes all vertices of $G$ and ensures the connectivity of the graph by securing at least one path between any two nodes. In contrast, $G^{\left( blue \right)} = \left( V^{\left( blue \right)}, E^{\left( blue \right)} \right)$, contains those vertices of $G$ that have free edges to link to other nodes and because these are fully utilized, the following also stands $E = E^{\left( red \right)} \cup E^{\left( blue \right)}$.

A joining peer $u$, partitions its neighbours into two subsets, the $B_u^{\left( red \right)}$ and $B_u^{\left( blue \right)}$. In order to populate the $B_u^{\left( red \right)}$ subset, peer $u$ samples peers uniformly and at random. Then, each of these selected peers discovers a routing path starting from itself towards the node with the smallest (or the largest) ID in the system.

\section{Structured overlays}
\emph{Structured} overlay networks provide a self-organising infrastructure for large-scale peer-to-peer applications. They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes through a deterministic mechanism. Some well-known implementations include CAN\cite{ratnasamy_can_2001}, Chord\cite{stoica_chord_2001}, Pastry\cite{antony_pastry_2001}, Tapestry\cite{zhao_tapestry_2001} and Kademlia\cite{maymounkov_kademlia_2002} which have been used in a wide range of distributed applications (TODO: cite some of these applications i.e OceanStore Far).

The main advantage of the DHT approach of structured p2p overlays over unstructured ones, is that they provide a guaranteed bound on the number of overlay routing hops that have to be taken in order for an object to be located. (TODO: provide the O complexity for each of the CAN($O(d(N^{\frac{1}{d}}))$), Chord($O(\log N)$), Pastry($O(\log N)$), Tapestry($O(\log N)$), Kademlia($O(\log N)$)).

Typically, node IDs are assigned \emph{randomly}, resulting in excellent load balancing, scalability and robustness. Unfortunatelly, this randomness has a negative impact on the \emph{routing locality} of the network. This means that even though the target node can be reached with logaritmic overlay hops, the distance traveled in the physical underlying network, during the overlay routing process, can be far from optimal. The ratio of the phyical distance, a query travels through the overlay to an object, and the minimal distance to that object (i.e. though IP) is known in the literature as \emph{stretch} or \emph{Relative Delay Penalty (RDP)}.

Xu et.al in \cite{xu_globstate_2003} state that there are three ways generaring proximity information
\begin{itemize}
 \item Expanding Ring Search,
 \item Heuristics, and
 \item Landmark Binning
\end{itemize}

Expanding ring search can be of two forms. First it can utilize the multicast infrastracture in the underlying network in order to emit its messages. Unfortunately such infrastructures are node widely deployed thus the implementations of this way of generating proximity information is limited to blindly flood the neighbourhood to obtain the reasonable results.

Heuristics are used in order to reduce the blindness of the expanding ring search and make realizations more efficient and effective. Unfortunatelly a common problem of all heuristic approaches is the local minimum pitfall in which the search might be caught into.

Landmark clustering is based on the view that nodes with similar distances to a set of predifined, well-known landmark nodes are pretty likely of being close to each other. But this approach has its weeknesses as well, such as the fact that is a rather coarse grained approximation, therefore not particularly well suited for differantiating nodes within close distance to each other.

Castro et.al in \cite{castro_proximitydht_2002, castro_topawareroute_2002} on the other hand, comment on the three basic approaches in exploiting proximity in DHT protocols suggested by Ratnasamy et.al. in \cite{ratnasamy_openq_2002}. These are:
\begin{itemize}
 \item Geographic Layout,
 \item Proximity Routing, and
 \item Proximity Neighbour Selection
\end{itemize}

According to the first, the node IDs are assigned in such a way, that nodes close by in the psysical network topology, be close in the node ID space as well. Implementations, that work relatively well with this approach, have been incorporated into CAN. Nodes measure the RTT between themselves and a set of landmarks in order to match the CAN space as much as possible to the psysical one. Unfortunately, the use of landmarks renders the protocol which uses it, not fully self-organizing which can furtherly lead to imbalanced node distribution. On other DHTs, such as Chord or Pastry, another problem emerges. To gain fault tolerant properties, these protocols, replicate key-value pairs on neighbouring (in the ID space) nodes. When a proximity-based node ID assignment has been used, the needed failure resilience is undermined by the fact that close by nodes are more likely to suffer collective failures.

Proximity routing does not require routing tables be built using any knowledge about network proximity. On the other hand it exploits such knowledge in order to choose the best next hop during routing a message. This approach dangerously balances between choosing the node that will further progress the routing towards the destination and choosing the closest entry in the routing table, in terms of network proximity. Thus, it is relatively less effective than geographical layout when applied to CAN(-like) implemetations. Moreover, the technique has been incorporated into a version of Chord (TODO: PChord | additioanlly check wheather this is true) but seems to increase the overhead of node joins and the size as well as maintenance cost of finger tables. \cite{dabek_cfs_2001} proposes a server selection scheme for the Chord DHT, on the domain of proximity routing selection. In \emph{CFS}, each node predicts the entire lookup latency as a function of the total number of nodes and the average overlay next routing peer. The problem is that it is very difficult to have a clear picture on the total number of nodes and the average hop latency from the local. This leads to rough estimations that concequently decreases overall performance.

Finally, the third approach, constructs the routing tables with proximity knowledge count in. Tapestry and Pastry's mechanisms of routing table maintenance try to minimize the distance to nodes appearing in a peer's routing table. The fact that routing is based on longest node ID prefix match there is not such thing as a message not progressing with each routing hop. \cite{castro_proximityp2p_2002} argues on how Pastry exploits proximity neighbour selection in order to create a scheme that is (more) location-aware compared to the other well-known DHTs (CAN, Chord).

\subsection{Landmark Binning}
The scheme proposed in \cite{ratnasamy_binning_2002} is based on the process of \emph{binning} close by nodes (in terms of network latency) into the same cluster. The authors set the following objectives for designing their proposed algorithm:
\begin{enumerate}
 \item simplicity (minimal support from any measuring infrastructure),
 \item scalablility (no global knowledge of the network),
 \item complete distribution (nodes need no communication-cooperation).
\end{enumerate}

\paragraph*{Binning scheme} The implementation requires a set of well-known \emph{landmark} machines spread accross the Internet. Every newly arriving node measures its distance from these landmarks and unilaterally decides to join a specific bin based on these results. In more detail the node measures its round-trip time to each of the landmarks and orders these measurements in a decreasing order. The ordering represents a ``bin'', in the sence of close-by nodes having the same landmark ordering and hance belong to the same ``bin''. This means that a landmark system consisting of $m$ such nodes results in $m!$ potentional different bins.

The algorithm can be considered scalable, as nodes need only compute distances to a small number of predefined nodes and thus without exchanging any information. A potential bottleneck could be the extra load that this ``ping''-like scheme imposes to the landmarks, especially when we need instant reaction from our topology when dealing with the dynamic nature of the p2p networks.

To answer the question as to whether the algorithm actually contributes possitivelly to the construction of an enhaned overlay, the paper defines the \emph{gain ratio} as the factor by which the latency reduces when someone communicates with a random node from the same bin than with one not in the bin. This is implemented with an inter-bin to an intra-bin latency ratio.

\paragraph*{Structured and Unstructured Binning} The \emph{binning} scheme can be incorporated within either a structured or an unstructured overlay construction algorithm. The authors provide an example for both these cases.

Assuming a structured approach based on CAN\cite{ratnasamy_can_2001} and $m$ landmark nodes. The coordinate space is then partitioned into $m!$ equally sized portions, each corresonding to a single ordering of the landmarks. To do this, the first dimension is divided into $m$ areas each of which is furtherly divided (second dimension) into $m - 1$ sections and so on. Having set this $m$ dimensional space, at joining time, a node measures the delay to the set of landmarks in order to determine its associated bin and thus position itself in that portion of the coordinate space associated with its landmark ordering. Even though this scheme can

For unstructured overlays the paper assumes \emph{a set of $n$ nodes where each node picks any $k$ neighbour nodes so that the average routing latency on the resultant overlay is low (assuming shortest path routing)}. According to the proposed heuristic algorithm called \emph{BinShort-Long}, a node picks its neighbours by choosing its $\frac{k}{2}$ closest\footnote{If the node's bin is not large enough for it to pick these $\frac{k}{2}$ neighbours, it picks the required nodes from the bin that matches the most in terms of landmark ordering.} ones (named \emph{short links}), using the \emph{binning} scheme and the rest $\frac{k}{2}$ randomly (\emph{long links}). The former set produces well-connected \emph{pockets} of nearby nodes while the later preserves the connectivity of the graph, both yielding a proximity factor of $\alpha = 0.5$ in an attempt to preserve the benefitial properties of unstructured topologies\cite{merugu_str2unstr_2003}.

(TODO double check validity) One disadvantage of this landmark scheme is related to the additional burden imposed to the landmark sites. The authors claim though that the algorithm requires so little work by the landmarks (maybe just echo to ping messages) that could in effect, act as ``unsuspecting participants''. Even if this is the case, the fact that it is not fully distributed, renders the protocol's scalability directly volnerable to any system size increase as well as usuitable for highly dynamic networks such as ad-hoc networks. Moreover, fixed points in a network are inherently more exposed to malicius attacks. The most significant downside of the algorithm though is that it can lead to an extremely uneven overlay ID distribution causing load inbalances and hot spots. Lastly, the scheme is coarse grained when it comes to distinguishing relatively close nodes\footnote{In the worst case, all nodes could ve clustered into a single bin.}.

\subsection{Global Soft-State}
The approach argued by Xu et.al in \cite{xu_globstate_2003} is to build a global map to help choose shorter routing paths, combining the landmark binning method and small scale distance probes to reveal the proximity properties of the underlying network to the overlay. This global view of the state is made available to all nodes in order to help them find the best way to route their messages. The authors focus on two aspects in order to accomplish their goal.

Initially, is the \emph{generation} of proximity information and then its \emph{effective exploitation}. For the first, a hybrid approach is proposed, which uses landmark clustering as a preprossesing step in order to select a number of potential nearest neighbour candidates and then refine the selection by incorporating an RTT scheme to ultimately choose the closest node. For the second, the algorithm chooses a different path from the classic gossiping approaches for constructing and maintaining the overlay. It is based on landmark-clustering-based strategic placement of proximity information on the overlay enabling any node to access such information using a landmark number that reflects its physical position in the network. For various logical regions\footnote{This might be a high-order zone in the eCAN\cite{xu_ecan_2002} context or a set of nodes sharing a particular prefix in overlays such as Pastry.} maps of physical information are built and published where each node may appear in a maximum of $log\left( N \right)$ such maps.. To dynamicaly adapt to changing network conditions, a node subscribes to relevant \emph{soft states} that utilize a notification system in order to initialize any necessary neighbour re-selection.

Maintaining several host states at different layers, makes any content migration costly. Additionaly, the method does not make any continuing effort to remap the overlay structure after a node successfuly joins, in order to adapt its state to any occurance of condition change. Although this approach greatly reduces the routin latency to far nodes, it is unable to dynamicaly identify nodes that are close to routers and gatways in order to construct the secondary overlay. Nevertheless, static recognition of such nodes is currently done based on BGP reports and prechosen landmarks, sucrificing the self-organising attribute of traditional DHTs.

\subsection{Self-Adaptive Topology Matching}
Ren et.all's goal while designing the \emph{SAT-Match} algorithm \cite{ren_satmatch_2004} was to create a protocol that would be decentralised and scalable, fine-grained in detecting changes and adaptively reform in response to the dynamic nature of a classic p2p environment and last, but certainly not least, of low incuring cost. The method is in a nutshell, a two-phased iterative process that focuses on local optimizations in order to collectively achieve a global full overlay optimization.  The iteration is finised when the node detects that it is physicaly close enough to its neigbours so that no additional optimization is needed. The paper defines \emph{stretch}, $S = \frac{\bar{L_l}}{\bar{L_p}}$, as a way quantify the \emph{topology match degree} of the constructed overlay, where $\bar{L_l}$ is the average logical link latency while the $\bar{L_p}$ is the average physical link latency.

\paragraph{Probing phase} \emph{SAT-Match} uses a small TTL value for the probing messages in order to reduce redunduncy\cite{jiang_lightflood_2008}. This process begins as soon as the node joins the network using a DHT mechanism. Each probing message contains information about the source and a small TTL value. The recipient of such a message returns information about itself to the source and forwards the probing message to its neighbours if the TTL is non-zero. The nodes been covered, are refered to as $TTL-k$ neighbourhood of the source node\footnote{Especialy, $TTL-1$ neighbourhood referes to the source's direct neighbours}. The list of responses is used to measure RTT to these nodes and sort the list ascending RTT order.

\paragraph{Jump phase} Blindly selecting the peer with the smallest RTT as neighbour is, generally, not the right desision to make in order to achieve global \emph{stretch} reduction, This is because, in a structured scheme, when a node jumps to connect to a physically close node, it may need to connect to other distant nodes to maintain the structure's integrity, thus creating an overall increase in the overlay's \emph{stretch}. The two nodes with the smallest RTT is then used in order to select one zone to jump in this phase. The algorithm is as follows: The source node $S$ calculates the stretch change of its $TTL-1$ neighbourhood and that of the $TTL-1$ neighbourhood of the first of the previously selected peers. These calculations are made as if the jump has been made. If the stretch reduction is over a predefined threshold the jump is performed, otherwise the second selected candidate is picked and the same computations are performed. If again, the threshold is not met, then no jump is ultimately done. In case of a jump, this is performed as a combination of \emph{leave} and \emph{join} operations, in the CAN context.

Moreover the algorithm takes several issues into consiteration in order to furtherly improve the resulted overlay. For example when multiple nodes try to jump simultaneously into a regeon, then the logical link brakes from one attempt may result in inacurate computation of the gain factor, for an other. This situation is identified as \emph{contention} and the nodes use an exponatial back-off algorithm to avoid it. An other problem is the uneseccary traffic incuerred by the probing phase in a region that after several jumps has settled to a stable state. In these cases it is more likely for jump attempts to be proven worthless. The algorithm doubles the probing period of such nodes, every time a jump is not taken.

The authors claim that this continuously adaptive mechanism achieves global topology matching optimization in a sufficiently large scope. This also secures the fast adaptation to frequent network changes. It also considered lightweight and can easily be embedded into current p2p systems, as well as effectively combined with other techniques, such as landmark binning.

\subsection{Delay Aware P2P System}
A new \emph{Delay Aware P2P System (DAPS)} is introduced in \cite{zhang_daps_2005}. Its main goal is to reduce the time $L$ of a look-up request by dividing the routing tables of peers into several sectors in increasing delay. The source node emitting the query message defines a delay boundary or the pruning factor $L_t$ in the paper's context. Request messages will be forwarded only to nodes whose delay less than or equal to $L_t$. With the clustered routing tables and the loose organisation the overlay network of \emph{DAPS} is between structured and unstructured.

\subsection{Mithos}
Waldvogel and Rinaldi proposed a protocol\cite{waldvogel_mythos_2003} which incorporates a directed incremental probing to find near optimal node placement.
\paragraph{Neighbour detection.} During bootstraping, the new comming node needs to know how to contact at least one of the existing members. A subset of these nodes will be used as the first set of candidate neighbours. Then, iteratively, each of the peers in the cadidate set is asked for their neighbours in order for each of the later to be probed for their network distance from the new coming one. The closest node is then used as the new candidate neighbour and the process is repeated until no further improvement is detected. Due to the fact that the algorithm may ultimately reach a local minima instead of a global one, \emph{Mithos}' approach is to first probe all neighbours within a two hop distance from the current minimum before concluding the process.
\paragraph{ID assignment.} After finding its first neighbour, the newcoming peer must be assigned its ID, a very crusial selection in order not to create many local minima that will prevent efficient future neighbourhood location. \emph{Mithos} uses information gathered during the previous step, in order to compute the ID, which inccurs no further communication costs to the algorithm. This information includes
\begin{inparaenum}[\itshape i\upshape)]
  \item the two closest nodes and their neigbours, and 
  \item their coresponding distances.
\end{inparaenum}
Using the above, it then assigns coordinates to the new-coming node, so that Euclidian distances between the node and all known hosts predict the network latency between them\cite{cox_vivaldi_2004}. These synthetic coordinates as explicitly used as the node's ID and as soon as it has been established, distances can be computed in the ID space, no longer requiring physical measurements.
\paragraph{Link establishement.} The last step of the algorithm is the interconnection among neighbours. \emph{Mithos} uses a \emph{quadrant}-based mechanism according to which each node establishes a link to the closest neighbour in each quadrant. During forwarding, the next hop is performed towards a neighbour in the same quadrant as the final destination\footnote{An implementation of this could be backed by a $d$-bit vector indexing (where $d$ is the number of dimensions) into the routing table. Then the next hop is identified by computing the difference between the current node's vector and that of the destination's.}. The problem is that after the ID assignment process, the new-coming node may not know of other neighbours in all quadrant and if even if it does, this cannot ensure that they are the nearest available. Thus, the node first identifies neighbours in all quadrants using a mechanism based on ideas similar to a perimeter walk\footnote{Used in Greedy Perimeter Stateless Routing (GPSR) protocol.} and then using parallel path processing improves the results by taking into account further geometric properties of node relationships.

(TODO review this part)In order to avoid local minima during neighbour detection, extensive probing must be undertaken. In simulation, unfortunately, only very small-sized overlay topologies (of 200 to 1000 nodes) have been used and thus no safe conclusions can be made as for the behaviour of an extensively large, real-world p2p deployment of the scheme.

\subsection{DHT-PNS}
The work in \cite{hancong_pnsbased_2006} propose a proximity neighbour selection scheme on top of the Chord DHT. Using the Vivaldi protocol\cite{cox_vivaldi_2004} each node is assigned synthetic $2$-dimension coordinates that can be used to derive network latency using Euclidian distances in the id space and without the need of explicit probing. Then the algorithm performs the two steps described in the following paragraphs.
\paragraph{Space mapping} The space is partitioned using a \emph{concentric circle clustering scheme} where succesive cycles of radiuses $\rho$, $2\rho$, $3\rho$ and so on, are constructed. Then the formed annuluses are divided into $2\chi-1$ \emph{sectors}, where $\chi$ denots the sequence number of the annulus starting from $\chi = 1$ for the centre cycle. It is proved in the paper, that this way each sector occupies the same area\footnote{The same area as does the center cycle.} and assuming uniform node distribution, this characteristic, favours a more load balanced clustering operation. Every sector in this $2-d$ coordinate space is mapped to a unique \emph{region} in the DHT space forming a multi-layer node identifier space. Thus, any nodes that belong to the same sector, are mapped to the same region as well, preserving their proximity relationship unveiled by the use of the Vivaldi protocol.
\paragraph{Routing table optimization} The system described in the previous section allows any node $\alpha$ to obtain its DHT's $key_{\alpha}$ and its region's $key_r$, in a fully distributed manner, just by applying a consistent hash function. Using a $Get\left( key_r \right)$ RPC call in Chord, the node can obtain the region's master node, called \emph{Cluster Node (CN)} which is responsible for clustering the nodes belonging in the same sector or region with that of $\alpha$. On the other hand, a $Put\left( key_r, info\right)$ RPC call, registers $\alpha$ to its corresponding region and publishes its information, through the special peer CN. $\alpha$ can, additionaly, ask CN for other nodes that have previously joined the region in order to add them into its neighbour set for future routing table optimization. Even in the case when no neighbour is detected in the current region, the search is expanded to adjasent regions and towards an upper layer identifier space until one is found or the first layer reached.

\subsection{Quasi-Chord}
The approach that is proposed by Sun and Zhang in \cite{sun_quasi_2008} confronts the topology mismatch problem, in the Chord DHT context, that is created by the fact that no consiteration for the underlying physical network topology is taken into account during the construction of the identifier cycle. To construct a \emph{Quasi-Chord} network three steps are needed. First, each host acquires its coordinates in the geometric space utilizing the \emph{global network position (GNP)} protocol \cite{ng_gnp_2001}. Second, using the \emph{Cantor space filling curve} the $2$-dimensional space is converted to a $1$-dimensional one, used in the last, third, step to build the Quasi-Chord circle according to the host's Cantor value. In the following paragraphs, these steps are furtherly discussed.
\paragraph{GNP coordinates} First of all the host must be positioned in the geometric space. The algoritm models the P2P network with a well defined distance function in such a way, it can predict, with high accuracy, the distance between any two points in the space by just evaluating the output of the distance function on the coordinates of these points. This is accomplished with the \emph{global network position (GNP)} protocol which\begin{inparaenum}[\itshape i\upshape)]
  \item creates a reference set of $N$ landmark nodes so as to minimize the error of ICMP measured distances and coordinate computed ones between them, and then
  \item each host is able to measure its round-trip times to the $N$ Landmarks in order to compute its own coordinates.
\end{inparaenum}
\paragraph{Cantor SPF} After the $2$-d coordinates are set, in the next stage of the algorithn, each participating peer is assigned a Cantor value, according to the application of a \emph{space filling curve} on the coordinate space. (TODO: add figure). This results in the conversion of the $2$-dimensional space to a $1$-dimensional Cantor space that can more easily be more mapped to the Chord hierarchy.
\paragraph{Quasi-Chord construction} As can intuitively be infered by observing the Cantor chart, close-by nodes in the physical layer, are more likely to have similar Cantor values. This attribute is exploited in order to construct a topology-aware identifier space for the Chord DHT. The cycle is constructed by sorting nodes in ascending order. Each host maintains 2 finger tables, one for clockwise and one for counter-clockwise stepping. This helps with the connectivity of the network because its not allowed to connect the first node with the last one since this will incur heavy traffic to the later\footnote{After all their Cantor values denote that they are actually the furthest of each other.}.

(TODO: doublecheck!) The disadvantage of the algorithm is that the coordinate assignment in the first stage is backed by a not fully distributed landmark-based algorithm. Moreover the Quasi-Chord model build-up is making an indirect assumption of a maximum number of allowable hosts since it is constructed. Last but not least the doubling of the required routing information which needs to be created and maintained is an additional negative point to the efficiency and the scallability of the algorithm.

\subsection{LAPTOP}
Laptop \cite{wu_laptop_2007}, introduced by Wu et al., organises the overlay into a tree based hierarchy with main focus on child-to-father relationships in order to reduce hops during message routing as well as minimize maintenance overhead. Additionally, a caching scheme is also incorporated so as to furtherly reduce routing table update cost. The authors argue that Laptop overlay network is bounded by $O\left( log_d N \right)$ hops in a balanced overlay tree, where $N$ is the number of nodes, and $d$ is the maximum degree of each node. It utilizes a geographical layout approach  and constructs a geographical layout in a self-organizing and efficient fashion, by estimating the round trip time (RTT) to a small number of nodes in the overlay network in order to make them roughly aware of their physical distances among them.

The protocol is based on four definitions.
\begin{enumerate}
 \item The amplitude of all possible measured RTTs is devided into intervals. Each node measures its distance to its parent and is assigned a label $L_i$ where $i$ denotes the configurable RTT interval in which the measured distance falls into. A special kind of node, the root, is initially assigned the $L_1$ label and maintains (as all $L_1$ nodes do) a list of other $L_1$ nodes in the overlay.
 \item Any node can have children with level lower than theirs, except an $L_{max}$ node which can only have $L_{max}$ level children and only in the case when its parent has reached its maximum degree.
 \item Each node is assigned an address in a dotted format (e.g 1.3.4). Each octet ranges from $1$ to $d$, where $d$ is the maximum degree of the nodes. The assignment process is done by appending a unique octet to the address of its parent. Root node is assigned address 1.
 \item For any descendant node $Y$ of a node $X$, the measured distance among each other, must always be less than the lower bound of the RTT interval denoted by $X$'s label.
\end{enumerate}

The routing scheme is similar to the IP's longest-prefix matching scheme. At each forwarding hop, any message travels up the tree until the first common uncestor of source and destination node is reached and then starts descending to arrive to its target. During tree traversals, special entries in the routing tables, called \emph{routing cache}, are maintained in order to increase routing efficiency and achieve finer load balance. Caching enables a node to forward a message to a better longest-prefix match than that of its direct ancestor making a large, quicker and more cost effective step through the overlay and toward the destination. To improve scalability, the number of children nodes and the size of the routing cache are limited.

In terms of overlay maintenance, Laptop incorporates a simple \emph{heartbeat}-based technique where only the parent node is responsible for monitoring its children.

At join process the newcomming node is assigned its level label as well as its address by its parent node. Additionaly it initializes its routing table (with normal and caching entries) as it traverses the overlay in search for its parents node.
The newcommer $N$ locates the root node and the later responds with a list of $L_1$ nodes. $N$ then probes each of the $L_1$ nodes in search for the closest one. If the measured RTT to the closest $L_1$ falls into the first interval then the newcommer becomes a $L_1$ node as well. Otherwise node $N$ sets the closest $L_1$ node as its potential parent node. This potential parent becomes the actual parent if it does not have any other children. If it has, $N$ gets a list of these $L_i$ nodes and by measuring the RTT to each of them tries to spot a new potential parent in order to repeat the above process.

During a gracefull departure, the refered node checkes for children in the overlay. If it does not have any, it simply notifies its parent and leaves. If it has, it selects the child node with the lowest RTT to it in order to take its place so that the locality property is preserved.

In case of an arbitrary failure, the children of the failed node are detecting their parent's absent when the later stops acknowledging their heartbeat messages. They start emmiting special messages to their grand parent node\footnote{The address of the grandparent node is stored during join process}. In case of no response from the grandparent node, children invoke the joining procedure to detect their new parent node.The parent of the failed node, aggregates the notifications from the above children nodes for a period of time, and then chooses the one with the lowest level label as the takeover node. Potential ties are broken by favouring the lower RTT. The parent of the failed node, finally informs accordingly all the children about the change in the hierarchy.

\subsection{IP-based Clustering (IPBC)}
\emph{IP-based clustering} \cite{karwaczynski_ipbc_2007} is a proximity neighbour selection technique that is based on a longest IP prefix matching scheme in order to quantify the proximity among peer nodes. The relation of the IP address and the physical location of a node in the topology of the Internet, is intuitevely realized by operation the network layer of the Internet stack. For example, nodes in the same local subnetwork share the same 3 bytes of their address. Reports in \cite{freedman_iploc_2005} state that almost $97\%$ of prefixes longer than 3 bytes correspond to address assignments at a single geographic location. Moreover, observing the \emph{IANA IPv4 Address Space Registry} we can infer that, since blocks of consecutive octet prefixes are assigned to the same Regional Internet Registries (RIRs), nodes inside will be physicaly close to each other.

The implementation of this paper is based on the above sceptic in order to create proximity information. This information is stored in a decentralised manner, within the overlay itself, just like any other resource would have been. In order to be published in the overlay, each node is first assigned a unique identifier and subsequently generates a key by hashing a fixed-length prefix of its IP. Authors argue on the prefix length to be used, as there is a thin balance between reducing the propability of finding closer nodes by adopting a long prefix and overload nodes that are responsible for information published by many peers when chosing to hash a short one. Their verdict is for the use of a 16-bit wide prefix for real world systems deployed in an Internet scale. In either case, ID and IP is then stored in the DHT using this generated key. This way, any node, at any time (either at join time or during peer lifetime for topology adaptation) can acquire information about close-by nodes just by quering the DHT for a specific key. Moreover, the algorithm takes care of the freshness of the proximity information in two ways. First, the advertising nodes themselves periodically update their advertisements or when they voluntarily leave the overlay, they explicitly remove their data. On the other hand, in case of a failure each publication is assigned an expiration time, and thus ultimately removed by the DHT maintenance mechanisms.

\subsection{CHOord considering Proximity on IPv6 (CHOP6)}
\cite{morimoto_chop6_2007} roughly estimates the proximity among nodes by exploiting the IPv6 address format as well as RTT information if this is available.The first is achieved by introducing a 64-bit ID scheme in which the least significant bit part\footnote{The exact range is a predefined system variable} is the IPv6 global routing prefix and thus enabling a longest prefix match scheme. The observation in which the protocol is based roots from the IPv6 address block assignment. Specicaly, block of /16 or /23 in size are assigned to Regional Internet Registries (RIRs) by Internet Assigned Numbers Authority (IANA). An RIR divides the signed address blockes into smaller address blocks which are furtherly assigned to Natinal Internet Registries (NIRs). That way, it is possible to estimate a node's geographical location by simply examining the upper 32-bits of its IPv6 address. Moreover, similar to Chord, CHOP6 uses a finger table, whose entries hold more than one candidate nodes. There are three cases in which a node chooses the next hop according to information it posseses.
\begin{itemize}
 \item When no information about RTT is available candidate next hops, the sender just selects a node in the finger table entry which shares the longest prefix with the destination.
 \item There is another possibility that after some communication with other nodes, the source node should know of the RTTs to some of the nodes in finger table entry. In this case, the source node chooses the one with the smallest RTT with propability $p$. If there is a node with no measured RTT then the sender can select such a node with probability $1 - p$
 \item In this last case the source node has already communicated with all candidate node in the finger table entry. Thus, the node selects the node whose RTT is the smallest with probability $p$, the one with the second smallest with probability $q$ and so on, where $0 \leq \ldots < r < q < p < 1$ and $p+q+r+\ldots \leq 1$
\end{itemize}

\subsection{Proximity in Kademlia}
In \cite{kaune_pkad_2008}, Kaune et al., studied the Kademlia DHT, in order to build a proximity aware scheme that would work in the context of iterative lookup algorithms. The protocols discussed, focus on both the overlay and underlay tiers in the sense that in the first improves the routing performance according to a cost metric that is provided by the later. This metric is tailored to the needs of a specific aplication. For example, routing could be optimised to maximise the within-ISP traffic or reduce the lookup latencies or even avoid contacting untrustworthy subnetworks.

Two algorithms are presented for the overlay optimization. One is \emph{Proximity Neighbour Selection (PNS)} and one for \emph{Proximity Route Selection (PRS)}. The first aims at keeping peers with the least contact cost in the routing table. As Kademlia constantly learns new peers (incoming requests, through iterative lookup) no special algorithm for searching more cost effective peers is necessary and simply choose the best peers seen so far. The later (PRS) aims at choosing the best next hop during routing a message. Due to the fact that the routing in Kademlia is iterative, it is the initiator of the lookup that chooses each next hop from a set of candidate nodes. The ``vanilla'' Kedemlia always chooses the closest node with respect to the XOR metric but PRS Kademlia chooses the one with the smallest underlay metric cost. Thus, as in all such approaches, there is a trade-off between the overlay and underlay distances.

An underlay metric provides information about the underlay network. The paper distingushes between three kinds of possibilities to quantify it:
\begin{itemize}
 \item Using measurements gained by previous lookups
 \item Using measurements acquired by other peers or jointly calculated with others
 \item Using a local database to look up information
\end{itemize}
In the paper, for locality of traffic and subsequently for reducing the lookup latency, a clustering scheme has been implemented exploiting information stored in a geographic information database (GeoIP,MaxMind) about peers. The goal of such a metric is to constrain the largest portion of communication within the limits of a peer's ISP.
Another incorporated approach is Vivaldi\cite{cox_vivaldi_2004}. At first all coordinates are random but as peers start to communicate they calculate RTTs as well, gradually updating their own coordinates. Vivaldi creates a system where a peer knowing the coordinates of a another, can approximate communication latency without the need of additional prompting.

\subsection{Cone}
Cone\cite{wang_cone_2007} uses a two-layered identifier space. The first, named Chord-layer identifier, denoted as $Id_{Chord}$, is the same as in ``vanilla'' Chord. The second is the Cone-layer identifier, $Id_{Cone}$ which is constructed by two component identifiers. The first, known as \emph{group identifier (gid)} denotes a relevant group the node belongs to while the second, namely \emph{local identifier (lid)} indicates the local identifier within the group. The group concept, which is introduced here, is a way of dividing nodes according to a common $Id_{Chord}$ prefix.
The structure of a Cone overlay, retains the Chord's circular topology. The differce lies on the fact that, now, two rings are created. A big ring, where nodes with the same \emph{gid} are arranged at each position. Each of these positions are a smaller ring for the particular group's \emph{lid}s. The routing is achieved in both clockwise and anti-clockwise direction in the big-ring. For this reason two routing tables are maintained, namely \emph{front} and \emph{back} finger tables, respectively. Entries in these tables, display physical network proximity with the current node. Moreover, a third table called \emph{group} table maintains information about other online peers within the current node's group in a way that entries are now close in the ID space.

Routing in Cone, comprises of the inter-group algoritm and the intra-group algorithm. First the group of nodes to which the desired key lies is detected, exploiting physical proximity information (front and back finger tables). Second during the next and last hop, the message is forwarded to the exact node that contains the desired key.

Cone uses Landmark+RTTs to generate proximity information and exploits this information using proximity neighbour selection.

\subsection{DynaMO}
In their work in \cite{winter_dynamo_2004}, Winter et al, try to build an overlay network, called \emph{DynaMO}, that tries to consider not only the physical proximity of peers but their mobility attributes as well. DynaMO is based on a Pastry overlay network. This was an intentional choice from the authors because Pastry's built-in locality heuristics are thoroughly analised in the literature \cite{castro_proximityp2p_2002} providing good backround against which to test and compare their results.  In order to adapt to mobile, ad-hoc networks, though, special care has been put to the maintenance of an even overlay ID distribution so that hot-spots are avoided. Pastry assignes them in a randomised fashion and then tries to consider proximity through the joining process and routing table maintenance. On the other hand, DynaMO dictates a newcoming node to gather information concerning its physical neighbourhood and uses it to assign itself an appropriate overlay ID.

DynaMO tries to capitalize an observation of Pastry's routing mechanism. Each node's table consists of rows equal to the number of digits of the overlay IDs and columns equal to the ID's base. As we go down the table rows, the matching prefix between the current node's ID and the row's entries increases by one. Thus, the leaf set contains the closest nodes in the ID space. Additionally as the prefix match increases by one, the result is exponentially less candidates that can fill the tables entries as the row increases. This leads to the observation \cite{antony_pastry_2001, castro_proximityp2p_2002} that from overlay routing hop to overlay routing hop the physical distance between nodes is likely to increase leading to a dominating last routing step in terms of the overall plysical routing path distance travelled during a key lookup. Since the last routing step is usually taken from the leaf set, DynaMO focuses on making this last step as physicaly close as possible. In this context, two approaches are considered namely \emph{Random Landmarking (RLM)} and \emph{Closest Neighbour Prefix Assignment (CNPA)}.

Random Landmarking uses the overlay lookup mechanism to locate nodes responsible for a fixed set of carefully chosen\footnote{In a way that the ID space is divided into equal portions} landmark keys. This means that when a node is assigned a landmark key then it plays the role of a temporary landmark for the network. Any joing node will measure its distance to that and all other landmarks and assign to itself an ID consisting of a prefix taken from the closest landmark and a remainder that could be assigned randomnly or generated by mechanism that takes into account physical neighbourhood. The legth of the prefix can be determined as $prefix_length=|log_b k|$, where $b$ is the ID base and $k$ the number of landmarks. The above scheme results in physically close nodes, forming regions of common ID prefix that are likely to be close to each other in the ID space and thus bringing a node's leaf set closer to itself. Additionally, the advantage of dynamic landmark nodes is that the failed ones can instantly be replaced by new responsibles, that share similar physical attributes to the failed and thus preserving the balance of the network.

RLM may incur more traffic especially on landmark nodes, load that in some cases is not acceptable. CPNA on the other hand takes advantage of Pastry's specification that a new coming node is bootstraped by a physicaly close node. The new comer then assumes the ID prefix of that neighbour while the rest is generated similarly to RLM. Unfortunately, less overhead comes at the expense of being more coarse-grained.

Both algorithms, are protected against the formation of physical landmark clusters or imbalanced ID distribution\footnote{More common during the initial formation of an overlay network} , by introducing the \emph{landmark gravitation range} as a theshold over which landmark keys are reassinged (for the RLM approach) or unutilized ID prefix ranges are detected and used (for the CPNA scheme) in order to balance the distribution of regions in the overlay.

\subsection{PChord}
\emph{PChord}\cite{hong_pchord_2005} is a based on the Chord DHT which adds proximity routing into its routing mechanism. The main modification over the ``vanilla'' Chord is the inclusion of a \emph{proximity list} into its routing table so that the next hop is decided based not only by considering best progress towards the key, but on the physical proximity of the candidate nodes as well. The list is, at join time, empty but as the PChord node starts to interact with other nodes it applies a heuristic mechanism to fill up the list. Entries are dynamicaly added or removed as the network state is constantly changing. Because, it is guaranteed that at each routing step there is progress towards the target node in the ID space, PChord will result in lower hop number than Chord, for each hop is larger or at least equal in key space in PChord than in Chord. Additionaly, passing through proximity links in the underlay means reduced routing cost. Moreover, if the number of entries in the proximity list is the same as the number of network partitions, PChord prevents hops from jumping back to the same network the current node belongs to.

\subsection{AChord}
\emph{AChord}\cite{dao_achord_2006} uses IPv6's anycast functionality in order to
\begin{itemize}
 \item releave the protocol from complex joining procedures, and
 \item achieve high accuracy of network proximity giving high routing efficiency,
\end{itemize}
all with a simple and lightweight mechanism that requires few changes to Chord without affecting its own advantageous characteristics. Anycast delivers a message comming from the outside of an anycast group to the physicaly closest node in that group. AChord organizes all nodes participating in the overlay network into an anycast group. Any joining node comes outside of that group and, thus, is automaticaly forwarded to the physicaly nearest node in order to bootstrap, avoiding
\begin{inparaenum}[\itshape i\upshape)]
  \item the need to, explicitly, maintain such nodes, and
  \item the effort of finding a way of locating the physically nearest among them.
\end{inparaenum}
The ID of the new comming node is computed based on the ID of the bootstrap node and bootstrap's predecessor in a way that its ID will position itself between the formentioned.

After joining, the finger table is build the same way as in the Chord protocol. Moreover, additional nodes are maintained in a structure called \emph{neighbourship table} which stores information about the closest known nodes\footnote{As a first entry, at bootstrap, the node stores information about the bootstrap node.}. The routing decision is made by using both the neighbourship and the finger table.

\subsection{Chord6}
\emph{Chord6} \cite{xiong_chord6_2005} is a Chord variant that tries to exploit the hierarchical features of IPv6 in order to create a substrate that reduces interdomain trafic and does that in a cost efficient way. Chord6 bareley modifies the original Chord protocol, only in the part of identifier definition. This, renders the aproach easily portable to other DHTs such as CAN, Pastry and Tapestry. In Chord6 the identifier contains two pares: the higher bits are obtained by hashing the  node's IPv6 address prefix of specific length, while the remaining lower bitss are the hash value of the rest of that IPv6 address. This way nodes in a domain will be mapped onto a contihuous key space on the overlay network. This guarantees that messages are routed in a way that they do not hop in and out of domains many time, thus minimizing overall routing cost. (TODO: Recheck validity) However, even though nodes in the same domain would have close identifiers, nodes in two close domains may have very different ones.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
Peer-to-peer architecture has been in the center of research attention in the last decade. Especially the decentralized unstructured genre exploits the advantages of loose coupling and self organization of computing nodes to form application-layer networks on top of the physical, best-effort infrastructure of the Internet that exhibit intresting properties. Scalability problems arouse quickly, though, because of the inefficient construction of this overlay that was built with no concern for the underlying physical network that causes a great deal of redundant traffic. The problem was identified by the research community as the topology mismatch problem between the overlay and the corresponding underlying physical network and a great deal of effort has been set towards alleviating it. Some fruits of this effort have been gathered and presented in this survey. Measurement of link cost through latency or RTT and deletion of inefficient established connections are the key concepts of almost all approaches. Others, manage to address the problem through hierarchical peer clustering (e.g. best IP matching). Additionally, some protocols focus on specific problems that furtherly arise, such as overlay partitioning, search scope reduction or convergence speed, to name just a few. Another desire is to form a protocol that could be applied to both decentralized unstructured and decentralized structured networks. Unfortunately no approach has equally addressed these problems in order to form a robust solution. So the field seems to be, still, fertile for any, new, clever idea.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{plain}
\bibliography{bib/p2p-overlays}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2013-11-10 distrproj.tex
16.03.2010 (Vassilis) Restructuring initial approach.

@@
 %opening
-\title{Alleviating the Topology Mismatch Problem with Location-Aware Overlays: A Survey}
+% TODO: Contribution of this paper
+% Q&A: 'Overlay' or 'Peer-to-Peer'/'P2P'
+\title{Alleviating the Topology Mismatch Problem in Unstructured and Structured Overlay Networks: A Survey}
 \author{
@@
 \begin{document}
-
 \maketitle
@@
 \begin{abstract}
-Peer-to-peer systems is a rapidly growing application architecture.
-The overlay network abstraction on top of a best-effort infrastructure provides a powerfull tool to realize ever-wanted exotic features like anonymity, robustness, load balancing, QoS at the application layer and many more.
+% TODO: Contribution of this paper
+Peer-to-peer (p2p) systems is a rapidly growing application architecture.
+The overlay network abstraction on top of a best-effort infrastructure provides a powerfull tool to realize ever-wanted exotic features like anonymity, high availability and robustness, load balancing, QoS at the application layer and many more.
 Unfortunately weaknesses of early effectuations constrained applications from unleashing the full potential of the paradigm.
-One of them is the topology mismatch problem between the overlay network and the physical underlying one.
-This survey is on the recent research developments in alleviating the problem, in unstructured schemes, using location-aware overlay construction algorithms.
+One of them is the topology mismatch problem between the overlay and the physical underlying networks, a problem that imposes unecessary stress to the network resources.
+This survey investigates on the recent research developments towards the alleviation of the problem, in both unstructured and structures p2p schemes.
 \end{abstract}
@@
 \section{Introduction}
-\emph{Peer-to-peer} systems gained significant attention due to their plethora of unique features in supporting file sharing among huge numbers of networked users, where each such \emph{peer} acts both as a resource provider and a resource consumer.
-Such serverless systems, proved to be able to achieve outstanding aggregate resource capacities as participants join the system\footnote{In peer-to-peer file-sharing systems there has been observed mitigation of this self-scaling property by the \emph{free-rider} problem\cite{adar_gnutellafreeriders_2000}}, without requiring additional expenditure for infrastructure.
-
-As these systems evolved, three main architectures have been emerged:
+
+% TODO: Contribution of this survey
+% UNDER CONSTRUCTION
+
+% TODO: Structure of this paper
+% UNDER CONSTRUCTION!
+This survey begins with a quick review of the conception of the overlay network and the available architectures of p2p systems.
+It continues with the identification of the phenomenon of the topology mismatch between overlay and underlying networks and its impact on the utilization of network resources.
+Ultimately, it reviews the most important resourses found in the literature that attempt to tackle the problem as it proposes a taxonomic scheme, using a set of supertype-subtype relationships based on unique characteristics they have and/or specific goals they target.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% OK
+\section{Overlay Networks and P2P Systems}
+An \emph{overlay network} is an abstract, logical interconnection of entities formed on top of another network.
+It consists of a set of nodes, the computing elements, that are being connected by virtual (logical) point-to-point links.
+For example, the \emph{Asymmetric Digital Subscriber Line (ADSL)} is, actually, an overlay network on top of the \emph{Public Switched Telephone Network (PSTN)}.
+
+The abstraction of overlay networks has been proposed as a way to implement efficient, fully distributed, application layer services on top a best-effort IP layer forming a widely ranged family of protocols collectively known as \emph{peer-to-peer} or \emph{p2p} for short.
+P2p network systems gained significant attention, in recent years, due to their plethora of unique features in supporting file sharing among huge numbers of (inter-) networked users, where each, such, \emph{peer} could act both as a resource provider and a resource consumer.
+This lead to the introduction of the \emph{servent}-concept \cite{gnutella}, a portmanteau that blends the notions of \emph{server} and \emph{client} to denote the twofold role of the participants in a p2p network.
+Such serverless systems, proved to be able to achieve outstanding aggregate resource capacities as participants join the system\footnote{Unfortunatelly, there has been observed mitigation of this self-scaling property by the undesirable behaviour of participating users, usually answered in popular file-sharing systems; from \emph{free-riding} (i.e.
+in Napster and Gnutella networks \cite{saroiu_measurefileshare_2002, adar_gnutellafreeriders_2000, hughes_gnutellafreeride_2005}), to the distribution of illegal content and/or other \emph{socio-technical} issues \cite{hughes_socp2p_2008}.}, without requiring additional expenditure for infrastructure.
+
+As p2p systems evolved, three main architectures have been emerged, namely:
 \begin{itemize}
@@
 The central index server was queried by the users and it returned pointers to the actual content.
-Thus, by centralizing search while distributing downloads, Napster achieved a high functional design that was widely acknowledged, at the time, as ``the fastest growing Internet application ever''.
-Of course, the architecture is not without flaws.
-The centralized search facility is its ``Achilles' heel'' as it can be prooven a bottleneck for the system's scalability as well as a central point of failure and vulnerable to \emph{Denial-of-Service (DoS)} attacks.
-
-This centralized approach, has been eventually, replaced by \emph{decentralized}, that distributed both the search and the download capabilities.
-In the \emph{decentralized unstructured} architectures, file placement is random, which means there is no correlation with the network topology whatsoever \cite{yang_improvep2psearch_2002}.
-The most important properties of such systems are that they:
-\begin{itemize}
-  \item support inherent heterogeneity of peers,
-  \item are highly resilient to peer failures, and
-  \item incur low maintenance overhead at peer arrivals and departures.
-\end{itemize}
-In the literature, they can be also found under the term \emph{broadcast-based} systems, because they use \emph{message flooding} among peers (Gnutella\cite{gnutella}) or among superpeers (KaZaA\cite{kazaa}) to propagate queries but with no specific destination.
-Each (super)peer receiving a query
-\begin{inparaenum}[\itshape i\upshape)]
-  \item decreases the incoming query message's lifetime which is measured in \emph{time-to-live (TTL)} hops,
-  \item back-propagates its results, and
-  \item if the lifetime of the incoming message did not expired (TTL did not reach zero), it forwards the message to all its neighbours except from the incoming link.
-\end{inparaenum}
-Early incarnations of the Gnutella protocol, though were considered far from scalable \cite{ritter_gnucantscale_2001} because:
-\begin{itemize}
-  \item randomly choosing logical neighbours without knowledge of the underlying network causes mismatch between physical and application-level topologies, and/or
-  \item queries may be flooded to multiple paths merging in one peer while in such case one of the paths would have been enough, and/or
-  \item peers may forward the same message to each other before they receive the query messages from each other
-\end{itemize}
-More generally, studies upon peer-to-peer traffic \cite{seroiu_analysiscds_2002, sen_analyzep2ptraffic_2004} with measurements on some popular systems, such as FastTrack, Gnutella and DirectConnect have shown that it contributes the largest portion of the overall Internet traffic.
-\cite{matei_mapgnutella_2002} showed that, even given that 95 percent of any two nodes are less than 7 hops away from each other, a flooding-based algorithm, like the one used by Gnutella, can generate 330TB/month in a 50.000 node network!
-
-Recently \emph{decentralized structured} schemes have been proposed \cite{ratnasamy_can_2001, stoica_chord_2001, zhao_tapestry_2001, antony_pastry_2001}.
-These, form an application-level topology that provides a hash table look-up and routing service, called \emph{Distributed Hash Table (DHT)} that requires only $O \left ( log n \right )$ steps for a hash-table-like \texttt{look-up()} search that also guarantees perfect recall even if only a single copy of the searched data exists in the system\footnote{In comparison to Gnutella, that requires $O \left ( n \right )$ to reliably locate a specific file}.
-
-Studying the behaviour of the various peer-to-peer schemes \cite{matei_mapgnutella_2002, lv_randomwalks_2002, merugu_str2unstr_2003} showed that the ad-hoc network topology of unstructured overlay networks that preserve \emph{Power Law} and \emph{Small World} characteristics\footnote{Power Law describes the node degree while the Small World describes characteristics of path length and clustering coefficient.
+Thus, by centralizing search while distributing downloads, Napster achieved a highly functional design that was widely acknowledged, at the time, as ``the fastest growing Internet application ever''.
+This centralized search facility, though, renders this scheme not fully distributed and can be proven its ``Achilles' heel'' in terms of being a scalability bottleneck, a central point of failure and vulnerable to malicious acts (e.g.
+\emph{Denial-of-Service (DoS) attacks}).
+
+This centralized approach, has been eventually, replaced by architectures that distributed both the search and the download capabilities.
+In these \emph{decentralized unstructured} architectures, file placement is random, which means there is no correlation with the network topology whatsoever \cite{yang_improvep2psearch_2002}.
+The most important properties of such systems are that they support inherent heterogeneity of peers, are highly resilient to peer failures, and incur low maintenance overhead at handling the dynamics of peer participation \cite{stutzbach_churn_2006}.
+In the literature, they are also known as \emph{broadcast-based} systems, because they use \emph{message flooding} among peers (i.e.
+Gnutella\cite{gnutella}) or among super-peers (i.e KaZaA\cite{kazaa}) to propagate queries.
+Each (super-) peer receiving a query decreases the incoming query message's lifetime which is measured in \emph{time-to-live (TTL)} hops, back-propagates its results, and  if the lifetime of the incoming message did not expired (TTL did not reach zero), it forwards the message to all its neighbours except from the one it receives it from.
+% TODO: maybe enumerate applications of decentralized unstructured p2ps.
+
+Recently, \emph{decentralized structured} schemes have been proposed in order to provide a self-organising infrastructure for large-scale p2p applications \cite{ratnasamy_can_2001, stoica_chord_2001, antony_pastry_2001, zhao_tapestry_2001, maymounkov_kademlia_2002}.
+They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes through a deterministic mechanism.
+The main advantage of the DHT approach of decentralized structured p2p networks over decentralized unstructured ones, is that they provide a guaranteed bound on the number of overlay routing hops that have to be taken in order for an object to be located even in the case where only a single copy exists in the system.
+This is $O \left ( log n \right )$ compared to Gnutella, that requires $O \left ( n \right )$ to reliably locate a specific object.
+% TODO: maybe enumerate applications of decentralized structured p2ps. (OceanStore, Far)
+
+% TODO: enumerate applications of overlay networks and p2p systems regardless of architecture
+% UNDER CONSTRUCTION
+% Q&A: Need more??
+Initial efforts targeted on the implementation of one-to-many addressing schemes that could replace \emph{IP multicast} in providing higher quality of streaming media through \emph{quality of service (QoS)} guarantees.
+IP multicast as well as other proposals such as \emph{IntServ} and \emph{DiffServ} \cite{cisco_diffserv_2005} have not seen wide acceptance (yet?), largely because they require changes to all routers of the network.
+On the other hand, an overlay network can be incrementally deployed on end-systems running the overlay protocol software, without cooperation from the ISPs.
+Academic research includes \cite{chu_esm_2000, jannotti_overcast_2000, kwon_tag_2002} among others.
+Overlay networks were also implemented in order to back the routing of messages to destinations whose address is not known in advance.
+This resulted in the appearance of well kwown peer-to-peer protocols such as \cite{gnutella} and \cite{maymounkov_kademlia_2002} widely used for digital content sharing among their network nodes.
+Other special perpose overlay networks include \cite{anderson_ron_2001} for resilient routing; \cite{subramanian_overqos_2004} for quality of service guarantees; and \cite{clarke_freenet_2001} for anonymity; to name just a few.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% start
+%TODO: Maybe this section should be reviewed or deleted or go to another section
+%Studying the behaviour of the various peer-to-peer schemes \cite{matei_mapgnutella_2002, lv_randomwalks_2002, merugu_str2unstr_2003} showed that the ad-hoc network topology of unstructured overlay networks that preserve \emph{Power Law} and \emph{Small World} characteristics\footnote{Power Law describes the node degree while the Small World describes characteristics of path length and clustering coefficient.
 The clustering coefficient for a node $\upsilon$ in a graph $G = \left( V, E \right)$ is defined as the ratio of the existing connections between $\upsilon$'s neighbouring nodes to $\gamma \times \left( \gamma - 1 \right)$, where $\gamma$ is the number of neighbouring nodes of $\upsilon$.
@@
 Particularly:
-\begin{itemize}
-  \item Peer-to-peer clients are extremely \emph{transient}.
+%\begin{itemize}
+%  \item Peer-to-peer clients are extremely \emph{transient}.
 Unstructured systems can have high maintenance traffic in delivering messages, updating the mapping, discovering failures and replicating lost data or pointers, making them insufficient on highly volatile networks.
-  \item \emph{Keyword searches} versus \emph{exact-match queries}.
+%  \item \emph{Keyword searches} versus \emph{exact-match queries}.
 In DHTs there is a tight control between the data placement and the topology of the network.
 For this reason it is hard to efficiently support partially matched queries while Gnutella and other similar systems effortlessly support keyword searches and other complex queries since the mechanism is realized locally, on a node-by-node basis.
-  \item Popular content is located at multiple peers and thus it is more likely for a flooding-based search to return results.
+%  \item Popular content is located at multiple peers and thus it is more likely for a flooding-based search to return results.
 DHTs, on the other hand, fit better in the systems which require ability to reliably locate content, even in the extreme case that only a single-copy exists in the network.
-\end{itemize}
-For this reasons, efforts have been placed for optimizing the efficiency of decentralized unstructured peer-to-peer networks. Research mainly focuses on
-\begin{inparaenum}[\itshape i\upshape)]
-  \item reducing unnecessary, redundant communication traffic, and
-  \item exploiting physical locality to reduce communication response.
-\end{inparaenum}
-The goal can be achieved at, both, the application-level network as well as the underlying physical one.
+%\end{itemize}
+% end
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% TODO: how 
+%For this reasons, efforts have been placed for optimizing the efficiency of decentralized unstructured peer-to-peer networks. Research mainly focuses on
+%\begin{inparaenum}[\itshape i\upshape)]
+%  \item reducing unnecessary, redundant communication traffic, and
+%  \item exploiting physical locality to reduce communication response.
+%\end{inparaenum}
+%The goal can be achieved at, both, the application-level network as well as the underlying physical one.
 In the first case by refining the message relay techniques, while in the second one, by adaptively reconstructing the application network to map as well as possible to the the physical network.
-
-This survey begins with a quick review of the conception of the overlay network.
-It continues with the phenomenon of the topology mismatch between the overlay and the underlying networks and ultimately end up presenting a series of distinguished research work that has been published in the recent literature concerning the alleviation of such a problematic state, focusing on decentralized unstructured peer-to-peer architectures.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Overlay Networks}
-An \emph{overlay} network is an abstract, logical interconnection formed on top of another network.
-It consists of a set of nodes, the computing elements, that are being connected by virtual (logical) point-to-point links.
-For example, the Asymmetric Digital Subscriber Line (ADSL) is, actually, an overlay network on top of the Public Switched Telephone Network (PSTN).
-
-The abstraction of overlay networks has been proposed as a way to implement efficient, fully distributed, application layer, services on top of the best-effort IP layer forming a widely ranged family of protocols collectively known as \emph{peer-to-peer} or \emph{p2p} for short.
-
-Initial efforts targeted on the implementation of one-to-many addressing schemes that could replace \emph{IP multicast} in providing higher quality of streaming media through \emph{quality of service (QoS)} guarantees.
-IP multicast as well as other proposals such as \emph{IntServ} and \emph{DiffServ}\cite{cisco_diffserv_2005} have not seen wide acceptance (yet?), largely because they require changes to all routers of the network.
-On the other hand, an overlay network can be incrementally deployed on end-systems running the overlay protocol software, without cooperation from the ISPs.
-Academic research includes \cite{chu_esm_2000}, \cite{jannotti_overcast_2000} and \cite{kwon_tag_2002} among others.
-
-Overlay networks were also implemented in order to back the routing of messages to destinations whose address is not known in advance.
-This resulted in the appearance of well kwown peer-to-peer protocols such as \cite{gnutella} and \cite{maymounkov_kademlia_2002} widely used for digital content sharing among their network nodes.
-
-Other special perpose overlay networks include \cite{anderson_ron_2001} for resilient routing; \cite{subramanian_overqos_2004} for quality of service guarantees; and \cite{clarke_freenet_2001} for anonymity; to name just a few.
 
@@
 \section{The Topology Mismatch Problem}
+
+% TODO: Define the problem
+% UNDER CONSTRUCTION
+% Q&A is it too late??
 An overlay network is able to control the sequence of peers, a message traverses, before reaching its destination but it is completely unaware of how the actual packets are switched at the underlying infrastructure along the overlay path.
-For example, a single logical point-to-point link on the overlay, most of the time corresponds to multiple physical links, in the underlying layer.
-
+For example, a single logical point-to-point link on the overlay, most of the time, corresponds to multiple physical links, in the underlying layer.
 Additionally a link in the underlying network often serves the mapping of several overlay paths causing increase of the physical link's \emph{stress}\cite{chu_esm_2002}.
@@
 
+% TODO: Argue on the problem
+% UNDER CONSTRUCTION
+Early incarnations of the Gnutella protocol, though were considered far from scalable \cite{ritter_gnucantscale_2001} because randomly choosing logical neighbours without knowledge of the underlying network causes mismatch between physical and application-level topologies.
+Additionaly, queries may be flooded to multiple paths merging in one peer while in such case one of the paths would have been enough.
+Nevertheless, peers may forward the same message to each other before they receive the query messages from each other.
+
+In structured schemes, typically, node IDs are assigned \emph{randomly}, resulting in excellent load balancing, scalability and robustness.
+Unfortunatelly, this randomness has a negative impact on the \emph{routing locality} of the network.
+This means that even though the target node can be reached with logaritmic overlay hops, the distance traveled in the physical underlying network, during the overlay routing process, can be far from optimal.
+The ratio of the phyical distance, a query travels through the overlay to an object, and the minimal distance to that object (i.e.
+though IP) is known in the literature as \emph{stretch} or \emph{Relative Delay Penalty (RDP)}.
+
+Studies on peer-to-peer traffic \cite{seroiu_analysiscds_2002, sen_analyzep2ptraffic_2004} with measurements on some popular systems, such as FastTrack, Gnutella and DirectConnect have shown that it contributes the largest portion of the overall Internet traffic.
+\cite{matei_mapgnutella_2002} showed that, even given that 95\% of any two nodes are less than 7 hops away from each other, a flooding-based algorithm, like the one used by Gnutella, can generate 330TB/month in a 50.000 node network!
+
 The problem of constructing an optimal overlay is known to be \emph{NP-hard}\cite{chawathe_scattercast_2000}.
@@
 
+% TODO: Taxomize the approaches that have been proposed in the literature
+% for unstructured and structured overlay networks
+% UNDER CONSTRUCTION
+% VERY IMPORTANT SECTION
+\section{Taxonomy}
+
+% TODO: Unstructured
+% UNDER CONSTRUCTION
 Several protocols for dealing with the problems derived from the topology mismatch problem in decentralised unstructured peer-to-peer systems have been proposed in the literature.
@@
 
+% TODO: Structured
+% UNDER CONSTRUCTION
+Xu et.al in \cite{xu_globstate_2003} state that there are three ways generaring proximity information
+\begin{itemize}
+ \item Expanding Ring Search,
+ \item Heuristics, and
+ \item Landmark Binning
+\end{itemize}
+
+Expanding ring search can be of two forms.
+First it can utilize the multicast infrastracture in the underlying network in order to emit its messages.
+Unfortunately such infrastructures are node widely deployed thus the implementations of this way of generating proximity information is limited to blindly flood the neighbourhood to obtain the reasonable results.
+
+Heuristics are used in order to reduce the blindness of the expanding ring search and make realizations more efficient and effective.
+Unfortunatelly a common problem of all heuristic approaches is the local minimum pitfall in which the search might be caught into.
+
+Landmark clustering is based on the view that nodes with similar distances to a set of predifined, well-known landmark nodes are pretty likely of being close to each other.
+But this approach has its weeknesses as well, such as the fact that is a rather coarse grained approximation, therefore not particularly well suited for differantiating nodes within close distance to each other.
+
+Castro et.al in \cite{castro_proximitydht_2002, castro_topawareroute_2002} on the other hand, comment on the three basic approaches in exploiting proximity in DHT protocols suggested by Ratnasamy et.al.
+in \cite{ratnasamy_openq_2002}.
+These are:
+\begin{itemize}
+ \item Geographic Layout,
+ \item Proximity Routing, and
+ \item Proximity Neighbour Selection
+\end{itemize}
+
+According to the first, the node IDs are assigned in such a way, that nodes close by in the psysical network topology, be close in the node ID space as well.
+Implementations, that work relatively well with this approach, have been incorporated into CAN.
+Nodes measure the RTT between themselves and a set of landmarks in order to match the CAN space as much as possible to the psysical one.
+Unfortunately, the use of landmarks renders the protocol which uses it, not fully self-organizing which can furtherly lead to imbalanced node distribution.
+On other DHTs, such as Chord or Pastry, another problem emerges.
+To gain fault tolerant properties, these protocols, replicate key-value pairs on neighbouring (in the ID space) nodes.
+When a proximity-based node ID assignment has been used, the needed failure resilience is undermined by the fact that close by nodes are more likely to suffer collective failures.
+
+Proximity routing does not require routing tables be built using any knowledge about network proximity.
+On the other hand it exploits such knowledge in order to choose the best next hop during routing a message.
+This approach dangerously balances between choosing the node that will further progress the routing towards the destination and choosing the closest entry in the routing table, in terms of network proximity.
+Thus, it is relatively less effective than geographical layout when applied to CAN(-like) implemetations.
+Moreover, the technique has been incorporated into a version of Chord (TODO: PChord | additioanlly check wheather this is true) but seems to increase the overhead of node joins and the size as well as maintenance cost of finger tables.
+\cite{dabek_cfs_2001} proposes a server selection scheme for the Chord DHT, on the domain of proximity routing selection.
+In \emph{CFS}, each node predicts the entire lookup latency as a function of the total number of nodes and the average overlay next routing peer.
+The problem is that it is very difficult to have a clear picture on the total number of nodes and the average hop latency from the local.
+This leads to rough estimations that concequently decreases overall performance.
+
+Finally, the third approach, constructs the routing tables with proximity knowledge count in.
+Tapestry and Pastry's mechanisms of routing table maintenance try to minimize the distance to nodes appearing in a peer's routing table.
+The fact that routing is based on longest node ID prefix match there is not such thing as a message not progressing with each routing hop.
+\cite{castro_proximityp2p_2002} argues on how Pastry exploits proximity neighbour selection in order to create a scheme that is (more) location-aware compared to the other well-known DHTs (CAN, Chord).
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% UNSTRUCTURED
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Narada}
+
+\paragraph*{}
 \emph{End System Multicast} \cite{chu_esm_2000, chu_esm_2001, chu_esm_2002} as well as other  proposes an architecture, where end-systems implement all multicast related functionality (including membership management and packet replication).
@@
 
+\paragraph*{}
 What Narada really does is to construct overlay spanning trees in a two-step process:
@@
 
+\paragraph*{}
 Narada incrementally improves the quality of the mesh by adding and dropping overlay links.
@@
 
+\paragraph*{}
 Narada protocol achieves relatively good performance for small and medium sized groups involving tens to hundreds of members.
@@
 \subsection{Gia}
+
+\paragraph*{}
 \emph{Gia} \cite{chawathe_gia_2003} is a decentralized unstructured peer-to-peer file sharing system with the following characteristics:
@@
 
+\paragraph*{}
 More specifically, there are four key components in the design of \emph{Gia} which are summarized bellow:
@@
 \subsection{Adaptive Overlay Topology Optimization}
+
+\paragraph*{}
 \emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{liu_auto_2003} is an algorithm for building an overlay multicast tree among each source node and its direct logical neighbours in order to alleviate the mismatching problem while providing a larger query coverage range.
@@
 
-\subsubsection*{Selective Flooding (SF)}
+\paragraph*{Selective Flooding (SF)}
 Instead of flooding to all neighbours, SF builds an overlay multicast tree among each peer and its immediate logical neighbours.
@@
 
-\subsubsection*{Active Topology (AT)}
+\paragraph*{Active Topology (AT)}
 At this step the overlay topology is reorganized.
@@
 \subsection{Location-aware Topology Matching}
+
+\paragraph*{}
 \emph{Location-aware topology matching} or \emph{LTM} \cite{liu_ltm_2004}, is an algorithm of building an efficient overlay by disconnecting low productive connections and choosing physically closer nodes as logical neighbours while still retaining the search scope and reducing response time of queries.
 This is done by issuing detector messages by peers, to a small region so that receivers of the detector can record relative delay information and cut inefficient or redundant logical links as well as add closer nodes as direct neighbours.
-
 Three operations are defined in LTMband are furtherly discussed in the following paragraphs.
 
-\subsubsection*{TTL2-detector flooding}
+\paragraph*{TTL2-detector flooding}
 \emph{TTL2-detector} messages are designed based on the specification of the Gnutella protocol.
@@
 
-\subsubsection*{Low productive connection cutting}
+\paragraph*{Low productive connection cutting}
 There are three cases for any peer $P$ who receives $d(i, S, v)$ multiple times:
@@
 
-\subsubsection*{Source probing}
+\paragraph*{Source probing}
 For a peer $P\in(N^2(S) - N(S))$ who receives only one $d(i, S, 0)$, the cost of $PS$ is obtained (with list look-up or probing).
-Then $P$ compares it with the cost form each hop and if $PS$ has the largest cost, $P$ will not keep this connection, while otherwise the connection will be created.
+Then $P$ compares it with the cost from each hop and if $PS$ has the largest cost, $P$ will not keep this connection, while otherwise the connection will be created.
 
@@
 
+\paragraph{}
 LTM disadvantages are
@@
 \subsection{Adaptive Connection Establishment}
+
+\paragraph{}
 \emph{Adaptive Connection Establishment (ACE)} \cite{liu_ace_2004} builds an overlay multicast tree among each source node and the peers within a certain diameter from the source peer and optimizes the neighbour connections that are not in that tree.
 
+\paragraph{}
 ACE indicates three phases for its algorithm:
@@
 
+\paragraph{}
 Simulations in \cite{liu_acesims_2004} show that the average scope of each query to cover the same scope of nodes is reduced by about 65 percent without losing any autonomy feature, while the average response time can be reduced by 35 percent.
@@
 \subsection{Hops Adaptive Neighbour Discovery}
+
+\paragraph{}
 Paper \cite{chen_hand_2006} proposes a new algorithm named \emph{Hops Adaptive Neighbour Discovery (HAND)} which uses a fully distributed triple hop adjustment strategy to address the topology mismatch problem.
-
 The advantages of the algorithm compared to other approaches are that
@@
 
+\paragraph{}
 The algorithm's ultimate goal is the optimal overlay, the \emph{Logical Communication Network (LCN)} in the paper's context.
@@
 
+\paragraph{}
 The mismatching detection is done in the following way.
@@
 
+\paragraph{}
 Measurements conducted for evaluation perposes showed that in a static environment the algorithm can effectively decrease traffic cost by about 77 percent and shorten the query response time by about 49 percent in less than two minutes.
@@
 \subsection{Distributed Cycle Minimization Protocol}
+
+\paragraph{}
 \emph{Distributed Cycle Minimization Protocol (DCMP)} which is introduced in \cite{zhu_dcmp_2008} is a dynamic, fully decentralized protocol that promises significant reduction of duplicate messages.
@@
 
+\paragraph{}
 The first step after detecting a duplicate message by some peer is to gather information from all peers in the cycle using a new type of control message called \emph{Information Collecting Message} or \emph{ICM}.
@@
 
+\paragraph{}
 Suppose $A$ detected the duplicate as depicted in Figure~\ref{figure:dcmp}.
@@
 
+\paragraph{}
 Experiments in \cite{zhu_dcmp_2008} showed that DCMP incurs a lower delay, returns more results and decreases the number of duplicate messages by 22\%, compared to LTM.
@@
 \subsection{Two-Hop-Away Neighbour Comparison and Selection}
+
+\paragraph{}
 Work in \cite{liu_thancs_2005, liu_thancs_2008} proposes a distributed heuristic called \emph{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} that
@@
 
+\paragraph{}
 THANCS is considered a \emph{local search method}, in the sense that it targets in finding a locally optimum solution, by exploiting knowledge within a 2-hop radius.
@@
 
-\paragraph*{Piggybacking neighbour distance on queries.}
+\paragraph{Piggybacking neighbour distance on queries}
 Using network delay as a metric for measuring the distance, each peer probes distances with its immediate naighbours and stores information locally.
@@
 
-\paragraph*{Neighbour comparison and selection.}
+\paragraph{Neighbour comparison and selection}
 Figure~\ref{figure:thancs} illustrates this component of the THANCS algorithm.
@@
 
+\paragraph{}
 In a static environment THANCS has been proven to be effective; optimizing 45 percent out of the 60 percent of mismatched paths, constructing a nearly optimal overlay.
@@
 
+\paragraph{}
 The proposed architecture, called \emph{mOverlay}, is a \emph{clustering approach} which creates a two-level hierarchical network, where on the top level we have connections between groups while on the bottom we have connections between hosts inside groups, thus attempting to recreate \emph{Small-World}-like properties for the constructed overlay network.
@@
 
-\paragraph{Locating process.} A new coming host, $Q$, first connects to a globally known host cache called the \emph{rendezvous point (RP)} in order to retrieve the starting point in the overlay, say $A$ in group $1$.
+\paragraph{Locating process} A new coming host, $Q$, first connects to a globally known host cache called the \emph{rendezvous point (RP)} in order to retrieve the starting point in the overlay, say $A$ in group $1$.
 Host $Q$ then, measures its distance to host $A$.
@@
 
-\paragraph{General overlay operations.} A set of additional protocols, are also introduced, similar to those found in traditional unstructred networks, but modified focusing on scalability and robustness.
+\paragraph{General overlay operations} A set of additional protocols, are also introduced, similar to those found in traditional unstructred networks, but modified focusing on scalability and robustness.
 For example a protocol for \emph{group formation} is introduced that exploits the inherent characteristic of proximity, in the overlay, in order to efficiently detect the neighbouring groups of a newly formated group from the set of adjasent groups of its closest neighbour.
@@
 \subsection{Distributed Domain Name Order}
+
+\paragraph{}
 \cite{zeinalipour-yazti_ddno_2005} proposes the \emph{Distributed Domain Name Order (DDNO)} technique which makes unstructured overlay networks, topologically aware.
@@
 
-\subsubsection*{Discovering \emph{random} neighbours}
+\paragraph{Discovering \emph{random} neighbours}
 Initially, when a new coming node wants to join a network $N$, uses a \emph{hostcache} mechanism to provide it with a list of nodes with which to establish its $\frac{d}{2}$ \emph{random} connections.
 
-\subsubsection*{Discovering \emph{sibling} neighbours}
+\paragraph{Discovering \emph{sibling} neighbours}
 In order to discover the rest $\frac{d}{2}$, \emph{sibling}, nodes, the \emph{lookupDN} procedure is initiated.
@@
 
+\paragraph{}
 Figure~\ref{figure:ddno_lookupdn} shows an $l$ message emitted by node $n$ which travels along the path [$a$, $b$, $c$, $e$, $b$, $d$].
@@
 
+\paragraph{}
 Unfortunately, use of DDNO results in networks which have a uniform distribution of node degrees, losing the valuable properties of Power Law and Small World networks, as mentioned previously in this survey.
@@
 \subsection{Critical Topology-Aware Grouping}
+
+\paragraph{}
 \emph{Critical Topology-Aware Grouping (CTAG)} presented in \cite{zhao_ctag_2006}, is a grouping algorithm that tries to exploit low-cost and low-delay communication of physically close-by peers.
@@
 
-\subsubsection*{Bootstrapping grouping}
+\paragraph{Bootstrapping grouping}
 The \emph{Gnutella Web Caching} mechanism has been modified in order for a new coming node to choose the closest \emph{GWC} in order to retrieve the node list for bootstrapping.
 
-\subsubsection*{Dynamic revision}
+\paragraph{Dynamic revision}
 Similarly to the bootstrapping phase, \emph{AM} metric is used to store hosts' addresses.
@@
 \subsection{Peer-exchange Routing Optimization Protocols}
+
+\paragraph{}
 Work presented in \cite{qiu_prop_2007} introduces \emph{peer-exchange}\footnote{A series of exchanges of neighbours between two peers.
@@
 
-\subsection{T2MC (unstrucured)}
+\subsection{T2MC}
+
+\paragraph{}
 \emph{T2MC}\cite{shi_t2mc_2008} exploits some properties of the Internet paradigm and clusters nodes belonging to the same ISP without any centralised control or predefined system parameterization.
 The algorithm considers the dynamic nature of peers and exploits the stable nature of routers in order to build a topology location relationship among end peers.
+
+\paragraph{}
 Some special routers split the physical network into autonomous system domains.
@@
 This is achieved by a 2-Means Classification, defined by the following steps:
-1.the peer chooses the minimum and maximum Latency results from the Traceroute for initializing the cntroinds of two sets ``first'' and ``second''.
-2.The peer calculates, for all hops allong its tracerouted path, the absolute distance to the centroids of both sets and assigns the routers to that centroid with which it has the smaller absolute distance.
-3.The peer calculates the latncy mean and variance value of two sets
-4.If the variance is larger than a predefined threshold then the algorithm takes a loop from step 2 picking the two latency mean values as new centroids of sets ``first'' and ``second''.
+\begin{enumerate}
+ \item the peer chooses the minimum and maximum Latency results from the Traceroute for initializing the cntroinds of two sets ``first'' and ``second''.
+ \item The peer calculates, for all hops allong its tracerouted path, the absolute distance to the centroids of both sets and assigns the routers to that centroid with which it has the smaller absolute distance.
+ \item The peer calculates the latncy mean and variance value of two sets
+ \item If the variance is larger than a predefined threshold then the algorithm takes a loop from step 2 picking the two latency mean values as new centroids of sets ``first'' and ``second''.
+\end{enumerate}
 Ultimately peer will end up with two sets having the minimum intra-set variance.
-
 Finally the peer chooses the router from ``second'' set with the minimum hops attribute and sets it as a threshold.
@@
 As Edge routers can provide more valuable information than other members of the ``near'' set, T2MC was designed to prioritize interaction of peers and edge gateways.
-TODO: figure t2mc.jpeg (a) stars represent edge routers (b) hop leaps
+
+% TODO: figure t2mc.jpeg (a) stars represent edge routers (b) hop leaps
+
+\paragraph{}
 The use of Traceroute as a tool for implementing the distance measuring infrastructure raise concearns about its efficiency and scalability.
@@
 \subsection{Unnamed Unstructured!!}
-(TODO: to be reviewed)
+% TODO: to be reviewed
+
+\paragraph{}
 In \cite{hsiao_redblue_2009}, Hsiao et al, claim to construct topology-aware unstructured overlays that \emph{guarantee} performance qualities in terms of
@@
 
-\section{Structured overlays}
-\emph{Structured} overlay networks provide a self-organising infrastructure for large-scale peer-to-peer applications.
-They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes through a deterministic mechanism.
-Some well-known implementations include CAN\cite{ratnasamy_can_2001}, Chord\cite{stoica_chord_2001}, Pastry\cite{antony_pastry_2001}, Tapestry\cite{zhao_tapestry_2001} and Kademlia\cite{maymounkov_kademlia_2002} which have been used in a wide range of distributed applications (TODO: cite some of these applications i.e OceanStore Far).
-
-The main advantage of the DHT approach of structured p2p overlays over unstructured ones, is that they provide a guaranteed bound on the number of overlay routing hops that have to be taken in order for an object to be located.
-(TODO: provide the O complexity for each of the CAN($O(d(N^{\frac{1}{d}}))$), Chord($O(\log N)$), Pastry($O(\log N)$), Tapestry($O(\log N)$), Kademlia($O(\log N)$)).
-
-Typically, node IDs are assigned \emph{randomly}, resulting in excellent load balancing, scalability and robustness.
-Unfortunatelly, this randomness has a negative impact on the \emph{routing locality} of the network.
-This means that even though the target node can be reached with logaritmic overlay hops, the distance traveled in the physical underlying network, during the overlay routing process, can be far from optimal.
-The ratio of the phyical distance, a query travels through the overlay to an object, and the minimal distance to that object (i.e.
-though IP) is known in the literature as \emph{stretch} or \emph{Relative Delay Penalty (RDP)}.
-
-Xu et.al in \cite{xu_globstate_2003} state that there are three ways generaring proximity information
-\begin{itemize}
- \item Expanding Ring Search,
- \item Heuristics, and
- \item Landmark Binning
-\end{itemize}
-
-Expanding ring search can be of two forms.
-First it can utilize the multicast infrastracture in the underlying network in order to emit its messages.
-Unfortunately such infrastructures are node widely deployed thus the implementations of this way of generating proximity information is limited to blindly flood the neighbourhood to obtain the reasonable results.
-
-Heuristics are used in order to reduce the blindness of the expanding ring search and make realizations more efficient and effective.
-Unfortunatelly a common problem of all heuristic approaches is the local minimum pitfall in which the search might be caught into.
-
-Landmark clustering is based on the view that nodes with similar distances to a set of predifined, well-known landmark nodes are pretty likely of being close to each other.
-But this approach has its weeknesses as well, such as the fact that is a rather coarse grained approximation, therefore not particularly well suited for differantiating nodes within close distance to each other.
-
-Castro et.al in \cite{castro_proximitydht_2002, castro_topawareroute_2002} on the other hand, comment on the three basic approaches in exploiting proximity in DHT protocols suggested by Ratnasamy et.al.
-in \cite{ratnasamy_openq_2002}.
-These are:
-\begin{itemize}
- \item Geographic Layout,
- \item Proximity Routing, and
- \item Proximity Neighbour Selection
-\end{itemize}
-
-According to the first, the node IDs are assigned in such a way, that nodes close by in the psysical network topology, be close in the node ID space as well.
-Implementations, that work relatively well with this approach, have been incorporated into CAN.
-Nodes measure the RTT between themselves and a set of landmarks in order to match the CAN space as much as possible to the psysical one.
-Unfortunately, the use of landmarks renders the protocol which uses it, not fully self-organizing which can furtherly lead to imbalanced node distribution.
-On other DHTs, such as Chord or Pastry, another problem emerges.
-To gain fault tolerant properties, these protocols, replicate key-value pairs on neighbouring (in the ID space) nodes.
-When a proximity-based node ID assignment has been used, the needed failure resilience is undermined by the fact that close by nodes are more likely to suffer collective failures.
-
-Proximity routing does not require routing tables be built using any knowledge about network proximity.
-On the other hand it exploits such knowledge in order to choose the best next hop during routing a message.
-This approach dangerously balances between choosing the node that will further progress the routing towards the destination and choosing the closest entry in the routing table, in terms of network proximity.
-Thus, it is relatively less effective than geographical layout when applied to CAN(-like) implemetations.
-Moreover, the technique has been incorporated into a version of Chord (TODO: PChord | additioanlly check wheather this is true) but seems to increase the overhead of node joins and the size as well as maintenance cost of finger tables.
-\cite{dabek_cfs_2001} proposes a server selection scheme for the Chord DHT, on the domain of proximity routing selection.
-In \emph{CFS}, each node predicts the entire lookup latency as a function of the total number of nodes and the average overlay next routing peer.
-The problem is that it is very difficult to have a clear picture on the total number of nodes and the average hop latency from the local.
-This leads to rough estimations that concequently decreases overall performance.
-
-Finally, the third approach, constructs the routing tables with proximity knowledge count in.
-Tapestry and Pastry's mechanisms of routing table maintenance try to minimize the distance to nodes appearing in a peer's routing table.
-The fact that routing is based on longest node ID prefix match there is not such thing as a message not progressing with each routing hop.
-\cite{castro_proximityp2p_2002} argues on how Pastry exploits proximity neighbour selection in order to create a scheme that is (more) location-aware compared to the other well-known DHTs (CAN, Chord).
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% STRUCTURED
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\paragraph*{Binning scheme} The implementation requires a set of well-known \emph{landmark} machines spread accross the Internet.
+\paragraph{Binning scheme}
+The implementation requires a set of well-known \emph{landmark} machines spread accross the Internet.
 Every newly arriving node measures its distance from these landmarks and unilaterally decides to join a specific bin based on these results.
@@
 
-\paragraph*{Structured and Unstructured Binning} The \emph{binning} scheme can be incorporated within either a structured or an unstructured overlay construction algorithm.
-The authors provide an example for both these cases.
-
+\paragraph{Structured and Unstructured Binning}
+The \emph{binning} scheme can be incorporated within either a structured or an unstructured overlay construction algorithm. The authors provide an example for both these cases.
+
+\paragraph{}
 Assuming a structured approach based on CAN\cite{ratnasamy_can_2001} and $m$ landmark nodes.
@@
 
-(TODO double check validity) One disadvantage of this landmark scheme is related to the additional burden imposed to the landmark sites.
+\paragraph{}
+% TODO double check validity
+One disadvantage of this landmark scheme is related to the additional burden imposed to the landmark sites.
 The authors claim though that the algorithm requires so little work by the landmarks (maybe just echo to ping messages) that could in effect, act as ``unsuspecting participants''.
@@
 \subsection{Global Soft-State}
+
+\paragraph{}
 The approach argued by Xu et.al in \cite{xu_globstate_2003} is to build a global map to help choose shorter routing paths, combining the landmark binning method and small scale distance probes to reveal the proximity properties of the underlying network to the overlay.
@@
 
+\paragraph{}
 Initially, is the \emph{generation} of proximity information and then its \emph{effective exploitation}.
@@
 
+\paragraph{}
 Maintaining several host states at different layers, makes any content migration costly.
@@
 \subsection{Self-Adaptive Topology Matching}
+
+\paragraph{}
 Ren et.all's goal while designing the \emph{SAT-Match} algorithm \cite{ren_satmatch_2004} was to create a protocol that would be decentralised and scalable, fine-grained in detecting changes and adaptively reform in response to the dynamic nature of a classic p2p environment and last, but certainly not least, of low incuring cost.
@@
 
-\paragraph{Probing phase} \emph{SAT-Match} uses a small TTL value for the probing messages in order to reduce redunduncy\cite{jiang_lightflood_2008}.
+\paragraph{Probing phase}
+\emph{SAT-Match} uses a small TTL value for the probing messages in order to reduce redunduncy\cite{jiang_lightflood_2008}.
 This process begins as soon as the node joins the network using a DHT mechanism.
@@
 
-\paragraph{Jump phase} Blindly selecting the peer with the smallest RTT as neighbour is, generally, not the right desision to make in order to achieve global \emph{stretch} reduction, This is because, in a structured scheme, when a node jumps to connect to a physically close node, it may need to connect to other distant nodes to maintain the structure's integrity, thus creating an overall increase in the overlay's \emph{stretch}.
+\paragraph{Jump phase}
+Blindly selecting the peer with the smallest RTT as neighbour is, generally, not the right desision to make in order to achieve global \emph{stretch} reduction, This is because, in a structured scheme, when a node jumps to connect to a physically close node, it may need to connect to other distant nodes to maintain the structure's integrity, thus creating an overall increase in the overlay's \emph{stretch}.
 The two nodes with the smallest RTT is then used in order to select one zone to jump in this phase.
@@
 
+\paragraph{}
 Moreover the algorithm takes several issues into consiteration in order to furtherly improve the resulted overlay.
@@
 
+\paragraph{}
 The authors claim that this continuously adaptive mechanism achieves global topology matching optimization in a sufficiently large scope.
@@
 \subsection{Mithos}
+
+\paragraph{}
 Waldvogel and Rinaldi proposed a protocol\cite{waldvogel_mythos_2003} which incorporates a directed incremental probing to find near optimal node placement.
-\paragraph{Neighbour detection.} During bootstraping, the new comming node needs to know how to contact at least one of the existing members.
+
+\paragraph{Neighbour detection}
+During bootstraping, the new comming node needs to know how to contact at least one of the existing members.
 A subset of these nodes will be used as the first set of candidate neighbours.
@@
 Due to the fact that the algorithm may ultimately reach a local minima instead of a global one, \emph{Mithos}' approach is to first probe all neighbours within a two hop distance from the current minimum before concluding the process.
-\paragraph{ID assignment.} After finding its first neighbour, the newcoming peer must be assigned its ID, a very crusial selection in order not to create many local minima that will prevent efficient future neighbourhood location.
+
+\paragraph{ID assignment}
+After finding its first neighbour, the newcoming peer must be assigned its ID, a very crusial selection in order not to create many local minima that will prevent efficient future neighbourhood location.
 \emph{Mithos} uses information gathered during the previous step, in order to compute the ID, which inccurs no further communication costs to the algorithm.
@@
 These synthetic coordinates as explicitly used as the node's ID and as soon as it has been established, distances can be computed in the ID space, no longer requiring physical measurements.
-\paragraph{Link establishement.} The last step of the algorithm is the interconnection among neighbours.
+
+\paragraph{Link establishement}
+The last step of the algorithm is the interconnection among neighbours.
 \emph{Mithos} uses a \emph{quadrant}-based mechanism according to which each node establishes a link to the closest neighbour in each quadrant.
@@
 
-(TODO review this part)In order to avoid local minima during neighbour detection, extensive probing must be undertaken.
+\paragraph{}
+% TODO review this part
+In order to avoid local minima during neighbour detection, extensive probing must be undertaken.
 In simulation, unfortunately, only very small-sized overlay topologies (of 200 to 1000 nodes) have been used and thus no safe conclusions can be made as for the behaviour of an extensively large, real-world p2p deployment of the scheme.
@@
 \subsection{DHT-PNS}
+
+\paragraph{}
 The work in \cite{hancong_pnsbased_2006} propose a proximity neighbour selection scheme on top of the Chord DHT.
@@
 Then the algorithm performs the two steps described in the following paragraphs.
-\paragraph{Space mapping} The space is partitioned using a \emph{concentric circle clustering scheme} where succesive cycles of radiuses $\rho$, $2\rho$, $3\rho$ and so on, are constructed.
+
+\paragraph{Space mapping}
+The space is partitioned using a \emph{concentric circle clustering scheme} where succesive cycles of radiuses $\rho$, $2\rho$, $3\rho$ and so on, are constructed.
 Then the formed annuluses are divided into $2\chi-1$ \emph{sectors}, where $\chi$ denots the sequence number of the annulus starting from $\chi = 1$ for the centre cycle.
@@
 Thus, any nodes that belong to the same sector, are mapped to the same region as well, preserving their proximity relationship unveiled by the use of the Vivaldi protocol.
-\paragraph{Routing table optimization} The system described in the previous section allows any node $\alpha$ to obtain its DHT's $key_{\alpha}$ and its region's $key_r$, in a fully distributed manner, just by applying a consistent hash function.
+
+\paragraph{Routing table optimization}
+The system described in the previous section allows any node $\alpha$ to obtain its DHT's $key_{\alpha}$ and its region's $key_r$, in a fully distributed manner, just by applying a consistent hash function.
 Using a $Get\left( key_r \right)$ RPC call in Chord, the node can obtain the region's master node, called \emph{Cluster Node (CN)} which is responsible for clustering the nodes belonging in the same sector or region with that of $\alpha$.
@@
 \subsection{Quasi-Chord}
+
+\paragraph{}
 The approach that is proposed by Sun and Zhang in \cite{sun_quasi_2008} confronts the topology mismatch problem, in the Chord DHT context, that is created by the fact that no consiteration for the underlying physical network topology is taken into account during the construction of the identifier cycle.
@@
 In the following paragraphs, these steps are furtherly discussed.
-\paragraph{GNP coordinates} First of all the host must be positioned in the geometric space.
+
+\paragraph{GNP coordinates}
+First of all the host must be positioned in the geometric space.
 The algoritm models the P2P network with a well defined distance function in such a way, it can predict, with high accuracy, the distance between any two points in the space by just evaluating the output of the distance function on the coordinates of these points.
@@
 \end{inparaenum}
-\paragraph{Cantor SPF} After the $2$-d coordinates are set, in the next stage of the algorithn, each participating peer is assigned a Cantor value, according to the application of a \emph{space filling curve} on the coordinate space.
+
+\paragraph{Cantor SPF}
+After the $2$-d coordinates are set, in the next stage of the algorithn, each participating peer is assigned a Cantor value, according to the application of a \emph{space filling curve} on the coordinate space.
 (TODO: add figure).
 This results in the conversion of the $2$-dimensional space to a $1$-dimensional Cantor space that can more easily be more mapped to the Chord hierarchy.
-\paragraph{Quasi-Chord construction} As can intuitively be infered by observing the Cantor chart, close-by nodes in the physical layer, are more likely to have similar Cantor values.
+
+\paragraph{Quasi-Chord construction}
+As can intuitively be infered by observing the Cantor chart, close-by nodes in the physical layer, are more likely to have similar Cantor values.
 This attribute is exploited in order to construct a topology-aware identifier space for the Chord DHT.
@@
 
-(TODO: doublecheck!) The disadvantage of the algorithm is that the coordinate assignment in the first stage is backed by a not fully distributed landmark-based algorithm.
+\paragraph{}
+% TODO: doublecheck!
+The disadvantage of the algorithm is that the coordinate assignment in the first stage is backed by a not fully distributed landmark-based algorithm.
 Moreover the Quasi-Chord model build-up is making an indirect assumption of a maximum number of allowable hosts since it is constructed.
@@
 \subsection{LAPTOP}
+
+\paragraph{}
 Laptop \cite{wu_laptop_2007}, introduced by Wu et al., organises the overlay into a tree based hierarchy with main focus on child-to-father relationships in order to reduce hops during message routing as well as minimize maintenance overhead.
@@
 
+\paragraph{}
 The protocol is based on four definitions.
@@
 
+\paragraph{}
 The routing scheme is similar to the IP's longest-prefix matching scheme.
@@
 
+\paragraph{}
 During a gracefull departure, the refered node checkes for children in the overlay.
@@
 \subsection{IP-based Clustering (IPBC)}
+
+\paragraph{}
 \emph{IP-based clustering} \cite{karwaczynski_ipbc_2007} is a proximity neighbour selection technique that is based on a longest IP prefix matching scheme in order to quantify the proximity among peer nodes.
@@
 
-The implementation of this paper is based on the above sceptic in order to create proximity information.
+\paragraph{}
+The implementation in this paper is based on the above observation in order to create proximity information.
 This information is stored in a decentralised manner, within the overlay itself, just like any other resource would have been.
@@
 \subsection{CHOord considering Proximity on IPv6 (CHOP6)}
+
+\paragraph{}
 \cite{morimoto_chop6_2007} roughly estimates the proximity among nodes by exploiting the IPv6 address format as well as RTT information if this is available.The first is achieved by introducing a 64-bit ID scheme in which the least significant bit part\footnote{The exact range is a predefined system variable} is the IPv6 global routing prefix and thus enabling a longest prefix match scheme.
@@
 \subsection{Proximity in Kademlia}
+
+\paragraph{}
 In \cite{kaune_pkad_2008}, Kaune et al., studied the Kademlia DHT, in order to build a proximity aware scheme that would work in the context of iterative lookup algorithms.
@@
 
+\paragraph{}
 Two algorithms are presented for the overlay optimization.
@@
 \subsection{Cone}
+
+\paragraph{}
 Cone\cite{wang_cone_2007} uses a two-layered identifier space.
@@
 
+\paragraph{}
 Routing in Cone, comprises of the inter-group algoritm and the intra-group algorithm.
@@
 \subsection{DynaMO}
+
+\paragraph{}
 In their work in \cite{winter_dynamo_2004}, Winter et al, try to build an overlay network, called \emph{DynaMO}, that tries to consider not only the physical proximity of peers but their mobility attributes as well.
@@
 
+\paragraph{}
 Random Landmarking uses the overlay lookup mechanism to locate nodes responsible for a fixed set of carefully chosen\footnote{In a way that the ID space is divided into equal portions} landmark keys.
@@
 
+\paragraph{}
 RLM may incur more traffic especially on landmark nodes, load that in some cases is not acceptable.
@@
 
+\paragraph{}
 Both algorithms, are protected against the formation of physical landmark clusters or imbalanced ID distribution\footnote{More common during the initial formation of an overlay network} , by introducing the \emph{landmark gravitation range} as a theshold over which landmark keys are reassinged (for the RLM approach) or unutilized ID prefix ranges are detected and used (for the CPNA scheme) in order to balance the distribution of regions in the overlay.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% TODO: Should be rewritten
 \section{Conclusions}

### 2013-11-10 mard-survey.tex
22.02.2012 (Hüseyin) TeX version of the PDF-only versions of the previous commits. Figures removed. Adding tabular data (work in progress) and first attempt for algorithm comparison.

%\documentclass[a4paper,10pt]{article}
\documentclass[acmcsur]{acmtrans2m}
\usepackage[utf-8]{inputenc}

\usepackage{paralist}

\usepackage{graphicx}

\usepackage{colortbl}
\usepackage{color,soul}
\usepackage{longtable}
\usepackage{pdflscape}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{definition}[theorem]{Definition}

%\textwidth      6.5in
%\textheight     9.2in
%\oddsidemargin  -0.5cm
%\evensidemargin -0.5cm
%\parindent      0pt              
%\topmargin      -0.3in
%\headsep        20pt
%
%\parskip        5pt
%\parindent      10pt

%opening
% TODO: Contribution of this paper
% Q&A: 'Overlay' or 'Peer-to-Peer'/'P2P'
\title{Alleviating the Topology Mismatch Problem in Unstructured and Structured Overlay Networks: A Survey}
\author{
Vassilis Moustakas$^1$, H\"useyin Akcan$^2$, Mema Roussopoulou$^1$ and Alex Delis$^1$\\
$^1$Department of Informatics and Telecommunications,\\
National and Kapodistrian University of Athens\\
\mbox{\texttt{\{b.moystakas, mema, ad\}@di.uoa.gr}}\\
$^2$Department of Software Engineering,\\
Izmir University of Economics, Izmir, Turkey \\
\mbox{\texttt{huseyin.akcan@ieu.edu.tr}}
}

\begin{document}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% TODO: Contribution of this paper
Peer-to-peer (p2p) systems is a rapidly growing application architecture. The overlay network abstraction on top of a best-effort infrastructure provides a powerful tool to realize ever-wanted exotic features like anonymity, high availability and robustness, load balancing, QoS at the application layer and many more. Unfortunately weaknesses of early effectuations constrained applications from unleashing the full potential of the paradigm. One of them is the topology mismatch problem between the overlay and the physical underlying networks, a problem that imposes unnecessary stress to the network resources. This survey investigates on the recent research developments towards the alleviation of the problem, in both unstructured and structures p2p schemes.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

% TODO: Contribution of this survey
% UNDER CONSTRUCTION

% TODO: Structure of this paper
% UNDER CONSTRUCTION!
This survey begins with a quick review of the conception of the overlay network and the available architectures of p2p systems. It continues with the identification of the phenomenon of the topology mismatch between overlay and underlying networks and its impact on the utilization of network resources. Ultimately, it reviews the most important resources found in the literature that attempt to tackle the problem as it proposes a taxonomic scheme, using a set of supertype-subtype relationships based on unique characteristics they have and/or specific goals they target.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% OK
\section{Overlay Networks and P2P Systems}

\begin{figure}[h!]
\centering
  \includegraphics[scale=0.5]{img/p2p.pdf}
\caption{An example overlay network.}
\label{figure:overlay}
\end{figure}

An \emph{overlay network} is an abstract, logical interconnection of entities
formed on top of another network. It consists of a set of nodes, the computing
elements, that are being connected by virtual (logical) point-to-point links.
For example, the \emph{Asymmetric Digital Subscriber Line (ADSL)} is, actually,
an overlay network on top of the \emph{Public Switched Telephone Network
(PSTN)}. Figure \ref{figure:overlay} illustrates an example physical ethernet network
and the corresponding overlay network formed by four nodes. 
Nodes $A$ and $B$ are on the same local network, while $C$ and $D$ are in
different networks. The top layer represents the overlay network formed by these
nodes, which is a reflection of the application layer connections among these
nodes. As the topology of the network can change based on the application, in
this example, it is easy to observe that even though nodes $A$ and $B$ are in the same
network, their communication goes through node $C$ on another network,
demonstrating a non-optimal mapping between the physical and the overlay
networks.
An overlay network is generally useful when
abstraction of the network layer from the application layer is important for
applications, particularly when the underlying network structure is subject to
frequent changes.%, which is a common case for ad-hoc networks.

The abstraction of overlay networks has been proposed as a way to implement
efficient, fully distributed, application layer services on top of a best-effort
IP layer forming a widely ranged family of protocols collectively known as
\emph{peer-to-peer} or \emph{p2p} for short. P2p network systems gained
significant attention, in recent years, due to their plethora of unique features
in supporting file sharing among huge numbers of (inter-) networked computers
called \emph{peers}, where each, such, \emph{peer} could act both as a resource
provider and a resource consumer. This changed the traditional
\emph{client-server} model dominating the internet and lead to the introduction
of the \emph{servent}-concept \cite{gnutella}, a portmanteau that blends the
notions of \emph{server} and \emph{client} to denote the twofold role of the
participants in a p2p network. Such serverless systems, proved to be able to
achieve outstanding aggregate resource capacities as participants join the
system
%\footnote{Unfortunatelly, there has been observed mitigation of this
%self-scaling property by the undesirable behaviour of participating users,
%usually answered in popular file-sharing systems; from \emph{free-riding} (i.e.
%in Napster and Gnutella networks \cite{saroiu_measurefileshare_2002,
%adar_gnutellafreeriders_2000, hughes_gnutellafreeride_2005}), to the
%distribution of illegal content and/or other \emph{socio-technical} issues
%\cite{hughes_socp2p_2008}.}, 
without requiring additional expenditure for infrastructure.

As p2p systems evolved, three main architectures have been emerged, namely:
\begin{itemize}
  \item \emph{centralized}
  \item \emph{decentralized unstructured}
  \item \emph{decentralized structured}
\end{itemize}

\emph{Centralized} architectures were the first to recognize that requests for popular content need not be sent to a central server but instead could be handled by the many hosts that already posses the content. \emph{Napster}, for example, maintained a \emph{central index server} based on file lists provided by participating peers. The central index server was queried by the users and it returned pointers to the actual content. Thus, by centralizing search while distributing downloads, Napster achieved a highly functional design that was widely acknowledged, at the time, as ``the fastest growing Internet application ever''. This centralized search facility, though, renders this scheme not fully distributed and can be proven its ``Achilles' heel'' in terms of being a scalability bottleneck, a central point of failure and vulnerable to malicious acts (e.g. \emph{Denial-of-Service (DoS) attacks}).

The centralized approach has been eventually replaced by architectures that distributed both the search and the download capabilities. In these \emph{decentralized unstructured} architectures, file placement is random, which means there is no correlation with the network topology whatsoever \cite{yang_improvep2psearch_2002}. The most important properties of such systems are that they support inherent heterogeneity of peers, are highly resilient to peer failures, and incur low maintenance overhead at handling the dynamics of peer participation \cite{stutzbach_churn_2006}. In the literature, they are also known as \emph{broadcast-based} systems, because they use \emph{message flooding} among peers (i.e. Gnutella\cite{gnutella}) or among super-peers (i.e KaZaA\cite{kazaa}) to propagate queries. 
Even though the unstructured approach became highly popular among file sharing applications, they do have a certain disadvantage in locating rare objects due to the unstructured nature and the limitations of the flooding approach. 
%Each (super-) peer receiving a query decreases the incoming query message's lifetime which is measured in \emph{time-to-live (TTL)} hops, back-propagates its results, and  if the lifetime of the incoming message did not expired (TTL did not reach zero), it forwards the message to all its neighbours except from the one it receives it from.
% TODO: maybe enumerate applications of decentralized unstructured p2ps.

Recently, \emph{decentralized structured} schemes have been proposed in order to provide a self-organising infrastructure for large-scale p2p applications \cite{ratnasamy_can_2001,stoica_chord_2001,antony_pastry_2001,zhao_tapestry_2001,maymounkov_kademlia_2002}. They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes through a deterministic mechanism. The main advantage of the DHT approach of decentralized structured p2p networks over decentralized unstructured ones, is that they provide a guaranteed bound on the number of overlay routing hops that have to be taken in order for an object to be located even in the case where only a single copy exists in the system. This is $O \left ( log n \right )$ compared to Gnutella, that requires $O \left ( n \right )$ to reliably locate a specific object.
% TODO: maybe enumerate applications of decentralized structured p2ps. (OceanStore, Far)

%{\sethlcolor{yellow}\hl{HA: Give the brief advantages and disadvantages of structured and unstructured
%approaches here!}}

% TODO: enumerate applications of overlay networks and p2p systems regardless of architecture
% UNDER CONSTRUCTION
% Q&A: Need more??
Initial efforts targeted on the implementation of one-to-many addressing schemes
that could replace \emph{IP multicast} in providing higher quality of streaming
media through \emph{quality of service (QoS)} guarantees. IP multicast as well
as other proposals such as \emph{IntServ} and \emph{DiffServ}
\cite{cisco_diffserv_2005} have not seen wide acceptance (yet?), largely because
they require changes to all routers of the network. On the other hand, an
overlay network can be incrementally deployed on end-systems running the overlay
protocol software, without cooperation from the ISPs. Academic research includes
\cite{chu_esm_2000,jannotti_overcast_2000,kwon_tag_2002} among others.  Overlay
networks were also implemented in order to back the routing of messages to
destinations whose address is not known in advance. This resulted in the
appearance of well known peer-to-peer protocols such as \cite{gnutella} and
\cite{maymounkov_kademlia_2002} widely used for digital content sharing among
their network nodes.  Other special purpose overlay networks include
\cite{anderson_ron_2001} for resilient routing; \cite{subramanian_overqos_2004}
for quality of service guarantees; and \cite{clarke_freenet_2001} for anonymity;
to name just a few.

The main focus of this paper is the topology mismatch problem, therefore the
available approaches in overlay networks are briefly summarized, and interested
readers are pointed to the available surveys on overlay networks
\cite{TheotokisS04,LuaCPSL05}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% start
%TODO: Maybe this section should be reviewed or deleted or go to another section
%Studying the behaviour of the various peer-to-peer schemes \cite{matei_mapgnutella_2002, lv_randomwalks_2002, merugu_str2unstr_2003} showed that the ad-hoc network topology of unstructured overlay networks that preserve \emph{Power Law} and \emph{Small World} characteristics\footnote{Power Law describes the node degree while the Small World describes characteristics of path length and clustering coefficient. The clustering coefficient for a node $\upsilon$ in a graph $G = \left( V, E \right)$ is defined as the ratio of the existing connections between $\upsilon$'s neighbouring nodes to $\gamma \times \left( \gamma - 1 \right)$, where $\gamma$ is the number of neighbouring nodes of $\upsilon$. High cluster coefficient means that neighbouring nodes of any node $\upsilon$ likely connect one another.} \cite{faloutsos_powerlaw_1999, saroiu_measurefileshare_2002} offer a more promising approach. Particularly:
%\begin{itemize}
%  \item Peer-to-peer clients are extremely \emph{transient}. Unstructured systems can have high maintenance traffic in delivering messages, updating the mapping, discovering failures and replicating lost data or pointers, making them insufficient on highly volatile networks.
%  \item \emph{Keyword searches} versus \emph{exact-match queries}. In DHTs there is a tight control between the data placement and the topology of the network. For this reason it is hard to efficiently support partially matched queries while Gnutella and other similar systems effortlessly support keyword searches and other complex queries since the mechanism is realized locally, on a node-by-node basis.
%  \item Popular content is located at multiple peers and thus it is more likely for a flooding-based search to return results. DHTs, on the other hand, fit better in the systems which require ability to reliably locate content, even in the extreme case that only a single-copy exists in the network.
%\end{itemize}
% end
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% TODO: how 
%For this reasons, efforts have been placed for optimizing the efficiency of decentralized unstructured peer-to-peer networks. Research mainly focuses on
%\begin{inparaenum}[\itshape i\upshape)]
%  \item reducing unnecessary, redundant communication traffic, and
%  \item exploiting physical locality to reduce communication response.
%\end{inparaenum}
%The goal can be achieved at, both, the application-level network as well as the underlying physical one. In the first case by refining the message relay techniques, while in the second one, by adaptively reconstructing the application network to map as well as possible to the the physical network.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Topology Mismatch Problem}

% TODO: Define the problem
% UNDER CONSTRUCTION
% Q&A is it too late??

%{\sethlcolor{yellow}\hl{ {\cite{hsiao_redblue_2009}} states that:
%Liu {\cite{liu_thancs_2008}}  concluded that more than 70 percent of communication
%paths in an unstructured overlay do not exploit the physical network topology
%(i.e., the Internet topology), resulting in a topology mismatch between the
%overlay and the underlying network.  } }

One of the major issues that defines the efficiency of the overlay network is
the mapping of the physical links and the overlay paths. Ideally, overlay
networks are expected to achieve an optimal mapping among the overlay paths and
the underlying physical links, and avoid the inefficient mapping that is shown
in Figure \ref{figure:overlay}. The problem of constructing an optimal overlay
is referred to as the \emph{topology mismatch problem}, and formally defined as
follows:

\begin{definition}
    Let $V = \{v_1, ..., v_n\}$ be a set of points denoting the network nodes,
    $\{v_i, v_j\} \in E$ be the set of unicast distances between nodes $v_i$ and
    $v_j$, $G=(V,E)$ be a complete distance graph over $V$. The topology
    mismatch problem is to construct a minimal spanning tree,  where 
    node degree is restricted to a constant ($k\geq 2$) by the bandwidth of each node $v_i$.
\end{definition}

Constructing degree constrained spanning tree is known to be NP-Hard
\cite{NPBOOK}, as well as the topology mismatch problem
\cite{chawathe_scattercast_2000}. 
 Moreover, the internet is an interesting environment
where end-to-end latencies demonstrate triangle inequality violations (TIVs),
which further complicates the problem.
These delay space TIV is a consequence of the Internet's structure and routing
policies and thus will remain a property of the Internet for the foreseeable
future \cite{zheng_irprtt_2005}. TIVs affect network coordinate
\cite{cox_vivaldi_2004,wong_meridian_2005} and positioning \cite{ng_gnp_2001}
and makes difficult the construction of delay aware overlays.  A number of
heuristic approaches have been proposed in an effort to finding an algorithm
that performs reasonably and consistently well.

A topology unaware overlay network is able to control the sequence of peers a
message traverses before reaching its destination, but it is completely unaware
of how the actual packets are switched at the underlying infrastructure along
the overlay path. For example, a single logical point-to-point link on the
overlay, most of the time, corresponds to multiple physical links in the
underlying layer. Additionally, a link in the underlying network often serves
the mapping of several overlay paths causing increase of the traffic on the
physical link, which is also called the link's
\emph{stress}\cite{chu_esm_2002}.  Furthermore, the stochastic behaviour of
peers randomly joining and leaving the peer-to-peer network all cause the
overlay to physical mapping to create such unnecessary, redundant maintenance
traffic that impacts the efficiency of the network in terms of average response
time.

%{\sethlcolor{yellow}\hl{
%HA: Motivation to solve topology mismatch problem:
%
%Most of the traffic (find refs!! see Location aware topology matching paper
%introduction for refs [31][29]) in the internet today is generated by P2P
%networks. The inefficient use of the network resources by the P2P protocols,
%especially the problem known as topology mismatch problem, causes unnecessary
%traffic load and a heavy stress
%on the internet infrastructure in general, and effects other users of the
%internet than the P2P nodes.
%}}

\begin{figure}
\centering
  \includegraphics[scale=0.8]{img/phys.pdf}
\caption{Interconnection of nodes in the physical level.}
\label{figure:phys}
\end{figure}

\begin{figure}
\centering
  \includegraphics[scale=0.8]{img/over1.pdf}
  \hspace{3ex}
  \vline
  \hspace{3ex}
  \includegraphics[scale=0.8]{img/over2.pdf}
\caption{HA: Put subfigure, and a new caption here}
\label{figure:over1}
\end{figure}

For example, assume nodes $A$, $B$, $C$ and $D$ are connected through the
physical network shown in Figure~\ref{figure:phys}, where the network costs are
given in milliseconds, and peer $A$ sends a message to peer $D$. If these peers
participate in an overlay network according to one of the setups of
Figures~\ref{figure:over1}~(left) and~\ref{figure:over1}~(right) then users will
yield different performances. In the first overlay shown in Figure
\ref{figure:over1}~(left), the message will traverse the following sequence of
links in the physical layer (marked with their costs): $3 \rightarrow 7
\rightarrow 1 \rightarrow 2 \rightarrow 2 \rightarrow 1 \rightarrow 7
\rightarrow 8 \rightarrow 5 \rightarrow 5 \rightarrow 4$. In the second overlay
shown in Figure ~\ref{figure:over1}~(right), the path will be: $3 \rightarrow 8
\rightarrow 4$. The total cost of the first path is $45 ms$, while the second
only costs a mere $15 ms$. Therefore, we can conclude that the second overlay is
more congruent with the underlying physical network than the first one and thus
more efficient.

% TODO: Argue on the problem
% UNDER CONSTRUCTION
Early incarnations of the overlay protocols, however, did not make use of the optimal
mapping with the physical network topology. Early Gnutella protocol, though were
considered far from scalable \cite{ritter_gnucantscale_2001}, randomly choose
logical neighbours without knowledge of the underlying network, effectively
causing mismatch between physical and application-level topologies.
Additionally, queries may be flooded to multiple paths merging in one peer while
in such case one of the paths would have been enough. 
%Nevertheless, peers may
%forward the same message to each other before they receive the query messages
%from each other.

Similar problems are observed in decentralized structured schemes also.
Typically, node IDs are assigned \emph{randomly}, resulting in excellent load
balancing, scalability and robustness.  Unfortunately, this randomness has a
negative impact on the \emph{routing locality} of the network. This means that
even though the target node can be reached with logarithmic overlay hops, the
distance traveled in the physical underlying network, during the overlay routing
process, can be far from optimal.  The ratio of the physical distance, a query
travels through the overlay to an object, and the minimal distance to that
object (i.e. though IP) is known in the literature as \emph{stretch} or
\emph{Relative Delay Penalty (RDP)} \cite{chu_esm_2000}.

Studies on peer-to-peer traffic reveals the effect of the peer-to-peer traffic
on the overall internet.  Measurements
\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004} on some popular
systems, such as FastTrack, Gnutella and DirectConnect have shown that the
peer-to-peer traffic contributes the largest portion of the overall Internet
traffic.  \cite{matei_mapgnutella_2002} showed that, even given that 95\% of any
two nodes are less than 7 hops away from each other, a flooding-based algorithm,
like the one used by Gnutella, can generate 330TB/month in a 50.000 node
network!  Therefore, the efficiency of the overlay networks has serious
consequences on the overall well-being of the modern internet.

% TODO: Taxomize the approaches that have been proposed in the literature
% for unstructured and structured overlay networks
% UNDER CONSTRUCTION
% VERY IMPORTANT SECTION
%\section{Taxonomy}
%
%% TODO: Unstructured
%% UNDER CONSTRUCTION
%Several overlay protocols for dealing with the problems derived from the
%topology mismatch problem have been proposed in the literature. Each
%implementation has different goals but all share some common ground on the
%aspect of the network functionality they focus on, in order to achieve them. In
%this section, the general taxonomy used throughout this paper to categorize the
%algorithms based on their functionality is presented.  The brief description of
%each category and sub-category is stated below.
%
%{\sethlcolor{yellow}\hl{CITE this paper {\cite{liu_ltm_2004}}}}
%
%\subsection{Forwarding based}
%
%Forwarding based methodology is generally adapted by distributed unstructured
%overlay systems. In this approach, a peer selects only a subset of its
%neighbours to re-broadcast query messages. The selection is made using one or a
%combination of various statistical metrics. Examples of such metrics are the
%number of responses received by a neighbour or the connection latency of the
%link between the nodes, etc. The forwarding based approach enhances search
%efficiency but has also several drawbacks. First, the search scope is reduced
%drastically.  Expanding the search scope, on the other hand, is no easy task
%because the overhead of forming multicast trees is proportional to the multicast
%group size.  Second, forwarding based schemes do not consider dynamic joining
%and leaving of peers so they do not scale well on dynamic environments.
%
%\subsection{cache based} 
%
%Caching based protocols are effectively used to reduce traffic costs and response
%times. The caching policy varies depending on the way protocol handles the index and the
%content. Centralized P2P systems
%use central index servers, while local caching systems, such as KazaA, use super peers
%to cache indices in a distributed way. Content caching is also possible in P2P
%systems, where nodes cache the forwarded content for further retrievals.
%Although caching has the above mentioned advantages,  duplication
%of messages still exist, which limits the scalability of these approaches.
%Therefore, cache based approaches are analyzed in the following categories:
%  \begin{itemize}
%    \item \emph{data index caching},
%    \item \emph{content index caching},
%    \item \emph{centralized}, and
%    \item \emph{local}.
%  \end{itemize}
%
%\subsection{overlay optimization based} 
%
%The overlay optimization based protocols modify the topology of the P2P network
%using various techniques. These approaches include creating spanning trees using 
%connection graphs, creating cluster of physically close nodes, or using latency
%information to detect proximity. The brief description of each category is presented
%below:
%
%  \begin{itemize}
%    \item \emph{Spanning tree based}. These approaches construct of a rich graph
%    based on the network connections and build minimum spanning tree on the
%    graphs, causing large traffic overhead to the system
%    \cite{chu_esm_2000,chu_esm_2002}.
%
%    \item \emph{Cluster based}. These approaches select to link physically
%    closer nodes with each other, therefore shrink the search scope
%    significantly while mapping accuracy is not always guaranteed.
%
%    \item \emph{Minimum latency first}. Use of latency as a metric to calculate
%    distance among peers. They require global latency information
%    ``landmarks''\footnote{Measuring latency between peers and stable Internet
%    servers.}.
%
%  \end{itemize}
%
%
%% TODO: Structured
%% UNDER CONSTRUCTION
%\subsection{proximity based}
%Xu \textit{et al.} \cite{xu_globstate_2003} state that there are three ways generating proximity information
%\begin{itemize}
% \item Expanding Ring Search\\
%Expanding ring search can be of two forms. First it can utilize the multicast
%infrastructure in the underlying network in order to emit its messages.
%Unfortunately such infrastructures are not widely deployed thus the
%implementations of this way of generating proximity information is limited to
%blindly flooding the neighbourhood to obtain reasonable results.
%
% \item Heuristics\\
%Heuristics are used in order to reduce the blindness of the expanding ring
%search and make realizations more efficient and effective. Unfortunately a
%common problem of all heuristic approaches is the local minimum pitfall in which
%the search might be caught into.
%
% \item Landmark Binning\\
%Landmark clustering is based on the view that nodes with similar distances to a
%set of predefined well-known landmark nodes are pretty likely of being close to
%each other. But this approach has its weaknesses as well, such as the fact that
%is a rather coarse grained approximation, therefore not particularly well suited
%for differentiating nodes within close distance to each other.
%
%\end{itemize}
%
%Castro \textit{et al.} \cite{castro_proximitydht_2002,castro_topawareroute_2002} on
%the other hand, comment on the three basic approaches in exploiting proximity
%in DHT protocols suggested by Ratnasamy \textit{et al.} \cite{ratnasamy_openq_2002}.
%These are:
%
%\begin{itemize}
% \item Geographic Layout\\
%The node IDs are assigned in such a way that nodes close by in the physical
%network topology, be close in the node ID space as well. Implementations that
%work relatively well with this approach have been incorporated into CAN. Nodes
%measure the RTT between themselves and a set of landmarks in order to match the
%CAN space as much as possible to the physical one. Unfortunately, the approach
%requires well known landmark servers and for that matter is  not fully
%self-organizing which can further lead to imbalanced node distribution. On
%other DHTs, such as Chord or Pastry, another problem emerges.  To gain fault
%tolerant properties, these protocols, replicate key-value pairs on neighbouring
%(in the ID space) nodes. When a proximity-based node ID assignment has been
%used, the needed failure resilience is undermined by the fact that close by
%nodes are more likely to suffer collective failures.
%
% \item Proximity Routing\\
%Proximity routing does not require routing tables be built using any knowledge
%about network proximity. On the other hand it exploits such knowledge in order
%to choose the best next hop during routing a message. This approach 
%balances between choosing the node that will further progress the routing
%towards the destination and choosing the closest entry in the routing table, in
%terms of network proximity. Thus, it is relatively less effective than
%geographical layout when applied to CAN(-like) implementations. Moreover, the
%technique has been incorporated into a version of Chord (TODO: PChord |
%additionally check whether this is true) causing an increase on the overhead of
%node joins and the size as well as maintenance cost of finger tables.
%\cite{dabek_cfs_2001} proposes a server selection scheme for the Chord DHT, on
%the domain of proximity routing selection. In \emph{CFS}, each node predicts
%the entire lookup latency as a function of the total number of nodes and the
%average overlay next routing peer. The problem is that it is very difficult to
%have a clear picture on the total number of nodes and the average hop latency
%from the local. This leads to rough estimations that consequently decreases
%overall performance.
%
% \item Proximity Neighbour Selection\\
%Finally, the third approach, constructs the routing tables using proximity
%knowledge. Tapestry and Pastry's mechanisms of routing table
%maintenance try to minimize the distance to nodes appearing in a peer's routing
%table. Since routing is based on longest node ID prefix match, messages 
%are gradually forwarded to nearby nodes at each routing step.
%\cite{castro_proximityp2p_2002} argues on how Pastry exploits proximity
%neighbour selection in order to create a scheme that is (more) location-aware
%compared to the other well-known DHTs (CAN, Chord).
%
%\end{itemize}
%
%{\sethlcolor{yellow}\hl{
%HA: Possible criteria: based on which protocol (eg Gnutella, CAN etc), peer
%selection (topology selection, cluster, cache etc), supports dynamic update,
%runtimes}}
%%\textbf{Algorithm} & \textbf{Overlay structure} & \textbf{Forwarding} & \textbf{Cache} & \textbf{Overlay optimization} & \textbf{Proximity information} & \textbf{Base protocol} & \textbf{Dynamic update} & \textbf{Runtime} \\
%
%
%{\sethlcolor{yellow}\hl{
%HA: Possible criteria2: Overlay optimization structure, base protocol (eg
%Gnutella, CAN etc), dynamic update, runtimes, scalability}}
%
%{\sethlcolor{yellow}\hl{
%HA: maybe add also the year of publication, and see if there is a pattern in
%terms of the method and the year??}}

Having understood the nature of the problem that peer-to-peer architectures face,
Section \ref{sec:unstructured} and \ref{sec:structured} discuss the recent academic work that has been conducted
in the field. What follows does not claim
to be a thorough citation of all known protocols that are available out there,
but instead, a careful selection of those that left a distinctive fingerprint
contribution in the efforts of the research community to alleviate the topology
mismatch problem.

%{\sethlcolor{yellow}\hl{
%Create a theoretical section. This is an Approximation alg for Top-mis problem
%
%A Near-Optimal Algorithm Attacking
%the Topology Mismatch Problem
%in Unstructured Peer-to-Peer Networks
%}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Unstructured Decentralized Algorithms}\label{sec:unstructured}

In this section, the unstructured decentralized P2P algorithms are analysed based on their architecture and
base methodology to tackle the topology mismatch problem. The
unstructured decentralized algorithms are visited and categorized based on their
use of the overlay structure, messaging between nodes, and on the techniques
proposed to detect proximity to optimize the topology. 
The base methods that unstructured P2P algorithms use to tackle the topology
mismatch problem can be categorized as Broadcast, Cache, Overlay Optimization, 
and Landmark Based. In this section, these methodologies are discussed in
detail and the algorithms in each category are highlighted.

%{\sethlcolor{yellow}\hl{
%HA: TALK ABOUT GNUTELLA HERE!!!!}}

\subsection{Broadcast Based}

In the naive Gnutella protocol, each received query request is forwarded to all the
neighbours, which clearly generates unnecessary traffic over the network. The
algorithms categorized in the \textit{broadcast based} methodology prefers
re-broadcasting the query messages to a selected subset of their neighbours,
where the selection criteria varies between algorithms.
The selection is made using one or a
combination of various statistical metrics, such as the reliability or the latency
of the link between nodes, etc.
The forwarding based approach enhances search
efficiency but has also the drawbacks of having drastically reduced search scope
due to the selective re-broadcasting policy, and scalability problems on large
P2P networks.
%Expanding the search scope, on the other hand, is no easy task
%because the overhead of forming multicast trees is proportional to the multicast group size.  
%forwarding based schemes do not consider dynamic joining
%and leaving of peers so they do not scale well on dynamic environments.

The forwarding based P2P algorithms and their details are presented below.

\paragraph*{\bf Improving search in peer-to-peer networks}

\cite{yang_improvep2psearch_2002} proposes an easy to implement and practical
solution to the inefficiency problem caused by blind flooding in Gnutella like
P2P file sharing protocols. The paper replaces blind flooding with three
approaches, namely \textit{iterative deepening}, \textit{directed BFS}, and
\textit{local indices}. In \textit{iterative deepening}, the search is performed
on a BFS tree with multiple preset depths. The depth limit is iteratively
increased by the source node for each query based on the quality of the results.
The source node, after examining the results, may issue a new request by
increasing the depth limit, which will trigger the nodes at the last depth
level to resume the search. The iterative approach avoids to start the search
from scratch at each iteration and reduces the load of the nodes on the upper
levels of the tree. The major drawback of \textit{iterative deepening} is the
delay between successive iterations, as the source node needs to examine the
results at each iteration before deciding to quit or resume the query.  The
\textit{directed BFS} tries to avoid this delay by forwarding the query messages
only to a selected set of neighbours, in which the selection criteria varies
from the number of results received previously, distance in terms of hops,
bandwidth, or the query load of the neighbour. In \textit{local indices}, each
node sustains a local data index of the nodes within a radius of $r$ hops of
itself and uses this local index to remotely query the neighbour nodes without
generating additional traffic.  \textit{Local indices} greatly reduces the
aggregate bandwidth usage of the network and improves query efficiency, however,
updating the indices in cases with frequent node joins and leaves introduces a
serious overhead to the system if the radius is kept broad.

\paragraph*{\bf Gia}

\emph{Gia} \cite{chawathe_gia_2003} is an effort to solve the scalability
problem of the unstructured P2P file sharing systems, in particular Gnutella.
The main novelty in the design of Gia is the replacement of blind flooding
approach of Gnutella with random walks \cite{lv_randomwalks_2002}. Although
random walks is a step in the right direction, issuing only a single copy of the
query within the network reduces the search scope, thus affects the success rate
of the query.  In order to overcome this limitation, Gia introduces a
token-based flow control algorithm, which is essentially an intelligent flow
control algorithm that gradually redirects the queries to nodes with more chance
of answering. Gia also acknowledges the heterogenity in peer bandwidth,
processing power, disk speed, etc, of the nodes in P2P networks and uses this
heterogenity when connecting nodes to each other, such that by using a topology
adaptation algorithm, Gia ensures that high capacity nodes have high degrees and
low capacity nodes are within short proximity of high capacity nodes. In order
to prevent overloading of nodes with query requests, Gia uses a token-based flow
control algorithm in which each node announces the number of query requests it
can handle in terms of tokens to its neighbours, so that neighbours only forward
query requests to nodes that they previously received tokens from. Although the
topology adaptation algorithm Gia uses improves the scalability of the network,
it does not help much in solving the topology mismatch problem, since it does
not consider the underlying physical topology.

\paragraph*{\bf Distributed Cycle Minimization Protocol}

The flooding based query approach of Gnutella creates many duplicate packets in
the network. Nodes receive the same query request from various neighbours, which
they generally discard when identified as duplicate. One other reason for
duplicate messages is the cycles in the forwarding paths. These duplicate
packets affect the performance of the P2P network negatively by increasing the
overall resource usage. Moreover, the more active nodes, that have high
capacity, high bandwidth, or contribute more to the network, suffer more from
the overhead caused by the duplicate packets. Therefore, \emph{Distributed Cycle
Minimization Protocol (DCMP)} \cite{zhu_dcmp_2008} is introduced as a method to
remove the cycles and eliminate the duplicate packets, without sacrificing the
connectivity. In DCMP, once a cycle is detected, the most powerful node in the
cycle is elected as the \emph{Gate Peer} and the cycle is cut from a special
link such that the distance of all the nodes within the cycle to the \emph{Gate
Peer} is minimized. The process is managed by using two special message types,
\emph{Information Collection (IC)} and \emph{Cut Message (CM)}. One disadvantage
of DCMP is that since the distance a forwarded message can travel is limited
with the TTL value, which is practically $7$ in Gnutella, the DCMP cannot detect
cycles of length larger than $7$. Even though cycle elimination improves the
network performance, it does not solve the topology mismatch problem.

%\subsection{Cache Based}
%
%Caching based protocols are effectively used to reduce traffic costs and response
%times. The caching policy varies depending on the way protocol handles the index and the
%content. Centralized P2P systems
%use central index servers, while local caching systems, such as KazaA, use super peers
%to cache indices in a distributed way. Content caching is also possible in P2P
%systems, where nodes cache the forwarded content for further retrievals.
%Although caching has the above mentioned advantages,  duplication
%of messages still exist, which limits the scalability of these approaches.
%Therefore, cache based approaches are analyzed in the following categories:
%  \begin{itemize}
%    \item \emph{data index caching},
%    \item \emph{content index caching},
%    \item \emph{centralized}, and
%    \item \emph{local}.
%  \end{itemize}
%

\subsection{Cache Based}
Caching is a widely used methodology to exploit the locality to minimize the
redundant transfer of data, which is successfully used in web servers and file servers.
As peers in a P2P system also work as servers, intuitively it is expected that
P2P file sharing systems can also benefit from caching. However, one of the most
important characteristic properties of P2P systems is that nodes
join and leave frequently, and the lifetime of a query is relatively short
compared to web servers, which makes caching in P2P networks nontrivial. There are
usually two levels of caching possible in file sharing systems, the indices or
pointers to data can be cached, or the data itself can be cached. Caching is
already implemented and successfully used in some commercial P2P systems. One
popular example is KazaA, which uses caching of indices in its super peers.
We present the other protocols that use caching and their details below.

\paragraph*{\bf Replication Strategies in Unstructured Peer-to-Peer Networks}

\cite{Cohen02} aim to improve the inefficient blind search algorithm by
replicating the data in a peer to peer network. The main intuition behind the
idea of replication, or using cached copies, is that as the number of copies for
each item increases in the network, it would be easier for a search algorithm,
even a random one, to locate these items. In order to analyse the feasibility of
such replication approach, the authors investigate three different replication
strategies, namely uniform, proportional, and square root allocation. In the
uniform model, the copies of items are uniformly replicated in the network,
while in the proportional model, the items are replicated based on their query
rate, so that frequently queried items are replicated more. The square root
allocation is a strategy proposed by the authors, which is a model between the
uniform and proportional allocation.  In order to evaluate the outcomes of the
replication models, the authors compare the overall costs of successful and
unsuccessful searches in the network. The results can be summarized as follows;
the uniform allocation model minimizes the maximum search size, therefore
reduces the time spent on an unsuccessful search. The proportional model, on
the other hand, promotes the more frequently queried items by replicating them
more, therefore decreases the search time for popular items, but suffers when
locating the rare items. The authors also claim that the expected successful
search size is the same for uniform and proportional models, and any approach
between them would behave much better. Therefore they propose the square root
allocation approach, which is a replication strategy that minimizes the expected
search size of successful queries in P2P networks.

\paragraph*{\bf Tracing a large-scale Peer to Peer System: an hour in the life of
Gnutella}

\cite{Markatos02} analysis the network traces of the Gnutella network and
detects that query requests in Gnutella has locality, and proposes a caching
algorithm that exploits this locality. The analysis of the trace data reveals
other important facts about the structure of the Gnutella network and the query
data.  One significant observation is that the geographic locations of clients
do not have a correlation with the number of query requests they receive. This
is an obvious result of the topology mismatch problem caused by the overlay
structure of the Gnutella network. Gnutella traffic is observed to be bursty
both for query requests and query responses, even in longer intervals, and nine
out of ten queries do not generate any response due to the inefficient design of
the Gnutella network. When developing a caching system to exploit locality,
applying an approach similar to web caching does not fit well with P2P systems.
The caches in P2P systems not only have to consider the query string, but also
the TTL value, the source of the query, and the time of the query as well.  In
general, even though optimum caching is hard to achieve, it is reported that
caching improves the overall performance of the Gnutella system.

\subsection{Overlay Optimization Based}

The overlay optimization based protocols modify the topology of the P2P network
using various techniques. The two commonly used approach that is investigated in
this section are creating spanning trees using 
connection graphs, and creating cluster of physically close nodes. 
%The brief description of each category is presented below:
Spanning tree based approaches construct rich graphs based on the network
connections and build minimum spanning trees on the graphs. Even though the
spanning tree provides efficient query performance, the construction
and the update costs of the spanning tree, especially in dynamic environments
with many nodes joining and leaving the network, cause large traffic
overhead to the system \cite{chu_esm_2000,chu_esm_2002}.

The cluster based approaches on the other hand select to link physically closer
nodes with each other, therefore shrink the search scope significantly. However,
commonly used methods for determining the correct physical positions of nodes
over the internet does not always return reliable results, therefore mapping
accuracy is not always guaranteed.

Below, the details of the algorithms that use the spanning tree based or cluster
based topology optimization methods are presented.

\paragraph*{\bf Narada}

Narada \cite{chu_esm_2000,chu_esm_2001,chu_esm_2002} is a generic protocol to
design self adapting overlay networks, that can achieve application layer
multicast without requiring IP multicast at the network layer. Although Narada
is not originally designed as a peer-to-peer file sharing application protocol,
it has become one of the pioneering works in the overlay networks area.  Even
though the design of the IP multicast protocol is finalized a while ago, and
some of the routers on the internet today already support this protocol, the IP
multicast idea did not take off as anticipated. One of the main reasons for this
delay is that the IP multicast violates the stateless design of the current
internet. Although the multicast is not widely used, and the future of the
protocol is not bright, there are multiple applications that need this kind of
service, such as IP TVs, P2P networks, streaming services, etc. The application
layer multicast is then a valid alternative for these applications.  The authors
claim that, although Narada is not as efficient as IP multicast, still it
provides reasonable performance, without the additional cost of implementing
multicast services on the current internet infrastructure. The main reason for
the inefficiency is due to the Topology mismatch problem.  Narada tries to solve
this problem by building a richer connected graph\footnote{ENarada
\cite{li_enarada_2008} used Gossip protocol for the construction}, called a
mesh, and building per source minimum spanning trees.  Narada also maintains the
graph and the trees dynamically updated on node joins and leaves. Moreover, Narada tries
to optimize the physical link stress, the overall resource usage and the
relative delay among end systems. The main limitation of Narada is that although
it works reasonably well for small groups, it does not scale well for larger
networks, therefore it is not suitable for P2P file sharing applications which
can easily scale to millions of nodes.

\paragraph*{\bf Adaptive Overlay Topology Optimization and Adaptive Connection
Establishment}

\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{liu_auto_2003} is one
of the first attempts, along with Narada, from the P2P research community to
address the topology mismatch problem. AOTO is a distributed algorithm
that uses \emph{Selective Flooding} and \emph{Active Topology} to optimize the
overlay network topology. The \emph{Selective Flooding} algorithm creates a minimum
spanning tree over the overlay and uses this tree instead of
flooding the whole network, without shrinking the search scope. 
In \emph{Active Topology}, the physical locations of nodes are estimated based
on the network delay among them and physically close nodes are connected as
neighbours to revise the overlay topology as close as possible to the physical
network topology. Overall, AOTO reports a $55\%$ performance improvement in
terms of the traffic load. In \emph{Adaptive Connection Establishment (ACE)}
\cite{liu_ace_2004}, the authors extend the idea by introducing optimizations
based on the depths of the minimum spanning trees. As the network delay is not always a reliable estimation
method to detect physical locations of the nodes, the algorithm still suffers
from the discrepancies caused by mislocated nodes.

\paragraph*{\bf Location-aware Topology Matching}

Peer-to-peer algorithms designed without considering the topology mismatch
problem generates unnecessary heavy traffic on the internet infrastructure, due
to the popular use of these P2P programs for file sharing. \emph{Location-aware
topology matching (LTM)} \cite{liu_ltm_2004} proposes a method to
optimize the overlay structure of the P2P network based on the physical
topology. In doing so, LTM issues a special message called
\textit{TTL2-detector} to detect the latency of $N^2$ neighbours around each
node. The latency information gathered is later used to evolve the overlay
network into a more efficient one, without reducing the search scope. Each node
examines the latency information with the neighbours and low productive
connections are dropped and replaced by more efficient ones, thus reducing the
latency on the overall network, and eliminating some potential inefficiencies,
such as unnecessary crossing of messages over autonomous system (AS) boundaries, etc. Although the
LTM approach improves the overall efficiency of the P2P network, since it uses
the latency to detect close by nodes, does not use the real physical topology
information, therefore does not offer a guaranteed bound to the
topology mismatch problem.

\paragraph*{\bf Scalable Bipartite Overlay}

\emph{Scalable Bipartite Overlay (SBO)}
\cite{liu_bipartite_IPDPS,liu_bipartite_2007} improves the overhead of creating
and maintaining a minimum spanning tree cost by randomly dividing the nodes into
two groups, red and white, and assigning different tasks to different groups.
The white peers measure distances to neighbours by using the network delay as
the metric and reports the results to red peers. The red peers, equipped with the
information of all two hop ($N^2$) neighbours, creates a minimum spanning tree
of these neighbours and assigns efficient forwarding paths. The white peers can
further optimize their positions within the overlay if need be.

%\paragraph*{\textbf{Two-Hop-Away Neighbour Comparison and Selection\\}}
%
%\paragraph{}
%Work in \cite{liu_thancs_2005,liu_thancs_2008} proposes a distributed heuristic called \emph{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} that
%\begin{inparaenum}[\itshape i\upshape)]
%  \item is completely distributed and needs no global knowledge,
%  \item presents trivial overhead compared to the query cost savings
%  \item its convergent speed of the algorithm is fast enough (faster than minimum spanning tree approaches) so that is effective to dynamic environments, and
%  \item does not shrink the search scope.
%\end{inparaenum}
%
%\paragraph{}
%THANCS is considered a \emph{local search method}, in the sense that it targets in finding a locally optimum solution, by exploiting knowledge within a 2-hop radius. The algorithm consists of two main components: \emph{piggybacking neighbour distance on queries} and \emph{neighbour comparison and selection} which are furtherly discussed bellow.
%
%\paragraph{Piggybacking neighbour distance on queries}
%Using network delay as a metric for measuring the distance, each peer probes distances with its immediate naighbours and stores information locally. For this reason a special query message type, \emph{Piggy Message (PM)}, is introduced. It is 6 bytes long and includes two fields: Neighbour IP Address and Neighbour Distance. A peer $P$ constructs a PM for its neighbour $Q$, which contains $Q$'s IP address and $Q$'s distance from $P$. When $P$ receives a query from $Q$, this PM will be piggybacked by the query that goes to all other neighbours of peer $P$. Upon receiving such a query message, each of the other neighbours will detach the PM, record the $PQ$ distance and process the query. This PM will not be further forwarded. In selecting which incoming queries  should piggyback a PM the paper proposes the \emph{pure propability-based (PPB)} and the \emph{new neighbour triggered (NNT)} policies.
%
%\paragraph{Neighbour comparison and selection}
%Figure~\ref{figure:thancs} illustrates this component of the THANCS algorithm. A peer $S$ probes the distance to all known unprobed $N^2(S)$\footnote{$N^2(S)$ denotes the set of peers being two hops away from $S$, while $N(S)$ denotes the set of direct logical neighbours of $S$.}. The distance of $SP$ is known to $S$. Upon receiving a PM from node $P$ with the distance of $PQ$, $S$ follows one of the following:
%\begin{itemize}
%  \item $Q \in \left( N(S) \cap N^2(S) \right)$, i.e. $Q$ is direct neighbour of $S$. In this case $S$ will compare cost of $SQ$, $SP$ and $PQ$. If the most costly connection is one of $SQ$ or $SP$ the corresponding link will be put into a \emph{will-cut list}\footnote{Links are not immediately disconnected when put in the will-cut list. There are useless for forwarding but are kept active, for some time, in order to serve query responses that are traveling to the source peer along the inverse search path.}. If the most costly connection is $PQ$ then $S$ will do nothing, as the fully distributed nature of the algorithm will give a chance to either $P$ or $Q$ to cut the connection.
%  \item $Q \in \left( N^2(S) - N(S) \right)$, i.e. $Q$ is a two-hop-away neighbour of $S$. If $S$ hasn't probed $Q$ before\footnote{A distance cache is maintained and looked up in such a case.}, it probes and stores the result in the \emph{distance cache}. Having the distance of $SQ$, $S$ compares costs of $SQ$, $SP$ and $PQ$. If $SQ$ is the most costly, $S$ will not establish the connection. If $SP$ is the most costly, $S$ will establish connection $SQ$ and put $SP$ in the will-cut list. If $PQ$ is the longest, $S$ will keep  the connection with both $P$ and $Q$, expecting that $P$ or $Q$ will eventually disconnect link $PQ$, later.
%\end{itemize}
%
%\begin{figure}
%\centering
%  \includegraphics[scale=0.4]{img/thancs.jpeg}
%\caption{Probing two-hop-away neighbours}
%\label{figure:thancs}
%\end{figure}
%
%\paragraph{}
%In a static environment THANCS has been proven to be effective; optimizing 45 percent out of the 60 percent of mismatched paths, constructing a nearly optimal overlay. This leads to a 60 percent reduction in traffic cost as well as a 40 percent decrease in query response time. In dynamic environments (Gnutella 0.6/Limewire super-peer-like and Ion flat-like), THANCS saves up to 70 percent of the traffic cost in the super-peer topology and 55 percent for the flat one. Average response time is also decreased by 60 and 45 percent, respectively. Generally, THANCS has similar performance to LTM, without needing synchronization. SBO, incurring half the  overhead of AOTO, reduces the traffic cost the most, while THANCS has lower response time and converges faster than SBO. THANCS is, thus, more suitable for a more dynamic environment. In addition, THANCS is easy to implement and its operation overhead is trivial, compared with the other three approaches. This design, however, has the limitation of not being easily extend to also support non-flooding-based systems.

\paragraph*{\bf Peer-exchange Routing Optimization Protocols}

\cite{qiu_prop_2007} introduces two protocols called Peer-exchange Routing
Optimization Protocols (PROP) to adjust the neighbourhood graph of the overlay
network in order to reduce the overall link latency of the network. The PROP
algorithms are based on the exchange of neighbours among peers, which is
triggered by the mutual benefit of both peers to reduce the network delay.  In
PROP-G (Generic), peers exchange all their neighbours with another peer, while
in PROP-O (Optimized) only selected number of neighbours are exchanged among
peers. The PROP-G is a generic protocol that guarantees the connectivity of the
overlay graph during exchanges, therefore can be applied to both unstructured
and structured overlay networks.

\paragraph*{\bf T2MC}

T2MC \cite{shi_t2mc_2008} uses traceroute logs to detect the AS boundaries and
cluster close by peers with each other to reduce the redundant multiple message
passes between the AS boundaries. T2MC uses a customized k-mean classification
algorithm with $k=2$ to perform the classification, and exploits the stable
structure of the internet routers to guide clustering.
Even though traceroute provides detailed information about the network
structure, use of traceroute creates overhead to the overall network structure.
For this same reason, it is not unusual for network administrators to disable
this support on their network routers, which may affect the performance of the
T2MC algorithm.

%\paragraph*{\textbf{T2MC\\}}

%\paragraph{}
%\emph{T2MC}\cite{shi_t2mc_2008} exploits some properties of the Internet paradigm and clusters nodes belonging to the same ISP without any centralised control or predefined system parameterization. The algorithm considers the dynamic nature of peers and exploits the stable nature of routers in order to build a topology location relationship among end peers.
%
%\paragraph{}
%Some special routers split the physical network into autonomous system domains. Using a Traceroute mechanism T2MC searches for latency leaps among the path to a host in order to form ``near'' and ``remote'' router clusters. This is achieved by a 2-Means Classification, defined by the following steps:
%\begin{enumerate}
% \item the peer chooses the minimum and maximum Latency results from the Traceroute for initializing the cntroinds of two sets ``first'' and ``second''.
% \item The peer calculates, for all hops allong its tracerouted path, the absolute distance to the centroids of both sets and assigns the routers to that centroid with which it has the smaller absolute distance.
% \item The peer calculates the latncy mean and variance value of two sets
% \item If the variance is larger than a predefined threshold then the algorithm takes a loop from step 2 picking the two latency mean values as new centroids of sets ``first'' and ``second''.
%\end{enumerate}
%Ultimately peer will end up with two sets having the minimum intra-set variance. Finally the peer chooses the router from ``second'' set with the minimum hops attribute and sets it as a threshold. The selected router and all others whose hops attribute is larger than the threshold are classified as ``remote'' router cluster. The remaining are classified as ``near''. From the ``near'' class, the peer chooses the one with maximum hops attribute as its edge router, and registers it along with the all the ``near'' cluster into the DHT of the p2p overlay. As new peers join the network, those that share the same edge router or any of the members of the ``near'' router clustersm ther would gather to form a ``close'' peer cluster. As Edge routers can provide more valuable information than other members of the ``near'' set, T2MC was designed to prioritize interaction of peers and edge gateways.
%
%% TODO: figure t2mc.jpeg (a) stars represent edge routers (b) hop leaps
%
%\paragraph{}
%The use of Traceroute as a tool for implementing the distance measuring infrastructure raise concearns about its efficiency and scalability. Being, primarilly, a network diagnostic utility, it is concearned too heavy weighted and intrucive for use in a larger scale\cite{ratnasamy_binning_2002}. Additionally, disabling ICMP is a common administrative policy for edge sites to enforce security, while dumping BGP routing tables\cite{krishnamurthy_bgpclust_2000} is not directly available to the application layer.

\paragraph*{\textbf{Unnamed Unstructured!!\\}}
% TODO: to be reviewed

\paragraph{}
In \cite{hsiao_redblue_2009}, Hsiao et al, claim to construct topology-aware unstructured overlays that \emph{guarantee} performance qualities in terms of
\begin{inparaenum}[\itshape i\upshape)]
  \item the expected communication latency among any two overlay peers regardless of the network size, and
  \item the broadcasting scope of each participating peer.
\end{inparaenum}

The algorithm constructs an undirected graph $G = \left( V, E \right)$ comprised by two subgraphs. The first, namely $G^{\left( red \right)} = \left( V^{\left( red \right)}, E^{\left( red \right)} \right)$ in the paper's context, includes all vertices of $G$ and ensures the connectivity of the graph by securing at least one path between any two nodes. In contrast, $G^{\left( blue \right)} = \left( V^{\left( blue \right)}, E^{\left( blue \right)} \right)$, contains those vertices of $G$ that have free edges to link to other nodes and because these are fully utilized, the following also stands $E = E^{\left( red \right)} \cup E^{\left( blue \right)}$.

A joining peer $u$, partitions its neighbours into two subsets, the $B_u^{\left( red \right)}$ and $B_u^{\left( blue \right)}$. In order to populate the $B_u^{\left( red \right)}$ subset, peer $u$ samples peers uniformly and at random. Then, each of these selected peers discovers a routing path starting from itself towards the node with the smallest (or the largest) ID in the system.

\paragraph*{\bf Distributed Domain Name Order}

%{\sethlcolor{yellow}\hl{HA:  In paper/introduction nice advantages of unstructured networks:
%
%Unstructured P2P networks o?er a number of
%important advantages: (i) An unstructured network
%imposes very small demands on individual nodes,
%and more speci?cally it allows nodes to join or leave
%the network without signi?cantly a?ecting the sys-
%tem performance. (ii) Unstructured networks are
%appropriate for content-based retrieval (e.g., key-
%word searches) as opposed to object identi?er loca-
%tion of structured overlays. (iii) Finally unstructured
%networks can easily accommodate nodes of vary-
%ing power. Consequently, they scale to very large
%sizes and they o?er more robust performance in
%the presence of node failures and connection
%unreliability.
%}}

\emph{Distributed Domain Name Order (DDNO)} \cite{zeinalipour-yazti_ddno_2005}
uses the domain names to detect topologically close nodes based on the
assumption that nodes within the same domain are topologically close to each
other. Half of the possible connections of a node is used to connect to these
local peers, and the other half is used to randomly connect to the peers
anywhere on the network.  \textit{DDNO} is a heuristic approach to solve the
topology mismatch problem, with local connections to improve the efficiency and
reduce the delay on the network.  The random connections on the other hand help
ensure the connectivity in the network and avoid partitioning. \textit{DDNO}
uses \textit{Split-Hash} and \textit{dnMatch} algorithms to detect locality by
using domain names.

\subsection{Landmark Based Proximity}\label{sec:landmark}

Detecting the proximity of a node using landmarks, also known as \textit{Landmark
Clustering}, is based on the view that nodes with similar distances to a
set of predefined well-known landmark nodes are pretty likely of being close to
each other. But this approach has its weaknesses as well, such as the fact that
is a rather coarse grained approximation, therefore not particularly well suited
for detecting the correct positions of nodes within close distance to each other.

\paragraph*{\bf Landmark Binning}\label{sec:landmark_binning}

Landmark Binning \cite{ratnasamy_binning_2002} is an approach to partition close
by nodes into bins based on their distance to well known anchor nodes within the
internet. In order to detect locality, nodes use network latency (round trip
time) as a measurement technique. The network latency, even though not always
accurate, is selected in this work because its non-intrusive, transparent and
easy to apply structure. In order for the binning to work, authors assume that
there are a few well known anchor servers with known physical locations on the
internet. The authors estimate that around 12 servers will do the job. The nodes
measure their distance to all these servers and record the latency values for
each. Later, the tuple of these latency values represents the bin of each node.
Nodes with similar latency values are assumed to be close to each other, or at
least within the same AS. The landmark binning is designed as a generic model
that can work both on structured and unstructured P2P networks, and
\cite{ratnasamy_binning_2002} has the detailed description of the algorithm for
both topologies. As the main operation of the algorithm is independent from the
underlying model, and does not change for structured and unstructured networks,
we classify and give the details of the algorithm in this decentralized
unstructured section, and omit the duplicate work in the decentralized
structured section. The landmark binning is also proposed as a good candidate to
work on content distribution networks. The major disadvantage of landmark
binning is to install and maintain landmark servers on different AS, all over
the world.  A typical P2P network usually has a couple of million nodes
connected at any time, which introduces possible scalability problems in terms
of the landmark servers.  The authors claim that latency estimation does not use
much network resources and in order to avoid the possible scalability problems
the authors propose to replace single landmark servers with clusters of servers
within the same physical area. However, this approach does not reduce the
possible high network traffic flow through these landmark servers. One other
possible problem is the incorrect binning caused by inaccuracy in delay
measurements methods, as the network latency is not a proven accurate method for
location estimation.

\paragraph*{\bf mOverlay}
mOverlay \cite{zhang_moverlay_2004} addresses the scalability problems that
affect static landmark servers by introducing the dynamic landmarks. The
nodes are divided into groups based on their distances to these groups and the
neighbour nodes in these groups behave as dynamic landmarks. The mOverlay
protocol is composed of two main parts, initially adding a new node to the
overlay by selecting the closest group, and maintaining the overlay afterwards.
Finding the correct closest group is the most important part of the overlay
construction. The authors also formally prove that any new node can reach its
group by performing at most $O(logN)$ communications within the network.

\renewcommand\arraystretch{1.4}% (MyValue=1.0 is for standard)

%\begin{figure}[h!]
\hspace{-3ex}
\begin{center}
\footnotesize
%\begin{tabular}{
\begin{landscape}
\begin{longtable}{
|>{\columncolor[gray]{.7}}m{0.1\columnwidth}
|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
|>{\columncolor[gray]{.9}}m{0.2\columnwidth}
|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
|}
\caption{Decentralized Unstructured Algorithms} \label{fig:unstruct_compare_table} \\
\hline
\rowcolor[gray]{.5}
\textbf{Algorithm} & \textbf{Arch.} & \textbf{Overlay structure} & \textbf{Base protocol} &
\textbf{Dynamic update} & \textbf{Runtime} & \textbf{Scalability} & \textbf{cites}\\
\hline
\endfirsthead
\multicolumn{4}{c}%
{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
\hline
\rowcolor[gray]{.5}
\textbf{Algorithm} & \textbf{Arch.} & \textbf{Overlay structure} & \textbf{Base protocol} &
\textbf{Dynamic update} & \textbf{Runtime} & \textbf{Scalability} & \textbf{cites}\\
\hline
\endhead
\hline \multicolumn{4}{r}{\textit{Continued on next page}} \\
\endfoot
\hline
\endlastfoot
\textbf{Narada} & Decentralized unstructured & \textbf{Overlay optimization
based}. Creates a mesh (richer connected graph) and builds minimum spanning
trees on this mesh & & Yes. Supports incremental evolving & & Small and sparse
groups & 2000  \\

\hline
\textbf{Gia} & Decentralized unstructured & \textbf{Forwarding based} Replaces
Gnutella flooding with random walk, and introduces KaZaA style supernodes. Uses
dynamic topology adaptation protocol &
 Gnutella & Yes & & Better than Gnutella & 974  \\

\hline
\textbf{Adaptive Overlay Topology Optimization} & Decentralized unstructured & \textbf{Overlay optimization
based}. Creates overlay multicast tree with Selective Flooding protocol&
Gnutella & Yes, using Active Topology protocol & & Better than Gnutella & 36 \\

\hline
\textbf{Location-aware Topology Matching} & Decentralized unstructured &
\textbf{Overlay Optimization Based}. Uses \textit{TTL2-detector flooding}, \textit{low productive
connection cutting}, and \textit{source peer probing}. & Gnutella & Yes & & Better than Gnutella & 204 \\

\hline
\textbf{Replication Strategies in Unstructured P2P Networks} & Decentralized unstructured &
\textbf{Cache Based}. Uses uniform, proportional and square root allocation
strategies to replicate data. & Gnutella & Yes & & Better than Gnutella & 580 \\

\hline
\textbf{Tracing a large-scale Peer to Peer System: an hour in the life of Gnutella.} & Decentralized unstructured &
\textbf{Cache Based}. Proposes a caching algorithm based on the traces of the Gnutella traffic & Gnutella &
Yes & & Better than Gnutella & 199 \\

\hline
\textbf{Improving search in P2P networks} & Decentralized unstructured &
\textbf{Forwarding Based}. Uses \textit{iterative deepening}, \textit{directed
BFS}, and \textit{local indices} to improve efficiency. & Gnutella & Yes & & Better than Gnutella & 380 \\

\hline
\textbf{Distributed Cycle Minimization Protocol} & Decentralized unstructured &
\textbf{Forwarding based} Uses a decentralized cycle elimination protocol  &  & Yes &  &  & 9 \\

\hline
\textbf{Scalable Bipartite Overlay} & Decentralized unstructured &
\textbf{Overlay optimization based} Uses bipartite partition graph and builds
local minimum spanning trees  & Gnutella & Yes &  & Better than Gnutella & 77 \\

\hline
\textbf{Adaptive Connection Establishment} & Decentralized unstructured &
\textbf{Overlay optimization based} Forms Neighbour Cost Tables, builds local
minimum spanning trees and perform local optimizations & Adaptive Overlay
Topology Optimization (AOTO), Gnutella & Yes &  & Better than Gnutella & 110 \\

\hline
\textbf{Hops Adaptive Neighbour Discovery} &  &
 &  &  &  &  & 1 \\

\hline
\textbf{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} & Decentralized unstructured &
\textbf{Overlay optimization based} Uses piggybacking to discover neighbour
distances and selects neighbours  & Gnutella & Yes &  &  & 36 \\

\hline
\textbf{mOverlay} & Decentralized unstructured &
\textbf{Overlay optimization based} Uses dynamic landmarks to find node locality
& & Yes &  & Due to dynamic landmarks and grouping, more scalable than tree-based or mesh-based protocols & 123 \\

\hline
\textbf{Distributed Domain Name Order (DDNO)} & Decentralized unstructured &
\textbf{Overlay optimization based} Connects half of the nodes connections to
the nodes in the same domain and the other half to random nodes, therefore
supports locality and topological connection  & & Yes &  & Yes, by using super
peers & 5 \\

\hline
\textbf{Peer-exchange Routing Optimization Protocols} & Works with both decentralized
structured and unstructured &
\textbf{Overlay optimization based} Optimizes overlay by the exchange of
neighbors among peers  & & Yes &  & Yes & 13 \\

\hline
\textbf{MAY OMIT- T2MC} & Decentralized unstructured &
\textbf{Cluster based---{\sethlcolor{yellow}\hl{ IS IT ALSO PROXIMITY BASED?}}} Uses traceroute results for clustering the
nodes  & &  &  &  & 1 \\

\hline
\textbf{Unnamed-unstructured} & Decentralized unstructured &
\textbf{Overlay optimization based} Minimizes the communication delay and
maximizes the broadcasting range & & Yes &  & Better than THANCS and mOverlay & 7\\

\hline
\textbf{Landmark Binning} & Can work with both decentralized structured and unstructured &
\textbf{Proximity based, landmark binning} Uses network latency to partition
nodes into bins & & Yes &  &  & 817 \\

\hline
%\end{tabular}
\end{longtable}
\end{landscape}
\end{center}
\vspace{-2.5ex}
\vspace{-2.5ex}
%\end{figure}

\subsection{Discussion on Unstructured Decentralized Algorithms}

Earlier in section \ref{sec:unstructured} we presented the state of the art unstructured
decentralized P2P algorithms in four different categories based on their
architecture; broadcast based, cache based, overlay optimization based, and
landmark based. In this section, we present a final discussion of all the
methods including the advantages, disadvantages,
and novelties presented in order to tackle the topology mismatch problem.
Table \ref{fig:unstruct_compare_table} presents an overview of all the
unstructured decentralized algorithms described earlier in this section.

The broadcast based approaches in general propose intelligent neighbour
selection algorithms in order to replace the inefficient blind
flooding algorithm of Gnutella.
Instead of forwarding the queries to all the neighbours, intelligently selecting
neighbours with high probability of answering the query, or neighbours that have
some specific features such as high bandwidth, high capacity, or low latency in
practice reduces the aggregate resource usage of the P2P network and improves
the overall performance of the system. However, since the broadcast based algorithms
proposed do not consider the underlying physical topology when selecting the
neighbours, or optimizing the overlay structure, the algorithms do not offer a
solution to the topology mismatch problem. Although in practice the algorithms
work, none can guarantee that a single link is not used by more than necessary,
or the created topology resembles the underlying physical topology. Broadcast
based methods can be used in conjunction with other approaches to improve the
quality of the P2P systems.

The success of caching, as a well studied method in the client server internet
applications, also made it a good candidate for the peer-to-peer networks. Even
though caching improves the performance, and reduces the overall resource usage
of P2P systems, the design of caches is non trivial compared to the web caching
systems. Due to the unique structure of the P2P overlays, each node being a
server and a client simultaneously, two important problems has to be solved when
designing caches. First, the lifetime of a query is short, as the nodes join and
leave frequently. Second, the result of a single query string is not always the
same, and changes based on the TTL value, source of the query, the lifetime of
the query, etc. So, in order to develop a successful caching system in P2P
systems, these parameters also have to be considered. Even though the state of
the art P2P algorithms using caching methods reduce the resource usage of the
network, since the physical network topology is not considered by any algorithm,
the proposed systems do not provide direct ways to overcome the topology
mismatch problem caused by the overlay network.

The overlay optimization based algorithms can be analyzed in
two main topics, the minimum spanning tree based ones, and the clustering based ones.
Even though the standardization of the IP multicast protocol is completed a
while ago, and some vendors deployed routers supporting the protocol, in
practice the standard is not widely embraced as predicted, mostly because the
protocol violates the stateless design of the internet routers. The research
community, as an alternative proposed application layer multicast frameworks,
such as Narada, to simulate the same multicast protocol without requiring IP
multicast services. The application layer multicast, however, is not as
efficient as IP multicast, and suffers from the topology mismatch problem.
Narada, and its followers (AOTO, LTM, SBO) try to solve the topology mismatch
problem by building a richer connected graph and forming minimum spanning trees
over this graph that can efficiently route messages among peers. The application
level multicast protocol, initiated by Narada, is a generic protocol that can be
applied to P2P file sharing, as well as content distribution networks. The
original Narada algorithm is designed for small groups therefore it is not
scalable. The followers try to solve the scalability problem by introducing
various methods including forming minimum spanning trees for each nodes' $N^2$
neighbours, partitioning the graph into two random groups where each group is
responsible of different tasks, or performing local optimizations dynamically on
the overlay graph. The advantage of the minimum spanning tree based approaches
is that they maintain the connectivity on the network in an efficient way, while
still not shrinking the search space. However, building and maintaining a
minimum spanning tree creates a huge overhead for the network. 
One other popular approach used in overlay optimization is the cluster based
approaches. In cluster based approach, nodes use various methods to detect
physically close nodes to form clusters. T2MC uses traceroute logs and DDNO uses
domain names to cluster close by nodes. However, the accuracy of the methods
used to form clusters directly affect the success rate of the algorithm. Traceroute for
example is a heavy weight protocol to frequently use on the network, and most of
the network vendors for this reason do not allow traceroute calls. The major problem with
cluster based approaches is that the limited connectivity within the local
domains shrink the search scope dramatically, which negatively affects the
search performance of the P2P system. DDNO addresses this limitation by allowing
half of each nodes connections to be to random nodes over the network, which
balances the efficiency of the clustering approach with improved connectivity.

In landmark based proximity method, nodes use network delay (round trip
time) as a distance measurement method to position themselves based on distance
measurements to apriori known servers on the internet. The servers behave as
landmark points, and nodes use an estimation method to discover their positions.
The landmark servers are either used by nodes to directly calculate their
positions, or indirectly used to cluster nodes into groups or bins based on
their estimated distances to well known landmark servers, due to the assumption
that nodes with similar distances to a set of landmarks are physically close to
each other over the network. The landmark based protocol has two important
drawbacks. Firstly the network delay is not a reliable distance estimation
method. For example, based on the load on the network the delay to certain
nodes or networks can change from time to time, which will eventually affect the
distance measurements and wrong measurements will lead to wrong estimated positions for the
nodes, or incorrect and non optimal clusterings of the nodes. The second drawback of
using landmark servers is the cost of installing and maintaining many landmark servers over the internet
for various AS domains. As popular P2P file sharing applications usually have
millions of peers connected at any time, it is not false to assume that the hardware and the network costs of
maintaining these landmark servers will be quite high. 
A possible solution to the scalability problem of the static landmark servers is to use ordinary nodes as dynamic
landmarks once they estimate their own positions. Even though this approach scales
much better than static landmark servers, still the measurement accuracy problem affects
the overall performance of the system.

\section{Structured Decentralized Algorithms}\label{sec:structured}

In this section, the algorithms proposed for the structured decentralized
architectures are analysed and categorized based on their use of the proximity
information to optimize the underlying structure. In structured P2P algorithms,
the construction of the routing tables among nodes determines the efficiency of
the algorithms.  Therefore, routing tables that represent the underlying
physical structure well can achieve much better performance. For this reason,
the proposed methods for topology mismatch problem generally use different
levels of proximity information to optimize the routing table close to the
physical network. We analyze the structured decentralized algorithms based on
the categories presented in
\cite{castro_proximitydht_2002,castro_topawareroute_2002,ratnasamy_openq_2002}.
In this section, the details of the methodologies used and the algorithms for
each methodology are discussed in details.

%{\sethlcolor{yellow}\hl{HA:  Talk about Tapestry, Pastry, CAN and Chord briefly here!}}

\subsection{Geographic Layout}

Geographic layout is used as a method to optimize the routing table of the
structured algorithms to represent the physical network topology as close as
possible. In this method, physically close nodes are detected, and they are
positioned closely in the overlay structure. In order for this method to work,
there should be means to detect physical proximity of internet nodes. One
popular method is to use the well known internet servers, or servers dedicated
for this sole purpose, as landmark nodes, similarly as discussed previously in
Section \ref{sec:landmark} (\textit{Landmark Based Proximity}).  The inaccuracy in
the positioning using landmarks, and the cost of deploying and maintaining these
landmarks are the major disadvantages of these systems. Even though managing the
overlay structure based on the geographic layout of the nodes improves the query
efficiency of the system, on the other hand, it tends to create hotspots, and
the needed failure resilience is undermined by the fact that close by nodes are
more likely to suffer collective failures.

\paragraph*{\bf Global Soft-State}

\textit{Global Soft-State} \cite{xu_globstate_2003} builds a global map to help
choose shorter routing paths, combining the landmark binning method and small
scale distance probes to reveal the proximity properties of the underlying
network to the overlay. This global view of the state is made available to all
nodes in order to help them find the best way to route their messages.
\textit{Global Soft-State} operates in two main stages, generation and using of
the proximity information. For generating the proximity information a hybrid
approach is proposed, which uses landmark clustering as a preprocessing step in
order to select a number of potential nearest neighbour candidates and then
refine the selection by incorporating an RTT scheme to ultimately choose the
closest node. For using the proximity information, the algorithm chooses a
different path from the classic gossiping approaches for constructing and
maintaining the overlay. It is based on landmark clustering based strategic
placement of proximity information on the overlay enabling any node to access
such information using a landmark number that reflects its physical position in
the network. For various logical regions\footnote{This might be a high-order
zone in the eCAN\cite{xu_ecan_2002} context or a set of nodes sharing a
particular prefix in overlays such as Pastry.} maps of physical information are
built and published where each node may appear in a maximum of $log\left( N
\right)$ such maps. To dynamically adapt to changing network conditions, a node
subscribes to relevant \emph{soft states} that utilize a notification system in
order to initialize any necessary neighbour re-selection.

Maintaining several host states at different layers, makes any content migration
costly. Additionally, the method does not make any continuing effort to remap the
overlay structure after a node successfully joins, in order to adapt its state to
any occurrence of condition change. Although this approach greatly reduces the
routing latency to far nodes, it is unable to dynamically identify nodes that are
close to routers and gateways in order to construct the secondary overlay.
Nevertheless, static recognition of such nodes is currently done based on BGP
reports and pre-chosen landmarks, sacrificing the self-organising attribute of
traditional DHTs.

\paragraph*{\bf Mithos}

\textit{Mithos} \cite{waldvogel_mythos_2003} is a P2P protocol which
incorporates a directed incremental probing to find near optimal node placement,
and classified by its authors as an integration of geographic layout
and proximity routing overlay optimization methods.

The bootstrap phase of \textit{Mithos} starts with a subset of existing members
as the first set of candidates, and while the iteratively closest nodes are
detected by probing the neighborhood, the candidate neighbor list is updated. In
order to avoid a local minima \textit{Mithos} probes all the neighbours within a
two hop distance from the current minimum before concluding the process.

After finding its first neighbour, the newcomming peer is assigned an ID using
information gathered during the iterative neighbour selection phase. Virtual
coordinates are assigned to the newcomming peer by using the distances of two
closest nodes and their neighbours, so that Euclidean distances between the node
and all known hosts predict the network latency between
them\cite{cox_vivaldi_2004}.  The benefit of these synthetic coordinates are
that they are explicitly used as the node's ID, and distance computations among
nodes can be done in ID space without requiring physical measurements.

The last step of the algorithm is the interconnection among neighbours.
\emph{Mithos} uses a \emph{quadrant}-based mechanism according to which each
node establishes a link to the closest neighbour in each quadrant. During
forwarding, the next hop is performed towards a neighbour in the same quadrant
as the final destination. The newcoming node may not know of other neighbours
in all quadrants, therefore, the node first identifies neighbours in all
quadrants using a mechanism based on ideas similar to a perimeter
walk\footnote{Used in Greedy Perimeter Stateless Routing (GPSR) protocol.} and
then improves the results using parallel path processing by taking into account
further geometric properties of node relationships.

One major limitation of \textit{Mithos} is that the protocol cannot handle
dynamic arrivals and removals from the network, as is happens in mobile
networks.

\paragraph*{\bf LAPTOP}

\textit{LAPTOP} \cite{wu_laptop_2007},  organises the overlay into a tree based
hierarchy with main focus on reducing hops during message routing as well as
minimizing maintenance overhead. Additionally, a caching scheme is also
incorporated so as to further reduce routing table update costs. The authors
theoretically show that in \textit{LAPTOP} routing path length is bounded by $O(log_d
N)$ and node joining and leaving in the overlay network is bounded by $O\left( d
log_d N \right)$ hops in a balanced overlay tree, where $N$ is the number of
nodes, and $d$ is the maximum degree of each node. \textit{LAPTOP} implements
the geographical layout approach  and constructs a geographical layout in a
self-organizing and efficient fashion, by estimating the round trip time (RTT)
to a small number of nodes in the overlay network in order to make them roughly
aware of their physical distances among them.

Each node is assigned an address in a dotted format (e.g 1.3.4). Each octet
ranges from $1$ to $d$, where $d$ is the maximum degree of the nodes. The
assignment process is done by appending a unique octet to the address of each
nodes
parent, while root node is assigned address 1.  The routing scheme is similar to the
longest-prefix IP matching scheme. At each forwarding hop, any message travels
up the tree until the first common ancestor of source and destination node is
reached and then starts descending to arrive to its target. During tree
traversals, special entries in the routing tables, called \emph{routing cache},
are maintained in order to increase routing efficiency and achieve finer load
balance. Caching enables a node to forward a message to a better longest-prefix
match than that of its direct ancestor making a large, quicker and more cost
effective step through the overlay and toward the destination. To improve
scalability, the number of children nodes and the size of the routing cache are
limited.  In terms of overlay maintenance, \textit{LAPTOP} incorporates a simple
\emph{heartbeat}-based technique where each parent node is responsible for
monitoring its children.  At join process the newcomming node is assigned its
level label as well as its address by its parent node. Additionally it
initializes its routing table (with normal and caching entries) as it traverses
the overlay in search for its parents node.  During a graceful departure, the
referred node checks for children in the overlay. If it does not have any, it
simply notifies its parent and leaves. If it has, it selects the child node with
the lowest RTT to it in order to take its place so that the locality property is
preserved.

\subsection{Proximity Routing}
Proximity routing does not require routing tables to be built using any knowledge
about network proximity. On the other hand it exploits such knowledge in order
to choose the best next hop during routing a message. This approach 
balances between choosing the node that will further progress the routing
towards the destination and choosing the closest entry in the routing table, in
terms of network proximity. Thus, it is relatively less effective than
geographical layout when applied to CAN(-like) implementations. Moreover, the
technique has been incorporated into a version of Chord causing an increase on the overhead of
node joins and the size as well as maintenance cost of finger tables.
%
%\cite{dabek_cfs_2001} proposes a server selection scheme for the Chord DHT, on
%the domain of proximity routing selection. In \emph{CFS}, each node predicts
%the entire lookup latency as a function of the total number of nodes and the
%average overlay next routing peer. The problem is that it is very difficult to
%have a clear picture on the total number of nodes and the average hop latency
%from the local. This leads to rough estimations that consequently decreases
%overall performance.
%
The methods that use proximity routing and their details are listed below.

\paragraph*{\bf Proximity in Kademlia}
Before discussing {\em Proximity in Kademlia}, it would be helpful to briefly
summarize the highlights about the {\em Kademlia} protocol.
{\em Kademlia} \cite{maymounkov_kademlia_2002} is a distributed hash table (DHT) based
protocol for P2P networks, which uses an iteration based lookup algorithm. The
protocol uses the standard $160$ bit ID system for nodes and locates the nodes in
a prefix binary tree, where IDs are used as prefixes. The iterative lookup
operations are done over this prefix binary tree, which converges to logarithmic lookup
times. The ID's are assigned randomly, therefore in Kademlia,
there is no proximity control, which results in inefficient use of the underlying
network during lookup and retrieve operations.

{\em Proximity in Kademlia} \cite{kaune_pkad_2008} introduces proximity controls
over the base Kademlia protocol to optimize the underlying network usage of the
protocol. The authors define an abstract {\em underlay metric} that calculates
the suitability of establishing a communication link between peers as a cost
function, based on the used proximity criteria (RTT, ISP locality, etc.).
{\em Proximity in Kademlia} adapted both {\em proximity routing} and {\em
proximity neighbour selection} overlay optimization algorithms for controlling proximity. Therefore,
we categorize the protocol only in this section, and skip the discussion in the
further {\em proximity neighbour selection} section (Sec.\ref{sec:pns}). {\em
Proximity in Kademlia} also uses the MaxMind GeoIP database\footnote{MaxMind
Geolocation Technology. http://www.maxmind.com} to detect the
proximity information of a given peer based on its region, country, or ISP
location, and use this information to form clusters of nearby peers. One
other method {\em Proximity in Kademlia} uses is the Vivaldi \cite{cox_vivaldi_2004} protocol. However, authors report that
clustering approach performs better than Vivaldi protocol based on their
experiments. Authors also report that {\em Proximity routing} protocol
successfully worked with the Kademlia protocol and improved the locality of the
connections over peers.

\paragraph*{\bf CHOord considering Proximity on IPv6 (CHOP6)}

\textit{CHOP6} \cite{morimoto_chop6_2007} is designed based on the
\textit{Chord} protocol. \textit{CHOP6} roughly estimates the proximity among
nodes by exploiting the IPv6 address format and RTT information if available.
The proximity estimation is achieved by introducing a 64-bit ID scheme in which
the least significant bit part is the IPv6 global routing prefix and thus
enabling a longest prefix match scheme. The protocol is designed based on the
observation that it is possible to estimate a node's geographical location by
simply examining the upper 32-bits of its IPv6 address. Moreover, similar to
\textit{Chord}, \textit{CHOP6} uses a finger table, whose entries hold more than
one candidate node.

\paragraph*{\bf Chord6}
\emph{Chord6} \cite{xiong_chord6_2005} is another \textit{Chord} variant that
tries to exploit the hierarchical features of IPv6 in order to create a
substrate that reduces interdomain traffic between service providers.
\textit{Chord6} is based on the original Chord protocol, and the main difference
from the original protocol is in the identifier definition. Therefore, the
approach can be easily portable to other DHTs such as CAN, Pastry and Tapestry.
In Chord6 the identifier contains two parts: the higher bits are obtained by
hashing the node's IPv6 address prefix of specific length, while the remaining
lower bits are the hash value of the rest of that IPv6 address. As a result of
this assignment, nodes in a domain will be mapped onto a continuous key space on
the overlay network, which avoids unnecessary message forwarding across
different service providers, thus minimizing overall routing cost.

\subsection{Proximity Neighbour Selection}\label{sec:pns}
\textit{Proximity Neighbour Selection} constructs the routing tables using proximity
knowledge. The proximity information used in this method is different than the
landmark based systems described in the \textit{Geographic Layout} section, as TTL values between
nodes, or directly node ID prefixes are used to detect proximity.  Tapestry,
Pastry, and CAN successfully implemented the proximity into their algorithms by
using this approach.  The routing protocol in Pastry is based on longest node ID
prefix matching, while CAN uses RTT values to detect close by nodes.
\cite{castro_proximityp2p_2002}  reports that proximity neighbour selection as
an effective proximity based method.
%
Proximity Neighbor Selection based algorithms are described in detail below.

\paragraph*{\bf DHT-PNS}

Chord-DHT-PNS \cite{hancong_pnsbased_2006} implements proximity neighbour
selection on top of the Chord DHT. The main purpose is to use proximity
information to group physically close by nodes as neighbours in the DHT table.
In order to detect proximity, virtual network coordinates of peers are used, by
using the Vivaldi protocol \cite{cox_vivaldi_2004}. The virtual coordinates are
then used by the nodes to map to identifier space in the DHT. The space is
partitioned using a \emph{concentric circle clustering scheme} where successive
cycles of radiuses $\rho$, $2\rho$, $3\rho$ and so on, are constructed. Then the
formed annuluses are divided into $2\chi-1$ \emph{sectors}, where $\chi$ denotes
the sequence number of the annulus starting from $\chi = 1$ for the centre
cycle. It is proved in the paper, that this way each sector occupies the same
area as does the center cycle. Assuming uniform node distribution, this
characteristic, favours a more load balanced clustering operation. Every sector
in this $2D$ coordinate space is mapped to a unique \emph{region} in the DHT
space forming a multi-layer node identifier space.  Thus, any nodes that belong
to the same sector, are mapped to the same region as well, preserving their
proximity relationship unveiled by the use of the Vivaldi protocol. The
individual pieces are mapped to the identifier space uniquely, allowing
logarithmic lookup operations with high probability on Chord.

\paragraph*{\bf IP-based Clustering (IPBC)}
Proximity neighbour selection algorithms use probing and other measures
to detect proximity. However, such methods are either are not precise enough, or
creates overload in the network. \emph{IP-based clustering}
\cite{karwaczynski_ipbc_2007} is a proximity neighbour selection based
algorithm in which the authors propose to use the IP address prefixes (16 bit
for IPv4) in order to detect proximity.  \cite{freedman_iploc_2005} states that
$97\%$ of prefixes larger than $24$ bits belong to a single geographical
location. However, using less number of bits creates less precise results and
more number of bits increase the burden on the network and reduce the possible
number of neighbours. Therefore, a careful selection is required in terms of
performance/accuracy tradeoff. The {\em IP-based Clustering} generates a key and
its prefix and stores its prefix in the DHT itself, so that any newly joining
nodes with the same IP prefix can query the prefix and identify all the
neighbours with the same prefix relatively easily. Nodes periodically update
their entries in the DHT and remove their entry when they leave, or the entry is
timed out with no further activity detected from the node.

\paragraph*{\bf Cone}

\textit{Cone}\cite{wang_cone_2007} extends the Chord using proximity neighbour selection
topology optimization algorithm. The proximity information is generated using
landmarks and RTT based distances to landmarks.  \textit{Cone} uses a two-layered
identifier space. The first, named Chord-layer identifier, denoted as
$Id_{Chord}$, is the same as in Chord. The second is the Cone-layer
identifier, $Id_{Cone}$ which is constructed by two component identifiers. The
first, known as \emph{group identifier (gid)} denotes a relevant group the node
belongs to while the second, namely \emph{local identifier (lid)} indicates the
local identifier within the group. The group concept, which is introduced here,
is a way of dividing nodes according to a common $Id_{Chord}$ prefix.  The
structure of a Cone overlay, retains the Chord's circular topology. The difference
lies on the fact that, now, two rings are created. A big ring, where nodes with
the same \emph{gid} are arranged at each position. Each of these positions are a
smaller ring for the particular group's \emph{lid}s. The routing is achieved in
both clockwise and counter-clockwise directions in the big-ring, for which two
routing tables are maintained, namely \emph{front} and \emph{back} finger
tables. Entries in these tables, display physical network
proximity with the current node. Moreover, a third table called \emph{group}
table maintains information about other online peers within the current node's
group in a way that entries are now close in the ID space.

\paragraph*{\bf SAT-Match: Self-Adaptive Topology Matching}

\textit{SAT-Match} \cite{ren_satmatch_2004} is a protocol that specifically
tries to solve the topology mismatch problem by mapping the overlay network as
close as possible to the physical one. Even though no theoretical proof is
given, valid results are obtained and demonstrated through experimentations.
Similar to the other approaches in the proximity neighbour selection,
\textit{SAT-Match} uses probing to detect close by peers and obtain proximity
information. However, as an additional feature, \textit{SAT-Match} peers can do
selective jumps to adjust their locations in the DHT, if it reduces the stretch
of its one-hop neighbourhood. The stretch is defined as the ratio between the
average logical and physical latency of a link. It is reported that, due to the
selective jumps, the \textit{SAT-Match} achieves $40\%$ reduction in link
stretch, and when used with the \textit{Landmark Binning} approach (see Sec.
\ref{sec:landmark_binning}), the reduction rate increases up to $60\%$.
For dynamic environments, with frequent node arrival and leave,
\textit{SAT-Match} scales much better than \textit{Mithos}, due to its self
adaptation mechanism and selective jumps.

\emph{SAT-Match} uses a small TTL value for the probing messages in order to
reduce redundancy\cite{jiang_lightflood_2008}. This process begins as soon as
the node joins the network using a DHT mechanism. Each probing message contains
information about the source and a small TTL value. The recipient of such a
message returns information about itself to the source and forwards the probing
message to its neighbours if the TTL is non-zero. The discovered nodes are
referred to as $TTL-k$ neighbourhood of the source node based on the $TTL$ distance
to the source node.

Blindly selecting the peer with the smallest RTT as neighbour is, generally, not
the optimal decision to make in order to achieve global \emph{stretch} reduction.
In a structured scheme, when a node jumps to connect to a
physically close node, it may need to connect to other distant nodes to maintain
the structure's integrity, thus creating an overall increase in the overlay's
\emph{stretch}. The two nodes with the smallest RTT is then used in order to
select one zone to jump in this phase. The algorithm is as follows: The source
node $S$ calculates the stretch change of its $TTL-1$ neighbourhood and that of
the $TTL-1$ neighbourhood of the first of the previously selected peers. These
calculations are made as if the jump has been made. If the stretch reduction is
over a predefined threshold the jump is performed, otherwise the second selected
candidate is picked and the same computations are performed. If again, the
threshold is not met, then no jump is ultimately done. In case of a jump, this
is performed as a combination of \emph{leave} and \emph{join} operations, in the
CAN context.

%\begin{figure}[h!]
\hspace{-3ex}
\begin{center}
\footnotesize
\begin{landscape}
\begin{longtable}{
|>{\columncolor[gray]{.7}}m{0.1\columnwidth}
|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
|>{\columncolor[gray]{.9}}m{0.2\columnwidth}
|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
|}
\caption{Decentralized Structured Algorithms}\label{fig:struct_compare_table}\\
\hline
\rowcolor[gray]{.5}
\textbf{Algorithm} & \textbf{Arch.} & \textbf{Overlay structure} & \textbf{Base protocol} &
\textbf{Dynamic update} & \textbf{Runtime} & \textbf{Scalability} & \textbf{cites}\\
\hline
\endfirsthead
\multicolumn{4}{c}%
{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
\hline
\rowcolor[gray]{.5}
\textbf{Algorithm} & \textbf{Arch.} & \textbf{Overlay structure} & \textbf{Base protocol} &
\textbf{Dynamic update} & \textbf{Runtime} & \textbf{Scalability} & \textbf{cites}\\
\hline
\endhead
\hline \multicolumn{4}{r}{\textit{Continued on next page}} \\
\endfoot
\hline
\endlastfoot

\hline
\textbf{Global Softstate} & decentralized structured &
\textbf{Proximity based, landmark clustering} Uses first landmark clustering
then measures RTTs to identify close nodes & & Yes &  &  & 195 \\

\hline
\textbf{Mithos} & decentralized structured &
\textbf{Proximity based} Uses nodes as topology landmarks and directed
incremental probing to optimize topology & & Yes &  & Scales well as all
operations are local ??? & 172 \\

\hline
\textbf{Self-Adaptive Topology Matching} & decentralized structured &
\textbf{Proximity based} Uses lightweight probing and
selective jumps to optimize the topology & CAN & Yes &  & Better than Mithos
& 40 \\

\hline
\textbf{Delay Aware P2P System} & &
\textbf{} & &  &  &  & 1 \\

\hline
\textbf{VERSION OF CHORD - DHT-PNS} & decentralized structured &
\textbf{Proximity based} Uses Proximity Neighbour Selection and the Vivaldi
system & Chord  & Yes &  &  & 5 \\

\hline
\textbf{MAY OMIT- VERSION OF CHORD - Quasi-Chord} & decentralized structured &
\textbf{Proximity based}  & Chord  & Yes &  &  & 0 \\

\hline
\textbf{LAPTOP} & decentralized structured &
\textbf{Geographic layout based} Hierarchical overlay structure  &   & Yes &  & routing path length $\log{_d N}$, 
join/leave overhead $d\log{_d N}$& 4 \\

\hline
\textbf{IP-Based Clustering} & decentralized structured &
\textbf{Proximity based} Proximity neighbour selection based on longest common
prefix of IP addresses &   & Yes &  &  & 1 \\

\hline
\textbf{CHOord considering Proximity on IPv6} & decentralized structured &
\textbf{Proximity based} Uses IPv6 address format to provide proximity &  Chord
& Yes &  & Better than Chord & 1 \\

\hline
\textbf{Proximity in Kademlia} & decentralized structured &
\textbf{Proximity based} Applies  proximity neighbour selection (PNS) and proximity route selection (PRS)
to Kademlia & Kademlia & Yes &  &  & 13 \\

\hline
\textbf{Cone} & decentralized structured &
\textbf{Proximity based} Uses proximity neighbour selection (PNS) & Chord & Yes
&  & Better than Chord & 3 \\

\hline
\textbf{Dynamo} &  & &  & 
&  & & 3 \\

\hline
\textbf{MAY OMIT-BADLY WRITTEN-PChord} & decentralized structured &
\textbf{Proximity based}  &  Chord
& Yes &  & & 16 \\

\hline
\textbf{AChord} & &  &  
&  &  & & 7 \\

\hline
\textbf{Chord6} & decentralized structured &
\textbf{Proximity based} Uses IPv6 hierarchical address format to cluster
topologically close nodes & Chord & Yes
&  &  & 14 \\

\hline
\end{longtable}
\end{landscape}
\end{center}
\vspace{-2.5ex}
\vspace{-2.5ex}
%\label{fig:struct_compare_table}
%\end{figure}

\subsection{Discussion on Structured Decentralized Algorithms}

In section \ref{sec:structured},  we presented the state of the art structured
decentralized P2P algorithms in three different categories based on their
handling of the network structure to tackle the topology mismatch problem:
geographic layout based, proximity routing based, and proximity neighbour
selection based. In this section, we present a final discussion of all the
methods including the advantages, disadvantages,
and novelties related to solve the topology mismatch problem.
Table \ref{fig:struct_compare_table} presents an overview of all the
structured decentralized algorithms described earlier in this section.

Structured P2P network algorithms use a global distributed hash table or a
prefix tree structure to uniquely lookup peers or their data in the overlay
network. As all the data is kept within the overlay, each node behaves as a
client and a server, therefore nodes join and leave according to rules
determined by the integrity of the global data structure. The main advantage of
the structured P2P topology is that by the help of the global data structure,
peers or their data can be found within the network even if there is only a
single copy of that item present. However, each node join and leave creates
maintenance overhead for the network due to updates required by the global data
structure, and for networks with frequent node arrivals and departures the
topology uses valuable network resources just to update the global structure.
Nodes join the network by using a key value, which determines the location and
the neighbourhood of the new node within the network. However, assigning
random key values to the newly inserted nodes creates non-optimal matching with
the underlying physical network topology, therefore, increasing the overhead of
the network even more. One solution for handling the topology mismatch problem
is to consider the proximity of the peers when generating the key and joining
the node to the network, so that nodes within the same network domains are
selected as peers, or neighbours, during the overlay topology construction. In
this chapter, we have described three such approaches to optimize the topology
matching problem: geographic layout, proximity routing and proximity neighbour
selection.

In geographic layout, nodes try to estimate the geographic positions of the
peers, and construct the DHT considering the proximity of the peers. Landmark
servers and RTT measurements are two popular methods, which can also be used in 
conjunction, to discover physically close by peers over the network. However, as
discussed in Section \ref{sec:landmark_binning}, these methods do not always
give reliable estimates for the node positions over the internet. The landmark
servers are not self-organizing and have maintenance overheads. To serve a
P2P network with millions of peers, multiple landmark servers
distributed over the whole world is required, which is hard to manage. The RTT
measurements also can measure the delay between peers, but it is a greedy method,
which can result non-optimal overlay topologies especially if close by nodes
have low bandwidth connections among themselves.

Proximity routing does not discover or store proximity information for the
peers, however, during package forwarding, nodes that have lower latency to the
destination, or with a closer key to the destination is selected. As no
proximity information is used, the system is based on a greedy approach and it
usually selects low latency paths over the overlay, which maps to suboptimal
longer paths on the physical topology. For these obvious reasons, proximity
neighbour selection is generally considered as a superior method to proximity
routing, however, joint uses of these two protocols are also possible.

Proximity neighbour selection approach is similar to proximity routing, with an
exception that during forwarding, the proximity information of the peers are
also considered. Therefore, for each application, proximity information is
extracted and used during routing. Depending on the application, the proximity
information can be the node sharing the same prefix with the current node, or
TTL to detect close by neighbours. Tapstry and Pastry are two popular methods
implementing the proximity neighbour selection among others mentioned above in the
proximity neighbour selection section.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{The Algorithms}
%Having understood the nature of the problem that peer-to-peer architecture face, this section starts the discussion of the academic work that has been conducted in the field the last few years. What follows this introduction, does not claim to be a thorough citation of all known protocols that are available out there, but instead, a carefull selection of those that left a distinctive fingerprint contribution in the efforts of the research community to alleviate the topology mismatch problem.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% UNSTRUCTURED
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\subsection{Narada}
%
%\paragraph*{}
%\emph{End System Multicast} \cite{chu_esm_2000,chu_esm_2001,chu_esm_2002} as well as other  proposes an architecture, where end-systems implement all multicast related functionality (including membership management and packet replication). In \emph{Narada}, the protocol developed in the paper's context, end-systems
%\begin{inparaenum}[\itshape i\upshape)]
%  \item self-organize, in a fully distributed manner,
%  \item into an efficient overlay structure, both from the network and the application perspective,
%  \item with self improving properties,
%  \item adapting to network dynamics\footnote{Long-term variations in Internet path characteristics (i.e. bandwidth, latency e.t.c.)}.
%\end{inparaenum}
%
%\paragraph*{}
%What Narada really does is to construct overlay spanning trees in a two-step process:
%\begin{enumerate}
%  \item It builds a richer connected graph \footnote{ENarada \cite{li_enarada_2008} used Gossip protocol for the construction}, termed a \emph{mesh}, while trying to ensure a couple of desired performance properties:
%    \begin{itemize}
%      \item path quality between any pair of members should be comparable to the quality of the unicast path between them, and
%      \item each member should have a limited number of neighbours on the mesh.
%    \end{itemize}
%
%  \item It computes minimum spanning trees out of the mesh, each rooted at the corresponding source member.
%\end{enumerate}
%
%\paragraph*{}
%Narada incrementally improves the quality of the mesh by adding and dropping overlay links. Members probe each other at random and new links are added depending on what is the gain in doing so. On the other hand, participants, continuously  monitor the utility of existing links, and drop them if they are found to be not usefull.
%
%\paragraph*{}
%Narada protocol achieves relatively good performance for small and medium sized groups involving tens to hundreds of members. This is not, by no means, what a contemporary peer-to-peer system needs to be characterized as scalable, but this work is considered pioneering\footnote{Another similar approach is \emph{Scattercast}\cite{chawathe_scattercast_2000}}, in that it was the first to have conducted a detailed Internet evaluation to analyse and demonstrate the feasibility of an overlay-based, application layer, multicast architecture that can dynamically adapt to bandwidth and latency properties of the physical underlying infrastructure.
%
%\subsection{Gia}
%
%\paragraph*{}
%\emph{Gia} \cite{chawathe_gia_2003} is a decentralized unstructured peer-to-peer file sharing system with the following characteristics:
%\begin{inparaenum}[\itshape i\upshape)]
%  \item replaces Gnutella's flooding with random walks \cite{lv_randomwalks_2002},
%  \item recognizes the importance of overlay network's topology, while using the random walks, and exploits it by incorporating a topology adaptation algorithm,
%  \item introduces a token-based flow control mechanism, and
%  \item detects the significant heterogeneity in peer bandwidth, processing power, disk speed e.t.c. and takes advantage of it.
%\end{inparaenum}
%
%\paragraph*{}
%More specifically, there are four key components in the design of \emph{Gia} which are summarized bellow:
%\begin{enumerate}
%  \item A \emph{dynamic topology adaptation} protocol that puts participating nodes within short reach of high capacity nodes so that these \emph{high-degree} nodes, which due to their high connectivity receive most of the queries, actually have the capacity to handle them.
%  \item An \emph{active flow control} scheme is used to avoid overloaded hot-spots. Heterogeneity is detected and flow control tokens are given to nodes based on the available capacity.
%  \item Every node maintains pointers to to the content that is offered by their immediate neighbours, creating a \emph{one-hop replication} of pointers, scheme.
%  \item A \emph{search protocol} based on random walks, that is biased towards directing queries to high-capacity nodes that are typically best able to answer these queries.
%\end{enumerate}
%
%\subsection{Adaptive Overlay Topology Optimization}
%
%\paragraph*{}
%\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{liu_auto_2003} is an algorithm for building an overlay multicast tree among each source node and its direct logical neighbours in order to alleviate the mismatching problem while providing a larger query coverage range. It includes the steps of
%\begin{inparaenum}[\itshape i\upshape)]
%  \item \emph{Selective Flooding (SF)}, and
%  \item \emph{Active Topology (AT)}
%\end{inparaenum}
%which are furtherlly discussed in the following paragraphs.
%
%\paragraph*{Selective Flooding (SF)}
%Instead of flooding to all neighbours, SF builds an overlay multicast tree among each peer and its immediate logical neighbours. This reduces overall message burden to the network by making some neighbours non-flooding. The tree is formed using a minimum spanning tree algorithm. For this reason a peer needs to know the costs to all its logical neighbours (e.g. the network delay) as well as between any pair of neighbours. Additional, cost probing, messages need to be added to the original messages defined by the Gnutella protocol.
%
%LTM's SF effectiveness has be proven to be detached from the different physical or overlay topologies. On the other hand, SF is more effective with large number of logical neighbours. It can reach an average optimization rate of 87.4 percent on a logical topology with an average of 30 logical neighbours. 
%
%\paragraph*{Active Topology (AT)}
%At this step the overlay topology is reorganized. Each peer, independently, makes optimizations in the overlay network to alleviate topology mismatching by replacing non-flooding neighbours, with closer nodes. The approach incorporated by the paper is called \emph{Randomized AT} algorithm which picks up a candidate peer at random among the non-flooding neighbour's neighbours. Whenever a new neighbour cost table is received or there is a change of neighbours, the source peer has to re-calculate the multicast tree and apply the randomized AT algorithm.
%
%Different numbers of average logical neighbours has little to do with the effectiveness of AT. If the source has $n$ non-flooding peers, there are $n$ potential neighbour replacements. The overhead to exhaust all $n$ possible replacements can be high, so in practice, after each replacement the source peer can decide whether it needs to find another candidate peer. This is done by computing the cost improvement ratio greater than some predefined termination threshold. The larger the threshold, the slower, in the number of optimization steps, the reduction of the normalized average distance. As a whole the average response time is significantly reduced when more optimization steps taken.
%
%\subsection{Location-aware Topology Matching}
%
%\paragraph*{}
%\emph{Location-aware topology matching} or \emph{LTM} \cite{liu_ltm_2004}, is an algorithm of building an efficient overlay by disconnecting low productive connections and choosing physically closer nodes as logical neighbours while still retaining the search scope and reducing response time of queries. This is done by issuing detector messages by peers, to a small region so that receivers of the detector can record relative delay information and cut inefficient or redundant logical links as well as add closer nodes as direct neighbours. Three operations are defined in LTMband are furtherly discussed in the following paragraphs.
%
%\paragraph*{TTL2-detector flooding}
%\emph{TTL2-detector} messages are designed based on the specification of the Gnutella protocol. Suppose that $d(i, S, v)$ denotes the TTL2-detector which has the message ID of $i$ with TTL value of $v$ and is initiated by $S$. In addition to Gnutella's unified 23-byte header, such a message has a body of two types:
%\begin{itemize}
%  \item short format $d(i, S, 1)$ if $S$ is a direct neighbour of the receiver
%  \item long format $d(i, S, 0)$ if $S$ is within a two-hop distance from the receiver
%\end{itemize}
%This way a peer can calculate link costs from within a one- or two-hop area around it and thus reach peers in $N(S)$ and $N^2(S)$ sets, respectively, in the paper's context. Each peer floods a TTL-detector periodically while it calculates link cost to another peer by receiving other peer's detector messages and checking:
%\begin{itemize}
%  \item the Source Timestamp field and its own clock upon time of receipt for one-hop messages
%  \item the Source Timestamp, TTL1 Timestamp and its own clock upon receipt for two-hop messages
%\end{itemize}
%Of course peer clocks must be synchronized in order to do such checking.
%
%\paragraph*{Low productive connection cutting}
%There are three cases for any peer $P$ who receives $d(i, S, v)$ multiple times:
%\begin{inparaenum}[\itshape i\upshape)]
%  \item $P$ receives both $d(i, S, 1)$ and $d(i, S, 0)$
%  \item $P$ receives multiple $d(i, S, 0)$s from different paths, and P randomly chooses to process one
%  \item $P$ receives one $d(i, S, 1)$ and multiple $d(i, S, 0)$s, and $P$ processes $d(i, S, 1)$ and one randomly selected $d(i, S, 0)$
%\end{inparaenum}
%If the link with the largest cost is found and is a direct neighbour then the connection is put in a will-cut list and stays there for a certain period of time. If it is not, then it is handled by other peers. After that period, connections are cut and recorded to $P$'s cut-list.
%
%\paragraph*{Source probing}
%For a peer $P\in(N^2(S) - N(S))$ who receives only one $d(i, S, 0)$, the cost of $PS$ is obtained (with list look-up or probing). Then $P$ compares it with the cost from each hop and if $PS$ has the largest cost, $P$ will not keep this connection, while otherwise the connection will be created.
%
%\paragraph{}
%Supposing $n$ is the number of peers, $c_n$ is the average number of neighbours and $c_e$ is the average cost of logical links, then in the flooding-based search the traffic incurred by one query from an arbitrary peer in a peer-to-peer network is $O(n)$. As observed in the Gnutella network \cite{sripanidkulchai_gnutella_2001}, each peer issues $0.3$ queries per minute in average, thus the per minute traffic incurred by the network with $n$ peers is $O(n^2)$. Because each $d(i, S, v)$ has a TTL of $2$ in each source peer, the traffic for one time LTM optimization in all peers is at most $2nc_n^2c_e$. If each peer uses LTM $k$ times per minute, the total traffic incurred is $2knc_n^2c_e$. Simulation shows the best value for $k$ is $2$ or $3$. So, the traffic overhead caused by LTM to the network is $O(n)$.
%
%TTLj-detectors, with $j > 2$, would detect and break cycles with more than 4 links. LTM though, does not use such detectors because detector-flood traffic would increase significantly, and cut links between two end-peers, could cause queries initiated by them to traverse a path much more expensive than the cost on the the cut link.
%
%\paragraph{}
%LTM disadvantages are
%\begin{inparaenum}[\itshape i\upshape)]
%  \item disagreement of measured delay due to unsynchronized clocks causes problems when deciding the cut positions, which can influence the network connectivity, and
%  \item the network delay metric mainly focuses on disabling the connections between peers physically far away without considering the shortcuts created by powerful peers.
%\end{inparaenum}
%
%\begin{figure}
%\centering
%  \includegraphics[scale=0.4]{img/ltm_multid.jpeg}
%\caption{Peer $P$ receives $d(i, S, v)$ multiple times}
%\label{figure:ltm_multid}
%\end{figure}
%
%\begin{figure}
%\centering
%  \includegraphics[scale=0.4]{img/ltm_example.jpeg}
%\caption{A full example of LTM}
%\label{figure:ltm_example}
%\end{figure}
%
%\subsection{Scalable Bipartite Overlay}
%\emph{Scalable Bipartite Overlay (SBO)} \cite{liu_bipartite_2007} employs an efficient strategy to select query forwarding paths and logical neighbours. The topology construction and optimization of SBO consist of four phases:
%\begin{enumerate}[\itshape i\upshape)]
%  \item When a new peer is joining the network, it will randomly take an initial colour; say red or white. This colour is not to be changed for the time the peer stays connected. If it leaves and then rejoins it will, again, have to pick a colour, randomly. Thus all peers are separated into two groups, red or white. Then the bootstrap host will provide, the joining peer, with a list of active peers, including their colour information, in order for the later to establish connections to different colour peers. This way, all peers form a bipartite overlay.
%  \item White peers probe distances with their immediate (red) neighbours, form a cost table and send this table to their corresponding red neighbours.
%  \item Each red peer builds a minimum spanning tree using the obtained neighbour cost tables. These are two-hop diameter trees so white peers do not need to build one. The links that are part of a minimum spanning tree are called \emph{forwarding connections (FC)} while the rest \emph{non-forwarding (NFC)}.
%  \item Having a minimum spanning tree with two hops a red peer is able to send its queries within that range. Some white peers, though, have become non-forwarding neighbours. In this phase, such a (white) neighbour will try to find another red peer being two hops away from its current red neighbour to replace the later as its new neighbour.
%\end{enumerate}
%
%\begin{figure}
%\centering
%  \includegraphics[scale=0.4]{img/sbo_efficient_forward.jpeg}
%\caption{A red peer $P$ has a small overlay topology of $N(P)$ and $N^2(P)$ and computes the efficient forwarding paths}
%\label{figure:sbo_efficient_forward}
%\end{figure}
%
%\begin{figure}
%\centering
%  \includegraphics[scale=0.4]{img/sbo_neighbour_replace.jpeg}
%\caption{Neighbor replacement}
%\label{figure:sbo_neighbour_replace}
%\end{figure}
%
%In a static enviromnent LTM may reduce traffic cost by around 80 to 85 percent while SBO reduces traffic cost between 85 and 90 percent. However, LTM  is proved to converge in around 2-3 steps while SBO needs 4-5 steps. Moreover LTM reduces response time by more than 60 percent in 3 steps while SBO needs 8. In a dynamic environment (10 minute average peer lifetime, 0.3 queries/sec by each peer) SBO and LTM reduce the average traffic cost per query (including the overhead due to the optimization steps) by 85 and 80 percent, respectively. Moreover LTM reduces the response time per query to 30 percent while SBO to 35 percent.
%
%\subsection{Adaptive Connection Establishment}
%
%\paragraph{}
%\emph{Adaptive Connection Establishment (ACE)} \cite{liu_ace_2004} builds an overlay multicast tree among each source node and the peers within a certain diameter from the source peer and optimizes the neighbour connections that are not in that tree.
%
%\paragraph{}
%ACE indicates three phases for its algorithm:
%\begin{enumerate}[\itshape i\upshape)]
%  \item Calculate cost between nodes using network delay as a metric. Each peer probes the costs with its immediate logical neighbours and forms a \emph{neighbour cost table (NCT)} using a special routing message type. Two neighbouring peers exchange their NCTs in order for every peer to obtain the cost between any pair of its local neighbours forming a small overlay topology.
%  \item Based on obtained NCTs a minimum spanning tree among each peer and its immediate neighbours is built (Figure~\ref{figure:ace_phase2}).
%  \item Physically far away neighbours are replaced by physically close neighbours. In ACE a peer, say $S$, probes the distance between one of its non-flooding neighbour's neighbour, say $G$ and $H$ respectively. If the link to neighbour's neighbour is smaller than that to the neighbour, the later is cut. If this is not the case but $S$ finds that cost of $GH$ is even larger than that of the $SH$, $S$ will keep $H$ as a new neighbour. Obviously, if $SH$ is larger than $SG$ and $GH$, the connection will not be established and $S$ will continue probing another neighbour's neighbour. The algorithm described is conducted within $1$-neighbour closure (among its source peer and all its direct neighbours) but the optimization scope can be enlarged. The larger the scope, the better topology matching improvement but also the greater the computational overhead (Figure~\ref{figure:ace_phase3}).
%\end{enumerate}
%
%\paragraph{}
%Simulations in \cite{liu_acesims_2004} show that the average scope of each query to cover the same scope of nodes is reduced by about 65 percent without losing any autonomy feature, while the average response time can be reduced by 35 percent. Larger diameter topologies lead to better topology optimization rate but also to higher communication and computation overhead. It was also found that it is more effective in higher connectivity dense topologies. Compared to LTM, it comes short of convergence speed. In \cite{ni_mismatch_2004} shows reduction of both total traffic (90 percent) and response time (80 percent) to message queries without shrinking the search scope. SBO, on the other hand, achieves approximately 85 percent reduction on traffic cost and about 60 percent reduction on query response time. Last but not least, it is concluded that work must be done on incorporating a more sophisticated selection policy for candidate non-flooding peers.
%
%\begin{figure}
%\centering
%  \includegraphics[scale=0.4]{img/ace_phase2.jpeg}
%\caption{Second phase in ACE}
%\label{figure:ace_phase2}
%\end{figure}
%
%\begin{figure}
%\centering
%  \includegraphics[scale=0.4]{img/ace_phase3.jpeg}
%\caption{Third phase in ACE}
%\label{figure:ace_phase3}
%\end{figure}
%
%\subsection{Hops Adaptive Neighbour Discovery}
%
%\paragraph{}
%Paper \cite{chen_hand_2006} proposes a new algorithm named \emph{Hops Adaptive Neighbour Discovery (HAND)} which uses a fully distributed triple hop adjustment strategy to address the topology mismatch problem. The advantages of the algorithm compared to other approaches are that
%\begin{inparaenum}[\itshape i\upshape)]
%  \item it does not need any clock synchronization,
%  \item it is a fully distributed algorithm making it robust and reliable in decentralized systems,
%  \item the traffic overhead of the triple hop adjustment is very low,
%  \item it is applicable to dynamic peer-to-peer environments, and
%  \item maintains lower query response time.
%\end{inparaenum}
%
%\paragraph{}
%The algorithm's ultimate goal is the optimal overlay, the \emph{Logical Communication Network (LCN)} in the paper's context. The key concept of the algorithm is that a graph $G^{*}$ that describes an LCN and a graph $G$ that describes the current overlay are matched only if all peer hop sequences $(v_1, v_2, \ldots, v_k)$ in $G$ exist in $G^{*}$ and in the same order. In practice triple sequences $(v_1, v_2, v_3)$ are used.
%
%\paragraph{}
%The mismatching detection is done in the following way. Suppose we want to verify peer sequence $v_2-v_1-v_3$ (see Figure~\ref{figure:hand_com_overlay}). A pair of probing messages are sent from $v_1$ to $v_2$ and $v_3$. Suppose delays of $(v_1,v_2)$ and $(v_1,v_3)$ are are denoted as $x$ and $z$, respectively. When the probing message arrives to $v_2$ it forwards it directly to $v_3$. Similarly, when the probing message arrives to $v_3$, it forwards it directly to $v_2$. These last steps are performed in order to obtain delays of $(v_2,v_3)$ and $(v_3,v_2)$ physical paths, respectively, denoted by $y$.
%\begin{itemize}
%  \item If $y=z-x\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched and should be adjusted to $v_1-v_2-v_3$ by deleting edge $(v_1,v_3)$ and adding a new $(v_2,v_3)$ (see Figure~\ref{figure:hand_matchtriple_lemma3}).
%  \item if $y=x-z\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched and should be adjusted to $v_1-v_3-v_2$ by deleting edge $(v_1,v_2)$ and adding a new $(v_3,v_2)$ (see Figure~\ref{figure:hand_matchtriple_lemma4}).
%\end{itemize}
%$\varepsilon$ is a small positive real number denoting additional delays caused by possible forwarding and jitter delays.
%
%\begin{figure}
%\centering
%  \includegraphics[scale=0.4]{img/hand_com_overlay.jpeg}
%\caption{The communication in overlay}
%\label{figure:hand_com_overlay}
%\end{figure}
%
%\begin{figure}
%\centering
%  \includegraphics[scale=0.4]{img/hand_matchtriple_lemma3.jpeg}
%\caption{The matching triple in paper's Lemma 3}
%\label{figure:hand_matchtriple_lemma3}
%\end{figure}
%
%\begin{figure}
%\centering
%  \includegraphics[scale=0.4]{img/hand_matchtriple_lemma4.jpeg}
%\caption{The matching triple in paper's Lemma 4}
%\label{figure:hand_matchtriple_lemma4}
%\end{figure}
%
%\paragraph{}
%Measurements conducted for evaluation perposes showed that in a static environment the algorithm can effectively decrease traffic cost by about 77 percent and shorten the query response time by about 49 percent in less than two minutes. In a dynamic environment it shows similar behaviour and with the size of the overlay network having little impact on the effectiveness of the algorithm. Compared to LTM both algorithms have almost the same traffic reduction rate, however on the response time reduction rate HAND has a higher one by about 4 percent. The traffic overhead of HAND is much less than that of LTM by an average of 55 percent.
%
%\subsection{Distributed Cycle Minimization Protocol}
%
%\paragraph{}
%\emph{Distributed Cycle Minimization Protocol (DCMP)} which is introduced in \cite{zhu_dcmp_2008} is a dynamic, fully decentralized protocol that promises significant reduction of duplicate messages. To achieve that, it uses the slightly different approach of eliminating unnecessary cycles, while retaining the connectivity of the network and preserving fault resilience and load balancing properties of unstructured peer-to-peer schemas by avoiding the creation of a hierarchical organization.
%
%DCMP aims at cutting the cycle paths at strategic locations. Any peer that detects a duplicate message can initiate the cutting process which consists of the following steps:
%\begin{enumerate}
%  \item Peers in the cycle elect a leader called \emph{GatePeer}\footnote{GatePeers are important for maintaining the connectivity and optimal structure of the network while peers join and leave randomly.}.
%  \item The cycle is cut at a well-defind point with respect to the GatePeer
%\end{enumerate}
%
%\paragraph{}
%The first step after detecting a duplicate message by some peer is to gather information from all peers in the cycle using a new type of control message called \emph{Information Collecting Message} or \emph{ICM}. ICM contains:
%\begin{inparaenum}[\itshape i\upshape)]
%  \item a \emph{GUID}\footnote{Globally Unique IDentifier assigned to every query message generated by any node.} field same as the one of the duplicate message,
%  \item \emph{DetectionID} field which represents the direction of the connection where the duplicate was identified\footnote{This ensures the uniqueness of the ICM messages because as it travels through many cyclic paths, multiple peers will detect the duplicates and initiate an ICM message.}, and
%  \item \emph{Node Information Vector (NIV)} which contains information (bandwidth, CPU power, etc) about peers that propagated the ICM.
%\end{inparaenum}
%
%\paragraph{}
%Suppose $A$ detected the duplicate as depicted in Figure~\ref{figure:dcmp}. It then emits an ICM to $B$ and $F$, that initially contains information only about itself. Each peer that receives the ICM, appends its information and propagates it along the reverse path of the original message. Since two copies of ICM are sent, at some point, a peer, say peer $D$, will receive a duplicate ICM. Using the information in the NIVs of the ICMs, $D$, decides to cut (for example) the EF connection. To inform the other peers about its decision, it emits a \emph{Cut Message (CM)} which contains the GUID and DetectionID of the corresponding ICM and an additional field that identifies the connection to be cut. $D$, then, forwards the CM in the reverse directions from where the ICM arrived. Similarly CMs received by any peer are propagated toward the reverse path of the corresponding ICM. Eventually, either peer $E$ or peer $F$ will receive the CM and cut the connection, thus eliminating the cycle.
%
%Receiving a duplicate ICM denotes the existence of a cycle. The opposite is not true though. For example, if the cycle contains $2 \times TTL$ edges, it will not be detected, because the ICM messages will be discarded before they locate it. There is a tradeoff between preserving the connectivity of the network and minimizing the duplicates that makes such a possibility to be safely ignored.
%
%\begin{figure}
%\centering
%  \includegraphics[scale=0.4]{img/dcmp.jpeg}
%\caption{Cycle elimination methods in DCMP}
%\label{figure:dcmp}
%\end{figure}
%
%\paragraph{}
%Experiments in \cite{zhu_dcmp_2008} showed that DCMP incurs a lower delay, returns more results and decreases the number of duplicate messages by 22\%, compared to LTM. Additionally DCMP has one to two orders of magnitude less overhead, because it adopts a more ``LAZY'' approach than broadcasting control messages periodically like LTM does.
%
%\subsection{Two-Hop-Away Neighbour Comparison and Selection}
%
%\paragraph{}
%Work in \cite{liu_thancs_2005,liu_thancs_2008} proposes a distributed heuristic called \emph{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} that
%\begin{inparaenum}[\itshape i\upshape)]
%  \item is completely distributed and needs no global knowledge,
%  \item presents trivial overhead compared to the query cost savings
%  \item its convergent speed of the algorithm is fast enough (faster than minimum spanning tree approaches) so that is effective to dynamic environments, and
%  \item does not shrink the search scope.
%\end{inparaenum}
%
%\paragraph{}
%THANCS is considered a \emph{local search method}, in the sense that it targets in finding a locally optimum solution, by exploiting knowledge within a 2-hop radius. The algorithm consists of two main components: \emph{piggybacking neighbour distance on queries} and \emph{neighbour comparison and selection} which are furtherly discussed bellow.
%
%\paragraph{Piggybacking neighbour distance on queries}
%Using network delay as a metric for measuring the distance, each peer probes distances with its immediate naighbours and stores information locally. For this reason a special query message type, \emph{Piggy Message (PM)}, is introduced. It is 6 bytes long and includes two fields: Neighbour IP Address and Neighbour Distance. A peer $P$ constructs a PM for its neighbour $Q$, which contains $Q$'s IP address and $Q$'s distance from $P$. When $P$ receives a query from $Q$, this PM will be piggybacked by the query that goes to all other neighbours of peer $P$. Upon receiving such a query message, each of the other neighbours will detach the PM, record the $PQ$ distance and process the query. This PM will not be further forwarded. In selecting which incoming queries  should piggyback a PM the paper proposes the \emph{pure propability-based (PPB)} and the \emph{new neighbour triggered (NNT)} policies.
%
%\paragraph{Neighbour comparison and selection}
%Figure~\ref{figure:thancs} illustrates this component of the THANCS algorithm. A peer $S$ probes the distance to all known unprobed $N^2(S)$\footnote{$N^2(S)$ denotes the set of peers being two hops away from $S$, while $N(S)$ denotes the set of direct logical neighbours of $S$.}. The distance of $SP$ is known to $S$. Upon receiving a PM from node $P$ with the distance of $PQ$, $S$ follows one of the following:
%\begin{itemize}
%  \item $Q \in \left( N(S) \cap N^2(S) \right)$, i.e. $Q$ is direct neighbour of $S$. In this case $S$ will compare cost of $SQ$, $SP$ and $PQ$. If the most costly connection is one of $SQ$ or $SP$ the corresponding link will be put into a \emph{will-cut list}\footnote{Links are not immediately disconnected when put in the will-cut list. There are useless for forwarding but are kept active, for some time, in order to serve query responses that are traveling to the source peer along the inverse search path.}. If the most costly connection is $PQ$ then $S$ will do nothing, as the fully distributed nature of the algorithm will give a chance to either $P$ or $Q$ to cut the connection.
%  \item $Q \in \left( N^2(S) - N(S) \right)$, i.e. $Q$ is a two-hop-away neighbour of $S$. If $S$ hasn't probed $Q$ before\footnote{A distance cache is maintained and looked up in such a case.}, it probes and stores the result in the \emph{distance cache}. Having the distance of $SQ$, $S$ compares costs of $SQ$, $SP$ and $PQ$. If $SQ$ is the most costly, $S$ will not establish the connection. If $SP$ is the most costly, $S$ will establish connection $SQ$ and put $SP$ in the will-cut list. If $PQ$ is the longest, $S$ will keep  the connection with both $P$ and $Q$, expecting that $P$ or $Q$ will eventually disconnect link $PQ$, later.
%\end{itemize}
%
%\begin{figure}
%\centering
%  \includegraphics[scale=0.4]{img/thancs.jpeg}
%\caption{Probing two-hop-away neighbours}
%\label{figure:thancs}
%\end{figure}
%
%\paragraph{}
%In a static environment THANCS has been proven to be effective; optimizing 45 percent out of the 60 percent of mismatched paths, constructing a nearly optimal overlay. This leads to a 60 percent reduction in traffic cost as well as a 40 percent decrease in query response time. In dynamic environments (Gnutella 0.6/Limewire super-peer-like and Ion flat-like), THANCS saves up to 70 percent of the traffic cost in the super-peer topology and 55 percent for the flat one. Average response time is also decreased by 60 and 45 percent, respectively. Generally, THANCS has similar performance to LTM, without needing synchronization. SBO, incurring half the  overhead of AOTO, reduces the traffic cost the most, while THANCS has lower response time and converges faster than SBO. THANCS is, thus, more suitable for a more dynamic environment. In addition, THANCS is easy to implement and its operation overhead is trivial, compared with the other three approaches. This design, however, has the limitation of not being easily extend to also support non-flooding-based systems.
%
%\subsection{mOverlay}
%Zhang et al. in \cite{zhang_moverlay_2004} focus on two aspects of the overlay construction. First, to achieve \emph{efficiency} of communication, a protocol should minimize unecessary long-distance hops and redundant traffic. Second, to be \emph{scalable} the overlay should be constructed in a distributed fashion and maintenance cost (data and locality management) should be minimized. Both of the above observations, are positively affected by the exploitation of node proximity at the underlying network. To accomplish it, the paper introduces the \emph{group} concept, according to which non-static (dynamic) landmarks are used to compute proximity, resulting in a highly robust and scalable overlay construction with reduced maintenance cost as well. The dynamic nature of landmarks help towards load-balance, meaning that hot-spots are avoided.
%
%\paragraph{}
%The proposed architecture, called \emph{mOverlay}, is a \emph{clustering approach} which creates a two-level hierarchical network, where on the top level we have connections between groups while on the bottom we have connections between hosts inside groups, thus attempting to recreate \emph{Small-World}-like properties for the constructed overlay network. The notion of a \emph{group}, in the paper's context, is a set of hosts that are close to each other with respect to any position $P$ in the underlying network. The distance between hosts can be 
%\begin{inparaenum}[\itshape i\upshape)]
%  \item network latency,
%  \item round-trip time,
%  \item minimum bandwidth on the links along the path connecting the two nodes, or
%  \item some other user-defined cost metric between the two nodes.
%\end{inparaenum}
%These nodes are refered to as neighbouring. Similarly groups can exchange messages with their neighbouring ones\footnote{Groups nearby, in the underlying physical network.}.
%
%According to the \emph{grouping criterion} set by the authors, a new host $Q$ belongs to some group $A$ if the distance between $Q$ and group $A$'s neighboring groups is the same as the distance between group $A$ and it's neighbouring groups. The neighbour groups thus, play the role of the dynamic landmarks and for accuracy and performance reasons the nearest of them are chosen to play that role.
%
%\paragraph{Locating process} A new coming host, $Q$, first connects to a globally known host cache called the \emph{rendezvous point (RP)} in order to retrieve the starting point in the overlay, say $A$ in group $1$. Host $Q$ then, measures its distance to host $A$. At the same time, the later, sends information about the neighbour groups of group $1$ back to host $Q$. This list is called \emph{candidate group list}, and the newcoming host sequnentially measures its distance to each of them in seek for the closest one. If the \emph{grouping criterion} is met, host $Q$ belongs to group $1$. If not, a boot host from the closest group is found and the algorithm is re-run until the criterion is met or after a predefined number of repetitions. In the later case, $Q$ creates a new group comprising itself only. The above protocol does not favour hotspots as it spreads the probability of visiting a group across the whole overlay and limits the overhead in the level of $O \left ( log N \right )$.
%
%\paragraph{General overlay operations} A set of additional protocols, are also introduced, similar to those found in traditional unstructred networks, but modified focusing on scalability and robustness. For example a protocol for \emph{group formation} is introduced that exploits the inherent characteristic of proximity, in the overlay, in order to efficiently detect the neighbouring groups of a newly formated group from the set of adjasent groups of its closest neighbour. Additionaly, during \emph{group joining} the coresponding protocol denotes the exchange of important information for group maintenance. This can be furtherly improved by \emph{information sharing} between nodes of the same group, functionality handled by a dedicated flood-like protocol\footnote{Since nodes that belong to the same group are physically close this can be achieved at a minimum price.}. Moreover, another set of distributed protocols handle the \emph{information update}. The information that needs update, in the proposed architecture, is
%\begin{inparaenum}[\itshape i\upshape)]
%  \item the host cache, when a new node joins, and
%  \item the neighbours of groups, when a close-by group is generated.
%\end{inparaenum}
%Finally, in case of \emph{host failure} or \emph{host departure} the system is able to maintain its stability since there are defined operations for periodical host cache update and group leader selection if one leaves or dies.
%
%\subsection{Distributed Domain Name Order}
%{\sethlcolor{yellow}\hl{HA:  In paper/introduction nice advantages of unstructured networks:
%
%Unstructured P2P networks o?er a number of
%important advantages: (i) An unstructured network
%imposes very small demands on individual nodes,
%and more speci?cally it allows nodes to join or leave
%the network without signi?cantly a?ecting the sys-
%tem performance. (ii) Unstructured networks are
%appropriate for content-based retrieval (e.g., key-
%word searches) as opposed to object identi?er loca-
%tion of structured overlays. (iii) Finally unstructured
%networks can easily accommodate nodes of vary-
%ing power. Consequently, they scale to very large
%sizes and they o?er more robust performance in
%the presence of node failures and connection
%unreliability.
%}}
%
%\paragraph{}
%\cite{zeinalipour-yazti_ddno_2005} proposes the \emph{Distributed Domain Name Order (DDNO)} technique which makes unstructured overlay networks, topologically aware. It results in a flat overlay topology but with some changes it can be utilized in \emph{superpeer} environments. In DDNO, a node of degree $d$, tries to connect to $\frac{d}{2}$ nodes that belong to the same domain (\emph{sibling} connections) and another $\frac{d}{2}$ of random nodes (\emph{random} connections). Connecting to \emph{sibling} nodes ensures the reduction of long distance message travelling and thus resulting to better performance, while \emph{random} nodes keep the structure connected.
%
%\paragraph{Discovering \emph{random} neighbours}
%Initially, when a new coming node wants to join a network $N$, uses a \emph{hostcache} mechanism to provide it with a list of nodes with which to establish its $\frac{d}{2}$ \emph{random} connections.
%
%\paragraph{Discovering \emph{sibling} neighbours}
%In order to discover the rest $\frac{d}{2}$, \emph{sibling}, nodes, the \emph{lookupDN} procedure is initiated. In this phase, a special $l$ walker message is multicasted, by the newcomer, in search of a node that knows\footnote{A special structure is used called ZoneCache that contains information about which nodes are reachable in an $r$-hop radius.} how to guide $n$ to a \emph{sibling} node.
%
%\paragraph{}
%Figure~\ref{figure:ddno_lookupdn} shows an $l$ message emitted by node $n$ which travels along the path [$a$, $b$, $c$, $e$, $b$, $d$]. At node $d$, $l$ finds information to make a decision on which neighbour to follow next\footnote{Contrary to the random choices, made at the previous hops.}. Ultimately $l$ arrives at $m$, which is a \emph{sibling} node to $n$. $m$ then issues a broadcast message to all its own \emph{siblings}. Each of the receiving nodes, $m$ included, will respond to $n$ with a designated $l^{'}$ message if it is willing to accept new connections and out of these answers, node $n$, will attempt to establish its, remaining, $\frac{d}{2}$ \emph{sibling} connections.
%
%\paragraph{}
%Unfortunately, use of DDNO results in networks which have a uniform distribution of node degrees, losing the valuable properties of Power Law and Small World networks, as mentioned previously in this survey.
%
%\begin{figure}
%\centering
%  \includegraphics[scale=0.375]{img/ddno_lookupdn.jpeg}
%\caption{Domain name look-up in a \emph{DDNO} topology}
%\label{figure:ddno_lookupdn}
%\end{figure}
%
%%\subsection{Critical Topology-Aware Grouping}
%%
%%\paragraph{}
%%\emph{Critical Topology-Aware Grouping (CTAG)} presented in \cite{zhao_ctag_2006}, is a grouping algorithm that tries to exploit low-cost and low-delay communication of physically close-by peers. The grouping strategy is based on the \emph{IANA}\footnote{Internet Assigned Numbers Authority} and the respective \emph{Regionanal Internet Registry (RIR)}'s IP assignment strategies, according to which nodes within the same organization are always addressed from the same block. The paper proposes the \emph{Adjacency Measurement (AM)} technique which uses the longest matching IP segment criterion to calculate node proximity.
%%
%%Observations made in various studies (e.g. \cite{matei_mapgnutella_2002}) concerning the distribution of nodes among \emph{Internet Service Providers (ISPs)} and \emph{Autonomous Systems (ASs)} have shown that only $2$ to $5$ percent of Gnutella connections link peers within a single AS while more than $40$ percent of all Gnutella peers are located within the top 10 ASs. Similarly, measurements in \cite{zeinalipour-yazti_gnudc_2002} used a $244,000$ IPs test-bed and results have shown that $45$ percent of the nodes belonged to only $10$ large ISPs and $58$ percent belong to only $20$. Such results mean that most overlay generated traffic crosses AS borders increasing topology mismatch cost.
%%
%%\emph{CTAG} focuses on both the construction of the overlay as well as dynamically revising it during node interaction, phases called \emph{bootstrapping grouping} and \emph{dynamic revision}, respectively in the paper's context.
%%
%%\paragraph{Bootstrapping grouping}
%%The \emph{Gnutella Web Caching} mechanism has been modified in order for a new coming node to choose the closest \emph{GWC} in order to retrieve the node list for bootstrapping.
%%
%%\paragraph{Dynamic revision}
%%Similarly to the bootstrapping phase, \emph{AM} metric is used to store hosts' addresses. read from \emph{X-Try} headers during handshake or from \emph{QueryHit} messages. Additionally, when a node reaches the max neighbour connections, node disconnects the neighbours with the lowest \emph{AM}.
%
%\subsection{Peer-exchange Routing Optimization Protocols}
%
%\paragraph{}
%Work presented in \cite{qiu_prop_2007} introduces \emph{peer-exchange}\footnote{A series of exchanges of neighbours between two peers. One exchange can be viewed as a pair of cut-add operations.} as a basic operation to adaptively adjust the connections of the overlay network, and efficiently reduce the average logical link latency of the whole system. There are two points that differentiate this scheme from others:
%\begin{inparaenum}[\itshape i\upshape)]
%  \item it utilizes the collaboration of two peers to optimize their neighbourhood environment, than simply letting each node to ``selfishly'' choose its own strategy, and
%  \item can be deployed effortlessly on both unstructured and structured peer-to-peer systems.
%\end{inparaenum}
%
%\emph{PROP-G}\footnote{\emph{G} stands for \emph{generic}.} is the simple and direct variant of \emph{peer-exchange} in which two neighbours swap all their neighbours. As depicted in figure~\ref{figure:prop-g} the mechanism can be viewed as, nodes, exchanging overlay ``positions''. Thus, the overall overlay topology is not affected by the this operation.
%
%\begin{figure}
%\centering
%  \includegraphics[scale=0.4]{img/prop-g.jpeg}
%\caption{PROP-G in which a neighbours are exchanged.}
%\label{figure:prop-g}
%\end{figure}
%
%The alternative protocol proposed, called \emph{PROP-O}\footnote{\emph{O} stands for optimized}, ensures that the degree of each participating node remains the same by selectively choose the same number of neighbours for the exchange. With this mechanism, 2 nodes exchange the same number of neighbours in order not to change the degree
%
%\begin{figure}
%\centering
%  \includegraphics[scale=0.4]{img/prop-o.jpeg}
%\caption{PROP-O exchanging \emph{m} neighbours.}
%\label{figure:prop-o}
%\end{figure}
%
%\subsection{T2MC}
%
%\paragraph{}
%\emph{T2MC}\cite{shi_t2mc_2008} exploits some properties of the Internet paradigm and clusters nodes belonging to the same ISP without any centralised control or predefined system parameterization. The algorithm considers the dynamic nature of peers and exploits the stable nature of routers in order to build a topology location relationship among end peers.
%
%\paragraph{}
%Some special routers split the physical network into autonomous system domains. Using a Traceroute mechanism T2MC searches for latency leaps among the path to a host in order to form ``near'' and ``remote'' router clusters. This is achieved by a 2-Means Classification, defined by the following steps:
%\begin{enumerate}
% \item the peer chooses the minimum and maximum Latency results from the Traceroute for initializing the cntroinds of two sets ``first'' and ``second''.
% \item The peer calculates, for all hops allong its tracerouted path, the absolute distance to the centroids of both sets and assigns the routers to that centroid with which it has the smaller absolute distance.
% \item The peer calculates the latncy mean and variance value of two sets
% \item If the variance is larger than a predefined threshold then the algorithm takes a loop from step 2 picking the two latency mean values as new centroids of sets ``first'' and ``second''.
%\end{enumerate}
%Ultimately peer will end up with two sets having the minimum intra-set variance. Finally the peer chooses the router from ``second'' set with the minimum hops attribute and sets it as a threshold. The selected router and all others whose hops attribute is larger than the threshold are classified as ``remote'' router cluster. The remaining are classified as ``near''. From the ``near'' class, the peer chooses the one with maximum hops attribute as its edge router, and registers it along with the all the ``near'' cluster into the DHT of the p2p overlay. As new peers join the network, those that share the same edge router or any of the members of the ``near'' router clustersm ther would gather to form a ``close'' peer cluster. As Edge routers can provide more valuable information than other members of the ``near'' set, T2MC was designed to prioritize interaction of peers and edge gateways.
%
%% TODO: figure t2mc.jpeg (a) stars represent edge routers (b) hop leaps
%
%\paragraph{}
%The use of Traceroute as a tool for implementing the distance measuring infrastructure raise concearns about its efficiency and scalability. Being, primarilly, a network diagnostic utility, it is concearned too heavy weighted and intrucive for use in a larger scale\cite{ratnasamy_binning_2002}. Additionally, disabling ICMP is a common administrative policy for edge sites to enforce security, while dumping BGP routing tables\cite{krishnamurthy_bgpclust_2000} is not directly available to the application layer.
%
%\subsection{Unnamed Unstructured!!}
%% TODO: to be reviewed
%
%\paragraph{}
%In \cite{hsiao_redblue_2009}, Hsiao et al, claim to construct topology-aware unstructured overlays that \emph{guarantee} performance qualities in terms of
%\begin{inparaenum}[\itshape i\upshape)]
%  \item the expected communication latency among any two overlay peers regardless of the network size, and
%  \item the broadcasting scope of each participating peer.
%\end{inparaenum}
%
%The algorithm constructs an undirected graph $G = \left( V, E \right)$ comprised by two subgraphs. The first, namely $G^{\left( red \right)} = \left( V^{\left( red \right)}, E^{\left( red \right)} \right)$ in the paper's context, includes all vertices of $G$ and ensures the connectivity of the graph by securing at least one path between any two nodes. In contrast, $G^{\left( blue \right)} = \left( V^{\left( blue \right)}, E^{\left( blue \right)} \right)$, contains those vertices of $G$ that have free edges to link to other nodes and because these are fully utilized, the following also stands $E = E^{\left( red \right)} \cup E^{\left( blue \right)}$.
%
%A joining peer $u$, partitions its neighbours into two subsets, the $B_u^{\left( red \right)}$ and $B_u^{\left( blue \right)}$. In order to populate the $B_u^{\left( red \right)}$ subset, peer $u$ samples peers uniformly and at random. Then, each of these selected peers discovers a routing path starting from itself towards the node with the smallest (or the largest) ID in the system.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% STRUCTURED
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\subsection{Landmark Binning}
%The scheme proposed in \cite{ratnasamy_binning_2002} is based on the process of \emph{binning} close by nodes (in terms of network latency) into the same cluster. The authors set the following objectives for designing their proposed algorithm:
%\begin{enumerate}
% \item simplicity (minimal support from any measuring infrastructure),
% \item scalablility (no global knowledge of the network),
% \item complete distribution (nodes need no communication-cooperation).
%\end{enumerate}
%
%\paragraph{Binning scheme}
%The implementation requires a set of well-known \emph{landmark} machines spread accross the Internet. Every newly arriving node measures its distance from these landmarks and unilaterally decides to join a specific bin based on these results. In more detail the node measures its round-trip time to each of the landmarks and orders these measurements in a decreasing order. The ordering represents a ``bin'', in the sence of close-by nodes having the same landmark ordering and hance belong to the same ``bin''. This means that a landmark system consisting of $m$ such nodes results in $m!$ potentional different bins.
%
%The algorithm can be considered scalable, as nodes need only compute distances to a small number of predefined nodes and thus without exchanging any information. A potential bottleneck could be the extra load that this ``ping''-like scheme imposes to the landmarks, especially when we need instant reaction from our topology when dealing with the dynamic nature of the p2p networks.
%
%To answer the question as to whether the algorithm actually contributes possitivelly to the construction of an enhaned overlay, the paper defines the \emph{gain ratio} as the factor by which the latency reduces when someone communicates with a random node from the same bin than with one not in the bin. This is implemented with an inter-bin to an intra-bin latency ratio.
%
%\paragraph{Structured and Unstructured Binning}
%The \emph{binning} scheme can be incorporated within either a structured or an unstructured overlay construction algorithm. The authors provide an example for both these cases.
%
%\paragraph{}
%Assuming a structured approach based on CAN\cite{ratnasamy_can_2001} and $m$ landmark nodes. The coordinate space is then partitioned into $m!$ equally sized portions, each corresonding to a single ordering of the landmarks. To do this, the first dimension is divided into $m$ areas each of which is furtherly divided (second dimension) into $m - 1$ sections and so on. Having set this $m$ dimensional space, at joining time, a node measures the delay to the set of landmarks in order to determine its associated bin and thus position itself in that portion of the coordinate space associated with its landmark ordering. Even though this scheme can 
%
%For unstructured overlays the paper assumes \emph{a set of $n$ nodes where each node picks any $k$ neighbour nodes so that the average routing latency on the resultant overlay is low (assuming shortest path routing)}. According to the proposed heuristic algorithm called \emph{BinShort-Long}, a node picks its neighbours by choosing its $\frac{k}{2}$ closest\footnote{If the node's bin is not large enough for it to pick these $\frac{k}{2}$ neighbours, it picks the required nodes from the bin that matches the most in terms of landmark ordering.} ones (named \emph{short links}), using the \emph{binning} scheme and the rest $\frac{k}{2}$ randomly (\emph{long links}). The former set produces well-connected \emph{pockets} of nearby nodes while the later preserves the connectivity of the graph, both yielding a proximity factor of $\alpha = 0.5$ in an attempt to preserve the benefitial properties of unstructured topologies\cite{merugu_str2unstr_2003}.
%
%\paragraph{}
%% TODO double check validity
%One disadvantage of this landmark scheme is related to the additional burden imposed to the landmark sites. The authors claim though that the algorithm requires so little work by the landmarks (maybe just echo to ping messages) that could in effect, act as ``unsuspecting participants''. Even if this is the case, the fact that it is not fully distributed, renders the protocol's scalability directly volnerable to any system size increase as well as usuitable for highly dynamic networks such as ad-hoc networks. Moreover, fixed points in a network are inherently more exposed to malicius attacks. The most significant downside of the algorithm though is that it can lead to an extremely uneven overlay ID distribution causing load inbalances and hot spots. Lastly, the scheme is coarse grained when it comes to distinguishing relatively close nodes\footnote{In the worst case, all nodes could ve clustered into a single bin.}.
%
%\subsection{Global Soft-State}
%{\sethlcolor{yellow}\hl{HA: in paper introduction, discussion about
%disadvantages of top-aware CAN:
%
%Techniques to exploit topology information in overlay
%routing include geographic layout, proximity routing and
%proximity neighbor selection [3]. With geographic layout
%such as topology-aware CAN [12], the overlay structure is
%constrained by underlying network topology. This tech-
%nique, unfortunately, can create uneven distribution of
%nodes in the overlay, increasing the chances of overloading
%nodes and rendering the maintenance cost formidable. Our
%study shows that, for a typical 10,000-node topology-aware
%CAN, 5% nodes occupy 85-98% of the entire Cartesian
%space, and some nodes have to maintain 450-1500 neigh-
%bors. In Proximity routing, physical topology is not consid-
%ered when constructing the overlay.
%
%[snip]
%Studies [14] have shown that triangle inequal-
%ity may not hold in Internet topology. In fact, study from
%Pastry has shown that the proximity approximation is much
%worse when using the Mercator topology that is based on
%the real measurements of the Internet [3].
%
%RELATED WORK:
%
%Miguel Castro et al [3] divide techniques used to
%exploit network proximity into three categories: geographic
%layout, proximity routing and proximity neighbor selection.
%Proximity neighbor selection is superior in terms of load
%balancing and proximity approximation. The existing algo-
%
%}}
%
%\paragraph{}
%The approach argued by Xu et.al in \cite{xu_globstate_2003} is to build a global map to help choose shorter routing paths, combining the landmark binning method and small scale distance probes to reveal the proximity properties of the underlying network to the overlay. This global view of the state is made available to all nodes in order to help them find the best way to route their messages. The authors focus on two aspects in order to accomplish their goal.
%
%\paragraph{}
%Initially, is the \emph{generation} of proximity information and then its \emph{effective exploitation}. For the first, a hybrid approach is proposed, which uses landmark clustering as a preprossesing step in order to select a number of potential nearest neighbour candidates and then refine the selection by incorporating an RTT scheme to ultimately choose the closest node. For the second, the algorithm chooses a different path from the classic gossiping approaches for constructing and maintaining the overlay. It is based on landmark-clustering-based strategic placement of proximity information on the overlay enabling any node to access such information using a landmark number that reflects its physical position in the network. For various logical regions\footnote{This might be a high-order zone in the eCAN\cite{xu_ecan_2002} context or a set of nodes sharing a particular prefix in overlays such as Pastry.} maps of physical information are built and published where each node may appear in a maximum of $log\left( N \right)$ such maps.. To dynamicaly adapt to changing network conditions, a node subscribes to relevant \emph{soft states} that utilize a notification system in order to initialize any necessary neighbour re-selection.
%
%\paragraph{}
%Maintaining several host states at different layers, makes any content migration costly. Additionaly, the method does not make any continuing effort to remap the overlay structure after a node successfuly joins, in order to adapt its state to any occurance of condition change. Although this approach greatly reduces the routin latency to far nodes, it is unable to dynamicaly identify nodes that are close to routers and gatways in order to construct the secondary overlay. Nevertheless, static recognition of such nodes is currently done based on BGP reports and prechosen landmarks, sucrificing the self-organising attribute of traditional DHTs.
%
%\subsection{Self-Adaptive Topology Matching}
%
%\paragraph{}
%Ren et.all's goal while designing the \emph{SAT-Match} algorithm \cite{ren_satmatch_2004} was to create a protocol that would be decentralised and scalable, fine-grained in detecting changes and adaptively reform in response to the dynamic nature of a classic p2p environment and last, but certainly not least, of low incuring cost. The method is in a nutshell, a two-phased iterative process that focuses on local optimizations in order to collectively achieve a global full overlay optimization.  The iteration is finised when the node detects that it is physicaly close enough to its neigbours so that no additional optimization is needed. The paper defines \emph{stretch}, $S = \frac{\bar{L_l}}{\bar{L_p}}$, as a way quantify the \emph{topology match degree} of the constructed overlay, where $\bar{L_l}$ is the average logical link latency while the $\bar{L_p}$ is the average physical link latency.
%
%\paragraph{Probing phase}
%\emph{SAT-Match} uses a small TTL value for the probing messages in order to reduce redunduncy\cite{jiang_lightflood_2008}. This process begins as soon as the node joins the network using a DHT mechanism. Each probing message contains information about the source and a small TTL value. The recipient of such a message returns information about itself to the source and forwards the probing message to its neighbours if the TTL is non-zero. The nodes been covered, are refered to as $TTL-k$ neighbourhood of the source node\footnote{Especialy, $TTL-1$ neighbourhood referes to the source's direct neighbours}. The list of responses is used to measure RTT to these nodes and sort the list ascending RTT order.
%
%\paragraph{Jump phase}
%Blindly selecting the peer with the smallest RTT as neighbour is, generally, not the right desision to make in order to achieve global \emph{stretch} reduction, This is because, in a structured scheme, when a node jumps to connect to a physically close node, it may need to connect to other distant nodes to maintain the structure's integrity, thus creating an overall increase in the overlay's \emph{stretch}. The two nodes with the smallest RTT is then used in order to select one zone to jump in this phase. The algorithm is as follows: The source node $S$ calculates the stretch change of its $TTL-1$ neighbourhood and that of the $TTL-1$ neighbourhood of the first of the previously selected peers. These calculations are made as if the jump has been made. If the stretch reduction is over a predefined threshold the jump is performed, otherwise the second selected candidate is picked and the same computations are performed. If again, the threshold is not met, then no jump is ultimately done. In case of a jump, this is performed as a combination of \emph{leave} and \emph{join} operations, in the CAN context.
%
%\paragraph{}
%Moreover the algorithm takes several issues into consiteration in order to furtherly improve the resulted overlay. For example when multiple nodes try to jump simultaneously into a regeon, then the logical link brakes from one attempt may result in inacurate computation of the gain factor, for an other. This situation is identified as \emph{contention} and the nodes use an exponatial back-off algorithm to avoid it. An other problem is the uneseccary traffic incuerred by the probing phase in a region that after several jumps has settled to a stable state. In these cases it is more likely for jump attempts to be proven worthless. The algorithm doubles the probing period of such nodes, every time a jump is not taken.
%
%\paragraph{}
%The authors claim that this continuously adaptive mechanism achieves global topology matching optimization in a sufficiently large scope. This also secures the fast adaptation to frequent network changes. It also considered lightweight and can easily be embedded into current p2p systems, as well as effectively combined with other techniques, such as landmark binning.
%
%\subsection{Delay Aware P2P System}
%A new \emph{Delay Aware P2P System (DAPS)} is introduced in \cite{zhang_daps_2005}. Its main goal is to reduce the time $L$ of a look-up request by dividing the routing tables of peers into several sectors in increasing delay. The source node emitting the query message defines a delay boundary or the pruning factor $L_t$ in the paper's context. Request messages will be forwarded only to nodes whose delay less than or equal to $L_t$. With the clustered routing tables and the loose organisation the overlay network of \emph{DAPS} is between structured and unstructured.
%
%\subsection{Mithos}
%
%\paragraph{}
%Waldvogel and Rinaldi proposed a protocol\cite{waldvogel_mythos_2003} which incorporates a directed incremental probing to find near optimal node placement.
%
%\paragraph{Neighbour detection}
%During bootstraping, the new comming node needs to know how to contact at least one of the existing members. A subset of these nodes will be used as the first set of candidate neighbours. Then, iteratively, each of the peers in the cadidate set is asked for their neighbours in order for each of the later to be probed for their network distance from the new coming one. The closest node is then used as the new candidate neighbour and the process is repeated until no further improvement is detected. Due to the fact that the algorithm may ultimately reach a local minima instead of a global one, \emph{Mithos}' approach is to first probe all neighbours within a two hop distance from the current minimum before concluding the process.
%
%\paragraph{ID assignment}
%After finding its first neighbour, the newcoming peer must be assigned its ID, a very crusial selection in order not to create many local minima that will prevent efficient future neighbourhood location. \emph{Mithos} uses information gathered during the previous step, in order to compute the ID, which inccurs no further communication costs to the algorithm. This information includes
%\begin{inparaenum}[\itshape i\upshape)]
%  \item the two closest nodes and their neigbours, and 
%  \item their coresponding distances.
%\end{inparaenum}
%Using the above, it then assigns coordinates to the new-coming node, so that Euclidian distances between the node and all known hosts predict the network latency between them\cite{cox_vivaldi_2004}. These synthetic coordinates as explicitly used as the node's ID and as soon as it has been established, distances can be computed in the ID space, no longer requiring physical measurements.
%
%\paragraph{Link establishement}
%The last step of the algorithm is the interconnection among neighbours. \emph{Mithos} uses a \emph{quadrant}-based mechanism according to which each node establishes a link to the closest neighbour in each quadrant. During forwarding, the next hop is performed towards a neighbour in the same quadrant as the final destination\footnote{An implementation of this could be backed by a $d$-bit vector indexing (where $d$ is the number of dimensions) into the routing table. Then the next hop is identified by computing the difference between the current node's vector and that of the destination's.}. The problem is that after the ID assignment process, the new-coming node may not know of other neighbours in all quadrant and if even if it does, this cannot ensure that they are the nearest available. Thus, the node first identifies neighbours in all quadrants using a mechanism based on ideas similar to a perimeter walk\footnote{Used in Greedy Perimeter Stateless Routing (GPSR) protocol.} and then using parallel path processing improves the results by taking into account further geometric properties of node relationships.
%
%\paragraph{}
%% TODO review this part
%In order to avoid local minima during neighbour detection, extensive probing must be undertaken. In simulation, unfortunately, only very small-sized overlay topologies (of 200 to 1000 nodes) have been used and thus no safe conclusions can be made as for the behaviour of an extensively large, real-world p2p deployment of the scheme. 
%
%\subsection{DHT-PNS}
%
%\paragraph{}
%The work in \cite{hancong_pnsbased_2006} propose a proximity neighbour selection scheme on top of the Chord DHT. Using the Vivaldi protocol\cite{cox_vivaldi_2004} each node is assigned synthetic $2$-dimension coordinates that can be used to derive network latency using Euclidian distances in the id space and without the need of explicit probing. Then the algorithm performs the two steps described in the following paragraphs.
%
%\paragraph{Space mapping}
%The space is partitioned using a \emph{concentric circle clustering scheme} where succesive cycles of radiuses $\rho$, $2\rho$, $3\rho$ and so on, are constructed. Then the formed annuluses are divided into $2\chi-1$ \emph{sectors}, where $\chi$ denots the sequence number of the annulus starting from $\chi = 1$ for the centre cycle. It is proved in the paper, that this way each sector occupies the same area\footnote{The same area as does the center cycle.} and assuming uniform node distribution, this characteristic, favours a more load balanced clustering operation. Every sector in this $2-d$ coordinate space is mapped to a unique \emph{region} in the DHT space forming a multi-layer node identifier space. Thus, any nodes that belong to the same sector, are mapped to the same region as well, preserving their proximity relationship unveiled by the use of the Vivaldi protocol.
%
%\paragraph{Routing table optimization}
%The system described in the previous section allows any node $\alpha$ to obtain its DHT's $key_{\alpha}$ and its region's $key_r$, in a fully distributed manner, just by applying a consistent hash function. Using a $Get\left( key_r \right)$ RPC call in Chord, the node can obtain the region's master node, called \emph{Cluster Node (CN)} which is responsible for clustering the nodes belonging in the same sector or region with that of $\alpha$. On the other hand, a $Put\left( key_r, info\right)$ RPC call, registers $\alpha$ to its corresponding region and publishes its information, through the special peer CN. $\alpha$ can, additionaly, ask CN for other nodes that have previously joined the region in order to add them into its neighbour set for future routing table optimization. Even in the case when no neighbour is detected in the current region, the search is expanded to adjasent regions and towards an upper layer identifier space until one is found or the first layer reached.
%
%\subsection{Quasi-Chord}
%
%\paragraph{}
%The approach that is proposed by Sun and Zhang in \cite{sun_quasi_2008} confronts the topology mismatch problem, in the Chord DHT context, that is created by the fact that no consiteration for the underlying physical network topology is taken into account during the construction of the identifier cycle. To construct a \emph{Quasi-Chord} network three steps are needed. First, each host acquires its coordinates in the geometric space utilizing the \emph{global network position (GNP)} protocol \cite{ng_gnp_2001}. Second, using the \emph{Cantor space filling curve} the $2$-dimensional space is converted to a $1$-dimensional one, used in the last, third, step to build the Quasi-Chord circle according to the host's Cantor value. In the following paragraphs, these steps are furtherly discussed.
%
%\paragraph{GNP coordinates}
%First of all the host must be positioned in the geometric space. The algoritm models the P2P network with a well defined distance function in such a way, it can predict, with high accuracy, the distance between any two points in the space by just evaluating the output of the distance function on the coordinates of these points. This is accomplished with the \emph{global network position (GNP)} protocol which\begin{inparaenum}[\itshape i\upshape)]
%  \item creates a reference set of $N$ landmark nodes so as to minimize the error of ICMP measured distances and coordinate computed ones between them, and then
%  \item each host is able to measure its round-trip times to the $N$ Landmarks in order to compute its own coordinates.
%\end{inparaenum}
%
%\paragraph{Cantor SPF}
%After the $2$-d coordinates are set, in the next stage of the algorithn, each participating peer is assigned a Cantor value, according to the application of a \emph{space filling curve} on the coordinate space. (TODO: add figure). This results in the conversion of the $2$-dimensional space to a $1$-dimensional Cantor space that can more easily be more mapped to the Chord hierarchy.
%
%\paragraph{Quasi-Chord construction}
%As can intuitively be infered by observing the Cantor chart, close-by nodes in the physical layer, are more likely to have similar Cantor values. This attribute is exploited in order to construct a topology-aware identifier space for the Chord DHT. The cycle is constructed by sorting nodes in ascending order. Each host maintains 2 finger tables, one for clockwise and one for counter-clockwise stepping. This helps with the connectivity of the network because its not allowed to connect the first node with the last one since this will incur heavy traffic to the later\footnote{After all their Cantor values denote that they are actually the furthest of each other.}.
%
%\paragraph{}
%% TODO: doublecheck!
%The disadvantage of the algorithm is that the coordinate assignment in the first stage is backed by a not fully distributed landmark-based algorithm. Moreover the Quasi-Chord model build-up is making an indirect assumption of a maximum number of allowable hosts since it is constructed. Last but not least the doubling of the required routing information which needs to be created and maintained is an additional negative point to the efficiency and the scallability of the algorithm.
%
%\subsection{LAPTOP}
%{\sethlcolor{yellow}\hl{HA - two important issues for decentralized structured P2P:
%P2P applications. However, the design of a decentralized but structured P2P network has to
%overcome two critical issues. The first issue is the long routing latency. Several proximity
%schemes  have been proposed to avert long routing latency in current structured P2P
%networks, but they require a high-complexity procedure to periodically maintain the routing
%table (e.g. Pastry system) or they need pre-chosen landmarks to construct the overlay.
%However, the P2P system is by its very nature unstable since nodes join and leave frequently.
%For instance, the study of Gnutella shows around approximately 1200 membership changes
%per minute in a 100 000 nodes P2P system. Another proximity scheme needs some pre-chosen
%landmarks or a complete BGP routing table support. As a result, they both increase the
%difficulty of the P2P system deployment.
%The second issue is system maintenance overhead. The existing structured P2P networks allow
%nodes to keep some nearby nodes in their routing tables in order to achieve efficient routing. The
%}}
%
%\paragraph{}
%Laptop \cite{wu_laptop_2007}, introduced by Wu et al., organises the overlay into a tree based hierarchy with main focus on child-to-father relationships in order to reduce hops during message routing as well as minimize maintenance overhead. Additionally, a caching scheme is also incorporated so as to furtherly reduce routing table update cost. The authors argue that Laptop overlay network is bounded by $O\left( log_d N \right)$ hops in a balanced overlay tree, where $N$ is the number of nodes, and $d$ is the maximum degree of each node. It utilizes a geographical layout approach  and constructs a geographical layout in a self-organizing and efficient fashion, by estimating the round trip time (RTT) to a small number of nodes in the overlay network in order to make them roughly aware of their physical distances among them.
%
%\paragraph{}
%The protocol is based on four definitions.
%\begin{enumerate}
% \item The amplitude of all possible measured RTTs is devided into intervals. Each node measures its distance to its parent and is assigned a label $L_i$ where $i$ denotes the configurable RTT interval in which the measured distance falls into. A special kind of node, the root, is initially assigned the $L_1$ label and maintains (as all $L_1$ nodes do) a list of other $L_1$ nodes in the overlay.
% \item Any node can have children with level lower than theirs, except an $L_{max}$ node which can only have $L_{max}$ level children and only in the case when its parent has reached its maximum degree.
% \item Each node is assigned an address in a dotted format (e.g 1.3.4). Each octet ranges from $1$ to $d$, where $d$ is the maximum degree of the nodes. The assignment process is done by appending a unique octet to the address of its parent. Root node is assigned address 1.
% \item For any descendant node $Y$ of a node $X$, the measured distance among each other, must always be less than the lower bound of the RTT interval denoted by $X$'s label.
%\end{enumerate}
%
%\paragraph{}
%The routing scheme is similar to the IP's longest-prefix matching scheme. At each forwarding hop, any message travels up the tree until the first common uncestor of source and destination node is reached and then starts descending to arrive to its target. During tree traversals, special entries in the routing tables, called \emph{routing cache}, are maintained in order to increase routing efficiency and achieve finer load balance. Caching enables a node to forward a message to a better longest-prefix match than that of its direct ancestor making a large, quicker and more cost effective step through the overlay and toward the destination. To improve scalability, the number of children nodes and the size of the routing cache are limited.
%
%In terms of overlay maintenance, Laptop incorporates a simple \emph{heartbeat}-based technique where only the parent node is responsible for monitoring its children.
%
%At join process the newcomming node is assigned its level label as well as its address by its parent node. Additionaly it initializes its routing table (with normal and caching entries) as it traverses the overlay in search for its parents node.
%The newcommer $N$ locates the root node and the later responds with a list of $L_1$ nodes. $N$ then probes each of the $L_1$ nodes in search for the closest one. If the measured RTT to the closest $L_1$ falls into the first interval then the newcommer becomes a $L_1$ node as well. Otherwise node $N$ sets the closest $L_1$ node as its potential parent node. This potential parent becomes the actual parent if it does not have any other children. If it has, $N$ gets a list of these $L_i$ nodes and by measuring the RTT to each of them tries to spot a new potential parent in order to repeat the above process.
%
%\paragraph{}
%During a gracefull departure, the refered node checkes for children in the overlay. If it does not have any, it simply notifies its parent and leaves. If it has, it selects the child node with the lowest RTT to it in order to take its place so that the locality property is preserved.
%
%In case of an arbitrary failure, the children of the failed node are detecting their parent's absent when the later stops acknowledging their heartbeat messages. They start emmiting special messages to their grand parent node\footnote{The address of the grandparent node is stored during join process}. In case of no response from the grandparent node, children invoke the joining procedure to detect their new parent node.The parent of the failed node, aggregates the notifications from the above children nodes for a period of time, and then chooses the one with the lowest level label as the takeover node. Potential ties are broken by favouring the lower RTT. The parent of the failed node, finally informs accordingly all the children about the change in the hierarchy.
%
%\subsection{IP-based Clustering (IPBC)}
%
%\paragraph{}
%\emph{IP-based clustering} \cite{karwaczynski_ipbc_2007} is a proximity neighbour selection technique that is based on a longest IP prefix matching scheme in order to quantify the proximity among peer nodes. The relation of the IP address and the physical location of a node in the topology of the Internet, is intuitevely realized by operation the network layer of the Internet stack. For example, nodes in the same local subnetwork share the same 3 bytes of their address. Reports in \cite{freedman_iploc_2005} state that almost $97\%$ of prefixes longer than 3 bytes correspond to address assignments at a single geographic location. Moreover, observing the \emph{IANA IPv4 Address Space Registry} we can infer that, since blocks of consecutive octet prefixes are assigned to the same Regional Internet Registries (RIRs), nodes inside will be physicaly close to each other.
%
%\paragraph{}
%The implementation in this paper is based on the above observation in order to create proximity information. This information is stored in a decentralised manner, within the overlay itself, just like any other resource would have been. In order to be published in the overlay, each node is first assigned a unique identifier and subsequently generates a key by hashing a fixed-length prefix of its IP. Authors argue on the prefix length to be used, as there is a thin balance between reducing the propability of finding closer nodes by adopting a long prefix and overload nodes that are responsible for information published by many peers when chosing to hash a short one. Their verdict is for the use of a 16-bit wide prefix for real world systems deployed in an Internet scale. In either case, ID and IP is then stored in the DHT using this generated key. This way, any node, at any time (either at join time or during peer lifetime for topology adaptation) can acquire information about close-by nodes just by quering the DHT for a specific key. Moreover, the algorithm takes care of the freshness of the proximity information in two ways. First, the advertising nodes themselves periodically update their advertisements or when they voluntarily leave the overlay, they explicitly remove their data. On the other hand, in case of a failure each publication is assigned an expiration time, and thus ultimately removed by the DHT maintenance mechanisms.
%
%\subsection{CHOord considering Proximity on IPv6 (CHOP6)}
%
%\paragraph{}
%\cite{morimoto_chop6_2007} roughly estimates the proximity among nodes by exploiting the IPv6 address format as well as RTT information if this is available.The first is achieved by introducing a 64-bit ID scheme in which the least significant bit part\footnote{The exact range is a predefined system variable} is the IPv6 global routing prefix and thus enabling a longest prefix match scheme. The observation in which the protocol is based roots from the IPv6 address block assignment. Specicaly, block of /16 or /23 in size are assigned to Regional Internet Registries (RIRs) by Internet Assigned Numbers Authority (IANA). An RIR divides the signed address blockes into smaller address blocks which are furtherly assigned to Natinal Internet Registries (NIRs). That way, it is possible to estimate a node's geographical location by simply examining the upper 32-bits of its IPv6 address. Moreover, similar to Chord, CHOP6 uses a finger table, whose entries hold more than one candidate nodes. There are three cases in which a node chooses the next hop according to information it posseses.
%\begin{itemize}
% \item When no information about RTT is available candidate next hops, the sender just selects a node in the finger table entry which shares the longest prefix with the destination.
% \item There is another possibility that after some communication with other nodes, the source node should know of the RTTs to some of the nodes in finger table entry. In this case, the source node chooses the one with the smallest RTT with propability $p$. If there is a node with no measured RTT then the sender can select such a node with probability $1 - p$
% \item In this last case the source node has already communicated with all candidate node in the finger table entry. Thus, the node selects the node whose RTT is the smallest with probability $p$, the one with the second smallest with probability $q$ and so on, where $0 \leq \ldots < r < q < p < 1$ and $p+q+r+\ldots \leq 1$
%\end{itemize}
%
%\subsection{Proximity in Kademlia}
%
%\paragraph{}
%In \cite{kaune_pkad_2008}, Kaune et al., studied the Kademlia DHT, in order to build a proximity aware scheme that would work in the context of iterative lookup algorithms. The protocols discussed, focus on both the overlay and underlay tiers in the sense that in the first improves the routing performance according to a cost metric that is provided by the later. This metric is tailored to the needs of a specific aplication. For example, routing could be optimised to maximise the within-ISP traffic or reduce the lookup latencies or even avoid contacting untrustworthy subnetworks.
%
%\paragraph{}
%Two algorithms are presented for the overlay optimization. One is \emph{Proximity Neighbour Selection (PNS)} and one for \emph{Proximity Route Selection (PRS)}. The first aims at keeping peers with the least contact cost in the routing table. As Kademlia constantly learns new peers (incoming requests, through iterative lookup) no special algorithm for searching more cost effective peers is necessary and simply choose the best peers seen so far. The later (PRS) aims at choosing the best next hop during routing a message. Due to the fact that the routing in Kademlia is iterative, it is the initiator of the lookup that chooses each next hop from a set of candidate nodes. The ``vanilla'' Kedemlia always chooses the closest node with respect to the XOR metric but PRS Kademlia chooses the one with the smallest underlay metric cost. Thus, as in all such approaches, there is a trade-off between the overlay and underlay distances.
%
%An underlay metric provides information about the underlay network. The paper distingushes between three kinds of possibilities to quantify it:
%\begin{itemize}
% \item Using measurements gained by previous lookups
% \item Using measurements acquired by other peers or jointly calculated with others
% \item Using a local database to look up information
%\end{itemize}
%In the paper, for locality of traffic and subsequently for reducing the lookup latency, a clustering scheme has been implemented exploiting information stored in a geographic information database (GeoIP,MaxMind) about peers. The goal of such a metric is to constrain the largest portion of communication within the limits of a peer's ISP.
%Another incorporated approach is Vivaldi\cite{cox_vivaldi_2004}. At first all coordinates are random but as peers start to communicate they calculate RTTs as well, gradually updating their own coordinates. Vivaldi creates a system where a peer knowing the coordinates of a another, can approximate communication latency without the need of additional prompting.
%
%\subsection{Cone}
%
%\paragraph{}
%Cone\cite{wang_cone_2007} uses a two-layered identifier space. The first, named Chord-layer identifier, denoted as $Id_{Chord}$, is the same as in ``vanilla'' Chord. The second is the Cone-layer identifier, $Id_{Cone}$ which is constructed by two component identifiers. The first, known as \emph{group identifier (gid)} denotes a relevant group the node belongs to while the second, namely \emph{local identifier (lid)} indicates the local identifier within the group. The group concept, which is introduced here, is a way of dividing nodes according to a common $Id_{Chord}$ prefix.
%The structure of a Cone overlay, retains the Chord's circular topology. The differce lies on the fact that, now, two rings are created. A big ring, where nodes with the same \emph{gid} are arranged at each position. Each of these positions are a smaller ring for the particular group's \emph{lid}s. The routing is achieved in both clockwise and anti-clockwise direction in the big-ring. For this reason two routing tables are maintained, namely \emph{front} and \emph{back} finger tables, respectively. Entries in these tables, display physical network proximity with the current node. Moreover, a third table called \emph{group} table maintains information about other online peers within the current node's group in a way that entries are now close in the ID space.
%
%\paragraph{}
%Routing in Cone, comprises of the inter-group algoritm and the intra-group algorithm. First the group of nodes to which the desired key lies is detected, exploiting physical proximity information (front and back finger tables). Second during the next and last hop, the message is forwarded to the exact node that contains the desired key.
%
%Cone uses Landmark+RTTs to generate proximity information and exploits this information using proximity neighbour selection.
%
%\subsection{DynaMO}
%
%\paragraph{}
%In their work in \cite{winter_dynamo_2004}, Winter et al, try to build an overlay network, called \emph{DynaMO}, that tries to consider not only the physical proximity of peers but their mobility attributes as well. DynaMO is based on a Pastry overlay network. This was an intentional choice from the authors because Pastry's built-in locality heuristics are thoroughly analised in the literature \cite{castro_proximityp2p_2002} providing good backround against which to test and compare their results.  In order to adapt to mobile, ad-hoc networks, though, special care has been put to the maintenance of an even overlay ID distribution so that hot-spots are avoided. Pastry assignes them in a randomised fashion and then tries to consider proximity through the joining process and routing table maintenance. On the other hand, DynaMO dictates a newcoming node to gather information concerning its physical neighbourhood and uses it to assign itself an appropriate overlay ID.
%
%DynaMO tries to capitalize an observation of Pastry's routing mechanism. Each node's table consists of rows equal to the number of digits of the overlay IDs and columns equal to the ID's base. As we go down the table rows, the matching prefix between the current node's ID and the row's entries increases by one. Thus, the leaf set contains the closest nodes in the ID space. Additionally as the prefix match increases by one, the result is exponentially less candidates that can fill the tables entries as the row increases. This leads to the observation \cite{antony_pastry_2001,castro_proximityp2p_2002} that from overlay routing hop to overlay routing hop the physical distance between nodes is likely to increase leading to a dominating last routing step in terms of the overall plysical routing path distance travelled during a key lookup. Since the last routing step is usually taken from the leaf set, DynaMO focuses on making this last step as physicaly close as possible. In this context, two approaches are considered namely \emph{Random Landmarking (RLM)} and \emph{Closest Neighbour Prefix Assignment (CNPA)}.
%
%\paragraph{}
%Random Landmarking uses the overlay lookup mechanism to locate nodes responsible for a fixed set of carefully chosen\footnote{In a way that the ID space is divided into equal portions} landmark keys. This means that when a node is assigned a landmark key then it plays the role of a temporary landmark for the network. Any joing node will measure its distance to that and all other landmarks and assign to itself an ID consisting of a prefix taken from the closest landmark and a remainder that could be assigned randomnly or generated by mechanism that takes into account physical neighbourhood. The legth of the prefix can be determined as $prefix_length=|log_b k|$, where $b$ is the ID base and $k$ the number of landmarks. The above scheme results in physically close nodes, forming regions of common ID prefix that are likely to be close to each other in the ID space and thus bringing a node's leaf set closer to itself. Additionally, the advantage of dynamic landmark nodes is that the failed ones can instantly be replaced by new responsibles, that share similar physical attributes to the failed and thus preserving the balance of the network.
%
%\paragraph{}
%RLM may incur more traffic especially on landmark nodes, load that in some cases is not acceptable. CPNA on the other hand takes advantage of Pastry's specification that a new coming node is bootstraped by a physicaly close node. The new comer then assumes the ID prefix of that neighbour while the rest is generated similarly to RLM. Unfortunately, less overhead comes at the expense of being more coarse-grained.
%
%\paragraph{}
%Both algorithms, are protected against the formation of physical landmark clusters or imbalanced ID distribution\footnote{More common during the initial formation of an overlay network} , by introducing the \emph{landmark gravitation range} as a theshold over which landmark keys are reassinged (for the RLM approach) or unutilized ID prefix ranges are detected and used (for the CPNA scheme) in order to balance the distribution of regions in the overlay.
%
%\subsection{PChord}
%\emph{PChord}\cite{hong_pchord_2005} is a based on the Chord DHT which adds proximity routing into its routing mechanism. The main modification over the ``vanilla'' Chord is the inclusion of a \emph{proximity list} into its routing table so that the next hop is decided based not only by considering best progress towards the key, but on the physical proximity of the candidate nodes as well. The list is, at join time, empty but as the PChord node starts to interact with other nodes it applies a heuristic mechanism to fill up the list. Entries are dynamicaly added or removed as the network state is constantly changing. Because, it is guaranteed that at each routing step there is progress towards the target node in the ID space, PChord will result in lower hop number than Chord, for each hop is larger or at least equal in key space in PChord than in Chord. Additionaly, passing through proximity links in the underlay means reduced routing cost. Moreover, if the number of entries in the proximity list is the same as the number of network partitions, PChord prevents hops from jumping back to the same network the current node belongs to.
%
%\subsection{AChord}
%\emph{AChord}\cite{dao_achord_2006} uses IPv6's anycast functionality in order to
%\begin{itemize}
% \item releave the protocol from complex joining procedures, and
% \item achieve high accuracy of network proximity giving high routing efficiency,
%\end{itemize}
%all with a simple and lightweight mechanism that requires few changes to Chord without affecting its own advantageous characteristics. Anycast delivers a message comming from the outside of an anycast group to the physicaly closest node in that group. AChord organizes all nodes participating in the overlay network into an anycast group. Any joining node comes outside of that group and, thus, is automaticaly forwarded to the physicaly nearest node in order to bootstrap, avoiding
%\begin{inparaenum}[\itshape i\upshape)]
%  \item the need to, explicitly, maintain such nodes, and
%  \item the effort of finding a way of locating the physically nearest among them.
%\end{inparaenum}
%The ID of the new comming node is computed based on the ID of the bootstrap node and bootstrap's predecessor in a way that its ID will position itself between the formentioned.
%
%After joining, the finger table is build the same way as in the Chord protocol. Moreover, additional nodes are maintained in a structure called \emph{neighbourship table} which stores information about the closest known nodes\footnote{As a first entry, at bootstrap, the node stores information about the bootstrap node.}. The routing decision is made by using both the neighbourship and the finger table.
%
%\subsection{Chord6}
%\emph{Chord6} \cite{xiong_chord6_2005} is a Chord variant that tries to exploit the hierarchical features of IPv6 in order to create a substrate that reduces interdomain trafic and does that in a cost efficient way. Chord6 bareley modifies the original Chord protocol, only in the part of identifier definition. This, renders the aproach easily portable to other DHTs such as CAN, Pastry and Tapestry. In Chord6 the identifier contains two pares: the higher bits are obtained by hashing the  node's IPv6 address prefix of specific length, while the remaining lower bitss are the hash value of the rest of that IPv6 address. This way nodes in a domain will be mapped onto a contihuous key space on the overlay network. This guarantees that messages are routed in a way that they do not hop in and out of domains many time, thus minimizing overall routing cost. (TODO: Recheck validity) However, even though nodes in the same domain would have close identifiers, nodes in two close domains may have very different ones.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TODO: Should be rewritten
\section{Conclusions}
Peer-to-peer architecture has been in the center of research attention in the
last decade. Especially the decentralized unstructured genre exploits the
advantages of loose coupling and self organization of computing nodes to form
application-layer networks on top of the physical, best-effort infrastructure of
the Internet that exhibit interesting properties. Scalability problems arouse
quickly, though, because of the inefficient construction of this overlay that
was built with no concern for the underlying physical network that causes a
great deal of redundant traffic. The problem was identified by the research
community as the topology mismatch problem between the overlay and the
corresponding underlying physical network and a great deal of effort has been
set towards alleviating it. Some fruits of this effort have been gathered and
presented in this survey. Measurement of link cost through latency or RTT and
deletion of inefficient established connections are the key concepts of almost
all approaches. Others, manage to address the problem through hierarchical peer
clustering (e.g. best IP matching). Additionally, some protocols focus on
specific problems that furtherly arise, such as overlay partitioning, search
scope reduction or convergence speed, to name just a few. Another desire is to
form a protocol that could be applied to both decentralized unstructured and
decentralized structured networks. Unfortunately no approach has equally
addressed these problems in order to form a robust solution. So the field seems
to be, still, fertile for any, new, clever idea.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{acmtrans}
%\bibliographystyle{abbrv}
\bibliography{bib/p2p-overlays}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2014-01-21 mard-survey.tex
2012.03.23 (Vassilis) Adding an abstract. Introduction rewritten to use first person active voice. Introduction restructured for better coherence. Better integration of the ACM LaTeX class. Bibliography not rendered is a know issue.

@@
 %\documentclass[a4paper,10pt]{article}
-\documentclass[acmcsur]{acmtrans2m}
+\documentclass[acmcsur,acmnow]{acmtrans2m}
+%\documentclass[acmtocl,acmnow]{acmtrans2m}
+
+%\acmVolume{V}
+%\acmNumber{N}
+%\acmYear{YY}
+%\acmMonth{Month}
+
+% packages
 \usepackage[utf-8]{inputenc}
-
 \usepackage{paralist}
-
+\usepackage{subfigure}
 \usepackage{graphicx}
-
 \usepackage{colortbl}
@@
 \usepackage{pdflscape}
+
+% definitions
 \newtheorem{theorem}{Theorem}[section]
-\newtheorem{definition}[theorem]{Definition}
+\newtheorem{conjecture}[theorem]{Conjecture}
+\newtheorem{corollary}[theorem]{Corollary}
+\newtheorem{proposition}[theorem]{Proposition}
+\newtheorem{lemma}[theorem]{Lemma}
+\newdef{definition}[theorem]{Definition}
+\newdef{remark}[theorem]{Remark}
+
+%\newtheorem{theorem}{Theorem}[section]
+%\newtheorem{definition}[theorem]{Definition}
 
@@
 
-%opening
-% TODO: Contribution of this paper
-% Q&A: 'Overlay' or 'Peer-to-Peer'/'P2P'
+\markboth{}{}{}
+
 \title{Alleviating the Topology Mismatch Problem in Unstructured and Structured Overlay Networks: A Survey}
+
 \author{
-Vassilis Moustakas$^1$, H\"useyin Akcan$^2$, Mema Roussopoulou$^1$ and Alex Delis$^1$\\
-$^1$Department of Informatics and Telecommunications,\\
-National and Kapodistrian University of Athens\\
-\mbox{\texttt{\{b.moystakas, mema, ad\}@di.uoa.gr}}\\
-$^2$Department of Software Engineering,\\
-Izmir University of Economics, Izmir, Turkey \\
-\mbox{\texttt{huseyin.akcan@ieu.edu.tr}}
+VASSILIS S. MOUSTAKAS\\\emph{National and Kapodistrian University of Athens}\\
+H\"USEYIN AKCAN\\\emph{Izmir University of Economics}\\
+MEMA ROUSSOPOULOS\\\emph{National and Kapodistrian University of Athens}\\
+AND\\
+ALEX DELIS\\\emph{National and Kapodistrian University of Athens}\\
+%  Vassilis Moustakas$^1$, H\"useyin Akcan$^2$, Mema Roussopoulou$^1$ and Alex Delis$^1$\\
+%  $^1$Department of Informatics and Telecommunications,\\
+%  National and Kapodistrian University of Athens\\
+%  \mbox{\texttt{\{b.moystakas, mema, ad\}@di.uoa.gr}}\\
+%  $^2$Department of Software Engineering,\\
+%  Izmir University of Economics, Izmir, Turkey \\
+%  \mbox{\texttt{huseyin.akcan@ieu.edu.tr}}
 }
 
+\begin{abstract}
+Peer-to-peer (P2P) systems is a rapidly growing application architecture.
+The overlay network abstraction on top of a best-effort infrastructure provides a powerful tool to realize a plethora of ever-wanted exotic features like anonymity, high availability and robustness, load balancing, quality-of-service (QoS) at the application layer and many more.
+Unfortunately weaknesses of early effectuations constrained applications from unleashing the full potential of the paradigm.
+One of them is the topology mismatch problem between the overlay and the physical underlying network, a problem that imposes a huge amount of unnecessary stress to the network resources and, thus, impede applications from scaling.
+In this survey we investigate on the recent research developments towards the alleviation of this problem, in both unstructured and structured overlay architectures.
+We visit the various algorithms available in the lietrature, focusing on the presentation of how each of them tries to tackle the problem.
+Additionally, we provide a finer grained categorisation of the approaches based on the generic type of methodology they use in order to accomplish this goal.
+% TODO: Review this after the comparison sections are fixed.
+Finally we discuss the advantages, disadvantages and novelties introduced by the algotithms and offer pictorial comparisons for them in order to further strengthen and organise understanding of the information and knowledge cited, in an attempt to make this survey a one-stop-shop reference for anyone interested in the topic, from researchers and students, to system architects and developers.
+\end{abstract}
+
+\category{}{}{}
+
+\terms{Topology mismatch, distributed, peer-to-peer, P2P, overlay}
+
+\keywords{Topology mismatch problem, distributed algorithm, peer-to-peer architecture, overlay network }
+
 \begin{document}
+
+\begin{bottomstuff}
+SAY SOMETHING COOL HERE...
+\end{bottomstuff}
+
 \maketitle
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{abstract}
-% TODO: Contribution of this paper
-Peer-to-peer (p2p) systems is a rapidly growing application architecture.
-The overlay network abstraction on top of a best-effort infrastructure provides a powerful tool to realize ever-wanted exotic features like anonymity, high availability and robustness, load balancing, QoS at the application layer and many more.
-Unfortunately weaknesses of early effectuations constrained applications from unleashing the full potential of the paradigm.
-One of them is the topology mismatch problem between the overlay and the physical underlying networks, a problem that imposes unnecessary stress to the network resources.
-This survey investigates on the recent research developments towards the alleviation of the problem, in both unstructured and structures p2p schemes.
-\end{abstract}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
-
-% TODO: Contribution of this survey
-% UNDER CONSTRUCTION
-
-% TODO: Structure of this paper
-% UNDER CONSTRUCTION!
-This survey begins with a quick review of the conception of the overlay network and the available architectures of p2p systems.
-It continues with the identification of the phenomenon of the topology mismatch between overlay and underlying networks and its impact on the utilization of network resources.
-Ultimately, it reviews the most important resources found in the literature that attempt to tackle the problem as it proposes a taxonomic scheme, using a set of supertype-subtype relationships based on unique characteristics they have and/or specific goals they target.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% OK
-\section{Overlay Networks and P2P Systems}
-
-\begin{figure}[h!]
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Distributed architectures have witnessed a significant increase of interest during the last decade, as the number of interconnected computers increased, the available digital information kept proliferating, at an exponetial rate each year, and problems to be addressed by IT became more and more demanding.
+Why is there a correlation, and it is not just a coincidence?
+Distributed computation suggests a simple paradigm\footnote{Of course, as always, this is the case only after someone thought about it first!}.
+Get a set of autonomous computing entities and put them to cooperate with each other in order to fullfil a well defined task.
+In the World of the Internet and the World Wide Web this also offers a thrilling promise.
+To harness the immence power scattered across the Internet and get amply rewarded by exploiting the data, storage and computational resources out there.
+
+In order to achieve that, computers collaborate by exchanging messages via an application-level network where they can address each other in the specific context of the task at hand.
+This communication structure is refered to as an \emph{overlay network} which is an abstract, logical inter-connection formed on top of another network\footnote{The Internet itself is an overalay on top of the telephone network.}.
+It consists of a set of nodes, the computing elements, that are being connected by virtual (logical) point-to-point links, perhaps corresponding to multiple physical links.
+Figure~\ref{figure:overlay} illustrates a simple four-node overlay contructed upon a wide area network.
+
+\begin{figure}
 \centering
@@
 
-An \emph{overlay network} is an abstract, logical interconnection of entities
-formed on top of another network. It consists of a set of nodes, the computing
-elements, that are being connected by virtual (logical) point-to-point links.
-For example, the \emph{Asymmetric Digital Subscriber Line (ADSL)} is, actually,
-an overlay network on top of the \emph{Public Switched Telephone Network
-(PSTN)}. Figure \ref{figure:overlay} illustrates an example physical ethernet network
-and the corresponding overlay network formed by four nodes. 
-Nodes $A$ and $B$ are on the same local network, while $C$ and $D$ are in
-different networks. The top layer represents the overlay network formed by these
-nodes, which is a reflection of the application layer connections among these
-nodes. As the topology of the network can change based on the application, in
-this example, it is easy to observe that even though nodes $A$ and $B$ are in the same
-network, their communication goes through node $C$ on another network,
-demonstrating a non-optimal mapping between the physical and the overlay
-networks.
-An overlay network is generally useful when
-abstraction of the network layer from the application layer is important for
-applications, particularly when the underlying network structure is subject to
-frequent changes.%, which is a common case for ad-hoc networks.
-
-The abstraction of overlay networks has been proposed as a way to implement
-efficient, fully distributed, application layer services on top of a best-effort
-IP layer forming a widely ranged family of protocols collectively known as
-\emph{peer-to-peer} or \emph{p2p} for short. P2p network systems gained
-significant attention, in recent years, due to their plethora of unique features
-in supporting file sharing among huge numbers of (inter-) networked computers
-called \emph{peers}, where each, such, \emph{peer} could act both as a resource
-provider and a resource consumer. This changed the traditional
-\emph{client-server} model dominating the internet and lead to the introduction
-of the \emph{servent}-concept \cite{gnutella}, a portmanteau that blends the
-notions of \emph{server} and \emph{client} to denote the twofold role of the
-participants in a p2p network. Such serverless systems, proved to be able to
-achieve outstanding aggregate resource capacities as participants join the
-system
-%\footnote{Unfortunatelly, there has been observed mitigation of this
-%self-scaling property by the undesirable behaviour of participating users,
-%usually answered in popular file-sharing systems; from \emph{free-riding} (i.e.
-%in Napster and Gnutella networks \cite{saroiu_measurefileshare_2002,
-%adar_gnutellafreeriders_2000, hughes_gnutellafreeride_2005}), to the
-%distribution of illegal content and/or other \emph{socio-technical} issues
-%\cite{hughes_socp2p_2008}.}, 
-without requiring additional expenditure for infrastructure.
-
-As p2p systems evolved, three main architectures have been emerged, namely:
-\begin{itemize}
-  \item \emph{centralized}
-  \item \emph{decentralized unstructured}
-  \item \emph{decentralized structured}
-\end{itemize}
-
-\emph{Centralized} architectures were the first to recognize that requests for popular content need not be sent to a central server but instead could be handled by the many hosts that already posses the content.
-\emph{Napster}, for example, maintained a \emph{central index server} based on file lists provided by participating peers.
+The abstraction of the overlay network has been proposed as a way to implement efficient, fully distributed, application-layer services.
+For example, backing the routing of messages to destinations that are not known in advance or building one-to-many addressing schemes
+% TODO find citation
+%[Narada, Adobe's Real Time Media Flow Protocol (RTMFP)]
+that would provide higher quality-of-service (QoS) guarantees.
+\emph{IP multicast}, as well as other proposals such as \emph{IntServ} and \emph{DiffServ} \cite{cisco_diffserv_2005}, have not seen wide acceptance (yet?), largely due to the fact they require support from the physical infrastructure.
+On the other hand, an overlay network can be deployed on end-systems running the overlay protocol software, without cooperation from the ISPs.
+%Academic research includes \cite{chu_esm_2000,jannotti_overcast_2000,kwon_tag_2002} among others.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Peer-to-Peer (P2P) Architectures in a Nutshell}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Overlay networks on top of the best-effort IP layer, formed a widely ranged family of protocols collectively known as \emph{peer-to-peer} or \emph{P2P} for short.
+P2P systems consist of a number of networked computers called \emph{peers}, where each, such, \emph{peer} could act both as a resource provider and a resource consumer.
+This changed the traditional \emph{client-server} model dominating the Internet and lead to the introduction of the \emph{servent} concept \cite{gnutella}, a \emph{portmanteau} that blends the notions of server and client to denote the twofold role of the P2P network participants.
+Such serverless systems, proved to be able to achieve outstanding aggregate resource capacities as more and more participants join the system without requiring additional expenditure for infrastructure\footnote{
+        Unfortunatelly, there has been observed mitigation of this self-scaling property by the undesirable behaviour of participating users, usually answered in popular file-sharing systems; from \emph{free-riding} \cite{saroiu_measurefileshare_2002, adar_gnutellafreeriders_2000, hughes_gnutellafreeride_2005} to the distribution of illegal content and/or other \emph{socio-technical} issues \cite{hughes_socp2p_2008}.
+}.
+P2p network systems gained extra attention whithin the distributed systems super-set, circa 2000, due to the fact that they succesfully supported popular applications that offered file sharing among vast numbers of participating users.
+But their development didn't stay still.
+On the contrary, during the last decade, research on P2P systems was constant and fruitfull generating a wide range of popular protocols, networks and applications.
+As P2P systems evolved three main architectures have been emerged, namely
+\begin{inparaenum}[\itshape i\upshape)]
+  \item \emph{centralized},
+  \item \emph{decentralized unstructured}, and
+  \item \emph{decentralized structured}.
+\end{inparaenum}
+
+\emph{Centralized} architectures (see Figure~\ref{figure:p2p-archs:centralised}) were the first to recognize that requests for resourses (e.g. CPU cycles \cite{seti} or popular content 
+% TODO find citation for Napster
+\cite{}) need not be sent to a dedicated server but instead could be handled by the many hosts that already posses them.
+This schema, though, is not fully distributed since it dictates a centralised searching facility, a characteristic that can be proven its ``Achilles' heel'' in terms of being a scalability bottleneck, a central point of failure and vulnerable to malicious acts (e.g.
+\emph{Denial-of-Service (DoS) attacks}).
+
+The centralized approach has been eventually replaced by architectures that distributed both search and download capabilities (see Figures~\ref{figure:p2p-archs:unstructured-full} and \ref{figure:p2p-archs:unstructured-hierarchical}).
+In these \emph{decentralized} architectures, file placement is random, which means there is no correlation with the network topology whatsoever \cite{yang_improvep2psearch_2002}.
+For this reason they are also refered to as \emph{unstructured}.
+The most important properties of such systems are that they support inherent heterogeneity of peers, are highly resilient to peer failures, and incur low maintenance overhead at handling the dynamics of peer participation \cite{stutzbach_churn_2006}.
+In the literature, they are also known as \emph{broadcast-based} systems, because they use \emph{message flooding} among peers or super-peers to propagate queries
+%\footnote{The life of each query message is constrained by an integer, refered to as \emph{time-to-live (TTL)}.
+When a querry is received by a (super-) peer, this value is decreased by one unit.
+Then, the receiver, back-propagates its results and if the TTL has not reached zero the node forwards the query message to all its neighbours except from the one it received it from.}
+.
+
+Recently, \emph{decentralized structured} schemes have been proposed (see Figure~\ref{figure:p2p-archs:structured}) in order to provide a self-organising infrastructure for large-scale P2P applications \cite{ratnasamy_can_2001, stoica_chord_2001, antony_pastry_2001, zhao_tapestry_2001, maymounkov_kademlia_2002, rgrk_bamboo_2004}.
+They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes through a deterministic mechanism.
+Even though unstructured networks became highly popular among file sharing applications, they do have a certain limitation in locating rare objects.
+The DHT approach of structured networks on the other hand, provides a guaranteed bound on the number of overlay routing hops that have to be taken in order for an object to be located, even in the case where only a single copy exists in the system.
+This means $O \left ( log n \right )$ hops, compared to unstructured networks that require $O \left ( n \right )$ to reliably locate a specific object\footnote{Unfortunatelly this efficiency does not come without cost.
+Structured overlays realise what is called object identifier location as opposed to content-based retrieval of unstructured ones.
+This means that the later can support more versatile object location mechanisms like keyword and partial matching searches.}.
+
+% TODO: Change these pictures to match look-n-feel
+\begin{figure}[ht]
+\centering
+\subfigure[Centralised (SETI@Home or Napster).] {
+  \includegraphics[scale=0.3]{img/centralised.png}
+  \label{figure:p2p-archs:centralised}
+}\qquad\qquad
+\subfigure[Fully unstructured (Gnutella v.0.4).] {
+  \includegraphics[scale=0.3]{img/unstructured-full.png}
+  \label{figure:p2p-archs:unstructured-full}
+}\qquad\qquad
+\subfigure[Hierarchical unstructured (Gnutella v.0.6 or FastTrack).] {
+  \includegraphics[scale=0.3]{img/unstructured-hierarchical.png}
+  \label{figure:p2p-archs:unstructured-hierarchical}
+}\qquad\qquad
+\subfigure[Structured (Chord and Kademlia).] {
+  \includegraphics[scale=0.3]{img/structured.png}
+  \label{figure:p2p-archs:structured}
+}
+\caption{The various peer-to-peer architectures.}
+\label{figure:p2p-archs}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Applications and Implications}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The most successful incarnation of the centrilised approach is without a doubt the popular file-sharing application \emph{Napster}\footnote{Pioneering, though, was the famous \emph{SETI@Home Project} \cite{seti}, which pools spare processing power of participants to search the universe for the existence of extra terrestrial intelligence!}
+% TODO find citation for Napster again
+\cite{}.
+Napster maintained a \emph{central index server} based on file lists provided by participating peers.
 The central index server was queried by the users and it returned pointers to the actual content.
-Thus, by centralizing search while distributing downloads, Napster achieved a highly functional design that was widely acknowledged, at the time, as ``the fastest growing Internet application ever''.
-This centralized search facility, though, renders this scheme not fully distributed and can be proven its ``Achilles' heel'' in terms of being a scalability bottleneck, a central point of failure and vulnerable to malicious acts (e.g.
-\emph{Denial-of-Service (DoS) attacks}).
-
-The centralized approach has been eventually replaced by architectures that distributed both the search and the download capabilities.
-In these \emph{decentralized unstructured} architectures, file placement is random, which means there is no correlation with the network topology whatsoever \cite{yang_improvep2psearch_2002}.
-The most important properties of such systems are that they support inherent heterogeneity of peers, are highly resilient to peer failures, and incur low maintenance overhead at handling the dynamics of peer participation \cite{stutzbach_churn_2006}.
-In the literature, they are also known as \emph{broadcast-based} systems, because they use \emph{message flooding} among peers (i.e.
-Gnutella\cite{gnutella}) or among super-peers (i.e KaZaA\cite{kazaa}) to propagate queries.
-Even though the unstructured approach became highly popular among file sharing applications, they do have a certain disadvantage in locating rare objects due to the unstructured nature and the limitations of the flooding approach.
-%Each (super-) peer receiving a query decreases the incoming query message's lifetime which is measured in \emph{time-to-live (TTL)} hops, back-propagates its results, and  if the lifetime of the incoming message did not expired (TTL did not reach zero), it forwards the message to all its neighbours except from the one it receives it from.
-% TODO: maybe enumerate applications of decentralized unstructured p2ps.
-
-Recently, \emph{decentralized structured} schemes have been proposed in order to provide a self-organising infrastructure for large-scale p2p applications \cite{ratnasamy_can_2001,stoica_chord_2001,antony_pastry_2001,zhao_tapestry_2001,maymounkov_kademlia_2002}.
-They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes through a deterministic mechanism.
-The main advantage of the DHT approach of decentralized structured p2p networks over decentralized unstructured ones, is that they provide a guaranteed bound on the number of overlay routing hops that have to be taken in order for an object to be located even in the case where only a single copy exists in the system.
-This is $O \left ( log n \right )$ compared to Gnutella, that requires $O \left ( n \right )$ to reliably locate a specific object.
-% TODO: maybe enumerate applications of decentralized structured p2ps. (OceanStore, Far)
-
-%{\sethlcolor{yellow}\hl{HA: Give the brief advantages and disadvantages of structured and unstructured
-%approaches here!}}
-
-% TODO: enumerate applications of overlay networks and p2p systems regardless of architecture
-% UNDER CONSTRUCTION
-% Q&A: Need more??
-Initial efforts targeted on the implementation of one-to-many addressing schemes
-that could replace \emph{IP multicast} in providing higher quality of streaming
-media through \emph{quality of service (QoS)} guarantees. IP multicast as well
-as other proposals such as \emph{IntServ} and \emph{DiffServ}
-\cite{cisco_diffserv_2005} have not seen wide acceptance (yet?), largely because
-they require changes to all routers of the network. On the other hand, an
-overlay network can be incrementally deployed on end-systems running the overlay
-protocol software, without cooperation from the ISPs. Academic research includes
-\cite{chu_esm_2000,jannotti_overcast_2000,kwon_tag_2002} among others.  Overlay
-networks were also implemented in order to back the routing of messages to
-destinations whose address is not known in advance. This resulted in the
-appearance of well known peer-to-peer protocols such as \cite{gnutella} and
-\cite{maymounkov_kademlia_2002} widely used for digital content sharing among
-their network nodes.  Other special purpose overlay networks include
-\cite{anderson_ron_2001} for resilient routing; \cite{subramanian_overqos_2004}
-for quality of service guarantees; and \cite{clarke_freenet_2001} for anonymity;
-to name just a few.
-
-The main focus of this paper is the topology mismatch problem, therefore the
-available approaches in overlay networks are briefly summarized, and interested
-readers are pointed to the available surveys on overlay networks
-\cite{TheotokisS04,LuaCPSL05}.
+Thus, by centralizing search while distributing downloads, Napster achieved a highly functional design that was widely acknowledged, at the time, as ``the fastest growing Internet application ever'' that toped some 26 million users and 80 million songs \cite{jmm_naptopusage_2001}.
+
+Napster was ultimatelly brought down after Recording Industry Association of America (RIAA) filed a lawsuit for copyright infringement of the intellectual property of artists whose music was distributed by the network\footnote{
+      The sueing spree was initiated by heavy metal giants \emph{Metallica} that feared Napster was undermining record sales.
+Opposed opinions instantly emerged  with best example the british alternative rock band \emph{Radiohead} whose album \emph{Kid A} had also been leaked to Napster three months before reaching music store shelves \cite{rm_radioheadkida_2000}.
+Despite music critics labeling the album ``experimental'' and ``challenging'' (it had no official singles nor music videos for any of its songs), in October 2000 Kid A captured the number one spot on the Billboard 200 sales chart in its debut week.
+Unlike Metallica, Madonna and Dr.Dre that led litigation with Napster, Radiohead had never hit the top20 in the US before.
+}
+\cite{wiki_amnaptrial_2001}.
+This triggered a shift towards decentralised networks like the classic \emph{Gnutella} and, its ``rival'' at the time, \emph{FastTrack}.
+Gnutella, in its first realisation (v.0.4 \cite{gnutellav0.4}), was a fully decentralised approach.
+Unfortunatelly, network resourses tended to be overwhelmed by message flooding and if queries were constrained to live for some (small) number of hops, this ultimatelly resulted to a shrinked search scope, leading to a death spiral that prevented the network from scaling.
+The solution came with a slightly more structured network by introducing the notions of \emph{ultra-peers} and \emph{leaf-nodes} (v.0.6 \cite{gnutella}).
+With \emph{Limewire}, the most widely used software that implemented the protocol, the network was highly successful and was reported to have reached connecting more that 3 million users concurrently by the end of 2006 \cite{rsr_gnutellaevol_2006}.
+Fasttrack and its implementing software KaZaA \cite{kazaa}, on the other hand, went with the hierarchical approach (\emph{super-nodes} in their context) from the very begining, dynamically assigning indexing functions to special peers in order to increase search performace and scaling potential.
+FastTrack network had an average of 3 million concurrently connected nodes in 2003 \cite{lkr_kazaameasure_2005}.
+It is estimated that the total number of users was greater than that of Napster at its peak.
+With Limewire discontinued/shutdown in 2010 and its successor, \emph{Frostwire}, moving to the Bittorent protocol in 2011, seem to very well signal the end of the once-mighty Gnutella network.
+KaZaA continues as a propriatery music store solution.
+In the meanwhile, Fastrack design team had already introduced a VoIP and text messaging application based on the design principles of KaZaA's hierarchical P2P nature, that was meant to change how people communicate around the world.
+This app was no other than Skype \cite{skype}.
+Despite its wide usage, little is known for the technical aspects of its overlay network, mainly due to its proprietary nature.
+Studies, though, confirm that it shares strong structural similarities to its file-sharing predecessor (ordinary hosts and super-nodes in Skype's terminology) but they significantly differentiate in terms of traffic behaviour \cite{g_voipskypesec_2005, gdj_skypestudy_2006}.
+Skype was introduced in 2003 and from then on it witnessed frenzy-rythmed increase to its popularity, with more that 650 milion registered users in 2011  \cite{skypetotalusers} and reported to have broken the barrier of 35 million skypers concurrently online on the the 5th of March 2012 \cite{skypesymusers}.
+On October 2011 Microsoft Corp.
+and Skype Global S.à.r.l.
+anounnced the completion of Microsoft's acquisition of Skype for \$$8.5$ billion \cite{skypemicrosoft}.
+
+File sharing was the most popular application area for structured architectures as well. Overnet
+% TODO find citation
+\cite{},
+for example, was created for sharing large files (e.g., movies and CD images).
+It implemented the Kademlia tree-like partitioning algorithm.
+Other implementations of Kademlia are the Kad Network (eMule/aMule  clients), BitTorrent for trackerless torrernts (official BitTorrent client, $\mu$Torrent, BitComet, Transmission and BitSpirit all share compatibility with the Mainline DHT which realises the Kademlia protocol) and the Gnutella DHT (first implemented by LimeWire as the Mojito DHT).
+Lately, RetroShare \cite{retroshare} is the new trend for file-sharing which incorporates an implementation of the Kademlia DHT to create a Friend-to-Friend (F2F) network with some attributes of a darknet as well.
+But the potential of structured architectures was not exhausted only to the above.
+Structured P2P networks became popular in distributed systems because they could scale smoothly and have fast and reliable object location mechanisms.
+Thus, research and development was very active and productive resulting in a variety of diverse applications such like distributed search engines \cite{yaci}, distributed data storage systems \cite{kbc_oceanstore_2000, bdet_fsdfs_2000, dkkms_cfs_2001, dr_pastutility_2001, abc_farsite_2002, mmfc_ivy_2002, arla, agebh_dks_2003}, Web caches, archives, and publishing systems \cite{ird_squirrel_2002, bags_youserv_2002, wrc_publius_2000, wm_tangler_2001}, messaging applications \cite{threedegrees}, event notification infrastructures \cite{rkcd_scribe_2001, cdkr_scribe_2002, agebh_dks_2003}, naming services \cite{cmm_chorddns_2002}, censor-resistent stores \cite{cswh_freenet_2001} and lately even cloud-based platforms \cite{mgpj_cloudsnap_2011}.
+
+%%%%%%%% NEW TRENDS %%%%%%%%
+%File sharing Limewire Kazzaa The file-sharing landscape is slowly adjusting in response to the continued push for more anti-piracy tools, the final Pirate Bay verdict, and the raids and arrests in the Megaupload case, SOPA, ACTA.
+This leads to users trying to find new solutions for anonymous, uncensored, decentralised file sharing solutions like Freenet, Tribler and RetroShare where connections are made only between trusted peers — sometimes called "friends" and the respective formed network F2F.
+
+% TODO: READ Dalesa published at eAsia 2009, http://www.e-asia.org/2009/ET_Presentations/Wathsala.pdf
+
+The huge numbers of users, for the various applications we presented above, reveals the tremendous popularity P2P architectures have enjoyed.
+This, though, does not come without any toll.
+Measurements on some popular systems, such as FastTrack, Gnutella and DirectConnect have shown that the P2P traffic contributes the largest portion of the overall Internet traffic \cite{seroiu_analysiscds_2002, sen_analyzep2ptraffic_2004}.
+Even with later studies, this observation still holds on.
+Using deep packet inspection techniques deployed at selected customer sites \emph{Ipoque} and \emph{CacheLogic} were able to track down application usage \cite{cachelogic, ipoque2007, ipoque2009}.
+CacheLogic reports that more than 60\% of the traffic is identified as P2P, with Ipoque verifying this observation reporting similar results for 2007, while observing some small decline for 2008/2009 (but still remaining dominant compared to other applications).
+Apart from its participation in overall usage terms, P2P is already a huge contibutor to the overall data exchanged across the Internet\footnote{
+    Mainly because file-sharing P2P is used to exchange big files of HD video or high sampled audio content compared to simple Web content distributed via, for example, HTTP.
+}.
+\cite{multinteligence} claims that P2P traffic in 2007 was around 1.6 petabytes per month and the potential is to be quintubled by the end of 2012 reaching almost 8 petabytes per month.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The Topology Mismatch Problem}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+One of the major issues that defines the efficiency of an overlay network is the mapping of the overlay structure to the underlying infrastructure.
+Remember, in Figure~\ref{figure:overlay} Nodes $A$ and $B$ are on the same local network, while $C$ and $D$ are in different networks.
+The top layer represents the overlay interconnection formed by these nodes at a higher level.
+Link connections there, can change as needed by the running environment, without any particular constraint by the underlying physical topology.
+Let's have a closer look with the help of a simple example.
+Assume nodes $A$, $B$, $C$ and $D$ are connected through the physical network shown in Figure~\ref{figure:phys}, where the network costs are given in milliseconds, and peer $A$ sends a message to peer $D$.
+
+\begin{figure}
+\centering
+  \includegraphics[scale=0.8]{img/phys.pdf}
+\caption{Interconnection of nodes in the physical level.}
+\label{figure:phys}
+\end{figure}
+
+If these peers participate in an overlay network according to one of the setups of Figure~\ref{figure:overlay-confs} then users will expierience different performances.
+In the overlay depicted in Figure~\ref{figure:overlay-1}, the message will traverse the following sequence of links in the physical layer (marked with their costs): $3 \rightarrow 7 \rightarrow 1 \rightarrow 2 \rightarrow 2 \rightarrow 1 \rightarrow 7 \rightarrow 8 \rightarrow 5 \rightarrow 5 \rightarrow 4$.
+In the alternative overlay of Figure ~\ref{figure:overlay-2}, the path will be: $3 \rightarrow 8 \rightarrow 4$.
+The total cost of the first sequnce is $45 ms$, while the second yields a mere $15 ms$.
+Therefore, we can conclude that the second overlay is more congruent with the underlying physical network than the first one and, thus, is more efficient.
+
+\begin{figure}[ht]
+\centering
+\subfigure[Nodes B and C have direct overlay connection.] {
+  \includegraphics[scale=0.8]{img/overlay-1.pdf}
+  \label{figure:overlay-1}
+}\qquad\qquad
+\subfigure[Nodes A and D have direct overlay connection.] {
+  \includegraphics[scale=0.8]{img/overlay-2.pdf}
+  \label{figure:overlay-2}
+}
+\caption{Two different overlay connection configuration.}
+\label{figure:overlay-confs}
+\end{figure}
+
+Ideally, overlay networks are expected to achieve an optimal mapping among the overlay paths and the underlying physical links, and avoid inefficient states.
+The problem of constructing an optimal overlay is referred to as the \emph{topology mismatch problem}, and formally defined as follows:
+\begin{definition}
+    Let $V = \{v_1, ..., v_n\}$ be a set of points denoting the network nodes, $\{v_i, v_j\} \in E$ be the set of unicast distances between nodes $v_i$ and $v_j$, $G=(V,E)$ be a complete distance graph over $V$.
+The topology mismatch problem is to construct a minimal spanning tree,  where node degree is restricted to a constant ($k\geq 2$) by the bandwidth of each node $v_i$.
+\end{definition}
+
+Early incarnations of the overlay protocols, however, did not make use of the optimal mapping with the physical network topology.
+Gnutella protocol for example was considered far from scalable \cite{ritter_gnucantscale_2001}.
+Since every peer can chοse its logical neighbours without knowledge of the underlying network had as a result a mismatch between physical and application-level topologies.
+Additionally, queries may be flooded to multiple paths merging in one peer while
+in such case one of the paths would have been enough.
+Nevertheless, peers may forward the same message to each other before they receive the query messages from each other.
+\cite{matei_mapgnutella_2002} showed that, even given that 95\% of any two nodes are less than 7 hops away from each other, a flooding-based algorithm, like the one used by Gnutella, can generate 330TB/month in a 50.000 node network!
+Therefore, P2P traffic can have a serious effect on the overall well-being of the Internet as a whole.
+
+Similar problems are observed in decentralized structured schemes also.
+Typically, node IDs are assigned \emph{randomly}, resulting in excellent load balancing, scalability and robustness.
+Unfortunately, this randomness has a negative impact on the \emph{routing locality} of the network.
+This means that
+even though the target node can be reached with logarithmic overlay hops, the distance traveled in the physical underlying network, during the overlay routing
+process, can be far from optimal.
+The ratio of the physical distance, a query travels through the overlay to an object, and the minimal distance to that object (i.e.
+through IP) is known in the literature as \emph{stretch} or \emph{Relative Delay Penalty (RDP)} \cite{chu_esm_2000}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Motivation and our Goal for this Survey}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+A topology unaware overlay network is able to control the sequence of peers a message traverses before reaching its destination, but it is completely unaware of how the actual packets are switched at the underlying infrastructure along the overlay path.
+For example, a single logical point-to-point link on the overlay, most of the time, corresponds to multiple physical links in the underlying layer.
+Additionally, a link in the underlying network often serves the mapping of several overlay paths causing increase of the traffic on the physical link, which is also called the link's \emph{stress}\cite{chu_esm_2002}.
+Furthermore, the network's \emph{churn}, that is the stochastic behaviour of peers randomly joining and leaving the P2P network, all cause the overlay to physical mapping to create such unnecessary, redundant maintenance traffic that impacts the efficiency of the network in terms of average response time.
+
+The topology mismatch problem is exacerbated by the high amounts of traffic generated by the highly popular P2P applications.
+This creats the killer combination for the network infrastructure, ultimately effecting all Internet users.
+Solving it, on the other hand, is no easy task.
+Topology mismatch is known to be an NP-Hard problem \cite{NPBOOK,chawathe_scattercast_2000}.
+Moreover, the Internet itself is an interesting environment where end-to-end latencies demonstrate what is called \emph{triangle inequality violations (TIVs)}, which further complicate the problem.
+These delay space TIV is a consequence of the Internet's structure and routing policies and thus will remain a property of the Internet for the foreseeable future \cite{zheng_irprtt_2005}.
+TIVs affect network coordinate \cite{cox_vivaldi_2004,wong_meridian_2005} and positioning \cite{ng_gnp_2001} and makes the construction of location/delay aware overlays difficult.
+An extensive array of heuristic approaches have been proposed in an effort to find an algorithm that performs reasonably and consistently well.
+The purpose of this survey is to gather the extensive knowledge that has been produced recently in this research field and organise the available information ultimately helping the reader sift through the myriad of these efforts, understand the different approaches and identify though a taxonomy the advantages and disadvantages of each of them.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Survey Outline}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% TODO: Review this after the content and article structure is fixed.
+In the introduction we had the oportunity to recognise the importance of overly networks in contemporary application architectures.
+Also we quickly reviewed the basic trends in both overlay construction\footnote{
+    The main focus of this paper is the topology mismatch problem, therefore the available approaches in overlay networks are briefly summarized, and interested readers are pointed to the available surveys on overlay networks \cite{TheotokisS04,LuaCPSL05}.
+}
+and application development.
+Last but not least, we identified and formaly defined the phenomenon of the topology mismatch problem and its impact on the utilization of network resources.
+The rest of the survey\ldots
+
+Sections~\ref{sec:unstructured} and \ref{sec:structured} discuss the recent academic work that has been conducted in the field.
+
+Ultimately, it reviews the most important resources found in the literature that attempt to tackle the problem as it proposes a taxonomic scheme, using a set of supertype-subtype relationships based on unique characteristics they have and/or specific goals they target.
+What follows does not claim to be a thorough citation of all known protocols that are available out there, but instead, a careful selection of those that left a distinctive fingerprint contribution in the efforts of the research community to alleviate the topology mismatch problem.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{The Topology Mismatch Problem}
-
-% TODO: Define the problem
-% UNDER CONSTRUCTION
-% Q&A is it too late??
-
-%{\sethlcolor{yellow}\hl{ {\cite{hsiao_redblue_2009}} states that:
-%Liu {\cite{liu_thancs_2008}}  concluded that more than 70 percent of communication
-%paths in an unstructured overlay do not exploit the physical network topology
-%(i.e., the Internet topology), resulting in a topology mismatch between the
-%overlay and the underlying network.  } }
-
-One of the major issues that defines the efficiency of the overlay network is
-the mapping of the physical links and the overlay paths. Ideally, overlay
-networks are expected to achieve an optimal mapping among the overlay paths and
-the underlying physical links, and avoid the inefficient mapping that is shown
-in Figure \ref{figure:overlay}. The problem of constructing an optimal overlay
-is referred to as the \emph{topology mismatch problem}, and formally defined as
-follows:
-
-\begin{definition}
-    Let $V = \{v_1, ..., v_n\}$ be a set of points denoting the network nodes,
-    $\{v_i, v_j\} \in E$ be the set of unicast distances between nodes $v_i$ and
-    $v_j$, $G=(V,E)$ be a complete distance graph over $V$. The topology
-    mismatch problem is to construct a minimal spanning tree,  where 
-    node degree is restricted to a constant ($k\geq 2$) by the bandwidth of each node $v_i$.
-\end{definition}
-
-Constructing degree constrained spanning tree is known to be NP-Hard
-\cite{NPBOOK}, as well as the topology mismatch problem
-\cite{chawathe_scattercast_2000}. 
- Moreover, the internet is an interesting environment
-where end-to-end latencies demonstrate triangle inequality violations (TIVs),
-which further complicates the problem.
-These delay space TIV is a consequence of the Internet's structure and routing
-policies and thus will remain a property of the Internet for the foreseeable
-future \cite{zheng_irprtt_2005}. TIVs affect network coordinate
-\cite{cox_vivaldi_2004,wong_meridian_2005} and positioning \cite{ng_gnp_2001}
-and makes difficult the construction of delay aware overlays.  A number of
-heuristic approaches have been proposed in an effort to finding an algorithm
-that performs reasonably and consistently well.
-
-A topology unaware overlay network is able to control the sequence of peers a
-message traverses before reaching its destination, but it is completely unaware
-of how the actual packets are switched at the underlying infrastructure along
-the overlay path. For example, a single logical point-to-point link on the
-overlay, most of the time, corresponds to multiple physical links in the
-underlying layer. Additionally, a link in the underlying network often serves
-the mapping of several overlay paths causing increase of the traffic on the
-physical link, which is also called the link's
-\emph{stress}\cite{chu_esm_2002}.  Furthermore, the stochastic behaviour of
-peers randomly joining and leaving the peer-to-peer network all cause the
-overlay to physical mapping to create such unnecessary, redundant maintenance
-traffic that impacts the efficiency of the network in terms of average response
-time.
-
-%{\sethlcolor{yellow}\hl{
-%HA: Motivation to solve topology mismatch problem:
-%
-%Most of the traffic (find refs!! see Location aware topology matching paper
-%introduction for refs [31][29]) in the internet today is generated by P2P
-%networks. The inefficient use of the network resources by the P2P protocols,
-%especially the problem known as topology mismatch problem, causes unnecessary
-%traffic load and a heavy stress
-%on the internet infrastructure in general, and effects other users of the
-%internet than the P2P nodes.
-%}}
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.8]{img/phys.pdf}
-\caption{Interconnection of nodes in the physical level.}
-\label{figure:phys}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.8]{img/over1.pdf}
-  \hspace{3ex}
-  \vline
-  \hspace{3ex}
-  \includegraphics[scale=0.8]{img/over2.pdf}
-\caption{HA: Put subfigure, and a new caption here}
-\label{figure:over1}
-\end{figure}
-
-For example, assume nodes $A$, $B$, $C$ and $D$ are connected through the
-physical network shown in Figure~\ref{figure:phys}, where the network costs are
-given in milliseconds, and peer $A$ sends a message to peer $D$. If these peers
-participate in an overlay network according to one of the setups of
-Figures~\ref{figure:over1}~(left) and~\ref{figure:over1}~(right) then users will
-yield different performances. In the first overlay shown in Figure
-\ref{figure:over1}~(left), the message will traverse the following sequence of
-links in the physical layer (marked with their costs): $3 \rightarrow 7
-\rightarrow 1 \rightarrow 2 \rightarrow 2 \rightarrow 1 \rightarrow 7
-\rightarrow 8 \rightarrow 5 \rightarrow 5 \rightarrow 4$. In the second overlay
-shown in Figure ~\ref{figure:over1}~(right), the path will be: $3 \rightarrow 8
-\rightarrow 4$. The total cost of the first path is $45 ms$, while the second
-only costs a mere $15 ms$. Therefore, we can conclude that the second overlay is
-more congruent with the underlying physical network than the first one and thus
-more efficient.
-
-% TODO: Argue on the problem
-% UNDER CONSTRUCTION
-Early incarnations of the overlay protocols, however, did not make use of the optimal
-mapping with the physical network topology. Early Gnutella protocol, though were
-considered far from scalable \cite{ritter_gnucantscale_2001}, randomly choose
-logical neighbours without knowledge of the underlying network, effectively
-causing mismatch between physical and application-level topologies.
-Additionally, queries may be flooded to multiple paths merging in one peer while
-in such case one of the paths would have been enough. 
-%Nevertheless, peers may
-%forward the same message to each other before they receive the query messages
-%from each other.
-
-Similar problems are observed in decentralized structured schemes also.
-Typically, node IDs are assigned \emph{randomly}, resulting in excellent load
-balancing, scalability and robustness.  Unfortunately, this randomness has a
-negative impact on the \emph{routing locality} of the network. This means that
-even though the target node can be reached with logarithmic overlay hops, the
-distance traveled in the physical underlying network, during the overlay routing
-process, can be far from optimal.  The ratio of the physical distance, a query
-travels through the overlay to an object, and the minimal distance to that
-object (i.e. though IP) is known in the literature as \emph{stretch} or
-\emph{Relative Delay Penalty (RDP)} \cite{chu_esm_2000}.
-
-Studies on peer-to-peer traffic reveals the effect of the peer-to-peer traffic
-on the overall internet.  Measurements
-\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004} on some popular
-systems, such as FastTrack, Gnutella and DirectConnect have shown that the
-peer-to-peer traffic contributes the largest portion of the overall Internet
-traffic.  \cite{matei_mapgnutella_2002} showed that, even given that 95\% of any
-two nodes are less than 7 hops away from each other, a flooding-based algorithm,
-like the one used by Gnutella, can generate 330TB/month in a 50.000 node
-network!  Therefore, the efficiency of the overlay networks has serious
-consequences on the overall well-being of the modern internet.
+%\section{The Topology Mismatch Problem}
+
+%\section{ADDITIONAL MATERIAL}
+
+%%%%%%%
+%%%%%%% TODO: NEEDS REWORDING
+%%%%%%%
+%%%%%%%
+%%%%%%%
+%Bindal et al. [14] proposed to improve the traffic
+%locality in BitTorrent systems via biased neighbor selection
+%enforced by packet control mechanisms on border gateway
+%routers. More recently, Ono [15] proposed to leverage the
+%existing CDN infrastructure to estimate the distance between
+%BitTorrent peers before making peer connections. P4P [16],
+%instead, aimed to provide a general framework for P2P
+%applications by developing ISP-application interfaces so that
+%application software can use the network layer information,
+%such as congestion status and network topology, for better
+%performance and resource utilization.
 
@@
 
-Having understood the nature of the problem that peer-to-peer architectures face,
-Section \ref{sec:unstructured} and \ref{sec:structured} discuss the recent academic work that has been conducted
-in the field. What follows does not claim
-to be a thorough citation of all known protocols that are available out there,
-but instead, a careful selection of those that left a distinctive fingerprint
-contribution in the efforts of the research community to alleviate the topology
-mismatch problem.
-
-%{\sethlcolor{yellow}\hl{
-%Create a theoretical section. This is an Approximation alg for Top-mis problem
-%
-%A Near-Optimal Algorithm Attacking
-%the Topology Mismatch Problem
-%in Unstructured Peer-to-Peer Networks
-%}}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-In this section, the unstructured decentralized P2P algorithms are analysed based on their architecture and
-base methodology to tackle the topology mismatch problem. The
-unstructured decentralized algorithms are visited and categorized based on their
-use of the overlay structure, messaging between nodes, and on the techniques
-proposed to detect proximity to optimize the topology. 
-The base methods that unstructured P2P algorithms use to tackle the topology
-mismatch problem can be categorized as Broadcast, Cache, Overlay Optimization, 
-and Landmark Based. In this section, these methodologies are discussed in
-detail and the algorithms in each category are highlighted.
+This section is dedicated to the unstructured P2P architectures.
+In the next paragraphs, algorithms for tackling the topology mismatch problem on such networks are visited, analysed and categorised based on the use of the overlay structure, messaging approach for peer communication and techniques proposed for detecting proximity in order to optimise overlay topology.
+In short, the methods that are highlighted here are divided into
+\begin{inparaenum}[\itshape i\upshape)]
+  \item broadcast-based,
+  \item cache-based,
+  \item overlay optimisation-based, and
+  \item landmark-based
+\end{inparaenum}
+ approaches.
+
+% TODO: READ A Near-Optimal Algorithm Attacking the Topology Mismatch Problem in Unstructured Peer-to-Peer Networks (This is an Approximation alg for Top-mis problem)
 
@@
 
-In the naive Gnutella protocol, each received query request is forwarded to all the
-neighbours, which clearly generates unnecessary traffic over the network. The
-algorithms categorized in the \textit{broadcast based} methodology prefers
-re-broadcasting the query messages to a selected subset of their neighbours,
-where the selection criteria varies between algorithms.
-The selection is made using one or a
-combination of various statistical metrics, such as the reliability or the latency
-of the link between nodes, etc.
-The forwarding based approach enhances search
-efficiency but has also the drawbacks of having drastically reduced search scope
-due to the selective re-broadcasting policy, and scalability problems on large
-P2P networks.
+In the first realisation of the Gnutella protocol, each query request received by a peer is then forwarded to all its
+neighbours, clearly generating unnecessary traffic over the network.
+The algorithms highlighted in this paragraph are referred to as broadcast-based because they enhance the "naive" broadcasting approach described above and prefer forwarding messages to a selected subset of a node's outgoing links.
+The selection criteria vary depending on the algorithm but they mainly use one or multiple statistical metrics that span from the obvious ones, like the latency of the link between the nodes, to more sophisticated alternatives (that may even take into account application level requirements) such as the reliability of the neighbour.
+The forwarding-based approach has the advantage of enhancing search efficiency but on the other hand suffer drawbacks like the drastic reduction of the search scope because of the restrictions of the broadcasting policies as well as limitations when scaling to large networks.
+
 %Expanding the search scope, on the other hand, is no easy task
@@
 
-The forwarding based P2P algorithms and their details are presented below.
-
 \paragraph*{\bf Improving search in peer-to-peer networks}
@@
 \cite{yang_improvep2psearch_2002} proposes an easy to implement and practical
-solution to the inefficiency problem caused by blind flooding in Gnutella like
+solution to the inefficiency problem caused by blind flooding in Gnutella-like
 P2P file sharing protocols. The paper replaces blind flooding with three
@@
 bandwidth, or the query load of the neighbour. In \textit{local indices}, each
-node sustains a local data index of the nodes within a radius of $r$ hops of
+node sustains a local data index of the nodes within a radius of $r$ hops from
 itself and uses this local index to remotely query the neighbour nodes without
@@
 problem of the unstructured P2P file sharing systems, in particular Gnutella.
-The main novelty in the design of Gia is the replacement of blind flooding
+The main novelty in the design of Gia is the replacement of the blind flooding
 approach of Gnutella with random walks \cite{lv_randomwalks_2002}. Although
 random walks is a step in the right direction, issuing only a single copy of the
-query within the network reduces the search scope, thus affects the success rate
+query within the network reduces the search scope, thus affecting negatively the success rate
 of the query.  In order to overcome this limitation, Gia introduces a
-token-based flow control algorithm, which is essentially an intelligent flow
-control algorithm that gradually redirects the queries to nodes with more chance
-of answering. Gia also acknowledges the heterogenity in peer bandwidth,
+token-based flow control mechanism, which is essentially an intelligent flow
+control algorithm that gradually redirects the queries to nodes which are more likely to answer.
+In order to prevent overloading of nodes with query requests, Gia uses a token-based flow control algorithm in which each node announces the number of query requests it
+can handle in terms of tokens to its neighbours, so that neighbours only forward
+query requests to nodes that they previously received tokens from. Gia also acknowledges the heterogenity in peer bandwidth,
 processing power, disk speed, etc, of the nodes in P2P networks and uses this
-heterogenity when connecting nodes to each other, such that by using a topology
+information when connecting nodes to each other and by using a topology
 adaptation algorithm, Gia ensures that high capacity nodes have high degrees and
-low capacity nodes are within short proximity of high capacity nodes. In order
-to prevent overloading of nodes with query requests, Gia uses a token-based flow
-control algorithm in which each node announces the number of query requests it
-can handle in terms of tokens to its neighbours, so that neighbours only forward
-query requests to nodes that they previously received tokens from. Although the
+low capacity nodes are within short proximity to high capacity ones. Although the
 topology adaptation algorithm Gia uses improves the scalability of the network,
@@
 
-The flooding based query approach of Gnutella creates many duplicate packets in
-the network. Nodes receive the same query request from various neighbours, which
-they generally discard when identified as duplicate. One other reason for
-duplicate messages is the cycles in the forwarding paths. These duplicate
-packets affect the performance of the P2P network negatively by increasing the
-overall resource usage. Moreover, the more active nodes, that have high
-capacity, high bandwidth, or contribute more to the network, suffer more from
-the overhead caused by the duplicate packets. Therefore, \emph{Distributed Cycle
-Minimization Protocol (DCMP)} \cite{zhu_dcmp_2008} is introduced as a method to
-remove the cycles and eliminate the duplicate packets, without sacrificing the
+We have already mentioned Gnutella's inefficient broadcasting mechanism in flooding messages across the network with no actual targeting.
+This leads, to nodes that are far from the path towards the query request gratification, to spend their valuable resources to receive, queue and process messages that they shouldn't have received in the first place and eventually simply forward them blindly to continue this vicious cycle.
+Except from that, though, an other case of wasting network resources is when a node that is on the right path for the query, receives the same message multiple times from different neighbours.
+The reason for these duplicate messages is cycles in the forwarding paths.
+Actually, this duplicate packet problem seems to hurt more the active nodes; those with higher capacities and bandwidth that contribute the most to the network.
+Distributed Cycle Minimization Protocol (DCMP) \cite{zhu_dcmp_2008} introduces a method to
+remove the cycles and eliminate the duplicate packets, without sacrificing overlay
 connectivity. In DCMP, once a cycle is detected, the most powerful node in the
-cycle is elected as the \emph{Gate Peer} and the cycle is cut from a special
-link such that the distance of all the nodes within the cycle to the \emph{Gate
-Peer} is minimized. The process is managed by using two special message types,
-\emph{Information Collection (IC)} and \emph{Cut Message (CM)}. One disadvantage
-of DCMP is that since the distance a forwarded message can travel is limited
-with the TTL value, which is practically $7$ in Gnutella, the DCMP cannot detect
-cycles of length larger than $7$. Even though cycle elimination improves the
-network performance, it does not solve the topology mismatch problem.
+cycle is elected as the \emph{Gate Peer} and the cycle is then destroyed by cutting that special link which will result in the minimisation of the distance to the Gate Peer for all the nodes that participated in that cycle.
+The process is managed by using two special message types namely \emph{Information Collection (IC)} and \emph{Cut Message (CM)}.
+One disadvantage of DCMP is that since the distance a forwarded message can travel is limited
+by the TTL value, which is practically $7$ in most cases DCMP cannot detect
+cycles formed by more than 7 peers. Even though cycle elimination improves the
+network performance, it does not actually solve the topology mismatch problem.
 
@@
 \subsection{Cache Based}
-Caching is a widely used methodology to exploit the locality to minimize the
-redundant transfer of data, which is successfully used in web servers and file servers.
-As peers in a P2P system also work as servers, intuitively it is expected that
-P2P file sharing systems can also benefit from caching. However, one of the most
-important characteristic properties of P2P systems is that nodes
-join and leave frequently, and the lifetime of a query is relatively short
-compared to web servers, which makes caching in P2P networks nontrivial. There are
-usually two levels of caching possible in file sharing systems, the indices or
-pointers to data can be cached, or the data itself can be cached. Caching is
-already implemented and successfully used in some commercial P2P systems. One
-popular example is KazaA, which uses caching of indices in its super peers.
-We present the other protocols that use caching and their details below.
+Caching is a widely used technique to exploit locality and minimise
+redundant transfer of data, initially adopted by web and file server application environments with great success.
+Since peers in a P2P system also work as servers, it is intuitively expected that P2P file sharing systems can also benefit from caching.
+However, two of the main characteristic properties of P2P systems, nodes frequently joining and leaving and query lifetime relatively short (at least compared to those needed to be served by web servers) , makes caching in P2P networks a non-trivial issue.
+There are
+usually two levels of caching possible in file sharing systems, caching indices or pointers to data or caching the data itself.
+Caching is already implemented and successfully used in some commercial P2P systems.
+One popular example is KazaA, which uses caching of indices on its super-peer level.
+In the paragraphs below, caching-based algorithms are presented and explained in more depth.
 
@@
 \cite{Cohen02} aim to improve the inefficient blind search algorithm by
-replicating the data in a peer to peer network. The main intuition behind the
+replicating the data in a P2P network. The main intuition behind the
 idea of replication, or using cached copies, is that as the number of copies for
@@
 even a random one, to locate these items. In order to analyse the feasibility of
-such replication approach, the authors investigate three different replication
+such an approach, the authors investigate three different replication
 strategies, namely uniform, proportional, and square root allocation. In the
@@
 rate, so that frequently queried items are replicated more. The square root
-allocation is a strategy proposed by the authors, which is a model between the
-uniform and proportional allocation.  In order to evaluate the outcomes of the
-replication models, the authors compare the overall costs of successful and
-unsuccessful searches in the network. The results can be summarized as follows;
-the uniform allocation model minimizes the maximum search size, therefore
-reduces the time spent on an unsuccessful search. The proportional model, on
-the other hand, promotes the more frequently queried items by replicating them
+allocation is a strategy proposed in this paper, which is a model between the
+uniform and proportional allocation.  In order to evaluate the outcome of the
+replication models, the overall costs of successful and unsuccessful searches in the network are compared.
+Results argue that the uniform allocation model minimizes the maximum search size, therefore reduces the time spent on an unsuccessful search.
+The proportional model, on the other hand, promotes the more frequently queried items by replicating them
 more, therefore decreases the search time for popular items, but suffers when
-locating the rare items. The authors also claim that the expected successful
+needing to locate the rare items. The authors also claim that the expected successful
 search size is the same for uniform and proportional models, and any approach
@@
 
-\cite{Markatos02} analysis the network traces of the Gnutella network and
-detects that query requests in Gnutella has locality, and proposes a caching
-algorithm that exploits this locality. The analysis of the trace data reveals
-other important facts about the structure of the Gnutella network and the query
-data.  One significant observation is that the geographic locations of clients
-do not have a correlation with the number of query requests they receive. This
-is an obvious result of the topology mismatch problem caused by the overlay
-structure of the Gnutella network. Gnutella traffic is observed to be bursty
+\cite{Markatos02} analyses Gnutella network traffic traces and by concluding there is locality among query requests, proposes a caching algorithm that tries to exploit these findings .
+The analysis of the trace data reveals other important facts about the structure of the Gnutella network and the query data.
+One significant observation is that the geographic locations of clients do not have a correlation with the number of query requests they receive.
+This is an obvious result of the topology mismatch problem caused by the overlay structure of the Gnutella network.
+Gnutella's traffic is observed to be bursty
 both for query requests and query responses, even in longer intervals, and nine
@@
 applying an approach similar to web caching does not fit well with P2P systems.
-The caches in P2P systems not only have to consider the query string, but also
+Caches in P2P systems not only have to consider the query string, but also
 the TTL value, the source of the query, and the time of the query as well.  In
 general, even though optimum caching is hard to achieve, it is reported that
-caching improves the overall performance of the Gnutella system.
+it improves the overall performance of the Gnutella network.
 
@@
 this section are creating spanning trees using 
-connection graphs, and creating cluster of physically close nodes. 
+connection graphs, and creating clusters of physically close nodes. 
 %The brief description of each category is presented below:
 Spanning tree based approaches construct rich graphs based on the network
-connections and build minimum spanning trees on the graphs. Even though the
-spanning tree provides efficient query performance, the construction
+connections and build minimum spanning trees on these graphs. Even though spanning trees provides efficient query performance, the construction
 and the update costs of the spanning tree, especially in dynamic environments
 with many nodes joining and leaving the network, cause large traffic
-overhead to the system \cite{chu_esm_2000,chu_esm_2002}.
-
-The cluster based approaches on the other hand select to link physically closer
-nodes with each other, therefore shrink the search scope significantly. However,
-commonly used methods for determining the correct physical positions of nodes
-over the internet does not always return reliable results, therefore mapping
-accuracy is not always guaranteed.
-
-Below, the details of the algorithms that use the spanning tree based or cluster
-based topology optimization methods are presented.
+overhead to the corresponding network \cite{chu_esm_2000,chu_esm_2002}.
+The cluster based approaches on the other hand select to link physically closer nodes with each other, therefore significantly shrinking the search scope of the system.
+Added to that,
+commonly used methods for correct determination of physical node positions across Internet do not always return reliable results, therefore mapping
+accuracy is not guaranteed. Bellow, the details of the algorithms that use the spanning tree based or cluster based topology optimization methods are presented.
 
@@
 
-Narada \cite{chu_esm_2000,chu_esm_2001,chu_esm_2002} is a generic protocol to
-design self adapting overlay networks, that can achieve application layer
-multicast without requiring IP multicast at the network layer. Although Narada
-is not originally designed as a peer-to-peer file sharing application protocol,
-it has become one of the pioneering works in the overlay networks area.  Even
-though the design of the IP multicast protocol is finalized a while ago, and
-some of the routers on the internet today already support this protocol, the IP
-multicast idea did not take off as anticipated. One of the main reasons for this
-delay is that the IP multicast violates the stateless design of the current
-internet. Although the multicast is not widely used, and the future of the
-protocol is not bright, there are multiple applications that need this kind of
-service, such as IP TVs, P2P networks, streaming services, etc. The application
-layer multicast is then a valid alternative for these applications.  The authors
-claim that, although Narada is not as efficient as IP multicast, still it
-provides reasonable performance, without the additional cost of implementing
-multicast services on the current internet infrastructure. The main reason for
-the inefficiency is due to the Topology mismatch problem.  Narada tries to solve
-this problem by building a richer connected graph\footnote{ENarada
-\cite{li_enarada_2008} used Gossip protocol for the construction}, called a
-mesh, and building per source minimum spanning trees.  Narada also maintains the
-graph and the trees dynamically updated on node joins and leaves. Moreover, Narada tries
-to optimize the physical link stress, the overall resource usage and the
-relative delay among end systems. The main limitation of Narada is that although
-it works reasonably well for small groups, it does not scale well for larger
-networks, therefore it is not suitable for P2P file sharing applications which
-can easily scale to millions of nodes.
+Narada \cite{chu_esm_2000,chu_esm_2001,chu_esm_2002} is a generic protocol for creating self-adapting overlay networks, that can achieve application-layer multicast communication without requiring IP multicast infrastructure at the network layer.
+IP multicast is the term reffering to the method of sending IP datagrams to a group of receivers in a single transmission.
+Even though the method is available some years now, IP multicast has not taken off as anticipated since many believe that it violates the stateless design of the current Internet.
+For this reason it is widely deployed in contained environments like enterprises (i.e.
+IPTV applications - distance learning, video conferencing), commercial stock exchanges, and multimedia content delivery networks.
+Narada tries to break these constraints.
+Although it was not originally designed as a P2P file sharing protocol, it has become one of the pioneering works in the overlay networks area.
+The authors claim that, although Narada is not as efficient as IP multicast, still it provides reasonable performance, without the additional cost of implementing multicast services on the current Internet infrastructure.
+Additionally, authors realised the inefficiency caused by the topology mismatch problem and tried to encounter it by building a richer connected graph, called a mesh, and building per source minimum spanning trees \footnote{ENarada \cite{li_enarada_2008} used Gossip protocol for the construction}.
+They also keep the
+graph and the trees dynamically updated while nodes continue to join and leave the network.
+Moreover, the protocol tries to ease the physical link stress, the overall resource usage and the relative delay among end systems.
+Unfortunately, the most imporpant limitation of Narada is that although it works reasonably well for small groups, it does not scale well for larger networks, therefore it is not suitable for, potentially very large, P2P file sharing application networks.
 
@@
 
-\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{liu_auto_2003} is one
-of the first attempts, along with Narada, from the P2P research community to
-address the topology mismatch problem. AOTO is a distributed algorithm
-that uses \emph{Selective Flooding} and \emph{Active Topology} to optimize the
-overlay network topology. The \emph{Selective Flooding} algorithm creates a minimum
-spanning tree over the overlay and uses this tree instead of
-flooding the whole network, without shrinking the search scope. 
-In \emph{Active Topology}, the physical locations of nodes are estimated based
-on the network delay among them and physically close nodes are connected as
-neighbours to revise the overlay topology as close as possible to the physical
-network topology. Overall, AOTO reports a $55\%$ performance improvement in
-terms of the traffic load. In \emph{Adaptive Connection Establishment (ACE)}
-\cite{liu_ace_2004}, the authors extend the idea by introducing optimizations
-based on the depths of the minimum spanning trees. As the network delay is not always a reliable estimation
-method to detect physical locations of the nodes, the algorithm still suffers
-from the discrepancies caused by mislocated nodes.
+\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{liu_auto_2003} is one of the first attempts, along with Narada, from the P2P research community to address the topology mismatch problem.
+AOTO is a distributed algorithm that uses \emph{Selective Flooding} and \emph{Active Topology} to optimize the overlay network topology.
+The \emph{Selective Flooding} algorithm creates a minimum spanning tree over the overlay and uses this tree instead of flooding the whole network, without shrinking the search scope.
+In \emph{Active Topology}, the physical locations of nodes are estimated based on the network delay among them and physically close nodes are connected as neighbours to revise the overlay topology as close as possible to the physical network topology.
+Overall, AOTO reports a $55\%$ performance improvement in terms of the traffic load.
+In \emph{Adaptive Connection Establishment (ACE)} \cite{liu_ace_2004}, the authors extend the idea by introducing optimizations based on the depths of the minimum spanning trees.
+But since the network delay is not always a reliable estimation method to detect the physical location of peers, the algorithm still suffers from the discrepancies caused by mislocated nodes.
 
@@
 
-Peer-to-peer algorithms designed without considering the topology mismatch
-problem generates unnecessary heavy traffic on the internet infrastructure, due
-to the popular use of these P2P programs for file sharing. \emph{Location-aware
-topology matching (LTM)} \cite{liu_ltm_2004} proposes a method to
-optimize the overlay structure of the P2P network based on the physical
-topology. In doing so, LTM issues a special message called
-\textit{TTL2-detector} to detect the latency of $N^2$ neighbours around each
-node. The latency information gathered is later used to evolve the overlay
-network into a more efficient one, without reducing the search scope. Each node
-examines the latency information with the neighbours and low productive
-connections are dropped and replaced by more efficient ones, thus reducing the
-latency on the overall network, and eliminating some potential inefficiencies,
-such as unnecessary crossing of messages over autonomous system (AS) boundaries, etc. Although the
-LTM approach improves the overall efficiency of the P2P network, since it uses
-the latency to detect close by nodes, does not use the real physical topology
-information, therefore does not offer a guaranteed bound to the
-topology mismatch problem.
+\emph{Location-aware Topology Matching (LTM)} \cite{liu_ltm_2004} proposes a method to optimize the overlay structure of the P2P network based on the physical topology.
+To achieve this, the algorithm issues a special message called \textit{TTL2-detector} to detect the latency of two-hop neighbours ($N^2$) around each node.
+The latency information gathered is later used to evolve the overlay network into a more efficient one, without reducing the search scope.
+Each node examines the latency information with the neighbours and low productive connections are dropped and replaced by more efficient ones, thus reducing the latency on the overall network, and eliminating some potential inefficiencies, such as unnecessary crossing of messages over autonomous system (AS) boundaries, etc.
+Although LTM improves the overall efficiency of the P2P network, it does not actually use real physical topology information, just latency metrics, therefore it does not offer a guaranteed safety net for the topology mismatch problem.
 
@@
 
-\emph{Scalable Bipartite Overlay (SBO)}
-\cite{liu_bipartite_IPDPS,liu_bipartite_2007} improves the overhead of creating
-and maintaining a minimum spanning tree cost by randomly dividing the nodes into
-two groups, red and white, and assigning different tasks to different groups.
-The white peers measure distances to neighbours by using the network delay as
-the metric and reports the results to red peers. The red peers, equipped with the
-information of all two hop ($N^2$) neighbours, creates a minimum spanning tree
+\emph{Scalable Bipartite Overlay (SBO)} \cite{liu_bipartite_IPDPS,liu_bipartite_2007} reduces the overhead of creating and maintaining a minimum spanning tree cost by randomly dividing the nodes into two groups, namely red and white and assigning different tasks to different groups.
+The white peers measure distances to neighbours by using the network delay as the metric and reports the results to red peers.
+The red peers, equipped with the
+information of all two-hop neighbours ($N^2$), creates a minimum spanning tree
 of these neighbours and assigns efficient forwarding paths. The white peers can
-further optimize their positions within the overlay if need be.
+further optimize their positions within the overlay if this is required.
 
@@
 
-\cite{qiu_prop_2007} introduces two protocols called Peer-exchange Routing
-Optimization Protocols (PROP) to adjust the neighbourhood graph of the overlay
-network in order to reduce the overall link latency of the network. The PROP
-algorithms are based on the exchange of neighbours among peers, which is
-triggered by the mutual benefit of both peers to reduce the network delay.  In
-PROP-G (Generic), peers exchange all their neighbours with another peer, while
-in PROP-O (Optimized) only selected number of neighbours are exchanged among
-peers. The PROP-G is a generic protocol that guarantees the connectivity of the
-overlay graph during exchanges, therefore can be applied to both unstructured
-and structured overlay networks.
+\cite{qiu_prop_2007} introduces two protocols called \emph{Peer-exchange Routing
+Optimization Protocols (PROP)} to adjust the neighbourhood graph of an overlay
+network in order to reduce its overall link latency.
+PROP algorithms are based on the exchange of neighbours among peers, driven by the mutual benefit of the participants to reduce the network delay among their connections.
+In PROP-G (Generic), a peer can exchange all of its neighbours with another peer, while in PROP-O (Optimized) two peers exchange a selected subset of their neighbours.
+PROP-G is a generic protocol that guarantees the connectivity of the overlay graph during exchanges, therefore can be applied to both unstructured and structured overlay networks.
 
@@
 
-T2MC \cite{shi_t2mc_2008} uses traceroute logs to detect the AS boundaries and
-cluster close by peers with each other to reduce the redundant multiple message
-passes between the AS boundaries. T2MC uses a customized k-mean classification
+T2MC \cite{shi_t2mc_2008} uses traceroute logs to detect different autonomous systems and
+cluster close by peers, thus reducing redundant traffic that crosses autonomous system boundaries. T2MC uses a customized k-mean classification
 algorithm with $k=2$ to perform the classification, and exploits the stable
-structure of the internet routers to guide clustering.
+structure of the Internet routers to guide clustering.
 Even though traceroute provides detailed information about the network
-structure, use of traceroute creates overhead to the overall network structure.
-For this same reason, it is not unusual for network administrators to disable
-this support on their network routers, which may affect the performance of the
-T2MC algorithm.
+structure, its excessive use creates additional overhead that stresses further the overall network infrastructure.
+This forces network administrators to disable its support from the routers they manage, thus jeopardizing the effectiveness of the T2MC algorithm.
 
@@
 
-\paragraph*{\textbf{Unnamed Unstructured!!\\}}
+\paragraph*{\textbf{Resolving the Topology Mismatch Problem in Unstructured Peer-to-Peer Networks (?)}}
 % TODO: to be reviewed
 
-\paragraph{}
 In \cite{hsiao_redblue_2009}, Hsiao et al, claim to construct topology-aware unstructured overlays that \emph{guarantee} performance qualities in terms of
@@
   \item the expected communication latency among any two overlay peers regardless of the network size, and
-  \item the broadcasting scope of each participating peer.
+  \item the broadcasting scope of each participating peer
 \end{inparaenum}
-
+.
 The algorithm constructs an undirected graph $G = \left( V, E \right)$ comprised by two subgraphs.
@@
 In contrast, $G^{\left( blue \right)} = \left( V^{\left( blue \right)}, E^{\left( blue \right)} \right)$, contains those vertices of $G$ that have free edges to link to other nodes and because these are fully utilized, the following also stands $E = E^{\left( red \right)} \cup E^{\left( blue \right)}$.
-
 A joining peer $u$, partitions its neighbours into two subsets, the $B_u^{\left( red \right)}$ and $B_u^{\left( blue \right)}$.
@@
 \paragraph*{\bf Distributed Domain Name Order}
-
-%{\sethlcolor{yellow}\hl{HA:  In paper/introduction nice advantages of unstructured networks:
-%
-%Unstructured P2P networks o?er a number of
-%important advantages: (i) An unstructured network
-%imposes very small demands on individual nodes,
-%and more speci?cally it allows nodes to join or leave
-%the network without signi?cantly a?ecting the sys-
-%tem performance. (ii) Unstructured networks are
-%appropriate for content-based retrieval (e.g., key-
-%word searches) as opposed to object identi?er loca-
-%tion of structured overlays. (iii) Finally unstructured
-%networks can easily accommodate nodes of vary-
-%ing power. Consequently, they scale to very large
-%sizes and they o?er more robust performance in
-%the presence of node failures and connection
-%unreliability.
-%}}
 
@@
 Clustering}, is based on the view that nodes with similar distances to a
-set of predefined well-known landmark nodes are pretty likely of being close to
-each other. But this approach has its weaknesses as well, such as the fact that
+set of predefined well-known landmark nodes are pretty likely being close to
+each other. But this approach has its weaknesses itself, like the fact that
 is a rather coarse grained approximation, therefore not particularly well suited
@@
 Landmark Binning \cite{ratnasamy_binning_2002} is an approach to partition close
-by nodes into bins based on their distance to well known anchor nodes within the
-internet. In order to detect locality, nodes use network latency (round trip
+by nodes into bins based on their distance to well known anchor nodes across the Internet. In order to detect locality, nodes use network latency (round trip
 time) as a measurement technique. The network latency, even though not always
-accurate, is selected in this work because its non-intrusive, transparent and
-easy to apply structure. In order for the binning to work, authors assume that
-there are a few well known anchor servers with known physical locations on the
-internet. The authors estimate that around 12 servers will do the job. The nodes
-measure their distance to all these servers and record the latency values for
-each. Later, the tuple of these latency values represents the bin of each node.
-Nodes with similar latency values are assumed to be close to each other, or at
-least within the same AS. The landmark binning is designed as a generic model
-that can work both on structured and unstructured P2P networks, and
-\cite{ratnasamy_binning_2002} has the detailed description of the algorithm for
-both topologies. As the main operation of the algorithm is independent from the
-underlying model, and does not change for structured and unstructured networks,
-we classify and give the details of the algorithm in this decentralized
-unstructured section, and omit the duplicate work in the decentralized
-structured section. The landmark binning is also proposed as a good candidate to
-work on content distribution networks. The major disadvantage of landmark
-binning is to install and maintain landmark servers on different AS, all over
-the world.  A typical P2P network usually has a couple of million nodes
-connected at any time, which introduces possible scalability problems in terms
-of the landmark servers.  The authors claim that latency estimation does not use
-much network resources and in order to avoid the possible scalability problems
-the authors propose to replace single landmark servers with clusters of servers
+accurate, is selected in this work due to its non-intrusive, transparent and
+easy to apply behaviour.
+In order for the binning to work, few well known anchor servers with known physical locations on the Internet.
+Authors estimate that a number of 12 such servers can prove sufficient for this.
+The nodes measure their distance to all these servers and record the latency values for each.
+Later, the tuple of these latency values represents the bin of each node.
+Nodes with similar latency values are considered close to each other, or at least within the same autonomous system.
+The operation of the method is independent of the model incorporated by the overlay network and thus it can be applied with no significant changes to both structured and unstructured P2P systems.
+\cite{ratnasamy_binning_2002} has the detailed description of such algorithm for both architectures.
+Landmark binning is also considered as a good candidate for content distribution networks in particular.
+The major disadvantage of landmark
+binning is to install and maintain landmark servers on different autonomous systems, all over the world.  A typical P2P network usually has a couple of million nodes
+connected at any given time, thus rendering landmark servers an important parameter when designing for large scales.  The authors claim that latency estimation does not drain
+much from the network resources but in case scalability problems arise, they claim the solution is replacing single landmark servers with clusters of servers
 within the same physical area. However, this approach does not reduce the
-possible high network traffic flow through these landmark servers. One other
-possible problem is the incorrect binning caused by inaccuracy in delay
-measurements methods, as the network latency is not a proven accurate method for
-location estimation.
+possibility of excessive network traffic flow through these landmark servers. One other
+possible problem is the incorrect binning caused by the use of inaccurate methods for delay measurement, since tracking down, for example, network latencies is far from proven to be safe for location estimation.
+
+% TODO: MAYBE THANCS deserves some more attention here
 
 \paragraph*{\bf mOverlay}
-mOverlay \cite{zhang_moverlay_2004} addresses the scalability problems that
-affect static landmark servers by introducing the dynamic landmarks. The
-nodes are divided into groups based on their distances to these groups and the
-neighbour nodes in these groups behave as dynamic landmarks. The mOverlay
-protocol is composed of two main parts, initially adding a new node to the
-overlay by selecting the closest group, and maintaining the overlay afterwards.
-Finding the correct closest group is the most important part of the overlay
-construction. The authors also formally prove that any new node can reach its
-group by performing at most $O(logN)$ communications within the network.
+
+mOverlay \cite{zhang_moverlay_2004} tries to addresses the scalability issues that might be imposed to networks when using static landmark servers by introducing the dynamic landmarks.
+The authors introduce the notion of a group which is a set of  peer considered close to each other with respect to any  position in the underlying network, proximity being some  user-defined cost metric (i.e.
+round-trip-time).
+mOverlay, tries to recreate small-world-like properties to the overlay network by creating a a two-level hierarchical structure, where on the top level we have connections between groups while on the bottom we have connections between peers inside groups.
+Finding the correct closest group is the most important part  of the overlay construction.
+Nodes are grouped based on their  distance to the groups already in the network, rendering the  later the (dynamic) landmarks for the process.
+The authors formally prove that any new node can reach its group by exchanging at most $O(logN)$ messages within the network.
+Finally mOverlay also considers a second important function in its protocol which is the constant maintenance of the overlay.
 
@@
 
-\subsection{Discussion on Unstructured Decentralized Algorithms}
-
-Earlier in section \ref{sec:unstructured} we presented the state of the art unstructured
-decentralized P2P algorithms in four different categories based on their
-architecture; broadcast based, cache based, overlay optimization based, and
-landmark based. In this section, we present a final discussion of all the
-methods including the advantages, disadvantages,
-and novelties presented in order to tackle the topology mismatch problem.
-Table \ref{fig:unstruct_compare_table} presents an overview of all the
-unstructured decentralized algorithms described earlier in this section.
-
-The broadcast based approaches in general propose intelligent neighbour
-selection algorithms in order to replace the inefficient blind
-flooding algorithm of Gnutella.
-Instead of forwarding the queries to all the neighbours, intelligently selecting
-neighbours with high probability of answering the query, or neighbours that have
-some specific features such as high bandwidth, high capacity, or low latency in
-practice reduces the aggregate resource usage of the P2P network and improves
-the overall performance of the system. However, since the broadcast based algorithms
-proposed do not consider the underlying physical topology when selecting the
-neighbours, or optimizing the overlay structure, the algorithms do not offer a
-solution to the topology mismatch problem. Although in practice the algorithms
-work, none can guarantee that a single link is not used by more than necessary,
-or the created topology resembles the underlying physical topology. Broadcast
-based methods can be used in conjunction with other approaches to improve the
-quality of the P2P systems.
-
-The success of caching, as a well studied method in the client server internet
-applications, also made it a good candidate for the peer-to-peer networks. Even
+\subsection{Discussion on the Algrithms for Unstructured Architectures}
+
+In this section we presented the state of the art for the algorithms applied to unstructured P2P systems.
+We categorised it based on their philosophy for alleviating the topology mismatch problem into four major categories namely
+\begin{inparaenum}[\itshape i\upshape)]
+  \item broadcast-based,
+  \item cache-based,
+  \item overlay optimisation-based and
+  \item landmark-based approaches
+\end{inparaenum}
+.
+Here we present a overall discussion on the above categories trying to point out the advantages, disadvantages and novelties introduced by each and by all of them.
+A quick overview can be seen in Table \ref{fig:unstruct_compare_table}.
+
+Broadcast-based approaches in general propose intelligent neighbour
+selection in order to replace the inefficient blind flooding first used in the Gnutella protocol.
+Instead of forwarding the queries to all their outgoing links they dictate the intelligent selection of neighbours with either higher probability to answer a query than others or neighbours  that expose specific features such as high bandwidth, high capacity, or low latency.
+This, in practice reduces the aggregate resource usage of the P2P network and improves the overall performance of the system.
+On the other hand, the algorithms do not offer a real solution to the topology mismatch problem.
+Thus, none can guarantee that overlay and underlying topologies are aligned with each other let alone quantify the mismatch and try to alleviate.
+Broadcast-based methods can be used in conjunction with other approaches to improve the quality of the P2P systems.
+
+The success of caching, as a well studied method in client-server Internet
+applications, also made it a good candidate for the P2P networks. Even
 though caching improves the performance, and reduces the overall resource usage
-of P2P systems, the design of caches is non trivial compared to the web caching
-systems. Due to the unique structure of the P2P overlays, each node being a
-server and a client simultaneously, two important problems has to be solved when
-designing caches. First, the lifetime of a query is short, as the nodes join and
+of P2P systems, the design of caches in this context is non-trivial compared to the web-based caching.
+Due to the unique, inherent characteristic of all P2P systems that, each node can act as both a server and a client, two important issues have to be considered when designing caches.
+First, the lifetime of a query is short, as the nodes join and
 leave frequently. Second, the result of a single query string is not always the
-same, and changes based on the TTL value, source of the query, the lifetime of
-the query, etc. So, in order to develop a successful caching system in P2P
-systems, these parameters also have to be considered. Even though the state of
-the art P2P algorithms using caching methods reduce the resource usage of the
-network, since the physical network topology is not considered by any algorithm,
-the proposed systems do not provide direct ways to overcome the topology
-mismatch problem caused by the overlay network.
-
-The overlay optimization based algorithms can be analyzed in
-two main topics, the minimum spanning tree based ones, and the clustering based ones.
-Even though the standardization of the IP multicast protocol is completed a
-while ago, and some vendors deployed routers supporting the protocol, in
-practice the standard is not widely embraced as predicted, mostly because the
-protocol violates the stateless design of the internet routers. The research
-community, as an alternative proposed application layer multicast frameworks,
-such as Narada, to simulate the same multicast protocol without requiring IP
-multicast services. The application layer multicast, however, is not as
-efficient as IP multicast, and suffers from the topology mismatch problem.
-Narada, and its followers (AOTO, LTM, SBO) try to solve the topology mismatch
-problem by building a richer connected graph and forming minimum spanning trees
-over this graph that can efficiently route messages among peers. The application
-level multicast protocol, initiated by Narada, is a generic protocol that can be
-applied to P2P file sharing, as well as content distribution networks. The
-original Narada algorithm is designed for small groups therefore it is not
-scalable. The followers try to solve the scalability problem by introducing
-various methods including forming minimum spanning trees for each nodes' $N^2$
-neighbours, partitioning the graph into two random groups where each group is
+same, as it is dependent on the source of the query, the TTL value set for the messages, the current interconnection of peers and the high volatility of the environment.
+So, in order to develop a successful caching system for a P2P architecture, these parameters also have to be considered.
+Even though the state of the art P2P algorithms using caching methods reduce the resource usage of the network, they also fail to consistently address the probelem of the topology mismatch.
+
+Overlay-optimisation-based  can further be devided into minimum-spanning-tree-based and clustering-based.
+Researchers initially started considering alternatives to IP multicast for the application layer, which led Narada, that simulates the same one-to-many communication approach requiring IP multicast services.
+Application layer multicast, however, is not as efficient as the one enabled by IP and still struggles to ease the difficulties imposed by the topology mismatch problem.
+Narada, and its followers (AOTO, LTM, SBO) try to solve the topology mismatch problem by building a richer connected graph and forming minimum spanning trees over this graph that can efficiently route messages among peers.
+Application level multicast protocols, pioneered by Narada, are generic protocols that can be applied to both P2P file sharing systems, as well as content distribution networks.
+The original Narada algorithm is designed for small groups therefore it is not scalable.
+Its followers try to solve the scalability problem by introducing
+various methods including forming minimum spanning trees for the two-hop neighbours ($N^2$) of each node, partitioning the graph into two random groups where each group is
 responsible of different tasks, or performing local optimizations dynamically on
-the overlay graph. The advantage of the minimum spanning tree based approaches
-is that they maintain the connectivity on the network in an efficient way, while
-still not shrinking the search space. However, building and maintaining a
-minimum spanning tree creates a huge overhead for the network. 
-One other popular approach used in overlay optimization is the cluster based
-approaches. In cluster based approach, nodes use various methods to detect
+the overlay graph. The advantage of building minimum spanning trees is that they maintain the connectivity on the network in an efficient manner, while
+still preserving the overall search scope. However, building and maintaining a
+minimum spanning tree creates additional overhead for the network. 
+The alternative method used in overlay optimization is the cluster based
+approache. According to that, nodes use various methods to detect
 physically close nodes to form clusters. T2MC uses traceroute logs and DDNO uses
@@
 used to form clusters directly affect the success rate of the algorithm. Traceroute for
-example is a heavy weight protocol to frequently use on the network, and most of
-the network vendors for this reason do not allow traceroute calls. The major problem with
+example is a heavy weight utility to be frequently used on the network, let alone the fact that many network equipment vendors do not allow traceroute calls. The major problem with
 cluster based approaches is that the limited connectivity within the local
@@
 
-In landmark based proximity method, nodes use network delay (round trip
-time) as a distance measurement method to position themselves based on distance
-measurements to apriori known servers on the internet. The servers behave as
-landmark points, and nodes use an estimation method to discover their positions.
-The landmark servers are either used by nodes to directly calculate their
-positions, or indirectly used to cluster nodes into groups or bins based on
-their estimated distances to well known landmark servers, due to the assumption
-that nodes with similar distances to a set of landmarks are physically close to
-each other over the network. The landmark based protocol has two important
-drawbacks. Firstly the network delay is not a reliable distance estimation
-method. For example, based on the load on the network the delay to certain
-nodes or networks can change from time to time, which will eventually affect the
-distance measurements and wrong measurements will lead to wrong estimated positions for the
-nodes, or incorrect and non optimal clusterings of the nodes. The second drawback of
-using landmark servers is the cost of installing and maintaining many landmark servers over the internet
-for various AS domains. As popular P2P file sharing applications usually have
-millions of peers connected at any time, it is not false to assume that the hardware and the network costs of
-maintaining these landmark servers will be quite high. 
-A possible solution to the scalability problem of the static landmark servers is to use ordinary nodes as dynamic
-landmarks once they estimate their own positions. Even though this approach scales
-much better than static landmark servers, still the measurement accuracy problem affects
-the overall performance of the system.
+In landmark-based algorithms, nodes use network delay (round trip
+time) as a distance measurement method to position themselves with respect to ``a priori'' known servers (the landmark servers) on the Internet.
+Landmark servers are used by nodes to estimate their positions based on the intutive assumption that nodes with similar distances to a set of landmarks, are physically close to each other, as well, over the network.
+The landmark based protocol has two important drawbacks.
+Firstly the network delay is not a reliable distance estimation method.
+For example, based on the load on the network the delay to certain nodes or networks can change from time to time, which will eventually affect the distance measurements and wrong measurements will lead to wrong estimated positions for the nodes, or incorrect and non optimal clusterings of the nodes.
+The second drawback of using landmark servers is the cost of installing and maintaining this kind of infrastructure across the whole Internet and for all the different autonomous system domains.
+As popular P2P file sharing applications usually have millions of peers connected at any time, it is not false to assume that the network costs of maintaining these landmark servers will be quite high.
+A possible solution to the scalability problem of the static landmark servers is to use ordinary nodes as dynamic landmarks once they estimate their own positions.
+Even though this approach scales much better than static landmark servers, still the measurement accuracy problem affects the overall performance of the system.
 
@@
 
-In this section, the algorithms proposed for the structured decentralized
-architectures are analysed and categorized based on their use of the proximity
-information to optimize the underlying structure. In structured P2P algorithms,
-the construction of the routing tables among nodes determines the efficiency of
-the algorithms.  Therefore, routing tables that represent the underlying
-physical structure well can achieve much better performance. For this reason,
-the proposed methods for topology mismatch problem generally use different
-levels of proximity information to optimize the routing table close to the
-physical network. We analyze the structured decentralized algorithms based on
-the categories presented in
-\cite{castro_proximitydht_2002,castro_topawareroute_2002,ratnasamy_openq_2002}.
-In this section, the details of the methodologies used and the algorithms for
-each methodology are discussed in details.
+This section focuses on the structured P2P architectures.
+The following paragraphs contain a thorough discussion and analysis of the algorithms proposed in the literature for dealing with the topology mismatch problem in such architectures.
+In structured P2P systems, the construction of the routing tables in every participating node determines the efficiency of message forwarding.
+Therefore, routing tables that represent the underlying physical structure well can achieve increased performance.
+For this reason, the proposed methods for topology mismatch problem basically use different levels of proximity information to optimize them towards this direction.
+The algorithms are here categorised based the previous work done by \cite{castro_proximitydht_2002,castro_topawareroute_2002,ratnasamy_openq_2002}.
 
@@
 
-Geographic layout is used as a method to optimize the routing table of the
-structured algorithms to represent the physical network topology as close as
-possible. In this method, physically close nodes are detected, and they are
-positioned closely in the overlay structure. In order for this method to work,
-there should be means to detect physical proximity of internet nodes. One
-popular method is to use the well known internet servers, or servers dedicated
-for this sole purpose, as landmark nodes, similarly as discussed previously in
-Section \ref{sec:landmark} (\textit{Landmark Based Proximity}).  The inaccuracy in
-the positioning using landmarks, and the cost of deploying and maintaining these
-landmarks are the major disadvantages of these systems. Even though managing the
-overlay structure based on the geographic layout of the nodes improves the query
-efficiency of the system, on the other hand, it tends to create hotspots, and
-the needed failure resilience is undermined by the fact that close by nodes are
-more likely to suffer collective failures.
+Inside geographic layout category fall those kind of methods which practically position physically close by nodes together in the application space as well.
+This is done by adjusting and maintaining the routing tables of all involved peers accordingly by exploiting proximity information that describe the overall geographic positioning of the peer nodes.
+One popular method for detecting and constructing a geographic view of the underlying network is the use of well established servers dedicated to landmarking the physical space, pretty much as discussed in Section \ref{sec:landmark} of landmark based proximity in the unstructured P2P networks.
+The same disadvantages of difficult and inaccurate positioning and the cost of deploying and maintaining these landmarks apply in this case as well.
+% TODO: maybe a discussion sentence bellow
+Even though managing the overlay structure based on the geographic layout of the nodes improves the query efficiency of the system, on the other hand, it tends to create hotspots, and the needed failure resilience is undermined by the fact that close by nodes are more likely to suffer collective failures.
 
@@
 nodes in order to help them find the best way to route their messages.
-\textit{Global Soft-State} operates in two main stages, generation and using of
+\textit{Global Soft-State} operates in two main stages, generation and use of
 the proximity information. For generating the proximity information a hybrid
@@
 order to select a number of potential nearest neighbour candidates and then
-refine the selection by incorporating an RTT scheme to ultimately choose the
+refine the selection by incorporating a round-trip-time-based scheme to ultimately choose the
 closest node. For using the proximity information, the algorithm chooses a
 different path from the classic gossiping approaches for constructing and
-maintaining the overlay. It is based on landmark clustering based strategic
+maintaining the overlay. It is based on landmark clustering for strategic
 placement of proximity information on the overlay enabling any node to access
@@
 overlay structure after a node successfully joins, in order to adapt its state to
-any occurrence of condition change. Although this approach greatly reduces the
+any condition change. Although this approach greatly reduces the
 routing latency to far nodes, it is unable to dynamically identify nodes that are
@@
 \textit{Mithos} \cite{waldvogel_mythos_2003} is a P2P protocol which
-incorporates a directed incremental probing to find near optimal node placement,
-and classified by its authors as an integration of geographic layout
+incorporates directed incremental probing to find near optimal node placement,
+and is classified by its authors as an integration of geographic layout
 and proximity routing overlay optimization methods.
@@
 the geographical layout approach  and constructs a geographical layout in a
-self-organizing and efficient fashion, by estimating the round trip time (RTT)
-to a small number of nodes in the overlay network in order to make them roughly
+self-organizing and efficient fashion, by estimating the round-trip-times to a small number of nodes in the overlay network in order to make them roughly
 aware of their physical distances among them.
@@
 longest-prefix IP matching scheme. At each forwarding hop, any message travels
-up the tree until the first common ancestor of source and destination node is
+up the tree until the first common ancestor of both source and destination nodes is
 reached and then starts descending to arrive to its target. During tree
@@
 balance. Caching enables a node to forward a message to a better longest-prefix
-match than that of its direct ancestor making a large, quicker and more cost
+match than that of its direct ancestor making a larger, quicker and more cost
 effective step through the overlay and toward the destination. To improve
@@
 initializes its routing table (with normal and caching entries) as it traverses
-the overlay in search for its parents node.  During a graceful departure, the
+the overlay in search for its parent node.  During a graceful departure, the
 referred node checks for children in the overlay. If it does not have any, it
 simply notifies its parent and leaves. If it has, it selects the child node with
-the lowest RTT to it in order to take its place so that the locality property is
+the lowest round-trip-time to it in order to take its place so that the locality property is
 preserved.
@@
 towards the destination and choosing the closest entry in the routing table, in
-terms of network proximity. Thus, it is relatively less effective than
+terms of network proximity.
+% TODO: looks like discussion material
+Thus, it is relatively less effective than
 geographical layout when applied to CAN(-like) implementations. Moreover, the
@@
 knowledge. The proximity information used in this method is different than the
-landmark based systems described in the \textit{Geographic Layout} section, as TTL values between
+landmark based systems described in the \textit{Geographic Layout} section, as time-to-live values between
 nodes, or directly node ID prefixes are used to detect proximity.  Tapestry,
@@
 Chord-DHT-PNS \cite{hancong_pnsbased_2006} implements proximity neighbour
-selection on top of the Chord DHT. The main purpose is to use proximity
+selection on top of the Chord DHT. The main goal is to use proximity
 information to group physically close by nodes as neighbours in the DHT table.
@@
 \paragraph*{\bf IP-based Clustering (IPBC)}
-Proximity neighbour selection algorithms use probing and other measures
-to detect proximity. However, such methods are either are not precise enough, or
-creates overload in the network. \emph{IP-based clustering}
+Proximity neighbour selection algorithms use probing and other techniques
+to detect proximity. However, such methods are either not precise enough, or
+overload the network. \emph{IP-based clustering}
 \cite{karwaczynski_ipbc_2007} is a proximity neighbour selection based
@@
 performance/accuracy tradeoff. The {\em IP-based Clustering} generates a key and
-its prefix and stores its prefix in the DHT itself, so that any newly joining
-nodes with the same IP prefix can query the prefix and identify all the
-neighbours with the same prefix relatively easily. Nodes periodically update
-their entries in the DHT and remove their entry when they leave, or the entry is
-timed out with no further activity detected from the node.
+its prefix and stores the later in the DHT itself, so that any newly joining
+nodes with the same IP prefix can query it and identify all the
+neighbours with the same one relatively easily. Nodes periodically update
+their entries in the DHT by removing those who leave, or those who has been timed out with no further activity detected from the node.
 
@@
 
-\textit{Cone}\cite{wang_cone_2007} extends the Chord using proximity neighbour selection
+\textit{Cone}\cite{wang_cone_2007} extends Chord using proximity neighbour selection
 topology optimization algorithm. The proximity information is generated using
-landmarks and RTT based distances to landmarks.  \textit{Cone} uses a two-layered
+landmarks and round-trip-time-based distances to landmarks.  \textit{Cone} uses a two-layered
 identifier space. The first, named Chord-layer identifier, denoted as
@@
 proximity with the current node. Moreover, a third table called \emph{group}
-table maintains information about other online peers within the current node's
+maintains information about other online peers within the current node's
 group in a way that entries are now close in the ID space.
@@
 information. However, as an additional feature, \textit{SAT-Match} peers can do
-selective jumps to adjust their locations in the DHT, if it reduces the stretch
+selective jumps to adjust their positioning in the DHT, if it reduces the stretch
 of its one-hop neighbourhood. The stretch is defined as the ratio between the
@@
 the structure's integrity, thus creating an overall increase in the overlay's
-\emph{stretch}. The two nodes with the smallest RTT is then used in order to
+\emph{stretch}. The two nodes with the smallest RTT are then used in order to
 select one zone to jump in this phase. The algorithm is as follows: The source
@@
 
-\subsection{Discussion on Structured Decentralized Algorithms}
+\subsection{Discussion on the Algrithms for Structured Architectures}
+In this section we presented the state of the art for the algorithms applied to structured P2P systems.
+We categorised it based on their philosophy they incorporated in order to alleviate the topology mismatch problem into overall three categories following the previous work from \cite{castro_proximitydht_2002,castro_topawareroute_2002,ratnasamy_openq_2002} namely
+\begin{inparaenum}[\itshape i\upshape)]
+  \item geographic-layout-based,
+  \item proximity-routing-based, and
+  \item proximity neighbour selection based
+\end{inparaenum}
+.
+Following is a discussion on the above categories trying to point out the advantages, disadvantages and novelties introduced by each and by all of them.
+A quick overview can be seen in Table \ref{fig:struct_compare_table}.
 
@@
 %
-%\paragraph{General overlay operations} A set of additional protocols, are also introduced, similar to those found in traditional unstructred networks, but modified focusing on scalability and robustness.
+%\paragraph{General overlay operations} A set of additional protocols, are also introduced, similar to those found in traditional unstructured networks, but modified focusing on scalability and robustness.
 For example a protocol for \emph{group formation} is introduced that exploits the inherent characteristic of proximity, in the overlay, in order to efficiently detect the neighbouring groups of a newly formated group from the set of adjasent groups of its closest neighbour.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{received}
+Received Month Year;
+revised Month Year; accepted Month Year
+\end{received}
 

### 2014-01-21 mard-survey.tex
2012.05.18 (Hüseyin) Updating table to be more compatible with the text. Figures added/updated.

@@
 
+\begin{figure*}
+\centering
+  \includegraphics[scale=0.4]{img/algorithms/iterative_deepening}
+\caption{Iterative deepening with three levels.}
+\label{fig:iterative_deepening}
+\end{figure*}
+
+\begin{figure*}
+\centering
+  \includegraphics[scale=0.4]{img/algorithms/directed_bfs}
+  \includegraphics[scale=0.4]{img/algorithms/local_indices}
+\caption{(left) directed BFS. (right) local indices with radius size equal to $2$.}
+\label{fig:directed_bfs}
+\end{figure*}
+
 \cite{yang_improvep2psearch_2002} proposes an easy to implement and practical
@@
 approaches, namely \textit{iterative deepening}, \textit{directed BFS}, and
-\textit{local indices}. In \textit{iterative deepening}, the search is performed
+\textit{local indices}. In \textit{iterative deepening}, as shown in Figure \ref{fig:iterative_deepening},
+the search is performed
 on a BFS tree with multiple preset depths. The depth limit is iteratively
@@
 results at each iteration before deciding to quit or resume the query.  The
-\textit{directed BFS} tries to avoid this delay by forwarding the query messages
+\textit{directed BFS}, as shown in Figure \ref{fig:directed_bfs}(left), tries to avoid this delay by forwarding the query messages
 only to a selected set of neighbours, in which the selection criteria varies
 from the number of results received previously, distance in terms of hops,
-bandwidth, or the query load of the neighbour. In \textit{local indices}, each
+bandwidth, or the query load of the neighbour. In \textit{local indices}, as shown in Figure \ref{fig:directed_bfs}(right), each
 node sustains a local data index of the nodes within a radius of $r$ hops from
@@
 
+\begin{figure*}
+\centering
+  \includegraphics[scale=0.4]{img/algorithms/narada}
+  \includegraphics[scale=0.4]{img/algorithms/narada2}
+  \includegraphics[scale=0.4]{img/algorithms/narada3}
+\caption{(left) underlying network with edge costs. (middle) Peer A sends a message to the rest using regular broadcast. (right) Peer A sends a message to the rest using end system multicast.}
+\label{fig:narada}
+\end{figure*}
+
 Narada \cite{chu_esm_2000,chu_esm_2001,chu_esm_2002} is a generic protocol for creating self-adapting overlay networks, that can achieve application-layer multicast communication without requiring IP multicast infrastructure at the network layer.
@@
 The authors claim that, although Narada is not as efficient as IP multicast, still it provides reasonable performance, without the additional cost of implementing multicast services on the current Internet infrastructure.
-Additionally, authors realised the inefficiency caused by the topology mismatch problem and tried to encounter it by building a richer connected graph, called a mesh, and building per source minimum spanning trees \footnote{ENarada \cite{li_enarada_2008} used Gossip protocol for the construction}.
+Additionally, authors realised the inefficiency caused by the topology mismatch problem and tried to encounter it by building a richer connected graph, called a mesh, and building per source minimum spanning trees \footnote{ENarada \cite{li_enarada_2008} used Gossip protocol for the construction} as shown in Figure \ref{fig:narada}.
 They also keep the
@@
 
+\begin{figure*}
+\centering
+  \includegraphics[scale=0.4]{img/algorithms/landmark_binning}
+\caption{Example 2D coordinate overlay and a sample routing path from node D to (x,y).}
+\label{fig:landmark_binning}
+\end{figure*}
+
 Landmark Binning \cite{ratnasamy_binning_2002} is an approach to partition close
-by nodes into bins based on their distance to well known anchor nodes across the Internet. In order to detect locality, nodes use network latency (round trip
+by nodes into bins based on their distance to well known anchor nodes across the Internet, as shown in Figure \ref{fig:landmark_binning}.
+In order to detect locality, nodes use network latency (round trip
 time) as a measurement technique. The network latency, even though not always
@@
 %\begin{tabular}{
-\begin{landscape}
+%\begin{landscape}
 \begin{longtable}{
-|>{\columncolor[gray]{.7}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
+|>{\columncolor[gray]{.7}}m{0.2\columnwidth}
+|>{\columncolor[gray]{.9}}m{0.4\columnwidth}
+|>{\columncolor[gray]{.8}}m{0.2\columnwidth}
 |>{\columncolor[gray]{.9}}m{0.2\columnwidth}
-|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
+%|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
+%|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
+%|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
+%|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
 |}
@@
 \rowcolor[gray]{.5}
-\textbf{Algorithm} & \textbf{Arch.} & \textbf{Overlay structure} & \textbf{Base protocol} &
-\textbf{Dynamic update} & \textbf{Runtime} & \textbf{Scalability} & \textbf{cites}\\
+\textbf{Algorithm} &  \textbf{Overlay structure} & \textbf{Base protocol} &
+ \textbf{Scalability}\\
 \hline
@@
 \endlastfoot
-\textbf{Narada} & Decentralized unstructured & \textbf{Overlay optimization
+\textbf{Narada} & \textbf{Overlay optimization
 based}. Creates a mesh (richer connected graph) and builds minimum spanning
-trees on this mesh & & Yes. Supports incremental evolving & & Small and sparse
-groups & 2000  \\
-
-\hline
-\textbf{Gia} & Decentralized unstructured & \textbf{Forwarding based} Replaces
+trees on this mesh & & Small and sparse groups \\
+
+\hline
+\textbf{Gia} & \textbf{Broadcast based} Replaces
 Gnutella flooding with random walk, and introduces KaZaA style supernodes. Uses
 dynamic topology adaptation protocol &
- Gnutella & Yes & & Better than Gnutella & 974  \\
-
-\hline
-\textbf{Adaptive Overlay Topology Optimization} & Decentralized unstructured & \textbf{Overlay optimization
+ Gnutella &  Better than Gnutella  \\
+
+\hline
+\textbf{Adaptive Overlay Topology Optimization} & \textbf{Overlay optimization
 based}. Creates overlay multicast tree with Selective Flooding protocol&
-Gnutella & Yes, using Active Topology protocol & & Better than Gnutella & 36 \\
-
-\hline
-\textbf{Location-aware Topology Matching} & Decentralized unstructured &
+Gnutella &  Better than Gnutella \\
+
+\hline
+\textbf{Location-aware Topology Matching} & 
 \textbf{Overlay Optimization Based}. Uses \textit{TTL2-detector flooding}, \textit{low productive
-connection cutting}, and \textit{source peer probing}. & Gnutella & Yes & & Better than Gnutella & 204 \\
-
-\hline
-\textbf{Replication Strategies in Unstructured P2P Networks} & Decentralized unstructured &
+connection cutting}, and \textit{source peer probing}. & Gnutella &  Better than Gnutella \\
+
+\hline
+\textbf{Replication Strategies in Unstructured P2P Networks} &
 \textbf{Cache Based}. Uses uniform, proportional and square root allocation
-strategies to replicate data. & Gnutella & Yes & & Better than Gnutella & 580 \\
-
-\hline
-\textbf{Tracing a large-scale Peer to Peer System: an hour in the life of Gnutella.} & Decentralized unstructured &
-\textbf{Cache Based}. Proposes a caching algorithm based on the traces of the Gnutella traffic & Gnutella &
-Yes & & Better than Gnutella & 199 \\
-
-\hline
-\textbf{Improving search in P2P networks} & Decentralized unstructured &
-\textbf{Forwarding Based}. Uses \textit{iterative deepening}, \textit{directed
-BFS}, and \textit{local indices} to improve efficiency. & Gnutella & Yes & & Better than Gnutella & 380 \\
-
-\hline
-\textbf{Distributed Cycle Minimization Protocol} & Decentralized unstructured &
-\textbf{Forwarding based} Uses a decentralized cycle elimination protocol  &  & Yes &  &  & 9 \\
-
-\hline
-\textbf{Scalable Bipartite Overlay} & Decentralized unstructured &
+strategies to replicate data. & Gnutella &  Better than Gnutella \\
+
+\hline
+\textbf{Tracing a large-scale Peer to Peer System: an hour in the life of Gnutella.} &
+\textbf{Cache Based}. Proposes a caching algorithm based on the traces of the Gnutella traffic & Gnutella & Better than Gnutella \\
+
+\hline
+\textbf{Improving search in P2P networks} &
+\textbf{Broadcast Based}. Uses \textit{iterative deepening}, \textit{directed
+BFS}, and \textit{local indices} to improve efficiency. & Gnutella &  Better than Gnutella \\
+
+\hline
+\textbf{Distributed Cycle Minimization Protocol} & 
+\textbf{Broadcast based} Uses a decentralized cycle elimination protocol  &  &  \\
+
+\hline
+\textbf{Scalable Bipartite Overlay} & 
 \textbf{Overlay optimization based} Uses bipartite partition graph and builds
-local minimum spanning trees  & Gnutella & Yes &  & Better than Gnutella & 77 \\
-
-\hline
-\textbf{Adaptive Connection Establishment} & Decentralized unstructured &
+local minimum spanning trees  & Gnutella  & Better than Gnutella \\
+
+\hline
+\textbf{Adaptive Connection Establishment} & 
 \textbf{Overlay optimization based} Forms Neighbour Cost Tables, builds local
 minimum spanning trees and perform local optimizations & Adaptive Overlay
-Topology Optimization (AOTO), Gnutella & Yes &  & Better than Gnutella & 110 \\
-
-\hline
-\textbf{Hops Adaptive Neighbour Discovery} &  &
- &  &  &  &  & 1 \\
-
-\hline
-\textbf{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} & Decentralized unstructured &
+Topology Optimization (AOTO), Gnutella & Better than Gnutella \\
+
+\hline
+\textbf{Hops Adaptive Neighbour Discovery} &  & &  \\
+
+\hline
+\textbf{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} &
 \textbf{Overlay optimization based} Uses piggybacking to discover neighbour
-distances and selects neighbours  & Gnutella & Yes &  &  & 36 \\
-
-\hline
-\textbf{mOverlay} & Decentralized unstructured &
-\textbf{Overlay optimization based} Uses dynamic landmarks to find node locality
-& & Yes &  & Due to dynamic landmarks and grouping, more scalable than tree-based or mesh-based protocols & 123 \\
-
-\hline
-\textbf{Distributed Domain Name Order (DDNO)} & Decentralized unstructured &
+distances and selects neighbours  & Gnutella  & \\
+
+\hline
+\textbf{mOverlay} &\textbf{Landmark based proximity} Uses dynamic landmarks to find node locality
+& & Due to dynamic landmarks and grouping, more scalable than tree-based or mesh-based protocols \\
+
+\hline
+\textbf{Distributed Domain Name Order (DDNO)} &
 \textbf{Overlay optimization based} Connects half of the nodes connections to
 the nodes in the same domain and the other half to random nodes, therefore
-supports locality and topological connection  & & Yes &  & Yes, by using super
-peers & 5 \\
-
-\hline
-\textbf{Peer-exchange Routing Optimization Protocols} & Works with both decentralized
-structured and unstructured &
-\textbf{Overlay optimization based} Optimizes overlay by the exchange of
-neighbors among peers  & & Yes &  & Yes & 13 \\
-
-\hline
-\textbf{MAY OMIT- T2MC} & Decentralized unstructured &
-\textbf{Cluster based---{\sethlcolor{yellow}\hl{ IS IT ALSO PROXIMITY BASED?}}} Uses traceroute results for clustering the
-nodes  & &  &  &  & 1 \\
-
-\hline
-\textbf{Unnamed-unstructured} & Decentralized unstructured &
+supports locality and topological connection  &  & Yes, by using super
+peers \\
+
+\hline
+\textbf{Peer-exchange Routing Optimization Protocols} & \textbf{Overlay optimization based} Optimizes overlay by the exchange of
+neighbors among peers  & Can work with both decentralized structured and
+unstructured architecture & Yes \\
+
+\hline
+\textbf{MAY OMIT- T2MC} &
+\textbf{Overlay optimization based} Uses traceroute results for clustering the
+nodes  & & \\
+
+\hline
+\textbf{Unnamed-unstructured} & 
 \textbf{Overlay optimization based} Minimizes the communication delay and
-maximizes the broadcasting range & & Yes &  & Better than THANCS and mOverlay & 7\\
-
-\hline
-\textbf{Landmark Binning} & Can work with both decentralized structured and unstructured &
-\textbf{Proximity based, landmark binning} Uses network latency to partition
-nodes into bins & & Yes &  &  & 817 \\
+maximizes the broadcasting range & & Better than THANCS and mOverlay \\
+
+\hline
+\textbf{Landmark Binning} & \textbf{Landmark based proximity} Uses network latency to partition
+nodes into bins & Can work with both decentralized structured and unstructured architecture & \\
 
@@
 \end{longtable}
-\end{landscape}
+%\end{landscape}
 \end{center}
@@
 
+\begin{figure*}
+\centering
+  \includegraphics[scale=0.8]{img/algorithms/global_softstate}
+\caption{(left) Position of $p$ in the space of $3$ landmarks. (right) Position of $p$ on the map.}
+\label{fig:global_softstate}
+\end{figure*}
+
 \textit{Global Soft-State} \cite{xu_globstate_2003} builds a global map to help
@@
 scale distance probes to reveal the proximity properties of the underlying
-network to the overlay. This global view of the state is made available to all
+network to the overlay. Figure \ref{fig:global_softstate} presents the landmark space and the mapping of the position of the nodes. This global view of the state is made available to all
 nodes in order to help them find the best way to route their messages.
@@
 
+\begin{figure*}
+\centering
+  \includegraphics[scale=0.6]{img/algorithms/SAT_match}
+  \hspace{3ex}
+  \includegraphics[scale=0.6]{img/algorithms/SAT_match2}
+\caption{(left) Example physical topology of nodes. (middle) A jumps to B in order to match the logical topology to the physical. (right) State of the logical topology after the jump.}
+\label{fig:SAT_match}
+\end{figure*}
+
 \textit{SAT-Match} \cite{ren_satmatch_2004} is a protocol that specifically
@@
 selective jumps to adjust their positioning in the DHT, if it reduces the stretch
-of its one-hop neighbourhood. The stretch is defined as the ratio between the
+of its one-hop neighbourhood. Figure \ref{fig:SAT_match} demonstrates the selective jump on an example topology.
+The stretch is defined as the ratio between the
 average logical and physical latency of a link. It is reported that, due to the
@@
 \footnotesize
-\begin{landscape}
+%\begin{landscape}
 \begin{longtable}{
-|>{\columncolor[gray]{.7}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
+|>{\columncolor[gray]{.7}}m{0.2\columnwidth}
+|>{\columncolor[gray]{.9}}m{0.4\columnwidth}
+|>{\columncolor[gray]{.8}}m{0.2\columnwidth}
 |>{\columncolor[gray]{.9}}m{0.2\columnwidth}
-|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
+%|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
+%|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
+%|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
+%|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
 |}
@@
 \rowcolor[gray]{.5}
-\textbf{Algorithm} & \textbf{Arch.} & \textbf{Overlay structure} & \textbf{Base protocol} &
-\textbf{Dynamic update} & \textbf{Runtime} & \textbf{Scalability} & \textbf{cites}\\
+\textbf{Algorithm} & \textbf{Overlay structure} & \textbf{Base protocol} & \textbf{Scalability} \\
 \hline
@@
 \hline
-\textbf{Global Softstate} & decentralized structured &
-\textbf{Proximity based, landmark clustering} Uses first landmark clustering
-then measures RTTs to identify close nodes & & Yes &  &  & 195 \\
-
-\hline
-\textbf{Mithos} & decentralized structured &
-\textbf{Proximity based} Uses nodes as topology landmarks and directed
-incremental probing to optimize topology & & Yes &  & Scales well as all
-operations are local ??? & 172 \\
-
-\hline
-\textbf{Self-Adaptive Topology Matching} & decentralized structured &
-\textbf{Proximity based} Uses lightweight probing and
-selective jumps to optimize the topology & CAN & Yes &  & Better than Mithos
-& 40 \\
-
-\hline
-\textbf{Delay Aware P2P System} & &
-\textbf{} & &  &  &  & 1 \\
-
-\hline
-\textbf{VERSION OF CHORD - DHT-PNS} & decentralized structured &
-\textbf{Proximity based} Uses Proximity Neighbour Selection and the Vivaldi
-system & Chord  & Yes &  &  & 5 \\
-
-\hline
-\textbf{MAY OMIT- VERSION OF CHORD - Quasi-Chord} & decentralized structured &
-\textbf{Proximity based}  & Chord  & Yes &  &  & 0 \\
-
-\hline
-\textbf{LAPTOP} & decentralized structured &
-\textbf{Geographic layout based} Hierarchical overlay structure  &   & Yes &  & routing path length $\log{_d N}$, 
-join/leave overhead $d\log{_d N}$& 4 \\
-
-\hline
-\textbf{IP-Based Clustering} & decentralized structured &
-\textbf{Proximity based} Proximity neighbour selection based on longest common
-prefix of IP addresses &   & Yes &  &  & 1 \\
-
-\hline
-\textbf{CHOord considering Proximity on IPv6} & decentralized structured &
-\textbf{Proximity based} Uses IPv6 address format to provide proximity &  Chord
-& Yes &  & Better than Chord & 1 \\
-
-\hline
-\textbf{Proximity in Kademlia} & decentralized structured &
-\textbf{Proximity based} Applies  proximity neighbour selection (PNS) and proximity route selection (PRS)
-to Kademlia & Kademlia & Yes &  &  & 13 \\
-
-\hline
-\textbf{Cone} & decentralized structured &
-\textbf{Proximity based} Uses proximity neighbour selection (PNS) & Chord & Yes
-&  & Better than Chord & 3 \\
-
-\hline
-\textbf{Dynamo} &  & &  & 
-&  & & 3 \\
-
-\hline
-\textbf{MAY OMIT-BADLY WRITTEN-PChord} & decentralized structured &
-\textbf{Proximity based}  &  Chord
-& Yes &  & & 16 \\
-
-\hline
-\textbf{AChord} & &  &  
-&  &  & & 7 \\
-
-\hline
-\textbf{Chord6} & decentralized structured &
-\textbf{Proximity based} Uses IPv6 hierarchical address format to cluster
-topologically close nodes & Chord & Yes
-&  &  & 14 \\
+\textbf{Global Softstate} & 
+\textbf{Geographic layout} Uses first landmark clustering
+then measures RTTs to identify close nodes &  &  \\
+
+\hline
+\textbf{Mithos} &
+\textbf{Geographic layout} Uses nodes as topology landmarks and directed
+incremental probing to optimize topology & & Scales well as all
+operations are local ??? \\
+
+\hline
+\textbf{Self-Adaptive Topology Matching} &
+\textbf{Proximity neighbour selection} Uses lightweight probing and
+selective jumps to optimize the topology & CAN &  Better than Mithos \\
+
+\hline
+\textbf{Delay Aware P2P System} & \textbf{} & & \\
+
+\hline
+\textbf{VERSION OF CHORD - DHT-PNS} &
+\textbf{Proximity neighbour selection} Uses Proximity Neighbour Selection and the Vivaldi
+system & Chord  &  \\
+
+\hline
+\textbf{MAY OMIT- VERSION OF CHORD - Quasi-Chord} &
+\textbf{Proximity based}  & Chord  &  \\
+
+\hline
+\textbf{LAPTOP} & 
+\textbf{Geographic layout} Hierarchical overlay structure  &  & routing path length $\log{_d N}$, 
+join/leave overhead $d\log{_d N}$ \\
+
+\hline
+\textbf{IP-Based Clustering} & 
+\textbf{Proximity neighbour selection} Proximity neighbour selection based on longest common
+prefix of IP addresses &    &  \\
+
+\hline
+\textbf{CHOord considering Proximity on IPv6} & 
+\textbf{Proximity routing} Uses IPv6 address format to provide proximity &  Chord
+ & Better than Chord \\
+
+\hline
+\textbf{Proximity in Kademlia} & 
+\textbf{Proximity routing} Applies  proximity neighbour selection (PNS) and proximity route selection (PRS)
+to Kademlia & Kademlia &   \\
+
+\hline
+\textbf{Cone} & 
+\textbf{Proximity neighbour selection} Uses proximity neighbour selection (PNS) & Chord  & Better than Chord \\
+
+\hline
+\textbf{Dynamo} &  &  &  \\
+
+\hline
+\textbf{MAY OMIT-BADLY WRITTEN-PChord} & 
+\textbf{Proximity based}  &  Chord  & \\
+
+\hline
+\textbf{AChord} &  & & \\
+
+\hline
+\textbf{Chord6} & 
+\textbf{Proximity routing} Uses IPv6 hierarchical address format to cluster
+topologically close nodes & Chord  &  \\
 

### 2014-01-21 conclusion.tex
2012.05.29 (Vassilis) Monolithic file broken down to multiple smaller ones (per section) for better parallel contribution. Additional material on more recent algorithms for unstructured networks. Now also covering the BitTorrent protocol. Merged Huseyin's figures from previous contribution and code refactored for better referencing. Fixed bibliography bug (references appearing as ?).

--- 2013-11-10 distrproj.tex
@@
-\documentclass[a4paper,10pt]{article}
-\usepackage[utf-8]{inputenc}
+% TODO: Should be reviewed
+\section{Conclusion}
+\label{section:conclusion}
 
-\usepackage{paralist}
-
-\usepackage{graphicx}
-
-%opening
-\title{Alleviating the Topology Mismatch Problem with Location-Aware Overlays: A Survey}
-\author{
-Vassilis Moustakas\\
-Department of Informatics and Telecommunications,\\
-National and Kapodistrian University of Athens\\
-\mbox{\texttt{b.moystakas@di.uoa.gr}}
-}
-
-\begin{document}
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{abstract}
-Peer-to-peer systems is a rapidly growing application architecture.
-The overlay network abstraction on top of a best-effort infrastructure provides a powerfull tool to realize ever-wanted exotic features like anonymity, robustness, load balancing, QoS at the application layer and many more.
-Unfortunately weaknesses of early effectuations constrained applications from unleashing the full potential of the paradigm.
-One of them is the topology mismatch problem between the overlay network and the physical underlying one.
-This survey is on the recent research developments in alleviating the problem, in unstructured schemes, using location-aware overlay construction algorithms.
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-\emph{Peer-to-peer} systems gained significant attention due to their plethora of unique features in supporting file sharing among huge numbers of networked users, where each such \emph{peer} acts both as a resource provider and a resource consumer.
-Such serverless systems, proved to be able to achieve outstanding aggregate resource capacities as participants join the system\footnote{In peer-to-peer file-sharing systems there has been observed mitigation of this self-scaling property by the \emph{free-rider} problem\cite{adar_gnutellafreeriders_2000}}, without requiring additional expenditure for infrastructure.
-
-As these systems evolved, three main architectures have been emerged:
-\begin{itemize}
-  \item \emph{centralized}
-  \item \emph{decentralized structured}
-  \item \emph{decentralized unstructured}
-\end{itemize}
-
-\emph{Centralized} architectures were the first to recognize that requests for popular content need not be sent to a central server but instead could be handled by the many hosts that already posses the content.
-\emph{Napster}, for example, maintained a \emph{central index server} based on file lists provided by participating peers.
-The central index server was queried by the users and it returned pointers to the actual content.
-Thus, by centralizing search while distributing downloads, Napster achieved a high functional design that was widely acknowledged, at the time, as ``the fastest growing Internet application ever''.
-Of course, the architecture is not without flaws.
-The centralized search facility is its ``Achilles' heel'' as it can be prooven a bottleneck for the system's scalability as well as a central point of failure and vulnerable to \emph{Denial-of-Service (DoS)} attacks.
-
-This centralized approach, has been eventually, replaced by \emph{decentralized}, that distributed both the search and the download capabilities.
-In the \emph{decentralized unstructured} architectures, file placement is random, which means there is no correlation with the network topology whatsoever \cite{yang_improvep2psearch_2002}.
-The most important properties of such systems are that they:
-\begin{itemize}
-  \item support inherent heterogeneity of peers,
-  \item are highly resilient to peer failures, and
-  \item incur low maintenance overhead at peer arrivals and departures.
-\end{itemize}
-In the literature, they can be also found under the term \emph{broadcast-based} systems, because they use \emph{message flooding} among peers (Gnutella\cite{gnutella}) or among superpeers (KaZaA\cite{kazaa}) to propagate queries but with no specific destination.
-Each (super)peer receiving a query
-\begin{inparaenum}[\itshape i\upshape)]
-  \item decreases the incoming query message's lifetime which is measured in \emph{time-to-live (TTL)} hops,
-  \item back-propagates its results, and
-  \item if the lifetime of the incoming message did not expired (TTL did not reach zero), it forwards the message to all its neighbours except from the incoming link.
-\end{inparaenum}
-Early incarnations of the Gnutella protocol, though were considered far from scalable \cite{ritter_gnucantscale_2001} because:
-\begin{itemize}
-  \item randomly choosing logical neighbours without knowledge of the underlying network causes mismatch between physical and application-level topologies, and/or
-  \item queries may be flooded to multiple paths merging in one peer while in such case one of the paths would have been enough, and/or
-  \item peers may forward the same message to each other before they receive the query messages from each other
-\end{itemize}
-More generally, studies upon peer-to-peer traffic \cite{seroiu_analysiscds_2002, sen_analyzep2ptraffic_2004} with measurements on some popular systems, such as FastTrack, Gnutella and DirectConnect have shown that it contributes the largest portion of the overall Internet traffic.
-\cite{matei_mapgnutella_2002} showed that, even given that 95 percent of any two nodes are less than 7 hops away from each other, a flooding-based algorithm, like the one used by Gnutella, can generate 330TB/month in a 50.000 node network!
-
-Recently \emph{decentralized structured} schemes have been proposed \cite{ratnasamy_can_2001, stoica_chord_2001, zhao_tapestry_2001, antony_pastry_2001}.
-These, form an application-level topology that provides a hash table look-up and routing service, called \emph{Distributed Hash Table (DHT)} that requires only $O \left ( log n \right )$ steps for a hash-table-like \texttt{look-up()} search that also guarantees perfect recall even if only a single copy of the searched data exists in the system\footnote{In comparison to Gnutella, that requires $O \left ( n \right )$ to reliably locate a specific file}.
-
-Studying the behaviour of the various peer-to-peer schemes \cite{matei_mapgnutella_2002, lv_randomwalks_2002, merugu_str2unstr_2003} showed that the ad-hoc network topology of unstructured overlay networks that preserve \emph{Power Law} and \emph{Small World} characteristics\footnote{Power Law describes the node degree while the Small World describes characteristics of path length and clustering coefficient.
-The clustering coefficient for a node $\upsilon$ in a graph $G = \left( V, E \right)$ is defined as the ratio of the existing connections between $\upsilon$'s neighbouring nodes to $\gamma \times \left( \gamma - 1 \right)$, where $\gamma$ is the number of neighbouring nodes of $\upsilon$.
-High cluster coefficient means that neighbouring nodes of any node $\upsilon$ likely connect one another.} \cite{faloutsos_powerlaw_1999, saroiu_measurefileshare_2002} offer a more promising approach.
-Particularly:
-\begin{itemize}
-  \item Peer-to-peer clients are extremely \emph{transient}.
-Unstructured systems can have high maintenance traffic in delivering messages, updating the mapping, discovering failures and replicating lost data or pointers, making them insufficient on highly volatile networks.
-  \item \emph{Keyword searches} versus \emph{exact-match queries}.
-In DHTs there is a tight control between the data placement and the topology of the network.
-For this reason it is hard to efficiently support partially matched queries while Gnutella and other similar systems effortlessly support keyword searches and other complex queries since the mechanism is realized locally, on a node-by-node basis.
-  \item Popular content is located at multiple peers and thus it is more likely for a flooding-based search to return results.
-DHTs, on the other hand, fit better in the systems which require ability to reliably locate content, even in the extreme case that only a single-copy exists in the network.
-\end{itemize}
-For this reasons, efforts have been placed for optimizing the efficiency of decentralized unstructured peer-to-peer networks. Research mainly focuses on
-\begin{inparaenum}[\itshape i\upshape)]
-  \item reducing unnecessary, redundant communication traffic, and
-  \item exploiting physical locality to reduce communication response.
-\end{inparaenum}
-The goal can be achieved at, both, the application-level network as well as the underlying physical one.
-In the first case by refining the message relay techniques, while in the second one, by adaptively reconstructing the application network to map as well as possible to the the physical network.
-
-This survey begins with a quick review of the conception of the overlay network.
-It continues with the phenomenon of the topology mismatch between the overlay and the underlying networks and ultimately end up presenting a series of distinguished research work that has been published in the recent literature concerning the alleviation of such a problematic state, focusing on decentralized unstructured peer-to-peer architectures.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Overlay Networks}
-An \emph{overlay} network is an abstract, logical interconnection formed on top of another network.
-It consists of a set of nodes, the computing elements, that are being connected by virtual (logical) point-to-point links.
-For example, the Asymmetric Digital Subscriber Line (ADSL) is, actually, an overlay network on top of the Public Switched Telephone Network (PSTN).
-
-The abstraction of overlay networks has been proposed as a way to implement efficient, fully distributed, application layer, services on top of the best-effort IP layer forming a widely ranged family of protocols collectively known as \emph{peer-to-peer} or \emph{p2p} for short.
-
-Initial efforts targeted on the implementation of one-to-many addressing schemes that could replace \emph{IP multicast} in providing higher quality of streaming media through \emph{quality of service (QoS)} guarantees.
-IP multicast as well as other proposals such as \emph{IntServ} and \emph{DiffServ}\cite{cisco_diffserv_2005} have not seen wide acceptance (yet?), largely because they require changes to all routers of the network.
-On the other hand, an overlay network can be incrementally deployed on end-systems running the overlay protocol software, without cooperation from the ISPs.
-Academic research includes \cite{chu_esm_2000}, \cite{jannotti_overcast_2000} and \cite{kwon_tag_2002} among others.
-
-Overlay networks were also implemented in order to back the routing of messages to destinations whose address is not known in advance.
-This resulted in the appearance of well kwown peer-to-peer protocols such as \cite{gnutella} and \cite{maymounkov_kademlia_2002} widely used for digital content sharing among their network nodes.
-
-Other special perpose overlay networks include \cite{anderson_ron_2001} for resilient routing; \cite{subramanian_overqos_2004} for quality of service guarantees; and \cite{clarke_freenet_2001} for anonymity; to name just a few.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{The Topology Mismatch Problem}
-An overlay network is able to control the sequence of peers, a message traverses, before reaching its destination but it is completely unaware of how the actual packets are switched at the underlying infrastructure along the overlay path.
-For example, a single logical point-to-point link on the overlay, most of the time corresponds to multiple physical links, in the underlying layer.
-
-Additionally a link in the underlying network often serves the mapping of several overlay paths causing increase of the physical link's \emph{stress}\cite{chu_esm_2002}.
-Furthermore, the stochastic behaviour of peers randomly joining and leaving the peer-to-peer network all cause the overlay to physical mapping to create such unnecessary, redundant traffic that impacts the efficiency of the network in terms of average response time.
-The phenomenon is reffered to as the \emph{topology mismatch problem}.
-
-For example, assume nodes $A$, $B$, $C$ and $D$ are connected through the physical network shown in Figure~\ref{figure:phys} and peer $A$ sends a message to peer $C$ (through $B$ of course).
-If these peers participate in an overlay network according to one of the setups of Figures~\ref{figure:over1} and~\ref{figure:over2} then users will yield different performances.
-In the fist overlay, the message will traverse the following sequence of links in the physical layer: $3 \rightarrow 7 \rightarrow 1 \rightarrow 2 \rightarrow 2 \rightarrow 1 \rightarrow 7 \rightarrow 8 \rightarrow 5 \rightarrow 5 \rightarrow 4$.
-In the second overlay: $3 \rightarrow 8 \rightarrow 4$.
-For the first case the cost is $45 ms$ while in the second $15 ms$.
-Therefore the second overlay is more congruent with the underlying physical network than the other one and thus more efficient.
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.4]{img/phys.jpeg}
-\caption{Interconnection of nodes in the physical level.}
-\label{figure:phys}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.4]{img/over1.jpeg}
-\caption{An overlay setup}
-\label{figure:over1}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.4]{img/over2.jpeg}
-\caption{An (other) overlay setup}
-\label{figure:over2}
-\end{figure}
-
-The problem of constructing an optimal overlay is known to be \emph{NP-hard}\cite{chawathe_scattercast_2000}.
-Moreover, The Internet is an interesting environment where end-to-end latencies demonstrate triangle inequality violations (TIVs).
-These delay space TIV is a consequence of the Internet's structure and routing policies and thus will remain a property of the Internet for the foreseeable future \cite{zheng_irprtt_2005}.
-TIVs affect network coordinate \cite{cox_vivaldi_2004, wong_meridian_2005} and positioning \cite{ng_gnp_2001} and makes difficult the construction of delay aware overlays.
-A number of heuristic approaches have been proposed in an effort to finding an algorithm that performes reasonably and consistently well.
-
-Several protocols for dealing with the problems derived from the topology mismatch problem in decentralised unstructured peer-to-peer systems have been proposed in the literature.
-Each implementation has different goals but all share some common ground on the aspect of the network functionality they focus on, in order to achieve them.
-Thus, algorithms can be categorized into three basic types:
-\begin{itemize}
-  \item \emph{forwarding-based}.
-A peer selects only a subset of its neighbours to re-broadcast query messages.
-The selection is made using one or a combination of various statistical metrics.
-Examples of such metrics, could be the number of responses received by a neighbour or the connection latency of the link between the nodes e.t.c.
-These approaches enhance search efficiency but have also several drawbacks.
-First they drastically reduce the search scope.
-Expanding the search scope, on the other hand, is no easy task because the overhead of forming multicast trees is proportional to the multicast group size.
-Also such schemes do not consider dynamic joining and leaving of peers so they do not scale well on dynamic environments.
-  \item \emph{cache based}. These approaches can be further divided into:
-  \begin{itemize}
-    \item \emph{data index caching}, and
-    \item \emph{content index caching}.
-  \end{itemize}
-  Additionally caching policies can be divided into:
-  \begin{itemize}
-    \item \emph{centralized}, and
-    \item \emph{local}.
-  \end{itemize}
-  Centralized peer-to-peer systems provide centralized index servers to keep indices of shared files of all peer (e.g.
-Napster).
-In the local caching policy, each or some peers maintain information about other peers (like the superpeer scheme of Kazaa).
-Caching can significantly reduce traffic costs and response time but duplication of messages still exist, limiting scalability of such approaches.
-  \item \emph{overlay optimization}. This approach can be subdivided furtherlly into:
-  \begin{itemize}
-    \item \emph{Spanning tree based}.
-Construction of a rich graph and the build its minimum spanning tree.
-These approaches incur large traffic overhead to the system \cite{chu_esm_2000, chu_esm_2002}.
-    \item \emph{Cluster based}. Selection to link physically closer nodes with each other. These approaches can shrink the search scope significantly while mapping accuracy is not always guaranteed.
-    \item \emph{Minimum latency first}.
-Use of latency as a metric to calculate distance among peers.
-They require global latency information ``landmarks''\footnote{Measuring latency between peers and stable Internet servers.}.
-  \end{itemize}
-\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{The Algorithms}
-Having understood the nature of the problem that peer-to-peer architecture face, this section starts the discussion of the academic work that has been conducted in the field the last few years.
-What follows this introduction, does not claim to be a thorough citation of all known protocols that are available out there, but instead, a carefull selection of those that left a distinctive fingerprint contribution in the efforts of the research community to alleviate the topology mismatch problem.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Narada}
-\emph{End System Multicast} \cite{chu_esm_2000, chu_esm_2001, chu_esm_2002} as well as other  proposes an architecture, where end-systems implement all multicast related functionality (including membership management and packet replication).
-In \emph{Narada}, the protocol developed in the paper's context, end-systems
-\begin{inparaenum}[\itshape i\upshape)]
-  \item self-organize, in a fully distributed manner,
-  \item into an efficient overlay structure, both from the network and the application perspective,
-  \item with self improving properties,
-  \item adapting to network dynamics\footnote{Long-term variations in Internet path characteristics (i.e. bandwidth, latency e.t.c.)}.
-\end{inparaenum}
-
-What Narada really does is to construct overlay spanning trees in a two-step process:
-\begin{enumerate}
-  \item It builds a richer connected graph \footnote{ENarada \cite{li_enarada_2008} used Gossip protocol for the construction}, termed a \emph{mesh}, while trying to ensure a couple of desired performance properties:
-    \begin{itemize}
-      \item path quality between any pair of members should be comparable to the quality of the unicast path between them, and
-      \item each member should have a limited number of neighbours on the mesh.
-    \end{itemize}
-
-  \item It computes minimum spanning trees out of the mesh, each rooted at the corresponding source member.
-\end{enumerate}
-
-Narada incrementally improves the quality of the mesh by adding and dropping overlay links.
-Members probe each other at random and new links are added depending on what is the gain in doing so.
-On the other hand, participants, continuously  monitor the utility of existing links, and drop them if they are found to be not usefull.
-
-Narada protocol achieves relatively good performance for small and medium sized groups involving tens to hundreds of members.
-This is not, by no means, what a contemporary peer-to-peer system needs to be characterized as scalable, but this work is considered pioneering\footnote{Another similar approach is \emph{Scattercast}\cite{chawathe_scattercast_2000}}, in that it was the first to have conducted a detailed Internet evaluation to analyse and demonstrate the feasibility of an overlay-based, application layer, multicast architecture that can dynamically adapt to bandwidth and latency properties of the physical underlying infrastructure.
-
-\subsection{Gia}
-\emph{Gia} \cite{chawathe_gia_2003} is a decentralized unstructured peer-to-peer file sharing system with the following characteristics:
-\begin{inparaenum}[\itshape i\upshape)]
-  \item replaces Gnutella's flooding with random walks \cite{lv_randomwalks_2002},
-  \item recognizes the importance of overlay network's topology, while using the random walks, and exploits it by incorporating a topology adaptation algorithm,
-  \item introduces a token-based flow control mechanism, and
-  \item detects the significant heterogeneity in peer bandwidth, processing power, disk speed e.t.c. and takes advantage of it.
-\end{inparaenum}
-
-More specifically, there are four key components in the design of \emph{Gia} which are summarized bellow:
-\begin{enumerate}
-  \item A \emph{dynamic topology adaptation} protocol that puts participating nodes within short reach of high capacity nodes so that these \emph{high-degree} nodes, which due to their high connectivity receive most of the queries, actually have the capacity to handle them.
-  \item An \emph{active flow control} scheme is used to avoid overloaded hot-spots. Heterogeneity is detected and flow control tokens are given to nodes based on the available capacity.
-  \item Every node maintains pointers to to the content that is offered by their immediate neighbours, creating a \emph{one-hop replication} of pointers, scheme.
-  \item A \emph{search protocol} based on random walks, that is biased towards directing queries to high-capacity nodes that are typically best able to answer these queries.
-\end{enumerate}
-
-\subsection{Adaptive Overlay Topology Optimization}
-\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{liu_auto_2003} is an algorithm for building an overlay multicast tree among each source node and its direct logical neighbours in order to alleviate the mismatching problem while providing a larger query coverage range.
-It includes the steps of
-\begin{inparaenum}[\itshape i\upshape)]
-  \item \emph{Selective Flooding (SF)}, and
-  \item \emph{Active Topology (AT)}
-\end{inparaenum}
-which are furtherlly discussed in the following paragraphs.
-
-\subsubsection*{Selective Flooding (SF)}
-Instead of flooding to all neighbours, SF builds an overlay multicast tree among each peer and its immediate logical neighbours.
-This reduces overall message burden to the network by making some neighbours non-flooding.
-The tree is formed using a minimum spanning tree algorithm.
-For this reason a peer needs to know the costs to all its logical neighbours (e.g.
-the network delay) as well as between any pair of neighbours.
-Additional, cost probing, messages need to be added to the original messages defined by the Gnutella protocol.
-
-LTM's SF effectiveness has be proven to be detached from the different physical or overlay topologies.
-On the other hand, SF is more effective with large number of logical neighbours.
-It can reach an average optimization rate of 87.4 percent on a logical topology with an average of 30 logical neighbours.
-
-\subsubsection*{Active Topology (AT)}
-At this step the overlay topology is reorganized.
-Each peer, independently, makes optimizations in the overlay network to alleviate topology mismatching by replacing non-flooding neighbours, with closer nodes.
-The approach incorporated by the paper is called \emph{Randomized AT} algorithm which picks up a candidate peer at random among the non-flooding neighbour's neighbours.
-Whenever a new neighbour cost table is received or there is a change of neighbours, the source peer has to re-calculate the multicast tree and apply the randomized AT algorithm.
-
-Different numbers of average logical neighbours has little to do with the effectiveness of AT.
-If the source has $n$ non-flooding peers, there are $n$ potential neighbour replacements.
-The overhead to exhaust all $n$ possible replacements can be high, so in practice, after each replacement the source peer can decide whether it needs to find another candidate peer.
-This is done by computing the cost improvement ratio greater than some predefined termination threshold.
-The larger the threshold, the slower, in the number of optimization steps, the reduction of the normalized average distance.
-As a whole the average response time is significantly reduced when more optimization steps taken.
-
-\subsection{Location-aware Topology Matching}
-\emph{Location-aware topology matching} or \emph{LTM} \cite{liu_ltm_2004}, is an algorithm of building an efficient overlay by disconnecting low productive connections and choosing physically closer nodes as logical neighbours while still retaining the search scope and reducing response time of queries.
-This is done by issuing detector messages by peers, to a small region so that receivers of the detector can record relative delay information and cut inefficient or redundant logical links as well as add closer nodes as direct neighbours.
-
-Three operations are defined in LTMband are furtherly discussed in the following paragraphs.
-
-\subsubsection*{TTL2-detector flooding}
-\emph{TTL2-detector} messages are designed based on the specification of the Gnutella protocol.
-Suppose that $d(i, S, v)$ denotes the TTL2-detector which has the message ID of $i$ with TTL value of $v$ and is initiated by $S$.
-In addition to Gnutella's unified 23-byte header, such a message has a body of two types:
-\begin{itemize}
-  \item short format $d(i, S, 1)$ if $S$ is a direct neighbour of the receiver
-  \item long format $d(i, S, 0)$ if $S$ is within a two-hop distance from the receiver
-\end{itemize}
-This way a peer can calculate link costs from within a one- or two-hop area around it and thus reach peers in $N(S)$ and $N^2(S)$ sets, respectively, in the paper's context.
-Each peer floods a TTL-detector periodically while it calculates link cost to another peer by receiving other peer's detector messages and checking:
-\begin{itemize}
-  \item the Source Timestamp field and its own clock upon time of receipt for one-hop messages
-  \item the Source Timestamp, TTL1 Timestamp and its own clock upon receipt for two-hop messages
-\end{itemize}
-Of course peer clocks must be synchronized in order to do such checking.
-
-\subsubsection*{Low productive connection cutting}
-There are three cases for any peer $P$ who receives $d(i, S, v)$ multiple times:
-\begin{inparaenum}[\itshape i\upshape)]
-  \item $P$ receives both $d(i, S, 1)$ and $d(i, S, 0)$
-  \item $P$ receives multiple $d(i, S, 0)$s from different paths, and P randomly chooses to process one
-  \item $P$ receives one $d(i, S, 1)$ and multiple $d(i, S, 0)$s, and $P$ processes $d(i, S, 1)$ and one randomly selected $d(i, S, 0)$
-\end{inparaenum}
-If the link with the largest cost is found and is a direct neighbour then the connection is put in a will-cut list and stays there for a certain period of time.
-If it is not, then it is handled by other peers.
-After that period, connections are cut and recorded to $P$'s cut-list.
-
-\subsubsection*{Source probing}
-For a peer $P\in(N^2(S) - N(S))$ who receives only one $d(i, S, 0)$, the cost of $PS$ is obtained (with list look-up or probing).
-Then $P$ compares it with the cost form each hop and if $PS$ has the largest cost, $P$ will not keep this connection, while otherwise the connection will be created.
-
-\paragraph{}
-Supposing $n$ is the number of peers, $c_n$ is the average number of neighbours and $c_e$ is the average cost of logical links, then in the flooding-based search the traffic incurred by one query from an arbitrary peer in a peer-to-peer network is $O(n)$.
-As observed in the Gnutella network \cite{sripanidkulchai_gnutella_2001}, each peer issues $0.3$ queries per minute in average, thus the per minute traffic incurred by the network with $n$ peers is $O(n^2)$.
-Because each $d(i, S, v)$ has a TTL of $2$ in each source peer, the traffic for one time LTM optimization in all peers is at most $2nc_n^2c_e$.
-If each peer uses LTM $k$ times per minute, the total traffic incurred is $2knc_n^2c_e$.
-Simulation shows the best value for $k$ is $2$ or $3$.
-So, the traffic overhead caused by LTM to the network is $O(n)$.
-
-TTLj-detectors, with $j > 2$, would detect and break cycles with more than 4 links.
-LTM though, does not use such detectors because detector-flood traffic would increase significantly, and cut links between two end-peers, could cause queries initiated by them to traverse a path much more expensive than the cost on the the cut link.
-
-LTM disadvantages are
-\begin{inparaenum}[\itshape i\upshape)]
-  \item disagreement of measured delay due to unsynchronized clocks causes problems when deciding the cut positions, which can influence the network connectivity, and
-  \item the network delay metric mainly focuses on disabling the connections between peers physically far away without considering the shortcuts created by powerful peers.
-\end{inparaenum}
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.4]{img/ltm_multid.jpeg}
-\caption{Peer $P$ receives $d(i, S, v)$ multiple times}
-\label{figure:ltm_multid}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.4]{img/ltm_example.jpeg}
-\caption{A full example of LTM}
-\label{figure:ltm_example}
-\end{figure}
-
-\subsection{Scalable Bipartite Overlay}
-\emph{Scalable Bipartite Overlay (SBO)} \cite{liu_bipartite_2007} employs an efficient strategy to select query forwarding paths and logical neighbours.
-The topology construction and optimization of SBO consist of four phases:
-\begin{enumerate}[\itshape i\upshape)]
-  \item When a new peer is joining the network, it will randomly take an initial colour; say red or white.
-This colour is not to be changed for the time the peer stays connected.
-If it leaves and then rejoins it will, again, have to pick a colour, randomly.
-Thus all peers are separated into two groups, red or white.
-Then the bootstrap host will provide, the joining peer, with a list of active peers, including their colour information, in order for the later to establish connections to different colour peers.
-This way, all peers form a bipartite overlay.
-  \item White peers probe distances with their immediate (red) neighbours, form a cost table and send this table to their corresponding red neighbours.
-  \item Each red peer builds a minimum spanning tree using the obtained neighbour cost tables.
-These are two-hop diameter trees so white peers do not need to build one.
-The links that are part of a minimum spanning tree are called \emph{forwarding connections (FC)} while the rest \emph{non-forwarding (NFC)}.
-  \item Having a minimum spanning tree with two hops a red peer is able to send its queries within that range.
-Some white peers, though, have become non-forwarding neighbours.
-In this phase, such a (white) neighbour will try to find another red peer being two hops away from its current red neighbour to replace the later as its new neighbour.
-\end{enumerate}
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.4]{img/sbo_efficient_forward.jpeg}
-\caption{A red peer $P$ has a small overlay topology of $N(P)$ and $N^2(P)$ and computes the efficient forwarding paths}
-\label{figure:sbo_efficient_forward}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.4]{img/sbo_neighbour_replace.jpeg}
-\caption{Neighbor replacement}
-\label{figure:sbo_neighbour_replace}
-\end{figure}
-
-In a static enviromnent LTM may reduce traffic cost by around 80 to 85 percent while SBO reduces traffic cost between 85 and 90 percent.
-However, LTM  is proved to converge in around 2-3 steps while SBO needs 4-5 steps.
-Moreover LTM reduces response time by more than 60 percent in 3 steps while SBO needs 8.
-In a dynamic environment (10 minute average peer lifetime, 0.3 queries/sec by each peer) SBO and LTM reduce the average traffic cost per query (including the overhead due to the optimization steps) by 85 and 80 percent, respectively.
-Moreover LTM reduces the response time per query to 30 percent while SBO to 35 percent.
-
-\subsection{Adaptive Connection Establishment}
-\emph{Adaptive Connection Establishment (ACE)} \cite{liu_ace_2004} builds an overlay multicast tree among each source node and the peers within a certain diameter from the source peer and optimizes the neighbour connections that are not in that tree.
-
-ACE indicates three phases for its algorithm:
-\begin{enumerate}[\itshape i\upshape)]
-  \item Calculate cost between nodes using network delay as a metric.
-Each peer probes the costs with its immediate logical neighbours and forms a \emph{neighbour cost table (NCT)} using a special routing message type.
-Two neighbouring peers exchange their NCTs in order for every peer to obtain the cost between any pair of its local neighbours forming a small overlay topology.
-  \item Based on obtained NCTs a minimum spanning tree among each peer and its immediate neighbours is built (Figure~\ref{figure:ace_phase2}).
-  \item Physically far away neighbours are replaced by physically close neighbours.
-In ACE a peer, say $S$, probes the distance between one of its non-flooding neighbour's neighbour, say $G$ and $H$ respectively.
-If the link to neighbour's neighbour is smaller than that to the neighbour, the later is cut.
-If this is not the case but $S$ finds that cost of $GH$ is even larger than that of the $SH$, $S$ will keep $H$ as a new neighbour.
-Obviously, if $SH$ is larger than $SG$ and $GH$, the connection will not be established and $S$ will continue probing another neighbour's neighbour.
-The algorithm described is conducted within $1$-neighbour closure (among its source peer and all its direct neighbours) but the optimization scope can be enlarged.
-The larger the scope, the better topology matching improvement but also the greater the computational overhead (Figure~\ref{figure:ace_phase3}).
-\end{enumerate}
-
-Simulations in \cite{liu_acesims_2004} show that the average scope of each query to cover the same scope of nodes is reduced by about 65 percent without losing any autonomy feature, while the average response time can be reduced by 35 percent.
-Larger diameter topologies lead to better topology optimization rate but also to higher communication and computation overhead.
-It was also found that it is more effective in higher connectivity dense topologies.
-Compared to LTM, it comes short of convergence speed.
-In \cite{ni_mismatch_2004} shows reduction of both total traffic (90 percent) and response time (80 percent) to message queries without shrinking the search scope.
-SBO, on the other hand, achieves approximately 85 percent reduction on traffic cost and about 60 percent reduction on query response time.
-Last but not least, it is concluded that work must be done on incorporating a more sophisticated selection policy for candidate non-flooding peers.
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.4]{img/ace_phase2.jpeg}
-\caption{Second phase in ACE}
-\label{figure:ace_phase2}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.4]{img/ace_phase3.jpeg}
-\caption{Third phase in ACE}
-\label{figure:ace_phase3}
-\end{figure}
-
-\subsection{Hops Adaptive Neighbour Discovery}
-Paper \cite{chen_hand_2006} proposes a new algorithm named \emph{Hops Adaptive Neighbour Discovery (HAND)} which uses a fully distributed triple hop adjustment strategy to address the topology mismatch problem.
-
-The advantages of the algorithm compared to other approaches are that
-\begin{inparaenum}[\itshape i\upshape)]
-  \item it does not need any clock synchronization,
-  \item it is a fully distributed algorithm making it robust and reliable in decentralized systems,
-  \item the traffic overhead of the triple hop adjustment is very low,
-  \item it is applicable to dynamic peer-to-peer environments, and
-  \item maintains lower query response time.
-\end{inparaenum}
-
-The algorithm's ultimate goal is the optimal overlay, the \emph{Logical Communication Network (LCN)} in the paper's context.
-The key concept of the algorithm is that a graph $G^{*}$ that describes an LCN and a graph $G$ that describes the current overlay are matched only if all peer hop sequences $(v_1, v_2, \ldots, v_k)$ in $G$ exist in $G^{*}$ and in the same order.
-In practice triple sequences $(v_1, v_2, v_3)$ are used.
-
-The mismatching detection is done in the following way.
-Suppose we want to verify peer sequence $v_2-v_1-v_3$ (see Figure~\ref{figure:hand_com_overlay}).
-A pair of probing messages are sent from $v_1$ to $v_2$ and $v_3$.
-Suppose delays of $(v_1,v_2)$ and $(v_1,v_3)$ are are denoted as $x$ and $z$, respectively.
-When the probing message arrives to $v_2$ it forwards it directly to $v_3$.
-Similarly, when the probing message arrives to $v_3$, it forwards it directly to $v_2$.
-These last steps are performed in order to obtain delays of $(v_2,v_3)$ and $(v_3,v_2)$ physical paths, respectively, denoted by $y$.
-\begin{itemize}
-  \item If $y=z-x\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched and should be adjusted to $v_1-v_2-v_3$ by deleting edge $(v_1,v_3)$ and adding a new $(v_2,v_3)$ (see Figure~\ref{figure:hand_matchtriple_lemma3}).
-  \item if $y=x-z\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched and should be adjusted to $v_1-v_3-v_2$ by deleting edge $(v_1,v_2)$ and adding a new $(v_3,v_2)$ (see Figure~\ref{figure:hand_matchtriple_lemma4}).
-\end{itemize}
-$\varepsilon$ is a small positive real number denoting additional delays caused by possible forwarding and jitter delays.
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.4]{img/hand_com_overlay.jpeg}
-\caption{The communication in overlay}
-\label{figure:hand_com_overlay}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.4]{img/hand_matchtriple_lemma3.jpeg}
-\caption{The matching triple in paper's Lemma 3}
-\label{figure:hand_matchtriple_lemma3}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.4]{img/hand_matchtriple_lemma4.jpeg}
-\caption{The matching triple in paper's Lemma 4}
-\label{figure:hand_matchtriple_lemma4}
-\end{figure}
-
-Measurements conducted for evaluation perposes showed that in a static environment the algorithm can effectively decrease traffic cost by about 77 percent and shorten the query response time by about 49 percent in less than two minutes.
-In a dynamic environment it shows similar behaviour and with the size of the overlay network having little impact on the effectiveness of the algorithm.
-Compared to LTM both algorithms have almost the same traffic reduction rate, however on the response time reduction rate HAND has a higher one by about 4 percent.
-The traffic overhead of HAND is much less than that of LTM by an average of 55 percent.
-
-\subsection{Distributed Cycle Minimization Protocol}
-\emph{Distributed Cycle Minimization Protocol (DCMP)} which is introduced in \cite{zhu_dcmp_2008} is a dynamic, fully decentralized protocol that promises significant reduction of duplicate messages.
-To achieve that, it uses the slightly different approach of eliminating unnecessary cycles, while retaining the connectivity of the network and preserving fault resilience and load balancing properties of unstructured peer-to-peer schemas by avoiding the creation of a hierarchical organization.
-
-DCMP aims at cutting the cycle paths at strategic locations. Any peer that detects a duplicate message can initiate the cutting process which consists of the following steps:
-\begin{enumerate}
-  \item Peers in the cycle elect a leader called \emph{GatePeer}\footnote{GatePeers are important for maintaining the connectivity and optimal structure of the network while peers join and leave randomly.}.
-  \item The cycle is cut at a well-defind point with respect to the GatePeer
-\end{enumerate}
-
-The first step after detecting a duplicate message by some peer is to gather information from all peers in the cycle using a new type of control message called \emph{Information Collecting Message} or \emph{ICM}.
-ICM contains:
-\begin{inparaenum}[\itshape i\upshape)]
-  \item a \emph{GUID}\footnote{Globally Unique IDentifier assigned to every query message generated by any node.} field same as the one of the duplicate message,
-  \item \emph{DetectionID} field which represents the direction of the connection where the duplicate was identified\footnote{This ensures the uniqueness of the ICM messages because as it travels through many cyclic paths, multiple peers will detect the duplicates and initiate an ICM message.}, and
-  \item \emph{Node Information Vector (NIV)} which contains information (bandwidth, CPU power, etc) about peers that propagated the ICM.
-\end{inparaenum}
-
-Suppose $A$ detected the duplicate as depicted in Figure~\ref{figure:dcmp}.
-It then emits an ICM to $B$ and $F$, that initially contains information only about itself.
-Each peer that receives the ICM, appends its information and propagates it along the reverse path of the original message.
-Since two copies of ICM are sent, at some point, a peer, say peer $D$, will receive a duplicate ICM.
-Using the information in the NIVs of the ICMs, $D$, decides to cut (for example) the EF connection.
-To inform the other peers about its decision, it emits a \emph{Cut Message (CM)} which contains the GUID and DetectionID of the corresponding ICM and an additional field that identifies the connection to be cut.
-$D$, then, forwards the CM in the reverse directions from where the ICM arrived.
-Similarly CMs received by any peer are propagated toward the reverse path of the corresponding ICM.
-Eventually, either peer $E$ or peer $F$ will receive the CM and cut the connection, thus eliminating the cycle.
-
-Receiving a duplicate ICM denotes the existence of a cycle.
-The opposite is not true though.
-For example, if the cycle contains $2 \times TTL$ edges, it will not be detected, because the ICM messages will be discarded before they locate it.
-There is a tradeoff between preserving the connectivity of the network and minimizing the duplicates that makes such a possibility to be safely ignored.
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.4]{img/dcmp.jpeg}
-\caption{Cycle elimination methods in DCMP}
-\label{figure:dcmp}
-\end{figure}
-
-Experiments in \cite{zhu_dcmp_2008} showed that DCMP incurs a lower delay, returns more results and decreases the number of duplicate messages by 22\%, compared to LTM.
-Additionally DCMP has one to two orders of magnitude less overhead, because it adopts a more ``LAZY'' approach than broadcasting control messages periodically like LTM does.
-
-\subsection{Two-Hop-Away Neighbour Comparison and Selection}
-Work in \cite{liu_thancs_2005, liu_thancs_2008} proposes a distributed heuristic called \emph{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} that
-\begin{inparaenum}[\itshape i\upshape)]
-  \item is completely distributed and needs no global knowledge,
-  \item presents trivial overhead compared to the query cost savings
-  \item its convergent speed of the algorithm is fast enough (faster than minimum spanning tree approaches) so that is effective to dynamic environments, and
-  \item does not shrink the search scope.
-\end{inparaenum}
-
-THANCS is considered a \emph{local search method}, in the sense that it targets in finding a locally optimum solution, by exploiting knowledge within a 2-hop radius.
-The algorithm consists of two main components: \emph{piggybacking neighbour distance on queries} and \emph{neighbour comparison and selection} which are furtherly discussed bellow.
-
-\paragraph*{Piggybacking neighbour distance on queries.}
-Using network delay as a metric for measuring the distance, each peer probes distances with its immediate naighbours and stores information locally.
-For this reason a special query message type, \emph{Piggy Message (PM)}, is introduced.
-It is 6 bytes long and includes two fields: Neighbour IP Address and Neighbour Distance.
-A peer $P$ constructs a PM for its neighbour $Q$, which contains $Q$'s IP address and $Q$'s distance from $P$.
-When $P$ receives a query from $Q$, this PM will be piggybacked by the query that goes to all other neighbours of peer $P$.
-Upon receiving such a query message, each of the other neighbours will detach the PM, record the $PQ$ distance and process the query.
-This PM will not be further forwarded.
-In selecting which incoming queries  should piggyback a PM the paper proposes the \emph{pure propability-based (PPB)} and the \emph{new neighbour triggered (NNT)} policies.
-
-\paragraph*{Neighbour comparison and selection.}
-Figure~\ref{figure:thancs} illustrates this component of the THANCS algorithm.
-A peer $S$ probes the distance to all known unprobed $N^2(S)$\footnote{$N^2(S)$ denotes the set of peers being two hops away from $S$, while $N(S)$ denotes the set of direct logical neighbours of $S$.}.
-The distance of $SP$ is known to $S$.
-Upon receiving a PM from node $P$ with the distance of $PQ$, $S$ follows one of the following:
-\begin{itemize}
-  \item $Q \in \left( N(S) \cap N^2(S) \right)$, i.e.
-$Q$ is direct neighbour of $S$.
-In this case $S$ will compare cost of $SQ$, $SP$ and $PQ$.
-If the most costly connection is one of $SQ$ or $SP$ the corresponding link will be put into a \emph{will-cut list}\footnote{Links are not immediately disconnected when put in the will-cut list.
-There are useless for forwarding but are kept active, for some time, in order to serve query responses that are traveling to the source peer along the inverse search path.}.
-If the most costly connection is $PQ$ then $S$ will do nothing, as the fully distributed nature of the algorithm will give a chance to either $P$ or $Q$ to cut the connection.
-  \item $Q \in \left( N^2(S) - N(S) \right)$, i.e.
-$Q$ is a two-hop-away neighbour of $S$.
-If $S$ hasn't probed $Q$ before\footnote{A distance cache is maintained and looked up in such a case.}, it probes and stores the result in the \emph{distance cache}.
-Having the distance of $SQ$, $S$ compares costs of $SQ$, $SP$ and $PQ$.
-If $SQ$ is the most costly, $S$ will not establish the connection.
-If $SP$ is the most costly, $S$ will establish connection $SQ$ and put $SP$ in the will-cut list.
-If $PQ$ is the longest, $S$ will keep  the connection with both $P$ and $Q$, expecting that $P$ or $Q$ will eventually disconnect link $PQ$, later.
-\end{itemize}
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.4]{img/thancs.jpeg}
-\caption{Probing two-hop-away neighbours}
-\label{figure:thancs}
-\end{figure}
-
-In a static environment THANCS has been proven to be effective; optimizing 45 percent out of the 60 percent of mismatched paths, constructing a nearly optimal overlay.
-This leads to a 60 percent reduction in traffic cost as well as a 40 percent decrease in query response time.
-In dynamic environments (Gnutella 0.6/Limewire super-peer-like and Ion flat-like), THANCS saves up to 70 percent of the traffic cost in the super-peer topology and 55 percent for the flat one.
-Average response time is also decreased by 60 and 45 percent, respectively.
-Generally, THANCS has similar performance to LTM, without needing synchronization.
-SBO, incurring half the  overhead of AOTO, reduces the traffic cost the most, while THANCS has lower response time and converges faster than SBO.
-THANCS is, thus, more suitable for a more dynamic environment.
-In addition, THANCS is easy to implement and its operation overhead is trivial, compared with the other three approaches.
-This design, however, has the limitation of not being easily extend to also support non-flooding-based systems.
-
-\subsection{mOverlay}
-Zhang et al.
-in \cite{zhang_moverlay_2004} focus on two aspects of the overlay construction.
-First, to achieve \emph{efficiency} of communication, a protocol should minimize unecessary long-distance hops and redundant traffic.
-Second, to be \emph{scalable} the overlay should be constructed in a distributed fashion and maintenance cost (data and locality management) should be minimized.
-Both of the above observations, are positively affected by the exploitation of node proximity at the underlying network.
-To accomplish it, the paper introduces the \emph{group} concept, according to which non-static (dynamic) landmarks are used to compute proximity, resulting in a highly robust and scalable overlay construction with reduced maintenance cost as well.
-The dynamic nature of landmarks help towards load-balance, meaning that hot-spots are avoided.
-
-The proposed architecture, called \emph{mOverlay}, is a \emph{clustering approach} which creates a two-level hierarchical network, where on the top level we have connections between groups while on the bottom we have connections between hosts inside groups, thus attempting to recreate \emph{Small-World}-like properties for the constructed overlay network.
-The notion of a \emph{group}, in the paper's context, is a set of hosts that are close to each other with respect to any position $P$ in the underlying network.
-The distance between hosts can be
-\begin{inparaenum}[\itshape i\upshape)]
-  \item network latency,
-  \item round-trip time,
-  \item minimum bandwidth on the links along the path connecting the two nodes, or
-  \item some other user-defined cost metric between the two nodes.
-\end{inparaenum}
-These nodes are refered to as neighbouring. Similarly groups can exchange messages with their neighbouring ones\footnote{Groups nearby, in the underlying physical network.}.
-
-According to the \emph{grouping criterion} set by the authors, a new host $Q$ belongs to some group $A$ if the distance between $Q$ and group $A$'s neighboring groups is the same as the distance between group $A$ and it's neighbouring groups.
-The neighbour groups thus, play the role of the dynamic landmarks and for accuracy and performance reasons the nearest of them are chosen to play that role.
-
-\paragraph{Locating process.} A new coming host, $Q$, first connects to a globally known host cache called the \emph{rendezvous point (RP)} in order to retrieve the starting point in the overlay, say $A$ in group $1$.
-Host $Q$ then, measures its distance to host $A$.
-At the same time, the later, sends information about the neighbour groups of group $1$ back to host $Q$.
-This list is called \emph{candidate group list}, and the newcoming host sequnentially measures its distance to each of them in seek for the closest one.
-If the \emph{grouping criterion} is met, host $Q$ belongs to group $1$.
-If not, a boot host from the closest group is found and the algorithm is re-run until the criterion is met or after a predefined number of repetitions.
-In the later case, $Q$ creates a new group comprising itself only.
-The above protocol does not favour hotspots as it spreads the probability of visiting a group across the whole overlay and limits the overhead in the level of $O \left ( log N \right )$.
-
-\paragraph{General overlay operations.} A set of additional protocols, are also introduced, similar to those found in traditional unstructred networks, but modified focusing on scalability and robustness.
-For example a protocol for \emph{group formation} is introduced that exploits the inherent characteristic of proximity, in the overlay, in order to efficiently detect the neighbouring groups of a newly formated group from the set of adjasent groups of its closest neighbour.
-Additionaly, during \emph{group joining} the coresponding protocol denotes the exchange of important information for group maintenance.
-This can be furtherly improved by \emph{information sharing} between nodes of the same group, functionality handled by a dedicated flood-like protocol\footnote{Since nodes that belong to the same group are physically close this can be achieved at a minimum price.}.
-Moreover, another set of distributed protocols handle the \emph{information update}.
-The information that needs update, in the proposed architecture, is
-\begin{inparaenum}[\itshape i\upshape)]
-  \item the host cache, when a new node joins, and
-  \item the neighbours of groups, when a close-by group is generated.
-\end{inparaenum}
-Finally, in case of \emph{host failure} or \emph{host departure} the system is able to maintain its stability since there are defined operations for periodical host cache update and group leader selection if one leaves or dies.
-
-\subsection{Distributed Domain Name Order}
-\cite{zeinalipour-yazti_ddno_2005} proposes the \emph{Distributed Domain Name Order (DDNO)} technique which makes unstructured overlay networks, topologically aware.
-It results in a flat overlay topology but with some changes it can be utilized in \emph{superpeer} environments.
-In DDNO, a node of degree $d$, tries to connect to $\frac{d}{2}$ nodes that belong to the same domain (\emph{sibling} connections) and another $\frac{d}{2}$ of random nodes (\emph{random} connections).
-Connecting to \emph{sibling} nodes ensures the reduction of long distance message travelling and thus resulting to better performance, while \emph{random} nodes keep the structure connected.
-
-\subsubsection*{Discovering \emph{random} neighbours}
-Initially, when a new coming node wants to join a network $N$, uses a \emph{hostcache} mechanism to provide it with a list of nodes with which to establish its $\frac{d}{2}$ \emph{random} connections.
-
-\subsubsection*{Discovering \emph{sibling} neighbours}
-In order to discover the rest $\frac{d}{2}$, \emph{sibling}, nodes, the \emph{lookupDN} procedure is initiated.
-In this phase, a special $l$ walker message is multicasted, by the newcomer, in search of a node that knows\footnote{A special structure is used called ZoneCache that contains information about which nodes are reachable in an $r$-hop radius.} how to guide $n$ to a \emph{sibling} node.
-
-Figure~\ref{figure:ddno_lookupdn} shows an $l$ message emitted by node $n$ which travels along the path [$a$, $b$, $c$, $e$, $b$, $d$].
-At node $d$, $l$ finds information to make a decision on which neighbour to follow next\footnote{Contrary to the random choices, made at the previous hops.}.
-Ultimately $l$ arrives at $m$, which is a \emph{sibling} node to $n$.
-$m$ then issues a broadcast message to all its own \emph{siblings}.
-Each of the receiving nodes, $m$ included, will respond to $n$ with a designated $l^{'}$ message if it is willing to accept new connections and out of these answers, node $n$, will attempt to establish its, remaining, $\frac{d}{2}$ \emph{sibling} connections.
-
-Unfortunately, use of DDNO results in networks which have a uniform distribution of node degrees, losing the valuable properties of Power Law and Small World networks, as mentioned previously in this survey.
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.375]{img/ddno_lookupdn.jpeg}
-\caption{Domain name look-up in a \emph{DDNO} topology}
-\label{figure:ddno_lookupdn}
-\end{figure}
-
-\subsection{Critical Topology-Aware Grouping}
-\emph{Critical Topology-Aware Grouping (CTAG)} presented in \cite{zhao_ctag_2006}, is a grouping algorithm that tries to exploit low-cost and low-delay communication of physically close-by peers.
-The grouping strategy is based on the \emph{IANA}\footnote{Internet Assigned Numbers Authority} and the respective \emph{Regionanal Internet Registry (RIR)}'s IP assignment strategies, according to which nodes within the same organization are always addressed from the same block.
-The paper proposes the \emph{Adjacency Measurement (AM)} technique which uses the longest matching IP segment criterion to calculate node proximity.
-
-Observations made in various studies (e.g.
-\cite{matei_mapgnutella_2002}) concerning the distribution of nodes among \emph{Internet Service Providers (ISPs)} and \emph{Autonomous Systems (ASs)} have shown that only $2$ to $5$ percent of Gnutella connections link peers within a single AS while more than $40$ percent of all Gnutella peers are located within the top 10 ASs.
-Similarly, measurements in \cite{zeinalipour-yazti_gnudc_2002} used a $244,000$ IPs test-bed and results have shown that $45$ percent of the nodes belonged to only $10$ large ISPs and $58$ percent belong to only $20$.
-Such results mean that most overlay generated traffic crosses AS borders increasing topology mismatch cost.
-
-\emph{CTAG} focuses on both the construction of the overlay as well as dynamically revising it during node interaction, phases called \emph{bootstrapping grouping} and \emph{dynamic revision}, respectively in the paper's context.
-
-\subsubsection*{Bootstrapping grouping}
-The \emph{Gnutella Web Caching} mechanism has been modified in order for a new coming node to choose the closest \emph{GWC} in order to retrieve the node list for bootstrapping.
-
-\subsubsection*{Dynamic revision}
-Similarly to the bootstrapping phase, \emph{AM} metric is used to store hosts' addresses.
-read from \emph{X-Try} headers during handshake or from \emph{QueryHit} messages.
-Additionally, when a node reaches the max neighbour connections, node disconnects the neighbours with the lowest \emph{AM}.
-
-\subsection{Peer-exchange Routing Optimization Protocols}
-Work presented in \cite{qiu_prop_2007} introduces \emph{peer-exchange}\footnote{A series of exchanges of neighbours between two peers.
-One exchange can be viewed as a pair of cut-add operations.} as a basic operation to adaptively adjust the connections of the overlay network, and efficiently reduce the average logical link latency of the whole system.
-There are two points that differentiate this scheme from others:
-\begin{inparaenum}[\itshape i\upshape)]
-  \item it utilizes the collaboration of two peers to optimize their neighbourhood environment, than simply letting each node to ``selfishly'' choose its own strategy, and
-  \item can be deployed effortlessly on both unstructured and structured peer-to-peer systems.
-\end{inparaenum}
-
-\emph{PROP-G}\footnote{\emph{G} stands for \emph{generic}.} is the simple and direct variant of \emph{peer-exchange} in which two neighbours swap all their neighbours.
-As depicted in figure~\ref{figure:prop-g} the mechanism can be viewed as, nodes, exchanging overlay ``positions''.
-Thus, the overall overlay topology is not affected by the this operation.
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.4]{img/prop-g.jpeg}
-\caption{PROP-G in which a neighbours are exchanged.}
-\label{figure:prop-g}
-\end{figure}
-
-The alternative protocol proposed, called \emph{PROP-O}\footnote{\emph{O} stands for optimized}, ensures that the degree of each participating node remains the same by selectively choose the same number of neighbours for the exchange.
-With this mechanism, 2 nodes exchange the same number of neighbours in order not to change the degree
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.4]{img/prop-o.jpeg}
-\caption{PROP-O exchanging \emph{m} neighbours.}
-\label{figure:prop-o}
-\end{figure}
-
-\subsection{T2MC (unstrucured)}
-\emph{T2MC}\cite{shi_t2mc_2008} exploits some properties of the Internet paradigm and clusters nodes belonging to the same ISP without any centralised control or predefined system parameterization.
-The algorithm considers the dynamic nature of peers and exploits the stable nature of routers in order to build a topology location relationship among end peers.
-Some special routers split the physical network into autonomous system domains.
-Using a Traceroute mechanism T2MC searches for latency leaps among the path to a host in order to form ``near'' and ``remote'' router clusters.
-This is achieved by a 2-Means Classification, defined by the following steps:
-1.the peer chooses the minimum and maximum Latency results from the Traceroute for initializing the cntroinds of two sets ``first'' and ``second''.
-2.The peer calculates, for all hops allong its tracerouted path, the absolute distance to the centroids of both sets and assigns the routers to that centroid with which it has the smaller absolute distance.
-3.The peer calculates the latncy mean and variance value of two sets
-4.If the variance is larger than a predefined threshold then the algorithm takes a loop from step 2 picking the two latency mean values as new centroids of sets ``first'' and ``second''.
-Ultimately peer will end up with two sets having the minimum intra-set variance.
-
-Finally the peer chooses the router from ``second'' set with the minimum hops attribute and sets it as a threshold.
-The selected router and all others whose hops attribute is larger than the threshold are classified as ``remote'' router cluster.
-The remaining are classified as ``near''.
-From the ``near'' class, the peer chooses the one with maximum hops attribute as its edge router, and registers it along with the all the ``near'' cluster into the DHT of the p2p overlay.
-As new peers join the network, those that share the same edge router or any of the members of the ``near'' router clustersm ther would gather to form a ``close'' peer cluster.
-As Edge routers can provide more valuable information than other members of the ``near'' set, T2MC was designed to prioritize interaction of peers and edge gateways.
-TODO: figure t2mc.jpeg (a) stars represent edge routers (b) hop leaps
-The use of Traceroute as a tool for implementing the distance measuring infrastructure raise concearns about its efficiency and scalability.
-Being, primarilly, a network diagnostic utility, it is concearned too heavy weighted and intrucive for use in a larger scale\cite{ratnasamy_binning_2002}.
-Additionally, disabling ICMP is a common administrative policy for edge sites to enforce security, while dumping BGP routing tables\cite{krishnamurthy_bgpclust_2000} is not directly available to the application layer.
-
-\subsection{Unnamed Unstructured!!}
-(TODO: to be reviewed)
-In \cite{hsiao_redblue_2009}, Hsiao et al, claim to construct topology-aware unstructured overlays that \emph{guarantee} performance qualities in terms of
-\begin{inparaenum}[\itshape i\upshape)]
-  \item the expected communication latency among any two overlay peers regardless of the network size, and
-  \item the broadcasting scope of each participating peer.
-\end{inparaenum}
-
-The algorithm constructs an undirected graph $G = \left( V, E \right)$ comprised by two subgraphs.
-The first, namely $G^{\left( red \right)} = \left( V^{\left( red \right)}, E^{\left( red \right)} \right)$ in the paper's context, includes all vertices of $G$ and ensures the connectivity of the graph by securing at least one path between any two nodes.
-In contrast, $G^{\left( blue \right)} = \left( V^{\left( blue \right)}, E^{\left( blue \right)} \right)$, contains those vertices of $G$ that have free edges to link to other nodes and because these are fully utilized, the following also stands $E = E^{\left( red \right)} \cup E^{\left( blue \right)}$.
-
-A joining peer $u$, partitions its neighbours into two subsets, the $B_u^{\left( red \right)}$ and $B_u^{\left( blue \right)}$.
-In order to populate the $B_u^{\left( red \right)}$ subset, peer $u$ samples peers uniformly and at random.
-Then, each of these selected peers discovers a routing path starting from itself towards the node with the smallest (or the largest) ID in the system.
-
-\section{Structured overlays}
-\emph{Structured} overlay networks provide a self-organising infrastructure for large-scale peer-to-peer applications.
-They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes through a deterministic mechanism.
-Some well-known implementations include CAN\cite{ratnasamy_can_2001}, Chord\cite{stoica_chord_2001}, Pastry\cite{antony_pastry_2001}, Tapestry\cite{zhao_tapestry_2001} and Kademlia\cite{maymounkov_kademlia_2002} which have been used in a wide range of distributed applications (TODO: cite some of these applications i.e OceanStore Far).
-
-The main advantage of the DHT approach of structured p2p overlays over unstructured ones, is that they provide a guaranteed bound on the number of overlay routing hops that have to be taken in order for an object to be located.
-(TODO: provide the O complexity for each of the CAN($O(d(N^{\frac{1}{d}}))$), Chord($O(\log N)$), Pastry($O(\log N)$), Tapestry($O(\log N)$), Kademlia($O(\log N)$)).
-
-Typically, node IDs are assigned \emph{randomly}, resulting in excellent load balancing, scalability and robustness.
-Unfortunatelly, this randomness has a negative impact on the \emph{routing locality} of the network.
-This means that even though the target node can be reached with logaritmic overlay hops, the distance traveled in the physical underlying network, during the overlay routing process, can be far from optimal.
-The ratio of the phyical distance, a query travels through the overlay to an object, and the minimal distance to that object (i.e.
-though IP) is known in the literature as \emph{stretch} or \emph{Relative Delay Penalty (RDP)}.
-
-Xu et.al in \cite{xu_globstate_2003} state that there are three ways generaring proximity information
-\begin{itemize}
- \item Expanding Ring Search,
- \item Heuristics, and
- \item Landmark Binning
-\end{itemize}
-
-Expanding ring search can be of two forms.
-First it can utilize the multicast infrastracture in the underlying network in order to emit its messages.
-Unfortunately such infrastructures are node widely deployed thus the implementations of this way of generating proximity information is limited to blindly flood the neighbourhood to obtain the reasonable results.
-
-Heuristics are used in order to reduce the blindness of the expanding ring search and make realizations more efficient and effective.
-Unfortunatelly a common problem of all heuristic approaches is the local minimum pitfall in which the search might be caught into.
-
-Landmark clustering is based on the view that nodes with similar distances to a set of predifined, well-known landmark nodes are pretty likely of being close to each other.
-But this approach has its weeknesses as well, such as the fact that is a rather coarse grained approximation, therefore not particularly well suited for differantiating nodes within close distance to each other.
-
-Castro et.al in \cite{castro_proximitydht_2002, castro_topawareroute_2002} on the other hand, comment on the three basic approaches in exploiting proximity in DHT protocols suggested by Ratnasamy et.al.
-in \cite{ratnasamy_openq_2002}.
-These are:
-\begin{itemize}
- \item Geographic Layout,
- \item Proximity Routing, and
- \item Proximity Neighbour Selection
-\end{itemize}
-
-According to the first, the node IDs are assigned in such a way, that nodes close by in the psysical network topology, be close in the node ID space as well.
-Implementations, that work relatively well with this approach, have been incorporated into CAN.
-Nodes measure the RTT between themselves and a set of landmarks in order to match the CAN space as much as possible to the psysical one.
-Unfortunately, the use of landmarks renders the protocol which uses it, not fully self-organizing which can furtherly lead to imbalanced node distribution.
-On other DHTs, such as Chord or Pastry, another problem emerges.
-To gain fault tolerant properties, these protocols, replicate key-value pairs on neighbouring (in the ID space) nodes.
-When a proximity-based node ID assignment has been used, the needed failure resilience is undermined by the fact that close by nodes are more likely to suffer collective failures.
-
-Proximity routing does not require routing tables be built using any knowledge about network proximity.
-On the other hand it exploits such knowledge in order to choose the best next hop during routing a message.
-This approach dangerously balances between choosing the node that will further progress the routing towards the destination and choosing the closest entry in the routing table, in terms of network proximity.
-Thus, it is relatively less effective than geographical layout when applied to CAN(-like) implemetations.
-Moreover, the technique has been incorporated into a version of Chord (TODO: PChord | additioanlly check wheather this is true) but seems to increase the overhead of node joins and the size as well as maintenance cost of finger tables.
-\cite{dabek_cfs_2001} proposes a server selection scheme for the Chord DHT, on the domain of proximity routing selection.
-In \emph{CFS}, each node predicts the entire lookup latency as a function of the total number of nodes and the average overlay next routing peer.
-The problem is that it is very difficult to have a clear picture on the total number of nodes and the average hop latency from the local.
-This leads to rough estimations that concequently decreases overall performance.
-
-Finally, the third approach, constructs the routing tables with proximity knowledge count in.
-Tapestry and Pastry's mechanisms of routing table maintenance try to minimize the distance to nodes appearing in a peer's routing table.
-The fact that routing is based on longest node ID prefix match there is not such thing as a message not progressing with each routing hop.
-\cite{castro_proximityp2p_2002} argues on how Pastry exploits proximity neighbour selection in order to create a scheme that is (more) location-aware compared to the other well-known DHTs (CAN, Chord).
-
-\subsection{Landmark Binning}
-The scheme proposed in \cite{ratnasamy_binning_2002} is based on the process of \emph{binning} close by nodes (in terms of network latency) into the same cluster.
-The authors set the following objectives for designing their proposed algorithm:
-\begin{enumerate}
- \item simplicity (minimal support from any measuring infrastructure),
- \item scalablility (no global knowledge of the network),
- \item complete distribution (nodes need no communication-cooperation).
-\end{enumerate}
-
-\paragraph*{Binning scheme} The implementation requires a set of well-known \emph{landmark} machines spread accross the Internet.
-Every newly arriving node measures its distance from these landmarks and unilaterally decides to join a specific bin based on these results.
-In more detail the node measures its round-trip time to each of the landmarks and orders these measurements in a decreasing order.
-The ordering represents a ``bin'', in the sence of close-by nodes having the same landmark ordering and hance belong to the same ``bin''.
-This means that a landmark system consisting of $m$ such nodes results in $m!$ potentional different bins.
-
-The algorithm can be considered scalable, as nodes need only compute distances to a small number of predefined nodes and thus without exchanging any information.
-A potential bottleneck could be the extra load that this ``ping''-like scheme imposes to the landmarks, especially when we need instant reaction from our topology when dealing with the dynamic nature of the p2p networks.
-
-To answer the question as to whether the algorithm actually contributes possitivelly to the construction of an enhaned overlay, the paper defines the \emph{gain ratio} as the factor by which the latency reduces when someone communicates with a random node from the same bin than with one not in the bin.
-This is implemented with an inter-bin to an intra-bin latency ratio.
-
-\paragraph*{Structured and Unstructured Binning} The \emph{binning} scheme can be incorporated within either a structured or an unstructured overlay construction algorithm.
-The authors provide an example for both these cases.
-
-Assuming a structured approach based on CAN\cite{ratnasamy_can_2001} and $m$ landmark nodes.
-The coordinate space is then partitioned into $m!$ equally sized portions, each corresonding to a single ordering of the landmarks.
-To do this, the first dimension is divided into $m$ areas each of which is furtherly divided (second dimension) into $m - 1$ sections and so on.
-Having set this $m$ dimensional space, at joining time, a node measures the delay to the set of landmarks in order to determine its associated bin and thus position itself in that portion of the coordinate space associated with its landmark ordering.
-Even though this scheme can
-
-For unstructured overlays the paper assumes \emph{a set of $n$ nodes where each node picks any $k$ neighbour nodes so that the average routing latency on the resultant overlay is low (assuming shortest path routing)}.
-According to the proposed heuristic algorithm called \emph{BinShort-Long}, a node picks its neighbours by choosing its $\frac{k}{2}$ closest\footnote{If the node's bin is not large enough for it to pick these $\frac{k}{2}$ neighbours, it picks the required nodes from the bin that matches the most in terms of landmark ordering.} ones (named \emph{short links}), using the \emph{binning} scheme and the rest $\frac{k}{2}$ randomly (\emph{long links}).
-The former set produces well-connected \emph{pockets} of nearby nodes while the later preserves the connectivity of the graph, both yielding a proximity factor of $\alpha = 0.5$ in an attempt to preserve the benefitial properties of unstructured topologies\cite{merugu_str2unstr_2003}.
-
-(TODO double check validity) One disadvantage of this landmark scheme is related to the additional burden imposed to the landmark sites.
-The authors claim though that the algorithm requires so little work by the landmarks (maybe just echo to ping messages) that could in effect, act as ``unsuspecting participants''.
-Even if this is the case, the fact that it is not fully distributed, renders the protocol's scalability directly volnerable to any system size increase as well as usuitable for highly dynamic networks such as ad-hoc networks.
-Moreover, fixed points in a network are inherently more exposed to malicius attacks.
-The most significant downside of the algorithm though is that it can lead to an extremely uneven overlay ID distribution causing load inbalances and hot spots.
-Lastly, the scheme is coarse grained when it comes to distinguishing relatively close nodes\footnote{In the worst case, all nodes could ve clustered into a single bin.}.
-
-\subsection{Global Soft-State}
-The approach argued by Xu et.al in \cite{xu_globstate_2003} is to build a global map to help choose shorter routing paths, combining the landmark binning method and small scale distance probes to reveal the proximity properties of the underlying network to the overlay.
-This global view of the state is made available to all nodes in order to help them find the best way to route their messages.
-The authors focus on two aspects in order to accomplish their goal.
-
-Initially, is the \emph{generation} of proximity information and then its \emph{effective exploitation}.
-For the first, a hybrid approach is proposed, which uses landmark clustering as a preprossesing step in order to select a number of potential nearest neighbour candidates and then refine the selection by incorporating an RTT scheme to ultimately choose the closest node.
-For the second, the algorithm chooses a different path from the classic gossiping approaches for constructing and maintaining the overlay.
-It is based on landmark-clustering-based strategic placement of proximity information on the overlay enabling any node to access such information using a landmark number that reflects its physical position in the network.
-For various logical regions\footnote{This might be a high-order zone in the eCAN\cite{xu_ecan_2002} context or a set of nodes sharing a particular prefix in overlays such as Pastry.} maps of physical information are built and published where each node may appear in a maximum of $log\left( N \right)$ such maps..
-To dynamicaly adapt to changing network conditions, a node subscribes to relevant \emph{soft states} that utilize a notification system in order to initialize any necessary neighbour re-selection.
-
-Maintaining several host states at different layers, makes any content migration costly.
-Additionaly, the method does not make any continuing effort to remap the overlay structure after a node successfuly joins, in order to adapt its state to any occurance of condition change.
-Although this approach greatly reduces the routin latency to far nodes, it is unable to dynamicaly identify nodes that are close to routers and gatways in order to construct the secondary overlay.
-Nevertheless, static recognition of such nodes is currently done based on BGP reports and prechosen landmarks, sucrificing the self-organising attribute of traditional DHTs.
-
-\subsection{Self-Adaptive Topology Matching}
-Ren et.all's goal while designing the \emph{SAT-Match} algorithm \cite{ren_satmatch_2004} was to create a protocol that would be decentralised and scalable, fine-grained in detecting changes and adaptively reform in response to the dynamic nature of a classic p2p environment and last, but certainly not least, of low incuring cost.
-The method is in a nutshell, a two-phased iterative process that focuses on local optimizations in order to collectively achieve a global full overlay optimization.
-The iteration is finised when the node detects that it is physicaly close enough to its neigbours so that no additional optimization is needed.
-The paper defines \emph{stretch}, $S = \frac{\bar{L_l}}{\bar{L_p}}$, as a way quantify the \emph{topology match degree} of the constructed overlay, where $\bar{L_l}$ is the average logical link latency while the $\bar{L_p}$ is the average physical link latency.
-
-\paragraph{Probing phase} \emph{SAT-Match} uses a small TTL value for the probing messages in order to reduce redunduncy\cite{jiang_lightflood_2008}.
-This process begins as soon as the node joins the network using a DHT mechanism.
-Each probing message contains information about the source and a small TTL value.
-The recipient of such a message returns information about itself to the source and forwards the probing message to its neighbours if the TTL is non-zero.
-The nodes been covered, are refered to as $TTL-k$ neighbourhood of the source node\footnote{Especialy, $TTL-1$ neighbourhood referes to the source's direct neighbours}.
-The list of responses is used to measure RTT to these nodes and sort the list ascending RTT order.
-
-\paragraph{Jump phase} Blindly selecting the peer with the smallest RTT as neighbour is, generally, not the right desision to make in order to achieve global \emph{stretch} reduction, This is because, in a structured scheme, when a node jumps to connect to a physically close node, it may need to connect to other distant nodes to maintain the structure's integrity, thus creating an overall increase in the overlay's \emph{stretch}.
-The two nodes with the smallest RTT is then used in order to select one zone to jump in this phase.
-The algorithm is as follows: The source node $S$ calculates the stretch change of its $TTL-1$ neighbourhood and that of the $TTL-1$ neighbourhood of the first of the previously selected peers.
-These calculations are made as if the jump has been made.
-If the stretch reduction is over a predefined threshold the jump is performed, otherwise the second selected candidate is picked and the same computations are performed.
-If again, the threshold is not met, then no jump is ultimately done.
-In case of a jump, this is performed as a combination of \emph{leave} and \emph{join} operations, in the CAN context.
-
-Moreover the algorithm takes several issues into consiteration in order to furtherly improve the resulted overlay.
-For example when multiple nodes try to jump simultaneously into a regeon, then the logical link brakes from one attempt may result in inacurate computation of the gain factor, for an other.
-This situation is identified as \emph{contention} and the nodes use an exponatial back-off algorithm to avoid it.
-An other problem is the uneseccary traffic incuerred by the probing phase in a region that after several jumps has settled to a stable state.
-In these cases it is more likely for jump attempts to be proven worthless.
-The algorithm doubles the probing period of such nodes, every time a jump is not taken.
-
-The authors claim that this continuously adaptive mechanism achieves global topology matching optimization in a sufficiently large scope.
-This also secures the fast adaptation to frequent network changes.
-It also considered lightweight and can easily be embedded into current p2p systems, as well as effectively combined with other techniques, such as landmark binning.
-
-\subsection{Delay Aware P2P System}
-A new \emph{Delay Aware P2P System (DAPS)} is introduced in \cite{zhang_daps_2005}.
-Its main goal is to reduce the time $L$ of a look-up request by dividing the routing tables of peers into several sectors in increasing delay.
-The source node emitting the query message defines a delay boundary or the pruning factor $L_t$ in the paper's context.
-Request messages will be forwarded only to nodes whose delay less than or equal to $L_t$.
-With the clustered routing tables and the loose organisation the overlay network of \emph{DAPS} is between structured and unstructured.
-
-\subsection{Mithos}
-Waldvogel and Rinaldi proposed a protocol\cite{waldvogel_mythos_2003} which incorporates a directed incremental probing to find near optimal node placement.
-\paragraph{Neighbour detection.} During bootstraping, the new comming node needs to know how to contact at least one of the existing members.
-A subset of these nodes will be used as the first set of candidate neighbours.
-Then, iteratively, each of the peers in the cadidate set is asked for their neighbours in order for each of the later to be probed for their network distance from the new coming one.
-The closest node is then used as the new candidate neighbour and the process is repeated until no further improvement is detected.
-Due to the fact that the algorithm may ultimately reach a local minima instead of a global one, \emph{Mithos}' approach is to first probe all neighbours within a two hop distance from the current minimum before concluding the process.
-\paragraph{ID assignment.} After finding its first neighbour, the newcoming peer must be assigned its ID, a very crusial selection in order not to create many local minima that will prevent efficient future neighbourhood location.
-\emph{Mithos} uses information gathered during the previous step, in order to compute the ID, which inccurs no further communication costs to the algorithm.
-This information includes
-\begin{inparaenum}[\itshape i\upshape)]
-  \item the two closest nodes and their neigbours, and 
-  \item their coresponding distances.
-\end{inparaenum}
-Using the above, it then assigns coordinates to the new-coming node, so that Euclidian distances between the node and all known hosts predict the network latency between them\cite{cox_vivaldi_2004}.
-These synthetic coordinates as explicitly used as the node's ID and as soon as it has been established, distances can be computed in the ID space, no longer requiring physical measurements.
-\paragraph{Link establishement.} The last step of the algorithm is the interconnection among neighbours.
-\emph{Mithos} uses a \emph{quadrant}-based mechanism according to which each node establishes a link to the closest neighbour in each quadrant.
-During forwarding, the next hop is performed towards a neighbour in the same quadrant as the final destination\footnote{An implementation of this could be backed by a $d$-bit vector indexing (where $d$ is the number of dimensions) into the routing table.
-Then the next hop is identified by computing the difference between the current node's vector and that of the destination's.}.
-The problem is that after the ID assignment process, the new-coming node may not know of other neighbours in all quadrant and if even if it does, this cannot ensure that they are the nearest available.
-Thus, the node first identifies neighbours in all quadrants using a mechanism based on ideas similar to a perimeter walk\footnote{Used in Greedy Perimeter Stateless Routing (GPSR) protocol.} and then using parallel path processing improves the results by taking into account further geometric properties of node relationships.
-
-(TODO review this part)In order to avoid local minima during neighbour detection, extensive probing must be undertaken.
-In simulation, unfortunately, only very small-sized overlay topologies (of 200 to 1000 nodes) have been used and thus no safe conclusions can be made as for the behaviour of an extensively large, real-world p2p deployment of the scheme.
-
-\subsection{DHT-PNS}
-The work in \cite{hancong_pnsbased_2006} propose a proximity neighbour selection scheme on top of the Chord DHT.
-Using the Vivaldi protocol\cite{cox_vivaldi_2004} each node is assigned synthetic $2$-dimension coordinates that can be used to derive network latency using Euclidian distances in the id space and without the need of explicit probing.
-Then the algorithm performs the two steps described in the following paragraphs.
-\paragraph{Space mapping} The space is partitioned using a \emph{concentric circle clustering scheme} where succesive cycles of radiuses $\rho$, $2\rho$, $3\rho$ and so on, are constructed.
-Then the formed annuluses are divided into $2\chi-1$ \emph{sectors}, where $\chi$ denots the sequence number of the annulus starting from $\chi = 1$ for the centre cycle.
-It is proved in the paper, that this way each sector occupies the same area\footnote{The same area as does the center cycle.} and assuming uniform node distribution, this characteristic, favours a more load balanced clustering operation.
-Every sector in this $2-d$ coordinate space is mapped to a unique \emph{region} in the DHT space forming a multi-layer node identifier space.
-Thus, any nodes that belong to the same sector, are mapped to the same region as well, preserving their proximity relationship unveiled by the use of the Vivaldi protocol.
-\paragraph{Routing table optimization} The system described in the previous section allows any node $\alpha$ to obtain its DHT's $key_{\alpha}$ and its region's $key_r$, in a fully distributed manner, just by applying a consistent hash function.
-Using a $Get\left( key_r \right)$ RPC call in Chord, the node can obtain the region's master node, called \emph{Cluster Node (CN)} which is responsible for clustering the nodes belonging in the same sector or region with that of $\alpha$.
-On the other hand, a $Put\left( key_r, info\right)$ RPC call, registers $\alpha$ to its corresponding region and publishes its information, through the special peer CN.
-$\alpha$ can, additionaly, ask CN for other nodes that have previously joined the region in order to add them into its neighbour set for future routing table optimization.
-Even in the case when no neighbour is detected in the current region, the search is expanded to adjasent regions and towards an upper layer identifier space until one is found or the first layer reached.
-
-\subsection{Quasi-Chord}
-The approach that is proposed by Sun and Zhang in \cite{sun_quasi_2008} confronts the topology mismatch problem, in the Chord DHT context, that is created by the fact that no consiteration for the underlying physical network topology is taken into account during the construction of the identifier cycle.
-To construct a \emph{Quasi-Chord} network three steps are needed.
-First, each host acquires its coordinates in the geometric space utilizing the \emph{global network position (GNP)} protocol \cite{ng_gnp_2001}.
-Second, using the \emph{Cantor space filling curve} the $2$-dimensional space is converted to a $1$-dimensional one, used in the last, third, step to build the Quasi-Chord circle according to the host's Cantor value.
-In the following paragraphs, these steps are furtherly discussed.
-\paragraph{GNP coordinates} First of all the host must be positioned in the geometric space.
-The algoritm models the P2P network with a well defined distance function in such a way, it can predict, with high accuracy, the distance between any two points in the space by just evaluating the output of the distance function on the coordinates of these points.
-This is accomplished with the \emph{global network position (GNP)} protocol which\begin{inparaenum}[\itshape i\upshape)]
-  \item creates a reference set of $N$ landmark nodes so as to minimize the error of ICMP measured distances and coordinate computed ones between them, and then
-  \item each host is able to measure its round-trip times to the $N$ Landmarks in order to compute its own coordinates.
-\end{inparaenum}
-\paragraph{Cantor SPF} After the $2$-d coordinates are set, in the next stage of the algorithn, each participating peer is assigned a Cantor value, according to the application of a \emph{space filling curve} on the coordinate space.
-(TODO: add figure).
-This results in the conversion of the $2$-dimensional space to a $1$-dimensional Cantor space that can more easily be more mapped to the Chord hierarchy.
-\paragraph{Quasi-Chord construction} As can intuitively be infered by observing the Cantor chart, close-by nodes in the physical layer, are more likely to have similar Cantor values.
-This attribute is exploited in order to construct a topology-aware identifier space for the Chord DHT.
-The cycle is constructed by sorting nodes in ascending order.
-Each host maintains 2 finger tables, one for clockwise and one for counter-clockwise stepping.
-This helps with the connectivity of the network because its not allowed to connect the first node with the last one since this will incur heavy traffic to the later\footnote{After all their Cantor values denote that they are actually the furthest of each other.}.
-
-(TODO: doublecheck!) The disadvantage of the algorithm is that the coordinate assignment in the first stage is backed by a not fully distributed landmark-based algorithm.
-Moreover the Quasi-Chord model build-up is making an indirect assumption of a maximum number of allowable hosts since it is constructed.
-Last but not least the doubling of the required routing information which needs to be created and maintained is an additional negative point to the efficiency and the scallability of the algorithm.
-
-\subsection{LAPTOP}
-Laptop \cite{wu_laptop_2007}, introduced by Wu et al., organises the overlay into a tree based hierarchy with main focus on child-to-father relationships in order to reduce hops during message routing as well as minimize maintenance overhead.
-Additionally, a caching scheme is also incorporated so as to furtherly reduce routing table update cost.
-The authors argue that Laptop overlay network is bounded by $O\left( log_d N \right)$ hops in a balanced overlay tree, where $N$ is the number of nodes, and $d$ is the maximum degree of each node.
-It utilizes a geographical layout approach  and constructs a geographical layout in a self-organizing and efficient fashion, by estimating the round trip time (RTT) to a small number of nodes in the overlay network in order to make them roughly aware of their physical distances among them.
-
-The protocol is based on four definitions.
-\begin{enumerate}
- \item The amplitude of all possible measured RTTs is devided into intervals.
-Each node measures its distance to its parent and is assigned a label $L_i$ where $i$ denotes the configurable RTT interval in which the measured distance falls into.
-A special kind of node, the root, is initially assigned the $L_1$ label and maintains (as all $L_1$ nodes do) a list of other $L_1$ nodes in the overlay.
- \item Any node can have children with level lower than theirs, except an $L_{max}$ node which can only have $L_{max}$ level children and only in the case when its parent has reached its maximum degree.
- \item Each node is assigned an address in a dotted format (e.g 1.3.4).
-Each octet ranges from $1$ to $d$, where $d$ is the maximum degree of the nodes.
-The assignment process is done by appending a unique octet to the address of its parent.
-Root node is assigned address 1.
- \item For any descendant node $Y$ of a node $X$, the measured distance among each other, must always be less than the lower bound of the RTT interval denoted by $X$'s label.
-\end{enumerate}
-
-The routing scheme is similar to the IP's longest-prefix matching scheme.
-At each forwarding hop, any message travels up the tree until the first common uncestor of source and destination node is reached and then starts descending to arrive to its target.
-During tree traversals, special entries in the routing tables, called \emph{routing cache}, are maintained in order to increase routing efficiency and achieve finer load balance.
-Caching enables a node to forward a message to a better longest-prefix match than that of its direct ancestor making a large, quicker and more cost effective step through the overlay and toward the destination.
-To improve scalability, the number of children nodes and the size of the routing cache are limited.
-
-In terms of overlay maintenance, Laptop incorporates a simple \emph{heartbeat}-based technique where only the parent node is responsible for monitoring its children.
-
-At join process the newcomming node is assigned its level label as well as its address by its parent node.
-Additionaly it initializes its routing table (with normal and caching entries) as it traverses the overlay in search for its parents node.
-The newcommer $N$ locates the root node and the later responds with a list of $L_1$ nodes.
-$N$ then probes each of the $L_1$ nodes in search for the closest one.
-If the measured RTT to the closest $L_1$ falls into the first interval then the newcommer becomes a $L_1$ node as well.
-Otherwise node $N$ sets the closest $L_1$ node as its potential parent node.
-This potential parent becomes the actual parent if it does not have any other children.
-If it has, $N$ gets a list of these $L_i$ nodes and by measuring the RTT to each of them tries to spot a new potential parent in order to repeat the above process.
-
-During a gracefull departure, the refered node checkes for children in the overlay.
-If it does not have any, it simply notifies its parent and leaves.
-If it has, it selects the child node with the lowest RTT to it in order to take its place so that the locality property is preserved.
-
-In case of an arbitrary failure, the children of the failed node are detecting their parent's absent when the later stops acknowledging their heartbeat messages.
-They start emmiting special messages to their grand parent node\footnote{The address of the grandparent node is stored during join process}.
-In case of no response from the grandparent node, children invoke the joining procedure to detect their new parent node.The parent of the failed node, aggregates the notifications from the above children nodes for a period of time, and then chooses the one with the lowest level label as the takeover node.
-Potential ties are broken by favouring the lower RTT.
-The parent of the failed node, finally informs accordingly all the children about the change in the hierarchy.
-
-\subsection{IP-based Clustering (IPBC)}
-\emph{IP-based clustering} \cite{karwaczynski_ipbc_2007} is a proximity neighbour selection technique that is based on a longest IP prefix matching scheme in order to quantify the proximity among peer nodes.
-The relation of the IP address and the physical location of a node in the topology of the Internet, is intuitevely realized by operation the network layer of the Internet stack.
-For example, nodes in the same local subnetwork share the same 3 bytes of their address.
-Reports in \cite{freedman_iploc_2005} state that almost $97\%$ of prefixes longer than 3 bytes correspond to address assignments at a single geographic location.
-Moreover, observing the \emph{IANA IPv4 Address Space Registry} we can infer that, since blocks of consecutive octet prefixes are assigned to the same Regional Internet Registries (RIRs), nodes inside will be physicaly close to each other.
-
-The implementation of this paper is based on the above sceptic in order to create proximity information.
-This information is stored in a decentralised manner, within the overlay itself, just like any other resource would have been.
-In order to be published in the overlay, each node is first assigned a unique identifier and subsequently generates a key by hashing a fixed-length prefix of its IP.
-Authors argue on the prefix length to be used, as there is a thin balance between reducing the propability of finding closer nodes by adopting a long prefix and overload nodes that are responsible for information published by many peers when chosing to hash a short one.
-Their verdict is for the use of a 16-bit wide prefix for real world systems deployed in an Internet scale.
-In either case, ID and IP is then stored in the DHT using this generated key.
-This way, any node, at any time (either at join time or during peer lifetime for topology adaptation) can acquire information about close-by nodes just by quering the DHT for a specific key.
-Moreover, the algorithm takes care of the freshness of the proximity information in two ways.
-First, the advertising nodes themselves periodically update their advertisements or when they voluntarily leave the overlay, they explicitly remove their data.
-On the other hand, in case of a failure each publication is assigned an expiration time, and thus ultimately removed by the DHT maintenance mechanisms.
-
-\subsection{CHOord considering Proximity on IPv6 (CHOP6)}
-\cite{morimoto_chop6_2007} roughly estimates the proximity among nodes by exploiting the IPv6 address format as well as RTT information if this is available.The first is achieved by introducing a 64-bit ID scheme in which the least significant bit part\footnote{The exact range is a predefined system variable} is the IPv6 global routing prefix and thus enabling a longest prefix match scheme.
-The observation in which the protocol is based roots from the IPv6 address block assignment.
-Specicaly, block of /16 or /23 in size are assigned to Regional Internet Registries (RIRs) by Internet Assigned Numbers Authority (IANA).
-An RIR divides the signed address blockes into smaller address blocks which are furtherly assigned to Natinal Internet Registries (NIRs).
-That way, it is possible to estimate a node's geographical location by simply examining the upper 32-bits of its IPv6 address.
-Moreover, similar to Chord, CHOP6 uses a finger table, whose entries hold more than one candidate nodes.
-There are three cases in which a node chooses the next hop according to information it posseses.
-\begin{itemize}
- \item When no information about RTT is available candidate next hops, the sender just selects a node in the finger table entry which shares the longest prefix with the destination.
- \item There is another possibility that after some communication with other nodes, the source node should know of the RTTs to some of the nodes in finger table entry.
-In this case, the source node chooses the one with the smallest RTT with propability $p$.
-If there is a node with no measured RTT then the sender can select such a node with probability $1 - p$
- \item In this last case the source node has already communicated with all candidate node in the finger table entry.
-Thus, the node selects the node whose RTT is the smallest with probability $p$, the one with the second smallest with probability $q$ and so on, where $0 \leq \ldots < r < q < p < 1$ and $p+q+r+\ldots \leq 1$
-\end{itemize}
-
-\subsection{Proximity in Kademlia}
-In \cite{kaune_pkad_2008}, Kaune et al., studied the Kademlia DHT, in order to build a proximity aware scheme that would work in the context of iterative lookup algorithms.
-The protocols discussed, focus on both the overlay and underlay tiers in the sense that in the first improves the routing performance according to a cost metric that is provided by the later.
-This metric is tailored to the needs of a specific aplication.
-For example, routing could be optimised to maximise the within-ISP traffic or reduce the lookup latencies or even avoid contacting untrustworthy subnetworks.
-
-Two algorithms are presented for the overlay optimization.
-One is \emph{Proximity Neighbour Selection (PNS)} and one for \emph{Proximity Route Selection (PRS)}.
-The first aims at keeping peers with the least contact cost in the routing table.
-As Kademlia constantly learns new peers (incoming requests, through iterative lookup) no special algorithm for searching more cost effective peers is necessary and simply choose the best peers seen so far.
-The later (PRS) aims at choosing the best next hop during routing a message.
-Due to the fact that the routing in Kademlia is iterative, it is the initiator of the lookup that chooses each next hop from a set of candidate nodes.
-The ``vanilla'' Kedemlia always chooses the closest node with respect to the XOR metric but PRS Kademlia chooses the one with the smallest underlay metric cost.
-Thus, as in all such approaches, there is a trade-off between the overlay and underlay distances.
-
-An underlay metric provides information about the underlay network. The paper distingushes between three kinds of possibilities to quantify it:
-\begin{itemize}
- \item Using measurements gained by previous lookups
- \item Using measurements acquired by other peers or jointly calculated with others
- \item Using a local database to look up information
-\end{itemize}
-In the paper, for locality of traffic and subsequently for reducing the lookup latency, a clustering scheme has been implemented exploiting information stored in a geographic information database (GeoIP,MaxMind) about peers.
-The goal of such a metric is to constrain the largest portion of communication within the limits of a peer's ISP.
-Another incorporated approach is Vivaldi\cite{cox_vivaldi_2004}.
-At first all coordinates are random but as peers start to communicate they calculate RTTs as well, gradually updating their own coordinates.
-Vivaldi creates a system where a peer knowing the coordinates of a another, can approximate communication latency without the need of additional prompting.
-
-\subsection{Cone}
-Cone\cite{wang_cone_2007} uses a two-layered identifier space.
-The first, named Chord-layer identifier, denoted as $Id_{Chord}$, is the same as in ``vanilla'' Chord.
-The second is the Cone-layer identifier, $Id_{Cone}$ which is constructed by two component identifiers.
-The first, known as \emph{group identifier (gid)} denotes a relevant group the node belongs to while the second, namely \emph{local identifier (lid)} indicates the local identifier within the group.
-The group concept, which is introduced here, is a way of dividing nodes according to a common $Id_{Chord}$ prefix.
-The structure of a Cone overlay, retains the Chord's circular topology.
-The differce lies on the fact that, now, two rings are created.
-A big ring, where nodes with the same \emph{gid} are arranged at each position.
-Each of these positions are a smaller ring for the particular group's \emph{lid}s.
-The routing is achieved in both clockwise and anti-clockwise direction in the big-ring.
-For this reason two routing tables are maintained, namely \emph{front} and \emph{back} finger tables, respectively.
-Entries in these tables, display physical network proximity with the current node.
-Moreover, a third table called \emph{group} table maintains information about other online peers within the current node's group in a way that entries are now close in the ID space.
-
-Routing in Cone, comprises of the inter-group algoritm and the intra-group algorithm.
-First the group of nodes to which the desired key lies is detected, exploiting physical proximity information (front and back finger tables).
-Second during the next and last hop, the message is forwarded to the exact node that contains the desired key.
-
-Cone uses Landmark+RTTs to generate proximity information and exploits this information using proximity neighbour selection.
-
-\subsection{DynaMO}
-In their work in \cite{winter_dynamo_2004}, Winter et al, try to build an overlay network, called \emph{DynaMO}, that tries to consider not only the physical proximity of peers but their mobility attributes as well.
-DynaMO is based on a Pastry overlay network.
-This was an intentional choice from the authors because Pastry's built-in locality heuristics are thoroughly analised in the literature \cite{castro_proximityp2p_2002} providing good backround against which to test and compare their results.
-In order to adapt to mobile, ad-hoc networks, though, special care has been put to the maintenance of an even overlay ID distribution so that hot-spots are avoided.
-Pastry assignes them in a randomised fashion and then tries to consider proximity through the joining process and routing table maintenance.
-On the other hand, DynaMO dictates a newcoming node to gather information concerning its physical neighbourhood and uses it to assign itself an appropriate overlay ID.
-
-DynaMO tries to capitalize an observation of Pastry's routing mechanism.
-Each node's table consists of rows equal to the number of digits of the overlay IDs and columns equal to the ID's base.
-As we go down the table rows, the matching prefix between the current node's ID and the row's entries increases by one.
-Thus, the leaf set contains the closest nodes in the ID space.
-Additionally as the prefix match increases by one, the result is exponentially less candidates that can fill the tables entries as the row increases.
-This leads to the observation \cite{antony_pastry_2001, castro_proximityp2p_2002} that from overlay routing hop to overlay routing hop the physical distance between nodes is likely to increase leading to a dominating last routing step in terms of the overall plysical routing path distance travelled during a key lookup.
-Since the last routing step is usually taken from the leaf set, DynaMO focuses on making this last step as physicaly close as possible.
-In this context, two approaches are considered namely \emph{Random Landmarking (RLM)} and \emph{Closest Neighbour Prefix Assignment (CNPA)}.
-
-Random Landmarking uses the overlay lookup mechanism to locate nodes responsible for a fixed set of carefully chosen\footnote{In a way that the ID space is divided into equal portions} landmark keys.
-This means that when a node is assigned a landmark key then it plays the role of a temporary landmark for the network.
-Any joing node will measure its distance to that and all other landmarks and assign to itself an ID consisting of a prefix taken from the closest landmark and a remainder that could be assigned randomnly or generated by mechanism that takes into account physical neighbourhood.
-The legth of the prefix can be determined as $prefix_length=|log_b k|$, where $b$ is the ID base and $k$ the number of landmarks.
-The above scheme results in physically close nodes, forming regions of common ID prefix that are likely to be close to each other in the ID space and thus bringing a node's leaf set closer to itself.
-Additionally, the advantage of dynamic landmark nodes is that the failed ones can instantly be replaced by new responsibles, that share similar physical attributes to the failed and thus preserving the balance of the network.
-
-RLM may incur more traffic especially on landmark nodes, load that in some cases is not acceptable.
-CPNA on the other hand takes advantage of Pastry's specification that a new coming node is bootstraped by a physicaly close node.
-The new comer then assumes the ID prefix of that neighbour while the rest is generated similarly to RLM.
-Unfortunately, less overhead comes at the expense of being more coarse-grained.
-
-Both algorithms, are protected against the formation of physical landmark clusters or imbalanced ID distribution\footnote{More common during the initial formation of an overlay network} , by introducing the \emph{landmark gravitation range} as a theshold over which landmark keys are reassinged (for the RLM approach) or unutilized ID prefix ranges are detected and used (for the CPNA scheme) in order to balance the distribution of regions in the overlay.
-
-\subsection{PChord}
-\emph{PChord}\cite{hong_pchord_2005} is a based on the Chord DHT which adds proximity routing into its routing mechanism.
-The main modification over the ``vanilla'' Chord is the inclusion of a \emph{proximity list} into its routing table so that the next hop is decided based not only by considering best progress towards the key, but on the physical proximity of the candidate nodes as well.
-The list is, at join time, empty but as the PChord node starts to interact with other nodes it applies a heuristic mechanism to fill up the list.
-Entries are dynamicaly added or removed as the network state is constantly changing.
-Because, it is guaranteed that at each routing step there is progress towards the target node in the ID space, PChord will result in lower hop number than Chord, for each hop is larger or at least equal in key space in PChord than in Chord.
-Additionaly, passing through proximity links in the underlay means reduced routing cost.
-Moreover, if the number of entries in the proximity list is the same as the number of network partitions, PChord prevents hops from jumping back to the same network the current node belongs to.
-
-\subsection{AChord}
-\emph{AChord}\cite{dao_achord_2006} uses IPv6's anycast functionality in order to
-\begin{itemize}
- \item releave the protocol from complex joining procedures, and
- \item achieve high accuracy of network proximity giving high routing efficiency,
-\end{itemize}
-all with a simple and lightweight mechanism that requires few changes to Chord without affecting its own advantageous characteristics.
-Anycast delivers a message comming from the outside of an anycast group to the physicaly closest node in that group.
-AChord organizes all nodes participating in the overlay network into an anycast group.
-Any joining node comes outside of that group and, thus, is automaticaly forwarded to the physicaly nearest node in order to bootstrap, avoiding
-\begin{inparaenum}[\itshape i\upshape)]
-  \item the need to, explicitly, maintain such nodes, and
-  \item the effort of finding a way of locating the physically nearest among them.
-\end{inparaenum}
-The ID of the new comming node is computed based on the ID of the bootstrap node and bootstrap's predecessor in a way that its ID will position itself between the formentioned.
-
-After joining, the finger table is build the same way as in the Chord protocol.
-Moreover, additional nodes are maintained in a structure called \emph{neighbourship table} which stores information about the closest known nodes\footnote{As a first entry, at bootstrap, the node stores information about the bootstrap node.}.
-The routing decision is made by using both the neighbourship and the finger table.
-
-\subsection{Chord6}
-\emph{Chord6} \cite{xiong_chord6_2005} is a Chord variant that tries to exploit the hierarchical features of IPv6 in order to create a substrate that reduces interdomain trafic and does that in a cost efficient way.
-Chord6 bareley modifies the original Chord protocol, only in the part of identifier definition.
-This, renders the aproach easily portable to other DHTs such as CAN, Pastry and Tapestry.
-In Chord6 the identifier contains two pares: the higher bits are obtained by hashing the  node's IPv6 address prefix of specific length, while the remaining lower bitss are the hash value of the rest of that IPv6 address.
-This way nodes in a domain will be mapped onto a contihuous key space on the overlay network.
-This guarantees that messages are routed in a way that they do not hop in and out of domains many time, thus minimizing overall routing cost.
-(TODO: Recheck validity) However, even though nodes in the same domain would have close identifiers, nodes in two close domains may have very different ones.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-Peer-to-peer architecture has been in the center of research attention in the last decade.
-Especially the decentralized unstructured genre exploits the advantages of loose coupling and self organization of computing nodes to form application-layer networks on top of the physical, best-effort infrastructure of the Internet that exhibit intresting properties.
-Scalability problems arouse quickly, though, because of the inefficient construction of this overlay that was built with no concern for the underlying physical network that causes a great deal of redundant traffic.
-The problem was identified by the research community as the topology mismatch problem between the overlay and the corresponding underlying physical network and a great deal of effort has been set towards alleviating it.
-Some fruits of this effort have been gathered and presented in this survey.
-Measurement of link cost through latency or RTT and deletion of inefficient established connections are the key concepts of almost all approaches.
-Others, manage to address the problem through hierarchical peer clustering (e.g.
-best IP matching).
-Additionally, some protocols focus on specific problems that furtherly arise, such as overlay partitioning, search scope reduction or convergence speed, to name just a few.
-Another desire is to form a protocol that could be applied to both decentralized unstructured and decentralized structured networks.
-Unfortunately no approach has equally addressed these problems in order to form a robust solution.
-So the field seems to be, still, fertile for any, new, clever idea.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliographystyle{plain}
-\bibliography{bib/p2p-overlays}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\end{document}
+Peer-to-peer architecture has been in the center of research attention in the
+last decade. Especially the decentralized unstructured genre exploits the
+advantages of loose coupling and self organization of computing nodes to form
+application-layer networks on top of the physical, best-effort infrastructure of
+the Internet that exhibit interesting properties. Scalability problems arouse
+quickly, though, because of the inefficient construction of this overlay that
+was built with no concern for the underlying physical network that causes a
+great deal of redundant traffic. The problem was identified by the research
+community as the topology mismatch problem between the overlay and the
+corresponding underlying physical network and a great deal of effort has been
+set towards alleviating it. Some fruits of this effort have been gathered and
+presented in this survey. Measurement of link cost through latency or RTT and
+deletion of inefficient established connections are the key concepts of almost
+all approaches. Others, manage to address the problem through hierarchical peer
+clustering (e.g. best IP matching). Additionally, some protocols focus on
+specific problems that furtherly arise, such as overlay partitioning, search
+scope reduction or convergence speed, to name just a few. Another desire is to
+form a protocol that could be applied to both decentralized unstructured and
+decentralized structured networks. Unfortunately no approach has equally
+addressed these problems in order to form a robust solution. So the field seems
+to be, still, fertile for any, new, clever idea.

### 2014-01-21 intro.tex
2012.05.29 (Vassilis) Monolithic file broken down to multiple smaller ones (per section) for better parallel contribution. Additional material on more recent algorithms for unstructured networks. Now also covering the BitTorrent protocol. Merged Huseyin's figures from previous contribution and code refactored for better referencing. Fixed bibliography bug (references appearing as ?).

--- 2014-01-21 mard-survey.tex
@@
-%\documentclass[a4paper,10pt]{article}
-\documentclass[acmcsur,acmnow]{acmtrans2m}
-%\documentclass[acmtocl,acmnow]{acmtrans2m}
-
-%\acmVolume{V}
-%\acmNumber{N}
-%\acmYear{YY}
-%\acmMonth{Month}
-
-% packages
-\usepackage[utf-8]{inputenc}
-\usepackage{paralist}
-\usepackage{subfigure}
-\usepackage{graphicx}
-\usepackage{colortbl}
-\usepackage{color,soul}
-\usepackage{longtable}
-\usepackage{pdflscape}
-
-% definitions
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{conjecture}[theorem]{Conjecture}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{lemma}[theorem]{Lemma}
-\newdef{definition}[theorem]{Definition}
-\newdef{remark}[theorem]{Remark}
-
-%\newtheorem{theorem}{Theorem}[section]
-%\newtheorem{definition}[theorem]{Definition}
-
-%\textwidth      6.5in
-%\textheight     9.2in
-%\oddsidemargin  -0.5cm
-%\evensidemargin -0.5cm
-%\parindent      0pt              
-%\topmargin      -0.3in
-%\headsep        20pt
-%
-%\parskip        5pt
-%\parindent      10pt
-
-\markboth{}{}{}
-
-\title{Alleviating the Topology Mismatch Problem in Unstructured and Structured Overlay Networks: A Survey}
-
-\author{
-VASSILIS S. MOUSTAKAS\\\emph{National and Kapodistrian University of Athens}\\
-H\"USEYIN AKCAN\\\emph{Izmir University of Economics}\\
-MEMA ROUSSOPOULOS\\\emph{National and Kapodistrian University of Athens}\\
-AND\\
-ALEX DELIS\\\emph{National and Kapodistrian University of Athens}\\
-%  Vassilis Moustakas$^1$, H\"useyin Akcan$^2$, Mema Roussopoulou$^1$ and Alex Delis$^1$\\
-%  $^1$Department of Informatics and Telecommunications,\\
-%  National and Kapodistrian University of Athens\\
-%  \mbox{\texttt{\{b.moystakas, mema, ad\}@di.uoa.gr}}\\
-%  $^2$Department of Software Engineering,\\
-%  Izmir University of Economics, Izmir, Turkey \\
-%  \mbox{\texttt{huseyin.akcan@ieu.edu.tr}}
-}
-
-\begin{abstract}
-Peer-to-peer (P2P) systems is a rapidly growing application architecture.
-The overlay network abstraction on top of a best-effort infrastructure provides a powerful tool to realize a plethora of ever-wanted exotic features like anonymity, high availability and robustness, load balancing, quality-of-service (QoS) at the application layer and many more.
-Unfortunately weaknesses of early effectuations constrained applications from unleashing the full potential of the paradigm.
-One of them is the topology mismatch problem between the overlay and the physical underlying network, a problem that imposes a huge amount of unnecessary stress to the network resources and, thus, impede applications from scaling.
-In this survey we investigate on the recent research developments towards the alleviation of this problem, in both unstructured and structured overlay architectures.
-We visit the various algorithms available in the lietrature, focusing on the presentation of how each of them tries to tackle the problem.
-Additionally, we provide a finer grained categorisation of the approaches based on the generic type of methodology they use in order to accomplish this goal.
-% TODO: Review this after the comparison sections are fixed.
-Finally we discuss the advantages, disadvantages and novelties introduced by the algotithms and offer pictorial comparisons for them in order to further strengthen and organise understanding of the information and knowledge cited, in an attempt to make this survey a one-stop-shop reference for anyone interested in the topic, from researchers and students, to system architects and developers.
-\end{abstract}
-
-\category{}{}{}
-
-\terms{Topology mismatch, distributed, peer-to-peer, P2P, overlay}
-
-\keywords{Topology mismatch problem, distributed algorithm, peer-to-peer architecture, overlay network }
-
-\begin{document}
-
-\begin{bottomstuff}
-SAY SOMETHING COOL HERE...
-\end{bottomstuff}
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Distributed architectures have witnessed a significant increase of interest during the last decade, as the number of interconnected computers increased, the available digital information kept proliferating, at an exponetial rate each year, and problems to be addressed by IT became more and more demanding.
-Why is there a correlation, and it is not just a coincidence?
-Distributed computation suggests a simple paradigm\footnote{Of course, as always, this is the case only after someone thought about it first!}.
-Get a set of autonomous computing entities and put them to cooperate with each other in order to fullfil a well defined task.
-In the World of the Internet and the World Wide Web this also offers a thrilling promise.
-To harness the immence power scattered across the Internet and get amply rewarded by exploiting the data, storage and computational resources out there.
-
-In order to achieve that, computers collaborate by exchanging messages via an application-level network where they can address each other in the specific context of the task at hand.
-This communication structure is refered to as an \emph{overlay network} which is an abstract, logical inter-connection formed on top of another network\footnote{The Internet itself is an overalay on top of the telephone network.}.
-It consists of a set of nodes, the computing elements, that are being connected by virtual (logical) point-to-point links, perhaps corresponding to multiple physical links.
-Figure~\ref{figure:overlay} illustrates a simple four-node overlay contructed upon a wide area network.
+\label{section:intro}
+
+Distributed architectures have witnessed a significant increase of interest
+during the last decade, as the number of interconnected computers increased, the
+available digital information kept proliferating, at an exponetial rate each
+year, and problems to be addressed by IT became more and more demanding. Why is
+there a correlation, and it is not just a coincidence? Distributed computation
+suggests a simple paradigm\footnote{Of course, as always, this is the case only
+after someone thought about it first!}. Get a set of autonomous computing
+entities and put them to cooperate with each other in order to fullfil a well
+defined task. In the World of the Internet and the World Wide Web this also
+offers a thrilling promise. To harness the immence power scattered across the
+Internet and get amply rewarded by exploiting the data, storage and
+computational resources out there.
+
+In order to achieve that, computers collaborate by exchanging messages via an
+application-level network where they can address each other in the specific
+context of the task at hand. This communication structure is refered to as an
+\emph{overlay network} which is an abstract, logical inter-connection formed on
+top of another network\footnote{The Internet itself is an overalay on top of the
+telephone network.}. It consists of a set of nodes, the computing elements, that
+are being connected by virtual (logical) point-to-point links, perhaps
+corresponding to multiple physical links. Figure~\ref{figure:overlay}
+illustrates a simple four-node overlay contructed upon a wide area network.
 
@@
 
-The abstraction of the overlay network has been proposed as a way to implement efficient, fully distributed, application-layer services.
-For example, backing the routing of messages to destinations that are not known in advance or building one-to-many addressing schemes
-% TODO find citation
-%[Narada, Adobe's Real Time Media Flow Protocol (RTMFP)]
-that would provide higher quality-of-service (QoS) guarantees.
-\emph{IP multicast}, as well as other proposals such as \emph{IntServ} and \emph{DiffServ} \cite{cisco_diffserv_2005}, have not seen wide acceptance (yet?), largely due to the fact they require support from the physical infrastructure.
-On the other hand, an overlay network can be deployed on end-systems running the overlay protocol software, without cooperation from the ISPs.
-%Academic research includes \cite{chu_esm_2000,jannotti_overcast_2000,kwon_tag_2002} among others.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The abstraction of the overlay network has been proposed as a way to implement
+efficient, fully distributed, application-layer services. For example, backing
+the routing of messages to destinations that are not known in advance or
+building one-to-many addressing schemes
+% TODO find citation for Narada & Adobe's Real Time Media Flow Protocol (RTMFP)
+\cite{} that would provide higher quality-of-service (QoS) guarantees. \emph{IP
+multicast}, as well as other proposals such as \emph{IntServ} and
+\emph{DiffServ} \cite{cisco_diffserv_2005}, have not seen wide acceptance
+(yet?), largely due to the fact they require support from the physical
+infrastructure. On the other hand, an overlay network can be deployed on
+end-systems running the overlay protocol software, without cooperation from the
+ISPs.
+%Academic research includes
+%\cite{chu_esm_2000,jannotti_overcast_2000,kwon_tag_2002} among others.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Peer-to-Peer (P2P) Architectures in a Nutshell}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Overlay networks on top of the best-effort IP layer, formed a widely ranged family of protocols collectively known as \emph{peer-to-peer} or \emph{P2P} for short.
-P2P systems consist of a number of networked computers called \emph{peers}, where each, such, \emph{peer} could act both as a resource provider and a resource consumer.
-This changed the traditional \emph{client-server} model dominating the Internet and lead to the introduction of the \emph{servent} concept \cite{gnutella}, a \emph{portmanteau} that blends the notions of server and client to denote the twofold role of the P2P network participants.
-Such serverless systems, proved to be able to achieve outstanding aggregate resource capacities as more and more participants join the system without requiring additional expenditure for infrastructure\footnote{
-        Unfortunatelly, there has been observed mitigation of this self-scaling property by the undesirable behaviour of participating users, usually answered in popular file-sharing systems; from \emph{free-riding} \cite{saroiu_measurefileshare_2002, adar_gnutellafreeriders_2000, hughes_gnutellafreeride_2005} to the distribution of illegal content and/or other \emph{socio-technical} issues \cite{hughes_socp2p_2008}.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Overlay networks on top of the best-effort IP layer, formed a widely ranged
+family of protocols collectively known as \emph{peer-to-peer} or \emph{P2P} for
+short. P2P systems consist of a number of networked computers called
+\emph{peers}, where each, such, \emph{peer} could act both as a resource
+provider and a resource consumer. This changed the traditional
+\emph{client-server} model dominating the Internet and lead to the introduction
+of the \emph{servent} concept \cite{gnutella}, a \emph{portmanteau} that blends
+the notions of server and client to denote the twofold role of the P2P network
+participants. Such serverless systems, proved to be able to achieve outstanding
+aggregate resource capacities as more and more participants join the system
+without requiring additional expenditure for infrastructure\footnote{
+  Unfortunatelly, there has been observed mitigation of this self-scaling
+  property by the undesirable behaviour of participating users, usually answered
+  in popular file-sharing systems; from \emph{free-riding}
+
+\cite{saroiu_measurefileshare_2002,adar_gnutellafreeriders_2000,hughes_gnutellafreeride_2005}
+to the distribution of illegal content and/or other \emph{socio-technical}
+issues \cite{hughes_socp2p_2008}.
 }.
-P2p network systems gained extra attention whithin the distributed systems super-set, circa 2000, due to the fact that they succesfully supported popular applications that offered file sharing among vast numbers of participating users.
-But their development didn't stay still.
-On the contrary, during the last decade, research on P2P systems was constant and fruitfull generating a wide range of popular protocols, networks and applications.
-As P2P systems evolved three main architectures have been emerged, namely
+P2p network systems gained extra attention whithin the distributed systems
+super-set, circa 2000, due to the fact that they succesfully supported popular
+applications that offered file sharing among vast numbers of participating
+users. But their development didn't stay still. On the contrary, during the last
+decade, research on P2P systems was constant and fruitfull generating a wide
+range of popular protocols, networks and applications. As P2P systems evolved
+three main architectures have been emerged, namely
 \begin{inparaenum}[\itshape i\upshape)]
@@
 
-\emph{Centralized} architectures (see Figure~\ref{figure:p2p-archs:centralised}) were the first to recognize that requests for resourses (e.g. CPU cycles \cite{seti} or popular content 
+\emph{Centralized} architectures (see Figure~\ref{figure:p2p-archs:centralised})
+were the first to recognize that requests for resourses (e.g. CPU cycles
+\cite{seti} or popular content
 % TODO find citation for Napster
-\cite{}) need not be sent to a dedicated server but instead could be handled by the many hosts that already posses them.
-This schema, though, is not fully distributed since it dictates a centralised searching facility, a characteristic that can be proven its ``Achilles' heel'' in terms of being a scalability bottleneck, a central point of failure and vulnerable to malicious acts (e.g.
+\cite{}) need not be sent to a dedicated server but instead could be handled by
+the many hosts that already posses them. This schema, though, is not fully
+distributed since it dictates a centralised searching facility, a characteristic
+that can be proven its ``Achilles' heel'' in terms of being a scalability
+bottleneck, a central point of failure and vulnerable to malicious acts (e.g.
 \emph{Denial-of-Service (DoS) attacks}).
 
-The centralized approach has been eventually replaced by architectures that distributed both search and download capabilities (see Figures~\ref{figure:p2p-archs:unstructured-full} and \ref{figure:p2p-archs:unstructured-hierarchical}).
-In these \emph{decentralized} architectures, file placement is random, which means there is no correlation with the network topology whatsoever \cite{yang_improvep2psearch_2002}.
-For this reason they are also refered to as \emph{unstructured}.
-The most important properties of such systems are that they support inherent heterogeneity of peers, are highly resilient to peer failures, and incur low maintenance overhead at handling the dynamics of peer participation \cite{stutzbach_churn_2006}.
-In the literature, they are also known as \emph{broadcast-based} systems, because they use \emph{message flooding} among peers or super-peers to propagate queries
-%\footnote{The life of each query message is constrained by an integer, refered to as \emph{time-to-live (TTL)}.
-When a querry is received by a (super-) peer, this value is decreased by one unit.
-Then, the receiver, back-propagates its results and if the TTL has not reached zero the node forwards the query message to all its neighbours except from the one it received it from.}
-.
-
-Recently, \emph{decentralized structured} schemes have been proposed (see Figure~\ref{figure:p2p-archs:structured}) in order to provide a self-organising infrastructure for large-scale P2P applications \cite{ratnasamy_can_2001, stoica_chord_2001, antony_pastry_2001, zhao_tapestry_2001, maymounkov_kademlia_2002, rgrk_bamboo_2004}.
-They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes through a deterministic mechanism.
-Even though unstructured networks became highly popular among file sharing applications, they do have a certain limitation in locating rare objects.
-The DHT approach of structured networks on the other hand, provides a guaranteed bound on the number of overlay routing hops that have to be taken in order for an object to be located, even in the case where only a single copy exists in the system.
-This means $O \left ( log n \right )$ hops, compared to unstructured networks that require $O \left ( n \right )$ to reliably locate a specific object\footnote{Unfortunatelly this efficiency does not come without cost.
-Structured overlays realise what is called object identifier location as opposed to content-based retrieval of unstructured ones.
-This means that the later can support more versatile object location mechanisms like keyword and partial matching searches.}.
+The centralized approach has been eventually replaced by architectures that
+distributed both search and download capabilities (see
+Figures~\ref{figure:p2p-archs:unstructured-full} and
+\ref{figure:p2p-archs:unstructured-hierarchical}). In these \emph{decentralized}
+architectures, file placement is random, which means there is no correlation
+with the network topology whatsoever \cite{yang_improvep2psearch_2002}. For this
+reason they are also refered to as \emph{unstructured}. The most important
+properties of such systems are that they support inherent heterogeneity of
+peers, are highly resilient to peer failures, and incur low maintenance overhead
+at handling the dynamics of peer participation \cite{stutzbach_churn_2006}. In
+the literature, they are also known as \emph{broadcast-based} systems, because
+they use \emph{message flooding} among peers or super-peers to propagate
+queries.
+%\footnote{The life of each query message is constrained by an integer, refered
+%to as \emph{time-to-live (TTL)}. When a querry is received by a (super-) peer,
+%this value is decreased by one unit. Then, the receiver, back-propagates its
+%results and if the TTL has not reached zero the node forwards the query message
+%to all its neighbours except from the one it received it from.}
+
+Recently, \emph{decentralized structured} schemes have been proposed (see
+Figure~\ref{figure:p2p-archs:structured}) in order to provide a self-organising
+infrastructure for large-scale P2P applications
+\cite{ratnasamy_can_2001,stoica_chord_2001,antony_pastry_2001,zhao_tapestry_2001,maymounkov_kademlia_2002,rgrk_bamboo_2004}.
+They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes
+through a deterministic mechanism. Even though unstructured networks became
+highly popular among file sharing applications, they do have a certain
+limitation in locating rare objects. The DHT approach of structured networks on
+the other hand, provides a guaranteed bound on the number of overlay routing
+hops that have to be taken in order for an object to be located, even in the
+case where only a single copy exists in the system.  This means
+$O \left ( log n \right )$ hops, compared to unstructured networks that require
+$O \left ( n \right )$ to reliably locate a specific object\footnote{
+  Unfortunatelly this efficiency does not come without cost. Structured overlays
+  realise what is called object identifier location as opposed to content-based
+  retrieval of unstructured ones. This means that the later can support more
+  versatile object location mechanisms like keyword and partial matching
+  searches.}.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Applications and Implications}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The most successful incarnation of the centrilised approach is without a doubt the popular file-sharing application \emph{Napster}\footnote{Pioneering, though, was the famous \emph{SETI@Home Project} \cite{seti}, which pools spare processing power of participants to search the universe for the existence of extra terrestrial intelligence!}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The most successful incarnation of the centrilised approach is without a doubt
+the popular file-sharing application \emph{Napster}\footnote{Pioneering, though,
+was the famous \emph{SETI@Home Project} \cite{seti}, which pools spare
+processing power of participants to search the universe for the existence of
+extra terrestrial intelligence!}
 % TODO find citation for Napster again
 \cite{}.
-Napster maintained a \emph{central index server} based on file lists provided by participating peers.
-The central index server was queried by the users and it returned pointers to the actual content.
-Thus, by centralizing search while distributing downloads, Napster achieved a highly functional design that was widely acknowledged, at the time, as ``the fastest growing Internet application ever'' that toped some 26 million users and 80 million songs \cite{jmm_naptopusage_2001}.
-
-Napster was ultimatelly brought down after Recording Industry Association of America (RIAA) filed a lawsuit for copyright infringement of the intellectual property of artists whose music was distributed by the network\footnote{
-      The sueing spree was initiated by heavy metal giants \emph{Metallica} that feared Napster was undermining record sales.
-Opposed opinions instantly emerged  with best example the british alternative rock band \emph{Radiohead} whose album \emph{Kid A} had also been leaked to Napster three months before reaching music store shelves \cite{rm_radioheadkida_2000}.
-Despite music critics labeling the album ``experimental'' and ``challenging'' (it had no official singles nor music videos for any of its songs), in October 2000 Kid A captured the number one spot on the Billboard 200 sales chart in its debut week.
-Unlike Metallica, Madonna and Dr.Dre that led litigation with Napster, Radiohead had never hit the top20 in the US before.
+Napster maintained a \emph{central index server} based on file lists provided by
+participating peers. The central index server was queried by the users and it
+returned pointers to the actual content. Thus, by centralizing search while
+distributing downloads, Napster achieved a highly functional design that was
+widely acknowledged, at the time, as ``the fastest growing Internet application
+ever'' that toped some 26 million users and 80 million songs
+\cite{jmm_naptopusage_2001}.
+
+Napster was ultimatelly brought down after Recording Industry Association of
+America (RIAA) filed a lawsuit for copyright infringement of the intellectual
+property of artists whose music was distributed by the network\footnote{
+  The sueing spree was initiated by heavy metal giants \emph{Metallica} that
+  feared Napster was undermining record sales. Opposed opinions instantly
+  emerged with best example the british alternative rock band \emph{Radiohead}
+  whose album \emph{Kid A} had also been leaked to Napster three months before
+  reaching music store shelves \cite{rm_radioheadkida_2000}. Despite music
+  critics labeling the album ``experimental'' and ``challenging'' (it had no
+  official singles nor music videos for any of its songs), in October 2000 Kid A
+  captured the number one spot on the Billboard 200 sales chart in its debut
+  week. Unlike Metallica, Madonna and Dr.Dre that led litigation with Napster,
+  Radiohead had never hit the top20 in the US before.
 }
-\cite{wiki_amnaptrial_2001}.
-This triggered a shift towards decentralised networks like the classic \emph{Gnutella} and, its ``rival'' at the time, \emph{FastTrack}.
-Gnutella, in its first realisation (v.0.4 \cite{gnutellav0.4}), was a fully decentralised approach.
-Unfortunatelly, network resourses tended to be overwhelmed by message flooding and if queries were constrained to live for some (small) number of hops, this ultimatelly resulted to a shrinked search scope, leading to a death spiral that prevented the network from scaling.
-The solution came with a slightly more structured network by introducing the notions of \emph{ultra-peers} and \emph{leaf-nodes} (v.0.6 \cite{gnutella}).
-With \emph{Limewire}, the most widely used software that implemented the protocol, the network was highly successful and was reported to have reached connecting more that 3 million users concurrently by the end of 2006 \cite{rsr_gnutellaevol_2006}.
-Fasttrack and its implementing software KaZaA \cite{kazaa}, on the other hand, went with the hierarchical approach (\emph{super-nodes} in their context) from the very begining, dynamically assigning indexing functions to special peers in order to increase search performace and scaling potential.
-FastTrack network had an average of 3 million concurrently connected nodes in 2003 \cite{lkr_kazaameasure_2005}.
-It is estimated that the total number of users was greater than that of Napster at its peak.
-With Limewire discontinued/shutdown in 2010 and its successor, \emph{Frostwire}, moving to the Bittorent protocol in 2011, seem to very well signal the end of the once-mighty Gnutella network.
-KaZaA continues as a propriatery music store solution.
-In the meanwhile, Fastrack design team had already introduced a VoIP and text messaging application based on the design principles of KaZaA's hierarchical P2P nature, that was meant to change how people communicate around the world.
-This app was no other than Skype \cite{skype}.
-Despite its wide usage, little is known for the technical aspects of its overlay network, mainly due to its proprietary nature.
-Studies, though, confirm that it shares strong structural similarities to its file-sharing predecessor (ordinary hosts and super-nodes in Skype's terminology) but they significantly differentiate in terms of traffic behaviour \cite{g_voipskypesec_2005, gdj_skypestudy_2006}.
-Skype was introduced in 2003 and from then on it witnessed frenzy-rythmed increase to its popularity, with more that 650 milion registered users in 2011  \cite{skypetotalusers} and reported to have broken the barrier of 35 million skypers concurrently online on the the 5th of March 2012 \cite{skypesymusers}.
-On October 2011 Microsoft Corp.
-and Skype Global S.à.r.l.
-anounnced the completion of Microsoft's acquisition of Skype for \$$8.5$ billion \cite{skypemicrosoft}.
-
-File sharing was the most popular application area for structured architectures as well. Overnet
-% TODO find citation
+\cite{wiki_amnaptrial_2001}. This triggered a shift towards decentralised
+networks like the classic \emph{Gnutella} and, its ``rival'' at the time,
+\emph{FastTrack}. Gnutella, in its first realisation (v.0.4 \cite{gnutellav04}),
+was a fully decentralised approach. Unfortunatelly, network resourses tended to
+be overwhelmed by message flooding and if queries were constrained to live for
+some (small) number of hops, this ultimatelly resulted to a shrinked search
+scope, leading to a death spiral that prevented the network from scaling. The
+solution came with a slightly more structured network by introducing the notions
+of \emph{ultra-peers} and \emph{leaf-nodes} (v.0.6 \cite{gnutella}). With
+\emph{Limewire}, the most widely used software that implemented the protocol,
+the network was highly successful and was reported to have reached connecting
+more that 3 million users concurrently by the end of 2006
+\cite{rsr_gnutellaevol_2006}. Fasttrack and its implementing software KaZaA
+\cite{kazaa}, on the other hand, went with the hierarchical approach
+(\emph{super-nodes} in their context) from the very begining, dynamically
+assigning indexing functions to special peers in order to increase search
+performace and scaling potential. FastTrack network had an average of 3 million
+concurrently connected nodes in 2003 \cite{lkxr_kazaameasure_2005}. It is
+estimated that the total number of users was greater than that of Napster at its
+peak. With Limewire discontinued/shutdown in 2010 and its successor,
+\emph{Frostwire}, moving to the Bittorent protocol in 2011, seem to very well
+signal the end of the once-mighty Gnutella network. KaZaA continues as a
+propriatery music store solution. In the meanwhile, Fastrack design team had
+already introduced a VoIP and text messaging application based on the design
+principles of KaZaA's hierarchical P2P nature, that was meant to change how
+people communicate around the world. This app was no other than Skype
+\cite{skype}. Despite its wide usage, little is known for the technical aspects
+of its overlay network, mainly due to its proprietary nature. Studies, though,
+confirm that it shares strong structural similarities to its file-sharing
+predecessor (ordinary hosts and super-nodes in Skype's terminology) but they
+significantly differentiate in terms of traffic behaviour
+\cite{g_voipskypesec_2005,gdj_skypestudy_2006}. Skype was introduced in 2003
+and from then on it witnessed frenzy-rythmed increase to its popularity, with
+more that 650 milion registered users in 2011 \cite{skypetotalusers} and
+reported to have broken the barrier of 35 million skypers concurrently online on
+the the 5th of March 2012 \cite{skypesymusers}. On October 2011 Microsoft Corp.
+and Skype Global S.à.r.l. anounnced the completion of Microsoft's acquisition of
+Skype for \$$8.5$ billion \cite{skypemicrosoft}.
+
+File sharing was the most popular application area for structured architectures
+as well. Overnet
+% TODO find citation for Overnet
 \cite{},
 for example, was created for sharing large files (e.g., movies and CD images).
-It implemented the Kademlia tree-like partitioning algorithm.
-Other implementations of Kademlia are the Kad Network (eMule/aMule  clients), BitTorrent for trackerless torrernts (official BitTorrent client, $\mu$Torrent, BitComet, Transmission and BitSpirit all share compatibility with the Mainline DHT which realises the Kademlia protocol) and the Gnutella DHT (first implemented by LimeWire as the Mojito DHT).
-Lately, RetroShare \cite{retroshare} is the new trend for file-sharing which incorporates an implementation of the Kademlia DHT to create a Friend-to-Friend (F2F) network with some attributes of a darknet as well.
-But the potential of structured architectures was not exhausted only to the above.
-Structured P2P networks became popular in distributed systems because they could scale smoothly and have fast and reliable object location mechanisms.
-Thus, research and development was very active and productive resulting in a variety of diverse applications such like distributed search engines \cite{yaci}, distributed data storage systems \cite{kbc_oceanstore_2000, bdet_fsdfs_2000, dkkms_cfs_2001, dr_pastutility_2001, abc_farsite_2002, mmfc_ivy_2002, arla, agebh_dks_2003}, Web caches, archives, and publishing systems \cite{ird_squirrel_2002, bags_youserv_2002, wrc_publius_2000, wm_tangler_2001}, messaging applications \cite{threedegrees}, event notification infrastructures \cite{rkcd_scribe_2001, cdkr_scribe_2002, agebh_dks_2003}, naming services \cite{cmm_chorddns_2002}, censor-resistent stores \cite{cswh_freenet_2001} and lately even cloud-based platforms \cite{mgpj_cloudsnap_2011}.
+It implemented the Kademlia tree-like partitioning algorithm. Other
+implementations of Kademlia are the Kad Network (eMule/aMule  clients),
+BitTorrent for trackerless torrernts (official BitTorrent client, $\mu$Torrent,
+BitComet, Transmission and BitSpirit all share compatibility with the Mainline
+DHT which realises the Kademlia protocol) and the Gnutella DHT (first
+implemented by LimeWire as the Mojito DHT). Lately, RetroShare \cite{retroshare}
+is the new trend for file-sharing which incorporates an implementation of the
+Kademlia DHT to create a Friend-to-Friend (F2F) network with some attributes of
+a darknet as well. But the potential of structured architectures was not
+exhausted only to the above. Structured P2P networks became popular in
+distributed systems because they could scale smoothly and have fast and reliable
+object location mechanisms. Thus, research and development was very active and
+productive resulting in a variety of diverse applications such like distributed
+search engines \cite{yaci}, distributed data storage systems
+\cite{kbc_oceanstore_2000,bdet_fsdfs_2000,dkkms_cfs_2001,dr_pastutility_2001,abc_farsite_2002,mmfc_ivy_2002,arla,agebh_dks_2003}, Web caches, archives,
+and publishing systems \cite{ird_squirrel_2002,bags_youserv_2002,wrc_publius_2000,wm_tangler_2001},
+messaging applications \cite{threedegrees}, event notification infrastructures
+\cite{rkcd_scribe_2001,cdkr_scribe_2002,agebh_dks_2003}, naming services
+\cite{cmm_chorddns_2002}, censor-resistent stores \cite{cswh_freenet_2001} and
+lately even cloud-based platforms \cite{mgpj_cloudsnap_2011}.
 
 %%%%%%%% NEW TRENDS %%%%%%%%
-%File sharing Limewire Kazzaa The file-sharing landscape is slowly adjusting in response to the continued push for more anti-piracy tools, the final Pirate Bay verdict, and the raids and arrests in the Megaupload case, SOPA, ACTA.
-This leads to users trying to find new solutions for anonymous, uncensored, decentralised file sharing solutions like Freenet, Tribler and RetroShare where connections are made only between trusted peers — sometimes called "friends" and the respective formed network F2F.
+%File sharing Limewire Kazzaa The file-sharing landscape is slowly adjusting in
+%response to the continued push for more anti-piracy tools, the final Pirate Bay
+%verdict, and the raids and arrests in the Megaupload case, SOPA, ACTA. This
+%leads to users trying to find new solutions for anonymous, uncensored,
+%decentralised file sharing solutions like Freenet, Tribler and RetroShare where
+%connections are made only between trusted peers — sometimes called "friends"
+%and the respective formed network F2F.
 
@@
 
-The huge numbers of users, for the various applications we presented above, reveals the tremendous popularity P2P architectures have enjoyed.
-This, though, does not come without any toll.
-Measurements on some popular systems, such as FastTrack, Gnutella and DirectConnect have shown that the P2P traffic contributes the largest portion of the overall Internet traffic \cite{seroiu_analysiscds_2002, sen_analyzep2ptraffic_2004}.
-Even with later studies, this observation still holds on.
-Using deep packet inspection techniques deployed at selected customer sites \emph{Ipoque} and \emph{CacheLogic} were able to track down application usage \cite{cachelogic, ipoque2007, ipoque2009}.
-CacheLogic reports that more than 60\% of the traffic is identified as P2P, with Ipoque verifying this observation reporting similar results for 2007, while observing some small decline for 2008/2009 (but still remaining dominant compared to other applications).
-Apart from its participation in overall usage terms, P2P is already a huge contibutor to the overall data exchanged across the Internet\footnote{
-    Mainly because file-sharing P2P is used to exchange big files of HD video or high sampled audio content compared to simple Web content distributed via, for example, HTTP.
+The huge numbers of users, for the various applications we presented above,
+reveals the tremendous popularity P2P architectures have enjoyed. This, though,
+does not come without any toll. Measurements on some popular systems, such as
+FastTrack, Gnutella, DirectConnect and BitTorrent have shown that the P2P
+traffic contributes the largest portion of the overall Internet traffic
+\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}.
+Even with later studies, this observation still holds on. Using deep packet
+inspection techniques deployed at selected customer sites \emph{Ipoque} and
+\emph{CacheLogic} were able to track down application usage \cite{cachelogic,ipoque2007,ipoque2009}.
+CacheLogic reports that more than 60\% of the traffic is identified as P2P, with
+Ipoque verifying this observation reporting similar results for 2007, while
+observing some small decline for 2008/2009 (but still remaining dominant
+compared to other applications). Apart from its participation in overall usage
+terms, P2P is already a huge contibutor to the overall data exchanged across the
+Internet\footnote{
+  Mainly because file-sharing P2P is used to exchange big files of HD video or
+  high sampled audio content compared to simple Web content distributed via, for
+  example, HTTP.
 }.
-\cite{multinteligence} claims that P2P traffic in 2007 was around 1.6 petabytes per month and the potential is to be quintubled by the end of 2012 reaching almost 8 petabytes per month.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\cite{multinteligence} claims that P2P traffic in 2007 was around 1.6 petabytes
+per month and the potential is to be quintubled by the end of 2012 reaching
+almost 8 petabytes per month.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{The Topology Mismatch Problem}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-One of the major issues that defines the efficiency of an overlay network is the mapping of the overlay structure to the underlying infrastructure.
-Remember, in Figure~\ref{figure:overlay} Nodes $A$ and $B$ are on the same local network, while $C$ and $D$ are in different networks.
-The top layer represents the overlay interconnection formed by these nodes at a higher level.
-Link connections there, can change as needed by the running environment, without any particular constraint by the underlying physical topology.
-Let's have a closer look with the help of a simple example.
-Assume nodes $A$, $B$, $C$ and $D$ are connected through the physical network shown in Figure~\ref{figure:phys}, where the network costs are given in milliseconds, and peer $A$ sends a message to peer $D$.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+One of the major issues that defines the efficiency of an overlay network is the
+mapping of the overlay structure to the underlying infrastructure. Remember, in
+Figure~\ref{figure:overlay} Nodes $A$ and $B$ are on the same local network,
+while $C$ and $D$ are in different networks. The top layer represents the
+overlay interconnection formed by these nodes at a higher level. Link
+connections there, can change as needed by the running environment, without any
+particular constraint by the underlying physical topology. Let's have a closer
+look with the help of a simple example. Assume nodes $A$, $B$, $C$ and $D$ are
+connected through the physical network shown in Figure~\ref{figure:phys}, where
+the network costs are given in milliseconds, and peer $A$ sends a message to
+peer $D$.
 
@@
 
-If these peers participate in an overlay network according to one of the setups of Figure~\ref{figure:overlay-confs} then users will expierience different performances.
-In the overlay depicted in Figure~\ref{figure:overlay-1}, the message will traverse the following sequence of links in the physical layer (marked with their costs): $3 \rightarrow 7 \rightarrow 1 \rightarrow 2 \rightarrow 2 \rightarrow 1 \rightarrow 7 \rightarrow 8 \rightarrow 5 \rightarrow 5 \rightarrow 4$.
-In the alternative overlay of Figure ~\ref{figure:overlay-2}, the path will be: $3 \rightarrow 8 \rightarrow 4$.
-The total cost of the first sequnce is $45 ms$, while the second yields a mere $15 ms$.
-Therefore, we can conclude that the second overlay is more congruent with the underlying physical network than the first one and, thus, is more efficient.
+If these peers participate in an overlay network according to one of the setups
+of Figure~\ref{figure:overlay-confs} then users will expierience different
+performances. In the overlay depicted in Figure~\ref{figure:overlay-1}, the
+message will traverse the following sequence of links in the physical layer
+(marked with their costs): $3 \rightarrow 7 \rightarrow 1 \rightarrow 2
+\rightarrow 2 \rightarrow 1 \rightarrow 7 \rightarrow 8 \rightarrow 5
+\rightarrow 5 \rightarrow 4$. In the alternative overlay of Figure
+~\ref{figure:overlay-2}, the path will be: $3 \rightarrow 8 \rightarrow 4$. The
+total cost of the first sequnce is $45 ms$, while the second yields a mere $15
+ms$. Therefore, we can conclude that the second overlay is more congruent with
+the underlying physical network than the first one and, thus, is more efficient.
 
@@
 
-Ideally, overlay networks are expected to achieve an optimal mapping among the overlay paths and the underlying physical links, and avoid inefficient states.
-The problem of constructing an optimal overlay is referred to as the \emph{topology mismatch problem}, and formally defined as follows:
+Ideally, overlay networks are expected to achieve an optimal mapping among the
+overlay paths and the underlying physical links, and avoid inefficient states.
+The problem of constructing an optimal overlay is referred to as the
+\emph{topology mismatch problem}, and formally defined as follows:
 \begin{definition}
-    Let $V = \{v_1, ..., v_n\}$ be a set of points denoting the network nodes, $\{v_i, v_j\} \in E$ be the set of unicast distances between nodes $v_i$ and $v_j$, $G=(V,E)$ be a complete distance graph over $V$.
-The topology mismatch problem is to construct a minimal spanning tree,  where node degree is restricted to a constant ($k\geq 2$) by the bandwidth of each node $v_i$.
+Let $V = \{v_1, ..., v_n\}$ be a set of points denoting the network nodes,
+$\{v_i, v_j\} \in E$ be the set of unicast distances between nodes $v_i$ and
+$v_j$, $G=(V,E)$ be a complete distance graph over $V$. The topology mismatch
+problem is to construct a minimal spanning tree,  where node degree is
+restricted to a constant ($k\geq 2$) by the bandwidth of each node $v_i$.
 \end{definition}
 
-Early incarnations of the overlay protocols, however, did not make use of the optimal mapping with the physical network topology.
-Gnutella protocol for example was considered far from scalable \cite{ritter_gnucantscale_2001}.
-Since every peer can chοse its logical neighbours without knowledge of the underlying network had as a result a mismatch between physical and application-level topologies.
-Additionally, queries may be flooded to multiple paths merging in one peer while
-in such case one of the paths would have been enough.
-Nevertheless, peers may forward the same message to each other before they receive the query messages from each other.
-\cite{matei_mapgnutella_2002} showed that, even given that 95\% of any two nodes are less than 7 hops away from each other, a flooding-based algorithm, like the one used by Gnutella, can generate 330TB/month in a 50.000 node network!
-Therefore, P2P traffic can have a serious effect on the overall well-being of the Internet as a whole.
+Early incarnations of the overlay protocols, however, did not make use of the
+optimal mapping with the physical network topology. Gnutella protocol for
+example was considered far from scalable \cite{ritter_gnucantscale_2001}. Since
+every peer can chοse its logical neighbours without knowledge of the underlying
+network had as a result a mismatch between physical and application-level
+topologies. Additionally, queries may be flooded to multiple paths merging in
+one peer while in such case one of the paths would have been enough.
+Nevertheless, peers may forward the same message to each other before they
+receive the query messages from each other. \cite{matei_mapgnutella_2002} showed
+that, even given that 95\% of any two nodes are less than 7 hops away from each
+other, a flooding-based algorithm, like the one used by Gnutella, can generate
+330TB/month in a 50.000 node network! Therefore, P2P traffic can have a serious
+effect on the overall well-being of the Internet as a whole.
 
 Similar problems are observed in decentralized structured schemes also.
-Typically, node IDs are assigned \emph{randomly}, resulting in excellent load balancing, scalability and robustness.
-Unfortunately, this randomness has a negative impact on the \emph{routing locality} of the network.
-This means that
-even though the target node can be reached with logarithmic overlay hops, the distance traveled in the physical underlying network, during the overlay routing
-process, can be far from optimal.
-The ratio of the physical distance, a query travels through the overlay to an object, and the minimal distance to that object (i.e.
-through IP) is known in the literature as \emph{stretch} or \emph{Relative Delay Penalty (RDP)} \cite{chu_esm_2000}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Typically, node IDs are assigned \emph{randomly}, resulting in excellent load
+balancing, scalability and robustness. Unfortunately, this randomness has a
+negative impact on the \emph{routing locality} of the network. This means that
+even though the target node can be reached with logarithmic overlay hops, the
+distance traveled in the physical underlying network, during the overlay routing
+process, can be far from optimal.  The ratio of the physical distance, a query
+travels through the overlay to an object, and the minimal distance to that
+object (i.e. through IP) is known in the literature as \emph{stretch} or
+\emph{Relative Delay Penalty (RDP)} \cite{chu_esm_2000}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Motivation and our Goal for this Survey}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-A topology unaware overlay network is able to control the sequence of peers a message traverses before reaching its destination, but it is completely unaware of how the actual packets are switched at the underlying infrastructure along the overlay path.
-For example, a single logical point-to-point link on the overlay, most of the time, corresponds to multiple physical links in the underlying layer.
-Additionally, a link in the underlying network often serves the mapping of several overlay paths causing increase of the traffic on the physical link, which is also called the link's \emph{stress}\cite{chu_esm_2002}.
-Furthermore, the network's \emph{churn}, that is the stochastic behaviour of peers randomly joining and leaving the P2P network, all cause the overlay to physical mapping to create such unnecessary, redundant maintenance traffic that impacts the efficiency of the network in terms of average response time.
-
-The topology mismatch problem is exacerbated by the high amounts of traffic generated by the highly popular P2P applications.
-This creats the killer combination for the network infrastructure, ultimately effecting all Internet users.
-Solving it, on the other hand, is no easy task.
-Topology mismatch is known to be an NP-Hard problem \cite{NPBOOK,chawathe_scattercast_2000}.
-Moreover, the Internet itself is an interesting environment where end-to-end latencies demonstrate what is called \emph{triangle inequality violations (TIVs)}, which further complicate the problem.
-These delay space TIV is a consequence of the Internet's structure and routing policies and thus will remain a property of the Internet for the foreseeable future \cite{zheng_irprtt_2005}.
-TIVs affect network coordinate \cite{cox_vivaldi_2004,wong_meridian_2005} and positioning \cite{ng_gnp_2001} and makes the construction of location/delay aware overlays difficult.
-An extensive array of heuristic approaches have been proposed in an effort to find an algorithm that performs reasonably and consistently well.
-The purpose of this survey is to gather the extensive knowledge that has been produced recently in this research field and organise the available information ultimately helping the reader sift through the myriad of these efforts, understand the different approaches and identify though a taxonomy the advantages and disadvantages of each of them.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+A topology unaware overlay network is able to control the sequence of peers a
+message traverses before reaching its destination, but it is completely unaware
+of how the actual packets are switched at the underlying infrastructure along
+the overlay path. For example, a single logical point-to-point link on the
+overlay, most of the time, corresponds to multiple physical links in the
+underlying layer. Additionally, a link in the underlying network often serves
+the mapping of several overlay paths causing increase of the traffic on the
+physical link, which is also called the link's \emph{stress}\cite{chu_esm_2002}.
+Furthermore, the network's \emph{churn}, that is the stochastic behaviour of
+peers randomly joining and leaving the P2P network, all cause the overlay to
+physical mapping to create such unnecessary, redundant maintenance traffic that
+impacts the efficiency of the network in terms of average response time.
+
+The topology mismatch problem is exacerbated by the high amounts of traffic
+generated by the highly popular P2P applications. This creats the killer
+combination for the network infrastructure, ultimately effecting all Internet
+users. Solving it, on the other hand, is no easy task. Topology mismatch is
+known to be an NP-Hard problem \cite{NPBOOK,chawathe_scattercast_2000}.
+Moreover, the Internet itself is an interesting environment where end-to-end
+latencies demonstrate what is called \emph{triangle inequality violations
+(TIVs)}, which further complicate the problem. These delay space TIV is a
+consequence of the Internet's structure and routing policies and thus will
+remain a property of the Internet for the foreseeable future
+\cite{zheng_irprtt_2005}. TIVs affect network coordinate
+\cite{cox_vivaldi_2004,wong_meridian_2005} and positioning \cite{ng_gnp_2001}
+and makes the construction of location/delay aware overlays difficult. An
+extensive array of heuristic approaches have been proposed in an effort to find
+an algorithm that performs reasonably and consistently well. The purpose of this
+survey is to gather the extensive knowledge that has been produced recently in
+this research field and organise the available information ultimately helping
+the reader sift through the myriad of these efforts, understand the different
+approaches and identify though a taxonomy the advantages and disadvantages of
+each of them.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Survey Outline}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 % TODO: Review this after the content and article structure is fixed.
-In the introduction we had the oportunity to recognise the importance of overly networks in contemporary application architectures.
-Also we quickly reviewed the basic trends in both overlay construction\footnote{
-    The main focus of this paper is the topology mismatch problem, therefore the available approaches in overlay networks are briefly summarized, and interested readers are pointed to the available surveys on overlay networks \cite{TheotokisS04,LuaCPSL05}.
+In the introduction we had the oportunity to recognise the importance of overly
+networks in contemporary application architectures. Also we quickly reviewed the
+basic trends in both overlay construction\footnote{
+  The main focus of this paper is the topology mismatch problem, therefore the
+  available approaches in overlay networks are briefly summarized, and
+  interested readers are pointed to the available surveys on overlay networks
+  \cite{TheotokisS04,LuaCPSL05}.
 }
-and application development.
-Last but not least, we identified and formaly defined the phenomenon of the topology mismatch problem and its impact on the utilization of network resources.
-The rest of the survey\ldots
-
-Sections~\ref{sec:unstructured} and \ref{sec:structured} discuss the recent academic work that has been conducted in the field.
-
-Ultimately, it reviews the most important resources found in the literature that attempt to tackle the problem as it proposes a taxonomic scheme, using a set of supertype-subtype relationships based on unique characteristics they have and/or specific goals they target.
-What follows does not claim to be a thorough citation of all known protocols that are available out there, but instead, a careful selection of those that left a distinctive fingerprint contribution in the efforts of the research community to alleviate the topology mismatch problem.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% start
-%TODO: Maybe this section should be reviewed or deleted or go to another section
-%Studying the behaviour of the various peer-to-peer schemes \cite{matei_mapgnutella_2002, lv_randomwalks_2002, merugu_str2unstr_2003} showed that the ad-hoc network topology of unstructured overlay networks that preserve \emph{Power Law} and \emph{Small World} characteristics\footnote{Power Law describes the node degree while the Small World describes characteristics of path length and clustering coefficient.
-The clustering coefficient for a node $\upsilon$ in a graph $G = \left( V, E \right)$ is defined as the ratio of the existing connections between $\upsilon$'s neighbouring nodes to $\gamma \times \left( \gamma - 1 \right)$, where $\gamma$ is the number of neighbouring nodes of $\upsilon$.
-High cluster coefficient means that neighbouring nodes of any node $\upsilon$ likely connect one another.} \cite{faloutsos_powerlaw_1999, saroiu_measurefileshare_2002} offer a more promising approach.
-Particularly:
-%\begin{itemize}
-%  \item Peer-to-peer clients are extremely \emph{transient}.
-Unstructured systems can have high maintenance traffic in delivering messages, updating the mapping, discovering failures and replicating lost data or pointers, making them insufficient on highly volatile networks.
-%  \item \emph{Keyword searches} versus \emph{exact-match queries}.
-In DHTs there is a tight control between the data placement and the topology of the network.
-For this reason it is hard to efficiently support partially matched queries while Gnutella and other similar systems effortlessly support keyword searches and other complex queries since the mechanism is realized locally, on a node-by-node basis.
-%  \item Popular content is located at multiple peers and thus it is more likely for a flooding-based search to return results.
-DHTs, on the other hand, fit better in the systems which require ability to reliably locate content, even in the extreme case that only a single-copy exists in the network.
-%\end{itemize}
-% end
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% TODO: how 
-%For this reasons, efforts have been placed for optimizing the efficiency of decentralized unstructured peer-to-peer networks. Research mainly focuses on
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item reducing unnecessary, redundant communication traffic, and
-%  \item exploiting physical locality to reduce communication response.
-%\end{inparaenum}
-%The goal can be achieved at, both, the application-level network as well as the underlying physical one.
-In the first case by refining the message relay techniques, while in the second one, by adaptively reconstructing the application network to map as well as possible to the the physical network.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{The Topology Mismatch Problem}
-
-%\section{ADDITIONAL MATERIAL}
-
-%%%%%%%
-%%%%%%% TODO: NEEDS REWORDING
-%%%%%%%
-%%%%%%%
-%%%%%%%
-%Bindal et al. [14] proposed to improve the traffic
-%locality in BitTorrent systems via biased neighbor selection
-%enforced by packet control mechanisms on border gateway
-%routers. More recently, Ono [15] proposed to leverage the
-%existing CDN infrastructure to estimate the distance between
-%BitTorrent peers before making peer connections. P4P [16],
-%instead, aimed to provide a general framework for P2P
-%applications by developing ISP-application interfaces so that
-%application software can use the network layer information,
-%such as congestion status and network topology, for better
-%performance and resource utilization.
-
-% TODO: Taxomize the approaches that have been proposed in the literature
-% for unstructured and structured overlay networks
-% UNDER CONSTRUCTION
-% VERY IMPORTANT SECTION
-%\section{Taxonomy}
-%
-%% TODO: Unstructured
-%% UNDER CONSTRUCTION
-%Several overlay protocols for dealing with the problems derived from the
-%topology mismatch problem have been proposed in the literature. Each
-%implementation has different goals but all share some common ground on the
-%aspect of the network functionality they focus on, in order to achieve them. In
-%this section, the general taxonomy used throughout this paper to categorize the
-%algorithms based on their functionality is presented.  The brief description of
-%each category and sub-category is stated below.
-%
-%{\sethlcolor{yellow}\hl{CITE this paper {\cite{liu_ltm_2004}}}}
-%
-%\subsection{Forwarding based}
-%
-%Forwarding based methodology is generally adapted by distributed unstructured
-%overlay systems. In this approach, a peer selects only a subset of its
-%neighbours to re-broadcast query messages. The selection is made using one or a
-%combination of various statistical metrics. Examples of such metrics are the
-%number of responses received by a neighbour or the connection latency of the
-%link between the nodes, etc. The forwarding based approach enhances search
-%efficiency but has also several drawbacks. First, the search scope is reduced
-%drastically.  Expanding the search scope, on the other hand, is no easy task
-%because the overhead of forming multicast trees is proportional to the multicast
-%group size.  Second, forwarding based schemes do not consider dynamic joining
-%and leaving of peers so they do not scale well on dynamic environments.
-%
-%\subsection{cache based} 
-%
-%Caching based protocols are effectively used to reduce traffic costs and response
-%times. The caching policy varies depending on the way protocol handles the index and the
-%content. Centralized P2P systems
-%use central index servers, while local caching systems, such as KazaA, use super peers
-%to cache indices in a distributed way. Content caching is also possible in P2P
-%systems, where nodes cache the forwarded content for further retrievals.
-%Although caching has the above mentioned advantages,  duplication
-%of messages still exist, which limits the scalability of these approaches.
-%Therefore, cache based approaches are analyzed in the following categories:
-%  \begin{itemize}
-%    \item \emph{data index caching},
-%    \item \emph{content index caching},
-%    \item \emph{centralized}, and
-%    \item \emph{local}.
-%  \end{itemize}
-%
-%\subsection{overlay optimization based} 
-%
-%The overlay optimization based protocols modify the topology of the P2P network
-%using various techniques. These approaches include creating spanning trees using 
-%connection graphs, creating cluster of physically close nodes, or using latency
-%information to detect proximity. The brief description of each category is presented
-%below:
-%
-%  \begin{itemize}
-%    \item \emph{Spanning tree based}. These approaches construct of a rich graph
-%    based on the network connections and build minimum spanning tree on the
-%    graphs, causing large traffic overhead to the system
-%    \cite{chu_esm_2000,chu_esm_2002}.
-%
-%    \item \emph{Cluster based}. These approaches select to link physically
-%    closer nodes with each other, therefore shrink the search scope
-%    significantly while mapping accuracy is not always guaranteed.
-%
-%    \item \emph{Minimum latency first}. Use of latency as a metric to calculate
-%    distance among peers. They require global latency information
-%    ``landmarks''\footnote{Measuring latency between peers and stable Internet
-%    servers.}.
-%
-%  \end{itemize}
-%
-%
-%% TODO: Structured
-%% UNDER CONSTRUCTION
-%\subsection{proximity based}
-%Xu \textit{et al.} \cite{xu_globstate_2003} state that there are three ways generating proximity information
-%\begin{itemize}
-% \item Expanding Ring Search\\
-%Expanding ring search can be of two forms. First it can utilize the multicast
-%infrastructure in the underlying network in order to emit its messages.
-%Unfortunately such infrastructures are not widely deployed thus the
-%implementations of this way of generating proximity information is limited to
-%blindly flooding the neighbourhood to obtain reasonable results.
-%
-% \item Heuristics\\
-%Heuristics are used in order to reduce the blindness of the expanding ring
-%search and make realizations more efficient and effective. Unfortunately a
-%common problem of all heuristic approaches is the local minimum pitfall in which
-%the search might be caught into.
-%
-% \item Landmark Binning\\
-%Landmark clustering is based on the view that nodes with similar distances to a
-%set of predefined well-known landmark nodes are pretty likely of being close to
-%each other. But this approach has its weaknesses as well, such as the fact that
-%is a rather coarse grained approximation, therefore not particularly well suited
-%for differentiating nodes within close distance to each other.
-%
-%\end{itemize}
-%
-%Castro \textit{et al.} \cite{castro_proximitydht_2002,castro_topawareroute_2002} on
-%the other hand, comment on the three basic approaches in exploiting proximity
-%in DHT protocols suggested by Ratnasamy \textit{et al.} \cite{ratnasamy_openq_2002}.
-%These are:
-%
-%\begin{itemize}
-% \item Geographic Layout\\
-%The node IDs are assigned in such a way that nodes close by in the physical
-%network topology, be close in the node ID space as well. Implementations that
-%work relatively well with this approach have been incorporated into CAN. Nodes
-%measure the RTT between themselves and a set of landmarks in order to match the
-%CAN space as much as possible to the physical one. Unfortunately, the approach
-%requires well known landmark servers and for that matter is  not fully
-%self-organizing which can further lead to imbalanced node distribution. On
-%other DHTs, such as Chord or Pastry, another problem emerges.  To gain fault
-%tolerant properties, these protocols, replicate key-value pairs on neighbouring
-%(in the ID space) nodes. When a proximity-based node ID assignment has been
-%used, the needed failure resilience is undermined by the fact that close by
-%nodes are more likely to suffer collective failures.
-%
-% \item Proximity Routing\\
-%Proximity routing does not require routing tables be built using any knowledge
-%about network proximity. On the other hand it exploits such knowledge in order
-%to choose the best next hop during routing a message. This approach 
-%balances between choosing the node that will further progress the routing
-%towards the destination and choosing the closest entry in the routing table, in
-%terms of network proximity. Thus, it is relatively less effective than
-%geographical layout when applied to CAN(-like) implementations. Moreover, the
-%technique has been incorporated into a version of Chord (TODO: PChord |
-%additionally check whether this is true) causing an increase on the overhead of
-%node joins and the size as well as maintenance cost of finger tables.
-%\cite{dabek_cfs_2001} proposes a server selection scheme for the Chord DHT, on
-%the domain of proximity routing selection. In \emph{CFS}, each node predicts
-%the entire lookup latency as a function of the total number of nodes and the
-%average overlay next routing peer. The problem is that it is very difficult to
-%have a clear picture on the total number of nodes and the average hop latency
-%from the local. This leads to rough estimations that consequently decreases
-%overall performance.
-%
-% \item Proximity Neighbour Selection\\
-%Finally, the third approach, constructs the routing tables using proximity
-%knowledge. Tapestry and Pastry's mechanisms of routing table
-%maintenance try to minimize the distance to nodes appearing in a peer's routing
-%table. Since routing is based on longest node ID prefix match, messages 
-%are gradually forwarded to nearby nodes at each routing step.
-%\cite{castro_proximityp2p_2002} argues on how Pastry exploits proximity
-%neighbour selection in order to create a scheme that is (more) location-aware
-%compared to the other well-known DHTs (CAN, Chord).
-%
-%\end{itemize}
-%
-%{\sethlcolor{yellow}\hl{
-%HA: Possible criteria: based on which protocol (eg Gnutella, CAN etc), peer
-%selection (topology selection, cluster, cache etc), supports dynamic update,
-%runtimes}}
-%%\textbf{Algorithm} & \textbf{Overlay structure} & \textbf{Forwarding} & \textbf{Cache} & \textbf{Overlay optimization} & \textbf{Proximity information} & \textbf{Base protocol} & \textbf{Dynamic update} & \textbf{Runtime} \\
-%
-%
-%{\sethlcolor{yellow}\hl{
-%HA: Possible criteria2: Overlay optimization structure, base protocol (eg
-%Gnutella, CAN etc), dynamic update, runtimes, scalability}}
-%
-%{\sethlcolor{yellow}\hl{
-%HA: maybe add also the year of publication, and see if there is a pattern in
-%terms of the method and the year??}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Unstructured Decentralized Algorithms}\label{sec:unstructured}
-
-This section is dedicated to the unstructured P2P architectures.
-In the next paragraphs, algorithms for tackling the topology mismatch problem on such networks are visited, analysed and categorised based on the use of the overlay structure, messaging approach for peer communication and techniques proposed for detecting proximity in order to optimise overlay topology.
-In short, the methods that are highlighted here are divided into
-\begin{inparaenum}[\itshape i\upshape)]
-  \item broadcast-based,
-  \item cache-based,
-  \item overlay optimisation-based, and
-  \item landmark-based
-\end{inparaenum}
- approaches.
-
-% TODO: READ A Near-Optimal Algorithm Attacking the Topology Mismatch Problem in Unstructured Peer-to-Peer Networks (This is an Approximation alg for Top-mis problem)
-
-%{\sethlcolor{yellow}\hl{
-%HA: TALK ABOUT GNUTELLA HERE!!!!}}
-
-\subsection{Broadcast Based}
-
-In the first realisation of the Gnutella protocol, each query request received by a peer is then forwarded to all its
-neighbours, clearly generating unnecessary traffic over the network.
-The algorithms highlighted in this paragraph are referred to as broadcast-based because they enhance the "naive" broadcasting approach described above and prefer forwarding messages to a selected subset of a node's outgoing links.
-The selection criteria vary depending on the algorithm but they mainly use one or multiple statistical metrics that span from the obvious ones, like the latency of the link between the nodes, to more sophisticated alternatives (that may even take into account application level requirements) such as the reliability of the neighbour.
-The forwarding-based approach has the advantage of enhancing search efficiency but on the other hand suffer drawbacks like the drastic reduction of the search scope because of the restrictions of the broadcasting policies as well as limitations when scaling to large networks.
-
-%Expanding the search scope, on the other hand, is no easy task
-%because the overhead of forming multicast trees is proportional to the multicast group size.  
-%forwarding based schemes do not consider dynamic joining
-%and leaving of peers so they do not scale well on dynamic environments.
-
-\paragraph*{\bf Improving search in peer-to-peer networks}
-
-\cite{yang_improvep2psearch_2002} proposes an easy to implement and practical
-solution to the inefficiency problem caused by blind flooding in Gnutella-like
-P2P file sharing protocols. The paper replaces blind flooding with three
-approaches, namely \textit{iterative deepening}, \textit{directed BFS}, and
-\textit{local indices}. In \textit{iterative deepening}, the search is performed
-on a BFS tree with multiple preset depths. The depth limit is iteratively
-increased by the source node for each query based on the quality of the results.
-The source node, after examining the results, may issue a new request by
-increasing the depth limit, which will trigger the nodes at the last depth
-level to resume the search. The iterative approach avoids to start the search
-from scratch at each iteration and reduces the load of the nodes on the upper
-levels of the tree. The major drawback of \textit{iterative deepening} is the
-delay between successive iterations, as the source node needs to examine the
-results at each iteration before deciding to quit or resume the query.  The
-\textit{directed BFS} tries to avoid this delay by forwarding the query messages
-only to a selected set of neighbours, in which the selection criteria varies
-from the number of results received previously, distance in terms of hops,
-bandwidth, or the query load of the neighbour. In \textit{local indices}, each
-node sustains a local data index of the nodes within a radius of $r$ hops from
-itself and uses this local index to remotely query the neighbour nodes without
-generating additional traffic.  \textit{Local indices} greatly reduces the
-aggregate bandwidth usage of the network and improves query efficiency, however,
-updating the indices in cases with frequent node joins and leaves introduces a
-serious overhead to the system if the radius is kept broad.
-
-\paragraph*{\bf Gia}
-
-\emph{Gia} \cite{chawathe_gia_2003} is an effort to solve the scalability
-problem of the unstructured P2P file sharing systems, in particular Gnutella.
-The main novelty in the design of Gia is the replacement of the blind flooding
-approach of Gnutella with random walks \cite{lv_randomwalks_2002}. Although
-random walks is a step in the right direction, issuing only a single copy of the
-query within the network reduces the search scope, thus affecting negatively the success rate
-of the query.  In order to overcome this limitation, Gia introduces a
-token-based flow control mechanism, which is essentially an intelligent flow
-control algorithm that gradually redirects the queries to nodes which are more likely to answer.
-In order to prevent overloading of nodes with query requests, Gia uses a token-based flow control algorithm in which each node announces the number of query requests it
-can handle in terms of tokens to its neighbours, so that neighbours only forward
-query requests to nodes that they previously received tokens from. Gia also acknowledges the heterogenity in peer bandwidth,
-processing power, disk speed, etc, of the nodes in P2P networks and uses this
-information when connecting nodes to each other and by using a topology
-adaptation algorithm, Gia ensures that high capacity nodes have high degrees and
-low capacity nodes are within short proximity to high capacity ones. Although the
-topology adaptation algorithm Gia uses improves the scalability of the network,
-it does not help much in solving the topology mismatch problem, since it does
-not consider the underlying physical topology.
-
-\paragraph*{\bf Distributed Cycle Minimization Protocol}
-
-We have already mentioned Gnutella's inefficient broadcasting mechanism in flooding messages across the network with no actual targeting.
-This leads, to nodes that are far from the path towards the query request gratification, to spend their valuable resources to receive, queue and process messages that they shouldn't have received in the first place and eventually simply forward them blindly to continue this vicious cycle.
-Except from that, though, an other case of wasting network resources is when a node that is on the right path for the query, receives the same message multiple times from different neighbours.
-The reason for these duplicate messages is cycles in the forwarding paths.
-Actually, this duplicate packet problem seems to hurt more the active nodes; those with higher capacities and bandwidth that contribute the most to the network.
-Distributed Cycle Minimization Protocol (DCMP) \cite{zhu_dcmp_2008} introduces a method to
-remove the cycles and eliminate the duplicate packets, without sacrificing overlay
-connectivity. In DCMP, once a cycle is detected, the most powerful node in the
-cycle is elected as the \emph{Gate Peer} and the cycle is then destroyed by cutting that special link which will result in the minimisation of the distance to the Gate Peer for all the nodes that participated in that cycle.
-The process is managed by using two special message types namely \emph{Information Collection (IC)} and \emph{Cut Message (CM)}.
-One disadvantage of DCMP is that since the distance a forwarded message can travel is limited
-by the TTL value, which is practically $7$ in most cases DCMP cannot detect
-cycles formed by more than 7 peers. Even though cycle elimination improves the
-network performance, it does not actually solve the topology mismatch problem.
-
-%\subsection{Cache Based}
-%
-%Caching based protocols are effectively used to reduce traffic costs and response
-%times. The caching policy varies depending on the way protocol handles the index and the
-%content. Centralized P2P systems
-%use central index servers, while local caching systems, such as KazaA, use super peers
-%to cache indices in a distributed way. Content caching is also possible in P2P
-%systems, where nodes cache the forwarded content for further retrievals.
-%Although caching has the above mentioned advantages,  duplication
-%of messages still exist, which limits the scalability of these approaches.
-%Therefore, cache based approaches are analyzed in the following categories:
-%  \begin{itemize}
-%    \item \emph{data index caching},
-%    \item \emph{content index caching},
-%    \item \emph{centralized}, and
-%    \item \emph{local}.
-%  \end{itemize}
-%
-
-\subsection{Cache Based}
-Caching is a widely used technique to exploit locality and minimise
-redundant transfer of data, initially adopted by web and file server application environments with great success.
-Since peers in a P2P system also work as servers, it is intuitively expected that P2P file sharing systems can also benefit from caching.
-However, two of the main characteristic properties of P2P systems, nodes frequently joining and leaving and query lifetime relatively short (at least compared to those needed to be served by web servers) , makes caching in P2P networks a non-trivial issue.
-There are
-usually two levels of caching possible in file sharing systems, caching indices or pointers to data or caching the data itself.
-Caching is already implemented and successfully used in some commercial P2P systems.
-One popular example is KazaA, which uses caching of indices on its super-peer level.
-In the paragraphs below, caching-based algorithms are presented and explained in more depth.
-
-\paragraph*{\bf Replication Strategies in Unstructured Peer-to-Peer Networks}
-
-\cite{Cohen02} aim to improve the inefficient blind search algorithm by
-replicating the data in a P2P network. The main intuition behind the
-idea of replication, or using cached copies, is that as the number of copies for
-each item increases in the network, it would be easier for a search algorithm,
-even a random one, to locate these items. In order to analyse the feasibility of
-such an approach, the authors investigate three different replication
-strategies, namely uniform, proportional, and square root allocation. In the
-uniform model, the copies of items are uniformly replicated in the network,
-while in the proportional model, the items are replicated based on their query
-rate, so that frequently queried items are replicated more. The square root
-allocation is a strategy proposed in this paper, which is a model between the
-uniform and proportional allocation.  In order to evaluate the outcome of the
-replication models, the overall costs of successful and unsuccessful searches in the network are compared.
-Results argue that the uniform allocation model minimizes the maximum search size, therefore reduces the time spent on an unsuccessful search.
-The proportional model, on the other hand, promotes the more frequently queried items by replicating them
-more, therefore decreases the search time for popular items, but suffers when
-needing to locate the rare items. The authors also claim that the expected successful
-search size is the same for uniform and proportional models, and any approach
-between them would behave much better. Therefore they propose the square root
-allocation approach, which is a replication strategy that minimizes the expected
-search size of successful queries in P2P networks.
-
-\paragraph*{\bf Tracing a large-scale Peer to Peer System: an hour in the life of
-Gnutella}
-
-\cite{Markatos02} analyses Gnutella network traffic traces and by concluding there is locality among query requests, proposes a caching algorithm that tries to exploit these findings .
-The analysis of the trace data reveals other important facts about the structure of the Gnutella network and the query data.
-One significant observation is that the geographic locations of clients do not have a correlation with the number of query requests they receive.
-This is an obvious result of the topology mismatch problem caused by the overlay structure of the Gnutella network.
-Gnutella's traffic is observed to be bursty
-both for query requests and query responses, even in longer intervals, and nine
-out of ten queries do not generate any response due to the inefficient design of
-the Gnutella network. When developing a caching system to exploit locality,
-applying an approach similar to web caching does not fit well with P2P systems.
-Caches in P2P systems not only have to consider the query string, but also
-the TTL value, the source of the query, and the time of the query as well.  In
-general, even though optimum caching is hard to achieve, it is reported that
-it improves the overall performance of the Gnutella network.
-
-\subsection{Overlay Optimization Based}
-
-The overlay optimization based protocols modify the topology of the P2P network
-using various techniques. The two commonly used approach that is investigated in
-this section are creating spanning trees using 
-connection graphs, and creating clusters of physically close nodes. 
-%The brief description of each category is presented below:
-Spanning tree based approaches construct rich graphs based on the network
-connections and build minimum spanning trees on these graphs. Even though spanning trees provides efficient query performance, the construction
-and the update costs of the spanning tree, especially in dynamic environments
-with many nodes joining and leaving the network, cause large traffic
-overhead to the corresponding network \cite{chu_esm_2000,chu_esm_2002}.
-The cluster based approaches on the other hand select to link physically closer nodes with each other, therefore significantly shrinking the search scope of the system.
-Added to that,
-commonly used methods for correct determination of physical node positions across Internet do not always return reliable results, therefore mapping
-accuracy is not guaranteed. Bellow, the details of the algorithms that use the spanning tree based or cluster based topology optimization methods are presented.
-
-\paragraph*{\bf Narada}
-
-Narada \cite{chu_esm_2000,chu_esm_2001,chu_esm_2002} is a generic protocol for creating self-adapting overlay networks, that can achieve application-layer multicast communication without requiring IP multicast infrastructure at the network layer.
-IP multicast is the term reffering to the method of sending IP datagrams to a group of receivers in a single transmission.
-Even though the method is available some years now, IP multicast has not taken off as anticipated since many believe that it violates the stateless design of the current Internet.
-For this reason it is widely deployed in contained environments like enterprises (i.e.
-IPTV applications - distance learning, video conferencing), commercial stock exchanges, and multimedia content delivery networks.
-Narada tries to break these constraints.
-Although it was not originally designed as a P2P file sharing protocol, it has become one of the pioneering works in the overlay networks area.
-The authors claim that, although Narada is not as efficient as IP multicast, still it provides reasonable performance, without the additional cost of implementing multicast services on the current Internet infrastructure.
-Additionally, authors realised the inefficiency caused by the topology mismatch problem and tried to encounter it by building a richer connected graph, called a mesh, and building per source minimum spanning trees \footnote{ENarada \cite{li_enarada_2008} used Gossip protocol for the construction}.
-They also keep the
-graph and the trees dynamically updated while nodes continue to join and leave the network.
-Moreover, the protocol tries to ease the physical link stress, the overall resource usage and the relative delay among end systems.
-Unfortunately, the most imporpant limitation of Narada is that although it works reasonably well for small groups, it does not scale well for larger networks, therefore it is not suitable for, potentially very large, P2P file sharing application networks.
-
-\paragraph*{\bf Adaptive Overlay Topology Optimization and Adaptive Connection
-Establishment}
-
-\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{liu_auto_2003} is one of the first attempts, along with Narada, from the P2P research community to address the topology mismatch problem.
-AOTO is a distributed algorithm that uses \emph{Selective Flooding} and \emph{Active Topology} to optimize the overlay network topology.
-The \emph{Selective Flooding} algorithm creates a minimum spanning tree over the overlay and uses this tree instead of flooding the whole network, without shrinking the search scope.
-In \emph{Active Topology}, the physical locations of nodes are estimated based on the network delay among them and physically close nodes are connected as neighbours to revise the overlay topology as close as possible to the physical network topology.
-Overall, AOTO reports a $55\%$ performance improvement in terms of the traffic load.
-In \emph{Adaptive Connection Establishment (ACE)} \cite{liu_ace_2004}, the authors extend the idea by introducing optimizations based on the depths of the minimum spanning trees.
-But since the network delay is not always a reliable estimation method to detect the physical location of peers, the algorithm still suffers from the discrepancies caused by mislocated nodes.
-
-\paragraph*{\bf Location-aware Topology Matching}
-
-\emph{Location-aware Topology Matching (LTM)} \cite{liu_ltm_2004} proposes a method to optimize the overlay structure of the P2P network based on the physical topology.
-To achieve this, the algorithm issues a special message called \textit{TTL2-detector} to detect the latency of two-hop neighbours ($N^2$) around each node.
-The latency information gathered is later used to evolve the overlay network into a more efficient one, without reducing the search scope.
-Each node examines the latency information with the neighbours and low productive connections are dropped and replaced by more efficient ones, thus reducing the latency on the overall network, and eliminating some potential inefficiencies, such as unnecessary crossing of messages over autonomous system (AS) boundaries, etc.
-Although LTM improves the overall efficiency of the P2P network, it does not actually use real physical topology information, just latency metrics, therefore it does not offer a guaranteed safety net for the topology mismatch problem.
-
-\paragraph*{\bf Scalable Bipartite Overlay}
-
-\emph{Scalable Bipartite Overlay (SBO)} \cite{liu_bipartite_IPDPS,liu_bipartite_2007} reduces the overhead of creating and maintaining a minimum spanning tree cost by randomly dividing the nodes into two groups, namely red and white and assigning different tasks to different groups.
-The white peers measure distances to neighbours by using the network delay as the metric and reports the results to red peers.
-The red peers, equipped with the
-information of all two-hop neighbours ($N^2$), creates a minimum spanning tree
-of these neighbours and assigns efficient forwarding paths. The white peers can
-further optimize their positions within the overlay if this is required.
-
-%\paragraph*{\textbf{Two-Hop-Away Neighbour Comparison and Selection\\}}
-%
-%\paragraph{}
-%Work in \cite{liu_thancs_2005,liu_thancs_2008} proposes a distributed heuristic called \emph{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} that
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item is completely distributed and needs no global knowledge,
-%  \item presents trivial overhead compared to the query cost savings
-%  \item its convergent speed of the algorithm is fast enough (faster than minimum spanning tree approaches) so that is effective to dynamic environments, and
-%  \item does not shrink the search scope.
-%\end{inparaenum}
-%
-%\paragraph{}
-%THANCS is considered a \emph{local search method}, in the sense that it targets in finding a locally optimum solution, by exploiting knowledge within a 2-hop radius.
-The algorithm consists of two main components: \emph{piggybacking neighbour distance on queries} and \emph{neighbour comparison and selection} which are furtherly discussed bellow.
-%
-%\paragraph{Piggybacking neighbour distance on queries}
-%Using network delay as a metric for measuring the distance, each peer probes distances with its immediate naighbours and stores information locally.
-For this reason a special query message type, \emph{Piggy Message (PM)}, is introduced.
-It is 6 bytes long and includes two fields: Neighbour IP Address and Neighbour Distance.
-A peer $P$ constructs a PM for its neighbour $Q$, which contains $Q$'s IP address and $Q$'s distance from $P$.
-When $P$ receives a query from $Q$, this PM will be piggybacked by the query that goes to all other neighbours of peer $P$.
-Upon receiving such a query message, each of the other neighbours will detach the PM, record the $PQ$ distance and process the query.
-This PM will not be further forwarded.
-In selecting which incoming queries  should piggyback a PM the paper proposes the \emph{pure propability-based (PPB)} and the \emph{new neighbour triggered (NNT)} policies.
-%
-%\paragraph{Neighbour comparison and selection}
-%Figure~\ref{figure:thancs} illustrates this component of the THANCS algorithm.
-A peer $S$ probes the distance to all known unprobed $N^2(S)$\footnote{$N^2(S)$ denotes the set of peers being two hops away from $S$, while $N(S)$ denotes the set of direct logical neighbours of $S$.}.
-The distance of $SP$ is known to $S$.
-Upon receiving a PM from node $P$ with the distance of $PQ$, $S$ follows one of the following:
-%\begin{itemize}
-%  \item $Q \in \left( N(S) \cap N^2(S) \right)$, i.e.
-$Q$ is direct neighbour of $S$.
-In this case $S$ will compare cost of $SQ$, $SP$ and $PQ$.
-If the most costly connection is one of $SQ$ or $SP$ the corresponding link will be put into a \emph{will-cut list}\footnote{Links are not immediately disconnected when put in the will-cut list.
-There are useless for forwarding but are kept active, for some time, in order to serve query responses that are traveling to the source peer along the inverse search path.}.
-If the most costly connection is $PQ$ then $S$ will do nothing, as the fully distributed nature of the algorithm will give a chance to either $P$ or $Q$ to cut the connection.
-%  \item $Q \in \left( N^2(S) - N(S) \right)$, i.e.
-$Q$ is a two-hop-away neighbour of $S$.
-If $S$ hasn't probed $Q$ before\footnote{A distance cache is maintained and looked up in such a case.}, it probes and stores the result in the \emph{distance cache}.
-Having the distance of $SQ$, $S$ compares costs of $SQ$, $SP$ and $PQ$.
-If $SQ$ is the most costly, $S$ will not establish the connection.
-If $SP$ is the most costly, $S$ will establish connection $SQ$ and put $SP$ in the will-cut list.
-If $PQ$ is the longest, $S$ will keep  the connection with both $P$ and $Q$, expecting that $P$ or $Q$ will eventually disconnect link $PQ$, later.
-%\end{itemize}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/thancs.jpeg}
-%\caption{Probing two-hop-away neighbours}
-%\label{figure:thancs}
-%\end{figure}
-%
-%\paragraph{}
-%In a static environment THANCS has been proven to be effective; optimizing 45 percent out of the 60 percent of mismatched paths, constructing a nearly optimal overlay.
-This leads to a 60 percent reduction in traffic cost as well as a 40 percent decrease in query response time.
-In dynamic environments (Gnutella 0.6/Limewire super-peer-like and Ion flat-like), THANCS saves up to 70 percent of the traffic cost in the super-peer topology and 55 percent for the flat one.
-Average response time is also decreased by 60 and 45 percent, respectively.
-Generally, THANCS has similar performance to LTM, without needing synchronization.
-SBO, incurring half the  overhead of AOTO, reduces the traffic cost the most, while THANCS has lower response time and converges faster than SBO.
-THANCS is, thus, more suitable for a more dynamic environment.
-In addition, THANCS is easy to implement and its operation overhead is trivial, compared with the other three approaches.
-This design, however, has the limitation of not being easily extend to also support non-flooding-based systems.
-
-\paragraph*{\bf Peer-exchange Routing Optimization Protocols}
-
-\cite{qiu_prop_2007} introduces two protocols called \emph{Peer-exchange Routing
-Optimization Protocols (PROP)} to adjust the neighbourhood graph of an overlay
-network in order to reduce its overall link latency.
-PROP algorithms are based on the exchange of neighbours among peers, driven by the mutual benefit of the participants to reduce the network delay among their connections.
-In PROP-G (Generic), a peer can exchange all of its neighbours with another peer, while in PROP-O (Optimized) two peers exchange a selected subset of their neighbours.
-PROP-G is a generic protocol that guarantees the connectivity of the overlay graph during exchanges, therefore can be applied to both unstructured and structured overlay networks.
-
-\paragraph*{\bf T2MC}
-
-T2MC \cite{shi_t2mc_2008} uses traceroute logs to detect different autonomous systems and
-cluster close by peers, thus reducing redundant traffic that crosses autonomous system boundaries. T2MC uses a customized k-mean classification
-algorithm with $k=2$ to perform the classification, and exploits the stable
-structure of the Internet routers to guide clustering.
-Even though traceroute provides detailed information about the network
-structure, its excessive use creates additional overhead that stresses further the overall network infrastructure.
-This forces network administrators to disable its support from the routers they manage, thus jeopardizing the effectiveness of the T2MC algorithm.
-
-%\paragraph*{\textbf{T2MC\\}}
-
-%\paragraph{}
-%\emph{T2MC}\cite{shi_t2mc_2008} exploits some properties of the Internet paradigm and clusters nodes belonging to the same ISP without any centralised control or predefined system parameterization.
-The algorithm considers the dynamic nature of peers and exploits the stable nature of routers in order to build a topology location relationship among end peers.
-%
-%\paragraph{}
-%Some special routers split the physical network into autonomous system domains.
-Using a Traceroute mechanism T2MC searches for latency leaps among the path to a host in order to form ``near'' and ``remote'' router clusters.
-This is achieved by a 2-Means Classification, defined by the following steps:
-%\begin{enumerate}
-% \item the peer chooses the minimum and maximum Latency results from the Traceroute for initializing the cntroinds of two sets ``first'' and ``second''.
-% \item The peer calculates, for all hops allong its tracerouted path, the absolute distance to the centroids of both sets and assigns the routers to that centroid with which it has the smaller absolute distance.
-% \item The peer calculates the latncy mean and variance value of two sets
-% \item If the variance is larger than a predefined threshold then the algorithm takes a loop from step 2 picking the two latency mean values as new centroids of sets ``first'' and ``second''.
-%\end{enumerate}
-%Ultimately peer will end up with two sets having the minimum intra-set variance.
-Finally the peer chooses the router from ``second'' set with the minimum hops attribute and sets it as a threshold.
-The selected router and all others whose hops attribute is larger than the threshold are classified as ``remote'' router cluster.
-The remaining are classified as ``near''.
-From the ``near'' class, the peer chooses the one with maximum hops attribute as its edge router, and registers it along with the all the ``near'' cluster into the DHT of the p2p overlay.
-As new peers join the network, those that share the same edge router or any of the members of the ``near'' router clustersm ther would gather to form a ``close'' peer cluster.
-As Edge routers can provide more valuable information than other members of the ``near'' set, T2MC was designed to prioritize interaction of peers and edge gateways.
-%
-%% TODO: figure t2mc.jpeg (a) stars represent edge routers (b) hop leaps
-%
-%\paragraph{}
-%The use of Traceroute as a tool for implementing the distance measuring infrastructure raise concearns about its efficiency and scalability.
-Being, primarilly, a network diagnostic utility, it is concearned too heavy weighted and intrucive for use in a larger scale\cite{ratnasamy_binning_2002}.
-Additionally, disabling ICMP is a common administrative policy for edge sites to enforce security, while dumping BGP routing tables\cite{krishnamurthy_bgpclust_2000} is not directly available to the application layer.
-
-\paragraph*{\textbf{Resolving the Topology Mismatch Problem in Unstructured Peer-to-Peer Networks (?)}}
-% TODO: to be reviewed
-
-In \cite{hsiao_redblue_2009}, Hsiao et al, claim to construct topology-aware unstructured overlays that \emph{guarantee} performance qualities in terms of
-\begin{inparaenum}[\itshape i\upshape)]
-  \item the expected communication latency among any two overlay peers regardless of the network size, and
-  \item the broadcasting scope of each participating peer
-\end{inparaenum}
-.
-The algorithm constructs an undirected graph $G = \left( V, E \right)$ comprised by two subgraphs.
-The first, namely $G^{\left( red \right)} = \left( V^{\left( red \right)}, E^{\left( red \right)} \right)$ in the paper's context, includes all vertices of $G$ and ensures the connectivity of the graph by securing at least one path between any two nodes.
-In contrast, $G^{\left( blue \right)} = \left( V^{\left( blue \right)}, E^{\left( blue \right)} \right)$, contains those vertices of $G$ that have free edges to link to other nodes and because these are fully utilized, the following also stands $E = E^{\left( red \right)} \cup E^{\left( blue \right)}$.
-A joining peer $u$, partitions its neighbours into two subsets, the $B_u^{\left( red \right)}$ and $B_u^{\left( blue \right)}$.
-In order to populate the $B_u^{\left( red \right)}$ subset, peer $u$ samples peers uniformly and at random.
-Then, each of these selected peers discovers a routing path starting from itself towards the node with the smallest (or the largest) ID in the system.
-
-\paragraph*{\bf Distributed Domain Name Order}
-
-\emph{Distributed Domain Name Order (DDNO)} \cite{zeinalipour-yazti_ddno_2005}
-uses the domain names to detect topologically close nodes based on the
-assumption that nodes within the same domain are topologically close to each
-other. Half of the possible connections of a node is used to connect to these
-local peers, and the other half is used to randomly connect to the peers
-anywhere on the network.  \textit{DDNO} is a heuristic approach to solve the
-topology mismatch problem, with local connections to improve the efficiency and
-reduce the delay on the network.  The random connections on the other hand help
-ensure the connectivity in the network and avoid partitioning. \textit{DDNO}
-uses \textit{Split-Hash} and \textit{dnMatch} algorithms to detect locality by
-using domain names.
-
-\subsection{Landmark Based Proximity}\label{sec:landmark}
-
-Detecting the proximity of a node using landmarks, also known as \textit{Landmark
-Clustering}, is based on the view that nodes with similar distances to a
-set of predefined well-known landmark nodes are pretty likely being close to
-each other. But this approach has its weaknesses itself, like the fact that
-is a rather coarse grained approximation, therefore not particularly well suited
-for detecting the correct positions of nodes within close distance to each other.
-
-\paragraph*{\bf Landmark Binning}\label{sec:landmark_binning}
-
-Landmark Binning \cite{ratnasamy_binning_2002} is an approach to partition close
-by nodes into bins based on their distance to well known anchor nodes across the Internet. In order to detect locality, nodes use network latency (round trip
-time) as a measurement technique. The network latency, even though not always
-accurate, is selected in this work due to its non-intrusive, transparent and
-easy to apply behaviour.
-In order for the binning to work, few well known anchor servers with known physical locations on the Internet.
-Authors estimate that a number of 12 such servers can prove sufficient for this.
-The nodes measure their distance to all these servers and record the latency values for each.
-Later, the tuple of these latency values represents the bin of each node.
-Nodes with similar latency values are considered close to each other, or at least within the same autonomous system.
-The operation of the method is independent of the model incorporated by the overlay network and thus it can be applied with no significant changes to both structured and unstructured P2P systems.
-\cite{ratnasamy_binning_2002} has the detailed description of such algorithm for both architectures.
-Landmark binning is also considered as a good candidate for content distribution networks in particular.
-The major disadvantage of landmark
-binning is to install and maintain landmark servers on different autonomous systems, all over the world.  A typical P2P network usually has a couple of million nodes
-connected at any given time, thus rendering landmark servers an important parameter when designing for large scales.  The authors claim that latency estimation does not drain
-much from the network resources but in case scalability problems arise, they claim the solution is replacing single landmark servers with clusters of servers
-within the same physical area. However, this approach does not reduce the
-possibility of excessive network traffic flow through these landmark servers. One other
-possible problem is the incorrect binning caused by the use of inaccurate methods for delay measurement, since tracking down, for example, network latencies is far from proven to be safe for location estimation.
-
-% TODO: MAYBE THANCS deserves some more attention here
-
-\paragraph*{\bf mOverlay}
-
-mOverlay \cite{zhang_moverlay_2004} tries to addresses the scalability issues that might be imposed to networks when using static landmark servers by introducing the dynamic landmarks.
-The authors introduce the notion of a group which is a set of  peer considered close to each other with respect to any  position in the underlying network, proximity being some  user-defined cost metric (i.e.
-round-trip-time).
-mOverlay, tries to recreate small-world-like properties to the overlay network by creating a a two-level hierarchical structure, where on the top level we have connections between groups while on the bottom we have connections between peers inside groups.
-Finding the correct closest group is the most important part  of the overlay construction.
-Nodes are grouped based on their  distance to the groups already in the network, rendering the  later the (dynamic) landmarks for the process.
-The authors formally prove that any new node can reach its group by exchanging at most $O(logN)$ messages within the network.
-Finally mOverlay also considers a second important function in its protocol which is the constant maintenance of the overlay.
-
-\renewcommand\arraystretch{1.4}% (MyValue=1.0 is for standard)
-
-%\begin{figure}[h!]
-\hspace{-3ex}
-\begin{center}
-\footnotesize
-%\begin{tabular}{
-\begin{landscape}
-\begin{longtable}{
-|>{\columncolor[gray]{.7}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.2\columnwidth}
-|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
-|}
-\caption{Decentralized Unstructured Algorithms} \label{fig:unstruct_compare_table} \\
-\hline
-\rowcolor[gray]{.5}
-\textbf{Algorithm} & \textbf{Arch.} & \textbf{Overlay structure} & \textbf{Base protocol} &
-\textbf{Dynamic update} & \textbf{Runtime} & \textbf{Scalability} & \textbf{cites}\\
-\hline
-\endfirsthead
-\multicolumn{4}{c}%
-{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
-\hline
-\rowcolor[gray]{.5}
-\textbf{Algorithm} & \textbf{Arch.} & \textbf{Overlay structure} & \textbf{Base protocol} &
-\textbf{Dynamic update} & \textbf{Runtime} & \textbf{Scalability} & \textbf{cites}\\
-\hline
-\endhead
-\hline \multicolumn{4}{r}{\textit{Continued on next page}} \\
-\endfoot
-\hline
-\endlastfoot
-\textbf{Narada} & Decentralized unstructured & \textbf{Overlay optimization
-based}. Creates a mesh (richer connected graph) and builds minimum spanning
-trees on this mesh & & Yes. Supports incremental evolving & & Small and sparse
-groups & 2000  \\
-
-\hline
-\textbf{Gia} & Decentralized unstructured & \textbf{Forwarding based} Replaces
-Gnutella flooding with random walk, and introduces KaZaA style supernodes. Uses
-dynamic topology adaptation protocol &
- Gnutella & Yes & & Better than Gnutella & 974  \\
-
-\hline
-\textbf{Adaptive Overlay Topology Optimization} & Decentralized unstructured & \textbf{Overlay optimization
-based}. Creates overlay multicast tree with Selective Flooding protocol&
-Gnutella & Yes, using Active Topology protocol & & Better than Gnutella & 36 \\
-
-\hline
-\textbf{Location-aware Topology Matching} & Decentralized unstructured &
-\textbf{Overlay Optimization Based}. Uses \textit{TTL2-detector flooding}, \textit{low productive
-connection cutting}, and \textit{source peer probing}. & Gnutella & Yes & & Better than Gnutella & 204 \\
-
-\hline
-\textbf{Replication Strategies in Unstructured P2P Networks} & Decentralized unstructured &
-\textbf{Cache Based}. Uses uniform, proportional and square root allocation
-strategies to replicate data. & Gnutella & Yes & & Better than Gnutella & 580 \\
-
-\hline
-\textbf{Tracing a large-scale Peer to Peer System: an hour in the life of Gnutella.} & Decentralized unstructured &
-\textbf{Cache Based}. Proposes a caching algorithm based on the traces of the Gnutella traffic & Gnutella &
-Yes & & Better than Gnutella & 199 \\
-
-\hline
-\textbf{Improving search in P2P networks} & Decentralized unstructured &
-\textbf{Forwarding Based}. Uses \textit{iterative deepening}, \textit{directed
-BFS}, and \textit{local indices} to improve efficiency. & Gnutella & Yes & & Better than Gnutella & 380 \\
-
-\hline
-\textbf{Distributed Cycle Minimization Protocol} & Decentralized unstructured &
-\textbf{Forwarding based} Uses a decentralized cycle elimination protocol  &  & Yes &  &  & 9 \\
-
-\hline
-\textbf{Scalable Bipartite Overlay} & Decentralized unstructured &
-\textbf{Overlay optimization based} Uses bipartite partition graph and builds
-local minimum spanning trees  & Gnutella & Yes &  & Better than Gnutella & 77 \\
-
-\hline
-\textbf{Adaptive Connection Establishment} & Decentralized unstructured &
-\textbf{Overlay optimization based} Forms Neighbour Cost Tables, builds local
-minimum spanning trees and perform local optimizations & Adaptive Overlay
-Topology Optimization (AOTO), Gnutella & Yes &  & Better than Gnutella & 110 \\
-
-\hline
-\textbf{Hops Adaptive Neighbour Discovery} &  &
- &  &  &  &  & 1 \\
-
-\hline
-\textbf{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} & Decentralized unstructured &
-\textbf{Overlay optimization based} Uses piggybacking to discover neighbour
-distances and selects neighbours  & Gnutella & Yes &  &  & 36 \\
-
-\hline
-\textbf{mOverlay} & Decentralized unstructured &
-\textbf{Overlay optimization based} Uses dynamic landmarks to find node locality
-& & Yes &  & Due to dynamic landmarks and grouping, more scalable than tree-based or mesh-based protocols & 123 \\
-
-\hline
-\textbf{Distributed Domain Name Order (DDNO)} & Decentralized unstructured &
-\textbf{Overlay optimization based} Connects half of the nodes connections to
-the nodes in the same domain and the other half to random nodes, therefore
-supports locality and topological connection  & & Yes &  & Yes, by using super
-peers & 5 \\
-
-\hline
-\textbf{Peer-exchange Routing Optimization Protocols} & Works with both decentralized
-structured and unstructured &
-\textbf{Overlay optimization based} Optimizes overlay by the exchange of
-neighbors among peers  & & Yes &  & Yes & 13 \\
-
-\hline
-\textbf{MAY OMIT- T2MC} & Decentralized unstructured &
-\textbf{Cluster based---{\sethlcolor{yellow}\hl{ IS IT ALSO PROXIMITY BASED?}}} Uses traceroute results for clustering the
-nodes  & &  &  &  & 1 \\
-
-\hline
-\textbf{Unnamed-unstructured} & Decentralized unstructured &
-\textbf{Overlay optimization based} Minimizes the communication delay and
-maximizes the broadcasting range & & Yes &  & Better than THANCS and mOverlay & 7\\
-
-\hline
-\textbf{Landmark Binning} & Can work with both decentralized structured and unstructured &
-\textbf{Proximity based, landmark binning} Uses network latency to partition
-nodes into bins & & Yes &  &  & 817 \\
-
-\hline
-%\end{tabular}
-\end{longtable}
-\end{landscape}
-\end{center}
-\vspace{-2.5ex}
-\vspace{-2.5ex}
-%\end{figure}
-
-\subsection{Discussion on the Algrithms for Unstructured Architectures}
-
-In this section we presented the state of the art for the algorithms applied to unstructured P2P systems.
-We categorised it based on their philosophy for alleviating the topology mismatch problem into four major categories namely
-\begin{inparaenum}[\itshape i\upshape)]
-  \item broadcast-based,
-  \item cache-based,
-  \item overlay optimisation-based and
-  \item landmark-based approaches
-\end{inparaenum}
-.
-Here we present a overall discussion on the above categories trying to point out the advantages, disadvantages and novelties introduced by each and by all of them.
-A quick overview can be seen in Table \ref{fig:unstruct_compare_table}.
-
-Broadcast-based approaches in general propose intelligent neighbour
-selection in order to replace the inefficient blind flooding first used in the Gnutella protocol.
-Instead of forwarding the queries to all their outgoing links they dictate the intelligent selection of neighbours with either higher probability to answer a query than others or neighbours  that expose specific features such as high bandwidth, high capacity, or low latency.
-This, in practice reduces the aggregate resource usage of the P2P network and improves the overall performance of the system.
-On the other hand, the algorithms do not offer a real solution to the topology mismatch problem.
-Thus, none can guarantee that overlay and underlying topologies are aligned with each other let alone quantify the mismatch and try to alleviate.
-Broadcast-based methods can be used in conjunction with other approaches to improve the quality of the P2P systems.
-
-The success of caching, as a well studied method in client-server Internet
-applications, also made it a good candidate for the P2P networks. Even
-though caching improves the performance, and reduces the overall resource usage
-of P2P systems, the design of caches in this context is non-trivial compared to the web-based caching.
-Due to the unique, inherent characteristic of all P2P systems that, each node can act as both a server and a client, two important issues have to be considered when designing caches.
-First, the lifetime of a query is short, as the nodes join and
-leave frequently. Second, the result of a single query string is not always the
-same, as it is dependent on the source of the query, the TTL value set for the messages, the current interconnection of peers and the high volatility of the environment.
-So, in order to develop a successful caching system for a P2P architecture, these parameters also have to be considered.
-Even though the state of the art P2P algorithms using caching methods reduce the resource usage of the network, they also fail to consistently address the probelem of the topology mismatch.
-
-Overlay-optimisation-based  can further be devided into minimum-spanning-tree-based and clustering-based.
-Researchers initially started considering alternatives to IP multicast for the application layer, which led Narada, that simulates the same one-to-many communication approach requiring IP multicast services.
-Application layer multicast, however, is not as efficient as the one enabled by IP and still struggles to ease the difficulties imposed by the topology mismatch problem.
-Narada, and its followers (AOTO, LTM, SBO) try to solve the topology mismatch problem by building a richer connected graph and forming minimum spanning trees over this graph that can efficiently route messages among peers.
-Application level multicast protocols, pioneered by Narada, are generic protocols that can be applied to both P2P file sharing systems, as well as content distribution networks.
-The original Narada algorithm is designed for small groups therefore it is not scalable.
-Its followers try to solve the scalability problem by introducing
-various methods including forming minimum spanning trees for the two-hop neighbours ($N^2$) of each node, partitioning the graph into two random groups where each group is
-responsible of different tasks, or performing local optimizations dynamically on
-the overlay graph. The advantage of building minimum spanning trees is that they maintain the connectivity on the network in an efficient manner, while
-still preserving the overall search scope. However, building and maintaining a
-minimum spanning tree creates additional overhead for the network. 
-The alternative method used in overlay optimization is the cluster based
-approache. According to that, nodes use various methods to detect
-physically close nodes to form clusters. T2MC uses traceroute logs and DDNO uses
-domain names to cluster close by nodes. However, the accuracy of the methods
-used to form clusters directly affect the success rate of the algorithm. Traceroute for
-example is a heavy weight utility to be frequently used on the network, let alone the fact that many network equipment vendors do not allow traceroute calls. The major problem with
-cluster based approaches is that the limited connectivity within the local
-domains shrink the search scope dramatically, which negatively affects the
-search performance of the P2P system. DDNO addresses this limitation by allowing
-half of each nodes connections to be to random nodes over the network, which
-balances the efficiency of the clustering approach with improved connectivity.
-
-In landmark-based algorithms, nodes use network delay (round trip
-time) as a distance measurement method to position themselves with respect to ``a priori'' known servers (the landmark servers) on the Internet.
-Landmark servers are used by nodes to estimate their positions based on the intutive assumption that nodes with similar distances to a set of landmarks, are physically close to each other, as well, over the network.
-The landmark based protocol has two important drawbacks.
-Firstly the network delay is not a reliable distance estimation method.
-For example, based on the load on the network the delay to certain nodes or networks can change from time to time, which will eventually affect the distance measurements and wrong measurements will lead to wrong estimated positions for the nodes, or incorrect and non optimal clusterings of the nodes.
-The second drawback of using landmark servers is the cost of installing and maintaining this kind of infrastructure across the whole Internet and for all the different autonomous system domains.
-As popular P2P file sharing applications usually have millions of peers connected at any time, it is not false to assume that the network costs of maintaining these landmark servers will be quite high.
-A possible solution to the scalability problem of the static landmark servers is to use ordinary nodes as dynamic landmarks once they estimate their own positions.
-Even though this approach scales much better than static landmark servers, still the measurement accuracy problem affects the overall performance of the system.
-
-\section{Structured Decentralized Algorithms}\label{sec:structured}
-
-This section focuses on the structured P2P architectures.
-The following paragraphs contain a thorough discussion and analysis of the algorithms proposed in the literature for dealing with the topology mismatch problem in such architectures.
-In structured P2P systems, the construction of the routing tables in every participating node determines the efficiency of message forwarding.
-Therefore, routing tables that represent the underlying physical structure well can achieve increased performance.
-For this reason, the proposed methods for topology mismatch problem basically use different levels of proximity information to optimize them towards this direction.
-The algorithms are here categorised based the previous work done by \cite{castro_proximitydht_2002,castro_topawareroute_2002,ratnasamy_openq_2002}.
-
-%{\sethlcolor{yellow}\hl{HA:  Talk about Tapestry, Pastry, CAN and Chord briefly here!}}
-
-\subsection{Geographic Layout}
-
-Inside geographic layout category fall those kind of methods which practically position physically close by nodes together in the application space as well.
-This is done by adjusting and maintaining the routing tables of all involved peers accordingly by exploiting proximity information that describe the overall geographic positioning of the peer nodes.
-One popular method for detecting and constructing a geographic view of the underlying network is the use of well established servers dedicated to landmarking the physical space, pretty much as discussed in Section \ref{sec:landmark} of landmark based proximity in the unstructured P2P networks.
-The same disadvantages of difficult and inaccurate positioning and the cost of deploying and maintaining these landmarks apply in this case as well.
-% TODO: maybe a discussion sentence bellow
-Even though managing the overlay structure based on the geographic layout of the nodes improves the query efficiency of the system, on the other hand, it tends to create hotspots, and the needed failure resilience is undermined by the fact that close by nodes are more likely to suffer collective failures.
-
-\paragraph*{\bf Global Soft-State}
-
-\textit{Global Soft-State} \cite{xu_globstate_2003} builds a global map to help
-choose shorter routing paths, combining the landmark binning method and small
-scale distance probes to reveal the proximity properties of the underlying
-network to the overlay. This global view of the state is made available to all
-nodes in order to help them find the best way to route their messages.
-\textit{Global Soft-State} operates in two main stages, generation and use of
-the proximity information. For generating the proximity information a hybrid
-approach is proposed, which uses landmark clustering as a preprocessing step in
-order to select a number of potential nearest neighbour candidates and then
-refine the selection by incorporating a round-trip-time-based scheme to ultimately choose the
-closest node. For using the proximity information, the algorithm chooses a
-different path from the classic gossiping approaches for constructing and
-maintaining the overlay. It is based on landmark clustering for strategic
-placement of proximity information on the overlay enabling any node to access
-such information using a landmark number that reflects its physical position in
-the network. For various logical regions\footnote{This might be a high-order
-zone in the eCAN\cite{xu_ecan_2002} context or a set of nodes sharing a
-particular prefix in overlays such as Pastry.} maps of physical information are
-built and published where each node may appear in a maximum of $log\left( N
-\right)$ such maps. To dynamically adapt to changing network conditions, a node
-subscribes to relevant \emph{soft states} that utilize a notification system in
-order to initialize any necessary neighbour re-selection.
-
-Maintaining several host states at different layers, makes any content migration
-costly. Additionally, the method does not make any continuing effort to remap the
-overlay structure after a node successfully joins, in order to adapt its state to
-any condition change. Although this approach greatly reduces the
-routing latency to far nodes, it is unable to dynamically identify nodes that are
-close to routers and gateways in order to construct the secondary overlay.
-Nevertheless, static recognition of such nodes is currently done based on BGP
-reports and pre-chosen landmarks, sacrificing the self-organising attribute of
-traditional DHTs.
-
-\paragraph*{\bf Mithos}
-
-\textit{Mithos} \cite{waldvogel_mythos_2003} is a P2P protocol which
-incorporates directed incremental probing to find near optimal node placement,
-and is classified by its authors as an integration of geographic layout
-and proximity routing overlay optimization methods.
-
-The bootstrap phase of \textit{Mithos} starts with a subset of existing members
-as the first set of candidates, and while the iteratively closest nodes are
-detected by probing the neighborhood, the candidate neighbor list is updated. In
-order to avoid a local minima \textit{Mithos} probes all the neighbours within a
-two hop distance from the current minimum before concluding the process.
-
-After finding its first neighbour, the newcomming peer is assigned an ID using
-information gathered during the iterative neighbour selection phase. Virtual
-coordinates are assigned to the newcomming peer by using the distances of two
-closest nodes and their neighbours, so that Euclidean distances between the node
-and all known hosts predict the network latency between
-them\cite{cox_vivaldi_2004}.  The benefit of these synthetic coordinates are
-that they are explicitly used as the node's ID, and distance computations among
-nodes can be done in ID space without requiring physical measurements.
-
-The last step of the algorithm is the interconnection among neighbours.
-\emph{Mithos} uses a \emph{quadrant}-based mechanism according to which each
-node establishes a link to the closest neighbour in each quadrant. During
-forwarding, the next hop is performed towards a neighbour in the same quadrant
-as the final destination. The newcoming node may not know of other neighbours
-in all quadrants, therefore, the node first identifies neighbours in all
-quadrants using a mechanism based on ideas similar to a perimeter
-walk\footnote{Used in Greedy Perimeter Stateless Routing (GPSR) protocol.} and
-then improves the results using parallel path processing by taking into account
-further geometric properties of node relationships.
-
-One major limitation of \textit{Mithos} is that the protocol cannot handle
-dynamic arrivals and removals from the network, as is happens in mobile
-networks.
-
-\paragraph*{\bf LAPTOP}
-
-\textit{LAPTOP} \cite{wu_laptop_2007},  organises the overlay into a tree based
-hierarchy with main focus on reducing hops during message routing as well as
-minimizing maintenance overhead. Additionally, a caching scheme is also
-incorporated so as to further reduce routing table update costs. The authors
-theoretically show that in \textit{LAPTOP} routing path length is bounded by $O(log_d
-N)$ and node joining and leaving in the overlay network is bounded by $O\left( d
-log_d N \right)$ hops in a balanced overlay tree, where $N$ is the number of
-nodes, and $d$ is the maximum degree of each node. \textit{LAPTOP} implements
-the geographical layout approach  and constructs a geographical layout in a
-self-organizing and efficient fashion, by estimating the round-trip-times to a small number of nodes in the overlay network in order to make them roughly
-aware of their physical distances among them.
-
-Each node is assigned an address in a dotted format (e.g 1.3.4). Each octet
-ranges from $1$ to $d$, where $d$ is the maximum degree of the nodes. The
-assignment process is done by appending a unique octet to the address of each
-nodes
-parent, while root node is assigned address 1.  The routing scheme is similar to the
-longest-prefix IP matching scheme. At each forwarding hop, any message travels
-up the tree until the first common ancestor of both source and destination nodes is
-reached and then starts descending to arrive to its target. During tree
-traversals, special entries in the routing tables, called \emph{routing cache},
-are maintained in order to increase routing efficiency and achieve finer load
-balance. Caching enables a node to forward a message to a better longest-prefix
-match than that of its direct ancestor making a larger, quicker and more cost
-effective step through the overlay and toward the destination. To improve
-scalability, the number of children nodes and the size of the routing cache are
-limited.  In terms of overlay maintenance, \textit{LAPTOP} incorporates a simple
-\emph{heartbeat}-based technique where each parent node is responsible for
-monitoring its children.  At join process the newcomming node is assigned its
-level label as well as its address by its parent node. Additionally it
-initializes its routing table (with normal and caching entries) as it traverses
-the overlay in search for its parent node.  During a graceful departure, the
-referred node checks for children in the overlay. If it does not have any, it
-simply notifies its parent and leaves. If it has, it selects the child node with
-the lowest round-trip-time to it in order to take its place so that the locality property is
-preserved.
-
-\subsection{Proximity Routing}
-Proximity routing does not require routing tables to be built using any knowledge
-about network proximity. On the other hand it exploits such knowledge in order
-to choose the best next hop during routing a message. This approach 
-balances between choosing the node that will further progress the routing
-towards the destination and choosing the closest entry in the routing table, in
-terms of network proximity.
-% TODO: looks like discussion material
-Thus, it is relatively less effective than
-geographical layout when applied to CAN(-like) implementations. Moreover, the
-technique has been incorporated into a version of Chord causing an increase on the overhead of
-node joins and the size as well as maintenance cost of finger tables.
-%
-%\cite{dabek_cfs_2001} proposes a server selection scheme for the Chord DHT, on
-%the domain of proximity routing selection. In \emph{CFS}, each node predicts
-%the entire lookup latency as a function of the total number of nodes and the
-%average overlay next routing peer. The problem is that it is very difficult to
-%have a clear picture on the total number of nodes and the average hop latency
-%from the local. This leads to rough estimations that consequently decreases
-%overall performance.
-%
-The methods that use proximity routing and their details are listed below.
-
-\paragraph*{\bf Proximity in Kademlia}
-Before discussing {\em Proximity in Kademlia}, it would be helpful to briefly
-summarize the highlights about the {\em Kademlia} protocol.
-{\em Kademlia} \cite{maymounkov_kademlia_2002} is a distributed hash table (DHT) based
-protocol for P2P networks, which uses an iteration based lookup algorithm. The
-protocol uses the standard $160$ bit ID system for nodes and locates the nodes in
-a prefix binary tree, where IDs are used as prefixes. The iterative lookup
-operations are done over this prefix binary tree, which converges to logarithmic lookup
-times. The ID's are assigned randomly, therefore in Kademlia,
-there is no proximity control, which results in inefficient use of the underlying
-network during lookup and retrieve operations.
-
-{\em Proximity in Kademlia} \cite{kaune_pkad_2008} introduces proximity controls
-over the base Kademlia protocol to optimize the underlying network usage of the
-protocol. The authors define an abstract {\em underlay metric} that calculates
-the suitability of establishing a communication link between peers as a cost
-function, based on the used proximity criteria (RTT, ISP locality, etc.).
-{\em Proximity in Kademlia} adapted both {\em proximity routing} and {\em
-proximity neighbour selection} overlay optimization algorithms for controlling proximity. Therefore,
-we categorize the protocol only in this section, and skip the discussion in the
-further {\em proximity neighbour selection} section (Sec.\ref{sec:pns}). {\em
-Proximity in Kademlia} also uses the MaxMind GeoIP database\footnote{MaxMind
-Geolocation Technology. http://www.maxmind.com} to detect the
-proximity information of a given peer based on its region, country, or ISP
-location, and use this information to form clusters of nearby peers. One
-other method {\em Proximity in Kademlia} uses is the Vivaldi \cite{cox_vivaldi_2004} protocol. However, authors report that
-clustering approach performs better than Vivaldi protocol based on their
-experiments. Authors also report that {\em Proximity routing} protocol
-successfully worked with the Kademlia protocol and improved the locality of the
-connections over peers.
-
-\paragraph*{\bf CHOord considering Proximity on IPv6 (CHOP6)}
-
-\textit{CHOP6} \cite{morimoto_chop6_2007} is designed based on the
-\textit{Chord} protocol. \textit{CHOP6} roughly estimates the proximity among
-nodes by exploiting the IPv6 address format and RTT information if available.
-The proximity estimation is achieved by introducing a 64-bit ID scheme in which
-the least significant bit part is the IPv6 global routing prefix and thus
-enabling a longest prefix match scheme. The protocol is designed based on the
-observation that it is possible to estimate a node's geographical location by
-simply examining the upper 32-bits of its IPv6 address. Moreover, similar to
-\textit{Chord}, \textit{CHOP6} uses a finger table, whose entries hold more than
-one candidate node.
-
-\paragraph*{\bf Chord6}
-\emph{Chord6} \cite{xiong_chord6_2005} is another \textit{Chord} variant that
-tries to exploit the hierarchical features of IPv6 in order to create a
-substrate that reduces interdomain traffic between service providers.
-\textit{Chord6} is based on the original Chord protocol, and the main difference
-from the original protocol is in the identifier definition. Therefore, the
-approach can be easily portable to other DHTs such as CAN, Pastry and Tapestry.
-In Chord6 the identifier contains two parts: the higher bits are obtained by
-hashing the node's IPv6 address prefix of specific length, while the remaining
-lower bits are the hash value of the rest of that IPv6 address. As a result of
-this assignment, nodes in a domain will be mapped onto a continuous key space on
-the overlay network, which avoids unnecessary message forwarding across
-different service providers, thus minimizing overall routing cost.
-
-\subsection{Proximity Neighbour Selection}\label{sec:pns}
-\textit{Proximity Neighbour Selection} constructs the routing tables using proximity
-knowledge. The proximity information used in this method is different than the
-landmark based systems described in the \textit{Geographic Layout} section, as time-to-live values between
-nodes, or directly node ID prefixes are used to detect proximity.  Tapestry,
-Pastry, and CAN successfully implemented the proximity into their algorithms by
-using this approach.  The routing protocol in Pastry is based on longest node ID
-prefix matching, while CAN uses RTT values to detect close by nodes.
-\cite{castro_proximityp2p_2002}  reports that proximity neighbour selection as
-an effective proximity based method.
-%
-Proximity Neighbor Selection based algorithms are described in detail below.
-
-\paragraph*{\bf DHT-PNS}
-
-Chord-DHT-PNS \cite{hancong_pnsbased_2006} implements proximity neighbour
-selection on top of the Chord DHT. The main goal is to use proximity
-information to group physically close by nodes as neighbours in the DHT table.
-In order to detect proximity, virtual network coordinates of peers are used, by
-using the Vivaldi protocol \cite{cox_vivaldi_2004}. The virtual coordinates are
-then used by the nodes to map to identifier space in the DHT. The space is
-partitioned using a \emph{concentric circle clustering scheme} where successive
-cycles of radiuses $\rho$, $2\rho$, $3\rho$ and so on, are constructed. Then the
-formed annuluses are divided into $2\chi-1$ \emph{sectors}, where $\chi$ denotes
-the sequence number of the annulus starting from $\chi = 1$ for the centre
-cycle. It is proved in the paper, that this way each sector occupies the same
-area as does the center cycle. Assuming uniform node distribution, this
-characteristic, favours a more load balanced clustering operation. Every sector
-in this $2D$ coordinate space is mapped to a unique \emph{region} in the DHT
-space forming a multi-layer node identifier space.  Thus, any nodes that belong
-to the same sector, are mapped to the same region as well, preserving their
-proximity relationship unveiled by the use of the Vivaldi protocol. The
-individual pieces are mapped to the identifier space uniquely, allowing
-logarithmic lookup operations with high probability on Chord.
-
-\paragraph*{\bf IP-based Clustering (IPBC)}
-Proximity neighbour selection algorithms use probing and other techniques
-to detect proximity. However, such methods are either not precise enough, or
-overload the network. \emph{IP-based clustering}
-\cite{karwaczynski_ipbc_2007} is a proximity neighbour selection based
-algorithm in which the authors propose to use the IP address prefixes (16 bit
-for IPv4) in order to detect proximity.  \cite{freedman_iploc_2005} states that
-$97\%$ of prefixes larger than $24$ bits belong to a single geographical
-location. However, using less number of bits creates less precise results and
-more number of bits increase the burden on the network and reduce the possible
-number of neighbours. Therefore, a careful selection is required in terms of
-performance/accuracy tradeoff. The {\em IP-based Clustering} generates a key and
-its prefix and stores the later in the DHT itself, so that any newly joining
-nodes with the same IP prefix can query it and identify all the
-neighbours with the same one relatively easily. Nodes periodically update
-their entries in the DHT by removing those who leave, or those who has been timed out with no further activity detected from the node.
-
-\paragraph*{\bf Cone}
-
-\textit{Cone}\cite{wang_cone_2007} extends Chord using proximity neighbour selection
-topology optimization algorithm. The proximity information is generated using
-landmarks and round-trip-time-based distances to landmarks.  \textit{Cone} uses a two-layered
-identifier space. The first, named Chord-layer identifier, denoted as
-$Id_{Chord}$, is the same as in Chord. The second is the Cone-layer
-identifier, $Id_{Cone}$ which is constructed by two component identifiers. The
-first, known as \emph{group identifier (gid)} denotes a relevant group the node
-belongs to while the second, namely \emph{local identifier (lid)} indicates the
-local identifier within the group. The group concept, which is introduced here,
-is a way of dividing nodes according to a common $Id_{Chord}$ prefix.  The
-structure of a Cone overlay, retains the Chord's circular topology. The difference
-lies on the fact that, now, two rings are created. A big ring, where nodes with
-the same \emph{gid} are arranged at each position. Each of these positions are a
-smaller ring for the particular group's \emph{lid}s. The routing is achieved in
-both clockwise and counter-clockwise directions in the big-ring, for which two
-routing tables are maintained, namely \emph{front} and \emph{back} finger
-tables. Entries in these tables, display physical network
-proximity with the current node. Moreover, a third table called \emph{group}
-maintains information about other online peers within the current node's
-group in a way that entries are now close in the ID space.
-
-\paragraph*{\bf SAT-Match: Self-Adaptive Topology Matching}
-
-\textit{SAT-Match} \cite{ren_satmatch_2004} is a protocol that specifically
-tries to solve the topology mismatch problem by mapping the overlay network as
-close as possible to the physical one. Even though no theoretical proof is
-given, valid results are obtained and demonstrated through experimentations.
-Similar to the other approaches in the proximity neighbour selection,
-\textit{SAT-Match} uses probing to detect close by peers and obtain proximity
-information. However, as an additional feature, \textit{SAT-Match} peers can do
-selective jumps to adjust their positioning in the DHT, if it reduces the stretch
-of its one-hop neighbourhood. The stretch is defined as the ratio between the
-average logical and physical latency of a link. It is reported that, due to the
-selective jumps, the \textit{SAT-Match} achieves $40\%$ reduction in link
-stretch, and when used with the \textit{Landmark Binning} approach (see Sec.
-\ref{sec:landmark_binning}), the reduction rate increases up to $60\%$.
-For dynamic environments, with frequent node arrival and leave,
-\textit{SAT-Match} scales much better than \textit{Mithos}, due to its self
-adaptation mechanism and selective jumps.
-
-\emph{SAT-Match} uses a small TTL value for the probing messages in order to
-reduce redundancy\cite{jiang_lightflood_2008}. This process begins as soon as
-the node joins the network using a DHT mechanism. Each probing message contains
-information about the source and a small TTL value. The recipient of such a
-message returns information about itself to the source and forwards the probing
-message to its neighbours if the TTL is non-zero. The discovered nodes are
-referred to as $TTL-k$ neighbourhood of the source node based on the $TTL$ distance
-to the source node.
-
-Blindly selecting the peer with the smallest RTT as neighbour is, generally, not
-the optimal decision to make in order to achieve global \emph{stretch} reduction.
-In a structured scheme, when a node jumps to connect to a
-physically close node, it may need to connect to other distant nodes to maintain
-the structure's integrity, thus creating an overall increase in the overlay's
-\emph{stretch}. The two nodes with the smallest RTT are then used in order to
-select one zone to jump in this phase. The algorithm is as follows: The source
-node $S$ calculates the stretch change of its $TTL-1$ neighbourhood and that of
-the $TTL-1$ neighbourhood of the first of the previously selected peers. These
-calculations are made as if the jump has been made. If the stretch reduction is
-over a predefined threshold the jump is performed, otherwise the second selected
-candidate is picked and the same computations are performed. If again, the
-threshold is not met, then no jump is ultimately done. In case of a jump, this
-is performed as a combination of \emph{leave} and \emph{join} operations, in the
-CAN context.
-
-%\begin{figure}[h!]
-\hspace{-3ex}
-\begin{center}
-\footnotesize
-\begin{landscape}
-\begin{longtable}{
-|>{\columncolor[gray]{.7}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.2\columnwidth}
-|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
-|}
-\caption{Decentralized Structured Algorithms}\label{fig:struct_compare_table}\\
-\hline
-\rowcolor[gray]{.5}
-\textbf{Algorithm} & \textbf{Arch.} & \textbf{Overlay structure} & \textbf{Base protocol} &
-\textbf{Dynamic update} & \textbf{Runtime} & \textbf{Scalability} & \textbf{cites}\\
-\hline
-\endfirsthead
-\multicolumn{4}{c}%
-{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
-\hline
-\rowcolor[gray]{.5}
-\textbf{Algorithm} & \textbf{Arch.} & \textbf{Overlay structure} & \textbf{Base protocol} &
-\textbf{Dynamic update} & \textbf{Runtime} & \textbf{Scalability} & \textbf{cites}\\
-\hline
-\endhead
-\hline \multicolumn{4}{r}{\textit{Continued on next page}} \\
-\endfoot
-\hline
-\endlastfoot
-
-\hline
-\textbf{Global Softstate} & decentralized structured &
-\textbf{Proximity based, landmark clustering} Uses first landmark clustering
-then measures RTTs to identify close nodes & & Yes &  &  & 195 \\
-
-\hline
-\textbf{Mithos} & decentralized structured &
-\textbf{Proximity based} Uses nodes as topology landmarks and directed
-incremental probing to optimize topology & & Yes &  & Scales well as all
-operations are local ??? & 172 \\
-
-\hline
-\textbf{Self-Adaptive Topology Matching} & decentralized structured &
-\textbf{Proximity based} Uses lightweight probing and
-selective jumps to optimize the topology & CAN & Yes &  & Better than Mithos
-& 40 \\
-
-\hline
-\textbf{Delay Aware P2P System} & &
-\textbf{} & &  &  &  & 1 \\
-
-\hline
-\textbf{VERSION OF CHORD - DHT-PNS} & decentralized structured &
-\textbf{Proximity based} Uses Proximity Neighbour Selection and the Vivaldi
-system & Chord  & Yes &  &  & 5 \\
-
-\hline
-\textbf{MAY OMIT- VERSION OF CHORD - Quasi-Chord} & decentralized structured &
-\textbf{Proximity based}  & Chord  & Yes &  &  & 0 \\
-
-\hline
-\textbf{LAPTOP} & decentralized structured &
-\textbf{Geographic layout based} Hierarchical overlay structure  &   & Yes &  & routing path length $\log{_d N}$, 
-join/leave overhead $d\log{_d N}$& 4 \\
-
-\hline
-\textbf{IP-Based Clustering} & decentralized structured &
-\textbf{Proximity based} Proximity neighbour selection based on longest common
-prefix of IP addresses &   & Yes &  &  & 1 \\
-
-\hline
-\textbf{CHOord considering Proximity on IPv6} & decentralized structured &
-\textbf{Proximity based} Uses IPv6 address format to provide proximity &  Chord
-& Yes &  & Better than Chord & 1 \\
-
-\hline
-\textbf{Proximity in Kademlia} & decentralized structured &
-\textbf{Proximity based} Applies  proximity neighbour selection (PNS) and proximity route selection (PRS)
-to Kademlia & Kademlia & Yes &  &  & 13 \\
-
-\hline
-\textbf{Cone} & decentralized structured &
-\textbf{Proximity based} Uses proximity neighbour selection (PNS) & Chord & Yes
-&  & Better than Chord & 3 \\
-
-\hline
-\textbf{Dynamo} &  & &  & 
-&  & & 3 \\
-
-\hline
-\textbf{MAY OMIT-BADLY WRITTEN-PChord} & decentralized structured &
-\textbf{Proximity based}  &  Chord
-& Yes &  & & 16 \\
-
-\hline
-\textbf{AChord} & &  &  
-&  &  & & 7 \\
-
-\hline
-\textbf{Chord6} & decentralized structured &
-\textbf{Proximity based} Uses IPv6 hierarchical address format to cluster
-topologically close nodes & Chord & Yes
-&  &  & 14 \\
-
-\hline
-\end{longtable}
-\end{landscape}
-\end{center}
-\vspace{-2.5ex}
-\vspace{-2.5ex}
-%\label{fig:struct_compare_table}
-%\end{figure}
-
-\subsection{Discussion on the Algrithms for Structured Architectures}
-In this section we presented the state of the art for the algorithms applied to structured P2P systems.
-We categorised it based on their philosophy they incorporated in order to alleviate the topology mismatch problem into overall three categories following the previous work from \cite{castro_proximitydht_2002,castro_topawareroute_2002,ratnasamy_openq_2002} namely
-\begin{inparaenum}[\itshape i\upshape)]
-  \item geographic-layout-based,
-  \item proximity-routing-based, and
-  \item proximity neighbour selection based
-\end{inparaenum}
-.
-Following is a discussion on the above categories trying to point out the advantages, disadvantages and novelties introduced by each and by all of them.
-A quick overview can be seen in Table \ref{fig:struct_compare_table}.
-
-In section \ref{sec:structured},  we presented the state of the art structured
-decentralized P2P algorithms in three different categories based on their
-handling of the network structure to tackle the topology mismatch problem:
-geographic layout based, proximity routing based, and proximity neighbour
-selection based. In this section, we present a final discussion of all the
-methods including the advantages, disadvantages,
-and novelties related to solve the topology mismatch problem.
-Table \ref{fig:struct_compare_table} presents an overview of all the
-structured decentralized algorithms described earlier in this section.
-
-Structured P2P network algorithms use a global distributed hash table or a
-prefix tree structure to uniquely lookup peers or their data in the overlay
-network. As all the data is kept within the overlay, each node behaves as a
-client and a server, therefore nodes join and leave according to rules
-determined by the integrity of the global data structure. The main advantage of
-the structured P2P topology is that by the help of the global data structure,
-peers or their data can be found within the network even if there is only a
-single copy of that item present. However, each node join and leave creates
-maintenance overhead for the network due to updates required by the global data
-structure, and for networks with frequent node arrivals and departures the
-topology uses valuable network resources just to update the global structure.
-Nodes join the network by using a key value, which determines the location and
-the neighbourhood of the new node within the network. However, assigning
-random key values to the newly inserted nodes creates non-optimal matching with
-the underlying physical network topology, therefore, increasing the overhead of
-the network even more. One solution for handling the topology mismatch problem
-is to consider the proximity of the peers when generating the key and joining
-the node to the network, so that nodes within the same network domains are
-selected as peers, or neighbours, during the overlay topology construction. In
-this chapter, we have described three such approaches to optimize the topology
-matching problem: geographic layout, proximity routing and proximity neighbour
-selection.
-
-In geographic layout, nodes try to estimate the geographic positions of the
-peers, and construct the DHT considering the proximity of the peers. Landmark
-servers and RTT measurements are two popular methods, which can also be used in 
-conjunction, to discover physically close by peers over the network. However, as
-discussed in Section \ref{sec:landmark_binning}, these methods do not always
-give reliable estimates for the node positions over the internet. The landmark
-servers are not self-organizing and have maintenance overheads. To serve a
-P2P network with millions of peers, multiple landmark servers
-distributed over the whole world is required, which is hard to manage. The RTT
-measurements also can measure the delay between peers, but it is a greedy method,
-which can result non-optimal overlay topologies especially if close by nodes
-have low bandwidth connections among themselves.
-
-Proximity routing does not discover or store proximity information for the
-peers, however, during package forwarding, nodes that have lower latency to the
-destination, or with a closer key to the destination is selected. As no
-proximity information is used, the system is based on a greedy approach and it
-usually selects low latency paths over the overlay, which maps to suboptimal
-longer paths on the physical topology. For these obvious reasons, proximity
-neighbour selection is generally considered as a superior method to proximity
-routing, however, joint uses of these two protocols are also possible.
-
-Proximity neighbour selection approach is similar to proximity routing, with an
-exception that during forwarding, the proximity information of the peers are
-also considered. Therefore, for each application, proximity information is
-extracted and used during routing. Depending on the application, the proximity
-information can be the node sharing the same prefix with the current node, or
-TTL to detect close by neighbours. Tapstry and Pastry are two popular methods
-implementing the proximity neighbour selection among others mentioned above in the
-proximity neighbour selection section.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{The Algorithms}
-%Having understood the nature of the problem that peer-to-peer architecture face, this section starts the discussion of the academic work that has been conducted in the field the last few years.
-What follows this introduction, does not claim to be a thorough citation of all known protocols that are available out there, but instead, a carefull selection of those that left a distinctive fingerprint contribution in the efforts of the research community to alleviate the topology mismatch problem.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% UNSTRUCTURED
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%\subsection{Narada}
-%
-%\paragraph*{}
-%\emph{End System Multicast} \cite{chu_esm_2000,chu_esm_2001,chu_esm_2002} as well as other  proposes an architecture, where end-systems implement all multicast related functionality (including membership management and packet replication).
-In \emph{Narada}, the protocol developed in the paper's context, end-systems
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item self-organize, in a fully distributed manner,
-%  \item into an efficient overlay structure, both from the network and the application perspective,
-%  \item with self improving properties,
-%  \item adapting to network dynamics\footnote{Long-term variations in Internet path characteristics (i.e. bandwidth, latency e.t.c.)}.
-%\end{inparaenum}
-%
-%\paragraph*{}
-%What Narada really does is to construct overlay spanning trees in a two-step process:
-%\begin{enumerate}
-%  \item It builds a richer connected graph \footnote{ENarada \cite{li_enarada_2008} used Gossip protocol for the construction}, termed a \emph{mesh}, while trying to ensure a couple of desired performance properties:
-%    \begin{itemize}
-%      \item path quality between any pair of members should be comparable to the quality of the unicast path between them, and
-%      \item each member should have a limited number of neighbours on the mesh.
-%    \end{itemize}
-%
-%  \item It computes minimum spanning trees out of the mesh, each rooted at the corresponding source member.
-%\end{enumerate}
-%
-%\paragraph*{}
-%Narada incrementally improves the quality of the mesh by adding and dropping overlay links.
-Members probe each other at random and new links are added depending on what is the gain in doing so.
-On the other hand, participants, continuously  monitor the utility of existing links, and drop them if they are found to be not usefull.
-%
-%\paragraph*{}
-%Narada protocol achieves relatively good performance for small and medium sized groups involving tens to hundreds of members.
-This is not, by no means, what a contemporary peer-to-peer system needs to be characterized as scalable, but this work is considered pioneering\footnote{Another similar approach is \emph{Scattercast}\cite{chawathe_scattercast_2000}}, in that it was the first to have conducted a detailed Internet evaluation to analyse and demonstrate the feasibility of an overlay-based, application layer, multicast architecture that can dynamically adapt to bandwidth and latency properties of the physical underlying infrastructure.
-%
-%\subsection{Gia}
-%
-%\paragraph*{}
-%\emph{Gia} \cite{chawathe_gia_2003} is a decentralized unstructured peer-to-peer file sharing system with the following characteristics:
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item replaces Gnutella's flooding with random walks \cite{lv_randomwalks_2002},
-%  \item recognizes the importance of overlay network's topology, while using the random walks, and exploits it by incorporating a topology adaptation algorithm,
-%  \item introduces a token-based flow control mechanism, and
-%  \item detects the significant heterogeneity in peer bandwidth, processing power, disk speed e.t.c. and takes advantage of it.
-%\end{inparaenum}
-%
-%\paragraph*{}
-%More specifically, there are four key components in the design of \emph{Gia} which are summarized bellow:
-%\begin{enumerate}
-%  \item A \emph{dynamic topology adaptation} protocol that puts participating nodes within short reach of high capacity nodes so that these \emph{high-degree} nodes, which due to their high connectivity receive most of the queries, actually have the capacity to handle them.
-%  \item An \emph{active flow control} scheme is used to avoid overloaded hot-spots. Heterogeneity is detected and flow control tokens are given to nodes based on the available capacity.
-%  \item Every node maintains pointers to to the content that is offered by their immediate neighbours, creating a \emph{one-hop replication} of pointers, scheme.
-%  \item A \emph{search protocol} based on random walks, that is biased towards directing queries to high-capacity nodes that are typically best able to answer these queries.
-%\end{enumerate}
-%
-%\subsection{Adaptive Overlay Topology Optimization}
-%
-%\paragraph*{}
-%\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{liu_auto_2003} is an algorithm for building an overlay multicast tree among each source node and its direct logical neighbours in order to alleviate the mismatching problem while providing a larger query coverage range.
-It includes the steps of
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item \emph{Selective Flooding (SF)}, and
-%  \item \emph{Active Topology (AT)}
-%\end{inparaenum}
-%which are furtherlly discussed in the following paragraphs.
-%
-%\paragraph*{Selective Flooding (SF)}
-%Instead of flooding to all neighbours, SF builds an overlay multicast tree among each peer and its immediate logical neighbours.
-This reduces overall message burden to the network by making some neighbours non-flooding.
-The tree is formed using a minimum spanning tree algorithm.
-For this reason a peer needs to know the costs to all its logical neighbours (e.g.
-the network delay) as well as between any pair of neighbours.
-Additional, cost probing, messages need to be added to the original messages defined by the Gnutella protocol.
-%
-%LTM's SF effectiveness has be proven to be detached from the different physical or overlay topologies.
-On the other hand, SF is more effective with large number of logical neighbours.
-It can reach an average optimization rate of 87.4 percent on a logical topology with an average of 30 logical neighbours.
-%
-%\paragraph*{Active Topology (AT)}
-%At this step the overlay topology is reorganized.
-Each peer, independently, makes optimizations in the overlay network to alleviate topology mismatching by replacing non-flooding neighbours, with closer nodes.
-The approach incorporated by the paper is called \emph{Randomized AT} algorithm which picks up a candidate peer at random among the non-flooding neighbour's neighbours.
-Whenever a new neighbour cost table is received or there is a change of neighbours, the source peer has to re-calculate the multicast tree and apply the randomized AT algorithm.
-%
-%Different numbers of average logical neighbours has little to do with the effectiveness of AT.
-If the source has $n$ non-flooding peers, there are $n$ potential neighbour replacements.
-The overhead to exhaust all $n$ possible replacements can be high, so in practice, after each replacement the source peer can decide whether it needs to find another candidate peer.
-This is done by computing the cost improvement ratio greater than some predefined termination threshold.
-The larger the threshold, the slower, in the number of optimization steps, the reduction of the normalized average distance.
-As a whole the average response time is significantly reduced when more optimization steps taken.
-%
-%\subsection{Location-aware Topology Matching}
-%
-%\paragraph*{}
-%\emph{Location-aware topology matching} or \emph{LTM} \cite{liu_ltm_2004}, is an algorithm of building an efficient overlay by disconnecting low productive connections and choosing physically closer nodes as logical neighbours while still retaining the search scope and reducing response time of queries.
-This is done by issuing detector messages by peers, to a small region so that receivers of the detector can record relative delay information and cut inefficient or redundant logical links as well as add closer nodes as direct neighbours.
-Three operations are defined in LTMband are furtherly discussed in the following paragraphs.
-%
-%\paragraph*{TTL2-detector flooding}
-%\emph{TTL2-detector} messages are designed based on the specification of the Gnutella protocol.
-Suppose that $d(i, S, v)$ denotes the TTL2-detector which has the message ID of $i$ with TTL value of $v$ and is initiated by $S$.
-In addition to Gnutella's unified 23-byte header, such a message has a body of two types:
-%\begin{itemize}
-%  \item short format $d(i, S, 1)$ if $S$ is a direct neighbour of the receiver
-%  \item long format $d(i, S, 0)$ if $S$ is within a two-hop distance from the receiver
-%\end{itemize}
-%This way a peer can calculate link costs from within a one- or two-hop area around it and thus reach peers in $N(S)$ and $N^2(S)$ sets, respectively, in the paper's context.
-Each peer floods a TTL-detector periodically while it calculates link cost to another peer by receiving other peer's detector messages and checking:
-%\begin{itemize}
-%  \item the Source Timestamp field and its own clock upon time of receipt for one-hop messages
-%  \item the Source Timestamp, TTL1 Timestamp and its own clock upon receipt for two-hop messages
-%\end{itemize}
-%Of course peer clocks must be synchronized in order to do such checking.
-%
-%\paragraph*{Low productive connection cutting}
-%There are three cases for any peer $P$ who receives $d(i, S, v)$ multiple times:
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item $P$ receives both $d(i, S, 1)$ and $d(i, S, 0)$
-%  \item $P$ receives multiple $d(i, S, 0)$s from different paths, and P randomly chooses to process one
-%  \item $P$ receives one $d(i, S, 1)$ and multiple $d(i, S, 0)$s, and $P$ processes $d(i, S, 1)$ and one randomly selected $d(i, S, 0)$
-%\end{inparaenum}
-%If the link with the largest cost is found and is a direct neighbour then the connection is put in a will-cut list and stays there for a certain period of time.
-If it is not, then it is handled by other peers.
-After that period, connections are cut and recorded to $P$'s cut-list.
-%
-%\paragraph*{Source probing}
-%For a peer $P\in(N^2(S) - N(S))$ who receives only one $d(i, S, 0)$, the cost of $PS$ is obtained (with list look-up or probing).
-Then $P$ compares it with the cost from each hop and if $PS$ has the largest cost, $P$ will not keep this connection, while otherwise the connection will be created.
-%
-%\paragraph{}
-%Supposing $n$ is the number of peers, $c_n$ is the average number of neighbours and $c_e$ is the average cost of logical links, then in the flooding-based search the traffic incurred by one query from an arbitrary peer in a peer-to-peer network is $O(n)$.
-As observed in the Gnutella network \cite{sripanidkulchai_gnutella_2001}, each peer issues $0.3$ queries per minute in average, thus the per minute traffic incurred by the network with $n$ peers is $O(n^2)$.
-Because each $d(i, S, v)$ has a TTL of $2$ in each source peer, the traffic for one time LTM optimization in all peers is at most $2nc_n^2c_e$.
-If each peer uses LTM $k$ times per minute, the total traffic incurred is $2knc_n^2c_e$.
-Simulation shows the best value for $k$ is $2$ or $3$.
-So, the traffic overhead caused by LTM to the network is $O(n)$.
-%
-%TTLj-detectors, with $j > 2$, would detect and break cycles with more than 4 links.
-LTM though, does not use such detectors because detector-flood traffic would increase significantly, and cut links between two end-peers, could cause queries initiated by them to traverse a path much more expensive than the cost on the the cut link.
-%
-%\paragraph{}
-%LTM disadvantages are
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item disagreement of measured delay due to unsynchronized clocks causes problems when deciding the cut positions, which can influence the network connectivity, and
-%  \item the network delay metric mainly focuses on disabling the connections between peers physically far away without considering the shortcuts created by powerful peers.
-%\end{inparaenum}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/ltm_multid.jpeg}
-%\caption{Peer $P$ receives $d(i, S, v)$ multiple times}
-%\label{figure:ltm_multid}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/ltm_example.jpeg}
-%\caption{A full example of LTM}
-%\label{figure:ltm_example}
-%\end{figure}
-%
-%\subsection{Scalable Bipartite Overlay}
-%\emph{Scalable Bipartite Overlay (SBO)} \cite{liu_bipartite_2007} employs an efficient strategy to select query forwarding paths and logical neighbours.
-The topology construction and optimization of SBO consist of four phases:
-%\begin{enumerate}[\itshape i\upshape)]
-%  \item When a new peer is joining the network, it will randomly take an initial colour; say red or white.
-This colour is not to be changed for the time the peer stays connected.
-If it leaves and then rejoins it will, again, have to pick a colour, randomly.
-Thus all peers are separated into two groups, red or white.
-Then the bootstrap host will provide, the joining peer, with a list of active peers, including their colour information, in order for the later to establish connections to different colour peers.
-This way, all peers form a bipartite overlay.
-%  \item White peers probe distances with their immediate (red) neighbours, form a cost table and send this table to their corresponding red neighbours.
-%  \item Each red peer builds a minimum spanning tree using the obtained neighbour cost tables.
-These are two-hop diameter trees so white peers do not need to build one.
-The links that are part of a minimum spanning tree are called \emph{forwarding connections (FC)} while the rest \emph{non-forwarding (NFC)}.
-%  \item Having a minimum spanning tree with two hops a red peer is able to send its queries within that range.
-Some white peers, though, have become non-forwarding neighbours.
-In this phase, such a (white) neighbour will try to find another red peer being two hops away from its current red neighbour to replace the later as its new neighbour.
-%\end{enumerate}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/sbo_efficient_forward.jpeg}
-%\caption{A red peer $P$ has a small overlay topology of $N(P)$ and $N^2(P)$ and computes the efficient forwarding paths}
-%\label{figure:sbo_efficient_forward}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/sbo_neighbour_replace.jpeg}
-%\caption{Neighbor replacement}
-%\label{figure:sbo_neighbour_replace}
-%\end{figure}
-%
-%In a static enviromnent LTM may reduce traffic cost by around 80 to 85 percent while SBO reduces traffic cost between 85 and 90 percent.
-However, LTM  is proved to converge in around 2-3 steps while SBO needs 4-5 steps.
-Moreover LTM reduces response time by more than 60 percent in 3 steps while SBO needs 8.
-In a dynamic environment (10 minute average peer lifetime, 0.3 queries/sec by each peer) SBO and LTM reduce the average traffic cost per query (including the overhead due to the optimization steps) by 85 and 80 percent, respectively.
-Moreover LTM reduces the response time per query to 30 percent while SBO to 35 percent.
-%
-%\subsection{Adaptive Connection Establishment}
-%
-%\paragraph{}
-%\emph{Adaptive Connection Establishment (ACE)} \cite{liu_ace_2004} builds an overlay multicast tree among each source node and the peers within a certain diameter from the source peer and optimizes the neighbour connections that are not in that tree.
-%
-%\paragraph{}
-%ACE indicates three phases for its algorithm:
-%\begin{enumerate}[\itshape i\upshape)]
-%  \item Calculate cost between nodes using network delay as a metric.
-Each peer probes the costs with its immediate logical neighbours and forms a \emph{neighbour cost table (NCT)} using a special routing message type.
-Two neighbouring peers exchange their NCTs in order for every peer to obtain the cost between any pair of its local neighbours forming a small overlay topology.
-%  \item Based on obtained NCTs a minimum spanning tree among each peer and its immediate neighbours is built (Figure~\ref{figure:ace_phase2}).
-%  \item Physically far away neighbours are replaced by physically close neighbours.
-In ACE a peer, say $S$, probes the distance between one of its non-flooding neighbour's neighbour, say $G$ and $H$ respectively.
-If the link to neighbour's neighbour is smaller than that to the neighbour, the later is cut.
-If this is not the case but $S$ finds that cost of $GH$ is even larger than that of the $SH$, $S$ will keep $H$ as a new neighbour.
-Obviously, if $SH$ is larger than $SG$ and $GH$, the connection will not be established and $S$ will continue probing another neighbour's neighbour.
-The algorithm described is conducted within $1$-neighbour closure (among its source peer and all its direct neighbours) but the optimization scope can be enlarged.
-The larger the scope, the better topology matching improvement but also the greater the computational overhead (Figure~\ref{figure:ace_phase3}).
-%\end{enumerate}
-%
-%\paragraph{}
-%Simulations in \cite{liu_acesims_2004} show that the average scope of each query to cover the same scope of nodes is reduced by about 65 percent without losing any autonomy feature, while the average response time can be reduced by 35 percent.
-Larger diameter topologies lead to better topology optimization rate but also to higher communication and computation overhead.
-It was also found that it is more effective in higher connectivity dense topologies.
-Compared to LTM, it comes short of convergence speed.
-In \cite{ni_mismatch_2004} shows reduction of both total traffic (90 percent) and response time (80 percent) to message queries without shrinking the search scope.
-SBO, on the other hand, achieves approximately 85 percent reduction on traffic cost and about 60 percent reduction on query response time.
-Last but not least, it is concluded that work must be done on incorporating a more sophisticated selection policy for candidate non-flooding peers.
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/ace_phase2.jpeg}
-%\caption{Second phase in ACE}
-%\label{figure:ace_phase2}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/ace_phase3.jpeg}
-%\caption{Third phase in ACE}
-%\label{figure:ace_phase3}
-%\end{figure}
-%
-%\subsection{Hops Adaptive Neighbour Discovery}
-%
-%\paragraph{}
-%Paper \cite{chen_hand_2006} proposes a new algorithm named \emph{Hops Adaptive Neighbour Discovery (HAND)} which uses a fully distributed triple hop adjustment strategy to address the topology mismatch problem.
-The advantages of the algorithm compared to other approaches are that
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item it does not need any clock synchronization,
-%  \item it is a fully distributed algorithm making it robust and reliable in decentralized systems,
-%  \item the traffic overhead of the triple hop adjustment is very low,
-%  \item it is applicable to dynamic peer-to-peer environments, and
-%  \item maintains lower query response time.
-%\end{inparaenum}
-%
-%\paragraph{}
-%The algorithm's ultimate goal is the optimal overlay, the \emph{Logical Communication Network (LCN)} in the paper's context.
-The key concept of the algorithm is that a graph $G^{*}$ that describes an LCN and a graph $G$ that describes the current overlay are matched only if all peer hop sequences $(v_1, v_2, \ldots, v_k)$ in $G$ exist in $G^{*}$ and in the same order.
-In practice triple sequences $(v_1, v_2, v_3)$ are used.
-%
-%\paragraph{}
-%The mismatching detection is done in the following way.
-Suppose we want to verify peer sequence $v_2-v_1-v_3$ (see Figure~\ref{figure:hand_com_overlay}).
-A pair of probing messages are sent from $v_1$ to $v_2$ and $v_3$.
-Suppose delays of $(v_1,v_2)$ and $(v_1,v_3)$ are are denoted as $x$ and $z$, respectively.
-When the probing message arrives to $v_2$ it forwards it directly to $v_3$.
-Similarly, when the probing message arrives to $v_3$, it forwards it directly to $v_2$.
-These last steps are performed in order to obtain delays of $(v_2,v_3)$ and $(v_3,v_2)$ physical paths, respectively, denoted by $y$.
-%\begin{itemize}
-%  \item If $y=z-x\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched and should be adjusted to $v_1-v_2-v_3$ by deleting edge $(v_1,v_3)$ and adding a new $(v_2,v_3)$ (see Figure~\ref{figure:hand_matchtriple_lemma3}).
-%  \item if $y=x-z\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched and should be adjusted to $v_1-v_3-v_2$ by deleting edge $(v_1,v_2)$ and adding a new $(v_3,v_2)$ (see Figure~\ref{figure:hand_matchtriple_lemma4}).
-%\end{itemize}
-%$\varepsilon$ is a small positive real number denoting additional delays caused by possible forwarding and jitter delays.
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/hand_com_overlay.jpeg}
-%\caption{The communication in overlay}
-%\label{figure:hand_com_overlay}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/hand_matchtriple_lemma3.jpeg}
-%\caption{The matching triple in paper's Lemma 3}
-%\label{figure:hand_matchtriple_lemma3}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/hand_matchtriple_lemma4.jpeg}
-%\caption{The matching triple in paper's Lemma 4}
-%\label{figure:hand_matchtriple_lemma4}
-%\end{figure}
-%
-%\paragraph{}
-%Measurements conducted for evaluation perposes showed that in a static environment the algorithm can effectively decrease traffic cost by about 77 percent and shorten the query response time by about 49 percent in less than two minutes.
-In a dynamic environment it shows similar behaviour and with the size of the overlay network having little impact on the effectiveness of the algorithm.
-Compared to LTM both algorithms have almost the same traffic reduction rate, however on the response time reduction rate HAND has a higher one by about 4 percent.
-The traffic overhead of HAND is much less than that of LTM by an average of 55 percent.
-%
-%\subsection{Distributed Cycle Minimization Protocol}
-%
-%\paragraph{}
-%\emph{Distributed Cycle Minimization Protocol (DCMP)} which is introduced in \cite{zhu_dcmp_2008} is a dynamic, fully decentralized protocol that promises significant reduction of duplicate messages.
-To achieve that, it uses the slightly different approach of eliminating unnecessary cycles, while retaining the connectivity of the network and preserving fault resilience and load balancing properties of unstructured peer-to-peer schemas by avoiding the creation of a hierarchical organization.
-%
-%DCMP aims at cutting the cycle paths at strategic locations. Any peer that detects a duplicate message can initiate the cutting process which consists of the following steps:
-%\begin{enumerate}
-%  \item Peers in the cycle elect a leader called \emph{GatePeer}\footnote{GatePeers are important for maintaining the connectivity and optimal structure of the network while peers join and leave randomly.}.
-%  \item The cycle is cut at a well-defind point with respect to the GatePeer
-%\end{enumerate}
-%
-%\paragraph{}
-%The first step after detecting a duplicate message by some peer is to gather information from all peers in the cycle using a new type of control message called \emph{Information Collecting Message} or \emph{ICM}.
-ICM contains:
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item a \emph{GUID}\footnote{Globally Unique IDentifier assigned to every query message generated by any node.} field same as the one of the duplicate message,
-%  \item \emph{DetectionID} field which represents the direction of the connection where the duplicate was identified\footnote{This ensures the uniqueness of the ICM messages because as it travels through many cyclic paths, multiple peers will detect the duplicates and initiate an ICM message.}, and
-%  \item \emph{Node Information Vector (NIV)} which contains information (bandwidth, CPU power, etc) about peers that propagated the ICM.
-%\end{inparaenum}
-%
-%\paragraph{}
-%Suppose $A$ detected the duplicate as depicted in Figure~\ref{figure:dcmp}.
-It then emits an ICM to $B$ and $F$, that initially contains information only about itself.
-Each peer that receives the ICM, appends its information and propagates it along the reverse path of the original message.
-Since two copies of ICM are sent, at some point, a peer, say peer $D$, will receive a duplicate ICM.
-Using the information in the NIVs of the ICMs, $D$, decides to cut (for example) the EF connection.
-To inform the other peers about its decision, it emits a \emph{Cut Message (CM)} which contains the GUID and DetectionID of the corresponding ICM and an additional field that identifies the connection to be cut.
-$D$, then, forwards the CM in the reverse directions from where the ICM arrived.
-Similarly CMs received by any peer are propagated toward the reverse path of the corresponding ICM.
-Eventually, either peer $E$ or peer $F$ will receive the CM and cut the connection, thus eliminating the cycle.
-%
-%Receiving a duplicate ICM denotes the existence of a cycle.
-The opposite is not true though.
-For example, if the cycle contains $2 \times TTL$ edges, it will not be detected, because the ICM messages will be discarded before they locate it.
-There is a tradeoff between preserving the connectivity of the network and minimizing the duplicates that makes such a possibility to be safely ignored.
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/dcmp.jpeg}
-%\caption{Cycle elimination methods in DCMP}
-%\label{figure:dcmp}
-%\end{figure}
-%
-%\paragraph{}
-%Experiments in \cite{zhu_dcmp_2008} showed that DCMP incurs a lower delay, returns more results and decreases the number of duplicate messages by 22\%, compared to LTM.
-Additionally DCMP has one to two orders of magnitude less overhead, because it adopts a more ``LAZY'' approach than broadcasting control messages periodically like LTM does.
-%
-%\subsection{Two-Hop-Away Neighbour Comparison and Selection}
-%
-%\paragraph{}
-%Work in \cite{liu_thancs_2005,liu_thancs_2008} proposes a distributed heuristic called \emph{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} that
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item is completely distributed and needs no global knowledge,
-%  \item presents trivial overhead compared to the query cost savings
-%  \item its convergent speed of the algorithm is fast enough (faster than minimum spanning tree approaches) so that is effective to dynamic environments, and
-%  \item does not shrink the search scope.
-%\end{inparaenum}
-%
-%\paragraph{}
-%THANCS is considered a \emph{local search method}, in the sense that it targets in finding a locally optimum solution, by exploiting knowledge within a 2-hop radius.
-The algorithm consists of two main components: \emph{piggybacking neighbour distance on queries} and \emph{neighbour comparison and selection} which are furtherly discussed bellow.
-%
-%\paragraph{Piggybacking neighbour distance on queries}
-%Using network delay as a metric for measuring the distance, each peer probes distances with its immediate naighbours and stores information locally.
-For this reason a special query message type, \emph{Piggy Message (PM)}, is introduced.
-It is 6 bytes long and includes two fields: Neighbour IP Address and Neighbour Distance.
-A peer $P$ constructs a PM for its neighbour $Q$, which contains $Q$'s IP address and $Q$'s distance from $P$.
-When $P$ receives a query from $Q$, this PM will be piggybacked by the query that goes to all other neighbours of peer $P$.
-Upon receiving such a query message, each of the other neighbours will detach the PM, record the $PQ$ distance and process the query.
-This PM will not be further forwarded.
-In selecting which incoming queries  should piggyback a PM the paper proposes the \emph{pure propability-based (PPB)} and the \emph{new neighbour triggered (NNT)} policies.
-%
-%\paragraph{Neighbour comparison and selection}
-%Figure~\ref{figure:thancs} illustrates this component of the THANCS algorithm.
-A peer $S$ probes the distance to all known unprobed $N^2(S)$\footnote{$N^2(S)$ denotes the set of peers being two hops away from $S$, while $N(S)$ denotes the set of direct logical neighbours of $S$.}.
-The distance of $SP$ is known to $S$.
-Upon receiving a PM from node $P$ with the distance of $PQ$, $S$ follows one of the following:
-%\begin{itemize}
-%  \item $Q \in \left( N(S) \cap N^2(S) \right)$, i.e.
-$Q$ is direct neighbour of $S$.
-In this case $S$ will compare cost of $SQ$, $SP$ and $PQ$.
-If the most costly connection is one of $SQ$ or $SP$ the corresponding link will be put into a \emph{will-cut list}\footnote{Links are not immediately disconnected when put in the will-cut list.
-There are useless for forwarding but are kept active, for some time, in order to serve query responses that are traveling to the source peer along the inverse search path.}.
-If the most costly connection is $PQ$ then $S$ will do nothing, as the fully distributed nature of the algorithm will give a chance to either $P$ or $Q$ to cut the connection.
-%  \item $Q \in \left( N^2(S) - N(S) \right)$, i.e.
-$Q$ is a two-hop-away neighbour of $S$.
-If $S$ hasn't probed $Q$ before\footnote{A distance cache is maintained and looked up in such a case.}, it probes and stores the result in the \emph{distance cache}.
-Having the distance of $SQ$, $S$ compares costs of $SQ$, $SP$ and $PQ$.
-If $SQ$ is the most costly, $S$ will not establish the connection.
-If $SP$ is the most costly, $S$ will establish connection $SQ$ and put $SP$ in the will-cut list.
-If $PQ$ is the longest, $S$ will keep  the connection with both $P$ and $Q$, expecting that $P$ or $Q$ will eventually disconnect link $PQ$, later.
-%\end{itemize}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/thancs.jpeg}
-%\caption{Probing two-hop-away neighbours}
-%\label{figure:thancs}
-%\end{figure}
-%
-%\paragraph{}
-%In a static environment THANCS has been proven to be effective; optimizing 45 percent out of the 60 percent of mismatched paths, constructing a nearly optimal overlay.
-This leads to a 60 percent reduction in traffic cost as well as a 40 percent decrease in query response time.
-In dynamic environments (Gnutella 0.6/Limewire super-peer-like and Ion flat-like), THANCS saves up to 70 percent of the traffic cost in the super-peer topology and 55 percent for the flat one.
-Average response time is also decreased by 60 and 45 percent, respectively.
-Generally, THANCS has similar performance to LTM, without needing synchronization.
-SBO, incurring half the  overhead of AOTO, reduces the traffic cost the most, while THANCS has lower response time and converges faster than SBO.
-THANCS is, thus, more suitable for a more dynamic environment.
-In addition, THANCS is easy to implement and its operation overhead is trivial, compared with the other three approaches.
-This design, however, has the limitation of not being easily extend to also support non-flooding-based systems.
-%
-%\subsection{mOverlay}
-%Zhang et al.
-in \cite{zhang_moverlay_2004} focus on two aspects of the overlay construction.
-First, to achieve \emph{efficiency} of communication, a protocol should minimize unecessary long-distance hops and redundant traffic.
-Second, to be \emph{scalable} the overlay should be constructed in a distributed fashion and maintenance cost (data and locality management) should be minimized.
-Both of the above observations, are positively affected by the exploitation of node proximity at the underlying network.
-To accomplish it, the paper introduces the \emph{group} concept, according to which non-static (dynamic) landmarks are used to compute proximity, resulting in a highly robust and scalable overlay construction with reduced maintenance cost as well.
-The dynamic nature of landmarks help towards load-balance, meaning that hot-spots are avoided.
-%
-%\paragraph{}
-%The proposed architecture, called \emph{mOverlay}, is a \emph{clustering approach} which creates a two-level hierarchical network, where on the top level we have connections between groups while on the bottom we have connections between hosts inside groups, thus attempting to recreate \emph{Small-World}-like properties for the constructed overlay network.
-The notion of a \emph{group}, in the paper's context, is a set of hosts that are close to each other with respect to any position $P$ in the underlying network.
-The distance between hosts can be
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item network latency,
-%  \item round-trip time,
-%  \item minimum bandwidth on the links along the path connecting the two nodes, or
-%  \item some other user-defined cost metric between the two nodes.
-%\end{inparaenum}
-%These nodes are refered to as neighbouring. Similarly groups can exchange messages with their neighbouring ones\footnote{Groups nearby, in the underlying physical network.}.
-%
-%According to the \emph{grouping criterion} set by the authors, a new host $Q$ belongs to some group $A$ if the distance between $Q$ and group $A$'s neighboring groups is the same as the distance between group $A$ and it's neighbouring groups.
-The neighbour groups thus, play the role of the dynamic landmarks and for accuracy and performance reasons the nearest of them are chosen to play that role.
-%
-%\paragraph{Locating process} A new coming host, $Q$, first connects to a globally known host cache called the \emph{rendezvous point (RP)} in order to retrieve the starting point in the overlay, say $A$ in group $1$.
-Host $Q$ then, measures its distance to host $A$.
-At the same time, the later, sends information about the neighbour groups of group $1$ back to host $Q$.
-This list is called \emph{candidate group list}, and the newcoming host sequnentially measures its distance to each of them in seek for the closest one.
-If the \emph{grouping criterion} is met, host $Q$ belongs to group $1$.
-If not, a boot host from the closest group is found and the algorithm is re-run until the criterion is met or after a predefined number of repetitions.
-In the later case, $Q$ creates a new group comprising itself only.
-The above protocol does not favour hotspots as it spreads the probability of visiting a group across the whole overlay and limits the overhead in the level of $O \left ( log N \right )$.
-%
-%\paragraph{General overlay operations} A set of additional protocols, are also introduced, similar to those found in traditional unstructured networks, but modified focusing on scalability and robustness.
-For example a protocol for \emph{group formation} is introduced that exploits the inherent characteristic of proximity, in the overlay, in order to efficiently detect the neighbouring groups of a newly formated group from the set of adjasent groups of its closest neighbour.
-Additionaly, during \emph{group joining} the coresponding protocol denotes the exchange of important information for group maintenance.
-This can be furtherly improved by \emph{information sharing} between nodes of the same group, functionality handled by a dedicated flood-like protocol\footnote{Since nodes that belong to the same group are physically close this can be achieved at a minimum price.}.
-Moreover, another set of distributed protocols handle the \emph{information update}.
-The information that needs update, in the proposed architecture, is
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item the host cache, when a new node joins, and
-%  \item the neighbours of groups, when a close-by group is generated.
-%\end{inparaenum}
-%Finally, in case of \emph{host failure} or \emph{host departure} the system is able to maintain its stability since there are defined operations for periodical host cache update and group leader selection if one leaves or dies.
-%
-%\subsection{Distributed Domain Name Order}
-%{\sethlcolor{yellow}\hl{HA:  In paper/introduction nice advantages of unstructured networks:
-%
-%Unstructured P2P networks o?er a number of
-%important advantages: (i) An unstructured network
-%imposes very small demands on individual nodes,
-%and more speci?cally it allows nodes to join or leave
-%the network without signi?cantly a?ecting the sys-
-%tem performance. (ii) Unstructured networks are
-%appropriate for content-based retrieval (e.g., key-
-%word searches) as opposed to object identi?er loca-
-%tion of structured overlays. (iii) Finally unstructured
-%networks can easily accommodate nodes of vary-
-%ing power. Consequently, they scale to very large
-%sizes and they o?er more robust performance in
-%the presence of node failures and connection
-%unreliability.
-%}}
-%
-%\paragraph{}
-%\cite{zeinalipour-yazti_ddno_2005} proposes the \emph{Distributed Domain Name Order (DDNO)} technique which makes unstructured overlay networks, topologically aware.
-It results in a flat overlay topology but with some changes it can be utilized in \emph{superpeer} environments.
-In DDNO, a node of degree $d$, tries to connect to $\frac{d}{2}$ nodes that belong to the same domain (\emph{sibling} connections) and another $\frac{d}{2}$ of random nodes (\emph{random} connections).
-Connecting to \emph{sibling} nodes ensures the reduction of long distance message travelling and thus resulting to better performance, while \emph{random} nodes keep the structure connected.
-%
-%\paragraph{Discovering \emph{random} neighbours}
-%Initially, when a new coming node wants to join a network $N$, uses a \emph{hostcache} mechanism to provide it with a list of nodes with which to establish its $\frac{d}{2}$ \emph{random} connections.
-%
-%\paragraph{Discovering \emph{sibling} neighbours}
-%In order to discover the rest $\frac{d}{2}$, \emph{sibling}, nodes, the \emph{lookupDN} procedure is initiated.
-In this phase, a special $l$ walker message is multicasted, by the newcomer, in search of a node that knows\footnote{A special structure is used called ZoneCache that contains information about which nodes are reachable in an $r$-hop radius.} how to guide $n$ to a \emph{sibling} node.
-%
-%\paragraph{}
-%Figure~\ref{figure:ddno_lookupdn} shows an $l$ message emitted by node $n$ which travels along the path [$a$, $b$, $c$, $e$, $b$, $d$].
-At node $d$, $l$ finds information to make a decision on which neighbour to follow next\footnote{Contrary to the random choices, made at the previous hops.}.
-Ultimately $l$ arrives at $m$, which is a \emph{sibling} node to $n$.
-$m$ then issues a broadcast message to all its own \emph{siblings}.
-Each of the receiving nodes, $m$ included, will respond to $n$ with a designated $l^{'}$ message if it is willing to accept new connections and out of these answers, node $n$, will attempt to establish its, remaining, $\frac{d}{2}$ \emph{sibling} connections.
-%
-%\paragraph{}
-%Unfortunately, use of DDNO results in networks which have a uniform distribution of node degrees, losing the valuable properties of Power Law and Small World networks, as mentioned previously in this survey.
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.375]{img/ddno_lookupdn.jpeg}
-%\caption{Domain name look-up in a \emph{DDNO} topology}
-%\label{figure:ddno_lookupdn}
-%\end{figure}
-%
-%%\subsection{Critical Topology-Aware Grouping}
-%%
-%%\paragraph{}
-%%\emph{Critical Topology-Aware Grouping (CTAG)} presented in \cite{zhao_ctag_2006}, is a grouping algorithm that tries to exploit low-cost and low-delay communication of physically close-by peers.
-The grouping strategy is based on the \emph{IANA}\footnote{Internet Assigned Numbers Authority} and the respective \emph{Regionanal Internet Registry (RIR)}'s IP assignment strategies, according to which nodes within the same organization are always addressed from the same block.
-The paper proposes the \emph{Adjacency Measurement (AM)} technique which uses the longest matching IP segment criterion to calculate node proximity.
-%%
-%%Observations made in various studies (e.g.
-\cite{matei_mapgnutella_2002}) concerning the distribution of nodes among \emph{Internet Service Providers (ISPs)} and \emph{Autonomous Systems (ASs)} have shown that only $2$ to $5$ percent of Gnutella connections link peers within a single AS while more than $40$ percent of all Gnutella peers are located within the top 10 ASs.
-Similarly, measurements in \cite{zeinalipour-yazti_gnudc_2002} used a $244,000$ IPs test-bed and results have shown that $45$ percent of the nodes belonged to only $10$ large ISPs and $58$ percent belong to only $20$.
-Such results mean that most overlay generated traffic crosses AS borders increasing topology mismatch cost.
-%%
-%%\emph{CTAG} focuses on both the construction of the overlay as well as dynamically revising it during node interaction, phases called \emph{bootstrapping grouping} and \emph{dynamic revision}, respectively in the paper's context.
-%%
-%%\paragraph{Bootstrapping grouping}
-%%The \emph{Gnutella Web Caching} mechanism has been modified in order for a new coming node to choose the closest \emph{GWC} in order to retrieve the node list for bootstrapping.
-%%
-%%\paragraph{Dynamic revision}
-%%Similarly to the bootstrapping phase, \emph{AM} metric is used to store hosts' addresses.
-read from \emph{X-Try} headers during handshake or from \emph{QueryHit} messages.
-Additionally, when a node reaches the max neighbour connections, node disconnects the neighbours with the lowest \emph{AM}.
-%
-%\subsection{Peer-exchange Routing Optimization Protocols}
-%
-%\paragraph{}
-%Work presented in \cite{qiu_prop_2007} introduces \emph{peer-exchange}\footnote{A series of exchanges of neighbours between two peers.
-One exchange can be viewed as a pair of cut-add operations.} as a basic operation to adaptively adjust the connections of the overlay network, and efficiently reduce the average logical link latency of the whole system.
-There are two points that differentiate this scheme from others:
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item it utilizes the collaboration of two peers to optimize their neighbourhood environment, than simply letting each node to ``selfishly'' choose its own strategy, and
-%  \item can be deployed effortlessly on both unstructured and structured peer-to-peer systems.
-%\end{inparaenum}
-%
-%\emph{PROP-G}\footnote{\emph{G} stands for \emph{generic}.} is the simple and direct variant of \emph{peer-exchange} in which two neighbours swap all their neighbours.
-As depicted in figure~\ref{figure:prop-g} the mechanism can be viewed as, nodes, exchanging overlay ``positions''.
-Thus, the overall overlay topology is not affected by the this operation.
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/prop-g.jpeg}
-%\caption{PROP-G in which a neighbours are exchanged.}
-%\label{figure:prop-g}
-%\end{figure}
-%
-%The alternative protocol proposed, called \emph{PROP-O}\footnote{\emph{O} stands for optimized}, ensures that the degree of each participating node remains the same by selectively choose the same number of neighbours for the exchange.
-With this mechanism, 2 nodes exchange the same number of neighbours in order not to change the degree
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/prop-o.jpeg}
-%\caption{PROP-O exchanging \emph{m} neighbours.}
-%\label{figure:prop-o}
-%\end{figure}
-%
-%\subsection{T2MC}
-%
-%\paragraph{}
-%\emph{T2MC}\cite{shi_t2mc_2008} exploits some properties of the Internet paradigm and clusters nodes belonging to the same ISP without any centralised control or predefined system parameterization.
-The algorithm considers the dynamic nature of peers and exploits the stable nature of routers in order to build a topology location relationship among end peers.
-%
-%\paragraph{}
-%Some special routers split the physical network into autonomous system domains.
-Using a Traceroute mechanism T2MC searches for latency leaps among the path to a host in order to form ``near'' and ``remote'' router clusters.
-This is achieved by a 2-Means Classification, defined by the following steps:
-%\begin{enumerate}
-% \item the peer chooses the minimum and maximum Latency results from the Traceroute for initializing the cntroinds of two sets ``first'' and ``second''.
-% \item The peer calculates, for all hops allong its tracerouted path, the absolute distance to the centroids of both sets and assigns the routers to that centroid with which it has the smaller absolute distance.
-% \item The peer calculates the latncy mean and variance value of two sets
-% \item If the variance is larger than a predefined threshold then the algorithm takes a loop from step 2 picking the two latency mean values as new centroids of sets ``first'' and ``second''.
-%\end{enumerate}
-%Ultimately peer will end up with two sets having the minimum intra-set variance.
-Finally the peer chooses the router from ``second'' set with the minimum hops attribute and sets it as a threshold.
-The selected router and all others whose hops attribute is larger than the threshold are classified as ``remote'' router cluster.
-The remaining are classified as ``near''.
-From the ``near'' class, the peer chooses the one with maximum hops attribute as its edge router, and registers it along with the all the ``near'' cluster into the DHT of the p2p overlay.
-As new peers join the network, those that share the same edge router or any of the members of the ``near'' router clustersm ther would gather to form a ``close'' peer cluster.
-As Edge routers can provide more valuable information than other members of the ``near'' set, T2MC was designed to prioritize interaction of peers and edge gateways.
-%
-%% TODO: figure t2mc.jpeg (a) stars represent edge routers (b) hop leaps
-%
-%\paragraph{}
-%The use of Traceroute as a tool for implementing the distance measuring infrastructure raise concearns about its efficiency and scalability.
-Being, primarilly, a network diagnostic utility, it is concearned too heavy weighted and intrucive for use in a larger scale\cite{ratnasamy_binning_2002}.
-Additionally, disabling ICMP is a common administrative policy for edge sites to enforce security, while dumping BGP routing tables\cite{krishnamurthy_bgpclust_2000} is not directly available to the application layer.
-%
-%\subsection{Unnamed Unstructured!!}
-%% TODO: to be reviewed
-%
-%\paragraph{}
-%In \cite{hsiao_redblue_2009}, Hsiao et al, claim to construct topology-aware unstructured overlays that \emph{guarantee} performance qualities in terms of
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item the expected communication latency among any two overlay peers regardless of the network size, and
-%  \item the broadcasting scope of each participating peer.
-%\end{inparaenum}
-%
-%The algorithm constructs an undirected graph $G = \left( V, E \right)$ comprised by two subgraphs.
-The first, namely $G^{\left( red \right)} = \left( V^{\left( red \right)}, E^{\left( red \right)} \right)$ in the paper's context, includes all vertices of $G$ and ensures the connectivity of the graph by securing at least one path between any two nodes.
-In contrast, $G^{\left( blue \right)} = \left( V^{\left( blue \right)}, E^{\left( blue \right)} \right)$, contains those vertices of $G$ that have free edges to link to other nodes and because these are fully utilized, the following also stands $E = E^{\left( red \right)} \cup E^{\left( blue \right)}$.
-%
-%A joining peer $u$, partitions its neighbours into two subsets, the $B_u^{\left( red \right)}$ and $B_u^{\left( blue \right)}$.
-In order to populate the $B_u^{\left( red \right)}$ subset, peer $u$ samples peers uniformly and at random.
-Then, each of these selected peers discovers a routing path starting from itself towards the node with the smallest (or the largest) ID in the system.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% STRUCTURED
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%\subsection{Landmark Binning}
-%The scheme proposed in \cite{ratnasamy_binning_2002} is based on the process of \emph{binning} close by nodes (in terms of network latency) into the same cluster.
-The authors set the following objectives for designing their proposed algorithm:
-%\begin{enumerate}
-% \item simplicity (minimal support from any measuring infrastructure),
-% \item scalablility (no global knowledge of the network),
-% \item complete distribution (nodes need no communication-cooperation).
-%\end{enumerate}
-%
-%\paragraph{Binning scheme}
-%The implementation requires a set of well-known \emph{landmark} machines spread accross the Internet.
-Every newly arriving node measures its distance from these landmarks and unilaterally decides to join a specific bin based on these results.
-In more detail the node measures its round-trip time to each of the landmarks and orders these measurements in a decreasing order.
-The ordering represents a ``bin'', in the sence of close-by nodes having the same landmark ordering and hance belong to the same ``bin''.
-This means that a landmark system consisting of $m$ such nodes results in $m!$ potentional different bins.
-%
-%The algorithm can be considered scalable, as nodes need only compute distances to a small number of predefined nodes and thus without exchanging any information.
-A potential bottleneck could be the extra load that this ``ping''-like scheme imposes to the landmarks, especially when we need instant reaction from our topology when dealing with the dynamic nature of the p2p networks.
-%
-%To answer the question as to whether the algorithm actually contributes possitivelly to the construction of an enhaned overlay, the paper defines the \emph{gain ratio} as the factor by which the latency reduces when someone communicates with a random node from the same bin than with one not in the bin.
-This is implemented with an inter-bin to an intra-bin latency ratio.
-%
-%\paragraph{Structured and Unstructured Binning}
-%The \emph{binning} scheme can be incorporated within either a structured or an unstructured overlay construction algorithm. The authors provide an example for both these cases.
-%
-%\paragraph{}
-%Assuming a structured approach based on CAN\cite{ratnasamy_can_2001} and $m$ landmark nodes.
-The coordinate space is then partitioned into $m!$ equally sized portions, each corresonding to a single ordering of the landmarks.
-To do this, the first dimension is divided into $m$ areas each of which is furtherly divided (second dimension) into $m - 1$ sections and so on.
-Having set this $m$ dimensional space, at joining time, a node measures the delay to the set of landmarks in order to determine its associated bin and thus position itself in that portion of the coordinate space associated with its landmark ordering.
-Even though this scheme can
-%
-%For unstructured overlays the paper assumes \emph{a set of $n$ nodes where each node picks any $k$ neighbour nodes so that the average routing latency on the resultant overlay is low (assuming shortest path routing)}.
-According to the proposed heuristic algorithm called \emph{BinShort-Long}, a node picks its neighbours by choosing its $\frac{k}{2}$ closest\footnote{If the node's bin is not large enough for it to pick these $\frac{k}{2}$ neighbours, it picks the required nodes from the bin that matches the most in terms of landmark ordering.} ones (named \emph{short links}), using the \emph{binning} scheme and the rest $\frac{k}{2}$ randomly (\emph{long links}).
-The former set produces well-connected \emph{pockets} of nearby nodes while the later preserves the connectivity of the graph, both yielding a proximity factor of $\alpha = 0.5$ in an attempt to preserve the benefitial properties of unstructured topologies\cite{merugu_str2unstr_2003}.
-%
-%\paragraph{}
-%% TODO double check validity
-%One disadvantage of this landmark scheme is related to the additional burden imposed to the landmark sites.
-The authors claim though that the algorithm requires so little work by the landmarks (maybe just echo to ping messages) that could in effect, act as ``unsuspecting participants''.
-Even if this is the case, the fact that it is not fully distributed, renders the protocol's scalability directly volnerable to any system size increase as well as usuitable for highly dynamic networks such as ad-hoc networks.
-Moreover, fixed points in a network are inherently more exposed to malicius attacks.
-The most significant downside of the algorithm though is that it can lead to an extremely uneven overlay ID distribution causing load inbalances and hot spots.
-Lastly, the scheme is coarse grained when it comes to distinguishing relatively close nodes\footnote{In the worst case, all nodes could ve clustered into a single bin.}.
-%
-%\subsection{Global Soft-State}
-%{\sethlcolor{yellow}\hl{HA: in paper introduction, discussion about
-%disadvantages of top-aware CAN:
-%
-%Techniques to exploit topology information in overlay
-%routing include geographic layout, proximity routing and
-%proximity neighbor selection [3]. With geographic layout
-%such as topology-aware CAN [12], the overlay structure is
-%constrained by underlying network topology. This tech-
-%nique, unfortunately, can create uneven distribution of
-%nodes in the overlay, increasing the chances of overloading
-%nodes and rendering the maintenance cost formidable. Our
-%study shows that, for a typical 10,000-node topology-aware
-%CAN, 5% nodes occupy 85-98% of the entire Cartesian
-%space, and some nodes have to maintain 450-1500 neigh-
-%bors. In Proximity routing, physical topology is not consid-
-%ered when constructing the overlay.
-%
-%[snip]
-%Studies [14] have shown that triangle inequal-
-%ity may not hold in Internet topology. In fact, study from
-%Pastry has shown that the proximity approximation is much
-%worse when using the Mercator topology that is based on
-%the real measurements of the Internet [3].
-%
-%RELATED WORK:
-%
-%Miguel Castro et al [3] divide techniques used to
-%exploit network proximity into three categories: geographic
-%layout, proximity routing and proximity neighbor selection.
-%Proximity neighbor selection is superior in terms of load
-%balancing and proximity approximation. The existing algo-
-%
-%}}
-%
-%\paragraph{}
-%The approach argued by Xu et.al in \cite{xu_globstate_2003} is to build a global map to help choose shorter routing paths, combining the landmark binning method and small scale distance probes to reveal the proximity properties of the underlying network to the overlay.
-This global view of the state is made available to all nodes in order to help them find the best way to route their messages.
-The authors focus on two aspects in order to accomplish their goal.
-%
-%\paragraph{}
-%Initially, is the \emph{generation} of proximity information and then its \emph{effective exploitation}.
-For the first, a hybrid approach is proposed, which uses landmark clustering as a preprossesing step in order to select a number of potential nearest neighbour candidates and then refine the selection by incorporating an RTT scheme to ultimately choose the closest node.
-For the second, the algorithm chooses a different path from the classic gossiping approaches for constructing and maintaining the overlay.
-It is based on landmark-clustering-based strategic placement of proximity information on the overlay enabling any node to access such information using a landmark number that reflects its physical position in the network.
-For various logical regions\footnote{This might be a high-order zone in the eCAN\cite{xu_ecan_2002} context or a set of nodes sharing a particular prefix in overlays such as Pastry.} maps of physical information are built and published where each node may appear in a maximum of $log\left( N \right)$ such maps..
-To dynamicaly adapt to changing network conditions, a node subscribes to relevant \emph{soft states} that utilize a notification system in order to initialize any necessary neighbour re-selection.
-%
-%\paragraph{}
-%Maintaining several host states at different layers, makes any content migration costly.
-Additionaly, the method does not make any continuing effort to remap the overlay structure after a node successfuly joins, in order to adapt its state to any occurance of condition change.
-Although this approach greatly reduces the routin latency to far nodes, it is unable to dynamicaly identify nodes that are close to routers and gatways in order to construct the secondary overlay.
-Nevertheless, static recognition of such nodes is currently done based on BGP reports and prechosen landmarks, sucrificing the self-organising attribute of traditional DHTs.
-%
-%\subsection{Self-Adaptive Topology Matching}
-%
-%\paragraph{}
-%Ren et.all's goal while designing the \emph{SAT-Match} algorithm \cite{ren_satmatch_2004} was to create a protocol that would be decentralised and scalable, fine-grained in detecting changes and adaptively reform in response to the dynamic nature of a classic p2p environment and last, but certainly not least, of low incuring cost.
-The method is in a nutshell, a two-phased iterative process that focuses on local optimizations in order to collectively achieve a global full overlay optimization.
-The iteration is finised when the node detects that it is physicaly close enough to its neigbours so that no additional optimization is needed.
-The paper defines \emph{stretch}, $S = \frac{\bar{L_l}}{\bar{L_p}}$, as a way quantify the \emph{topology match degree} of the constructed overlay, where $\bar{L_l}$ is the average logical link latency while the $\bar{L_p}$ is the average physical link latency.
-%
-%\paragraph{Probing phase}
-%\emph{SAT-Match} uses a small TTL value for the probing messages in order to reduce redunduncy\cite{jiang_lightflood_2008}.
-This process begins as soon as the node joins the network using a DHT mechanism.
-Each probing message contains information about the source and a small TTL value.
-The recipient of such a message returns information about itself to the source and forwards the probing message to its neighbours if the TTL is non-zero.
-The nodes been covered, are refered to as $TTL-k$ neighbourhood of the source node\footnote{Especialy, $TTL-1$ neighbourhood referes to the source's direct neighbours}.
-The list of responses is used to measure RTT to these nodes and sort the list ascending RTT order.
-%
-%\paragraph{Jump phase}
-%Blindly selecting the peer with the smallest RTT as neighbour is, generally, not the right desision to make in order to achieve global \emph{stretch} reduction, This is because, in a structured scheme, when a node jumps to connect to a physically close node, it may need to connect to other distant nodes to maintain the structure's integrity, thus creating an overall increase in the overlay's \emph{stretch}.
-The two nodes with the smallest RTT is then used in order to select one zone to jump in this phase.
-The algorithm is as follows: The source node $S$ calculates the stretch change of its $TTL-1$ neighbourhood and that of the $TTL-1$ neighbourhood of the first of the previously selected peers.
-These calculations are made as if the jump has been made.
-If the stretch reduction is over a predefined threshold the jump is performed, otherwise the second selected candidate is picked and the same computations are performed.
-If again, the threshold is not met, then no jump is ultimately done.
-In case of a jump, this is performed as a combination of \emph{leave} and \emph{join} operations, in the CAN context.
-%
-%\paragraph{}
-%Moreover the algorithm takes several issues into consiteration in order to furtherly improve the resulted overlay.
-For example when multiple nodes try to jump simultaneously into a regeon, then the logical link brakes from one attempt may result in inacurate computation of the gain factor, for an other.
-This situation is identified as \emph{contention} and the nodes use an exponatial back-off algorithm to avoid it.
-An other problem is the uneseccary traffic incuerred by the probing phase in a region that after several jumps has settled to a stable state.
-In these cases it is more likely for jump attempts to be proven worthless.
-The algorithm doubles the probing period of such nodes, every time a jump is not taken.
-%
-%\paragraph{}
-%The authors claim that this continuously adaptive mechanism achieves global topology matching optimization in a sufficiently large scope.
-This also secures the fast adaptation to frequent network changes.
-It also considered lightweight and can easily be embedded into current p2p systems, as well as effectively combined with other techniques, such as landmark binning.
-%
-%\subsection{Delay Aware P2P System}
-%A new \emph{Delay Aware P2P System (DAPS)} is introduced in \cite{zhang_daps_2005}.
-Its main goal is to reduce the time $L$ of a look-up request by dividing the routing tables of peers into several sectors in increasing delay.
-The source node emitting the query message defines a delay boundary or the pruning factor $L_t$ in the paper's context.
-Request messages will be forwarded only to nodes whose delay less than or equal to $L_t$.
-With the clustered routing tables and the loose organisation the overlay network of \emph{DAPS} is between structured and unstructured.
-%
-%\subsection{Mithos}
-%
-%\paragraph{}
-%Waldvogel and Rinaldi proposed a protocol\cite{waldvogel_mythos_2003} which incorporates a directed incremental probing to find near optimal node placement.
-%
-%\paragraph{Neighbour detection}
-%During bootstraping, the new comming node needs to know how to contact at least one of the existing members.
-A subset of these nodes will be used as the first set of candidate neighbours.
-Then, iteratively, each of the peers in the cadidate set is asked for their neighbours in order for each of the later to be probed for their network distance from the new coming one.
-The closest node is then used as the new candidate neighbour and the process is repeated until no further improvement is detected.
-Due to the fact that the algorithm may ultimately reach a local minima instead of a global one, \emph{Mithos}' approach is to first probe all neighbours within a two hop distance from the current minimum before concluding the process.
-%
-%\paragraph{ID assignment}
-%After finding its first neighbour, the newcoming peer must be assigned its ID, a very crusial selection in order not to create many local minima that will prevent efficient future neighbourhood location.
-\emph{Mithos} uses information gathered during the previous step, in order to compute the ID, which inccurs no further communication costs to the algorithm.
-This information includes
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item the two closest nodes and their neigbours, and 
-%  \item their coresponding distances.
-%\end{inparaenum}
-%Using the above, it then assigns coordinates to the new-coming node, so that Euclidian distances between the node and all known hosts predict the network latency between them\cite{cox_vivaldi_2004}.
-These synthetic coordinates as explicitly used as the node's ID and as soon as it has been established, distances can be computed in the ID space, no longer requiring physical measurements.
-%
-%\paragraph{Link establishement}
-%The last step of the algorithm is the interconnection among neighbours.
-\emph{Mithos} uses a \emph{quadrant}-based mechanism according to which each node establishes a link to the closest neighbour in each quadrant.
-During forwarding, the next hop is performed towards a neighbour in the same quadrant as the final destination\footnote{An implementation of this could be backed by a $d$-bit vector indexing (where $d$ is the number of dimensions) into the routing table.
-Then the next hop is identified by computing the difference between the current node's vector and that of the destination's.}.
-The problem is that after the ID assignment process, the new-coming node may not know of other neighbours in all quadrant and if even if it does, this cannot ensure that they are the nearest available.
-Thus, the node first identifies neighbours in all quadrants using a mechanism based on ideas similar to a perimeter walk\footnote{Used in Greedy Perimeter Stateless Routing (GPSR) protocol.} and then using parallel path processing improves the results by taking into account further geometric properties of node relationships.
-%
-%\paragraph{}
-%% TODO review this part
-%In order to avoid local minima during neighbour detection, extensive probing must be undertaken.
-In simulation, unfortunately, only very small-sized overlay topologies (of 200 to 1000 nodes) have been used and thus no safe conclusions can be made as for the behaviour of an extensively large, real-world p2p deployment of the scheme.
-%
-%\subsection{DHT-PNS}
-%
-%\paragraph{}
-%The work in \cite{hancong_pnsbased_2006} propose a proximity neighbour selection scheme on top of the Chord DHT.
-Using the Vivaldi protocol\cite{cox_vivaldi_2004} each node is assigned synthetic $2$-dimension coordinates that can be used to derive network latency using Euclidian distances in the id space and without the need of explicit probing.
-Then the algorithm performs the two steps described in the following paragraphs.
-%
-%\paragraph{Space mapping}
-%The space is partitioned using a \emph{concentric circle clustering scheme} where succesive cycles of radiuses $\rho$, $2\rho$, $3\rho$ and so on, are constructed.
-Then the formed annuluses are divided into $2\chi-1$ \emph{sectors}, where $\chi$ denots the sequence number of the annulus starting from $\chi = 1$ for the centre cycle.
-It is proved in the paper, that this way each sector occupies the same area\footnote{The same area as does the center cycle.} and assuming uniform node distribution, this characteristic, favours a more load balanced clustering operation.
-Every sector in this $2-d$ coordinate space is mapped to a unique \emph{region} in the DHT space forming a multi-layer node identifier space.
-Thus, any nodes that belong to the same sector, are mapped to the same region as well, preserving their proximity relationship unveiled by the use of the Vivaldi protocol.
-%
-%\paragraph{Routing table optimization}
-%The system described in the previous section allows any node $\alpha$ to obtain its DHT's $key_{\alpha}$ and its region's $key_r$, in a fully distributed manner, just by applying a consistent hash function.
-Using a $Get\left( key_r \right)$ RPC call in Chord, the node can obtain the region's master node, called \emph{Cluster Node (CN)} which is responsible for clustering the nodes belonging in the same sector or region with that of $\alpha$.
-On the other hand, a $Put\left( key_r, info\right)$ RPC call, registers $\alpha$ to its corresponding region and publishes its information, through the special peer CN.
-$\alpha$ can, additionaly, ask CN for other nodes that have previously joined the region in order to add them into its neighbour set for future routing table optimization.
-Even in the case when no neighbour is detected in the current region, the search is expanded to adjasent regions and towards an upper layer identifier space until one is found or the first layer reached.
-%
-%\subsection{Quasi-Chord}
-%
-%\paragraph{}
-%The approach that is proposed by Sun and Zhang in \cite{sun_quasi_2008} confronts the topology mismatch problem, in the Chord DHT context, that is created by the fact that no consiteration for the underlying physical network topology is taken into account during the construction of the identifier cycle.
-To construct a \emph{Quasi-Chord} network three steps are needed.
-First, each host acquires its coordinates in the geometric space utilizing the \emph{global network position (GNP)} protocol \cite{ng_gnp_2001}.
-Second, using the \emph{Cantor space filling curve} the $2$-dimensional space is converted to a $1$-dimensional one, used in the last, third, step to build the Quasi-Chord circle according to the host's Cantor value.
-In the following paragraphs, these steps are furtherly discussed.
-%
-%\paragraph{GNP coordinates}
-%First of all the host must be positioned in the geometric space.
-The algoritm models the P2P network with a well defined distance function in such a way, it can predict, with high accuracy, the distance between any two points in the space by just evaluating the output of the distance function on the coordinates of these points.
-This is accomplished with the \emph{global network position (GNP)} protocol which\begin{inparaenum}[\itshape i\upshape)]
-%  \item creates a reference set of $N$ landmark nodes so as to minimize the error of ICMP measured distances and coordinate computed ones between them, and then
-%  \item each host is able to measure its round-trip times to the $N$ Landmarks in order to compute its own coordinates.
-%\end{inparaenum}
-%
-%\paragraph{Cantor SPF}
-%After the $2$-d coordinates are set, in the next stage of the algorithn, each participating peer is assigned a Cantor value, according to the application of a \emph{space filling curve} on the coordinate space.
-(TODO: add figure).
-This results in the conversion of the $2$-dimensional space to a $1$-dimensional Cantor space that can more easily be more mapped to the Chord hierarchy.
-%
-%\paragraph{Quasi-Chord construction}
-%As can intuitively be infered by observing the Cantor chart, close-by nodes in the physical layer, are more likely to have similar Cantor values.
-This attribute is exploited in order to construct a topology-aware identifier space for the Chord DHT.
-The cycle is constructed by sorting nodes in ascending order.
-Each host maintains 2 finger tables, one for clockwise and one for counter-clockwise stepping.
-This helps with the connectivity of the network because its not allowed to connect the first node with the last one since this will incur heavy traffic to the later\footnote{After all their Cantor values denote that they are actually the furthest of each other.}.
-%
-%\paragraph{}
-%% TODO: doublecheck!
-%The disadvantage of the algorithm is that the coordinate assignment in the first stage is backed by a not fully distributed landmark-based algorithm.
-Moreover the Quasi-Chord model build-up is making an indirect assumption of a maximum number of allowable hosts since it is constructed.
-Last but not least the doubling of the required routing information which needs to be created and maintained is an additional negative point to the efficiency and the scallability of the algorithm.
-%
-%\subsection{LAPTOP}
-%{\sethlcolor{yellow}\hl{HA - two important issues for decentralized structured P2P:
-%P2P applications. However, the design of a decentralized but structured P2P network has to
-%overcome two critical issues. The first issue is the long routing latency. Several proximity
-%schemes  have been proposed to avert long routing latency in current structured P2P
-%networks, but they require a high-complexity procedure to periodically maintain the routing
-%table (e.g. Pastry system) or they need pre-chosen landmarks to construct the overlay.
-%However, the P2P system is by its very nature unstable since nodes join and leave frequently.
-%For instance, the study of Gnutella shows around approximately 1200 membership changes
-%per minute in a 100 000 nodes P2P system. Another proximity scheme needs some pre-chosen
-%landmarks or a complete BGP routing table support. As a result, they both increase the
-%difficulty of the P2P system deployment.
-%The second issue is system maintenance overhead. The existing structured P2P networks allow
-%nodes to keep some nearby nodes in their routing tables in order to achieve efficient routing. The
-%}}
-%
-%\paragraph{}
-%Laptop \cite{wu_laptop_2007}, introduced by Wu et al., organises the overlay into a tree based hierarchy with main focus on child-to-father relationships in order to reduce hops during message routing as well as minimize maintenance overhead.
-Additionally, a caching scheme is also incorporated so as to furtherly reduce routing table update cost.
-The authors argue that Laptop overlay network is bounded by $O\left( log_d N \right)$ hops in a balanced overlay tree, where $N$ is the number of nodes, and $d$ is the maximum degree of each node.
-It utilizes a geographical layout approach  and constructs a geographical layout in a self-organizing and efficient fashion, by estimating the round trip time (RTT) to a small number of nodes in the overlay network in order to make them roughly aware of their physical distances among them.
-%
-%\paragraph{}
-%The protocol is based on four definitions.
-%\begin{enumerate}
-% \item The amplitude of all possible measured RTTs is devided into intervals.
-Each node measures its distance to its parent and is assigned a label $L_i$ where $i$ denotes the configurable RTT interval in which the measured distance falls into.
-A special kind of node, the root, is initially assigned the $L_1$ label and maintains (as all $L_1$ nodes do) a list of other $L_1$ nodes in the overlay.
-% \item Any node can have children with level lower than theirs, except an $L_{max}$ node which can only have $L_{max}$ level children and only in the case when its parent has reached its maximum degree.
-% \item Each node is assigned an address in a dotted format (e.g 1.3.4).
-Each octet ranges from $1$ to $d$, where $d$ is the maximum degree of the nodes.
-The assignment process is done by appending a unique octet to the address of its parent.
-Root node is assigned address 1.
-% \item For any descendant node $Y$ of a node $X$, the measured distance among each other, must always be less than the lower bound of the RTT interval denoted by $X$'s label.
-%\end{enumerate}
-%
-%\paragraph{}
-%The routing scheme is similar to the IP's longest-prefix matching scheme.
-At each forwarding hop, any message travels up the tree until the first common uncestor of source and destination node is reached and then starts descending to arrive to its target.
-During tree traversals, special entries in the routing tables, called \emph{routing cache}, are maintained in order to increase routing efficiency and achieve finer load balance.
-Caching enables a node to forward a message to a better longest-prefix match than that of its direct ancestor making a large, quicker and more cost effective step through the overlay and toward the destination.
-To improve scalability, the number of children nodes and the size of the routing cache are limited.
-%
-%In terms of overlay maintenance, Laptop incorporates a simple \emph{heartbeat}-based technique where only the parent node is responsible for monitoring its children.
-%
-%At join process the newcomming node is assigned its level label as well as its address by its parent node.
-Additionaly it initializes its routing table (with normal and caching entries) as it traverses the overlay in search for its parents node.
-%The newcommer $N$ locates the root node and the later responds with a list of $L_1$ nodes.
-$N$ then probes each of the $L_1$ nodes in search for the closest one.
-If the measured RTT to the closest $L_1$ falls into the first interval then the newcommer becomes a $L_1$ node as well.
-Otherwise node $N$ sets the closest $L_1$ node as its potential parent node.
-This potential parent becomes the actual parent if it does not have any other children.
-If it has, $N$ gets a list of these $L_i$ nodes and by measuring the RTT to each of them tries to spot a new potential parent in order to repeat the above process.
-%
-%\paragraph{}
-%During a gracefull departure, the refered node checkes for children in the overlay.
-If it does not have any, it simply notifies its parent and leaves.
-If it has, it selects the child node with the lowest RTT to it in order to take its place so that the locality property is preserved.
-%
-%In case of an arbitrary failure, the children of the failed node are detecting their parent's absent when the later stops acknowledging their heartbeat messages.
-They start emmiting special messages to their grand parent node\footnote{The address of the grandparent node is stored during join process}.
-In case of no response from the grandparent node, children invoke the joining procedure to detect their new parent node.The parent of the failed node, aggregates the notifications from the above children nodes for a period of time, and then chooses the one with the lowest level label as the takeover node.
-Potential ties are broken by favouring the lower RTT.
-The parent of the failed node, finally informs accordingly all the children about the change in the hierarchy.
-%
-%\subsection{IP-based Clustering (IPBC)}
-%
-%\paragraph{}
-%\emph{IP-based clustering} \cite{karwaczynski_ipbc_2007} is a proximity neighbour selection technique that is based on a longest IP prefix matching scheme in order to quantify the proximity among peer nodes.
-The relation of the IP address and the physical location of a node in the topology of the Internet, is intuitevely realized by operation the network layer of the Internet stack.
-For example, nodes in the same local subnetwork share the same 3 bytes of their address.
-Reports in \cite{freedman_iploc_2005} state that almost $97\%$ of prefixes longer than 3 bytes correspond to address assignments at a single geographic location.
-Moreover, observing the \emph{IANA IPv4 Address Space Registry} we can infer that, since blocks of consecutive octet prefixes are assigned to the same Regional Internet Registries (RIRs), nodes inside will be physicaly close to each other.
-%
-%\paragraph{}
-%The implementation in this paper is based on the above observation in order to create proximity information.
-This information is stored in a decentralised manner, within the overlay itself, just like any other resource would have been.
-In order to be published in the overlay, each node is first assigned a unique identifier and subsequently generates a key by hashing a fixed-length prefix of its IP.
-Authors argue on the prefix length to be used, as there is a thin balance between reducing the propability of finding closer nodes by adopting a long prefix and overload nodes that are responsible for information published by many peers when chosing to hash a short one.
-Their verdict is for the use of a 16-bit wide prefix for real world systems deployed in an Internet scale.
-In either case, ID and IP is then stored in the DHT using this generated key.
-This way, any node, at any time (either at join time or during peer lifetime for topology adaptation) can acquire information about close-by nodes just by quering the DHT for a specific key.
-Moreover, the algorithm takes care of the freshness of the proximity information in two ways.
-First, the advertising nodes themselves periodically update their advertisements or when they voluntarily leave the overlay, they explicitly remove their data.
-On the other hand, in case of a failure each publication is assigned an expiration time, and thus ultimately removed by the DHT maintenance mechanisms.
-%
-%\subsection{CHOord considering Proximity on IPv6 (CHOP6)}
-%
-%\paragraph{}
-%\cite{morimoto_chop6_2007} roughly estimates the proximity among nodes by exploiting the IPv6 address format as well as RTT information if this is available.The first is achieved by introducing a 64-bit ID scheme in which the least significant bit part\footnote{The exact range is a predefined system variable} is the IPv6 global routing prefix and thus enabling a longest prefix match scheme.
-The observation in which the protocol is based roots from the IPv6 address block assignment.
-Specicaly, block of /16 or /23 in size are assigned to Regional Internet Registries (RIRs) by Internet Assigned Numbers Authority (IANA).
-An RIR divides the signed address blockes into smaller address blocks which are furtherly assigned to Natinal Internet Registries (NIRs).
-That way, it is possible to estimate a node's geographical location by simply examining the upper 32-bits of its IPv6 address.
-Moreover, similar to Chord, CHOP6 uses a finger table, whose entries hold more than one candidate nodes.
-There are three cases in which a node chooses the next hop according to information it posseses.
-%\begin{itemize}
-% \item When no information about RTT is available candidate next hops, the sender just selects a node in the finger table entry which shares the longest prefix with the destination.
-% \item There is another possibility that after some communication with other nodes, the source node should know of the RTTs to some of the nodes in finger table entry.
-In this case, the source node chooses the one with the smallest RTT with propability $p$.
-If there is a node with no measured RTT then the sender can select such a node with probability $1 - p$
-% \item In this last case the source node has already communicated with all candidate node in the finger table entry.
-Thus, the node selects the node whose RTT is the smallest with probability $p$, the one with the second smallest with probability $q$ and so on, where $0 \leq \ldots < r < q < p < 1$ and $p+q+r+\ldots \leq 1$
-%\end{itemize}
-%
-%\subsection{Proximity in Kademlia}
-%
-%\paragraph{}
-%In \cite{kaune_pkad_2008}, Kaune et al., studied the Kademlia DHT, in order to build a proximity aware scheme that would work in the context of iterative lookup algorithms.
-The protocols discussed, focus on both the overlay and underlay tiers in the sense that in the first improves the routing performance according to a cost metric that is provided by the later.
-This metric is tailored to the needs of a specific aplication.
-For example, routing could be optimised to maximise the within-ISP traffic or reduce the lookup latencies or even avoid contacting untrustworthy subnetworks.
-%
-%\paragraph{}
-%Two algorithms are presented for the overlay optimization.
-One is \emph{Proximity Neighbour Selection (PNS)} and one for \emph{Proximity Route Selection (PRS)}.
-The first aims at keeping peers with the least contact cost in the routing table.
-As Kademlia constantly learns new peers (incoming requests, through iterative lookup) no special algorithm for searching more cost effective peers is necessary and simply choose the best peers seen so far.
-The later (PRS) aims at choosing the best next hop during routing a message.
-Due to the fact that the routing in Kademlia is iterative, it is the initiator of the lookup that chooses each next hop from a set of candidate nodes.
-The ``vanilla'' Kedemlia always chooses the closest node with respect to the XOR metric but PRS Kademlia chooses the one with the smallest underlay metric cost.
-Thus, as in all such approaches, there is a trade-off between the overlay and underlay distances.
-%
-%An underlay metric provides information about the underlay network. The paper distingushes between three kinds of possibilities to quantify it:
-%\begin{itemize}
-% \item Using measurements gained by previous lookups
-% \item Using measurements acquired by other peers or jointly calculated with others
-% \item Using a local database to look up information
-%\end{itemize}
-%In the paper, for locality of traffic and subsequently for reducing the lookup latency, a clustering scheme has been implemented exploiting information stored in a geographic information database (GeoIP,MaxMind) about peers.
-The goal of such a metric is to constrain the largest portion of communication within the limits of a peer's ISP.
-%Another incorporated approach is Vivaldi\cite{cox_vivaldi_2004}.
-At first all coordinates are random but as peers start to communicate they calculate RTTs as well, gradually updating their own coordinates.
-Vivaldi creates a system where a peer knowing the coordinates of a another, can approximate communication latency without the need of additional prompting.
-%
-%\subsection{Cone}
-%
-%\paragraph{}
-%Cone\cite{wang_cone_2007} uses a two-layered identifier space.
-The first, named Chord-layer identifier, denoted as $Id_{Chord}$, is the same as in ``vanilla'' Chord.
-The second is the Cone-layer identifier, $Id_{Cone}$ which is constructed by two component identifiers.
-The first, known as \emph{group identifier (gid)} denotes a relevant group the node belongs to while the second, namely \emph{local identifier (lid)} indicates the local identifier within the group.
-The group concept, which is introduced here, is a way of dividing nodes according to a common $Id_{Chord}$ prefix.
-%The structure of a Cone overlay, retains the Chord's circular topology.
-The differce lies on the fact that, now, two rings are created.
-A big ring, where nodes with the same \emph{gid} are arranged at each position.
-Each of these positions are a smaller ring for the particular group's \emph{lid}s.
-The routing is achieved in both clockwise and anti-clockwise direction in the big-ring.
-For this reason two routing tables are maintained, namely \emph{front} and \emph{back} finger tables, respectively.
-Entries in these tables, display physical network proximity with the current node.
-Moreover, a third table called \emph{group} table maintains information about other online peers within the current node's group in a way that entries are now close in the ID space.
-%
-%\paragraph{}
-%Routing in Cone, comprises of the inter-group algoritm and the intra-group algorithm.
-First the group of nodes to which the desired key lies is detected, exploiting physical proximity information (front and back finger tables).
-Second during the next and last hop, the message is forwarded to the exact node that contains the desired key.
-%
-%Cone uses Landmark+RTTs to generate proximity information and exploits this information using proximity neighbour selection.
-%
-%\subsection{DynaMO}
-%
-%\paragraph{}
-%In their work in \cite{winter_dynamo_2004}, Winter et al, try to build an overlay network, called \emph{DynaMO}, that tries to consider not only the physical proximity of peers but their mobility attributes as well.
-DynaMO is based on a Pastry overlay network.
-This was an intentional choice from the authors because Pastry's built-in locality heuristics are thoroughly analised in the literature \cite{castro_proximityp2p_2002} providing good backround against which to test and compare their results.
-In order to adapt to mobile, ad-hoc networks, though, special care has been put to the maintenance of an even overlay ID distribution so that hot-spots are avoided.
-Pastry assignes them in a randomised fashion and then tries to consider proximity through the joining process and routing table maintenance.
-On the other hand, DynaMO dictates a newcoming node to gather information concerning its physical neighbourhood and uses it to assign itself an appropriate overlay ID.
-%
-%DynaMO tries to capitalize an observation of Pastry's routing mechanism.
-Each node's table consists of rows equal to the number of digits of the overlay IDs and columns equal to the ID's base.
-As we go down the table rows, the matching prefix between the current node's ID and the row's entries increases by one.
-Thus, the leaf set contains the closest nodes in the ID space.
-Additionally as the prefix match increases by one, the result is exponentially less candidates that can fill the tables entries as the row increases.
-This leads to the observation \cite{antony_pastry_2001,castro_proximityp2p_2002} that from overlay routing hop to overlay routing hop the physical distance between nodes is likely to increase leading to a dominating last routing step in terms of the overall plysical routing path distance travelled during a key lookup.
-Since the last routing step is usually taken from the leaf set, DynaMO focuses on making this last step as physicaly close as possible.
-In this context, two approaches are considered namely \emph{Random Landmarking (RLM)} and \emph{Closest Neighbour Prefix Assignment (CNPA)}.
-%
-%\paragraph{}
-%Random Landmarking uses the overlay lookup mechanism to locate nodes responsible for a fixed set of carefully chosen\footnote{In a way that the ID space is divided into equal portions} landmark keys.
-This means that when a node is assigned a landmark key then it plays the role of a temporary landmark for the network.
-Any joing node will measure its distance to that and all other landmarks and assign to itself an ID consisting of a prefix taken from the closest landmark and a remainder that could be assigned randomnly or generated by mechanism that takes into account physical neighbourhood.
-The legth of the prefix can be determined as $prefix_length=|log_b k|$, where $b$ is the ID base and $k$ the number of landmarks.
-The above scheme results in physically close nodes, forming regions of common ID prefix that are likely to be close to each other in the ID space and thus bringing a node's leaf set closer to itself.
-Additionally, the advantage of dynamic landmark nodes is that the failed ones can instantly be replaced by new responsibles, that share similar physical attributes to the failed and thus preserving the balance of the network.
-%
-%\paragraph{}
-%RLM may incur more traffic especially on landmark nodes, load that in some cases is not acceptable.
-CPNA on the other hand takes advantage of Pastry's specification that a new coming node is bootstraped by a physicaly close node.
-The new comer then assumes the ID prefix of that neighbour while the rest is generated similarly to RLM.
-Unfortunately, less overhead comes at the expense of being more coarse-grained.
-%
-%\paragraph{}
-%Both algorithms, are protected against the formation of physical landmark clusters or imbalanced ID distribution\footnote{More common during the initial formation of an overlay network} , by introducing the \emph{landmark gravitation range} as a theshold over which landmark keys are reassinged (for the RLM approach) or unutilized ID prefix ranges are detected and used (for the CPNA scheme) in order to balance the distribution of regions in the overlay.
-%
-%\subsection{PChord}
-%\emph{PChord}\cite{hong_pchord_2005} is a based on the Chord DHT which adds proximity routing into its routing mechanism.
-The main modification over the ``vanilla'' Chord is the inclusion of a \emph{proximity list} into its routing table so that the next hop is decided based not only by considering best progress towards the key, but on the physical proximity of the candidate nodes as well.
-The list is, at join time, empty but as the PChord node starts to interact with other nodes it applies a heuristic mechanism to fill up the list.
-Entries are dynamicaly added or removed as the network state is constantly changing.
-Because, it is guaranteed that at each routing step there is progress towards the target node in the ID space, PChord will result in lower hop number than Chord, for each hop is larger or at least equal in key space in PChord than in Chord.
-Additionaly, passing through proximity links in the underlay means reduced routing cost.
-Moreover, if the number of entries in the proximity list is the same as the number of network partitions, PChord prevents hops from jumping back to the same network the current node belongs to.
-%
-%\subsection{AChord}
-%\emph{AChord}\cite{dao_achord_2006} uses IPv6's anycast functionality in order to
-%\begin{itemize}
-% \item releave the protocol from complex joining procedures, and
-% \item achieve high accuracy of network proximity giving high routing efficiency,
-%\end{itemize}
-%all with a simple and lightweight mechanism that requires few changes to Chord without affecting its own advantageous characteristics.
-Anycast delivers a message comming from the outside of an anycast group to the physicaly closest node in that group.
-AChord organizes all nodes participating in the overlay network into an anycast group.
-Any joining node comes outside of that group and, thus, is automaticaly forwarded to the physicaly nearest node in order to bootstrap, avoiding
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item the need to, explicitly, maintain such nodes, and
-%  \item the effort of finding a way of locating the physically nearest among them.
-%\end{inparaenum}
-%The ID of the new comming node is computed based on the ID of the bootstrap node and bootstrap's predecessor in a way that its ID will position itself between the formentioned.
-%
-%After joining, the finger table is build the same way as in the Chord protocol.
-Moreover, additional nodes are maintained in a structure called \emph{neighbourship table} which stores information about the closest known nodes\footnote{As a first entry, at bootstrap, the node stores information about the bootstrap node.}.
-The routing decision is made by using both the neighbourship and the finger table.
-%
-%\subsection{Chord6}
-%\emph{Chord6} \cite{xiong_chord6_2005} is a Chord variant that tries to exploit the hierarchical features of IPv6 in order to create a substrate that reduces interdomain trafic and does that in a cost efficient way.
-Chord6 bareley modifies the original Chord protocol, only in the part of identifier definition.
-This, renders the aproach easily portable to other DHTs such as CAN, Pastry and Tapestry.
-In Chord6 the identifier contains two pares: the higher bits are obtained by hashing the  node's IPv6 address prefix of specific length, while the remaining lower bitss are the hash value of the rest of that IPv6 address.
-This way nodes in a domain will be mapped onto a contihuous key space on the overlay network.
-This guarantees that messages are routed in a way that they do not hop in and out of domains many time, thus minimizing overall routing cost.
-(TODO: Recheck validity) However, even though nodes in the same domain would have close identifiers, nodes in two close domains may have very different ones.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% TODO: Should be rewritten
-\section{Conclusions}
-Peer-to-peer architecture has been in the center of research attention in the
-last decade. Especially the decentralized unstructured genre exploits the
-advantages of loose coupling and self organization of computing nodes to form
-application-layer networks on top of the physical, best-effort infrastructure of
-the Internet that exhibit interesting properties. Scalability problems arouse
-quickly, though, because of the inefficient construction of this overlay that
-was built with no concern for the underlying physical network that causes a
-great deal of redundant traffic. The problem was identified by the research
-community as the topology mismatch problem between the overlay and the
-corresponding underlying physical network and a great deal of effort has been
-set towards alleviating it. Some fruits of this effort have been gathered and
-presented in this survey. Measurement of link cost through latency or RTT and
-deletion of inefficient established connections are the key concepts of almost
-all approaches. Others, manage to address the problem through hierarchical peer
-clustering (e.g. best IP matching). Additionally, some protocols focus on
-specific problems that furtherly arise, such as overlay partitioning, search
-scope reduction or convergence speed, to name just a few. Another desire is to
-form a protocol that could be applied to both decentralized unstructured and
-decentralized structured networks. Unfortunately no approach has equally
-addressed these problems in order to form a robust solution. So the field seems
-to be, still, fertile for any, new, clever idea.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliographystyle{acmtrans}
-%\bibliographystyle{abbrv}
-\bibliography{bib/p2p-overlays}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{received}
-Received Month Year;
-revised Month Year; accepted Month Year
-\end{received}
-
-\end{document}
+and application development. Last but not least, we identified and formaly
+defined the phenomenon of the topology mismatch problem and its impact on the
+utilization of network resources. The rest of the survey\ldots
+
+Sections~\ref{sec:unstructured} and \ref{sec:structured} discuss the recent
+academic work that has been conducted in the field.
+
+Ultimately, it reviews the most important resources found in the literature that
+attempt to tackle the problem as it proposes a taxonomic scheme, using a set of
+supertype-subtype relationships based on unique characteristics they have and/or
+specific goals they target. What follows does not claim to be a thorough
+citation of all known protocols that are available out there, but instead, a
+careful selection of those that left a distinctive fingerprint contribution in
+the efforts of the research community to alleviate the topology mismatch
+problem.

### 2014-01-21 mard-survey.tex
2012.05.29 (Vassilis) Monolithic file broken down to multiple smaller ones (per section) for better parallel contribution. Additional material on more recent algorithms for unstructured networks. Now also covering the BitTorrent protocol. Merged Huseyin's figures from previous contribution and code refactored for better referencing. Fixed bibliography bug (references appearing as ?).

--- 2013-11-10 mard-survey.tex
@@
 
-%opening
-% TODO: Contribution of this paper
-% Q&A: 'Overlay' or 'Peer-to-Peer'/'P2P'
-\title{Alleviating the Topology Mismatch Problem in Unstructured and Structured Overlay Networks: A Survey}
+\markboth{}{}{}
+
+\title{Alleviating the Topology Mismatch Problem in Unstructured and Structured
+Overlay Networks: A Survey}
+
 \author{
-Vassilis Moustakas$^1$, H\"useyin Akcan$^2$, Mema Roussopoulou$^1$ and Alex Delis$^1$\\
-$^1$Department of Informatics and Telecommunications,\\
-National and Kapodistrian University of Athens\\
-\mbox{\texttt{\{b.moystakas, mema, ad\}@di.uoa.gr}}\\
-$^2$Department of Software Engineering,\\
-Izmir University of Economics, Izmir, Turkey \\
-\mbox{\texttt{huseyin.akcan@ieu.edu.tr}}
+VASSILIS S. MOUSTAKAS\\\emph{Department of Informatics and Telecommunications}\\\emph{National and Kapodistrian University of Athens}\\
+H\"USEYIN AKCAN\\\emph{Department of Software Engineering}\\\emph{Izmir University of Economics}\\
+MEMA ROUSSOPOULOS\\\emph{Department of Informatics and Telecommunications}\\\emph{National and Kapodistrian University of Athens}\\
+AND\\
+ALEX DELIS\\\emph{Department of Informatics and Telecommunications}\\\emph{National and Kapodistrian University of Athens}\\
+%  Vassilis Moustakas$^1$, H\"useyin Akcan$^2$, Mema Roussopoulou$^1$ and Alex
+%Delis$^1$\\
+%  $^1$Department of Informatics and Telecommunications,\\
+%  National and Kapodistrian University of Athens\\
+%  \mbox{\texttt{\{b.moystakas, mema, ad\}@di.uoa.gr}}\\
+%  $^2$Department of Software Engineering,\\
+%  Izmir University of Economics, Izmir, Turkey \\
+%  \mbox{\texttt{huseyin.akcan@ieu.edu.tr}}
 }
 
+\begin{abstract}
+Peer-to-peer (P2P) systems is a rapidly growing application architecture. The
+overlay network abstraction on top of a best-effort infrastructure provides a
+powerful tool to realize a plethora of ever-wanted exotic features like
+anonymity, high availability and robustness, load balancing, quality-of-service
+(QoS) at the application layer and many more. Unfortunately weaknesses of early
+effectuations constrained applications from unleashing the full potential of the
+paradigm. One of them is the topology mismatch problem between the overlay and
+the physical underlying network, a problem that imposes a huge amount of
+unnecessary stress to the network resources and, thus, impede applications from
+scaling. In this survey we investigate on the recent research developments
+towards the alleviation of this problem, in both unstructured and structured
+overlay architectures. We visit the various algorithms available in the
+lietrature, focusing on the presentation of how each of them tries to tackle the
+problem. Additionally, we provide a finer grained categorisation of the
+approaches based on the generic type of methodology they use in order to
+accomplish this goal.
+% TODO: Review this after the comparison sections are fixed.
+Finally we discuss the advantages, disadvantages and novelties introduced by the
+algotithms and offer pictorial comparisons for them in order to further
+strengthen and organise understanding of the information and knowledge cited, in
+an attempt to make this survey a one-stop-shop reference for anyone interested
+in the topic, from researchers and students, to system architects and
+developers.
+\end{abstract}
+
+\category{}{}{}
+
+\terms{Topology mismatch, distributed, peer-to-peer, P2P, overlay}
+
+\keywords{Topology mismatch problem, distributed algorithm, peer-to-peer
+architecture, overlay network }
+
 \begin{document}
+
+\begin{bottomstuff}
+SAY SOMETHING COOL HERE...
+\end{bottomstuff}
+
 \maketitle
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{abstract}
-% TODO: Contribution of this paper
-Peer-to-peer (p2p) systems is a rapidly growing application architecture.
-The overlay network abstraction on top of a best-effort infrastructure provides a powerful tool to realize ever-wanted exotic features like anonymity, high availability and robustness, load balancing, QoS at the application layer and many more.
-Unfortunately weaknesses of early effectuations constrained applications from unleashing the full potential of the paradigm.
-One of them is the topology mismatch problem between the overlay and the physical underlying networks, a problem that imposes unnecessary stress to the network resources.
-This survey investigates on the recent research developments towards the alleviation of the problem, in both unstructured and structures p2p schemes.
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-
-% TODO: Contribution of this survey
-% UNDER CONSTRUCTION
-
-% TODO: Structure of this paper
-% UNDER CONSTRUCTION!
-This survey begins with a quick review of the conception of the overlay network and the available architectures of p2p systems.
-It continues with the identification of the phenomenon of the topology mismatch between overlay and underlying networks and its impact on the utilization of network resources.
-Ultimately, it reviews the most important resources found in the literature that attempt to tackle the problem as it proposes a taxonomic scheme, using a set of supertype-subtype relationships based on unique characteristics they have and/or specific goals they target.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% OK
-\section{Overlay Networks and P2P Systems}
-
-\begin{figure}[h!]
-\centering
-  \includegraphics[scale=0.5]{img/p2p.pdf}
-\caption{An example overlay network.}
-\label{figure:overlay}
-\end{figure}
-
-An \emph{overlay network} is an abstract, logical interconnection of entities
-formed on top of another network. It consists of a set of nodes, the computing
-elements, that are being connected by virtual (logical) point-to-point links.
-For example, the \emph{Asymmetric Digital Subscriber Line (ADSL)} is, actually,
-an overlay network on top of the \emph{Public Switched Telephone Network
-(PSTN)}. Figure \ref{figure:overlay} illustrates an example physical ethernet network
-and the corresponding overlay network formed by four nodes. 
-Nodes $A$ and $B$ are on the same local network, while $C$ and $D$ are in
-different networks. The top layer represents the overlay network formed by these
-nodes, which is a reflection of the application layer connections among these
-nodes. As the topology of the network can change based on the application, in
-this example, it is easy to observe that even though nodes $A$ and $B$ are in the same
-network, their communication goes through node $C$ on another network,
-demonstrating a non-optimal mapping between the physical and the overlay
-networks.
-An overlay network is generally useful when
-abstraction of the network layer from the application layer is important for
-applications, particularly when the underlying network structure is subject to
-frequent changes.%, which is a common case for ad-hoc networks.
-
-The abstraction of overlay networks has been proposed as a way to implement
-efficient, fully distributed, application layer services on top of a best-effort
-IP layer forming a widely ranged family of protocols collectively known as
-\emph{peer-to-peer} or \emph{p2p} for short. P2p network systems gained
-significant attention, in recent years, due to their plethora of unique features
-in supporting file sharing among huge numbers of (inter-) networked computers
-called \emph{peers}, where each, such, \emph{peer} could act both as a resource
-provider and a resource consumer. This changed the traditional
-\emph{client-server} model dominating the internet and lead to the introduction
-of the \emph{servent}-concept \cite{gnutella}, a portmanteau that blends the
-notions of \emph{server} and \emph{client} to denote the twofold role of the
-participants in a p2p network. Such serverless systems, proved to be able to
-achieve outstanding aggregate resource capacities as participants join the
-system
-%\footnote{Unfortunatelly, there has been observed mitigation of this
-%self-scaling property by the undesirable behaviour of participating users,
-%usually answered in popular file-sharing systems; from \emph{free-riding} (i.e.
-%in Napster and Gnutella networks \cite{saroiu_measurefileshare_2002,
-%adar_gnutellafreeriders_2000, hughes_gnutellafreeride_2005}), to the
-%distribution of illegal content and/or other \emph{socio-technical} issues
-%\cite{hughes_socp2p_2008}.}, 
-without requiring additional expenditure for infrastructure.
-
-As p2p systems evolved, three main architectures have been emerged, namely:
-\begin{itemize}
-  \item \emph{centralized}
-  \item \emph{decentralized unstructured}
-  \item \emph{decentralized structured}
-\end{itemize}
-
-\emph{Centralized} architectures were the first to recognize that requests for popular content need not be sent to a central server but instead could be handled by the many hosts that already posses the content.
-\emph{Napster}, for example, maintained a \emph{central index server} based on file lists provided by participating peers.
-The central index server was queried by the users and it returned pointers to the actual content.
-Thus, by centralizing search while distributing downloads, Napster achieved a highly functional design that was widely acknowledged, at the time, as ``the fastest growing Internet application ever''.
-This centralized search facility, though, renders this scheme not fully distributed and can be proven its ``Achilles' heel'' in terms of being a scalability bottleneck, a central point of failure and vulnerable to malicious acts (e.g.
-\emph{Denial-of-Service (DoS) attacks}).
-
-The centralized approach has been eventually replaced by architectures that distributed both the search and the download capabilities.
-In these \emph{decentralized unstructured} architectures, file placement is random, which means there is no correlation with the network topology whatsoever \cite{yang_improvep2psearch_2002}.
-The most important properties of such systems are that they support inherent heterogeneity of peers, are highly resilient to peer failures, and incur low maintenance overhead at handling the dynamics of peer participation \cite{stutzbach_churn_2006}.
-In the literature, they are also known as \emph{broadcast-based} systems, because they use \emph{message flooding} among peers (i.e.
-Gnutella\cite{gnutella}) or among super-peers (i.e KaZaA\cite{kazaa}) to propagate queries.
-Even though the unstructured approach became highly popular among file sharing applications, they do have a certain disadvantage in locating rare objects due to the unstructured nature and the limitations of the flooding approach.
-%Each (super-) peer receiving a query decreases the incoming query message's lifetime which is measured in \emph{time-to-live (TTL)} hops, back-propagates its results, and  if the lifetime of the incoming message did not expired (TTL did not reach zero), it forwards the message to all its neighbours except from the one it receives it from.
-% TODO: maybe enumerate applications of decentralized unstructured p2ps.
-
-Recently, \emph{decentralized structured} schemes have been proposed in order to provide a self-organising infrastructure for large-scale p2p applications \cite{ratnasamy_can_2001,stoica_chord_2001,antony_pastry_2001,zhao_tapestry_2001,maymounkov_kademlia_2002}.
-They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes through a deterministic mechanism.
-The main advantage of the DHT approach of decentralized structured p2p networks over decentralized unstructured ones, is that they provide a guaranteed bound on the number of overlay routing hops that have to be taken in order for an object to be located even in the case where only a single copy exists in the system.
-This is $O \left ( log n \right )$ compared to Gnutella, that requires $O \left ( n \right )$ to reliably locate a specific object.
-% TODO: maybe enumerate applications of decentralized structured p2ps. (OceanStore, Far)
-
-%{\sethlcolor{yellow}\hl{HA: Give the brief advantages and disadvantages of structured and unstructured
-%approaches here!}}
-
-% TODO: enumerate applications of overlay networks and p2p systems regardless of architecture
-% UNDER CONSTRUCTION
-% Q&A: Need more??
-Initial efforts targeted on the implementation of one-to-many addressing schemes
-that could replace \emph{IP multicast} in providing higher quality of streaming
-media through \emph{quality of service (QoS)} guarantees. IP multicast as well
-as other proposals such as \emph{IntServ} and \emph{DiffServ}
-\cite{cisco_diffserv_2005} have not seen wide acceptance (yet?), largely because
-they require changes to all routers of the network. On the other hand, an
-overlay network can be incrementally deployed on end-systems running the overlay
-protocol software, without cooperation from the ISPs. Academic research includes
-\cite{chu_esm_2000,jannotti_overcast_2000,kwon_tag_2002} among others.  Overlay
-networks were also implemented in order to back the routing of messages to
-destinations whose address is not known in advance. This resulted in the
-appearance of well known peer-to-peer protocols such as \cite{gnutella} and
-\cite{maymounkov_kademlia_2002} widely used for digital content sharing among
-their network nodes.  Other special purpose overlay networks include
-\cite{anderson_ron_2001} for resilient routing; \cite{subramanian_overqos_2004}
-for quality of service guarantees; and \cite{clarke_freenet_2001} for anonymity;
-to name just a few.
-
-The main focus of this paper is the topology mismatch problem, therefore the
-available approaches in overlay networks are briefly summarized, and interested
-readers are pointed to the available surveys on overlay networks
-\cite{TheotokisS04,LuaCPSL05}.
+\input{intro}
 
@@
 %TODO: Maybe this section should be reviewed or deleted or go to another section
-%Studying the behaviour of the various peer-to-peer schemes \cite{matei_mapgnutella_2002, lv_randomwalks_2002, merugu_str2unstr_2003} showed that the ad-hoc network topology of unstructured overlay networks that preserve \emph{Power Law} and \emph{Small World} characteristics\footnote{Power Law describes the node degree while the Small World describes characteristics of path length and clustering coefficient.
-The clustering coefficient for a node $\upsilon$ in a graph $G = \left( V, E \right)$ is defined as the ratio of the existing connections between $\upsilon$'s neighbouring nodes to $\gamma \times \left( \gamma - 1 \right)$, where $\gamma$ is the number of neighbouring nodes of $\upsilon$.
-High cluster coefficient means that neighbouring nodes of any node $\upsilon$ likely connect one another.} \cite{faloutsos_powerlaw_1999, saroiu_measurefileshare_2002} offer a more promising approach.
-Particularly:
+%Studying the behaviour of the various peer-to-peer schemes
+%\cite{matei_mapgnutella_2002, lv_randomwalks_2002, merugu_str2unstr_2003}
+%showed that the ad-hoc network topology of unstructured overlay networks that
+%preserve \emph{Power Law} and \emph{Small World} characteristics\footnote{Power
+%Law describes the node degree while the Small World describes characteristics
+%of path length and clustering coefficient. The clustering coefficient for a
+%node $\upsilon$ in a graph $G = \left( V, E \right)$ is defined as the ratio of
+%the existing connections between $\upsilon$'s neighbouring nodes to $\gamma
+%\times \left( \gamma - 1 \right)$, where $\gamma$ is the number of neighbouring
+%nodes of $\upsilon$. High cluster coefficient means that neighbouring nodes of
+%any node $\upsilon$ likely connect one another.} \cite{faloutsos_powerlaw_1999,
+%saroiu_measurefileshare_2002} offer a more promising approach. Particularly:
 %\begin{itemize}
-%  \item Peer-to-peer clients are extremely \emph{transient}.
-Unstructured systems can have high maintenance traffic in delivering messages, updating the mapping, discovering failures and replicating lost data or pointers, making them insufficient on highly volatile networks.
-%  \item \emph{Keyword searches} versus \emph{exact-match queries}.
-In DHTs there is a tight control between the data placement and the topology of the network.
-For this reason it is hard to efficiently support partially matched queries while Gnutella and other similar systems effortlessly support keyword searches and other complex queries since the mechanism is realized locally, on a node-by-node basis.
-%  \item Popular content is located at multiple peers and thus it is more likely for a flooding-based search to return results.
-DHTs, on the other hand, fit better in the systems which require ability to reliably locate content, even in the extreme case that only a single-copy exists in the network.
+%  \item Peer-to-peer clients are extremely \emph{transient}. Unstructured
+%systems can have high maintenance traffic in delivering messages, updating the
+%mapping, discovering failures and replicating lost data or pointers, making
+%them insufficient on highly volatile networks.
+%  \item \emph{Keyword searches} versus \emph{exact-match queries}. In DHTs
+%there is a tight control between the data placement and the topology of the
+%network. For this reason it is hard to efficiently support partially matched
+%queries while Gnutella and other similar systems effortlessly support keyword
+%searches and other complex queries since the mechanism is realized locally, on
+%a node-by-node basis.
+%  \item Popular content is located at multiple peers and thus it is more likely
+%for a flooding-based search to return results. DHTs, on the other hand, fit
+%better in the systems which require ability to reliably locate content, even in
+%the extreme case that only a single-copy exists in the network.
 %\end{itemize}
@@
 % TODO: how 
-%For this reasons, efforts have been placed for optimizing the efficiency of decentralized unstructured peer-to-peer networks. Research mainly focuses on
+%For this reasons, efforts have been placed for optimizing the efficiency of
+%decentralized unstructured peer-to-peer networks. Research mainly focuses on
 %\begin{inparaenum}[\itshape i\upshape)]
@@
 %\end{inparaenum}
-%The goal can be achieved at, both, the application-level network as well as the underlying physical one.
-In the first case by refining the message relay techniques, while in the second one, by adaptively reconstructing the application network to map as well as possible to the the physical network.
+%The goal can be achieved at, both, the application-level network as well as the
+%underlying physical one. In the first case by refining the message relay
+%techniques, while in the second one, by adaptively reconstructing the
+%application network to map as well as possible to the the physical network.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{The Topology Mismatch Problem}
-
-% TODO: Define the problem
-% UNDER CONSTRUCTION
-% Q&A is it too late??
-
-%{\sethlcolor{yellow}\hl{ {\cite{hsiao_redblue_2009}} states that:
-%Liu {\cite{liu_thancs_2008}}  concluded that more than 70 percent of communication
-%paths in an unstructured overlay do not exploit the physical network topology
-%(i.e., the Internet topology), resulting in a topology mismatch between the
-%overlay and the underlying network.  } }
-
-One of the major issues that defines the efficiency of the overlay network is
-the mapping of the physical links and the overlay paths. Ideally, overlay
-networks are expected to achieve an optimal mapping among the overlay paths and
-the underlying physical links, and avoid the inefficient mapping that is shown
-in Figure \ref{figure:overlay}. The problem of constructing an optimal overlay
-is referred to as the \emph{topology mismatch problem}, and formally defined as
-follows:
-
-\begin{definition}
-    Let $V = \{v_1, ..., v_n\}$ be a set of points denoting the network nodes,
-    $\{v_i, v_j\} \in E$ be the set of unicast distances between nodes $v_i$ and
-    $v_j$, $G=(V,E)$ be a complete distance graph over $V$. The topology
-    mismatch problem is to construct a minimal spanning tree,  where 
-    node degree is restricted to a constant ($k\geq 2$) by the bandwidth of each node $v_i$.
-\end{definition}
-
-Constructing degree constrained spanning tree is known to be NP-Hard
-\cite{NPBOOK}, as well as the topology mismatch problem
-\cite{chawathe_scattercast_2000}. 
- Moreover, the internet is an interesting environment
-where end-to-end latencies demonstrate triangle inequality violations (TIVs),
-which further complicates the problem.
-These delay space TIV is a consequence of the Internet's structure and routing
-policies and thus will remain a property of the Internet for the foreseeable
-future \cite{zheng_irprtt_2005}. TIVs affect network coordinate
-\cite{cox_vivaldi_2004,wong_meridian_2005} and positioning \cite{ng_gnp_2001}
-and makes difficult the construction of delay aware overlays.  A number of
-heuristic approaches have been proposed in an effort to finding an algorithm
-that performs reasonably and consistently well.
-
-A topology unaware overlay network is able to control the sequence of peers a
-message traverses before reaching its destination, but it is completely unaware
-of how the actual packets are switched at the underlying infrastructure along
-the overlay path. For example, a single logical point-to-point link on the
-overlay, most of the time, corresponds to multiple physical links in the
-underlying layer. Additionally, a link in the underlying network often serves
-the mapping of several overlay paths causing increase of the traffic on the
-physical link, which is also called the link's
-\emph{stress}\cite{chu_esm_2002}.  Furthermore, the stochastic behaviour of
-peers randomly joining and leaving the peer-to-peer network all cause the
-overlay to physical mapping to create such unnecessary, redundant maintenance
-traffic that impacts the efficiency of the network in terms of average response
-time.
-
-%{\sethlcolor{yellow}\hl{
-%HA: Motivation to solve topology mismatch problem:
-%
-%Most of the traffic (find refs!! see Location aware topology matching paper
-%introduction for refs [31][29]) in the internet today is generated by P2P
-%networks. The inefficient use of the network resources by the P2P protocols,
-%especially the problem known as topology mismatch problem, causes unnecessary
-%traffic load and a heavy stress
-%on the internet infrastructure in general, and effects other users of the
-%internet than the P2P nodes.
-%}}
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.8]{img/phys.pdf}
-\caption{Interconnection of nodes in the physical level.}
-\label{figure:phys}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.8]{img/over1.pdf}
-  \hspace{3ex}
-  \vline
-  \hspace{3ex}
-  \includegraphics[scale=0.8]{img/over2.pdf}
-\caption{HA: Put subfigure, and a new caption here}
-\label{figure:over1}
-\end{figure}
-
-For example, assume nodes $A$, $B$, $C$ and $D$ are connected through the
-physical network shown in Figure~\ref{figure:phys}, where the network costs are
-given in milliseconds, and peer $A$ sends a message to peer $D$. If these peers
-participate in an overlay network according to one of the setups of
-Figures~\ref{figure:over1}~(left) and~\ref{figure:over1}~(right) then users will
-yield different performances. In the first overlay shown in Figure
-\ref{figure:over1}~(left), the message will traverse the following sequence of
-links in the physical layer (marked with their costs): $3 \rightarrow 7
-\rightarrow 1 \rightarrow 2 \rightarrow 2 \rightarrow 1 \rightarrow 7
-\rightarrow 8 \rightarrow 5 \rightarrow 5 \rightarrow 4$. In the second overlay
-shown in Figure ~\ref{figure:over1}~(right), the path will be: $3 \rightarrow 8
-\rightarrow 4$. The total cost of the first path is $45 ms$, while the second
-only costs a mere $15 ms$. Therefore, we can conclude that the second overlay is
-more congruent with the underlying physical network than the first one and thus
-more efficient.
-
-% TODO: Argue on the problem
-% UNDER CONSTRUCTION
-Early incarnations of the overlay protocols, however, did not make use of the optimal
-mapping with the physical network topology. Early Gnutella protocol, though were
-considered far from scalable \cite{ritter_gnucantscale_2001}, randomly choose
-logical neighbours without knowledge of the underlying network, effectively
-causing mismatch between physical and application-level topologies.
-Additionally, queries may be flooded to multiple paths merging in one peer while
-in such case one of the paths would have been enough. 
-%Nevertheless, peers may
-%forward the same message to each other before they receive the query messages
-%from each other.
-
-Similar problems are observed in decentralized structured schemes also.
-Typically, node IDs are assigned \emph{randomly}, resulting in excellent load
-balancing, scalability and robustness.  Unfortunately, this randomness has a
-negative impact on the \emph{routing locality} of the network. This means that
-even though the target node can be reached with logarithmic overlay hops, the
-distance traveled in the physical underlying network, during the overlay routing
-process, can be far from optimal.  The ratio of the physical distance, a query
-travels through the overlay to an object, and the minimal distance to that
-object (i.e. though IP) is known in the literature as \emph{stretch} or
-\emph{Relative Delay Penalty (RDP)} \cite{chu_esm_2000}.
-
-Studies on peer-to-peer traffic reveals the effect of the peer-to-peer traffic
-on the overall internet.  Measurements
-\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004} on some popular
-systems, such as FastTrack, Gnutella and DirectConnect have shown that the
-peer-to-peer traffic contributes the largest portion of the overall Internet
-traffic.  \cite{matei_mapgnutella_2002} showed that, even given that 95\% of any
-two nodes are less than 7 hops away from each other, a flooding-based algorithm,
-like the one used by Gnutella, can generate 330TB/month in a 50.000 node
-network!  Therefore, the efficiency of the overlay networks has serious
-consequences on the overall well-being of the modern internet.
+%\section{The Topology Mismatch Problem}
+
+%%%%%%%
+%%%%%%% TODO: NEEDS REWORDING
+%%%%%%% Additional Material
+%%%%%%%
+%%%%%%%
+%%%%%%%
+% P4P [16],instead, aimed to provide a general framework for P2P
+%applications by developing ISP-application interfaces so that
+%application software can use the network layer information,
+%such as congestion status and network topology, for better
+%performance and resource utilization.
+
+%P4P [37] project attempts to address
+%the problem through custom trackers, both for ISPs and P2P
+%systems, using an interface based on a primal-dual decomposition
+%of an optimization problem. This interface design simplifies
+%the realization of traffic-engineering objectives from each parties’
+%perspective and ensures the extensibility of the approach. Through
+%simulation-based studies and limited experimental deployments,
+%these collaborative approaches have been shown to effectively
+%reduce network costs while minimally impacting application performance.
+%A clear advantage of these proposals is that they
+%allow ISPs to incorporate aggregated traffic policies in their tracker
+%recommendations (e.g., a particular traffic balance ratio between
+%peering providers). However, all of them require deployments
+%of oracles for each participating ISP and their effectiveness is
+%ultimately predicated on their adoption by P2P applications and a
+%trust relationship between P2P users and their ISPs.
 
@@
 %
-%Caching based protocols are effectively used to reduce traffic costs and response
-%times. The caching policy varies depending on the way protocol handles the index and the
-%content. Centralized P2P systems
-%use central index servers, while local caching systems, such as KazaA, use super peers
+%Caching based protocols are effectively used to reduce traffic costs and
+%response times. The caching policy varies depending on the way protocol handles
+%the index and the content. Centralized P2P systems use central index servers,
+% while local caching systems, such as KazaA, use super peers
 %to cache indices in a distributed way. Content caching is also possible in P2P
@@
 %connection graphs, creating cluster of physically close nodes, or using latency
-%information to detect proximity. The brief description of each category is presented
-%below:
+%information to detect proximity. The brief description of each category is
+%presented below:
 %
@@
 %runtimes}}
-%%\textbf{Algorithm} & \textbf{Overlay structure} & \textbf{Forwarding} & \textbf{Cache} & \textbf{Overlay optimization} & \textbf{Proximity information} & \textbf{Base protocol} & \textbf{Dynamic update} & \textbf{Runtime} \\
+%%\textbf{Algorithm} & \textbf{Overlay structure} & \textbf{Forwarding} &
+%\textbf{Cache} & \textbf{Overlay optimization} & \textbf{Proximity information}
+%& \textbf{Base protocol} & \textbf{Dynamic update} & \textbf{Runtime} \\
 %
@@
 
-Having understood the nature of the problem that peer-to-peer architectures face,
-Section \ref{sec:unstructured} and \ref{sec:structured} discuss the recent academic work that has been conducted
-in the field. What follows does not claim
-to be a thorough citation of all known protocols that are available out there,
-but instead, a careful selection of those that left a distinctive fingerprint
-contribution in the efforts of the research community to alleviate the topology
-mismatch problem.
-
-%{\sethlcolor{yellow}\hl{
-%Create a theoretical section. This is an Approximation alg for Top-mis problem
-%
-%A Near-Optimal Algorithm Attacking
-%the Topology Mismatch Problem
-%in Unstructured Peer-to-Peer Networks
-%}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Unstructured Decentralized Algorithms}\label{sec:unstructured}
-
-In this section, the unstructured decentralized P2P algorithms are analysed based on their architecture and
-base methodology to tackle the topology mismatch problem. The
-unstructured decentralized algorithms are visited and categorized based on their
-use of the overlay structure, messaging between nodes, and on the techniques
-proposed to detect proximity to optimize the topology. 
-The base methods that unstructured P2P algorithms use to tackle the topology
-mismatch problem can be categorized as Broadcast, Cache, Overlay Optimization, 
-and Landmark Based. In this section, these methodologies are discussed in
-detail and the algorithms in each category are highlighted.
-
-%{\sethlcolor{yellow}\hl{
-%HA: TALK ABOUT GNUTELLA HERE!!!!}}
-
-\subsection{Broadcast Based}
-
-In the naive Gnutella protocol, each received query request is forwarded to all the
-neighbours, which clearly generates unnecessary traffic over the network. The
-algorithms categorized in the \textit{broadcast based} methodology prefers
-re-broadcasting the query messages to a selected subset of their neighbours,
-where the selection criteria varies between algorithms.
-The selection is made using one or a
-combination of various statistical metrics, such as the reliability or the latency
-of the link between nodes, etc.
-The forwarding based approach enhances search
-efficiency but has also the drawbacks of having drastically reduced search scope
-due to the selective re-broadcasting policy, and scalability problems on large
-P2P networks.
-%Expanding the search scope, on the other hand, is no easy task
-%because the overhead of forming multicast trees is proportional to the multicast group size.  
-%forwarding based schemes do not consider dynamic joining
-%and leaving of peers so they do not scale well on dynamic environments.
-
-The forwarding based P2P algorithms and their details are presented below.
-
-\paragraph*{\bf Improving search in peer-to-peer networks}
-
-\cite{yang_improvep2psearch_2002} proposes an easy to implement and practical
-solution to the inefficiency problem caused by blind flooding in Gnutella like
-P2P file sharing protocols. The paper replaces blind flooding with three
-approaches, namely \textit{iterative deepening}, \textit{directed BFS}, and
-\textit{local indices}. In \textit{iterative deepening}, the search is performed
-on a BFS tree with multiple preset depths. The depth limit is iteratively
-increased by the source node for each query based on the quality of the results.
-The source node, after examining the results, may issue a new request by
-increasing the depth limit, which will trigger the nodes at the last depth
-level to resume the search. The iterative approach avoids to start the search
-from scratch at each iteration and reduces the load of the nodes on the upper
-levels of the tree. The major drawback of \textit{iterative deepening} is the
-delay between successive iterations, as the source node needs to examine the
-results at each iteration before deciding to quit or resume the query.  The
-\textit{directed BFS} tries to avoid this delay by forwarding the query messages
-only to a selected set of neighbours, in which the selection criteria varies
-from the number of results received previously, distance in terms of hops,
-bandwidth, or the query load of the neighbour. In \textit{local indices}, each
-node sustains a local data index of the nodes within a radius of $r$ hops of
-itself and uses this local index to remotely query the neighbour nodes without
-generating additional traffic.  \textit{Local indices} greatly reduces the
-aggregate bandwidth usage of the network and improves query efficiency, however,
-updating the indices in cases with frequent node joins and leaves introduces a
-serious overhead to the system if the radius is kept broad.
-
-\paragraph*{\bf Gia}
-
-\emph{Gia} \cite{chawathe_gia_2003} is an effort to solve the scalability
-problem of the unstructured P2P file sharing systems, in particular Gnutella.
-The main novelty in the design of Gia is the replacement of blind flooding
-approach of Gnutella with random walks \cite{lv_randomwalks_2002}. Although
-random walks is a step in the right direction, issuing only a single copy of the
-query within the network reduces the search scope, thus affects the success rate
-of the query.  In order to overcome this limitation, Gia introduces a
-token-based flow control algorithm, which is essentially an intelligent flow
-control algorithm that gradually redirects the queries to nodes with more chance
-of answering. Gia also acknowledges the heterogenity in peer bandwidth,
-processing power, disk speed, etc, of the nodes in P2P networks and uses this
-heterogenity when connecting nodes to each other, such that by using a topology
-adaptation algorithm, Gia ensures that high capacity nodes have high degrees and
-low capacity nodes are within short proximity of high capacity nodes. In order
-to prevent overloading of nodes with query requests, Gia uses a token-based flow
-control algorithm in which each node announces the number of query requests it
-can handle in terms of tokens to its neighbours, so that neighbours only forward
-query requests to nodes that they previously received tokens from. Although the
-topology adaptation algorithm Gia uses improves the scalability of the network,
-it does not help much in solving the topology mismatch problem, since it does
-not consider the underlying physical topology.
-
-\paragraph*{\bf Distributed Cycle Minimization Protocol}
-
-The flooding based query approach of Gnutella creates many duplicate packets in
-the network. Nodes receive the same query request from various neighbours, which
-they generally discard when identified as duplicate. One other reason for
-duplicate messages is the cycles in the forwarding paths. These duplicate
-packets affect the performance of the P2P network negatively by increasing the
-overall resource usage. Moreover, the more active nodes, that have high
-capacity, high bandwidth, or contribute more to the network, suffer more from
-the overhead caused by the duplicate packets. Therefore, \emph{Distributed Cycle
-Minimization Protocol (DCMP)} \cite{zhu_dcmp_2008} is introduced as a method to
-remove the cycles and eliminate the duplicate packets, without sacrificing the
-connectivity. In DCMP, once a cycle is detected, the most powerful node in the
-cycle is elected as the \emph{Gate Peer} and the cycle is cut from a special
-link such that the distance of all the nodes within the cycle to the \emph{Gate
-Peer} is minimized. The process is managed by using two special message types,
-\emph{Information Collection (IC)} and \emph{Cut Message (CM)}. One disadvantage
-of DCMP is that since the distance a forwarded message can travel is limited
-with the TTL value, which is practically $7$ in Gnutella, the DCMP cannot detect
-cycles of length larger than $7$. Even though cycle elimination improves the
-network performance, it does not solve the topology mismatch problem.
-
-%\subsection{Cache Based}
-%
-%Caching based protocols are effectively used to reduce traffic costs and response
-%times. The caching policy varies depending on the way protocol handles the index and the
-%content. Centralized P2P systems
-%use central index servers, while local caching systems, such as KazaA, use super peers
-%to cache indices in a distributed way. Content caching is also possible in P2P
-%systems, where nodes cache the forwarded content for further retrievals.
-%Although caching has the above mentioned advantages,  duplication
-%of messages still exist, which limits the scalability of these approaches.
-%Therefore, cache based approaches are analyzed in the following categories:
+\input{unstructured}
+
+\input{structured}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\section{The Algorithms}
+%Having understood the nature of the problem that peer-to-peer architecture face, this section starts the discussion of the academic work that has been conducted in the field the last few years.
+What follows this introduction, does not claim to be a thorough citation of all known protocols that are available out there, but instead, a carefull selection of those that left a distinctive fingerprint contribution in the efforts of the research community to alleviate the topology mismatch problem.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% UNSTRUCTURED
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%\subsection{Narada}
+%
+%\paragraph*{}
+%\emph{End System Multicast} \cite{chu_esm_2000,chu_esm_2001,chu_esm_2002} as well as other  proposes an architecture, where end-systems implement all multicast related functionality (including membership management and packet replication).
+In \emph{Narada}, the protocol developed in the paper's context, end-systems
+%\begin{inparaenum}[\itshape i\upshape)]
+%  \item self-organize, in a fully distributed manner,
+%  \item into an efficient overlay structure, both from the network and the application perspective,
+%  \item with self improving properties,
+%  \item adapting to network dynamics\footnote{Long-term variations in Internet path characteristics (i.e. bandwidth, latency e.t.c.)}.
+%\end{inparaenum}
+%
+%\paragraph*{}
+%What Narada really does is to construct overlay spanning trees in a two-step process:
+%\begin{enumerate}
+%  \item It builds a richer connected graph \footnote{ENarada \cite{li_enarada_2008} used Gossip protocol for the construction}, termed a \emph{mesh}, while trying to ensure a couple of desired performance properties:
 %    \begin{itemize}
-%    \item \emph{data index caching},
-%    \item \emph{content index caching},
-%    \item \emph{centralized}, and
-%    \item \emph{local}.
+%      \item path quality between any pair of members should be comparable to the quality of the unicast path between them, and
+%      \item each member should have a limited number of neighbours on the mesh.
 %    \end{itemize}
 %
-
-\subsection{Cache Based}
-Caching is a widely used methodology to exploit the locality to minimize the
-redundant transfer of data, which is successfully used in web servers and file servers.
-As peers in a P2P system also work as servers, intuitively it is expected that
-P2P file sharing systems can also benefit from caching. However, one of the most
-important characteristic properties of P2P systems is that nodes
-join and leave frequently, and the lifetime of a query is relatively short
-compared to web servers, which makes caching in P2P networks nontrivial. There are
-usually two levels of caching possible in file sharing systems, the indices or
-pointers to data can be cached, or the data itself can be cached. Caching is
-already implemented and successfully used in some commercial P2P systems. One
-popular example is KazaA, which uses caching of indices in its super peers.
-We present the other protocols that use caching and their details below.
-
-\paragraph*{\bf Replication Strategies in Unstructured Peer-to-Peer Networks}
-
-\cite{Cohen02} aim to improve the inefficient blind search algorithm by
-replicating the data in a peer to peer network. The main intuition behind the
-idea of replication, or using cached copies, is that as the number of copies for
-each item increases in the network, it would be easier for a search algorithm,
-even a random one, to locate these items. In order to analyse the feasibility of
-such replication approach, the authors investigate three different replication
-strategies, namely uniform, proportional, and square root allocation. In the
-uniform model, the copies of items are uniformly replicated in the network,
-while in the proportional model, the items are replicated based on their query
-rate, so that frequently queried items are replicated more. The square root
-allocation is a strategy proposed by the authors, which is a model between the
-uniform and proportional allocation.  In order to evaluate the outcomes of the
-replication models, the authors compare the overall costs of successful and
-unsuccessful searches in the network. The results can be summarized as follows;
-the uniform allocation model minimizes the maximum search size, therefore
-reduces the time spent on an unsuccessful search. The proportional model, on
-the other hand, promotes the more frequently queried items by replicating them
-more, therefore decreases the search time for popular items, but suffers when
-locating the rare items. The authors also claim that the expected successful
-search size is the same for uniform and proportional models, and any approach
-between them would behave much better. Therefore they propose the square root
-allocation approach, which is a replication strategy that minimizes the expected
-search size of successful queries in P2P networks.
-
-\paragraph*{\bf Tracing a large-scale Peer to Peer System: an hour in the life of
-Gnutella}
-
-\cite{Markatos02} analysis the network traces of the Gnutella network and
-detects that query requests in Gnutella has locality, and proposes a caching
-algorithm that exploits this locality. The analysis of the trace data reveals
-other important facts about the structure of the Gnutella network and the query
-data.  One significant observation is that the geographic locations of clients
-do not have a correlation with the number of query requests they receive. This
-is an obvious result of the topology mismatch problem caused by the overlay
-structure of the Gnutella network. Gnutella traffic is observed to be bursty
-both for query requests and query responses, even in longer intervals, and nine
-out of ten queries do not generate any response due to the inefficient design of
-the Gnutella network. When developing a caching system to exploit locality,
-applying an approach similar to web caching does not fit well with P2P systems.
-The caches in P2P systems not only have to consider the query string, but also
-the TTL value, the source of the query, and the time of the query as well.  In
-general, even though optimum caching is hard to achieve, it is reported that
-caching improves the overall performance of the Gnutella system.
-
-\subsection{Overlay Optimization Based}
-
-The overlay optimization based protocols modify the topology of the P2P network
-using various techniques. The two commonly used approach that is investigated in
-this section are creating spanning trees using 
-connection graphs, and creating cluster of physically close nodes. 
-%The brief description of each category is presented below:
-Spanning tree based approaches construct rich graphs based on the network
-connections and build minimum spanning trees on the graphs. Even though the
-spanning tree provides efficient query performance, the construction
-and the update costs of the spanning tree, especially in dynamic environments
-with many nodes joining and leaving the network, cause large traffic
-overhead to the system \cite{chu_esm_2000,chu_esm_2002}.
-
-The cluster based approaches on the other hand select to link physically closer
-nodes with each other, therefore shrink the search scope significantly. However,
-commonly used methods for determining the correct physical positions of nodes
-over the internet does not always return reliable results, therefore mapping
-accuracy is not always guaranteed.
-
-Below, the details of the algorithms that use the spanning tree based or cluster
-based topology optimization methods are presented.
-
-\paragraph*{\bf Narada}
-
-Narada \cite{chu_esm_2000,chu_esm_2001,chu_esm_2002} is a generic protocol to
-design self adapting overlay networks, that can achieve application layer
-multicast without requiring IP multicast at the network layer. Although Narada
-is not originally designed as a peer-to-peer file sharing application protocol,
-it has become one of the pioneering works in the overlay networks area.  Even
-though the design of the IP multicast protocol is finalized a while ago, and
-some of the routers on the internet today already support this protocol, the IP
-multicast idea did not take off as anticipated. One of the main reasons for this
-delay is that the IP multicast violates the stateless design of the current
-internet. Although the multicast is not widely used, and the future of the
-protocol is not bright, there are multiple applications that need this kind of
-service, such as IP TVs, P2P networks, streaming services, etc. The application
-layer multicast is then a valid alternative for these applications.  The authors
-claim that, although Narada is not as efficient as IP multicast, still it
-provides reasonable performance, without the additional cost of implementing
-multicast services on the current internet infrastructure. The main reason for
-the inefficiency is due to the Topology mismatch problem.  Narada tries to solve
-this problem by building a richer connected graph\footnote{ENarada
-\cite{li_enarada_2008} used Gossip protocol for the construction}, called a
-mesh, and building per source minimum spanning trees.  Narada also maintains the
-graph and the trees dynamically updated on node joins and leaves. Moreover, Narada tries
-to optimize the physical link stress, the overall resource usage and the
-relative delay among end systems. The main limitation of Narada is that although
-it works reasonably well for small groups, it does not scale well for larger
-networks, therefore it is not suitable for P2P file sharing applications which
-can easily scale to millions of nodes.
-
-\paragraph*{\bf Adaptive Overlay Topology Optimization and Adaptive Connection
-Establishment}
-
-\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{liu_auto_2003} is one
-of the first attempts, along with Narada, from the P2P research community to
-address the topology mismatch problem. AOTO is a distributed algorithm
-that uses \emph{Selective Flooding} and \emph{Active Topology} to optimize the
-overlay network topology. The \emph{Selective Flooding} algorithm creates a minimum
-spanning tree over the overlay and uses this tree instead of
-flooding the whole network, without shrinking the search scope. 
-In \emph{Active Topology}, the physical locations of nodes are estimated based
-on the network delay among them and physically close nodes are connected as
-neighbours to revise the overlay topology as close as possible to the physical
-network topology. Overall, AOTO reports a $55\%$ performance improvement in
-terms of the traffic load. In \emph{Adaptive Connection Establishment (ACE)}
-\cite{liu_ace_2004}, the authors extend the idea by introducing optimizations
-based on the depths of the minimum spanning trees. As the network delay is not always a reliable estimation
-method to detect physical locations of the nodes, the algorithm still suffers
-from the discrepancies caused by mislocated nodes.
-
-\paragraph*{\bf Location-aware Topology Matching}
-
-Peer-to-peer algorithms designed without considering the topology mismatch
-problem generates unnecessary heavy traffic on the internet infrastructure, due
-to the popular use of these P2P programs for file sharing. \emph{Location-aware
-topology matching (LTM)} \cite{liu_ltm_2004} proposes a method to
-optimize the overlay structure of the P2P network based on the physical
-topology. In doing so, LTM issues a special message called
-\textit{TTL2-detector} to detect the latency of $N^2$ neighbours around each
-node. The latency information gathered is later used to evolve the overlay
-network into a more efficient one, without reducing the search scope. Each node
-examines the latency information with the neighbours and low productive
-connections are dropped and replaced by more efficient ones, thus reducing the
-latency on the overall network, and eliminating some potential inefficiencies,
-such as unnecessary crossing of messages over autonomous system (AS) boundaries, etc. Although the
-LTM approach improves the overall efficiency of the P2P network, since it uses
-the latency to detect close by nodes, does not use the real physical topology
-information, therefore does not offer a guaranteed bound to the
-topology mismatch problem.
-
-\paragraph*{\bf Scalable Bipartite Overlay}
-
-\emph{Scalable Bipartite Overlay (SBO)}
-\cite{liu_bipartite_IPDPS,liu_bipartite_2007} improves the overhead of creating
-and maintaining a minimum spanning tree cost by randomly dividing the nodes into
-two groups, red and white, and assigning different tasks to different groups.
-The white peers measure distances to neighbours by using the network delay as
-the metric and reports the results to red peers. The red peers, equipped with the
-information of all two hop ($N^2$) neighbours, creates a minimum spanning tree
-of these neighbours and assigns efficient forwarding paths. The white peers can
-further optimize their positions within the overlay if need be.
-
-%\paragraph*{\textbf{Two-Hop-Away Neighbour Comparison and Selection\\}}
+%  \item It computes minimum spanning trees out of the mesh, each rooted at the corresponding source member.
+%\end{enumerate}
+%
+%\paragraph*{}
+%Narada incrementally improves the quality of the mesh by adding and dropping overlay links.
+Members probe each other at random and new links are added depending on what is the gain in doing so.
+On the other hand, participants, continuously  monitor the utility of existing links, and drop them if they are found to be not usefull.
+%
+%\paragraph*{}
+%Narada protocol achieves relatively good performance for small and medium sized groups involving tens to hundreds of members.
+This is not, by no means, what a contemporary peer-to-peer system needs to be characterized as scalable, but this work is considered pioneering\footnote{Another similar approach is \emph{Scattercast}\cite{chawathe_scattercast_2000}}, in that it was the first to have conducted a detailed Internet evaluation to analyse and demonstrate the feasibility of an overlay-based, application layer, multicast architecture that can dynamically adapt to bandwidth and latency properties of the physical underlying infrastructure.
+%
+%\subsection{Gia}
+%
+%\paragraph*{}
+%\emph{Gia} \cite{chawathe_gia_2003} is a decentralized unstructured peer-to-peer file sharing system with the following characteristics:
+%\begin{inparaenum}[\itshape i\upshape)]
+%  \item replaces Gnutella's flooding with random walks \cite{lv_randomwalks_2002},
+%  \item recognizes the importance of overlay network's topology, while using the random walks, and exploits it by incorporating a topology adaptation algorithm,
+%  \item introduces a token-based flow control mechanism, and
+%  \item detects the significant heterogeneity in peer bandwidth, processing power, disk speed e.t.c. and takes advantage of it.
+%\end{inparaenum}
+%
+%\paragraph*{}
+%More specifically, there are four key components in the design of \emph{Gia} which are summarized bellow:
+%\begin{enumerate}
+%  \item A \emph{dynamic topology adaptation} protocol that puts participating nodes within short reach of high capacity nodes so that these \emph{high-degree} nodes, which due to their high connectivity receive most of the queries, actually have the capacity to handle them.
+%  \item An \emph{active flow control} scheme is used to avoid overloaded hot-spots. Heterogeneity is detected and flow control tokens are given to nodes based on the available capacity.
+%  \item Every node maintains pointers to to the content that is offered by their immediate neighbours, creating a \emph{one-hop replication} of pointers, scheme.
+%  \item A \emph{search protocol} based on random walks, that is biased towards directing queries to high-capacity nodes that are typically best able to answer these queries.
+%\end{enumerate}
+%
+%\subsection{Adaptive Overlay Topology Optimization}
+%
+%\paragraph*{}
+%\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{liu_auto_2003} is an algorithm for building an overlay multicast tree among each source node and its direct logical neighbours in order to alleviate the mismatching problem while providing a larger query coverage range.
+It includes the steps of
+%\begin{inparaenum}[\itshape i\upshape)]
+%  \item \emph{Selective Flooding (SF)}, and
+%  \item \emph{Active Topology (AT)}
+%\end{inparaenum}
+%which are furtherlly discussed in the following paragraphs.
+%
+%\paragraph*{Selective Flooding (SF)}
+%Instead of flooding to all neighbours, SF builds an overlay multicast tree among each peer and its immediate logical neighbours.
+This reduces overall message burden to the network by making some neighbours non-flooding.
+The tree is formed using a minimum spanning tree algorithm.
+For this reason a peer needs to know the costs to all its logical neighbours (e.g.
+the network delay) as well as between any pair of neighbours.
+Additional, cost probing, messages need to be added to the original messages defined by the Gnutella protocol.
+%
+%LTM's SF effectiveness has be proven to be detached from the different physical or overlay topologies.
+On the other hand, SF is more effective with large number of logical neighbours.
+It can reach an average optimization rate of 87.4 percent on a logical topology with an average of 30 logical neighbours.
+%
+%\paragraph*{Active Topology (AT)}
+%At this step the overlay topology is reorganized.
+Each peer, independently, makes optimizations in the overlay network to alleviate topology mismatching by replacing non-flooding neighbours, with closer nodes.
+The approach incorporated by the paper is called \emph{Randomized AT} algorithm which picks up a candidate peer at random among the non-flooding neighbour's neighbours.
+Whenever a new neighbour cost table is received or there is a change of neighbours, the source peer has to re-calculate the multicast tree and apply the randomized AT algorithm.
+%
+%Different numbers of average logical neighbours has little to do with the effectiveness of AT.
+If the source has $n$ non-flooding peers, there are $n$ potential neighbour replacements.
+The overhead to exhaust all $n$ possible replacements can be high, so in practice, after each replacement the source peer can decide whether it needs to find another candidate peer.
+This is done by computing the cost improvement ratio greater than some predefined termination threshold.
+The larger the threshold, the slower, in the number of optimization steps, the reduction of the normalized average distance.
+As a whole the average response time is significantly reduced when more optimization steps taken.
+%
+%\subsection{Location-aware Topology Matching}
+%
+%\paragraph*{}
+%\emph{Location-aware topology matching} or \emph{LTM} \cite{liu_ltm_2004}, is an algorithm of building an efficient overlay by disconnecting low productive connections and choosing physically closer nodes as logical neighbours while still retaining the search scope and reducing response time of queries.
+This is done by issuing detector messages by peers, to a small region so that receivers of the detector can record relative delay information and cut inefficient or redundant logical links as well as add closer nodes as direct neighbours.
+Three operations are defined in LTMband are furtherly discussed in the following paragraphs.
+%
+%\paragraph*{TTL2-detector flooding}
+%\emph{TTL2-detector} messages are designed based on the specification of the Gnutella protocol.
+Suppose that $d(i, S, v)$ denotes the TTL2-detector which has the message ID of $i$ with TTL value of $v$ and is initiated by $S$.
+In addition to Gnutella's unified 23-byte header, such a message has a body of two types:
+%\begin{itemize}
+%  \item short format $d(i, S, 1)$ if $S$ is a direct neighbour of the receiver
+%  \item long format $d(i, S, 0)$ if $S$ is within a two-hop distance from the receiver
+%\end{itemize}
+%This way a peer can calculate link costs from within a one- or two-hop area around it and thus reach peers in $N(S)$ and $N^2(S)$ sets, respectively, in the paper's context.
+Each peer floods a TTL-detector periodically while it calculates link cost to another peer by receiving other peer's detector messages and checking:
+%\begin{itemize}
+%  \item the Source Timestamp field and its own clock upon time of receipt for one-hop messages
+%  \item the Source Timestamp, TTL1 Timestamp and its own clock upon receipt for two-hop messages
+%\end{itemize}
+%Of course peer clocks must be synchronized in order to do such checking.
+%
+%\paragraph*{Low productive connection cutting}
+%There are three cases for any peer $P$ who receives $d(i, S, v)$ multiple times:
+%\begin{inparaenum}[\itshape i\upshape)]
+%  \item $P$ receives both $d(i, S, 1)$ and $d(i, S, 0)$
+%  \item $P$ receives multiple $d(i, S, 0)$s from different paths, and P randomly chooses to process one
+%  \item $P$ receives one $d(i, S, 1)$ and multiple $d(i, S, 0)$s, and $P$ processes $d(i, S, 1)$ and one randomly selected $d(i, S, 0)$
+%\end{inparaenum}
+%If the link with the largest cost is found and is a direct neighbour then the connection is put in a will-cut list and stays there for a certain period of time.
+If it is not, then it is handled by other peers.
+After that period, connections are cut and recorded to $P$'s cut-list.
+%
+%\paragraph*{Source probing}
+%For a peer $P\in(N^2(S) - N(S))$ who receives only one $d(i, S, 0)$, the cost of $PS$ is obtained (with list look-up or probing).
+Then $P$ compares it with the cost from each hop and if $PS$ has the largest cost, $P$ will not keep this connection, while otherwise the connection will be created.
+%
+%\paragraph{}
+%Supposing $n$ is the number of peers, $c_n$ is the average number of neighbours and $c_e$ is the average cost of logical links, then in the flooding-based search the traffic incurred by one query from an arbitrary peer in a peer-to-peer network is $O(n)$.
+As observed in the Gnutella network \cite{sripanidkulchai_gnutella_2001}, each peer issues $0.3$ queries per minute in average, thus the per minute traffic incurred by the network with $n$ peers is $O(n^2)$.
+Because each $d(i, S, v)$ has a TTL of $2$ in each source peer, the traffic for one time LTM optimization in all peers is at most $2nc_n^2c_e$.
+If each peer uses LTM $k$ times per minute, the total traffic incurred is $2knc_n^2c_e$.
+Simulation shows the best value for $k$ is $2$ or $3$.
+So, the traffic overhead caused by LTM to the network is $O(n)$.
+%
+%TTLj-detectors, with $j > 2$, would detect and break cycles with more than 4 links.
+LTM though, does not use such detectors because detector-flood traffic would increase significantly, and cut links between two end-peers, could cause queries initiated by them to traverse a path much more expensive than the cost on the the cut link.
+%
+%\paragraph{}
+%LTM disadvantages are
+%\begin{inparaenum}[\itshape i\upshape)]
+%  \item disagreement of measured delay due to unsynchronized clocks causes problems when deciding the cut positions, which can influence the network connectivity, and
+%  \item the network delay metric mainly focuses on disabling the connections between peers physically far away without considering the shortcuts created by powerful peers.
+%\end{inparaenum}
+%
+%\begin{figure}
+%\centering
+%  \includegraphics[scale=0.4]{img/ltm_multid.jpeg}
+%\caption{Peer $P$ receives $d(i, S, v)$ multiple times}
+%\label{figure:ltm_multid}
+%\end{figure}
+%
+%\begin{figure}
+%\centering
+%  \includegraphics[scale=0.4]{img/ltm_example.jpeg}
+%\caption{A full example of LTM}
+%\label{figure:ltm_example}
+%\end{figure}
+%
+%\subsection{Scalable Bipartite Overlay}
+%\emph{Scalable Bipartite Overlay (SBO)} \cite{liu_bipartite_2007} employs an efficient strategy to select query forwarding paths and logical neighbours.
+The topology construction and optimization of SBO consist of four phases:
+%\begin{enumerate}[\itshape i\upshape)]
+%  \item When a new peer is joining the network, it will randomly take an initial colour; say red or white.
+This colour is not to be changed for the time the peer stays connected.
+If it leaves and then rejoins it will, again, have to pick a colour, randomly.
+Thus all peers are separated into two groups, red or white.
+Then the bootstrap host will provide, the joining peer, with a list of active peers, including their colour information, in order for the later to establish connections to different colour peers.
+This way, all peers form a bipartite overlay.
+%  \item White peers probe distances with their immediate (red) neighbours, form a cost table and send this table to their corresponding red neighbours.
+%  \item Each red peer builds a minimum spanning tree using the obtained neighbour cost tables.
+These are two-hop diameter trees so white peers do not need to build one.
+The links that are part of a minimum spanning tree are called \emph{forwarding connections (FC)} while the rest \emph{non-forwarding (NFC)}.
+%  \item Having a minimum spanning tree with two hops a red peer is able to send its queries within that range.
+Some white peers, though, have become non-forwarding neighbours.
+In this phase, such a (white) neighbour will try to find another red peer being two hops away from its current red neighbour to replace the later as its new neighbour.
+%\end{enumerate}
+%
+%\begin{figure}
+%\centering
+%  \includegraphics[scale=0.4]{img/sbo_efficient_forward.jpeg}
+%\caption{A red peer $P$ has a small overlay topology of $N(P)$ and $N^2(P)$ and computes the efficient forwarding paths}
+%\label{figure:sbo_efficient_forward}
+%\end{figure}
+%
+%\begin{figure}
+%\centering
+%  \includegraphics[scale=0.4]{img/sbo_neighbour_replace.jpeg}
+%\caption{Neighbor replacement}
+%\label{figure:sbo_neighbour_replace}
+%\end{figure}
+%
+%In a static enviromnent LTM may reduce traffic cost by around 80 to 85 percent while SBO reduces traffic cost between 85 and 90 percent.
+However, LTM  is proved to converge in around 2-3 steps while SBO needs 4-5 steps.
+Moreover LTM reduces response time by more than 60 percent in 3 steps while SBO needs 8.
+In a dynamic environment (10 minute average peer lifetime, 0.3 queries/sec by each peer) SBO and LTM reduce the average traffic cost per query (including the overhead due to the optimization steps) by 85 and 80 percent, respectively.
+Moreover LTM reduces the response time per query to 30 percent while SBO to 35 percent.
+%
+%\subsection{Adaptive Connection Establishment}
+%
+%\paragraph{}
+%\emph{Adaptive Connection Establishment (ACE)} \cite{liu_ace_2004} builds an overlay multicast tree among each source node and the peers within a certain diameter from the source peer and optimizes the neighbour connections that are not in that tree.
+%
+%\paragraph{}
+%ACE indicates three phases for its algorithm:
+%\begin{enumerate}[\itshape i\upshape)]
+%  \item Calculate cost between nodes using network delay as a metric.
+Each peer probes the costs with its immediate logical neighbours and forms a \emph{neighbour cost table (NCT)} using a special routing message type.
+Two neighbouring peers exchange their NCTs in order for every peer to obtain the cost between any pair of its local neighbours forming a small overlay topology.
+%  \item Based on obtained NCTs a minimum spanning tree among each peer and its immediate neighbours is built (Figure~\ref{figure:ace_phase2}).
+%  \item Physically far away neighbours are replaced by physically close neighbours.
+In ACE a peer, say $S$, probes the distance between one of its non-flooding neighbour's neighbour, say $G$ and $H$ respectively.
+If the link to neighbour's neighbour is smaller than that to the neighbour, the later is cut.
+If this is not the case but $S$ finds that cost of $GH$ is even larger than that of the $SH$, $S$ will keep $H$ as a new neighbour.
+Obviously, if $SH$ is larger than $SG$ and $GH$, the connection will not be established and $S$ will continue probing another neighbour's neighbour.
+The algorithm described is conducted within $1$-neighbour closure (among its source peer and all its direct neighbours) but the optimization scope can be enlarged.
+The larger the scope, the better topology matching improvement but also the greater the computational overhead (Figure~\ref{figure:ace_phase3}).
+%\end{enumerate}
+%
+%\paragraph{}
+%Simulations in \cite{liu_acesims_2004} show that the average scope of each query to cover the same scope of nodes is reduced by about 65 percent without losing any autonomy feature, while the average response time can be reduced by 35 percent.
+Larger diameter topologies lead to better topology optimization rate but also to higher communication and computation overhead.
+It was also found that it is more effective in higher connectivity dense topologies.
+Compared to LTM, it comes short of convergence speed.
+In \cite{ni_mismatch_2004} shows reduction of both total traffic (90 percent) and response time (80 percent) to message queries without shrinking the search scope.
+SBO, on the other hand, achieves approximately 85 percent reduction on traffic cost and about 60 percent reduction on query response time.
+Last but not least, it is concluded that work must be done on incorporating a more sophisticated selection policy for candidate non-flooding peers.
+%
+%\begin{figure}
+%\centering
+%  \includegraphics[scale=0.4]{img/ace_phase2.jpeg}
+%\caption{Second phase in ACE}
+%\label{figure:ace_phase2}
+%\end{figure}
+%
+%\begin{figure}
+%\centering
+%  \includegraphics[scale=0.4]{img/ace_phase3.jpeg}
+%\caption{Third phase in ACE}
+%\label{figure:ace_phase3}
+%\end{figure}
+%
+%\subsection{Hops Adaptive Neighbour Discovery}
+%
+%\paragraph{}
+%Paper \cite{chen_hand_2006} proposes a new algorithm named \emph{Hops Adaptive Neighbour Discovery (HAND)} which uses a fully distributed triple hop adjustment strategy to address the topology mismatch problem.
+The advantages of the algorithm compared to other approaches are that
+%\begin{inparaenum}[\itshape i\upshape)]
+%  \item it does not need any clock synchronization,
+%  \item it is a fully distributed algorithm making it robust and reliable in decentralized systems,
+%  \item the traffic overhead of the triple hop adjustment is very low,
+%  \item it is applicable to dynamic peer-to-peer environments, and
+%  \item maintains lower query response time.
+%\end{inparaenum}
+%
+%\paragraph{}
+%The algorithm's ultimate goal is the optimal overlay, the \emph{Logical Communication Network (LCN)} in the paper's context.
+The key concept of the algorithm is that a graph $G^{*}$ that describes an LCN and a graph $G$ that describes the current overlay are matched only if all peer hop sequences $(v_1, v_2, \ldots, v_k)$ in $G$ exist in $G^{*}$ and in the same order.
+In practice triple sequences $(v_1, v_2, v_3)$ are used.
+%
+%\paragraph{}
+%The mismatching detection is done in the following way.
+Suppose we want to verify peer sequence $v_2-v_1-v_3$ (see Figure~\ref{figure:hand_com_overlay}).
+A pair of probing messages are sent from $v_1$ to $v_2$ and $v_3$.
+Suppose delays of $(v_1,v_2)$ and $(v_1,v_3)$ are are denoted as $x$ and $z$, respectively.
+When the probing message arrives to $v_2$ it forwards it directly to $v_3$.
+Similarly, when the probing message arrives to $v_3$, it forwards it directly to $v_2$.
+These last steps are performed in order to obtain delays of $(v_2,v_3)$ and $(v_3,v_2)$ physical paths, respectively, denoted by $y$.
+%\begin{itemize}
+%  \item If $y=z-x\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched and should be adjusted to $v_1-v_2-v_3$ by deleting edge $(v_1,v_3)$ and adding a new $(v_2,v_3)$ (see Figure~\ref{figure:hand_matchtriple_lemma3}).
+%  \item if $y=x-z\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched and should be adjusted to $v_1-v_3-v_2$ by deleting edge $(v_1,v_2)$ and adding a new $(v_3,v_2)$ (see Figure~\ref{figure:hand_matchtriple_lemma4}).
+%\end{itemize}
+%$\varepsilon$ is a small positive real number denoting additional delays caused by possible forwarding and jitter delays.
+%
+%\begin{figure}
+%\centering
+%  \includegraphics[scale=0.4]{img/hand_com_overlay.jpeg}
+%\caption{The communication in overlay}
+%\label{figure:hand_com_overlay}
+%\end{figure}
+%
+%\begin{figure}
+%\centering
+%  \includegraphics[scale=0.4]{img/hand_matchtriple_lemma3.jpeg}
+%\caption{The matching triple in paper's Lemma 3}
+%\label{figure:hand_matchtriple_lemma3}
+%\end{figure}
+%
+%\begin{figure}
+%\centering
+%  \includegraphics[scale=0.4]{img/hand_matchtriple_lemma4.jpeg}
+%\caption{The matching triple in paper's Lemma 4}
+%\label{figure:hand_matchtriple_lemma4}
+%\end{figure}
+%
+%\paragraph{}
+%Measurements conducted for evaluation perposes showed that in a static environment the algorithm can effectively decrease traffic cost by about 77 percent and shorten the query response time by about 49 percent in less than two minutes.
+In a dynamic environment it shows similar behaviour and with the size of the overlay network having little impact on the effectiveness of the algorithm.
+Compared to LTM both algorithms have almost the same traffic reduction rate, however on the response time reduction rate HAND has a higher one by about 4 percent.
+The traffic overhead of HAND is much less than that of LTM by an average of 55 percent.
+%
+%\subsection{Distributed Cycle Minimization Protocol}
+%
+%\paragraph{}
+%\emph{Distributed Cycle Minimization Protocol (DCMP)} which is introduced in \cite{zhu_dcmp_2008} is a dynamic, fully decentralized protocol that promises significant reduction of duplicate messages.
+To achieve that, it uses the slightly different approach of eliminating unnecessary cycles, while retaining the connectivity of the network and preserving fault resilience and load balancing properties of unstructured peer-to-peer schemas by avoiding the creation of a hierarchical organization.
+%
+%DCMP aims at cutting the cycle paths at strategic locations. Any peer that detects a duplicate message can initiate the cutting process which consists of the following steps:
+%\begin{enumerate}
+%  \item Peers in the cycle elect a leader called \emph{GatePeer}\footnote{GatePeers are important for maintaining the connectivity and optimal structure of the network while peers join and leave randomly.}.
+%  \item The cycle is cut at a well-defind point with respect to the GatePeer
+%\end{enumerate}
+%
+%\paragraph{}
+%The first step after detecting a duplicate message by some peer is to gather information from all peers in the cycle using a new type of control message called \emph{Information Collecting Message} or \emph{ICM}.
+ICM contains:
+%\begin{inparaenum}[\itshape i\upshape)]
+%  \item a \emph{GUID}\footnote{Globally Unique IDentifier assigned to every query message generated by any node.} field same as the one of the duplicate message,
+%  \item \emph{DetectionID} field which represents the direction of the connection where the duplicate was identified\footnote{This ensures the uniqueness of the ICM messages because as it travels through many cyclic paths, multiple peers will detect the duplicates and initiate an ICM message.}, and
+%  \item \emph{Node Information Vector (NIV)} which contains information (bandwidth, CPU power, etc) about peers that propagated the ICM.
+%\end{inparaenum}
+%
+%\paragraph{}
+%Suppose $A$ detected the duplicate as depicted in Figure~\ref{figure:dcmp}.
+It then emits an ICM to $B$ and $F$, that initially contains information only about itself.
+Each peer that receives the ICM, appends its information and propagates it along the reverse path of the original message.
+Since two copies of ICM are sent, at some point, a peer, say peer $D$, will receive a duplicate ICM.
+Using the information in the NIVs of the ICMs, $D$, decides to cut (for example) the EF connection.
+To inform the other peers about its decision, it emits a \emph{Cut Message (CM)} which contains the GUID and DetectionID of the corresponding ICM and an additional field that identifies the connection to be cut.
+$D$, then, forwards the CM in the reverse directions from where the ICM arrived.
+Similarly CMs received by any peer are propagated toward the reverse path of the corresponding ICM.
+Eventually, either peer $E$ or peer $F$ will receive the CM and cut the connection, thus eliminating the cycle.
+%
+%Receiving a duplicate ICM denotes the existence of a cycle.
+The opposite is not true though.
+For example, if the cycle contains $2 \times TTL$ edges, it will not be detected, because the ICM messages will be discarded before they locate it.
+There is a tradeoff between preserving the connectivity of the network and minimizing the duplicates that makes such a possibility to be safely ignored.
+%
+%\begin{figure}
+%\centering
+%  \includegraphics[scale=0.4]{img/dcmp.jpeg}
+%\caption{Cycle elimination methods in DCMP}
+%\label{figure:dcmp}
+%\end{figure}
+%
+%\paragraph{}
+%Experiments in \cite{zhu_dcmp_2008} showed that DCMP incurs a lower delay, returns more results and decreases the number of duplicate messages by 22\%, compared to LTM.
+Additionally DCMP has one to two orders of magnitude less overhead, because it adopts a more ``LAZY'' approach than broadcasting control messages periodically like LTM does.
+%
+%\subsection{Two-Hop-Away Neighbour Comparison and Selection}
 %
@@
 This design, however, has the limitation of not being easily extend to also support non-flooding-based systems.
-
-\paragraph*{\bf Peer-exchange Routing Optimization Protocols}
-
-\cite{qiu_prop_2007} introduces two protocols called Peer-exchange Routing
-Optimization Protocols (PROP) to adjust the neighbourhood graph of the overlay
-network in order to reduce the overall link latency of the network. The PROP
-algorithms are based on the exchange of neighbours among peers, which is
-triggered by the mutual benefit of both peers to reduce the network delay.  In
-PROP-G (Generic), peers exchange all their neighbours with another peer, while
-in PROP-O (Optimized) only selected number of neighbours are exchanged among
-peers. The PROP-G is a generic protocol that guarantees the connectivity of the
-overlay graph during exchanges, therefore can be applied to both unstructured
-and structured overlay networks.
-
-\paragraph*{\bf T2MC}
-
-T2MC \cite{shi_t2mc_2008} uses traceroute logs to detect the AS boundaries and
-cluster close by peers with each other to reduce the redundant multiple message
-passes between the AS boundaries. T2MC uses a customized k-mean classification
-algorithm with $k=2$ to perform the classification, and exploits the stable
-structure of the internet routers to guide clustering.
-Even though traceroute provides detailed information about the network
-structure, use of traceroute creates overhead to the overall network structure.
-For this same reason, it is not unusual for network administrators to disable
-this support on their network routers, which may affect the performance of the
-T2MC algorithm.
-
-%\paragraph*{\textbf{T2MC\\}}
-
-%\paragraph{}
-%\emph{T2MC}\cite{shi_t2mc_2008} exploits some properties of the Internet paradigm and clusters nodes belonging to the same ISP without any centralised control or predefined system parameterization.
-The algorithm considers the dynamic nature of peers and exploits the stable nature of routers in order to build a topology location relationship among end peers.
-%
-%\paragraph{}
-%Some special routers split the physical network into autonomous system domains.
-Using a Traceroute mechanism T2MC searches for latency leaps among the path to a host in order to form ``near'' and ``remote'' router clusters.
-This is achieved by a 2-Means Classification, defined by the following steps:
-%\begin{enumerate}
-% \item the peer chooses the minimum and maximum Latency results from the Traceroute for initializing the cntroinds of two sets ``first'' and ``second''.
-% \item The peer calculates, for all hops allong its tracerouted path, the absolute distance to the centroids of both sets and assigns the routers to that centroid with which it has the smaller absolute distance.
-% \item The peer calculates the latncy mean and variance value of two sets
-% \item If the variance is larger than a predefined threshold then the algorithm takes a loop from step 2 picking the two latency mean values as new centroids of sets ``first'' and ``second''.
-%\end{enumerate}
-%Ultimately peer will end up with two sets having the minimum intra-set variance.
-Finally the peer chooses the router from ``second'' set with the minimum hops attribute and sets it as a threshold.
-The selected router and all others whose hops attribute is larger than the threshold are classified as ``remote'' router cluster.
-The remaining are classified as ``near''.
-From the ``near'' class, the peer chooses the one with maximum hops attribute as its edge router, and registers it along with the all the ``near'' cluster into the DHT of the p2p overlay.
-As new peers join the network, those that share the same edge router or any of the members of the ``near'' router clustersm ther would gather to form a ``close'' peer cluster.
-As Edge routers can provide more valuable information than other members of the ``near'' set, T2MC was designed to prioritize interaction of peers and edge gateways.
-%
-%% TODO: figure t2mc.jpeg (a) stars represent edge routers (b) hop leaps
-%
-%\paragraph{}
-%The use of Traceroute as a tool for implementing the distance measuring infrastructure raise concearns about its efficiency and scalability.
-Being, primarilly, a network diagnostic utility, it is concearned too heavy weighted and intrucive for use in a larger scale\cite{ratnasamy_binning_2002}.
-Additionally, disabling ICMP is a common administrative policy for edge sites to enforce security, while dumping BGP routing tables\cite{krishnamurthy_bgpclust_2000} is not directly available to the application layer.
-
-\paragraph*{\textbf{Unnamed Unstructured!!\\}}
-% TODO: to be reviewed
-
-\paragraph{}
-In \cite{hsiao_redblue_2009}, Hsiao et al, claim to construct topology-aware unstructured overlays that \emph{guarantee} performance qualities in terms of
-\begin{inparaenum}[\itshape i\upshape)]
-  \item the expected communication latency among any two overlay peers regardless of the network size, and
-  \item the broadcasting scope of each participating peer.
-\end{inparaenum}
-
-The algorithm constructs an undirected graph $G = \left( V, E \right)$ comprised by two subgraphs.
-The first, namely $G^{\left( red \right)} = \left( V^{\left( red \right)}, E^{\left( red \right)} \right)$ in the paper's context, includes all vertices of $G$ and ensures the connectivity of the graph by securing at least one path between any two nodes.
-In contrast, $G^{\left( blue \right)} = \left( V^{\left( blue \right)}, E^{\left( blue \right)} \right)$, contains those vertices of $G$ that have free edges to link to other nodes and because these are fully utilized, the following also stands $E = E^{\left( red \right)} \cup E^{\left( blue \right)}$.
-
-A joining peer $u$, partitions its neighbours into two subsets, the $B_u^{\left( red \right)}$ and $B_u^{\left( blue \right)}$.
-In order to populate the $B_u^{\left( red \right)}$ subset, peer $u$ samples peers uniformly and at random.
-Then, each of these selected peers discovers a routing path starting from itself towards the node with the smallest (or the largest) ID in the system.
-
-\paragraph*{\bf Distributed Domain Name Order}
-
+%
+%\subsection{mOverlay}
+%Zhang et al.
+in \cite{zhang_moverlay_2004} focus on two aspects of the overlay construction.
+First, to achieve \emph{efficiency} of communication, a protocol should minimize unecessary long-distance hops and redundant traffic.
+Second, to be \emph{scalable} the overlay should be constructed in a distributed fashion and maintenance cost (data and locality management) should be minimized.
+Both of the above observations, are positively affected by the exploitation of node proximity at the underlying network.
+To accomplish it, the paper introduces the \emph{group} concept, according to which non-static (dynamic) landmarks are used to compute proximity, resulting in a highly robust and scalable overlay construction with reduced maintenance cost as well.
+The dynamic nature of landmarks help towards load-balance, meaning that hot-spots are avoided.
+%
+%\paragraph{}
+%The proposed architecture, called \emph{mOverlay}, is a \emph{clustering approach} which creates a two-level hierarchical network, where on the top level we have connections between groups while on the bottom we have connections between hosts inside groups, thus attempting to recreate \emph{Small-World}-like properties for the constructed overlay network.
+The notion of a \emph{group}, in the paper's context, is a set of hosts that are close to each other with respect to any position $P$ in the underlying network.
+The distance between hosts can be
+%\begin{inparaenum}[\itshape i\upshape)]
+%  \item network latency,
+%  \item round-trip time,
+%  \item minimum bandwidth on the links along the path connecting the two nodes, or
+%  \item some other user-defined cost metric between the two nodes.
+%\end{inparaenum}
+%These nodes are refered to as neighbouring. Similarly groups can exchange messages with their neighbouring ones\footnote{Groups nearby, in the underlying physical network.}.
+%
+%According to the \emph{grouping criterion} set by the authors, a new host $Q$ belongs to some group $A$ if the distance between $Q$ and group $A$'s neighboring groups is the same as the distance between group $A$ and it's neighbouring groups.
+The neighbour groups thus, play the role of the dynamic landmarks and for accuracy and performance reasons the nearest of them are chosen to play that role.
+%
+%\paragraph{Locating process} A new coming host, $Q$, first connects to a globally known host cache called the \emph{rendezvous point (RP)} in order to retrieve the starting point in the overlay, say $A$ in group $1$.
+Host $Q$ then, measures its distance to host $A$.
+At the same time, the later, sends information about the neighbour groups of group $1$ back to host $Q$.
+This list is called \emph{candidate group list}, and the newcoming host sequnentially measures its distance to each of them in seek for the closest one.
+If the \emph{grouping criterion} is met, host $Q$ belongs to group $1$.
+If not, a boot host from the closest group is found and the algorithm is re-run until the criterion is met or after a predefined number of repetitions.
+In the later case, $Q$ creates a new group comprising itself only.
+The above protocol does not favour hotspots as it spreads the probability of visiting a group across the whole overlay and limits the overhead in the level of $O \left ( log N \right )$.
+%
+%\paragraph{General overlay operations} A set of additional protocols, are also introduced, similar to those found in traditional unstructred networks, but modified focusing on scalability and robustness.
+For example a protocol for \emph{group formation} is introduced that exploits the inherent characteristic of proximity, in the overlay, in order to efficiently detect the neighbouring groups of a newly formated group from the set of adjasent groups of its closest neighbour.
+Additionaly, during \emph{group joining} the coresponding protocol denotes the exchange of important information for group maintenance.
+This can be furtherly improved by \emph{information sharing} between nodes of the same group, functionality handled by a dedicated flood-like protocol\footnote{Since nodes that belong to the same group are physically close this can be achieved at a minimum price.}.
+Moreover, another set of distributed protocols handle the \emph{information update}.
+The information that needs update, in the proposed architecture, is
+%\begin{inparaenum}[\itshape i\upshape)]
+%  \item the host cache, when a new node joins, and
+%  \item the neighbours of groups, when a close-by group is generated.
+%\end{inparaenum}
+%Finally, in case of \emph{host failure} or \emph{host departure} the system is able to maintain its stability since there are defined operations for periodical host cache update and group leader selection if one leaves or dies.
+%
+%\subsection{Distributed Domain Name Order}
 %{\sethlcolor{yellow}\hl{HA:  In paper/introduction nice advantages of unstructured networks:
@@
 %}}
-
-\emph{Distributed Domain Name Order (DDNO)} \cite{zeinalipour-yazti_ddno_2005}
-uses the domain names to detect topologically close nodes based on the
-assumption that nodes within the same domain are topologically close to each
-other. Half of the possible connections of a node is used to connect to these
-local peers, and the other half is used to randomly connect to the peers
-anywhere on the network.  \textit{DDNO} is a heuristic approach to solve the
-topology mismatch problem, with local connections to improve the efficiency and
-reduce the delay on the network.  The random connections on the other hand help
-ensure the connectivity in the network and avoid partitioning. \textit{DDNO}
-uses \textit{Split-Hash} and \textit{dnMatch} algorithms to detect locality by
-using domain names.
-
-\subsection{Landmark Based Proximity}\label{sec:landmark}
-
-Detecting the proximity of a node using landmarks, also known as \textit{Landmark
-Clustering}, is based on the view that nodes with similar distances to a
-set of predefined well-known landmark nodes are pretty likely of being close to
-each other. But this approach has its weaknesses as well, such as the fact that
-is a rather coarse grained approximation, therefore not particularly well suited
-for detecting the correct positions of nodes within close distance to each other.
-
-\paragraph*{\bf Landmark Binning}\label{sec:landmark_binning}
-
-Landmark Binning \cite{ratnasamy_binning_2002} is an approach to partition close
-by nodes into bins based on their distance to well known anchor nodes within the
-internet. In order to detect locality, nodes use network latency (round trip
-time) as a measurement technique. The network latency, even though not always
-accurate, is selected in this work because its non-intrusive, transparent and
-easy to apply structure. In order for the binning to work, authors assume that
-there are a few well known anchor servers with known physical locations on the
-internet. The authors estimate that around 12 servers will do the job. The nodes
-measure their distance to all these servers and record the latency values for
-each. Later, the tuple of these latency values represents the bin of each node.
-Nodes with similar latency values are assumed to be close to each other, or at
-least within the same AS. The landmark binning is designed as a generic model
-that can work both on structured and unstructured P2P networks, and
-\cite{ratnasamy_binning_2002} has the detailed description of the algorithm for
-both topologies. As the main operation of the algorithm is independent from the
-underlying model, and does not change for structured and unstructured networks,
-we classify and give the details of the algorithm in this decentralized
-unstructured section, and omit the duplicate work in the decentralized
-structured section. The landmark binning is also proposed as a good candidate to
-work on content distribution networks. The major disadvantage of landmark
-binning is to install and maintain landmark servers on different AS, all over
-the world.  A typical P2P network usually has a couple of million nodes
-connected at any time, which introduces possible scalability problems in terms
-of the landmark servers.  The authors claim that latency estimation does not use
-much network resources and in order to avoid the possible scalability problems
-the authors propose to replace single landmark servers with clusters of servers
-within the same physical area. However, this approach does not reduce the
-possible high network traffic flow through these landmark servers. One other
-possible problem is the incorrect binning caused by inaccuracy in delay
-measurements methods, as the network latency is not a proven accurate method for
-location estimation.
-
-\paragraph*{\bf mOverlay}
-mOverlay \cite{zhang_moverlay_2004} addresses the scalability problems that
-affect static landmark servers by introducing the dynamic landmarks. The
-nodes are divided into groups based on their distances to these groups and the
-neighbour nodes in these groups behave as dynamic landmarks. The mOverlay
-protocol is composed of two main parts, initially adding a new node to the
-overlay by selecting the closest group, and maintaining the overlay afterwards.
-Finding the correct closest group is the most important part of the overlay
-construction. The authors also formally prove that any new node can reach its
-group by performing at most $O(logN)$ communications within the network.
-
-\renewcommand\arraystretch{1.4}% (MyValue=1.0 is for standard)
-
-%\begin{figure}[h!]
-\hspace{-3ex}
-\begin{center}
-\footnotesize
-%\begin{tabular}{
-\begin{landscape}
-\begin{longtable}{
-|>{\columncolor[gray]{.7}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.2\columnwidth}
-|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
-|}
-\caption{Decentralized Unstructured Algorithms} \label{fig:unstruct_compare_table} \\
-\hline
-\rowcolor[gray]{.5}
-\textbf{Algorithm} & \textbf{Arch.} & \textbf{Overlay structure} & \textbf{Base protocol} &
-\textbf{Dynamic update} & \textbf{Runtime} & \textbf{Scalability} & \textbf{cites}\\
-\hline
-\endfirsthead
-\multicolumn{4}{c}%
-{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
-\hline
-\rowcolor[gray]{.5}
-\textbf{Algorithm} & \textbf{Arch.} & \textbf{Overlay structure} & \textbf{Base protocol} &
-\textbf{Dynamic update} & \textbf{Runtime} & \textbf{Scalability} & \textbf{cites}\\
-\hline
-\endhead
-\hline \multicolumn{4}{r}{\textit{Continued on next page}} \\
-\endfoot
-\hline
-\endlastfoot
-\textbf{Narada} & Decentralized unstructured & \textbf{Overlay optimization
-based}. Creates a mesh (richer connected graph) and builds minimum spanning
-trees on this mesh & & Yes. Supports incremental evolving & & Small and sparse
-groups & 2000  \\
-
-\hline
-\textbf{Gia} & Decentralized unstructured & \textbf{Forwarding based} Replaces
-Gnutella flooding with random walk, and introduces KaZaA style supernodes. Uses
-dynamic topology adaptation protocol &
- Gnutella & Yes & & Better than Gnutella & 974  \\
-
-\hline
-\textbf{Adaptive Overlay Topology Optimization} & Decentralized unstructured & \textbf{Overlay optimization
-based}. Creates overlay multicast tree with Selective Flooding protocol&
-Gnutella & Yes, using Active Topology protocol & & Better than Gnutella & 36 \\
-
-\hline
-\textbf{Location-aware Topology Matching} & Decentralized unstructured &
-\textbf{Overlay Optimization Based}. Uses \textit{TTL2-detector flooding}, \textit{low productive
-connection cutting}, and \textit{source peer probing}. & Gnutella & Yes & & Better than Gnutella & 204 \\
-
-\hline
-\textbf{Replication Strategies in Unstructured P2P Networks} & Decentralized unstructured &
-\textbf{Cache Based}. Uses uniform, proportional and square root allocation
-strategies to replicate data. & Gnutella & Yes & & Better than Gnutella & 580 \\
-
-\hline
-\textbf{Tracing a large-scale Peer to Peer System: an hour in the life of Gnutella.} & Decentralized unstructured &
-\textbf{Cache Based}. Proposes a caching algorithm based on the traces of the Gnutella traffic & Gnutella &
-Yes & & Better than Gnutella & 199 \\
-
-\hline
-\textbf{Improving search in P2P networks} & Decentralized unstructured &
-\textbf{Forwarding Based}. Uses \textit{iterative deepening}, \textit{directed
-BFS}, and \textit{local indices} to improve efficiency. & Gnutella & Yes & & Better than Gnutella & 380 \\
-
-\hline
-\textbf{Distributed Cycle Minimization Protocol} & Decentralized unstructured &
-\textbf{Forwarding based} Uses a decentralized cycle elimination protocol  &  & Yes &  &  & 9 \\
-
-\hline
-\textbf{Scalable Bipartite Overlay} & Decentralized unstructured &
-\textbf{Overlay optimization based} Uses bipartite partition graph and builds
-local minimum spanning trees  & Gnutella & Yes &  & Better than Gnutella & 77 \\
-
-\hline
-\textbf{Adaptive Connection Establishment} & Decentralized unstructured &
-\textbf{Overlay optimization based} Forms Neighbour Cost Tables, builds local
-minimum spanning trees and perform local optimizations & Adaptive Overlay
-Topology Optimization (AOTO), Gnutella & Yes &  & Better than Gnutella & 110 \\
-
-\hline
-\textbf{Hops Adaptive Neighbour Discovery} &  &
- &  &  &  &  & 1 \\
-
-\hline
-\textbf{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} & Decentralized unstructured &
-\textbf{Overlay optimization based} Uses piggybacking to discover neighbour
-distances and selects neighbours  & Gnutella & Yes &  &  & 36 \\
-
-\hline
-\textbf{mOverlay} & Decentralized unstructured &
-\textbf{Overlay optimization based} Uses dynamic landmarks to find node locality
-& & Yes &  & Due to dynamic landmarks and grouping, more scalable than tree-based or mesh-based protocols & 123 \\
-
-\hline
-\textbf{Distributed Domain Name Order (DDNO)} & Decentralized unstructured &
-\textbf{Overlay optimization based} Connects half of the nodes connections to
-the nodes in the same domain and the other half to random nodes, therefore
-supports locality and topological connection  & & Yes &  & Yes, by using super
-peers & 5 \\
-
-\hline
-\textbf{Peer-exchange Routing Optimization Protocols} & Works with both decentralized
-structured and unstructured &
-\textbf{Overlay optimization based} Optimizes overlay by the exchange of
-neighbors among peers  & & Yes &  & Yes & 13 \\
-
-\hline
-\textbf{MAY OMIT- T2MC} & Decentralized unstructured &
-\textbf{Cluster based---{\sethlcolor{yellow}\hl{ IS IT ALSO PROXIMITY BASED?}}} Uses traceroute results for clustering the
-nodes  & &  &  &  & 1 \\
-
-\hline
-\textbf{Unnamed-unstructured} & Decentralized unstructured &
-\textbf{Overlay optimization based} Minimizes the communication delay and
-maximizes the broadcasting range & & Yes &  & Better than THANCS and mOverlay & 7\\
-
-\hline
-\textbf{Landmark Binning} & Can work with both decentralized structured and unstructured &
-\textbf{Proximity based, landmark binning} Uses network latency to partition
-nodes into bins & & Yes &  &  & 817 \\
-
-\hline
-%\end{tabular}
-\end{longtable}
-\end{landscape}
-\end{center}
-\vspace{-2.5ex}
-\vspace{-2.5ex}
-%\end{figure}
-
-\subsection{Discussion on Unstructured Decentralized Algorithms}
-
-Earlier in section \ref{sec:unstructured} we presented the state of the art unstructured
-decentralized P2P algorithms in four different categories based on their
-architecture; broadcast based, cache based, overlay optimization based, and
-landmark based. In this section, we present a final discussion of all the
-methods including the advantages, disadvantages,
-and novelties presented in order to tackle the topology mismatch problem.
-Table \ref{fig:unstruct_compare_table} presents an overview of all the
-unstructured decentralized algorithms described earlier in this section.
-
-The broadcast based approaches in general propose intelligent neighbour
-selection algorithms in order to replace the inefficient blind
-flooding algorithm of Gnutella.
-Instead of forwarding the queries to all the neighbours, intelligently selecting
-neighbours with high probability of answering the query, or neighbours that have
-some specific features such as high bandwidth, high capacity, or low latency in
-practice reduces the aggregate resource usage of the P2P network and improves
-the overall performance of the system. However, since the broadcast based algorithms
-proposed do not consider the underlying physical topology when selecting the
-neighbours, or optimizing the overlay structure, the algorithms do not offer a
-solution to the topology mismatch problem. Although in practice the algorithms
-work, none can guarantee that a single link is not used by more than necessary,
-or the created topology resembles the underlying physical topology. Broadcast
-based methods can be used in conjunction with other approaches to improve the
-quality of the P2P systems.
-
-The success of caching, as a well studied method in the client server internet
-applications, also made it a good candidate for the peer-to-peer networks. Even
-though caching improves the performance, and reduces the overall resource usage
-of P2P systems, the design of caches is non trivial compared to the web caching
-systems. Due to the unique structure of the P2P overlays, each node being a
-server and a client simultaneously, two important problems has to be solved when
-designing caches. First, the lifetime of a query is short, as the nodes join and
-leave frequently. Second, the result of a single query string is not always the
-same, and changes based on the TTL value, source of the query, the lifetime of
-the query, etc. So, in order to develop a successful caching system in P2P
-systems, these parameters also have to be considered. Even though the state of
-the art P2P algorithms using caching methods reduce the resource usage of the
-network, since the physical network topology is not considered by any algorithm,
-the proposed systems do not provide direct ways to overcome the topology
-mismatch problem caused by the overlay network.
-
-The overlay optimization based algorithms can be analyzed in
-two main topics, the minimum spanning tree based ones, and the clustering based ones.
-Even though the standardization of the IP multicast protocol is completed a
-while ago, and some vendors deployed routers supporting the protocol, in
-practice the standard is not widely embraced as predicted, mostly because the
-protocol violates the stateless design of the internet routers. The research
-community, as an alternative proposed application layer multicast frameworks,
-such as Narada, to simulate the same multicast protocol without requiring IP
-multicast services. The application layer multicast, however, is not as
-efficient as IP multicast, and suffers from the topology mismatch problem.
-Narada, and its followers (AOTO, LTM, SBO) try to solve the topology mismatch
-problem by building a richer connected graph and forming minimum spanning trees
-over this graph that can efficiently route messages among peers. The application
-level multicast protocol, initiated by Narada, is a generic protocol that can be
-applied to P2P file sharing, as well as content distribution networks. The
-original Narada algorithm is designed for small groups therefore it is not
-scalable. The followers try to solve the scalability problem by introducing
-various methods including forming minimum spanning trees for each nodes' $N^2$
-neighbours, partitioning the graph into two random groups where each group is
-responsible of different tasks, or performing local optimizations dynamically on
-the overlay graph. The advantage of the minimum spanning tree based approaches
-is that they maintain the connectivity on the network in an efficient way, while
-still not shrinking the search space. However, building and maintaining a
-minimum spanning tree creates a huge overhead for the network. 
-One other popular approach used in overlay optimization is the cluster based
-approaches. In cluster based approach, nodes use various methods to detect
-physically close nodes to form clusters. T2MC uses traceroute logs and DDNO uses
-domain names to cluster close by nodes. However, the accuracy of the methods
-used to form clusters directly affect the success rate of the algorithm. Traceroute for
-example is a heavy weight protocol to frequently use on the network, and most of
-the network vendors for this reason do not allow traceroute calls. The major problem with
-cluster based approaches is that the limited connectivity within the local
-domains shrink the search scope dramatically, which negatively affects the
-search performance of the P2P system. DDNO addresses this limitation by allowing
-half of each nodes connections to be to random nodes over the network, which
-balances the efficiency of the clustering approach with improved connectivity.
-
-In landmark based proximity method, nodes use network delay (round trip
-time) as a distance measurement method to position themselves based on distance
-measurements to apriori known servers on the internet. The servers behave as
-landmark points, and nodes use an estimation method to discover their positions.
-The landmark servers are either used by nodes to directly calculate their
-positions, or indirectly used to cluster nodes into groups or bins based on
-their estimated distances to well known landmark servers, due to the assumption
-that nodes with similar distances to a set of landmarks are physically close to
-each other over the network. The landmark based protocol has two important
-drawbacks. Firstly the network delay is not a reliable distance estimation
-method. For example, based on the load on the network the delay to certain
-nodes or networks can change from time to time, which will eventually affect the
-distance measurements and wrong measurements will lead to wrong estimated positions for the
-nodes, or incorrect and non optimal clusterings of the nodes. The second drawback of
-using landmark servers is the cost of installing and maintaining many landmark servers over the internet
-for various AS domains. As popular P2P file sharing applications usually have
-millions of peers connected at any time, it is not false to assume that the hardware and the network costs of
-maintaining these landmark servers will be quite high. 
-A possible solution to the scalability problem of the static landmark servers is to use ordinary nodes as dynamic
-landmarks once they estimate their own positions. Even though this approach scales
-much better than static landmark servers, still the measurement accuracy problem affects
-the overall performance of the system.
-
-\section{Structured Decentralized Algorithms}\label{sec:structured}
-
-In this section, the algorithms proposed for the structured decentralized
-architectures are analysed and categorized based on their use of the proximity
-information to optimize the underlying structure. In structured P2P algorithms,
-the construction of the routing tables among nodes determines the efficiency of
-the algorithms.  Therefore, routing tables that represent the underlying
-physical structure well can achieve much better performance. For this reason,
-the proposed methods for topology mismatch problem generally use different
-levels of proximity information to optimize the routing table close to the
-physical network. We analyze the structured decentralized algorithms based on
-the categories presented in
-\cite{castro_proximitydht_2002,castro_topawareroute_2002,ratnasamy_openq_2002}.
-In this section, the details of the methodologies used and the algorithms for
-each methodology are discussed in details.
-
-%{\sethlcolor{yellow}\hl{HA:  Talk about Tapestry, Pastry, CAN and Chord briefly here!}}
-
-\subsection{Geographic Layout}
-
-Geographic layout is used as a method to optimize the routing table of the
-structured algorithms to represent the physical network topology as close as
-possible. In this method, physically close nodes are detected, and they are
-positioned closely in the overlay structure. In order for this method to work,
-there should be means to detect physical proximity of internet nodes. One
-popular method is to use the well known internet servers, or servers dedicated
-for this sole purpose, as landmark nodes, similarly as discussed previously in
-Section \ref{sec:landmark} (\textit{Landmark Based Proximity}).  The inaccuracy in
-the positioning using landmarks, and the cost of deploying and maintaining these
-landmarks are the major disadvantages of these systems. Even though managing the
-overlay structure based on the geographic layout of the nodes improves the query
-efficiency of the system, on the other hand, it tends to create hotspots, and
-the needed failure resilience is undermined by the fact that close by nodes are
-more likely to suffer collective failures.
-
-\paragraph*{\bf Global Soft-State}
-
-\textit{Global Soft-State} \cite{xu_globstate_2003} builds a global map to help
-choose shorter routing paths, combining the landmark binning method and small
-scale distance probes to reveal the proximity properties of the underlying
-network to the overlay. This global view of the state is made available to all
-nodes in order to help them find the best way to route their messages.
-\textit{Global Soft-State} operates in two main stages, generation and using of
-the proximity information. For generating the proximity information a hybrid
-approach is proposed, which uses landmark clustering as a preprocessing step in
-order to select a number of potential nearest neighbour candidates and then
-refine the selection by incorporating an RTT scheme to ultimately choose the
-closest node. For using the proximity information, the algorithm chooses a
-different path from the classic gossiping approaches for constructing and
-maintaining the overlay. It is based on landmark clustering based strategic
-placement of proximity information on the overlay enabling any node to access
-such information using a landmark number that reflects its physical position in
-the network. For various logical regions\footnote{This might be a high-order
-zone in the eCAN\cite{xu_ecan_2002} context or a set of nodes sharing a
-particular prefix in overlays such as Pastry.} maps of physical information are
-built and published where each node may appear in a maximum of $log\left( N
-\right)$ such maps. To dynamically adapt to changing network conditions, a node
-subscribes to relevant \emph{soft states} that utilize a notification system in
-order to initialize any necessary neighbour re-selection.
-
-Maintaining several host states at different layers, makes any content migration
-costly. Additionally, the method does not make any continuing effort to remap the
-overlay structure after a node successfully joins, in order to adapt its state to
-any occurrence of condition change. Although this approach greatly reduces the
-routing latency to far nodes, it is unable to dynamically identify nodes that are
-close to routers and gateways in order to construct the secondary overlay.
-Nevertheless, static recognition of such nodes is currently done based on BGP
-reports and pre-chosen landmarks, sacrificing the self-organising attribute of
-traditional DHTs.
-
-\paragraph*{\bf Mithos}
-
-\textit{Mithos} \cite{waldvogel_mythos_2003} is a P2P protocol which
-incorporates a directed incremental probing to find near optimal node placement,
-and classified by its authors as an integration of geographic layout
-and proximity routing overlay optimization methods.
-
-The bootstrap phase of \textit{Mithos} starts with a subset of existing members
-as the first set of candidates, and while the iteratively closest nodes are
-detected by probing the neighborhood, the candidate neighbor list is updated. In
-order to avoid a local minima \textit{Mithos} probes all the neighbours within a
-two hop distance from the current minimum before concluding the process.
-
-After finding its first neighbour, the newcomming peer is assigned an ID using
-information gathered during the iterative neighbour selection phase. Virtual
-coordinates are assigned to the newcomming peer by using the distances of two
-closest nodes and their neighbours, so that Euclidean distances between the node
-and all known hosts predict the network latency between
-them\cite{cox_vivaldi_2004}.  The benefit of these synthetic coordinates are
-that they are explicitly used as the node's ID, and distance computations among
-nodes can be done in ID space without requiring physical measurements.
-
-The last step of the algorithm is the interconnection among neighbours.
-\emph{Mithos} uses a \emph{quadrant}-based mechanism according to which each
-node establishes a link to the closest neighbour in each quadrant. During
-forwarding, the next hop is performed towards a neighbour in the same quadrant
-as the final destination. The newcoming node may not know of other neighbours
-in all quadrants, therefore, the node first identifies neighbours in all
-quadrants using a mechanism based on ideas similar to a perimeter
-walk\footnote{Used in Greedy Perimeter Stateless Routing (GPSR) protocol.} and
-then improves the results using parallel path processing by taking into account
-further geometric properties of node relationships.
-
-One major limitation of \textit{Mithos} is that the protocol cannot handle
-dynamic arrivals and removals from the network, as is happens in mobile
-networks.
-
-\paragraph*{\bf LAPTOP}
-
-\textit{LAPTOP} \cite{wu_laptop_2007},  organises the overlay into a tree based
-hierarchy with main focus on reducing hops during message routing as well as
-minimizing maintenance overhead. Additionally, a caching scheme is also
-incorporated so as to further reduce routing table update costs. The authors
-theoretically show that in \textit{LAPTOP} routing path length is bounded by $O(log_d
-N)$ and node joining and leaving in the overlay network is bounded by $O\left( d
-log_d N \right)$ hops in a balanced overlay tree, where $N$ is the number of
-nodes, and $d$ is the maximum degree of each node. \textit{LAPTOP} implements
-the geographical layout approach  and constructs a geographical layout in a
-self-organizing and efficient fashion, by estimating the round trip time (RTT)
-to a small number of nodes in the overlay network in order to make them roughly
-aware of their physical distances among them.
-
-Each node is assigned an address in a dotted format (e.g 1.3.4). Each octet
-ranges from $1$ to $d$, where $d$ is the maximum degree of the nodes. The
-assignment process is done by appending a unique octet to the address of each
-nodes
-parent, while root node is assigned address 1.  The routing scheme is similar to the
-longest-prefix IP matching scheme. At each forwarding hop, any message travels
-up the tree until the first common ancestor of source and destination node is
-reached and then starts descending to arrive to its target. During tree
-traversals, special entries in the routing tables, called \emph{routing cache},
-are maintained in order to increase routing efficiency and achieve finer load
-balance. Caching enables a node to forward a message to a better longest-prefix
-match than that of its direct ancestor making a large, quicker and more cost
-effective step through the overlay and toward the destination. To improve
-scalability, the number of children nodes and the size of the routing cache are
-limited.  In terms of overlay maintenance, \textit{LAPTOP} incorporates a simple
-\emph{heartbeat}-based technique where each parent node is responsible for
-monitoring its children.  At join process the newcomming node is assigned its
-level label as well as its address by its parent node. Additionally it
-initializes its routing table (with normal and caching entries) as it traverses
-the overlay in search for its parents node.  During a graceful departure, the
-referred node checks for children in the overlay. If it does not have any, it
-simply notifies its parent and leaves. If it has, it selects the child node with
-the lowest RTT to it in order to take its place so that the locality property is
-preserved.
-
-\subsection{Proximity Routing}
-Proximity routing does not require routing tables to be built using any knowledge
-about network proximity. On the other hand it exploits such knowledge in order
-to choose the best next hop during routing a message. This approach 
-balances between choosing the node that will further progress the routing
-towards the destination and choosing the closest entry in the routing table, in
-terms of network proximity. Thus, it is relatively less effective than
-geographical layout when applied to CAN(-like) implementations. Moreover, the
-technique has been incorporated into a version of Chord causing an increase on the overhead of
-node joins and the size as well as maintenance cost of finger tables.
-%
-%\cite{dabek_cfs_2001} proposes a server selection scheme for the Chord DHT, on
-%the domain of proximity routing selection. In \emph{CFS}, each node predicts
-%the entire lookup latency as a function of the total number of nodes and the
-%average overlay next routing peer. The problem is that it is very difficult to
-%have a clear picture on the total number of nodes and the average hop latency
-%from the local. This leads to rough estimations that consequently decreases
-%overall performance.
-%
-The methods that use proximity routing and their details are listed below.
-
-\paragraph*{\bf Proximity in Kademlia}
-Before discussing {\em Proximity in Kademlia}, it would be helpful to briefly
-summarize the highlights about the {\em Kademlia} protocol.
-{\em Kademlia} \cite{maymounkov_kademlia_2002} is a distributed hash table (DHT) based
-protocol for P2P networks, which uses an iteration based lookup algorithm. The
-protocol uses the standard $160$ bit ID system for nodes and locates the nodes in
-a prefix binary tree, where IDs are used as prefixes. The iterative lookup
-operations are done over this prefix binary tree, which converges to logarithmic lookup
-times. The ID's are assigned randomly, therefore in Kademlia,
-there is no proximity control, which results in inefficient use of the underlying
-network during lookup and retrieve operations.
-
-{\em Proximity in Kademlia} \cite{kaune_pkad_2008} introduces proximity controls
-over the base Kademlia protocol to optimize the underlying network usage of the
-protocol. The authors define an abstract {\em underlay metric} that calculates
-the suitability of establishing a communication link between peers as a cost
-function, based on the used proximity criteria (RTT, ISP locality, etc.).
-{\em Proximity in Kademlia} adapted both {\em proximity routing} and {\em
-proximity neighbour selection} overlay optimization algorithms for controlling proximity. Therefore,
-we categorize the protocol only in this section, and skip the discussion in the
-further {\em proximity neighbour selection} section (Sec.\ref{sec:pns}). {\em
-Proximity in Kademlia} also uses the MaxMind GeoIP database\footnote{MaxMind
-Geolocation Technology. http://www.maxmind.com} to detect the
-proximity information of a given peer based on its region, country, or ISP
-location, and use this information to form clusters of nearby peers. One
-other method {\em Proximity in Kademlia} uses is the Vivaldi \cite{cox_vivaldi_2004} protocol. However, authors report that
-clustering approach performs better than Vivaldi protocol based on their
-experiments. Authors also report that {\em Proximity routing} protocol
-successfully worked with the Kademlia protocol and improved the locality of the
-connections over peers.
-
-\paragraph*{\bf CHOord considering Proximity on IPv6 (CHOP6)}
-
-\textit{CHOP6} \cite{morimoto_chop6_2007} is designed based on the
-\textit{Chord} protocol. \textit{CHOP6} roughly estimates the proximity among
-nodes by exploiting the IPv6 address format and RTT information if available.
-The proximity estimation is achieved by introducing a 64-bit ID scheme in which
-the least significant bit part is the IPv6 global routing prefix and thus
-enabling a longest prefix match scheme. The protocol is designed based on the
-observation that it is possible to estimate a node's geographical location by
-simply examining the upper 32-bits of its IPv6 address. Moreover, similar to
-\textit{Chord}, \textit{CHOP6} uses a finger table, whose entries hold more than
-one candidate node.
-
-\paragraph*{\bf Chord6}
-\emph{Chord6} \cite{xiong_chord6_2005} is another \textit{Chord} variant that
-tries to exploit the hierarchical features of IPv6 in order to create a
-substrate that reduces interdomain traffic between service providers.
-\textit{Chord6} is based on the original Chord protocol, and the main difference
-from the original protocol is in the identifier definition. Therefore, the
-approach can be easily portable to other DHTs such as CAN, Pastry and Tapestry.
-In Chord6 the identifier contains two parts: the higher bits are obtained by
-hashing the node's IPv6 address prefix of specific length, while the remaining
-lower bits are the hash value of the rest of that IPv6 address. As a result of
-this assignment, nodes in a domain will be mapped onto a continuous key space on
-the overlay network, which avoids unnecessary message forwarding across
-different service providers, thus minimizing overall routing cost.
-
-\subsection{Proximity Neighbour Selection}\label{sec:pns}
-\textit{Proximity Neighbour Selection} constructs the routing tables using proximity
-knowledge. The proximity information used in this method is different than the
-landmark based systems described in the \textit{Geographic Layout} section, as TTL values between
-nodes, or directly node ID prefixes are used to detect proximity.  Tapestry,
-Pastry, and CAN successfully implemented the proximity into their algorithms by
-using this approach.  The routing protocol in Pastry is based on longest node ID
-prefix matching, while CAN uses RTT values to detect close by nodes.
-\cite{castro_proximityp2p_2002}  reports that proximity neighbour selection as
-an effective proximity based method.
-%
-Proximity Neighbor Selection based algorithms are described in detail below.
-
-\paragraph*{\bf DHT-PNS}
-
-Chord-DHT-PNS \cite{hancong_pnsbased_2006} implements proximity neighbour
-selection on top of the Chord DHT. The main purpose is to use proximity
-information to group physically close by nodes as neighbours in the DHT table.
-In order to detect proximity, virtual network coordinates of peers are used, by
-using the Vivaldi protocol \cite{cox_vivaldi_2004}. The virtual coordinates are
-then used by the nodes to map to identifier space in the DHT. The space is
-partitioned using a \emph{concentric circle clustering scheme} where successive
-cycles of radiuses $\rho$, $2\rho$, $3\rho$ and so on, are constructed. Then the
-formed annuluses are divided into $2\chi-1$ \emph{sectors}, where $\chi$ denotes
-the sequence number of the annulus starting from $\chi = 1$ for the centre
-cycle. It is proved in the paper, that this way each sector occupies the same
-area as does the center cycle. Assuming uniform node distribution, this
-characteristic, favours a more load balanced clustering operation. Every sector
-in this $2D$ coordinate space is mapped to a unique \emph{region} in the DHT
-space forming a multi-layer node identifier space.  Thus, any nodes that belong
-to the same sector, are mapped to the same region as well, preserving their
-proximity relationship unveiled by the use of the Vivaldi protocol. The
-individual pieces are mapped to the identifier space uniquely, allowing
-logarithmic lookup operations with high probability on Chord.
-
-\paragraph*{\bf IP-based Clustering (IPBC)}
-Proximity neighbour selection algorithms use probing and other measures
-to detect proximity. However, such methods are either are not precise enough, or
-creates overload in the network. \emph{IP-based clustering}
-\cite{karwaczynski_ipbc_2007} is a proximity neighbour selection based
-algorithm in which the authors propose to use the IP address prefixes (16 bit
-for IPv4) in order to detect proximity.  \cite{freedman_iploc_2005} states that
-$97\%$ of prefixes larger than $24$ bits belong to a single geographical
-location. However, using less number of bits creates less precise results and
-more number of bits increase the burden on the network and reduce the possible
-number of neighbours. Therefore, a careful selection is required in terms of
-performance/accuracy tradeoff. The {\em IP-based Clustering} generates a key and
-its prefix and stores its prefix in the DHT itself, so that any newly joining
-nodes with the same IP prefix can query the prefix and identify all the
-neighbours with the same prefix relatively easily. Nodes periodically update
-their entries in the DHT and remove their entry when they leave, or the entry is
-timed out with no further activity detected from the node.
-
-\paragraph*{\bf Cone}
-
-\textit{Cone}\cite{wang_cone_2007} extends the Chord using proximity neighbour selection
-topology optimization algorithm. The proximity information is generated using
-landmarks and RTT based distances to landmarks.  \textit{Cone} uses a two-layered
-identifier space. The first, named Chord-layer identifier, denoted as
-$Id_{Chord}$, is the same as in Chord. The second is the Cone-layer
-identifier, $Id_{Cone}$ which is constructed by two component identifiers. The
-first, known as \emph{group identifier (gid)} denotes a relevant group the node
-belongs to while the second, namely \emph{local identifier (lid)} indicates the
-local identifier within the group. The group concept, which is introduced here,
-is a way of dividing nodes according to a common $Id_{Chord}$ prefix.  The
-structure of a Cone overlay, retains the Chord's circular topology. The difference
-lies on the fact that, now, two rings are created. A big ring, where nodes with
-the same \emph{gid} are arranged at each position. Each of these positions are a
-smaller ring for the particular group's \emph{lid}s. The routing is achieved in
-both clockwise and counter-clockwise directions in the big-ring, for which two
-routing tables are maintained, namely \emph{front} and \emph{back} finger
-tables. Entries in these tables, display physical network
-proximity with the current node. Moreover, a third table called \emph{group}
-table maintains information about other online peers within the current node's
-group in a way that entries are now close in the ID space.
-
-\paragraph*{\bf SAT-Match: Self-Adaptive Topology Matching}
-
-\textit{SAT-Match} \cite{ren_satmatch_2004} is a protocol that specifically
-tries to solve the topology mismatch problem by mapping the overlay network as
-close as possible to the physical one. Even though no theoretical proof is
-given, valid results are obtained and demonstrated through experimentations.
-Similar to the other approaches in the proximity neighbour selection,
-\textit{SAT-Match} uses probing to detect close by peers and obtain proximity
-information. However, as an additional feature, \textit{SAT-Match} peers can do
-selective jumps to adjust their locations in the DHT, if it reduces the stretch
-of its one-hop neighbourhood. The stretch is defined as the ratio between the
-average logical and physical latency of a link. It is reported that, due to the
-selective jumps, the \textit{SAT-Match} achieves $40\%$ reduction in link
-stretch, and when used with the \textit{Landmark Binning} approach (see Sec.
-\ref{sec:landmark_binning}), the reduction rate increases up to $60\%$.
-For dynamic environments, with frequent node arrival and leave,
-\textit{SAT-Match} scales much better than \textit{Mithos}, due to its self
-adaptation mechanism and selective jumps.
-
-\emph{SAT-Match} uses a small TTL value for the probing messages in order to
-reduce redundancy\cite{jiang_lightflood_2008}. This process begins as soon as
-the node joins the network using a DHT mechanism. Each probing message contains
-information about the source and a small TTL value. The recipient of such a
-message returns information about itself to the source and forwards the probing
-message to its neighbours if the TTL is non-zero. The discovered nodes are
-referred to as $TTL-k$ neighbourhood of the source node based on the $TTL$ distance
-to the source node.
-
-Blindly selecting the peer with the smallest RTT as neighbour is, generally, not
-the optimal decision to make in order to achieve global \emph{stretch} reduction.
-In a structured scheme, when a node jumps to connect to a
-physically close node, it may need to connect to other distant nodes to maintain
-the structure's integrity, thus creating an overall increase in the overlay's
-\emph{stretch}. The two nodes with the smallest RTT is then used in order to
-select one zone to jump in this phase. The algorithm is as follows: The source
-node $S$ calculates the stretch change of its $TTL-1$ neighbourhood and that of
-the $TTL-1$ neighbourhood of the first of the previously selected peers. These
-calculations are made as if the jump has been made. If the stretch reduction is
-over a predefined threshold the jump is performed, otherwise the second selected
-candidate is picked and the same computations are performed. If again, the
-threshold is not met, then no jump is ultimately done. In case of a jump, this
-is performed as a combination of \emph{leave} and \emph{join} operations, in the
-CAN context.
-
-%\begin{figure}[h!]
-\hspace{-3ex}
-\begin{center}
-\footnotesize
-\begin{landscape}
-\begin{longtable}{
-|>{\columncolor[gray]{.7}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.2\columnwidth}
-|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
-|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
-|}
-\caption{Decentralized Structured Algorithms}\label{fig:struct_compare_table}\\
-\hline
-\rowcolor[gray]{.5}
-\textbf{Algorithm} & \textbf{Arch.} & \textbf{Overlay structure} & \textbf{Base protocol} &
-\textbf{Dynamic update} & \textbf{Runtime} & \textbf{Scalability} & \textbf{cites}\\
-\hline
-\endfirsthead
-\multicolumn{4}{c}%
-{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
-\hline
-\rowcolor[gray]{.5}
-\textbf{Algorithm} & \textbf{Arch.} & \textbf{Overlay structure} & \textbf{Base protocol} &
-\textbf{Dynamic update} & \textbf{Runtime} & \textbf{Scalability} & \textbf{cites}\\
-\hline
-\endhead
-\hline \multicolumn{4}{r}{\textit{Continued on next page}} \\
-\endfoot
-\hline
-\endlastfoot
-
-\hline
-\textbf{Global Softstate} & decentralized structured &
-\textbf{Proximity based, landmark clustering} Uses first landmark clustering
-then measures RTTs to identify close nodes & & Yes &  &  & 195 \\
-
-\hline
-\textbf{Mithos} & decentralized structured &
-\textbf{Proximity based} Uses nodes as topology landmarks and directed
-incremental probing to optimize topology & & Yes &  & Scales well as all
-operations are local ??? & 172 \\
-
-\hline
-\textbf{Self-Adaptive Topology Matching} & decentralized structured &
-\textbf{Proximity based} Uses lightweight probing and
-selective jumps to optimize the topology & CAN & Yes &  & Better than Mithos
-& 40 \\
-
-\hline
-\textbf{Delay Aware P2P System} & &
-\textbf{} & &  &  &  & 1 \\
-
-\hline
-\textbf{VERSION OF CHORD - DHT-PNS} & decentralized structured &
-\textbf{Proximity based} Uses Proximity Neighbour Selection and the Vivaldi
-system & Chord  & Yes &  &  & 5 \\
-
-\hline
-\textbf{MAY OMIT- VERSION OF CHORD - Quasi-Chord} & decentralized structured &
-\textbf{Proximity based}  & Chord  & Yes &  &  & 0 \\
-
-\hline
-\textbf{LAPTOP} & decentralized structured &
-\textbf{Geographic layout based} Hierarchical overlay structure  &   & Yes &  & routing path length $\log{_d N}$, 
-join/leave overhead $d\log{_d N}$& 4 \\
-
-\hline
-\textbf{IP-Based Clustering} & decentralized structured &
-\textbf{Proximity based} Proximity neighbour selection based on longest common
-prefix of IP addresses &   & Yes &  &  & 1 \\
-
-\hline
-\textbf{CHOord considering Proximity on IPv6} & decentralized structured &
-\textbf{Proximity based} Uses IPv6 address format to provide proximity &  Chord
-& Yes &  & Better than Chord & 1 \\
-
-\hline
-\textbf{Proximity in Kademlia} & decentralized structured &
-\textbf{Proximity based} Applies  proximity neighbour selection (PNS) and proximity route selection (PRS)
-to Kademlia & Kademlia & Yes &  &  & 13 \\
-
-\hline
-\textbf{Cone} & decentralized structured &
-\textbf{Proximity based} Uses proximity neighbour selection (PNS) & Chord & Yes
-&  & Better than Chord & 3 \\
-
-\hline
-\textbf{Dynamo} &  & &  & 
-&  & & 3 \\
-
-\hline
-\textbf{MAY OMIT-BADLY WRITTEN-PChord} & decentralized structured &
-\textbf{Proximity based}  &  Chord
-& Yes &  & & 16 \\
-
-\hline
-\textbf{AChord} & &  &  
-&  &  & & 7 \\
-
-\hline
-\textbf{Chord6} & decentralized structured &
-\textbf{Proximity based} Uses IPv6 hierarchical address format to cluster
-topologically close nodes & Chord & Yes
-&  &  & 14 \\
-
-\hline
-\end{longtable}
-\end{landscape}
-\end{center}
-\vspace{-2.5ex}
-\vspace{-2.5ex}
-%\label{fig:struct_compare_table}
-%\end{figure}
-
-\subsection{Discussion on Structured Decentralized Algorithms}
-
-In section \ref{sec:structured},  we presented the state of the art structured
-decentralized P2P algorithms in three different categories based on their
-handling of the network structure to tackle the topology mismatch problem:
-geographic layout based, proximity routing based, and proximity neighbour
-selection based. In this section, we present a final discussion of all the
-methods including the advantages, disadvantages,
-and novelties related to solve the topology mismatch problem.
-Table \ref{fig:struct_compare_table} presents an overview of all the
-structured decentralized algorithms described earlier in this section.
-
-Structured P2P network algorithms use a global distributed hash table or a
-prefix tree structure to uniquely lookup peers or their data in the overlay
-network. As all the data is kept within the overlay, each node behaves as a
-client and a server, therefore nodes join and leave according to rules
-determined by the integrity of the global data structure. The main advantage of
-the structured P2P topology is that by the help of the global data structure,
-peers or their data can be found within the network even if there is only a
-single copy of that item present. However, each node join and leave creates
-maintenance overhead for the network due to updates required by the global data
-structure, and for networks with frequent node arrivals and departures the
-topology uses valuable network resources just to update the global structure.
-Nodes join the network by using a key value, which determines the location and
-the neighbourhood of the new node within the network. However, assigning
-random key values to the newly inserted nodes creates non-optimal matching with
-the underlying physical network topology, therefore, increasing the overhead of
-the network even more. One solution for handling the topology mismatch problem
-is to consider the proximity of the peers when generating the key and joining
-the node to the network, so that nodes within the same network domains are
-selected as peers, or neighbours, during the overlay topology construction. In
-this chapter, we have described three such approaches to optimize the topology
-matching problem: geographic layout, proximity routing and proximity neighbour
-selection.
-
-In geographic layout, nodes try to estimate the geographic positions of the
-peers, and construct the DHT considering the proximity of the peers. Landmark
-servers and RTT measurements are two popular methods, which can also be used in 
-conjunction, to discover physically close by peers over the network. However, as
-discussed in Section \ref{sec:landmark_binning}, these methods do not always
-give reliable estimates for the node positions over the internet. The landmark
-servers are not self-organizing and have maintenance overheads. To serve a
-P2P network with millions of peers, multiple landmark servers
-distributed over the whole world is required, which is hard to manage. The RTT
-measurements also can measure the delay between peers, but it is a greedy method,
-which can result non-optimal overlay topologies especially if close by nodes
-have low bandwidth connections among themselves.
-
-Proximity routing does not discover or store proximity information for the
-peers, however, during package forwarding, nodes that have lower latency to the
-destination, or with a closer key to the destination is selected. As no
-proximity information is used, the system is based on a greedy approach and it
-usually selects low latency paths over the overlay, which maps to suboptimal
-longer paths on the physical topology. For these obvious reasons, proximity
-neighbour selection is generally considered as a superior method to proximity
-routing, however, joint uses of these two protocols are also possible.
-
-Proximity neighbour selection approach is similar to proximity routing, with an
-exception that during forwarding, the proximity information of the peers are
-also considered. Therefore, for each application, proximity information is
-extracted and used during routing. Depending on the application, the proximity
-information can be the node sharing the same prefix with the current node, or
-TTL to detect close by neighbours. Tapstry and Pastry are two popular methods
-implementing the proximity neighbour selection among others mentioned above in the
-proximity neighbour selection section.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{The Algorithms}
-%Having understood the nature of the problem that peer-to-peer architecture face, this section starts the discussion of the academic work that has been conducted in the field the last few years.
-What follows this introduction, does not claim to be a thorough citation of all known protocols that are available out there, but instead, a carefull selection of those that left a distinctive fingerprint contribution in the efforts of the research community to alleviate the topology mismatch problem.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% UNSTRUCTURED
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%\subsection{Narada}
-%
-%\paragraph*{}
-%\emph{End System Multicast} \cite{chu_esm_2000,chu_esm_2001,chu_esm_2002} as well as other  proposes an architecture, where end-systems implement all multicast related functionality (including membership management and packet replication).
-In \emph{Narada}, the protocol developed in the paper's context, end-systems
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item self-organize, in a fully distributed manner,
-%  \item into an efficient overlay structure, both from the network and the application perspective,
-%  \item with self improving properties,
-%  \item adapting to network dynamics\footnote{Long-term variations in Internet path characteristics (i.e. bandwidth, latency e.t.c.)}.
-%\end{inparaenum}
-%
-%\paragraph*{}
-%What Narada really does is to construct overlay spanning trees in a two-step process:
-%\begin{enumerate}
-%  \item It builds a richer connected graph \footnote{ENarada \cite{li_enarada_2008} used Gossip protocol for the construction}, termed a \emph{mesh}, while trying to ensure a couple of desired performance properties:
-%    \begin{itemize}
-%      \item path quality between any pair of members should be comparable to the quality of the unicast path between them, and
-%      \item each member should have a limited number of neighbours on the mesh.
-%    \end{itemize}
-%
-%  \item It computes minimum spanning trees out of the mesh, each rooted at the corresponding source member.
-%\end{enumerate}
-%
-%\paragraph*{}
-%Narada incrementally improves the quality of the mesh by adding and dropping overlay links.
-Members probe each other at random and new links are added depending on what is the gain in doing so.
-On the other hand, participants, continuously  monitor the utility of existing links, and drop them if they are found to be not usefull.
-%
-%\paragraph*{}
-%Narada protocol achieves relatively good performance for small and medium sized groups involving tens to hundreds of members.
-This is not, by no means, what a contemporary peer-to-peer system needs to be characterized as scalable, but this work is considered pioneering\footnote{Another similar approach is \emph{Scattercast}\cite{chawathe_scattercast_2000}}, in that it was the first to have conducted a detailed Internet evaluation to analyse and demonstrate the feasibility of an overlay-based, application layer, multicast architecture that can dynamically adapt to bandwidth and latency properties of the physical underlying infrastructure.
-%
-%\subsection{Gia}
-%
-%\paragraph*{}
-%\emph{Gia} \cite{chawathe_gia_2003} is a decentralized unstructured peer-to-peer file sharing system with the following characteristics:
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item replaces Gnutella's flooding with random walks \cite{lv_randomwalks_2002},
-%  \item recognizes the importance of overlay network's topology, while using the random walks, and exploits it by incorporating a topology adaptation algorithm,
-%  \item introduces a token-based flow control mechanism, and
-%  \item detects the significant heterogeneity in peer bandwidth, processing power, disk speed e.t.c. and takes advantage of it.
-%\end{inparaenum}
-%
-%\paragraph*{}
-%More specifically, there are four key components in the design of \emph{Gia} which are summarized bellow:
-%\begin{enumerate}
-%  \item A \emph{dynamic topology adaptation} protocol that puts participating nodes within short reach of high capacity nodes so that these \emph{high-degree} nodes, which due to their high connectivity receive most of the queries, actually have the capacity to handle them.
-%  \item An \emph{active flow control} scheme is used to avoid overloaded hot-spots. Heterogeneity is detected and flow control tokens are given to nodes based on the available capacity.
-%  \item Every node maintains pointers to to the content that is offered by their immediate neighbours, creating a \emph{one-hop replication} of pointers, scheme.
-%  \item A \emph{search protocol} based on random walks, that is biased towards directing queries to high-capacity nodes that are typically best able to answer these queries.
-%\end{enumerate}
-%
-%\subsection{Adaptive Overlay Topology Optimization}
-%
-%\paragraph*{}
-%\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{liu_auto_2003} is an algorithm for building an overlay multicast tree among each source node and its direct logical neighbours in order to alleviate the mismatching problem while providing a larger query coverage range.
-It includes the steps of
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item \emph{Selective Flooding (SF)}, and
-%  \item \emph{Active Topology (AT)}
-%\end{inparaenum}
-%which are furtherlly discussed in the following paragraphs.
-%
-%\paragraph*{Selective Flooding (SF)}
-%Instead of flooding to all neighbours, SF builds an overlay multicast tree among each peer and its immediate logical neighbours.
-This reduces overall message burden to the network by making some neighbours non-flooding.
-The tree is formed using a minimum spanning tree algorithm.
-For this reason a peer needs to know the costs to all its logical neighbours (e.g.
-the network delay) as well as between any pair of neighbours.
-Additional, cost probing, messages need to be added to the original messages defined by the Gnutella protocol.
-%
-%LTM's SF effectiveness has be proven to be detached from the different physical or overlay topologies.
-On the other hand, SF is more effective with large number of logical neighbours.
-It can reach an average optimization rate of 87.4 percent on a logical topology with an average of 30 logical neighbours.
-%
-%\paragraph*{Active Topology (AT)}
-%At this step the overlay topology is reorganized.
-Each peer, independently, makes optimizations in the overlay network to alleviate topology mismatching by replacing non-flooding neighbours, with closer nodes.
-The approach incorporated by the paper is called \emph{Randomized AT} algorithm which picks up a candidate peer at random among the non-flooding neighbour's neighbours.
-Whenever a new neighbour cost table is received or there is a change of neighbours, the source peer has to re-calculate the multicast tree and apply the randomized AT algorithm.
-%
-%Different numbers of average logical neighbours has little to do with the effectiveness of AT.
-If the source has $n$ non-flooding peers, there are $n$ potential neighbour replacements.
-The overhead to exhaust all $n$ possible replacements can be high, so in practice, after each replacement the source peer can decide whether it needs to find another candidate peer.
-This is done by computing the cost improvement ratio greater than some predefined termination threshold.
-The larger the threshold, the slower, in the number of optimization steps, the reduction of the normalized average distance.
-As a whole the average response time is significantly reduced when more optimization steps taken.
-%
-%\subsection{Location-aware Topology Matching}
-%
-%\paragraph*{}
-%\emph{Location-aware topology matching} or \emph{LTM} \cite{liu_ltm_2004}, is an algorithm of building an efficient overlay by disconnecting low productive connections and choosing physically closer nodes as logical neighbours while still retaining the search scope and reducing response time of queries.
-This is done by issuing detector messages by peers, to a small region so that receivers of the detector can record relative delay information and cut inefficient or redundant logical links as well as add closer nodes as direct neighbours.
-Three operations are defined in LTMband are furtherly discussed in the following paragraphs.
-%
-%\paragraph*{TTL2-detector flooding}
-%\emph{TTL2-detector} messages are designed based on the specification of the Gnutella protocol.
-Suppose that $d(i, S, v)$ denotes the TTL2-detector which has the message ID of $i$ with TTL value of $v$ and is initiated by $S$.
-In addition to Gnutella's unified 23-byte header, such a message has a body of two types:
-%\begin{itemize}
-%  \item short format $d(i, S, 1)$ if $S$ is a direct neighbour of the receiver
-%  \item long format $d(i, S, 0)$ if $S$ is within a two-hop distance from the receiver
-%\end{itemize}
-%This way a peer can calculate link costs from within a one- or two-hop area around it and thus reach peers in $N(S)$ and $N^2(S)$ sets, respectively, in the paper's context.
-Each peer floods a TTL-detector periodically while it calculates link cost to another peer by receiving other peer's detector messages and checking:
-%\begin{itemize}
-%  \item the Source Timestamp field and its own clock upon time of receipt for one-hop messages
-%  \item the Source Timestamp, TTL1 Timestamp and its own clock upon receipt for two-hop messages
-%\end{itemize}
-%Of course peer clocks must be synchronized in order to do such checking.
-%
-%\paragraph*{Low productive connection cutting}
-%There are three cases for any peer $P$ who receives $d(i, S, v)$ multiple times:
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item $P$ receives both $d(i, S, 1)$ and $d(i, S, 0)$
-%  \item $P$ receives multiple $d(i, S, 0)$s from different paths, and P randomly chooses to process one
-%  \item $P$ receives one $d(i, S, 1)$ and multiple $d(i, S, 0)$s, and $P$ processes $d(i, S, 1)$ and one randomly selected $d(i, S, 0)$
-%\end{inparaenum}
-%If the link with the largest cost is found and is a direct neighbour then the connection is put in a will-cut list and stays there for a certain period of time.
-If it is not, then it is handled by other peers.
-After that period, connections are cut and recorded to $P$'s cut-list.
-%
-%\paragraph*{Source probing}
-%For a peer $P\in(N^2(S) - N(S))$ who receives only one $d(i, S, 0)$, the cost of $PS$ is obtained (with list look-up or probing).
-Then $P$ compares it with the cost from each hop and if $PS$ has the largest cost, $P$ will not keep this connection, while otherwise the connection will be created.
-%
-%\paragraph{}
-%Supposing $n$ is the number of peers, $c_n$ is the average number of neighbours and $c_e$ is the average cost of logical links, then in the flooding-based search the traffic incurred by one query from an arbitrary peer in a peer-to-peer network is $O(n)$.
-As observed in the Gnutella network \cite{sripanidkulchai_gnutella_2001}, each peer issues $0.3$ queries per minute in average, thus the per minute traffic incurred by the network with $n$ peers is $O(n^2)$.
-Because each $d(i, S, v)$ has a TTL of $2$ in each source peer, the traffic for one time LTM optimization in all peers is at most $2nc_n^2c_e$.
-If each peer uses LTM $k$ times per minute, the total traffic incurred is $2knc_n^2c_e$.
-Simulation shows the best value for $k$ is $2$ or $3$.
-So, the traffic overhead caused by LTM to the network is $O(n)$.
-%
-%TTLj-detectors, with $j > 2$, would detect and break cycles with more than 4 links.
-LTM though, does not use such detectors because detector-flood traffic would increase significantly, and cut links between two end-peers, could cause queries initiated by them to traverse a path much more expensive than the cost on the the cut link.
-%
-%\paragraph{}
-%LTM disadvantages are
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item disagreement of measured delay due to unsynchronized clocks causes problems when deciding the cut positions, which can influence the network connectivity, and
-%  \item the network delay metric mainly focuses on disabling the connections between peers physically far away without considering the shortcuts created by powerful peers.
-%\end{inparaenum}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/ltm_multid.jpeg}
-%\caption{Peer $P$ receives $d(i, S, v)$ multiple times}
-%\label{figure:ltm_multid}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/ltm_example.jpeg}
-%\caption{A full example of LTM}
-%\label{figure:ltm_example}
-%\end{figure}
-%
-%\subsection{Scalable Bipartite Overlay}
-%\emph{Scalable Bipartite Overlay (SBO)} \cite{liu_bipartite_2007} employs an efficient strategy to select query forwarding paths and logical neighbours.
-The topology construction and optimization of SBO consist of four phases:
-%\begin{enumerate}[\itshape i\upshape)]
-%  \item When a new peer is joining the network, it will randomly take an initial colour; say red or white.
-This colour is not to be changed for the time the peer stays connected.
-If it leaves and then rejoins it will, again, have to pick a colour, randomly.
-Thus all peers are separated into two groups, red or white.
-Then the bootstrap host will provide, the joining peer, with a list of active peers, including their colour information, in order for the later to establish connections to different colour peers.
-This way, all peers form a bipartite overlay.
-%  \item White peers probe distances with their immediate (red) neighbours, form a cost table and send this table to their corresponding red neighbours.
-%  \item Each red peer builds a minimum spanning tree using the obtained neighbour cost tables.
-These are two-hop diameter trees so white peers do not need to build one.
-The links that are part of a minimum spanning tree are called \emph{forwarding connections (FC)} while the rest \emph{non-forwarding (NFC)}.
-%  \item Having a minimum spanning tree with two hops a red peer is able to send its queries within that range.
-Some white peers, though, have become non-forwarding neighbours.
-In this phase, such a (white) neighbour will try to find another red peer being two hops away from its current red neighbour to replace the later as its new neighbour.
-%\end{enumerate}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/sbo_efficient_forward.jpeg}
-%\caption{A red peer $P$ has a small overlay topology of $N(P)$ and $N^2(P)$ and computes the efficient forwarding paths}
-%\label{figure:sbo_efficient_forward}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/sbo_neighbour_replace.jpeg}
-%\caption{Neighbor replacement}
-%\label{figure:sbo_neighbour_replace}
-%\end{figure}
-%
-%In a static enviromnent LTM may reduce traffic cost by around 80 to 85 percent while SBO reduces traffic cost between 85 and 90 percent.
-However, LTM  is proved to converge in around 2-3 steps while SBO needs 4-5 steps.
-Moreover LTM reduces response time by more than 60 percent in 3 steps while SBO needs 8.
-In a dynamic environment (10 minute average peer lifetime, 0.3 queries/sec by each peer) SBO and LTM reduce the average traffic cost per query (including the overhead due to the optimization steps) by 85 and 80 percent, respectively.
-Moreover LTM reduces the response time per query to 30 percent while SBO to 35 percent.
-%
-%\subsection{Adaptive Connection Establishment}
-%
-%\paragraph{}
-%\emph{Adaptive Connection Establishment (ACE)} \cite{liu_ace_2004} builds an overlay multicast tree among each source node and the peers within a certain diameter from the source peer and optimizes the neighbour connections that are not in that tree.
-%
-%\paragraph{}
-%ACE indicates three phases for its algorithm:
-%\begin{enumerate}[\itshape i\upshape)]
-%  \item Calculate cost between nodes using network delay as a metric.
-Each peer probes the costs with its immediate logical neighbours and forms a \emph{neighbour cost table (NCT)} using a special routing message type.
-Two neighbouring peers exchange their NCTs in order for every peer to obtain the cost between any pair of its local neighbours forming a small overlay topology.
-%  \item Based on obtained NCTs a minimum spanning tree among each peer and its immediate neighbours is built (Figure~\ref{figure:ace_phase2}).
-%  \item Physically far away neighbours are replaced by physically close neighbours.
-In ACE a peer, say $S$, probes the distance between one of its non-flooding neighbour's neighbour, say $G$ and $H$ respectively.
-If the link to neighbour's neighbour is smaller than that to the neighbour, the later is cut.
-If this is not the case but $S$ finds that cost of $GH$ is even larger than that of the $SH$, $S$ will keep $H$ as a new neighbour.
-Obviously, if $SH$ is larger than $SG$ and $GH$, the connection will not be established and $S$ will continue probing another neighbour's neighbour.
-The algorithm described is conducted within $1$-neighbour closure (among its source peer and all its direct neighbours) but the optimization scope can be enlarged.
-The larger the scope, the better topology matching improvement but also the greater the computational overhead (Figure~\ref{figure:ace_phase3}).
-%\end{enumerate}
-%
-%\paragraph{}
-%Simulations in \cite{liu_acesims_2004} show that the average scope of each query to cover the same scope of nodes is reduced by about 65 percent without losing any autonomy feature, while the average response time can be reduced by 35 percent.
-Larger diameter topologies lead to better topology optimization rate but also to higher communication and computation overhead.
-It was also found that it is more effective in higher connectivity dense topologies.
-Compared to LTM, it comes short of convergence speed.
-In \cite{ni_mismatch_2004} shows reduction of both total traffic (90 percent) and response time (80 percent) to message queries without shrinking the search scope.
-SBO, on the other hand, achieves approximately 85 percent reduction on traffic cost and about 60 percent reduction on query response time.
-Last but not least, it is concluded that work must be done on incorporating a more sophisticated selection policy for candidate non-flooding peers.
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/ace_phase2.jpeg}
-%\caption{Second phase in ACE}
-%\label{figure:ace_phase2}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/ace_phase3.jpeg}
-%\caption{Third phase in ACE}
-%\label{figure:ace_phase3}
-%\end{figure}
-%
-%\subsection{Hops Adaptive Neighbour Discovery}
-%
-%\paragraph{}
-%Paper \cite{chen_hand_2006} proposes a new algorithm named \emph{Hops Adaptive Neighbour Discovery (HAND)} which uses a fully distributed triple hop adjustment strategy to address the topology mismatch problem.
-The advantages of the algorithm compared to other approaches are that
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item it does not need any clock synchronization,
-%  \item it is a fully distributed algorithm making it robust and reliable in decentralized systems,
-%  \item the traffic overhead of the triple hop adjustment is very low,
-%  \item it is applicable to dynamic peer-to-peer environments, and
-%  \item maintains lower query response time.
-%\end{inparaenum}
-%
-%\paragraph{}
-%The algorithm's ultimate goal is the optimal overlay, the \emph{Logical Communication Network (LCN)} in the paper's context.
-The key concept of the algorithm is that a graph $G^{*}$ that describes an LCN and a graph $G$ that describes the current overlay are matched only if all peer hop sequences $(v_1, v_2, \ldots, v_k)$ in $G$ exist in $G^{*}$ and in the same order.
-In practice triple sequences $(v_1, v_2, v_3)$ are used.
-%
-%\paragraph{}
-%The mismatching detection is done in the following way.
-Suppose we want to verify peer sequence $v_2-v_1-v_3$ (see Figure~\ref{figure:hand_com_overlay}).
-A pair of probing messages are sent from $v_1$ to $v_2$ and $v_3$.
-Suppose delays of $(v_1,v_2)$ and $(v_1,v_3)$ are are denoted as $x$ and $z$, respectively.
-When the probing message arrives to $v_2$ it forwards it directly to $v_3$.
-Similarly, when the probing message arrives to $v_3$, it forwards it directly to $v_2$.
-These last steps are performed in order to obtain delays of $(v_2,v_3)$ and $(v_3,v_2)$ physical paths, respectively, denoted by $y$.
-%\begin{itemize}
-%  \item If $y=z-x\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched and should be adjusted to $v_1-v_2-v_3$ by deleting edge $(v_1,v_3)$ and adding a new $(v_2,v_3)$ (see Figure~\ref{figure:hand_matchtriple_lemma3}).
-%  \item if $y=x-z\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched and should be adjusted to $v_1-v_3-v_2$ by deleting edge $(v_1,v_2)$ and adding a new $(v_3,v_2)$ (see Figure~\ref{figure:hand_matchtriple_lemma4}).
-%\end{itemize}
-%$\varepsilon$ is a small positive real number denoting additional delays caused by possible forwarding and jitter delays.
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/hand_com_overlay.jpeg}
-%\caption{The communication in overlay}
-%\label{figure:hand_com_overlay}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/hand_matchtriple_lemma3.jpeg}
-%\caption{The matching triple in paper's Lemma 3}
-%\label{figure:hand_matchtriple_lemma3}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/hand_matchtriple_lemma4.jpeg}
-%\caption{The matching triple in paper's Lemma 4}
-%\label{figure:hand_matchtriple_lemma4}
-%\end{figure}
-%
-%\paragraph{}
-%Measurements conducted for evaluation perposes showed that in a static environment the algorithm can effectively decrease traffic cost by about 77 percent and shorten the query response time by about 49 percent in less than two minutes.
-In a dynamic environment it shows similar behaviour and with the size of the overlay network having little impact on the effectiveness of the algorithm.
-Compared to LTM both algorithms have almost the same traffic reduction rate, however on the response time reduction rate HAND has a higher one by about 4 percent.
-The traffic overhead of HAND is much less than that of LTM by an average of 55 percent.
-%
-%\subsection{Distributed Cycle Minimization Protocol}
-%
-%\paragraph{}
-%\emph{Distributed Cycle Minimization Protocol (DCMP)} which is introduced in \cite{zhu_dcmp_2008} is a dynamic, fully decentralized protocol that promises significant reduction of duplicate messages.
-To achieve that, it uses the slightly different approach of eliminating unnecessary cycles, while retaining the connectivity of the network and preserving fault resilience and load balancing properties of unstructured peer-to-peer schemas by avoiding the creation of a hierarchical organization.
-%
-%DCMP aims at cutting the cycle paths at strategic locations. Any peer that detects a duplicate message can initiate the cutting process which consists of the following steps:
-%\begin{enumerate}
-%  \item Peers in the cycle elect a leader called \emph{GatePeer}\footnote{GatePeers are important for maintaining the connectivity and optimal structure of the network while peers join and leave randomly.}.
-%  \item The cycle is cut at a well-defind point with respect to the GatePeer
-%\end{enumerate}
-%
-%\paragraph{}
-%The first step after detecting a duplicate message by some peer is to gather information from all peers in the cycle using a new type of control message called \emph{Information Collecting Message} or \emph{ICM}.
-ICM contains:
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item a \emph{GUID}\footnote{Globally Unique IDentifier assigned to every query message generated by any node.} field same as the one of the duplicate message,
-%  \item \emph{DetectionID} field which represents the direction of the connection where the duplicate was identified\footnote{This ensures the uniqueness of the ICM messages because as it travels through many cyclic paths, multiple peers will detect the duplicates and initiate an ICM message.}, and
-%  \item \emph{Node Information Vector (NIV)} which contains information (bandwidth, CPU power, etc) about peers that propagated the ICM.
-%\end{inparaenum}
-%
-%\paragraph{}
-%Suppose $A$ detected the duplicate as depicted in Figure~\ref{figure:dcmp}.
-It then emits an ICM to $B$ and $F$, that initially contains information only about itself.
-Each peer that receives the ICM, appends its information and propagates it along the reverse path of the original message.
-Since two copies of ICM are sent, at some point, a peer, say peer $D$, will receive a duplicate ICM.
-Using the information in the NIVs of the ICMs, $D$, decides to cut (for example) the EF connection.
-To inform the other peers about its decision, it emits a \emph{Cut Message (CM)} which contains the GUID and DetectionID of the corresponding ICM and an additional field that identifies the connection to be cut.
-$D$, then, forwards the CM in the reverse directions from where the ICM arrived.
-Similarly CMs received by any peer are propagated toward the reverse path of the corresponding ICM.
-Eventually, either peer $E$ or peer $F$ will receive the CM and cut the connection, thus eliminating the cycle.
-%
-%Receiving a duplicate ICM denotes the existence of a cycle.
-The opposite is not true though.
-For example, if the cycle contains $2 \times TTL$ edges, it will not be detected, because the ICM messages will be discarded before they locate it.
-There is a tradeoff between preserving the connectivity of the network and minimizing the duplicates that makes such a possibility to be safely ignored.
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/dcmp.jpeg}
-%\caption{Cycle elimination methods in DCMP}
-%\label{figure:dcmp}
-%\end{figure}
-%
-%\paragraph{}
-%Experiments in \cite{zhu_dcmp_2008} showed that DCMP incurs a lower delay, returns more results and decreases the number of duplicate messages by 22\%, compared to LTM.
-Additionally DCMP has one to two orders of magnitude less overhead, because it adopts a more ``LAZY'' approach than broadcasting control messages periodically like LTM does.
-%
-%\subsection{Two-Hop-Away Neighbour Comparison and Selection}
-%
-%\paragraph{}
-%Work in \cite{liu_thancs_2005,liu_thancs_2008} proposes a distributed heuristic called \emph{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} that
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item is completely distributed and needs no global knowledge,
-%  \item presents trivial overhead compared to the query cost savings
-%  \item its convergent speed of the algorithm is fast enough (faster than minimum spanning tree approaches) so that is effective to dynamic environments, and
-%  \item does not shrink the search scope.
-%\end{inparaenum}
-%
-%\paragraph{}
-%THANCS is considered a \emph{local search method}, in the sense that it targets in finding a locally optimum solution, by exploiting knowledge within a 2-hop radius.
-The algorithm consists of two main components: \emph{piggybacking neighbour distance on queries} and \emph{neighbour comparison and selection} which are furtherly discussed bellow.
-%
-%\paragraph{Piggybacking neighbour distance on queries}
-%Using network delay as a metric for measuring the distance, each peer probes distances with its immediate naighbours and stores information locally.
-For this reason a special query message type, \emph{Piggy Message (PM)}, is introduced.
-It is 6 bytes long and includes two fields: Neighbour IP Address and Neighbour Distance.
-A peer $P$ constructs a PM for its neighbour $Q$, which contains $Q$'s IP address and $Q$'s distance from $P$.
-When $P$ receives a query from $Q$, this PM will be piggybacked by the query that goes to all other neighbours of peer $P$.
-Upon receiving such a query message, each of the other neighbours will detach the PM, record the $PQ$ distance and process the query.
-This PM will not be further forwarded.
-In selecting which incoming queries  should piggyback a PM the paper proposes the \emph{pure propability-based (PPB)} and the \emph{new neighbour triggered (NNT)} policies.
-%
-%\paragraph{Neighbour comparison and selection}
-%Figure~\ref{figure:thancs} illustrates this component of the THANCS algorithm.
-A peer $S$ probes the distance to all known unprobed $N^2(S)$\footnote{$N^2(S)$ denotes the set of peers being two hops away from $S$, while $N(S)$ denotes the set of direct logical neighbours of $S$.}.
-The distance of $SP$ is known to $S$.
-Upon receiving a PM from node $P$ with the distance of $PQ$, $S$ follows one of the following:
-%\begin{itemize}
-%  \item $Q \in \left( N(S) \cap N^2(S) \right)$, i.e.
-$Q$ is direct neighbour of $S$.
-In this case $S$ will compare cost of $SQ$, $SP$ and $PQ$.
-If the most costly connection is one of $SQ$ or $SP$ the corresponding link will be put into a \emph{will-cut list}\footnote{Links are not immediately disconnected when put in the will-cut list.
-There are useless for forwarding but are kept active, for some time, in order to serve query responses that are traveling to the source peer along the inverse search path.}.
-If the most costly connection is $PQ$ then $S$ will do nothing, as the fully distributed nature of the algorithm will give a chance to either $P$ or $Q$ to cut the connection.
-%  \item $Q \in \left( N^2(S) - N(S) \right)$, i.e.
-$Q$ is a two-hop-away neighbour of $S$.
-If $S$ hasn't probed $Q$ before\footnote{A distance cache is maintained and looked up in such a case.}, it probes and stores the result in the \emph{distance cache}.
-Having the distance of $SQ$, $S$ compares costs of $SQ$, $SP$ and $PQ$.
-If $SQ$ is the most costly, $S$ will not establish the connection.
-If $SP$ is the most costly, $S$ will establish connection $SQ$ and put $SP$ in the will-cut list.
-If $PQ$ is the longest, $S$ will keep  the connection with both $P$ and $Q$, expecting that $P$ or $Q$ will eventually disconnect link $PQ$, later.
-%\end{itemize}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/thancs.jpeg}
-%\caption{Probing two-hop-away neighbours}
-%\label{figure:thancs}
-%\end{figure}
-%
-%\paragraph{}
-%In a static environment THANCS has been proven to be effective; optimizing 45 percent out of the 60 percent of mismatched paths, constructing a nearly optimal overlay.
-This leads to a 60 percent reduction in traffic cost as well as a 40 percent decrease in query response time.
-In dynamic environments (Gnutella 0.6/Limewire super-peer-like and Ion flat-like), THANCS saves up to 70 percent of the traffic cost in the super-peer topology and 55 percent for the flat one.
-Average response time is also decreased by 60 and 45 percent, respectively.
-Generally, THANCS has similar performance to LTM, without needing synchronization.
-SBO, incurring half the  overhead of AOTO, reduces the traffic cost the most, while THANCS has lower response time and converges faster than SBO.
-THANCS is, thus, more suitable for a more dynamic environment.
-In addition, THANCS is easy to implement and its operation overhead is trivial, compared with the other three approaches.
-This design, however, has the limitation of not being easily extend to also support non-flooding-based systems.
-%
-%\subsection{mOverlay}
-%Zhang et al.
-in \cite{zhang_moverlay_2004} focus on two aspects of the overlay construction.
-First, to achieve \emph{efficiency} of communication, a protocol should minimize unecessary long-distance hops and redundant traffic.
-Second, to be \emph{scalable} the overlay should be constructed in a distributed fashion and maintenance cost (data and locality management) should be minimized.
-Both of the above observations, are positively affected by the exploitation of node proximity at the underlying network.
-To accomplish it, the paper introduces the \emph{group} concept, according to which non-static (dynamic) landmarks are used to compute proximity, resulting in a highly robust and scalable overlay construction with reduced maintenance cost as well.
-The dynamic nature of landmarks help towards load-balance, meaning that hot-spots are avoided.
-%
-%\paragraph{}
-%The proposed architecture, called \emph{mOverlay}, is a \emph{clustering approach} which creates a two-level hierarchical network, where on the top level we have connections between groups while on the bottom we have connections between hosts inside groups, thus attempting to recreate \emph{Small-World}-like properties for the constructed overlay network.
-The notion of a \emph{group}, in the paper's context, is a set of hosts that are close to each other with respect to any position $P$ in the underlying network.
-The distance between hosts can be
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item network latency,
-%  \item round-trip time,
-%  \item minimum bandwidth on the links along the path connecting the two nodes, or
-%  \item some other user-defined cost metric between the two nodes.
-%\end{inparaenum}
-%These nodes are refered to as neighbouring. Similarly groups can exchange messages with their neighbouring ones\footnote{Groups nearby, in the underlying physical network.}.
-%
-%According to the \emph{grouping criterion} set by the authors, a new host $Q$ belongs to some group $A$ if the distance between $Q$ and group $A$'s neighboring groups is the same as the distance between group $A$ and it's neighbouring groups.
-The neighbour groups thus, play the role of the dynamic landmarks and for accuracy and performance reasons the nearest of them are chosen to play that role.
-%
-%\paragraph{Locating process} A new coming host, $Q$, first connects to a globally known host cache called the \emph{rendezvous point (RP)} in order to retrieve the starting point in the overlay, say $A$ in group $1$.
-Host $Q$ then, measures its distance to host $A$.
-At the same time, the later, sends information about the neighbour groups of group $1$ back to host $Q$.
-This list is called \emph{candidate group list}, and the newcoming host sequnentially measures its distance to each of them in seek for the closest one.
-If the \emph{grouping criterion} is met, host $Q$ belongs to group $1$.
-If not, a boot host from the closest group is found and the algorithm is re-run until the criterion is met or after a predefined number of repetitions.
-In the later case, $Q$ creates a new group comprising itself only.
-The above protocol does not favour hotspots as it spreads the probability of visiting a group across the whole overlay and limits the overhead in the level of $O \left ( log N \right )$.
-%
-%\paragraph{General overlay operations} A set of additional protocols, are also introduced, similar to those found in traditional unstructred networks, but modified focusing on scalability and robustness.
-For example a protocol for \emph{group formation} is introduced that exploits the inherent characteristic of proximity, in the overlay, in order to efficiently detect the neighbouring groups of a newly formated group from the set of adjasent groups of its closest neighbour.
-Additionaly, during \emph{group joining} the coresponding protocol denotes the exchange of important information for group maintenance.
-This can be furtherly improved by \emph{information sharing} between nodes of the same group, functionality handled by a dedicated flood-like protocol\footnote{Since nodes that belong to the same group are physically close this can be achieved at a minimum price.}.
-Moreover, another set of distributed protocols handle the \emph{information update}.
-The information that needs update, in the proposed architecture, is
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item the host cache, when a new node joins, and
-%  \item the neighbours of groups, when a close-by group is generated.
-%\end{inparaenum}
-%Finally, in case of \emph{host failure} or \emph{host departure} the system is able to maintain its stability since there are defined operations for periodical host cache update and group leader selection if one leaves or dies.
-%
-%\subsection{Distributed Domain Name Order}
-%{\sethlcolor{yellow}\hl{HA:  In paper/introduction nice advantages of unstructured networks:
-%
-%Unstructured P2P networks o?er a number of
-%important advantages: (i) An unstructured network
-%imposes very small demands on individual nodes,
-%and more speci?cally it allows nodes to join or leave
-%the network without signi?cantly a?ecting the sys-
-%tem performance. (ii) Unstructured networks are
-%appropriate for content-based retrieval (e.g., key-
-%word searches) as opposed to object identi?er loca-
-%tion of structured overlays. (iii) Finally unstructured
-%networks can easily accommodate nodes of vary-
-%ing power. Consequently, they scale to very large
-%sizes and they o?er more robust performance in
-%the presence of node failures and connection
-%unreliability.
-%}}
-%
-%\paragraph{}
-%\cite{zeinalipour-yazti_ddno_2005} proposes the \emph{Distributed Domain Name Order (DDNO)} technique which makes unstructured overlay networks, topologically aware.
+%
+%\paragraph{}
+%\cite{z-yk_ddno_2005} proposes the \emph{Distributed Domain Name Order (DDNO)} technique which makes unstructured overlay networks, topologically aware.
 It results in a flat overlay topology but with some changes it can be utilized in \emph{superpeer} environments.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% TODO: Should be rewritten
-\section{Conclusions}
-Peer-to-peer architecture has been in the center of research attention in the
-last decade. Especially the decentralized unstructured genre exploits the
-advantages of loose coupling and self organization of computing nodes to form
-application-layer networks on top of the physical, best-effort infrastructure of
-the Internet that exhibit interesting properties. Scalability problems arouse
-quickly, though, because of the inefficient construction of this overlay that
-was built with no concern for the underlying physical network that causes a
-great deal of redundant traffic. The problem was identified by the research
-community as the topology mismatch problem between the overlay and the
-corresponding underlying physical network and a great deal of effort has been
-set towards alleviating it. Some fruits of this effort have been gathered and
-presented in this survey. Measurement of link cost through latency or RTT and
-deletion of inefficient established connections are the key concepts of almost
-all approaches. Others, manage to address the problem through hierarchical peer
-clustering (e.g. best IP matching). Additionally, some protocols focus on
-specific problems that furtherly arise, such as overlay partitioning, search
-scope reduction or convergence speed, to name just a few. Another desire is to
-form a protocol that could be applied to both decentralized unstructured and
-decentralized structured networks. Unfortunately no approach has equally
-addressed these problems in order to form a robust solution. So the field seems
-to be, still, fertile for any, new, clever idea.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\input{conclusion}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \bibliographystyle{acmtrans}
-%\bibliographystyle{abbrv}
-\bibliography{bib/p2p-overlays}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\bibliography{mard-survey}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{received}
+Received Month Year;
+revised Month Year; accepted Month Year
+\end{received}
 

### 2014-01-21 structured.tex
2012.05.29 (Vassilis) Monolithic file broken down to multiple smaller ones (per section) for better parallel contribution. Additional material on more recent algorithms for unstructured networks. Now also covering the BitTorrent protocol. Merged Huseyin's figures from previous contribution and code refactored for better referencing. Fixed bibliography bug (references appearing as ?).

\section{Decentralized Structured Algorithms}
\label{section:structured}

This section focuses on the structured P2P architectures. The following
paragraphs contain a thorough discussion and analysis of the algorithms proposed
in the literature for dealing with the topology mismatch problem in such
architectures. In structured P2P systems, the construction of the routing tables
in every participating node determines the efficiency of message forwarding.
Therefore, routing tables that represent the underlying physical structure well
can achieve increased performance. For this reason, the proposed methods for
topology mismatch problem basically use different levels of proximity
information to optimize them towards this direction. The algorithms are here
categorised based the previous work done by
\cite{castro_proximitydht_2002,castro_topawareroute_2002,ratnasamy_openq_2002}.

%{\sethlcolor{yellow}\hl{HA:  Talk about Tapestry, Pastry, CAN and Chord briefly here!}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Geographic Layout}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Inside geographic layout category fall those kind of methods which practically
position physically close by nodes together in the application space as well.
This is done by adjusting and maintaining the routing tables of all involved
peers accordingly by exploiting proximity information that describe the overall
geographic positioning of the peer nodes. One popular method for detecting and
constructing a geographic view of the underlying network is the use of well
established servers dedicated to landmarking the physical space, pretty much as
discussed in Section \ref{sec:landmark} of landmark based proximity in the
unstructured P2P networks. The same disadvantages of difficult and inaccurate
positioning and the cost of deploying and maintaining these landmarks apply in
this case as well.
% TODO: maybe a discussion sentence bellow
Even though managing the overlay structure based on the geographic layout of the
nodes improves the query efficiency of the system, on the other hand, it tends
to create hotspots, and the needed failure resilience is undermined by the fact
that close by nodes are more likely to suffer collective failures.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Global Soft-State}

\begin{figure*}
\centering
  \includegraphics[scale=0.8]{img/algorithms/global_softstate}
\caption{(left) Position of $p$ in the space of $3$ landmarks. (right) Position
of $p$ on the map.}
\label{fig:global_softstate}
\end{figure*}

\textit{Global Soft-State} \cite{xu_globstate_2003} builds a global map to help
choose shorter routing paths, combining the landmark binning method and small
scale distance probes to reveal the proximity properties of the underlying
network to the overlay. Figure \ref{fig:global_softstate} presents the landmark
space and the mapping of the position of the nodes. This global view of the
state is made available to all nodes in order to help them find the best way to
route their messages. \textit{Global Soft-State} operates in two main stages,
generation and use of the proximity information. For generating the proximity
information a hybrid approach is proposed, which uses landmark clustering as a
preprocessing step in order to select a number of potential nearest neighbour
candidates and then refine the selection by incorporating a
round-trip-time-based scheme to ultimately choose the closest node. For using
the proximity information, the algorithm chooses a different path from the
classic gossiping approaches for constructing and maintaining the overlay. It is
based on landmark clustering for strategic placement of proximity information on
the overlay enabling any node to access such information using a landmark number
that reflects its physical position in the network. For various logical
regions\footnote{This might be a high-order zone in the eCAN\cite{xu_ecan_2002}
context or a set of nodes sharing a particular prefix in overlays such as
Pastry.} maps of physical information are built and published where each node
may appear in a maximum of $log\left( N \right)$ such maps. To dynamically adapt
to changing network conditions, a node subscribes to relevant \emph{soft states}
that utilize a notification system in order to initialize any necessary
neighbour re-selection.

Maintaining several host states at different layers, makes any content migration
costly. Additionally, the method does not make any continuing effort to remap
the overlay structure after a node successfully joins, in order to adapt its
state to any condition change. Although this approach greatly reduces the
routing latency to far nodes, it is unable to dynamically identify nodes that
are close to routers and gateways in order to construct the secondary overlay.
Nevertheless, static recognition of such nodes is currently done based on BGP
reports and pre-chosen landmarks, sacrificing the self-organising attribute of
traditional DHTs.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Mithos}
\textit{Mithos} \cite{waldvogel_mythos_2003} is a P2P protocol which
incorporates directed incremental probing to find near optimal node placement,
and is classified by its authors as an integration of geographic layout and
proximity routing overlay optimization methods.

The bootstrap phase of \textit{Mithos} starts with a subset of existing members
as the first set of candidates, and while the iteratively closest nodes are
detected by probing the neighborhood, the candidate neighbor list is updated. In
order to avoid a local minima \textit{Mithos} probes all the neighbours within a
two hop distance from the current minimum before concluding the process.

After finding its first neighbour, the newcomming peer is assigned an ID using
information gathered during the iterative neighbour selection phase. Virtual
coordinates are assigned to the newcomming peer by using the distances of two
closest nodes and their neighbours, so that Euclidean distances between the node
and all known hosts predict the network latency between
them\cite{cox_vivaldi_2004}.  The benefit of these synthetic coordinates are
that they are explicitly used as the node's ID, and distance computations among
nodes can be done in ID space without requiring physical measurements.

The last step of the algorithm is the interconnection among neighbours.
\emph{Mithos} uses a \emph{quadrant}-based mechanism according to which each
node establishes a link to the closest neighbour in each quadrant. During
forwarding, the next hop is performed towards a neighbour in the same quadrant
as the final destination. The newcoming node may not know of other neighbours
in all quadrants, therefore, the node first identifies neighbours in all
quadrants using a mechanism based on ideas similar to a perimeter
walk\footnote{Used in Greedy Perimeter Stateless Routing (GPSR) protocol.} and
then improves the results using parallel path processing by taking into account
further geometric properties of node relationships.

One major limitation of \textit{Mithos} is that the protocol cannot handle
dynamic arrivals and removals from the network, as is happens in mobile
networks.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf LAPTOP}
\emph{LAPTOP} \cite{wu_laptop_2007},  organises the overlay into a tree-based
hierarchy with main focus on reducing hops during message routing as well as
minimizing maintenance overhead. Additionally, a caching scheme is also
incorporated so as to further reduce routing table update costs. The authors
theoretically show that in \textit{LAPTOP} routing path length is bounded by
$O(log_d N)$ and node joining and leaving in the overlay network is bounded by
$O\left( d log_d N \right)$ hops in a balanced overlay tree, where $N$ is the
number of nodes, and $d$ is the maximum degree of each node. \textit{LAPTOP}
implements the geographical layout approach  and constructs a geographical
layout in a self-organizing and efficient fashion, by estimating the
round-trip-times to a small number of nodes in the overlay network in order to
make them roughly aware of their physical distances among them.

Each node is assigned an address in a dotted format (e.g 1.3.4). Each octet
ranges from $1$ to $d$, where $d$ is the maximum degree of the nodes. The
assignment process is done by appending a unique octet to the address of each
nodes
parent, while root node is assigned address 1.  The routing scheme is similar to
the longest-prefix IP matching scheme. At each forwarding hop, any message
travels up the tree until the first common ancestor of both source and
destination nodes is reached and then starts descending to arrive to its target.
During tree traversals, special entries in the routing tables, called
\emph{routing cache}, are maintained in order to increase routing efficiency and
achieve finer load balance. Caching enables a node to forward a message to a
better longest-prefix match than that of its direct ancestor making a larger,
quicker and more cost effective step through the overlay and toward the
destination. To improve scalability, the number of children nodes and the size
of the routing cache are limited.  In terms of overlay maintenance, LAPTOP
incorporates a simple \emph{heartbeat}-based technique where each parent node is
responsible for monitoring its children.  At join process the newcomming node is
assigned its level label as well as its address by its parent node. Additionally
it initializes its routing table (with normal and caching entries) as it
traverses the overlay in search for its parent node.  During a graceful
departure, the referred node checks for children in the overlay. If it does not
have any, it simply notifies its parent and leaves. If it has, it selects the
child node with the lowest round-trip-time to it in order to take its place so
that the locality property is preserved.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Proximity Routing}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Proximity routing does not require routing tables to be built using any
knowledge about network proximity. On the other hand it exploits such knowledge
in order to choose the best next hop during routing a message. This approach
balances between choosing the node that will further progress the routing
towards the destination and choosing the closest entry in the routing table, in
terms of network proximity.
% TODO: looks like discussion material
Thus, it is relatively less effective than
geographical layout when applied to CAN(-like) implementations. Moreover, the
technique has been incorporated into a version of Chord causing an increase on
the overhead of node joins and the size as well as maintenance cost of finger
tables.
%
%\cite{dabek_cfs_2001} proposes a server selection scheme for the Chord DHT, on
%the domain of proximity routing selection. In \emph{CFS}, each node predicts
%the entire lookup latency as a function of the total number of nodes and the
%average overlay next routing peer. The problem is that it is very difficult to
%have a clear picture on the total number of nodes and the average hop latency
%from the local. This leads to rough estimations that consequently decreases
%overall performance.
%
The methods that use proximity routing and their details are listed below.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Proximity in Kademlia}
Before discussing {\em Proximity in Kademlia}, it would be helpful to briefly
summarize the highlights about the {\em Kademlia} protocol. {\em Kademlia}
\cite{maymounkov_kademlia_2002} is a distributed hash table (DHT) based
protocol for P2P networks, which uses an iteration based lookup algorithm. The
protocol uses the standard $160$ bit ID system for nodes and locates the nodes
in a prefix binary tree, where IDs are used as prefixes. The iterative lookup
operations are done over this prefix binary tree, which converges to logarithmic
lookup times. The ID's are assigned randomly, therefore in Kademlia, there is no
proximity control, which results in inefficient use of the underlying network
during lookup and retrieve operations.

{\em Proximity in Kademlia} \cite{kaune_pkad_2008} introduces proximity controls
over the base Kademlia protocol to optimize the underlying network usage of the
protocol. The authors define an abstract {\em underlay metric} that calculates
the suitability of establishing a communication link between peers as a cost
function, based on the used proximity criteria (RTT, ISP locality, etc.).
{\em Proximity in Kademlia} adapted both {\em proximity routing} and {\em
proximity neighbour selection} overlay optimization algorithms for controlling
proximity. Therefore, we categorize the protocol only in this section, and skip
the discussion in the further {\em proximity neighbour selection} section
(Sec.\ref{sec:pns}). {\em Proximity in Kademlia} also uses the MaxMind GeoIP
database\footnote{MaxMind Geolocation Technology. http://www.maxmind.com} to
detect the proximity information of a given peer based on its region, country,
or ISP location, and use this information to form clusters of nearby peers. One
other method {\em Proximity in Kademlia} uses is the Vivaldi
\cite{cox_vivaldi_2004} protocol. However, authors report that clustering
approach performs better than Vivaldi protocol based on their experiments.
Authors also report that {\em Proximity routing} protocol successfully worked
with the Kademlia protocol and improved the locality of the connections over
peers.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf CHOord considering Proximity on IPv6 (CHOP6)}
\textit{CHOP6} \cite{morimoto_chop6_2007} is designed based on the
\textit{Chord} protocol. \textit{CHOP6} roughly estimates the proximity among
nodes by exploiting the IPv6 address format and RTT information if available.
The proximity estimation is achieved by introducing a 64-bit ID scheme in which
the least significant bit part is the IPv6 global routing prefix and thus
enabling a longest prefix match scheme. The protocol is designed based on the
observation that it is possible to estimate a node's geographical location by
simply examining the upper 32-bits of its IPv6 address. Moreover, similar to
\textit{Chord}, \textit{CHOP6} uses a finger table, whose entries hold more than
one candidate node.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Chord6}
\emph{Chord6} \cite{xiong_chord6_2005} is another \textit{Chord} variant that
tries to exploit the hierarchical features of IPv6 in order to create a
substrate that reduces interdomain traffic between service providers.
\textit{Chord6} is based on the original Chord protocol, and the main difference
from the original protocol is in the identifier definition. Therefore, the
approach can be easily portable to other DHTs such as CAN, Pastry and Tapestry.
In Chord6 the identifier contains two parts: the higher bits are obtained by
hashing the node's IPv6 address prefix of specific length, while the remaining
lower bits are the hash value of the rest of that IPv6 address. As a result of
this assignment, nodes in a domain will be mapped onto a continuous key space on
the overlay network, which avoids unnecessary message forwarding across
different service providers, thus minimizing overall routing cost.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Proximity Neighbour Selection}\label{sec:pns}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textit{Proximity Neighbour Selection} constructs the routing tables using
proximity knowledge. The proximity information used in this method is different
than the landmark based systems described in the \textit{Geographic Layout}
section, as time-to-live values between nodes, or directly node ID prefixes are
used to detect proximity.  Tapestry, Pastry, and CAN successfully implemented
the proximity into their algorithms by using this approach.  The routing
protocol in Pastry is based on longest node ID prefix matching, while CAN uses
RTT values to detect close by nodes. \cite{castro_proximityp2p_2002}  reports
that proximity neighbour selection as an effective proximity based method.
Proximity Neighbor Selection based algorithms are described in detail below.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf DHT-PNS}
Chord-DHT-PNS \cite{hancong_pnsbased_2006} implements proximity neighbour
selection on top of the Chord DHT. The main goal is to use proximity information
to group physically close by nodes as neighbours in the DHT table. In order to
detect proximity, virtual network coordinates of peers are used, by using the
Vivaldi protocol \cite{cox_vivaldi_2004}. The virtual coordinates are then used
by the nodes to map to identifier space in the DHT. The space is partitioned
using a \emph{concentric circle clustering scheme} where successive cycles of
radiuses $\rho$, $2\rho$, $3\rho$ and so on, are constructed. Then the formed
annuluses are divided into $2\chi-1$ \emph{sectors}, where $\chi$ denotes the
sequence number of the annulus starting from $\chi = 1$ for the centre cycle. It
is proved in the paper, that this way each sector occupies the same area as does
the center cycle. Assuming uniform node distribution, this characteristic,
favours a more load balanced clustering operation. Every sector in this $2D$
coordinate space is mapped to a unique \emph{region} in the DHT space forming a
multi-layer node identifier space.  Thus, any nodes that belong to the same
sector, are mapped to the same region as well, preserving their proximity
relationship unveiled by the use of the Vivaldi protocol. The individual pieces
are mapped to the identifier space uniquely, allowing logarithmic lookup
operations with high probability on Chord.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf IP-based Clustering (IPBC)}
Proximity neighbour selection algorithms use probing and other techniques
to detect proximity. However, such methods are either not precise enough, or
overload the network. \emph{IP-based clustering}
\cite{karwaczynski_ipbc_2007} is a proximity neighbour selection based
algorithm in which the authors propose to use the IP address prefixes (16 bit
for IPv4) in order to detect proximity.  \cite{freedman_iploc_2005} states that
$97\%$ of prefixes larger than $24$ bits belong to a single geographical
location. However, using less number of bits creates less precise results and
more number of bits increase the burden on the network and reduce the possible
number of neighbours. Therefore, a careful selection is required in terms of
performance/accuracy tradeoff. The {\em IP-based Clustering} generates a key and
its prefix and stores the later in the DHT itself, so that any newly joining
nodes with the same IP prefix can query it and identify all the
neighbours with the same one relatively easily. Nodes periodically update
their entries in the DHT by removing those who leave, or those who has been
timed out with no further activity detected from the node.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Cone}
\textit{Cone}\cite{wang_cone_2007} extends Chord using proximity neighbour
selection topology optimization algorithm. The proximity information is
generated using landmarks and round-trip-time-based distances to landmarks.
\textit{Cone} uses a two-layered identifier space. The first, named Chord-layer
identifier, denoted as $Id_{Chord}$, is the same as in Chord. The second is the
Cone-layer identifier, $Id_{Cone}$ which is constructed by two component
identifiers. The first, known as \emph{group identifier (gid)} denotes a
relevant group the node belongs to while the second, namely \emph{local
identifier (lid)} indicates the local identifier within the group. The group
concept, which is introduced here, is a way of dividing nodes according to a
common $Id_{Chord}$ prefix.  The structure of a Cone overlay, retains the
Chord's circular topology. The difference lies on the fact that, now, two rings
are created. A big ring, where nodes with the same \emph{gid} are arranged at
each position. Each of these positions are a smaller ring for the particular
group's \emph{lid}s. The routing is achieved in both clockwise and
counter-clockwise directions in the big-ring, for which two routing tables are
maintained, namely \emph{front} and \emph{back} finger tables. Entries in these
tables, display physical network proximity with the current node. Moreover, a
third table called \emph{group} maintains information about other online peers
within the current node's group in a way that entries are now close in the ID
space.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf SAT-Match: Self-Adaptive Topology Matching}

\begin{figure}[ht]
\centering
\subfigure[Example physical topology of nodes.] {
  \includegraphics[scale=0.6]{img/algorithms/SAT_match}
  \label{figure:sat_match:before}
}\qquad\qquad
\subfigure[(left) A jumps to B in order to match the logical topology to the
physical. (right) State of the logical topology after the jump.] {
  \includegraphics[scale=0.6]{img/algorithms/SAT_match2}
  \label{figure:sat_match:after}
}
\caption{The SAT-Match algorithm.}
\label{figure:sat_match}
\end{figure}

\textit{SAT-Match} \cite{ren_satmatch_2004} is a protocol that specifically
tries to solve the topology mismatch problem by mapping the overlay network as
close as possible to the physical one. Even though no theoretical proof is
given, valid results are obtained and demonstrated through experimentations.
Similar to the other approaches in the proximity neighbour selection,
\textit{SAT-Match} uses probing to detect close by peers and obtain proximity
information. However, as an additional feature, \textit{SAT-Match} peers can do
selective jumps to adjust their positioning in the DHT, if it reduces the stretch
of its one-hop neighbourhood. Figure~\ref{figure:sat_match} demonstrates the
selective jump on an example topology. The stretch is defined as the ratio
between the average logical and physical latency of a link. It is reported that,
due to the selective jumps, the \textit{SAT-Match} achieves $40\%$ reduction in
link stretch, and when used with the \textit{Landmark Binning} approach (see
Sec. \ref{sec:landmark_binning}), the reduction rate increases up to $60\%$.
For dynamic environments, with frequent node arrival and leave,
\textit{SAT-Match} scales much better than \textit{Mithos}, due to its self
adaptation mechanism and selective jumps.

\emph{SAT-Match} uses a small TTL value for the probing messages in order to
reduce redundancy\cite{jiang_lightflood_2008}. This process begins as soon as
the node joins the network using a DHT mechanism. Each probing message contains
information about the source and a small TTL value. The recipient of such a
message returns information about itself to the source and forwards the probing
message to its neighbours if the TTL is non-zero. The discovered nodes are
referred to as $TTL-k$ neighbourhood of the source node based on the $TTL$
distance to the source node.

Blindly selecting the peer with the smallest RTT as neighbour is, generally, not
the optimal decision to make in order to achieve global \emph{stretch} reduction.
In a structured scheme, when a node jumps to connect to a physically close node,
it may need to connect to other distant nodes to maintain the structure's
integrity, thus creating an overall increase in the overlay's \emph{stretch}.
The two nodes with the smallest RTT are then used in order to select one zone to
jump in this phase. The algorithm is as follows: The source node $S$ calculates
the stretch change of its $TTL-1$ neighbourhood and that of the $TTL-1$
neighbourhood of the first of the previously selected peers. These calculations
are made as if the jump has been made. If the stretch reduction is over a
predefined threshold the jump is performed, otherwise the second selected
candidate is picked and the same computations are performed. If again, the
threshold is not met, then no jump is ultimately done. In case of a jump, this
is performed as a combination of \emph{leave} and \emph{join} operations, in the
CAN context.

%\begin{figure}[h!]
\hspace{-3ex}
\begin{center}
\footnotesize
%\begin{landscape}
\begin{longtable}{
|>{\columncolor[gray]{.7}}m{0.2\columnwidth}
|>{\columncolor[gray]{.9}}m{0.4\columnwidth}
|>{\columncolor[gray]{.8}}m{0.2\columnwidth}
|>{\columncolor[gray]{.9}}m{0.2\columnwidth}
%|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
%|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
%|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
%|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
|}
\caption{Decentralized Structured Algorithms}\label{fig:struct_compare_table}\\
\hline
\rowcolor[gray]{.5}
\textbf{Algorithm} & \textbf{Overlay structure} & \textbf{Base protocol} & \textbf{Scalability} \\
\hline
\endfirsthead
\multicolumn{4}{c}%
{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
\hline
\rowcolor[gray]{.5}
\textbf{Algorithm} & \textbf{Overlay structure} & \textbf{Base protocol} & \textbf{Scalability} \\
\hline
\endhead
\hline \multicolumn{4}{r}{\textit{Continued on next page}} \\
\endfoot
\hline
\endlastfoot

\hline
\textbf{Global Softstate} &
\textbf{Geographic layout} Uses first landmark clustering
then measures RTTs to identify close nodes &  &  \\

\hline
\textbf{Mithos} &
\textbf{Geographic layout} Uses nodes as topology landmarks and directed
incremental probing to optimize topology & & Scales well as all
operations are local ??? \\

\hline
\textbf{Self-Adaptive Topology Matching} &
\textbf{Proximity neighbour selection} Uses lightweight probing and
selective jumps to optimize the topology & CAN &  Better than Mithos \\

\hline
\textbf{Delay Aware P2P System} & \textbf{} & & \\

\hline
\textbf{VERSION OF CHORD - DHT-PNS} &
\textbf{Proximity neighbour selection} Uses Proximity Neighbour Selection and the Vivaldi
system & Chord  &  \\

\hline
\textbf{MAY OMIT- VERSION OF CHORD - Quasi-Chord} &
\textbf{Proximity based}  & Chord  &  \\

\hline
\textbf{LAPTOP} &
\textbf{Geographic layout} Hierarchical overlay structure  &  & routing path length $\log{_d N}$,
join/leave overhead $d\log{_d N}$ \\

\hline
\textbf{IP-Based Clustering} &
\textbf{Proximity neighbour selection} Proximity neighbour selection based on longest common
prefix of IP addresses &    &  \\

\hline
\textbf{CHOord considering Proximity on IPv6} &
\textbf{Proximity routing} Uses IPv6 address format to provide proximity &  Chord
 & Better than Chord \\

\hline
\textbf{Proximity in Kademlia} &
\textbf{Proximity routing} Applies  proximity neighbour selection (PNS) and proximity route selection (PRS)
to Kademlia & Kademlia &   \\

\hline
\textbf{Cone} &
\textbf{Proximity neighbour selection} Uses proximity neighbour selection (PNS) & Chord  & Better than Chord \\

\hline
\textbf{Dynamo} &  &  &  \\

\hline
\textbf{MAY OMIT-BADLY WRITTEN-PChord} &
\textbf{Proximity based}  &  Chord  & \\

\hline
\textbf{AChord} &  & & \\

\hline
\textbf{Chord6} &
\textbf{Proximity routing} Uses IPv6 hierarchical address format to cluster
topologically close nodes & Chord  &  \\

\hline
\end{longtable}
%\end{landscape}
\end{center}
\vspace{-2.5ex}
\vspace{-2.5ex}
%\label{fig:struct_compare_table}
%\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Discussion on the Algrithms for Structured Architectures}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this section we presented the state of the art for the algorithms applied to
structured P2P systems. We categorised it based on their philosophy they
incorporated in order to alleviate the topology mismatch problem into overall
three categories following the previous work from
\cite{castro_proximitydht_2002,castro_topawareroute_2002,ratnasamy_openq_2002}
namely
\begin{inparaenum}[\itshape i\upshape)]
  \item geographic-layout-based,
  \item proximity-routing-based, and
  \item proximity neighbour selection based
\end{inparaenum}
. Following is a discussion on the above categories trying to point out the
advantages, disadvantages and novelties introduced by each and by all of them.
A quick overview can be seen in Table \ref{fig:struct_compare_table}.

In Section~\ref{section:structured}, we presented the state of the art
structured decentralized P2P algorithms in three different categories based on
their handling of the network structure to tackle the topology mismatch problem:
geographic layout based, proximity routing based, and proximity neighbour
selection based. In this section, we present a final discussion of all the
methods including the advantages, disadvantages, and novelties related to solve
the topology mismatch problem. Table~\ref{fig:struct_compare_table} presents an
overview of all the structured decentralized algorithms described earlier in
this section.

Structured P2P network algorithms use a global distributed hash table or a
prefix tree structure to uniquely lookup peers or their data in the overlay
network. As all the data is kept within the overlay, each node behaves as a
client and a server, therefore nodes join and leave according to rules
determined by the integrity of the global data structure. The main advantage of
the structured P2P topology is that by the help of the global data structure,
peers or their data can be found within the network even if there is only a
single copy of that item present. However, each node join and leave creates
maintenance overhead for the network due to updates required by the global data
structure, and for networks with frequent node arrivals and departures the
topology uses valuable network resources just to update the global structure.
Nodes join the network by using a key value, which determines the location and
the neighbourhood of the new node within the network. However, assigning
random key values to the newly inserted nodes creates non-optimal matching with
the underlying physical network topology, therefore, increasing the overhead of
the network even more. One solution for handling the topology mismatch problem
is to consider the proximity of the peers when generating the key and joining
the node to the network, so that nodes within the same network domains are
selected as peers, or neighbours, during the overlay topology construction. In
this chapter, we have described three such approaches to optimize the topology
matching problem: geographic layout, proximity routing and proximity neighbour
selection.

In geographic layout, nodes try to estimate the geographic positions of the
peers, and construct the DHT considering the proximity of the peers. Landmark
servers and RTT measurements are two popular methods, which can also be used in
conjunction, to discover physically close by peers over the network. However, as
discussed in Section \ref{sec:landmark_binning}, these methods do not always
give reliable estimates for the node positions over the internet. The landmark
servers are not self-organizing and have maintenance overheads. To serve a
P2P network with millions of peers, multiple landmark servers
distributed over the whole world is required, which is hard to manage. The RTT
measurements also can measure the delay between peers, but it is a greedy method,
which can result non-optimal overlay topologies especially if close by nodes
have low bandwidth connections among themselves.

Proximity routing does not discover or store proximity information for the
peers, however, during package forwarding, nodes that have lower latency to the
destination, or with a closer key to the destination is selected. As no
proximity information is used, the system is based on a greedy approach and it
usually selects low latency paths over the overlay, which maps to suboptimal
longer paths on the physical topology. For these obvious reasons, proximity
neighbour selection is generally considered as a superior method to proximity
routing, however, joint uses of these two protocols are also possible.

Proximity neighbour selection approach is similar to proximity routing, with an
exception that during forwarding, the proximity information of the peers are
also considered. Therefore, for each application, proximity information is
extracted and used during routing. Depending on the application, the proximity
information can be the node sharing the same prefix with the current node, or
TTL to detect close by neighbours. Tapstry and Pastry are two popular methods
implementing the proximity neighbour selection among others mentioned above in the
proximity neighbour selection section.

### 2014-01-21 unstructured.tex
2012.05.29 (Vassilis) Monolithic file broken down to multiple smaller ones (per section) for better parallel contribution. Additional material on more recent algorithms for unstructured networks. Now also covering the BitTorrent protocol. Merged Huseyin's figures from previous contribution and code refactored for better referencing. Fixed bibliography bug (references appearing as ?).

\section{Decentralized Unstructured Algorithms}
\label{section:unstructured}

This section is dedicated to the unstructured P2P architectures. In the next
paragraphs,algorithms for tackling the topology mismatch problem on such
networks are visited, analysed and categorised based on the use of the overlay
structure, messaging approach for peer communication and techniques proposed for
detecting proximity in order to optimise overlay topology. In short, the methods
that are highlighted here are divided into
\begin{inparaenum}[\itshape i\upshape)]
  \item broadcast-based,
  \item cache-based,
  \item overlay optimisation-based, and
  \item landmark-based
\end{inparaenum}
 approaches.

% TODO: READ A Near-Optimal Algorithm Attacking the Topology Mismatch Problem in
%Unstructured Peer-to-Peer Networks (This is an Approximation alg for Top-mis
%problem)

%{\sethlcolor{yellow}\hl{
%HA: TALK ABOUT GNUTELLA HERE!!!!}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Broadcast Based}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In the first realisation of the Gnutella protocol, each query request received
by a peer is then forwarded to all its neighbours, clearly generating
unnecessary traffic over the network. The algorithms highlighted in this
paragraph are referred to as broadcast-based because they enhance the "naive"
broadcasting approach described above and prefer forwarding messages to a
selected subset of a node's outgoing links. The selection criteria vary
depending on the algorithm but they mainly use one or multiple statistical
metrics that span from the obvious ones, like the latency of the link between
the nodes, to more sophisticated alternatives (that may even take into account
application level requirements) such as the reliability of the neighbour. The
forwarding-based approach has the advantage of enhancing search efficiency but
on the other hand suffer drawbacks like the drastic reduction of the search
scope because of the restrictions of the broadcasting policies as well as
limitations when scaling to large networks.

%Expanding the search scope, on the other hand, is no easy task
%because the overhead of forming multicast trees is proportional to the
%multicast group size.
%forwarding based schemes do not consider dynamic joining
%and leaving of peers so they do not scale well on dynamic environments.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Improving search in peer-to-peer networks}

\begin{figure}[ht]
\centering
\subfigure[Iterative deepening with three levels.] {
  \includegraphics[scale=0.4]{img/algorithms/iterative_deepening}
  \label{figure:dbfs:iterdeep}
}\qquad\qquad
\subfigure[Directed BFS.] {
  \includegraphics[scale=0.4]{img/algorithms/directed_bfs}
  \label{figure:dbfs:dbfs}
}\qquad\qquad
\subfigure[local indices with radius size equal to $2$.] {
  \includegraphics[scale=0.4]{img/algorithms/local_indices}
  \label{figure:dbfs:localindx}
}
\caption{}
\label{figure:dbfs}
\end{figure}

\cite{yang_improvep2psearch_2002} proposes an easy to implement and practical
solution to the inefficiency problem caused by blind flooding in Gnutella-like
P2P file sharing protocols. The paper replaces blind flooding with three
approaches, namely \textit{iterative deepening}, \textit{directed BFS}, and
\textit{local indices}. In \textit{iterative deepening}, as shown in
Figure~\ref{figure:dbfs:iterdeep}, the search is performed on a BFS tree with
multiple preset depths. The depth limit is iteratively increased by the source
node for each query based on the quality of the results. The source node, after
examining the results, may issue a new request by increasing the depth limit,
which will trigger the nodes at the last depth level to resume the search. The
iterative approach avoids to start the search from scratch at each iteration and
reduces the load of the nodes on the upper levels of the tree. The major
drawback of \textit{iterative deepening} is the delay between successive
iterations, as the source node needs to examine the results at each iteration
before deciding to quit or resume the query.  The \textit{directed BFS}, as
shown in Figure~\ref{figure:dbfs:dbfs}, tries to avoid this delay by
forwarding the query messages only to a selected set of neighbours, in which the
selection criteria varies from the number of results received previously,
distance in terms of hops, bandwidth, or the query load of the neighbour. In
\textit{local indices}, as shown in Figure~\ref{figure:dbfs:localindx}, each
node sustains a local data index of the nodes within a radius of $r$ hops from
itself and uses this local index to remotely query the neighbour nodes without
generating additional traffic. \textit{Local indices} greatly reduces the
aggregate bandwidth usage of the network and improves query efficiency, however,
updating the indices in cases with frequent node joins and leaves introduces a
serious overhead to the system if the radius is kept broad.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Gia}
\emph{Gia} \cite{chawathe_gia_2003} is an effort to solve the scalability
problem of the unstructured P2P file sharing systems, in particular Gnutella.
The main novelty in the design of Gia is the replacement of the blind flooding
approach of Gnutella with random walks \cite{lv_randomwalks_2002}. Although
random walks is a step in the right direction, issuing only a single copy of the
query within the network reduces the search scope, thus affecting negatively the
success rate of the query.  In order to overcome this limitation, Gia introduces
a token-based flow control mechanism, which is essentially an intelligent flow
control algorithm that gradually redirects the queries to nodes which are more
likely to answer. In order to prevent overloading of nodes with query requests,
Gia uses a token-based flow control algorithm in which each node announces the
number of query requests it can handle in terms of tokens to its neighbours, so
that neighbours only forward query requests to nodes that they previously
received tokens from. Gia also acknowledges the heterogenity in peer bandwidth,
processing power, disk speed, etc, of the nodes in P2P networks and uses this
information when connecting nodes to each other and by using a topology
adaptation algorithm, Gia ensures that high capacity nodes have high degrees and
low capacity nodes are within short proximity to high capacity ones. Although
the topology adaptation algorithm Gia uses improves the scalability of the
network, it does not help much in solving the topology mismatch problem, since
it does not consider the underlying physical topology.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Distributed Cycle Minimization Protocol}
We have already mentioned Gnutella's inefficient broadcasting mechanism in
flooding messages across the network with no actual targeting. This leads, to
nodes that are far from the path towards the query request gratification, to
spend their valuable resources to receive, queue and process messages that they
shouldn't have received in the first place and eventually simply forward them
blindly to continue this vicious cycle. Except from that, though, an other case
of wasting network resources is when a node that is on the right path for the
query, receives the same message multiple times from different neighbours. The
reason for these duplicate messages is cycles in the forwarding paths. Actually,
this duplicate packet problem seems to hurt more the active nodes; those with
higher capacities and bandwidth that contribute the most to the network.
Distributed Cycle Minimization Protocol (DCMP) \cite{zhu_dcmp_2008} introduces a
method to
remove the cycles and eliminate the duplicate packets, without sacrificing
overlay connectivity. In DCMP, once a cycle is detected, the most powerful node
in the cycle is elected as the \emph{Gate Peer} and the cycle is then destroyed
by cutting that special link which will result in the minimisation of the
distance to the Gate Peer for all the nodes that participated in that cycle. The
process is managed by using two special message types namely \emph{Information
Collection (IC)} and \emph{Cut Message (CM)}. One disadvantage of DCMP is that
since the distance a forwarded message can travel is limited by the TTL value,
which is practically $7$ in most cases DCMP cannot detect cycles formed by more
than 7 peers. Even though cycle elimination improves the network performance, it
does not actually solve the topology mismatch problem.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cache Based}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\subsection{Cache Based}
%
%Caching based protocols are effectively used to reduce traffic costs and
%response
%times. The caching policy varies depending on the way protocol handles the
%index and the
%content. Centralized P2P systems
%use central index servers, while local caching systems, such as KazaA, use
%super peers
%to cache indices in a distributed way. Content caching is also possible in P2P
%systems, where nodes cache the forwarded content for further retrievals.
%Although caching has the above mentioned advantages,  duplication
%of messages still exist, which limits the scalability of these approaches.
%Therefore, cache based approaches are analyzed in the following categories:
%  \begin{itemize}
%    \item \emph{data index caching},
%    \item \emph{content index caching},
%    \item \emph{centralized}, and
%    \item \emph{local}.
%  \end{itemize}
%
Caching is a widely used technique to exploit locality and minimise redundant
transfer of data, initially adopted by web and file server application
environments with great success. Since peers in a P2P system also work as
servers, it is intuitively expected that P2P file sharing systems can also
benefit from caching. However, two of the main characteristic properties of P2P
systems, nodes frequently joining and leaving and query lifetime relatively
short (at least compared to those needed to be served by web servers) , makes
caching in P2P networks a non-trivial issue. There are usually two levels of
caching possible in file sharing systems, caching indices or pointers to data or
caching the data itself. Caching is already implemented and successfully used in
some commercial P2P systems. One popular example is KazaA, which uses caching of
indices on its super-peer level. In the paragraphs below, caching-based
algorithms are presented and explained in more depth.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Replication Strategies in Unstructured Peer-to-Peer Networks}
\cite{Cohen02} aim to improve the inefficient blind search algorithm by
replicating the data in a P2P network. The main intuition behind the
idea of replication, or using cached copies, is that as the number of copies for
each item increases in the network, it would be easier for a search algorithm,
even a random one, to locate these items. In order to analyse the feasibility of
such an approach, the authors investigate three different replication
strategies, namely uniform, proportional, and square root allocation. In the
uniform model, the copies of items are uniformly replicated in the network,
while in the proportional model, the items are replicated based on their query
rate, so that frequently queried items are replicated more. The square root
allocation is a strategy proposed in this paper, which is a model between the
uniform and proportional allocation.  In order to evaluate the outcome of the
replication models, the overall costs of successful and unsuccessful searches in
the network are compared. Results argue that the uniform allocation model
minimizes the maximum search size, therefore reduces the time spent on an
unsuccessful search. The proportional model, on the other hand, promotes the
more frequently queried items by replicating them more, therefore decreases the
search time for popular items, but suffers when needing to locate the rare
items. The authors also claim that the expected successful search size is the
same for uniform and proportional models, and any approach between them would
behave much better. Therefore they propose the square root allocation approach,
which is a replication strategy that minimizes the expected search size of
successful queries in P2P networks.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Tracing a large-scale Peer to Peer System: an hour in the life
of Gnutella}
\cite{Markatos02} analyses Gnutella network traffic traces and by concluding
there is locality among query requests, proposes a caching algorithm that tries
to exploit these findings . The analysis of the trace data reveals other
important facts about the structure of the Gnutella network and the query data.
One significant observation is that the geographic locations of clients do not
have a correlation with the number of query requests they receive. This is an
obvious result of the topology mismatch problem caused by the overlay structure
of the Gnutella network. Gnutella's traffic is observed to be bursty both for
query requests and query responses, even in longer intervals. It is observed
that a peer receives 50 query messages per second on average! Moreover nine out
of ten queries do not generate any response due to the inefficient design of
the Gnutella network. When developing a caching system to exploit locality,
applying an approach similar to web caching does not fit well with P2P systems.
Caches in P2P systems not only have to consider the query string, but also
the TTL value, the source of the query, and the time of the query as well.  In
general, even though optimum caching is hard to achieve, it is reported that
it improves the overall performance of the Gnutella network.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Overlay Optimization Based}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The overlay optimization based protocols modify the topology of the P2P network
using various techniques. The two commonly used approach that is investigated in
this section are creating spanning trees using connection graphs, and creating
clusters of physically close nodes. Spanning tree based approaches construct
rich graphs based on the network connections and build minimum spanning trees on
these graphs. Even though spanning trees provides efficient query performance,
the construction and the update costs of the spanning tree, especially in
dynamic environments with many nodes joining and leaving the network, cause
large traffic overhead to the corresponding
network \cite{chu_esm_2000,chu_esm_2002}. The cluster based approaches on the
other hand select to link physically closer nodes with each other, therefore
significantly shrinking the search scope of the system. Added to that, commonly
used methods for correct determination of physical node positions across
Internet do not always return reliable results, therefore mapping accuracy is
not guaranteed. Bellow, the details of the algorithms that use the spanning tree
based or cluster based topology optimization methods are presented.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Narada}

\begin{figure}[ht]
\centering
\subfigure[Underlying network with edge costs.] {
  \includegraphics[scale=0.4]{img/algorithms/narada}
  \label{figure:narada:underlying}
}\qquad\qquad
\subfigure[Peer A sends a message to the rest using regular broadcast.] {
  \includegraphics[scale=0.4]{img/algorithms/narada2}
  \label{figure:narada:regu}
}\qquad\qquad
\subfigure[Peer A sends a message to the rest using end system multicast.] {
  \includegraphics[scale=0.4]{img/algorithms/narada3}
  \label{figure:narada:multicast}
}
\caption{A visualisation of the Narada protocol.}
\label{figure:narada}
\end{figure}

Narada \cite{chu_esm_2000,chu_esm_2001,chu_esm_2002} is a generic protocol for
creating self-adapting overlay networks, that can achieve application-layer
multicast communication without requiring IP multicast infrastructure at the
network layer. IP multicast is the term reffering to the method of sending IP
datagrams to a group of receivers in a single transmission. Even though the
method is available some years now, IP multicast has not taken off as
anticipated since many believe that it violates the stateless design of the
current Internet. For this reason it is widely deployed in contained
environments like enterprises (i.e. IPTV applications - distance learning, video
conferencing), commercial stock exchanges, and multimedia content delivery
networks. Narada tries to break these constraints. Although it was not
originally designed as a P2P file sharing protocol, it has become one of the
pioneering works in the overlay networks area. The authors claim that, although
Narada is not as efficient as IP multicast, still it provides reasonable
performance, without the additional cost of implementing multicast services on
the current Internet infrastructure. Additionally, authors realised the
inefficiency caused by the topology mismatch problem and tried to encounter it
by building a richer connected graph, called a mesh, and building per source
minimum spanning trees \footnote{ENarada \cite{li_enarada_2008} used Gossip
protocol for the construction} as shown in Figure~\ref{figure:narada}. They also
keep the graph and the trees dynamically updated while nodes continue to join
and leave the network. Moreover, the protocol tries to ease the physical link
stress, the overall resource usage and the relative delay among end systems.
Unfortunately, the most imporpant limitation of Narada is that although it works
reasonably well for small groups, it does not scale well for larger networks,
therefore it is not suitable for, potentially very large, P2P file sharing
application networks.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Adaptive Overlay Topology Optimization and Adaptive Connection
Establishment}
\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{liu_auto_2003} is one
of the first attempts, along with Narada, from the P2P research community to
address the topology mismatch problem. AOTO is a distributed algorithm that uses
\emph{Selective Flooding} and \emph{Active Topology} to optimize the overlay
network topology. The \emph{Selective Flooding} algorithm creates a minimum
spanning tree over the overlay and uses this tree instead of flooding the whole
network, without shrinking the search scope. In \emph{Active Topology}, the
physical locations of nodes are estimated based on the network delay among them
and physically close nodes are connected as neighbours to revise the overlay
topology as close as possible to the physical network topology. Overall, AOTO
reports a $55\%$ performance improvement in terms of the traffic load. In
\emph{Adaptive Connection Establishment (ACE)} \cite{liu_ace_2004}, the authors
extend the idea by introducing optimizations based on the depths of the minimum
spanning trees. But since the network delay is not always a reliable estimation
method to detect the physical location of peers, the algorithm still suffers
from the discrepancies caused by mislocated nodes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Location-aware Topology Matching}
\emph{Location-aware Topology Matching (LTM)} \cite{liu_ltm_2004} proposes a
method to optimize the overlay structure of the P2P network based on the
physical topology. To achieve this, the algorithm issues a special message
called \textit{TTL2-detector} to detect the latency of two-hop neighbours
($N^2$) around each node. The latency information gathered is later used to
evolve the overlay network into a more efficient one, without reducing the
search scope. Each node examines the latency information with the neighbours and
low productive connections are dropped and replaced by more efficient ones, thus
reducing the latency on the overall network, and eliminating some potential
inefficiencies, such as unnecessary crossing of messages over autonomous system
(AS) boundaries, etc. Although LTM improves the overall efficiency of the P2P
network, it does not actually use real physical topology information, just
latency metrics, therefore it does not offer a guaranteed safety net for the
topology mismatch problem.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Scalable Bipartite Overlay}
\emph{Scalable Bipartite Overlay
(SBO)} \cite{liu_bipartite_IPDPS,liu_bipartite_2007} reduces the overhead of
creating and maintaining a minimum spanning tree cost by randomly dividing the
nodes into two groups, namely red and white and assigning different tasks to
different groups. The white peers measure distances to neighbours by using the
network delay as the metric and reports the results to red peers. The red peers,
equipped with the information of all two-hop neighbours ($N^2$), creates a
minimum spanning tree of these neighbours and assigns efficient forwarding
paths. The white peers can further optimize their positions within the overlay
if this is required.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TODO: DOUBLE CHECK IF THIS ALGORITHM GOES IN THIS SECTION/SUBSECTION
\paragraph*{\bf Adaptive Peer Selection}
Bernstein et al., contributed on the use of machine learning for building peer
selection strategies from past experience \cite{bflz_adaptpeersel_2003}. A
decision tree is used to rate peers based on information about connection
characteristics that is collected. These can be load, bandwidth, and past
uploading experience. Then a by Markov decision process is incorporated as a
mechanism to shape the policy for switching among the peers. The problem
with this approach is that the peer selection algorithm is slow due to the
learning process and the complexity of the method.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Innocuous Topology Aware (ITA) Overlay Construction}
\cite{prfm_ita_2009} intoduces an algorithm for Innocuous Topology Aware
construction of unstructured P2P networks. The paper's proposition is twoforld.
Overlay construction and search method. For the overlay construction it employs
the notion of \emph{short} and \emph{long} connections. Assuming $N$ is the
number of all network participants, $\alpha \leq 1 $ a system-wide magic number
and $x$ an $\alpha$-related latency threashold, the bootstrapping peer randomly
selects $\alpha \ast N$ close (latency bellow $x$) and
$\left( 1 - \alpha \right) \ast N$ distant (latency above $x$) nodes as its
neighbours. The search method has two steps. First, the quering node initiates a
flood to its distant neighbourhood with $TTL = 1$. Then, peers that receive a
query over a long link, start a local flood with $TTL = ttl$, were $ttl$ is a
system defined parameter.

The goal of the overlay construction is to create a network that exposes low
clustering, a beneficial characteristic of random graphs which can lead to a
larger coverage of peers with the same number of messages and reduced
duplication. This means more efficient information lookup that exposes low or no
negative impact whatsoever to the rest of the mechanisms employed in
unstructured P2P systems\footnote{For example the 1-hop replication and the
dynamic querying mechanisms of later versions of Gnutella.}. The paper reports
50\% reduction in communication latency ammong peers by cutting off some 20\% of
the IP-level message traffic generation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf EGOIST}
EGOIST \cite{egoist_2008} is a set of algorithms to construct and manage overlay
networks. It utilises a selfish approach in the sense that every participating
peer continuously updates its neigbours so as to minimise the sum of distances
to all destinations under shortest-path routing. In EGOIST, a newlly arriving
peer, randomly connects to an already participating node through a bootstrap
server. Being connected, it starts receiving info throught the link-state
mechanism and thus, after some time, it has a complete picture of the overlay
graph. Then it estimates the delay to all other nodes in order to determine its
potential neighbours and ultimatelly connect to the overlay using some
policy\footnote{For example, minimisation of the average delay to all its
neighbours.}. It is obvious that there is extensive resource usage for updating
the wiring of the all nodes in the system. The authors claim that they can
reduce the load imposed by the monitoring and announcing process from $O(n^2)$
to only $O(kn)$, where $n$ is the number of all nodes in the network and $k$ is
the number of links a node chooses to establish.

% TODO: CHECK IF THE FOLLOWING BITTORRENT ALGORITMS BELONG IN THIS SUBSECTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Biased Neighbour Selection}
Bindal et al. \cite{rpc_bitbiased_2006} claim to strengthen BitTorrent
protocol\cite{c_bittorrent_2003} locality by selecting most of a peer's
neighbours to be out of the same ISP, while preserving the near optimal download
performance of the protocol itself. BitTorrent systems employ most of the time
mechanisms that can be proven very aggresive to an ISP's networking and
accounting, basically, due to their lack of knowlegde of AS boarder limits.
BitTorrent specification, by default allows for each peer, 35 connections.
Biased neighbour selection experiments on some number $k$ which denotes a number
of neighbouring peers which are not on the same ISP than the peer at hand.
Consequently, the remaining $35 - k$ peers are from the same ISP. These $k$
nodes are used in order to preserve a more extended, global view of a network,
but in the same localize the load within the limits of a single ISP. This
prevents clients from exchanging traffic through an expensive transit link, if
there are alternative local connections that could offer the same, faster and at
no additional cost for the ISP. Implementation can either be done through
modifications on the tracker side, to identify ISP locallity, or through P2P
traffic shaping devices installed on ISP's edge routers. The paper also proposes
combination of biased neighbour selection along with bandwidth throttling and
some caching approach for near optimal results. Unfortunatelly, implementation
need, in some extend, contribution (exposure of ISP AS mappings) or
infrastructure changes (installation of P2P traffic shaping devices) on the ISP
level itself, making the adoption difficult in the general case.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Ono}
Choffnes and Bustamante propose Ono \cite{cb_ono_2008}, a protocol for managing
BitTorrent traffic so that it significantly reduces cross-ISP traffic and in
the same time enhances downloading rates by favouring connections within the
borders of a single autonomous system. Contrary to biased peer selection
proposed in \cite{rpc_bitbiased_2006}, this work can lead to improved
performance with no cooperation between ISPs and their subscribers, no
additional infrastructure and no network topology information. The selection
approach is landmark-based and leverages existing CDN infrastructure for peer
distance estimation. CDNs already use both static (i.e. geographical) and
dynamic (network measurement systems) information for their replica selection,
so the authors claim that peers which exhibit similar redirection behaviour, are
very likely they are close to the replica servers and as a consequence to each
other. The redirection behaviour is modeled in terms of \emph{ratio map} in the
article's parlance, a vector of (replica-server,ratio) tuples, where ratio is
the percentage of times CDN redirects the peer to the specific replica-server.
The bootstrapping phase consists of the peer performing DNS lookup to CDN names
in order to build its redirection information. The interval for polling DNSs
starts from 30 seconds and increases by one minute every time redirection info
to the CDN is found to remain unchanged after a lookup. On the other hand, the
interval is halved whenever is found to have been changed. In order not to avoid
the bootstrapping phase if the ratio map is sufficiently fresh, the protocol
persists the info after the end of a BitTorrent session.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Locality-Awareness in BitTorrent-like P2P Applications}
In \cite{lclx_bitlocal_2009} the authors study and compare three different
approaches in injecting locality awareness in BitTorrent-like applications. The
first acts at a \emph{macroscopic-level} and targets neighbour selection. When a
peer asks the tracker to join, the later sorts the swarm peers according to
their distance to the requesting peer in AS hop count and send it the first
(e.g. 50) peers in the list. The second manipulates the chocking/unchocking
BitTorrent mechanisms at an \emph{intermediate-level}. A peer unchockes its 4
closest in terms of AS hop count interested neighbours. The same applies also
when the peer turns to the seeding state\footnote{Thus this approach favours
least distance, contrary to the original BitTorrent implementation which favours
uploading speed.}. On a \emph{microscopic-level}, the rare-first chunk picking
policy is substituted by the locality-fist policy. The distance value of a piece
is calculated as the mean value of the distances of the peers that posses it.
The paper conducts several measurements on the efficiency of the above scheme to
reach the conclusion that the major design desision for all locality aware P2P
systems is the trade-off of optimising inter-AS traffic, which their solution
achieves, and of the fairness among peers, a section that their locality-based
approach does not do well in comparison to the random approach of the standard
BitTorrent protocol.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf TopBT}
\cite{rtlcgz_topbt_2010} proposed an approach for enhancing proximity awareness
in the BitTorrent protocol without any need of any infrastructure installed. It
suggests that a good peer selection metric should take into account both the
downloading speed and the network topology. Thus, it proposes the use of the
ration of download rate to upload rate, $\frac{d}{u}$, divided by link-level or
AS-level routing hops, $l$ or $a$ respectively, in an attempt to form a
comprehensive metric to select peers with high download rate, low reciprocal
upload demand and low routing hops. The TopBT metric is applied on serveral
parts of the original BitTorrent protocol in order to inject topology awareness
for better peer selection; more specifically on the peer list the tracker
returns when someone first tries to connect to the swarm, on the initial
connection establishment on the connection replacement and of course the
unchoking mechanism.

%A peer that runs the TopBT protocol evaluates its neighbours by periodicaly
%emitting pings or traceroutes in order to uchock those peer that exhibit less
%hops to reach and higher download rates. \emph{Link-hops} are measured by using
%the TTL value that the originator receives as a response from the remote
%peer\footnote{Initial TTL values are known for the different operating systems
%(e.g. 64 for Linux or 128 for Windows NT/2000/XP) so the originator can
%calculate the hops using the value of the TTL on the response message it
%receives.}. Also the peer, when offline, builds a table that maps IPs to
%\emph{AS-hops}, using BGP dumps.

Through experimentation on hundreds of PlanetLab as well as residential hosts,
the authors claim more than 25\% traffic reduction and 15\% faster downloads,
both with a lightweight cost when compared to popular BitTorrent
implementations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Underlying Topology-Aware Peer Selection (UTAPS)}
UTAPS \cite{lcy_utaps_2008} is another peer selection strategy for the
BitTorrent protocol. Its first step is to collect the needed information in
order to construct a picture of the underlying topology and the second is to
make the peer selection, per se, based on this knowledge. For the first part,
the algorithm uses network tomography, a technique which suggests probing only
from a large network's end points in order to infer its internal characteristics
\cite{chny_tomography_2002}. Upon a new peer arrival, the tracker traceroutes it
to get some knowledge (IP address, routers involved, RTT and hops conducted
e.t.c). The more the peers in the swarm the better picture a tracker has for the
underlying infrastructure, which can then return to the newcomming peer in the
form of the bootstrapping peer list. The peers furtherly enhance this picture by
tracerouting the returned peers themselves and reporting back to the tracker.
For the selection part, a family of heuristics are proposed, which in general
target those peers that expose low RTT and are within certain hop-count away of
measured routers. The author's claim that peers running UTAPS instead of random
peer selection can achieve better downloading rates with reduced burden on the
underlying ISP infrastructure though their technique is coarse-grained and their
evaluation configuration was far too small for extracting further safe
conclusions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{ \bf An Effective Network-Aware Peer Selection Algorithm in
BitTorrent}
Qin et al., in their work in \cite{qlzg_biteffpeersel_2009}, propose a
clustering approach that differentiates peers in a swarm into local, intra-ISP
and inter-ISP neighbours. The classification is done through the routers they
use in order to communicate with the tracker. For this, a newly arriving
BitTorrent peer traceroutes the tracker and sends the results in a form of a
vector of triples that contain the IP and the hop counter for the router on the
traceroute path as well as the link latency of the hop that was made in order to
arrive to this router. The tracker uses this information in order classify the
router through a $k$-means classification algorithm, with $k = 3$. For peer
selection, the authors propose a biased approach where the tracker returns $c$
close peers and $d = N - c$ distant peers, where $N$ is the length of the
returned list. In choosing $c$, the tracker employs an iterative search process
first in the peer's local neighbourhood. If not enough peers are available
there, it goes to intra-ISP and subsequently, if needed, to the inter-ISP
cluster. The evaluation of the algorithm is done in a similar artificial,
contained laboratory enviromnent as in \cite{lcy_utaps_2008} and it shows up to
5\% faster downloading times, and up to 22\% reduction of the total cross-ISP
traffic.

%\paragraph*{\textbf{Two-Hop-Away Neighbour Comparison and Selection\\}}
%
%\paragraph{}
%Work in \cite{liu_thancs_2005,liu_thancs_2008} proposes a distributed heuristic
%called \emph{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} that
%\begin{inparaenum}[\itshape i\upshape)]
%  \item is completely distributed and needs no global knowledge,
%  \item presents trivial overhead compared to the query cost savings
%  \item its convergent speed of the algorithm is fast enough (faster than
%minimum spanning tree approaches) so that is effective to dynamic environments,
%and
%  \item does not shrink the search scope.
%\end{inparaenum}
%
%\paragraph{}
%THANCS is considered a \emph{local search method}, in the sense that it targets in finding a locally optimum solution, by exploiting knowledge within a 2-hop radius. The algorithm consists of two main components: \emph{piggybacking neighbour distance on queries} and \emph{neighbour comparison and selection} which are furtherly discussed bellow.
%
%\paragraph{Piggybacking neighbour distance on queries}
%Using network delay as a metric for measuring the distance, each peer probes distances with its immediate naighbours and stores information locally. For this reason a special query message type, \emph{Piggy Message (PM)}, is introduced. It is 6 bytes long and includes two fields: Neighbour IP Address and Neighbour Distance. A peer $P$ constructs a PM for its neighbour $Q$, which contains $Q$'s IP address and $Q$'s distance from $P$. When $P$ receives a query from $Q$, this PM will be piggybacked by the query that goes to all other neighbours of peer $P$. Upon receiving such a query message, each of the other neighbours will detach the PM, record the $PQ$ distance and process the query. This PM will not be further forwarded. In selecting which incoming queries  should piggyback a PM the paper proposes the \emph{pure propability-based (PPB)} and the \emph{new neighbour triggered (NNT)} policies.
%
%\paragraph{Neighbour comparison and selection}
%Figure~\ref{figure:thancs} illustrates this component of the THANCS algorithm. A peer $S$ probes the distance to all known unprobed $N^2(S)$\footnote{$N^2(S)$ denotes the set of peers being two hops away from $S$, while $N(S)$ denotes the set of direct logical neighbours of $S$.}. The distance of $SP$ is known to $S$. Upon receiving a PM from node $P$ with the distance of $PQ$, $S$ follows one of the following:
%\begin{itemize}
%  \item $Q \in \left( N(S) \cap N^2(S) \right)$, i.e. $Q$ is direct neighbour of $S$. In this case $S$ will compare cost of $SQ$, $SP$ and $PQ$. If the most costly connection is one of $SQ$ or $SP$ the corresponding link will be put into a \emph{will-cut list}\footnote{Links are not immediately disconnected when put in the will-cut list. There are useless for forwarding but are kept active, for some time, in order to serve query responses that are traveling to the source peer along the inverse search path.}. If the most costly connection is $PQ$ then $S$ will do nothing, as the fully distributed nature of the algorithm will give a chance to either $P$ or $Q$ to cut the connection.
%  \item $Q \in \left( N^2(S) - N(S) \right)$, i.e. $Q$ is a two-hop-away neighbour of $S$. If $S$ hasn't probed $Q$ before\footnote{A distance cache is maintained and looked up in such a case.}, it probes and stores the result in the \emph{distance cache}. Having the distance of $SQ$, $S$ compares costs of $SQ$, $SP$ and $PQ$. If $SQ$ is the most costly, $S$ will not establish the connection. If $SP$ is the most costly, $S$ will establish connection $SQ$ and put $SP$ in the will-cut list. If $PQ$ is the longest, $S$ will keep  the connection with both $P$ and $Q$, expecting that $P$ or $Q$ will eventually disconnect link $PQ$, later.
%\end{itemize}
%
%\begin{figure}
%\centering
%  \includegraphics[scale=0.4]{img/thancs.jpeg}
%\caption{Probing two-hop-away neighbours}
%\label{figure:thancs}
%\end{figure}
%
%\paragraph{}
%In a static environment THANCS has been proven to be effective; optimizing 45 percent out of the 60 percent of mismatched paths, constructing a nearly optimal overlay. This leads to a 60 percent reduction in traffic cost as well as a 40 percent decrease in query response time. In dynamic environments (Gnutella 0.6/Limewire super-peer-like and Ion flat-like), THANCS saves up to 70 percent of the traffic cost in the super-peer topology and 55 percent for the flat one. Average response time is also decreased by 60 and 45 percent, respectively. Generally, THANCS has similar performance to LTM, without needing synchronization. SBO, incurring half the  overhead of AOTO, reduces the traffic cost the most, while THANCS has lower response time and converges faster than SBO. THANCS is, thus, more suitable for a more dynamic environment. In addition, THANCS is easy to implement and its operation overhead is trivial, compared with the other three approaches. This design, however, has the limitation of not being easily extend to also support non-flooding-based systems.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Peer-exchange Routing Optimization Protocols}
\cite{qiu_prop_2007} introduces two protocols called \emph{Peer-exchange Routing
Optimization Protocols (PROP)} to adjust the neighbourhood graph of an overlay
network in order to reduce its overall link latency. PROP algorithms are based
on the exchange of neighbours among peers, driven by the mutual benefit of the
participants to reduce the network delay among their connections.  In PROP-G
(Generic), a peer can exchange all of its neighbours with another peer, while in
PROP-O (Optimized) two peers exchange a selected subset of their neighbours.
PROP-G is a generic protocol that guarantees the connectivity of the overlay
graph during exchanges, therefore can be applied to both unstructured and
structured overlay networks.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf T2MC}
T2MC \cite{shi_t2mc_2008} uses traceroute logs to detect different autonomous
systems and cluster close by peers, thus reducing redundant traffic that crosses
autonomous system boundaries. T2MC uses a customized k-mean classification
algorithm with $k=2$ to perform the classification, and exploits the stable
structure of the Internet routers to guide clustering. Even though traceroute
provides detailed information about the network structure, its excessive use
creates additional overhead that stresses further the overall network
infrastructure. This forces network administrators to disable its support from
the routers they manage, thus jeopardizing the effectiveness of the T2MC
algorithm.

%\paragraph*{\textbf{T2MC\\}}

%\paragraph{}
%\emph{T2MC}\cite{shi_t2mc_2008} exploits some properties of the Internet
%paradigm and clusters nodes belonging to the same ISP without any centralised
%control or predefined system parameterization. The algorithm considers the
%dynamic nature of peers and exploits the stable nature of routers in order to
%build a topology location relationship among end peers.
%
%\paragraph{}
%Some special routers split the physical network into autonomous system domains.
%Using a Traceroute mechanism T2MC searches for latency leaps among the path to
%a host in order to form ``near'' and ``remote'' router clusters. This is
%achieved by a 2-Means Classification, defined by the following steps:
%\begin{enumerate}
% \item the peer chooses the minimum and maximum Latency results from the
%Traceroute for initializing the cntroinds of two sets ``first'' and ``second''.
% \item The peer calculates, for all hops allong its tracerouted path, the
%absolute distance to the centroids of both sets and assigns the routers to that
%centroid with which it has the smaller absolute distance.
% \item The peer calculates the latncy mean and variance value of two sets
% \item If the variance is larger than a predefined threshold then the algorithm
%takes a loop from step 2 picking the two latency mean values as new centroids
%of sets ``first'' and ``second''.
%\end{enumerate}
%Ultimately peer will end up with two sets having the minimum intra-set
%variance. Finally the peer chooses the router from ``second'' set with the
%minimum hops attribute and sets it as a threshold. The selected router and all
%others whose hops attribute is larger than the threshold are classified as
%``remote'' router cluster. The remaining are classified as ``near''. From the
%``near'' class, the peer chooses the one with maximum hops attribute as its
%edge router, and registers it along with the all the ``near'' cluster into the
%DHT of the p2p overlay. As new peers join the network, those that share the
%same edge router or any of the members of the ``near'' router clustersm ther
%would gather to form a ``close'' peer cluster. As Edge routers can provide more
%valuable information than other members of the ``near'' set, T2MC was designed
%to prioritize interaction of peers and edge gateways.
%
%% TODO: figure t2mc.jpeg (a) stars represent edge routers (b) hop leaps
%
%\paragraph{}
%The use of Traceroute as a tool for implementing the distance measuring
%infrastructure raise concearns about its efficiency and scalability. Being,
%primarilly, a network diagnostic utility, it is concearned too heavy weighted
%and intrucive for use in a larger scale\cite{ratnasamy_binning_2002}.
%Additionally, disabling ICMP is a common administrative policy for edge sites
%to enforce security, while dumping BGP routing
%tables\cite{krishnamurthy_bgpclust_2000} is not directly available to the
%application layer.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Resolving the Topology Mismatch Problem in Unstructured
Peer-to-Peer Networks}
% TODO: to be reviewed

In \cite{hsiao_redblue_2009}, Hsiao et al, claim to construct topology-aware
unstructured overlays that \emph{guarantee} performance qualities in terms of
\begin{inparaenum}[\itshape i\upshape)]
  \item the expected communication latency among any two overlay peers
regardless of the network size, and
  \item the broadcasting scope of each participating peer
\end{inparaenum}
. The algorithm constructs an undirected graph $G = \left( V, E \right)$
comprised by two subgraphs. The first, namely $G^{\left( red \right)} = \left(
V^{\left( red \right)}, E^{\left( red \right)} \right)$ in the paper's context,
includes all vertices of $G$ and ensures the connectivity of the graph by
securing at least one path between any two nodes. In contrast, $G^{\left( blue
\right)} = \left( V^{\left( blue \right)}, E^{\left( blue \right)} \right)$,
contains those vertices of $G$ that have free edges to link to other nodes and
because these are fully utilized, the following also stands $E = E^{\left( red
\right)} \cup E^{\left( blue \right)}$. A joining peer $u$, partitions its
neighbours into two subsets, the $B_u^{\left( red \right)}$ and $B_u^{\left(
blue \right)}$. In order to populate the $B_u^{\left( red \right)}$ subset, peer
$u$ samples peers uniformly and at random. Then, each of these selected peers
discovers a routing path starting from itself towards the node with the smallest
(or the largest) ID in the system.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Distributed Domain Name Order}
\emph{Distributed Domain Name Order (DDNO)} \cite{z-yk_ddno_2005}
uses the domain names to detect topologically close nodes based on the
assumption that nodes within the same domain are topologically close to each
other. Half of the possible connections of a node is used to connect to these
local peers, and the other half is used to randomly connect to the peers
anywhere on the network. \textit{DDNO} is a heuristic approach to solve the
topology mismatch problem, with local connections to improve the efficiency and
reduce the delay on the network.  The random connections on the other hand help
ensure the connectivity in the network and avoid partitioning. \textit{DDNO}
uses \textit{Split-Hash} and \textit{dnMatch} algorithms to detect locality by
using domain names.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Landmark Based Proximity}\label{sec:landmark}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Detecting the proximity of a node using landmarks, also known as
\textit{Landmark Clustering}, is based on the view that nodes with similar
distances to a set of predefined well-known landmark nodes are pretty likely
being close to each other. But this approach has its weaknesses itself, like the
fact that is a rather coarse grained approximation, therefore not particularly
well suited for detecting the correct positions of nodes within close distance
to each other.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Landmark Binning}\label{sec:landmark_binning}
\begin{figure*}
\centering
  \includegraphics[scale=0.4]{img/algorithms/landmark_binning}
\caption{Example 2D coordinate overlay and a sample routing path from node D to
(x,y).}
\label{fig:landmark_binning}
\end{figure*}
Landmark Binning \cite{ratnasamy_binning_2002} is an approach to partition close
by nodes into bins based on their distance to well known anchor nodes across the
Internet, as shown in Figure \ref{fig:landmark_binning}. In order to detect
locality, nodes use network latency (i.e. RTT) as a measurement technique. The
network latency, even though not always accurate, is selected in this work due
to its non-intrusive, transparent and easy to apply behaviour. In order for the
binning to work, few well known anchor servers with known physical locations on
the Internet. Authors estimate that a number of 12 such servers can prove
sufficient for this. The nodes measure their distance to all these servers and
record the latency values for each. Later, the tuple of these latency values
represents the bin of each node. Nodes with similar latency values are
considered close to each other, or at least within the same autonomous system.
The operation of the method is independent of the model incorporated by the
overlay network and thus it can be applied with no significant changes to both
structured and unstructured P2P systems. \cite{ratnasamy_binning_2002} has the
detailed description of such algorithm for both architectures. Landmark binning
is also considered as a good candidate for content distribution networks in
particular. The major disadvantage of landmark binning is to install and
maintain landmark servers on different autonomous systems, all over the world.
A typical P2P network usually has a couple of million nodes connected at any
given time, thus rendering landmark servers an important parameter when
designing for large scales.  The authors claim that latency estimation does not
drain much from the network resources but in case scalability problems arise,
they claim the solution is replacing single landmark servers with clusters of
servers within the same physical area. However, this approach does not reduce
the possibility of excessive network traffic flow through these landmark
servers. One other possible problem is the incorrect binning caused by the use
of inaccurate methods for delay measurement, since tracking down, for example,
network latencies is far from proven to be safe for location estimation.

% TODO: MAYBE THANCS deserves some more attention here

\paragraph*{\bf mOverlay}
mOverlay \cite{zhang_moverlay_2004} tries to addresses the scalability
issues that might be imposed to networks when using static landmark servers by
introducing the dynamic landmarks. The authors introduce the notion of a group
which is a set of  peer considered close to each other with respect to any
position in the underlying network, proximity being some  user-defined cost
metric (i.e. round-trip-time). mOverlay, tries to recreate small-world-like
properties to the overlay network by creating a a two-level hierarchical
structure, where on the top level we have connections between groups while on
the bottom we have connections between peers inside groups. Finding the correct
closest group is the most important part  of the overlay construction. Nodes are
grouped based on their  distance to the groups already in the network, rendering
the  later the (dynamic) landmarks for the process. The authors formally prove
that any new node can reach its group by exchanging at most $O(logN)$ messages
within the network. Finally mOverlay also considers a second important function
in its protocol which is the constant maintenance of the overlay.

\renewcommand\arraystretch{1.4}% (MyValue=1.0 is for standard)

%\begin{figure}[h!]
\hspace{-3ex}
\begin{center}
\footnotesize
%\begin{tabular}{
%\begin{landscape}
\begin{longtable}{
|>{\columncolor[gray]{.7}}m{0.2\columnwidth}
|>{\columncolor[gray]{.9}}m{0.4\columnwidth}
|>{\columncolor[gray]{.8}}m{0.2\columnwidth}
|>{\columncolor[gray]{.9}}m{0.2\columnwidth}
%|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
%|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
%|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
%|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
|}
\caption{Decentralized Unstructured Algorithms} \label{fig:unstruct_compare_table} \\
\hline
\rowcolor[gray]{.5}
\textbf{Algorithm} &  \textbf{Overlay structure} & \textbf{Base protocol} &
 \textbf{Scalability}\\
\hline
\endfirsthead
\multicolumn{4}{c}%
{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
\hline
\rowcolor[gray]{.5}
\textbf{Algorithm} &  \textbf{Overlay structure} & \textbf{Base protocol} &
 \textbf{Scalability}\\
\hline
\endhead
\hline \multicolumn{4}{r}{\textit{Continued on next page}} \\
\endfoot
\hline
\endlastfoot
\textbf{Narada} & \textbf{Overlay optimization
based}. Creates a mesh (richer connected graph) and builds minimum spanning
trees on this mesh & & Small and sparse groups \\

\hline
\textbf{Gia} & \textbf{Broadcast based} Replaces
Gnutella flooding with random walk, and introduces KaZaA style supernodes. Uses
dynamic topology adaptation protocol &
 Gnutella &  Better than Gnutella  \\

\hline
\textbf{Adaptive Overlay Topology Optimization} & \textbf{Overlay optimization
based}. Creates overlay multicast tree with Selective Flooding protocol&
Gnutella &  Better than Gnutella \\

\hline
\textbf{Location-aware Topology Matching} &
\textbf{Overlay Optimization Based}. Uses \textit{TTL2-detector flooding}, \textit{low productive
connection cutting}, and \textit{source peer probing}. & Gnutella &  Better than Gnutella \\

\hline
\textbf{Replication Strategies in Unstructured P2P Networks} &
\textbf{Cache Based}. Uses uniform, proportional and square root allocation
strategies to replicate data. & Gnutella &  Better than Gnutella \\

\hline
\textbf{Tracing a large-scale Peer to Peer System: an hour in the life of Gnutella.} &
\textbf{Cache Based}. Proposes a caching algorithm based on the traces of the Gnutella traffic & Gnutella & Better than Gnutella \\

\hline
\textbf{Improving search in P2P networks} &
\textbf{Broadcast Based}. Uses \textit{iterative deepening}, \textit{directed
BFS}, and \textit{local indices} to improve efficiency. & Gnutella &  Better than Gnutella \\

\hline
\textbf{Distributed Cycle Minimization Protocol} &
\textbf{Broadcast based} Uses a decentralized cycle elimination protocol  &  &  \\

\hline
\textbf{Scalable Bipartite Overlay} &
\textbf{Overlay optimization based} Uses bipartite partition graph and builds
local minimum spanning trees  & Gnutella  & Better than Gnutella \\

\hline
\textbf{Adaptive Connection Establishment} &
\textbf{Overlay optimization based} Forms Neighbour Cost Tables, builds local
minimum spanning trees and perform local optimizations & Adaptive Overlay
Topology Optimization (AOTO), Gnutella & Better than Gnutella \\

\hline
\textbf{Hops Adaptive Neighbour Discovery} &  & &  \\

\hline
\textbf{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} &
\textbf{Overlay optimization based} Uses piggybacking to discover neighbour
distances and selects neighbours  & Gnutella  & \\

\hline
\textbf{mOverlay} &\textbf{Landmark based proximity} Uses dynamic landmarks to find node locality
& & Due to dynamic landmarks and grouping, more scalable than tree-based or mesh-based protocols \\

\hline
\textbf{Distributed Domain Name Order (DDNO)} &
\textbf{Overlay optimization based} Connects half of the nodes connections to
the nodes in the same domain and the other half to random nodes, therefore
supports locality and topological connection  &  & Yes, by using super
peers \\

\hline
\textbf{Peer-exchange Routing Optimization Protocols} & \textbf{Overlay optimization based} Optimizes overlay by the exchange of
neighbors among peers  & Can work with both decentralized structured and
unstructured architecture & Yes \\

\hline
\textbf{MAY OMIT- T2MC} &
\textbf{Overlay optimization based} Uses traceroute results for clustering the
nodes  & & \\

\hline
\textbf{Unnamed-unstructured} &
\textbf{Overlay optimization based} Minimizes the communication delay and
maximizes the broadcasting range & & Better than THANCS and mOverlay \\

\hline
\textbf{Landmark Binning} & \textbf{Landmark based proximity} Uses network latency to partition
nodes into bins & Can work with both decentralized structured and unstructured architecture & \\

\hline
%\end{tabular}
\end{longtable}
%\end{landscape}
\end{center}
\vspace{-2.5ex}
\vspace{-2.5ex}
%\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Discussion on the Algrithms for Unstructured Architectures}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this section we presented the state of the art for the algorithms applied to
unstructured P2P systems. We categorised it based on their philosophy for
alleviating the topology mismatch problem into four major categories namely
\begin{inparaenum}[\itshape i\upshape)]
  \item broadcast-based,
  \item cache-based,
  \item overlay optimisation-based and
  \item landmark-based approaches
\end{inparaenum}
. Here we present a overall discussion on the above categories trying to point
out the advantages, disadvantages and novelties introduced by each and by all of
them. A quick overview can be seen in Table \ref{fig:unstruct_compare_table}.

Broadcast-based approaches in general propose intelligent neighbour selection in
order to replace the inefficient blind flooding first used in the Gnutella
protocol. Instead of forwarding the queries to all their outgoing links they
dictate the intelligent selection of neighbours with either higher probability
to answer a query than others or neighbours  that expose specific features such
as high bandwidth, high capacity, or low latency. This, in practice reduces the
aggregate resource usage of the P2P network and improves the overall performance
of the system. On the other hand, the algorithms do not offer a real solution to
the topology mismatch problem. Thus, none can guarantee that overlay and
underlying topologies are aligned with each other let alone quantify the
mismatch and try to alleviate. Broadcast-based methods can be used in
conjunction with other approaches to improve the quality of the P2P systems.

The success of caching, as a well studied method in client-server Internet
applications, also made it a good candidate for the P2P networks. Even
though caching improves the performance, and reduces the overall resource usage
of P2P systems, the design of caches in this context is non-trivial compared to
the web-based caching. Due to the unique, inherent characteristic of all P2P
systems that, each node can act as both a server and a client, two important
issues have to be considered when designing caches. First, the lifetime of a
query is short, as the nodes join and leave frequently. Second, the result of a
single query string is not always the same, as it is dependent on the source of
the query, the TTL value set for the messages, the current interconnection of
peers and the high volatility of the environment. So, in order to develop a
successful caching system for a P2P architecture, these parameters also have to
be considered. Even though the state of the art P2P algorithms using caching
methods reduce the resource usage of the network, they also fail to consistently
address the probelem of the topology mismatch.

Overlay-optimisation-based  can further be devided into
minimum-spanning-tree-based and clustering-based. Researchers initially started
considering alternatives to IP multicast for the application layer, which led
Narada, that simulates the same one-to-many communication approach requiring IP
multicast services. Application layer multicast, however, is not as efficient as
the one enabled by IP and still struggles to ease the difficulties imposed by
the topology mismatch problem. Narada, and its followers (AOTO, LTM, SBO) try to
solve the topology mismatch problem by building a richer connected graph and
forming minimum spanning trees over this graph that can efficiently route
messages among peers. Application level multicast protocols, pioneered by
Narada, are generic protocols that can be applied to both P2P file sharing
systems, as well as content distribution networks. The original Narada algorithm
is designed for small groups therefore it is not scalable. Its followers try to
solve the scalability problem by introducing various methods including forming
minimum spanning trees for the two-hop neighbours ($N^2$) of each node,
partitioning the graph into two random groups where each group is responsible of
different tasks, or performing local optimizations dynamically on the overlay
graph. The advantage of building minimum spanning trees is that they maintain
the connectivity on the network in an efficient manner, while still preserving
the overall search scope. However, building and maintaining a minimum spanning
tree creates additional overhead for the network. The alternative method used in
overlay optimization is the cluster based approache. According to that, nodes
use various methods to detect physically close nodes to form clusters. T2MC uses
traceroute logs and DDNO uses domain names to cluster close by nodes. However,
the accuracy of the methods used to form clusters directly affect the success
rate of the algorithm. Traceroute for example is a heavy weight utility to be
frequently used on the network, let alone the fact that many network equipment
vendors do not allow traceroute calls. The major problem with cluster based
approaches is that the limited connectivity within the local domains shrink the
search scope dramatically, which negatively affects the search performance of
the P2P system. DDNO addresses this limitation by allowing half of each nodes
connections to be to random nodes over the network, which balances the
efficiency of the clustering approach with improved connectivity.

In landmark-based algorithms, nodes use network delay (e.g. RTT) as a
distance measurement method to position themselves with respect to ``a priori''
known servers (the landmark servers) on the Internet. Landmark servers are used
by nodes to estimate their positions based on the intutive assumption that nodes
with similar distances to a set of landmarks, are physically close to each
other, as well, over the network. The landmark based protocol has two important
drawbacks. Firstly the network delay is not a reliable distance estimation
method. For example, based on the load on the network the delay to certain nodes
or networks can change from time to time, which will eventually affect the
distance measurements and wrong measurements will lead to wrong estimated
positions for the nodes, or incorrect and non optimal clusterings of the nodes.
The second drawback of using landmark servers is the cost of installing and
maintaining this kind of infrastructure across the whole Internet and for all
the different autonomous system domains. As popular P2P file sharing
applications usually have millions of peers connected at any time, it is not
false to assume that the network costs of maintaining these landmark servers
will be quite high. A possible solution to the scalability problem of the static
landmark servers is to use ordinary nodes as dynamic landmarks once they
estimate their own positions. Even though this approach scales much better than
static landmark servers, still the measurement accuracy problem affects the
overall performance of the system.

### 2014-01-21 mard-survey.tex
2012.06.11 (Vassilis) Content revisited. Duplicate text has been removed.

@@
 \author{
-VASSILIS S. MOUSTAKAS\\\emph{Department of Informatics and Telecommunications}\\\emph{National and Kapodistrian University of Athens}\\
-H\"USEYIN AKCAN\\\emph{Department of Software Engineering}\\\emph{Izmir University of Economics}\\
-MEMA ROUSSOPOULOS\\\emph{Department of Informatics and Telecommunications}\\\emph{National and Kapodistrian University of Athens}\\
+VASSILIS S. MOUSTAKAS\\
+\texttt{b.moystakas@di.uoa.gr}\\
+\emph{Department of Informatics and Telecommunications}\\
+\emph{National and Kapodistrian University of Athens}\\
+H\"USEYIN AKCAN\\
+\texttt{huseyin.akcan@ieu.edu.tr}\\
+\emph{Department of Software Engineering}\\
+\emph{Izmir University of Economics}\\
+MEMA ROUSSOPOULOS\\
+\texttt{mema@di.uoa.gr}\\
+\emph{Department of Informatics and Telecommunications}\\
+\emph{National and Kapodistrian University of Athens}\\
 AND\\
-ALEX DELIS\\\emph{Department of Informatics and Telecommunications}\\\emph{National and Kapodistrian University of Athens}\\
-%  Vassilis Moustakas$^1$, H\"useyin Akcan$^2$, Mema Roussopoulou$^1$ and Alex
-%Delis$^1$\\
-%  $^1$Department of Informatics and Telecommunications,\\
-%  National and Kapodistrian University of Athens\\
-%  \mbox{\texttt{\{b.moystakas, mema, ad\}@di.uoa.gr}}\\
-%  $^2$Department of Software Engineering,\\
-%  Izmir University of Economics, Izmir, Turkey \\
-%  \mbox{\texttt{huseyin.akcan@ieu.edu.tr}}
+ALEX DELIS\\
+\texttt{ad@di.uoa.gr}\\
+\emph{Department of Informatics and Telecommunications}\\
+\emph{National and Kapodistrian University of Athens}
 }
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% start
-%TODO: Maybe this section should be reviewed or deleted or go to another section
-%Studying the behaviour of the various peer-to-peer schemes
-%\cite{matei_mapgnutella_2002, lv_randomwalks_2002, merugu_str2unstr_2003}
-%showed that the ad-hoc network topology of unstructured overlay networks that
-%preserve \emph{Power Law} and \emph{Small World} characteristics\footnote{Power
-%Law describes the node degree while the Small World describes characteristics
-%of path length and clustering coefficient. The clustering coefficient for a
-%node $\upsilon$ in a graph $G = \left( V, E \right)$ is defined as the ratio of
-%the existing connections between $\upsilon$'s neighbouring nodes to $\gamma
-%\times \left( \gamma - 1 \right)$, where $\gamma$ is the number of neighbouring
-%nodes of $\upsilon$. High cluster coefficient means that neighbouring nodes of
-%any node $\upsilon$ likely connect one another.} \cite{faloutsos_powerlaw_1999,
-%saroiu_measurefileshare_2002} offer a more promising approach. Particularly:
-%\begin{itemize}
-%  \item Peer-to-peer clients are extremely \emph{transient}. Unstructured
-%systems can have high maintenance traffic in delivering messages, updating the
-%mapping, discovering failures and replicating lost data or pointers, making
-%them insufficient on highly volatile networks.
-%  \item \emph{Keyword searches} versus \emph{exact-match queries}. In DHTs
-%there is a tight control between the data placement and the topology of the
-%network. For this reason it is hard to efficiently support partially matched
-%queries while Gnutella and other similar systems effortlessly support keyword
-%searches and other complex queries since the mechanism is realized locally, on
-%a node-by-node basis.
-%  \item Popular content is located at multiple peers and thus it is more likely
-%for a flooding-based search to return results. DHTs, on the other hand, fit
-%better in the systems which require ability to reliably locate content, even in
-%the extreme case that only a single-copy exists in the network.
-%\end{itemize}
-% end
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% TODO: how 
-%For this reasons, efforts have been placed for optimizing the efficiency of
-%decentralized unstructured peer-to-peer networks. Research mainly focuses on
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item reducing unnecessary, redundant communication traffic, and
-%  \item exploiting physical locality to reduce communication response.
-%\end{inparaenum}
-%The goal can be achieved at, both, the application-level network as well as the
-%underlying physical one. In the first case by refining the message relay
-%techniques, while in the second one, by adaptively reconstructing the
-%application network to map as well as possible to the the physical network.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{The Topology Mismatch Problem}
-
-%%%%%%%
-%%%%%%% TODO: NEEDS REWORDING
-%%%%%%% Additional Material
-%%%%%%%
-%%%%%%%
-%%%%%%%
-% P4P [16],instead, aimed to provide a general framework for P2P
-%applications by developing ISP-application interfaces so that
-%application software can use the network layer information,
-%such as congestion status and network topology, for better
-%performance and resource utilization.
-
-%P4P [37] project attempts to address
-%the problem through custom trackers, both for ISPs and P2P
-%systems, using an interface based on a primal-dual decomposition
-%of an optimization problem. This interface design simplifies
-%the realization of traffic-engineering objectives from each parties’
-%perspective and ensures the extensibility of the approach. Through
-%simulation-based studies and limited experimental deployments,
-%these collaborative approaches have been shown to effectively
-%reduce network costs while minimally impacting application performance.
-%A clear advantage of these proposals is that they
-%allow ISPs to incorporate aggregated traffic policies in their tracker
-%recommendations (e.g., a particular traffic balance ratio between
-%peering providers). However, all of them require deployments
-%of oracles for each participating ISP and their effectiveness is
-%ultimately predicated on their adoption by P2P applications and a
-%trust relationship between P2P users and their ISPs.
-
-% TODO: Taxomize the approaches that have been proposed in the literature
-% for unstructured and structured overlay networks
-% UNDER CONSTRUCTION
-% VERY IMPORTANT SECTION
-%\section{Taxonomy}
-%
-%% TODO: Unstructured
-%% UNDER CONSTRUCTION
-%Several overlay protocols for dealing with the problems derived from the
-%topology mismatch problem have been proposed in the literature. Each
-%implementation has different goals but all share some common ground on the
-%aspect of the network functionality they focus on, in order to achieve them. In
-%this section, the general taxonomy used throughout this paper to categorize the
-%algorithms based on their functionality is presented.  The brief description of
-%each category and sub-category is stated below.
-%
-%{\sethlcolor{yellow}\hl{CITE this paper {\cite{liu_ltm_2004}}}}
-%
-%\subsection{Forwarding based}
-%
-%Forwarding based methodology is generally adapted by distributed unstructured
-%overlay systems. In this approach, a peer selects only a subset of its
-%neighbours to re-broadcast query messages. The selection is made using one or a
-%combination of various statistical metrics. Examples of such metrics are the
-%number of responses received by a neighbour or the connection latency of the
-%link between the nodes, etc. The forwarding based approach enhances search
-%efficiency but has also several drawbacks. First, the search scope is reduced
-%drastically.  Expanding the search scope, on the other hand, is no easy task
-%because the overhead of forming multicast trees is proportional to the multicast
-%group size.  Second, forwarding based schemes do not consider dynamic joining
-%and leaving of peers so they do not scale well on dynamic environments.
-%
-%\subsection{cache based} 
-%
-%Caching based protocols are effectively used to reduce traffic costs and
-%response times. The caching policy varies depending on the way protocol handles
-%the index and the content. Centralized P2P systems use central index servers,
-% while local caching systems, such as KazaA, use super peers
-%to cache indices in a distributed way. Content caching is also possible in P2P
-%systems, where nodes cache the forwarded content for further retrievals.
-%Although caching has the above mentioned advantages,  duplication
-%of messages still exist, which limits the scalability of these approaches.
-%Therefore, cache based approaches are analyzed in the following categories:
-%  \begin{itemize}
-%    \item \emph{data index caching},
-%    \item \emph{content index caching},
-%    \item \emph{centralized}, and
-%    \item \emph{local}.
-%  \end{itemize}
-%
-%\subsection{overlay optimization based} 
-%
-%The overlay optimization based protocols modify the topology of the P2P network
-%using various techniques. These approaches include creating spanning trees using 
-%connection graphs, creating cluster of physically close nodes, or using latency
-%information to detect proximity. The brief description of each category is
-%presented below:
-%
-%  \begin{itemize}
-%    \item \emph{Spanning tree based}. These approaches construct of a rich graph
-%    based on the network connections and build minimum spanning tree on the
-%    graphs, causing large traffic overhead to the system
-%    \cite{chu_esm_2000,chu_esm_2002}.
-%
-%    \item \emph{Cluster based}. These approaches select to link physically
-%    closer nodes with each other, therefore shrink the search scope
-%    significantly while mapping accuracy is not always guaranteed.
-%
-%    \item \emph{Minimum latency first}. Use of latency as a metric to calculate
-%    distance among peers. They require global latency information
-%    ``landmarks''\footnote{Measuring latency between peers and stable Internet
-%    servers.}.
-%
-%  \end{itemize}
-%
-%
-%% TODO: Structured
-%% UNDER CONSTRUCTION
-%\subsection{proximity based}
-%Xu \textit{et al.} \cite{xu_globstate_2003} state that there are three ways generating proximity information
-%\begin{itemize}
-% \item Expanding Ring Search\\
-%Expanding ring search can be of two forms. First it can utilize the multicast
-%infrastructure in the underlying network in order to emit its messages.
-%Unfortunately such infrastructures are not widely deployed thus the
-%implementations of this way of generating proximity information is limited to
-%blindly flooding the neighbourhood to obtain reasonable results.
-%
-% \item Heuristics\\
-%Heuristics are used in order to reduce the blindness of the expanding ring
-%search and make realizations more efficient and effective. Unfortunately a
-%common problem of all heuristic approaches is the local minimum pitfall in which
-%the search might be caught into.
-%
-% \item Landmark Binning\\
-%Landmark clustering is based on the view that nodes with similar distances to a
-%set of predefined well-known landmark nodes are pretty likely of being close to
-%each other. But this approach has its weaknesses as well, such as the fact that
-%is a rather coarse grained approximation, therefore not particularly well suited
-%for differentiating nodes within close distance to each other.
-%
-%\end{itemize}
-%
-%Castro \textit{et al.} \cite{castro_proximitydht_2002,castro_topawareroute_2002} on
-%the other hand, comment on the three basic approaches in exploiting proximity
-%in DHT protocols suggested by Ratnasamy \textit{et al.} \cite{ratnasamy_openq_2002}.
-%These are:
-%
-%\begin{itemize}
-% \item Geographic Layout\\
-%The node IDs are assigned in such a way that nodes close by in the physical
-%network topology, be close in the node ID space as well. Implementations that
-%work relatively well with this approach have been incorporated into CAN. Nodes
-%measure the RTT between themselves and a set of landmarks in order to match the
-%CAN space as much as possible to the physical one. Unfortunately, the approach
-%requires well known landmark servers and for that matter is  not fully
-%self-organizing which can further lead to imbalanced node distribution. On
-%other DHTs, such as Chord or Pastry, another problem emerges.  To gain fault
-%tolerant properties, these protocols, replicate key-value pairs on neighbouring
-%(in the ID space) nodes. When a proximity-based node ID assignment has been
-%used, the needed failure resilience is undermined by the fact that close by
-%nodes are more likely to suffer collective failures.
-%
-% \item Proximity Routing\\
-%Proximity routing does not require routing tables be built using any knowledge
-%about network proximity. On the other hand it exploits such knowledge in order
-%to choose the best next hop during routing a message. This approach 
-%balances between choosing the node that will further progress the routing
-%towards the destination and choosing the closest entry in the routing table, in
-%terms of network proximity. Thus, it is relatively less effective than
-%geographical layout when applied to CAN(-like) implementations. Moreover, the
-%technique has been incorporated into a version of Chord (TODO: PChord |
-%additionally check whether this is true) causing an increase on the overhead of
-%node joins and the size as well as maintenance cost of finger tables.
-%\cite{dabek_cfs_2001} proposes a server selection scheme for the Chord DHT, on
-%the domain of proximity routing selection. In \emph{CFS}, each node predicts
-%the entire lookup latency as a function of the total number of nodes and the
-%average overlay next routing peer. The problem is that it is very difficult to
-%have a clear picture on the total number of nodes and the average hop latency
-%from the local. This leads to rough estimations that consequently decreases
-%overall performance.
-%
 % \item Proximity Neighbour Selection\\
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{The Algorithms}
-%Having understood the nature of the problem that peer-to-peer architecture face, this section starts the discussion of the academic work that has been conducted in the field the last few years.
-What follows this introduction, does not claim to be a thorough citation of all known protocols that are available out there, but instead, a carefull selection of those that left a distinctive fingerprint contribution in the efforts of the research community to alleviate the topology mismatch problem.
 %
+% TODO: SOME INFORMATION ON ISP-ASSISTED TOPOLOGY AWARENESS
+%       ATTENTION: NEEDS REWORDING
+%
+% P4P [16],instead, aimed to provide a general framework for P2P
+%applications by developing ISP-application interfaces so that
+%application software can use the network layer information,
+%such as congestion status and network topology, for better
+%performance and resource utilization.
+
+%P4P [37] project attempts to address
+%the problem through custom trackers, both for ISPs and P2P
+%systems, using an interface based on a primal-dual decomposition
+%of an optimization problem. This interface design simplifies
+%the realization of traffic-engineering objectives from each parties’
+%perspective and ensures the extensibility of the approach. Through
+%simulation-based studies and limited experimental deployments,
+%these collaborative approaches have been shown to effectively
+%reduce network costs while minimally impacting application performance.
+%A clear advantage of these proposals is that they
+%allow ISPs to incorporate aggregated traffic policies in their tracker
+%recommendations (e.g., a particular traffic balance ratio between
+%peering providers). However, all of them require deployments
+%of oracles for each participating ISP and their effectiveness is
+%ultimately predicated on their adoption by P2P applications and a
+%trust relationship between P2P users and their ISPs.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% UNSTRUCTURED
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%\subsection{Narada}
-%
-%\paragraph*{}
-%\emph{End System Multicast} \cite{chu_esm_2000,chu_esm_2001,chu_esm_2002} as well as other  proposes an architecture, where end-systems implement all multicast related functionality (including membership management and packet replication).
-In \emph{Narada}, the protocol developed in the paper's context, end-systems
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item self-organize, in a fully distributed manner,
-%  \item into an efficient overlay structure, both from the network and the application perspective,
-%  \item with self improving properties,
-%  \item adapting to network dynamics\footnote{Long-term variations in Internet path characteristics (i.e. bandwidth, latency e.t.c.)}.
-%\end{inparaenum}
-%
-%\paragraph*{}
-%What Narada really does is to construct overlay spanning trees in a two-step process:
-%\begin{enumerate}
-%  \item It builds a richer connected graph \footnote{ENarada \cite{li_enarada_2008} used Gossip protocol for the construction}, termed a \emph{mesh}, while trying to ensure a couple of desired performance properties:
-%    \begin{itemize}
-%      \item path quality between any pair of members should be comparable to the quality of the unicast path between them, and
-%      \item each member should have a limited number of neighbours on the mesh.
-%    \end{itemize}
-%
-%  \item It computes minimum spanning trees out of the mesh, each rooted at the corresponding source member.
-%\end{enumerate}
-%
-%\paragraph*{}
-%Narada incrementally improves the quality of the mesh by adding and dropping overlay links.
-Members probe each other at random and new links are added depending on what is the gain in doing so.
-On the other hand, participants, continuously  monitor the utility of existing links, and drop them if they are found to be not usefull.
-%
-%\paragraph*{}
-%Narada protocol achieves relatively good performance for small and medium sized groups involving tens to hundreds of members.
-This is not, by no means, what a contemporary peer-to-peer system needs to be characterized as scalable, but this work is considered pioneering\footnote{Another similar approach is \emph{Scattercast}\cite{chawathe_scattercast_2000}}, in that it was the first to have conducted a detailed Internet evaluation to analyse and demonstrate the feasibility of an overlay-based, application layer, multicast architecture that can dynamically adapt to bandwidth and latency properties of the physical underlying infrastructure.
-%
-%\subsection{Gia}
-%
-%\paragraph*{}
-%\emph{Gia} \cite{chawathe_gia_2003} is a decentralized unstructured peer-to-peer file sharing system with the following characteristics:
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item replaces Gnutella's flooding with random walks \cite{lv_randomwalks_2002},
-%  \item recognizes the importance of overlay network's topology, while using the random walks, and exploits it by incorporating a topology adaptation algorithm,
-%  \item introduces a token-based flow control mechanism, and
-%  \item detects the significant heterogeneity in peer bandwidth, processing power, disk speed e.t.c. and takes advantage of it.
-%\end{inparaenum}
-%
-%\paragraph*{}
-%More specifically, there are four key components in the design of \emph{Gia} which are summarized bellow:
-%\begin{enumerate}
-%  \item A \emph{dynamic topology adaptation} protocol that puts participating nodes within short reach of high capacity nodes so that these \emph{high-degree} nodes, which due to their high connectivity receive most of the queries, actually have the capacity to handle them.
-%  \item An \emph{active flow control} scheme is used to avoid overloaded hot-spots. Heterogeneity is detected and flow control tokens are given to nodes based on the available capacity.
-%  \item Every node maintains pointers to to the content that is offered by their immediate neighbours, creating a \emph{one-hop replication} of pointers, scheme.
-%  \item A \emph{search protocol} based on random walks, that is biased towards directing queries to high-capacity nodes that are typically best able to answer these queries.
-%\end{enumerate}
-%
-%\subsection{Adaptive Overlay Topology Optimization}
-%
-%\paragraph*{}
-%\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{liu_auto_2003} is an algorithm for building an overlay multicast tree among each source node and its direct logical neighbours in order to alleviate the mismatching problem while providing a larger query coverage range.
-It includes the steps of
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item \emph{Selective Flooding (SF)}, and
-%  \item \emph{Active Topology (AT)}
-%\end{inparaenum}
-%which are furtherlly discussed in the following paragraphs.
-%
-%\paragraph*{Selective Flooding (SF)}
-%Instead of flooding to all neighbours, SF builds an overlay multicast tree among each peer and its immediate logical neighbours.
-This reduces overall message burden to the network by making some neighbours non-flooding.
-The tree is formed using a minimum spanning tree algorithm.
-For this reason a peer needs to know the costs to all its logical neighbours (e.g.
-the network delay) as well as between any pair of neighbours.
-Additional, cost probing, messages need to be added to the original messages defined by the Gnutella protocol.
-%
-%LTM's SF effectiveness has be proven to be detached from the different physical or overlay topologies.
-On the other hand, SF is more effective with large number of logical neighbours.
-It can reach an average optimization rate of 87.4 percent on a logical topology with an average of 30 logical neighbours.
-%
-%\paragraph*{Active Topology (AT)}
-%At this step the overlay topology is reorganized.
-Each peer, independently, makes optimizations in the overlay network to alleviate topology mismatching by replacing non-flooding neighbours, with closer nodes.
-The approach incorporated by the paper is called \emph{Randomized AT} algorithm which picks up a candidate peer at random among the non-flooding neighbour's neighbours.
-Whenever a new neighbour cost table is received or there is a change of neighbours, the source peer has to re-calculate the multicast tree and apply the randomized AT algorithm.
-%
-%Different numbers of average logical neighbours has little to do with the effectiveness of AT.
-If the source has $n$ non-flooding peers, there are $n$ potential neighbour replacements.
-The overhead to exhaust all $n$ possible replacements can be high, so in practice, after each replacement the source peer can decide whether it needs to find another candidate peer.
-This is done by computing the cost improvement ratio greater than some predefined termination threshold.
-The larger the threshold, the slower, in the number of optimization steps, the reduction of the normalized average distance.
-As a whole the average response time is significantly reduced when more optimization steps taken.
-%
-%\subsection{Location-aware Topology Matching}
-%
-%\paragraph*{}
-%\emph{Location-aware topology matching} or \emph{LTM} \cite{liu_ltm_2004}, is an algorithm of building an efficient overlay by disconnecting low productive connections and choosing physically closer nodes as logical neighbours while still retaining the search scope and reducing response time of queries.
-This is done by issuing detector messages by peers, to a small region so that receivers of the detector can record relative delay information and cut inefficient or redundant logical links as well as add closer nodes as direct neighbours.
-Three operations are defined in LTMband are furtherly discussed in the following paragraphs.
-%
-%\paragraph*{TTL2-detector flooding}
-%\emph{TTL2-detector} messages are designed based on the specification of the Gnutella protocol.
-Suppose that $d(i, S, v)$ denotes the TTL2-detector which has the message ID of $i$ with TTL value of $v$ and is initiated by $S$.
-In addition to Gnutella's unified 23-byte header, such a message has a body of two types:
-%\begin{itemize}
-%  \item short format $d(i, S, 1)$ if $S$ is a direct neighbour of the receiver
-%  \item long format $d(i, S, 0)$ if $S$ is within a two-hop distance from the receiver
-%\end{itemize}
-%This way a peer can calculate link costs from within a one- or two-hop area around it and thus reach peers in $N(S)$ and $N^2(S)$ sets, respectively, in the paper's context.
-Each peer floods a TTL-detector periodically while it calculates link cost to another peer by receiving other peer's detector messages and checking:
-%\begin{itemize}
-%  \item the Source Timestamp field and its own clock upon time of receipt for one-hop messages
-%  \item the Source Timestamp, TTL1 Timestamp and its own clock upon receipt for two-hop messages
-%\end{itemize}
-%Of course peer clocks must be synchronized in order to do such checking.
-%
-%\paragraph*{Low productive connection cutting}
-%There are three cases for any peer $P$ who receives $d(i, S, v)$ multiple times:
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item $P$ receives both $d(i, S, 1)$ and $d(i, S, 0)$
-%  \item $P$ receives multiple $d(i, S, 0)$s from different paths, and P randomly chooses to process one
-%  \item $P$ receives one $d(i, S, 1)$ and multiple $d(i, S, 0)$s, and $P$ processes $d(i, S, 1)$ and one randomly selected $d(i, S, 0)$
-%\end{inparaenum}
-%If the link with the largest cost is found and is a direct neighbour then the connection is put in a will-cut list and stays there for a certain period of time.
-If it is not, then it is handled by other peers.
-After that period, connections are cut and recorded to $P$'s cut-list.
-%
-%\paragraph*{Source probing}
-%For a peer $P\in(N^2(S) - N(S))$ who receives only one $d(i, S, 0)$, the cost of $PS$ is obtained (with list look-up or probing).
-Then $P$ compares it with the cost from each hop and if $PS$ has the largest cost, $P$ will not keep this connection, while otherwise the connection will be created.
-%
-%\paragraph{}
-%Supposing $n$ is the number of peers, $c_n$ is the average number of neighbours and $c_e$ is the average cost of logical links, then in the flooding-based search the traffic incurred by one query from an arbitrary peer in a peer-to-peer network is $O(n)$.
-As observed in the Gnutella network \cite{sripanidkulchai_gnutella_2001}, each peer issues $0.3$ queries per minute in average, thus the per minute traffic incurred by the network with $n$ peers is $O(n^2)$.
-Because each $d(i, S, v)$ has a TTL of $2$ in each source peer, the traffic for one time LTM optimization in all peers is at most $2nc_n^2c_e$.
-If each peer uses LTM $k$ times per minute, the total traffic incurred is $2knc_n^2c_e$.
-Simulation shows the best value for $k$ is $2$ or $3$.
-So, the traffic overhead caused by LTM to the network is $O(n)$.
-%
-%TTLj-detectors, with $j > 2$, would detect and break cycles with more than 4 links.
-LTM though, does not use such detectors because detector-flood traffic would increase significantly, and cut links between two end-peers, could cause queries initiated by them to traverse a path much more expensive than the cost on the the cut link.
-%
-%\paragraph{}
-%LTM disadvantages are
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item disagreement of measured delay due to unsynchronized clocks causes problems when deciding the cut positions, which can influence the network connectivity, and
-%  \item the network delay metric mainly focuses on disabling the connections between peers physically far away without considering the shortcuts created by powerful peers.
-%\end{inparaenum}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/ltm_multid.jpeg}
-%\caption{Peer $P$ receives $d(i, S, v)$ multiple times}
-%\label{figure:ltm_multid}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/ltm_example.jpeg}
-%\caption{A full example of LTM}
-%\label{figure:ltm_example}
-%\end{figure}
-%
-%\subsection{Scalable Bipartite Overlay}
-%\emph{Scalable Bipartite Overlay (SBO)} \cite{liu_bipartite_2007} employs an efficient strategy to select query forwarding paths and logical neighbours.
-The topology construction and optimization of SBO consist of four phases:
-%\begin{enumerate}[\itshape i\upshape)]
-%  \item When a new peer is joining the network, it will randomly take an initial colour; say red or white.
-This colour is not to be changed for the time the peer stays connected.
-If it leaves and then rejoins it will, again, have to pick a colour, randomly.
-Thus all peers are separated into two groups, red or white.
-Then the bootstrap host will provide, the joining peer, with a list of active peers, including their colour information, in order for the later to establish connections to different colour peers.
-This way, all peers form a bipartite overlay.
-%  \item White peers probe distances with their immediate (red) neighbours, form a cost table and send this table to their corresponding red neighbours.
-%  \item Each red peer builds a minimum spanning tree using the obtained neighbour cost tables.
-These are two-hop diameter trees so white peers do not need to build one.
-The links that are part of a minimum spanning tree are called \emph{forwarding connections (FC)} while the rest \emph{non-forwarding (NFC)}.
-%  \item Having a minimum spanning tree with two hops a red peer is able to send its queries within that range.
-Some white peers, though, have become non-forwarding neighbours.
-In this phase, such a (white) neighbour will try to find another red peer being two hops away from its current red neighbour to replace the later as its new neighbour.
-%\end{enumerate}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/sbo_efficient_forward.jpeg}
-%\caption{A red peer $P$ has a small overlay topology of $N(P)$ and $N^2(P)$ and computes the efficient forwarding paths}
-%\label{figure:sbo_efficient_forward}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/sbo_neighbour_replace.jpeg}
-%\caption{Neighbor replacement}
-%\label{figure:sbo_neighbour_replace}
-%\end{figure}
-%
-%In a static enviromnent LTM may reduce traffic cost by around 80 to 85 percent while SBO reduces traffic cost between 85 and 90 percent.
-However, LTM  is proved to converge in around 2-3 steps while SBO needs 4-5 steps.
-Moreover LTM reduces response time by more than 60 percent in 3 steps while SBO needs 8.
-In a dynamic environment (10 minute average peer lifetime, 0.3 queries/sec by each peer) SBO and LTM reduce the average traffic cost per query (including the overhead due to the optimization steps) by 85 and 80 percent, respectively.
-Moreover LTM reduces the response time per query to 30 percent while SBO to 35 percent.
-%
-%\subsection{Adaptive Connection Establishment}
-%
-%\paragraph{}
-%\emph{Adaptive Connection Establishment (ACE)} \cite{liu_ace_2004} builds an overlay multicast tree among each source node and the peers within a certain diameter from the source peer and optimizes the neighbour connections that are not in that tree.
-%
-%\paragraph{}
-%ACE indicates three phases for its algorithm:
-%\begin{enumerate}[\itshape i\upshape)]
-%  \item Calculate cost between nodes using network delay as a metric.
-Each peer probes the costs with its immediate logical neighbours and forms a \emph{neighbour cost table (NCT)} using a special routing message type.
-Two neighbouring peers exchange their NCTs in order for every peer to obtain the cost between any pair of its local neighbours forming a small overlay topology.
-%  \item Based on obtained NCTs a minimum spanning tree among each peer and its immediate neighbours is built (Figure~\ref{figure:ace_phase2}).
-%  \item Physically far away neighbours are replaced by physically close neighbours.
-In ACE a peer, say $S$, probes the distance between one of its non-flooding neighbour's neighbour, say $G$ and $H$ respectively.
-If the link to neighbour's neighbour is smaller than that to the neighbour, the later is cut.
-If this is not the case but $S$ finds that cost of $GH$ is even larger than that of the $SH$, $S$ will keep $H$ as a new neighbour.
-Obviously, if $SH$ is larger than $SG$ and $GH$, the connection will not be established and $S$ will continue probing another neighbour's neighbour.
-The algorithm described is conducted within $1$-neighbour closure (among its source peer and all its direct neighbours) but the optimization scope can be enlarged.
-The larger the scope, the better topology matching improvement but also the greater the computational overhead (Figure~\ref{figure:ace_phase3}).
-%\end{enumerate}
-%
-%\paragraph{}
-%Simulations in \cite{liu_acesims_2004} show that the average scope of each query to cover the same scope of nodes is reduced by about 65 percent without losing any autonomy feature, while the average response time can be reduced by 35 percent.
-Larger diameter topologies lead to better topology optimization rate but also to higher communication and computation overhead.
-It was also found that it is more effective in higher connectivity dense topologies.
-Compared to LTM, it comes short of convergence speed.
-In \cite{ni_mismatch_2004} shows reduction of both total traffic (90 percent) and response time (80 percent) to message queries without shrinking the search scope.
-SBO, on the other hand, achieves approximately 85 percent reduction on traffic cost and about 60 percent reduction on query response time.
-Last but not least, it is concluded that work must be done on incorporating a more sophisticated selection policy for candidate non-flooding peers.
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/ace_phase2.jpeg}
-%\caption{Second phase in ACE}
-%\label{figure:ace_phase2}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/ace_phase3.jpeg}
-%\caption{Third phase in ACE}
-%\label{figure:ace_phase3}
-%\end{figure}
-%
-%\subsection{Hops Adaptive Neighbour Discovery}
-%
-%\paragraph{}
-%Paper \cite{chen_hand_2006} proposes a new algorithm named \emph{Hops Adaptive Neighbour Discovery (HAND)} which uses a fully distributed triple hop adjustment strategy to address the topology mismatch problem.
-The advantages of the algorithm compared to other approaches are that
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item it does not need any clock synchronization,
-%  \item it is a fully distributed algorithm making it robust and reliable in decentralized systems,
-%  \item the traffic overhead of the triple hop adjustment is very low,
-%  \item it is applicable to dynamic peer-to-peer environments, and
-%  \item maintains lower query response time.
-%\end{inparaenum}
-%
-%\paragraph{}
-%The algorithm's ultimate goal is the optimal overlay, the \emph{Logical Communication Network (LCN)} in the paper's context.
-The key concept of the algorithm is that a graph $G^{*}$ that describes an LCN and a graph $G$ that describes the current overlay are matched only if all peer hop sequences $(v_1, v_2, \ldots, v_k)$ in $G$ exist in $G^{*}$ and in the same order.
-In practice triple sequences $(v_1, v_2, v_3)$ are used.
-%
-%\paragraph{}
-%The mismatching detection is done in the following way.
-Suppose we want to verify peer sequence $v_2-v_1-v_3$ (see Figure~\ref{figure:hand_com_overlay}).
-A pair of probing messages are sent from $v_1$ to $v_2$ and $v_3$.
-Suppose delays of $(v_1,v_2)$ and $(v_1,v_3)$ are are denoted as $x$ and $z$, respectively.
-When the probing message arrives to $v_2$ it forwards it directly to $v_3$.
-Similarly, when the probing message arrives to $v_3$, it forwards it directly to $v_2$.
-These last steps are performed in order to obtain delays of $(v_2,v_3)$ and $(v_3,v_2)$ physical paths, respectively, denoted by $y$.
-%\begin{itemize}
-%  \item If $y=z-x\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched and should be adjusted to $v_1-v_2-v_3$ by deleting edge $(v_1,v_3)$ and adding a new $(v_2,v_3)$ (see Figure~\ref{figure:hand_matchtriple_lemma3}).
-%  \item if $y=x-z\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched and should be adjusted to $v_1-v_3-v_2$ by deleting edge $(v_1,v_2)$ and adding a new $(v_3,v_2)$ (see Figure~\ref{figure:hand_matchtriple_lemma4}).
-%\end{itemize}
-%$\varepsilon$ is a small positive real number denoting additional delays caused by possible forwarding and jitter delays.
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/hand_com_overlay.jpeg}
-%\caption{The communication in overlay}
-%\label{figure:hand_com_overlay}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/hand_matchtriple_lemma3.jpeg}
-%\caption{The matching triple in paper's Lemma 3}
-%\label{figure:hand_matchtriple_lemma3}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/hand_matchtriple_lemma4.jpeg}
-%\caption{The matching triple in paper's Lemma 4}
-%\label{figure:hand_matchtriple_lemma4}
-%\end{figure}
-%
-%\paragraph{}
-%Measurements conducted for evaluation perposes showed that in a static environment the algorithm can effectively decrease traffic cost by about 77 percent and shorten the query response time by about 49 percent in less than two minutes.
-In a dynamic environment it shows similar behaviour and with the size of the overlay network having little impact on the effectiveness of the algorithm.
-Compared to LTM both algorithms have almost the same traffic reduction rate, however on the response time reduction rate HAND has a higher one by about 4 percent.
-The traffic overhead of HAND is much less than that of LTM by an average of 55 percent.
-%
-%\subsection{Distributed Cycle Minimization Protocol}
-%
-%\paragraph{}
-%\emph{Distributed Cycle Minimization Protocol (DCMP)} which is introduced in \cite{zhu_dcmp_2008} is a dynamic, fully decentralized protocol that promises significant reduction of duplicate messages.
-To achieve that, it uses the slightly different approach of eliminating unnecessary cycles, while retaining the connectivity of the network and preserving fault resilience and load balancing properties of unstructured peer-to-peer schemas by avoiding the creation of a hierarchical organization.
-%
-%DCMP aims at cutting the cycle paths at strategic locations. Any peer that detects a duplicate message can initiate the cutting process which consists of the following steps:
-%\begin{enumerate}
-%  \item Peers in the cycle elect a leader called \emph{GatePeer}\footnote{GatePeers are important for maintaining the connectivity and optimal structure of the network while peers join and leave randomly.}.
-%  \item The cycle is cut at a well-defind point with respect to the GatePeer
-%\end{enumerate}
-%
-%\paragraph{}
-%The first step after detecting a duplicate message by some peer is to gather information from all peers in the cycle using a new type of control message called \emph{Information Collecting Message} or \emph{ICM}.
-ICM contains:
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item a \emph{GUID}\footnote{Globally Unique IDentifier assigned to every query message generated by any node.} field same as the one of the duplicate message,
-%  \item \emph{DetectionID} field which represents the direction of the connection where the duplicate was identified\footnote{This ensures the uniqueness of the ICM messages because as it travels through many cyclic paths, multiple peers will detect the duplicates and initiate an ICM message.}, and
-%  \item \emph{Node Information Vector (NIV)} which contains information (bandwidth, CPU power, etc) about peers that propagated the ICM.
-%\end{inparaenum}
-%
-%\paragraph{}
-%Suppose $A$ detected the duplicate as depicted in Figure~\ref{figure:dcmp}.
-It then emits an ICM to $B$ and $F$, that initially contains information only about itself.
-Each peer that receives the ICM, appends its information and propagates it along the reverse path of the original message.
-Since two copies of ICM are sent, at some point, a peer, say peer $D$, will receive a duplicate ICM.
-Using the information in the NIVs of the ICMs, $D$, decides to cut (for example) the EF connection.
-To inform the other peers about its decision, it emits a \emph{Cut Message (CM)} which contains the GUID and DetectionID of the corresponding ICM and an additional field that identifies the connection to be cut.
-$D$, then, forwards the CM in the reverse directions from where the ICM arrived.
-Similarly CMs received by any peer are propagated toward the reverse path of the corresponding ICM.
-Eventually, either peer $E$ or peer $F$ will receive the CM and cut the connection, thus eliminating the cycle.
-%
-%Receiving a duplicate ICM denotes the existence of a cycle.
-The opposite is not true though.
-For example, if the cycle contains $2 \times TTL$ edges, it will not be detected, because the ICM messages will be discarded before they locate it.
-There is a tradeoff between preserving the connectivity of the network and minimizing the duplicates that makes such a possibility to be safely ignored.
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/dcmp.jpeg}
-%\caption{Cycle elimination methods in DCMP}
-%\label{figure:dcmp}
-%\end{figure}
-%
-%\paragraph{}
-%Experiments in \cite{zhu_dcmp_2008} showed that DCMP incurs a lower delay, returns more results and decreases the number of duplicate messages by 22\%, compared to LTM.
-Additionally DCMP has one to two orders of magnitude less overhead, because it adopts a more ``LAZY'' approach than broadcasting control messages periodically like LTM does.
-%
-%\subsection{Two-Hop-Away Neighbour Comparison and Selection}
-%
-%\paragraph{}
-%Work in \cite{liu_thancs_2005,liu_thancs_2008} proposes a distributed heuristic called \emph{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} that
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item is completely distributed and needs no global knowledge,
-%  \item presents trivial overhead compared to the query cost savings
-%  \item its convergent speed of the algorithm is fast enough (faster than minimum spanning tree approaches) so that is effective to dynamic environments, and
-%  \item does not shrink the search scope.
-%\end{inparaenum}
-%
-%\paragraph{}
-%THANCS is considered a \emph{local search method}, in the sense that it targets in finding a locally optimum solution, by exploiting knowledge within a 2-hop radius.
-The algorithm consists of two main components: \emph{piggybacking neighbour distance on queries} and \emph{neighbour comparison and selection} which are furtherly discussed bellow.
-%
-%\paragraph{Piggybacking neighbour distance on queries}
-%Using network delay as a metric for measuring the distance, each peer probes distances with its immediate naighbours and stores information locally.
-For this reason a special query message type, \emph{Piggy Message (PM)}, is introduced.
-It is 6 bytes long and includes two fields: Neighbour IP Address and Neighbour Distance.
-A peer $P$ constructs a PM for its neighbour $Q$, which contains $Q$'s IP address and $Q$'s distance from $P$.
-When $P$ receives a query from $Q$, this PM will be piggybacked by the query that goes to all other neighbours of peer $P$.
-Upon receiving such a query message, each of the other neighbours will detach the PM, record the $PQ$ distance and process the query.
-This PM will not be further forwarded.
-In selecting which incoming queries  should piggyback a PM the paper proposes the \emph{pure propability-based (PPB)} and the \emph{new neighbour triggered (NNT)} policies.
-%
-%\paragraph{Neighbour comparison and selection}
-%Figure~\ref{figure:thancs} illustrates this component of the THANCS algorithm.
-A peer $S$ probes the distance to all known unprobed $N^2(S)$\footnote{$N^2(S)$ denotes the set of peers being two hops away from $S$, while $N(S)$ denotes the set of direct logical neighbours of $S$.}.
-The distance of $SP$ is known to $S$.
-Upon receiving a PM from node $P$ with the distance of $PQ$, $S$ follows one of the following:
-%\begin{itemize}
-%  \item $Q \in \left( N(S) \cap N^2(S) \right)$, i.e.
-$Q$ is direct neighbour of $S$.
-In this case $S$ will compare cost of $SQ$, $SP$ and $PQ$.
-If the most costly connection is one of $SQ$ or $SP$ the corresponding link will be put into a \emph{will-cut list}\footnote{Links are not immediately disconnected when put in the will-cut list.
-There are useless for forwarding but are kept active, for some time, in order to serve query responses that are traveling to the source peer along the inverse search path.}.
-If the most costly connection is $PQ$ then $S$ will do nothing, as the fully distributed nature of the algorithm will give a chance to either $P$ or $Q$ to cut the connection.
-%  \item $Q \in \left( N^2(S) - N(S) \right)$, i.e.
-$Q$ is a two-hop-away neighbour of $S$.
-If $S$ hasn't probed $Q$ before\footnote{A distance cache is maintained and looked up in such a case.}, it probes and stores the result in the \emph{distance cache}.
-Having the distance of $SQ$, $S$ compares costs of $SQ$, $SP$ and $PQ$.
-If $SQ$ is the most costly, $S$ will not establish the connection.
-If $SP$ is the most costly, $S$ will establish connection $SQ$ and put $SP$ in the will-cut list.
-If $PQ$ is the longest, $S$ will keep  the connection with both $P$ and $Q$, expecting that $P$ or $Q$ will eventually disconnect link $PQ$, later.
-%\end{itemize}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/thancs.jpeg}
-%\caption{Probing two-hop-away neighbours}
-%\label{figure:thancs}
-%\end{figure}
-%
-%\paragraph{}
-%In a static environment THANCS has been proven to be effective; optimizing 45 percent out of the 60 percent of mismatched paths, constructing a nearly optimal overlay.
-This leads to a 60 percent reduction in traffic cost as well as a 40 percent decrease in query response time.
-In dynamic environments (Gnutella 0.6/Limewire super-peer-like and Ion flat-like), THANCS saves up to 70 percent of the traffic cost in the super-peer topology and 55 percent for the flat one.
-Average response time is also decreased by 60 and 45 percent, respectively.
-Generally, THANCS has similar performance to LTM, without needing synchronization.
-SBO, incurring half the  overhead of AOTO, reduces the traffic cost the most, while THANCS has lower response time and converges faster than SBO.
-THANCS is, thus, more suitable for a more dynamic environment.
-In addition, THANCS is easy to implement and its operation overhead is trivial, compared with the other three approaches.
-This design, however, has the limitation of not being easily extend to also support non-flooding-based systems.
-%
-%\subsection{mOverlay}
-%Zhang et al.
-in \cite{zhang_moverlay_2004} focus on two aspects of the overlay construction.
-First, to achieve \emph{efficiency} of communication, a protocol should minimize unecessary long-distance hops and redundant traffic.
-Second, to be \emph{scalable} the overlay should be constructed in a distributed fashion and maintenance cost (data and locality management) should be minimized.
-Both of the above observations, are positively affected by the exploitation of node proximity at the underlying network.
-To accomplish it, the paper introduces the \emph{group} concept, according to which non-static (dynamic) landmarks are used to compute proximity, resulting in a highly robust and scalable overlay construction with reduced maintenance cost as well.
-The dynamic nature of landmarks help towards load-balance, meaning that hot-spots are avoided.
-%
-%\paragraph{}
-%The proposed architecture, called \emph{mOverlay}, is a \emph{clustering approach} which creates a two-level hierarchical network, where on the top level we have connections between groups while on the bottom we have connections between hosts inside groups, thus attempting to recreate \emph{Small-World}-like properties for the constructed overlay network.
-The notion of a \emph{group}, in the paper's context, is a set of hosts that are close to each other with respect to any position $P$ in the underlying network.
-The distance between hosts can be
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item network latency,
-%  \item round-trip time,
-%  \item minimum bandwidth on the links along the path connecting the two nodes, or
-%  \item some other user-defined cost metric between the two nodes.
-%\end{inparaenum}
-%These nodes are refered to as neighbouring. Similarly groups can exchange messages with their neighbouring ones\footnote{Groups nearby, in the underlying physical network.}.
-%
-%According to the \emph{grouping criterion} set by the authors, a new host $Q$ belongs to some group $A$ if the distance between $Q$ and group $A$'s neighboring groups is the same as the distance between group $A$ and it's neighbouring groups.
-The neighbour groups thus, play the role of the dynamic landmarks and for accuracy and performance reasons the nearest of them are chosen to play that role.
-%
-%\paragraph{Locating process} A new coming host, $Q$, first connects to a globally known host cache called the \emph{rendezvous point (RP)} in order to retrieve the starting point in the overlay, say $A$ in group $1$.
-Host $Q$ then, measures its distance to host $A$.
-At the same time, the later, sends information about the neighbour groups of group $1$ back to host $Q$.
-This list is called \emph{candidate group list}, and the newcoming host sequnentially measures its distance to each of them in seek for the closest one.
-If the \emph{grouping criterion} is met, host $Q$ belongs to group $1$.
-If not, a boot host from the closest group is found and the algorithm is re-run until the criterion is met or after a predefined number of repetitions.
-In the later case, $Q$ creates a new group comprising itself only.
-The above protocol does not favour hotspots as it spreads the probability of visiting a group across the whole overlay and limits the overhead in the level of $O \left ( log N \right )$.
-%
-%\paragraph{General overlay operations} A set of additional protocols, are also introduced, similar to those found in traditional unstructred networks, but modified focusing on scalability and robustness.
-For example a protocol for \emph{group formation} is introduced that exploits the inherent characteristic of proximity, in the overlay, in order to efficiently detect the neighbouring groups of a newly formated group from the set of adjasent groups of its closest neighbour.
-Additionaly, during \emph{group joining} the coresponding protocol denotes the exchange of important information for group maintenance.
-This can be furtherly improved by \emph{information sharing} between nodes of the same group, functionality handled by a dedicated flood-like protocol\footnote{Since nodes that belong to the same group are physically close this can be achieved at a minimum price.}.
-Moreover, another set of distributed protocols handle the \emph{information update}.
-The information that needs update, in the proposed architecture, is
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item the host cache, when a new node joins, and
-%  \item the neighbours of groups, when a close-by group is generated.
-%\end{inparaenum}
-%Finally, in case of \emph{host failure} or \emph{host departure} the system is able to maintain its stability since there are defined operations for periodical host cache update and group leader selection if one leaves or dies.
-%
-%\subsection{Distributed Domain Name Order}
-%{\sethlcolor{yellow}\hl{HA:  In paper/introduction nice advantages of unstructured networks:
-%
-%Unstructured P2P networks o?er a number of
-%important advantages: (i) An unstructured network
-%imposes very small demands on individual nodes,
-%and more speci?cally it allows nodes to join or leave
-%the network without signi?cantly a?ecting the sys-
-%tem performance. (ii) Unstructured networks are
-%appropriate for content-based retrieval (e.g., key-
-%word searches) as opposed to object identi?er loca-
-%tion of structured overlays. (iii) Finally unstructured
-%networks can easily accommodate nodes of vary-
-%ing power. Consequently, they scale to very large
-%sizes and they o?er more robust performance in
-%the presence of node failures and connection
-%unreliability.
-%}}
-%
-%\paragraph{}
-%\cite{z-yk_ddno_2005} proposes the \emph{Distributed Domain Name Order (DDNO)} technique which makes unstructured overlay networks, topologically aware.
-It results in a flat overlay topology but with some changes it can be utilized in \emph{superpeer} environments.
-In DDNO, a node of degree $d$, tries to connect to $\frac{d}{2}$ nodes that belong to the same domain (\emph{sibling} connections) and another $\frac{d}{2}$ of random nodes (\emph{random} connections).
-Connecting to \emph{sibling} nodes ensures the reduction of long distance message travelling and thus resulting to better performance, while \emph{random} nodes keep the structure connected.
-%
-%\paragraph{Discovering \emph{random} neighbours}
-%Initially, when a new coming node wants to join a network $N$, uses a \emph{hostcache} mechanism to provide it with a list of nodes with which to establish its $\frac{d}{2}$ \emph{random} connections.
-%
-%\paragraph{Discovering \emph{sibling} neighbours}
-%In order to discover the rest $\frac{d}{2}$, \emph{sibling}, nodes, the \emph{lookupDN} procedure is initiated.
-In this phase, a special $l$ walker message is multicasted, by the newcomer, in search of a node that knows\footnote{A special structure is used called ZoneCache that contains information about which nodes are reachable in an $r$-hop radius.} how to guide $n$ to a \emph{sibling} node.
-%
-%\paragraph{}
-%Figure~\ref{figure:ddno_lookupdn} shows an $l$ message emitted by node $n$ which travels along the path [$a$, $b$, $c$, $e$, $b$, $d$].
-At node $d$, $l$ finds information to make a decision on which neighbour to follow next\footnote{Contrary to the random choices, made at the previous hops.}.
-Ultimately $l$ arrives at $m$, which is a \emph{sibling} node to $n$.
-$m$ then issues a broadcast message to all its own \emph{siblings}.
-Each of the receiving nodes, $m$ included, will respond to $n$ with a designated $l^{'}$ message if it is willing to accept new connections and out of these answers, node $n$, will attempt to establish its, remaining, $\frac{d}{2}$ \emph{sibling} connections.
-%
-%\paragraph{}
-%Unfortunately, use of DDNO results in networks which have a uniform distribution of node degrees, losing the valuable properties of Power Law and Small World networks, as mentioned previously in this survey.
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.375]{img/ddno_lookupdn.jpeg}
-%\caption{Domain name look-up in a \emph{DDNO} topology}
-%\label{figure:ddno_lookupdn}
-%\end{figure}
-%
-%%\subsection{Critical Topology-Aware Grouping}
-%%
-%%\paragraph{}
-%%\emph{Critical Topology-Aware Grouping (CTAG)} presented in \cite{zhao_ctag_2006}, is a grouping algorithm that tries to exploit low-cost and low-delay communication of physically close-by peers.
-The grouping strategy is based on the \emph{IANA}\footnote{Internet Assigned Numbers Authority} and the respective \emph{Regionanal Internet Registry (RIR)}'s IP assignment strategies, according to which nodes within the same organization are always addressed from the same block.
-The paper proposes the \emph{Adjacency Measurement (AM)} technique which uses the longest matching IP segment criterion to calculate node proximity.
-%%
-%%Observations made in various studies (e.g.
-\cite{matei_mapgnutella_2002}) concerning the distribution of nodes among \emph{Internet Service Providers (ISPs)} and \emph{Autonomous Systems (ASs)} have shown that only $2$ to $5$ percent of Gnutella connections link peers within a single AS while more than $40$ percent of all Gnutella peers are located within the top 10 ASs.
-Similarly, measurements in \cite{zeinalipour-yazti_gnudc_2002} used a $244,000$ IPs test-bed and results have shown that $45$ percent of the nodes belonged to only $10$ large ISPs and $58$ percent belong to only $20$.
-Such results mean that most overlay generated traffic crosses AS borders increasing topology mismatch cost.
-%%
-%%\emph{CTAG} focuses on both the construction of the overlay as well as dynamically revising it during node interaction, phases called \emph{bootstrapping grouping} and \emph{dynamic revision}, respectively in the paper's context.
-%%
-%%\paragraph{Bootstrapping grouping}
-%%The \emph{Gnutella Web Caching} mechanism has been modified in order for a new coming node to choose the closest \emph{GWC} in order to retrieve the node list for bootstrapping.
-%%
-%%\paragraph{Dynamic revision}
-%%Similarly to the bootstrapping phase, \emph{AM} metric is used to store hosts' addresses.
-read from \emph{X-Try} headers during handshake or from \emph{QueryHit} messages.
-Additionally, when a node reaches the max neighbour connections, node disconnects the neighbours with the lowest \emph{AM}.
-%
-%\subsection{Peer-exchange Routing Optimization Protocols}
-%
-%\paragraph{}
-%Work presented in \cite{qiu_prop_2007} introduces \emph{peer-exchange}\footnote{A series of exchanges of neighbours between two peers.
-One exchange can be viewed as a pair of cut-add operations.} as a basic operation to adaptively adjust the connections of the overlay network, and efficiently reduce the average logical link latency of the whole system.
-There are two points that differentiate this scheme from others:
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item it utilizes the collaboration of two peers to optimize their neighbourhood environment, than simply letting each node to ``selfishly'' choose its own strategy, and
-%  \item can be deployed effortlessly on both unstructured and structured peer-to-peer systems.
-%\end{inparaenum}
-%
-%\emph{PROP-G}\footnote{\emph{G} stands for \emph{generic}.} is the simple and direct variant of \emph{peer-exchange} in which two neighbours swap all their neighbours.
-As depicted in figure~\ref{figure:prop-g} the mechanism can be viewed as, nodes, exchanging overlay ``positions''.
-Thus, the overall overlay topology is not affected by the this operation.
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/prop-g.jpeg}
-%\caption{PROP-G in which a neighbours are exchanged.}
-%\label{figure:prop-g}
-%\end{figure}
-%
-%The alternative protocol proposed, called \emph{PROP-O}\footnote{\emph{O} stands for optimized}, ensures that the degree of each participating node remains the same by selectively choose the same number of neighbours for the exchange.
-With this mechanism, 2 nodes exchange the same number of neighbours in order not to change the degree
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/prop-o.jpeg}
-%\caption{PROP-O exchanging \emph{m} neighbours.}
-%\label{figure:prop-o}
-%\end{figure}
-%
-%\subsection{T2MC}
-%
-%\paragraph{}
-%\emph{T2MC}\cite{shi_t2mc_2008} exploits some properties of the Internet paradigm and clusters nodes belonging to the same ISP without any centralised control or predefined system parameterization.
-The algorithm considers the dynamic nature of peers and exploits the stable nature of routers in order to build a topology location relationship among end peers.
-%
-%\paragraph{}
-%Some special routers split the physical network into autonomous system domains.
-Using a Traceroute mechanism T2MC searches for latency leaps among the path to a host in order to form ``near'' and ``remote'' router clusters.
-This is achieved by a 2-Means Classification, defined by the following steps:
-%\begin{enumerate}
-% \item the peer chooses the minimum and maximum Latency results from the Traceroute for initializing the cntroinds of two sets ``first'' and ``second''.
-% \item The peer calculates, for all hops allong its tracerouted path, the absolute distance to the centroids of both sets and assigns the routers to that centroid with which it has the smaller absolute distance.
-% \item The peer calculates the latncy mean and variance value of two sets
-% \item If the variance is larger than a predefined threshold then the algorithm takes a loop from step 2 picking the two latency mean values as new centroids of sets ``first'' and ``second''.
-%\end{enumerate}
-%Ultimately peer will end up with two sets having the minimum intra-set variance.
-Finally the peer chooses the router from ``second'' set with the minimum hops attribute and sets it as a threshold.
-The selected router and all others whose hops attribute is larger than the threshold are classified as ``remote'' router cluster.
-The remaining are classified as ``near''.
-From the ``near'' class, the peer chooses the one with maximum hops attribute as its edge router, and registers it along with the all the ``near'' cluster into the DHT of the p2p overlay.
-As new peers join the network, those that share the same edge router or any of the members of the ``near'' router clustersm ther would gather to form a ``close'' peer cluster.
-As Edge routers can provide more valuable information than other members of the ``near'' set, T2MC was designed to prioritize interaction of peers and edge gateways.
-%
-%% TODO: figure t2mc.jpeg (a) stars represent edge routers (b) hop leaps
-%
-%\paragraph{}
-%The use of Traceroute as a tool for implementing the distance measuring infrastructure raise concearns about its efficiency and scalability.
-Being, primarilly, a network diagnostic utility, it is concearned too heavy weighted and intrucive for use in a larger scale\cite{ratnasamy_binning_2002}.
-Additionally, disabling ICMP is a common administrative policy for edge sites to enforce security, while dumping BGP routing tables\cite{krishnamurthy_bgpclust_2000} is not directly available to the application layer.
-%
-%\subsection{Unnamed Unstructured!!}
-%% TODO: to be reviewed
-%
-%\paragraph{}
-%In \cite{hsiao_redblue_2009}, Hsiao et al, claim to construct topology-aware unstructured overlays that \emph{guarantee} performance qualities in terms of
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item the expected communication latency among any two overlay peers regardless of the network size, and
-%  \item the broadcasting scope of each participating peer.
-%\end{inparaenum}
-%
-%The algorithm constructs an undirected graph $G = \left( V, E \right)$ comprised by two subgraphs.
-The first, namely $G^{\left( red \right)} = \left( V^{\left( red \right)}, E^{\left( red \right)} \right)$ in the paper's context, includes all vertices of $G$ and ensures the connectivity of the graph by securing at least one path between any two nodes.
-In contrast, $G^{\left( blue \right)} = \left( V^{\left( blue \right)}, E^{\left( blue \right)} \right)$, contains those vertices of $G$ that have free edges to link to other nodes and because these are fully utilized, the following also stands $E = E^{\left( red \right)} \cup E^{\left( blue \right)}$.
-%
-%A joining peer $u$, partitions its neighbours into two subsets, the $B_u^{\left( red \right)}$ and $B_u^{\left( blue \right)}$.
-In order to populate the $B_u^{\left( red \right)}$ subset, peer $u$ samples peers uniformly and at random.
-Then, each of these selected peers discovers a routing path starting from itself towards the node with the smallest (or the largest) ID in the system.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% STRUCTURED
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%\subsection{Landmark Binning}
-%The scheme proposed in \cite{ratnasamy_binning_2002} is based on the process of \emph{binning} close by nodes (in terms of network latency) into the same cluster.
-The authors set the following objectives for designing their proposed algorithm:
-%\begin{enumerate}
-% \item simplicity (minimal support from any measuring infrastructure),
-% \item scalablility (no global knowledge of the network),
-% \item complete distribution (nodes need no communication-cooperation).
-%\end{enumerate}
-%
-%\paragraph{Binning scheme}
-%The implementation requires a set of well-known \emph{landmark} machines spread accross the Internet.
-Every newly arriving node measures its distance from these landmarks and unilaterally decides to join a specific bin based on these results.
-In more detail the node measures its round-trip time to each of the landmarks and orders these measurements in a decreasing order.
-The ordering represents a ``bin'', in the sence of close-by nodes having the same landmark ordering and hance belong to the same ``bin''.
-This means that a landmark system consisting of $m$ such nodes results in $m!$ potentional different bins.
-%
-%The algorithm can be considered scalable, as nodes need only compute distances to a small number of predefined nodes and thus without exchanging any information.
-A potential bottleneck could be the extra load that this ``ping''-like scheme imposes to the landmarks, especially when we need instant reaction from our topology when dealing with the dynamic nature of the p2p networks.
-%
-%To answer the question as to whether the algorithm actually contributes possitivelly to the construction of an enhaned overlay, the paper defines the \emph{gain ratio} as the factor by which the latency reduces when someone communicates with a random node from the same bin than with one not in the bin.
-This is implemented with an inter-bin to an intra-bin latency ratio.
-%
-%\paragraph{Structured and Unstructured Binning}
-%The \emph{binning} scheme can be incorporated within either a structured or an unstructured overlay construction algorithm. The authors provide an example for both these cases.
-%
-%\paragraph{}
-%Assuming a structured approach based on CAN\cite{ratnasamy_can_2001} and $m$ landmark nodes.
-The coordinate space is then partitioned into $m!$ equally sized portions, each corresonding to a single ordering of the landmarks.
-To do this, the first dimension is divided into $m$ areas each of which is furtherly divided (second dimension) into $m - 1$ sections and so on.
-Having set this $m$ dimensional space, at joining time, a node measures the delay to the set of landmarks in order to determine its associated bin and thus position itself in that portion of the coordinate space associated with its landmark ordering.
-Even though this scheme can
-%
-%For unstructured overlays the paper assumes \emph{a set of $n$ nodes where each node picks any $k$ neighbour nodes so that the average routing latency on the resultant overlay is low (assuming shortest path routing)}.
-According to the proposed heuristic algorithm called \emph{BinShort-Long}, a node picks its neighbours by choosing its $\frac{k}{2}$ closest\footnote{If the node's bin is not large enough for it to pick these $\frac{k}{2}$ neighbours, it picks the required nodes from the bin that matches the most in terms of landmark ordering.} ones (named \emph{short links}), using the \emph{binning} scheme and the rest $\frac{k}{2}$ randomly (\emph{long links}).
-The former set produces well-connected \emph{pockets} of nearby nodes while the later preserves the connectivity of the graph, both yielding a proximity factor of $\alpha = 0.5$ in an attempt to preserve the benefitial properties of unstructured topologies\cite{merugu_str2unstr_2003}.
-%
-%\paragraph{}
-%% TODO double check validity
-%One disadvantage of this landmark scheme is related to the additional burden imposed to the landmark sites.
-The authors claim though that the algorithm requires so little work by the landmarks (maybe just echo to ping messages) that could in effect, act as ``unsuspecting participants''.
-Even if this is the case, the fact that it is not fully distributed, renders the protocol's scalability directly volnerable to any system size increase as well as usuitable for highly dynamic networks such as ad-hoc networks.
-Moreover, fixed points in a network are inherently more exposed to malicius attacks.
-The most significant downside of the algorithm though is that it can lead to an extremely uneven overlay ID distribution causing load inbalances and hot spots.
-Lastly, the scheme is coarse grained when it comes to distinguishing relatively close nodes\footnote{In the worst case, all nodes could ve clustered into a single bin.}.
-%
-%\subsection{Global Soft-State}
-%{\sethlcolor{yellow}\hl{HA: in paper introduction, discussion about
-%disadvantages of top-aware CAN:
-%
-%Techniques to exploit topology information in overlay
-%routing include geographic layout, proximity routing and
-%proximity neighbor selection [3]. With geographic layout
-%such as topology-aware CAN [12], the overlay structure is
-%constrained by underlying network topology. This tech-
-%nique, unfortunately, can create uneven distribution of
-%nodes in the overlay, increasing the chances of overloading
-%nodes and rendering the maintenance cost formidable. Our
-%study shows that, for a typical 10,000-node topology-aware
-%CAN, 5% nodes occupy 85-98% of the entire Cartesian
-%space, and some nodes have to maintain 450-1500 neigh-
-%bors. In Proximity routing, physical topology is not consid-
-%ered when constructing the overlay.
-%
-%[snip]
-%Studies [14] have shown that triangle inequal-
-%ity may not hold in Internet topology. In fact, study from
-%Pastry has shown that the proximity approximation is much
-%worse when using the Mercator topology that is based on
-%the real measurements of the Internet [3].
-%
-%RELATED WORK:
-%
-%Miguel Castro et al [3] divide techniques used to
-%exploit network proximity into three categories: geographic
-%layout, proximity routing and proximity neighbor selection.
-%Proximity neighbor selection is superior in terms of load
-%balancing and proximity approximation. The existing algo-
-%
-%}}
-%
-%\paragraph{}
-%The approach argued by Xu et.al in \cite{xu_globstate_2003} is to build a global map to help choose shorter routing paths, combining the landmark binning method and small scale distance probes to reveal the proximity properties of the underlying network to the overlay.
-This global view of the state is made available to all nodes in order to help them find the best way to route their messages.
-The authors focus on two aspects in order to accomplish their goal.
-%
-%\paragraph{}
-%Initially, is the \emph{generation} of proximity information and then its \emph{effective exploitation}.
-For the first, a hybrid approach is proposed, which uses landmark clustering as a preprossesing step in order to select a number of potential nearest neighbour candidates and then refine the selection by incorporating an RTT scheme to ultimately choose the closest node.
-For the second, the algorithm chooses a different path from the classic gossiping approaches for constructing and maintaining the overlay.
-It is based on landmark-clustering-based strategic placement of proximity information on the overlay enabling any node to access such information using a landmark number that reflects its physical position in the network.
-For various logical regions\footnote{This might be a high-order zone in the eCAN\cite{xu_ecan_2002} context or a set of nodes sharing a particular prefix in overlays such as Pastry.} maps of physical information are built and published where each node may appear in a maximum of $log\left( N \right)$ such maps..
-To dynamicaly adapt to changing network conditions, a node subscribes to relevant \emph{soft states} that utilize a notification system in order to initialize any necessary neighbour re-selection.
-%
-%\paragraph{}
-%Maintaining several host states at different layers, makes any content migration costly.
-Additionaly, the method does not make any continuing effort to remap the overlay structure after a node successfuly joins, in order to adapt its state to any occurance of condition change.
-Although this approach greatly reduces the routin latency to far nodes, it is unable to dynamicaly identify nodes that are close to routers and gatways in order to construct the secondary overlay.
-Nevertheless, static recognition of such nodes is currently done based on BGP reports and prechosen landmarks, sucrificing the self-organising attribute of traditional DHTs.
-%
-%\subsection{Self-Adaptive Topology Matching}
-%
-%\paragraph{}
-%Ren et.all's goal while designing the \emph{SAT-Match} algorithm \cite{ren_satmatch_2004} was to create a protocol that would be decentralised and scalable, fine-grained in detecting changes and adaptively reform in response to the dynamic nature of a classic p2p environment and last, but certainly not least, of low incuring cost.
-The method is in a nutshell, a two-phased iterative process that focuses on local optimizations in order to collectively achieve a global full overlay optimization.
-The iteration is finised when the node detects that it is physicaly close enough to its neigbours so that no additional optimization is needed.
-The paper defines \emph{stretch}, $S = \frac{\bar{L_l}}{\bar{L_p}}$, as a way quantify the \emph{topology match degree} of the constructed overlay, where $\bar{L_l}$ is the average logical link latency while the $\bar{L_p}$ is the average physical link latency.
-%
-%\paragraph{Probing phase}
-%\emph{SAT-Match} uses a small TTL value for the probing messages in order to reduce redunduncy\cite{jiang_lightflood_2008}.
-This process begins as soon as the node joins the network using a DHT mechanism.
-Each probing message contains information about the source and a small TTL value.
-The recipient of such a message returns information about itself to the source and forwards the probing message to its neighbours if the TTL is non-zero.
-The nodes been covered, are refered to as $TTL-k$ neighbourhood of the source node\footnote{Especialy, $TTL-1$ neighbourhood referes to the source's direct neighbours}.
-The list of responses is used to measure RTT to these nodes and sort the list ascending RTT order.
-%
-%\paragraph{Jump phase}
-%Blindly selecting the peer with the smallest RTT as neighbour is, generally, not the right desision to make in order to achieve global \emph{stretch} reduction, This is because, in a structured scheme, when a node jumps to connect to a physically close node, it may need to connect to other distant nodes to maintain the structure's integrity, thus creating an overall increase in the overlay's \emph{stretch}.
-The two nodes with the smallest RTT is then used in order to select one zone to jump in this phase.
-The algorithm is as follows: The source node $S$ calculates the stretch change of its $TTL-1$ neighbourhood and that of the $TTL-1$ neighbourhood of the first of the previously selected peers.
-These calculations are made as if the jump has been made.
-If the stretch reduction is over a predefined threshold the jump is performed, otherwise the second selected candidate is picked and the same computations are performed.
-If again, the threshold is not met, then no jump is ultimately done.
-In case of a jump, this is performed as a combination of \emph{leave} and \emph{join} operations, in the CAN context.
-%
-%\paragraph{}
-%Moreover the algorithm takes several issues into consiteration in order to furtherly improve the resulted overlay.
-For example when multiple nodes try to jump simultaneously into a regeon, then the logical link brakes from one attempt may result in inacurate computation of the gain factor, for an other.
-This situation is identified as \emph{contention} and the nodes use an exponatial back-off algorithm to avoid it.
-An other problem is the uneseccary traffic incuerred by the probing phase in a region that after several jumps has settled to a stable state.
-In these cases it is more likely for jump attempts to be proven worthless.
-The algorithm doubles the probing period of such nodes, every time a jump is not taken.
-%
-%\paragraph{}
-%The authors claim that this continuously adaptive mechanism achieves global topology matching optimization in a sufficiently large scope.
-This also secures the fast adaptation to frequent network changes.
-It also considered lightweight and can easily be embedded into current p2p systems, as well as effectively combined with other techniques, such as landmark binning.
-%
-%\subsection{Delay Aware P2P System}
-%A new \emph{Delay Aware P2P System (DAPS)} is introduced in \cite{zhang_daps_2005}.
-Its main goal is to reduce the time $L$ of a look-up request by dividing the routing tables of peers into several sectors in increasing delay.
-The source node emitting the query message defines a delay boundary or the pruning factor $L_t$ in the paper's context.
-Request messages will be forwarded only to nodes whose delay less than or equal to $L_t$.
-With the clustered routing tables and the loose organisation the overlay network of \emph{DAPS} is between structured and unstructured.
-%
-%\subsection{Mithos}
-%
-%\paragraph{}
-%Waldvogel and Rinaldi proposed a protocol\cite{waldvogel_mythos_2003} which incorporates a directed incremental probing to find near optimal node placement.
-%
-%\paragraph{Neighbour detection}
-%During bootstraping, the new comming node needs to know how to contact at least one of the existing members.
-A subset of these nodes will be used as the first set of candidate neighbours.
-Then, iteratively, each of the peers in the cadidate set is asked for their neighbours in order for each of the later to be probed for their network distance from the new coming one.
-The closest node is then used as the new candidate neighbour and the process is repeated until no further improvement is detected.
-Due to the fact that the algorithm may ultimately reach a local minima instead of a global one, \emph{Mithos}' approach is to first probe all neighbours within a two hop distance from the current minimum before concluding the process.
-%
-%\paragraph{ID assignment}
-%After finding its first neighbour, the newcoming peer must be assigned its ID, a very crusial selection in order not to create many local minima that will prevent efficient future neighbourhood location.
-\emph{Mithos} uses information gathered during the previous step, in order to compute the ID, which inccurs no further communication costs to the algorithm.
-This information includes
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item the two closest nodes and their neigbours, and 
-%  \item their coresponding distances.
-%\end{inparaenum}
-%Using the above, it then assigns coordinates to the new-coming node, so that Euclidian distances between the node and all known hosts predict the network latency between them\cite{cox_vivaldi_2004}.
-These synthetic coordinates as explicitly used as the node's ID and as soon as it has been established, distances can be computed in the ID space, no longer requiring physical measurements.
-%
-%\paragraph{Link establishement}
-%The last step of the algorithm is the interconnection among neighbours.
-\emph{Mithos} uses a \emph{quadrant}-based mechanism according to which each node establishes a link to the closest neighbour in each quadrant.
-During forwarding, the next hop is performed towards a neighbour in the same quadrant as the final destination\footnote{An implementation of this could be backed by a $d$-bit vector indexing (where $d$ is the number of dimensions) into the routing table.
-Then the next hop is identified by computing the difference between the current node's vector and that of the destination's.}.
-The problem is that after the ID assignment process, the new-coming node may not know of other neighbours in all quadrant and if even if it does, this cannot ensure that they are the nearest available.
-Thus, the node first identifies neighbours in all quadrants using a mechanism based on ideas similar to a perimeter walk\footnote{Used in Greedy Perimeter Stateless Routing (GPSR) protocol.} and then using parallel path processing improves the results by taking into account further geometric properties of node relationships.
-%
-%\paragraph{}
-%% TODO review this part
-%In order to avoid local minima during neighbour detection, extensive probing must be undertaken.
-In simulation, unfortunately, only very small-sized overlay topologies (of 200 to 1000 nodes) have been used and thus no safe conclusions can be made as for the behaviour of an extensively large, real-world p2p deployment of the scheme.
-%
-%\subsection{DHT-PNS}
-%
-%\paragraph{}
-%The work in \cite{hancong_pnsbased_2006} propose a proximity neighbour selection scheme on top of the Chord DHT.
-Using the Vivaldi protocol\cite{cox_vivaldi_2004} each node is assigned synthetic $2$-dimension coordinates that can be used to derive network latency using Euclidian distances in the id space and without the need of explicit probing.
-Then the algorithm performs the two steps described in the following paragraphs.
-%
-%\paragraph{Space mapping}
-%The space is partitioned using a \emph{concentric circle clustering scheme} where succesive cycles of radiuses $\rho$, $2\rho$, $3\rho$ and so on, are constructed.
-Then the formed annuluses are divided into $2\chi-1$ \emph{sectors}, where $\chi$ denots the sequence number of the annulus starting from $\chi = 1$ for the centre cycle.
-It is proved in the paper, that this way each sector occupies the same area\footnote{The same area as does the center cycle.} and assuming uniform node distribution, this characteristic, favours a more load balanced clustering operation.
-Every sector in this $2-d$ coordinate space is mapped to a unique \emph{region} in the DHT space forming a multi-layer node identifier space.
-Thus, any nodes that belong to the same sector, are mapped to the same region as well, preserving their proximity relationship unveiled by the use of the Vivaldi protocol.
-%
-%\paragraph{Routing table optimization}
-%The system described in the previous section allows any node $\alpha$ to obtain its DHT's $key_{\alpha}$ and its region's $key_r$, in a fully distributed manner, just by applying a consistent hash function.
-Using a $Get\left( key_r \right)$ RPC call in Chord, the node can obtain the region's master node, called \emph{Cluster Node (CN)} which is responsible for clustering the nodes belonging in the same sector or region with that of $\alpha$.
-On the other hand, a $Put\left( key_r, info\right)$ RPC call, registers $\alpha$ to its corresponding region and publishes its information, through the special peer CN.
-$\alpha$ can, additionaly, ask CN for other nodes that have previously joined the region in order to add them into its neighbour set for future routing table optimization.
-Even in the case when no neighbour is detected in the current region, the search is expanded to adjasent regions and towards an upper layer identifier space until one is found or the first layer reached.
-%
-%\subsection{Quasi-Chord}
-%
-%\paragraph{}
-%The approach that is proposed by Sun and Zhang in \cite{sun_quasi_2008} confronts the topology mismatch problem, in the Chord DHT context, that is created by the fact that no consiteration for the underlying physical network topology is taken into account during the construction of the identifier cycle.
-To construct a \emph{Quasi-Chord} network three steps are needed.
-First, each host acquires its coordinates in the geometric space utilizing the \emph{global network position (GNP)} protocol \cite{ng_gnp_2001}.
-Second, using the \emph{Cantor space filling curve} the $2$-dimensional space is converted to a $1$-dimensional one, used in the last, third, step to build the Quasi-Chord circle according to the host's Cantor value.
-In the following paragraphs, these steps are furtherly discussed.
-%
-%\paragraph{GNP coordinates}
-%First of all the host must be positioned in the geometric space.
-The algoritm models the P2P network with a well defined distance function in such a way, it can predict, with high accuracy, the distance between any two points in the space by just evaluating the output of the distance function on the coordinates of these points.
-This is accomplished with the \emph{global network position (GNP)} protocol which\begin{inparaenum}[\itshape i\upshape)]
-%  \item creates a reference set of $N$ landmark nodes so as to minimize the error of ICMP measured distances and coordinate computed ones between them, and then
-%  \item each host is able to measure its round-trip times to the $N$ Landmarks in order to compute its own coordinates.
-%\end{inparaenum}
-%
-%\paragraph{Cantor SPF}
-%After the $2$-d coordinates are set, in the next stage of the algorithn, each participating peer is assigned a Cantor value, according to the application of a \emph{space filling curve} on the coordinate space.
-(TODO: add figure).
-This results in the conversion of the $2$-dimensional space to a $1$-dimensional Cantor space that can more easily be more mapped to the Chord hierarchy.
-%
-%\paragraph{Quasi-Chord construction}
-%As can intuitively be infered by observing the Cantor chart, close-by nodes in the physical layer, are more likely to have similar Cantor values.
-This attribute is exploited in order to construct a topology-aware identifier space for the Chord DHT.
-The cycle is constructed by sorting nodes in ascending order.
-Each host maintains 2 finger tables, one for clockwise and one for counter-clockwise stepping.
-This helps with the connectivity of the network because its not allowed to connect the first node with the last one since this will incur heavy traffic to the later\footnote{After all their Cantor values denote that they are actually the furthest of each other.}.
-%
-%\paragraph{}
-%% TODO: doublecheck!
-%The disadvantage of the algorithm is that the coordinate assignment in the first stage is backed by a not fully distributed landmark-based algorithm.
-Moreover the Quasi-Chord model build-up is making an indirect assumption of a maximum number of allowable hosts since it is constructed.
-Last but not least the doubling of the required routing information which needs to be created and maintained is an additional negative point to the efficiency and the scallability of the algorithm.
-%
-%\subsection{LAPTOP}
-%{\sethlcolor{yellow}\hl{HA - two important issues for decentralized structured P2P:
-%P2P applications. However, the design of a decentralized but structured P2P network has to
-%overcome two critical issues. The first issue is the long routing latency. Several proximity
-%schemes  have been proposed to avert long routing latency in current structured P2P
-%networks, but they require a high-complexity procedure to periodically maintain the routing
-%table (e.g. Pastry system) or they need pre-chosen landmarks to construct the overlay.
-%However, the P2P system is by its very nature unstable since nodes join and leave frequently.
-%For instance, the study of Gnutella shows around approximately 1200 membership changes
-%per minute in a 100 000 nodes P2P system. Another proximity scheme needs some pre-chosen
-%landmarks or a complete BGP routing table support. As a result, they both increase the
-%difficulty of the P2P system deployment.
-%The second issue is system maintenance overhead. The existing structured P2P networks allow
-%nodes to keep some nearby nodes in their routing tables in order to achieve efficient routing. The
-%}}
-%
-%\paragraph{}
-%Laptop \cite{wu_laptop_2007}, introduced by Wu et al., organises the overlay into a tree based hierarchy with main focus on child-to-father relationships in order to reduce hops during message routing as well as minimize maintenance overhead.
-Additionally, a caching scheme is also incorporated so as to furtherly reduce routing table update cost.
-The authors argue that Laptop overlay network is bounded by $O\left( log_d N \right)$ hops in a balanced overlay tree, where $N$ is the number of nodes, and $d$ is the maximum degree of each node.
-It utilizes a geographical layout approach  and constructs a geographical layout in a self-organizing and efficient fashion, by estimating the round trip time (RTT) to a small number of nodes in the overlay network in order to make them roughly aware of their physical distances among them.
-%
-%\paragraph{}
-%The protocol is based on four definitions.
-%\begin{enumerate}
-% \item The amplitude of all possible measured RTTs is devided into intervals.
-Each node measures its distance to its parent and is assigned a label $L_i$ where $i$ denotes the configurable RTT interval in which the measured distance falls into.
-A special kind of node, the root, is initially assigned the $L_1$ label and maintains (as all $L_1$ nodes do) a list of other $L_1$ nodes in the overlay.
-% \item Any node can have children with level lower than theirs, except an $L_{max}$ node which can only have $L_{max}$ level children and only in the case when its parent has reached its maximum degree.
-% \item Each node is assigned an address in a dotted format (e.g 1.3.4).
-Each octet ranges from $1$ to $d$, where $d$ is the maximum degree of the nodes.
-The assignment process is done by appending a unique octet to the address of its parent.
-Root node is assigned address 1.
-% \item For any descendant node $Y$ of a node $X$, the measured distance among each other, must always be less than the lower bound of the RTT interval denoted by $X$'s label.
-%\end{enumerate}
-%
-%\paragraph{}
-%The routing scheme is similar to the IP's longest-prefix matching scheme.
-At each forwarding hop, any message travels up the tree until the first common uncestor of source and destination node is reached and then starts descending to arrive to its target.
-During tree traversals, special entries in the routing tables, called \emph{routing cache}, are maintained in order to increase routing efficiency and achieve finer load balance.
-Caching enables a node to forward a message to a better longest-prefix match than that of its direct ancestor making a large, quicker and more cost effective step through the overlay and toward the destination.
-To improve scalability, the number of children nodes and the size of the routing cache are limited.
-%
-%In terms of overlay maintenance, Laptop incorporates a simple \emph{heartbeat}-based technique where only the parent node is responsible for monitoring its children.
-%
-%At join process the newcomming node is assigned its level label as well as its address by its parent node.
-Additionaly it initializes its routing table (with normal and caching entries) as it traverses the overlay in search for its parents node.
-%The newcommer $N$ locates the root node and the later responds with a list of $L_1$ nodes.
-$N$ then probes each of the $L_1$ nodes in search for the closest one.
-If the measured RTT to the closest $L_1$ falls into the first interval then the newcommer becomes a $L_1$ node as well.
-Otherwise node $N$ sets the closest $L_1$ node as its potential parent node.
-This potential parent becomes the actual parent if it does not have any other children.
-If it has, $N$ gets a list of these $L_i$ nodes and by measuring the RTT to each of them tries to spot a new potential parent in order to repeat the above process.
-%
-%\paragraph{}
-%During a gracefull departure, the refered node checkes for children in the overlay.
-If it does not have any, it simply notifies its parent and leaves.
-If it has, it selects the child node with the lowest RTT to it in order to take its place so that the locality property is preserved.
-%
-%In case of an arbitrary failure, the children of the failed node are detecting their parent's absent when the later stops acknowledging their heartbeat messages.
-They start emmiting special messages to their grand parent node\footnote{The address of the grandparent node is stored during join process}.
-In case of no response from the grandparent node, children invoke the joining procedure to detect their new parent node.The parent of the failed node, aggregates the notifications from the above children nodes for a period of time, and then chooses the one with the lowest level label as the takeover node.
-Potential ties are broken by favouring the lower RTT.
-The parent of the failed node, finally informs accordingly all the children about the change in the hierarchy.
-%
-%\subsection{IP-based Clustering (IPBC)}
-%
-%\paragraph{}
-%\emph{IP-based clustering} \cite{karwaczynski_ipbc_2007} is a proximity neighbour selection technique that is based on a longest IP prefix matching scheme in order to quantify the proximity among peer nodes.
-The relation of the IP address and the physical location of a node in the topology of the Internet, is intuitevely realized by operation the network layer of the Internet stack.
-For example, nodes in the same local subnetwork share the same 3 bytes of their address.
-Reports in \cite{freedman_iploc_2005} state that almost $97\%$ of prefixes longer than 3 bytes correspond to address assignments at a single geographic location.
-Moreover, observing the \emph{IANA IPv4 Address Space Registry} we can infer that, since blocks of consecutive octet prefixes are assigned to the same Regional Internet Registries (RIRs), nodes inside will be physicaly close to each other.
-%
-%\paragraph{}
-%The implementation in this paper is based on the above observation in order to create proximity information.
-This information is stored in a decentralised manner, within the overlay itself, just like any other resource would have been.
-In order to be published in the overlay, each node is first assigned a unique identifier and subsequently generates a key by hashing a fixed-length prefix of its IP.
-Authors argue on the prefix length to be used, as there is a thin balance between reducing the propability of finding closer nodes by adopting a long prefix and overload nodes that are responsible for information published by many peers when chosing to hash a short one.
-Their verdict is for the use of a 16-bit wide prefix for real world systems deployed in an Internet scale.
-In either case, ID and IP is then stored in the DHT using this generated key.
-This way, any node, at any time (either at join time or during peer lifetime for topology adaptation) can acquire information about close-by nodes just by quering the DHT for a specific key.
-Moreover, the algorithm takes care of the freshness of the proximity information in two ways.
-First, the advertising nodes themselves periodically update their advertisements or when they voluntarily leave the overlay, they explicitly remove their data.
-On the other hand, in case of a failure each publication is assigned an expiration time, and thus ultimately removed by the DHT maintenance mechanisms.
-%
-%\subsection{CHOord considering Proximity on IPv6 (CHOP6)}
-%
-%\paragraph{}
-%\cite{morimoto_chop6_2007} roughly estimates the proximity among nodes by exploiting the IPv6 address format as well as RTT information if this is available.The first is achieved by introducing a 64-bit ID scheme in which the least significant bit part\footnote{The exact range is a predefined system variable} is the IPv6 global routing prefix and thus enabling a longest prefix match scheme.
-The observation in which the protocol is based roots from the IPv6 address block assignment.
-Specicaly, block of /16 or /23 in size are assigned to Regional Internet Registries (RIRs) by Internet Assigned Numbers Authority (IANA).
-An RIR divides the signed address blockes into smaller address blocks which are furtherly assigned to Natinal Internet Registries (NIRs).
-That way, it is possible to estimate a node's geographical location by simply examining the upper 32-bits of its IPv6 address.
-Moreover, similar to Chord, CHOP6 uses a finger table, whose entries hold more than one candidate nodes.
-There are three cases in which a node chooses the next hop according to information it posseses.
-%\begin{itemize}
-% \item When no information about RTT is available candidate next hops, the sender just selects a node in the finger table entry which shares the longest prefix with the destination.
-% \item There is another possibility that after some communication with other nodes, the source node should know of the RTTs to some of the nodes in finger table entry.
-In this case, the source node chooses the one with the smallest RTT with propability $p$.
-If there is a node with no measured RTT then the sender can select such a node with probability $1 - p$
-% \item In this last case the source node has already communicated with all candidate node in the finger table entry.
-Thus, the node selects the node whose RTT is the smallest with probability $p$, the one with the second smallest with probability $q$ and so on, where $0 \leq \ldots < r < q < p < 1$ and $p+q+r+\ldots \leq 1$
-%\end{itemize}
-%
-%\subsection{Proximity in Kademlia}
-%
-%\paragraph{}
-%In \cite{kaune_pkad_2008}, Kaune et al., studied the Kademlia DHT, in order to build a proximity aware scheme that would work in the context of iterative lookup algorithms.
-The protocols discussed, focus on both the overlay and underlay tiers in the sense that in the first improves the routing performance according to a cost metric that is provided by the later.
-This metric is tailored to the needs of a specific aplication.
-For example, routing could be optimised to maximise the within-ISP traffic or reduce the lookup latencies or even avoid contacting untrustworthy subnetworks.
-%
-%\paragraph{}
-%Two algorithms are presented for the overlay optimization.
-One is \emph{Proximity Neighbour Selection (PNS)} and one for \emph{Proximity Route Selection (PRS)}.
-The first aims at keeping peers with the least contact cost in the routing table.
-As Kademlia constantly learns new peers (incoming requests, through iterative lookup) no special algorithm for searching more cost effective peers is necessary and simply choose the best peers seen so far.
-The later (PRS) aims at choosing the best next hop during routing a message.
-Due to the fact that the routing in Kademlia is iterative, it is the initiator of the lookup that chooses each next hop from a set of candidate nodes.
-The ``vanilla'' Kedemlia always chooses the closest node with respect to the XOR metric but PRS Kademlia chooses the one with the smallest underlay metric cost.
-Thus, as in all such approaches, there is a trade-off between the overlay and underlay distances.
-%
-%An underlay metric provides information about the underlay network. The paper distingushes between three kinds of possibilities to quantify it:
-%\begin{itemize}
-% \item Using measurements gained by previous lookups
-% \item Using measurements acquired by other peers or jointly calculated with others
-% \item Using a local database to look up information
-%\end{itemize}
-%In the paper, for locality of traffic and subsequently for reducing the lookup latency, a clustering scheme has been implemented exploiting information stored in a geographic information database (GeoIP,MaxMind) about peers.
-The goal of such a metric is to constrain the largest portion of communication within the limits of a peer's ISP.
-%Another incorporated approach is Vivaldi\cite{cox_vivaldi_2004}.
-At first all coordinates are random but as peers start to communicate they calculate RTTs as well, gradually updating their own coordinates.
-Vivaldi creates a system where a peer knowing the coordinates of a another, can approximate communication latency without the need of additional prompting.
-%
-%\subsection{Cone}
-%
-%\paragraph{}
-%Cone\cite{wang_cone_2007} uses a two-layered identifier space.
-The first, named Chord-layer identifier, denoted as $Id_{Chord}$, is the same as in ``vanilla'' Chord.
-The second is the Cone-layer identifier, $Id_{Cone}$ which is constructed by two component identifiers.
-The first, known as \emph{group identifier (gid)} denotes a relevant group the node belongs to while the second, namely \emph{local identifier (lid)} indicates the local identifier within the group.
-The group concept, which is introduced here, is a way of dividing nodes according to a common $Id_{Chord}$ prefix.
-%The structure of a Cone overlay, retains the Chord's circular topology.
-The differce lies on the fact that, now, two rings are created.
-A big ring, where nodes with the same \emph{gid} are arranged at each position.
-Each of these positions are a smaller ring for the particular group's \emph{lid}s.
-The routing is achieved in both clockwise and anti-clockwise direction in the big-ring.
-For this reason two routing tables are maintained, namely \emph{front} and \emph{back} finger tables, respectively.
-Entries in these tables, display physical network proximity with the current node.
-Moreover, a third table called \emph{group} table maintains information about other online peers within the current node's group in a way that entries are now close in the ID space.
-%
-%\paragraph{}
-%Routing in Cone, comprises of the inter-group algoritm and the intra-group algorithm.
-First the group of nodes to which the desired key lies is detected, exploiting physical proximity information (front and back finger tables).
-Second during the next and last hop, the message is forwarded to the exact node that contains the desired key.
-%
-%Cone uses Landmark+RTTs to generate proximity information and exploits this information using proximity neighbour selection.
-%
-%\subsection{DynaMO}
-%
-%\paragraph{}
-%In their work in \cite{winter_dynamo_2004}, Winter et al, try to build an overlay network, called \emph{DynaMO}, that tries to consider not only the physical proximity of peers but their mobility attributes as well.
-DynaMO is based on a Pastry overlay network.
-This was an intentional choice from the authors because Pastry's built-in locality heuristics are thoroughly analised in the literature \cite{castro_proximityp2p_2002} providing good backround against which to test and compare their results.
-In order to adapt to mobile, ad-hoc networks, though, special care has been put to the maintenance of an even overlay ID distribution so that hot-spots are avoided.
-Pastry assignes them in a randomised fashion and then tries to consider proximity through the joining process and routing table maintenance.
-On the other hand, DynaMO dictates a newcoming node to gather information concerning its physical neighbourhood and uses it to assign itself an appropriate overlay ID.
-%
-%DynaMO tries to capitalize an observation of Pastry's routing mechanism.
-Each node's table consists of rows equal to the number of digits of the overlay IDs and columns equal to the ID's base.
-As we go down the table rows, the matching prefix between the current node's ID and the row's entries increases by one.
-Thus, the leaf set contains the closest nodes in the ID space.
-Additionally as the prefix match increases by one, the result is exponentially less candidates that can fill the tables entries as the row increases.
-This leads to the observation \cite{antony_pastry_2001,castro_proximityp2p_2002} that from overlay routing hop to overlay routing hop the physical distance between nodes is likely to increase leading to a dominating last routing step in terms of the overall plysical routing path distance travelled during a key lookup.
-Since the last routing step is usually taken from the leaf set, DynaMO focuses on making this last step as physicaly close as possible.
-In this context, two approaches are considered namely \emph{Random Landmarking (RLM)} and \emph{Closest Neighbour Prefix Assignment (CNPA)}.
-%
-%\paragraph{}
-%Random Landmarking uses the overlay lookup mechanism to locate nodes responsible for a fixed set of carefully chosen\footnote{In a way that the ID space is divided into equal portions} landmark keys.
-This means that when a node is assigned a landmark key then it plays the role of a temporary landmark for the network.
-Any joing node will measure its distance to that and all other landmarks and assign to itself an ID consisting of a prefix taken from the closest landmark and a remainder that could be assigned randomnly or generated by mechanism that takes into account physical neighbourhood.
-The legth of the prefix can be determined as $prefix_length=|log_b k|$, where $b$ is the ID base and $k$ the number of landmarks.
-The above scheme results in physically close nodes, forming regions of common ID prefix that are likely to be close to each other in the ID space and thus bringing a node's leaf set closer to itself.
-Additionally, the advantage of dynamic landmark nodes is that the failed ones can instantly be replaced by new responsibles, that share similar physical attributes to the failed and thus preserving the balance of the network.
-%
-%\paragraph{}
-%RLM may incur more traffic especially on landmark nodes, load that in some cases is not acceptable.
-CPNA on the other hand takes advantage of Pastry's specification that a new coming node is bootstraped by a physicaly close node.
-The new comer then assumes the ID prefix of that neighbour while the rest is generated similarly to RLM.
-Unfortunately, less overhead comes at the expense of being more coarse-grained.
-%
-%\paragraph{}
-%Both algorithms, are protected against the formation of physical landmark clusters or imbalanced ID distribution\footnote{More common during the initial formation of an overlay network} , by introducing the \emph{landmark gravitation range} as a theshold over which landmark keys are reassinged (for the RLM approach) or unutilized ID prefix ranges are detected and used (for the CPNA scheme) in order to balance the distribution of regions in the overlay.
-%
-%\subsection{PChord}
-%\emph{PChord}\cite{hong_pchord_2005} is a based on the Chord DHT which adds proximity routing into its routing mechanism.
-The main modification over the ``vanilla'' Chord is the inclusion of a \emph{proximity list} into its routing table so that the next hop is decided based not only by considering best progress towards the key, but on the physical proximity of the candidate nodes as well.
-The list is, at join time, empty but as the PChord node starts to interact with other nodes it applies a heuristic mechanism to fill up the list.
-Entries are dynamicaly added or removed as the network state is constantly changing.
-Because, it is guaranteed that at each routing step there is progress towards the target node in the ID space, PChord will result in lower hop number than Chord, for each hop is larger or at least equal in key space in PChord than in Chord.
-Additionaly, passing through proximity links in the underlay means reduced routing cost.
-Moreover, if the number of entries in the proximity list is the same as the number of network partitions, PChord prevents hops from jumping back to the same network the current node belongs to.
-%
-%\subsection{AChord}
-%\emph{AChord}\cite{dao_achord_2006} uses IPv6's anycast functionality in order to
-%\begin{itemize}
-% \item releave the protocol from complex joining procedures, and
-% \item achieve high accuracy of network proximity giving high routing efficiency,
-%\end{itemize}
-%all with a simple and lightweight mechanism that requires few changes to Chord without affecting its own advantageous characteristics.
-Anycast delivers a message comming from the outside of an anycast group to the physicaly closest node in that group.
-AChord organizes all nodes participating in the overlay network into an anycast group.
-Any joining node comes outside of that group and, thus, is automaticaly forwarded to the physicaly nearest node in order to bootstrap, avoiding
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item the need to, explicitly, maintain such nodes, and
-%  \item the effort of finding a way of locating the physically nearest among them.
-%\end{inparaenum}
-%The ID of the new comming node is computed based on the ID of the bootstrap node and bootstrap's predecessor in a way that its ID will position itself between the formentioned.
-%
-%After joining, the finger table is build the same way as in the Chord protocol.
-Moreover, additional nodes are maintained in a structure called \emph{neighbourship table} which stores information about the closest known nodes\footnote{As a first entry, at bootstrap, the node stores information about the bootstrap node.}.
-The routing decision is made by using both the neighbourship and the finger table.
-%
-%\subsection{Chord6}
-%\emph{Chord6} \cite{xiong_chord6_2005} is a Chord variant that tries to exploit the hierarchical features of IPv6 in order to create a substrate that reduces interdomain trafic and does that in a cost efficient way.
-Chord6 bareley modifies the original Chord protocol, only in the part of identifier definition.
-This, renders the aproach easily portable to other DHTs such as CAN, Pastry and Tapestry.
-In Chord6 the identifier contains two pares: the higher bits are obtained by hashing the  node's IPv6 address prefix of specific length, while the remaining lower bitss are the hash value of the rest of that IPv6 address.
-This way nodes in a domain will be mapped onto a contihuous key space on the overlay network.
-This guarantees that messages are routed in a way that they do not hop in and out of domains many time, thus minimizing overall routing cost.
-(TODO: Recheck validity) However, even though nodes in the same domain would have close identifiers, nodes in two close domains may have very different ones.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2014-01-21 structured.tex
2012.06.11 (Vassilis) Content revisited. Duplicate text has been removed.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Geographic Layout}
+\subsection{Geographic Layout} \label{section:geographic_layout}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 that close by nodes are more likely to suffer collective failures.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% TODO: Geographic Layout
+%
+%\begin{itemize}
+% \item Geographic Layout\\
+%The node IDs are assigned in such a way that nodes close by in the physical
+%network topology, be close in the node ID space as well. Implementations that
+%work relatively well with this approach have been incorporated into CAN. Nodes
+%measure the RTT between themselves and a set of landmarks in order to match the
+%CAN space as much as possible to the physical one. Unfortunately, the approach
+%requires well known landmark servers and for that matter is  not fully
+%self-organizing which can further lead to imbalanced node distribution. On
+%other DHTs, such as Chord or Pastry, another problem emerges.  To gain fault
+%tolerant properties, these protocols, replicate key-value pairs on neighbouring
+%(in the ID space) nodes. When a proximity-based node ID assignment has been
+%used, the needed failure resilience is undermined by the fact that close by
+%nodes are more likely to suffer collective failures.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%
+% TODO: SOME DISCUSSION
+%
+% Xu \textit{et al.} \cite{xu_globstate_2003} state that there are three ways
+% generating proximity information
+%\begin{itemize}
+% \item Expanding Ring Search\\
+%Expanding ring search can be of two forms. First it can utilize the multicast
+%infrastructure in the underlying network in order to emit its messages.
+%Unfortunately such infrastructures are not widely deployed thus the
+%implementations of this way of generating proximity information is limited to
+%blindly flooding the neighbourhood to obtain reasonable results.
+%
+% \item Heuristics\\
+%Heuristics are used in order to reduce the blindness of the expanding ring
+%search and make realizations more efficient and effective. Unfortunately a
+%common problem of all heuristic approaches is the local minimum pitfall in
+%which
+%the search might be caught into.
+%
+% \item Landmark Binning\\
+%Landmark clustering is based on the view that nodes with similar distances to a
+%set of predefined well-known landmark nodes are pretty likely of being close to
+%each other. But this approach has its weaknesses as well, such as the fact that
+%is a rather coarse grained approximation, therefore not particularly well
+%suited
+%for differentiating nodes within close distance to each other.
+%
+%\end{itemize}
+%
+%
+
+%
+% TODO: IN PAPER INTRODUCTION A DISCUSSION ABOUT
+%TOPOLOGY-AWARE CAN
+%
+%Techniques to exploit topology information in overlay
+%routing include geographic layout, proximity routing and
+%proximity neighbor selection [3]. With geographic layout
+%such as topology-aware CAN [12], the overlay structure is
+%constrained by underlying network topology. This tech-
+%nique, unfortunately, can create uneven distribution of
+%nodes in the overlay, increasing the chances of overloading
+%nodes and rendering the maintenance cost formidable. Our
+%study shows that, for a typical 10,000-node topology-aware
+%CAN, 5% nodes occupy 85-98% of the entire Cartesian
+%space, and some nodes have to maintain 450-1500 neigh-
+%bors. In Proximity routing, physical topology is not consid-
+%ered when constructing the overlay.
+%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \paragraph*{\bf Mithos}
-\textit{Mithos} \cite{waldvogel_mythos_2003} is a P2P protocol which
-incorporates directed incremental probing to find near optimal node placement,
-and is classified by its authors as an integration of geographic layout and
-proximity routing overlay optimization methods.
+\emph{Mithos} \cite{waldvogel_mythos_2003} is a P2P protocol which incorporates
+directed incremental probing to find near optimal node placement, and is
+classified by its authors as an integration of geographic layout and proximity
+routing overlay optimization methods.
 
-The bootstrap phase of \textit{Mithos} starts with a subset of existing members
-as the first set of candidates, and while the iteratively closest nodes are
-detected by probing the neighborhood, the candidate neighbor list is updated. In
-order to avoid a local minima \textit{Mithos} probes all the neighbours within a
-two hop distance from the current minimum before concluding the process.
+The bootstrap phase of Mithos starts with a subset of existing members as the
+first set of candidates, and while the iteratively closest nodes are detected by
+probing the neighborhood, the candidate neighbor list is updated. In order to
+avoid a local minima, Mithos probes all the neighbours within a two-hop distance
+from the current minimum before concluding the process.
 
@@
 and all known hosts predict the network latency between
-them\cite{cox_vivaldi_2004}.  The benefit of these synthetic coordinates are
-that they are explicitly used as the node's ID, and distance computations among
-nodes can be done in ID space without requiring physical measurements.
+them\cite{cox_vivaldi_2004}. The benefit of these synthetic coordinates are that
+they are explicitly used as the node's ID, and distance computations among nodes
+can be done in ID space without requiring physical measurements.
 
-The last step of the algorithm is the interconnection among neighbours.
-\emph{Mithos} uses a \emph{quadrant}-based mechanism according to which each
-node establishes a link to the closest neighbour in each quadrant. During
-forwarding, the next hop is performed towards a neighbour in the same quadrant
-as the final destination. The newcoming node may not know of other neighbours
-in all quadrants, therefore, the node first identifies neighbours in all
-quadrants using a mechanism based on ideas similar to a perimeter
-walk\footnote{Used in Greedy Perimeter Stateless Routing (GPSR) protocol.} and
-then improves the results using parallel path processing by taking into account
-further geometric properties of node relationships.
+The last step of the algorithm is the interconnection among neighbours. Mithos
+uses a \emph{quadrant-based} mechanism according to which each node establishes
+a link to the closest neighbour in each quadrant. During forwarding, the next
+hop is performed towards a neighbour in the same quadrant as the final
+destination. The newcoming node may not know of other neighbours in all
+quadrants, therefore, the node first identifies neighbours in all quadrants
+using a mechanism based on ideas similar to a perimeter walk\footnote{Used in
+Greedy Perimeter Stateless Routing (GPSR) protocol.} and then improves the
+results using parallel path processing by taking into account further geometric
+properties of node relationships.
 
@@
 networks.
+
+%
+% TODO: review this part
+%
+%In order to avoid local minima during neighbour detection, extensive probing
+%  must be undertaken. In simulation, unfortunately, only very small-sized
+% overlay topologies (of 200 to 1000 nodes) have been used and thus no safe
+% conclusions can be made as for the behaviour of an extensively large,
+% real-world p2p deployment of the scheme.
 
@@
 incorporated so as to further reduce routing table update costs. The authors
-theoretically show that in \textit{LAPTOP} routing path length is bounded by
-$O(log_d N)$ and node joining and leaving in the overlay network is bounded by
+theoretically show that in LAPTOP routing path length is bounded by $O(log_d N)$
+and node joining and leaving in the overlay network is bounded by
 $O\left( d log_d N \right)$ hops in a balanced overlay tree, where $N$ is the
-number of nodes, and $d$ is the maximum degree of each node. \textit{LAPTOP}
-implements the geographical layout approach  and constructs a geographical
-layout in a self-organizing and efficient fashion, by estimating the
-round-trip-times to a small number of nodes in the overlay network in order to
-make them roughly aware of their physical distances among them.
+number of nodes, and $d$ is the maximum degree of each node. LAPTOP implements
+the geographical layout approach  and constructs its layout in a self-organizing
+and efficient fashion, by estimating the round-trip times to a small number of
+nodes in the overlay network in order to make them roughly aware of their
+physical distances among them.
 
-Each node is assigned an address in a dotted format (e.g 1.3.4). Each octet
-ranges from $1$ to $d$, where $d$ is the maximum degree of the nodes. The
-assignment process is done by appending a unique octet to the address of each
-nodes
-parent, while root node is assigned address 1.  The routing scheme is similar to
-the longest-prefix IP matching scheme. At each forwarding hop, any message
-travels up the tree until the first common ancestor of both source and
-destination nodes is reached and then starts descending to arrive to its target.
-During tree traversals, special entries in the routing tables, called
-\emph{routing cache}, are maintained in order to increase routing efficiency and
-achieve finer load balance. Caching enables a node to forward a message to a
-better longest-prefix match than that of its direct ancestor making a larger,
-quicker and more cost effective step through the overlay and toward the
-destination. To improve scalability, the number of children nodes and the size
-of the routing cache are limited.  In terms of overlay maintenance, LAPTOP
+Each node is assigned a dot-formated address (e.g 1.3.4). Each octet ranges from
+$1$ to $d$, where $d$ is the maximum degree of the nodes. The assignment process
+is done by appending a unique octet to the address of each nodes parent, while
+root node is assigned address $1$.
+
+%The protocol is based on four definitions.
+%\begin{enumerate}
+% \item The amplitude of all possible measured RTTs is devided into intervals.
+%  Each node measures its distance to its parent and is assigned a label $L_i$
+% where $i$ denotes the configurable RTT interval in which the measured distance
+% falls into. A special kind of node, the root, is initially assigned the $L_1$
+% label and maintains (as all $L_1$ nodes do) a list of other $L_1$ nodes in the
+% overlay.
+% \item Any node can have children with level lower than theirs, except an
+%  $L_{max}$ node which can only have $L_{max}$ level children and only in the
+% case when its parent has reached its maximum degree.
+% \item Each node is assigned a dot-formated address (e.g 1.3.4). Each octet
+% ranges from
+% $1$ to $d$, where $d$ is the maximum degree of the nodes. The assignment
+% process
+% is done by appending a unique octet to the address of each nodes parent, while
+% root node is assigned address $1$.
+% \item For any descendant node $Y$ of a node $X$, the measured distance among
+%  each other, must always be less than the lower bound of the RTT interval
+% denoted by $X$'s label.
+%\end{enumerate}
+%
+
+The routing scheme is similar to the longest-prefix IP matching scheme. At each
+forwarding hop, any message travels up the tree until the first common ancestor
+of both source and destination nodes is reached and then starts descending to
+arrive to its target. During tree traversals, special entries in the routing
+tables, called \emph{routing cache}, are maintained in order to increase routing
+efficiency and achieve finer load balance. Caching enables a node to forward a
+message to a better longest-prefix match than that of its direct ancestor making
+a larger, quicker and more cost effective step through the overlay and toward
+the destination. To improve scalability, the number of children nodes and the
+size of the routing cache are limited. In terms of overlay maintenance, LAPTOP
 incorporates a simple \emph{heartbeat}-based technique where each parent node is
-responsible for monitoring its children.  At join process the newcomming node is
-assigned its level label as well as its address by its parent node. Additionally
-it initializes its routing table (with normal and caching entries) as it
-traverses the overlay in search for its parent node.  During a graceful
-departure, the referred node checks for children in the overlay. If it does not
-have any, it simply notifies its parent and leaves. If it has, it selects the
-child node with the lowest round-trip-time to it in order to take its place so
-that the locality property is preserved.
+responsible for monitoring its children.
+
+%
+%The newcommer $N$ locates the root node and the later responds with a list of
+%  $L_1$ nodes. $N$ then probes each of the $L_1$ nodes in search for the
+% closest one. If the measured RTT to the closest $L_1$ falls into the first
+% interval then the newcommer becomes a $L_1$ node as well. Otherwise node $N$
+% sets the closest $L_1$ node as its potential parent node. This potential
+% parent becomes the actual parent if it does not have any other children. If it
+% has, $N$ gets a list of these $L_i$ nodes and by measuring the RTT to each of
+% them tries to spot a new potential parent in order to repeat the above
+% process.
+%
+
+% At join process the newcomming node is assigned its level label as well as its
+% address by its parent node. Additionally, it initializes its routing table
+%(with normal and caching entries) as it traverses the overlay in search for
+%its parent node.  During a graceful departure, the referred node checks for
+%children in the overlay. If it does not have any, it simply notifies its
+%parent and leaves. If it has, it selects the child node with the lowest
+%round-trip time to it in order to take its place so that the locality
+%property is preserved. On the other hand, in the case of an arbitrary
+%failure, after the children of the failed node detect their parent's absence
+%(throught the aformentioned heartbeat mechanism), they start emmiting special
+%messages to their grandparent node (the address of which is stored during
+%join process}. In case of no response from the grandparent node, the children
+%invoke the joining process to detect their new parent node.The parent of the
+%failed node, aggregates the notifications from the above children nodes for a
+%period of time, and then chooses the one with the lowest level label as the
+%takeover node. Potential ties are broken by favouring the lower RTT. The
+%parent of the failed node, finally informs accordingly all the children about
+%the change in the hierarchy.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% TODO: HERE STRUCTURED LANDMARK BINING APPROCH LIES
+%
+%Assuming a structured approach based on CAN\cite{ratnasamy_can_2001} and $m$
+% landmark nodes. The coordinate space is then partitioned into $m!$ equally
+%sized portions, each corresonding to a single ordering of the landmarks. To do
+%this, the first dimension is divided into $m$ areas each of which is furtherly
+%divided (second dimension) into $m - 1$ sections and so on. Having set this $m$
+%dimensional space, at joining time, a node measures the delay to the set of
+%landmarks in order to determine its associated bin and thus position itself in
+%that portion of the coordinate space associated with its landmark ordering.
+%Even though this scheme can
+%
 
@@
 \subsection{Proximity Routing}
+\label{section:pr}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 tables.
+
+% Moreover, the technique has been incorporated into a version of Chord
+% (TODO: PChord | additionally check whether this is true) causing an increase
+% on the overhead of node joins and the size as well as maintenance cost of
+% finger tables.
+
 %
@@
 \paragraph*{\bf Proximity in Kademlia}
-Before discussing {\em Proximity in Kademlia}, it would be helpful to briefly
-summarize the highlights about the {\em Kademlia} protocol. {\em Kademlia}
-\cite{maymounkov_kademlia_2002} is a distributed hash table (DHT) based
-protocol for P2P networks, which uses an iteration based lookup algorithm. The
-protocol uses the standard $160$ bit ID system for nodes and locates the nodes
-in a prefix binary tree, where IDs are used as prefixes. The iterative lookup
-operations are done over this prefix binary tree, which converges to logarithmic
-lookup times. The ID's are assigned randomly, therefore in Kademlia, there is no
-proximity control, which results in inefficient use of the underlying network
-during lookup and retrieve operations.
+% Before discussing {\em Proximity in Kademlia}, it would be helpful to briefly
+% summarize the highlights about the {\em Kademlia} protocol. {\em Kademlia}
+% \cite{maymounkov_kademlia_2002} is a distributed hash table (DHT) based
+% protocol for P2P networks, which uses an iteration based lookup algorithm. The
+% protocol uses the standard $160$ bit ID system for nodes and locates the nodes
+% in a prefix binary tree, where IDs are used as prefixes. The iterative lookup
+% operations are done over this prefix binary tree, which converges to
+%logarithmic
+% lookup times. The ID's are assigned randomly, therefore in Kademlia, there is
+%no
+% proximity control, which results in inefficient use of the underlying network
+% during lookup and retrieve operations.
 
-{\em Proximity in Kademlia} \cite{kaune_pkad_2008} introduces proximity controls
-over the base Kademlia protocol to optimize the underlying network usage of the
-protocol. The authors define an abstract {\em underlay metric} that calculates
-the suitability of establishing a communication link between peers as a cost
-function, based on the used proximity criteria (RTT, ISP locality, etc.).
-{\em Proximity in Kademlia} adapted both {\em proximity routing} and {\em
-proximity neighbour selection} overlay optimization algorithms for controlling
-proximity. Therefore, we categorize the protocol only in this section, and skip
-the discussion in the further {\em proximity neighbour selection} section
-(Sec.\ref{sec:pns}). {\em Proximity in Kademlia} also uses the MaxMind GeoIP
-database\footnote{MaxMind Geolocation Technology. http://www.maxmind.com} to
-detect the proximity information of a given peer based on its region, country,
-or ISP location, and use this information to form clusters of nearby peers. One
-other method {\em Proximity in Kademlia} uses is the Vivaldi
-\cite{cox_vivaldi_2004} protocol. However, authors report that clustering
-approach performs better than Vivaldi protocol based on their experiments.
-Authors also report that {\em Proximity routing} protocol successfully worked
-with the Kademlia protocol and improved the locality of the connections over
-peers.
+\emph{Proximity in Kademlia} \cite{kaune_pkad_2008} introduces proximity
+controls over the vanilla Kademlia protocol's random ID assignment, in order to
+optimize the underlying network usage of the protocol. The work in this paper
+can also be considered, as a more generic study over building proximity in the
+context of iterative lookup algorithms. Focus is given on both the overlay and
+the underlaying tiers in the sense that in the first improves the routing
+performance according to a cost metric that is provided by the later and
+quantifies the suitability of establishing a communication link between peers as
+a cost function, based on the used proximity criteria (RTT, ISP locality, etc.).
+%An underlay metric provides information about the underlay network. The paper
+%  distingushes between three kinds of possibilities to quantify it:
+%\begin{itemize}
+% \item Using measurements gained by previous lookups
+% \item Using measurements acquired by other peers or jointly calculated with
+%  others
+% \item Using a local database to look up information
+%\end{itemize}
+The paper proposes two approaches for overlay optimisation. One falling into
+{\em proximity routing} and the other on {\em proximity neighbour selection}
+(see Section~\ref{section:pns}) approaches overlay optimization techiques for
+exploiting proximity. For the first, the goal is in choosing the best next hop
+during routing a message. Due to the fact that the routing in Kademlia is
+iterative, it is the initiator of the lookup that chooses each next hop from a
+set of candidate nodes. Standard Kedemlia always chooses the closest node with
+respect to the XOR metric but with proximity routing selection, Kademlia
+chooses the one with the smallest underlay metric cost. Thus, as in all such
+approaches, there is a trade-off between the overlay and underlay distances.
+
+Toward the enhancement of traffic locality, the authors also use information
+from the MaxMind GeoIP database\footnote{MaxMind Geolocation Technology.
+http://www.maxmind.com} to retrieve proximity information of a given
+peer based on its region, country, or ISP location, and use this information to
+form clusters of nearby peers. An alternative method could also be the use of 
+Vivaldi coordinates \cite{cox_vivaldi_2004}. However, authors report that
+clustering approach performs better than Vivaldi protocol based on their
+experiments. Authors also report that {\em Proximity routing} protocol
+successfully worked with the Kademlia protocol and improved the locality of the
+connections over peers.
 
@@
 \paragraph*{\bf CHOord considering Proximity on IPv6 (CHOP6)}
-\textit{CHOP6} \cite{morimoto_chop6_2007} is designed based on the
-\textit{Chord} protocol. \textit{CHOP6} roughly estimates the proximity among
+\emph{CHOP6} \cite{morimoto_chop6_2007} roughly estimates the proximity among
 nodes by exploiting the IPv6 address format and RTT information if available.
@@
 simply examining the upper 32-bits of its IPv6 address. Moreover, similar to
-\textit{Chord}, \textit{CHOP6} uses a finger table, whose entries hold more than
+Chord, CHOP6 uses a finger table, whose entries hold more than
 one candidate node.
+
+%
+%\paragraph{}
+% There are three cases in which a node chooses the next hop according to
+% information it posseses.
+%\begin{itemize}
+% \item When no information about RTT is available candidate next hops, the
+%  sender just selects a node in the finger table entry which shares the longest
+% prefix with the destination.
+% \item There is another possibility that after some communication with other
+%  nodes, the source node should know of the RTTs to some of the nodes in finger
+% table entry. In this case, the source node chooses the one with the smallest
+% RTT with propability $p$. If there is a node with no measured RTT then the
+% sender can select such a node with probability $1 - p$
+% \item In this last case the source node has already communicated with all
+%  candidate node in the finger table entry. Thus, the node selects the node
+% whose RTT is the smallest with probability $p$, the one with the second
+% smallest with probability $q$ and so on, where $0 \leq \ldots < r < q < p < 1$
+% and $p+q+r+\ldots \leq 1$
+%\end{itemize}
+%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\paragraph*{\bf T2MC}
+T2MC \cite{shi_t2mc_2008} uses traceroute logs to detect different autonomous
+systems and cluster close by peers, thus reducing redundant traffic that crosses
+autonomous system boundaries. T2MC uses a customized k-mean classification
+algorithm with $k=2$ to perform the classification, and exploits the stable
+structure of the Internet routers to guide clustering. The peer chooses the
+minimum and maximum Latency results from the traceroute to initialize the
+centroinds of two sets. Then the peer calculates, for all hops allong its
+tracerouted path, the absolute distance to these centroids of both sets and
+assigns the routers to that centroid with which it has the smaller absolute
+distance. After that, the mean and variance values are calculated and if the
+variance is larger than a predefined threshold then the algorithm sets the two
+latency mean values as new centroids for the two sets and loops again.
+Ultimately peer will end up with two sets having the minimum intra-set variance.
+The peer, then, chooses the router from the second set with the minimum hops
+attribute and sets it as a threshold. The selected router and all others whose
+hops attribute is larger than the threshold are classified as \emph{remote}
+router cluster. The remaining are classified as \emph{near}. From the near
+class, the peer chooses the one with maximum hops attribute as its edge router,
+and registers it along with the all the near cluster into the DHT of the P@P
+overlay. As new peers join the network, those that share the same edge router
+or any of the members of the near router clustersm there would gather to form a
+close peer cluster. As Edge routers can provide more valuable information than
+other members of the near set, T2MC was designed to prioritize interaction of
+peers and edge gateways.
+
+%
+% TODO: SOME DISCUSSION
+%
+%The use of Traceroute as a tool for implementing the distance measuring
+%infrastructure raise concearns about its efficiency and scalability. Being,
+%primarilly, a network diagnostic utility, it is concearned too heavy weighted
+%and intrucive for use in a larger scale\cite{ratnasamy_binning_2002}.
+%Additionally, disabling ICMP is a common administrative policy for edge sites
+%to enforce security, while dumping BGP routing
+%tables\cite{krishnamurthy_bgpclust_2000} is not directly available to the
+%application layer.
+% Even though traceroute provides detailed information about the network
+% structure, its excessive use creates additional overhead that stresses further
+% the overall network infrastructure. This forces network administrators to
+% disable its support from the routers they manage, thus jeopardizing the
+% effectiveness of the T2MC algorithm.
+%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\paragraph*{\bf PChord}
+\emph{PChord}\cite{hong_pchord_2005} is a based on the Chord DHT which adds
+proximity routing into its routing mechanism. The main modification over the
+vanilla Chord is the inclusion of a \emph{proximity list} into its routing table
+so that the next hop is decided based not only by considering best progress
+towards the key, but on the physical proximity of the candidate nodes as well.
+The list is, at join time, empty but as the PChord node starts to interact with
+other nodes it applies a heuristic mechanism to fill up the list. Entries are
+dynamicaly added or removed as the network state changes. Due to the fact that,
+with each step, there is always progress towards the target node in the ID
+space, PChord will result in a reduced hop number than Chord, since each hop in
+PChord is larger or at least equal in the key space. Passing through proximity
+links in the underlaying network means reduced routing cost. Additionally, if
+the number of entries in the proximity list is the same as the number of network
+partitions, PChord prevents hops from jumping in and out the network partition
+the current node belongs to.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\paragraph*{\bf AChord}
+\emph{AChord} \cite{dao_achord_2006} uses IPv6's anycast mechanism in order to
+releave the protocol from complex joining procedures, and achieve high accuracy
+of network proximity giving high routing efficiency, Anycast delivers a message
+comming from the outside of an anycast group to the physicaly closest node in
+that group. AChord organizes all nodes participating in the overlay network into
+an anycast group. Any joining node comes outside of that group and, thus, is
+automaticaly forwarded to the physicaly nearest node in order to bootstrap,
+avoiding the need to, explicitly, maintain such nodes, and the effort of finding
+a way of locating the physically nearest among them. The ID of the new comming
+node is computed based on the ID of the bootstrap node and the bootstrap's
+predecessor in a way that the constructed ID will be positioned between the
+aformentioned two. After joining, the finger table is build the same way as in
+the Chord protocol. Moreover, additional nodes are maintained in a structure
+called \emph{neighbourship table} which stores information about the closest
+known node. The routing decision is made by using both the neighbourship and
+the finger table. The authors claim this to be a lightweight approach that
+requires few changes to Chord in order not to affecting its own advantageous
+characteristics.
 
@@
 \paragraph*{\bf Chord6}
-\emph{Chord6} \cite{xiong_chord6_2005} is another \textit{Chord} variant that
-tries to exploit the hierarchical features of IPv6 in order to create a
-substrate that reduces interdomain traffic between service providers.
-\textit{Chord6} is based on the original Chord protocol, and the main difference
-from the original protocol is in the identifier definition. Therefore, the
-approach can be easily portable to other DHTs such as CAN, Pastry and Tapestry.
-In Chord6 the identifier contains two parts: the higher bits are obtained by
-hashing the node's IPv6 address prefix of specific length, while the remaining
-lower bits are the hash value of the rest of that IPv6 address. As a result of
-this assignment, nodes in a domain will be mapped onto a continuous key space on
-the overlay network, which avoids unnecessary message forwarding across
-different service providers, thus minimizing overall routing cost.
+\emph{Chord6} \cite{xiong_chord6_2005} is another Chord variant that tries to
+exploit the hierarchical features of IPv6 in order to create a substrate that
+reduces interdomain traffic between service providers. The difference between
+Chord6 and Chord the identifier definition. Therefore, the approach can be
+easily portable to other DHTs such as CAN, Pastry and Tapestry. In Chord6 the
+identifier contains two parts: the higher bits are obtained by hashing the
+node's IPv6 address prefix of specific length, while the remaining lower bits
+are the hash value of the rest of that IPv6 address. As a result of this
+assignment, nodes in a domain will be mapped onto a continuous key space on the
+overlay network, which avoids unnecessary message forwarding across different
+service providers, thus minimising overall routing cost.
+
+%
+% TODO: SOME DISCUSSION
+%       RECHECK VALIDITY
+%
+% However, even though nodes in the same domain would have close identifiers,
+% nodes in two close domains may have very different ones.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Proximity Neighbour Selection}\label{sec:pns}
+\subsection{Proximity Neighbour Selection}
+\label{section:pns}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\textit{Proximity Neighbour Selection} constructs the routing tables using
+\emph{Proximity Neighbour Selection} constructs the routing tables using
 proximity knowledge. The proximity information used in this method is different
-than the landmark based systems described in the \textit{Geographic Layout}
-section, as time-to-live values between nodes, or directly node ID prefixes are
-used to detect proximity.  Tapestry, Pastry, and CAN successfully implemented
-the proximity into their algorithms by using this approach.  The routing
-protocol in Pastry is based on longest node ID prefix matching, while CAN uses
-RTT values to detect close by nodes. \cite{castro_proximityp2p_2002}  reports
-that proximity neighbour selection as an effective proximity based method.
-Proximity Neighbor Selection based algorithms are described in detail below.
+than the landmark based systems described in
+Section~\ref{section:geographic_layout}, as time-to-live values between nodes,
+or directly node ID prefixes are used to detect proximity. Tapestry, Pastry, and
+CAN successfully implemented the proximity into their algorithms by using this
+approach. The routing protocol in Pastry is based on longest node ID prefix
+matching, while CAN uses RTT values to detect close by nodes.
+\cite{castro_proximityp2p_2002} reports that proximity neighbour selection as an
+effective proximity based method. Proximity Neighbor Selection based algorithms
+are described in detail below.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\paragraph*{\bf Proximity in Kademlia}
+As already mentioned on the the corresponding paragraph when discussing
+proximity routing approaches (see Section~\ref{section:pr}), the paper
+proposes \emph{proximity neighbour selection} method, as well, for overlay
+optimisation. The goal, according to this, is to keep peers with the least
+contact cost in the routing table. As Kademlia constantly learns new peers
+(incoming requests, through iterative lookup) no special algorithm for
+searching more cost effective peers is necessary and simply choose the best
+peers seen so far. For more information on the Proximity in Kademlia please
+consult the paragraph in Section~\ref{section:pr}.
 
@@
 \paragraph*{\bf DHT-PNS}
-Chord-DHT-PNS \cite{hancong_pnsbased_2006} implements proximity neighbour
-selection on top of the Chord DHT. The main goal is to use proximity information
+The work in \cite{hancong_pnsbased_2006} propose a proximity neighbour selection
+scheme on top of the Chord DHT. The main goal is to use proximity information
 to group physically close by nodes as neighbours in the DHT table. In order to
@@
 the center cycle. Assuming uniform node distribution, this characteristic,
-favours a more load balanced clustering operation. Every sector in this $2D$
+favours a more load balanced clustering operation. Every sector in this $2d$
 coordinate space is mapped to a unique \emph{region} in the DHT space forming a
@@
 are mapped to the identifier space uniquely, allowing logarithmic lookup
-operations with high probability on Chord.
+operations with high probability on Chord. The system described in the previous
+section allows any node to obtain its DHT key and its region key, in a fully
+distributed manner, just by applying a consistent hash function. Using an RPC
+call in Chord, the node can obtain the region's master node, called
+\emph{Cluster Node (CN)} which is responsible for clustering the nodes belonging
+in the same sector or region with that of the node at hand. An additional RPC
+call, registers it to its corresponding region and publishes its information,
+through the special peer CN. The node can, additionaly, ask CN for other
+nodes that have previously joined the region in order to add them into its
+neighbour set for future routing table optimization. Even in the case when no
+neighbour is detected in the current region, the search is expanded to adjasent
+regions and towards an upper layer identifier space until one is found or the
+first layer reached.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\paragraph*{\bf Quasi-Chord}
+The approach that is proposed by Sun and Zhang in \cite{sun_quasi_2008}
+confronts the topology mismatch problem, in the Chord DHT context which, in its
+vanilla version, takes no consiteration for the underlying physical network
+topology during the construction of the identifier cycle. To construct a
+\emph{Quasi-Chord} network three steps are followed needed. First, each host
+acquires its coordinates in the geometric space utilizing the \emph{global
+network position (GNP)} protocol \cite{ng_gnp_2001}. Second, using the
+\emph{Cantor space filling curve}, the $2$-dimensional space is converted to a
+$1$-dimensional one, used in the last, third, step to build the Quasi-Chord
+circle according to the host's Cantor value. In the following paragraphs,
+these steps are furtherly discussed.
+
+%
+%\paragraph{GNP coordinates}
+%First of all the host must be positioned in the geometric space. The algoritm
+%  models the P2P network with a well defined distance function in such a way,
+% it can predict, with high accuracy, the distance between any two points in the
+% space by just evaluating the output of the distance function on the
+% coordinates of these points. This is accomplished with the \emph{global
+% network position (GNP)} protocol which\begin{inparaenum}[\itshape i\upshape)]
+%  \item creates a reference set of $N$ landmark nodes so as to minimize the
+%  error of ICMP measured distances and coordinate computed ones between them,
+% and then
+%  \item each host is able to measure its round-trip times to the $N$ Landmarks
+% in order to compute its own coordinates.
+%\end{inparaenum}
+%
+%\paragraph{Cantor SPF}
+%After the $2$-d coordinates are set, in the next stage of the algorithn, each
+%  participating peer is assigned a Cantor value, according to the application
+% of a \emph{space filling curve} on the coordinate space. (TODO: add figure).
+% This results in the conversion of the $2$-dimensional space to a
+% $1$-dimensional Cantor space that can more easily be more mapped to the Chord
+% hierarchy.
+%
+%\paragraph{Quasi-Chord construction}
+%As can intuitively be infered by observing the Cantor chart, close-by nodes in
+%  the physical layer, are more likely to have similar Cantor values. This
+% attribute is exploited in order to construct a topology-aware identifier space
+% for the Chord DHT. The cycle is constructed by sorting nodes in ascending
+% order. Each host maintains 2 finger tables, one for clockwise and one for
+% counter-clockwise stepping. This helps with the connectivity of the network
+% because its not allowed to connect the first node with the last one since this
+% will incur heavy traffic to the later\footnote{After all their Cantor values
+% denote that they are actually the furthest of each other.}.
+%
+
+%
+% TODO: SOME DISCUSSION
+%       NEEDS DOUBLECHECK
+%
+%The disadvantage of the algorithm is that the coordinate assignment in the
+%  first stage is backed by a not fully distributed landmark-based algorithm.
+% Moreover the Quasi-Chord model build-up is making an indirect assumption of a
+% maximum number of allowable hosts since it is constructed. Last but not least
+% the doubling of the required routing information which needs to be created and
+% maintained is an additional negative point to the efficiency and the
+% scallability of the algorithm.
+%
 
@@
 \paragraph*{\bf IP-based Clustering (IPBC)}
-Proximity neighbour selection algorithms use probing and other techniques
-to detect proximity. However, such methods are either not precise enough, or
-overload the network. \emph{IP-based clustering}
-\cite{karwaczynski_ipbc_2007} is a proximity neighbour selection based
-algorithm in which the authors propose to use the IP address prefixes (16 bit
-for IPv4) in order to detect proximity.  \cite{freedman_iploc_2005} states that
-$97\%$ of prefixes larger than $24$ bits belong to a single geographical
-location. However, using less number of bits creates less precise results and
-more number of bits increase the burden on the network and reduce the possible
-number of neighbours. Therefore, a careful selection is required in terms of
-performance/accuracy tradeoff. The {\em IP-based Clustering} generates a key and
-its prefix and stores the later in the DHT itself, so that any newly joining
-nodes with the same IP prefix can query it and identify all the
-neighbours with the same one relatively easily. Nodes periodically update
-their entries in the DHT by removing those who leave, or those who has been
-timed out with no further activity detected from the node.
+Proximity neighbour selection algorithms use probing and other techniques to
+detect proximity. However, such methods are either not precise enough, or
+overload the network. \emph{IP-based clustering} \cite{karwaczynski_ipbc_2007}
+is a proximity neighbour selection based algorithm in which the authors propose
+to use the IP address prefixes (16 bit for IPv4) in order to detect proximity.
+\cite{freedman_iploc_2005} states that $97\%$ of prefixes larger than $24$ bits
+belong to a single geographical location. However, using less number of bits
+creates less precise results and more number of bits increase the burden on the
+network and reduce the possible number of neighbours. Therefore, a careful
+selection is required in terms of performance/accuracy tradeoff. IP-based
+Clustering generates a key and its prefix and stores the later in the DHT
+itself, so that any newly joining nodes with the same IP prefix can query it and
+identify all the neighbours with the same one relatively easy. Nodes
+periodically update their entries in the DHT by removing those who leave, or
+those who has been timed out with no further activity detected from the node.
+
+% Moreover, observing the
+% \emph{IANA IPv4 Address Space Registry} we can infer that, since blocks of
+% consecutive octet prefixes are assigned to the same Regional Internet
+% Registries (RIRs), nodes inside will be physicaly close to each other.
+
+%
+% In order to be published in the overlay, each node is first assigned a
+% unique identifier and subsequently generates a key by hashing a fixed-length
+% prefix of its IP. Authors argue on the prefix length to be used, as there is a
+% thin balance between reducing the propability of finding closer nodes by
+% adopting a long prefix and overload nodes that are responsible for information
+% published by many peers when chosing to hash a short one. Their verdict is for
+% the use of a 16-bit wide prefix for real world systems deployed in an Internet
+% scale. In either case, ID and IP is then stored in the DHT using this
+% generated key. This way, any node, at any time (either at join time or during
+% peer lifetime for topology adaptation) can acquire information about close-by
+% nodes just by quering thde DHT for a specific key. Moreover, the algorithm
+% takes care of the freshness of the proximity information in two ways. First,
+% the advertising nodes themselves periodically update their advertisements or
+% when they voluntarily leave the overlay, they explicitly remove their data. On
+% the other hand, in case of a failure each publication is assigned an
+% expiration time, and thus ultimately removed by the DHT maintenance
+%mechanisms.
+%
 
@@
 \paragraph*{\bf Cone}
-\textit{Cone}\cite{wang_cone_2007} extends Chord using proximity neighbour
+\emph{Cone} \cite{wang_cone_2007} extends Chord using proximity neighbour
 selection topology optimization algorithm. The proximity information is
 generated using landmarks and round-trip-time-based distances to landmarks.
-\textit{Cone} uses a two-layered identifier space. The first, named Chord-layer
+Cone uses a two-layered identifier space. The first, named Chord-layer
 identifier, denoted as $Id_{Chord}$, is the same as in Chord. The second is the
@@
 space.
+
+%
+%Routing in Cone, comprises of the inter-group algoritm and the intra-group
+%  algorithm. First the group of nodes to which the desired key lies is
+% detected, exploiting physical proximity information (front and back finger
+% tables). Second during the next and last hop, the message is forwarded to the
+% exact node that contains the desired key.
+%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\paragraph*{\bf DynaMo}
+\emph{DynaMO} \cite{winter_dynamo_2004} tries to build an overlay that not only 
+considers physical proximity among peers but their mobility attributes as well.
+DynaMO is based on a Pastry overlay network and this was a preety concious
+choice, by the authors, basically because of Pastry's built-in locality
+heuristics which are thoroughly analysed in the in the literature
+\cite{castro_proximityp2p_2002} providing good backround against which to test
+and compare their results. In order to adapt to mobile, ad-hoc networks,
+though, special care has been put to the maintenance of an even overlay ID
+distribution so that hot-spots are avoided. DynaMO dictates a newcoming node to
+gather information concerning its physical neighbourhood and uses it to assign
+itself an appropriate overlay ID. DynaMO tries to capitalize on the fact that
+in Pastry each node's table consists of rows equal to the number of digits of
+the overlay IDs and columns equal to the ID's base. As we go down the table
+rows, the matching prefix between the current node's ID and the row's entries
+increases by one. Thus, the leaf set contains the closest nodes in the ID space.
+Additionally as the prefix match increases by one, the result is exponentially
+less candidates that can fill the tables entries as the row increases. This
+leads to the observation \cite{antony_pastry_2001,castro_proximityp2p_2002} that
+from overlay routing hop to overlay routing hop the physical distance between
+nodes is likely to increase leading to a dominating last routing step in terms
+of the overall plysical routing path distance travelled during a key lookup.
+Since the last routing step is usually taken from the leaf set, DynaMO focuses
+on making this last step as physicaly close as possible. In this context, two
+approaches are considered namely \emph{Random Landmarking (RLM)} and
+\emph{Closest Neighbour Prefix Assignment (CNPA)}.
+
+RLM lookup mechanism to locate nodes responsible for a fixed set of carefully
+chosen (in a way that the ID space is divided into equal portions) landmark
+keys. This means that when a node is assigned a landmark key then it plays the
+role of a temporary landmark for the network. Any joing node will measure its
+distance to that and all other landmarks and assign to itself an ID consisting
+of a prefix taken from the closest landmark and a remainder that could be
+assigned randomnly or generated by mechanism that takes into account physical
+neighbourhood. The legth of the prefix can be determined as
+$prefix_length=|log_b k|$, where $b$ is the ID base and $k$ the number of
+landmarks. The above scheme results in physically close nodes, forming regions
+of common ID prefix that are likely to be close to each other in the ID space
+and thus bringing a node's leaf set closer to itself. Additionally, the
+advantage of dynamic landmark nodes is that the failed ones can instantly be
+replaced by new responsibles, that share similar physical attributes to the
+failed and thus preserving the balance of the network.
+
+RLM may incur more traffic especially on landmark nodes, load that in some cases
+is not acceptable. CPNA on the other hand takes advantage of Pastry's
+specification that a new coming node is bootstraped by a physicaly close node.
+The new comer then assumes the ID prefix of that neighbour while the rest is
+generated similarly to RLM. Unfortunately, less overhead comes at the expense
+of being more coarse-grained.
+
+%
+% TODO: SOME DISCUSSION
+%
+%Both algorithms, are protected against the formation of physical landmark
+%  clusters or imbalanced ID distribution\footnote{More common during the
+% initial formation of an overlay network} , by introducing the \emph{landmark
+% gravitation range} as a theshold over which landmark keys are reassinged (for
+% the RLM approach) or unutilized ID prefix ranges are detected and used (for
+% the CPNA scheme) in order to balance the distribution of regions in the
+% overlay.
+%
 
@@
 
-\textit{SAT-Match} \cite{ren_satmatch_2004} is a protocol that specifically
-tries to solve the topology mismatch problem by mapping the overlay network as
-close as possible to the physical one. Even though no theoretical proof is
+\emph{SAT-Match} \cite{ren_satmatch_2004}. Even though no theoretical proof is
 given, valid results are obtained and demonstrated through experimentations.
@@
 \textit{SAT-Match} uses probing to detect close by peers and obtain proximity
-information. However, as an additional feature, \textit{SAT-Match} peers can do
-selective jumps to adjust their positioning in the DHT, if it reduces the stretch
-of its one-hop neighbourhood. Figure~\ref{figure:sat_match} demonstrates the
-selective jump on an example topology. The stretch is defined as the ratio
-between the average logical and physical latency of a link. It is reported that,
-due to the selective jumps, the \textit{SAT-Match} achieves $40\%$ reduction in
-link stretch, and when used with the \textit{Landmark Binning} approach (see
-Sec. \ref{sec:landmark_binning}), the reduction rate increases up to $60\%$.
-For dynamic environments, with frequent node arrival and leave,
-\textit{SAT-Match} scales much better than \textit{Mithos}, due to its self
-adaptation mechanism and selective jumps.
-
-\emph{SAT-Match} uses a small TTL value for the probing messages in order to
-reduce redundancy\cite{jiang_lightflood_2008}. This process begins as soon as
-the node joins the network using a DHT mechanism. Each probing message contains
+information. However, as an additional feature, SAT-Match peers can do
+selective jumps to adjust their positioning in the DHT, if it reduces the
+\emph{stretch} of its one-hop neighbourhood. The paper defines stretch
+$S = \frac{\bar{L_l}}{\bar{L_p}}$, where $\bar{L_l}$ is the average
+logical link latency while the $\bar{L_p}$ is the average physical
+link latency, and uses it a way quantify the \emph{topology
+match degree} of the constructed overlay. For the probing phase SAT-Match uses a
+small TTL value for the probing messages in order to reduce
+redundancy\cite{jiang_lightflood_2008}. This process begins as soon as the node
+joins the network using a DHT mechanism. Each probing message contains
 information about the source and a small TTL value. The recipient of such a
@@
 referred to as $TTL-k$ neighbourhood of the source node based on the $TTL$
-distance to the source node.
-
-Blindly selecting the peer with the smallest RTT as neighbour is, generally, not
-the optimal decision to make in order to achieve global \emph{stretch} reduction.
-In a structured scheme, when a node jumps to connect to a physically close node,
-it may need to connect to other distant nodes to maintain the structure's
-integrity, thus creating an overall increase in the overlay's \emph{stretch}.
-The two nodes with the smallest RTT are then used in order to select one zone to
-jump in this phase. The algorithm is as follows: The source node $S$ calculates
-the stretch change of its $TTL-1$ neighbourhood and that of the $TTL-1$
-neighbourhood of the first of the previously selected peers. These calculations
-are made as if the jump has been made. If the stretch reduction is over a
-predefined threshold the jump is performed, otherwise the second selected
+distance to the source node. Blindly selecting the peer with the smallest RTT
+as neighbour is, generally, not the optimal decision to make in order to achieve
+global stretch reduction. In a structured scheme, when a node jumps to connect
+to a physically close node, it may need to connect to other distant nodes to
+maintain the structure's integrity, thus creating an overall increase in the
+overlay's stretch. The two nodes with the smallest RTT are then used in order to
+select one zone to jump in this phase. The algorithm is as follows: The source
+node $S$ calculates the stretch change of its $TTL-1$ neighbourhood and that of
+the $TTL-1$ neighbourhood of the first of the previously selected peers. These
+calculations are made as if the jump has been made. If the stretch reduction is
+over a predefined threshold the jump is performed, otherwise the second selected
 candidate is picked and the same computations are performed. If again, the
@@
 is performed as a combination of \emph{leave} and \emph{join} operations, in the
-CAN context.
+CAN context. Figure~\ref{figure:sat_match} demonstrates the selective jump on an
+example topology.
+
+%
+%\paragraph{}
+%Moreover the algorithm takes several issues into consiteration in order to
+%  furtherly improve the resulted overlay. For example when multiple nodes try
+% to jump simultaneously into a regeon, then the logical link brakes from one
+% attempt may result in inacurate computation of the gain factor, for an other.
+% This situation is identified as \emph{contention} and the nodes use an
+% exponatial back-off algorithm to avoid it. An other problem is the uneseccary
+% traffic incuerred by the probing phase in a region that after several jumps
+% has settled to a stable state. In these cases it is more likely for jump
+% attempts to be proven worthless. The algorithm doubles the probing period of
+% such nodes, every time a jump is not taken.
+%
+%\paragraph{}
+%The authors claim that this continuously adaptive mechanism achieves global
+%  topology matching optimization in a sufficiently large scope. This also
+% secures the fast adaptation to frequent network changes. It also considered
+% lightweight and can easily be embedded into current p2p systems, as well
+% as effectively combined with other techniques, such as landmark binning.
+
+%
+% TODO: SOME DISCUSSION
+%
+% It is reported that,
+% due to the selective jumps, SAT-Match achieves $40\%$ reduction in
+% link stretch, and when used with the landmark binning (see
+% Sec. \ref{sec:landmark_binning}), the reduction rate increases up to $60\%$.
+% For dynamic environments, with frequent node arrival and leave, SAT-Match
+% scales much better than Mithos, due to its self adaptation mechanism and
+% selective jumps.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\paragraph*{ \bf Delay Aware P2P System}
+A new \emph{Delay Aware P2P System (DAPS)} is introduced in
+\cite{zhang_daps_2005}. Its main goal is to reduce the time $L$ of a look-up
+request by dividing the routing tables of peers into several sectors in
+increasing delay. The source node emitting the query message defines a delay
+boundary or the pruning factor $L_t$ in the paper's context. Request messages
+will be forwarded only to nodes whose delay less than or equal to $L_t$. With
+the clustered routing tables and the loose organisation the overlay network of
+\emph{DAPS} is between structured and unstructured.
 
@@
 \hline
-\textbf{Dynamo} &  &  &  \\
+\textbf{DynaMO} &  &  &  \\
 

### 2014-01-21 unstructured.tex
2012.06.11 (Vassilis) Content revisited. Duplicate text has been removed.

@@
 
-%{\sethlcolor{yellow}\hl{
-%HA: TALK ABOUT GNUTELLA HERE!!!!}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% TODO: HOW CAN UNSTRUCTURED SCHEMES BE REFINED
+%
+%For this reasons, efforts have been placed for optimizing the efficiency of
+%decentralized unstructured peer-to-peer networks. Research mainly focuses on
+%\begin{inparaenum}[\itshape i\upshape)]
+%  \item reducing unnecessary, redundant communication traffic, and
+%  \item exploiting physical locality to reduce communication response.
+%\end{inparaenum}
+%The goal can be achieved at, both, the application-level network as well as the
+%underlying physical one. In the first case by refining the message relay
+%techniques, while in the second one, by adaptively reconstructing the
+%application network to map as well as possible to the the physical network.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 }\qquad\qquad
-\subfigure[local indices with radius size equal to $2$.] {
+\subfigure[Local indices with radius size equal to $2$.] {
   \includegraphics[scale=0.4]{img/algorithms/local_indices}
@@
 }
-\caption{}
+\caption{Improving search in P2P networks}
 \label{figure:dbfs}
@@
 adaptation algorithm, Gia ensures that high capacity nodes have high degrees and
-low capacity nodes are within short proximity to high capacity ones. Although
-the topology adaptation algorithm Gia uses improves the scalability of the
-network, it does not help much in solving the topology mismatch problem, since
-it does not consider the underlying physical topology.
+low capacity nodes are within short proximity to high capacity ones.
+% TODO: THIS FOLLOWING REMARK SOUNDS LIKE THE ALGORITHM SHOULDN'T BE IN THIS
+%       SURVEY AFTER ALL!!
+Although the topology adaptation algorithm Gia uses improves the scalability of
+the network, it does not help much in solving the topology mismatch problem,
+since it does not consider the underlying physical topology.
+
+%
+%More specifically, there are four key components in the design of
+%\emph{Gia} which are summarized bellow:
+%\begin{enumerate}
+%  \item A \emph{dynamic topology adaptation} protocol that puts
+%participating nodes within short reach of high capacity nodes so that these
+%\emph{high-degree} nodes, which due to their high connectivity receive most of
+%the queries, actually have the capacity to handle them.
+%  \item An \emph{active flow control} scheme is used to avoid overloaded
+% hot-spots. Heterogeneity is detected and flow control tokens are given to
+%nodes based on the available capacity.
+%  \item Every node maintains pointers to to the content that is offered by
+% their immediate neighbours, creating a \emph{one-hop replication} of pointers,
+%scheme.
+%  \item A \emph{search protocol} based on random walks, that is biased towards
+% directing queries to high-capacity nodes that are typically best able to
+%answer these queries.
+%\end{enumerate}
+%
 
@@
 \paragraph*{\bf Distributed Cycle Minimization Protocol}
-We have already mentioned Gnutella's inefficient broadcasting mechanism in
-flooding messages across the network with no actual targeting. This leads, to
-nodes that are far from the path towards the query request gratification, to
-spend their valuable resources to receive, queue and process messages that they
-shouldn't have received in the first place and eventually simply forward them
-blindly to continue this vicious cycle. Except from that, though, an other case
-of wasting network resources is when a node that is on the right path for the
-query, receives the same message multiple times from different neighbours. The
-reason for these duplicate messages is cycles in the forwarding paths. Actually,
-this duplicate packet problem seems to hurt more the active nodes; those with
-higher capacities and bandwidth that contribute the most to the network.
-Distributed Cycle Minimization Protocol (DCMP) \cite{zhu_dcmp_2008} introduces a
-method to
-remove the cycles and eliminate the duplicate packets, without sacrificing
-overlay connectivity. In DCMP, once a cycle is detected, the most powerful node
-in the cycle is elected as the \emph{Gate Peer} and the cycle is then destroyed
-by cutting that special link which will result in the minimisation of the
-distance to the Gate Peer for all the nodes that participated in that cycle. The
-process is managed by using two special message types namely \emph{Information
+We have already discussed that blind flooding mechanism broadcasts messages with
+no real targeting, many times to a direction far from the path that can
+ultimatelly fullfil the query request. Another problem with topology unaware
+systems is the duplication of messages due to cycles, even along the correct
+forwarding path. The authors of \cite{zhu_dcmp_2008} fucused on that exact
+deficiency of overlay networks and introduced Distributed Cycle Minimization
+Protocol (DCMP), a dynamic, fully distributed method to remove cycles without
+sacrificing overlay connectivity, resilience and other important properties of
+unstructured architectures by avoiding a hierarchical organisation of peers. In
+DCMP, once a cycle is detected, the most powerful node in that cycle is
+elected as the \emph{Gate Peer} and the cycle is then broken at a strategic
+place so that it will result in the minimisation of the distance between the
+Gate Peer and all other nodes that participated in that cycle. The process
+is managed by using two special message types namely \emph{Information
 Collection (IC)} and \emph{Cut Message (CM)}. One disadvantage of DCMP is that
@@
 does not actually solve the topology mismatch problem.
+
+%
+%\subsection{Distributed Cycle Minimization Protocol}
+%
+%\paragraph{}
+%The first step after detecting a duplicate message by some peer is to gather
+% information from all peers in the cycle using a new type of control message
+%called \emph{Information Collecting Message} or \emph{ICM}. ICM contains:
+%\begin{inparaenum}[\itshape i\upshape)]
+%  \item a \emph{GUID}\footnote{Globally Unique IDentifier assigned to every
+% query message generated by any node.} field same as the one of the duplicate
+%message,
+%  \item \emph{DetectionID} field which represents the direction of the
+% connection where the duplicate was identified\footnote{This ensures the
+%uniqueness of the ICM messages because as it travels through many cyclic paths,
+%multiple peers will detect the duplicates and initiate an ICM message.}, and
+%  \item \emph{Node Information Vector (NIV)} which contains information
+% (bandwidth, CPU power, etc) about peers that propagated the ICM.
+%\end{inparaenum}
+%
+%\paragraph{}
+%Suppose $A$ detected the duplicate as depicted in Figure~\ref{figure:dcmp}. It
+% then emits an ICM to $B$ and $F$, that initially contains information only
+%about itself. Each peer that receives the ICM, appends its information and
+%propagates it along the reverse path of the original message. Since two copies
+%of ICM are sent, at some point, a peer, say peer $D$, will receive a duplicate
+%ICM. Using the information in the NIVs of the ICMs, $D$, decides to cut (for
+%example) the EF connection. To inform the other peers about its decision, it
+%emits a \emph{Cut Message (CM)} which contains the GUID and DetectionID of the
+%corresponding ICM and an additional field that identifies the connection to be
+%cut. $D$, then, forwards the CM in the reverse directions from where the ICM
+%arrived. Similarly CMs received by any peer are propagated toward the reverse
+%path of the corresponding ICM. Eventually, either peer $E$ or peer $F$ will
+%receive the CM and cut the connection, thus eliminating the cycle.
+%
+%Receiving a duplicate ICM denotes the existence of a cycle. The opposite is not
+% true though. For example, if the cycle contains $2 \times TTL$ edges, it will
+%not be detected, because the ICM messages will be discarded before they locate
+%it. There is a tradeoff between preserving the connectivity of the network and
+%minimizing the duplicates that makes such a possibility to be safely ignored.
+%
+%\begin{figure}
+%\centering
+%  \includegraphics[scale=0.4]{img/dcmp.jpeg}
+%\caption{Cycle elimination methods in DCMP}
+%\label{figure:dcmp}
+%\end{figure}
+%
+%\paragraph{}
+
+%
+% TODO: SOME DISCUSSION
+%
+%Actually, the duplicate packet problem seems to hurt more the active nodes;
+%those with higher capacities and bandwidth that contribute the most to the
+%network.
+%Experiments in \cite{zhu_dcmp_2008} showed that DCMP incurs a lower delay,
+% returns more results and decreases the number of duplicate messages by 22\%,
+%compared to LTM. Additionally DCMP has one to two orders of magnitude less
+%overhead, because it adopts a more ``LAZY'' approach than broadcasting control
+%messages periodically like LTM does.
+%
 
@@
 
+% These approaches include creating spanning trees using
+%connection graphs, creating cluster of physically close nodes, or using latency
+%information to detect proximity. The brief description of each category is
+%presented below:
+%
+%  \begin{itemize}
+%    \item \emph{Spanning tree based}. These approaches construct of a rich
+%    graph  based on the network connections and build minimum spanning tree
+%    on the graphs, causing large traffic overhead to the system
+%    \cite{chu_esm_2000,chu_esm_2002}.
+%
+%    \item \emph{Cluster based}. These approaches select to link physically
+%    closer nodes with each other, therefore shrink the search scope
+%    significantly while mapping accuracy is not always guaranteed.
+%
+%    \item \emph{Minimum latency first}. Use of latency as a metric to calculate
+%    distance among peers. They require global latency information
+%    ``landmarks''\footnote{Measuring latency between peers and stable Internet
+%    servers.}.
+%
+%  \end{itemize}
+
 The overlay optimization based protocols modify the topology of the P2P network
@@
 
-Narada \cite{chu_esm_2000,chu_esm_2001,chu_esm_2002} is a generic protocol for
-creating self-adapting overlay networks, that can achieve application-layer
-multicast communication without requiring IP multicast infrastructure at the
-network layer. IP multicast is the term reffering to the method of sending IP
-datagrams to a group of receivers in a single transmission. Even though the
-method is available some years now, IP multicast has not taken off as
-anticipated since many believe that it violates the stateless design of the
-current Internet. For this reason it is widely deployed in contained
-environments like enterprises (i.e. IPTV applications - distance learning, video
-conferencing), commercial stock exchanges, and multimedia content delivery
-networks. Narada tries to break these constraints. Although it was not
-originally designed as a P2P file sharing protocol, it has become one of the
-pioneering works in the overlay networks area. The authors claim that, although
-Narada is not as efficient as IP multicast, still it provides reasonable
-performance, without the additional cost of implementing multicast services on
-the current Internet infrastructure. Additionally, authors realised the
-inefficiency caused by the topology mismatch problem and tried to encounter it
-by building a richer connected graph, called a mesh, and building per source
-minimum spanning trees \footnote{ENarada \cite{li_enarada_2008} used Gossip
-protocol for the construction} as shown in Figure~\ref{figure:narada}. They also
-keep the graph and the trees dynamically updated while nodes continue to join
-and leave the network. Moreover, the protocol tries to ease the physical link
-stress, the overall resource usage and the relative delay among end systems.
-Unfortunately, the most imporpant limitation of Narada is that although it works
-reasonably well for small groups, it does not scale well for larger networks,
-therefore it is not suitable for, potentially very large, P2P file sharing
-application networks.
+\emph{Narada} \cite{chu_esm_2000,chu_esm_2001,chu_esm_2002} is a generic
+protocol for creating self-adapting overlay networks, that can achieve
+application-layer multicast communication without requiring IP multicast
+infrastructure at the network layer\footnote{
+  IP multicast is the term reffering to the method of sending IP datagrams to a
+  group of receivers in a single transmission. Even though the method is
+  available some years now, IP multicast has not taken off as anticipated since
+  many believe that it violates the stateless design of the current Internet.
+  For this reason it is widely deployed in contained environments like
+  enterprises (i.e. IPTV applications - distance learning, video conferencing),
+  commercial stock exchanges, and multimedia content delivery networks.
+}. Although it was not originally designed as a P2P system nor can be considered
+scalable in the contemporary sense of the term, Narada is considered pioneering
+(maybe along with \emph{Scattercast} \cite{chawathe_scattercast_2000}) in that
+it was the first to consider the feasibility of overlay-based, application
+layer, services over the Internet that can take into account bandwidth and
+latency properties of the physical underlying infrastructure. Additionally,
+authors realised the inefficiency caused by the topology mismatch problem and
+tried to encounter it by building a richer connected graph, called a
+\emph{mesh}, and building per source minimum spanning trees\footnote{
+  ENarada \cite{li_enarada_2008} used Gossip protocol for the construction
+}
+as shown in Figure~\ref{figure:narada}. They also keep the graph and the trees
+dynamically updated while nodes continue to join and leave the network.
+Moreover, the protocol tries to ease the physical link stress, the overall
+resource usage and the relative delay among end systems. Unfortunately, the most
+imporpant limitation of Narada is that although it works reasonably well for
+small groups, it does not scale well for larger networks, therefore it is not
+suitable for, potentially very large, P2P file sharing application networks.
 
@@
 Establishment}
-\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{liu_auto_2003} is one
-of the first attempts, along with Narada, from the P2P research community to
-address the topology mismatch problem. AOTO is a distributed algorithm that uses
-\emph{Selective Flooding} and \emph{Active Topology} to optimize the overlay
-network topology. The \emph{Selective Flooding} algorithm creates a minimum
-spanning tree over the overlay and uses this tree instead of flooding the whole
-network, without shrinking the search scope. In \emph{Active Topology}, the
-physical locations of nodes are estimated based on the network delay among them
-and physically close nodes are connected as neighbours to revise the overlay
-topology as close as possible to the physical network topology. Overall, AOTO
-reports a $55\%$ performance improvement in terms of the traffic load. In
+\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{liu_aoto_2003} is one
+of the first attempts, along with Narada, to address the topology mismatch
+problem. AOTO is a distributed algorithm that works in two steps, namelly
+\emph{Selective Flooding} and \emph{Active Topology} to optimize the usage of
+the underlying physical resourses. In \emph{Selective Flooding} a minimum
+spanning tree is built among each peer and its immediate neighbours so that
+queries are not flooding the whole network and at the same time without
+shrinking the search scope. During \emph{Active Topology}, the overlay
+links are revised so that they depict the physical network topology as close as
+possible. This is done independently, by each peer by replacing non-flooding
+neighbours, with closer nodes. Picking a replacement out of these non-fooding
+neighbours follows a random policy (called \emph{Randomized AT} algorithm in the
+paper's context). For these steps a peer needs to keep track of communication
+costs to all its neighbours (e.g. the network delay) as well as between any
+pair of neighbours (meaning also that additional, cost probing, message types
+need to be added to the Gnutella protocol). Whenever a new neighbour cost table
+is received or there is a change of neighbours, the source peer has to
+re-calculate the multicast tree and apply the randomized AT algorithm.
+
+%\paragraph*{Selective Flooding (SF)}
+%
+% TODO: I DON'T REMEMBER WHY THE FOLLOWING IS MENTIONING LTM WHICH IS ANOTHER
+%       ALGORITM. MAYBE THIS CAN BE USED AS A PART OF DISCUSSION AND/OR
+%       COMPARISON OF AOTO AND LTM
+%LTM's SF effectiveness has be proven to be detached from the different physical
+% or overlay topologies. On the other hand, SF is more effective with large
+%number of logical neighbours. It can reach an average optimization rate of 87.4
+%percent on a logical topology with an average of 30 logical neighbours.
+%
+%\paragraph*{Active Topology (AT)}
+%
+%Different numbers of average logical neighbours has little to do with the
+% effectiveness of AT. If the source has $n$ non-flooding peers, there are $n$
+%potential neighbour replacements. The overhead to exhaust all $n$ possible
+%replacements can be high, so in practice, after each replacement the source
+%peer can decide whether it needs to find another candidate peer. This is done
+%by computing the cost improvement ratio greater than some predefined
+%termination threshold. The larger the threshold, the slower, in the number of
+%optimization steps, the reduction of the normalized average distance. As a
+%whole the average response time is significantly reduced when more optimization
+%steps taken.
+
+%
+% TODO: REMAKE AAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAA!!!!!!!!!!!!!!!!!!
+%
+ In
 \emph{Adaptive Connection Establishment (ACE)} \cite{liu_ace_2004}, the authors
@@
 method to optimize the overlay structure of the P2P network based on the
-physical topology. To achieve this, the algorithm issues a special message
-called \textit{TTL2-detector} to detect the latency of two-hop neighbours
-($N^2$) around each node. The latency information gathered is later used to
-evolve the overlay network into a more efficient one, without reducing the
-search scope. Each node examines the latency information with the neighbours and
-low productive connections are dropped and replaced by more efficient ones, thus
-reducing the latency on the overall network, and eliminating some potential
-inefficiencies, such as unnecessary crossing of messages over autonomous system
-(AS) boundaries, etc. Although LTM improves the overall efficiency of the P2P
-network, it does not actually use real physical topology information, just
-latency metrics, therefore it does not offer a guaranteed safety net for the
-topology mismatch problem.
+physical topology. To achieve this, peers issue special messages called
+\textit{TTL-detector}s with TTL values of 0 and 1 so that the receiving peers
+discover one- and two-hop neighbour sets ($N$ and $N^2$ respectively) around
+them and calculating communication costs. Calculation is achieved through
+timestamp checking, so clocks must be synchronised. The latency information
+gathered is later used to evolve the overlay network into a more efficient one,
+without reducing the search scope. Each node examines the latency information
+among his direct neighbours and those with larger latency are put in to a
+will-cut list where they remain for a certain period of time until when they are
+ultimatelly cut and recorded to the peer's cut list. Thus, low productive
+connections are dropped and replaced by more efficient ones, reducing the
+latency on the overall network. Although LTM improves the overall efficiency of
+the P2P network, it does not actually use real physical topology information,
+just latency metrics, therefore it does not offer a guaranteed safety net for
+the topology mismatch problem.
+
+%
+% TODO: SOME DISCUSSION
+%
+%\paragraph*{Low productive connection cutting}
+%There are three cases for any peer $P$ who receives $d(i, S, v)$ multiple
+% times:
+%\begin{inparaenum}[\itshape i\upshape)]
+%  \item $P$ receives both $d(i, S, 1)$ and $d(i, S, 0)$
+%  \item $P$ receives multiple $d(i, S, 0)$s from different paths, and P
+% randomly chooses to process one
+%  \item $P$ receives one $d(i, S, 1)$ and multiple $d(i, S, 0)$s, and $P$
+% processes $d(i, S, 1)$ and one randomly selected $d(i, S, 0)$
+%\end{inparaenum}
+%If the link with the largest cost is found and is a direct neighbour then the
+% connection is put in a will-cut list and stays there for a certain period of
+%time. If it is not, then it is handled by other peers. After that period,
+%connections are cut and recorded to $P$'s cut-list.
+%
+%\paragraph*{Source probing}
+%For a peer $P\in(N^2(S) - N(S))$ who receives only one $d(i, S, 0)$, the cost
+% of $PS$ is obtained (with list look-up or probing). Then $P$ compares it with
+%the cost from each hop and if $PS$ has the largest cost, $P$ will not keep this
+%connection, while otherwise the connection will be created.
+%
+%\paragraph{}
+%Supposing $n$ is the number of peers, $c_n$ is the average number of neighbours
+% and $c_e$ is the average cost of logical links, then in the flooding-based
+%search the traffic incurred by one query from an arbitrary peer in a
+%peer-to-peer network is $O(n)$. As observed in the Gnutella network
+%\cite{sripanidkulchai_gnutella_2001}, each peer issues $0.3$ queries per minute
+%in average, thus the per minute traffic incurred by the network with $n$ peers
+%is $O(n^2)$. Because each $d(i, S, v)$ has a TTL of $2$ in each source peer,
+%the traffic for one time LTM optimization in all peers is at most $2nc_n^2c_e$.
+%If each peer uses LTM $k$ times per minute, the total traffic incurred is
+%$2knc_n^2c_e$. Simulation shows the best value for $k$ is $2$ or $3$. So, the
+%traffic overhead caused by LTM to the network is $O(n)$.
+%
+%TTLj-detectors, with $j > 2$, would detect and break cycles with more than 4
+% links. LTM though, does not use such detectors because detector-flood traffic
+%would increase significantly, and cut links between two end-peers, could cause
+%queries initiated by them to traverse a path much more expensive than the cost
+%on the the cut link.
+%
+%\paragraph{}
+%LTM disadvantages are
+%\begin{inparaenum}[\itshape i\upshape)]
+%  \item disagreement of measured delay due to unsynchronized clocks causes
+% problems when deciding the cut positions, which can influence the network
+%connectivity, and
+%  \item the network delay metric mainly focuses on disabling the connections
+% between peers physically far away without considering the shortcuts created by
+%powerful peers.
+%\end{inparaenum}
+%
 
@@
 \paragraph*{\bf Scalable Bipartite Overlay}
-\emph{Scalable Bipartite Overlay
-(SBO)} \cite{liu_bipartite_IPDPS,liu_bipartite_2007} reduces the overhead of
+\emph{Scalable Bipartite Overlay (SBO)}
+\cite{liu_bipartite_IPDPS,liu_bipartite_2007} reduces the overhead of
 creating and maintaining a minimum spanning tree cost by randomly dividing the
 nodes into two groups, namely red and white and assigning different tasks to
-different groups. The white peers measure distances to neighbours by using the
-network delay as the metric and reports the results to red peers. The red peers,
-equipped with the information of all two-hop neighbours ($N^2$), creates a
-minimum spanning tree of these neighbours and assigns efficient forwarding
-paths. The white peers can further optimize their positions within the overlay
-if this is required.
+different groups. When a peer joins the network, is randomly assigned with an
+initial colour, say red or white (separating all peers into two groups). Then
+the peer that plays the role of the bbootstrap host provides the joining peer
+with a list of active peers along with their colour information. The later uses
+this list to establish connections to differently coloured peers. This way, all
+peers form a bipartite overlay. The white peers measure distances to
+neighbours (red only) by using the network delay as a metric and report their
+corresponding red neighbours. The red peers, equipped with the information of
+all two-hop neighbours ($N^2$), creates minimum spanning tree of these
+neighbours and assigns efficient forwarding paths. Having a minimum spanning
+tree with two hops, a red peer is able to send its queries within that range.
+Some white peers, though, have become non-forwarding neighbours. In this phase,
+such a (white) neighbour will try to find another red peer being two hops away
+from its current red neighbour to replace the later as its new neighbour. The
+white peers can further optimize their positions within the overlay if this is
+required.
+
+%
+% TODO: SOME DISCUSSION
+%
+%In a static enviromnent LTM may reduce traffic cost by around 80 to 85 percent
+% while SBO reduces traffic cost between 85 and 90 percent. However, LTM  is
+%proved to converge in around 2-3 steps while SBO needs 4-5 steps. Moreover LTM
+%reduces response time by more than 60 percent in 3 steps while SBO needs 8. In
+%a dynamic environment (10 minute average peer lifetime, 0.3 queries/sec by each
+%peer) SBO and LTM reduce the average traffic cost per query (including the
+%overhead due to the optimization steps) by 85 and 80 percent, respectively.
+%Moreover LTM reduces the response time per query to 30 percent while SBO to 35
+%percent.
+%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\paragraph*{\bf Adaptive Connection Establishment (ACE)}
+\emph{Adaptive Connection Establishment (ACE)} \cite{liu_ace_2004} builds an
+overlay multicast tree among each source node and the peers within a certain
+diameter from the source peer and optimizes the neighbour connections that are
+not in that tree. To achieve that it calculate cost between nodes using network
+delay as a metric. Each peer probes the costs with its immediate logical
+neighbours and forms a \emph{neighbour cost table (NCT)} using a special
+routing message type. Two neighbouring peers exchange their NCTs in order for
+every peer to obtain the cost between any pair of its local neighbours forming a
+small overlay topology. Moreover, based on obtained NCTs a minimum spanning tree
+among each peer and its immediate neighbours is built. Ultimately, physically
+far away neighbours are replaced by physically close ones. Specifically, a peer
+$S$ probes the distance between one of its non-flooding neighbour's neighbour,
+say $G$ and $H$ respectively. Assuming that the link to the neighbour's neigbour
+($SH$) is smaller that that of the neighbour ($SG$), the later is cut and a new
+connection is established. In the oposite situation where $SG$ is smaller than
+$SH$ then if $SH$ is smaller that $GH$ then $H$ is preserved as $S$'s new
+neighbour. In the case where $SH$ is larger than $SG$ and $GH$, no connection
+will be established and $S$ will continue probing another neighbour's neighbour.
+The above optimisation is conducted within $1$-neighbour closure (among its
+source peer and all its direct neighbours) but the scope can be extended. The
+larger the scope, the better topology matching improvement but also the greater
+the computational overhead.
+
+%
+% TODO: SOME DISCUSSION
+%
+%Simulations in \cite{liu_acesims_2004} show that the average scope of each
+% query to cover the same scope of nodes is reduced by about 65 percent without
+%losing any autonomy feature, while the average response time can be reduced by
+%35 percent. Larger diameter topologies lead to better topology optimization
+%rate but also to higher communication and computation overhead. It was also
+%found that it is more effective in higher connectivity dense topologies.
+%Compared to LTM, it comes short of convergence speed. In
+%\cite{ni_mismatch_2004} shows reduction of both total traffic (90 percent) and
+%response time (80 percent) to message queries without shrinking the search
+%scope. SBO, on the other hand, achieves approximately 85 percent reduction on
+%traffic cost and about 60 percent reduction on query response time. Last but
+%not least, it is concluded that work must be done on incorporating a more
+%sophisticated selection policy for candidate non-flooding peers.
+%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\paragraph*{\bf Hops Adaptive Neighbour Discovery (HAND)}
+\cite{chen_hand_2006} proposes the \emph{HAND} algorithm which uses a
+fully distributed triple hop adjustment strategy to address the topology
+mismatch problem. The approach's ultimate goal is to shape the current overlay
+graph $G$ into the \emph{Logical Communication Network (LCN)} $G^{*}$ which is
+how the optimal overlay is refered in the paper's context. A formal
+definition of achieving the above match is only if all peer hop sequences $(v_1,
+v_2, \ldots, v_k)$ in $G$ exist in $G^{*}$ and in the same order\footnote{In
+practice triple sequences $(v_1, v_2, v_3)$ are used.}. The mismatching
+detection is done in the following way. Suppose we want to verify peer a
+sequence, say $v_2-v_1-v_3$. A pair of probing messages are sent from $v_1$ to
+$v_2$ and $v_3$. Suppose delays of $(v_1,v_2)$ and $(v_1,v_3)$ are are denoted
+as $x$ and $z$, respectively. When the probing message arrives to $v_2$ it
+forwards it directly to $v_3$. Similarly, when the probing message arrives to
+$v_3$, it forwards it directly to $v_2$. These last steps are performed in order
+to obtain delays of $(v_2,v_3)$ and $(v_3,v_2)$ physical paths, respectively,
+denoted by $y$. If $y=z-x\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched
+and should be adjusted to $v_1-v_2-v_3$ by deleting edge $(v_1,v_3)$ and adding
+a new $(v_2,v_3)$. If $y=x-z\pm\varepsilon$, sequence $v_2-v_1-v_3$ is
+mismatched and should be adjusted to $v_1-v_3-v_2$ by deleting edge $(v_1,v_2)$
+and adding a new $(v_3,v_2)$, where $\varepsilon$ is a small positive real
+number denoting additional delays caused by possible forwarding and jitter
+delays. The advantages of the HAND algorithm compared to other approaches are
+that:
+\begin{inparaenum}[\itshape i\upshape)]
+  \item it does not need any clock synchronization,
+  \item it is a fully distributed algorithm making it robust and reliable in
+        decentralized systems,
+  \item the traffic overhead of the triple hop adjustment is very low,
+  \item it is applicable to dynamic peer-to-peer environments, and
+  \item maintains lower query response time.
+\end{inparaenum}
+
+%
+% TODO: SOME DISCUSSION
+%
+%Measurements conducted for evaluation perposes showed that in a static
+% environment the algorithm can effectively decrease traffic cost by about 77
+%percent and shorten the query response time by about 49 percent in less than
+%two minutes. In a dynamic environment it shows similar behaviour and with the
+%size of the overlay network having little impact on the effectiveness of the
+%algorithm. Compared to LTM both algorithms have almost the same traffic
+%reduction rate, however on the response time reduction rate HAND has a higher
+%one by about 4 percent. The traffic overhead of HAND is much less than that of
+%LTM by an average of 55 percent.
 
@@
 \paragraph*{\bf TopBT}
-\cite{rtlcgz_topbt_2010} proposed an approach for enhancing proximity awareness
+\cite{rtl_topbt_2010} proposed an approach for enhancing proximity awareness
 in the BitTorrent protocol without any need of any infrastructure installed. It
@@
 
-%\paragraph*{\textbf{Two-Hop-Away Neighbour Comparison and Selection\\}}
-%
-%\paragraph{}
-%Work in \cite{liu_thancs_2005,liu_thancs_2008} proposes a distributed heuristic
-%called \emph{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} that
-%\begin{inparaenum}[\itshape i\upshape)]
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\paragraph*{\bf Two-Hop-Away Neighbour Comparison and Selection}
+Work in \cite{liu_thancs_2005,liu_thancs_2008} proposes a distributed heuristic
+called \emph{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} that
+tries to minimise the overlay hop costs. THANCS is considered a \emph{local
+search method}, in the sense that it targets in finding a locally optimum
+solution, by exploiting knowledge within a 2-hop radius. The algorithm consists
+of two main components: \emph{piggybacking neighbour distance on queries} and
+\emph{neighbour comparison and selection} which are furtherly discussed bellow.
+The first component dictates peers to probe distances (using network delay
+distance measuring) with its immediate naighbours and stores information
+locally. This is done by introducing a special query message type, the
+\emph{Piggy Message (PM)} which includes information about the neigbour
+identification and measured distance. A peer $p$ constructs a PM for its
+neighbour $n$, which contains $n$'s IP address and the distance between them.
+When $p$ receives a query from $n$, this PM will be piggybacked to the query
+message that will be then forwarded to all other neighbours of peer $p$. Upon
+receiving such a query message, each of the other neighbours will detach the PM
+information (it will not be further forwarded), record the $pq$ distance and
+process the query as always. The paper proposes selection of which incoming
+queries should piggyback a PM by either \emph{pure propability-based (PPB)} or
+\emph{new neighbour triggered (NNT)} policies. The second component, namely
+neighbour comparison and selection, dictates peer $p$ to probe the distance to
+all his known, not yet probed, two-hop neighbours ($ N^2(p)$). The distance of
+$pn$ is known to $p$. Upon receiving a PM from node $n$ with the distance of
+$nq$, where $q$ is a direct neigbour of $n$, $p$ does follows one of the
+following two paths. First, if $q$ is a direct neighbour of $p$, then the later
+will check the cost of the involved links. If the most cost-intensive link is
+one of $pq$ or $pn$, then it is put into a \emph{will-cut} list. If it is $nq$
+then nothing is done (since either $n$ or $q$ will have the chance to handle it
+themselves). On the other hand, if $q$ is within an two-hop radius from $p$,
+the later probes the former (if it hasn't got the distance information yet) and
+again compares distances $pq$, $pn$ and $nq$. If $pq$ is the most
+cost-intensive, $p$ will not establish the connection. If it is $pn$, $p$ will
+establish connection $pq$ and put $pn$ in the will-cut list. If it is $nq$, $p$
+will keep the connection with both $n$ and $q$, expecting that $n$ or $q$ will
+eventually disconnect link $nq$, later.
+
+%
+% TODO: SOME DISCUSSION
+%
+% \begin{inparaenum}[\itshape i\upshape)]
 %   \item is completely distributed and needs no global knowledge,
@@
 %   \item its convergent speed of the algorithm is fast enough (faster than
-%minimum spanning tree approaches) so that is effective to dynamic environments,
-%and
+% minimum spanning tree approaches) so that is effective to dynamic
+% environments, and
 %   \item does not shrink the search scope.
-%\end{inparaenum}
-%
-%\paragraph{}
-%THANCS is considered a \emph{local search method}, in the sense that it targets in finding a locally optimum solution, by exploiting knowledge within a 2-hop radius.
-The algorithm consists of two main components: \emph{piggybacking neighbour distance on queries} and \emph{neighbour comparison and selection} which are furtherly discussed bellow.
-%
-%\paragraph{Piggybacking neighbour distance on queries}
-%Using network delay as a metric for measuring the distance, each peer probes distances with its immediate naighbours and stores information locally.
-For this reason a special query message type, \emph{Piggy Message (PM)}, is introduced.
-It is 6 bytes long and includes two fields: Neighbour IP Address and Neighbour Distance.
-A peer $P$ constructs a PM for its neighbour $Q$, which contains $Q$'s IP address and $Q$'s distance from $P$.
-When $P$ receives a query from $Q$, this PM will be piggybacked by the query that goes to all other neighbours of peer $P$.
-Upon receiving such a query message, each of the other neighbours will detach the PM, record the $PQ$ distance and process the query.
-This PM will not be further forwarded.
-In selecting which incoming queries  should piggyback a PM the paper proposes the \emph{pure propability-based (PPB)} and the \emph{new neighbour triggered (NNT)} policies.
-%
-%\paragraph{Neighbour comparison and selection}
-%Figure~\ref{figure:thancs} illustrates this component of the THANCS algorithm.
-A peer $S$ probes the distance to all known unprobed $N^2(S)$\footnote{$N^2(S)$ denotes the set of peers being two hops away from $S$, while $N(S)$ denotes the set of direct logical neighbours of $S$.}.
-The distance of $SP$ is known to $S$.
-Upon receiving a PM from node $P$ with the distance of $PQ$, $S$ follows one of the following:
-%\begin{itemize}
-%  \item $Q \in \left( N(S) \cap N^2(S) \right)$, i.e.
-$Q$ is direct neighbour of $S$.
-In this case $S$ will compare cost of $SQ$, $SP$ and $PQ$.
-If the most costly connection is one of $SQ$ or $SP$ the corresponding link will be put into a \emph{will-cut list}\footnote{Links are not immediately disconnected when put in the will-cut list.
-There are useless for forwarding but are kept active, for some time, in order to serve query responses that are traveling to the source peer along the inverse search path.}.
-If the most costly connection is $PQ$ then $S$ will do nothing, as the fully distributed nature of the algorithm will give a chance to either $P$ or $Q$ to cut the connection.
-%  \item $Q \in \left( N^2(S) - N(S) \right)$, i.e.
-$Q$ is a two-hop-away neighbour of $S$.
-If $S$ hasn't probed $Q$ before\footnote{A distance cache is maintained and looked up in such a case.}, it probes and stores the result in the \emph{distance cache}.
-Having the distance of $SQ$, $S$ compares costs of $SQ$, $SP$ and $PQ$.
-If $SQ$ is the most costly, $S$ will not establish the connection.
-If $SP$ is the most costly, $S$ will establish connection $SQ$ and put $SP$ in the will-cut list.
-If $PQ$ is the longest, $S$ will keep  the connection with both $P$ and $Q$, expecting that $P$ or $Q$ will eventually disconnect link $PQ$, later.
-%\end{itemize}
-%
-%\begin{figure}
-%\centering
-%  \includegraphics[scale=0.4]{img/thancs.jpeg}
-%\caption{Probing two-hop-away neighbours}
-%\label{figure:thancs}
-%\end{figure}
-%
-%\paragraph{}
-%In a static environment THANCS has been proven to be effective; optimizing 45 percent out of the 60 percent of mismatched paths, constructing a nearly optimal overlay.
-This leads to a 60 percent reduction in traffic cost as well as a 40 percent decrease in query response time.
-In dynamic environments (Gnutella 0.6/Limewire super-peer-like and Ion flat-like), THANCS saves up to 70 percent of the traffic cost in the super-peer topology and 55 percent for the flat one.
-Average response time is also decreased by 60 and 45 percent, respectively.
-Generally, THANCS has similar performance to LTM, without needing synchronization.
-SBO, incurring half the  overhead of AOTO, reduces the traffic cost the most, while THANCS has lower response time and converges faster than SBO.
-THANCS is, thus, more suitable for a more dynamic environment.
-In addition, THANCS is easy to implement and its operation overhead is trivial, compared with the other three approaches.
-This design, however, has the limitation of not being easily extend to also support non-flooding-based systems.
+% \end{inparaenum}
+%
+%In a static environment THANCS has been proven to be effective; optimizing 45
+% percent out of the 60 percent of mismatched paths, constructing a nearly
+%optimal overlay. This leads to a 60 percent reduction in traffic cost as well
+%as a 40 percent decrease in query response time. In dynamic environments
+%(Gnutella 0.6/Limewire super-peer-like and Ion flat-like), THANCS saves up to
+%70 percent of the traffic cost in the super-peer topology and 55 percent for
+%the flat one. Average response time is also decreased by 60 and 45 percent,
+%respectively. Generally, THANCS has similar performance to LTM, without needing
+%synchronization. SBO, incurring half the  overhead of AOTO, reduces the traffic
+%cost the most, while THANCS has lower response time and converges faster than
+%SBO. THANCS is, thus, more suitable for a more dynamic environment. In
+%addition, THANCS is easy to implement and its operation overhead is trivial,
+%compared with the other three approaches. This design, however, has the
+%limitation of not being easily extend to also support non-flooding-based
+%systems.
 
@@
 on the exchange of neighbours among peers, driven by the mutual benefit of the
-participants to reduce the network delay among their connections.  In PROP-G
-(Generic), a peer can exchange all of its neighbours with another peer, while in
-PROP-O (Optimized) two peers exchange a selected subset of their neighbours.
-PROP-G is a generic protocol that guarantees the connectivity of the overlay
-graph during exchanges, therefore can be applied to both unstructured and
-structured overlay networks.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf T2MC}
-T2MC \cite{shi_t2mc_2008} uses traceroute logs to detect different autonomous
-systems and cluster close by peers, thus reducing redundant traffic that crosses
-autonomous system boundaries. T2MC uses a customized k-mean classification
-algorithm with $k=2$ to perform the classification, and exploits the stable
-structure of the Internet routers to guide clustering. Even though traceroute
-provides detailed information about the network structure, its excessive use
-creates additional overhead that stresses further the overall network
-infrastructure. This forces network administrators to disable its support from
-the routers they manage, thus jeopardizing the effectiveness of the T2MC
-algorithm.
-
-%\paragraph*{\textbf{T2MC\\}}
-
-%\paragraph{}
-%\emph{T2MC}\cite{shi_t2mc_2008} exploits some properties of the Internet
-%paradigm and clusters nodes belonging to the same ISP without any centralised
-%control or predefined system parameterization. The algorithm considers the
-%dynamic nature of peers and exploits the stable nature of routers in order to
-%build a topology location relationship among end peers.
-%
-%\paragraph{}
-%Some special routers split the physical network into autonomous system domains.
-%Using a Traceroute mechanism T2MC searches for latency leaps among the path to
-%a host in order to form ``near'' and ``remote'' router clusters. This is
-%achieved by a 2-Means Classification, defined by the following steps:
-%\begin{enumerate}
-% \item the peer chooses the minimum and maximum Latency results from the
-%Traceroute for initializing the cntroinds of two sets ``first'' and ``second''.
-% \item The peer calculates, for all hops allong its tracerouted path, the
-%absolute distance to the centroids of both sets and assigns the routers to that
-%centroid with which it has the smaller absolute distance.
-% \item The peer calculates the latncy mean and variance value of two sets
-% \item If the variance is larger than a predefined threshold then the algorithm
-%takes a loop from step 2 picking the two latency mean values as new centroids
-%of sets ``first'' and ``second''.
-%\end{enumerate}
-%Ultimately peer will end up with two sets having the minimum intra-set
-%variance. Finally the peer chooses the router from ``second'' set with the
-%minimum hops attribute and sets it as a threshold. The selected router and all
-%others whose hops attribute is larger than the threshold are classified as
-%``remote'' router cluster. The remaining are classified as ``near''. From the
-%``near'' class, the peer chooses the one with maximum hops attribute as its
-%edge router, and registers it along with the all the ``near'' cluster into the
-%DHT of the p2p overlay. As new peers join the network, those that share the
-%same edge router or any of the members of the ``near'' router clustersm ther
-%would gather to form a ``close'' peer cluster. As Edge routers can provide more
-%valuable information than other members of the ``near'' set, T2MC was designed
-%to prioritize interaction of peers and edge gateways.
-%
-%% TODO: figure t2mc.jpeg (a) stars represent edge routers (b) hop leaps
-%
-%\paragraph{}
-%The use of Traceroute as a tool for implementing the distance measuring
-%infrastructure raise concearns about its efficiency and scalability. Being,
-%primarilly, a network diagnostic utility, it is concearned too heavy weighted
-%and intrucive for use in a larger scale\cite{ratnasamy_binning_2002}.
-%Additionally, disabling ICMP is a common administrative policy for edge sites
-%to enforce security, while dumping BGP routing
-%tables\cite{krishnamurthy_bgpclust_2000} is not directly available to the
-%application layer.
+participants to reduce the network delay among their connections. This
+``collaboration'', is what differentiats this approach from others by letting
+two peers to optimize their neighbourhood environment, than simply letting each
+node to ``selfishly'' choose its own strategy. In PROP-G (Generic), a peer can
+exchange all of its neighbours with another peer, while in PROP-O (Optimized)
+two peers exchange a selected subset of their neighbours. PROP-G is a generic
+protocol that guarantees the connectivity of the overlay graph during exchanges,
+therefore can be applied to both unstructured and structured overlay networks.
 
@@
 \paragraph*{\bf Distributed Domain Name Order}
-\emph{Distributed Domain Name Order (DDNO)} \cite{z-yk_ddno_2005}
-uses the domain names to detect topologically close nodes based on the
-assumption that nodes within the same domain are topologically close to each
-other. Half of the possible connections of a node is used to connect to these
-local peers, and the other half is used to randomly connect to the peers
-anywhere on the network. \textit{DDNO} is a heuristic approach to solve the
-topology mismatch problem, with local connections to improve the efficiency and
-reduce the delay on the network.  The random connections on the other hand help
-ensure the connectivity in the network and avoid partitioning. \textit{DDNO}
-uses \textit{Split-Hash} and \textit{dnMatch} algorithms to detect locality by
-using domain names.
+\emph{Distributed Domain Name Order (DDNO)} \cite{z-yk_ddno_2005} uses the
+domain names to detect topologically close nodes based on the assumption that
+nodes within the same domain are topologically close to each other. DDNO is
+a heuristic approach to solve the topology mismatch problem and uses
+\emph{Split-Hash} and \emph{dnMatch} algorithms to detect locality by using
+domain names. The result is a flat overlay topology but with some changes it
+can be utilized in super-peer architectures as well. According to the algorithm
+half of the possible connections of a node is used to connect to local peers
+(called \emph{sibling} connections), and the other half is used to randomly
+connect to the peers anywhere on the network (called \emph{random} connections).
+The former ensure the refuction of long distance travelling for messages, while
+the later secure th connectivity of the structure and prevent partitioning. Upon
+arrival the newcomer asks a \emph{hostcache} in order to establish its random
+connections. For establishing its sibling connections, the newcommer multicasts
+a special walker message in order to help it locate the right candidate. If it
+is found then the candidate as well as its own siblings will respond to the
+newcomer about their willingness to establish connections with it.
+
+\paragraph*{\bf Critical Topology-Aware Grouping}
+\emph{Critical Topology-Aware Grouping (CTAG)} presented in
+\cite{zhao_ctag_2006}, is a grouping algorithm. The grouping strategy is based
+on the \emph{Internet Assigned Numbers Authority (IANA)} and the respective
+\emph{Regionanal Internet Registry (RIR)}'s IP assignment strategies, according
+to which nodes within the same organization are always addressed from the same
+block. The paper proposes to calculate peer proximity using longest maching IP
+segment criterion, called \emph{Adjacency Measurement (AM)} in the technique
+which uses the longest matching IP segment criterion to calculate node
+proximity. CTAG focuses on both the construction of the overlay as well as its
+constant and dynamic adaptation. In the first, called \emph{bootstrapping
+grouping}, the \emph{Gnutella Web Caching} mechanism is modified in order for a
+newcomer to choose the closest cache to retrieve the bootstrapping list. In the
+second, called \emph{dynamic revision} in the paper's context, the AM metric is
+used to store hosts' addresses. read from \emph{X-Try} headers during handshake
+or from \emph{QueryHit} messages. Additionally, when a node reaches the max
+neighbour connections, node disconnects the neighbours with the lowestAM.
 
@@
 \end{figure*}
+
 Landmark Binning \cite{ratnasamy_binning_2002} is an approach to partition close
 by nodes into bins based on their distance to well known anchor nodes across the
-Internet, as shown in Figure \ref{fig:landmark_binning}. In order to detect
+Internet, as shown in Figure~\ref{fig:landmark_binning}. In order to detect
 locality, nodes use network latency (i.e. RTT) as a measurement technique. The
@@
 the Internet. Authors estimate that a number of 12 such servers can prove
-sufficient for this. The nodes measure their distance to all these servers and
-record the latency values for each. Later, the tuple of these latency values
-represents the bin of each node. Nodes with similar latency values are
-considered close to each other, or at least within the same autonomous system.
-The operation of the method is independent of the model incorporated by the
-overlay network and thus it can be applied with no significant changes to both
-structured and unstructured P2P systems. \cite{ratnasamy_binning_2002} has the
-detailed description of such algorithm for both architectures. Landmark binning
-is also considered as a good candidate for content distribution networks in
-particular. The major disadvantage of landmark binning is to install and
-maintain landmark servers on different autonomous systems, all over the world.
-A typical P2P network usually has a couple of million nodes connected at any
-given time, thus rendering landmark servers an important parameter when
-designing for large scales.  The authors claim that latency estimation does not
-drain much from the network resources but in case scalability problems arise,
-they claim the solution is replacing single landmark servers with clusters of
-servers within the same physical area. However, this approach does not reduce
-the possibility of excessive network traffic flow through these landmark
-servers. One other possible problem is the incorrect binning caused by the use
-of inaccurate methods for delay measurement, since tracking down, for example,
-network latencies is far from proven to be safe for location estimation.
-
-% TODO: MAYBE THANCS deserves some more attention here
-
+sufficient for this. Every newly arriving node measures its distance from these
+landmarks and unilaterally decides to join a specific bin based on these
+results. Specifically, the node measures its round-trip time to each of the
+landmarks and orders the values in a decreasing order. The ordering represents
+a \emph{bin}, in the sence of close-by nodes having the same landmark ordering
+and hance belong to the same bin. This means that a landmark system consisting
+of $m$ such nodes results in $m!$ potentional different bins.The operation of
+the method is independent of the model incorporated by the overlay network and
+thus it can be applied with no significant changes to both structured and
+unstructured P2P systems. \cite{ratnasamy_binning_2002} has the detailed
+description of such algorithm for both architectures. Landmark binning is also
+considered as a good candidate for content distribution networks in particular.
+The major disadvantage of landmark binning is to install and maintain landmark
+servers on different autonomous systems, all over the world. A typical P2P
+network usually has a couple of million nodes connected at any given time, thus
+rendering landmark servers an important parameter when designing for large
+scales.  The authors claim that latency estimation does not drain much from the
+network resources but in case scalability problems arise, they claim the
+solution is replacing single landmark servers with clusters of servers within
+the same physical area. However, this approach does not reduce the possibility
+of excessive network traffic flow through these landmark servers. One other
+possible problem is the incorrect binning caused by the use of inaccurate
+methods for delay measurement, since tracking down, for example, network
+latencies is far from proven to be safe for location estimation.
+
+%The algorithm can be considered scalable, as nodes need only compute distances
+% to a small number of predefined nodes and thus without exchanging any
+%information.
+%To answer the question as to whether the algorithm actually contributes
+% possitivelly to the construction of an enhaned overlay, the paper defines the
+%\emph{gain ratio} as the factor by which the latency reduces when someone
+%communicates with a random node from the same bin than with one not in the bin.
+%This is implemented with an inter-bin to an intra-bin latency ratio.
+
+%
+% TODO: LANDMARK BINING FOR UNSTRUCTURED OVERLAYS
+%
+%For unstructured overlays the paper assumes \emph{a set of $n$ nodes where each
+% node picks any $k$ neighbour nodes so that the average routing latency on the
+%resultant overlay is low (assuming shortest path routing)}. According to the
+%proposed heuristic algorithm called \emph{BinShort-Long}, a node picks its
+%neighbours by choosing its $\frac{k}{2}$ closest\footnote{If the node's bin is
+%not large enough for it to pick these $\frac{k}{2}$ neighbours, it picks the
+%required nodes from the bin that matches the most in terms of landmark
+%ordering.} ones (named \emph{short links}), using the \emph{binning} scheme and
+%the rest $\frac{k}{2}$ randomly (\emph{long links}). The former set produces
+%well-connected \emph{pockets} of nearby nodes while the later preserves the
+%connectivity of the graph, both yielding a proximity factor of $\alpha = 0.5$
+%in an attempt to preserve the benefitial properties of unstructured
+%topologies\cite{merugu_str2unstr_2003}.
+
+%
+% TODO: SOME DISCUSSION
+%
+%A potential bottleneck could be the extra load that this
+%``ping''-like scheme imposes to the landmarks, especially when we need instant
+%reaction from our topology when dealing with the dynamic nature of the p2p
+%networks.
+%
+%One disadvantage of this landmark scheme is related to the additional burden
+% imposed to the landmark sites. The authors claim though that the algorithm
+%requires so little work by the landmarks (maybe just echo to ping messages)
+%that could in effect, act as ``unsuspecting participants''. Even if this is the
+%case, the fact that it is not fully distributed, renders the protocol's
+%scalability directly volnerable to any system size increase as well as
+%usuitable for highly dynamic networks such as ad-hoc networks. Moreover, fixed
+%points in a network are inherently more exposed to malicius attacks. The most
+%significant downside of the algorithm though is that it can lead to an
+%extremely uneven overlay ID distribution causing load inbalances and hot spots.
+%Lastly, the scheme is coarse grained when it comes to distinguishing relatively
+%close nodes\footnote{In the worst case, all nodes could ve clustered into a
+%single bin.}.
+%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \paragraph*{\bf mOverlay}
-mOverlay \cite{zhang_moverlay_2004} tries to addresses the scalability
-issues that might be imposed to networks when using static landmark servers by
-introducing the dynamic landmarks. The authors introduce the notion of a group
-which is a set of  peer considered close to each other with respect to any
-position in the underlying network, proximity being some  user-defined cost
-metric (i.e. round-trip-time). mOverlay, tries to recreate small-world-like
-properties to the overlay network by creating a a two-level hierarchical
+\emph{mOverlay} \cite{zhang_moverlay_2004} tries to addresses the scalability
+issues that might be imposed to networks (i.e. load unbalance) when using static
+landmark servers by introducing dynamic landmarks. Zhang et al. introduce the
+notion of a \emph{group} which is a set of peers considered close to each other
+with respect to any position in the underlying network, proximity being defined
+by some user-defined cost metric like RTT, latency or something else. mOverlay
+is considered a clustering approach since it tries to recreate small-world-like
+properties to the overlay network by creating a two-level hierarchical
 structure, where on the top level we have connections between groups while on
@@
 that any new node can reach its group by exchanging at most $O(logN)$ messages
-within the network. Finally mOverlay also considers a second important function
+% within the network. Finally mOverlay also considers a second important
+function
 in its protocol which is the constant maintenance of the overlay.
+
+%
+%\paragraph{Locating process} A new coming host, $Q$, first connects to a
+%globally known host cache called the \emph{rendezvous point (RP)} in order to
+%retrieve the starting point in the overlay, say $A$ in group $1$. Host $Q$
+%then, measures its distance to host $A$. At the same time, the later, sends
+%information about the neighbour groups of group $1$ back to host $Q$. This list
+%is called \emph{candidate group list}, and the newcoming host sequnentially
+%measures its distance to each of them in seek for the closest one. If the
+%\emph{grouping criterion} is met, host $Q$ belongs to group $1$. If not, a boot
+%host from the closest group is found and the algorithm is re-run until the
+%criterion is met or after a predefined number of repetitions. In the later
+%case, $Q$ creates a new group comprising itself only. The above protocol does
+%not favour hotspots as it spreads the probability of visiting a group across
+%the whole overlay and limits the overhead in the level of $O \left ( log N
+%\right )$.
+%
+%\paragraph{General overlay operations} A set of additional protocols, are also
+% introduced, similar to those found in traditional unstructred networks, but
+%modified focusing on scalability and robustness. For example a protocol for
+%\emph{group formation} is introduced that exploits the inherent characteristic
+%of proximity, in the overlay, in order to efficiently detect the neighbouring
+%groups of a newly formated group from the set of adjasent groups of its closest
+%neighbour. Additionaly, during \emph{group joining} the coresponding protocol
+%denotes the exchange of important information for group maintenance. This can
+%be furtherly improved by \emph{information sharing} between nodes of the same
+%group, functionality handled by a dedicated flood-like protocol\footnote{Since
+%nodes that belong to the same group are physically close this can be achieved
+%at a minimum price.}. Moreover, another set of distributed protocols handle the
+%\emph{information update}. The information that needs update, in the proposed
+%architecture, is
+%\begin{inparaenum}[\itshape i\upshape)]
+%  \item the host cache, when a new node joins, and
+%  \item the neighbours of groups, when a close-by group is generated.
+%\end{inparaenum}
+%Finally, in case of \emph{host failure} or \emph{host departure} the system is
+% able to maintain its stability since there are defined operations for
+%periodical host cache update and group leader selection if one leaves or dies.
+%
 
@@
 \hline
-\textbf{MAY OMIT- T2MC} &
+\textbf{MAY OMIT - I CHANGED IT TO STRUCTURED SINCE THERE IS A REFERENCE FOR
+DHT (OF COURSE IT MIGHT POSIBLE TO BE APPLIED TO BOTH. MAYBE NEED TO CHECK) -
+T2MC} &
 \textbf{Overlay optimization based} Uses traceroute results for clustering the

### 2014-01-23 background.tex
2012.06.26 (Mema) Big introduction section split to a crispier 3-page intro and a background section. The latter also shortened and updated.

--- 2014-01-21 intro.tex
@@
-\section{Introduction}
-\label{section:intro}
+\section{Background}
+\label{section:background}
 
-Distributed architectures have witnessed a significant increase of interest
-during the last decade, as the number of interconnected computers increased, the
-available digital information kept proliferating, at an exponetial rate each
-year, and problems to be addressed by IT became more and more demanding. Why is
-there a correlation, and it is not just a coincidence? Distributed computation
-suggests a simple paradigm\footnote{Of course, as always, this is the case only
-after someone thought about it first!}. Get a set of autonomous computing
-entities and put them to cooperate with each other in order to fullfil a well
-defined task. In the World of the Internet and the World Wide Web this also
-offers a thrilling promise. To harness the immence power scattered across the
-Internet and get amply rewarded by exploiting the data, storage and
-computational resources out there.
-
-In order to achieve that, computers collaborate by exchanging messages via an
-application-level network where they can address each other in the specific
-context of the task at hand. This communication structure is refered to as an
-\emph{overlay network} which is an abstract, logical inter-connection formed on
-top of another network\footnote{The Internet itself is an overalay on top of the
-telephone network.}. It consists of a set of nodes, the computing elements, that
-are being connected by virtual (logical) point-to-point links, perhaps
-corresponding to multiple physical links. Figure~\ref{figure:overlay}
-illustrates a simple four-node overlay contructed upon a wide area network.
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.5]{img/p2p.pdf}
-\caption{An example overlay network.}
-\label{figure:overlay}
-\end{figure}
-
-The abstraction of the overlay network has been proposed as a way to implement
-efficient, fully distributed, application-layer services. For example, backing
-the routing of messages to destinations that are not known in advance or
-building one-to-many addressing schemes
+%The abstraction of the overlay network has been proposed as a way to implement
+%efficient, fully distributed, application-layer services such as  
+%routing messages to destinations that are not known in advance or building
+%one-to-many addressing schemes 
 % TODO find citation for Narada & Adobe's Real Time Media Flow Protocol (RTMFP)
-\cite{} that would provide higher quality-of-service (QoS) guarantees. \emph{IP
-multicast}, as well as other proposals such as \emph{IntServ} and
-\emph{DiffServ} \cite{cisco_diffserv_2005}, have not seen wide acceptance
-(yet?), largely due to the fact they require support from the physical
-infrastructure. On the other hand, an overlay network can be deployed on
-end-systems running the overlay protocol software, without cooperation from the
-ISPs.
+%\cite{} that provide quality-of-service (QoS) guarantees. \emph{IP
+%multicast}, as well as other proposals such as \emph{IntServ} and
+%\emph{DiffServ} \cite{cisco_diffserv_2005}, have still not seen wide deployment
+%largely due to the fact they require support from the IP
+%infrastructure. On the other hand, an overlay network can be deployed on
+%end-systems running the overlay protocol software, without the cooperation of ISPs.
 %Academic research includes
@@
 
+In this section, we provide some background information on the types
+of P2P overlay architectures that have been implemented and deployed.  We also
+describe and motivate more formally the topology mismatch problem tackled
+by many P2P researchers over the past several years.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Peer-to-Peer (P2P) Architectures in a Nutshell}
+\subsection{Peer-to-Peer (P2P) Overlay Architectures}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Overlay networks on top of the best-effort IP layer, formed a widely ranged
-family of protocols collectively known as \emph{peer-to-peer} or \emph{P2P} for
-short. P2P systems consist of a number of networked computers called
-\emph{peers}, where each, such, \emph{peer} could act both as a resource
-provider and a resource consumer. This changed the traditional
-\emph{client-server} model dominating the Internet and lead to the introduction
+Circa 2000, the first P2P file-sharing systems introduced the notion 
 of the \emph{servent} concept \cite{gnutella}, a \emph{portmanteau} that blends
 the notions of server and client to denote the twofold role of the P2P network
-participants. Such serverless systems, proved to be able to achieve outstanding
+participants. Such serverless systems have proven to be able to achieve outstanding
 aggregate resource capacities as more and more participants join the system
-without requiring additional expenditure for infrastructure\footnote{
-  Unfortunatelly, there has been observed mitigation of this self-scaling
-  property by the undesirable behaviour of participating users, usually answered
-  in popular file-sharing systems; from \emph{free-riding}
+without requiring additional expenditure for infrastructure.  While
+certain undesirable features such as 
+free-riding~\cite{saroiu_measurefileshare_2002,adar_gnutellafreeriders_2000,hughes_gnutellafreeride_2005},
+the distribution of illegal content and/or other \emph{socio-technical}
+issues \cite{hughes_socp2p_2008}, have arisen, P2P file-sharing systems 
+have continued
+to gain popularity because they have successfully supported
+popular
+applications amongst vast numbers of participating
+users.
 
-\cite{saroiu_measurefileshare_2002,adar_gnutellafreeriders_2000,hughes_gnutellafreeride_2005}
-to the distribution of illegal content and/or other \emph{socio-technical}
-issues \cite{hughes_socp2p_2008}.
-}.
-P2p network systems gained extra attention whithin the distributed systems
-super-set, circa 2000, due to the fact that they succesfully supported popular
-applications that offered file sharing among vast numbers of participating
-users. But their development didn't stay still. On the contrary, during the last
-decade, research on P2P systems was constant and fruitfull generating a wide
-range of popular protocols, networks and applications. As P2P systems evolved
+For over a decade, excitement over the immense potential of P2P systems
+has generated a flurry of research and development, resulting in a wide 
+range of popular protocols, networks, and applications. As P2P systems evolved
 three main architectures have been emerged, namely
@@
 
-\emph{Centralized} architectures (see Figure~\ref{figure:p2p-archs:centralised})
-were the first to recognize that requests for resourses (e.g. CPU cycles
-\cite{seti} or popular content
+Designers of \emph{centralized} architectures (see Figure~\ref{figure:p2p-archs:centralised})
+were the first to recognize that requests for resources (e.g. CPU 
+cycles~\cite{seti} or popular content
 % TODO find citation for Napster
 \cite{}) need not be sent to a dedicated server but instead could be handled by
-the many hosts that already posses them. This schema, though, is not fully
-distributed since it dictates a centralised searching facility, a characteristic
-that can be proven its ``Achilles' heel'' in terms of being a scalability
-bottleneck, a central point of failure and vulnerable to malicious acts (e.g.
-\emph{Denial-of-Service (DoS) attacks}).
+the many hosts that already posess the resources. The downside of
+this approach is it requires a centralized searching facility which becomes
+a single point of failure, a scalability bottleneck, and a target of malicious 
+attacks (e.g., \emph{Denial-of-Service (DoS) attacks}).
 
-The centralized approach has been eventually replaced by architectures that
-distributed both search and download capabilities (see
+The most successful incarnation of the centralized approach is
+the file-sharing application \emph{Napster}
+% TODO find citation for Napster again
+\cite{}.
+Napster maintained a \emph{central index server} based on file lists provided by
+participating peers. The central index server was queried by users and it
+returned pointers to the actual content. Thus, by centralizing search while
+distributing downloads, Napster achieved a highly functional design that 
+at its height, attracted 26 million users sharing 80 million 
+songs~\cite{jmm_naptopusage_2001}.  
+Napster's centralized nature was ultimately what brought it down.
+The Recording Industry Association of
+America (RIAA) 
+filed a lawsuit for copyright infringement of the intellectual
+property of artists whose music was distributed by the Napster application.
+RIAA was able to point to the centralized indexing component of
+Napster as the entity responsible for illegal song-trading.
+
+The centralized approach was quickly replaced by architectures that
+distribute both search and resource provision capabilities (see
 Figures~\ref{figure:p2p-archs:unstructured-full} and
 \ref{figure:p2p-archs:unstructured-hierarchical}). In these \emph{decentralized}
-architectures, file placement is random, which means there is no correlation
-with the network topology whatsoever \cite{yang_improvep2psearch_2002}. For this
-reason they are also refered to as \emph{unstructured}. The most important
+architectures, resource placement within the overlay topology is 
+random~\cite{yang_improvep2psearch_2002}. For this
+reason they are referred to as \emph{unstructured}. The most important
 properties of such systems are that they support inherent heterogeneity of
@@
 the literature, they are also known as \emph{broadcast-based} systems, because
-they use \emph{message flooding} among peers or super-peers to propagate
-queries.
-%\footnote{The life of each query message is constrained by an integer, refered
-%to as \emph{time-to-live (TTL)}. When a querry is received by a (super-) peer,
-%this value is decreased by one unit. Then, the receiver, back-propagates its
-%results and if the TTL has not reached zero the node forwards the query message
-%to all its neighbours except from the one it received it from.}
+they use message flooding among peers to propagate
+search queries. A well-known example of the fully dentralized unstructured 
+approach is Gnutella, in its first realisation (v.0.4) \cite{gnutellav04}.
 
-Recently, \emph{decentralized structured} schemes have been proposed (see
-Figure~\ref{figure:p2p-archs:structured}) in order to provide a self-organising
+A key disadvantage of the unstructured approach
+is that message flooding burdens the network since queries travel within the
+network randomly, visiting nodes that do not have the resource of interest
+and thus wasting bandwidth.  To save on bandwidth, unstructured networks
+typically limit how long a query travels within the overlay (via a
+time-to-live (TTL) parameter).   When a query is received by a node, it decreases
+the TTL by one unit.  If the TTL has not reached zero, the node
+forwards the query to its neighbors.  Otherwise, it drops the query.  While
+this approach prevents queries from visiting the entire network, it does
+not guarantee that the resource of interest, (e.g., a rare recording 
+of a musical group from
+the 1930s) will be found.
+
+To improve search scope and scalability and to lighten network load, 
+unstructured systems evolved into ``hierarchical'' systems that dynamically
+assign indexing functions to special peers,
+called \emph{ultra-peers}  or \emph{super-peers}.
+Peers with reliable network connections
+and/or high compute power became ultra/super-peers facilitating less 
+powerful peers (called \emph{leaf-nodes}) to find resources of interest.
+Well-known examples of these hierarchical unstructured systems include
+Gnutella v.0.6~\cite{gnutella}, Limewire[XXX], and KaZaA[XXX], and Skype[XXX]. 
+%% TODO: add references for the above line.
+While the hierarchical approach relieves network pressure as search
+queries are flooded over a much smaller subset of peers (the ultra/super-peers),
+the disadvantages of reduced search scope and inability to locate
+rare resources efficiently remain.
+
+To fully address the disadvantages of decentralized unstructured approaches,
+\emph{decentralized structured} schemes have been proposed (see
+Figure~\ref{figure:p2p-archs:structured}).  The key goal
+of these systems is to provide a self-organising
 infrastructure for large-scale P2P applications
@@
 They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes
-through a deterministic mechanism. Even though unstructured networks became
-highly popular among file sharing applications, they do have a certain
-limitation in locating rare objects. The DHT approach of structured networks on
-the other hand, provides a guaranteed bound on the number of overlay routing
-hops that have to be taken in order for an object to be located, even in the
-case where only a single copy exists in the system.  This means
+through a deterministic mechanism. 
+The DHT  
+provides a guaranteed bound on the number of overlay routing
+hops that have to be taken to locate a resource, even in the
+case where only a single copy of the resource exists in the system.  This means
 $O \left ( log n \right )$ hops, compared to unstructured networks that require
-$O \left ( n \right )$ to reliably locate a specific object\footnote{
-  Unfortunatelly this efficiency does not come without cost. Structured overlays
-  realise what is called object identifier location as opposed to content-based
-  retrieval of unstructured ones. This means that the later can support more
-  versatile object location mechanisms like keyword and partial matching
-  searches.}.
+$O \left ( n \right )$ to reliably locate a specific object.
+Unfortunately this efficiency does not come without cost. Structured overlays
+can only support exact-match queries (i.e., queries that identify resources
+by name) as opposed to content-based
+retrieval provided by unstructured overlays.  This means 
+that unstructured overlays can support more 
+versatile resource location queries such as keyword and partial matching
+searches.  Several popular file-sharing peer-to-peer applications have
+been based on DHTs including Overnet[], Kademlia[], the Kad Network[], 
+some flavors of BitTorrent, and many
+more.  Moreover, DHTs have sparked a flurry of research activity and 
+development for a large range of applications: from distributed search 
+engines to event notification infrastructures to cloud-based platforms.
+We cite several examples in the Introduction.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% TODO: SOME DISCUSSION ON P2P ARCHITECTURES
+%
+%\cite{matei_mapgnutella_2002, lv_randomwalks_2002, merugu_str2unstr_2003}
+%showed that the ad-hoc network topology of unstructured overlay networks that
+%preserve \emph{Power Law} and \emph{Small World} characteristics\footnote{Power
+%Law describes the node degree while the Small World describes characteristics
+%of path length and clustering coefficient. The clustering coefficient for a
+%node $\upsilon$ in a graph $G = \left( V, E \right)$ is defined as the ratio of
+%the existing connections between $\upsilon$'s neighbouring nodes to $\gamma
+%\times \left( \gamma - 1 \right)$, where $\gamma$ is the number of neighbouring
+%nodes of $\upsilon$. High cluster coefficient means that neighbouring nodes of
+%any node $\upsilon$ likely connect one another.} \cite{faloutsos_powerlaw_1999,
+%saroiu_measurefileshare_2002} offer a more promising approach. Particularly:
+%\begin{itemize}
+%  \item Peer-to-peer clients are extremely \emph{transient}. Unstructured
+%systems can have high maintenance traffic in delivering messages, updating the
+%mapping, discovering failures and replicating lost data or pointers, making
+%them insufficient on highly volatile networks.
+%  \item \emph{Keyword searches} versus \emph{exact-match queries}. In DHTs
+%there is a tight control between the data placement and the topology of the
+%network. For this reason it is hard to efficiently support partially matched
+%queries while Gnutella and other similar systems effortlessly support keyword
+%searches and other complex queries since the mechanism is realized locally, on
+%a node-by-node basis.
+%  \item Popular content is located at multiple peers and thus it is more likely
+%for a flooding-based search to return results. DHTs, on the other hand, fit
+%better in the systems which require ability to reliably locate content, even in
+%the extreme case that only a single-copy exists in the network.
+%\end{itemize}
+
+%
+% TODO: SOME DISCUSSION ON THE ADVANTAGES OF UNSTRUCTURED NETWORKS FROM
+%       \cite{z-yk_ddno_2005}
+%
+%Unstructured P2P networks o?er a number of
+%important advantages: (i) An unstructured network
+%imposes very small demands on individual nodes,
+%and more speci?cally it allows nodes to join or leave
+%the network without signi?cantly a?ecting the sys-
+%tem performance. (ii) Unstructured networks are
+%appropriate for content-based retrieval (e.g., key-
+%word searches) as opposed to object identi?er loca-
+%tion of structured overlays. (iii) Finally unstructured
+%networks can easily accommodate nodes of vary-
+%ing power. Consequently, they scale to very large
+%sizes and they o?er more robust performance in
+%the presence of node failures and connection
+%unreliability.
+%
+
+%
+% TODO: SOME DISCUSSION ON STRUCTURED OVERLAYS PROBABLY FROM
+%       \cite{wu_laptop_2007}
+% However, the design of a decentralized but structured P2P network has to
+% overcome two critical issues. The first issue is the long routing latency.
+% Several proximity schemes  have been proposed to avert long routing latency in
+% current structured P2P networks, but they require a high-complexity procedure
+% to periodically maintain the routing table (e.g. Pastry system) or they need
+% pre-chosen landmarks to construct the overlay. However, the P2P system is by
+% its very nature unstable since nodes join and leave frequently. For instance,
+% the study of Gnutella shows around approximately 1200 membership changes per
+% minute in a 100 000 nodes P2P system. Another proximity scheme needs some
+% pre-chosen landmarks or a complete BGP routing table support. As a result,
+% they both increase the difficulty of the P2P system deployment. The second
+% issue is system maintenance overhead. The existing structured P2P networks
+% allow nodes to keep some nearby nodes in their routing tables in order to
+% achieve efficient routing.
+%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Applications and Implications}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The most successful incarnation of the centrilised approach is without a doubt
-the popular file-sharing application \emph{Napster}\footnote{Pioneering, though,
-was the famous \emph{SETI@Home Project} \cite{seti}, which pools spare
-processing power of participants to search the universe for the existence of
-extra terrestrial intelligence!}
-% TODO find citation for Napster again
-\cite{}.
-Napster maintained a \emph{central index server} based on file lists provided by
-participating peers. The central index server was queried by the users and it
-returned pointers to the actual content. Thus, by centralizing search while
-distributing downloads, Napster achieved a highly functional design that was
-widely acknowledged, at the time, as ``the fastest growing Internet application
-ever'' that toped some 26 million users and 80 million songs
-\cite{jmm_naptopusage_2001}.
-
-Napster was ultimatelly brought down after Recording Industry Association of
-America (RIAA) filed a lawsuit for copyright infringement of the intellectual
-property of artists whose music was distributed by the network\footnote{
-  The sueing spree was initiated by heavy metal giants \emph{Metallica} that
-  feared Napster was undermining record sales. Opposed opinions instantly
-  emerged with best example the british alternative rock band \emph{Radiohead}
-  whose album \emph{Kid A} had also been leaked to Napster three months before
-  reaching music store shelves \cite{rm_radioheadkida_2000}. Despite music
-  critics labeling the album ``experimental'' and ``challenging'' (it had no
-  official singles nor music videos for any of its songs), in October 2000 Kid A
-  captured the number one spot on the Billboard 200 sales chart in its debut
-  week. Unlike Metallica, Madonna and Dr.Dre that led litigation with Napster,
-  Radiohead had never hit the top20 in the US before.
-}
-\cite{wiki_amnaptrial_2001}. This triggered a shift towards decentralised
-networks like the classic \emph{Gnutella} and, its ``rival'' at the time,
-\emph{FastTrack}. Gnutella, in its first realisation (v.0.4 \cite{gnutellav04}),
-was a fully decentralised approach. Unfortunatelly, network resourses tended to
-be overwhelmed by message flooding and if queries were constrained to live for
-some (small) number of hops, this ultimatelly resulted to a shrinked search
-scope, leading to a death spiral that prevented the network from scaling. The
-solution came with a slightly more structured network by introducing the notions
-of \emph{ultra-peers} and \emph{leaf-nodes} (v.0.6 \cite{gnutella}). With
-\emph{Limewire}, the most widely used software that implemented the protocol,
-the network was highly successful and was reported to have reached connecting
-more that 3 million users concurrently by the end of 2006
-\cite{rsr_gnutellaevol_2006}. Fasttrack and its implementing software KaZaA
-\cite{kazaa}, on the other hand, went with the hierarchical approach
-(\emph{super-nodes} in their context) from the very begining, dynamically
-assigning indexing functions to special peers in order to increase search
-performace and scaling potential. FastTrack network had an average of 3 million
-concurrently connected nodes in 2003 \cite{lkxr_kazaameasure_2005}. It is
-estimated that the total number of users was greater than that of Napster at its
-peak. With Limewire discontinued/shutdown in 2010 and its successor,
-\emph{Frostwire}, moving to the Bittorent protocol in 2011, seem to very well
-signal the end of the once-mighty Gnutella network. KaZaA continues as a
-propriatery music store solution. In the meanwhile, Fastrack design team had
-already introduced a VoIP and text messaging application based on the design
-principles of KaZaA's hierarchical P2P nature, that was meant to change how
-people communicate around the world. This app was no other than Skype
-\cite{skype}. Despite its wide usage, little is known for the technical aspects
-of its overlay network, mainly due to its proprietary nature. Studies, though,
-confirm that it shares strong structural similarities to its file-sharing
-predecessor (ordinary hosts and super-nodes in Skype's terminology) but they
-significantly differentiate in terms of traffic behaviour
-\cite{g_voipskypesec_2005,gdj_skypestudy_2006}. Skype was introduced in 2003
-and from then on it witnessed frenzy-rythmed increase to its popularity, with
-more that 650 milion registered users in 2011 \cite{skypetotalusers} and
-reported to have broken the barrier of 35 million skypers concurrently online on
-the the 5th of March 2012 \cite{skypesymusers}. On October 2011 Microsoft Corp.
-and Skype Global S.à.r.l. anounnced the completion of Microsoft's acquisition of
-Skype for \$$8.5$ billion \cite{skypemicrosoft}.
-
-File sharing was the most popular application area for structured architectures
-as well. Overnet
-% TODO find citation for Overnet
-\cite{},
-for example, was created for sharing large files (e.g., movies and CD images).
-It implemented the Kademlia tree-like partitioning algorithm. Other
-implementations of Kademlia are the Kad Network (eMule/aMule  clients),
-BitTorrent for trackerless torrernts (official BitTorrent client, $\mu$Torrent,
-BitComet, Transmission and BitSpirit all share compatibility with the Mainline
-DHT which realises the Kademlia protocol) and the Gnutella DHT (first
-implemented by LimeWire as the Mojito DHT). Lately, RetroShare \cite{retroshare}
-is the new trend for file-sharing which incorporates an implementation of the
-Kademlia DHT to create a Friend-to-Friend (F2F) network with some attributes of
-a darknet as well. But the potential of structured architectures was not
-exhausted only to the above. Structured P2P networks became popular in
-distributed systems because they could scale smoothly and have fast and reliable
-object location mechanisms. Thus, research and development was very active and
-productive resulting in a variety of diverse applications such like distributed
-search engines \cite{yaci}, distributed data storage systems
-\cite{kbc_oceanstore_2000,bdet_fsdfs_2000,dkkms_cfs_2001,dr_pastutility_2001,abc_farsite_2002,mmfc_ivy_2002,arla,agebh_dks_2003}, Web caches, archives,
-and publishing systems \cite{ird_squirrel_2002,bags_youserv_2002,wrc_publius_2000,wm_tangler_2001},
-messaging applications \cite{threedegrees}, event notification infrastructures
-\cite{rkcd_scribe_2001,cdkr_scribe_2002,agebh_dks_2003}, naming services
-\cite{cmm_chorddns_2002}, censor-resistent stores \cite{cswh_freenet_2001} and
-lately even cloud-based platforms \cite{mgpj_cloudsnap_2011}.
-
-%%%%%%%% NEW TRENDS %%%%%%%%
-%File sharing Limewire Kazzaa The file-sharing landscape is slowly adjusting in
-%response to the continued push for more anti-piracy tools, the final Pirate Bay
-%verdict, and the raids and arrests in the Megaupload case, SOPA, ACTA. This
-%leads to users trying to find new solutions for anonymous, uncensored,
-%decentralised file sharing solutions like Freenet, Tribler and RetroShare where
-%connections are made only between trusted peers — sometimes called "friends"
-%and the respective formed network F2F.
-
-% TODO: READ Dalesa published at eAsia 2009, http://www.e-asia.org/2009/ET_Presentations/Wathsala.pdf
-
-The huge numbers of users, for the various applications we presented above,
-reveals the tremendous popularity P2P architectures have enjoyed. This, though,
-does not come without any toll. Measurements on some popular systems, such as
-FastTrack, Gnutella, DirectConnect and BitTorrent have shown that the P2P
-traffic contributes the largest portion of the overall Internet traffic
-\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}.
-Even with later studies, this observation still holds on. Using deep packet
-inspection techniques deployed at selected customer sites \emph{Ipoque} and
-\emph{CacheLogic} were able to track down application usage \cite{cachelogic,ipoque2007,ipoque2009}.
-CacheLogic reports that more than 60\% of the traffic is identified as P2P, with
-Ipoque verifying this observation reporting similar results for 2007, while
-observing some small decline for 2008/2009 (but still remaining dominant
-compared to other applications). Apart from its participation in overall usage
-terms, P2P is already a huge contibutor to the overall data exchanged across the
-Internet\footnote{
-  Mainly because file-sharing P2P is used to exchange big files of HD video or
-  high sampled audio content compared to simple Web content distributed via, for
-  example, HTTP.
-}.
-\cite{multinteligence} claims that P2P traffic in 2007 was around 1.6 petabytes
-per month and the potential is to be quintubled by the end of 2012 reaching
-almost 8 petabytes per month.
 
@@
 If these peers participate in an overlay network according to one of the setups
-of Figure~\ref{figure:overlay-confs} then users will expierience different
+of Figure~\ref{figure:overlay-confs}, then users will experience different
 performances. In the overlay depicted in Figure~\ref{figure:overlay-1}, the
@@
 ~\ref{figure:overlay-2}, the path will be: $3 \rightarrow 8 \rightarrow 4$. The
-total cost of the first sequnce is $45 ms$, while the second yields a mere $15
+total cost of the first sequence is $45 ms$, while the second yields a mere $15
 ms$. Therefore, we can conclude that the second overlay is more congruent with
@@
 
-Ideally, overlay networks are expected to achieve an optimal mapping among the
-overlay paths and the underlying physical links, and avoid inefficient states.
+Ideally, an overlay network should be constructed to achieve an optimal 
+mapping among the
+overlay paths and the underlying IP network links to avoid inefficient states.
 The problem of constructing an optimal overlay is referred to as the
-\emph{topology mismatch problem}, and formally defined as follows:
+\emph{topology mismatch problem}, and is formally defined as follows:
 \begin{definition}
@@
 
-Early incarnations of the overlay protocols, however, did not make use of the
-optimal mapping with the physical network topology. Gnutella protocol for
-example was considered far from scalable \cite{ritter_gnucantscale_2001}. Since
-every peer can chοse its logical neighbours without knowledge of the underlying
-network had as a result a mismatch between physical and application-level
-topologies. Additionally, queries may be flooded to multiple paths merging in
-one peer while in such case one of the paths would have been enough.
-Nevertheless, peers may forward the same message to each other before they
+Early incarnations of the overlay protocols did not make use of the
+optimal mapping with the underlying network topology. Gnutella, for
+example, is considered far from scalable \cite{ritter_gnucantscale_2001} because
+every peer chooses its neighbors in the overlay without any knowledge
+of the underlying network, resulting in a mismatch between the two topologies.
+Queries may be flooded over multiple paths merging in
+one peer while one of the paths would have been enough.
+Moreover, peers may forward the same message to each other before they
 receive the query messages from each other. \cite{matei_mapgnutella_2002} showed
-that, even given that 95\% of any two nodes are less than 7 hops away from each
+that, even when 95\% of any two nodes are less than 7 hops away from each
 other, a flooding-based algorithm, like the one used by Gnutella, can generate
-330TB/month in a 50.000 node network! Therefore, P2P traffic can have a serious
-effect on the overall well-being of the Internet as a whole.
+330TB/month in a 50.000 node network. A topology mismatch can thus have
+a serious effect on Internet router performance.
 
-Similar problems are observed in decentralized structured schemes also.
+Similar problems are observed in decentralized structured schemes as well.
 Typically, node IDs are assigned \emph{randomly}, resulting in excellent load
-balancing, scalability and robustness. Unfortunately, this randomness has a
+balancing, scalability, and robustness for the overlay. Unfortunately, this randomness has a
 negative impact on the \emph{routing locality} of the network. This means that
 even though the target node can be reached with logarithmic overlay hops, the
-distance traveled in the physical underlying network, during the overlay routing
-process, can be far from optimal.  The ratio of the physical distance, a query
-travels through the overlay to an object, and the minimal distance to that
-object (i.e. through IP) is known in the literature as \emph{stretch} or
+distance traveled in the underlying network, during the overlay routing
+process, can be far from optimal.  The ratio of the actual IP network distance a query
+travels to reach an object of interest (via overlay routing) to
+the minimal IP network distance to that
+object (i.e. through IP) is known as \emph{stretch} or
 \emph{Relative Delay Penalty (RDP)} \cite{chu_esm_2000}.
@@
 the overlay path. For example, a single logical point-to-point link on the
-overlay, most of the time, corresponds to multiple physical links in the
-underlying layer. Additionally, a link in the underlying network often serves
-the mapping of several overlay paths causing increase of the traffic on the
-physical link, which is also called the link's \emph{stress}\cite{chu_esm_2002}.
+overlay typically corresponds to multiple physical links in the
+underlying layer. Additionally, a link in the underlying network often lies
+on several overlay paths causing increase of the traffic on the
+link, known as the link's \emph{stress}~\cite{chu_esm_2002}.
+XXX MR:  The following sentence is unclear: XXX
 Furthermore, the network's \emph{churn}, that is the stochastic behaviour of
@@
 
-The topology mismatch problem is exacerbated by the high amounts of traffic
-generated by the highly popular P2P applications. This creats the killer
-combination for the network infrastructure, ultimately effecting all Internet
-users. Solving it, on the other hand, is no easy task. Topology mismatch is
+The topology mismatch problem is exacerbated by the extremely large
+amounts of traffic
+generated by highly popular P2P applications. As noted in the Introduction,
+studies have found that more than 60\% of the traffic on the Internet is
+P2P.  Thus, solving the topology mismatch problem is a fundamental challenge
+in contributing to the overall
+efficiency of the Internet.  Unfortunately, topology mismatch is
 known to be an NP-Hard problem \cite{NPBOOK,chawathe_scattercast_2000}.
-Moreover, the Internet itself is an interesting environment where end-to-end
-latencies demonstrate what is called \emph{triangle inequality violations
-(TIVs)}, which further complicate the problem. These delay space TIV is a
+Adding to the complexity of the problem is the fact that end-to-end latencies 
+demonstrate \emph{triangle inequality violations
+(TIVs)} which are a 
 consequence of the Internet's structure and routing policies and thus will
 remain a property of the Internet for the foreseeable future
-\cite{zheng_irprtt_2005}. TIVs affect network coordinate
+\cite{zheng_irprtt_2005}. TIVs affect both network coordinate
 \cite{cox_vivaldi_2004,wong_meridian_2005} and positioning \cite{ng_gnp_2001}
-and makes the construction of location/delay aware overlays difficult. An
-extensive array of heuristic approaches have been proposed in an effort to find
-an algorithm that performs reasonably and consistently well. The purpose of this
-survey is to gather the extensive knowledge that has been produced recently in
-this research field and organise the available information ultimately helping
-the reader sift through the myriad of these efforts, understand the different
-approaches and identify though a taxonomy the advantages and disadvantages of
-each of them.
+systems 
+and makes the construction of IP-topology-aware overlays difficult.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Survey Outline}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Over the past several years, researchers have proposed and studied 
+an extensive array of heuristic solutions to the topology 
+mismatch problem.
+The purpose of this
+survey is to gather and organize the extensive knowledge that has been produced. 
+We hope to assist the reader
+to understand the different approaches that have been proposed to tackle
+the topology mismatch problem and to identify, through our taxonomy, the
+advantages and disadvantages of each of them.
 
-% TODO: Review this after the content and article structure is fixed.
-In the introduction we had the oportunity to recognise the importance of overly
-networks in contemporary application architectures. Also we quickly reviewed the
-basic trends in both overlay construction\footnote{
-  The main focus of this paper is the topology mismatch problem, therefore the
-  available approaches in overlay networks are briefly summarized, and
-  interested readers are pointed to the available surveys on overlay networks
-  \cite{TheotokisS04,LuaCPSL05}.
-}
-and application development. Last but not least, we identified and formaly
-defined the phenomenon of the topology mismatch problem and its impact on the
-utilization of network resources. The rest of the survey\ldots
-
-Sections~\ref{sec:unstructured} and \ref{sec:structured} discuss the recent
-academic work that has been conducted in the field.
-
-Ultimately, it reviews the most important resources found in the literature that
-attempt to tackle the problem as it proposes a taxonomic scheme, using a set of
-supertype-subtype relationships based on unique characteristics they have and/or
-specific goals they target. What follows does not claim to be a thorough
-citation of all known protocols that are available out there, but instead, a
-careful selection of those that left a distinctive fingerprint contribution in
-the efforts of the research community to alleviate the topology mismatch
-problem.
+%
+% TODO: SOME DISCUSSION FROM \cite{xu_globstate_2003) CONTRADICTS TO THE
+%TRIANGLE INEQUALITY ARGUMENT ABOVE
+%
+%[snip]
+%Studies [14] have shown that triangle inequality may not hold in Internet
+%topology. In fact, study from Pastry has shown that the proximity approximation
+% is much worse when using the Mercator topology that is based on the real
+% measurements of the Internet [3].
+%

### 2014-01-23 intro.tex
2012.06.11 (Vassilis) Content revisited. Duplicate text has been removed.
2012.06.26 (Mema) Big introduction section split to a crispier 3-page intro and a background section. The latter also shortened and updated.

@@
 
-Distributed architectures have witnessed a significant increase of interest
-during the last decade, as the number of interconnected computers increased, the
-available digital information kept proliferating, at an exponetial rate each
-year, and problems to be addressed by IT became more and more demanding. Why is
-there a correlation, and it is not just a coincidence? Distributed computation
-suggests a simple paradigm\footnote{Of course, as always, this is the case only
-after someone thought about it first!}. Get a set of autonomous computing
-entities and put them to cooperate with each other in order to fullfil a well
-defined task. In the World of the Internet and the World Wide Web this also
-offers a thrilling promise. To harness the immence power scattered across the
-Internet and get amply rewarded by exploiting the data, storage and
-computational resources out there.
+Peer-to-peer (P2P) networks are self-organizing distributed systems
+where participating nodes, called \emph{peers}, act as both resource
+providers and resource consumers, in contrast to the traditional \emph{client-server}
+model where nodes take on the role of pure servers or pure clients.  
+Peer-to-peer networks have enjoyed immense attention from Internet users,
+companies, and researchers and have been widely
+deployed for over a decade, primarily because of the great
+number of features they offer distributed applications that are built
+on top of them.  These features include:  high availability and robustness,
+load-balancing, quality of service, scalability, decentralized
+administration, and anonymity.
 
-In order to achieve that, computers collaborate by exchanging messages via an
-application-level network where they can address each other in the specific
-context of the task at hand. This communication structure is refered to as an
-\emph{overlay network} which is an abstract, logical inter-connection formed on
-top of another network\footnote{The Internet itself is an overalay on top of the
-telephone network.}. It consists of a set of nodes, the computing elements, that
-are being connected by virtual (logical) point-to-point links, perhaps
-corresponding to multiple physical links. Figure~\ref{figure:overlay}
-illustrates a simple four-node overlay contructed upon a wide area network.
+A myriad of applications have been built and deployed using the peer-to-peer
+paradigm including two widely-known ``killer'' applications:  file-sharing and telephony.
+P2P file-sharing was originally introduced by Napster (see Section 2)
+which by 2001 was widely acknowledged as the ``fastest growing Internet application
+ever" with 26 million users sharing 80 million songs at the height of its popularity.  
+Other file-sharing applications
+quickly followed suit including Gnutella and its descendent, Limewire, connecting
+3 million users concurrently by the end of 2006 and BitTorrent, connecting
+over 150 million users.
+P2P telephony
+has seen explosive growth with the advent of Skype.  Introduced in 2003, Skype
+has experienced explosive popularity with 
+more that 650 million users in 2011 \cite{skypetotalusers} and
+35 million skypers concurrently online on
+the 5th of March 2012 \cite{skypesymusers}.
+
+In addition to file-sharing and telephony, a large number of diverse applications have
+been built on top of peer-to-peer systems including:
+distributed search engines \cite{yaci}, distributed data storage systems
+\cite{kbc_oceanstore_2000,bdet_fsdfs_2000,dkkms_cfs_2001,dr_pastutility_2001,abc_farsite_2002,mmfc_ivy_2002,arla,agebh_dks_2003}, Web caches, archives,
+and publishing systems \cite{ird_squirrel_2002,bags_youserv_2002,wrc_publius_2000,wm_tangler_2001},
+messaging applications \cite{threedegrees}, event notification infrastructures
+\cite{rkcd_scribe_2001,cdkr_scribe_2002,agebh_dks_2003}, naming services
+\cite{cmm_chorddns_2002}, censor-resistant stores \cite{cswh_freenet_2001} and
+lately even cloud-based platforms \cite{mgpj_cloudsnap_2011}.
+
+Peer-to-peer systems are implemented using overlay networks.   
+An overlay network is a virtual network of nodes and logical links 
+that is built on top of an existing network with the aim of providing 
+a service (such as content distribution) that is not available 
+in the existing network.  Nodes in the overlay are connected via virtual 
+links, each of which corresponds to a path of possibly many physical 
+links in the underlying network.  
+Figure~\ref{figure:overlay}
+illustrates a simple four-node overlay constructed over a wide-area network.
+In the case of peer-to-peer networks, typically the participating nodes
+are user PCs sitting at the edge of the Internet.
 
@@
 
-The abstraction of the overlay network has been proposed as a way to implement
-efficient, fully distributed, application-layer services. For example, backing
-the routing of messages to destinations that are not known in advance or
-building one-to-many addressing schemes
-% TODO find citation for Narada & Adobe's Real Time Media Flow Protocol (RTMFP)
-\cite{} that would provide higher quality-of-service (QoS) guarantees. \emph{IP
-multicast}, as well as other proposals such as \emph{IntServ} and
-\emph{DiffServ} \cite{cisco_diffserv_2005}, have not seen wide acceptance
-(yet?), largely due to the fact they require support from the physical
-infrastructure. On the other hand, an overlay network can be deployed on
-end-systems running the overlay protocol software, without cooperation from the
-ISPs.
-%Academic research includes
-%\cite{chu_esm_2000,jannotti_overcast_2000,kwon_tag_2002} among others.
+One of the fundamental issues that defines the efficiency of an overlay 
+network is how 
+well the overlay network topology maps to the underlying network topology on which it rests.
+Consider two nodes that share a link in an overlay network which maps to multiple 
+hops in the underlying IP topology.  If the traffic patterns of the application are such
+that traffic flows heavily over the overlay link between the two nodes, then it is
+beneficial to construct the overlay topology such that the number of underlying IP links
+between these two nodes is minimized.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Peer-to-Peer (P2P) Architectures in a Nutshell}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Overlay networks on top of the best-effort IP layer, formed a widely ranged
-family of protocols collectively known as \emph{peer-to-peer} or \emph{P2P} for
-short. P2P systems consist of a number of networked computers called
-\emph{peers}, where each, such, \emph{peer} could act both as a resource
-provider and a resource consumer. This changed the traditional
-\emph{client-server} model dominating the Internet and lead to the introduction
-of the \emph{servent} concept \cite{gnutella}, a \emph{portmanteau} that blends
-the notions of server and client to denote the twofold role of the P2P network
-participants. Such serverless systems, proved to be able to achieve outstanding
-aggregate resource capacities as more and more participants join the system
-without requiring additional expenditure for infrastructure\footnote{
-  Unfortunatelly, there has been observed mitigation of this self-scaling
-  property by the undesirable behaviour of participating users, usually answered
-  in popular file-sharing systems; from \emph{free-riding}
+If the overlay network is constructed such that the topology does not match the 
+underlying IP topology well, the resulting \emph{topology mismatch} 
+can result in
+two significant problems.  First, application performance (e.g., search query latency
+or jitter) can be adversely affected since application traffic must flow over a larger number
+of IP hops than if the overlay were constructed more wisely.  Second, studies have shown
+that P2P traffic contributes the largest portion of overall Internet traffic
+\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}, with
+some reporting that more than 60\% of the traffic is P2P~\cite{cachelogic,ipoque2007,ipoque2009}  
+and others reporting that P2P traffic is expected to reach almost 8 petabytes 
+per month by the end of 2012~\cite{multinteligence}.  This is a significant burden for Internet Service Providers (ISPs) who 
+must transfer this traffic to destination nodes sitting at the edge of the Internet.
+If the overlay topology of a peer-to-peer system is not cleverly constructed, then 
+the burden of ISPs carrying P2P traffic could be even greater.
+For example, traffic might have to
+flow back and forth several times between two neighboring ISPs as it travels from its
+source node in the overlay to its destination node.
+Thus, it is fundamentally important that P2P overlay networks are constructed
+and maintained so that
+their topology matches as closely as possible the underlying IP topology.
 
-\cite{saroiu_measurefileshare_2002,adar_gnutellafreeriders_2000,hughes_gnutellafreeride_2005}
-to the distribution of illegal content and/or other \emph{socio-technical}
-issues \cite{hughes_socp2p_2008}.
-}.
-P2p network systems gained extra attention whithin the distributed systems
-super-set, circa 2000, due to the fact that they succesfully supported popular
-applications that offered file sharing among vast numbers of participating
-users. But their development didn't stay still. On the contrary, during the last
-decade, research on P2P systems was constant and fruitfull generating a wide
-range of popular protocols, networks and applications. As P2P systems evolved
-three main architectures have been emerged, namely
-\begin{inparaenum}[\itshape i\upshape)]
-  \item \emph{centralized},
-  \item \emph{decentralized unstructured}, and
-  \item \emph{decentralized structured}.
-\end{inparaenum}
+For over a decade, several tens of researchers have published profusely on the 
+topology mismatch problem.  In this paper, we survey the literature
+on this topic and provide a taxonomy of the proposed approaches and solutions.
+We point out synergies
+as well as similarities and differences in the published approaches.  
+Ultimately, our goal is to help readers sift through the vast
+amount of literature, to help them understand the advantages and 
+disadvantages of each approach, and to provide them with enough 
+perspective so that when the need arises, they are able to select 
+amongst the different approaches for their 
+particular application.
 
-\emph{Centralized} architectures (see Figure~\ref{figure:p2p-archs:centralised})
-were the first to recognize that requests for resourses (e.g. CPU cycles
-\cite{seti} or popular content
-% TODO find citation for Napster
-\cite{}) need not be sent to a dedicated server but instead could be handled by
-the many hosts that already posses them. This schema, though, is not fully
-distributed since it dictates a centralised searching facility, a characteristic
-that can be proven its ``Achilles' heel'' in terms of being a scalability
-bottleneck, a central point of failure and vulnerable to malicious acts (e.g.
-\emph{Denial-of-Service (DoS) attacks}).
-
-The centralized approach has been eventually replaced by architectures that
-distributed both search and download capabilities (see
-Figures~\ref{figure:p2p-archs:unstructured-full} and
-\ref{figure:p2p-archs:unstructured-hierarchical}). In these \emph{decentralized}
-architectures, file placement is random, which means there is no correlation
-with the network topology whatsoever \cite{yang_improvep2psearch_2002}. For this
-reason they are also refered to as \emph{unstructured}. The most important
-properties of such systems are that they support inherent heterogeneity of
-peers, are highly resilient to peer failures, and incur low maintenance overhead
-at handling the dynamics of peer participation \cite{stutzbach_churn_2006}. In
-the literature, they are also known as \emph{broadcast-based} systems, because
-they use \emph{message flooding} among peers or super-peers to propagate
-queries.
-%\footnote{The life of each query message is constrained by an integer, refered
-%to as \emph{time-to-live (TTL)}. When a querry is received by a (super-) peer,
-%this value is decreased by one unit. Then, the receiver, back-propagates its
-%results and if the TTL has not reached zero the node forwards the query message
-%to all its neighbours except from the one it received it from.}
-
-Recently, \emph{decentralized structured} schemes have been proposed (see
-Figure~\ref{figure:p2p-archs:structured}) in order to provide a self-organising
-infrastructure for large-scale P2P applications
-\cite{ratnasamy_can_2001,stoica_chord_2001,antony_pastry_2001,zhao_tapestry_2001,maymounkov_kademlia_2002,rgrk_bamboo_2004}.
-They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes
-through a deterministic mechanism. Even though unstructured networks became
-highly popular among file sharing applications, they do have a certain
-limitation in locating rare objects. The DHT approach of structured networks on
-the other hand, provides a guaranteed bound on the number of overlay routing
-hops that have to be taken in order for an object to be located, even in the
-case where only a single copy exists in the system.  This means
-$O \left ( log n \right )$ hops, compared to unstructured networks that require
-$O \left ( n \right )$ to reliably locate a specific object\footnote{
-  Unfortunatelly this efficiency does not come without cost. Structured overlays
-  realise what is called object identifier location as opposed to content-based
-  retrieval of unstructured ones. This means that the later can support more
-  versatile object location mechanisms like keyword and partial matching
-  searches.}.
-
-% TODO: Change these pictures to match look-n-feel
-\begin{figure}[ht]
-\centering
-\subfigure[Centralised (SETI@Home or Napster).] {
-  \includegraphics[scale=0.3]{img/centralised.png}
-  \label{figure:p2p-archs:centralised}
-}\qquad\qquad
-\subfigure[Fully unstructured (Gnutella v.0.4).] {
-  \includegraphics[scale=0.3]{img/unstructured-full.png}
-  \label{figure:p2p-archs:unstructured-full}
-}\qquad\qquad
-\subfigure[Hierarchical unstructured (Gnutella v.0.6 or FastTrack).] {
-  \includegraphics[scale=0.3]{img/unstructured-hierarchical.png}
-  \label{figure:p2p-archs:unstructured-hierarchical}
-}\qquad\qquad
-\subfigure[Structured (Chord and Kademlia).] {
-  \includegraphics[scale=0.3]{img/structured.png}
-  \label{figure:p2p-archs:structured}
-}
-\caption{The various peer-to-peer architectures.}
-\label{figure:p2p-archs}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Applications and Implications}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The most successful incarnation of the centrilised approach is without a doubt
-the popular file-sharing application \emph{Napster}\footnote{Pioneering, though,
-was the famous \emph{SETI@Home Project} \cite{seti}, which pools spare
-processing power of participants to search the universe for the existence of
-extra terrestrial intelligence!}
-% TODO find citation for Napster again
-\cite{}.
-Napster maintained a \emph{central index server} based on file lists provided by
-participating peers. The central index server was queried by the users and it
-returned pointers to the actual content. Thus, by centralizing search while
-distributing downloads, Napster achieved a highly functional design that was
-widely acknowledged, at the time, as ``the fastest growing Internet application
-ever'' that toped some 26 million users and 80 million songs
-\cite{jmm_naptopusage_2001}.
-
-Napster was ultimatelly brought down after Recording Industry Association of
-America (RIAA) filed a lawsuit for copyright infringement of the intellectual
-property of artists whose music was distributed by the network\footnote{
-  The sueing spree was initiated by heavy metal giants \emph{Metallica} that
-  feared Napster was undermining record sales. Opposed opinions instantly
-  emerged with best example the british alternative rock band \emph{Radiohead}
-  whose album \emph{Kid A} had also been leaked to Napster three months before
-  reaching music store shelves \cite{rm_radioheadkida_2000}. Despite music
-  critics labeling the album ``experimental'' and ``challenging'' (it had no
-  official singles nor music videos for any of its songs), in October 2000 Kid A
-  captured the number one spot on the Billboard 200 sales chart in its debut
-  week. Unlike Metallica, Madonna and Dr.Dre that led litigation with Napster,
-  Radiohead had never hit the top20 in the US before.
-}
-\cite{wiki_amnaptrial_2001}. This triggered a shift towards decentralised
-networks like the classic \emph{Gnutella} and, its ``rival'' at the time,
-\emph{FastTrack}. Gnutella, in its first realisation (v.0.4 \cite{gnutellav04}),
-was a fully decentralised approach. Unfortunatelly, network resourses tended to
-be overwhelmed by message flooding and if queries were constrained to live for
-some (small) number of hops, this ultimatelly resulted to a shrinked search
-scope, leading to a death spiral that prevented the network from scaling. The
-solution came with a slightly more structured network by introducing the notions
-of \emph{ultra-peers} and \emph{leaf-nodes} (v.0.6 \cite{gnutella}). With
-\emph{Limewire}, the most widely used software that implemented the protocol,
-the network was highly successful and was reported to have reached connecting
-more that 3 million users concurrently by the end of 2006
-\cite{rsr_gnutellaevol_2006}. Fasttrack and its implementing software KaZaA
-\cite{kazaa}, on the other hand, went with the hierarchical approach
-(\emph{super-nodes} in their context) from the very begining, dynamically
-assigning indexing functions to special peers in order to increase search
-performace and scaling potential. FastTrack network had an average of 3 million
-concurrently connected nodes in 2003 \cite{lkxr_kazaameasure_2005}. It is
-estimated that the total number of users was greater than that of Napster at its
-peak. With Limewire discontinued/shutdown in 2010 and its successor,
-\emph{Frostwire}, moving to the Bittorent protocol in 2011, seem to very well
-signal the end of the once-mighty Gnutella network. KaZaA continues as a
-propriatery music store solution. In the meanwhile, Fastrack design team had
-already introduced a VoIP and text messaging application based on the design
-principles of KaZaA's hierarchical P2P nature, that was meant to change how
-people communicate around the world. This app was no other than Skype
-\cite{skype}. Despite its wide usage, little is known for the technical aspects
-of its overlay network, mainly due to its proprietary nature. Studies, though,
-confirm that it shares strong structural similarities to its file-sharing
-predecessor (ordinary hosts and super-nodes in Skype's terminology) but they
-significantly differentiate in terms of traffic behaviour
-\cite{g_voipskypesec_2005,gdj_skypestudy_2006}. Skype was introduced in 2003
-and from then on it witnessed frenzy-rythmed increase to its popularity, with
-more that 650 milion registered users in 2011 \cite{skypetotalusers} and
-reported to have broken the barrier of 35 million skypers concurrently online on
-the the 5th of March 2012 \cite{skypesymusers}. On October 2011 Microsoft Corp.
-and Skype Global S.à.r.l. anounnced the completion of Microsoft's acquisition of
-Skype for \$$8.5$ billion \cite{skypemicrosoft}.
-
-File sharing was the most popular application area for structured architectures
-as well. Overnet
-% TODO find citation for Overnet
-\cite{},
-for example, was created for sharing large files (e.g., movies and CD images).
-It implemented the Kademlia tree-like partitioning algorithm. Other
-implementations of Kademlia are the Kad Network (eMule/aMule  clients),
-BitTorrent for trackerless torrernts (official BitTorrent client, $\mu$Torrent,
-BitComet, Transmission and BitSpirit all share compatibility with the Mainline
-DHT which realises the Kademlia protocol) and the Gnutella DHT (first
-implemented by LimeWire as the Mojito DHT). Lately, RetroShare \cite{retroshare}
-is the new trend for file-sharing which incorporates an implementation of the
-Kademlia DHT to create a Friend-to-Friend (F2F) network with some attributes of
-a darknet as well. But the potential of structured architectures was not
-exhausted only to the above. Structured P2P networks became popular in
-distributed systems because they could scale smoothly and have fast and reliable
-object location mechanisms. Thus, research and development was very active and
-productive resulting in a variety of diverse applications such like distributed
-search engines \cite{yaci}, distributed data storage systems
-\cite{kbc_oceanstore_2000,bdet_fsdfs_2000,dkkms_cfs_2001,dr_pastutility_2001,abc_farsite_2002,mmfc_ivy_2002,arla,agebh_dks_2003}, Web caches, archives,
-and publishing systems \cite{ird_squirrel_2002,bags_youserv_2002,wrc_publius_2000,wm_tangler_2001},
-messaging applications \cite{threedegrees}, event notification infrastructures
-\cite{rkcd_scribe_2001,cdkr_scribe_2002,agebh_dks_2003}, naming services
-\cite{cmm_chorddns_2002}, censor-resistent stores \cite{cswh_freenet_2001} and
-lately even cloud-based platforms \cite{mgpj_cloudsnap_2011}.
-
-%%%%%%%% NEW TRENDS %%%%%%%%
-%File sharing Limewire Kazzaa The file-sharing landscape is slowly adjusting in
-%response to the continued push for more anti-piracy tools, the final Pirate Bay
-%verdict, and the raids and arrests in the Megaupload case, SOPA, ACTA. This
-%leads to users trying to find new solutions for anonymous, uncensored,
-%decentralised file sharing solutions like Freenet, Tribler and RetroShare where
-%connections are made only between trusted peers — sometimes called "friends"
-%and the respective formed network F2F.
-
-% TODO: READ Dalesa published at eAsia 2009, http://www.e-asia.org/2009/ET_Presentations/Wathsala.pdf
-
-The huge numbers of users, for the various applications we presented above,
-reveals the tremendous popularity P2P architectures have enjoyed. This, though,
-does not come without any toll. Measurements on some popular systems, such as
-FastTrack, Gnutella, DirectConnect and BitTorrent have shown that the P2P
-traffic contributes the largest portion of the overall Internet traffic
-\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}.
-Even with later studies, this observation still holds on. Using deep packet
-inspection techniques deployed at selected customer sites \emph{Ipoque} and
-\emph{CacheLogic} were able to track down application usage \cite{cachelogic,ipoque2007,ipoque2009}.
-CacheLogic reports that more than 60\% of the traffic is identified as P2P, with
-Ipoque verifying this observation reporting similar results for 2007, while
-observing some small decline for 2008/2009 (but still remaining dominant
-compared to other applications). Apart from its participation in overall usage
-terms, P2P is already a huge contibutor to the overall data exchanged across the
-Internet\footnote{
-  Mainly because file-sharing P2P is used to exchange big files of HD video or
-  high sampled audio content compared to simple Web content distributed via, for
-  example, HTTP.
-}.
-\cite{multinteligence} claims that P2P traffic in 2007 was around 1.6 petabytes
-per month and the potential is to be quintubled by the end of 2012 reaching
-almost 8 petabytes per month.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The Topology Mismatch Problem}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-One of the major issues that defines the efficiency of an overlay network is the
-mapping of the overlay structure to the underlying infrastructure. Remember, in
-Figure~\ref{figure:overlay} Nodes $A$ and $B$ are on the same local network,
-while $C$ and $D$ are in different networks. The top layer represents the
-overlay interconnection formed by these nodes at a higher level. Link
-connections there, can change as needed by the running environment, without any
-particular constraint by the underlying physical topology. Let's have a closer
-look with the help of a simple example. Assume nodes $A$, $B$, $C$ and $D$ are
-connected through the physical network shown in Figure~\ref{figure:phys}, where
-the network costs are given in milliseconds, and peer $A$ sends a message to
-peer $D$.
-
-\begin{figure}
-\centering
-  \includegraphics[scale=0.8]{img/phys.pdf}
-\caption{Interconnection of nodes in the physical level.}
-\label{figure:phys}
-\end{figure}
-
-If these peers participate in an overlay network according to one of the setups
-of Figure~\ref{figure:overlay-confs} then users will expierience different
-performances. In the overlay depicted in Figure~\ref{figure:overlay-1}, the
-message will traverse the following sequence of links in the physical layer
-(marked with their costs): $3 \rightarrow 7 \rightarrow 1 \rightarrow 2
-\rightarrow 2 \rightarrow 1 \rightarrow 7 \rightarrow 8 \rightarrow 5
-\rightarrow 5 \rightarrow 4$. In the alternative overlay of Figure
-~\ref{figure:overlay-2}, the path will be: $3 \rightarrow 8 \rightarrow 4$. The
-total cost of the first sequnce is $45 ms$, while the second yields a mere $15
-ms$. Therefore, we can conclude that the second overlay is more congruent with
-the underlying physical network than the first one and, thus, is more efficient.
-
-\begin{figure}[ht]
-\centering
-\subfigure[Nodes B and C have direct overlay connection.] {
-  \includegraphics[scale=0.8]{img/overlay-1.pdf}
-  \label{figure:overlay-1}
-}\qquad\qquad
-\subfigure[Nodes A and D have direct overlay connection.] {
-  \includegraphics[scale=0.8]{img/overlay-2.pdf}
-  \label{figure:overlay-2}
-}
-\caption{Two different overlay connection configuration.}
-\label{figure:overlay-confs}
-\end{figure}
-
-Ideally, overlay networks are expected to achieve an optimal mapping among the
-overlay paths and the underlying physical links, and avoid inefficient states.
-The problem of constructing an optimal overlay is referred to as the
-\emph{topology mismatch problem}, and formally defined as follows:
-\begin{definition}
-Let $V = \{v_1, ..., v_n\}$ be a set of points denoting the network nodes,
-$\{v_i, v_j\} \in E$ be the set of unicast distances between nodes $v_i$ and
-$v_j$, $G=(V,E)$ be a complete distance graph over $V$. The topology mismatch
-problem is to construct a minimal spanning tree,  where node degree is
-restricted to a constant ($k\geq 2$) by the bandwidth of each node $v_i$.
-\end{definition}
-
-Early incarnations of the overlay protocols, however, did not make use of the
-optimal mapping with the physical network topology. Gnutella protocol for
-example was considered far from scalable \cite{ritter_gnucantscale_2001}. Since
-every peer can chοse its logical neighbours without knowledge of the underlying
-network had as a result a mismatch between physical and application-level
-topologies. Additionally, queries may be flooded to multiple paths merging in
-one peer while in such case one of the paths would have been enough.
-Nevertheless, peers may forward the same message to each other before they
-receive the query messages from each other. \cite{matei_mapgnutella_2002} showed
-that, even given that 95\% of any two nodes are less than 7 hops away from each
-other, a flooding-based algorithm, like the one used by Gnutella, can generate
-330TB/month in a 50.000 node network! Therefore, P2P traffic can have a serious
-effect on the overall well-being of the Internet as a whole.
-
-Similar problems are observed in decentralized structured schemes also.
-Typically, node IDs are assigned \emph{randomly}, resulting in excellent load
-balancing, scalability and robustness. Unfortunately, this randomness has a
-negative impact on the \emph{routing locality} of the network. This means that
-even though the target node can be reached with logarithmic overlay hops, the
-distance traveled in the physical underlying network, during the overlay routing
-process, can be far from optimal.  The ratio of the physical distance, a query
-travels through the overlay to an object, and the minimal distance to that
-object (i.e. through IP) is known in the literature as \emph{stretch} or
-\emph{Relative Delay Penalty (RDP)} \cite{chu_esm_2000}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Motivation and our Goal for this Survey}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-A topology unaware overlay network is able to control the sequence of peers a
-message traverses before reaching its destination, but it is completely unaware
-of how the actual packets are switched at the underlying infrastructure along
-the overlay path. For example, a single logical point-to-point link on the
-overlay, most of the time, corresponds to multiple physical links in the
-underlying layer. Additionally, a link in the underlying network often serves
-the mapping of several overlay paths causing increase of the traffic on the
-physical link, which is also called the link's \emph{stress}\cite{chu_esm_2002}.
-Furthermore, the network's \emph{churn}, that is the stochastic behaviour of
-peers randomly joining and leaving the P2P network, all cause the overlay to
-physical mapping to create such unnecessary, redundant maintenance traffic that
-impacts the efficiency of the network in terms of average response time.
-
-The topology mismatch problem is exacerbated by the high amounts of traffic
-generated by the highly popular P2P applications. This creats the killer
-combination for the network infrastructure, ultimately effecting all Internet
-users. Solving it, on the other hand, is no easy task. Topology mismatch is
-known to be an NP-Hard problem \cite{NPBOOK,chawathe_scattercast_2000}.
-Moreover, the Internet itself is an interesting environment where end-to-end
-latencies demonstrate what is called \emph{triangle inequality violations
-(TIVs)}, which further complicate the problem. These delay space TIV is a
-consequence of the Internet's structure and routing policies and thus will
-remain a property of the Internet for the foreseeable future
-\cite{zheng_irprtt_2005}. TIVs affect network coordinate
-\cite{cox_vivaldi_2004,wong_meridian_2005} and positioning \cite{ng_gnp_2001}
-and makes the construction of location/delay aware overlays difficult. An
-extensive array of heuristic approaches have been proposed in an effort to find
-an algorithm that performs reasonably and consistently well. The purpose of this
-survey is to gather the extensive knowledge that has been produced recently in
-this research field and organise the available information ultimately helping
-the reader sift through the myriad of these efforts, understand the different
-approaches and identify though a taxonomy the advantages and disadvantages of
-each of them.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Survey Outline}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% TODO: Review this after the content and article structure is fixed.
-In the introduction we had the oportunity to recognise the importance of overly
-networks in contemporary application architectures. Also we quickly reviewed the
-basic trends in both overlay construction\footnote{
-  The main focus of this paper is the topology mismatch problem, therefore the
-  available approaches in overlay networks are briefly summarized, and
-  interested readers are pointed to the available surveys on overlay networks
-  \cite{TheotokisS04,LuaCPSL05}.
-}
-and application development. Last but not least, we identified and formaly
-defined the phenomenon of the topology mismatch problem and its impact on the
-utilization of network resources. The rest of the survey\ldots
-
-Sections~\ref{sec:unstructured} and \ref{sec:structured} discuss the recent
-academic work that has been conducted in the field.
-
-Ultimately, it reviews the most important resources found in the literature that
-attempt to tackle the problem as it proposes a taxonomic scheme, using a set of
-supertype-subtype relationships based on unique characteristics they have and/or
-specific goals they target. What follows does not claim to be a thorough
-citation of all known protocols that are available out there, but instead, a
-careful selection of those that left a distinctive fingerprint contribution in
-the efforts of the research community to alleviate the topology mismatch
-problem.
+The rest of the paper is as follows:  In Section~\ref{section:background}, 
+we provide some background information
+on peer-to-peer overlay architectures including centralized, decentralized unstructured
+and decentralized structured peer-to-peer systems.  We also present formally the
+problem of alleviating the topology mismatch between a peer-to-peer overlay and the
+underlying IP network.  In Sections~\ref{section:unstructured} and 
+\ref{section:structured}, we describe a taxonomy of
+the research efforts and solutions
+proposed for decentralized unstructured peer-to-peer systems 
+and decentralized
+structured peer-to-peer systems, respectively.  Finally, 
+in Section~\ref{section:conclusion}, we conclude.

### 2014-01-23 structured.tex
2012.07.12 (Vassilis) Merged introduction section. Sectioning according to the method an algorithm is using dropped (almost all use a combination of two or more). First attempt for the tabular data on an external file.

@@
-\section{Decentralized Structured Algorithms}
+\section{Decentralized Structured Architectures}
 \label{section:structured}
@@
 
-%{\sethlcolor{yellow}\hl{HA:  Talk about Tapestry, Pastry, CAN and Chord briefly here!}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Geographic Layout} \label{section:geographic_layout}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Inside geographic layout category fall those kind of methods which practically
-position physically close by nodes together in the application space as well.
-This is done by adjusting and maintaining the routing tables of all involved
-peers accordingly by exploiting proximity information that describe the overall
-geographic positioning of the peer nodes. One popular method for detecting and
-constructing a geographic view of the underlying network is the use of well
-established servers dedicated to landmarking the physical space, pretty much as
-discussed in Section \ref{sec:landmark} of landmark based proximity in the
-unstructured P2P networks. The same disadvantages of difficult and inaccurate
-positioning and the cost of deploying and maintaining these landmarks apply in
-this case as well.
-% TODO: maybe a discussion sentence bellow
-Even though managing the overlay structure based on the geographic layout of the
-nodes improves the query efficiency of the system, on the other hand, it tends
-to create hotspots, and the needed failure resilience is undermined by the fact
-that close by nodes are more likely to suffer collective failures.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% TODO: Geographic Layout
-%
-%\begin{itemize}
-% \item Geographic Layout\\
-%The node IDs are assigned in such a way that nodes close by in the physical
-%network topology, be close in the node ID space as well. Implementations that
-%work relatively well with this approach have been incorporated into CAN. Nodes
-%measure the RTT between themselves and a set of landmarks in order to match the
-%CAN space as much as possible to the physical one. Unfortunately, the approach
-%requires well known landmark servers and for that matter is  not fully
-%self-organizing which can further lead to imbalanced node distribution. On
-%other DHTs, such as Chord or Pastry, another problem emerges.  To gain fault
-%tolerant properties, these protocols, replicate key-value pairs on neighbouring
-%(in the ID space) nodes. When a proximity-based node ID assignment has been
-%used, the needed failure resilience is undermined by the fact that close by
-%nodes are more likely to suffer collective failures.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Global Soft-State}
+\subsection{Algorithms for Structured Architectures}
+
+%###############################################################################
+%###############################################################################
+%###############################################################################
+%       GEGRAPHIC LAYOUT
+%###############################################################################
+%###############################################################################
+%###############################################################################
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Global Soft-State}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Mithos}
+\subsubsection{Mithos}
 \emph{Mithos} \cite{waldvogel_mythos_2003} is a P2P protocol which incorporates
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf LAPTOP}
+\subsubsection{LAPTOP}
 \emph{LAPTOP} \cite{wu_laptop_2007}, organises the overlay into a tree-based
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Proximity Routing}
-\label{section:pr}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Proximity routing does not require routing tables to be built using any
-knowledge about network proximity. On the other hand it exploits such knowledge
-in order to choose the best next hop during routing a message. This approach
-balances between choosing the node that will further progress the routing
-towards the destination and choosing the closest entry in the routing table, in
-terms of network proximity.
-% TODO: looks like discussion material
-Thus, it is relatively less effective than
-geographical layout when applied to CAN(-like) implementations. Moreover, the
-technique has been incorporated into a version of Chord causing an increase on
-the overhead of node joins and the size as well as maintenance cost of finger
-tables.
-
-% Moreover, the technique has been incorporated into a version of Chord
-% (TODO: PChord | additionally check whether this is true) causing an increase
-% on the overhead of node joins and the size as well as maintenance cost of
-% finger tables.
-
-%
-%\cite{dabek_cfs_2001} proposes a server selection scheme for the Chord DHT, on
-%the domain of proximity routing selection. In \emph{CFS}, each node predicts
-%the entire lookup latency as a function of the total number of nodes and the
-%average overlay next routing peer. The problem is that it is very difficult to
-%have a clear picture on the total number of nodes and the average hop latency
-%from the local. This leads to rough estimations that consequently decreases
-%overall performance.
-%
-The methods that use proximity routing and their details are listed below.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Proximity in Kademlia}
+%###############################################################################
+%###############################################################################
+%###############################################################################
+%       PROXIMITY ROUTING
+%###############################################################################
+%###############################################################################
+%###############################################################################
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Proximity in Kademlia}
 % Before discussing {\em Proximity in Kademlia}, it would be helpful to briefly
@@
 {\em proximity routing} and the other on {\em proximity neighbour selection}
-(see Section~\ref{section:pns}) approaches overlay optimization techiques for
+approaches overlay optimization techiques for
 exploiting proximity. For the first, the goal is in choosing the best next hop
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf CHOord considering Proximity on IPv6 (CHOP6)}
+As already mentioned the paper
+proposes \emph{proximity neighbour selection} method, as well, for overlay
+optimisation. The goal, according to this, is to keep peers with the least
+contact cost in the routing table. As Kademlia constantly learns new peers
+(incoming requests, through iterative lookup) no special algorithm for
+searching more cost effective peers is necessary and simply choose the best
+peers seen so far.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{CHOord considering Proximity on IPv6 (CHOP6)}
 \emph{CHOP6} \cite{morimoto_chop6_2007} roughly estimates the proximity among
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf T2MC}
+\subsubsection{T2MC}
 T2MC \cite{shi_t2mc_2008} uses traceroute logs to detect different autonomous
@@
 class, the peer chooses the one with maximum hops attribute as its edge router,
-and registers it along with the all the near cluster into the DHT of the P@P
+and registers it along with the all the near cluster into the DHT of the P2P
 overlay. As new peers join the network, those that share the same edge router
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf PChord}
+\subsubsection{PChord}
 \emph{PChord}\cite{hong_pchord_2005} is a based on the Chord DHT which adds
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf AChord}
+\subsubsection{AChord}
 \emph{AChord} \cite{dao_achord_2006} uses IPv6's anycast mechanism in order to
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Chord6}
+\subsubsection{Chord6}
 \emph{Chord6} \cite{xiong_chord6_2005} is another Chord variant that tries to
@@
 % TODO: SOME DISCUSSION
-%       RECHECK VALIDITY
-%
 % However, even though nodes in the same domain would have close identifiers,
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Proximity Neighbour Selection}
-\label{section:pns}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\emph{Proximity Neighbour Selection} constructs the routing tables using
-proximity knowledge. The proximity information used in this method is different
-than the landmark based systems described in
-Section~\ref{section:geographic_layout}, as time-to-live values between nodes,
-or directly node ID prefixes are used to detect proximity. Tapestry, Pastry, and
-CAN successfully implemented the proximity into their algorithms by using this
-approach. The routing protocol in Pastry is based on longest node ID prefix
-matching, while CAN uses RTT values to detect close by nodes.
-\cite{castro_proximityp2p_2002} reports that proximity neighbour selection as an
-effective proximity based method. Proximity Neighbor Selection based algorithms
-are described in detail below.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Proximity in Kademlia}
-As already mentioned on the the corresponding paragraph when discussing
-proximity routing approaches (see Section~\ref{section:pr}), the paper
-proposes \emph{proximity neighbour selection} method, as well, for overlay
-optimisation. The goal, according to this, is to keep peers with the least
-contact cost in the routing table. As Kademlia constantly learns new peers
-(incoming requests, through iterative lookup) no special algorithm for
-searching more cost effective peers is necessary and simply choose the best
-peers seen so far. For more information on the Proximity in Kademlia please
-consult the paragraph in Section~\ref{section:pr}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf DHT-PNS}
+%###############################################################################
+%###############################################################################
+%###############################################################################
+%       PROXIMITY NEIGHBOUR SELECTION
+%###############################################################################
+%###############################################################################
+%###############################################################################
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{DHT-PNS}
 The work in \cite{hancong_pnsbased_2006} propose a proximity neighbour selection
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Quasi-Chord}
+\subsubsection{Quasi-Chord}
 The approach that is proposed by Sun and Zhang in \cite{sun_quasi_2008}
@@
 $1$-dimensional one, used in the last, third, step to build the Quasi-Chord
-circle according to the host's Cantor value. In the following paragraphs,
-these steps are furtherly discussed.
-
+circle according to the host's Cantor value.
+
+%In the following paragraphs, these steps are furtherly discussed.
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf IP-based Clustering (IPBC)}
+\subsubsection{IP-based Clustering (IPBC)}
 Proximity neighbour selection algorithms use probing and other techniques to
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Cone}
+\subsubsection{Cone}
 \emph{Cone} \cite{wang_cone_2007} extends Chord using proximity neighbour
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf DynaMo}
+\subsubsection{DynaMo}
 \emph{DynaMO} \cite{winter_dynamo_2004} tries to build an overlay that not only 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf SAT-Match: Self-Adaptive Topology Matching}
+\subsubsection{SAT-Match: Self-Adaptive Topology Matching}
 
@@
 %  furtherly improve the resulted overlay. For example when multiple nodes try
-% to jump simultaneously into a regeon, then the logical link brakes from one
+% to jump simultaneously into a region, then the logical link brakes from one
 % attempt may result in inacurate computation of the gain factor, for an other.
@@
 \emph{DAPS} is between structured and unstructured.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Discussion on the Algorithms for Structured Architectures}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Discussion on the Algrithms for Structured Architectures}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Geographic Layout} \label{section:geographic_layout}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% TODO: TAKEN FROM ALGORITHMS SUBSECTION. SHOULD BE MERGED WITH DISCUSSION
+
+Inside geographic layout category fall those kind of methods which practically
+position physically close by nodes together in the application space as well.
+This is done by adjusting and maintaining the routing tables of all involved
+peers accordingly by exploiting proximity information that describe the overall
+geographic positioning of the peer nodes. One popular method for detecting and
+constructing a geographic view of the underlying network is the use of well
+established servers dedicated to landmarking the physical space, pretty much as
+discussed in Section \ref{sec:landmark} of landmark based proximity in the
+unstructured P2P networks. The same disadvantages of difficult and inaccurate
+positioning and the cost of deploying and maintaining these landmarks apply in
+this case as well.
+% TODO: maybe a discussion sentence bellow
+Even though managing the overlay structure based on the geographic layout of the
+nodes improves the query efficiency of the system, on the other hand, it tends
+to create hotspots, and the needed failure resilience is undermined by the fact
+that close by nodes are more likely to suffer collective failures.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% TODO: Geographic Layout
+%
+%\begin{itemize}
+% \item Geographic Layout\\
+%The node IDs are assigned in such a way that nodes close by in the physical
+%network topology, be close in the node ID space as well. Implementations that
+%work relatively well with this approach have been incorporated into CAN. Nodes
+%measure the RTT between themselves and a set of landmarks in order to match the
+%CAN space as much as possible to the physical one. Unfortunately, the approach
+%requires well known landmark servers and for that matter is  not fully
+%self-organizing which can further lead to imbalanced node distribution. On
+%other DHTs, such as Chord or Pastry, another problem emerges.  To gain fault
+%tolerant properties, these protocols, replicate key-value pairs on neighbouring
+%(in the ID space) nodes. When a proximity-based node ID assignment has been
+%used, the needed failure resilience is undermined by the fact that close by
+%nodes are more likely to suffer collective failures.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Proximity Routing}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% TODO: TAKEN FROM ALGORITHMS SUBSECTION. SHOULD BE MERGED WITH DISCUSSION
+
+Proximity routing does not require routing tables to be built using any
+knowledge about network proximity. On the other hand it exploits such knowledge
+in order to choose the best next hop during routing a message. This approach
+balances between choosing the node that will further progress the routing
+towards the destination and choosing the closest entry in the routing table, in
+terms of network proximity.
+% TODO: looks like discussion material
+Thus, it is relatively less effective than
+geographical layout when applied to CAN(-like) implementations. Moreover, the
+technique has been incorporated into a version of Chord causing an increase on
+the overhead of node joins and the size as well as maintenance cost of finger
+tables.
+
+% Moreover, the technique has been incorporated into a version of Chord
+% (TODO: PChord | additionally check whether this is true) causing an increase
+% on the overhead of node joins and the size as well as maintenance cost of
+% finger tables.
+
+%
+%\cite{dabek_cfs_2001} proposes a server selection scheme for the Chord DHT, on
+%the domain of proximity routing selection. In \emph{CFS}, each node predicts
+%the entire lookup latency as a function of the total number of nodes and the
+%average overlay next routing peer. The problem is that it is very difficult to
+%have a clear picture on the total number of nodes and the average hop latency
+%from the local. This leads to rough estimations that consequently decreases
+%overall performance.
+%
+The methods that use proximity routing and their details are listed below.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Proximity Neighbour Selection}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% TODO: TAKEN FROM ALGORITHMS SUBSECTION. SHOULD BE MERGED WITH DISCUSSION
+
+\emph{Proximity Neighbour Selection} constructs the routing tables using
+proximity knowledge. The proximity information used in this method is different
+than the landmark based systems described in
+Section~\ref{section:geographic_layout}, as time-to-live values between nodes,
+or directly node ID prefixes are used to detect proximity. Tapestry, Pastry, and
+CAN successfully implemented the proximity into their algorithms by using this
+approach. The routing protocol in Pastry is based on longest node ID prefix
+matching, while CAN uses RTT values to detect close by nodes.
+\cite{castro_proximityp2p_2002} reports that proximity neighbour selection as an
+effective proximity based method. Proximity Neighbor Selection based algorithms
+are described in detail below.
 

### 2014-01-23 unstructured.tex
2012.07.12 (Vassilis) Merged introduction section. Sectioning according to the method an algorithm is using dropped (almost all use a combination of two or more). First attempt for the tabular data on an external file.

@@
-\section{Decentralized Unstructured Algorithms}
+\section{Decentralized Unstructured Architectures}
 \label{section:unstructured}
@@
 paragraphs,algorithms for tackling the topology mismatch problem on such
-networks are visited, analysed and categorised based on the use of the overlay
-structure, messaging approach for peer communication and techniques proposed for
-detecting proximity in order to optimise overlay topology. In short, the methods
-that are highlighted here are divided into
-\begin{inparaenum}[\itshape i\upshape)]
-  \item broadcast-based,
-  \item cache-based,
-  \item overlay optimisation-based, and
-  \item landmark-based
-\end{inparaenum}
- approaches.
+networks are discussed, analysed and categorised based on the use of the
+overlay structure, message forwarding approach for peer communication and
+techniques proposed for detecting proximity in order to optimise overlay
+topology.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Broadcast Based}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In the first realisation of the Gnutella protocol, each query request received
-by a peer is then forwarded to all its neighbours, clearly generating
-unnecessary traffic over the network. The algorithms highlighted in this
-paragraph are referred to as broadcast-based because they enhance the "naive"
-broadcasting approach described above and prefer forwarding messages to a
-selected subset of a node's outgoing links. The selection criteria vary
-depending on the algorithm but they mainly use one or multiple statistical
-metrics that span from the obvious ones, like the latency of the link between
-the nodes, to more sophisticated alternatives (that may even take into account
-application level requirements) such as the reliability of the neighbour. The
-forwarding-based approach has the advantage of enhancing search efficiency but
-on the other hand suffer drawbacks like the drastic reduction of the search
-scope because of the restrictions of the broadcasting policies as well as
-limitations when scaling to large networks.
-
-%Expanding the search scope, on the other hand, is no easy task
-%because the overhead of forming multicast trees is proportional to the
-%multicast group size.
-%forwarding based schemes do not consider dynamic joining
-%and leaving of peers so they do not scale well on dynamic environments.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Improving search in peer-to-peer networks}
+\subsection{Algorithms for Unstructured Architectures}
+
+%###############################################################################
+%###############################################################################
+%       BROADCAST OPTIMISATION
+%###############################################################################
+%###############################################################################
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Improving search in peer-to-peer networks}
 
@@
 \cite{yang_improvep2psearch_2002} proposes an easy to implement and practical
-solution to the inefficiency problem caused by blind flooding in Gnutella-like
-P2P file sharing protocols. The paper replaces blind flooding with three
-approaches, namely \textit{iterative deepening}, \textit{directed BFS}, and
-\textit{local indices}. In \textit{iterative deepening}, as shown in
-Figure~\ref{figure:dbfs:iterdeep}, the search is performed on a BFS tree with
-multiple preset depths. The depth limit is iteratively increased by the source
-node for each query based on the quality of the results. The source node, after
-examining the results, may issue a new request by increasing the depth limit,
-which will trigger the nodes at the last depth level to resume the search. The
-iterative approach avoids to start the search from scratch at each iteration and
-reduces the load of the nodes on the upper levels of the tree. The major
-drawback of \textit{iterative deepening} is the delay between successive
-iterations, as the source node needs to examine the results at each iteration
-before deciding to quit or resume the query.  The \textit{directed BFS}, as
-shown in Figure~\ref{figure:dbfs:dbfs}, tries to avoid this delay by
-forwarding the query messages only to a selected set of neighbours, in which the
-selection criteria varies from the number of results received previously,
-distance in terms of hops, bandwidth, or the query load of the neighbour. In
-\textit{local indices}, as shown in Figure~\ref{figure:dbfs:localindx}, each
-node sustains a local data index of the nodes within a radius of $r$ hops from
-itself and uses this local index to remotely query the neighbour nodes without
-generating additional traffic. \textit{Local indices} greatly reduces the
-aggregate bandwidth usage of the network and improves query efficiency, however,
-updating the indices in cases with frequent node joins and leaves introduces a
-serious overhead to the system if the radius is kept broad.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Gia}
+solution to the inefficiency of Gnutella's ``blind flooding'' approach. The
+paper introduces three different forwarding methods, namely \emph{iterative
+deepening}, \emph{directed BFS}, and \emph{local indices}.
+
+In iterative deepening (see Figure~\ref{figure:dbfs:iterdeep}), the search is
+performed on a BFS tree with multiple preset depths. The depth limit is
+iteratively increased by the source node for each query, based on the quality of
+the results. The source node may issue a new request with increased depth limit
+that will trigger the nodes at the last depth level to resume the search. The
+iterative approach avoids restarting the whole search process from scratch at
+each iteration and reduces the load of the nodes on the upper levels of the
+tree. Its major drawback is the delay between successive iterations, as the
+source node needs to examine the results at each iteration before deciding to
+quit or resume the query.
+
+The directed BFS, as shown in Figure~\ref{figure:dbfs:dbfs}, tries to avoid this
+delay by forwarding the query messages only to a selected set of neighbours, in
+which the selection criteria varies from the number of results received
+previously, distance in terms of hops, bandwidth, or the query load of the
+neighbour.
+
+% TODO: local indices is a cache based algorithm so maybe we have to put a
+% reference to the respective section???
+In local indices, as shown in Figure~\ref{figure:dbfs:localindx}, each node
+indexes data within a radius of $r$ hops and uses this local index to answer
+queries on behalf of them without generating additional traffic. Local indices
+greatly reduce the aggregate bandwidth usage of the network and improves query
+efficiency, however, updating the indices in cases with frequent node joins and
+leaves introduces a serious overhead to the system if the radius is kept broad.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Gia}
 \emph{Gia} \cite{chawathe_gia_2003} is an effort to solve the scalability
@@
 
+% TODO: from the following (commented out) design components of Gia has one
+% part that falls into topology adaptation, one part on forwarding optimisation
+% and one part on
+
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Distributed Cycle Minimization Protocol}
+\subsubsection{Distributed Cycle Minimization Protocol (DCMP)}
 We have already discussed that blind flooding mechanism broadcasts messages with
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Cache Based}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\subsection{Cache Based}
-%
-%Caching based protocols are effectively used to reduce traffic costs and
-%response
-%times. The caching policy varies depending on the way protocol handles the
-%index and the
-%content. Centralized P2P systems
-%use central index servers, while local caching systems, such as KazaA, use
-%super peers
-%to cache indices in a distributed way. Content caching is also possible in P2P
-%systems, where nodes cache the forwarded content for further retrievals.
-%Although caching has the above mentioned advantages,  duplication
-%of messages still exist, which limits the scalability of these approaches.
-%Therefore, cache based approaches are analyzed in the following categories:
-%  \begin{itemize}
-%    \item \emph{data index caching},
-%    \item \emph{content index caching},
-%    \item \emph{centralized}, and
-%    \item \emph{local}.
-%  \end{itemize}
-%
-Caching is a widely used technique to exploit locality and minimise redundant
-transfer of data, initially adopted by web and file server application
-environments with great success. Since peers in a P2P system also work as
-servers, it is intuitively expected that P2P file sharing systems can also
-benefit from caching. However, two of the main characteristic properties of P2P
-systems, nodes frequently joining and leaving and query lifetime relatively
-short (at least compared to those needed to be served by web servers) , makes
-caching in P2P networks a non-trivial issue. There are usually two levels of
-caching possible in file sharing systems, caching indices or pointers to data or
-caching the data itself. Caching is already implemented and successfully used in
-some commercial P2P systems. One popular example is KazaA, which uses caching of
-indices on its super-peer level. In the paragraphs below, caching-based
-algorithms are presented and explained in more depth.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Replication Strategies in Unstructured Peer-to-Peer Networks}
+%###############################################################################
+%###############################################################################
+%###############################################################################
+%       CACHING
+%###############################################################################
+%###############################################################################
+%###############################################################################
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Replication Strategies in Unstructured Peer-to-Peer Networks}
 \cite{Cohen02} aim to improve the inefficient blind search algorithm by
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Tracing a large-scale Peer to Peer System: an hour in the life
+\subsubsection{Tracing a large-scale Peer to Peer System: an hour in the
+life
 of Gnutella}
+% TODO: This looks better as an analysis for the unstructured networks in
+% general and specifically for caching, so it doesn't seem to contribute any
+% new approach.... At least we do not present anything here. Maybe we need to
+% revisit the paper itself once more
 \cite{Markatos02} analyses Gnutella network traffic traces and by concluding
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Overlay Optimization Based}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% These approaches include creating spanning trees using
-%connection graphs, creating cluster of physically close nodes, or using latency
-%information to detect proximity. The brief description of each category is
-%presented below:
-%
-%  \begin{itemize}
-%    \item \emph{Spanning tree based}. These approaches construct of a rich
-%    graph  based on the network connections and build minimum spanning tree
-%    on the graphs, causing large traffic overhead to the system
-%    \cite{chu_esm_2000,chu_esm_2002}.
-%
-%    \item \emph{Cluster based}. These approaches select to link physically
-%    closer nodes with each other, therefore shrink the search scope
-%    significantly while mapping accuracy is not always guaranteed.
-%
-%    \item \emph{Minimum latency first}. Use of latency as a metric to calculate
-%    distance among peers. They require global latency information
-%    ``landmarks''\footnote{Measuring latency between peers and stable Internet
-%    servers.}.
-%
-%  \end{itemize}
-
-The overlay optimization based protocols modify the topology of the P2P network
-using various techniques. The two commonly used approach that is investigated in
-this section are creating spanning trees using connection graphs, and creating
-clusters of physically close nodes. Spanning tree based approaches construct
-rich graphs based on the network connections and build minimum spanning trees on
-these graphs. Even though spanning trees provides efficient query performance,
-the construction and the update costs of the spanning tree, especially in
-dynamic environments with many nodes joining and leaving the network, cause
-large traffic overhead to the corresponding
-network \cite{chu_esm_2000,chu_esm_2002}. The cluster based approaches on the
-other hand select to link physically closer nodes with each other, therefore
-significantly shrinking the search scope of the system. Added to that, commonly
-used methods for correct determination of physical node positions across
-Internet do not always return reliable results, therefore mapping accuracy is
-not guaranteed. Bellow, the details of the algorithms that use the spanning tree
-based or cluster based topology optimization methods are presented.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Narada}
+%###############################################################################
+%###############################################################################
+%###############################################################################
+%       OVERLAY OPTIMISATION
+%###############################################################################
+%###############################################################################
+%###############################################################################
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Narada}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Adaptive Overlay Topology Optimization and Adaptive Connection
-Establishment}
+\subsubsection{Adaptive Overlay Topology Optimization (AOTO)}
 \emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{liu_aoto_2003} is one
@@
 %       COMPARISON OF AOTO AND LTM
+% TODO: EDIT: LTM SEEMS A MISTAKE HERE BETTER NOT USE IT FOR DISCUSSION (THIS
+%             IS AOTO)
 %LTM's SF effectiveness has be proven to be detached from the different physical
@@
 
-%
-% TODO: REMAKE AAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAA!!!!!!!!!!!!!!!!!!
-%
- In
-\emph{Adaptive Connection Establishment (ACE)} \cite{liu_ace_2004}, the authors
-extend the idea by introducing optimizations based on the depths of the minimum
-spanning trees. But since the network delay is not always a reliable estimation
-method to detect the physical location of peers, the algorithm still suffers
-from the discrepancies caused by mislocated nodes.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Location-aware Topology Matching}
-\emph{Location-aware Topology Matching (LTM)} \cite{liu_ltm_2004} proposes a
-method to optimize the overlay structure of the P2P network based on the
-physical topology. To achieve this, peers issue special messages called
-\textit{TTL-detector}s with TTL values of 0 and 1 so that the receiving peers
-discover one- and two-hop neighbour sets ($N$ and $N^2$ respectively) around
-them and calculating communication costs. Calculation is achieved through
-timestamp checking, so clocks must be synchronised. The latency information
-gathered is later used to evolve the overlay network into a more efficient one,
-without reducing the search scope. Each node examines the latency information
-among his direct neighbours and those with larger latency are put in to a
-will-cut list where they remain for a certain period of time until when they are
-ultimatelly cut and recorded to the peer's cut list. Thus, low productive
-connections are dropped and replaced by more efficient ones, reducing the
-latency on the overall network. Although LTM improves the overall efficiency of
-the P2P network, it does not actually use real physical topology information,
-just latency metrics, therefore it does not offer a guaranteed safety net for
-the topology mismatch problem.
-
-%
-% TODO: SOME DISCUSSION
-%
-%\paragraph*{Low productive connection cutting}
-%There are three cases for any peer $P$ who receives $d(i, S, v)$ multiple
-% times:
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item $P$ receives both $d(i, S, 1)$ and $d(i, S, 0)$
-%  \item $P$ receives multiple $d(i, S, 0)$s from different paths, and P
-% randomly chooses to process one
-%  \item $P$ receives one $d(i, S, 1)$ and multiple $d(i, S, 0)$s, and $P$
-% processes $d(i, S, 1)$ and one randomly selected $d(i, S, 0)$
-%\end{inparaenum}
-%If the link with the largest cost is found and is a direct neighbour then the
-% connection is put in a will-cut list and stays there for a certain period of
-%time. If it is not, then it is handled by other peers. After that period,
-%connections are cut and recorded to $P$'s cut-list.
-%
-%\paragraph*{Source probing}
-%For a peer $P\in(N^2(S) - N(S))$ who receives only one $d(i, S, 0)$, the cost
-% of $PS$ is obtained (with list look-up or probing). Then $P$ compares it with
-%the cost from each hop and if $PS$ has the largest cost, $P$ will not keep this
-%connection, while otherwise the connection will be created.
-%
-%\paragraph{}
-%Supposing $n$ is the number of peers, $c_n$ is the average number of neighbours
-% and $c_e$ is the average cost of logical links, then in the flooding-based
-%search the traffic incurred by one query from an arbitrary peer in a
-%peer-to-peer network is $O(n)$. As observed in the Gnutella network
-%\cite{sripanidkulchai_gnutella_2001}, each peer issues $0.3$ queries per minute
-%in average, thus the per minute traffic incurred by the network with $n$ peers
-%is $O(n^2)$. Because each $d(i, S, v)$ has a TTL of $2$ in each source peer,
-%the traffic for one time LTM optimization in all peers is at most $2nc_n^2c_e$.
-%If each peer uses LTM $k$ times per minute, the total traffic incurred is
-%$2knc_n^2c_e$. Simulation shows the best value for $k$ is $2$ or $3$. So, the
-%traffic overhead caused by LTM to the network is $O(n)$.
-%
-%TTLj-detectors, with $j > 2$, would detect and break cycles with more than 4
-% links. LTM though, does not use such detectors because detector-flood traffic
-%would increase significantly, and cut links between two end-peers, could cause
-%queries initiated by them to traverse a path much more expensive than the cost
-%on the the cut link.
-%
-%\paragraph{}
-%LTM disadvantages are
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item disagreement of measured delay due to unsynchronized clocks causes
-% problems when deciding the cut positions, which can influence the network
-%connectivity, and
-%  \item the network delay metric mainly focuses on disabling the connections
-% between peers physically far away without considering the shortcuts created by
-%powerful peers.
-%\end{inparaenum}
-%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Scalable Bipartite Overlay}
-\emph{Scalable Bipartite Overlay (SBO)}
-\cite{liu_bipartite_IPDPS,liu_bipartite_2007} reduces the overhead of
-creating and maintaining a minimum spanning tree cost by randomly dividing the
-nodes into two groups, namely red and white and assigning different tasks to
-different groups. When a peer joins the network, is randomly assigned with an
-initial colour, say red or white (separating all peers into two groups). Then
-the peer that plays the role of the bbootstrap host provides the joining peer
-with a list of active peers along with their colour information. The later uses
-this list to establish connections to differently coloured peers. This way, all
-peers form a bipartite overlay. The white peers measure distances to
-neighbours (red only) by using the network delay as a metric and report their
-corresponding red neighbours. The red peers, equipped with the information of
-all two-hop neighbours ($N^2$), creates minimum spanning tree of these
-neighbours and assigns efficient forwarding paths. Having a minimum spanning
-tree with two hops, a red peer is able to send its queries within that range.
-Some white peers, though, have become non-forwarding neighbours. In this phase,
-such a (white) neighbour will try to find another red peer being two hops away
-from its current red neighbour to replace the later as its new neighbour. The
-white peers can further optimize their positions within the overlay if this is
-required.
-
-%
-% TODO: SOME DISCUSSION
-%
-%In a static enviromnent LTM may reduce traffic cost by around 80 to 85 percent
-% while SBO reduces traffic cost between 85 and 90 percent. However, LTM  is
-%proved to converge in around 2-3 steps while SBO needs 4-5 steps. Moreover LTM
-%reduces response time by more than 60 percent in 3 steps while SBO needs 8. In
-%a dynamic environment (10 minute average peer lifetime, 0.3 queries/sec by each
-%peer) SBO and LTM reduce the average traffic cost per query (including the
-%overhead due to the optimization steps) by 85 and 80 percent, respectively.
-%Moreover LTM reduces the response time per query to 30 percent while SBO to 35
-%percent.
-%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Adaptive Connection Establishment (ACE)}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Adaptive Connection Establishment (ACE)}
 \emph{Adaptive Connection Establishment (ACE)} \cite{liu_ace_2004} builds an
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Hops Adaptive Neighbour Discovery (HAND)}
-\cite{chen_hand_2006} proposes the \emph{HAND} algorithm which uses a
-fully distributed triple hop adjustment strategy to address the topology
-mismatch problem. The approach's ultimate goal is to shape the current overlay
-graph $G$ into the \emph{Logical Communication Network (LCN)} $G^{*}$ which is
-how the optimal overlay is refered in the paper's context. A formal
-definition of achieving the above match is only if all peer hop sequences $(v_1,
-v_2, \ldots, v_k)$ in $G$ exist in $G^{*}$ and in the same order\footnote{In
-practice triple sequences $(v_1, v_2, v_3)$ are used.}. The mismatching
-detection is done in the following way. Suppose we want to verify peer a
-sequence, say $v_2-v_1-v_3$. A pair of probing messages are sent from $v_1$ to
-$v_2$ and $v_3$. Suppose delays of $(v_1,v_2)$ and $(v_1,v_3)$ are are denoted
-as $x$ and $z$, respectively. When the probing message arrives to $v_2$ it
-forwards it directly to $v_3$. Similarly, when the probing message arrives to
-$v_3$, it forwards it directly to $v_2$. These last steps are performed in order
-to obtain delays of $(v_2,v_3)$ and $(v_3,v_2)$ physical paths, respectively,
-denoted by $y$. If $y=z-x\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched
-and should be adjusted to $v_1-v_2-v_3$ by deleting edge $(v_1,v_3)$ and adding
-a new $(v_2,v_3)$. If $y=x-z\pm\varepsilon$, sequence $v_2-v_1-v_3$ is
-mismatched and should be adjusted to $v_1-v_3-v_2$ by deleting edge $(v_1,v_2)$
-and adding a new $(v_3,v_2)$, where $\varepsilon$ is a small positive real
-number denoting additional delays caused by possible forwarding and jitter
-delays. The advantages of the HAND algorithm compared to other approaches are
-that:
-\begin{inparaenum}[\itshape i\upshape)]
-  \item it does not need any clock synchronization,
-  \item it is a fully distributed algorithm making it robust and reliable in
-        decentralized systems,
-  \item the traffic overhead of the triple hop adjustment is very low,
-  \item it is applicable to dynamic peer-to-peer environments, and
-  \item maintains lower query response time.
-\end{inparaenum}
+%In \emph{Adaptive Connection Establishment (ACE)} \cite{liu_ace_2004}, the
+%authors extend the idea of AOTO by introducing optimizations based on the
+%depths of the minimum spanning trees. But since the network delay is not
+%always a reliable estimation method to detect the physical location of peers,
+%the algorithm still suffers from the discrepancies caused by mislocated nodes.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Location-aware Topology Matching (LTM)}
+\emph{Location-aware Topology Matching (LTM)} \cite{liu_ltm_2004} proposes a
+method to optimize the overlay structure of the P2P network based on the
+physical topology. To achieve this, peers issue special messages called
+\textit{TTL-detector}s with TTL values of 0 and 1 so that the receiving peers
+discover one- and two-hop neighbour sets ($N$ and $N^2$ respectively) around
+them and calculating communication costs. Calculation is achieved through
+timestamp checking, so clocks must be synchronised. The latency information
+gathered is later used to evolve the overlay network into a more efficient one,
+without reducing the search scope. Each node examines the latency information
+among his direct neighbours and those with larger latency are put in to a
+will-cut list where they remain for a certain period of time until when they are
+ultimatelly cut and recorded to the peer's cut list. Thus, low productive
+connections are dropped and replaced by more efficient ones, reducing the
+latency on the overall network. Although LTM improves the overall efficiency of
+the P2P network, it does not actually use real physical topology information,
+just latency metrics, therefore it does not offer a guaranteed safety net for
+the topology mismatch problem.
 
@@
 %
-%Measurements conducted for evaluation perposes showed that in a static
-% environment the algorithm can effectively decrease traffic cost by about 77
-%percent and shorten the query response time by about 49 percent in less than
-%two minutes. In a dynamic environment it shows similar behaviour and with the
-%size of the overlay network having little impact on the effectiveness of the
-%algorithm. Compared to LTM both algorithms have almost the same traffic
-%reduction rate, however on the response time reduction rate HAND has a higher
-%one by about 4 percent. The traffic overhead of HAND is much less than that of
-%LTM by an average of 55 percent.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% TODO: DOUBLE CHECK IF THIS ALGORITHM GOES IN THIS SECTION/SUBSECTION
-\paragraph*{\bf Adaptive Peer Selection}
-Bernstein et al., contributed on the use of machine learning for building peer
-selection strategies from past experience \cite{bflz_adaptpeersel_2003}. A
-decision tree is used to rate peers based on information about connection
-characteristics that is collected. These can be load, bandwidth, and past
-uploading experience. Then a by Markov decision process is incorporated as a
-mechanism to shape the policy for switching among the peers. The problem
-with this approach is that the peer selection algorithm is slow due to the
-learning process and the complexity of the method.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Innocuous Topology Aware (ITA) Overlay Construction}
-\cite{prfm_ita_2009} intoduces an algorithm for Innocuous Topology Aware
-construction of unstructured P2P networks. The paper's proposition is twoforld.
-Overlay construction and search method. For the overlay construction it employs
-the notion of \emph{short} and \emph{long} connections. Assuming $N$ is the
-number of all network participants, $\alpha \leq 1 $ a system-wide magic number
-and $x$ an $\alpha$-related latency threashold, the bootstrapping peer randomly
-selects $\alpha \ast N$ close (latency bellow $x$) and
-$\left( 1 - \alpha \right) \ast N$ distant (latency above $x$) nodes as its
-neighbours. The search method has two steps. First, the quering node initiates a
-flood to its distant neighbourhood with $TTL = 1$. Then, peers that receive a
-query over a long link, start a local flood with $TTL = ttl$, were $ttl$ is a
-system defined parameter.
-
-The goal of the overlay construction is to create a network that exposes low
-clustering, a beneficial characteristic of random graphs which can lead to a
-larger coverage of peers with the same number of messages and reduced
-duplication. This means more efficient information lookup that exposes low or no
-negative impact whatsoever to the rest of the mechanisms employed in
-unstructured P2P systems\footnote{For example the 1-hop replication and the
-dynamic querying mechanisms of later versions of Gnutella.}. The paper reports
-50\% reduction in communication latency ammong peers by cutting off some 20\% of
-the IP-level message traffic generation.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf EGOIST}
-EGOIST \cite{egoist_2008} is a set of algorithms to construct and manage overlay
-networks. It utilises a selfish approach in the sense that every participating
-peer continuously updates its neigbours so as to minimise the sum of distances
-to all destinations under shortest-path routing. In EGOIST, a newlly arriving
-peer, randomly connects to an already participating node through a bootstrap
-server. Being connected, it starts receiving info throught the link-state
-mechanism and thus, after some time, it has a complete picture of the overlay
-graph. Then it estimates the delay to all other nodes in order to determine its
-potential neighbours and ultimatelly connect to the overlay using some
-policy\footnote{For example, minimisation of the average delay to all its
-neighbours.}. It is obvious that there is extensive resource usage for updating
-the wiring of the all nodes in the system. The authors claim that they can
-reduce the load imposed by the monitoring and announcing process from $O(n^2)$
-to only $O(kn)$, where $n$ is the number of all nodes in the network and $k$ is
-the number of links a node chooses to establish.
-
-% TODO: CHECK IF THE FOLLOWING BITTORRENT ALGORITMS BELONG IN THIS SUBSECTION
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Biased Neighbour Selection}
-Bindal et al. \cite{rpc_bitbiased_2006} claim to strengthen BitTorrent
-protocol\cite{c_bittorrent_2003} locality by selecting most of a peer's
-neighbours to be out of the same ISP, while preserving the near optimal download
-performance of the protocol itself. BitTorrent systems employ most of the time
-mechanisms that can be proven very aggresive to an ISP's networking and
-accounting, basically, due to their lack of knowlegde of AS boarder limits.
-BitTorrent specification, by default allows for each peer, 35 connections.
-Biased neighbour selection experiments on some number $k$ which denotes a number
-of neighbouring peers which are not on the same ISP than the peer at hand.
-Consequently, the remaining $35 - k$ peers are from the same ISP. These $k$
-nodes are used in order to preserve a more extended, global view of a network,
-but in the same localize the load within the limits of a single ISP. This
-prevents clients from exchanging traffic through an expensive transit link, if
-there are alternative local connections that could offer the same, faster and at
-no additional cost for the ISP. Implementation can either be done through
-modifications on the tracker side, to identify ISP locallity, or through P2P
-traffic shaping devices installed on ISP's edge routers. The paper also proposes
-combination of biased neighbour selection along with bandwidth throttling and
-some caching approach for near optimal results. Unfortunatelly, implementation
-need, in some extend, contribution (exposure of ISP AS mappings) or
-infrastructure changes (installation of P2P traffic shaping devices) on the ISP
-level itself, making the adoption difficult in the general case.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Ono}
-Choffnes and Bustamante propose Ono \cite{cb_ono_2008}, a protocol for managing
-BitTorrent traffic so that it significantly reduces cross-ISP traffic and in
-the same time enhances downloading rates by favouring connections within the
-borders of a single autonomous system. Contrary to biased peer selection
-proposed in \cite{rpc_bitbiased_2006}, this work can lead to improved
-performance with no cooperation between ISPs and their subscribers, no
-additional infrastructure and no network topology information. The selection
-approach is landmark-based and leverages existing CDN infrastructure for peer
-distance estimation. CDNs already use both static (i.e. geographical) and
-dynamic (network measurement systems) information for their replica selection,
-so the authors claim that peers which exhibit similar redirection behaviour, are
-very likely they are close to the replica servers and as a consequence to each
-other. The redirection behaviour is modeled in terms of \emph{ratio map} in the
-article's parlance, a vector of (replica-server,ratio) tuples, where ratio is
-the percentage of times CDN redirects the peer to the specific replica-server.
-The bootstrapping phase consists of the peer performing DNS lookup to CDN names
-in order to build its redirection information. The interval for polling DNSs
-starts from 30 seconds and increases by one minute every time redirection info
-to the CDN is found to remain unchanged after a lookup. On the other hand, the
-interval is halved whenever is found to have been changed. In order not to avoid
-the bootstrapping phase if the ratio map is sufficiently fresh, the protocol
-persists the info after the end of a BitTorrent session.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Locality-Awareness in BitTorrent-like P2P Applications}
-In \cite{lclx_bitlocal_2009} the authors study and compare three different
-approaches in injecting locality awareness in BitTorrent-like applications. The
-first acts at a \emph{macroscopic-level} and targets neighbour selection. When a
-peer asks the tracker to join, the later sorts the swarm peers according to
-their distance to the requesting peer in AS hop count and send it the first
-(e.g. 50) peers in the list. The second manipulates the chocking/unchocking
-BitTorrent mechanisms at an \emph{intermediate-level}. A peer unchockes its 4
-closest in terms of AS hop count interested neighbours. The same applies also
-when the peer turns to the seeding state\footnote{Thus this approach favours
-least distance, contrary to the original BitTorrent implementation which favours
-uploading speed.}. On a \emph{microscopic-level}, the rare-first chunk picking
-policy is substituted by the locality-fist policy. The distance value of a piece
-is calculated as the mean value of the distances of the peers that posses it.
-The paper conducts several measurements on the efficiency of the above scheme to
-reach the conclusion that the major design desision for all locality aware P2P
-systems is the trade-off of optimising inter-AS traffic, which their solution
-achieves, and of the fairness among peers, a section that their locality-based
-approach does not do well in comparison to the random approach of the standard
-BitTorrent protocol.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf TopBT}
-\cite{rtl_topbt_2010} proposed an approach for enhancing proximity awareness
-in the BitTorrent protocol without any need of any infrastructure installed. It
-suggests that a good peer selection metric should take into account both the
-downloading speed and the network topology. Thus, it proposes the use of the
-ration of download rate to upload rate, $\frac{d}{u}$, divided by link-level or
-AS-level routing hops, $l$ or $a$ respectively, in an attempt to form a
-comprehensive metric to select peers with high download rate, low reciprocal
-upload demand and low routing hops. The TopBT metric is applied on serveral
-parts of the original BitTorrent protocol in order to inject topology awareness
-for better peer selection; more specifically on the peer list the tracker
-returns when someone first tries to connect to the swarm, on the initial
-connection establishment on the connection replacement and of course the
-unchoking mechanism.
-
-%A peer that runs the TopBT protocol evaluates its neighbours by periodicaly
-%emitting pings or traceroutes in order to uchock those peer that exhibit less
-%hops to reach and higher download rates. \emph{Link-hops} are measured by using
-%the TTL value that the originator receives as a response from the remote
-%peer\footnote{Initial TTL values are known for the different operating systems
-%(e.g. 64 for Linux or 128 for Windows NT/2000/XP) so the originator can
-%calculate the hops using the value of the TTL on the response message it
-%receives.}. Also the peer, when offline, builds a table that maps IPs to
-%\emph{AS-hops}, using BGP dumps.
-
-Through experimentation on hundreds of PlanetLab as well as residential hosts,
-the authors claim more than 25\% traffic reduction and 15\% faster downloads,
-both with a lightweight cost when compared to popular BitTorrent
-implementations.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Underlying Topology-Aware Peer Selection (UTAPS)}
-UTAPS \cite{lcy_utaps_2008} is another peer selection strategy for the
-BitTorrent protocol. Its first step is to collect the needed information in
-order to construct a picture of the underlying topology and the second is to
-make the peer selection, per se, based on this knowledge. For the first part,
-the algorithm uses network tomography, a technique which suggests probing only
-from a large network's end points in order to infer its internal characteristics
-\cite{chny_tomography_2002}. Upon a new peer arrival, the tracker traceroutes it
-to get some knowledge (IP address, routers involved, RTT and hops conducted
-e.t.c). The more the peers in the swarm the better picture a tracker has for the
-underlying infrastructure, which can then return to the newcomming peer in the
-form of the bootstrapping peer list. The peers furtherly enhance this picture by
-tracerouting the returned peers themselves and reporting back to the tracker.
-For the selection part, a family of heuristics are proposed, which in general
-target those peers that expose low RTT and are within certain hop-count away of
-measured routers. The author's claim that peers running UTAPS instead of random
-peer selection can achieve better downloading rates with reduced burden on the
-underlying ISP infrastructure though their technique is coarse-grained and their
-evaluation configuration was far too small for extracting further safe
-conclusions.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{ \bf An Effective Network-Aware Peer Selection Algorithm in
-BitTorrent}
-Qin et al., in their work in \cite{qlzg_biteffpeersel_2009}, propose a
-clustering approach that differentiates peers in a swarm into local, intra-ISP
-and inter-ISP neighbours. The classification is done through the routers they
-use in order to communicate with the tracker. For this, a newly arriving
-BitTorrent peer traceroutes the tracker and sends the results in a form of a
-vector of triples that contain the IP and the hop counter for the router on the
-traceroute path as well as the link latency of the hop that was made in order to
-arrive to this router. The tracker uses this information in order classify the
-router through a $k$-means classification algorithm, with $k = 3$. For peer
-selection, the authors propose a biased approach where the tracker returns $c$
-close peers and $d = N - c$ distant peers, where $N$ is the length of the
-returned list. In choosing $c$, the tracker employs an iterative search process
-first in the peer's local neighbourhood. If not enough peers are available
-there, it goes to intra-ISP and subsequently, if needed, to the inter-ISP
-cluster. The evaluation of the algorithm is done in a similar artificial,
-contained laboratory enviromnent as in \cite{lcy_utaps_2008} and it shows up to
-5\% faster downloading times, and up to 22\% reduction of the total cross-ISP
-traffic.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Two-Hop-Away Neighbour Comparison and Selection}
+%\paragraph*{Low productive connection cutting}
+%There are three cases for any peer $P$ who receives $d(i, S, v)$ multiple
+% times:
+%\begin{inparaenum}[\itshape i\upshape)]
+%  \item $P$ receives both $d(i, S, 1)$ and $d(i, S, 0)$
+%  \item $P$ receives multiple $d(i, S, 0)$s from different paths, and P
+% randomly chooses to process one
+%  \item $P$ receives one $d(i, S, 1)$ and multiple $d(i, S, 0)$s, and $P$
+% processes $d(i, S, 1)$ and one randomly selected $d(i, S, 0)$
+%\end{inparaenum}
+%If the link with the largest cost is found and is a direct neighbour then the
+% connection is put in a will-cut list and stays there for a certain period of
+%time. If it is not, then it is handled by other peers. After that period,
+%connections are cut and recorded to $P$'s cut-list.
+%
+%\paragraph*{Source probing}
+%For a peer $P\in(N^2(S) - N(S))$ who receives only one $d(i, S, 0)$, the cost
+% of $PS$ is obtained (with list look-up or probing). Then $P$ compares it with
+%the cost from each hop and if $PS$ has the largest cost, $P$ will not keep this
+%connection, while otherwise the connection will be created.
+%
+%\paragraph{}
+%Supposing $n$ is the number of peers, $c_n$ is the average number of neighbours
+% and $c_e$ is the average cost of logical links, then in the flooding-based
+%search the traffic incurred by one query from an arbitrary peer in a
+%peer-to-peer network is $O(n)$. As observed in the Gnutella network
+%\cite{sripanidkulchai_gnutella_2001}, each peer issues $0.3$ queries per minute
+%in average, thus the per minute traffic incurred by the network with $n$ peers
+%is $O(n^2)$. Because each $d(i, S, v)$ has a TTL of $2$ in each source peer,
+%the traffic for one time LTM optimization in all peers is at most $2nc_n^2c_e$.
+%If each peer uses LTM $k$ times per minute, the total traffic incurred is
+%$2knc_n^2c_e$. Simulation shows the best value for $k$ is $2$ or $3$. So, the
+%traffic overhead caused by LTM to the network is $O(n)$.
+%
+%TTLj-detectors, with $j > 2$, would detect and break cycles with more than 4
+% links. LTM though, does not use such detectors because detector-flood traffic
+%would increase significantly, and cut links between two end-peers, could cause
+%queries initiated by them to traverse a path much more expensive than the cost
+%on the the cut link.
+%
+%\paragraph{}
+%LTM disadvantages are
+%\begin{inparaenum}[\itshape i\upshape)]
+%  \item disagreement of measured delay due to unsynchronized clocks causes
+% problems when deciding the cut positions, which can influence the network
+%connectivity, and
+%  \item the network delay metric mainly focuses on disabling the connections
+% between peers physically far away without considering the shortcuts created by
+%powerful peers.
+%\end{inparaenum}
+%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Scalable Bipartite Overlay (SBO)}
+\emph{Scalable Bipartite Overlay (SBO)}
+\cite{liu_bipartite_IPDPS,liu_bipartite_2007} reduces the overhead of creating
+and maintaining a minimum spanning tree cost by randomly dividing the nodes into
+two groups, namely red and white and assigning different tasks to different
+groups. When a peer joins the network, is randomly assigned with an initial
+colour, say red or white (separating all peers into two groups). Then the peer
+that plays the role of the bbootstrap host provides the joining peer with a list
+of active peers along with their colour information. The later uses this list to
+establish connections to differently coloured peers. This way, all peers form a
+bipartite overlay. The white peers measure distances to neighbours (red only) by
+using the network delay as a metric and report their corresponding red
+neighbours. The red peers, equipped with the information of all two-hop
+neighbours ($N^2$), creates minimum spanning tree of these neighbours and
+assigns efficient forwarding paths. Having a minimum spanning tree with two
+hops, a red peer is able to send its queries within that range. Some white
+peers, though, have become non-forwarding neighbours. In this phase, such a
+(white) neighbour will try to find another red peer being two hops away from its
+current red neighbour to replace the later as its new neighbour. The white peers
+can further optimize their positions within the overlay if this is required.
+
+%
+% TODO: SOME DISCUSSION
+%
+%In a static enviromnent LTM may reduce traffic cost by around 80 to 85 percent
+% while SBO reduces traffic cost between 85 and 90 percent. However, LTM  is
+%proved to converge in around 2-3 steps while SBO needs 4-5 steps. Moreover LTM
+%reduces response time by more than 60 percent in 3 steps while SBO needs 8. In
+%a dynamic environment (10 minute average peer lifetime, 0.3 queries/sec by each
+%peer) SBO and LTM reduce the average traffic cost per query (including the
+%overhead due to the optimization steps) by 85 and 80 percent, respectively.
+%Moreover LTM reduces the response time per query to 30 percent while SBO to 35
+%percent.
+%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Two-Hop-Away Neighbour Comparison and Selection (THANCS)}
 Work in \cite{liu_thancs_2005,liu_thancs_2008} proposes a distributed heuristic
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Peer-exchange Routing Optimization Protocols}
+\subsubsection{Hops Adaptive Neighbour Discovery (HAND)}
+\cite{chen_hand_2006} proposes the \emph{HAND} algorithm which uses a
+fully distributed triple hop adjustment strategy to address the topology
+mismatch problem. The approach's ultimate goal is to shape the current overlay
+graph $G$ into the \emph{Logical Communication Network (LCN)} $G^{*}$ which is
+how the optimal overlay is refered in the paper's context. A formal
+definition of achieving the above match is only if all peer hop sequences $(v_1,
+v_2, \ldots, v_k)$ in $G$ exist in $G^{*}$ and in the same order\footnote{In
+practice triple sequences $(v_1, v_2, v_3)$ are used.}. The mismatching
+detection is done in the following way. Suppose we want to verify peer a
+sequence, say $v_2-v_1-v_3$. A pair of probing messages are sent from $v_1$ to
+$v_2$ and $v_3$. Suppose delays of $(v_1,v_2)$ and $(v_1,v_3)$ are are denoted
+as $x$ and $z$, respectively. When the probing message arrives to $v_2$ it
+forwards it directly to $v_3$. Similarly, when the probing message arrives to
+$v_3$, it forwards it directly to $v_2$. These last steps are performed in order
+to obtain delays of $(v_2,v_3)$ and $(v_3,v_2)$ physical paths, respectively,
+denoted by $y$. If $y=z-x\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched
+and should be adjusted to $v_1-v_2-v_3$ by deleting edge $(v_1,v_3)$ and adding
+a new $(v_2,v_3)$. If $y=x-z\pm\varepsilon$, sequence $v_2-v_1-v_3$ is
+mismatched and should be adjusted to $v_1-v_3-v_2$ by deleting edge $(v_1,v_2)$
+and adding a new $(v_3,v_2)$, where $\varepsilon$ is a small positive real
+number denoting additional delays caused by possible forwarding and jitter
+delays. The advantages of the HAND algorithm compared to other approaches are
+that:
+\begin{inparaenum}[\itshape i\upshape)]
+  \item it does not need any clock synchronization,
+  \item it is a fully distributed algorithm making it robust and reliable in
+        decentralized systems,
+  \item the traffic overhead of the triple hop adjustment is very low,
+  \item it is applicable to dynamic peer-to-peer environments, and
+  \item maintains lower query response time.
+\end{inparaenum}
+
+%
+% TODO: SOME DISCUSSION
+%
+%Measurements conducted for evaluation perposes showed that in a static
+% environment the algorithm can effectively decrease traffic cost by about 77
+%percent and shorten the query response time by about 49 percent in less than
+%two minutes. In a dynamic environment it shows similar behaviour and with the
+%size of the overlay network having little impact on the effectiveness of the
+%algorithm. Compared to LTM both algorithms have almost the same traffic
+%reduction rate, however on the response time reduction rate HAND has a higher
+%one by about 4 percent. The traffic overhead of HAND is much less than that of
+%LTM by an average of 55 percent.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% TODO: DOUBLE CHECK IF THIS ALGORITHM GOES IN THIS SECTION/SUBSECTION
+\subsubsection{Adaptive Peer Selection}
+Bernstein et al., contributed on the use of machine learning for building peer
+selection strategies from past experience \cite{bflz_adaptpeersel_2003}. A
+decision tree is used to rate peers based on information about connection
+characteristics that is collected. These can be load, bandwidth, and past
+uploading experience. Then a by Markov decision process is incorporated as a
+mechanism to shape the policy for switching among the peers. The problem
+with this approach is that the peer selection algorithm is slow due to the
+learning process and the complexity of the method.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Innocuous Topology Aware (ITA) Overlay Construction}
+\cite{prfm_ita_2009} intoduces an algorithm for Innocuous Topology Aware
+construction of unstructured P2P networks. The paper's proposition is twoforld.
+Overlay construction and search method. For the overlay construction it employs
+the notion of \emph{short} and \emph{long} connections. Assuming $N$ is the
+number of all network participants, $\alpha \leq 1 $ a system-wide magic number
+and $x$ an $\alpha$-related latency threashold, the bootstrapping peer randomly
+selects $\alpha \ast N$ close (latency bellow $x$) and
+$\left( 1 - \alpha \right) \ast N$ distant (latency above $x$) nodes as its
+neighbours. The search method has two steps. First, the quering node initiates a
+flood to its distant neighbourhood with $TTL = 1$. Then, peers that receive a
+query over a long link, start a local flood with $TTL = ttl$, were $ttl$ is a
+system defined parameter.
+
+The goal of the overlay construction is to create a network that exposes low
+clustering, a beneficial characteristic of random graphs which can lead to a
+larger coverage of peers with the same number of messages and reduced
+duplication. This means more efficient information lookup that exposes low or no
+negative impact whatsoever to the rest of the mechanisms employed in
+unstructured P2P systems\footnote{For example the 1-hop replication and the
+dynamic querying mechanisms of later versions of Gnutella.}. The paper reports
+50\% reduction in communication latency ammong peers by cutting off some 20\% of
+the IP-level message traffic generation.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{EGOIST}
+EGOIST \cite{egoist_2008} is a set of algorithms to construct and manage overlay
+networks. It utilises a selfish approach in the sense that every participating
+peer continuously updates its neigbours so as to minimise the sum of distances
+to all destinations under shortest-path routing. In EGOIST, a newlly arriving
+peer, randomly connects to an already participating node through a bootstrap
+server. Being connected, it starts receiving info throught the link-state
+mechanism and thus, after some time, it has a complete picture of the overlay
+graph. Then it estimates the delay to all other nodes in order to determine its
+potential neighbours and ultimatelly connect to the overlay using some
+policy\footnote{For example, minimisation of the average delay to all its
+neighbours.}. It is obvious that there is extensive resource usage for updating
+the wiring of the all nodes in the system. The authors claim that they can
+reduce the load imposed by the monitoring and announcing process from $O(n^2)$
+to only $O(kn)$, where $n$ is the number of all nodes in the network and $k$ is
+the number of links a node chooses to establish.
+
+% TODO: CHECK IF THE FOLLOWING BITTORRENT ALGORITMS BELONG IN THIS SUBSECTION
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Biased Neighbour Selection}
+Bindal et al. \cite{rpc_bitbiased_2006} claim to strengthen BitTorrent
+protocol\cite{c_bittorrent_2003} locality by selecting most of a peer's
+neighbours to be out of the same ISP, while preserving the near optimal download
+performance of the protocol itself. BitTorrent systems employ most of the time
+mechanisms that can be proven very aggresive to an ISP's networking and
+accounting, basically, due to their lack of knowlegde of AS boarder limits.
+BitTorrent specification, by default allows for each peer, 35 connections.
+Biased neighbour selection experiments on some number $k$ which denotes a number
+of neighbouring peers which are not on the same ISP than the peer at hand.
+Consequently, the remaining $35 - k$ peers are from the same ISP. These $k$
+nodes are used in order to preserve a more extended, global view of a network,
+but in the same localize the load within the limits of a single ISP. This
+prevents clients from exchanging traffic through an expensive transit link, if
+there are alternative local connections that could offer the same, faster and at
+no additional cost for the ISP. Implementation can either be done through
+modifications on the tracker side, to identify ISP locallity, or through P2P
+traffic shaping devices installed on ISP's edge routers. The paper also proposes
+combination of biased neighbour selection along with bandwidth throttling and
+some caching approach for near optimal results. Unfortunatelly, implementation
+need, in some extend, contribution (exposure of ISP AS mappings) or
+infrastructure changes (installation of P2P traffic shaping devices) on the ISP
+level itself, making the adoption difficult in the general case.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Ono}
+Choffnes and Bustamante propose Ono \cite{cb_ono_2008}, a protocol for managing
+BitTorrent traffic so that it significantly reduces cross-ISP traffic and in
+the same time enhances downloading rates by favouring connections within the
+borders of a single autonomous system. Contrary to biased peer selection
+proposed in \cite{rpc_bitbiased_2006}, this work can lead to improved
+performance with no cooperation between ISPs and their subscribers, no
+additional infrastructure and no network topology information. The selection
+approach is landmark-based and leverages existing CDN infrastructure for peer
+distance estimation. CDNs already use both static (i.e. geographical) and
+dynamic (network measurement systems) information for their replica selection,
+so the authors claim that peers which exhibit similar redirection behaviour, are
+very likely they are close to the replica servers and as a consequence to each
+other. The redirection behaviour is modeled in terms of \emph{ratio map} in the
+article's parlance, a vector of (replica-server,ratio) tuples, where ratio is
+the percentage of times CDN redirects the peer to the specific replica-server.
+The bootstrapping phase consists of the peer performing DNS lookup to CDN names
+in order to build its redirection information. The interval for polling DNSs
+starts from 30 seconds and increases by one minute every time redirection info
+to the CDN is found to remain unchanged after a lookup. On the other hand, the
+interval is halved whenever is found to have been changed. In order not to avoid
+the bootstrapping phase if the ratio map is sufficiently fresh, the protocol
+persists the info after the end of a BitTorrent session.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Locality-Awareness in BitTorrent-like P2P Applications}
+In \cite{lclx_bitlocal_2009} the authors study and compare three different
+approaches in injecting locality awareness in BitTorrent-like applications. The
+first acts at a \emph{macroscopic-level} and targets neighbour selection. When a
+peer asks the tracker to join, the later sorts the swarm peers according to
+their distance to the requesting peer in AS hop count and send it the first
+(e.g. 50) peers in the list. The second manipulates the chocking/unchocking
+BitTorrent mechanisms at an \emph{intermediate-level}. A peer unchockes its 4
+closest in terms of AS hop count interested neighbours. The same applies also
+when the peer turns to the seeding state\footnote{Thus this approach favours
+least distance, contrary to the original BitTorrent implementation which favours
+uploading speed.}. On a \emph{microscopic-level}, the rare-first chunk picking
+policy is substituted by the locality-fist policy. The distance value of a piece
+is calculated as the mean value of the distances of the peers that posses it.
+The paper conducts several measurements on the efficiency of the above scheme to
+reach the conclusion that the major design desision for all locality aware P2P
+systems is the trade-off of optimising inter-AS traffic, which their solution
+achieves, and of the fairness among peers, a section that their locality-based
+approach does not do well in comparison to the random approach of the standard
+BitTorrent protocol.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{TopBT}
+\cite{rtl_topbt_2010} proposed an approach for enhancing proximity awareness
+in the BitTorrent protocol without any need of any infrastructure installed. It
+suggests that a good peer selection metric should take into account both the
+downloading speed and the network topology. Thus, it proposes the use of the
+ratio of download rate to upload rate, $\frac{d}{u}$, divided by link-level or
+AS-level routing hops, $l$ or $a$ respectively, in an attempt to form a
+comprehensive metric to select peers with high download rate, low reciprocal
+upload demand and low routing hops. The TopBT metric is applied on serveral
+parts of the original BitTorrent protocol in order to inject topology awareness
+for better peer selection; more specifically on the peer list the tracker
+returns when someone first tries to connect to the swarm, on the initial
+connection establishment on the connection replacement and of course the
+unchoking mechanism.
+
+%A peer that runs the TopBT protocol evaluates its neighbours by periodicaly
+%emitting pings or traceroutes in order to uchock those peer that exhibit less
+%hops to reach and higher download rates. \emph{Link-hops} are measured by using
+%the TTL value that the originator receives as a response from the remote
+%peer\footnote{Initial TTL values are known for the different operating systems
+%(e.g. 64 for Linux or 128 for Windows NT/2000/XP) so the originator can
+%calculate the hops using the value of the TTL on the response message it
+%receives.}. Also the peer, when offline, builds a table that maps IPs to
+%\emph{AS-hops}, using BGP dumps.
+
+Through experimentation on hundreds of PlanetLab as well as residential hosts,
+the authors claim more than 25\% traffic reduction and 15\% faster downloads,
+both with a lightweight cost when compared to popular BitTorrent
+implementations.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Underlying Topology-Aware Peer Selection (UTAPS)}
+UTAPS \cite{lcy_utaps_2008} is another peer selection strategy for the
+BitTorrent protocol. Its first step is to collect the needed information in
+order to construct a picture of the underlying topology and the second is to
+make the peer selection, per se, based on this knowledge. For the first part,
+the algorithm uses network tomography, a technique which suggests probing only
+from a large network's end points in order to infer its internal characteristics
+\cite{chny_tomography_2002}. Upon a new peer arrival, the tracker traceroutes it
+to get some knowledge (IP address, routers involved, RTT and hops conducted
+e.t.c). The more the peers in the swarm the better picture a tracker has for the
+underlying infrastructure, which can then return to the newcomming peer in the
+form of the bootstrapping peer list. The peers furtherly enhance this picture by
+tracerouting the returned peers themselves and reporting back to the tracker.
+For the selection part, a family of heuristics are proposed, which in general
+target those peers that expose low RTT and are within certain hop-count away of
+measured routers. The author's claim that peers running UTAPS instead of random
+peer selection can achieve better downloading rates with reduced burden on the
+underlying ISP infrastructure though their technique is coarse-grained and their
+evaluation configuration was far too small for extracting further safe
+conclusions.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{An Effective Network-Aware Peer Selection Algorithm in
+BitTorrent}
+Qin et al., in their work in \cite{qlzg_biteffpeersel_2009}, propose a
+clustering approach that differentiates peers in a swarm into local, intra-ISP
+and inter-ISP neighbours. The classification is done through the routers they
+use in order to communicate with the tracker. For this, a newly arriving
+BitTorrent peer traceroutes the tracker and sends the results in a form of a
+vector of triples that contain the IP and the hop counter for the router on the
+traceroute path as well as the link latency of the hop that was made in order to
+arrive to this router. The tracker uses this information in order classify the
+router through a $k$-means classification algorithm, with $k = 3$. For peer
+selection, the authors propose a biased approach where the tracker returns $c$
+close peers and $d = N - c$ distant peers, where $N$ is the length of the
+returned list. In choosing $c$, the tracker employs an iterative search process
+first in the peer's local neighbourhood. If not enough peers are available
+there, it goes to intra-ISP and subsequently, if needed, to the inter-ISP
+cluster. The evaluation of the algorithm is done in a similar artificial,
+contained laboratory enviromnent as in \cite{lcy_utaps_2008} and it shows up to
+5\% faster downloading times, and up to 22\% reduction of the total cross-ISP
+traffic.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Peer-exchange Routing Optimization Protocols (PROP)}
 \cite{qiu_prop_2007} introduces two protocols called \emph{Peer-exchange Routing
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Resolving the Topology Mismatch Problem in Unstructured
+\subsubsection{Resolving the Topology Mismatch Problem in Unstructured
 Peer-to-Peer Networks}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Distributed Domain Name Order}
+\subsubsection{Distributed Domain Name Order (DDNO)}
 \emph{Distributed Domain Name Order (DDNO)} \cite{z-yk_ddno_2005} uses the
@@
 
-\paragraph*{\bf Critical Topology-Aware Grouping}
+\subsubsection{Critical Topology-Aware Grouping (CTAG)}
 \emph{Critical Topology-Aware Grouping (CTAG)} presented in
@@
 to which nodes within the same organization are always addressed from the same
-block. The paper proposes to calculate peer proximity using longest maching IP
-segment criterion, called \emph{Adjacency Measurement (AM)} in the technique
+block. The paper proposes the \emph{Adjacency Measurement (AM)}, a technique
 which uses the longest matching IP segment criterion to calculate node
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Landmark Based Proximity}\label{sec:landmark}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Detecting the proximity of a node using landmarks, also known as
-\textit{Landmark Clustering}, is based on the view that nodes with similar
-distances to a set of predefined well-known landmark nodes are pretty likely
-being close to each other. But this approach has its weaknesses itself, like the
-fact that is a rather coarse grained approximation, therefore not particularly
-well suited for detecting the correct positions of nodes within close distance
-to each other.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf Landmark Binning}\label{sec:landmark_binning}
+%###############################################################################
+%###############################################################################
+%###############################################################################
+%       LANDMARKING
+%###############################################################################
+%###############################################################################
+%###############################################################################
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Landmark Binning}\label{sec:landmark_binning}
 \begin{figure*}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{\bf mOverlay}
+\subsubsection{mOverlay}
 \emph{mOverlay} \cite{zhang_moverlay_2004} tries to addresses the scalability
@@
 that any new node can reach its group by exchanging at most $O(logN)$ messages
-% within the network. Finally mOverlay also considers a second important
-function
+within the network. Finally mOverlay also considers a second important function
 in its protocol which is the constant maintenance of the overlay.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Discussion on the Algorithms for Unstructured Architectures}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \renewcommand\arraystretch{1.4}% (MyValue=1.0 is for standard)
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Discussion on the Algrithms for Unstructured Architectures}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Discussion on the Algorithms for Unstructured Architectures}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+ In short, the methods highlighted here are divided into the following
+groups
+\begin{inparaenum}[\itshape i\upshape)]
+  \item broadcast optimisation
+  \item caching,
+  \item overlay optimisation, and
+  \item landmark -based
+\end{inparaenum}
+ approaches.
+
+%###############################################################################
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Broadcast Optimisation} 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% TODO: TAKEN FROM ALGORITHMS SUBSECTION. SHOULD BE MERGED WITH DISCUSSION
+
+In the first realisation of the Gnutella protocol, each query request received
+by a peer was then forwarded to all of its neighbours which, clearly, generated
+unnecessary traffic over the communication links of all the network. The
+algorithms that fall into the group discussed in this section are referred
+to as broadcast optimisation -based because they enhance the ``naive''
+broadcasting approach, described above, and prefer to forward the messages to a
+selected subset of a node's outgoing links. The selection criteria vary
+depending on the algorithm but they mainly use one or multiple statistical
+metrics that span from some obvious ones, like the latency of the link between
+the nodes, to more sophisticated alternatives that may even take into account
+application level requirements such as the reliability of the neighbour.
+These appoaches have the advantage of enhancing the search responsiveness but on
+the other hand suffer from the drastic reduction of the search scope thus
+limiting the scalability of the whole network.
+
+%Expanding the search scope, on the other hand, is no easy task
+%because the overhead of forming multicast trees is proportional to the
+%multicast group size.
+%forwarding based schemes do not consider dynamic joining
+%and leaving of peers so they do not scale well on dynamic environments.
+
+%###############################################################################
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Caching}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% TODO: TAKEN FROM ALGORITHMS SUBSECTION. SHOULD BE MERGED WITH DISCUSSION
+
+%\subsection{Cache Based}
+%
+%Caching based protocols are effectively used to reduce traffic costs and
+%response
+%times. The caching policy varies depending on the way protocol handles the
+%index and the
+%content. Centralized P2P systems
+%use central index servers, while local caching systems, such as KazaA, use
+%super peers
+%to cache indices in a distributed way. Content caching is also possible in P2P
+%systems, where nodes cache the forwarded content for further retrievals.
+%Although caching has the above mentioned advantages,  duplication
+%of messages still exist, which limits the scalability of these approaches.
+%Therefore, cache based approaches are analyzed in the following categories:
+%  \begin{itemize}
+%    \item \emph{data index caching},
+%    \item \emph{content index caching},
+%    \item \emph{centralized}, and
+%    \item \emph{local}.
+%  \end{itemize}
+%
+Caching is a widely used technique to exploit locality and minimise redundant
+transfer of data, initially adopted by web and file server application
+environments with great success. Since peers in a P2P system also work as
+servers, it is intuitively expected that P2P file sharing systems can also
+benefit from caching. However, two of the main characteristic properties of P2P
+systems, nodes frequently joining and leaving and query lifetime relatively
+short (at least compared to those needed to be served by web servers) , makes
+caching in P2P networks a non-trivial issue. There are usually two levels of
+caching possible in file sharing systems, caching indices or pointers to data or
+caching the data itself. Caching is already implemented and successfully used in
+some commercial P2P systems. One popular example is KazaA, which uses caching of
+indices on its super-peer level. In the paragraphs below, caching-based
+algorithms are presented and explained in more depth.
+
+%###############################################################################
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Overlay Optimization}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% TODO: TAKEN FROM ALGORITHMS SUBSECTION. SHOULD BE MERGED WITH DISCUSSION
+
+% These approaches include creating spanning trees using
+%connection graphs, creating cluster of physically close nodes, or using latency
+%information to detect proximity. The brief description of each category is
+%presented below:
+%
+%  \begin{itemize}
+%    \item \emph{Spanning tree based}. These approaches construct of a rich
+%    graph  based on the network connections and build minimum spanning tree
+%    on the graphs, causing large traffic overhead to the system
+%    \cite{chu_esm_2000,chu_esm_2002}.
+%
+%    \item \emph{Cluster based}. These approaches select to link physically
+%    closer nodes with each other, therefore shrink the search scope
+%    significantly while mapping accuracy is not always guaranteed.
+%
+%    \item \emph{Minimum latency first}. Use of latency as a metric to calculate
+%    distance among peers. They require global latency information
+%    ``landmarks''\footnote{Measuring latency between peers and stable Internet
+%    servers.}.
+%
+%  \end{itemize}
+
+The overlay optimization based protocols modify the topology of the P2P network
+using various techniques. The two commonly used approach that is investigated in
+this section are creating spanning trees using connection graphs, and creating
+clusters of physically close nodes. Spanning tree based approaches construct
+rich graphs based on the network connections and build minimum spanning trees on
+these graphs. Even though spanning trees provides efficient query performance,
+the construction and the update costs of the spanning tree, especially in
+dynamic environments with many nodes joining and leaving the network, cause
+large traffic overhead to the corresponding
+network \cite{chu_esm_2000,chu_esm_2002}. The cluster based approaches on the
+other hand select to link physically closer nodes with each other, therefore
+significantly shrinking the search scope of the system. Added to that, commonly
+used methods for correct determination of physical node positions across
+Internet do not always return reliable results, therefore mapping accuracy is
+not guaranteed. Bellow, the details of the algorithms that use the spanning tree
+based or cluster based topology optimization methods are presented.
+
+%###############################################################################
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Landmark Based Proximity}\label{sec:landmark}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Detecting the proximity of a node using landmarks, also known as
+\textit{Landmark Clustering}, is based on the view that nodes with similar
+distances to a set of predefined well-known landmark nodes are pretty likely
+being close to each other. But this approach has its weaknesses itself, like the
+fact that is a rather coarse grained approximation, therefore not particularly
+well suited for detecting the correct positions of nodes within close distance
+to each other.
+
+%###############################################################################
 

### 2014-01-23 structured.tex
2012.08.20 (Vassilis) Added LaTeX tables (known issue on wrapping text inside cells). Reshaped discussion material. Introducing rough, algorithm performance, estimation (high, medium, low) by tracking efficiency, control overhead and scalability. First attempt on plotting data in a systematic way (needs more work though).

@@
 
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% Unable to identify nodes that are close to routers but it can reduce routing
+% latency to far away node by approximatelly 50-75%.
+medium &
+% maintenance of node states, subscriptions and notifications need global
+% overview of the overlay resulting in high control overhead
+high &
+% Statically implemented sucrificing the the decentralised nature of DHTs and
+% thus poorly scalling.
+low
+\end{tabular}
+\end{center}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 closest nodes and their neighbours, so that Euclidean distances between the node
-and all known hosts predict the network latency between
-them\cite{cox_vivaldi_2004}. The benefit of these synthetic coordinates are that
+and all known hosts predict the network latency between them
+\cite{cox_vivaldi_2004}. The benefit of these synthetic coordinates are that
 they are explicitly used as the node's ID, and distance computations among nodes
@@
 One major limitation of \textit{Mithos} is that the protocol cannot handle
-dynamic arrivals and removals from the network, as is happens in mobile
+dynamic arrivals and departures from the network, as is happens in mobile
 networks.
@@
 % real-world p2p deployment of the scheme.
+
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% It has very good results in reducing comunication latency among peers but its
+% efficience is being constrained by the fact that the proximity is calculated
+% at bootstrap, assigning virtual coordinates that are used througout of a
+% nodes life cycle and thus the overlay is not adapting to a high-churn
+% environment.
+medium &
+% Incrimental probing is applied in quadradants and in two-hop distance at each
+% step that does not generate a lot o control overhead.
+medium &
+% It is constrained by the fact that the protocol does not respond to dynamic
+% peer arrivals and departures
+low
+\end{tabular}
+\end{center}
 
@@
 
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% It minimises routing length, but may suffer load imbalance since nodes higher
+% in the tree hierarchy are more likely to route most of the messages in the
+% overlay.
+medium &
+% Heartbeat approach incurs extra control overhead to the network. Fortunatelly
+% involves only the parent and the children of the departing/failing node.
+% nodes.
+low &
+% Seems to scale nicely from 10K- to 1M-node testbeds.
+high
+\end{tabular}
+\end{center}
+
 %###############################################################################
@@
 {\em proximity routing} and the other on {\em proximity neighbour selection}
-approaches overlay optimization techiques for
-exploiting proximity. For the first, the goal is in choosing the best next hop
-during routing a message. Due to the fact that the routing in Kademlia is
-iterative, it is the initiator of the lookup that chooses each next hop from a
-set of candidate nodes. Standard Kedemlia always chooses the closest node with
-respect to the XOR metric but with proximity routing selection, Kademlia
-chooses the one with the smallest underlay metric cost. Thus, as in all such
-approaches, there is a trade-off between the overlay and underlay distances.
+approaches overlay optimization techiques for exploiting proximity. For the
+first, the goal is in choosing the best next hop during routing a message. Due
+to the fact that the routing in Kademlia is iterative, it is the initiator of
+the lookup that chooses each next hop from a set of candidate nodes. Standard
+Kedemlia always chooses the closest node with respect to the XOR metric but with
+proximity routing selection, Kademlia chooses the one with the smallest underlay
+metric cost. Thus, as in all such approaches, there is a trade-off between the
+overlay and underlay distances.
 
@@
 from the MaxMind GeoIP database\footnote{MaxMind Geolocation Technology.
-http://www.maxmind.com} to retrieve proximity information of a given
-peer based on its region, country, or ISP location, and use this information to
-form clusters of nearby peers. An alternative method could also be the use of 
-Vivaldi coordinates \cite{cox_vivaldi_2004}. However, authors report that
-clustering approach performs better than Vivaldi protocol based on their
-experiments. Authors also report that {\em Proximity routing} protocol
-successfully worked with the Kademlia protocol and improved the locality of the
-connections over peers.
-
-As already mentioned the paper
-proposes \emph{proximity neighbour selection} method, as well, for overlay
-optimisation. The goal, according to this, is to keep peers with the least
-contact cost in the routing table. As Kademlia constantly learns new peers
-(incoming requests, through iterative lookup) no special algorithm for
-searching more cost effective peers is necessary and simply choose the best
-peers seen so far.
+http://www.maxmind.com} to retrieve proximity information of a given peer based
+on its region, country, or ISP location, and use this information to form
+clusters of nearby peers. An alternative method could also be the use of Vivaldi
+coordinates \cite{cox_vivaldi_2004}. However, authors report that clustering
+approach performs better than Vivaldi protocol based on their experiments.
+Authors also report that {\em Proximity routing} protocol successfully worked
+with the Kademlia protocol and improved the locality of the connections over
+peers.
+
+As already mentioned the paper proposes \emph{proximity neighbour selection}
+method, as well, for overlay optimisation. The goal, according to this, is to
+keep peers with the least contact cost in the routing table. As Kademlia
+constantly learns new peers (incoming requests, through iterative lookup) no
+special algorithm for searching more cost effective peers is necessary and
+simply choose the best peers seen so far.
+
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% Uses static information to cluster nodes which is not the most efficient
+% approach in dynamic systems
+medium &
+% 
+medium &
+%
+medium
+\end{tabular}
+\end{center}
 
@@
 %
+
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% Simple approach that exploits proximity but it is coarse grained and thus not
+% efficient.
+low &
+%
+low &
+%
+medium
+\end{tabular}
+\end{center}
 
@@
 %infrastructure raise concearns about its efficiency and scalability. Being,
-%primarilly, a network diagnostic utility, it is concearned too heavy weighted
-%and intrucive for use in a larger scale\cite{ratnasamy_binning_2002}.
-%Additionally, disabling ICMP is a common administrative policy for edge sites
-%to enforce security, while dumping BGP routing
-%tables\cite{krishnamurthy_bgpclust_2000} is not directly available to the
-%application layer.
+%primarilly, a network diagnostic utility,
 % Even though traceroute provides detailed information about the network
@@
 %
+
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% The efficiency of choosing close by peers is very good and reported almost
+% 65\% better than that of GNP (and thus generally of static landmarking as a
+% whole)
+high &
+% traceroute is a very heavy tool and thus incurs excessive burden to the
+% network resourses.
+high &
+% Traceroute is categorised as too heavy-weighted and intrucive for use in a
+% larger scale system \cite{ratnasamy_binning_2002}. Additionally, disabling
+% ICMP is a common administrative policy for edge sites to enforce security,
+% while dumping BGP routing tables\cite{krishnamurthy_bgpclust_2000} is not
+% directly available to the application layer.
+low
+\end{tabular}
+\end{center}
 
@@
 other nodes it applies a heuristic mechanism to fill up the list. Entries are
-dynamicaly added or removed as the network state changes. Due to the fact that,
-with each step, there is always progress towards the target node in the ID
-space, PChord will result in a reduced hop number than Chord, since each hop in
-PChord is larger or at least equal in the key space. Passing through proximity
-links in the underlaying network means reduced routing cost. Additionally, if
-the number of entries in the proximity list is the same as the number of network
-partitions, PChord prevents hops from jumping in and out the network partition
-the current node belongs to.
+dynamicaly added or removed as the network state changes. The overlay
+maintenance cost can be considered similar to Chord's since besides node
+join/leave and cost of finger table entries heartbeat, PChord adds heartbeat of
+nodes inside the proximity list but since this list always contains nodes with
+the lowest RTT the added cost is kept to a minimal.
+
+Due to the fact that, with each step, there is always progress towards the
+target node in the ID space, PChord will result in a reduced hop number than
+Chord, since each hop in PChord is larger or at least equal in the key space.
+Passing through proximity links in the underlaying network means reduced routing
+cost. Additionally, if the number of entries in the proximity list is the same
+as the number of network partitions, PChord prevents hops from jumping in and
+out the network partition the current node belongs to.
+
+% TODO: SOME DISCUSSION
+% TODO: NEEDS REWORDING
+% PChord’s main routing optimizations are of less overlay hops, passing
+% proximity links of the underly network and passing the same network partition
+% only once in the routing path, which directly results in the lower RDP of
+% routing efficiency. Meanwhile, PChord keeps light-weightmaintenance cost as
+% Chord does.
+
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% Small decrease in needed overlay hops per object publishing
+medium &
+% The are heartbeat mechanism for both finger table and proximity list adds a
+% burden to the network by default. But the overall overhead is decreased as
+% the proximity list becomes longer.
+medium &
+% Improvement of routing is achieved with more object pointers helping to
+% shorten the routing path of object querying. When the distance of query source
+% to target is far away both Chord and PChord achieve the average hop number to
+% not increase.
+high
+\end{tabular}
+\end{center}
 
@@
 
+% TODO: MAYBE REVIEW THE EVALUATION
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% good routing efficiency
+medium &
+% Releaves the system from joining overhead and topology aware ID construction
+% but doubles the information needed to be stored and maintained by the
+% introduction of neighbouring table
+medium &
+% assumes the operation of ideal underlying anycast mechanism (meaning that
+% outside messages are always routed to the nearest node which in the current
+% Internet, this is not always feasible. Moreover, is designed on the basis of
+% IPv6 protocol which is not widely operational yet.
+medium
+\end{tabular}
+\end{center}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 exploit the hierarchical features of IPv6 in order to create a substrate that
-reduces interdomain traffic between service providers. The difference between
+reduces inter-domain traffic between service providers. The difference between
 Chord6 and Chord the identifier definition. Therefore, the approach can be
@@
 
-%
-% TODO: SOME DISCUSSION
-% However, even though nodes in the same domain would have close identifiers,
-% nodes in two close domains may have very different ones.
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% Can reduce intra-domain average path length but even though nodes in the same
+% domain would have close identifiers, nodes in two close domains may have very
+% different ones reducing the overal efficience of the approach in general when
+% considering inter-domain average path length.
+low &
+% The construction of the identifier space is done at node join and in an
+% independent fashion involving the application of two hash functions meaning
+% resulting in low overhead stress to the network and its nodes.
+low &
+% Is designed on the basis of IPv6 protocol which is not widely operational yet.
+medium
+\end{tabular}
+\end{center}
 
@@
 
+% TODO: MAYBE REVIEW THE EVALUATION
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+%
+medium &
+%
+medium &
+%
+medium
+\end{tabular}
+\end{center}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Quasi-Chord}
-The approach that is proposed by Sun and Zhang in \cite{sun_quasi_2008}
-confronts the topology mismatch problem, in the Chord DHT context which, in its
-vanilla version, takes no consiteration for the underlying physical network
-topology during the construction of the identifier cycle. To construct a
-\emph{Quasi-Chord} network three steps are followed needed. First, each host
-acquires its coordinates in the geometric space utilizing the \emph{global
-network position (GNP)} protocol \cite{ng_gnp_2001}. Second, using the
-\emph{Cantor space filling curve}, the $2$-dimensional space is converted to a
-$1$-dimensional one, used in the last, third, step to build the Quasi-Chord
-circle according to the host's Cantor value.
+The approach that is proposed by Sun and Zhang in \cite{sun_quasi_2008}, tries
+to address the topology mismatch problem, at the construction phase of the
+Chord identifier cycle, on the the basis of underlying topology information.
+The proposed model constructs the \emph{Quasi-Chord} ovelay network in three
+steps. First, each host acquires its coordinates in the geometric space
+utilizing the \emph{global network position (GNP)} protocol \cite{ng_gnp_2001}.
+Second, using the \emph{Cantor space filling curve}, the $2$-dimensional space
+is converted to a $1$-dimensional one, used in the last, third, step to build
+the Quasi-Chord circle according to the host's Cantor value. In Quasi-Chord
+each node stores 2 finger tables; one clockwise and one counter-clockwise.
 
@@
 %
+
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% The algorithm shows to reduce end-to-end delay as well as not produce
+% duplicates of messages.
+medium &
+% The overhead is more compared to vanilla chord since mapping of 2-dimensional
+% space to 1-demensional with Cantor SPF adds some extra burden in time
+% complexity. Also the addition of the couter-clockwise finger table doubles the
+% data, per se, needed to be stored and maintained.
+high &
+% The GNP system is a static landmark approach making things not fully
+% distributed and thus constrain the model's scalability.
+medium
+\end{tabular}
+\end{center}
 
@@
 %
+
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% The efficiency of proximity detection through the prefix matching technique
+% is caught in a vicious cycle with the storage space needed for DHT object
+% publishing (how many bits will be used for either)
+low &
+% The overhead for detecting proximity is only comparison of prefixes
+low &
+%
+medium
+\end{tabular}
+\end{center}
 
@@
 %
+
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% The result of experiments show that the Cone, compared with Chord, has obvious
+% improved on the delay of route and the hops of overlay networks…
+medium &
+% The approach incorporates a static landmark + RTT to these landmarks in order
+% to exploit proximity information thus incurring low overall additional
+% operational overhead but imposing extra stress to the landmarks themselves 
+medium &
+% landmark based approaches are not full distributed thus not smoothly scalable.
+medium
+\end{tabular}
+\end{center}
 
@@
 %
+
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+%
+medium &
+% The paper suggests that this approach generates significantly lower overhead
+% than the vanilla Pastry algorithm
+low &
+% Dynamic landmarkin approach helps the scalability of the algotithm. The
+% authors claim that the temporary nature of RLM's landmark nodes does not
+% prevent them to maintain a good ratio in the presence of high landmark
+% degression rates.
+high
+\end{tabular}
+\end{center}
 
@@
 
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% SAT-Match seems to outperform landmark binning in the sense of providing a
+% higher grained differenciation of proximity, adding to the fact that
+% SAT-match is fully distributed. Also to further enhance the efficiency it can
+% work cooperatively with landmark binning itself.
+high &
+% distance probing and iterative computations add to the approach's overhead.
+% Also the fact that the selective jumps are not always performed the added
+% overhead to efficiency gained is not guarranteed in any configuration and
+% circumstance.
+medium &
+% The self-adaptation nature of the algorithm and the selective jump mechanisms
+% scale much better than, for example, Mithos.
+medium
+\end{tabular}
+\end{center}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+% TODO: MAYBE REVIEW THE EVALUATION
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+%
+low &
+%
+low &
+%
+medium
+\end{tabular}
+\end{center}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\begin{figure}[h!]
-\hspace{-3ex}
-\begin{center}
-\footnotesize
-%\begin{landscape}
-\begin{longtable}{
-|>{\columncolor[gray]{.7}}m{0.2\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.4\columnwidth}
-|>{\columncolor[gray]{.8}}m{0.2\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.2\columnwidth}
-%|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
-%|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
-%|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
-%|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
-|}
-\caption{Decentralized Structured Algorithms}\label{fig:struct_compare_table}\\
-\hline
-\rowcolor[gray]{.5}
-\textbf{Algorithm} & \textbf{Overlay structure} & \textbf{Base protocol} & \textbf{Scalability} \\
-\hline
-\endfirsthead
-\multicolumn{4}{c}%
-{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
-\hline
-\rowcolor[gray]{.5}
-\textbf{Algorithm} & \textbf{Overlay structure} & \textbf{Base protocol} & \textbf{Scalability} \\
-\hline
-\endhead
-\hline \multicolumn{4}{r}{\textit{Continued on next page}} \\
-\endfoot
-\hline
-\endlastfoot
-
-\hline
-\textbf{Global Softstate} &
-\textbf{Geographic layout} Uses first landmark clustering
-then measures RTTs to identify close nodes &  &  \\
-
-\hline
-\textbf{Mithos} &
-\textbf{Geographic layout} Uses nodes as topology landmarks and directed
-incremental probing to optimize topology & & Scales well as all
-operations are local ??? \\
-
-\hline
-\textbf{Self-Adaptive Topology Matching} &
-\textbf{Proximity neighbour selection} Uses lightweight probing and
-selective jumps to optimize the topology & CAN &  Better than Mithos \\
-
-\hline
-\textbf{Delay Aware P2P System} & \textbf{} & & \\
-
-\hline
-\textbf{VERSION OF CHORD - DHT-PNS} &
-\textbf{Proximity neighbour selection} Uses Proximity Neighbour Selection and the Vivaldi
-system & Chord  &  \\
-
-\hline
-\textbf{MAY OMIT- VERSION OF CHORD - Quasi-Chord} &
-\textbf{Proximity based}  & Chord  &  \\
-
-\hline
-\textbf{LAPTOP} &
-\textbf{Geographic layout} Hierarchical overlay structure  &  & routing path length $\log{_d N}$,
-join/leave overhead $d\log{_d N}$ \\
-
-\hline
-\textbf{IP-Based Clustering} &
-\textbf{Proximity neighbour selection} Proximity neighbour selection based on longest common
-prefix of IP addresses &    &  \\
-
-\hline
-\textbf{CHOord considering Proximity on IPv6} &
-\textbf{Proximity routing} Uses IPv6 address format to provide proximity &  Chord
- & Better than Chord \\
-
-\hline
-\textbf{Proximity in Kademlia} &
-\textbf{Proximity routing} Applies  proximity neighbour selection (PNS) and proximity route selection (PRS)
-to Kademlia & Kademlia &   \\
-
-\hline
-\textbf{Cone} &
-\textbf{Proximity neighbour selection} Uses proximity neighbour selection (PNS) & Chord  & Better than Chord \\
-
-\hline
-\textbf{DynaMO} &  &  &  \\
-
-\hline
-\textbf{MAY OMIT-BADLY WRITTEN-PChord} &
-\textbf{Proximity based}  &  Chord  & \\
-
-\hline
-\textbf{AChord} &  & & \\
-
-\hline
-\textbf{Chord6} &
-\textbf{Proximity routing} Uses IPv6 hierarchical address format to cluster
-topologically close nodes & Chord  &  \\
-
-\hline
-\end{longtable}
-%\end{landscape}
-\end{center}
-\vspace{-2.5ex}
-\vspace{-2.5ex}
-%\label{fig:struct_compare_table}
-%\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Geographic Layout} \label{section:geographic_layout}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% TODO: TAKEN FROM ALGORITHMS SUBSECTION. SHOULD BE MERGED WITH DISCUSSION
-
-Inside geographic layout category fall those kind of methods which practically
-position physically close by nodes together in the application space as well.
-This is done by adjusting and maintaining the routing tables of all involved
-peers accordingly by exploiting proximity information that describe the overall
-geographic positioning of the peer nodes. One popular method for detecting and
-constructing a geographic view of the underlying network is the use of well
-established servers dedicated to landmarking the physical space, pretty much as
-discussed in Section \ref{sec:landmark} of landmark based proximity in the
-unstructured P2P networks. The same disadvantages of difficult and inaccurate
-positioning and the cost of deploying and maintaining these landmarks apply in
-this case as well.
-% TODO: maybe a discussion sentence bellow
-Even though managing the overlay structure based on the geographic layout of the
-nodes improves the query efficiency of the system, on the other hand, it tends
-to create hotspots, and the needed failure resilience is undermined by the fact
-that close by nodes are more likely to suffer collective failures.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% TODO: Geographic Layout
-%
-%\begin{itemize}
-% \item Geographic Layout\\
-%The node IDs are assigned in such a way that nodes close by in the physical
-%network topology, be close in the node ID space as well. Implementations that
-%work relatively well with this approach have been incorporated into CAN. Nodes
-%measure the RTT between themselves and a set of landmarks in order to match the
-%CAN space as much as possible to the physical one. Unfortunately, the approach
-%requires well known landmark servers and for that matter is  not fully
-%self-organizing which can further lead to imbalanced node distribution. On
-%other DHTs, such as Chord or Pastry, another problem emerges.  To gain fault
-%tolerant properties, these protocols, replicate key-value pairs on neighbouring
-%(in the ID space) nodes. When a proximity-based node ID assignment has been
-%used, the needed failure resilience is undermined by the fact that close by
-%nodes are more likely to suffer collective failures.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Proximity Routing}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% TODO: TAKEN FROM ALGORITHMS SUBSECTION. SHOULD BE MERGED WITH DISCUSSION
-
-Proximity routing does not require routing tables to be built using any
-knowledge about network proximity. On the other hand it exploits such knowledge
-in order to choose the best next hop during routing a message. This approach
-balances between choosing the node that will further progress the routing
-towards the destination and choosing the closest entry in the routing table, in
-terms of network proximity.
-% TODO: looks like discussion material
-Thus, it is relatively less effective than
-geographical layout when applied to CAN(-like) implementations. Moreover, the
-technique has been incorporated into a version of Chord causing an increase on
-the overhead of node joins and the size as well as maintenance cost of finger
-tables.
-
-% Moreover, the technique has been incorporated into a version of Chord
-% (TODO: PChord | additionally check whether this is true) causing an increase
-% on the overhead of node joins and the size as well as maintenance cost of
-% finger tables.
-
-%
-%\cite{dabek_cfs_2001} proposes a server selection scheme for the Chord DHT, on
-%the domain of proximity routing selection. In \emph{CFS}, each node predicts
-%the entire lookup latency as a function of the total number of nodes and the
-%average overlay next routing peer. The problem is that it is very difficult to
-%have a clear picture on the total number of nodes and the average hop latency
-%from the local. This leads to rough estimations that consequently decreases
-%overall performance.
-%
-The methods that use proximity routing and their details are listed below.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Proximity Neighbour Selection}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% TODO: TAKEN FROM ALGORITHMS SUBSECTION. SHOULD BE MERGED WITH DISCUSSION
-
-\emph{Proximity Neighbour Selection} constructs the routing tables using
-proximity knowledge. The proximity information used in this method is different
-than the landmark based systems described in
-Section~\ref{section:geographic_layout}, as time-to-live values between nodes,
-or directly node ID prefixes are used to detect proximity. Tapestry, Pastry, and
-CAN successfully implemented the proximity into their algorithms by using this
-approach. The routing protocol in Pastry is based on longest node ID prefix
-matching, while CAN uses RTT values to detect close by nodes.
-\cite{castro_proximityp2p_2002} reports that proximity neighbour selection as an
-effective proximity based method. Proximity Neighbor Selection based algorithms
-are described in detail below.
-
-In this section we presented the state of the art for the algorithms applied to
-structured P2P systems. We categorised it based on their philosophy they
-incorporated in order to alleviate the topology mismatch problem into overall
-three categories following the previous work from
-\cite{castro_proximitydht_2002,castro_topawareroute_2002,ratnasamy_openq_2002}
-namely
-\begin{inparaenum}[\itshape i\upshape)]
-  \item geographic-layout-based,
-  \item proximity-routing-based, and
-  \item proximity neighbour selection based
-\end{inparaenum}
-. Following is a discussion on the above categories trying to point out the
-advantages, disadvantages and novelties introduced by each and by all of them.
-A quick overview can be seen in Table \ref{fig:struct_compare_table}.
-
-In Section~\ref{section:structured}, we presented the state of the art
-structured decentralized P2P algorithms in three different categories based on
-their handling of the network structure to tackle the topology mismatch problem:
-geographic layout based, proximity routing based, and proximity neighbour
-selection based. In this section, we present a final discussion of all the
-methods including the advantages, disadvantages, and novelties related to solve
-the topology mismatch problem. Table~\ref{fig:struct_compare_table} presents an
-overview of all the structured decentralized algorithms described earlier in
-this section.
-
 Structured P2P network algorithms use a global distributed hash table or a
@@
 the node to the network, so that nodes within the same network domains are
-selected as peers, or neighbours, during the overlay topology construction. In
-this chapter, we have described three such approaches to optimize the topology
-matching problem: geographic layout, proximity routing and proximity neighbour
-selection.
-
-In geographic layout, nodes try to estimate the geographic positions of the
-peers, and construct the DHT considering the proximity of the peers. Landmark
-servers and RTT measurements are two popular methods, which can also be used in
-conjunction, to discover physically close by peers over the network. However, as
-discussed in Section \ref{sec:landmark_binning}, these methods do not always
-give reliable estimates for the node positions over the internet. The landmark
-servers are not self-organizing and have maintenance overheads. To serve a
-P2P network with millions of peers, multiple landmark servers
-distributed over the whole world is required, which is hard to manage. The RTT
-measurements also can measure the delay between peers, but it is a greedy method,
-which can result non-optimal overlay topologies especially if close by nodes
-have low bandwidth connections among themselves.
-
-Proximity routing does not discover or store proximity information for the
-peers, however, during package forwarding, nodes that have lower latency to the
-destination, or with a closer key to the destination is selected. As no
-proximity information is used, the system is based on a greedy approach and it
-usually selects low latency paths over the overlay, which maps to suboptimal
-longer paths on the physical topology. For these obvious reasons, proximity
-neighbour selection is generally considered as a superior method to proximity
-routing, however, joint uses of these two protocols are also possible.
-
-Proximity neighbour selection approach is similar to proximity routing, with an
-exception that during forwarding, the proximity information of the peers are
-also considered. Therefore, for each application, proximity information is
-extracted and used during routing. Depending on the application, the proximity
-information can be the node sharing the same prefix with the current node, or
-TTL to detect close by neighbours. Tapstry and Pastry are two popular methods
-implementing the proximity neighbour selection among others mentioned above in the
-proximity neighbour selection section.
+selected as peers, or neighbours, during the overlay topology construction.
+
+This section gathered together some of the most distinctive research work that
+can be found in the literature for addressing the inefficient network resource
+utilisation in structured architectures imposed by the topology mismatch
+problem. A categorisation of the methods used for achieving their goals is the
+following based on previous work from
+\cite{castro_proximitydht_2002,castro_topawareroute_2002,ratnasamy_openq_2002}:
+\begin{inparaenum}[\itshape i\upshape)]
+  \item geographic-layout-based,
+  \item proximity-routing-based,
+  \item proximity neighbour selection based,
+\end{inparaenum}
+
+A quick overview can be obtained in Table~\ref{structured:table}. Topology
+adaptation, landmarking and caching/replication have been visited during the
+discussion on unstructured systems in Section~\ref{section:unstructured} so
+we will not elaborate more here. Again, we additionaly gather highlights and a
+rough estimation of each algorithm's positive and negative aspects regarding
+their implementation.
+
+\input{structured-table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Geographic layout} \label{section:geographic_layout}
+Inside geographic layout category fall those kind of methods which practically
+position physically close by nodes together in the application space as well.
+This is done by adjusting and maintaining the routing tables of all involved
+peers accordingly by exploiting proximity information that describe the overall
+geographic positioning of the peer nodes. Landmark servers and RTT measurements
+are two popular methods, which can also be used in conjunction, to discover
+physically close by peers over the network. However, as discussed in Section
+\ref{sec:landmark_binning}, these methods do not always give reliable estimates
+for the node positions over the internet. The landmark servers are not
+self-organizing and have maintenance overheads. To serve a P2P network with
+millions of peers, multiple landmark servers distributed over the whole world is
+required, which is hard to manage. The RTT measurements also can measure the
+delay between peers, but it is a greedy method, which can result non-optimal
+overlay topologies especially if close by nodes have low bandwidth connections
+among themselves.
+
+Even though managing the overlay structure based on the geographic layout of the
+nodes improves the query efficiency of the system, on the other hand, it tends
+to create hotspots, and the needed failure resilience is undermined by the fact
+that close by nodes are more likely to suffer collective failures.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% TODO: Geographic Layout
+%
+%\begin{itemize}
+% \item Geographic Layout\\
+%The node IDs are assigned in such a way that nodes close by in the physical
+%network topology, be close in the node ID space as well. Implementations that
+%work relatively well with this approach have been incorporated into CAN. Nodes
+%measure the RTT between themselves and a set of landmarks in order to match the
+%CAN space as much as possible to the physical one. Unfortunately, the approach
+%requires well known landmark servers and for that matter is  not fully
+%self-organizing which can further lead to imbalanced node distribution. On
+%other DHTs, such as Chord or Pastry, another problem emerges.  To gain fault
+%tolerant properties, these protocols, replicate key-value pairs on neighbouring
+%(in the ID space) nodes. When a proximity-based node ID assignment has been
+%used, the needed failure resilience is undermined by the fact that close by
+%nodes are more likely to suffer collective failures.
+%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Proximity Routing}
+Proximity routing does not require routing tables to be built using any
+knowledge about network proximity. On the other hand it exploits such knowledge
+in order to choose the best next hop during routing a message. This approach
+tries to balance between choosing the node that will further progress the
+routing towards the destination and choosing the closest entry in the routing
+table, in terms of network proximity. As this is very difficult to be achieved,
+the system might end up using a greedy approach that selects low latency paths
+over the overlay, which maps to suboptimal longer paths on the physical
+topology. Thus, in general, proximity neighbour selection is considered as a
+superior method to proximity routing, however, joint uses of these two protocols
+are also possible. Moreover, they are considered relatively less effective than
+geographical layout when applied to CAN(-like) implementations. Moreover, the
+technique has been incorporated into a version of Chord causing an increase on
+the overhead of node joins and the size as well as maintenance cost of finger
+tables.
+
+%
+%\cite{dabek_cfs_2001} proposes a server selection scheme for the Chord DHT, on
+%the domain of proximity routing selection. In \emph{CFS}, each node predicts
+%the entire lookup latency as a function of the total number of nodes and the
+%average overlay next routing peer. The problem is that it is very difficult to
+%have a clear picture on the total number of nodes and the average hop latency
+%from the local. This leads to rough estimations that consequently decreases
+%overall performance.
+%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Proximity neighbour selection}
+\emph{Proximity Neighbour Selection} constructs the routing tables using
+proximity knowledge. The proximity information used in this method is different
+than the landmark based systems described in
+Section~\ref{section:geographic_layout}, as time-to-live values between nodes,
+or directly node ID prefixes are used to detect proximity. Tapestry, Pastry, and
+CAN successfully implemented the proximity into their algorithms by using this
+approach. The routing protocol in Pastry is based on longest node ID prefix
+matching, while CAN uses RTT values to detect close by nodes.
+\cite{castro_proximityp2p_2002} reports that proximity neighbour selection as an
+effective proximity based method.

### 2014-01-23 unstructured.tex
2012.08.20 (Vassilis) Added LaTeX tables (known issue on wrapping text inside cells). Reshaped discussion material. Introducing rough, algorithm performance, estimation (high, medium, low) by tracking efficiency, control overhead and scalability. First attempt on plotting data in a systematic way (needs more work though).

@@
 
-%###############################################################################
-%###############################################################################
 %       BROADCAST OPTIMISATION
-%###############################################################################
-%###############################################################################
 
@@
 
-% TODO: local indices is a cache based algorithm so maybe we have to put a
-% reference to the respective section???
 In local indices, as shown in Figure~\ref{figure:dbfs:localindx}, each node
@@
 leaves introduces a serious overhead to the system if the radius is kept broad.
+
+% \begin{center}
+% \begin{tabular}{ccc}
+% \textbf{Efficiency} & \textbf{Overhead} & \textbf{Scalability} \\
+% \hline
+% Iterative Deepening & ? & ? & ?
+% Directed BFS & ? & ? & ?
+% Local Indices & ? & ? & ?
+% \end{tabular}
+% \end{center}
+\begin{center}
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{Iterative Deepening} &
+% needs recalculation in every iteration step
+% extra delay imposed in a high churn environment that prohibits the algorithm
+% to start from the last level of nodes and results in a restart of the whole
+% process
+low &
+% the process in not started from the begining at each iteration step
+low &
+% 
+medium \\
+\emph{Directed BFS} &
+% Better time to satisfaction compared to iterative deepening
+medium &
+% Better quality and quicker results mean more aggregate bandwidth and
+% processing power needs to be consumed since more nodes are involved (death
+% spirral).
+medium &
+% BFS is Gnutella's non efficient communcation method
+medium \\
+\emph{Local Indices} &
+% efficiency is enhanced greatly since data pointers are replicated acros the
+% network
+high &
+% updating indices is a time and resource consuming process.
+medium &
+% the scalability is constrained by the indices update in a highly dynamic
+% environment (high churn)
+medium \\
+\end{tabular}
+\end{center}
 
@@
 
-% TODO: from the following (commented out) design components of Gia has one
-% part that falls into topology adaptation, one part on forwarding optimisation
-% and one part on
-
 %
@@
 %
+
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% Drastically reduces the search scope.
+low &
+% Only one copy of the query is issued to the network.
+% A token based flow control mechanism redirects queries to nodes that are most
+% likely to fulfill them
+low &
+% Reduced search scope makes this not really scalable
+low
+\end{tabular}
+\end{center}
 
@@
 %network.
-%Experiments in \cite{zhu_dcmp_2008} showed that DCMP incurs a lower delay,
-% returns more results and decreases the number of duplicate messages by 22\%,
-%compared to LTM. Additionally DCMP has one to two orders of magnitude less
-%overhead, because it adopts a more ``LAZY'' approach than broadcasting control
-%messages periodically like LTM does.
-%
-
-%###############################################################################
-%###############################################################################
-%###############################################################################
+%
+
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% It return more result than LTM according to experiments in
+% \cite{zhu_dcmp_2008}
+medium &
+% Duplicates are an important fraction of the redundant overhead imposed by
+% topology mismatch and thus DCMP minimises this as possible.
+% The control overhead is also one to two orders of magnitude less than LTM due
+% to its ``lazy'' broadcasting of control messages (as opposed to LTM's periodic
+% approach) - \cite{zhu_dcmp_2008}
+low &
+% fully distributed method
+high
+\end{tabular}
+\end{center}
+
 %       CACHING
-%###############################################################################
-%###############################################################################
-%###############################################################################
 
@@
 same for uniform and proportional models, and any approach between them would
-behave much better. Therefore they propose the square root allocation approach,
-which is a replication strategy that minimizes the expected search size of
-successful queries in P2P networks.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Tracing a large-scale Peer to Peer System: an hour in the
-life
+always behave much better. Therefore they propose the square root allocation
+approach, which is a replication strategy that minimizes the expected search
+size of successful queries in P2P networks.
+
+\begin{center}
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{Uniform Replication} &
+% Proportional and Uniform are the worst “reasonable” strategies
+low &
+% When item is created, replicate its key in a fixed number of hosts.
+low &
+%
+medium \\
+\emph{Proportional Replication} &
+% Proportional and Uniform are the worst “reasonable” strategies
+low &
+% for each query, replicate the key in a fixed number of hosts (need to know or
+% estimate the query rate)
+medium &
+%
+low \\
+\emph{Square Root Replication Allocation} &
+% All (strictly) in-between strategies are (strictly) better than Uniform and
+% Proportional - replication theory
+medium &
+% Assuming that each query keeps track of the search size then each time a query
+% is finished the object is copied to a number of sites proportional to the
+% number of probes that the search required.
+medium &
+% e.g. path replication is easy to scale, just detect the path along which the
+% query ultimatelly got fulfilled
+medium \\
+\end{tabular}
+\end{center}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Tracing a large-scale Peer to Peer System: an hour in the life
 of Gnutella}
@@
 
-%###############################################################################
-%###############################################################################
-%###############################################################################
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+%
+? &
+%
+? &
+%
+?
+\end{tabular}
+\end{center}
+
 %       OVERLAY OPTIMISATION
-%###############################################################################
-%###############################################################################
-%###############################################################################
 
@@
 }. Although it was not originally designed as a P2P system nor can be considered
-scalable in the contemporary sense of the term, Narada is considered pioneering
+scalable in the contemporary sense of the term, Narada has been pioneering
 (maybe along with \emph{Scattercast} \cite{chawathe_scattercast_2000}) in that
@@
 resource usage and the relative delay among end systems. Unfortunately, the most
-imporpant limitation of Narada is that although it works reasonably well for
+important limitation of Narada is that although it works reasonably well for
 small groups, it does not scale well for larger networks, therefore it is not
 suitable for, potentially very large, P2P file sharing application networks.
+
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+%
+low &
+% The overhead of Narada is proportional to the multicast group size
+high &
+%
+low
+\end{tabular}
+\end{center}
 
@@
 
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+%
+medium &
+% tracking of communication costs to all peer's neighbours as well as between
+% any pair of neighbours. Whenever a new neighbour cost table is received or
+% there is a change of neighbours, the source peer has to re-calculate the
+% multicast tree and apply the randomized AT algorithm.
+high &
+% needs global knowledge
+medium
+\end{tabular}
+\end{center}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 diameter from the source peer and optimizes the neighbour connections that are
-not in that tree. To achieve that it calculate cost between nodes using network
+not in that tree. To achieve that it calculates cost between nodes using network
 delay as a metric. Each peer probes the costs with its immediate logical
@@
 %response time (80 percent) to message queries without shrinking the search
-%scope. SBO, on the other hand, achieves approximately 85 percent reduction on
-%traffic cost and about 60 percent reduction on query response time. Last but
+%scope.Last but
 %not least, it is concluded that work must be done on incorporating a more
@@
 %the algorithm still suffers from the discrepancies caused by mislocated nodes.
+
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% compared to LTM it has slow convergence speed.
+medium &
+% A little bit better that AOTO since the computation here is done within a
+% certain diameter from the source peer
+high &
+% Larger diameter topologies lead to better topology optimization
+% rate but also to higher communication and computation overhead.
+medium
+\end{tabular}
+\end{center}
 
@@
 %
+
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% Authors claim 75% reduction on traffic cost and about 65% reduction on query
+% response time.
+medium &
+%
+medium &
+%
+medium
+\end{tabular}
+\end{center}
 
@@
 %
+%\cite{ni_mismatch_2004} shows SBO, achieves approximately 85 percent reduction
+% on traffic cost and about 60 percent reduction on query response time.
+
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% SBO reduces traffic cost between 85 and 90 percent
+high &
+% compared to LTM it has slower convergence speed thus incurres more, total,
+% overhead.
+medium &
+%
+medium
+\end{tabular}
+\end{center}
 
@@
 
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% 
+medium &
+% trivial overhead compared to the query cost savings and its convergent speed
+% is faster than minimum spanning tree approaches so less overall overhead cost.
+low &
+% completely distributed and needs no global knowledge,
+high
+\end{tabular}
+\end{center}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% similar traffic reduction rate savings to LTM
+medium &
+% The traffic overhead of HAND is much less than that of LTM by an average of
+% 55 percent.
+low &
+% no clock synching needed, fully distributed and with low control overhead the
+% algorithm can be considered scalable.
+high
+\end{tabular}
+\end{center}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % TODO: DOUBLE CHECK IF THIS ALGORITHM GOES IN THIS SECTION/SUBSECTION
-\subsubsection{Adaptive Peer Selection}
+\subsubsection{Adaptive Peer Selection (APS)}
 Bernstein et al., contributed on the use of machine learning for building peer
@@
 learning process and the complexity of the method.
+
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% machine learning the best way to adapt the overlay 
+high &
+% high computation overhead for learning and very slow convergence speed for
+% the learning process.
+high &
+%
+low
+\end{tabular}
+\end{center}
 
@@
 the IP-level message traffic generation.
+
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% low clustering with larger coverage of peers with the same number of messages
+% and reduced duplication.
+medium &
+% construction at peer join, while searching using local ``floods''
+low &
+% the selection of neighbours is done at join time. No overlay adaptation means
+% low scalability in dynamic environments.
+medium
+\end{tabular}
+\end{center}
 
@@
 
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% peer neighbourhood selected to minimise, for example, average delay
+high &
+% link state info (heartbeat) and delay estimation to all other nodes
+% resulting to a complete picture of the overlay graph.
+high &
+% need global knowledge
+low
+\end{tabular}
+\end{center}
+
 % TODO: CHECK IF THE FOLLOWING BITTORRENT ALGORITMS BELONG IN THIS SUBSECTION
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Biased Neighbour Selection}
+\subsubsection{Biased Neighbour Selection (BNS)}
 Bindal et al. \cite{rpc_bitbiased_2006} claim to strengthen BitTorrent
@@
 level itself, making the adoption difficult in the general case.
+
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% inter-ISP cost minimised while retaining the near optimal performance of
+% BitTorrent download.
+high &
+% this info is taken at swarm join
+low &
+% trackers should maintain and update proximity info/proximity info is
+% retrieved from special hardware.
+medium/high
+\end{tabular}
+\end{center}
 
@@
 
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% Authors claim minimisation of inter-ISP traffic cost while enhancing the,
+% already, near optimal performance of BitTorrent download.
+high &
+% periodic DNS lookups and CDN redirection imposes medium control overhead to
+% the method.
+medium &
+% the use of CDNs make the approach not fully distributed thus less scalable
+low
+\end{tabular}
+\end{center}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% minimisation of inter-AS traffic but unfair among peers.
+medium &
+% info (AS hop distances) aqcuired at swarm join
+low &
+%
+high
+\end{tabular}
+\end{center}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% it takes into account both dl speed and network topology. The later though is
+% rathier coarse grained since it is taken into account on the AS level.
+medium &
+% the algoritm does not involve extensive calculations.
+low &
+% Does not need infrastructure and the calculations are distributed so it can
+% scale
+high
+\end{tabular}
+\end{center}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 make the peer selection, per se, based on this knowledge. For the first part,
-the algorithm uses network tomography, a technique which suggests probing only
+the algorithm uses network tomography, a techni  que which suggests probing only
 from a large network's end points in order to infer its internal characteristics
@@
 conclusions.
+
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% coarse grained approach of the network tomography doesn't provide a detailed
+% picture of the underlying network that the peers can then exploit
+medium &
+% tracerout and hop counts and calculation from both the tracker and the peers
+% (the later for all peers in the bootstrapping list) increases the control
+% overhead of the approach.
+medium &
+%
+medium
+\end{tabular}
+\end{center}
 
@@
 
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% It incoorporates clustering of peers to inter- and intra ISP which is a
+% rather coarse grained method.
+medium &
+% only on peer arrival proximities to routers are estimated and sent to the
+% tracker. Even though the tracker is backing the identification of a peer
+% being close or distant the actual computation is done in a distributed mannar
+% thus not incurring too much additional controlling overhead.
+medium &
+% TODO: how intensive with regards to computing power is an online k-means (with
+% k=3) classification???
+medium
+\end{tabular}
+\end{center}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 therefore can be applied to both unstructured and structured overlay networks.
+
+\begin{center}
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{PROP-G} &
+% The authors show that an Internet-like underlying physical topology has much
+% better performance
+medium &
+% Exchanging neighbours within a radius of 2 (TTL value) reduces stretch
+% significantly, while keeping the additional overhead at very reasonable levels
+low &
+% The effectiveness of the scheme is reduced while the system size becomes
+% larger, but as the number grows, this reduction becomes smaller
+medium \\
+\emph{PROP-O} &
+% The algorithm can better reclaim the heterogenity of peers (those with faster
+% response) to reduce the system's aggregate delay. 
+medium &
+% Same as above
+medium &
+% The queries are directed through fast nodes so it is important such nodes
+% be part of the network
+medium \\
+\end{tabular}
+\end{center}
+
+% TODO: SOME DISCUSSION
+% PROP-G can be applied to both structured and unstructured overlays.
 
@@
 
+% TODO: CHECK ONCE AGAIN
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+%
+? &
+%
+? &
+%
+?
+\end{tabular}
+\end{center}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% Clustering approaches reduce search scope (or at least increase search
+% satisfaction time). The algorithm tries to ease this problem by allowing some
+% connections to be random
+medium &
+% overhead of split-hash and dnMatch is of low cost (TODO: double check this).
+% Walker message is multicasted only when sibling connections must be
+% established.
+low &
+% it can be utilised in hierarchical architectures as well meaning increased
+% scalability. Tests show that an increase from 5K- to 10K-node system slightly
+% increases the average hops needed to satisfy a lookupDN message.
+medium
+\end{tabular}
+\end{center}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Critical Topology-Aware Grouping (CTAG)}
@@
 
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% The technique is partitions the system reducing the search efficiency
+low &
+% IP-based clustering involves low overhead
+low &
+% Has average scalability mainly to its low overhead. In larger applications the
+% clustering of the system prevents the system from scaling smoothly.
+medium
+\end{tabular}
+\end{center}
+
 %###############################################################################
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Landmark Binning}\label{sec:landmark_binning}
+\subsubsection{Landmark Binning (LM)}\label{sec:landmark_binning}
 \begin{figure*}
@@
 
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% The technique is coarse grained thus doesn't achieve optimal results
+% (especially in small networks)
+medium &
+% The overhead is confined to the communication of a node with, most probably,
+% 12 landmark servers at most.
+low &
+% The introduction of landmark servers renders the approach not fully
+% decentralised, thus preventing it from scalling smoothly. Communicating
+% and overloading landmark servers in high-churn systems is another scalability
+% concern.
+low
+\end{tabular}
+\end{center}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% The technique is coarse grained thus doesn't achieve optimal results
+% (especially in small networks)
+medium &
+% The algorithm is iterative through the available groups. At each group
+% probing of a candidate list must be performed. This process is done at
+% bootstrapping time so the overhead increases in high-churn systems
+medium &
+% The introduction of dynamic landmark servers renders this approach much more
+% scalable than the traditional static landmarking techniques 
+medium
+\end{tabular}
+\end{center}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\renewcommand\arraystretch{1.4}% (MyValue=1.0 is for standard)
-
-%\begin{figure}[h!]
-\hspace{-3ex}
-\begin{center}
-\footnotesize
-%\begin{tabular}{
-%\begin{landscape}
-\begin{longtable}{
-|>{\columncolor[gray]{.7}}m{0.2\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.4\columnwidth}
-|>{\columncolor[gray]{.8}}m{0.2\columnwidth}
-|>{\columncolor[gray]{.9}}m{0.2\columnwidth}
-%|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
-%|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
-%|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
-%|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
-|}
-\caption{Decentralized Unstructured Algorithms} \label{fig:unstruct_compare_table} \\
-\hline
-\rowcolor[gray]{.5}
-\textbf{Algorithm} &  \textbf{Overlay structure} & \textbf{Base protocol} &
- \textbf{Scalability}\\
-\hline
-\endfirsthead
-\multicolumn{4}{c}%
-{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
-\hline
-\rowcolor[gray]{.5}
-\textbf{Algorithm} &  \textbf{Overlay structure} & \textbf{Base protocol} &
- \textbf{Scalability}\\
-\hline
-\endhead
-\hline \multicolumn{4}{r}{\textit{Continued on next page}} \\
-\endfoot
-\hline
-\endlastfoot
-\textbf{Narada} & \textbf{Overlay optimization
-based}. Creates a mesh (richer connected graph) and builds minimum spanning
-trees on this mesh & & Small and sparse groups \\
-
-\hline
-\textbf{Gia} & \textbf{Broadcast based} Replaces
-Gnutella flooding with random walk, and introduces KaZaA style supernodes. Uses
-dynamic topology adaptation protocol &
- Gnutella &  Better than Gnutella  \\
-
-\hline
-\textbf{Adaptive Overlay Topology Optimization} & \textbf{Overlay optimization
-based}. Creates overlay multicast tree with Selective Flooding protocol&
-Gnutella &  Better than Gnutella \\
-
-\hline
-\textbf{Location-aware Topology Matching} &
-\textbf{Overlay Optimization Based}. Uses \textit{TTL2-detector flooding}, \textit{low productive
-connection cutting}, and \textit{source peer probing}. & Gnutella &  Better than Gnutella \\
-
-\hline
-\textbf{Replication Strategies in Unstructured P2P Networks} &
-\textbf{Cache Based}. Uses uniform, proportional and square root allocation
-strategies to replicate data. & Gnutella &  Better than Gnutella \\
-
-\hline
-\textbf{Tracing a large-scale Peer to Peer System: an hour in the life of Gnutella.} &
-\textbf{Cache Based}. Proposes a caching algorithm based on the traces of the Gnutella traffic & Gnutella & Better than Gnutella \\
-
-\hline
-\textbf{Improving search in P2P networks} &
-\textbf{Broadcast Based}. Uses \textit{iterative deepening}, \textit{directed
-BFS}, and \textit{local indices} to improve efficiency. & Gnutella &  Better than Gnutella \\
-
-\hline
-\textbf{Distributed Cycle Minimization Protocol} &
-\textbf{Broadcast based} Uses a decentralized cycle elimination protocol  &  &  \\
-
-\hline
-\textbf{Scalable Bipartite Overlay} &
-\textbf{Overlay optimization based} Uses bipartite partition graph and builds
-local minimum spanning trees  & Gnutella  & Better than Gnutella \\
-
-\hline
-\textbf{Adaptive Connection Establishment} &
-\textbf{Overlay optimization based} Forms Neighbour Cost Tables, builds local
-minimum spanning trees and perform local optimizations & Adaptive Overlay
-Topology Optimization (AOTO), Gnutella & Better than Gnutella \\
-
-\hline
-\textbf{Hops Adaptive Neighbour Discovery} &  & &  \\
-
-\hline
-\textbf{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} &
-\textbf{Overlay optimization based} Uses piggybacking to discover neighbour
-distances and selects neighbours  & Gnutella  & \\
-
-\hline
-\textbf{mOverlay} &\textbf{Landmark based proximity} Uses dynamic landmarks to find node locality
-& & Due to dynamic landmarks and grouping, more scalable than tree-based or mesh-based protocols \\
-
-\hline
-\textbf{Distributed Domain Name Order (DDNO)} &
-\textbf{Overlay optimization based} Connects half of the nodes connections to
-the nodes in the same domain and the other half to random nodes, therefore
-supports locality and topological connection  &  & Yes, by using super
-peers \\
-
-\hline
-\textbf{Peer-exchange Routing Optimization Protocols} & \textbf{Overlay optimization based} Optimizes overlay by the exchange of
-neighbors among peers  & Can work with both decentralized structured and
-unstructured architecture & Yes \\
-
-\hline
-\textbf{MAY OMIT - I CHANGED IT TO STRUCTURED SINCE THERE IS A REFERENCE FOR
-DHT (OF COURSE IT MIGHT POSIBLE TO BE APPLIED TO BOTH. MAYBE NEED TO CHECK) -
-T2MC} &
-\textbf{Overlay optimization based} Uses traceroute results for clustering the
-nodes  & & \\
-
-\hline
-\textbf{Unnamed-unstructured} &
-\textbf{Overlay optimization based} Minimizes the communication delay and
-maximizes the broadcasting range & & Better than THANCS and mOverlay \\
-
-\hline
-\textbf{Landmark Binning} & \textbf{Landmark based proximity} Uses network latency to partition
-nodes into bins & Can work with both decentralized structured and unstructured architecture & \\
-
-\hline
-%\end{tabular}
-\end{longtable}
-%\end{landscape}
-\end{center}
-\vspace{-2.5ex}
-\vspace{-2.5ex}
-%\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Discussion on the Algorithms for Unstructured Architectures}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
- In short, the methods highlighted here are divided into the following
-groups
-\begin{inparaenum}[\itshape i\upshape)]
-  \item broadcast optimisation
-  \item caching,
-  \item overlay optimisation, and
-  \item landmark -based
-\end{inparaenum}
- approaches.
-
-%###############################################################################
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Broadcast Optimisation} 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% TODO: TAKEN FROM ALGORITHMS SUBSECTION. SHOULD BE MERGED WITH DISCUSSION
-
+In this section we introduced the topology mismatching problem from the
+perspective of decentralised unstruThe introduction of landmark servers renders
+the approach not fully
+% decentralised, thus preventing it from scalling smoothly. Communicating
+% and overloading landmark servers in high-churn systems is another scalability
+% concern.ctured architectures and we visited a series
+of algorithms that have been proposed in the literature to tackle its negative
+influence to the networks' performance. Even though, at first glance, the
+algorithms seem to do things their own way actually we can distinguish four main
+methodologies that are being utilised in order for every algorithm to achive its
+intended result. These methodologies are:
+\begin{enumerate}[\itshape i\upshape)]
+  \item topology adaptation
+  \item forwarding optimisation
+  \item caching and replication, and
+  \item landmarking
+\end{enumerate}
+
+Table~\ref{unstructured:table} offers a summarised overview of the algorithms
+we saw earlier in the section and tracks down the afformentioned methodologies
+each of them incorporates along with their highlights and a rough estimation of
+positive and negative aspects of their implementation.
+
+\input{unstructured-table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Topology adaptation}
+Researchers initially started considering alternatives for services who
+implement one-to-many communication schemes, like IP multicast, on the
+application layer (i.e. Narada). Application level multicast protocols
+are generic protocols that can be applied to both P2P file sharing systems, as
+well as content distribution networks. Application layer multicast, however, is
+not as efficient as the one enabled by IP and still struggles to ease the
+difficulties imposed by the topology mismatch problem.
+
+The topology adaptation based protocols modify the topology of the P2P network
+using various techniques. The two commonly used approaches that have been
+investigated in this section are creating \emph{spanning tree}s using connection
+graphs, and creating \emph{clusters} of physically close nodes.
+
+Narada and subsequent algorithms like AOTO, LTM, SBO, try to the problem by
+building a richer connected graph and forming minimum spanning trees over this
+graph that can efficiently route messages among peers. Even though spanning
+trees provide efficient query performance, the construction and the update costs
+of the spanning tree, especially in dynamic environments with many nodes joining
+and leaving the network, cause large traffic overhead to the corresponding
+network \cite{chu_esm_2000,chu_esm_2002}. Thus, the original Narada algorithm is
+designed for small groups therefore it is not scalable. Its followers try to
+solve the scalability problem by introducing various methods including forming
+minimum spanning trees for the two-hop neighbours ($N^2$) of each node,
+partitioning the graph into two random groups where each group is responsible of
+different tasks, or performing local optimizations dynamically on the overlay
+graph. The advantage of building minimum spanning trees is that they maintain
+the connectivity on the network in an efficient manner, while still preserving
+the overall search scope. However, building and maintaining a minimum spanning
+tree creates additional overhead for the network.
+
+The cluster based approaches, on the other hand, select to link physically
+closer nodes with each other. T2MC, for example, uses traceroute logs and DDNO
+uses domain names to cluster close by nodes. Unfortunatelly, commonly used
+methods for proximity detection across Internet do not always return
+reliable results, therefore mapping accuracy is not guaranteed. Moreover,
+traceroute for example, is a heavy weight utility to be frequently used on the
+network, let alone the fact that many network equipment vendors do not allow
+traceroute calls at all. The most problematic aspect of clusterting, though, is
+its nature per se. Limited connectivity among the various local domains can
+significantly shrink the search scope, negatively affecting the query response
+time the P2P end-user experiences. Of course, techniques that try to round up
+such edges and balance the efficiency of the clustering approach with improved
+connectivity, have also been proposed, like DDNO, which attempts to address this
+limitation by forcing half of each node's connections to be with other, randomly
+selected, nodes over the network
+
+% These approaches include creating spanning trees using
+%connection graphs, creating cluster of physically close nodes, or using latency
+%information to detect proximity. The brief description of each category is
+%presented below:
+%
+%  \begin{itemize}
+%    \item \emph{Spanning tree based}. These approaches construct of a rich
+%    graph  based on the network connections and build minimum spanning tree
+%    on the graphs, causing large traffic overhead to the system
+%    \cite{chu_esm_2000,chu_esm_2002}.
+%
+%    \item \emph{Cluster based}. These approaches select to link physically
+%    closer nodes with each other, therefore shrink the search scope
+%    significantly while mapping accuracy is not always guaranteed.
+%
+%    \item \emph{Minimum latency first}. Use of latency as a metric to calculate
+%    distance among peers. They require global latency information
+%    ``landmarks''\footnote{Measuring latency between peers and stable Internet
+%    servers.}.
+%
+%  \end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Forwarding optimisation}
 In the first realisation of the Gnutella protocol, each query request received
-by a peer was then forwarded to all of its neighbours which, clearly, generated
-unnecessary traffic over the communication links of all the network. The
-algorithms that fall into the group discussed in this section are referred
-to as broadcast optimisation -based because they enhance the ``naive''
-broadcasting approach, described above, and prefer to forward the messages to a
-selected subset of a node's outgoing links. The selection criteria vary
-depending on the algorithm but they mainly use one or multiple statistical
-metrics that span from some obvious ones, like the latency of the link between
-the nodes, to more sophisticated alternatives that may even take into account
-application level requirements such as the reliability of the neighbour.
+by a peer was then forwarded to all of its neighbours thus generating
+unnecessary traffic over the network. Forwarding-optimisation-based approaches,
+propose intelligent neighbour selection for a message's next hop across the
+routing path, in order to replace that inefficient \emph{blind flooding} scheme
+of Gnutella's. The selection criteria vary depending on the algorithm, but they
+mainly use one or multiple statistical metrics that span from some obvious ones,
+like the latency of the link between the nodes or the exposure of specific
+features such as high processing bandwidth, high capacity, or low latency, to
+more sophisticated alternatives that may even take into account application
+level requirements such as the reliability of the neighbour.
 These appoaches have the advantage of enhancing the search responsiveness but on
 the other hand suffer from the drastic reduction of the search scope thus
-limiting the scalability of the whole network.
+limiting the scalability of the whole network. This, in practice reduces the
+aggregate resource usage of the P2P network and improves the overall performance
+of the system. On the other hand, the algorithms do not offer a real solution to
+the topology mismatch problem. Thus, none can guarantee that overlay and
+underlying topologies are aligned with each other let alone quantify the
+mismatch and try to alleviate. Broadcast-based methods can be used in
+conjunction with other approaches to improve the quality of the P2P systems.
 
@@
 
-%###############################################################################
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Caching}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% TODO: TAKEN FROM ALGORITHMS SUBSECTION. SHOULD BE MERGED WITH DISCUSSION
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Caching and replication}
+Caching is a widely used technique to exploit locality and minimise redundant
+transfer of data, initially and successfully adopted by web and file server
+application environments. Since peers in a P2P system also operate as
+servers, it is intuitively expected that P2P file sharing systems can also
+benefit from caching in improving performance and reducing overall resource
+usage. However, the design of caches in this context is non-trivial compared to
+the web-based caching. Exactly, because each node can act as both a server and a
+client, two important issues have to be considered at design time. First, the
+lifetime of a query is short, as the nodes join and leave frequently. Second,
+the result of a single query string is not always the same, as it is dependent
+on the source of the query, the TTL value set for the messages, the current
+interconnection of peers and the high volatility of the environment. So, in
+order to develop a successful caching system for a P2P architecture, these
+parameters also have to be considered. P2P caching/replication can be applied
+on two different levels, namely caching indices or pointers to data or caching
+the data itself. Successfull implementations have already been developed in
+some commercial P2P systems, like KazaA, which uses caching of indices on its
+super-peer level. Unfortunatelly, even though the state of the art in P2P
+algorithms using caching methods reduce the resource usage of the network, they
+also fail to consistently address the probelem of the topology mismatch.
 
@@
 %  \end{itemize}
-%
-Caching is a widely used technique to exploit locality and minimise redundant
-transfer of data, initially adopted by web and file server application
-environments with great success. Since peers in a P2P system also work as
-servers, it is intuitively expected that P2P file sharing systems can also
-benefit from caching. However, two of the main characteristic properties of P2P
-systems, nodes frequently joining and leaving and query lifetime relatively
-short (at least compared to those needed to be served by web servers) , makes
-caching in P2P networks a non-trivial issue. There are usually two levels of
-caching possible in file sharing systems, caching indices or pointers to data or
-caching the data itself. Caching is already implemented and successfully used in
-some commercial P2P systems. One popular example is KazaA, which uses caching of
-indices on its super-peer level. In the paragraphs below, caching-based
-algorithms are presented and explained in more depth.
-
-%###############################################################################
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Overlay Optimization}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% TODO: TAKEN FROM ALGORITHMS SUBSECTION. SHOULD BE MERGED WITH DISCUSSION
-
-% These approaches include creating spanning trees using
-%connection graphs, creating cluster of physically close nodes, or using latency
-%information to detect proximity. The brief description of each category is
-%presented below:
-%
-%  \begin{itemize}
-%    \item \emph{Spanning tree based}. These approaches construct of a rich
-%    graph  based on the network connections and build minimum spanning tree
-%    on the graphs, causing large traffic overhead to the system
-%    \cite{chu_esm_2000,chu_esm_2002}.
-%
-%    \item \emph{Cluster based}. These approaches select to link physically
-%    closer nodes with each other, therefore shrink the search scope
-%    significantly while mapping accuracy is not always guaranteed.
-%
-%    \item \emph{Minimum latency first}. Use of latency as a metric to calculate
-%    distance among peers. They require global latency information
-%    ``landmarks''\footnote{Measuring latency between peers and stable Internet
-%    servers.}.
-%
-%  \end{itemize}
-
-The overlay optimization based protocols modify the topology of the P2P network
-using various techniques. The two commonly used approach that is investigated in
-this section are creating spanning trees using connection graphs, and creating
-clusters of physically close nodes. Spanning tree based approaches construct
-rich graphs based on the network connections and build minimum spanning trees on
-these graphs. Even though spanning trees provides efficient query performance,
-the construction and the update costs of the spanning tree, especially in
-dynamic environments with many nodes joining and leaving the network, cause
-large traffic overhead to the corresponding
-network \cite{chu_esm_2000,chu_esm_2002}. The cluster based approaches on the
-other hand select to link physically closer nodes with each other, therefore
-significantly shrinking the search scope of the system. Added to that, commonly
-used methods for correct determination of physical node positions across
-Internet do not always return reliable results, therefore mapping accuracy is
-not guaranteed. Bellow, the details of the algorithms that use the spanning tree
-based or cluster based topology optimization methods are presented.
-
-%###############################################################################
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Landmark Based Proximity}\label{sec:landmark}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Detecting the proximity of a node using landmarks, also known as
-\textit{Landmark Clustering}, is based on the view that nodes with similar
-distances to a set of predefined well-known landmark nodes are pretty likely
-being close to each other. But this approach has its weaknesses itself, like the
-fact that is a rather coarse grained approximation, therefore not particularly
-well suited for detecting the correct positions of nodes within close distance
-to each other.
-
-%###############################################################################
-
-In this section we presented the state of the art for the algorithms applied to
-unstructured P2P systems. We categorised it based on their philosophy for
-alleviating the topology mismatch problem into four major categories namely
-\begin{inparaenum}[\itshape i\upshape)]
-  \item broadcast-based,
-  \item cache-based,
-  \item overlay optimisation-based and
-  \item landmark-based approaches
-\end{inparaenum}
-. Here we present a overall discussion on the above categories trying to point
-out the advantages, disadvantages and novelties introduced by each and by all of
-them. A quick overview can be seen in Table \ref{fig:unstruct_compare_table}.
-
-Broadcast-based approaches in general propose intelligent neighbour selection in
-order to replace the inefficient blind flooding first used in the Gnutella
-protocol. Instead of forwarding the queries to all their outgoing links they
-dictate the intelligent selection of neighbours with either higher probability
-to answer a query than others or neighbours  that expose specific features such
-as high bandwidth, high capacity, or low latency. This, in practice reduces the
-aggregate resource usage of the P2P network and improves the overall performance
-of the system. On the other hand, the algorithms do not offer a real solution to
-the topology mismatch problem. Thus, none can guarantee that overlay and
-underlying topologies are aligned with each other let alone quantify the
-mismatch and try to alleviate. Broadcast-based methods can be used in
-conjunction with other approaches to improve the quality of the P2P systems.
-
-The success of caching, as a well studied method in client-server Internet
-applications, also made it a good candidate for the P2P networks. Even
-though caching improves the performance, and reduces the overall resource usage
-of P2P systems, the design of caches in this context is non-trivial compared to
-the web-based caching. Due to the unique, inherent characteristic of all P2P
-systems that, each node can act as both a server and a client, two important
-issues have to be considered when designing caches. First, the lifetime of a
-query is short, as the nodes join and leave frequently. Second, the result of a
-single query string is not always the same, as it is dependent on the source of
-the query, the TTL value set for the messages, the current interconnection of
-peers and the high volatility of the environment. So, in order to develop a
-successful caching system for a P2P architecture, these parameters also have to
-be considered. Even though the state of the art P2P algorithms using caching
-methods reduce the resource usage of the network, they also fail to consistently
-address the probelem of the topology mismatch.
-
-Overlay-optimisation-based  can further be devided into
-minimum-spanning-tree-based and clustering-based. Researchers initially started
-considering alternatives to IP multicast for the application layer, which led
-Narada, that simulates the same one-to-many communication approach requiring IP
-multicast services. Application layer multicast, however, is not as efficient as
-the one enabled by IP and still struggles to ease the difficulties imposed by
-the topology mismatch problem. Narada, and its followers (AOTO, LTM, SBO) try to
-solve the topology mismatch problem by building a richer connected graph and
-forming minimum spanning trees over this graph that can efficiently route
-messages among peers. Application level multicast protocols, pioneered by
-Narada, are generic protocols that can be applied to both P2P file sharing
-systems, as well as content distribution networks. The original Narada algorithm
-is designed for small groups therefore it is not scalable. Its followers try to
-solve the scalability problem by introducing various methods including forming
-minimum spanning trees for the two-hop neighbours ($N^2$) of each node,
-partitioning the graph into two random groups where each group is responsible of
-different tasks, or performing local optimizations dynamically on the overlay
-graph. The advantage of building minimum spanning trees is that they maintain
-the connectivity on the network in an efficient manner, while still preserving
-the overall search scope. However, building and maintaining a minimum spanning
-tree creates additional overhead for the network. The alternative method used in
-overlay optimization is the cluster based approache. According to that, nodes
-use various methods to detect physically close nodes to form clusters. T2MC uses
-traceroute logs and DDNO uses domain names to cluster close by nodes. However,
-the accuracy of the methods used to form clusters directly affect the success
-rate of the algorithm. Traceroute for example is a heavy weight utility to be
-frequently used on the network, let alone the fact that many network equipment
-vendors do not allow traceroute calls. The major problem with cluster based
-approaches is that the limited connectivity within the local domains shrink the
-search scope dramatically, which negatively affects the search performance of
-the P2P system. DDNO addresses this limitation by allowing half of each nodes
-connections to be to random nodes over the network, which balances the
-efficiency of the clustering approach with improved connectivity.
-
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Landmarking}\label{sec:landmark}
 In landmark-based algorithms, nodes use network delay (e.g. RTT) as a
@@
 static landmark servers, still the measurement accuracy problem affects the
-overall performance of the system.
+overall performance of the system. Moreover,it can be characterised as a rather
+coarse-grained approximation, therefore not particularly well suited for
+detecting the correct positions of nodes within close distance to each other.

### 2014-01-26 background.tex
2012.08.20 (Vassilis) Added LaTeX tables (known issue on wrapping text inside cells). Reshaped discussion material. Introducing rough, algorithm performance, estimation (high, medium, low) by tracking efficiency, control overhead and scalability. First attempt on plotting data in a systematic way (needs more work though).
2012.10.08 (Vassilis) Text added explaining the three indicators (low, medium, high) for each of the criteria against which the algorithms are evaluated (efficiency, overhead, scalability). Filled in missing references and question marks in the tables and text. Intro and back

@@
 
-In this section, we provide some background information on the types
-of P2P overlay architectures that have been implemented and deployed.  We also
-describe and motivate more formally the topology mismatch problem tackled
-by many P2P researchers over the past several years.
+In this section, we provide some background information on the types of
+peer-to-peer architectures that have been implemented and deployed. We also
+describe and motivate more formally the topology mismatch problem tackled by
+many P2P researchers over the past several years.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Circa 2000, the first P2P file-sharing systems introduced the notion 
-of the \emph{servent} concept \cite{gnutella}, a \emph{portmanteau} that blends
-the notions of server and client to denote the twofold role of the P2P network
-participants. Such serverless systems have proven to be able to achieve outstanding
+Circa 2000, the first P2P file-sharing systems introduced the \emph{servent}
+concept \cite{gnutella}, a \emph{portmanteau} that blends the notions of server
+and client to denote the twofold role of a node participating in the P2P
+network. Such serverless systems have proven to be able to achieve outstanding
 aggregate resource capacities as more and more participants join the system
-without requiring additional expenditure for infrastructure.  While
-certain undesirable features such as 
+without requiring additional expenditure for infrastructure. While certain
+undesirable features such as 
 free-riding~\cite{saroiu_measurefileshare_2002,adar_gnutellafreeriders_2000,hughes_gnutellafreeride_2005},
-the distribution of illegal content and/or other \emph{socio-technical}
-issues \cite{hughes_socp2p_2008}, have arisen, P2P file-sharing systems 
-have continued
-to gain popularity because they have successfully supported
-popular
-applications amongst vast numbers of participating
+the distribution of illegal content and other \emph{socio-technical}
+issues \cite{hughes_socp2p_2008}, have arisen, peer-to-peer systems have
+continued to gain popularity mainly because they have successfully supported
+popular applications (i.e. file-sharing) amongst vast numbers of participating
 users.
 
-For over a decade, excitement over the immense potential of P2P systems
-has generated a flurry of research and development, resulting in a wide 
-range of popular protocols, networks, and applications. As P2P systems evolved
-three main architectures have been emerged, namely
+For over a decade, excitement over the immense potential of P2P systems has
+generated a flurry of research and development, resulting in a wide range of
+popular protocols, networks, and applications. As P2P systems evolved, three
+main architectures have emerged, namely
 \begin{inparaenum}[\itshape i\upshape)]
@@
 
-Designers of \emph{centralized} architectures (see Figure~\ref{figure:p2p-archs:centralised})
-were the first to recognize that requests for resources (e.g. CPU 
-cycles~\cite{seti} or popular content
-% TODO find citation for Napster
-\cite{}) need not be sent to a dedicated server but instead could be handled by
-the many hosts that already posess the resources. The downside of
-this approach is it requires a centralized searching facility which becomes
-a single point of failure, a scalability bottleneck, and a target of malicious 
-attacks (e.g., \emph{Denial-of-Service (DoS) attacks}).
+Designers of \emph{centralized} architectures
+(seeFigure~\ref{figure:p2p-archs:centralised}) were the first to recognise that
+requests for resources (e.g. CPU cycles or popular content) need
+not be sent to a dedicated server but instead could be handled by
+the many hosts that already posess the resources. The downside of this approach
+is that it requires a centralized searching facility which becomes a single
+point of failure, a scalability bottleneck, and a target of malicious attacks
+(e.g., \emph{Denial-of-Service (DoS) attacks}).
 
-The most successful incarnation of the centralized approach is
-the file-sharing application \emph{Napster}
-% TODO find citation for Napster again
-\cite{}.
+The most successful incarnation of the centralized approach is the file-sharing
+application \emph{Napster}. % TODO find citation for Napster again \cite{}
 Napster maintained a \emph{central index server} based on file lists provided by
@@
 returned pointers to the actual content. Thus, by centralizing search while
-distributing downloads, Napster achieved a highly functional design that 
-at its height, attracted 26 million users sharing 80 million 
-songs~\cite{jmm_naptopusage_2001}.  
-Napster's centralized nature was ultimately what brought it down.
-The Recording Industry Association of
-America (RIAA) 
-filed a lawsuit for copyright infringement of the intellectual
-property of artists whose music was distributed by the Napster application.
-RIAA was able to point to the centralized indexing component of
-Napster as the entity responsible for illegal song-trading.
+distributing downloads, Napster achieved a highly functional design that, at its
+height, attracted 26 million users sharing approximately 80 million
+songs~\cite{jmm_naptopusage_2001}. Napster's centralized nature was ultimately
+what brought it down after Recording Industry Association of America (RIAA) 
+filed a lawsuit for copyright infringement of the intellectual property of
+artists whose music was distributed by the Napster application. RIAA was able to
+point to the centralized indexing component of Napster as the entity responsible
+for illegal song-trading.
 
-The centralized approach was quickly replaced by architectures that
-distribute both search and resource provision capabilities (see
-Figures~\ref{figure:p2p-archs:unstructured-full} and
-\ref{figure:p2p-archs:unstructured-hierarchical}). In these \emph{decentralized}
+The centralized approach was quickly replaced by architectures that distribute
+both search and resource provision capabilities (see
+Figure~\ref{figure:p2p-archs:unstructured-full}). In these \emph{decentralized}
 architectures, resource placement within the overlay topology is
-random~\cite{yang_improvep2psearch_2002}. For this
-reason they are referred to as \emph{unstructured}. The most important
-properties of such systems are that they support inherent heterogeneity of
-peers, are highly resilient to peer failures, and incur low maintenance overhead
-at handling the dynamics of peer participation \cite{stutzbach_churn_2006}. In
-the literature, they are also known as \emph{broadcast-based} systems, because
-they use message flooding among peers to propagate
-search queries. A well-known example of the fully dentralized unstructured 
-approach is Gnutella, in its first realisation (v.0.4) \cite{gnutellav04}.
+random~\cite{YG-M2002}. For this reason they are referred to as
+\emph{unstructured}. The most important properties of such systems are that they
+support inherent heterogeneity of peers, are highly resilient to peer failures,
+and incur low maintenance overhead at handling the dynamics of peer
+participation \cite{stutzbach_churn_2006}. In the literature, they are also
+known as \emph{broadcast-based} systems, because they use message flooding among
+peers to propagate search queries. A well-known example of the fully dentralized
+unstructured approach is Gnutella, in its first realisation (v.0.4)
+\cite{gnutellav04}.
 
-A key disadvantage of the unstructured approach
-is that message flooding burdens the network since queries travel within the
-network randomly, visiting nodes that do not have the resource of interest
-and thus wasting bandwidth.  To save on bandwidth, unstructured networks
-typically limit how long a query travels within the overlay (via a
-time-to-live (TTL) parameter).   When a query is received by a node, it decreases
-the TTL by one unit.  If the TTL has not reached zero, the node
-forwards the query to its neighbors.  Otherwise, it drops the query.  While
-this approach prevents queries from visiting the entire network, it does
-not guarantee that the resource of interest, (e.g., a rare recording 
-of a musical group from
-the 1930s) will be found.
+A key disadvantage of the unstructured approach is that message flooding burdens
+the network since queries travel within the network randomly, visiting nodes
+that do not have the resource of interest and thus wasting bandwidth.  To save
+on bandwidth, unstructured networks typically limit how long a query travels
+within the overlay (via a time-to-live (TTL) parameter). When a query is
+received by a node, it decreases the TTL by one unit. If the TTL has not reached
+zero, the node forwards the query to its neighbours.  Otherwise, it drops the
+query. While this approach prevents queries from visiting the entire network, it
+does not guarantee that the resource of interest, (e.g., a rare recording 
+of a musical group from the 1930s) will be found.
 
@@
 unstructured systems evolved into ``hierarchical'' systems that dynamically
-assign indexing functions to special peers,
-called \emph{ultra-peers}  or \emph{super-peers}.
-Peers with reliable network connections
-and/or high compute power became ultra/super-peers facilitating less 
-powerful peers (called \emph{leaf-nodes}) to find resources of interest.
-Well-known examples of these hierarchical unstructured systems include
-Gnutella v.0.6~\cite{gnutella}, Limewire[XXX], and KaZaA[XXX], and Skype[XXX]. 
-%% TODO: add references for the above line.
-While the hierarchical approach relieves network pressure as search
-queries are flooded over a much smaller subset of peers (the ultra/super-peers),
-the disadvantages of reduced search scope and inability to locate
-rare resources efficiently remain.
+assign indexing functions to special peers, called \emph{ultra-peers}  or
+\emph{super-peers} (see
+Figure~\ref{figure:p2p-archs:unstructured-hierarchical}). Peers with reliable
+network connections and/or high compute power became ultra/super-peers
+facilitating less powerful peers (called \emph{leaf-nodes}) to find resources of
+interest. Well-known examples of these hierarchical unstructured systems include
+Gnutella v.0.6/Limewire~\cite{gnutella}, KaZaA~\cite{kazaa} and
+Skype~\cite{skype}. While the hierarchical approach relieves network pressure as
+search queries are flooded over a much smaller subset of peers (the
+ultra/super-peers), the disadvantages of reduced search scope and inability to
+locate rare resources efficiently remain.
 
@@
 \emph{decentralized structured} schemes have been proposed (see
-Figure~\ref{figure:p2p-archs:structured}).  The key goal
-of these systems is to provide a self-organising
-infrastructure for large-scale P2P applications
+Figure~\ref{figure:p2p-archs:structured}). The key goal of these systems is to
+provide a self-organising infrastructure for large-scale P2P applications
 \cite{ratnasamy_can_2001,stoica_chord_2001,antony_pastry_2001,zhao_tapestry_2001,maymounkov_kademlia_2002,rgrk_bamboo_2004}.
 They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes
-through a deterministic mechanism. 
-The DHT  
-provides a guaranteed bound on the number of overlay routing
-hops that have to be taken to locate a resource, even in the
-case where only a single copy of the resource exists in the system.  This means
-$O \left ( log n \right )$ hops, compared to unstructured networks that require
-$O \left ( n \right )$ to reliably locate a specific object.
-Unfortunately this efficiency does not come without cost. Structured overlays
-can only support exact-match queries (i.e., queries that identify resources
-by name) as opposed to content-based
-retrieval provided by unstructured overlays.  This means 
-that unstructured overlays can support more 
-versatile resource location queries such as keyword and partial matching
-searches.  Several popular file-sharing peer-to-peer applications have
-been based on DHTs including Overnet[], Kademlia[], the Kad Network[], 
-some flavors of BitTorrent, and many
-more.  Moreover, DHTs have sparked a flurry of research activity and 
-development for a large range of applications: from distributed search 
-engines to event notification infrastructures to cloud-based platforms.
-We cite several examples in the Introduction.
+through a deterministic mechanism. The DHT provides a guaranteed bound on the
+number of overlay routing hops that have to be taken to locate a resource, even
+in the case when only a single copy exists in the system.  This means
+$O \left( log n \right)$ hops, compared to unstructured networks that require
+$O \left( n \right)$ to reliably locate any object. Unfortunately this
+efficiency does not come without cost. Structured overlays can only support
+exact-match queries (i.e., queries that identify resources by name) as opposed
+to content-based retrieval provided by unstructured overlays. This means
+that unstructured overlays can support more versatile resource location queries
+such as keyword and partial matching searches. Several popular file-sharing
+peer-to-peer applications have been based on DHTs including
+Overnet/eDonkey~\cite{overnet}, Kademlia/Kad
+Network~\cite{maymounkov_kademlia_2002}, some flavors of
+BitTorrent~\cite{c_bittorrent_2003}, and many more.  Moreover, DHTs have sparked
+a flurry of research activity and development for a large range of applications:
+from distributed search engines to event notification infrastructures to
+cloud-based platforms.We cite several examples in the Introduction.
 
@@
 % TODO: SOME DISCUSSION ON STRUCTURED OVERLAYS PROBABLY FROM
-%       \cite{wu_laptop_2007}
+%       \cite{WLH2007}
 % However, the design of a decentralized but structured P2P network has to
@@
 mapping of the overlay structure to the underlying infrastructure. Remember, in
-Figure~\ref{figure:overlay} Nodes $A$ and $B$ are on the same local network,
+Figure~\ref{figure:overlay} nodes $A$ and $B$ are on the same local network,
 while $C$ and $D$ are in different networks. The top layer represents the
@@
 \rightarrow 2 \rightarrow 1 \rightarrow 7 \rightarrow 8 \rightarrow 5
-\rightarrow 5 \rightarrow 4$. In the alternative overlay of Figure
-~\ref{figure:overlay-2}, the path will be: $3 \rightarrow 8 \rightarrow 4$. The
-total cost of the first sequence is $45 ms$, while the second yields a mere $15
-ms$. Therefore, we can conclude that the second overlay is more congruent with
-the underlying physical network than the first one and, thus, is more efficient.
+\rightarrow 5 \rightarrow 4$. In the alternative overlay of
+Figure~\ref{figure:overlay-2}, the path will be: $3 \rightarrow 8 \rightarrow
+4$. The total cost of the first sequence is $45 ms$, while the second yields a
+mere $15 ms$. Therefore, we can conclude that the second overlay is more
+congruent with the underlying physical network than the first one and, thus, is
+more efficient.
 
@@
 Ideally, an overlay network should be constructed to achieve an optimal 
-mapping among the
-overlay paths and the underlying IP network links to avoid inefficient states.
-The problem of constructing an optimal overlay is referred to as the
-\emph{topology mismatch problem}, and is formally defined as follows:
+mapping among the overlay paths and the underlying IP network links to avoid
+inefficient states. The problem of constructing an optimal overlay is referred
+to as the \emph{topology mismatch problem}, and is formally defined as follows:
 \begin{definition}
@@
 
-Early incarnations of the overlay protocols did not make use of the
-optimal mapping with the underlying network topology. Gnutella, for
-example, is considered far from scalable \cite{ritter_gnucantscale_2001} because
-every peer chooses its neighbors in the overlay without any knowledge
-of the underlying network, resulting in a mismatch between the two topologies.
-Queries may be flooded over multiple paths merging in
-one peer while one of the paths would have been enough.
-Moreover, peers may forward the same message to each other before they
-receive the query messages from each other. \cite{matei_mapgnutella_2002} showed
-that, even when 95\% of any two nodes are less than 7 hops away from each
-other, a flooding-based algorithm, like the one used by Gnutella, can generate
-330TB/month in a 50.000 node network. A topology mismatch can thus have
-a serious effect on Internet router performance.
+Early incarnations of the overlay protocols did not make use of the optimal
+mapping with the underlying network topology. Gnutella, for example, is
+considered far from scalable \cite{ritter_gnucantscale_2001} because every peer
+chooses its neighbors in the overlay without any knowledge of the underlying
+network, resulting in a mismatch between the two topologies. Queries may be
+flooded over multiple paths merging in one peer while one of the paths would
+have been enough. Moreover, peers may forward the same message to each other
+before they receive the query messages from each other.
+\cite{matei_mapgnutella_2002} showed that, even when 95\% of any two nodes are
+less than 7 hops away from each other, a \emph{flooding-based} algorithm, like
+the one used by Gnutella, can generate 330TB/month in a 50.000 node network. A
+topology mismatch can thus have a serious effect on Internet router performance.
 
@@
 Typically, node IDs are assigned \emph{randomly}, resulting in excellent load
-balancing, scalability, and robustness for the overlay. Unfortunately, this randomness has a
-negative impact on the \emph{routing locality} of the network. This means that
-even though the target node can be reached with logarithmic overlay hops, the
-distance traveled in the underlying network, during the overlay routing
-process, can be far from optimal.  The ratio of the actual IP network distance a query
-travels to reach an object of interest (via overlay routing) to
-the minimal IP network distance to that
-object (i.e. through IP) is known as \emph{stretch} or
-\emph{Relative Delay Penalty (RDP)} \cite{chu_esm_2000}.
+balancing, scalability, and robustness for the overlay. Unfortunately, this
+randomness has a negative impact on the \emph{routing locality} of the network.
+This means that even though the target node can be reached with logarithmic
+overlay hops, the distance traveled in the underlying network, during the
+overlay routing process, can be far from optimal.  The ratio of the actual IP
+network distance a query travels to reach an object of interest (via overlay
+routing) to the minimal IP network distance to that object (i.e. through IP) is
+known as \emph{stretch} or \emph{Relative Delay Penalty (RDP)} \cite{CRZ2000}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Motivation and our Goal for this Survey}
+\subsection{Motivation and Goal for this Survey}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 on several overlay paths causing increase of the traffic on the
-link, known as the link's \emph{stress}~\cite{chu_esm_2002}.
-XXX MR:  The following sentence is unclear: XXX
-Furthermore, the network's \emph{churn}, that is the stochastic behaviour of
-peers randomly joining and leaving the P2P network, all cause the overlay to
-physical mapping to create such unnecessary, redundant maintenance traffic that
-impacts the efficiency of the network in terms of average response time.
+link, known as the link's \emph{stress}~\cite{CRSZ2002}. Even in the case
+of an approach that provides mechanisms to a newcoming node to detect close-by
+peers at join time, the stochastic behaviour with which peers join and leave
+the overlay network, called the network's churn, further undermines the effort
+for maintaing the best position in the network throughout the life-time of a
+peer.
 
-The topology mismatch problem is exacerbated by the extremely large
-amounts of traffic
-generated by highly popular P2P applications. As noted in the Introduction,
-studies have found that more than 60\% of the traffic on the Internet is
-P2P.  Thus, solving the topology mismatch problem is a fundamental challenge
-in contributing to the overall
-efficiency of the Internet.  Unfortunately, topology mismatch is
-known to be an NP-Hard problem \cite{NPBOOK,chawathe_scattercast_2000}.
-Adding to the complexity of the problem is the fact that end-to-end latencies 
-demonstrate \emph{triangle inequality violations
-(TIVs)} which are a 
-consequence of the Internet's structure and routing policies and thus will
-remain a property of the Internet for the foreseeable future
-\cite{zheng_irprtt_2005}. TIVs affect both network coordinate
+The topology mismatch problem is exacerbated by the extremely large amounts of
+traffic generated by highly popular P2P applications (refer to the discussion
+made in Section~\ref{section:intro}). Thus, solving the topology mismatch
+problem is a fundamental challenge in contributing to the overall efficiency of
+the Internet. Unfortunately, topology mismatch is known to be an NP-Hard problem
+\cite{NPBOOK,chawathe_scattercast_2000}. Adding to the complexity of the problem
+is the fact that end-to-end latencies demonstrate \emph{triangle inequality
+violations (TIVs)} which are a consequence of the Internet's structure and
+routing policies and thus will remain a property of the Internet for the
+foreseeable future \cite{zheng_irprtt_2005}. TIVs affect both network coordinate
 \cite{cox_vivaldi_2004,wong_meridian_2005} and positioning \cite{ng_gnp_2001}
-systems 
-and makes the construction of IP-topology-aware overlays difficult.
+systems and makes the construction of IP-topology-aware overlays difficult.
 
-Over the past several years, researchers have proposed and studied 
-an extensive array of heuristic solutions to the topology 
-mismatch problem.
-The purpose of this
-survey is to gather and organize the extensive knowledge that has been produced. 
-We hope to assist the reader
-to understand the different approaches that have been proposed to tackle
-the topology mismatch problem and to identify, through our taxonomy, the
-advantages and disadvantages of each of them.
+Over the past several years, researchers have proposed and studied an extensive
+array of heuristic solutions to the topology mismatch problem. The purpose of
+this survey is to gather and organize the extensive knowledge that has been
+produced. We hope to assist the reader to understand the different approaches
+that have been proposed to tackle the topology mismatch problem. In order to do
+so, we dedicate a discussion subsection after the presentation of the algorithms
+for both unstructred and structured architectures. This section offers:
+\begin{inparaenum}[\itshape i\upshape)]
+  \item further discussion that can help the reader identify, through
+a taxonomy, the pluses and minuses in the preceeding material,
+  \item a comprehensive tabular organisation of the key-point information, and
+  \item an attempt to give a pictorial comparison of the relative performance of
+the algorithms.
+\end{inparaenum}
+
+In discussion we offer the reader a deeper insight for the observations that
+pushed research toward some direction. Thus, we try to categorise these
+approaches so that we can organise directions making them comprehensible
+and helping the reader build critical thinking on the advantages and
+disadvantages of each approach.
+
+Tables provide a one-stop-shop overview for the algorithms so that can be a
+helpful tool both as a companion during reading the survey and a reference for
+future use.
+
+Pictorial comparison consists of bar charts that give a relative feeling of the
+perfomance of the various algorithms, discussed in the corresponding section, in
+terms of \emph{efficiency}, \emph{overhead} and \emph{scalability}; three
+important characteristics for every distributed algorithm. For all we used a
+coarse grained, three level, gradation to denote how well the algorithms do,
+namely \emph{low}, \emph{medium} and \emph{high}. Each category has its own
+interpretation of the above literal values. Efficiency means how well the
+algorithm performs its search and topology adaptation functions, without
+effecting other characteristics of the networks (e.g. BitTorrent's optimal
+dowload performace). To name few, we assume that efficieny is
+negatively effected when an algorithm reduces the search scope. Moreover, the
+minimum-spanning-tree-based tecniques have slow convergence speed in topology
+adaptation or the high clustering impacts load balancing and constrains search
+scope (or at least increases search satisfaction time), all reducing overall
+efficiency. With overhead we track the cost to compute and communicate peer
+join, message forwarding and topology maintenance. For example when exchanging
+cost tables only along neighbours the traffic generated can be considered
+trivial. Also duplicates minimisation is a plus and lazy or heart-beat-based
+overlay control mechanisms impose different amount of stress to the network.
+Nevertheless, caching/replication helps reducing the averall overhead by
+constraining the query from spreading too much. Minimum-spanning-tree-based
+approaches's low convergence speed we mentioned above for efficiency means
+additional computational and communicational overhead. % TODO: We check
+% complexity in the asymptotic sence, using \emph{Big O} notation % in order to
+% check how the algorithm responds t. We take into account topology % creation and
+% topology maintenance...
+% TODO: HELP NEEDED!
+With scalability we check the systems behaviour with respect to the number of
+participating peers. Low, medium and high scalability can be generally
+considered in the order of hundrends, tenths of thousands and millions,
+respectivelly. Fully distributed aproaches is a plus (no system-centric entities
+involved, no need for global knowledge) and topology adaptation streanghtens
+scalability in highly dynamic environments. Reduced search scope, on the other
+side, penaltises the scalability since the more peers comming in the network the
+more potentially unsatisfied queries there will be. Furthermore, clock
+synchonisation is a limiting factor to scalability of a distributed systems.
+Again to name a few.
 
 %
-% TODO: SOME DISCUSSION FROM \cite{xu_globstate_2003) CONTRADICTS TO THE
+% TODO: SOME DISCUSSION FROM \cite{XTZ2003) CONTRADICTS TO THE
 %TRIANGLE INEQUALITY ARGUMENT ABOVE

### 2014-01-26 intro.tex
2012.10.08 (Vassilis) Text added explaining the three indicators (low, medium, high) for each of the criteria against which the algorithms are evaluated (efficiency, overhead, scalability). Filled in missing references and question marks in the tables and text. Intro and background reworked. First page shows the full abstract and bar charts calibrated (orientation, axis labels etc). Text now conforms to ACM Computing Surveys style/format.

@@
 
-Peer-to-peer (P2P) networks are self-organizing distributed systems
-where participating nodes, called \emph{peers}, act as both resource
-providers and resource consumers, in contrast to the traditional \emph{client-server}
-model where nodes take on the role of pure servers or pure clients.  
-Peer-to-peer networks have enjoyed immense attention from Internet users,
-companies, and researchers and have been widely
-deployed for over a decade, primarily because of the great
-number of features they offer distributed applications that are built
-on top of them.  These features include:  high availability and robustness,
-load-balancing, quality of service, scalability, decentralized
-administration, and anonymity.
+\emph{Peer-to-peer (P2P)} networks are self-organizing distributed systems where
+participating nodes, called \emph{peers}, act as both resource providers and
+resource consumers, in contrast to the traditional \emph{client-server} model
+where nodes take on the role of pure servers or pure clients. Peer-to-peer
+networks have enjoyed immense attention from Internet users, companies, and
+researchers and have been widely deployed for over a decade, primarily because
+of the great number of features they offer distributed applications that are
+built on top of them. These features include:  high availability and robustness,
+load-balancing, quality of service, scalability, decentralized administration,
+and anonymity.
 
-A myriad of applications have been built and deployed using the peer-to-peer
-paradigm including two widely-known ``killer'' applications:  file-sharing and telephony.
-P2P file-sharing was originally introduced by Napster (see Section 2)
-which by 2001 was widely acknowledged as the ``fastest growing Internet application
-ever" with 26 million users sharing 80 million songs at the height of its popularity.  
-Other file-sharing applications
-quickly followed suit including Gnutella and its descendent, Limewire, connecting
-3 million users concurrently by the end of 2006 and BitTorrent, connecting
-over 150 million users.
-P2P telephony
-has seen explosive growth with the advent of Skype.  Introduced in 2003, Skype
-has experienced explosive popularity with 
-more that 650 million users in 2011 \cite{skypetotalusers} and
-35 million skypers concurrently online on
-the 5th of March 2012 \cite{skypesymusers}.
+Two ``killer'' applications of the peer-to-peer paradigm are, but not limited
+to, file-sharing and telephony. P2P file-sharing was originally introduced by
+Napster (see Section~\ref{section:background}), which by 2001 was widely
+acknowledged as the ``fastest growing Internet application ever" with 26 million
+users sharing 80 million songs at the height of its popularity. Other similar
+applications quickly followed suit including Gnutella and its descendent,
+Limewire, enjoying 3 million concurrently connected peers by the end of 2006
+and BitTorrent, connecting over 150 million users. P2P telephony has seen
+explosive growth with the advent of Skype. Introduced in 2003, Skype has
+experienced explosive popularity with more that 650 million users in 2011
+\cite{skypetotalusers} and 35 million skypers concurrently online on the 5th of
+March 2012 \cite{skypesymusers}.
 
-In addition to file-sharing and telephony, a large number of diverse applications have
-been built on top of peer-to-peer systems including:
+In addition to file-sharing and telephony, a large number of diverse
+applications have been built on top of peer-to-peer systems including:
 distributed search engines \cite{yaci}, distributed data storage systems
-\cite{kbc_oceanstore_2000,bdet_fsdfs_2000,dkkms_cfs_2001,dr_pastutility_2001,abc_farsite_2002,mmfc_ivy_2002,arla,agebh_dks_2003}, Web caches, archives,
-and publishing systems \cite{ird_squirrel_2002,bags_youserv_2002,wrc_publius_2000,wm_tangler_2001},
+\cite{kbc_oceanstore_2000,bdet_fsdfs_2000,dkkms_cfs_2001,dr_pastutility_2001,abc_farsite_2002,mmfc_ivy_2002,arla,agebh_dks_2003},
+Web caches, archives and publishing systems
+\cite{ird_squirrel_2002,bags_youserv_2002,wrc_publius_2000,wm_tangler_2001},
 messaging applications \cite{threedegrees}, event notification infrastructures
@@
 
-Peer-to-peer systems are implemented using overlay networks.   
-An overlay network is a virtual network of nodes and logical links 
-that is built on top of an existing network with the aim of providing 
-a service (such as content distribution) that is not available 
-in the existing network.  Nodes in the overlay are connected via virtual 
-links, each of which corresponds to a path of possibly many physical 
-links in the underlying network.  
-Figure~\ref{figure:overlay}
-illustrates a simple four-node overlay constructed over a wide-area network.
-In the case of peer-to-peer networks, typically the participating nodes
-are user PCs sitting at the edge of the Internet.
+Peer-to-peer systems are implemented using overlay networks. An \emph{overlay
+network} is a virtual system of nodes and logical links, that is built on top of
+an existing network with the aim of providing a service (such as content
+distribution) that is not available in the existing one. Nodes in the overlay,
+communicate via virtual connections, each of which may correspond to a path of
+possibly many physical links in the underlying network.
+Figure~\ref{figure:overlay} illustrates a simple four-node overlay constructed
+over a wide-area network.
 
@@
 
-One of the fundamental issues that defines the efficiency of an overlay 
-network is how 
-well the overlay network topology maps to the underlying network topology on which it rests.
-Consider two nodes that share a link in an overlay network which maps to multiple 
-hops in the underlying IP topology.  If the traffic patterns of the application are such
-that traffic flows heavily over the overlay link between the two nodes, then it is
-beneficial to construct the overlay topology such that the number of underlying IP links
-between these two nodes is minimized.
+One of the fundamental issues that defines the efficiency of an overlay network
+is how well the overlay maps to the underlying network topology on which it
+rests. Consider two nodes\footnote{In the case of peer-to-peer networks,
+typically the participating nodes are user PCs sitting at the edge of the
+Internet.} that share a link in an overlay network which maps to multiple hops
+in the underlying (say IP) topology. If the traffic patterns of the application
+are such that traffic flows heavily over the overlay link between the two nodes,
+then it is beneficial to construct the overlay topology such that the number of
+underlying IP links between these two nodes is minimized.
 
 If the overlay network is constructed such that the topology does not match the 
-underlying IP topology well, the resulting \emph{topology mismatch} 
-can result in
-two significant problems.  First, application performance (e.g., search query latency
-or jitter) can be adversely affected since application traffic must flow over a larger number
-of IP hops than if the overlay were constructed more wisely.  Second, studies have shown
-that P2P traffic contributes the largest portion of overall Internet traffic
+underlying IP topology well, the resulting \emph{topology mismatch} (See
+Section~\ref{section:background} for a more formal definition) can result in two
+significant problems. First, application performance (e.g. search query latency
+or jitter) can be adversely affected since application traffic must flow over a
+larger number of IP hops than if the overlay were constructed more wisely.
+Second, studies have shown that P2P applications contribute the largest portion
+of overall Internet traffic
 \cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}, with
-some reporting that more than 60\% of the traffic is P2P~\cite{cachelogic,ipoque2007,ipoque2009}  
-and others reporting that P2P traffic is expected to reach almost 8 petabytes 
-per month by the end of 2012~\cite{multinteligence}.  This is a significant burden for Internet Service Providers (ISPs) who 
-must transfer this traffic to destination nodes sitting at the edge of the Internet.
-If the overlay topology of a peer-to-peer system is not cleverly constructed, then 
-the burden of ISPs carrying P2P traffic could be even greater.
-For example, traffic might have to
-flow back and forth several times between two neighboring ISPs as it travels from its
-source node in the overlay to its destination node.
-Thus, it is fundamentally important that P2P overlay networks are constructed
-and maintained so that
+some reporting that more than 60\% of it, is
+P2P-related~\cite{cachelogic,ipoque2007,ipoque2009} and others claiming that it
+is expected to reach almost 8 petabytes per month by the end of
+2012~\cite{multinteligence}. This is a significant burden for Internet Service
+Providers (ISPs) who must handle all this traffic to destination nodes sitting
+at the edge of the Internet. If the overlay topology of a peer-to-peer system is
+not efficintly constructed, then this burden to the Internet's backbone
+infrastructure could be even greater. For example, traffic might have to flow
+back and forth several times between two neighboring ISPs as it travels from its
+source node in the overlay to its destination node. Thus, it is fundamentally
+important that P2P overlay networks are constructed and maintained so that
 their topology matches as closely as possible the underlying IP topology.
@@
 For over a decade, several tens of researchers have published profusely on the 
-topology mismatch problem.  In this paper, we survey the literature
-on this topic and provide a taxonomy of the proposed approaches and solutions.
-We point out synergies
-as well as similarities and differences in the published approaches.  
-Ultimately, our goal is to help readers sift through the vast
-amount of literature, to help them understand the advantages and 
-disadvantages of each approach, and to provide them with enough 
-perspective so that when the need arises, they are able to select 
-amongst the different approaches for their 
-particular application.
+topology mismatch problem.  In this paper, we survey this research area and
+provide a taxonomy of the proposed solutions. We point out synergies as well as
+similarities and differences in the published approaches. Ultimately, our goal
+is to help readers sift through the vast amount of literature, to help them
+understand the advantages and disadvantages of each of each work, and to provide
+them with enough perspective so that when the need arises, they are able to
+select amongst the different approaches for their particular application.
 
-The rest of the paper is as follows:  In Section~\ref{section:background}, 
-we provide some background information
-on peer-to-peer overlay architectures including centralized, decentralized unstructured
-and decentralized structured peer-to-peer systems.  We also present formally the
-problem of alleviating the topology mismatch between a peer-to-peer overlay and the
-underlying IP network.  In Sections~\ref{section:unstructured} and 
-\ref{section:structured}, we describe a taxonomy of
-the research efforts and solutions
-proposed for decentralized unstructured peer-to-peer systems 
-and decentralized
-structured peer-to-peer systems, respectively.  Finally, 
-in Section~\ref{section:conclusion}, we conclude.
+The rest of the paper is as follows: In Section~\ref{section:background}, we
+provide some background information on peer-to-peer overlay architectures
+including centralized, decentralized unstructured and decentralized structured
+peer-to-peer systems. We also formally define the problem of topology mismatch
+between an overlay and its underlying physical network and elaborate on the
+rationale behind our coarse grained evaluation of the techinques presented
+throughout this survey. In Sections~\ref{section:unstructured} and
+\ref{section:structured}, we present the various research efforts for
+decentralized unstructured and decentralized structured peer-to-peer systems,
+respectively. Finally, with Section~\ref{section:conclusion}, we
+conclude.

### 2014-01-26 mard-survey.tex
2012.06.26 (Mema) Big introduction section split to a crispier 3-page intro and a background section. The latter also shortened and updated.
2012.07.12 (Vassilis) Merged introduction section. Sectioning according to the method an algorithm is using dropped (almost all use a combination of two or more). First attempt for the tabular data on an external file.
2012.08.20 (Vassilis) Added LaTeX tables (known issue on wrapping text inside cells). Reshaped discussion material. Introducing rough, algorithm performance, estimation (high, medium, low) by tracking efficiency, control overhead and scalab

@@
 \usepackage{longtable}
+\usepackage{wasysym}
 \usepackage{pdflscape}
+%\usepackage{tikz}
+\usepackage{pgfplots}
 
@@
 Overlay Networks: A Survey}
+% TODO: We added here "P2P". I'm not sure, because for example Narada is not P2P
+% TODO: I deleted "P2P"
 
 \author{
-VASSILIS S. MOUSTAKAS\\
-\texttt{b.moystakas@di.uoa.gr}\\
-\emph{Department of Informatics and Telecommunications}\\
-\emph{National and Kapodistrian University of Athens}\\
-H\"USEYIN AKCAN\\
-\texttt{huseyin.akcan@ieu.edu.tr}\\
-\emph{Department of Software Engineering}\\
-\emph{Izmir University of Economics}\\
-MEMA ROUSSOPOULOS\\
-\texttt{mema@di.uoa.gr}\\
-\emph{Department of Informatics and Telecommunications}\\
-\emph{National and Kapodistrian University of Athens}\\
-AND\\
-ALEX DELIS\\
-\texttt{ad@di.uoa.gr}\\
-\emph{Department of Informatics and Telecommunications}\\
-\emph{National and Kapodistrian University of Athens}
+VASSILIS S. MOUSTAKAS$^1$, H\"USEYIN AKCAN$^2$, MEMA ROUSSOPOULOS$^1$ and ALEX
+DELIS$^1$\\
+$^1$\emph{National and Kapodistrian University of Athens}, \emph{Athens},
+\emph{Greece}\\
+$^2$\emph{Izmir University of Economics}, \emph{Izmir}, \emph{Turkey}
 }
@@
 
-\category{}{}{}
+\category{A.1}{General Literature}{Introductory and Survey}
+\category{C.2.4}{Distributed Systems}{Distributed Applications}
 
-\terms{Topology mismatch, distributed, peer-to-peer, P2P, overlay}
+\terms{Topology mismatch}
 
-\keywords{Topology mismatch problem, distributed algorithm, peer-to-peer
-architecture, overlay network }
+\keywords{Distributed systems, peer-to-peer architectures, overlay network,
+topology awareness}
 
@@
 \begin{bottomstuff}
-SAY SOMETHING COOL HERE...
+Authors' addresses: Department of Informatics and Telecommunications, National
+and Kapodistrian University of Athens, Athens, Greece. Department of Software
+Engineering, Izmir University of Economics, Izmir, Turkey.\\
+emails: \{b.moystakas,mema,ad\}@di.uoa.gr, huseyin.akcan@ieu.edu.tr
 \end{bottomstuff}
@@
 \input{intro}
+\input{background}
 

### 2014-01-26 structured.tex
2012.09.06 (Vassilis) Algorithm depiction reworked.
2012.10.08 (Vassilis) Text added explaining the three indicators (low, medium, high) for each of the criteria against which the algorithms are evaluated (efficiency, overhead, scalability). Filled in missing references and question marks in the tables and text. Intro and background reworked. First page shows the full abstract and bar charts calibrated (orientation, axis labels etc). Text now conforms to ACM Computing Surveys style/format.

@@
 information to optimize them towards this direction. The algorithms are here
-categorised based the previous work done by
-\cite{castro_proximitydht_2002,castro_topawareroute_2002,ratnasamy_openq_2002}.
+categorised based the previous work done by \cite{CDHR2002,CDCR2002,RSS2002}.
 
@@
 
-\textit{Global Soft-State} \cite{xu_globstate_2003} builds a global map to help
+\textit{Global Soft-State} \cite{XTZ2003} builds a global map to help
 choose shorter routing paths, combining the landmark binning method and small
@@
 %
-% Xu \textit{et al.} \cite{xu_globstate_2003} state that there are three ways
+% Xu \textit{et al.} \cite{XTZ2003} state that there are three ways
 % generating proximity information
@@
 high &
-% Statically implemented sucrificing the the decentralised nature of DHTs and
+% Statically implemented sacrificing the decentralised nature of DHTs and
 % thus poorly scalling.
@@
 \subsubsection{Mithos}
-\emph{Mithos} \cite{waldvogel_mythos_2003} is a P2P protocol which incorporates
-directed incremental probing to find near optimal node placement, and is
-classified by its authors as an integration of geographic layout and proximity
-routing overlay optimization methods.
+\emph{Mithos} \cite{WR2003} is a P2P protocol which incorporates directed
+incremental probing to find near optimal node placement, and is classified by
+its authors as an integration of geographic layout and proximity routing overlay
+optimization methods.
 
@@
 \subsubsection{LAPTOP}
-\emph{LAPTOP} \cite{wu_laptop_2007}, organises the overlay into a tree-based
-hierarchy with main focus on reducing hops during message routing as well as
-minimizing maintenance overhead. Additionally, a caching scheme is also
-incorporated so as to further reduce routing table update costs. The authors
-theoretically show that in LAPTOP routing path length is bounded by $O(log_d N)$
-and node joining and leaving in the overlay network is bounded by
-$O\left( d log_d N \right)$ hops in a balanced overlay tree, where $N$ is the
-number of nodes, and $d$ is the maximum degree of each node. LAPTOP implements
-the geographical layout approach  and constructs its layout in a self-organizing
-and efficient fashion, by estimating the round-trip times to a small number of
-nodes in the overlay network in order to make them roughly aware of their
-physical distances among them.
+\emph{LAPTOP} \cite{WLH2007}, organises the overlay into a tree-based hierarchy
+with main focus on reducing hops during message routing as well as minimizing
+maintenance overhead. Additionally, a caching scheme is also incorporated so as
+to further reduce routing table update costs. The authors theoretically show
+that in LAPTOP routing path length is bounded by $O(log_d N)$ and node joining
+and leaving in the overlay network is bounded by $O\left( d log_d N \right)$
+hops in a balanced overlay tree, where $N$ is the number of nodes, and $d$ is
+the maximum degree of each node. LAPTOP implements the geographical layout
+approach  and constructs its layout in a self-organizing and efficient fashion,
+by estimating the round-trip times to a small number of nodes in the overlay
+network in order to make them roughly aware of their physical distances among
+them.
 
@@
 
-\emph{Proximity in Kademlia} \cite{kaune_pkad_2008} introduces proximity
+\emph{Proximity in Kademlia} \cite{KLKP2008} introduces proximity
 controls over the vanilla Kademlia protocol's random ID assignment, in order to
@@
 \subsubsection{CHOord considering Proximity on IPv6 (CHOP6)}
-\emph{CHOP6} \cite{morimoto_chop6_2007} roughly estimates the proximity among
+\emph{CHOP6} \cite{MT2007} roughly estimates the proximity among
 nodes by exploiting the IPv6 address format and RTT information if available.
@@
 \subsubsection{T2MC}
-T2MC \cite{shi_t2mc_2008} uses traceroute logs to detect different autonomous
+T2MC \cite{SLCGZ2008} uses traceroute logs to detect different autonomous
 systems and cluster close by peers, thus reducing redundant traffic that crosses
@@
 % Traceroute is categorised as too heavy-weighted and intrucive for use in a
-% larger scale system \cite{ratnasamy_binning_2002}. Additionally, disabling
+% larger scale system \cite{RHMKS2002}. Additionally, disabling
 % ICMP is a common administrative policy for edge sites to enforce security,
@@
 \subsubsection{PChord}
-\emph{PChord}\cite{hong_pchord_2005} is a based on the Chord DHT which adds
+\emph{PChord}\cite{HLYW2005} is a based on the Chord DHT which adds
 proximity routing into its routing mechanism. The main modification over the
@@
 \subsubsection{AChord}
-\emph{AChord} \cite{dao_achord_2006} uses IPv6's anycast mechanism in order to
+\emph{AChord} \cite{DK2006} uses IPv6's anycast mechanism in order to
 releave the protocol from complex joining procedures, and achieve high accuracy
@@
 \subsubsection{Chord6}
-\emph{Chord6} \cite{xiong_chord6_2005} is another Chord variant that tries to
+\emph{Chord6} \cite{XZHL2005} is another Chord variant that tries to
 exploit the hierarchical features of IPv6 in order to create a substrate that
@@
 \subsubsection{DHT-PNS}
-The work in \cite{hancong_pnsbased_2006} propose a proximity neighbour selection
+The work in \cite{DLTZZ2006} propose a proximity neighbour selection
 scheme on top of the Chord DHT. The main goal is to use proximity information
@@
 \subsubsection{Quasi-Chord}
-The approach that is proposed by Sun and Zhang in \cite{sun_quasi_2008}, tries
+The approach that is proposed by Sun and Zhang in \cite{SZ2008}, tries
 to address the topology mismatch problem, at the construction phase of the
@@
 % space to 1-demensional with Cantor SPF adds some extra burden in time
-% complexity. Also the addition of the couter-clockwise finger table doubles the
-% data, per se, needed to be stored and maintained.
+% complexity. Also the addition of the counter-clockwise finger table doubles
+% the data, per se, needed to be stored and maintained.
 high &
@@
 detect proximity. However, such methods are either not precise enough, or
-overload the network. \emph{IP-based clustering} \cite{karwaczynski_ipbc_2007}
-is a proximity neighbour selection based algorithm in which the authors propose
-to use the IP address prefixes (16 bit for IPv4) in order to detect proximity.
+overload the network. \emph{IP-based clustering} \cite{KM2007} is a proximity
+neighbour selection based algorithm in which the authors propose to use the IP
+address prefixes (16 bit for IPv4) in order to detect proximity.
 \cite{freedman_iploc_2005} states that $97\%$ of prefixes larger than $24$ bits
@@
 \subsubsection{Cone}
-\emph{Cone} \cite{wang_cone_2007} extends Chord using proximity neighbour
+\emph{Cone} \cite{HY2007} extends Chord using proximity neighbour
 selection topology optimization algorithm. The proximity information is
@@
 \subsubsection{DynaMo}
-\emph{DynaMO} \cite{winter_dynamo_2004} tries to build an overlay that not only 
+\emph{DynaMO} \cite{WZS2004} tries to build an overlay that not only 
 considers physical proximity among peers but their mobility attributes as well.
@@
 heuristics which are thoroughly analysed in the in the literature
-\cite{castro_proximityp2p_2002} providing good backround against which to test
-and compare their results. In order to adapt to mobile, ad-hoc networks,
-though, special care has been put to the maintenance of an even overlay ID
-distribution so that hot-spots are avoided. DynaMO dictates a newcoming node to
-gather information concerning its physical neighbourhood and uses it to assign
-itself an appropriate overlay ID. DynaMO tries to capitalize on the fact that
-in Pastry each node's table consists of rows equal to the number of digits of
-the overlay IDs and columns equal to the ID's base. As we go down the table
-rows, the matching prefix between the current node's ID and the row's entries
-increases by one. Thus, the leaf set contains the closest nodes in the ID space.
-Additionally as the prefix match increases by one, the result is exponentially
-less candidates that can fill the tables entries as the row increases. This
-leads to the observation \cite{antony_pastry_2001,castro_proximityp2p_2002} that
-from overlay routing hop to overlay routing hop the physical distance between
-nodes is likely to increase leading to a dominating last routing step in terms
-of the overall plysical routing path distance travelled during a key lookup.
-Since the last routing step is usually taken from the leaf set, DynaMO focuses
-on making this last step as physicaly close as possible. In this context, two
-approaches are considered namely \emph{Random Landmarking (RLM)} and
-\emph{Closest Neighbour Prefix Assignment (CNPA)}.
+\cite{CDCR2002a} providing good backround against which to test and compare
+their results. In order to adapt to mobile, ad-hoc networks, though, special
+care has been put to the maintenance of an even overlay ID distribution so that
+hot-spots are avoided. DynaMO dictates a newcoming node to gather information
+concerning its physical neighbourhood and uses it to assign itself an
+appropriate overlay ID. DynaMO tries to capitalize on the fact that in Pastry
+each node's table consists of rows equal to the number of digits of the overlay
+IDs and columns equal to the ID's base. As we go down the table rows, the
+matching prefix between the current node's ID and the row's entries increases by
+one. Thus, the leaf set contains the closest nodes in the ID space. Additionally
+as the prefix match increases by one, the result is exponentially less
+candidates that can fill the tables entries as the row increases. This leads to
+the observation \cite{antony_pastry_2001,CDCR2002a} that from overlay routing
+hop to overlay routing hop the physical distance between nodes is likely to
+increase leading to a dominating last routing step in terms of the overall
+plysical routing path distance travelled during a key lookup. Since the last
+routing step is usually taken from the leaf set, DynaMO focuses on making this
+last step as physicaly close as possible. In this context, two approaches are
+considered namely \emph{Random Landmarking (RLM)} and \emph{Closest Neighbour
+Prefix Assignment (CNPA)}.
 
@@
 
-\emph{SAT-Match} \cite{ren_satmatch_2004}. Even though no theoretical proof is
+\emph{SAT-Match} \cite{RGJZ2004}. Even though no theoretical proof is
 given, valid results are obtained and demonstrated through experimentations.
@@
 small TTL value for the probing messages in order to reduce
-redundancy\cite{jiang_lightflood_2008}. This process begins as soon as the node
+redundancy \cite{jiang_lightflood_2008}. This process begins as soon as the node
 joins the network using a DHT mechanism. Each probing message contains
@@
 \paragraph*{ \bf Delay Aware P2P System}
-A new \emph{Delay Aware P2P System (DAPS)} is introduced in
-\cite{zhang_daps_2005}. Its main goal is to reduce the time $L$ of a look-up
-request by dividing the routing tables of peers into several sectors in
-increasing delay. The source node emitting the query message defines a delay
-boundary or the pruning factor $L_t$ in the paper's context. Request messages
-will be forwarded only to nodes whose delay less than or equal to $L_t$. With
-the clustered routing tables and the loose organisation the overlay network of
-\emph{DAPS} is between structured and unstructured.
+A new \emph{Delay Aware P2P System (DAPS)} is introduced in \cite{ZL2005}. Its
+main goal is to reduce the time $L$ of a look-up request by dividing the routing
+tables of peers into several sectors in increasing delay. The source node
+emitting the query message defines a delay boundary or the pruning factor $L_t$
+in the paper's context. Request messages will be forwarded only to nodes whose
+delay less than or equal to $L_t$. With the clustered routing tables and the
+loose organisation the overlay network of \emph{DAPS} is between structured and
+unstructured.
 
@@
 following based on previous work from
-\cite{castro_proximitydht_2002,castro_topawareroute_2002,ratnasamy_openq_2002}:
+\cite{CDHR2002,CDCR2002,RSS2002}:
 \begin{inparaenum}[\itshape i\upshape)]
@@
 \input{structured-table}
+
+\input{structured-plot}
 
@@
 matching, while CAN uses RTT values to detect close by nodes.
-\cite{castro_proximityp2p_2002} reports that proximity neighbour selection as an
+\cite{CDCR2002a} reports that proximity neighbour selection as an
 effective proximity based method.

### 2014-01-26 unstructured.tex
2012.09.06 (Vassilis) Algorithm depiction reworked.
2012.10.08 (Vassilis) Text added explaining the three indicators (low, medium, high) for each of the criteria against which the algorithms are evaluated (efficiency, overhead, scalability). Filled in missing references and question marks in the tables and text. Intro and background reworked. First page shows the full abstract and bar charts calibrated (orientation, axis labels etc). Text now conforms to ACM Computing Surveys style/format.

@@
 
-\cite{yang_improvep2psearch_2002} proposes an easy to implement and practical
-solution to the inefficiency of Gnutella's ``blind flooding'' approach. The
-paper introduces three different forwarding methods, namely \emph{iterative
-deepening}, \emph{directed BFS}, and \emph{local indices}.
+\cite{YG-M2002} proposes an easy to implement and practical solution to the
+inefficiency of Gnutella's ``blind flooding'' approach. The paper introduces
+three different forwarding methods, namely \emph{iterative deepening},
+\emph{directed BFS}, and \emph{local indices}.
 
@@
 \emph{Local Indices} &
-% efficiency is enhanced greatly since data pointers are replicated acros the
+% efficiency is enhanced greatly since data pointers are replicated across the
 % network
@@
 \subsubsection{Gia}
-\emph{Gia} \cite{chawathe_gia_2003} is an effort to solve the scalability
+\emph{Gia} \cite{CRBLS2003} is an effort to solve the scalability
 problem of the unstructured P2P file sharing systems, in particular Gnutella.
@@
 % A token based flow control mechanism redirects queries to nodes that are most
-% likely to fulfill them
-low &
-% Reduced search scope makes this not really scalable
-low
-\end{tabular}
-\end{center}
+% likely to fulfill them.
+low &
+% Even thought the low overhead scalability is reduced by the reducing of
+% search scope.
+medium
+\end{tabular}
+\end{center}
+
+% \begin{center}
+% \begin{tikzpicture}
+% \begin{axis}[
+%   ybar,
+%    symbolic x coords={
+%      Efficiency,
+%      Overhead,
+%      Scalability
+%    },
+%    symbolic y coords={lo, med, hi},
+%    x tick label style={rotate=45,anchor=east},
+%    xtick=data, ytick=data
+% ]
+
+% \addplot coordinates {
+%   (Efficiency,lo)
+%   (Overhead,lo)
+%   (Scalability,lo)
+% };
+
+% \end{axis}
+% \end{tikzpicture}
+% \end{center}
 
@@
 systems is the duplication of messages due to cycles, even along the correct
-forwarding path. The authors of \cite{zhu_dcmp_2008} fucused on that exact
-deficiency of overlay networks and introduced Distributed Cycle Minimization
-Protocol (DCMP), a dynamic, fully distributed method to remove cycles without
-sacrificing overlay connectivity, resilience and other important properties of
-unstructured architectures by avoiding a hierarchical organisation of peers. In
-DCMP, once a cycle is detected, the most powerful node in that cycle is
-elected as the \emph{Gate Peer} and the cycle is then broken at a strategic
-place so that it will result in the minimisation of the distance between the
-Gate Peer and all other nodes that participated in that cycle. The process
-is managed by using two special message types namely \emph{Information
-Collection (IC)} and \emph{Cut Message (CM)}. One disadvantage of DCMP is that
-since the distance a forwarded message can travel is limited by the TTL value,
-which is practically $7$ in most cases DCMP cannot detect cycles formed by more
-than 7 peers. Even though cycle elimination improves the network performance, it
-does not actually solve the topology mismatch problem.
+forwarding path. The authors of \cite{ZKB2008} fucused on that exact deficiency
+of overlay networks and introduced Distributed Cycle Minimization Protocol
+(DCMP), a dynamic, fully distributed method to remove cycles without sacrificing
+overlay connectivity, resilience and other important properties of unstructured
+architectures by avoiding a hierarchical organisation of peers. In DCMP, once a
+cycle is detected, the most powerful node in that cycle is elected as the
+\emph{Gate Peer} and the cycle is then broken at a strategic place so that it
+will result in the minimisation of the distance between the Gate Peer and all
+other nodes that participated in that cycle. The process is managed by using two
+special message types namely \emph{Information Collection (IC)} and \emph{Cut
+Message (CM)}. One disadvantage of DCMP is that since the distance a forwarded
+message can travel is limited by the TTL value, which is practically $7$ in most
+cases DCMP cannot detect cycles formed by more than 7 peers. Even though cycle
+elimination improves the network performance, it does not actually solve the
+topology mismatch problem.
 
@@
 % It return more result than LTM according to experiments in
-% \cite{zhu_dcmp_2008}
+% \cite{ZKB2008}
 medium &
@@
 % to its ``lazy'' broadcasting of control messages (as opposed to LTM's periodic
-% approach) - \cite{zhu_dcmp_2008}
+% approach) - \cite{ZKB2008}
 low &
@@
 \subsubsection{Replication Strategies in Unstructured Peer-to-Peer Networks}
-\cite{Cohen02} aim to improve the inefficient blind search algorithm by
+\cite{CS2002} aim to improve the inefficient blind search algorithm by
 replicating the data in a P2P network. The main intuition behind the
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Tracing a large-scale Peer to Peer System: an hour in the life
-of Gnutella}
+% \subsubsection{Tracing a large-scale Peer to Peer System: an hour in the life
+% of Gnutella}
 % TODO: This looks better as an analysis for the unstructured networks in
@@
 % revisit the paper itself once more
-\cite{Markatos02} analyses Gnutella network traffic traces and by concluding
-there is locality among query requests, proposes a caching algorithm that tries
-to exploit these findings . The analysis of the trace data reveals other
-important facts about the structure of the Gnutella network and the query data.
-One significant observation is that the geographic locations of clients do not
-have a correlation with the number of query requests they receive. This is an
-obvious result of the topology mismatch problem caused by the overlay structure
-of the Gnutella network. Gnutella's traffic is observed to be bursty both for
-query requests and query responses, even in longer intervals. It is observed
-that a peer receives 50 query messages per second on average! Moreover nine out
-of ten queries do not generate any response due to the inefficient design of
-the Gnutella network. When developing a caching system to exploit locality,
-applying an approach similar to web caching does not fit well with P2P systems.
-Caches in P2P systems not only have to consider the query string, but also
-the TTL value, the source of the query, and the time of the query as well.  In
-general, even though optimum caching is hard to achieve, it is reported that
-it improves the overall performance of the Gnutella network.
-
-\begin{center}
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
-%
-? &
-%
-? &
-%
-?
-\end{tabular}
-\end{center}
+% \cite{Markatos02} analyses Gnutella network traffic traces and by concluding
+% there is locality among query requests, proposes a caching algorithm that tries
+% to exploit these findings . The analysis of the trace data reveals other
+% important facts about the structure of the Gnutella network and the query data.
+% One significant observation is that the geographic locations of clients do not
+% have a correlation with the number of query requests they receive. This is an
+% obvious result of the topology mismatch problem caused by the overlay structure
+% of the Gnutella network. Gnutella's traffic is observed to be bursty both for
+% query requests and query responses, even in longer intervals. It is observed
+% that a peer receives 50 query messages per second on average! Moreover nine out
+% of ten queries do not generate any response due to the inefficient design of
+% the Gnutella network. When developing a caching system to exploit locality,
+% applying an approach similar to web caching does not fit well with P2P systems.
+% Caches in P2P systems not only have to consider the query string, but also
+% the TTL value, the source of the query, and the time of the query as well.  In
+% general, even though optimum caching is hard to achieve, it is reported that
+% it improves the overall performance of the Gnutella network.
+
+% \begin{center}
+% \begin{tabular}{ccc}
+% \emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+% \hline
+%
+% ? &
+%
+% ? &
+%
+% ?
+% \end{tabular}
+% \end{center}
 
@@
 
-\emph{Narada} \cite{chu_esm_2000,chu_esm_2001,chu_esm_2002} is a generic
+\emph{Narada} \cite{CRZ2000,CRSZ2001,CRSZ2002} is a generic
 protocol for creating self-adapting overlay networks, that can achieve
@@
 scalable in the contemporary sense of the term, Narada has been pioneering
-(maybe along with \emph{Scattercast} \cite{chawathe_scattercast_2000}) in that
+(maybe along with \emph{Scattercast} \cite{C2000}) in that
 it was the first to consider the feasibility of overlay-based, application
@@
 \emph{mesh}, and building per source minimum spanning trees\footnote{
-  ENarada \cite{li_enarada_2008} used Gossip protocol for the construction
+  ENarada \cite{LYL2008} used Gossip protocol for the
+construction
 }
@@
 \subsubsection{Adaptive Overlay Topology Optimization (AOTO)}
-\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{liu_aoto_2003} is one
+\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{LZXN2003} is one
 of the first attempts, along with Narada, to address the topology mismatch
@@
 \subsubsection{Adaptive Connection Establishment (ACE)}
-\emph{Adaptive Connection Establishment (ACE)} \cite{liu_ace_2004} builds an
+\emph{Adaptive Connection Establishment (ACE)} \cite{LZXN2004} builds an
 overlay multicast tree among each source node and the peers within a certain
@@
 
-%In \emph{Adaptive Connection Establishment (ACE)} \cite{liu_ace_2004}, the
+%In \emph{Adaptive Connection Establishment (ACE)} \cite{LZXN2004}, the
 %authors extend the idea of AOTO by introducing optimizations based on the
@@
 % certain diameter from the source peer
-high &
+medium &
 % Larger diameter topologies lead to better topology optimization
@@
 \subsubsection{Location-aware Topology Matching (LTM)}
-\emph{Location-aware Topology Matching (LTM)} \cite{liu_ltm_2004} proposes a
+\emph{Location-aware Topology Matching (LTM)} \cite{LLXNZ2004} proposes a
 method to optimize the overlay structure of the P2P network based on the
@@
 \emph{Scalable Bipartite Overlay (SBO)}
-\cite{liu_bipartite_IPDPS,liu_bipartite_2007} reduces the overhead of creating
+\cite{LXN2004,LXN2007} reduces the overhead of creating
 and maintaining a minimum spanning tree cost by randomly dividing the nodes into
@@
 \subsubsection{Two-Hop-Away Neighbour Comparison and Selection (THANCS)}
-Work in \cite{liu_thancs_2005,liu_thancs_2008} proposes a distributed heuristic
+Work in \cite{LNXE2005,L2008} proposes a distributed heuristic
 called \emph{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} that
@@
 \subsubsection{Hops Adaptive Neighbour Discovery (HAND)}
-\cite{chen_hand_2006} proposes the \emph{HAND} algorithm which uses a
+\cite{CLZHC2006} proposes the \emph{HAND} algorithm which uses a
 fully distributed triple hop adjustment strategy to address the topology
@@
 Bernstein et al., contributed on the use of machine learning for building peer
-selection strategies from past experience \cite{bflz_adaptpeersel_2003}. A
-decision tree is used to rate peers based on information about connection
-characteristics that is collected. These can be load, bandwidth, and past
-uploading experience. Then a by Markov decision process is incorporated as a
-mechanism to shape the policy for switching among the peers. The problem
-with this approach is that the peer selection algorithm is slow due to the
-learning process and the complexity of the method.
+selection strategies from past experience \cite{BFLZ2003}. A decision tree is
+used to rate peers based on information about connection characteristics that is
+collected. These can be load, bandwidth, and past uploading experience. Then a
+by Markov decision process is incorporated as a mechanism to shape the policy
+for switching among the peers. The problem with this approach is that the peer
+selection algorithm is slow due to the learning process and the complexity of
+the method.
 
@@
 \subsubsection{Innocuous Topology Aware (ITA) Overlay Construction}
-\cite{prfm_ita_2009} intoduces an algorithm for Innocuous Topology Aware
+\cite{PRFM2009} intoduces an algorithm for Innocuous Topology Aware
 construction of unstructured P2P networks. The paper's proposition is twoforld.
@@
 \subsubsection{EGOIST}
-EGOIST \cite{egoist_2008} is a set of algorithms to construct and manage overlay
+EGOIST \cite{SLLBBR2008} is a set of algorithms to construct and manage overlay
 networks. It utilises a selfish approach in the sense that every participating
@@
 \subsubsection{Biased Neighbour Selection (BNS)}
-Bindal et al. \cite{rpc_bitbiased_2006} claim to strengthen BitTorrent
+Bindal et al. \cite{BCCMSBZ2006} claim to strengthen BitTorrent
 protocol\cite{c_bittorrent_2003} locality by selecting most of a peer's
@@
 % retrieved from special hardware.
-medium/high
+medium
 \end{tabular}
@@
 \subsubsection{Ono}
-Choffnes and Bustamante propose Ono \cite{cb_ono_2008}, a protocol for managing
+Choffnes and Bustamante propose Ono \cite{CB2008}, a protocol for managing
 BitTorrent traffic so that it significantly reduces cross-ISP traffic and in
@@
 borders of a single autonomous system. Contrary to biased peer selection
-proposed in \cite{rpc_bitbiased_2006}, this work can lead to improved
+proposed in \cite{BCCMSBZ2006}, this work can lead to improved
 performance with no cooperation between ISPs and their subscribers, no
@@
 \subsubsection{Locality-Awareness in BitTorrent-like P2P Applications}
-In \cite{lclx_bitlocal_2009} the authors study and compare three different
+In \cite{LCLX2009} the authors study and compare three different
 approaches in injecting locality awareness in BitTorrent-like applications. The
@@
 \subsubsection{TopBT}
-\cite{rtl_topbt_2010} proposed an approach for enhancing proximity awareness
+\cite{RTLCGZ2010} proposed an approach for enhancing proximity awareness
 in the BitTorrent protocol without any need of any infrastructure installed. It
@@
 \subsubsection{Underlying Topology-Aware Peer Selection (UTAPS)}
-UTAPS \cite{lcy_utaps_2008} is another peer selection strategy for the
-BitTorrent protocol. Its first step is to collect the needed information in
-order to construct a picture of the underlying topology and the second is to
-make the peer selection, per se, based on this knowledge. For the first part,
-the algorithm uses network tomography, a techni  que which suggests probing only
+UTAPS \cite{LCY2008} is another peer selection strategy for the BitTorrent
+protocol. Its first step is to collect the needed information in order to
+construct a picture of the underlying topology and the second is to make the
+peer selection, per se, based on this knowledge. For the first part, the
+algorithm uses network tomography, a techni  que which suggests probing only
 from a large network's end points in order to infer its internal characteristics
@@
 medium &
-% tracerout and hop counts and calculation from both the tracker and the peers
+% traceroute and hop counts and calculation from both the tracker and the peers
 % (the later for all peers in the bootstrapping list) increases the control
@@
 BitTorrent}
-Qin et al., in their work in \cite{qlzg_biteffpeersel_2009}, propose a
+Qin et al., in their work in \cite{QLZG2009}, propose a
 clustering approach that differentiates peers in a swarm into local, intra-ISP
@@
 cluster. The evaluation of the algorithm is done in a similar artificial,
-contained laboratory enviromnent as in \cite{lcy_utaps_2008} and it shows up to
+contained laboratory enviromnent as in \cite{LCY2008} and it shows up to
 5\% faster downloading times, and up to 22\% reduction of the total cross-ISP
@@
 \subsubsection{Peer-exchange Routing Optimization Protocols (PROP)}
-\cite{qiu_prop_2007} introduces two protocols called \emph{Peer-exchange Routing
+\cite{QCYCZ2007} introduces two protocols called \emph{Peer-exchange Routing
 Optimization Protocols (PROP)} to adjust the neighbourhood graph of an overlay
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Resolving the Topology Mismatch Problem in Unstructured
-Peer-to-Peer Networks}
+% \subsubsection{Resolving the Topology Mismatch Problem in Unstructured
+% Peer-to-Peer Networks}
 % TODO: to be reviewed
 
-In \cite{hsiao_redblue_2009}, Hsiao et al, claim to construct topology-aware
-unstructured overlays that \emph{guarantee} performance qualities in terms of
-\begin{inparaenum}[\itshape i\upshape)]
-  \item the expected communication latency among any two overlay peers
-regardless of the network size, and
-  \item the broadcasting scope of each participating peer
-\end{inparaenum}
-. The algorithm constructs an undirected graph $G = \left( V, E \right)$
-comprised by two subgraphs. The first, namely $G^{\left( red \right)} = \left(
-V^{\left( red \right)}, E^{\left( red \right)} \right)$ in the paper's context,
-includes all vertices of $G$ and ensures the connectivity of the graph by
-securing at least one path between any two nodes. In contrast, $G^{\left( blue
-\right)} = \left( V^{\left( blue \right)}, E^{\left( blue \right)} \right)$,
-contains those vertices of $G$ that have free edges to link to other nodes and
-because these are fully utilized, the following also stands $E = E^{\left( red
-\right)} \cup E^{\left( blue \right)}$. A joining peer $u$, partitions its
-neighbours into two subsets, the $B_u^{\left( red \right)}$ and $B_u^{\left(
-blue \right)}$. In order to populate the $B_u^{\left( red \right)}$ subset, peer
-$u$ samples peers uniformly and at random. Then, each of these selected peers
-discovers a routing path starting from itself towards the node with the smallest
-(or the largest) ID in the system.
+% In \cite{hsiao_redblue_2009}, Hsiao et al, claim to construct topology-aware
+% unstructured overlays that \emph{guarantee} performance qualities in terms of
+% \begin{inparaenum}[\itshape i\upshape)]
+%   \item the expected communication latency among any two overlay peers
+% regardless of the network size, and
+%   \item the broadcasting scope of each participating peer
+% \end{inparaenum}
+% . The algorithm constructs an undirected graph $G = \left( V, E \right)$
+% comprised by two subgraphs. The first, namely $G^{\left( red \right)} = \left(
+% V^{\left( red \right)}, E^{\left( red \right)} \right)$ in the paper's context,
+% includes all vertices of $G$ and ensures the connectivity of the graph by
+% securing at least one path between any two nodes. In contrast, $G^{\left( blue
+% \right)} = \left( V^{\left( blue \right)}, E^{\left( blue \right)} \right)$,
+% contains those vertices of $G$ that have free edges to link to other nodes and
+% because these are fully utilized, the following also stands $E = E^{\left( red
+% \right)} \cup E^{\left( blue \right)}$. A joining peer $u$, partitions its
+% neighbours into two subsets, the $B_u^{\left( red \right)}$ and $B_u^{\left(
+% blue \right)}$. In order to populate the $B_u^{\left( red \right)}$ subset, peer
+% $u$ samples peers uniformly and at random. Then, each of these selected peers
+% discovers a routing path starting from itself towards the node with the smallest
+% (or the largest) ID in the system.
 
 % TODO: CHECK ONCE AGAIN
-\begin{center}
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
-%
-? &
-%
-? &
-%
-?
-\end{tabular}
-\end{center}
+% \begin{center}
+% \begin{tabular}{ccc}
+% \emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+% \hline
+%
+% ? &
+%
+% ? &
+%
+% ?
+% \end{tabular}
+% \end{center}
 
@@
 \subsubsection{Distributed Domain Name Order (DDNO)}
-\emph{Distributed Domain Name Order (DDNO)} \cite{z-yk_ddno_2005} uses the
-domain names to detect topologically close nodes based on the assumption that
-nodes within the same domain are topologically close to each other. DDNO is
-a heuristic approach to solve the topology mismatch problem and uses
+\emph{Distributed Domain Name Order (DDNO)} \cite{Z-YK2005} uses the domain
+names to detect topologically close nodes based on the assumption that nodes
+within the same domain are topologically close to each other. DDNO is a
+heuristic approach to solve the topology mismatch problem and uses
 \emph{Split-Hash} and \emph{dnMatch} algorithms to detect locality by using
@@
 \emph{Critical Topology-Aware Grouping (CTAG)} presented in
-\cite{zhao_ctag_2006}, is a grouping algorithm. The grouping strategy is based
+\cite{ZL2006}, is a grouping algorithm. The grouping strategy is
+based
 on the \emph{Internet Assigned Numbers Authority (IANA)} and the respective
@@
 \hline
-% The technique is partitions the system reducing the search efficiency
+% The technique partitions the system reducing the search efficiency
 low &
@@
 
-Landmark Binning \cite{ratnasamy_binning_2002} is an approach to partition close
+Landmark Binning \cite{RHMKS2002} is an approach to partition close
 by nodes into bins based on their distance to well known anchor nodes across the
@@
 thus it can be applied with no significant changes to both structured and
-unstructured P2P systems. \cite{ratnasamy_binning_2002} has the detailed
-description of such algorithm for both architectures. Landmark binning is also
-considered as a good candidate for content distribution networks in particular.
-The major disadvantage of landmark binning is to install and maintain landmark
-servers on different autonomous systems, all over the world. A typical P2P
-network usually has a couple of million nodes connected at any given time, thus
-rendering landmark servers an important parameter when designing for large
-scales.  The authors claim that latency estimation does not drain much from the
-network resources but in case scalability problems arise, they claim the
-solution is replacing single landmark servers with clusters of servers within
-the same physical area. However, this approach does not reduce the possibility
-of excessive network traffic flow through these landmark servers. One other
+unstructured P2P systems. \cite{RHMKS2002} has the detailed description of such
+algorithm for both architectures. Landmark binning is also considered as a good
+candidate for content distribution networks in particular. The major
+disadvantage of landmark binning is to install and maintain landmark servers on
+different autonomous systems, all over the world. A typical P2P network usually
+has a couple of million nodes connected at any given time, thus rendering
+landmark servers an important parameter when designing for large scales.  The
+authors claim that latency estimation does not drain much from the network
+resources but in case scalability problems arise, they claim the solution is
+replacing single landmark servers with clusters of servers within the same
+physical area. However, this approach does not reduce the possibility of
+excessive network traffic flow through these landmark servers. One other
 possible problem is the incorrect binning caused by the use of inaccurate
@@
 \subsubsection{mOverlay}
-\emph{mOverlay} \cite{zhang_moverlay_2004} tries to addresses the scalability
+\emph{mOverlay} \cite{ZZZSZ2004} tries to addresses the scalability
 issues that might be imposed to networks (i.e. load unbalance) when using static
@@
 
+\input{unstructured-plot}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 and leaving the network, cause large traffic overhead to the corresponding
-network \cite{chu_esm_2000,chu_esm_2002}. Thus, the original Narada algorithm is
+network \cite{CRZ2000,CRSZ2001,CRSZ2002}. Thus, the original Narada algorithm is
 designed for small groups therefore it is not scalable. Its followers try to

### 2014-01-27 background.tex
2012.10.20 (Vassilis) Images reworked. Text alligned with the new images.

@@
 \subfigure[Centralised (SETI@Home or Napster).] {
-  \includegraphics[scale=0.3]{img/centralised.png}
+  \includegraphics[scale=0.25]{img/pdf/centralised.pdf}
   \label{figure:p2p-archs:centralised}
@@
 \subfigure[Fully unstructured (Gnutella v.0.4).] {
-  \includegraphics[scale=0.3]{img/unstructured-full.png}
-  \label{figure:p2p-archs:unstructured-full}
+  \includegraphics[scale=0.25]{img/pdf/unstructured-full.pdf}
+  \label{figure:p2p-archs:unstructred-full}
 }\qquad\qquad
 \subfigure[Hierarchical unstructured (Gnutella v.0.6 or FastTrack).] {
-  \includegraphics[scale=0.3]{img/unstructured-hierarchical.png}
+  \includegraphics[scale=0.25]{img/pdf/unstructured-hierarchical.pdf}
   \label{figure:p2p-archs:unstructured-hierarchical}
@@
 \subfigure[Structured (Chord and Kademlia).] {
-  \includegraphics[scale=0.3]{img/structured.png}
+  \includegraphics[scale=0.25]{img/pdf/structured.pdf}
   \label{figure:p2p-archs:structured}
@@
 \centering
-  \includegraphics[scale=0.8]{img/phys.pdf}
+  \includegraphics[scale=0.7]{img/pdf/example-physical.pdf}
 \caption{Interconnection of nodes in the physical level.}
@@
 \subfigure[Nodes B and C have direct overlay connection.] {
-  \includegraphics[scale=0.8]{img/overlay-1.pdf}
+  \includegraphics[scale=0.5]{img/pdf/example-overlay-inefficient.pdf}
   \label{figure:overlay-1}
@@
 \subfigure[Nodes A and D have direct overlay connection.] {
-  \includegraphics[scale=0.8]{img/overlay-2.pdf}
+  \includegraphics[scale=0.5]{img/pdf/example-overlay-efficient.pdf}
   \label{figure:overlay-2}

### 2014-01-27 structured.tex
2012.10.20 (Vassilis) Images reworked. Text alligned with the new images.

@@
 
-\begin{figure*}
-\centering
-  \includegraphics[scale=0.8]{img/algorithms/global_softstate}
-\caption{(left) Position of $p$ in the space of $3$ landmarks. (right) Position
-of $p$ on the map.}
-\label{fig:global_softstate}
-\end{figure*}
+% \begin{figure*}
+% \centering
+%   \includegraphics[scale=0.8]{img/algorithms/global_softstate}
+% \caption{(left) Position of $p$ in the space of $3$ landmarks. (right) Position
+% of $p$ on the map.}
+% \label{fig:global_softstate}
+% \end{figure*}
 
@@
 scale distance probes to reveal the proximity properties of the underlying
-network to the overlay. Figure \ref{fig:global_softstate} presents the landmark
-space and the mapping of the position of the nodes. This global view of the
+network to the overlay. %Figure \ref{fig:global_softstate} presents the landmark
+%space and the mapping of the position of the nodes.
+This global view of the
 state is made available to all nodes in order to help them find the best way to
@@
 
-\begin{figure}[ht]
-\centering
-\subfigure[Example physical topology of nodes.] {
-  \includegraphics[scale=0.6]{img/algorithms/SAT_match}
-  \label{figure:sat_match:before}
-}\qquad\qquad
-\subfigure[(left) A jumps to B in order to match the logical topology to the
-physical. (right) State of the logical topology after the jump.] {
-  \includegraphics[scale=0.6]{img/algorithms/SAT_match2}
-  \label{figure:sat_match:after}
-}
-\caption{The SAT-Match algorithm.}
-\label{figure:sat_match}
-\end{figure}
+% \begin{figure}[ht]
+% \centering
+% \subfigure[Example physical topology of nodes.] {
+%   \includegraphics[scale=0.6]{img/algorithms/SAT_match}
+%   \label{figure:sat_match:before}
+% }\qquad\qquad
+% \subfigure[(left) A jumps to B in order to match the logical topology to the
+% physical. (right) State of the logical topology after the jump.] {
+%   \includegraphics[scale=0.6]{img/algorithms/SAT_match2}
+%   \label{figure:sat_match:after}
+% }
+% \caption{The SAT-Match algorithm.}
+% \label{figure:sat_match}
+% \end{figure}
 
@@
 is performed as a combination of \emph{leave} and \emph{join} operations, in the
-CAN context. Figure~\ref{figure:sat_match} demonstrates the selective jump on an
-example topology.
+CAN context.
+% Figure~\ref{figure:sat_match} demonstrates the selective jump on an
+% example topology.
 
@@
 % scale much better than, for example, Mithos.
-medium
-\end{tabular}
-\end{center}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{ \bf Delay Aware P2P System}
-A new \emph{Delay Aware P2P System (DAPS)} is introduced in \cite{ZL2005}. Its
-main goal is to reduce the time $L$ of a look-up request by dividing the routing
-tables of peers into several sectors in increasing delay. The source node
-emitting the query message defines a delay boundary or the pruning factor $L_t$
-in the paper's context. Request messages will be forwarded only to nodes whose
-delay less than or equal to $L_t$. With the clustered routing tables and the
-loose organisation the overlay network of \emph{DAPS} is between structured and
-unstructured.
-
-% TODO: MAYBE REVIEW THE EVALUATION
-\begin{center}
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
-%
-low &
-%
-low &
-%
 medium
@@
 \subsubsection{Geographic layout} \label{section:geographic_layout}
+
+\begin{figure}[ht]
+\centering
+  \includegraphics[scale=0.4]{img/pdf/geographic-layout.pdf}
+\caption{Successive nodes in the overlay circle can be chosen to be from the
+same or near autonomous systems.}
+\label{figure:geographic-layout}
+\end{figure}
+
 Inside geographic layout category fall those kind of methods which practically
-position physically close by nodes together in the application space as well.
-This is done by adjusting and maintaining the routing tables of all involved
-peers accordingly by exploiting proximity information that describe the overall
-geographic positioning of the peer nodes. Landmark servers and RTT measurements
-are two popular methods, which can also be used in conjunction, to discover
-physically close by peers over the network. However, as discussed in Section
-\ref{sec:landmark_binning}, these methods do not always give reliable estimates
-for the node positions over the internet. The landmark servers are not
-self-organizing and have maintenance overheads. To serve a P2P network with
-millions of peers, multiple landmark servers distributed over the whole world is
-required, which is hard to manage. The RTT measurements also can measure the
-delay between peers, but it is a greedy method, which can result non-optimal
-overlay topologies especially if close by nodes have low bandwidth connections
-among themselves.
+position physically close by nodes together in the application space as well
+(check Figure~\ref{figure:geographic-layout}). This is done by adjusting and
+maintaining the routing tables of all involved peers accordingly by exploiting
+proximity information that describe the overall geographic positioning of the
+peer nodes. Landmark servers and RTT measurements are two popular methods, which
+can also be used in conjunction, to discover physically close by peers over the
+network. However, as discussed in Section~\ref{sec:landmark_binning}, these
+methods do not always give reliable estimates for the node positions over the
+internet. The landmark servers are not self-organizing and have maintenance
+overheads. To serve a P2P network with millions of peers, multiple landmark
+servers distributed over the whole world is required, which is hard to manage.
+The RTT measurements also can measure the delay between peers, but it is a
+greedy method, which can result in non-optimal overlay topologies especially if
+close by nodes have low bandwidth connections among themselves.
 
@@
 \subsubsection{Proximity Routing}
+
+\begin{figure}[ht]
+\centering
+  \includegraphics[scale=0.4]{img/pdf/proximity-routing.pdf}
+\caption{Proximity routing takes proximity info into account during the
+routing process.}
+\label{figure:proximity-routing}
+\end{figure}
+
 Proximity routing does not require routing tables to be built using any
 knowledge about network proximity. On the other hand it exploits such knowledge
-in order to choose the best next hop during routing a message. This approach
-tries to balance between choosing the node that will further progress the
-routing towards the destination and choosing the closest entry in the routing
-table, in terms of network proximity. As this is very difficult to be achieved,
-the system might end up using a greedy approach that selects low latency paths
-over the overlay, which maps to suboptimal longer paths on the physical
-topology. Thus, in general, proximity neighbour selection is considered as a
-superior method to proximity routing, however, joint uses of these two protocols
-are also possible. Moreover, they are considered relatively less effective than
-geographical layout when applied to CAN(-like) implementations. Moreover, the
-technique has been incorporated into a version of Chord causing an increase on
-the overhead of node joins and the size as well as maintenance cost of finger
-tables.
+in order to choose the best next hop during routing a message as can be seen in
+Figure~\ref{figure:proximity-routing}. This approach tries to balance between
+choosing the node that will further progress the routing towards the destination
+and choosing the closest entry in the routing table, in terms of network
+proximity. As this is very difficult to be achieved, the system might end up
+using a greedy approach that selects low latency paths over the overlay, which
+maps to suboptimal longer paths on the physical topology. Thus, in general,
+proximity neighbour selection is considered as a superior method to proximity
+routing, however, joint uses of these two protocols are also possible. Moreover,
+they are considered relatively less effective than geographical layout when
+applied to CAN(-like) implementations. Moreover, the technique has been
+incorporated into a version of Chord causing an increase on the overhead of node
+joins and the size as well as maintenance cost of finger tables.
 
@@
 \subsubsection{Proximity neighbour selection}
+
+\begin{figure}[ht]
+\centering
+  \includegraphics[scale=0.4]{img/pdf/proximity-neighbour-selection.pdf}
+\caption{Proximity neighbour selection dictates that proximity is taken into
+acount during routing table maintenance.}
+\label{figure:proximity-neighbour-selection}
+\end{figure}
+
 \emph{Proximity Neighbour Selection} constructs the routing tables using
-proximity knowledge. The proximity information used in this method is different
-than the landmark based systems described in
-Section~\ref{section:geographic_layout}, as time-to-live values between nodes,
-or directly node ID prefixes are used to detect proximity. Tapestry, Pastry, and
-CAN successfully implemented the proximity into their algorithms by using this
-approach. The routing protocol in Pastry is based on longest node ID prefix
-matching, while CAN uses RTT values to detect close by nodes.
-\cite{CDCR2002a} reports that proximity neighbour selection as an
-effective proximity based method.
+proximity knowledge (see Figure~\ref{figure:proximity-neighbour-selection}). The
+proximity information used in this method is different from the the
+landmark-based systems described in Section~\ref{section:geographic_layout}.
+Here time-to-live values between nodes, or directly node ID prefixes are used to
+detect proximity. Tapestry, Pastry, and CAN successfully implemented the
+proximity into their algorithms by using this approach. The routing protocol in
+Pastry is based on longest node ID prefix matching, while CAN uses RTT values to
+detect close by nodes. \cite{CDCR2002a} reports that proximity neighbour
+selection as an effective proximity based method.

### 2014-01-27 unstructured.tex
2012.10.20 (Vassilis) Images reworked. Text alligned with the new images.

@@
 
-\begin{figure}[ht]
-\centering
-\subfigure[Iterative deepening with three levels.] {
-  \includegraphics[scale=0.4]{img/algorithms/iterative_deepening}
-  \label{figure:dbfs:iterdeep}
-}\qquad\qquad
-\subfigure[Directed BFS.] {
-  \includegraphics[scale=0.4]{img/algorithms/directed_bfs}
-  \label{figure:dbfs:dbfs}
-}\qquad\qquad
-\subfigure[Local indices with radius size equal to $2$.] {
-  \includegraphics[scale=0.4]{img/algorithms/local_indices}
-  \label{figure:dbfs:localindx}
-}
-\caption{Improving search in P2P networks}
-\label{figure:dbfs}
-\end{figure}
+% \begin{figure}[ht]
+% \centering
+% \subfigure[Iterative deepening with three levels.] {
+%   \includegraphics[scale=0.4]{img/algorithms/iterative_deepening}
+%   \label{figure:dbfs:iterdeep}
+% }\qquad\qquad
+% \subfigure[Directed BFS.] {
+%   \includegraphics[scale=0.4]{img/algorithms/directed_bfs}
+%   \label{figure:dbfs:dbfs}
+% }\qquad\qquad
+% \subfigure[Local indices with radius size equal to $2$.] {
+%   \includegraphics[scale=0.4]{img/algorithms/local_indices}
+%   \label{figure:dbfs:localindx}
+% }
+% \caption{Improving search in P2P networks}
+% \label{figure:dbfs}
+% \end{figure}
 
@@
 
-In iterative deepening (see Figure~\ref{figure:dbfs:iterdeep}), the search is
+In iterative deepening, %(see Figure~\ref{figure:dbfs:iterdeep})
+the search is
 performed on a BFS tree with multiple preset depths. The depth limit is
@@
 
-The directed BFS, as shown in Figure~\ref{figure:dbfs:dbfs}, tries to avoid this
-delay by forwarding the query messages only to a selected set of neighbours, in
-which the selection criteria varies from the number of results received
-previously, distance in terms of hops, bandwidth, or the query load of the
-neighbour.
-
-In local indices, as shown in Figure~\ref{figure:dbfs:localindx}, each node
+The directed BFS%, as shown in Figure~\ref{figure:dbfs:dbfs},
+tries to avoid this delay by forwarding the query messages only to a selected
+set of neighbours, in which the selection criteria varies from the number of
+results received previously, distance in terms of hops, bandwidth, or the query
+load of the neighbour.
+
+In local indices%, as shown in Figure~\ref{figure:dbfs:localindx},
+each node
 indexes data within a radius of $r$ hops and uses this local index to answer
@@
 medium \\
+\end{tabular}
+\end{center}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\paragraph*{ \bf Delay Aware P2P System}
+A new \emph{Delay Aware P2P System (DAPS)} is introduced in \cite{ZL2005}. Its
+main goal is to reduce the time of a look-up request by dividing the routing
+tables of peers into several sectors of increasing delay. The source node
+emitting the query message defines a delay boundary or the \emph{pruning
+factor} in the paper's context. Request messages are then forwarded only to
+nodes whose delay is less than or equal that boundary. With the clustered
+routing tables and the loose organisation of DAPS' overlay network it is
+considered by is between structured and unstructured.
+
+% TODO: MAYBE REVIEW THE EVALUATION
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+%
+low &
+%
+low &
+%
+medium
 \end{tabular}
@@
 
-\begin{figure}[ht]
-\centering
-\subfigure[Underlying network with edge costs.] {
-  \includegraphics[scale=0.4]{img/algorithms/narada}
-  \label{figure:narada:underlying}
-}\qquad\qquad
-\subfigure[Peer A sends a message to the rest using regular broadcast.] {
-  \includegraphics[scale=0.4]{img/algorithms/narada2}
-  \label{figure:narada:regu}
-}\qquad\qquad
-\subfigure[Peer A sends a message to the rest using end system multicast.] {
-  \includegraphics[scale=0.4]{img/algorithms/narada3}
-  \label{figure:narada:multicast}
-}
-\caption{A visualisation of the Narada protocol.}
-\label{figure:narada}
-\end{figure}
+% \begin{figure}[ht]
+% \centering
+% \subfigure[Underlying network with edge costs.] {
+%   \includegraphics[scale=0.4]{img/algorithms/narada}
+%   \label{figure:narada:underlying}
+% }\qquad\qquad
+% \subfigure[Peer A sends a message to the rest using regular broadcast.] {
+%   \includegraphics[scale=0.4]{img/algorithms/narada2}
+%   \label{figure:narada:regu}
+% }\qquad\qquad
+% \subfigure[Peer A sends a message to the rest using end system multicast.] {
+%   \includegraphics[scale=0.4]{img/algorithms/narada3}
+%   \label{figure:narada:multicast}
+% }
+% \caption{A visualisation of the Narada protocol.}
+% \label{figure:narada}
+% \end{figure}
 
@@
 construction
-}
-as shown in Figure~\ref{figure:narada}. They also keep the graph and the trees
+}.
+%as shown in Figure~\ref{figure:narada}.
+They also keep the graph and the trees
 dynamically updated while nodes continue to join and leave the network.
@@
 \subsubsection{Landmark Binning (LM)}\label{sec:landmark_binning}
-\begin{figure*}
-\centering
-  \includegraphics[scale=0.4]{img/algorithms/landmark_binning}
-\caption{Example 2D coordinate overlay and a sample routing path from node D to
-(x,y).}
-\label{fig:landmark_binning}
-\end{figure*}
+% \begin{figure*}
+% \centering
+%   \includegraphics[scale=0.4]{img/algorithms/landmark_binning}
+% \caption{Example 2D coordinate overlay and a sample routing path from node D to
+% (x,y).}
+% \label{fig:landmark_binning}
+% \end{figure*}
 
@@
 by nodes into bins based on their distance to well known anchor nodes across the
-Internet, as shown in Figure~\ref{fig:landmark_binning}. In order to detect
+Internet.%, as shown in Figure~\ref{fig:landmark_binning}.
+In order to detect
 locality, nodes use network latency (i.e. RTT) as a measurement technique. The
@@
 \subsubsection{Topology adaptation}
-Researchers initially started considering alternatives for services who
-implement one-to-many communication schemes, like IP multicast, on the
-application layer (i.e. Narada). Application level multicast protocols
-are generic protocols that can be applied to both P2P file sharing systems, as
-well as content distribution networks. Application layer multicast, however, is
-not as efficient as the one enabled by IP and still struggles to ease the
-difficulties imposed by the topology mismatch problem.
+\begin{figure}[ht]
+\centering
+\subfigure[Ineffictient overlay topology.] {
+  \includegraphics[scale=0.4]{img/pdf/topology-adaptation-before.pdf}
+  \label{figure:topology-adaptation:before}
+}\qquad\qquad
+\subfigure[Efficient overlay topology after adaptation.] {
+  \includegraphics[scale=0.4]{img/pdf/topology-adaptation-after.pdf}
+  \label{figure:topology-adaptation:after}
+}
+\caption{A simple example of how an overlay topology can be adapted.}
+\label{figure:topology-adaptation}
+\end{figure}
 
 The topology adaptation based protocols modify the topology of the P2P network
-using various techniques. The two commonly used approaches that have been
-investigated in this section are creating \emph{spanning tree}s using connection
-graphs, and creating \emph{clusters} of physically close nodes.
-
-Narada and subsequent algorithms like AOTO, LTM, SBO, try to the problem by
-building a richer connected graph and forming minimum spanning trees over this
-graph that can efficiently route messages among peers. Even though spanning
-trees provide efficient query performance, the construction and the update costs
-of the spanning tree, especially in dynamic environments with many nodes joining
-and leaving the network, cause large traffic overhead to the corresponding
-network \cite{CRZ2000,CRSZ2001,CRSZ2002}. Thus, the original Narada algorithm is
-designed for small groups therefore it is not scalable. Its followers try to
-solve the scalability problem by introducing various methods including forming
-minimum spanning trees for the two-hop neighbours ($N^2$) of each node,
-partitioning the graph into two random groups where each group is responsible of
-different tasks, or performing local optimizations dynamically on the overlay
-graph. The advantage of building minimum spanning trees is that they maintain
-the connectivity on the network in an efficient manner, while still preserving
-the overall search scope. However, building and maintaining a minimum spanning
-tree creates additional overhead for the network.
+using various techniques. The two commonly used approaches that can
+back such feature and were investigated in this section are creating
+\emph{spanning tree}s using connection graphs, and creating \emph{clusters} of
+physically close nodes.
+
+Narada and subsequent algorithms like AOTO, LTM, SBO, try to solve the problem
+by building a richer connected graph and forming minimum spanning trees over
+this graph that can efficiently route messages among peers. Even though spanning
+trees provide efficient query performance, their maintenance costs can be
+proven a real blocker. That is why the original Narada algorithm did not have
+any real topology adaptation mechanism meaning that it was designed for small
+groups (therefore not scalable). Its followers try to overcome these obstacles
+with clever tricks like forming minimum spanning trees for the two-hop
+neighbours ($N^2$) of each node, partitioning the graph into two random groups
+where each group is responsible of different tasks, or performing local
+optimizations dynamically on the overlay graph. The advantage of building
+minimum spanning trees is that they maintain the connectivity on the network in
+an efficient manner, while still preserving the overall search scope. However,
+their construction and the update costs, especially in dynamic, high-churn
+environments, cause large additional traffic overhead to the underlying
+network \cite{CRZ2000,CRSZ2001,CRSZ2002}.
 
@@
 reliable results, therefore mapping accuracy is not guaranteed. Moreover,
-traceroute for example, is a heavy weight utility to be frequently used on the
+traceroute, is a heavy-weight utility to be frequently used on the
 network, let alone the fact that many network equipment vendors do not allow
@@
 limitation by forcing half of each node's connections to be with other, randomly
-selected, nodes over the network
+selected, nodes over the network.
+
+Concluding, topology adaptation ultimatelly targets on the reduction of the
+average cost of the paths from one node to another. This way someone can deduce
+that the overlay network is better fit on top of the physical, underlying one.
+This deduction is fair in the general case but not always. For example an
+adaptive algorithm can exchange a slow edge in the overlay with a faster one
+thus reducing the average latency of the network communication. The pitfall here
+is that this new virtual link can traverse a fast AS-to-AS link meaning that
+even though is reduced message round-trip-time, it has more cost in terms of
+communication network economics and management.
 
@@
 \subsubsection{Forwarding optimisation}
-In the first realisation of the Gnutella protocol, each query request received
-by a peer was then forwarded to all of its neighbours thus generating
-unnecessary traffic over the network. Forwarding-optimisation-based approaches,
-propose intelligent neighbour selection for a message's next hop across the
-routing path, in order to replace that inefficient \emph{blind flooding} scheme
-of Gnutella's. The selection criteria vary depending on the algorithm, but they
-mainly use one or multiple statistical metrics that span from some obvious ones,
-like the latency of the link between the nodes or the exposure of specific
-features such as high processing bandwidth, high capacity, or low latency, to
-more sophisticated alternatives that may even take into account application
-level requirements such as the reliability of the neighbour.
-These appoaches have the advantage of enhancing the search responsiveness but on
-the other hand suffer from the drastic reduction of the search scope thus
-limiting the scalability of the whole network. This, in practice reduces the
-aggregate resource usage of the P2P network and improves the overall performance
-of the system. On the other hand, the algorithms do not offer a real solution to
-the topology mismatch problem. Thus, none can guarantee that overlay and
-underlying topologies are aligned with each other let alone quantify the
-mismatch and try to alleviate. Broadcast-based methods can be used in
-conjunction with other approaches to improve the quality of the P2P systems.
+
+\begin{figure}[ht]
+\centering
+\subfigure[Typical blind flooding.] {
+\includegraphics[scale=0.25]{img/pdf/forwarding-optimisation-before.pdf}
+  \label{figure:forwarding-optimisation:before}
+}\qquad\qquad
+\subfigure[A node can decide where to forward the messages.] {
+  \includegraphics[scale=0.25]{img/pdf/forwarding-optimisation-after.pdf}
+  \label{figure:forwarding-optimisation:after}
+}
+\caption{Forwarding optimisation in constraining message flooding.}
+\label{figure:forwarding-optimisation}
+\end{figure}
+
+% Researchers initially started considering alternatives for services who
+% implement one-to-many communication schemes, like IP multicast, on the
+% application layer (i.e. Narada). Application level multicast protocols
+% are generic protocols that can be applied to both P2P file sharing systems,
+% as well as content distribution networks. Application layer multicast, however,
+% is not as efficient as the one enabled by IP and still struggles to ease the
+% difficulties imposed by the topology mismatch problem.
+
+Forwarding optimisation can reduce the emitted messages flooding the overlay and
+as a concequence the overall burden, network resources have to handle. In the
+first realisation of the Gnutella protocol, each query request received by a
+peer, was then forwarded to all of its neighbours. This was the source of
+unecessary, sometimes redundant traffic. Forwarding-optimisation-based
+approaches, propose intelligent forwarding link selection for a message's next
+hop across the routing path. The selection criteria vary depending on the
+algorithm, but they mainly use one or more statistical heuristics. These can
+span from some obvious ones, like the latency of the link between the nodes or
+the exposure of specific features, such as high processing bandwidth, high
+capacity, or low latency, to more sophisticated alternatives that may even take
+into account application level requirements such as the reliability or
+friendship. Figure~\ref{figure:forwarding-optimisation} shows a simple example
+of how forwarding optimisation can work.
+Subfigure~\ref{figure:forwarding-optimisation:before} shows an algorithm that
+floods the entire network for search of an object the existence of which is
+denoted by black nodes. Alternativelly, a node can decide to forward its
+messages not to every output link but to a specific or subset of its output
+links. In
+Subfigure~\ref{figure:forwarding-optimisation:after}, the node in the middle,
+does not flood its neighbours. Instead it picks one of them. The dashed line on
+the right, shows that, for this routing process, the algorithm has rendered two
+potential forwarding paths as inefficient (e.g., the target nodes show
+overloading signs). These appoaches have the advantage of enhancing the search
+responsiveness and reducing the aggregate resource usage of the physical
+network. On the other hand, they suffer from drastic reduction of the search
+scope (in Subfigure~\ref{figure:forwarding-optimisation:after} the object on the
+far right is not reached) thus limiting the scalability of the whole network.
+Moreover they don't actually solve the topology mismatch problem since they
+don't give any guarantees that overlay and underlying topologies are aligned
+with each other let alone quantify the mismatch and try to alleviate. Like all
+other approaches, forwarding based methods are commonly applied in conjunction
+with other to improve the quality of the P2P systems.
 
@@
 \subsubsection{Caching and replication}
+
+\begin{figure}[ht]
+\centering
+\subfigure[Indexing can reduce the cost of the last hop.] {
+\includegraphics[scale=0.3]{img/pdf/replication-index.pdf}
+  \label{figure:replication:index}
+}\qquad\qquad
+\subfigure[Data replication allong the path of a successful query.] {
+  \includegraphics[scale=0.3]{img/pdf/replication-data.pdf}
+  \label{figure:replication:data}
+}
+\caption{Index and data replication strategies.}
+\label{figure:replication}
+\end{figure}
+
 Caching is a widely used technique to exploit locality and minimise redundant
@@
 parameters also have to be considered. P2P caching/replication can be applied
-on two different levels, namely caching indices or pointers to data or caching
-the data itself. Successfull implementations have already been developed in
-some commercial P2P systems, like KazaA, which uses caching of indices on its
-super-peer level. Unfortunatelly, even though the state of the art in P2P
-algorithms using caching methods reduce the resource usage of the network, they
-also fail to consistently address the probelem of the topology mismatch.
+on two different levels, namely caching indices or pointers to data (see
+Subfigure~\ref{figure:replication:index}) or caching
+the data itself (see
+Subfigure~\ref{figure:replication:data}). Successfull implementations have
+already been developed in some commercial P2P systems, like KazaA, which uses
+caching of indices on its super-peer level. Unfortunatelly, even though the
+state of the art in P2P algorithms using caching methods reduce the resource
+usage of the network, they also fail to consistently address typical problems
+of a topologicaly mismatched environment such as duplication of messages.
 
@@
 \subsubsection{Landmarking}\label{sec:landmark}
+
+\begin{figure}[ht]
+\centering
+  \includegraphics[scale=0.4]{img/pdf/landmarking.pdf}
+\caption{Landmark binning during node bootstrap.}
+\label{figure:landmarking}
+\end{figure}
+
 In landmark-based algorithms, nodes use network delay (e.g. RTT) as a
@@
 with similar distances to a set of landmarks, are physically close to each
-other, as well, over the network. The landmark based protocol has two important
-drawbacks. Firstly the network delay is not a reliable distance estimation
-method. For example, based on the load on the network the delay to certain nodes
-or networks can change from time to time, which will eventually affect the
-distance measurements and wrong measurements will lead to wrong estimated
-positions for the nodes, or incorrect and non optimal clusterings of the nodes.
-The second drawback of using landmark servers is the cost of installing and
-maintaining this kind of infrastructure across the whole Internet and for all
-the different autonomous system domains. As popular P2P file sharing
-applications usually have millions of peers connected at any time, it is not
-false to assume that the network costs of maintaining these landmark servers
-will be quite high. A possible solution to the scalability problem of the static
-landmark servers is to use ordinary nodes as dynamic landmarks once they
-estimate their own positions. Even though this approach scales much better than
-static landmark servers, still the measurement accuracy problem affects the
-overall performance of the system. Moreover,it can be characterised as a rather
-coarse-grained approximation, therefore not particularly well suited for
-detecting the correct positions of nodes within close distance to each other.
+other, as well, over the network. In Figure~\ref{figure:landmarking} the newlly
+arriving peer, measures its distance to an array of landmark point denoted as
+$Ln$ and and makes a sorted list of the peers, say in increasing order. Then in
+order to choose the already participating peers with which to connect to,
+compares its list to the list of the potential neighbour. Peers with similar
+measured distances to the these landmark point are very likely to be close to
+each other in the physical level as well.
+
+Landmark based protocols has four important drawbacks. Firstly the network
+delay is not a reliable distance estimation method. For example, based on the
+load on the network the delay to certain nodes or networks can change from time
+to time, which will eventually affect the distance measurements and wrong
+measurements will lead to wrong estimated positions for the nodes, or incorrect
+and non optimal clusterings of the nodes. Secondly, relying on predefined nodes
+make the whole paradigm not fully distributed and the landmark system prone to
+become a single point of failure. Moreover, the third drawback of using
+landmark servers is the cost of installing and maintaining this kind of
+infrastructure across the whole Internet and for all the different autonomous
+system domains. As popular P2P file sharing applications usually have millions
+of peers connected at any time, it is not false to assume that the network costs
+of maintaining these landmark servers will be quite high. A possible solution to
+the scalability problem of the static landmark servers is to use ordinary nodes
+as dynamic landmarks once they estimate their own positions. Even though this
+approach scales much better than static landmark servers, still the measurement
+accuracy problem affects the overall performance of the system. Moreover,it can
+be characterised as a rather coarse-grained approximation, therefore not
+particularly well suited for detecting the correct positions of nodes within
+close distance to each other.

### 2014-01-27 background.tex
2012.10.24 (Vassilis) Changed paper title. Text reworked (content clarifications added, typos corrected, broken references fixed).

@@
 and client to denote the twofold role of a node participating in the P2P
-network. Such serverless systems have proven to be able to achieve outstanding
+network. Such server-less systems have proven to be able to achieve outstanding
 aggregate resource capacities as more and more participants join the system
@@
 
-Designers of \emph{centralized} architectures
-(seeFigure~\ref{figure:p2p-archs:centralised}) were the first to recognise that
-requests for resources (e.g. CPU cycles or popular content) need
-not be sent to a dedicated server but instead could be handled by
-the many hosts that already posess the resources. The downside of this approach
-is that it requires a centralized searching facility which becomes a single
-point of failure, a scalability bottleneck, and a target of malicious attacks
-(e.g., \emph{Denial-of-Service (DoS) attacks}).
+Designers of \emph{centralized} architectures (see
+Figure~\ref{figure:p2p-archs:centralized}) were the first to recognize that
+requests for resources (e.g. CPU cycles or popular content) need not be sent to
+a dedicated server but instead could be handled by the many hosts that already
+posses the resources. The downside of this approach is that it requires a
+centralized searching facility which becomes a single point of failure, a
+scalability bottleneck, and a target of malicious attacks (e.g.,
+\emph{Denial-of-Service (DoS) attacks}).
 
 The most successful incarnation of the centralized approach is the file-sharing
-application \emph{Napster}. % TODO find citation for Napster again \cite{}
-Napster maintained a \emph{central index server} based on file lists provided by
-participating peers. The central index server was queried by users and it
-returned pointers to the actual content. Thus, by centralizing search while
-distributing downloads, Napster achieved a highly functional design that, at its
-height, attracted 26 million users sharing approximately 80 million
-songs~\cite{jmm_naptopusage_2001}. Napster's centralized nature was ultimately
-what brought it down after Recording Industry Association of America (RIAA) 
-filed a lawsuit for copyright infringement of the intellectual property of
-artists whose music was distributed by the Napster application. RIAA was able to
-point to the centralized indexing component of Napster as the entity responsible
-for illegal song-trading.
+application \emph{Napster}\cite{napster}. Napster maintained a \emph{central
+index server} based on file lists provided by participating peers. The central
+index server was queried by users and it returned pointers to the actual
+content. Thus, by centralizing search while distributing downloads,  Napster
+achieved a highly functional design that, at its height, attrated 26 million
+users sharing approximately 80 million songs\cite{jmm_naptopusage_2001}.
+Napter's  centralized nature was what ultimately brought it down after Recording
+Industry Association of America (RIAA) filed a lawsuit for copyright infringemnt
+of the intellectual property of artists whose music was distributed by the
+Napster application. RIAA was able to point to the centralized indexing
+component of Napster as the entity resposible for illegal song-trading.
 
@@
 known as \emph{broadcast-based} systems, because they use message flooding among
-peers to propagate search queries. A well-known example of the fully dentralized
-unstructured approach is Gnutella, in its first realisation (v.0.4)
+peers to propagate search queries. A well-known example of the fully
+decentralized
+unstructured approach is Gnutella, in its first realization (v.0.4)
 \cite{gnutellav04}.
@@
 received by a node, it decreases the TTL by one unit. If the TTL has not reached
-zero, the node forwards the query to its neighbours.  Otherwise, it drops the
+zero, the node forwards the query to its neighbors.  Otherwise, it drops the
 query. While this approach prevents queries from visiting the entire network, it
@@
 Figure~\ref{figure:p2p-archs:structured}). The key goal of these systems is to
-provide a self-organising infrastructure for large-scale P2P applications
+provide a self-organizing infrastructure for large-scale P2P applications
 \cite{ratnasamy_can_2001,stoica_chord_2001,antony_pastry_2001,zhao_tapestry_2001,maymounkov_kademlia_2002,rgrk_bamboo_2004}.
@@
 from distributed search engines to event notification infrastructures to
-cloud-based platforms.We cite several examples in the Introduction.
+cloud-based platforms. We cite several examples in Section~\ref{section:intro}.
 
-% TODO: Change these pictures to match look-n-feel
 \begin{figure}[ht]
 \centering
-\subfigure[Centralised (SETI@Home or Napster).] {
-  \includegraphics[scale=0.25]{img/pdf/centralised.pdf}
-  \label{figure:p2p-archs:centralised}
+\subfigure[Centralized (SETI@Home or Napster). Peers query the centralized
+entity (dotted lines) and then establish the connection to the data source.] {
+  \includegraphics[scale=0.25]{img/pdf/centralized.pdf}
+  \label{figure:p2p-archs:centralized}
 }\qquad\qquad
-\subfigure[Fully unstructured (Gnutella v.0.4).] {
+\subfigure[Fully unstructured (Gnutella v.0.4). Data is distributed randomly,
+the topology is also random and all participants are equal.] {
   \includegraphics[scale=0.25]{img/pdf/unstructured-full.pdf}
-  \label{figure:p2p-archs:unstructred-full}
+  \label{figure:p2p-archs:unstructured-full}
 }\qquad\qquad
-\subfigure[Hierarchical unstructured (Gnutella v.0.6 or FastTrack).] {
+\subfigure[Hierarchical unstructured (Gnutella v.0.6 or FastTrack). Certain
+peers (bigger circles) shoulder specific responsibilities (e.g., data
+indexing).] {
   \includegraphics[scale=0.25]{img/pdf/unstructured-hierarchical.pdf}
@@
 }\qquad\qquad
-\subfigure[Structured (Chord and Kademlia).] {
+\subfigure[Structured (Chord and Kademlia). Peers and data are placed into the
+overlay in a deterministic manner.] {
   \includegraphics[scale=0.25]{img/pdf/structured.pdf}
@@
 %node $\upsilon$ in a graph $G = \left( V, E \right)$ is defined as the ratio of
-%the existing connections between $\upsilon$'s neighbouring nodes to $\gamma
-%\times \left( \gamma - 1 \right)$, where $\gamma$ is the number of neighbouring
-%nodes of $\upsilon$. High cluster coefficient means that neighbouring nodes of
+%the existing connections between $\upsilon$'s neighboring nodes to $\gamma
+%\times \left( \gamma - 1 \right)$, where $\gamma$ is the number of neighboring
+%nodes of $\upsilon$. High cluster coefficient means that neighboring nodes of
 %any node $\upsilon$ likely connect one another.} \cite{faloutsos_powerlaw_1999,
@@
 Ideally, an overlay network should be constructed to achieve an optimal 
-mapping among the overlay paths and the underlying IP network links to avoid
-inefficient states. The problem of constructing an optimal overlay is referred
-to as the \emph{topology mismatch problem}, and is formally defined as follows:
+mapping among the overlay paths and the underlying network links to avoid
+inefficient states were redundant physical resources are utilised for the
+operation of a link on the overlay. The problem of constructing an optimal
+overlay is referred to as the \emph{topology mismatch problem}, and is formally
+defined as follows:
 \begin{definition}
@@
 link, known as the link's \emph{stress}~\cite{CRSZ2002}. Even in the case
-of an approach that provides mechanisms to a newcoming node to detect close-by
-peers at join time, the stochastic behaviour with which peers join and leave
+of an approach that provides mechanisms to a new-coming node to detect close-by
+peers at join time, the stochastic behavior with which peers join and leave
 the overlay network, called the network's churn, further undermines the effort
-for maintaing the best position in the network throughout the life-time of a
+for maintaining the best position in the network throughout the life-time of a
 peer.
@@
 the Internet. Unfortunately, topology mismatch is known to be an NP-Hard problem
-\cite{NPBOOK,chawathe_scattercast_2000}. Adding to the complexity of the problem
+\cite{NPBOOK,C2000}. Adding to the complexity of the problem
 is the fact that end-to-end latencies demonstrate \emph{triangle inequality
@@
 so, we dedicate a discussion subsection after the presentation of the algorithms
-for both unstructred and structured architectures. This section offers:
+for both unstructured and structured architectures. This section offers:
 \begin{inparaenum}[\itshape i\upshape)]
   \item further discussion that can help the reader identify, through
-a taxonomy, the pluses and minuses in the preceeding material,
-  \item a comprehensive tabular organisation of the key-point information, and
+a taxonomy, the pluses and minuses in the preceding material,
+  \item a comprehensive tabular organization of the key-point information, and
   \item an attempt to give a pictorial comparison of the relative performance of
@@
 In discussion we offer the reader a deeper insight for the observations that
-pushed research toward some direction. Thus, we try to categorise these
-approaches so that we can organise directions making them comprehensible
+pushed research toward some direction. Thus, we try to categorize these
+approaches so that we can organize directions making them comprehensible
 and helping the reader build critical thinking on the advantages and
@@
 Pictorial comparison consists of bar charts that give a relative feeling of the
-perfomance of the various algorithms, discussed in the corresponding section, in
+performance of the various algorithms, discussed in the corresponding section,
+in
 terms of \emph{efficiency}, \emph{overhead} and \emph{scalability}; three
@@
 effecting other characteristics of the networks (e.g. BitTorrent's optimal
-dowload performace). To name few, we assume that efficieny is
+download performance). To name few, we assume that efficiency is
 negatively effected when an algorithm reduces the search scope. Moreover, the
-minimum-spanning-tree-based tecniques have slow convergence speed in topology
+minimum-spanning-tree-based techniques have slow convergence speed in topology
 adaptation or the high clustering impacts load balancing and constrains search
@@
 join, message forwarding and topology maintenance. For example when exchanging
-cost tables only along neighbours the traffic generated can be considered
-trivial. Also duplicates minimisation is a plus and lazy or heart-beat-based
+cost tables only along neighbors the traffic generated can be considered
+trivial. Also duplicates minimization is a plus and lazy or heart-beat-based
 overlay control mechanisms impose different amount of stress to the network.
-Nevertheless, caching/replication helps reducing the averall overhead by
-constraining the query from spreading too much. Minimum-spanning-tree-based
-approaches's low convergence speed we mentioned above for efficiency means
-additional computational and communicational overhead. % TODO: We check
-% complexity in the asymptotic sence, using \emph{Big O} notation % in order to
+Nevertheless, caching/replication helps reducing the overall overhead by
+constraining the query from spreading too much. Low convergence speed of
+minimum-spanning-tree-based approaches, we mentioned above for efficiency,
+results also in additional computational and communicational overhead.
+% TODO: We check
+% complexity in the asymptotic sense, using \emph{Big O} notation % in order to
 % check how the algorithm responds t. We take into account topology % creation and
@@
 % TODO: HELP NEEDED!
-With scalability we check the systems behaviour with respect to the number of
+With scalability we check the systems behavior with respect to the number of
 participating peers. Low, medium and high scalability can be generally
-considered in the order of hundrends, tenths of thousands and millions,
-respectivelly. Fully distributed aproaches is a plus (no system-centric entities
-involved, no need for global knowledge) and topology adaptation streanghtens
+considered in the order of hundreds, tenths of thousands and millions,
+respectively. Fully distributed approaches is a plus (no system-centric entities
+involved, no need for global knowledge) and topology adaptation strengthens
 scalability in highly dynamic environments. Reduced search scope, on the other
-side, penaltises the scalability since the more peers comming in the network the
-more potentially unsatisfied queries there will be. Furthermore, clock
-synchonisation is a limiting factor to scalability of a distributed systems.
-Again to name a few.
+hand, penalizes the scalability since the more peers coming in the network the
+more potentially unsatisfied queries will there be. Furthermore, clock
+synchronization is a limiting factor to scalability of a distributed system.
 

### 2014-01-27 intro.tex
2012.10.20 (Vassilis) Images reworked. Text alligned with the new images.
2012.10.24 (Vassilis) Changed paper title. Text reworked (content clarifications added, typos corrected, broken references fixed).

@@
 users sharing 80 million songs at the height of its popularity. Other similar
-applications quickly followed suit including Gnutella and its descendent,
+applications quickly followed suit including Gnutella and its descendant,
 Limewire, enjoying 3 million concurrently connected peers by the end of 2006
 and BitTorrent, connecting over 150 million users. P2P telephony has seen
-explosive growth with the advent of Skype. Introduced in 2003, Skype has
-experienced explosive popularity with more that 650 million users in 2011
+explosive growth with the advent of Skype. Since its introduction in 2003,
+Skype has become extremely popular with more than 650 million users in 2011
 \cite{skypetotalusers} and 35 million skypers concurrently online on the 5th of
@@
 network} is a virtual system of nodes and logical links, that is built on top of
-an existing network with the aim of providing a service (such as content
-distribution) that is not available in the existing one. Nodes in the overlay,
-communicate via virtual connections, each of which may correspond to a path of
-possibly many physical links in the underlying network.
+an existing network intending to provide added value to it (e.g., offering QoS
+content distribution over the best effort Internet network). Nodes in the
+overlay communicate via virtual connections each of which may correspond to a
+path of possibly many physical links in the underlying network.
 Figure~\ref{figure:overlay} illustrates a simple four-node overlay constructed
@@
 \centering
-  \includegraphics[scale=0.5]{img/p2p.pdf}
+  \includegraphics[scale=0.7]{img/pdf/under-over-lay.pdf}
 \caption{An example overlay network.}
@@
 
-If the overlay network is constructed such that the topology does not match the 
-underlying IP topology well, the resulting \emph{topology mismatch} (See
+If the overlay network is constructed such that it does not match the underlying
+topology well, the resulting \emph{topology mismatch} (See
 Section~\ref{section:background} for a more formal definition) can result in two
-significant problems. First, application performance (e.g. search query latency
-or jitter) can be adversely affected since application traffic must flow over a
-larger number of IP hops than if the overlay were constructed more wisely.
-Second, studies have shown that P2P applications contribute the largest portion
-of overall Internet traffic
+significant problems. First, the performance of the application per se, can be
+adversely affected since application traffic must flow over a larger,
+redundunt, number of physical hops resulting in poor user experience like
+extensive search latency or jitter. Second, can cumulatively degrade the
+performance of the whole underlying infrastructure, effecting other applications
+as well. For example, studies have shown that the extreme popularity of P2P
+applications amply deployed everywhere nowadays, result in them contributing the
+largest portion of overall Internet traffic
 \cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}, with
@@
 at the edge of the Internet. If the overlay topology of a peer-to-peer system is
-not efficintly constructed, then this burden to the Internet's backbone
+not efficiently constructed, then this burden to the Internet's backbone
 infrastructure could be even greater. For example, traffic might have to flow
@@
 
-For over a decade, several tens of researchers have published profusely on the 
-topology mismatch problem.  In this paper, we survey this research area and
-provide a taxonomy of the proposed solutions. We point out synergies as well as
+For over a decade, several researchers have published profusely on the topology
+mismatch problem.  In this paper, we survey this research area and provide a
+taxonomy of the proposed solutions. We point out synergies as well as
 similarities and differences in the published approaches. Ultimately, our goal
@@
 
-The rest of the paper is as follows: In Section~\ref{section:background}, we
-provide some background information on peer-to-peer overlay architectures
-including centralized, decentralized unstructured and decentralized structured
-peer-to-peer systems. We also formally define the problem of topology mismatch
-between an overlay and its underlying physical network and elaborate on the
-rationale behind our coarse grained evaluation of the techinques presented
-throughout this survey. In Sections~\ref{section:unstructured} and
-\ref{section:structured}, we present the various research efforts for
-decentralized unstructured and decentralized structured peer-to-peer systems,
-respectively. Finally, with Section~\ref{section:conclusion}, we
-conclude.
+The rest of the paper is organized as follows: In
+Section~\ref{section:background}, we provide background information on
+peer-to-peer overlay architectures including centralized, decentralized
+unstructured and decentralized structured peer-to-peer systems. We also formally
+define the problem of topology mismatch between an overlay and its underlying
+physical network and elaborate on the rationale behind our coarse grained
+evaluation of the techniques presented throughout this survey. In
+Sections~\ref{section:unstructured} and \ref{section:structured}, we present the
+various research efforts for decentralized unstructured and decentralized
+structured peer-to-peer systems, respectively. Finally, with
+Section~\ref{section:conclusion}, we conclude the paper.

### 2014-01-27 mard-survey.tex
2012.10.20 (Vassilis) Images reworked. Text alligned with the new images.
2012.10.24 (Vassilis) Changed paper title. Text reworked (content clarifications added, typos corrected, broken references fixed).

@@
 \usepackage{pgfplots}
+\usetikzlibrary{patterns}
 
@@
 
-\title{Alleviating the Topology Mismatch Problem in Unstructured and Structured
-Overlay Networks: A Survey}
+\title{Alleviating Topology Mismatch in Distributed Overlay Networks: A Survey}
 % TODO: We added here "P2P". I'm not sure, because for example Narada is not P2P
@@
 overlay architectures. We visit the various algorithms available in the
-lietrature, focusing on the presentation of how each of them tries to tackle the
-problem. Additionally, we provide a finer grained categorisation of the
+literature, focusing on the presentation of how each of them tries to tackle the
+problem. Additionally, we provide a finer grained categorization of the
 approaches based on the generic type of methodology they use in order to
@@
 Finally we discuss the advantages, disadvantages and novelties introduced by the
-algotithms and offer pictorial comparisons for them in order to further
-strengthen and organise understanding of the information and knowledge cited, in
+algorithms and offer pictorial comparisons for them in order to further
+strengthen and organize understanding of the information and knowledge cited, in
 an attempt to make this survey a one-stop-shop reference for anyone interested
@@
 
-% \item Proximity Neighbour Selection\\
+% \item Proximity Neighbor Selection\\
 %Finally, the third approach, constructs the routing tables using proximity
@@
 %\cite{castro_proximityp2p_2002} argues on how Pastry exploits proximity
-%neighbour selection in order to create a scheme that is (more) location-aware
+%neighbor selection in order to create a scheme that is (more) location-aware
 %compared to the other well-known DHTs (CAN, Chord).

### 2014-01-27 structured.tex
2012.10.24 (Vassilis) Changed paper title. Text reworked (content clarifications added, typos corrected, broken references fixed).

@@
 information to optimize them towards this direction. The algorithms are here
-categorised based the previous work done by \cite{CDHR2002,CDCR2002,RSS2002}.
+categorized based on the previous work done by \cite{CDHR2002,CDCR2002,RSS2002}.
 
@@
 %###############################################################################
-%       GEGRAPHIC LAYOUT
+%       GEOGRAPHIC LAYOUT
 %###############################################################################
@@
 information a hybrid approach is proposed, which uses landmark clustering as a
-preprocessing step in order to select a number of potential nearest neighbour
+preprocessing step in order to select a number of potential nearest neighbor
 candidates and then refine the selection by incorporating a
@@
 context or a set of nodes sharing a particular prefix in overlays such as
-Pastry.} maps of physical information are built and published where each node
-may appear in a maximum of $log\left( N \right)$ such maps. To dynamically adapt
-to changing network conditions, a node subscribes to relevant \emph{soft states}
-that utilize a notification system in order to initialize any necessary
-neighbour re-selection.
+Pastry\cite{antony_pastry_2001}.} maps of physical information are built and
+published where each node may appear in a maximum of $log\left( N \right)$ such
+maps. To dynamically adapt to changing network conditions, a node subscribes to
+relevant \emph{soft states} that utilize a notification system in order to
+initialize any necessary neighbor re-selection.
 
@@
 Nevertheless, static recognition of such nodes is currently done based on BGP
-reports and pre-chosen landmarks, sacrificing the self-organising attribute of
+reports and pre-chosen landmarks, sacrificing the self-organizing attribute of
 traditional DHTs.
@@
 % \item Expanding Ring Search\\
-%Expanding ring search can be of two forms. First it can utilize the multicast
+%Expanding ring search can be of two forms. First it can utilize the multi-cast
 %infrastructure in the underlying network in order to emit its messages.
@@
 %implementations of this way of generating proximity information is limited to
-%blindly flooding the neighbourhood to obtain reasonable results.
+%blindly flooding the neighborhood to obtain reasonable results.
 %
@@
 % Unable to identify nodes that are close to routers but it can reduce routing
-% latency to far away node by approximatelly 50-75%.
+% latency to far away node by approximately 50-75%.
 medium &
@@
 high &
-% Statically implemented sacrificing the decentralised nature of DHTs and
-% thus poorly scalling.
+% Statically implemented sacrificing the decentralized nature of DHTs and
+% thus poorly scaling.
 low
@@
 probing the neighborhood, the candidate neighbor list is updated. In order to
-avoid a local minima, Mithos probes all the neighbours within a two-hop distance
+avoid a local minima, Mithos probes all the neighbors within a two-hop distance
 from the current minimum before concluding the process.
 
-After finding its first neighbour, the newcomming peer is assigned an ID using
-information gathered during the iterative neighbour selection phase. Virtual
-coordinates are assigned to the newcomming peer by using the distances of two
-closest nodes and their neighbours, so that Euclidean distances between the node
+After finding its first neighbor, the new coming peer is assigned an ID using
+information gathered during the iterative neighbor selection phase. Virtual
+coordinates are assigned to the new coming peer by using the distances of two
+closest nodes and their neighbors, so that Euclidean distances between the node
 and all known hosts predict the network latency between them
@@
 
-The last step of the algorithm is the interconnection among neighbours. Mithos
+The last step of the algorithm is the interconnection among neighbors. Mithos
 uses a \emph{quadrant-based} mechanism according to which each node establishes
-a link to the closest neighbour in each quadrant. During forwarding, the next
-hop is performed towards a neighbour in the same quadrant as the final
-destination. The newcoming node may not know of other neighbours in all
-quadrants, therefore, the node first identifies neighbours in all quadrants
+a link to the closest neighbor in each quadrant. During forwarding, the next
+hop is performed towards a neighbor in the same quadrant as the final
+destination. The new coming node may not know of other neighbors in all
+quadrants, therefore, the node first identifies neighbors in all quadrants
 using a mechanism based on ideas similar to a perimeter walk\footnote{Used in
@@
 %
-%In order to avoid local minima during neighbour detection, extensive probing
+%In order to avoid local minima during neighbor detection, extensive probing
 %  must be undertaken. In simulation, unfortunately, only very small-sized
 % overlay topologies (of 200 to 1000 nodes) have been used and thus no safe
-% conclusions can be made as for the behaviour of an extensively large,
+% conclusions can be made as for the behavior of an extensively large,
 % real-world p2p deployment of the scheme.
@@
 \hline
-% It has very good results in reducing comunication latency among peers but its
-% efficience is being constrained by the fact that the proximity is calculated
-% at bootstrap, assigning virtual coordinates that are used througout of a
+% It has very good results in reducing communication latency among peers but its
+% efficiency is being constrained by the fact that the proximity is calculated
+% at bootstrap, assigning virtual coordinates that are used throughout of a
 % nodes life cycle and thus the overlay is not adapting to a high-churn
@@
 medium &
-% Incrimental probing is applied in quadradants and in two-hop distance at each
+% Incremental probing is applied in quadrants and in two-hop distance at each
 % step that does not generate a lot o control overhead.
@@
 \subsubsection{LAPTOP}
-\emph{LAPTOP} \cite{WLH2007}, organises the overlay into a tree-based hierarchy
+\emph{LAPTOP} \cite{WLH2007}, organizes the overlay into a tree-based hierarchy
 with main focus on reducing hops during message routing as well as minimizing
@@
 %\begin{enumerate}
-% \item The amplitude of all possible measured RTTs is devided into intervals.
+% \item The amplitude of all possible measured RTTs is divided into intervals.
 %  Each node measures its distance to its parent and is assigned a label $L_i$
@@
 %
-%The newcommer $N$ locates the root node and the later responds with a list of
+%The new comer $N$ locates the root node and the later responds with a list of
 %  $L_1$ nodes. $N$ then probes each of the $L_1$ nodes in search for the
 % closest one. If the measured RTT to the closest $L_1$ falls into the first
-% interval then the newcommer becomes a $L_1$ node as well. Otherwise node $N$
+% interval then the new comer becomes a $L_1$ node as well. Otherwise node $N$
 % sets the closest $L_1$ node as its potential parent node. This potential
@@
 
-% At join process the newcomming node is assigned its level label as well as its
+% At join process the new coming node is assigned its level label as well as its
 % address by its parent node. Additionally, it initializes its routing table
@@
 %failure, after the children of the failed node detect their parent's absence
-%(throught the aformentioned heartbeat mechanism), they start emmiting special
+%(through the aforementioned heartbeat mechanism), they start emitting special
 %messages to their grandparent node (the address of which is stored during
@@
 %period of time, and then chooses the one with the lowest level label as the
-%takeover node. Potential ties are broken by favouring the lower RTT. The
+%takeover node. Potential ties are broken by favoring the lower RTT. The
 %parent of the failed node, finally informs accordingly all the children about
@@
 %
-% TODO: HERE STRUCTURED LANDMARK BINING APPROCH LIES
+% TODO: HERE STRUCTURED LANDMARK BINNING APPROACH LIES
 %
@@
 % landmark nodes. The coordinate space is then partitioned into $m!$ equally
-%sized portions, each corresonding to a single ordering of the landmarks. To do
-%this, the first dimension is divided into $m$ areas each of which is furtherly
+%sized portions, each corresponding to a single ordering of the landmarks. To do
+%this, the first dimension is divided into $m$ areas each of which is further
 %divided (second dimension) into $m - 1$ sections and so on. Having set this $m$
@@
 \hline
-% It minimises routing length, but may suffer load imbalance since nodes higher
+% It minimizes routing length, but may suffer load imbalance since nodes higher
 % in the tree hierarchy are more likely to route most of the messages in the
@@
 medium &
-% Heartbeat approach incurs extra control overhead to the network. Fortunatelly
+% Heartbeat approach incurs extra control overhead to the network. Fortunately
 % involves only the parent and the children of the departing/failing node.
@@
 %An underlay metric provides information about the underlay network. The paper
-%  distingushes between three kinds of possibilities to quantify it:
+%  distinguishes between three kinds of possibilities to quantify it:
 %\begin{itemize}
@@
 %\end{itemize}
-The paper proposes two approaches for overlay optimisation. One falling into
-{\em proximity routing} and the other on {\em proximity neighbour selection}
-approaches overlay optimization techiques for exploiting proximity. For the
+The paper proposes two approaches for overlay optimization. One falling into
+{\em proximity routing} and the other on {\em proximity neighbor selection}
+approaches overlay optimization techniques for exploiting proximity. For the
 first, the goal is in choosing the best next hop during routing a message. Due
@@
 the lookup that chooses each next hop from a set of candidate nodes. Standard
-Kedemlia always chooses the closest node with respect to the XOR metric but with
+Kademlia always chooses the closest node with respect to the XOR metric but with
 proximity routing selection, Kademlia chooses the one with the smallest underlay
@@
 
-Toward the enhancement of traffic locality, the authors also use information
+Towards the enhancement of traffic locality, the authors also use information
 from the MaxMind GeoIP database\footnote{MaxMind Geolocation Technology.
@@
 
-As already mentioned the paper proposes \emph{proximity neighbour selection}
-method, as well, for overlay optimisation. The goal, according to this, is to
+As already mentioned the paper proposes \emph{proximity neighbor selection}
+method, as well, for overlay optimization. The goal, according to this, is to
 keep peers with the least contact cost in the routing table. As Kademlia
@@
 simply examining the upper 32-bits of its IPv6 address. Moreover, similar to
-Chord, CHOP6 uses a finger table, whose entries hold more than
-one candidate node.
+Chord\cite{stoica_chord_2001}, CHOP6 uses a finger table, whose entries hold
+more than one candidate node.
 
@@
 % There are three cases in which a node chooses the next hop according to
-% information it posseses.
+% information it possesses.
 %\begin{itemize}
@@
 % table entry. In this case, the source node chooses the one with the smallest
-% RTT with propability $p$. If there is a node with no measured RTT then the
+% RTT with probability $p$. If there is a node with no measured RTT then the
 % sender can select such a node with probability $1 - p$
@@
 \subsubsection{T2MC}
-T2MC \cite{SLCGZ2008} uses traceroute logs to detect different autonomous
+T2MC \cite{SLCGZ2008} uses trace-route logs to detect different autonomous
 systems and cluster close by peers, thus reducing redundant traffic that crosses
@@
 structure of the Internet routers to guide clustering. The peer chooses the
-minimum and maximum Latency results from the traceroute to initialize the
-centroinds of two sets. Then the peer calculates, for all hops allong its
-tracerouted path, the absolute distance to these centroids of both sets and
-assigns the routers to that centroid with which it has the smaller absolute
-distance. After that, the mean and variance values are calculated and if the
-variance is larger than a predefined threshold then the algorithm sets the two
-latency mean values as new centroids for the two sets and loops again.
-Ultimately peer will end up with two sets having the minimum intra-set variance.
-The peer, then, chooses the router from the second set with the minimum hops
-attribute and sets it as a threshold. The selected router and all others whose
-hops attribute is larger than the threshold are classified as \emph{remote}
-router cluster. The remaining are classified as \emph{near}. From the near
-class, the peer chooses the one with maximum hops attribute as its edge router,
-and registers it along with the all the near cluster into the DHT of the P2P
-overlay. As new peers join the network, those that share the same edge router
-or any of the members of the near router clustersm there would gather to form a
-close peer cluster. As Edge routers can provide more valuable information than
-other members of the near set, T2MC was designed to prioritize interaction of
-peers and edge gateways.
+minimum and maximum Latency results from the trace-route to initialize the
+centroids of two sets. Then the peer calculates, for all hops along its
+trace-routed path, the peer calculates the absolute distance to these centroids
+of both sets and assigns the routers to that centroid with which it has the
+smaller absolute distance. After that, the mean and variance values are
+calculated and if the variance is larger than a predefined threshold then the
+algorithm sets the two latency mean values as new centroids for the two sets and
+loops again. Ultimately peer will end up with two sets having the minimum
+intra-set variance. The peer, then, chooses the router from the second set with
+the minimum hops attribute and sets it as a threshold. The selected router and
+all others whose hops attribute is larger than the threshold are classified as
+\emph{remote} router cluster. The remaining are classified as \emph{near}. From
+the near class, the peer chooses the one with maximum hops attribute as its edge
+router, and registers it along with all near clusters into the DHT of the
+P2P overlay. As new peers join the network, those that share the same edge
+router or any of the members of the near router clusters, would form a close
+peer cluster. As Edge routers can provide more valuable information than other
+members of the near set, T2MC was designed to prioritize interaction of peers
+and edge gateways.
 
@@
 %
-%The use of Traceroute as a tool for implementing the distance measuring
-%infrastructure raise concearns about its efficiency and scalability. Being,
-%primarilly, a network diagnostic utility,
-% Even though traceroute provides detailed information about the network
+%The use of Trace-route as a tool for implementing the distance measuring
+%infrastructure raise concerns about its efficiency and scalability. Being,
+%primarily, a network diagnostic utility,
+% Even though trace-route provides detailed information about the network
 % structure, its excessive use creates additional overhead that stresses further
@@
 high &
-% traceroute is a very heavy tool and thus incurs excessive burden to the
-% network resourses.
+% trace-route is a very heavy tool and thus incurs excessive burden to the
+% network resources.
 high &
-% Traceroute is categorised as too heavy-weighted and intrucive for use in a
+% Trace-route is categorized as too heavy-weighted and intrusive for use in a
 % larger scale system \cite{RHMKS2002}. Additionally, disabling
@@
 other nodes it applies a heuristic mechanism to fill up the list. Entries are
-dynamicaly added or removed as the network state changes. The overlay
-maintenance cost can be considered similar to Chord's since besides node
-join/leave and cost of finger table entries heartbeat, PChord adds heartbeat of
-nodes inside the proximity list but since this list always contains nodes with
-the lowest RTT the added cost is kept to a minimal.
+dynamically added or removed as the network state changes. The overlay
+maintenance cost can be considered similar to Chord's. Besides node join/leave
+and cost of heartbeat messages for connection verification, PChord adds
+heartbeat of nodes inside the proximity list but since this list always contains
+nodes with the lowest RTT the added cost is kept to a minimal.
 
@@
 % only once in the routing path, which directly results in the lower RDP of
-% routing efficiency. Meanwhile, PChord keeps light-weightmaintenance cost as
+% routing efficiency. Meanwhile, PChord keeps light-weight maintenance cost as
 % Chord does.
@@
 \subsubsection{AChord}
-\emph{AChord} \cite{DK2006} uses IPv6's anycast mechanism in order to
-releave the protocol from complex joining procedures, and achieve high accuracy
-of network proximity giving high routing efficiency, Anycast delivers a message
-comming from the outside of an anycast group to the physicaly closest node in
+\emph{AChord} \cite{DK2006} uses IPv6's any-cast mechanism in order to
+relieve the protocol from complex joining procedures, and achieve high accuracy
+of network proximity giving high routing efficiency, Any-cast delivers a message
+coming from the outside of an any-cast group to the physically closest node in
 that group. AChord organizes all nodes participating in the overlay network into
-an anycast group. Any joining node comes outside of that group and, thus, is
-automaticaly forwarded to the physicaly nearest node in order to bootstrap,
+an any-cast group. Any joining node comes outside of that group and, thus, is
+automatically forwarded to the physically nearest node in order to bootstrap,
 avoiding the need to, explicitly, maintain such nodes, and the effort of finding
-a way of locating the physically nearest among them. The ID of the new comming
+a way of locating the physically nearest among them. The ID of the new coming
 node is computed based on the ID of the bootstrap node and the bootstrap's
 predecessor in a way that the constructed ID will be positioned between the
-aformentioned two. After joining, the finger table is build the same way as in
+aforementioned two. After joining, the finger table is build the same way as in
 the Chord protocol. Moreover, additional nodes are maintained in a structure
-called \emph{neighbourship table} which stores information about the closest
-known node. The routing decision is made by using both the neighbourship and
-the finger table. The authors claim this to be a lightweight approach that
-requires few changes to Chord in order not to affecting its own advantageous
-characteristics.
+called \emph{adjacency table} which stores information about the closest known
+node. The routing decision is made by using both the adjacency and the finger
+table. The authors claim this to be a lightweight approach that requires few
+changes to Chord in order not to affecting its own advantageous characteristics.
 
@@
 medium &
-% Releaves the system from joining overhead and topology aware ID construction
+% Relieves the system from joining overhead and topology aware ID construction
 % but doubles the information needed to be stored and maintained by the
-% introduction of neighbouring table
-medium &
-% assumes the operation of ideal underlying anycast mechanism (meaning that
+% introduction of adjacency table
+medium &
+% assumes the operation of ideal underlying any-cast mechanism (meaning that
 % outside messages are always routed to the nearest node which in the current
@@
 reduces inter-domain traffic between service providers. The difference between
-Chord6 and Chord the identifier definition. Therefore, the approach can be
-easily portable to other DHTs such as CAN, Pastry and Tapestry. In Chord6 the
-identifier contains two parts: the higher bits are obtained by hashing the
+Chord6 and Chord relies on the identifier definition. Therefore, the approach
+can be easily portable to other DHTs such as CAN, Pastry and Tapestry. In Chord6
+the identifier contains two parts: the higher bits are obtained by hashing the
 node's IPv6 address prefix of specific length, while the remaining lower bits
@@
 overlay network, which avoids unnecessary message forwarding across different
-service providers, thus minimising overall routing cost.
+service providers, thus minimizing overall routing cost.
 
@@
 % domain would have close identifiers, nodes in two close domains may have very
-% different ones reducing the overal efficience of the approach in general when
+% different ones reducing the overall efficiency of the approach in general when
 % considering inter-domain average path length.
@@
 %###############################################################################
-%       PROXIMITY NEIGHBOUR SELECTION
+%       PROXIMITY NEIGHBOR SELECTION
 %###############################################################################
@@
 \subsubsection{DHT-PNS}
-The work in \cite{DLTZZ2006} propose a proximity neighbour selection
+The work in \cite{DLTZZ2006} propose a proximity neighbor selection
 scheme on top of the Chord DHT. The main goal is to use proximity information
-to group physically close by nodes as neighbours in the DHT table. In order to
+to group physically close by nodes as neighbors in the DHT table. In order to
 detect proximity, virtual network coordinates of peers are used, by using the
@@
 annuluses are divided into $2\chi-1$ \emph{sectors}, where $\chi$ denotes the
-sequence number of the annulus starting from $\chi = 1$ for the centre cycle. It
+sequence number of the annulus starting from $\chi = 1$ for the center cycle. It
 is proved in the paper, that this way each sector occupies the same area as does
 the center cycle. Assuming uniform node distribution, this characteristic,
-favours a more load balanced clustering operation. Every sector in this $2d$
+favors a more load balanced clustering operation. Every sector in this $2d$
 coordinate space is mapped to a unique \emph{region} in the DHT space forming a
@@
 call, registers it to its corresponding region and publishes its information,
-through the special peer CN. The node can, additionaly, ask CN for other
+through the special peer CN. The node can, additionally, ask CN for other
 nodes that have previously joined the region in order to add them into its
-neighbour set for future routing table optimization. Even in the case when no
-neighbour is detected in the current region, the search is expanded to adjasent
+neighbor set for future routing table optimization. Even in the case when no
+neighbor is detected in the current region, the search is expanded to adjacent
 regions and towards an upper layer identifier space until one is found or the
@@
 to address the topology mismatch problem, at the construction phase of the
-Chord identifier cycle, on the the basis of underlying topology information.
-The proposed model constructs the \emph{Quasi-Chord} ovelay network in three
+Chord identifier cycle, on the basis of underlying topology information.
+The proposed model constructs the \emph{Quasi-Chord} overlay network in three
 steps. First, each host acquires its coordinates in the geometric space
@@
 
-%In the following paragraphs, these steps are furtherly discussed.
+%In the following paragraphs, these steps are further discussed.
 %
 %\paragraph{GNP coordinates}
-%First of all the host must be positioned in the geometric space. The algoritm
+%First of all the host must be positioned in the geometric space. The algorithm
 %  models the P2P network with a well defined distance function in such a way,
@@
 %\paragraph{Cantor SPF}
-%After the $2$-d coordinates are set, in the next stage of the algorithn, each
+%After the $2$-d coordinates are set, in the next stage of the algorithm, each
 %  participating peer is assigned a Cantor value, according to the application
@@
 %\paragraph{Quasi-Chord construction}
-%As can intuitively be infered by observing the Cantor chart, close-by nodes in
+%As can intuitively be inferred by observing the Cantor chart, close-by nodes in
 %  the physical layer, are more likely to have similar Cantor values. This
@@
 % TODO: SOME DISCUSSION
-%       NEEDS DOUBLECHECK
+%       NEEDS DOUBLE-CHECK
 %
@@
 % maintained is an additional negative point to the efficiency and the
-% scallability of the algorithm.
+% scalability of the algorithm.
 %
@@
 % The overhead is more compared to vanilla chord since mapping of 2-dimensional
-% space to 1-demensional with Cantor SPF adds some extra burden in time
+% space to 1-dimensional with Cantor SPF adds some extra burden in time
 % complexity. Also the addition of the counter-clockwise finger table doubles
@@
 \subsubsection{IP-based Clustering (IPBC)}
-Proximity neighbour selection algorithms use probing and other techniques to
+Proximity neighbor selection algorithms use probing and other techniques to
 detect proximity. However, such methods are either not precise enough, or
 overload the network. \emph{IP-based clustering} \cite{KM2007} is a proximity
-neighbour selection based algorithm in which the authors propose to use the IP
+neighbor selection based algorithm in which the authors propose to use the IP
 address prefixes (16 bit for IPv4) in order to detect proximity.
@@
 creates less precise results and more number of bits increase the burden on the
-network and reduce the possible number of neighbours. Therefore, a careful
-selection is required in terms of performance/accuracy tradeoff. IP-based
+network and reduce the possible number of neighbors. Therefore, a careful
+selection is required in terms of performance/accuracy trade-off. IP-based
 Clustering generates a key and its prefix and stores the later in the DHT
 itself, so that any newly joining nodes with the same IP prefix can query it and
-identify all the neighbours with the same one relatively easy. Nodes
+identify all the neighbors with the same prefix relatively easy. Nodes
 periodically update their entries in the DHT by removing those who leave, or
@@
 % consecutive octet prefixes are assigned to the same Regional Internet
-% Registries (RIRs), nodes inside will be physicaly close to each other.
+% Registries (RIRs), nodes inside will be physically close to each other.
 
@@
 % prefix of its IP. Authors argue on the prefix length to be used, as there is a
-% thin balance between reducing the propability of finding closer nodes by
+% thin balance between reducing the probability of finding closer nodes by
 % adopting a long prefix and overload nodes that are responsible for information
-% published by many peers when chosing to hash a short one. Their verdict is for
+% published by many peers when choosing to hash a short one. Their verdict is
+for
 % the use of a 16-bit wide prefix for real world systems deployed in an Internet
@@
 % peer lifetime for topology adaptation) can acquire information about close-by
-% nodes just by quering thde DHT for a specific key. Moreover, the algorithm
+% nodes just by querying the DHT for a specific key. Moreover, the algorithm
 % takes care of the freshness of the proximity information in two ways. First,
@@
 \subsubsection{Cone}
-\emph{Cone} \cite{HY2007} extends Chord using proximity neighbour
+\emph{Cone} \cite{HY2007} extends Chord using proximity neighbor
 selection topology optimization algorithm. The proximity information is
@@
 tables, display physical network proximity with the current node. Moreover, a
-third table called \emph{group} maintains information about other online peers
+third table called \emph{group} maintains information about other on-line peers
 within the current node's group in a way that entries are now close in the ID
@@
 %
-%Routing in Cone, comprises of the inter-group algoritm and the intra-group
+%Routing in Cone, comprises of the inter-group algorithm and the intra-group
 %  algorithm. First the group of nodes to which the desired key lies is
@@
 considers physical proximity among peers but their mobility attributes as well.
-DynaMO is based on a Pastry overlay network and this was a preety concious
+DynaMO is based on a Pastry overlay network and this was a pretty conscious
 choice, by the authors, basically because of Pastry's built-in locality
-heuristics which are thoroughly analysed in the in the literature
-\cite{CDCR2002a} providing good backround against which to test and compare
-their results. In order to adapt to mobile, ad-hoc networks, though, special
-care has been put to the maintenance of an even overlay ID distribution so that
-hot-spots are avoided. DynaMO dictates a newcoming node to gather information
-concerning its physical neighbourhood and uses it to assign itself an
-appropriate overlay ID. DynaMO tries to capitalize on the fact that in Pastry
-each node's table consists of rows equal to the number of digits of the overlay
-IDs and columns equal to the ID's base. As we go down the table rows, the
-matching prefix between the current node's ID and the row's entries increases by
-one. Thus, the leaf set contains the closest nodes in the ID space. Additionally
-as the prefix match increases by one, the result is exponentially less
-candidates that can fill the tables entries as the row increases. This leads to
-the observation \cite{antony_pastry_2001,CDCR2002a} that from overlay routing
-hop to overlay routing hop the physical distance between nodes is likely to
-increase leading to a dominating last routing step in terms of the overall
-plysical routing path distance travelled during a key lookup. Since the last
-routing step is usually taken from the leaf set, DynaMO focuses on making this
-last step as physicaly close as possible. In this context, two approaches are
-considered namely \emph{Random Landmarking (RLM)} and \emph{Closest Neighbour
-Prefix Assignment (CNPA)}.
+heuristics which are thoroughly analyzed in the literature \cite{CDCR2002a}
+provide a good background against which to compare their own results. In
+order to adapt to mobile, ad-hoc networks, though, special care has been put to
+the maintenance of an even overlay ID distribution so that hot-spots are
+avoided. DynaMO dictates a new coming node to gather information concerning its
+physical neighborhood and uses it to assign itself an appropriate overlay ID.
+DynaMO tries to capitalize on the fact that in Pastry each node's table consists
+of rows equal to the number of digits of the overlay IDs and columns equal to
+the ID's base. As we go down the table rows, the matching prefix between the
+current node's ID and the row's entries increases by one. Thus, the leaf set
+contains the closest nodes in the ID space. Additionally, as the prefix match
+increases by one, the result is exponentially less candidates that can fill the
+tables entries as the row increases. This leads to the observation
+\cite{antony_pastry_2001,CDCR2002a} that from overlay routing hop to overlay
+routing hop the physical distance between nodes is likely to increase leading to
+a dominating last routing step in terms of the overall physical routing path
+distance traveled during a key lookup. Since the last routing step is usually
+taken from the leaf set, DynaMO focuses on making this last step as physically
+close as possible. In this context, two approaches are considered namely
+\emph{Random Landmarking (RLM)} and \emph{Closest Neighbor Prefix Assignment
+(CNPA)}.
 
@@
 keys. This means that when a node is assigned a landmark key then it plays the
-role of a temporary landmark for the network. Any joing node will measure its
-distance to that and all other landmarks and assign to itself an ID consisting
+role of a temporary landmark for the network. Any joining node will measure its
+distance to this landmark node and assign to itself an ID consisting
 of a prefix taken from the closest landmark and a remainder that could be
-assigned randomnly or generated by mechanism that takes into account physical
-neighbourhood. The legth of the prefix can be determined as
+assigned randomly or generated by mechanism that takes into account physical
+neighborhood. The length of the prefix can be determined as
 $prefix_length=|log_b k|$, where $b$ is the ID base and $k$ the number of
@@
 and thus bringing a node's leaf set closer to itself. Additionally, the
-advantage of dynamic landmark nodes is that the failed ones can instantly be
-replaced by new responsibles, that share similar physical attributes to the
-failed and thus preserving the balance of the network.
+advantage of dynamic landmarking is that in the case of failed landmark servers
+they can instantly be replaced by new substitutes, that share similar physical
+attributes to the failed and thus preserving the balance of the network.
 
@@
 is not acceptable. CPNA on the other hand takes advantage of Pastry's
-specification that a new coming node is bootstraped by a physicaly close node.
-The new comer then assumes the ID prefix of that neighbour while the rest is
+specification that a new coming node is boot-strapped by a physically close
+node.
+The new comer then assumes the ID prefix of that neighbor while the rest is
 generated similarly to RLM. Unfortunately, less overhead comes at the expense
@@
 % initial formation of an overlay network} , by introducing the \emph{landmark
-% gravitation range} as a theshold over which landmark keys are reassinged (for
+% gravitation range} as a threshold over which landmark keys are reassigned (for
 % the RLM approach) or unutilized ID prefix ranges are detected and used (for
@@
 low &
-% Dynamic landmarkin approach helps the scalability of the algotithm. The
+% Dynamic landmarking approach helps the scalability of the algorithm. The
 % authors claim that the temporary nature of RLM's landmark nodes does not
@@
 
-\emph{SAT-Match} \cite{RGJZ2004}. Even though no theoretical proof is
-given, valid results are obtained and demonstrated through experimentations.
-Similar to the other approaches in the proximity neighbour selection,
-\textit{SAT-Match} uses probing to detect close by peers and obtain proximity
-information. However, as an additional feature, SAT-Match peers can do
-selective jumps to adjust their positioning in the DHT, if it reduces the
-\emph{stretch} of its one-hop neighbourhood. The paper defines stretch
-$S = \frac{\bar{L_l}}{\bar{L_p}}$, where $\bar{L_l}$ is the average
-logical link latency while the $\bar{L_p}$ is the average physical
-link latency, and uses it a way quantify the \emph{topology
-match degree} of the constructed overlay. For the probing phase SAT-Match uses a
-small TTL value for the probing messages in order to reduce
-redundancy \cite{jiang_lightflood_2008}. This process begins as soon as the node
-joins the network using a DHT mechanism. Each probing message contains
-information about the source and a small TTL value. The recipient of such a
-message returns information about itself to the source and forwards the probing
-message to its neighbours if the TTL is non-zero. The discovered nodes are
-referred to as $TTL-k$ neighbourhood of the source node based on the $TTL$
+\emph{SAT-Match} \cite{RGJZ2004}, similarly to the other approaches in proximity
+neighbor selection, uses probing to detect close by peers and obtain proximity
+information. However, as an additional feature, SAT-Match peers can do selective
+jumps to adjust their positioning in the DHT, if it reduces the \emph{stretch}
+of its one-hop neighborhood. The paper defines stretch 
+$S = \frac{\bar{L_l}}{\bar{L_p}}$, where $\bar{L_l}$ is the average logical link
+latency while the $\bar{L_p}$ is the average physical link latency, and uses it
+a way quantify the \emph{topology match degree} of the constructed overlay. For
+the probing phase SAT-Match uses a small TTL value for the probing messages in
+order to reduce redundancy \cite{jiang_lightflood_2008}. This process begins as
+soon as the node joins the network using a DHT mechanism. Each probing message
+contains information about the source and a small TTL value. The recipient of
+such a message returns information about itself to the source and forwards the
+probing message to its neighbors if the TTL is non-zero. The discovered nodes
+are referred to as $TTL-k$ neighborhood of the source node based on the $TTL$
 distance to the source node. Blindly selecting the peer with the smallest RTT
-as neighbour is, generally, not the optimal decision to make in order to achieve
+as neighbor is, generally, not the optimal decision to make in order to achieve
 global stretch reduction. In a structured scheme, when a node jumps to connect
@@
 select one zone to jump in this phase. The algorithm is as follows: The source
-node $S$ calculates the stretch change of its $TTL-1$ neighbourhood and that of
-the $TTL-1$ neighbourhood of the first of the previously selected peers. These
+node $S$ calculates the stretch change of its $TTL-1$ neighborhood and that of
+the $TTL-1$ neighborhood of the first of the previously selected peers. These
 calculations are made as if the jump has been made. If the stretch reduction is
@@
 is performed as a combination of \emph{leave} and \emph{join} operations, in the
-CAN context.
+CAN context. Even though no theoretical proof is given, valid results are
+obtained and demonstrated through experimentations.
 % Figure~\ref{figure:sat_match} demonstrates the selective jump on an
@@
 %\paragraph{}
-%Moreover the algorithm takes several issues into consiteration in order to
-%  furtherly improve the resulted overlay. For example when multiple nodes try
+%Moreover the algorithm takes several issues into consideration in order to
+%  further improve the resulted overlay. For example when multiple nodes try
 % to jump simultaneously into a region, then the logical link brakes from one
-% attempt may result in inacurate computation of the gain factor, for an other.
+% attempt may result in inaccurate computation of the gain factor, for an other.
 % This situation is identified as \emph{contention} and the nodes use an
-% exponatial back-off algorithm to avoid it. An other problem is the uneseccary
-% traffic incuerred by the probing phase in a region that after several jumps
+% exponential back-off algorithm to avoid it. An other problem is the
+%unnecessary
+% traffic incurred by the probing phase in a region that after several jumps
 % has settled to a stable state. In these cases it is more likely for jump
@@
 % SAT-Match seems to outperform landmark binning in the sense of providing a
-% higher grained differenciation of proximity, adding to the fact that
+% higher grained differentiation of proximity, adding to the fact that
 % SAT-match is fully distributed. Also to further enhance the efficiency it can
@@
 % Also the fact that the selective jumps are not always performed the added
-% overhead to efficiency gained is not guarranteed in any configuration and
+% overhead to efficiency gained is not guaranteed in any configuration and
 % circumstance.
@@
 Nodes join the network by using a key value, which determines the location and
-the neighbourhood of the new node within the network. However, assigning
+the neighborhood of the new node within the network. However, assigning
 random key values to the newly inserted nodes creates non-optimal matching with
@@
 the node to the network, so that nodes within the same network domains are
-selected as peers, or neighbours, during the overlay topology construction.
+selected as peers, or neighbors, during the overlay topology construction.
 
@@
 can be found in the literature for addressing the inefficient network resource
-utilisation in structured architectures imposed by the topology mismatch
-problem. A categorisation of the methods used for achieving their goals is the
+utilization in structured architectures imposed by the topology mismatch
+problem. A categorization of the methods used for achieving their goals is the
 following based on previous work from
@@
   \item proximity-routing-based,
-  \item proximity neighbour selection based,
+  \item proximity neighbor selection based,
 \end{inparaenum}
@@
 discussion on unstructured systems in Section~\ref{section:unstructured} so
-we will not elaborate more here. Again, we additionaly gather highlights and a
+we will not elaborate more here. Again, we additionally gather highlights and a
 rough estimation of each algorithm's positive and negative aspects regarding
@@
 methods do not always give reliable estimates for the node positions over the
-internet. The landmark servers are not self-organizing and have maintenance
+Internet. The landmark servers are not self-organizing and have maintenance
 overheads. To serve a P2P network with millions of peers, multiple landmark
@@
 nodes improves the query efficiency of the system, on the other hand, it tends
-to create hotspots, and the needed failure resilience is undermined by the fact
+to create hot-spots, and the needed failure resilience is undermined by the fact
 that close by nodes are more likely to suffer collective failures.
@@
 %other DHTs, such as Chord or Pastry, another problem emerges.  To gain fault
-%tolerant properties, these protocols, replicate key-value pairs on neighbouring
+%tolerant properties, these protocols, replicate key-value pairs on neighboring
 %(in the ID space) nodes. When a proximity-based node ID assignment has been
@@
 maps to suboptimal longer paths on the physical topology. Thus, in general,
-proximity neighbour selection is considered as a superior method to proximity
+proximity neighbor selection is considered as a superior method to proximity
 routing, however, joint uses of these two protocols are also possible. Moreover,
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Proximity neighbour selection}
+\subsubsection{Proximity neighbor selection}
 
@@
 \centering
-  \includegraphics[scale=0.4]{img/pdf/proximity-neighbour-selection.pdf}
-\caption{Proximity neighbour selection dictates that proximity is taken into
-acount during routing table maintenance.}
+  \includegraphics[scale=0.4]{img/pdf/proximity-neighbor-selection.pdf}
+\caption{Proximity neighbor selection dictates that proximity is taken into
+account during routing table maintenance.}
 \label{figure:proximity-neighbour-selection}
@@
 
-\emph{Proximity Neighbour Selection} constructs the routing tables using
+\emph{Proximity Neighbor Selection} constructs the routing tables using
 proximity knowledge (see Figure~\ref{figure:proximity-neighbour-selection}). The
@@
 Pastry is based on longest node ID prefix matching, while CAN uses RTT values to
-detect close by nodes. \cite{CDCR2002a} reports that proximity neighbour
+detect close by nodes. \cite{CDCR2002a} reports that proximity neighbor
 selection as an effective proximity based method.

### 2014-01-27 unstructured.tex
2012.10.24 (Vassilis) Changed paper title. Text reworked (content clarifications added, typos corrected, broken references fixed).

@@
 This section is dedicated to the unstructured P2P architectures. In the next
-paragraphs,algorithms for tackling the topology mismatch problem on such
-networks are discussed, analysed and categorised based on the use of the
+paragraphs, algorithms for tackling the topology mismatch problem on such
+networks are discussed, analyzed and categorized based on the use of the
 overlay structure, message forwarding approach for peer communication and
-techniques proposed for detecting proximity in order to optimise overlay
+techniques proposed for detecting proximity in order to optimize overlay
 topology.
@@
 
-% BROADCAST OPTIMISATION
+% BROADCAST OPTIMIZATION
 
@@
 tries to avoid this delay by forwarding the query messages only to a selected
-set of neighbours, in which the selection criteria varies from the number of
+set of neighbors, in which the selection criteria varies from the number of
 results received previously, distance in terms of hops, bandwidth, or the query
-load of the neighbour.
+load of the neighbor.
 
@@
 low &
-% the process in not started from the begining at each iteration step
+% the process in not started from the beginning at each iteration step
 low &
@@
 medium &
-% BFS is Gnutella's non efficient communcation method
+% BFS is Gnutella's non efficient communication method
 medium \\
@@
 nodes whose delay is less than or equal that boundary. With the clustered
-routing tables and the loose organisation of DAPS' overlay network it is
+routing tables and the loose organization of DAPS' overlay network it is
 considered by is between structured and unstructured.
@@
 Gia uses a token-based flow control algorithm in which each node announces the
-number of query requests it can handle in terms of tokens to its neighbours, so
-that neighbours only forward query requests to nodes that they previously
-received tokens from. Gia also acknowledges the heterogenity in peer bandwidth,
+number of query requests it can handle in terms of tokens to its neighbors, so
+that neighbors only forward query requests to nodes that they previously
+received tokens from. Gia also acknowledges the heterogeneity in peer bandwidth,
 processing power, disk speed, etc, of the nodes in P2P networks and uses this
@@
 %  \item Every node maintains pointers to to the content that is offered by
-% their immediate neighbours, creating a \emph{one-hop replication} of pointers,
+% their immediate neighbors, creating a \emph{one-hop replication} of pointers,
 %scheme.
@@
 no real targeting, many times to a direction far from the path that can
-ultimatelly fullfil the query request. Another problem with topology unaware
+ultimately fulfill the query request. Another problem with topology unaware
 systems is the duplication of messages due to cycles, even along the correct
-forwarding path. The authors of \cite{ZKB2008} fucused on that exact deficiency
+forwarding path. The authors of \cite{ZKB2008} focused on that exact deficiency
 of overlay networks and introduced Distributed Cycle Minimization Protocol
@@
 overlay connectivity, resilience and other important properties of unstructured
-architectures by avoiding a hierarchical organisation of peers. In DCMP, once a
+architectures by avoiding a hierarchical organization of peers. In DCMP, once a
 cycle is detected, the most powerful node in that cycle is elected as the
 \emph{Gate Peer} and the cycle is then broken at a strategic place so that it
-will result in the minimisation of the distance between the Gate Peer and all
+will result in the minimization of the distance between the Gate Peer and all
 other nodes that participated in that cycle. The process is managed by using two
@@
 %\begin{inparaenum}[\itshape i\upshape)]
-%  \item a \emph{GUID}\footnote{Globally Unique IDentifier assigned to every
+%  \item a \emph{GUID}\footnote{Globally Unique Identifier assigned to every
 % query message generated by any node.} field same as the one of the duplicate
@@
 %not be detected, because the ICM messages will be discarded before they locate
-%it. There is a tradeoff between preserving the connectivity of the network and
+%it. There is a trade-off between preserving the connectivity of the network and
 %minimizing the duplicates that makes such a possibility to be safely ignored.
@@
 % Duplicates are an important fraction of the redundant overhead imposed by
-% topology mismatch and thus DCMP minimises this as possible.
+% topology mismatch and thus DCMP minimizes this as possible.
 % The control overhead is also one to two orders of magnitude less than LTM due
@@
 each item increases in the network, it would be easier for a search algorithm,
-even a random one, to locate these items. In order to analyse the feasibility of
+even a random one, to locate these items. In order to analyze the feasibility of
 such an approach, the authors investigate three different replication
@@
 rate, so that frequently queried items are replicated more. The square root
-allocation is a strategy proposed in this paper, which is a model between the
-uniform and proportional allocation.  In order to evaluate the outcome of the
-replication models, the overall costs of successful and unsuccessful searches in
-the network are compared. Results argue that the uniform allocation model
-minimizes the maximum search size, therefore reduces the time spent on an
+allocation is a strategy proposed by Cohen and Shenker, which is a model between
+the uniform and proportional allocation.  In order to evaluate the outcome of
+the replication models, the overall costs of successful and unsuccessful
+searches in the network are compared. Results argue that the uniform allocation
+model minimizes the maximum search size, therefore reduces the time spent on an
 unsuccessful search. The proportional model, on the other hand, promotes the
@@
 % e.g. path replication is easy to scale, just detect the path along which the
-% query ultimatelly got fulfilled
+% query ultimately got fulfilled
 medium \\
@@
 
-%       OVERLAY OPTIMISATION
+%       OVERLAY OPTIMIZATION
 
@@
 % }\qquad\qquad
-% \subfigure[Peer A sends a message to the rest using end system multicast.] {
+% \subfigure[Peer A sends a message to the rest using end system multi-cast.] {
 %   \includegraphics[scale=0.4]{img/algorithms/narada3}
@@
 % }
-% \caption{A visualisation of the Narada protocol.}
+% \caption{A visualization of the Narada protocol.}
 % \label{figure:narada}
@@
 protocol for creating self-adapting overlay networks, that can achieve
-application-layer multicast communication without requiring IP multicast
+application-layer multi-cast communication without requiring IP multi-cast
 infrastructure at the network layer\footnote{
-  IP multicast is the term reffering to the method of sending IP datagrams to a
+  IP multi-cast is the term referring to the method of sending IP datagrams to a
   group of receivers in a single transmission. Even though the method is
-  available some years now, IP multicast has not taken off as anticipated since
+  available some years now, IP multi-cast has not taken off as anticipated since
   many believe that it violates the stateless design of the current Internet.
@@
 latency properties of the physical underlying infrastructure. Additionally,
-authors realised the inefficiency caused by the topology mismatch problem and
+authors realized the inefficiency caused by the topology mismatch problem and
 tried to encounter it by building a richer connected graph, called a
@@
 of the first attempts, along with Narada, to address the topology mismatch
-problem. AOTO is a distributed algorithm that works in two steps, namelly
+problem. AOTO is a distributed algorithm that works in two steps, namely
 \emph{Selective Flooding} and \emph{Active Topology} to optimize the usage of
-the underlying physical resourses. In \emph{Selective Flooding} a minimum
-spanning tree is built among each peer and its immediate neighbours so that
+the underlying physical resources. In \emph{Selective Flooding} a minimum
+spanning tree is built among each peer and its immediate neighbors so that
 queries are not flooding the whole network and at the same time without
@@
 possible. This is done independently, by each peer by replacing non-flooding
-neighbours, with closer nodes. Picking a replacement out of these non-fooding
-neighbours follows a random policy (called \emph{Randomized AT} algorithm in the
+neighbors, with closer nodes. Picking a replacement out of these non-flooding
+neighbors follows a random policy (called \emph{Randomized AT} algorithm in the
 paper's context). For these steps a peer needs to keep track of communication
-costs to all its neighbours (e.g. the network delay) as well as between any
-pair of neighbours (meaning also that additional, cost probing, message types
-need to be added to the Gnutella protocol). Whenever a new neighbour cost table
-is received or there is a change of neighbours, the source peer has to
-re-calculate the multicast tree and apply the randomized AT algorithm.
+costs to all its neighbors (e.g. the network delay) as well as between any
+pair of neighbors (meaning also that additional, cost probing, message types
+need to be added to the Gnutella protocol). Whenever a new neighbor cost table
+is received or there is a change of neighbors, the source peer has to
+re-calculate the multi-cast tree and apply the randomized AT algorithm.
 
@@
 % or overlay topologies. On the other hand, SF is more effective with large
-%number of logical neighbours. It can reach an average optimization rate of 87.4
-%percent on a logical topology with an average of 30 logical neighbours.
+%number of logical neighbors. It can reach an average optimization rate of 87.4
+%percent on a logical topology with an average of 30 logical neighbors.
 %
@@
 %
-%Different numbers of average logical neighbours has little to do with the
+%Different numbers of average logical neighbors has little to do with the
 % effectiveness of AT. If the source has $n$ non-flooding peers, there are $n$
-%potential neighbour replacements. The overhead to exhaust all $n$ possible
+%potential neighbor replacements. The overhead to exhaust all $n$ possible
 %replacements can be high, so in practice, after each replacement the source
@@
 medium &
-% tracking of communication costs to all peer's neighbours as well as between
-% any pair of neighbours. Whenever a new neighbour cost table is received or
-% there is a change of neighbours, the source peer has to re-calculate the
-% multicast tree and apply the randomized AT algorithm.
+% tracking of communication costs to all peer's neighbors as well as between
+% any pair of neighbors. Whenever a new neighbor cost table is received or
+% there is a change of neighbors, the source peer has to re-calculate the
+% multi-cast tree and apply the randomized AT algorithm.
 high &
@@
 \emph{Adaptive Connection Establishment (ACE)} \cite{LZXN2004} builds an
-overlay multicast tree among each source node and the peers within a certain
-diameter from the source peer and optimizes the neighbour connections that are
-not in that tree. To achieve that it calculates cost between nodes using network
-delay as a metric. Each peer probes the costs with its immediate logical
-neighbours and forms a \emph{neighbour cost table (NCT)} using a special
-routing message type. Two neighbouring peers exchange their NCTs in order for
-every peer to obtain the cost between any pair of its local neighbours forming a
+overlay multi-cast tree among each source node and the peers within a certain
+diameter from the source peer and optimizes the neighbor connections that are
+not in that tree. To achieve that, it calculates cost between nodes using
+network delay as a metric. Each peer probes the costs with its immediate logical
+neighbors and forms a \emph{neighbor cost table (NCT)} using a special
+routing message type. Two neighboring peers exchange their NCTs in order for
+every peer to obtain the cost between any pair of its local neighbors forming a
 small overlay topology. Moreover, based on obtained NCTs a minimum spanning tree
-among each peer and its immediate neighbours is built. Ultimately, physically
-far away neighbours are replaced by physically close ones. Specifically, a peer
-$S$ probes the distance between one of its non-flooding neighbour's neighbour,
-say $G$ and $H$ respectively. Assuming that the link to the neighbour's neigbour
-($SH$) is smaller that that of the neighbour ($SG$), the later is cut and a new
-connection is established. In the oposite situation where $SG$ is smaller than
-$SH$ then if $SH$ is smaller that $GH$ then $H$ is preserved as $S$'s new
-neighbour. In the case where $SH$ is larger than $SG$ and $GH$, no connection
-will be established and $S$ will continue probing another neighbour's neighbour.
-The above optimisation is conducted within $1$-neighbour closure (among its
-source peer and all its direct neighbours) but the scope can be extended. The
-larger the scope, the better topology matching improvement but also the greater
-the computational overhead.
+among each peer and its immediate neighbors is built. Ultimately, physically
+far away neighbors are replaced by physically close ones. Specifically, a peer
+$S$ probes the distance between one of its non-flooding neighbor's neighbor,
+say $G$ and $H$ respectively. Assuming that the neighbor's neighbor ($SH$) is
+measured as being ``closer'' that the neighbor ($SG$), the link to the later
+is cut and a new connection is established. In the opposite situation where
+distance $SG$ is closer than $SH$ then if $SH$ is shorter that $GH$ then $H$ is
+preserved as $S$'s new neighbor. In the case where $SH$ is larger than $SG$ and
+$GH$, no connection will be established and $S$ will continue probing another
+neighbor's neighbor. The above optimization is conducted within $1$-neighbor
+closure (among its source peer and all its direct neighbors) but the scope can
+be extended. The larger the scope, the better topology matching improvement but
+also the greater the computational overhead.
 
@@
 %always a reliable estimation method to detect the physical location of peers,
-%the algorithm still suffers from the discrepancies caused by mislocated nodes.
+%the algorithm still suffers from the discrepancies caused by mis-located nodes.
 
@@
 \textit{TTL-detector}s with TTL values of 0 and 1 so that the receiving peers
-discover one- and two-hop neighbour sets ($N$ and $N^2$ respectively) around
+discover one- and two-hop neighbor sets ($N$ and $N^2$ respectively) around
 them and calculating communication costs. Calculation is achieved through
-timestamp checking, so clocks must be synchronised. The latency information
+time-stamp checking, so clocks must be synchronized. The latency information
 gathered is later used to evolve the overlay network into a more efficient one,
 without reducing the search scope. Each node examines the latency information
-among his direct neighbours and those with larger latency are put in to a
+among his direct neighbors and those with larger latency are put in to a
 will-cut list where they remain for a certain period of time until when they are
-ultimatelly cut and recorded to the peer's cut list. Thus, low productive
+ultimately cut and recorded to the peer's cut list. Thus, low productive
 connections are dropped and replaced by more efficient ones, reducing the
@@
 %\end{inparaenum}
-%If the link with the largest cost is found and is a direct neighbour then the
+%If the link with the largest cost is found and is a direct neighbor then the
 % connection is put in a will-cut list and stays there for a certain period of
@@
 %\paragraph{}
-%Supposing $n$ is the number of peers, $c_n$ is the average number of neighbours
+%Supposing $n$ is the number of peers, $c_n$ is the average number of neighbors
 % and $c_e$ is the average cost of logical links, then in the flooding-based
@@
 two groups, namely red and white and assigning different tasks to different
-groups. When a peer joins the network, is randomly assigned with an initial
-colour, say red or white (separating all peers into two groups). Then the peer
-that plays the role of the bbootstrap host provides the joining peer with a list
-of active peers along with their colour information. The later uses this list to
-establish connections to differently coloured peers. This way, all peers form a
-bipartite overlay. The white peers measure distances to neighbours (red only) by
+groups. When a peer joins the network, it is randomly assigned with an initial
+color, say red or white (separating all peers into two groups). Then the peer
+that plays the role of the boot-strap host provides the joining peer with a list
+of active peers along with their color information. The later uses this list to
+establish connections to differently colored peers. This way, all peers form a
+bipartite overlay. The white peers measure distances to neighbors (red only) by
 using the network delay as a metric and report their corresponding red
-neighbours. The red peers, equipped with the information of all two-hop
-neighbours ($N^2$), creates minimum spanning tree of these neighbours and
+neighbors. The red peers, equipped with the information of all two-hop
+neighbors ($N^2$), creates minimum spanning tree of these neighbors and
 assigns efficient forwarding paths. Having a minimum spanning tree with two
 hops, a red peer is able to send its queries within that range. Some white
-peers, though, have become non-forwarding neighbours. In this phase, such a
-(white) neighbour will try to find another red peer being two hops away from its
-current red neighbour to replace the later as its new neighbour. The white peers
+peers, though, have become non-forwarding neighbors. In this phase, such a
+(white) neighbor will try to find another red peer being two hops away from its
+current red neighbor to replace the later as its new neighbor. The white peers
 can further optimize their positions within the overlay if this is required.
@@
 %
-%In a static enviromnent LTM may reduce traffic cost by around 80 to 85 percent
+%In a static environment LTM may reduce traffic cost by around 80 to 85 percent
 % while SBO reduces traffic cost between 85 and 90 percent. However, LTM  is
@@
 high &
-% compared to LTM it has slower convergence speed thus incurres more, total,
+% compared to LTM it has slower convergence speed thus incurs more, total,
 % overhead.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Two-Hop-Away Neighbour Comparison and Selection (THANCS)}
+\subsubsection{Two-Hop-Away Neighbor Comparison and Selection (THANCS)}
 Work in \cite{LNXE2005,L2008} proposes a distributed heuristic
-called \emph{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} that
-tries to minimise the overlay hop costs. THANCS is considered a \emph{local
+called \emph{Two-Hop-Away Neighbor Comparison and Selection (THANCS)} that
+tries to minimize the overlay hop costs. THANCS is considered a \emph{local
 search method}, in the sense that it targets in finding a locally optimum
 solution, by exploiting knowledge within a 2-hop radius. The algorithm consists
-of two main components: \emph{piggybacking neighbour distance on queries} and
-\emph{neighbour comparison and selection} which are furtherly discussed bellow.
+of two main components: \emph{piggybacking neighbor distance on queries} and
+\emph{neighbor comparison and selection} which are further discussed bellow.
 The first component dictates peers to probe distances (using network delay
-distance measuring) with its immediate naighbours and stores information
+distance measuring) with its immediate neighbors and stores information
 locally. This is done by introducing a special query message type, the
-\emph{Piggy Message (PM)} which includes information about the neigbour
+\emph{Piggy Message (PM)} which includes information about the neighbor
 identification and measured distance. A peer $p$ constructs a PM for its
-neighbour $n$, which contains $n$'s IP address and the distance between them.
+neighbor $n$, which contains $n$'s IP address and the distance between them.
 When $p$ receives a query from $n$, this PM will be piggybacked to the query
-message that will be then forwarded to all other neighbours of peer $p$. Upon
-receiving such a query message, each of the other neighbours will detach the PM
+message that will be then forwarded to all other neighbors of peer $p$. Upon
+receiving such a query message, each of the other neighbors will detach the PM
 information (it will not be further forwarded), record the $pq$ distance and
-process the query as always. The paper proposes selection of which incoming
-queries should piggyback a PM by either \emph{pure propability-based (PPB)} or
-\emph{new neighbour triggered (NNT)} policies. The second component, namely
-neighbour comparison and selection, dictates peer $p$ to probe the distance to
-all his known, not yet probed, two-hop neighbours ($ N^2(p)$). The distance of
+process the query as always. Selection of which incoming queries should
+piggyback a PM is proposed by either \emph{pure probability-based (PPB)} or
+\emph{new neighbor triggered (NNT)} policies. The second component, namely
+neighbor comparison and selection, dictates peer $p$ to probe the distance to
+all his known, not yet probed, two-hop neighbors ($ N^2(p)$). The distance of
 $pn$ is known to $p$. Upon receiving a PM from node $n$ with the distance of
-$nq$, where $q$ is a direct neigbour of $n$, $p$ does follows one of the
-following two paths. First, if $q$ is a direct neighbour of $p$, then the later
-will check the cost of the involved links. If the most cost-intensive link is
-one of $pq$ or $pn$, then it is put into a \emph{will-cut} list. If it is $nq$
-then nothing is done (since either $n$ or $q$ will have the chance to handle it
-themselves). On the other hand, if $q$ is within an two-hop radius from $p$,
-the later probes the former (if it hasn't got the distance information yet) and
-again compares distances $pq$, $pn$ and $nq$. If $pq$ is the most
-cost-intensive, $p$ will not establish the connection. If it is $pn$, $p$ will
-establish connection $pq$ and put $pn$ in the will-cut list. If it is $nq$, $p$
-will keep the connection with both $n$ and $q$, expecting that $n$ or $q$ will
-eventually disconnect link $nq$, later.
+$nq$, where $q$ is a direct neighbor of $n$, $p$ chooses one of the following
+two paths. First, if $q$ is a direct neighbor of $p$, then the later will check
+the cost of the involved links. If the most cost-intensive link is one of $pq$
+or $pn$, then it is put into a \emph{will-cut} list. If it is $nq$ then nothing
+is done (since either $n$ or $q$ will have the chance to handle it themselves).
+On the other hand, if $q$ is within two-hop radius from $p$, the later probes
+the former (if it hasn't got the distance information yet) and again compares
+distances $pq$, $pn$ and $nq$. If $pq$ is the most cost-intensive, $p$ will not
+establish the connection. If it is $pn$, $p$ will establish connection $pq$ and
+put $pn$ in the will-cut list. If it is $nq$, $p$ will keep the connection with
+both $n$ and $q$, expecting that eventually either $n$ or $q$ will disconnect
+link $nq$, later.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Hops Adaptive Neighbour Discovery (HAND)}
+\subsubsection{Hops Adaptive Neighbor Discovery (HAND)}
 \cite{CLZHC2006} proposes the \emph{HAND} algorithm which uses a
@@
 graph $G$ into the \emph{Logical Communication Network (LCN)} $G^{*}$ which is
-how the optimal overlay is refered in the paper's context. A formal
+how the optimal overlay is referred in the paper's context. A formal
 definition of achieving the above match is only if all peer hop sequences $(v_1,
@@
 sequence, say $v_2-v_1-v_3$. A pair of probing messages are sent from $v_1$ to
-$v_2$ and $v_3$. Suppose delays of $(v_1,v_2)$ and $(v_1,v_3)$ are are denoted
-as $x$ and $z$, respectively. When the probing message arrives to $v_2$ it
-forwards it directly to $v_3$. Similarly, when the probing message arrives to
-$v_3$, it forwards it directly to $v_2$. These last steps are performed in order
-to obtain delays of $(v_2,v_3)$ and $(v_3,v_2)$ physical paths, respectively,
-denoted by $y$. If $y=z-x\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched
-and should be adjusted to $v_1-v_2-v_3$ by deleting edge $(v_1,v_3)$ and adding
-a new $(v_2,v_3)$. If $y=x-z\pm\varepsilon$, sequence $v_2-v_1-v_3$ is
-mismatched and should be adjusted to $v_1-v_3-v_2$ by deleting edge $(v_1,v_2)$
-and adding a new $(v_3,v_2)$, where $\varepsilon$ is a small positive real
-number denoting additional delays caused by possible forwarding and jitter
-delays. The advantages of the HAND algorithm compared to other approaches are
-that:
+$v_2$ and $v_3$. Suppose delays of $(v_1,v_2)$ and $(v_1,v_3)$ are denoted as
+$x$ and $z$, respectively. When the probing message arrives to $v_2$ it forwards
+it directly to $v_3$. Similarly, when the probing message arrives to $v_3$, it
+forwards it directly to $v_2$. These last steps are performed in order to obtain
+delays of $(v_2,v_3)$ and $(v_3,v_2)$ physical paths, respectively, denoted by
+$y$. If $y=z-x\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched and should
+be adjusted to $v_1-v_2-v_3$ by deleting edge $(v_1,v_3)$ and adding a new
+$(v_2,v_3)$. If $y=x-z\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched and
+should be adjusted to $v_1-v_3-v_2$ by deleting edge $(v_1,v_2)$ and adding a
+new $(v_3,v_2)$, where $\varepsilon$ is a small positive real number denoting
+additional delays caused by possible forwarding and jitter delays. The
+advantages of the HAND algorithm compared to other approaches are that:
 \begin{inparaenum}[\itshape i\upshape)]
@@
   \item it is applicable to dynamic peer-to-peer environments, and
-  \item maintains lower query response time.
+  \item maintains lower query response times.
 \end{inparaenum}
@@
 %
-%Measurements conducted for evaluation perposes showed that in a static
+%Measurements conducted for evaluation purposes showed that in a static
 % environment the algorithm can effectively decrease traffic cost by about 77
 %percent and shorten the query response time by about 49 percent in less than
-%two minutes. In a dynamic environment it shows similar behaviour and with the
+%two minutes. In a dynamic environment it shows similar behavior and with the
 %size of the overlay network having little impact on the effectiveness of the
@@
 low &
-% no clock synching needed, fully distributed and with low control overhead the
+% no clock syncing needed, fully distributed and with low control overhead the
 % algorithm can be considered scalable.
@@
 \subsubsection{Innocuous Topology Aware (ITA) Overlay Construction}
-\cite{PRFM2009} intoduces an algorithm for Innocuous Topology Aware
-construction of unstructured P2P networks. The paper's proposition is twoforld.
+\cite{PRFM2009} introduces an algorithm for Innocuous Topology Aware
+construction of unstructured P2P networks. The paper's proposition is twofold.
 Overlay construction and search method. For the overlay construction it employs
@@
 number of all network participants, $\alpha \leq 1 $ a system-wide magic number
-and $x$ an $\alpha$-related latency threashold, the bootstrapping peer randomly
+and $x$ an $\alpha$-related latency threshold, the boot-strapping peer randomly
 selects $\alpha \ast N$ close (latency bellow $x$) and
 $\left( 1 - \alpha \right) \ast N$ distant (latency above $x$) nodes as its
-neighbours. The search method has two steps. First, the quering node initiates a
-flood to its distant neighbourhood with $TTL = 1$. Then, peers that receive a
+neighbors. The search method has two steps. First, the querying node initiates a
+flood to its distant neighborhood with $TTL = 1$. Then, peers that receive a
 query over a long link, start a local flood with $TTL = ttl$, were $ttl$ is a
@@
 dynamic querying mechanisms of later versions of Gnutella.}. The paper reports
-50\% reduction in communication latency ammong peers by cutting off some 20\% of
+50\% reduction in communication latency among peers by cutting off some 20\% of
 the IP-level message traffic generation.
@@
 low &
-% the selection of neighbours is done at join time. No overlay adaptation means
+% the selection of neighbors is done at join time. No overlay adaptation means
 % low scalability in dynamic environments.
@@
 EGOIST \cite{SLLBBR2008} is a set of algorithms to construct and manage overlay
-networks. It utilises a selfish approach in the sense that every participating
-peer continuously updates its neigbours so as to minimise the sum of distances
-to all destinations under shortest-path routing. In EGOIST, a newlly arriving
+networks. It utilizes a selfish approach in the sense that every participating
+peer continuously updates its neighbors so as to minimize the sum of distances
+to all destinations under shortest-path routing. In EGOIST, a newly arriving
 peer, randomly connects to an already participating node through a bootstrap
-server. Being connected, it starts receiving info throught the link-state
-mechanism and thus, after some time, it has a complete picture of the overlay
-graph. Then it estimates the delay to all other nodes in order to determine its
-potential neighbours and ultimatelly connect to the overlay using some
-policy\footnote{For example, minimisation of the average delay to all its
-neighbours.}. It is obvious that there is extensive resource usage for updating
+server. Being connected, it starts receiving information throughout the
+link-state mechanism and thus, after some time, it has a complete picture of the
+overlay graph. Then it estimates the delay to all other nodes in order to
+determine its potential neighbors and ultimately connect to the overlay using
+some policy\footnote{For example, minimization of the average delay to all its
+neighbors.}. It is obvious that there is extensive resource usage for updating
 the wiring of the all nodes in the system. The authors claim that they can
@@
 \hline
-% peer neighbourhood selected to minimise, for example, average delay
+% peer neighborhood selected to minimize, for example, average delay
 high &
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Biased Neighbour Selection (BNS)}
+\subsubsection{Biased Neighbor Selection (BNS)}
 Bindal et al. \cite{BCCMSBZ2006} claim to strengthen BitTorrent
 protocol\cite{c_bittorrent_2003} locality by selecting most of a peer's
-neighbours to be out of the same ISP, while preserving the near optimal download
+neighbors to be out of the same ISP, while preserving the near optimal download
 performance of the protocol itself. BitTorrent systems employ most of the time
-mechanisms that can be proven very aggresive to an ISP's networking and
-accounting, basically, due to their lack of knowlegde of AS boarder limits.
-BitTorrent specification, by default allows for each peer, 35 connections.
-Biased neighbour selection experiments on some number $k$ which denotes a number
-of neighbouring peers which are not on the same ISP than the peer at hand.
-Consequently, the remaining $35 - k$ peers are from the same ISP. These $k$
-nodes are used in order to preserve a more extended, global view of a network,
-but in the same localize the load within the limits of a single ISP. This
-prevents clients from exchanging traffic through an expensive transit link, if
-there are alternative local connections that could offer the same, faster and at
-no additional cost for the ISP. Implementation can either be done through
-modifications on the tracker side, to identify ISP locallity, or through P2P
+mechanisms that can be proven very aggressive to an ISP's networking and
+accounting, basically, due to their lack of knowledge of Autonomous System
+boarder limits. BitTorrent specification, by default allows for each peer, 35
+connections. Biased neighbor selection experiments on some number $k$ which
+denotes a number of neighboring peers which are not on the same ISP than the
+peer at hand. Consequently, the remaining $35 - k$ peers are from the same ISP.
+These $k$ nodes are used in order to preserve a more extended, global view of a
+network, but in the same localize the load within the limits of a single ISP.
+This prevents clients from exchanging traffic through an expensive transit link,
+if there are alternative local connections that could offer the same, faster and
+at no additional cost for the ISP. Implementation can either be done through
+modifications on the tracker side, to identify ISP locality, or through P2P
 traffic shaping devices installed on ISP's edge routers. The paper also proposes
-combination of biased neighbour selection along with bandwidth throttling and
-some caching approach for near optimal results. Unfortunatelly, implementation
-need, in some extend, contribution (exposure of ISP AS mappings) or
-infrastructure changes (installation of P2P traffic shaping devices) on the ISP
-level itself, making the adoption difficult in the general case.
-
-\begin{center}
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
-% inter-ISP cost minimised while retaining the near optimal performance of
+combination of biased neighbor selection along with bandwidth throttling and
+some caching approach for near optimal results. Unfortunately, implementation
+need, in some extend, contribution (exposure of ISP Autonomous System mappings)
+or infrastructure changes (installation of P2P traffic shaping devices) on the
+ISP level itself, making the adoption difficult in the general case.
+
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% inter-ISP cost minimized while retaining the near optimal performance of
 % BitTorrent download.
@@
 BitTorrent traffic so that it significantly reduces cross-ISP traffic and in
-the same time enhances downloading rates by favouring connections within the
+the same time enhances downloading rates by favoring connections within the
 borders of a single autonomous system. Contrary to biased peer selection
-proposed in \cite{BCCMSBZ2006}, this work can lead to improved
-performance with no cooperation between ISPs and their subscribers, no
-additional infrastructure and no network topology information. The selection
-approach is landmark-based and leverages existing CDN infrastructure for peer
-distance estimation. CDNs already use both static (i.e. geographical) and
-dynamic (network measurement systems) information for their replica selection,
-so the authors claim that peers which exhibit similar redirection behaviour, are
-very likely they are close to the replica servers and as a consequence to each
-other. The redirection behaviour is modeled in terms of \emph{ratio map} in the
-article's parlance, a vector of (replica-server,ratio) tuples, where ratio is
-the percentage of times CDN redirects the peer to the specific replica-server.
-The bootstrapping phase consists of the peer performing DNS lookup to CDN names
-in order to build its redirection information. The interval for polling DNSs
-starts from 30 seconds and increases by one minute every time redirection info
-to the CDN is found to remain unchanged after a lookup. On the other hand, the
-interval is halved whenever is found to have been changed. In order not to avoid
-the bootstrapping phase if the ratio map is sufficiently fresh, the protocol
-persists the info after the end of a BitTorrent session.
-
-\begin{center}
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
-% Authors claim minimisation of inter-ISP traffic cost while enhancing the,
+proposed in \cite{BCCMSBZ2006}, this work can lead to improved performance with
+no cooperation between ISPs and their subscribers, no additional infrastructure
+and no network topology information. The selection approach is landmark-based
+and leverages existing CDN infrastructure for peer distance estimation. CDNs
+already use both static (i.e. geographical) and dynamic (network measurement
+systems) information for their replica selection, so the authors claim that
+peers which exhibit similar redirection behavior, are very likely they are close
+to the replica servers and as a consequence to each other. The redirection
+behavior is modeled in terms of \emph{ratio map} in the article's parlance, a
+vector of <replica-server,ratio> tuples, where ratio is the percentage of times
+CDN redirects the peer to the specific replica-server. The bootstrapping phase
+consists of the peer performing DNS lookup to CDN names in order to build its
+redirection information. The interval for polling DNSs starts from 30 seconds
+and increases by one minute every time redirection information to the CDN is
+found to remain unchanged after a lookup. On the other hand, the interval is
+halved whenever is found to have been changed. In order not to avoid the
+bootstrapping phase if the ratio map is sufficiently fresh, the protocol
+persists the information after the end of a BitTorrent session.
+
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% Authors claim minimization of inter-ISP traffic cost while enhancing the,
 % already, near optimal performance of BitTorrent download.
@@
 approaches in injecting locality awareness in BitTorrent-like applications. The
-first acts at a \emph{macroscopic-level} and targets neighbour selection. When a
+first acts at a \emph{macroscopic-level} and targets neighbor selection. When a
 peer asks the tracker to join, the later sorts the swarm peers according to
-their distance to the requesting peer in AS hop count and send it the first
+their distance to the requesting peer in Autonomous System hop count and send it
+the first
 (e.g. 50) peers in the list. The second manipulates the chocking/unchocking
 BitTorrent mechanisms at an \emph{intermediate-level}. A peer unchockes its 4
-closest in terms of AS hop count interested neighbours. The same applies also
-when the peer turns to the seeding state\footnote{Thus this approach favours
-least distance, contrary to the original BitTorrent implementation which favours
+closest in terms of Autonomous System hop count interested neighbors. The same
+applies also
+when the peer turns to the seeding state\footnote{Thus this approach favors
+least distance, contrary to the original BitTorrent implementation which favors
 uploading speed.}. On a \emph{microscopic-level}, the rare-first chunk picking
-policy is substituted by the locality-fist policy. The distance value of a piece
-is calculated as the mean value of the distances of the peers that posses it.
-The paper conducts several measurements on the efficiency of the above scheme to
-reach the conclusion that the major design desision for all locality aware P2P
-systems is the trade-off of optimising inter-AS traffic, which their solution
-achieves, and of the fairness among peers, a section that their locality-based
-approach does not do well in comparison to the random approach of the standard
-BitTorrent protocol.
-
-\begin{center}
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
-% minimisation of inter-AS traffic but unfair among peers.
-medium &
-% info (AS hop distances) aqcuired at swarm join
+policy is substituted by the locality-first policy. The distance value of a
+piece is calculated as the mean value of the distances of the peers that posses
+it. The paper conducts several measurements on the efficiency of the above
+scheme to reach the conclusion that the major design decision for all locality
+aware P2P systems is the trade-off of optimizing inter- Autonomous System
+traffic, which their solution achieves, and of the fairness among peers, a
+section that their locality-based approach does not do well in comparison to the
+random approach of the standard BitTorrent protocol.
+
+\begin{center}
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% minimization of inter-AS traffic but unfair among peers.
+medium &
+% info (AS hop distances) acquired at swarm join
 low &
@@
 ratio of download rate to upload rate, $\frac{d}{u}$, divided by link-level or
-AS-level routing hops, $l$ or $a$ respectively, in an attempt to form a
-comprehensive metric to select peers with high download rate, low reciprocal
-upload demand and low routing hops. The TopBT metric is applied on serveral
-parts of the original BitTorrent protocol in order to inject topology awareness
-for better peer selection; more specifically on the peer list the tracker
-returns when someone first tries to connect to the swarm, on the initial
+Autonomous System -level routing hops, $l$ or $a$ respectively, in an attempt to
+form a comprehensive metric to select peers with high download rate, low
+reciprocal upload demand and low routing hops. The TopBT metric is applied on
+several parts of the original BitTorrent protocol in order to inject topology
+awareness for better peer selection; more specifically on the peer list the
+tracker returns when someone first tries to connect to the swarm, on the initial
 connection establishment on the connection replacement and of course the
@@
 
-%A peer that runs the TopBT protocol evaluates its neighbours by periodicaly
-%emitting pings or traceroutes in order to uchock those peer that exhibit less
+%A peer that runs the TopBT protocol evaluates its neighbors by periodically
+%emitting pings or trace-routes in order to unchock those peer that exhibit less
 %hops to reach and higher download rates. \emph{Link-hops} are measured by using
@@
 %calculate the hops using the value of the TTL on the response message it
-%receives.}. Also the peer, when offline, builds a table that maps IPs to
+%receives.}. Also the peer, when off-line, builds a table that maps IPs to
 %\emph{AS-hops}, using BGP dumps.
@@
 % it takes into account both dl speed and network topology. The later though is
-% rathier coarse grained since it is taken into account on the AS level.
-medium &
-% the algoritm does not involve extensive calculations.
+% rather coarse grained since it is taken into account on the AS level.
+medium &
+% the algorithm does not involve extensive calculations.
 low &
@@
 peer selection, per se, based on this knowledge. For the first part, the
-algorithm uses network tomography, a techni  que which suggests probing only
+algorithm uses network tomography, a technique which suggests probing only
 from a large network's end points in order to infer its internal characteristics
-\cite{chny_tomography_2002}. Upon a new peer arrival, the tracker traceroutes it
-to get some knowledge (IP address, routers involved, RTT and hops conducted
+\cite{chny_tomography_2002}. Upon a new peer arrival, the tracker trace-routes
+it to get some knowledge (IP address, routers involved, RTT and hops conducted
 e.t.c). The more the peers in the swarm the better picture a tracker has for the
-underlying infrastructure, which can then return to the newcomming peer in the
-form of the bootstrapping peer list. The peers furtherly enhance this picture by
-tracerouting the returned peers themselves and reporting back to the tracker.
+underlying infrastructure, which can then return to the new coming peer in the
+form of the bootstrapping peer list. The peers further enhance this picture by
+trace-routing the returned peers themselves and reporting back to the tracker.
 For the selection part, a family of heuristics are proposed, which in general
@@
 medium &
-% traceroute and hop counts and calculation from both the tracker and the peers
+% trace-route and hop counts and calculation from both the tracker and the peers
 % (the later for all peers in the bootstrapping list) increases the control
@@
 clustering approach that differentiates peers in a swarm into local, intra-ISP
-and inter-ISP neighbours. The classification is done through the routers they
+and inter-ISP neighbors. The classification is done through the routers they
 use in order to communicate with the tracker. For this, a newly arriving
-BitTorrent peer traceroutes the tracker and sends the results in a form of a
+BitTorrent peer trace-routes the tracker and sends the results in a form of a
 vector of triples that contain the IP and the hop counter for the router on the
-traceroute path as well as the link latency of the hop that was made in order to
-arrive to this router. The tracker uses this information in order classify the
-router through a $k$-means classification algorithm, with $k = 3$. For peer
+trace-route path as well as the link latency of the hop that was made in order
+to arrive to this router. The tracker uses this information in order to classify
+the router through a $k$-means classification algorithm, with $k = 3$. For peer
 selection, the authors propose a biased approach where the tracker returns $c$
@@
 returned list. In choosing $c$, the tracker employs an iterative search process
-first in the peer's local neighbourhood. If not enough peers are available
+first in the peer's local neighborhood. If not enough peers are available
 there, it goes to intra-ISP and subsequently, if needed, to the inter-ISP
 cluster. The evaluation of the algorithm is done in a similar artificial,
-contained laboratory enviromnent as in \cite{LCY2008} and it shows up to
+contained laboratory environment as in \cite{LCY2008} and it shows up to
 5\% faster downloading times, and up to 22\% reduction of the total cross-ISP
@@
 \hline
-% It incoorporates clustering of peers to inter- and intra ISP which is a
+% It incorporates clustering of peers to inter- and intra ISP which is a
 % rather coarse grained method.
@@
 % tracker. Even though the tracker is backing the identification of a peer
-% being close or distant the actual computation is done in a distributed mannar
+% being close or distant the actual computation is done in a distributed manner
 % thus not incurring too much additional controlling overhead.
 medium &
-% TODO: how intensive with regards to computing power is an online k-means (with
+% TODO: how intensive with regards to computing power is an on-line k-means
+(with
 % k=3) classification???
@@
 \subsubsection{Peer-exchange Routing Optimization Protocols (PROP)}
-\cite{QCYCZ2007} introduces two protocols called \emph{Peer-exchange Routing
-Optimization Protocols (PROP)} to adjust the neighbourhood graph of an overlay
-network in order to reduce its overall link latency. PROP algorithms are based
-on the exchange of neighbours among peers, driven by the mutual benefit of the
-participants to reduce the network delay among their connections. This
-``collaboration'', is what differentiats this approach from others by letting
-two peers to optimize their neighbourhood environment, than simply letting each
-node to ``selfishly'' choose its own strategy. In PROP-G (Generic), a peer can
-exchange all of its neighbours with another peer, while in PROP-O (Optimized)
-two peers exchange a selected subset of their neighbours. PROP-G is a generic
-protocol that guarantees the connectivity of the overlay graph during exchanges,
-therefore can be applied to both unstructured and structured overlay networks.
+\cite{QCYCZ2007} introduces two protocols, one generic and one optimized,
+collectively referred to as \emph{Peer-exchange Routing Optimization Protocols
+(PROP)} to adjust the neighborhood graph of an overlay network in order to
+reduce its overall link latency. PROP algorithms are based on the exchange of
+neighbors among peers, driven by the mutual benefit of the participants to
+reduce the network delay among their connections. This ``collaboration'', is
+what differentiates this approach from others by letting two peers to optimize
+their neighborhood environment, than simply letting each node to ``selfishly''
+choose its own strategy. In PROP-G (Generic), a peer can exchange all of its
+neighbors with another peer, while in PROP-O (Optimized) two peers exchange a
+selected subset of their neighbors. PROP-G is a generic protocol that guarantees
+the connectivity of the overlay graph during exchanges, therefore can be applied
+to both unstructured and structured overlay networks.
 
@@
 medium &
-% Exchanging neighbours within a radius of 2 (TTL value) reduces stretch
+% Exchanging neighbors within a radius of 2 (TTL value) reduces stretch
 % significantly, while keeping the additional overhead at very reasonable levels
@@
 \emph{PROP-O} &
-% The algorithm can better reclaim the heterogenity of peers (those with faster
+% The algorithm can better reclaim the heterogeneity of peers (those with faster
 % response) to reduce the system's aggregate delay. 
@@
 % \right)} \cup E^{\left( blue \right)}$. A joining peer $u$, partitions its
-% neighbours into two subsets, the $B_u^{\left( red \right)}$ and $B_u^{\left(
+% neighbors into two subsets, the $B_u^{\left( red \right)}$ and $B_u^{\left(
 % blue \right)}$. In order to populate the $B_u^{\left( red \right)}$ subset, peer
@@
 connect to the peers anywhere on the network (called \emph{random} connections).
-The former ensure the refuction of long distance travelling for messages, while
-the later secure th connectivity of the structure and prevent partitioning. Upon
-arrival the newcomer asks a \emph{hostcache} in order to establish its random
-connections. For establishing its sibling connections, the newcommer multicasts
-a special walker message in order to help it locate the right candidate. If it
-is found then the candidate as well as its own siblings will respond to the
-newcomer about their willingness to establish connections with it.
+The former ensure the reduction of long distance traveling for messages, while
+the later secure the connectivity of the structure and prevent partitioning.
+Upon arrival the newcomer asks a \emph{host-cache} in order to establish its
+random connections. For establishing its sibling connections, the new comer
+multi-casts a special walker message in order to help it locate the right
+candidate. If it is found then the candidate as well as its own siblings will
+respond to the newcomer about their willingness to establish connections with
+it.
 
@@
 % overhead of split-hash and dnMatch is of low cost (TODO: double check this).
-% Walker message is multicasted only when sibling connections must be
+% Walker message is multi-casted only when sibling connections must be
 % established.
 low &
-% it can be utilised in hierarchical architectures as well meaning increased
+% it can be utilized in hierarchical architectures as well meaning increased
 % scalability. Tests show that an increase from 5K- to 10K-node system slightly
@@
 \subsubsection{Critical Topology-Aware Grouping (CTAG)}
-\emph{Critical Topology-Aware Grouping (CTAG)} presented in
-\cite{ZL2006}, is a grouping algorithm. The grouping strategy is
-based
-on the \emph{Internet Assigned Numbers Authority (IANA)} and the respective
-\emph{Regionanal Internet Registry (RIR)}'s IP assignment strategies, according
-to which nodes within the same organization are always addressed from the same
-block. The paper proposes the \emph{Adjacency Measurement (AM)}, a technique
-which uses the longest matching IP segment criterion to calculate node
-proximity. CTAG focuses on both the construction of the overlay as well as its
-constant and dynamic adaptation. In the first, called \emph{bootstrapping
-grouping}, the \emph{Gnutella Web Caching} mechanism is modified in order for a
-newcomer to choose the closest cache to retrieve the bootstrapping list. In the
-second, called \emph{dynamic revision} in the paper's context, the AM metric is
-used to store hosts' addresses. read from \emph{X-Try} headers during handshake
-or from \emph{QueryHit} messages. Additionally, when a node reaches the max
-neighbour connections, node disconnects the neighbours with the lowestAM.
+\emph{Critical Topology-Aware Grouping (CTAG)} presented in \cite{ZL2006}, is a
+grouping algorithm. The grouping strategy is based on the \emph{Internet
+Assigned Numbers Authority (IANA)} and the respective \emph{Regional Internet
+Registry (RIR)}'s IP assignment strategies, according to which nodes within the
+same organization are always addressed from the same block. The paper proposes
+the \emph{Adjacency Measurement (AM)}, a technique which uses the longest
+matching IP segment criterion to calculate node proximity. CTAG focuses on both
+the construction of the overlay as well as its constant and dynamic adaptation.
+In the first, called \emph{bootstrapping grouping}, the \emph{Gnutella Web
+Caching} mechanism is modified in order for a newcomer to choose the closest
+cache to retrieve the bootstrapping list. In the second, called \emph{dynamic
+revision} in the paper's context, the AM metric is used to store hosts'
+addresses. read from \emph{X-Try} headers during handshake or from
+\emph{QueryHit} messages. Additionally, when a node reaches the max neighbor
+connections, node disconnects the neighbors with the lowest AM.
 
@@
 Internet.%, as shown in Figure~\ref{fig:landmark_binning}.
-In order to detect
-locality, nodes use network latency (i.e. RTT) as a measurement technique. The
-network latency, even though not always accurate, is selected in this work due
-to its non-intrusive, transparent and easy to apply behaviour. In order for the
-binning to work, few well known anchor servers with known physical locations on
-the Internet. Authors estimate that a number of 12 such servers can prove
-sufficient for this. Every newly arriving node measures its distance from these
-landmarks and unilaterally decides to join a specific bin based on these
-results. Specifically, the node measures its round-trip time to each of the
-landmarks and orders the values in a decreasing order. The ordering represents
-a \emph{bin}, in the sence of close-by nodes having the same landmark ordering
-and hance belong to the same bin. This means that a landmark system consisting
-of $m$ such nodes results in $m!$ potentional different bins.The operation of
-the method is independent of the model incorporated by the overlay network and
-thus it can be applied with no significant changes to both structured and
-unstructured P2P systems. \cite{RHMKS2002} has the detailed description of such
-algorithm for both architectures. Landmark binning is also considered as a good
-candidate for content distribution networks in particular. The major
-disadvantage of landmark binning is to install and maintain landmark servers on
-different autonomous systems, all over the world. A typical P2P network usually
-has a couple of million nodes connected at any given time, thus rendering
-landmark servers an important parameter when designing for large scales.  The
-authors claim that latency estimation does not drain much from the network
-resources but in case scalability problems arise, they claim the solution is
-replacing single landmark servers with clusters of servers within the same
-physical area. However, this approach does not reduce the possibility of
-excessive network traffic flow through these landmark servers. One other
-possible problem is the incorrect binning caused by the use of inaccurate
-methods for delay measurement, since tracking down, for example, network
-latencies is far from proven to be safe for location estimation.
+In order to detect locality, nodes use network latency (i.e. RTT) as a
+measurement technique. The network latency, even though not always accurate, is
+selected in this work due to its non-intrusive, transparent and easy to apply
+behavior. In order for the binning to work, a few anchor servers with known
+physical locations need to be installed in strategic positions accross the the
+Internet. Authors estimate that a number of 12 such servers can prove sufficient
+for this. Every newly arriving node measures its distance from these landmarks
+and unilaterally decides to join a specific bin based on these results.
+Specifically, the node measures its round-trip time to each of the landmarks and
+orders the values in a decreasing order. The ordering represents a \emph{bin},
+in the sense of close-by nodes having the same landmark ordering and hence
+belong to the same bin. This means that a landmark system consisting of $m$ such
+nodes results in $m!$ potential different bins.The operation of the method is
+independent of the model incorporated by the overlay network and thus it can be
+applied with no significant changes to both structured and unstructured P2P
+systems. \cite{RHMKS2002} has the detailed description of such algorithm for
+both architectures. Landmark binning is also considered as a good candidate for
+content distribution networks in particular. The major disadvantage of landmark
+binning is to install and maintain landmark servers on different autonomous
+systems, all over the world. A typical P2P network usually has a couple of
+million nodes connected at any given time, thus rendering landmark servers an
+important parameter when designing for large scales.  The authors claim that
+latency estimation does not drain much from the network resources but in case
+scalability problems arise, they claim the solution is replacing single landmark
+servers with clusters of servers within the same physical area. However, this
+approach does not reduce the possibility of excessive network traffic flow
+through these landmark servers. One other possible problem is the incorrect
+binning caused by the use of inaccurate methods for delay measurement
+(latency is not proven to be a safe way to estimate distance).
 
@@
 %To answer the question as to whether the algorithm actually contributes
-% possitivelly to the construction of an enhaned overlay, the paper defines the
+% positively to the construction of an enhanced overlay, the paper defines the
 %\emph{gain ratio} as the factor by which the latency reduces when someone
@@
 %
-% TODO: LANDMARK BINING FOR UNSTRUCTURED OVERLAYS
+% TODO: LANDMARK BINNING FOR UNSTRUCTURED OVERLAYS
 %
 %For unstructured overlays the paper assumes \emph{a set of $n$ nodes where each
-% node picks any $k$ neighbour nodes so that the average routing latency on the
+% node picks any $k$ neighbor nodes so that the average routing latency on the
 %resultant overlay is low (assuming shortest path routing)}. According to the
 %proposed heuristic algorithm called \emph{BinShort-Long}, a node picks its
-%neighbours by choosing its $\frac{k}{2}$ closest\footnote{If the node's bin is
-%not large enough for it to pick these $\frac{k}{2}$ neighbours, it picks the
+%neighbors by choosing its $\frac{k}{2}$ closest\footnote{If the node's bin is
+%not large enough for it to pick these $\frac{k}{2}$ neighbors, it picks the
 %required nodes from the bin that matches the most in terms of landmark
@@
 %connectivity of the graph, both yielding a proximity factor of $\alpha = 0.5$
-%in an attempt to preserve the benefitial properties of unstructured
+%in an attempt to preserve the beneficial properties of unstructured
 %topologies\cite{merugu_str2unstr_2003}.
@@
 %case, the fact that it is not fully distributed, renders the protocol's
-%scalability directly volnerable to any system size increase as well as
-%usuitable for highly dynamic networks such as ad-hoc networks. Moreover, fixed
-%points in a network are inherently more exposed to malicius attacks. The most
+%scalability directly vulnerable to any system size increase as well as
+%suitable for highly dynamic networks such as ad-hoc networks. Moreover, fixed
+%points in a network are inherently more exposed to malicious attacks. The most
 %significant downside of the algorithm though is that it can lead to an
-%extremely uneven overlay ID distribution causing load inbalances and hot spots.
+%extremely uneven overlay ID distribution causing load unbalances and hot spots.
 %Lastly, the scheme is coarse grained when it comes to distinguishing relatively
@@
 % The introduction of landmark servers renders the approach not fully
-% decentralised, thus preventing it from scalling smoothly. Communicating
+% decentralized, thus preventing it from scaling smoothly. Communicating
 % and overloading landmark servers in high-churn systems is another scalability
@@
 \subsubsection{mOverlay}
-\emph{mOverlay} \cite{ZZZSZ2004} tries to addresses the scalability
-issues that might be imposed to networks (i.e. load unbalance) when using static
-landmark servers by introducing dynamic landmarks. Zhang et al. introduce the
-notion of a \emph{group} which is a set of peers considered close to each other
-with respect to any position in the underlying network, proximity being defined
-by some user-defined cost metric like RTT, latency or something else. mOverlay
-is considered a clustering approach since it tries to recreate small-world-like
+\emph{mOverlay} \cite{ZZZSZ2004} tries to addresses the scalability issues that
+might be imposed to networks (i.e. load unbalance) when using static landmark
+servers by introducing dynamic landmarks. Zhang et al. introduce the notion of a
+\emph{group} which is a set of peers considered close to each other with respect
+to any position in the underlying network, proximity being defined by some
+user-defined cost metric like RTT, latency or something else. mOverlay is
+considered a clustering approach since it tries to recreate small-world-like
 properties to the overlay network by creating a two-level hierarchical
@@
 grouped based on their distance to the groups already in the network, rendering
-the later the (dynamic) landmarks for the process. The authors formally prove
+the later (dynamic) landmarks for the process. The authors formally prove
 that any new node can reach its group by exchanging at most $O(logN)$ messages
@@
 %then, measures its distance to host $A$. At the same time, the later, sends
-%information about the neighbour groups of group $1$ back to host $Q$. This list
-%is called \emph{candidate group list}, and the newcoming host sequnentially
+%information about the neighbor groups of group $1$ back to host $Q$. This list
+%is called \emph{candidate group list}, and the new coming host sequentially
 %measures its distance to each of them in seek for the closest one. If the
@@
 %case, $Q$ creates a new group comprising itself only. The above protocol does
-%not favour hotspots as it spreads the probability of visiting a group across
+%not favor hot-spots as it spreads the probability of visiting a group across
 %the whole overlay and limits the overhead in the level of $O \left ( log N
@@
 %\paragraph{General overlay operations} A set of additional protocols, are also
-% introduced, similar to those found in traditional unstructred networks, but
+% introduced, similar to those found in traditional unstructured networks, but
 %modified focusing on scalability and robustness. For example a protocol for
 %\emph{group formation} is introduced that exploits the inherent characteristic
-%of proximity, in the overlay, in order to efficiently detect the neighbouring
-%groups of a newly formated group from the set of adjasent groups of its closest
-%neighbour. Additionaly, during \emph{group joining} the coresponding protocol
+%of proximity, in the overlay, in order to efficiently detect the neighboring
+%groups of a newly formated group from the set of adjacent groups of its closest
+%neighbor. Additionally, during \emph{group joining} the corresponding protocol
 %denotes the exchange of important information for group maintenance. This can
-%be furtherly improved by \emph{information sharing} between nodes of the same
+%be further improved by \emph{information sharing} between nodes of the same
 %group, functionality handled by a dedicated flood-like protocol\footnote{Since
@@
 %  \item the host cache, when a new node joins, and
-%  \item the neighbours of groups, when a close-by group is generated.
+%  \item the neighbors of groups, when a close-by group is generated.
 %\end{inparaenum}
@@
 
+% The introduction of landmark servers renders the approach not fully
+% decentralized, thus preventing it from scaling smoothly. Communicating
+% and overloading landmark servers in high-churn systems is another scalability
+% concern.
+
 In this section we introduced the topology mismatching problem from the
-perspective of decentralised unstruThe introduction of landmark servers renders
-the approach not fully
-% decentralised, thus preventing it from scalling smoothly. Communicating
-% and overloading landmark servers in high-churn systems is another scalability
-% concern.ctured architectures and we visited a series
+perspective of decentralized unstructured architectures and we visited a series
 of algorithms that have been proposed in the literature to tackle its negative
-influence to the networks' performance. Even though, at first glance, the
-algorithms seem to do things their own way actually we can distinguish four main
-methodologies that are being utilised in order for every algorithm to achive its
-intended result. These methodologies are:
+influence to networks' performance. Even though, at first glance, the algorithms
+seem to do things their own way actually we can distinguish four main
+methodologies that are being utilized in order for every algorithm to achieve
+its intended result. These methodologies are:
 \begin{enumerate}[\itshape i\upshape)]
   \item topology adaptation
-  \item forwarding optimisation
+  \item forwarding optimization
   \item caching and replication, and
@@
 
-Table~\ref{unstructured:table} offers a summarised overview of the algorithms
-we saw earlier in the section and tracks down the afformentioned methodologies
+Table~\ref{unstructured:table} offers a summarized overview of the algorithms
+we saw earlier in the section and tracks down the aforementioned methodologies
 each of them incorporates along with their highlights and a rough estimation of
 positive and negative aspects of their implementation.
-
-\input{unstructured-table}
-
-\input{unstructured-plot}
 
@@
 \centering
-\subfigure[Ineffictient overlay topology.] {
+\subfigure[Inefficient overlay topology.] {
   \includegraphics[scale=0.4]{img/pdf/topology-adaptation-before.pdf}
@@
 any real topology adaptation mechanism meaning that it was designed for small
-groups (therefore not scalable). Its followers try to overcome these obstacles
-with clever tricks like forming minimum spanning trees for the two-hop
-neighbours ($N^2$) of each node, partitioning the graph into two random groups
-where each group is responsible of different tasks, or performing local
-optimizations dynamically on the overlay graph. The advantage of building
+groups (therefore not scalable). Algorithms like AOTO, LTM or SBO try to
+overcome these obstacles with clever tricks like forming minimum spanning trees
+for the two-hop neighbors ($N^2$) of each node, partitioning the graph into two
+random groups where each group is responsible of different tasks, or performing
+local optimizations dynamically on the overlay graph. The advantage of building
 minimum spanning trees is that they maintain the connectivity on the network in
@@
 The cluster based approaches, on the other hand, select to link physically
-closer nodes with each other. T2MC, for example, uses traceroute logs and DDNO
-uses domain names to cluster close by nodes. Unfortunatelly, commonly used
+closer nodes with each other. T2MC, for example, uses trace-route logs and DDNO
+uses domain names to cluster close by nodes. Unfortunately, commonly used
 methods for proximity detection across Internet do not always return
 reliable results, therefore mapping accuracy is not guaranteed. Moreover,
-traceroute, is a heavy-weight utility to be frequently used on the
+trace-route, is a heavy-weight utility to be frequently used on the
 network, let alone the fact that many network equipment vendors do not allow
-traceroute calls at all. The most problematic aspect of clusterting, though, is
+trace-route calls at all. The most problematic aspect of clustering, though, is
 its nature per se. Limited connectivity among the various local domains can
@@
 
-Concluding, topology adaptation ultimatelly targets on the reduction of the
+Concluding, topology adaptation ultimately targets on the reduction of the
 average cost of the paths from one node to another. This way someone can deduce
@@
 thus reducing the average latency of the network communication. The pitfall here
-is that this new virtual link can traverse a fast AS-to-AS link meaning that
-even though is reduced message round-trip-time, it has more cost in terms of
-communication network economics and management.
+is that this new virtual link can traverse a fast Autonomous System -to-
+Autonomous System link meaning that even though is reduced message
+round-trip-time, it has more cost in terms of communication network economics
+and management.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Forwarding optimisation}
+\subsubsection{Forwarding optimization}
 
@@
 \subfigure[Typical blind flooding.] {
-\includegraphics[scale=0.25]{img/pdf/forwarding-optimisation-before.pdf}
+\includegraphics[scale=0.25]{img/pdf/forwarding-optimization-before.pdf}
   \label{figure:forwarding-optimisation:before}
@@
 \subfigure[A node can decide where to forward the messages.] {
-  \includegraphics[scale=0.25]{img/pdf/forwarding-optimisation-after.pdf}
+  \includegraphics[scale=0.25]{img/pdf/forwarding-optimization-after.pdf}
   \label{figure:forwarding-optimisation:after}
 }
-\caption{Forwarding optimisation in constraining message flooding.}
+\caption{Forwarding optimization in constraining message flooding.}
 \label{figure:forwarding-optimisation}
@@
 % Researchers initially started considering alternatives for services who
-% implement one-to-many communication schemes, like IP multicast, on the
-% application layer (i.e. Narada). Application level multicast protocols
+% implement one-to-many communication schemes, like IP multi-cast, on the
+% application layer (i.e. Narada). Application level multi-cast protocols
 % are generic protocols that can be applied to both P2P file sharing systems,
-% as well as content distribution networks. Application layer multicast, however,
+% as well as content distribution networks. Application layer multi-cast,
+however,
 % is not as efficient as the one enabled by IP and still struggles to ease the
@@
 
-Forwarding optimisation can reduce the emitted messages flooding the overlay and
-as a concequence the overall burden, network resources have to handle. In the
-first realisation of the Gnutella protocol, each query request received by a
-peer, was then forwarded to all of its neighbours. This was the source of
-unecessary, sometimes redundant traffic. Forwarding-optimisation-based
+Forwarding optimization can reduce the emitted messages flooding the overlay and
+as a consequence the overall burden, network resources have to handle. In the
+first realization of the Gnutella protocol, each query request received by a
+peer, was then forwarded to all of its neighbors. This was the source of
+unnecessary, sometimes redundant traffic. Forwarding-optimization-based
 approaches, propose intelligent forwarding link selection for a message's next
@@
 friendship. Figure~\ref{figure:forwarding-optimisation} shows a simple example
-of how forwarding optimisation can work.
-Subfigure~\ref{figure:forwarding-optimisation:before} shows an algorithm that
+of how forwarding optimization can work.
+Sub-figure~\ref{figure:forwarding-optimisation:before} shows an algorithm that
 floods the entire network for search of an object the existence of which is
-denoted by black nodes. Alternativelly, a node can decide to forward its
+denoted by black nodes. Alternatively, a node can decide to forward its
 messages not to every output link but to a specific or subset of its output
-links. In
-Subfigure~\ref{figure:forwarding-optimisation:after}, the node in the middle,
-does not flood its neighbours. Instead it picks one of them. The dashed line on
-the right, shows that, for this routing process, the algorithm has rendered two
-potential forwarding paths as inefficient (e.g., the target nodes show
-overloading signs). These appoaches have the advantage of enhancing the search
-responsiveness and reducing the aggregate resource usage of the physical
+links. In Sub-figure~\ref{figure:forwarding-optimisation:after}, the node in the
+middle, does not flood its neighbors. Instead it picks one of them. The dashed
+line on the right, shows that, for this routing process, the algorithm has
+rendered two potential forwarding paths as inefficient (e.g., the target nodes
+show overloading signs). These approaches have the advantage of enhancing the
+search responsiveness and reducing the aggregate resource usage of the physical
 network. On the other hand, they suffer from drastic reduction of the search
-scope (in Subfigure~\ref{figure:forwarding-optimisation:after} the object on the
-far right is not reached) thus limiting the scalability of the whole network.
-Moreover they don't actually solve the topology mismatch problem since they
-don't give any guarantees that overlay and underlying topologies are aligned
-with each other let alone quantify the mismatch and try to alleviate. Like all
-other approaches, forwarding based methods are commonly applied in conjunction
-with other to improve the quality of the P2P systems.
+scope (in Sub-figure~\ref{figure:forwarding-optimisation:after} the object on
+the far right is not reached) thus limiting the scalability of the whole
+network. Moreover they don't actually address the problem of topology mismatch
+since they don't give any guarantees that overlay and underlying topologies are
+aligned with each other let alone quantify the mismatch and try to alleviate.
+Like all other approaches, forwarding based methods are commonly applied in
+conjunction with other to improve the quality of the P2P systems.
 
 %Expanding the search scope, on the other hand, is no easy task
-%because the overhead of forming multicast trees is proportional to the
-%multicast group size.
+%because the overhead of forming multi-cast trees is proportional to the
+%multi-cast group size.
 %forwarding based schemes do not consider dynamic joining
@@
 }\qquad\qquad
-\subfigure[Data replication allong the path of a successful query.] {
+\subfigure[Data replication along the path of a successful query.] {
   \includegraphics[scale=0.3]{img/pdf/replication-data.pdf}
@@
 
-Caching is a widely used technique to exploit locality and minimise redundant
-transfer of data, initially and successfully adopted by web and file server
-application environments. Since peers in a P2P system also operate as
-servers, it is intuitively expected that P2P file sharing systems can also
-benefit from caching in improving performance and reducing overall resource
-usage. However, the design of caches in this context is non-trivial compared to
-the web-based caching. Exactly, because each node can act as both a server and a
-client, two important issues have to be considered at design time. First, the
-lifetime of a query is short, as the nodes join and leave frequently. Second,
-the result of a single query string is not always the same, as it is dependent
-on the source of the query, the TTL value set for the messages, the current
+Caching is a widely used technique to exploit locality and minimize redundant
+transfer of data, previously successfully adopted by web and file server
+application environments. Since peers in a P2P system also operate as servers,
+it is intuitively expected that P2P file sharing systems can also benefit from
+caching in improving performance and reducing overall resource usage. However,
+the design of caches in this context is non-trivial compared to the web-based
+caching. Due to the fact that each node can act as both a server and a client,
+two important issues have to be considered at design time. First, the lifetime
+of a query is short, as the nodes join and leave frequently. Second, the result
+of a single query string is not always the same, as it is dependent on the
+source of the query, the TTL value set for the messages, the current
 interconnection of peers and the high volatility of the environment. So, in
@@
 on two different levels, namely caching indices or pointers to data (see
-Subfigure~\ref{figure:replication:index}) or caching
-the data itself (see
-Subfigure~\ref{figure:replication:data}). Successfull implementations have
+Sub-figure~\ref{figure:replication:index}) or caching the data itself (see
+Sub-figure~\ref{figure:replication:data}). Successful implementations have
 already been developed in some commercial P2P systems, like KazaA, which uses
-caching of indices on its super-peer level. Unfortunatelly, even though the
+caching of indices on its super-peer level. Unfortunately, even though the
 state of the art in P2P algorithms using caching methods reduce the resource
 usage of the network, they also fail to consistently address typical problems
-of a topologicaly mismatched environment such as duplication of messages.
+of a topologically mismatched environment such as duplication of messages.
 
@@
 known servers (the landmark servers) on the Internet. Landmark servers are used
-by nodes to estimate their positions based on the intutive assumption that nodes
+by nodes to estimate their positions based on the intuitive assumption that
+nodes
 with similar distances to a set of landmarks, are physically close to each
-other, as well, over the network. In Figure~\ref{figure:landmarking} the newlly
+other, as well, over the network. In Figure~\ref{figure:landmarking} the newly
 arriving peer, measures its distance to an array of landmark point denoted as
@@
 order to choose the already participating peers with which to connect to,
-compares its list to the list of the potential neighbour. Peers with similar
-measured distances to the these landmark point are very likely to be close to
-each other in the physical level as well.
+compares its list to the list of the potential neighbor. Peers with similar
+measured distances to the these landmark points is very likely to be close to
+each other as well.
 
@@
 accuracy problem affects the overall performance of the system. Moreover,it can
-be characterised as a rather coarse-grained approximation, therefore not
+be characterized as a rather coarse-grained approximation, therefore not
 particularly well suited for detecting the correct positions of nodes within
 close distance to each other.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\input{unstructured-table}
+
+\input{unstructured-plot}

### 2014-01-27 structured.tex
2012.11.19 (Hüseyin) Resolving similarity issues.

@@
 the maximum degree of each node. LAPTOP implements the geographical layout
-approach  and constructs its layout in a self-organizing and efficient fashion,
-by estimating the round-trip times to a small number of nodes in the overlay
-network in order to make them roughly aware of their physical distances among
-them.
+approach  and constructs its layout in a self-organizing and efficient fashion.
+The physical distances are estimated with round trip times to a few existing
+nodes in the overlay network.
 
@@
 \subsubsection{T2MC}
-T2MC \cite{SLCGZ2008} uses trace-route logs to detect different autonomous
-systems and cluster close by peers, thus reducing redundant traffic that crosses
-autonomous system boundaries. T2MC uses a customized k-mean classification
-algorithm with $k=2$ to perform the classification, and exploits the stable
-structure of the Internet routers to guide clustering. The peer chooses the
-minimum and maximum Latency results from the trace-route to initialize the
-centroids of two sets. Then the peer calculates, for all hops along its
-trace-routed path, the peer calculates the absolute distance to these centroids
-of both sets and assigns the routers to that centroid with which it has the
-smaller absolute distance. After that, the mean and variance values are
-calculated and if the variance is larger than a predefined threshold then the
-algorithm sets the two latency mean values as new centroids for the two sets and
-loops again. Ultimately peer will end up with two sets having the minimum
-intra-set variance. The peer, then, chooses the router from the second set with
-the minimum hops attribute and sets it as a threshold. The selected router and
-all others whose hops attribute is larger than the threshold are classified as
-\emph{remote} router cluster. The remaining are classified as \emph{near}. From
-the near class, the peer chooses the one with maximum hops attribute as its edge
-router, and registers it along with all near clusters into the DHT of the
-P2P overlay. As new peers join the network, those that share the same edge
-router or any of the members of the near router clusters, would form a close
-peer cluster. As Edge routers can provide more valuable information than other
-members of the near set, T2MC was designed to prioritize interaction of peers
-and edge gateways.
+
+T2MC \cite{SLCGZ2008} uses traceroute logs to detect the AS boundaries and
+cluster close by peers with each other to reduce the redundant multiple message
+passes between the AS boundaries. T2MC uses a customized k-mean classification
+algorithm with k = 2 to perform the classification, and exploits the stable
+structure of the internet routers to guide clustering. Even though traceroute
+provides detailed information about the network structure, use of traceroute
+creates overhead to the overall network structure. For this same reason, it is
+not unusual for network administrators to disable this support on their network
+routers, which may affect the performance of the T2MC algorithm.
+
+%T2MC \cite{SLCGZ2008} uses trace-route logs to detect different autonomous
+%systems and cluster close by peers, thus reducing redundant traffic that crosses
+%autonomous system boundaries. T2MC uses a customized k-mean classification
+%algorithm with $k=2$ to perform the classification, and exploits the stable
+%structure of the Internet routers to guide clustering. The peer chooses the
+%minimum and maximum Latency results from the trace-route to initialize the
+%centroids of two sets. Then the peer calculates, for all hops along its
+%trace-routed path, the peer calculates the absolute distance to these centroids
+%of both sets and assigns the routers to that centroid with which it has the
+%smaller absolute distance. After that, the mean and variance values are
+%calculated and if the variance is larger than a predefined threshold then the
+%algorithm sets the two latency mean values as new centroids for the two sets and
+%loops again. Ultimately peer will end up with two sets having the minimum
+%intra-set variance. The peer, then, chooses the router from the second set with
+%the minimum hops attribute and sets it as a threshold. The selected router and
+%all others whose hops attribute is larger than the threshold are classified as
+%\emph{remote} router cluster. The remaining are classified as \emph{near}. From
+%the near class, the peer chooses the one with maximum hops attribute as its edge
+%router, and registers it along with all near clusters into the DHT of the
+%P2P overlay. As new peers join the network, those that share the same edge
+%router or any of the members of the near router clusters, would form a close
+%peer cluster. As Edge routers can provide more valuable information than other
+%members of the near set, T2MC was designed to prioritize interaction of peers
+%and edge gateways.
 

### 2014-01-27 unstructured.tex
2012.11.19 (Hüseyin) Resolving similarity issues.

@@
 pair of neighbors (meaning also that additional, cost probing, message types
-need to be added to the Gnutella protocol). Whenever a new neighbor cost table
-is received or there is a change of neighbors, the source peer has to
-re-calculate the multi-cast tree and apply the randomized AT algorithm.
+need to be added to the Gnutella protocol). The randomized AT algorithm is
+applied by the source peer everytime the neighbour list is updated or an updated
+neighbour cost table is received.
 
@@
 \subsubsection{Adaptive Connection Establishment (ACE)}
-\emph{Adaptive Connection Establishment (ACE)} \cite{LZXN2004} builds an
-overlay multi-cast tree among each source node and the peers within a certain
-diameter from the source peer and optimizes the neighbor connections that are
-not in that tree. To achieve that, it calculates cost between nodes using
-network delay as a metric. Each peer probes the costs with its immediate logical
+\emph{Adaptive Connection Establishment (ACE)} \cite{LZXN2004} uses the network
+delay as a metric to calculate the costs between nodes in a spanning tree.
+Each peer calculates the costs to its logical one-hop
 neighbors and forms a \emph{neighbor cost table (NCT)} using a special
@@
 small overlay topology. Moreover, based on obtained NCTs a minimum spanning tree
-among each peer and its immediate neighbors is built. Ultimately, physically
+among each peer and its one-hop neighbors is built. Ultimately, physically
 far away neighbors are replaced by physically close ones. Specifically, a peer
@@
 When $p$ receives a query from $n$, this PM will be piggybacked to the query
-message that will be then forwarded to all other neighbors of peer $p$. Upon
-receiving such a query message, each of the other neighbors will detach the PM
-information (it will not be further forwarded), record the $pq$ distance and
-process the query as always. Selection of which incoming queries should
-piggyback a PM is proposed by either \emph{pure probability-based (PPB)} or
-\emph{new neighbor triggered (NNT)} policies. The second component, namely
-neighbor comparison and selection, dictates peer $p$ to probe the distance to
-all his known, not yet probed, two-hop neighbors ($ N^2(p)$). The distance of
-$pn$ is known to $p$. Upon receiving a PM from node $n$ with the distance of
-$nq$, where $q$ is a direct neighbor of $n$, $p$ chooses one of the following
-two paths. First, if $q$ is a direct neighbor of $p$, then the later will check
-the cost of the involved links. If the most cost-intensive link is one of $pq$
-or $pn$, then it is put into a \emph{will-cut} list. If it is $nq$ then nothing
-is done (since either $n$ or $q$ will have the chance to handle it themselves).
-On the other hand, if $q$ is within two-hop radius from $p$, the later probes
-the former (if it hasn't got the distance information yet) and again compares
-distances $pq$, $pn$ and $nq$. If $pq$ is the most cost-intensive, $p$ will not
-establish the connection. If it is $pn$, $p$ will establish connection $pq$ and
-put $pn$ in the will-cut list. If it is $nq$, $p$ will keep the connection with
-both $n$ and $q$, expecting that eventually either $n$ or $q$ will disconnect
-link $nq$, later.
+message that will be then forwarded to all other neighbors of peer $p$. 
+Each neighbor will detach the PM information (it will not be further forwarded),
+record the $pq$ distance and process the query as always after receiving a query
+message. Selection of which incoming queries should piggyback a PM is proposed
+by either \emph{pure probability-based (PPB)} or \emph{new neighbor triggered
+(NNT)} policies. The second component, namely neighbor comparison and selection,
+dictates peer $p$ to probe the distance to all his known, not yet probed,
+two-hop neighbors ($ N^2(p)$). The distance of $pn$ is known to $p$. Upon
+receiving a PM from node $n$ with the distance of $nq$, where $q$ is a direct
+neighbor of $n$, $p$ chooses one of the following two paths. First, if $q$ is a
+direct neighbor of $p$, then the later will check the cost of the involved
+links. If the most cost-intensive link is one of $pq$ or $pn$, then it is put
+into a \emph{will-cut} list. If it is $nq$ then nothing is done (since either
+$n$ or $q$ will have the chance to handle it themselves).  On the other hand, if
+$q$ is within two-hop radius from $p$, the later probes the former (if it hasn't
+got the distance information yet) and again compares distances $pq$, $pn$ and
+$nq$. If $pq$ is the most cost-intensive, $p$ will not establish the connection.
+If it is $pn$, $p$ will establish connection $pq$ and put $pn$ in the will-cut
+list. If it is $nq$, $p$ will stay connected with both $n$ and $q$,
+such that eventually in later stages either $n$ or $q$ will disconnect link $nq$.
 
@@
   \item it does not need any clock synchronization,
-  \item it is a fully distributed algorithm making it robust and reliable in
+  \item it is a fully distributed algorithm making it suitable for
         decentralized systems,
-  \item the traffic overhead of the triple hop adjustment is very low,
-  \item it is applicable to dynamic peer-to-peer environments, and
+  \item the traffic overhead is low,
+  \item it can be used in dynamic peer-to-peer environments, and
   \item maintains lower query response times.
@@
 Choffnes and Bustamante propose Ono \cite{CB2008}, a protocol for managing
-BitTorrent traffic so that it significantly reduces cross-ISP traffic and in
-the same time enhances downloading rates by favoring connections within the
-borders of a single autonomous system. Contrary to biased peer selection
-proposed in \cite{BCCMSBZ2006}, this work can lead to improved performance with
-no cooperation between ISPs and their subscribers, no additional infrastructure
-and no network topology information. The selection approach is landmark-based
-and leverages existing CDN infrastructure for peer distance estimation. CDNs
-already use both static (i.e. geographical) and dynamic (network measurement
-systems) information for their replica selection, so the authors claim that
-peers which exhibit similar redirection behavior, are very likely they are close
-to the replica servers and as a consequence to each other. The redirection
-behavior is modeled in terms of \emph{ratio map} in the article's parlance, a
-vector of <replica-server,ratio> tuples, where ratio is the percentage of times
-CDN redirects the peer to the specific replica-server. The bootstrapping phase
-consists of the peer performing DNS lookup to CDN names in order to build its
-redirection information. The interval for polling DNSs starts from 30 seconds
-and increases by one minute every time redirection information to the CDN is
-found to remain unchanged after a lookup. On the other hand, the interval is
-halved whenever is found to have been changed. In order not to avoid the
-bootstrapping phase if the ratio map is sufficiently fresh, the protocol
-persists the information after the end of a BitTorrent session.
+BitTorrent traffic so that it significantly reduces cross-ISP traffic and in the
+same time enhances downloading rates by favoring connections within the borders
+of a single autonomous system. Contrary to biased peer selection proposed in
+\cite{BCCMSBZ2006}, this work can lead to improved performance without requiring
+any cooperation between ISPs subscribers, or requiring any additional
+infrastructure and network topology information. The selection approach is
+landmark-based and leverages existing CDN infrastructure for peer distance
+estimation. CDNs already use both static (i.e. geographical) and dynamic
+(network measurement systems) information for their replica selection, so the
+authors claim that peers which exhibit similar redirection behavior, are very
+likely they are close to the replica servers and as a consequence to each other.
+The redirection behavior is modeled in terms of \emph{ratio map} in the
+article's parlance, a vector of <replica-server,ratio> tuples, where ratio is
+the percentage of times CDN redirects the peer to the specific replica-server.
+The bootstrapping phase consists of the peer performing DNS lookup to CDN names
+in order to build its redirection information. The interval for polling DNSs
+starts from 30 seconds and increases by one minute every time redirection
+information to the CDN is found to remain unchanged after a lookup. On the other
+hand, the interval is halved whenever is found to have been changed. In order
+not to avoid the bootstrapping phase if the ratio map is sufficiently fresh, the
+protocol persists the information after the end of a BitTorrent session.
 

### 2014-01-27 background.tex
2012.12.23 (Mema) Text reworked.

@@
 issues \cite{hughes_socp2p_2008}, have arisen, peer-to-peer systems have
-continued to gain popularity mainly because they have successfully supported
+continued to gain popularity throughout the years primarily because they have successfully supported
 popular applications (i.e. file-sharing) amongst vast numbers of participating
@@
 a dedicated server but instead could be handled by the many hosts that already
-posses the resources. The downside of this approach is that it requires a
+posess the resources. The downside of this approach is that it requires a
 centralized searching facility which becomes a single point of failure, a
@@
 content. Thus, by centralizing search while distributing downloads,  Napster
-achieved a highly functional design that, at its height, attrated 26 million
+achieved a highly functional design that, at its height, attracted 26 million
 users sharing approximately 80 million songs\cite{jmm_naptopusage_2001}.
@@
 ultra/super-peers), the disadvantages of reduced search scope and inability to
-locate rare resources efficiently remain.
+locate rare resources remain.
 
@@
 through a deterministic mechanism. The DHT provides a guaranteed bound on the
-number of overlay routing hops that have to be taken to locate a resource, even
-in the case when only a single copy exists in the system.  This means
+number of overlay routing hops that have to be crossed to locate a resource, even
+in the case when only a single copy exists in the system.  This requires
 $O \left( log n \right)$ hops, compared to unstructured networks that require
@@
 }\qquad\qquad
-\subfigure[Structured (Chord and Kademlia). Peers and data are placed into the
+\subfigure[Structured (Chord and Kademlia). Peers and data are placed in the
 overlay in a deterministic manner.] {
@@
 One of the major issues that defines the efficiency of an overlay network is the
-mapping of the overlay structure to the underlying infrastructure. Remember, in
+mapping of the overlay structure to the underlying infrastructure. Recall,, in
 Figure~\ref{figure:overlay} nodes $A$ and $B$ are on the same local network,
@@
 connections there, can change as needed by the running environment, without any
-particular constraint by the underlying physical topology. Let's have a closer
-look with the help of a simple example. Assume nodes $A$, $B$, $C$ and $D$ are
-connected through the physical network shown in Figure~\ref{figure:phys}, where
+particular constraint by the underlying physical topology. Consider the following
+example. 
+Assume nodes $A$, $B$, $C$ and $D$ are
+connected via the physical network shown in Figure~\ref{figure:phys}, where
 the network costs are given in milliseconds, and peer $A$ sends a message to
@@
 
-If these peers participate in an overlay network according to one of the setups
-of Figure~\ref{figure:overlay-confs}, then users will experience different
-performances. In the overlay depicted in Figure~\ref{figure:overlay-1}, the
+Depending on the configuration of the overlay,
+peers will experience different performance.
+For example, in the overlay depicted in
+Figure~\ref{figure:overlay-1}, the
 message will traverse the following sequence of links in the physical layer
@@
 Ideally, an overlay network should be constructed to achieve an optimal 
-mapping among the overlay paths and the underlying network links to avoid
-inefficient states were redundant physical resources are utilised for the
-operation of a link on the overlay. The problem of constructing an optimal
+mapping between the overlay paths and the underlying network links to avoid
+inefficient states where redundant physical resources are utilised for the
+operation of the same link in the overlay. The problem of constructing an optimal
 overlay is referred to as the \emph{topology mismatch problem}, and is formally
@@
 flooded over multiple paths merging in one peer while one of the paths would
-have been enough. Moreover, peers may forward the same message to each other
+suffice.  Moreover, peers may forward the same message to each other
 before they receive the query messages from each other.
@@
 randomness has a negative impact on the \emph{routing locality} of the network.
-This means that even though the target node can be reached with logarithmic
-overlay hops, the distance traveled in the underlying network, during the
+This means that even though the target node can be reached with a logarithmic
+number of overlay hops, the distance traveled in the underlying network, during the
 overlay routing process, can be far from optimal.  The ratio of the actual IP
@@
   \item further discussion that can help the reader identify, through
-a taxonomy, the pluses and minuses in the preceding material,
+a taxonomy, the pluses and minuses  of the presented algorithms,
   \item a comprehensive tabular organization of the key-point information, and
@@
 
-In discussion we offer the reader a deeper insight for the observations that
+We hope to offer the reader a deeper insight for the observations that
 pushed research toward some direction. Thus, we try to categorize these
@@
 
-Tables provide a one-stop-shop overview for the algorithms so that can be a
-helpful tool both as a companion during reading the survey and a reference for
+Tables provide a one-stop-shop overview of the algorithms and serve as a 
+tool both as a companion to the reader while reading the survey and a reference for
 future use.
@@
 terms of \emph{efficiency}, \emph{overhead} and \emph{scalability}; three
-important characteristics for every distributed algorithm. For all we used a
-coarse grained, three level, gradation to denote how well the algorithms do,
-namely \emph{low}, \emph{medium} and \emph{high}. Each category has its own
+important characteristics for every distributed algorithm. For each of these
+criteria, we use a
+coarse-grained, three-level, gradation to denote how well the algorithms do,
+namely \emph{low}, \emph{medium} and \emph{high}. Each criterion has its own
 interpretation of the above literal values. Efficiency means how well the
 algorithm performs its search and topology adaptation functions, without
-effecting other characteristics of the networks (e.g. BitTorrent's optimal
-download performance). To name few, we assume that efficiency is
-negatively effected when an algorithm reduces the search scope. Moreover, the
-minimum-spanning-tree-based techniques have slow convergence speed in topology
-adaptation or the high clustering impacts load balancing and constrains search
-scope (or at least increases search satisfaction time), all reducing overall
-efficiency. With overhead we track the cost to compute and communicate peer
-join, message forwarding and topology maintenance. For example when exchanging
-cost tables only along neighbors the traffic generated can be considered
-trivial. Also duplicates minimization is a plus and lazy or heart-beat-based
-overlay control mechanisms impose different amount of stress to the network.
-Nevertheless, caching/replication helps reducing the overall overhead by
-constraining the query from spreading too much. Low convergence speed of
-minimum-spanning-tree-based approaches, we mentioned above for efficiency,
-results also in additional computational and communicational overhead.
+affecting other characteristics of the networks (e.g. BitTorrent's optimal
+download performance). We assume that efficiency is
+negatively effected when an algorithm reduces the search scope. 
+%Moreover, the
+
+%minimum-spanning-tree-based techniques have slow convergence speed in topology
+
+%adaptation or the high clustering impacts load balancing and constrains search
+
+%scope (or at least increases search satisfaction time), all reducing overall
+
+%efficiency. 
+For the overhead criterion, we track the cost to compute and communicate peer
+join, message forwarding, and topology maintenance. For example, when exchanging
+route tables only amongst neighbors, we consider the traffic generated to
+be low.
+%Also duplicates minimization is a plus and lazy or heart-beat-based
+%overlay control mechanisms impose different amount of stress to the network.
+%Nevertheless, caching/replication helps reducing the overall overhead by
+%constraining the query from spreading too much. Low convergence speed of
+%minimum-spanning-tree-based approaches, we mentioned above for efficiency,
+%results also in additional computational and communicational overhead.
 % TODO: We check
@@
 % TODO: HELP NEEDED!
-With scalability we check the systems behavior with respect to the number of
-participating peers. Low, medium and high scalability can be generally
-considered in the order of hundreds, tenths of thousands and millions,
-respectively. Fully distributed approaches is a plus (no system-centric entities
-involved, no need for global knowledge) and topology adaptation strengthens
-scalability in highly dynamic environments. Reduced search scope, on the other
-hand, penalizes the scalability since the more peers coming in the network the
-more potentially unsatisfied queries will there be. Furthermore, clock
-synchronization is a limiting factor to scalability of a distributed system.
+For the scalability criterion, we consider system behavior with respect to the number of
+participating peers. We consider systems with low, medium, and high scalability 
+to mean they can support 
+hundreds, tenths of thousands, and millions of nodes,
+respectively. 
+%Fully distributed approaches is a plus (no system-centric entities
+
+%involved, no need for global knowledge) and topology adaptation strengthens
+
+%scalability in highly dynamic environments. Reduced search scope, on the other
+
+%hand, penalizes the scalability since the more peers coming in the network the
+
+%more potentially unsatisfied queries will there be. Furthermore, clock
+
+%synchronization is a limiting factor to scalability of a distributed system.
 

### 2014-01-27 conclusion.tex
2012.12.23 (Mema) Text reworked.

@@
 
-Peer-to-peer architecture has been in the center of research attention in the
-last decade. Especially the decentralized unstructured genre exploits the
-advantages of loose coupling and self organization of computing nodes to form
-application-layer networks on top of the physical, best-effort infrastructure of
-the Internet that exhibit interesting properties. Scalability problems arouse
-quickly, though, because of the inefficient construction of this overlay that
-was built with no concern for the underlying physical network that causes a
-great deal of redundant traffic. The problem was identified by the research
-community as the topology mismatch problem between the overlay and the
-corresponding underlying physical network and a great deal of effort has been
-set towards alleviating it. Some fruits of this effort have been gathered and
-presented in this survey. Measurement of link cost through latency or RTT and
-deletion of inefficient established connections are the key concepts of almost
-all approaches. Others, manage to address the problem through hierarchical peer
-clustering (e.g. best IP matching). Additionally, some protocols focus on
-specific problems that furtherly arise, such as overlay partitioning, search
-scope reduction or convergence speed, to name just a few. Another desire is to
-form a protocol that could be applied to both decentralized unstructured and
-decentralized structured networks. Unfortunately no approach has equally
-addressed these problems in order to form a robust solution. So the field seems
-to be, still, fertile for any, new, clever idea.
+Computer systems design is typically about juggling a set of tradeoffs (e.g.,
+choosing higher accuracy at the cost of increased system overhead.)  
+In surveying more than
+a decade's worth of research efforts aimed at solving the topology mismatch
+problem in both unstructured and structured P2P networks, we find this
+tussle-of-tradeoffs property to hold.  With regards to the three criteria 
+we used (efficiency,
+overhead, and scalability), we find that none of the proposed solutions 
+is superior to the others on all three fronts.  This is not surprising.
+Nonetheless, by presenting 1) an
+analysis of each of the solutions, including what distinguishes it from the
+others and its advantages and disadvantages and 
+2) a pictorial
+comparison of how each solution fares compared with the others with regards
+to efficiency, overhead, and scalability, we hope to have provided P2P
+developers and researchers 
+with enough perspective so that depending on the needs of their particular
+application, they are able to select
+amongst these surveyed solutions or propose a new solution of their own.
+
+%Peer-to-peer architecture has been in the center of research attention in the
+%advantages of loose coupling and self organization of computing nodes to form
+%application-layer networks on top of the physical, best-effort infrastructure of
+%the Internet that exhibit interesting properties. Scalability problems arouse
+%quickly, though, because of the inefficient construction of this overlay that
+%was built with no concern for the underlying physical network that causes a
+%great deal of redundant traffic. The problem was identified by the research
+%community as the topology mismatch problem between the overlay and the
+%corresponding underlying physical network and a great deal of effort has been
+%set towards alleviating it. Some fruits of this effort have been gathered and
+%presented in this survey. Measurement of link cost through latency or RTT and
+%deletion of inefficient established connections are the key concepts of almost
+%all approaches. Others, manage to address the problem through hierarchical peer
+%clustering (e.g. best IP matching). Additionally, some protocols focus on
+%specific problems that furtherly arise, such as overlay partitioning, search
+%scope reduction or convergence speed, to name just a few. Another desire is to
+%form a protocol that could be applied to both decentralized unstructured and
+%decentralized structured networks. Unfortunately no approach has equally
+%addressed these problems in order to form a robust solution. So the field seems
+%to be, still, fertile for any, new, clever idea.

### 2014-01-27 intro.tex
2012.12.23 (Mema) Text reworked.

@@
 
-Two ``killer'' applications of the peer-to-peer paradigm are, but not limited
-to, file-sharing and telephony. P2P file-sharing was originally introduced by
+Two ``killer'' applications of the peer-to-peer paradigm are 
+file-sharing and telephony. P2P file-sharing was originally introduced by
 Napster (see Section~\ref{section:background}), which by 2001 was widely
 acknowledged as the ``fastest growing Internet application ever" with 26 million
-users sharing 80 million songs at the height of its popularity. Other similar
+users sharing 80 million songs at the height of its popularity. Other file-sharing 
 applications quickly followed suit including Gnutella and its descendant,
 Limewire, enjoying 3 million concurrently connected peers by the end of 2006
-and BitTorrent, connecting over 150 million users. P2P telephony has seen
+and BitTorrent, connecting over 150 million users. P2P telephony, on the
+other hand, has seen
 explosive growth with the advent of Skype. Since its introduction in 2003,
@@
 Internet.} that share a link in an overlay network which maps to multiple hops
-in the underlying (say IP) topology. If the traffic patterns of the application
+in the underlying IP topology. If the traffic patterns of the application
 are such that traffic flows heavily over the overlay link between the two nodes,
@@
 If the overlay network is constructed such that it does not match the underlying
-topology well, the resulting \emph{topology mismatch} (See
-Section~\ref{section:background} for a more formal definition) can result in two
+topology well, the resulting \emph{topology mismatch} 
+can result in two
 significant problems. First, the performance of the application per se, can be
@@
 redundunt, number of physical hops resulting in poor user experience like
-extensive search latency or jitter. Second, can cumulatively degrade the
-performance of the whole underlying infrastructure, effecting other applications
-as well. For example, studies have shown that the extreme popularity of P2P
-applications amply deployed everywhere nowadays, result in them contributing the
-largest portion of overall Internet traffic
+extensive search latency or jitter. Second, other applications running
+on the underlying network infrastructure are also adversely affected.
+Studies have shown that highly popular P2P applications contribute the largest
+portion of overall Internet traffic
 \cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}, with
-some reporting that more than 60\% of it, is
-P2P-related~\cite{cachelogic,ipoque2007,ipoque2009} and others claiming that it
-is expected to reach almost 8 petabytes per month by the end of
+some studies reporting that more than 60\% of Internet traffic is
+P2P-related~\cite{cachelogic,ipoque2007,ipoque2009} and others predicting that it
+will reach almost
+8 petabytes per month by the end of
 2012~\cite{multinteligence}. This is a significant burden for Internet Service
-Providers (ISPs) who must handle all this traffic to destination nodes sitting
+Providers (ISPs) who must route all of this traffic to destination nodes sitting
 at the edge of the Internet. If the overlay topology of a peer-to-peer system is
-not efficiently constructed, then this burden to the Internet's backbone
-infrastructure could be even greater. For example, traffic might have to flow
+not efficiently constructed, then the burden on the Internet's backbone
+infrastructure could be even greater because traffic might have to flow
 back and forth several times between two neighboring ISPs as it travels from its
@@
 is to help readers sift through the vast amount of literature, to help them
-understand the advantages and disadvantages of each of each work, and to provide
+understand the advantages and disadvantages of each work, and to provide
 them with enough perspective so that when the need arises, they are able to
@@
 define the problem of topology mismatch between an overlay and its underlying
-physical network and elaborate on the rationale behind our coarse grained
+physical network and present the rationale behind our 
 evaluation of the techniques presented throughout this survey. In
@@
 various research efforts for decentralized unstructured and decentralized
-structured peer-to-peer systems, respectively. Finally, with
+structured peer-to-peer systems, respectively. Finally, in 
 Section~\ref{section:conclusion}, we conclude the paper.

### 2014-01-27 mard-survey.tex
2012.12.23 (Mema) Text reworked.

@@
 \begin{abstract}
-Peer-to-peer (P2P) systems is a rapidly growing application architecture. The
-overlay network abstraction on top of a best-effort infrastructure provides a
-powerful tool to realize a plethora of ever-wanted exotic features like
-anonymity, high availability and robustness, load balancing, quality-of-service
-(QoS) at the application layer and many more. Unfortunately weaknesses of early
-effectuations constrained applications from unleashing the full potential of the
-paradigm. One of them is the topology mismatch problem between the overlay and
-the physical underlying network, a problem that imposes a huge amount of
-unnecessary stress to the network resources and, thus, impede applications from
-scaling. In this survey we investigate on the recent research developments
-towards the alleviation of this problem, in both unstructured and structured
-overlay architectures. We visit the various algorithms available in the
-literature, focusing on the presentation of how each of them tries to tackle the
-problem. Additionally, we provide a finer grained categorization of the
-approaches based on the generic type of methodology they use in order to
-accomplish this goal.
-% TODO: Review this after the comparison sections are fixed.
-Finally we discuss the advantages, disadvantages and novelties introduced by the
-algorithms and offer pictorial comparisons for them in order to further
-strengthen and organize understanding of the information and knowledge cited, in
-an attempt to make this survey a one-stop-shop reference for anyone interested
-in the topic, from researchers and students, to system architects and
-developers.
+Peer-to-peer (P2P) systems have enjoyed immense attention and have been
+widely deployed on the Internet for well over a decade.  P2P systems 
+are implemented via
+an overlay network abstraction on top of the best-effort IP infrastructure
+of the Internet. These systems offer a plethora
+of desirable features to distributed applications that are built on top of them
+including anonymity, high availability and robustness, load balancing,
+quality of service, scalability, and others.  Unfortunately, weaknesses of
+early deployments of P2P systems prevented applications from leveraging 
+these features.  One major weakness, identified early on, is the topology
+mismatch problem between the overlay network and the underlying IP topology.
+This mismatch can impose an extraordinary amount of unnecessary stress on
+network resources and can adversely affect the scalability of applications
+built on top of P2P systems.  
+In this paper, we survey over a decade's worth of research efforts
+aimed at alleviating the topology mismatch problem, in both unstructured
+and structured P2P systems.   
+We provide a fine-grained categorization of the
+proposed solutions and discuss the advantages, disadvantages, and novelties
+introduced by each solution.  Finally, we provide an analysis and pictorial 
+comparisons 
+of the proposed solutions in an attempt to provide a quick one-stop-shop
+reference for anyone interested in the topic, from researchers and students,
+to system architects and developers.
+
 \end{abstract}

### 2014-01-27 structured.tex
2012.12.23 (Mema) Text reworked.

@@
-\section{Decentralized Structured Architectures}
+\section{Structured P2P Networks}
 \label{section:structured}
 
-This section focuses on the structured P2P architectures. The following
-paragraphs contain a thorough discussion and analysis of the algorithms proposed
-in the literature for dealing with the topology mismatch problem in such
-architectures. In structured P2P systems, the construction of the routing tables
-in every participating node determines the efficiency of message forwarding.
-Therefore, routing tables that represent the underlying physical structure well
-can achieve increased performance. For this reason, the proposed methods for
-topology mismatch problem basically use different levels of proximity
-information to optimize them towards this direction. The algorithms are here
-categorized based on the previous work done by \cite{CDHR2002,CDCR2002,RSS2002}.
+This section presents a description and analysis of algorithms proposed
+to address the topology mismatch problem in structured P2P networks.
+In structured P2P networks, the construction of the routing tables
+held by participating nodes determines the efficiency of message forwarding.
+Routing tables that closely match the underlying IP topology
+achieve increased performance. As a result, the proposed solutions for
+the topology mismatch problem for structured P2P networks 
+use different levels of proximity
+information to optimize the routing tables used by peers.  
+The categorization of the algorithms we present here is based on the
+categorization suggested by previous 
+work~\cite{CDHR2002,CDCR2002,RSS2002}.
 
@@
 This global view of the
-state is made available to all nodes in order to help them find the best way to
-route their messages. \textit{Global Soft-State} operates in two main stages,
+state is made available to all nodes to help them find the best way to
+route their messages. \textit{Global Soft-State} operates in two phases: 
 generation and use of the proximity information. For generating the proximity
 information a hybrid approach is proposed, which uses landmark clustering as a
-preprocessing step in order to select a number of potential nearest neighbor
+preprocessing step to select a number of potential nearest neighbor
 candidates and then refine the selection by incorporating a
@@
 costly. Additionally, the method does not make any continuing effort to remap
-the overlay structure after a node successfully joins, in order to adapt its
+the overlay structure after a node successfully joins in order to adapt its
 state to any condition change. Although this approach greatly reduces the
@@
 first set of candidates, and while the iteratively closest nodes are detected by
-probing the neighborhood, the candidate neighbor list is updated. In order to
+probing the neighborhood, the candidate neighbor list is updated. To
 avoid a local minima, Mithos probes all the neighbors within a two-hop distance
@@
 
-After finding its first neighbor, the new coming peer is assigned an ID using
+After finding its first neighbor, the newcoming peer is assigned an ID using
 information gathered during the iterative neighbor selection phase. Virtual
-coordinates are assigned to the new coming peer by using the distances of two
+coordinates are assigned to the newcoming peer by using the distances of the two
 closest nodes and their neighbors, so that Euclidean distances between the node
@@
 The last step of the algorithm is the interconnection among neighbors. Mithos
-uses a \emph{quadrant-based} mechanism according to which each node establishes
+uses a \emph{quadrant-based} mechanism in which each node establishes
 a link to the closest neighbor in each quadrant. During forwarding, the next
@@
 using a mechanism based on ideas similar to a perimeter walk\footnote{Used in
-Greedy Perimeter Stateless Routing (GPSR) protocol.} and then improves the
+the Greedy Perimeter Stateless Routing (GPSR) protocol.} and then improves the
 results using parallel path processing by taking into account further geometric
@@
 One major limitation of \textit{Mithos} is that the protocol cannot handle
-dynamic arrivals and departures from the network, as is happens in mobile
-networks.
+dynamic arrivals and departures from the network.
 
@@
 \emph{LAPTOP} \cite{WLH2007}, organizes the overlay into a tree-based hierarchy
-with main focus on reducing hops during message routing as well as minimizing
+with the goal of reducing hops during message routing as well as minimizing
 maintenance overhead. Additionally, a caching scheme is also incorporated so as
 to further reduce routing table update costs. The authors theoretically show
-that in LAPTOP routing path length is bounded by $O(log_d N)$ and node joining
+that in LAPTOP, routing path length is bounded by $O(log_d N)$ and node joining
 and leaving in the overlay network is bounded by $O\left( d log_d N \right)$
@@
 $1$ to $d$, where $d$ is the maximum degree of the nodes. The assignment process
-is done by appending a unique octet to the address of each nodes parent, while
-root node is assigned address $1$.
+is done by appending a unique octet to the address of each node's parent, while
+the root node is assigned address $1$.
 
@@
 arrive to its target. During tree traversals, special entries in the routing
-tables, called \emph{routing cache}, are maintained in order to increase routing
+tables, called \emph{routing cache}, are maintained to increase routing
 efficiency and achieve finer load balance. Caching enables a node to forward a
@@
 a larger, quicker and more cost effective step through the overlay and toward
-the destination. To improve scalability, the number of children nodes and the
+the destination. To improve scalability, the number of child nodes and the
 size of the routing cache are limited. In terms of overlay maintenance, LAPTOP
@@
 \emph{Proximity in Kademlia} \cite{KLKP2008} introduces proximity
-controls over the vanilla Kademlia protocol's random ID assignment, in order to
+controls over the vanilla Kademlia protocol's random ID assignment to
 optimize the underlying network usage of the protocol. The work in this paper
-can also be considered, as a more generic study over building proximity in the
-context of iterative lookup algorithms. Focus is given on both the overlay and
-the underlaying tiers in the sense that in the first improves the routing
-performance according to a cost metric that is provided by the later and
-quantifies the suitability of establishing a communication link between peers as
+can also be considered as a more generic study of incorporating proximity into
+iterative lookup algorithms. 
+The overlay tier improves its routing
+performance using a cost metric that is provided by the underlying tier
+and quantifies the suitability of establishing a communication link between peers as
 a cost function, based on the used proximity criteria (RTT, ISP locality, etc.).
@@
 %\end{itemize}
-The paper proposes two approaches for overlay optimization. One falling into
-{\em proximity routing} and the other on {\em proximity neighbor selection}
-approaches overlay optimization techniques for exploiting proximity. For the
-first, the goal is in choosing the best next hop during routing a message. Due
+
+The paper proposes two approaches for overlay optimization: 
+{\em proximity routing} and {\em proximity neighbor selection}.
+In proximity routing, the goal
+is to choose the best next hop during routing a message. Due
 to the fact that the routing in Kademlia is iterative, it is the initiator of
@@
 Kademlia always chooses the closest node with respect to the XOR metric but with
-proximity routing selection, Kademlia chooses the one with the smallest underlay
+proximity routing selection, Kademlia chooses the node with the smallest underlay
 metric cost. Thus, as in all such approaches, there is a trade-off between the
@@
 on its region, country, or ISP location, and use this information to form
-clusters of nearby peers. An alternative method could also be the use of Vivaldi
-coordinates \cite{cox_vivaldi_2004}. However, authors report that clustering
-approach performs better than Vivaldi protocol based on their experiments.
+clusters of nearby peers. An alternative method could use Vivaldi
+coordinates \cite{cox_vivaldi_2004}. However, the authors report that their clustering
+approach performs better than the Vivaldi protocol based on their experiments.
 Authors also report that {\em Proximity routing} protocol successfully worked
-with the Kademlia protocol and improved the locality of the connections over
+with the Kademlia protocol and improved the locality of the connections between
 peers.
 
-As already mentioned the paper proposes \emph{proximity neighbor selection}
-method, as well, for overlay optimization. The goal, according to this, is to
+In \emph{proximity neighbor selection}, the goal is to
 keep peers with the least contact cost in the routing table. As Kademlia
-constantly learns new peers (incoming requests, through iterative lookup) no
-special algorithm for searching more cost effective peers is necessary and
-simply choose the best peers seen so far.
+continuously learns about new peers via incoming requests and an iterative
+lookup process, no 
+special algorithm for searching more cost-effective peers is necessary and
+it simply chooses the best peers seen so far.
 
@@
 T2MC \cite{SLCGZ2008} uses traceroute logs to detect the AS boundaries and
-cluster close by peers with each other to reduce the redundant multiple message
+cluster nearby peers together to reduce the redundant multiple message
 passes between the AS boundaries. T2MC uses a customized k-mean classification
@@
 provides detailed information about the network structure, use of traceroute
-creates overhead to the overall network structure. For this same reason, it is
+creates overhead on the overall network structure. For this same reason, it is
 not unusual for network administrators to disable this support on their network
@@
 \subsubsection{PChord}
-\emph{PChord}\cite{HLYW2005} is a based on the Chord DHT which adds
+\emph{PChord}\cite{HLYW2005} is based on the Chord DHT and adds
 proximity routing into its routing mechanism. The main modification over the
 vanilla Chord is the inclusion of a \emph{proximity list} into its routing table
-so that the next hop is decided based not only by considering best progress
-towards the key, but on the physical proximity of the candidate nodes as well.
+so that the next hop is decided based on both the best progress
+towards the key and the physical proximity of the candidate nodes as well.
 The list is, at join time, empty but as the PChord node starts to interact with
-other nodes it applies a heuristic mechanism to fill up the list. Entries are
+other nodes it applies a heuristic mechanism to fill the list. Entries are
 dynamically added or removed as the network state changes. The overlay
@@
 heartbeat of nodes inside the proximity list but since this list always contains
-nodes with the lowest RTT the added cost is kept to a minimal.
+nodes with the lowest RTT the added cost is kept to a minimum.
 
@@
 Chord, since each hop in PChord is larger or at least equal in the key space.
-Passing through proximity links in the underlaying network means reduced routing
+Passing through proximity links in the underlying network means reduced routing
 cost. Additionally, if the number of entries in the proximity list is the same
 as the number of network partitions, PChord prevents hops from jumping in and
-out the network partition the current node belongs to.
+out of the network partition the current node belongs to.
 
@@
 \subsubsection{AChord}
-\emph{AChord} \cite{DK2006} uses IPv6's any-cast mechanism in order to
+\emph{AChord} \cite{DK2006} uses IPv6's anycast mechanism to
 relieve the protocol from complex joining procedures, and achieve high accuracy
-of network proximity giving high routing efficiency, Any-cast delivers a message
-coming from the outside of an any-cast group to the physically closest node in
+of network proximity and thus, high routing efficiency. Anycast delivers a message
+coming from the outside of an anycast group to the physically closest node in
 that group. AChord organizes all nodes participating in the overlay network into
-an any-cast group. Any joining node comes outside of that group and, thus, is
+an anycast group. Any joining node comes outside of that group and, thus, is
 automatically forwarded to the physically nearest node in order to bootstrap,
 avoiding the need to, explicitly, maintain such nodes, and the effort of finding
-a way of locating the physically nearest among them. The ID of the new coming
+a way of locating the physically nearest among them. The ID of the newcoming
 node is computed based on the ID of the bootstrap node and the bootstrap's
 predecessor in a way that the constructed ID will be positioned between the
-aforementioned two. After joining, the finger table is build the same way as in
+aforementioned two. After joining, the finger table is built the same way as in
 the Chord protocol. Moreover, additional nodes are maintained in a structure
@@
 table. The authors claim this to be a lightweight approach that requires few
-changes to Chord in order not to affecting its own advantageous characteristics.
+changes to Chord while keeping its advantageous characteristics.
 
@@
 medium &
-% assumes the operation of ideal underlying any-cast mechanism (meaning that
+% assumes the operation of ideal underlying anycast mechanism (meaning that
 % outside messages are always routed to the nearest node which in the current
@@
 \emph{Chord6} \cite{XZHL2005} is another Chord variant that tries to
-exploit the hierarchical features of IPv6 in order to create a substrate that
+exploit the hierarchical features of IPv6 to create a substrate that
 reduces inter-domain traffic between service providers. The difference between
@@
 node's IPv6 address prefix of specific length, while the remaining lower bits
-are the hash value of the rest of that IPv6 address. As a result of this
+are the hash value of the rest of the IPv6 address. As a result of this
 assignment, nodes in a domain will be mapped onto a continuous key space on the
@@
 through the special peer CN. The node can, additionally, ask CN for other
-nodes that have previously joined the region in order to add them into its
+nodes that have previously joined the region to add them into its
 neighbor set for future routing table optimization. Even in the case when no
@@
 \cite{freedman_iploc_2005} states that $97\%$ of prefixes larger than $24$ bits
-belong to a single geographical location. However, using less number of bits
-creates less precise results and more number of bits increase the burden on the
-network and reduce the possible number of neighbors. Therefore, a careful
+belong to a single geographical location. However, using a smaller number of bits
+creates less precise results and a larger number of bits increases the burden on the
+network and reduces the possible number of neighbors. Therefore, a careful
 selection is required in terms of performance/accuracy trade-off. IP-based
-Clustering generates a key and its prefix and stores the later in the DHT
+clustering generates a key and its prefix and stores the prefix in the DHT
 itself, so that any newly joining nodes with the same IP prefix can query it and
-identify all the neighbors with the same prefix relatively easy. Nodes
+identify all the neighbors with the same prefix easily.  Nodes
 periodically update their entries in the DHT by removing those who leave, or
-those who has been timed out with no further activity detected from the node.
+those who have been timed out with no further activity detected from the node.
 
@@
 \subsubsection{Cone}
-\emph{Cone} \cite{HY2007} extends Chord using proximity neighbor
+\emph{Cone} \cite{HY2007} extends Chord using a proximity neighbor
 selection topology optimization algorithm. The proximity information is
@@
 identifier, denoted as $Id_{Chord}$, is the same as in Chord. The second is the
-Cone-layer identifier, $Id_{Cone}$ which is constructed by two component
+Cone-layer identifier, $Id_{Cone}$ which is constructed using two component
 identifiers. The first, known as \emph{group identifier (gid)} denotes a
-relevant group the node belongs to while the second, namely \emph{local
+relevant group to which the node belongs while the second, namely \emph{local
 identifier (lid)} indicates the local identifier within the group. The group
-concept, which is introduced here, is a way of dividing nodes according to a
+concept is a way of dividing nodes according to a
 common $Id_{Chord}$ prefix.  The structure of a Cone overlay, retains the
-Chord's circular topology. The difference lies on the fact that, now, two rings
+Chord's circular topology. The difference lies in the fact that, now, two rings
 are created. A big ring, where nodes with the same \emph{gid} are arranged at
 each position. Each of these positions are a smaller ring for the particular
-group's \emph{lid}s. The routing is achieved in both clockwise and
+group's \emph{lid}s. Routing is achieved in both clockwise and
 counter-clockwise directions in the big-ring, for which two routing tables are
@@
 \emph{DynaMO} \cite{WZS2004} tries to build an overlay that not only 
-considers physical proximity among peers but their mobility attributes as well.
-DynaMO is based on a Pastry overlay network and this was a pretty conscious
-choice, by the authors, basically because of Pastry's built-in locality
-heuristics which are thoroughly analyzed in the literature \cite{CDCR2002a}
-provide a good background against which to compare their own results. In
-order to adapt to mobile, ad-hoc networks, though, special care has been put to
-the maintenance of an even overlay ID distribution so that hot-spots are
-avoided. DynaMO dictates a new coming node to gather information concerning its
+considers physical proximity amongst peers but their mobility attributes as well.
+DynaMO is based on the Pastry overlay
+because of Pastry's built-in locality
+heuristics which are thoroughly analyzed in the literature \cite{CDCR2002a}.  
+To adapt to mobile, ad-hoc networks, though, special care has been put in 
+the maintenance of an even overlay ID distribution so that hotspots are
+avoided. DynaMO requires a newcoming node to gather information concerning its
 physical neighborhood and uses it to assign itself an appropriate overlay ID.
 DynaMO tries to capitalize on the fact that in Pastry each node's table consists
-of rows equal to the number of digits of the overlay IDs and columns equal to
+of rows equal to the number of digits in the overlay ID and columns equal to
 the ID's base. As we go down the table rows, the matching prefix between the
@@
 contains the closest nodes in the ID space. Additionally, as the prefix match
-increases by one, the result is exponentially less candidates that can fill the
+increases by one, the result is exponentially fewer candidates that can fill the
 tables entries as the row increases. This leads to the observation
 \cite{antony_pastry_2001,CDCR2002a} that from overlay routing hop to overlay
-routing hop the physical distance between nodes is likely to increase leading to
+routing hop, the physical distance between nodes is likely to increase leading to
 a dominating last routing step in terms of the overall physical routing path
@@
 
-RLM lookup mechanism to locate nodes responsible for a fixed set of carefully
+The RLM lookup mechanism is used to locate nodes responsible for a fixed set of carefully
 chosen (in a way that the ID space is divided into equal portions) landmark
-keys. This means that when a node is assigned a landmark key then it plays the
+keys. This means that when a node is assigned a landmark key, it plays the
 role of a temporary landmark for the network. Any joining node will measure its
@@
 of a prefix taken from the closest landmark and a remainder that could be
-assigned randomly or generated by mechanism that takes into account physical
+assigned randomly or generated by a mechanism that takes into account physical
 neighborhood. The length of the prefix can be determined as
@@
 they can instantly be replaced by new substitutes, that share similar physical
-attributes to the failed and thus preserving the balance of the network.
+attributes with the failed and thus preserve the balance of the network.
 
@@
 is not acceptable. CPNA on the other hand takes advantage of Pastry's
-specification that a new coming node is boot-strapped by a physically close
+specification that a newcoming node is bootstrapped by a physically close
 node.
-The new comer then assumes the ID prefix of that neighbor while the rest is
+The newcomer then assumes the ID prefix of that neighbor while the rest of the ID is 
 generated similarly to RLM. Unfortunately, less overhead comes at the expense
@@
 
-\emph{SAT-Match} \cite{RGJZ2004}, similarly to the other approaches in proximity
-neighbor selection, uses probing to detect close by peers and obtain proximity
+\emph{SAT-Match} \cite{RGJZ2004}, similar to the other approaches in proximity
+neighbor selection, uses probing to detect nearby peers and obtain proximity
 information. However, as an additional feature, SAT-Match peers can do selective
@@
 latency while the $\bar{L_p}$ is the average physical link latency, and uses it
-a way quantify the \emph{topology match degree} of the constructed overlay. For
+a way to quantify the \emph{topology match degree} of the constructed overlay. For
 the probing phase SAT-Match uses a small TTL value for the probing messages in
@@
 distance to the source node. Blindly selecting the peer with the smallest RTT
-as neighbor is, generally, not the optimal decision to make in order to achieve
+as neighbor is, generally, not the optimal decision to make to achieve
 global stretch reduction. In a structured scheme, when a node jumps to connect
@@
 maintain the structure's integrity, thus creating an overall increase in the
-overlay's stretch. The two nodes with the smallest RTT are then used in order to
+overlay's stretch. The two nodes with the smallest RTT are then used to
 select one zone to jump in this phase. The algorithm is as follows: The source
@@
 determined by the integrity of the global data structure. The main advantage of
-the structured P2P topology is that by the help of the global data structure,
+the structured P2P topology is that with the help of the global data structure,
 peers or their data can be found within the network even if there is only a
@@
 maintenance overhead for the network due to updates required by the global data
-structure, and for networks with frequent node arrivals and departures the
-topology uses valuable network resources just to update the global structure.
+structure, and for networks with frequent node arrivals and departures, the
+overlay uses valuable network resources just to update the global structure.
 Nodes join the network by using a key value, which determines the location and
@@
 discussion on unstructured systems in Section~\ref{section:unstructured} so
-we will not elaborate more here. Again, we additionally gather highlights and a
-rough estimation of each algorithm's positive and negative aspects regarding
-their implementation.
+we will not elaborate more here. As in the previous section, we provide the 
+key properties of each algorithm that differentiate it from the others and
+highlight 
+each algorithm's advantages and disadvantages.
 
@@
 
-Inside geographic layout category fall those kind of methods which practically
-position physically close by nodes together in the application space as well
+In the geographic layout category fall methods which practically
+position physically nearby nodes together in the application space as well
 (check Figure~\ref{figure:geographic-layout}). This is done by adjusting and
-maintaining the routing tables of all involved peers accordingly by exploiting
+maintaining the routing tables of all involved peers by exploiting
 proximity information that describe the overall geographic positioning of the
 peer nodes. Landmark servers and RTT measurements are two popular methods, which
-can also be used in conjunction, to discover physically close by peers over the
+can also be used in conjunction, to discover physically nearby peers over the
 network. However, as discussed in Section~\ref{sec:landmark_binning}, these
@@
 nodes improves the query efficiency of the system, on the other hand, it tends
-to create hot-spots, and the needed failure resilience is undermined by the fact
-that close by nodes are more likely to suffer collective failures.
+to create hotspots, and the needed failure resilience is undermined by the fact
+that neaerby nodes are more likely to suffer collective failures.
 
@@
   \includegraphics[scale=0.4]{img/pdf/proximity-routing.pdf}
-\caption{Proximity routing takes proximity info into account during the
+\caption{Proximity routing takes proximity information into account during the
 routing process.}
@@
 knowledge about network proximity. On the other hand it exploits such knowledge
-in order to choose the best next hop during routing a message as can be seen in
+to choose the best next hop when routing a message as can be seen in
 Figure~\ref{figure:proximity-routing}. This approach tries to balance between
@@
 and choosing the closest entry in the routing table, in terms of network
-proximity. As this is very difficult to be achieved, the system might end up
-using a greedy approach that selects low latency paths over the overlay, which
+proximity. As this is very difficult to achieve, the system might end up
+using a greedy approach that selects low latency paths on the overlay, which
 maps to suboptimal longer paths on the physical topology. Thus, in general,
@@
 applied to CAN(-like) implementations. Moreover, the technique has been
-incorporated into a version of Chord causing an increase on the overhead of node
+incorporated into a version of Chord causing an increase in the overhead of node
 joins and the size as well as maintenance cost of finger tables.
@@
 Here time-to-live values between nodes, or directly node ID prefixes are used to
-detect proximity. Tapestry, Pastry, and CAN successfully implemented the
-proximity into their algorithms by using this approach. The routing protocol in
+detect proximity. Tapestry, Pastry, and CAN successfully incorporated 
+proximity neighbor selection into their algorithms. The routing protocol in
 Pastry is based on longest node ID prefix matching, while CAN uses RTT values to
-detect close by nodes. \cite{CDCR2002a} reports that proximity neighbor
-selection as an effective proximity based method.
+detect nearby nodes. \cite{CDCR2002a} reports that proximity neighbor
+selection is an effective proximity based method.

### 2014-01-27 unstructured.tex
2012.12.23 (Mema) Text reworked.

@@
-\section{Decentralized Unstructured Architectures}
+\section{Unstructured P2P Networks}
 \label{section:unstructured}
 
-This section is dedicated to the unstructured P2P architectures. In the next
-paragraphs, algorithms for tackling the topology mismatch problem on such
-networks are discussed, analyzed and categorized based on the use of the
-overlay structure, message forwarding approach for peer communication and
+This section is dedicated to unstructured P2P networks. Algorithms
+for tackling the topology mismatch problem in such
+networks are discussed, analyzed, and categorized based on the use of the
+overlay structure, the message forwarding approach for peer communication, and
 techniques proposed for detecting proximity in order to optimize overlay
@@
 
-The directed BFS%, as shown in Figure~\ref{figure:dbfs:dbfs},
-tries to avoid this delay by forwarding the query messages only to a selected
+The directed BFS
+tries to avoid this delay by forwarding the query messages only to a select
 set of neighbors, in which the selection criteria varies from the number of
@@
 
-In local indices%, as shown in Figure~\ref{figure:dbfs:localindx},
+In local indices
 each node
@@
 efficiency, however, updating the indices in cases with frequent node joins and
-leaves introduces a serious overhead to the system if the radius is kept broad.
+leaves introduces a serious overhead into the system if the radius is kept broad.
 
@@
 factor} in the paper's context. Request messages are then forwarded only to
-nodes whose delay is less than or equal that boundary. With the clustered
-routing tables and the loose organization of DAPS' overlay network it is
-considered by is between structured and unstructured.
+nodes whose delay is less than or equal that boundary. 
+%With the clustered
+%routing tables and the loose organization of DAPS' overlay network it is
+%considered by is between structured and unstructured.
 
@@
 \subsubsection{Distributed Cycle Minimization Protocol (DCMP)}
-We have already discussed that blind flooding mechanism broadcasts messages with
-no real targeting, many times to a direction far from the path that can
+We have already discussed that blind flooding often causes messages to be routed in a direction
+that is far from the path that can
 ultimately fulfill the query request. Another problem with topology unaware
@@
 message can travel is limited by the TTL value, which is practically $7$ in most
-cases DCMP cannot detect cycles formed by more than 7 peers. Even though cycle
+cases, DCMP cannot detect cycles formed by more than 7 peers. Even though cycle
 elimination improves the network performance, it does not actually solve the
@@
 each item increases in the network, it would be easier for a search algorithm,
-even a random one, to locate these items. In order to analyze the feasibility of
+even a random one, to locate these items. To analyze the feasibility of
 such an approach, the authors investigate three different replication
@@
 allocation is a strategy proposed by Cohen and Shenker, which is a model between
-the uniform and proportional allocation.  In order to evaluate the outcome of
+the uniform and proportional allocation.  To evaluate the outcome of
 the replication models, the overall costs of successful and unsuccessful
@@
 unsuccessful search. The proportional model, on the other hand, promotes the
-more frequently queried items by replicating them more, therefore decreases the
+more frequently queried items by replicating them more and therefore decreases the
 search time for popular items, but suffers when needing to locate the rare
@@
 same for uniform and proportional models, and any approach between them would
-always behave much better. Therefore they propose the square root allocation
+always behave much better. Finally, they propose the square root allocation
 approach, which is a replication strategy that minimizes the expected search
@@
 % }\qquad\qquad
-% \subfigure[Peer A sends a message to the rest using end system multi-cast.] {
+% \subfigure[Peer A sends a message to the rest using end system multicast.] {
 %   \includegraphics[scale=0.4]{img/algorithms/narada3}
@@
 protocol for creating self-adapting overlay networks, that can achieve
-application-layer multi-cast communication without requiring IP multi-cast
+application-layer multicast communication without requiring IP multicast
 infrastructure at the network layer\footnote{
-  IP multi-cast is the term referring to the method of sending IP datagrams to a
+  IP multicast is the term referring to the method of sending IP datagrams to a
   group of receivers in a single transmission. Even though the method is
-  available some years now, IP multi-cast has not taken off as anticipated since
+  available some years now, IP multicast has not taken off as anticipated since
   many believe that it violates the stateless design of the current Internet.
-  For this reason it is widely deployed in contained environments like
+  For this reason, it is widely deployed in contained environments like
   enterprises (i.e. IPTV applications - distance learning, video conferencing),
   commercial stock exchanges, and multimedia content delivery networks.
-}. Although it was not originally designed as a P2P system nor can be considered
-scalable in the contemporary sense of the term, Narada has been pioneering
-(maybe along with \emph{Scattercast} \cite{C2000}) in that
+}. Although it was not originally designed as a P2P system, Narada is
+a pioneer in that 
+%(and the accompanying \emph{Scattercast} \cite{C2000})
 it was the first to consider the feasibility of overlay-based, application
@@
 authors realized the inefficiency caused by the topology mismatch problem and
-tried to encounter it by building a richer connected graph, called a
+tried to address it by building a richer connected graph, called a
 \emph{mesh}, and building per source minimum spanning trees\footnote{
@@
 dynamically updated while nodes continue to join and leave the network.
-Moreover, the protocol tries to ease the physical link stress, the overall
+The protocol aims to ease the physical link stress, the overall
 resource usage and the relative delay among end systems. Unfortunately, the most
@@
 \emph{Selective Flooding} and \emph{Active Topology} to optimize the usage of
-the underlying physical resources. In \emph{Selective Flooding} a minimum
+the underlying physical resources. In \emph{Selective Flooding}, a minimum
 spanning tree is built among each peer and its immediate neighbors so that
@@
 need to be added to the Gnutella protocol). The randomized AT algorithm is
-applied by the source peer everytime the neighbour list is updated or an updated
+applied by the source peer every time the neighbour list is updated or an updated
 neighbour cost table is received.
@@
 % there is a change of neighbors, the source peer has to re-calculate the
-% multi-cast tree and apply the randomized AT algorithm.
+% multicast tree and apply the randomized AT algorithm.
 high &
@@
 among each peer and its one-hop neighbors is built. Ultimately, physically
-far away neighbors are replaced by physically close ones. Specifically, a peer
-$S$ probes the distance between one of its non-flooding neighbor's neighbor,
+far away neighbors are replaced by physically close neighbors. Specifically, a peer
+$S$ probes the distance between one of its non-flooding neighbor's neighbors,
 say $G$ and $H$ respectively. Assuming that the neighbor's neighbor ($SH$) is
-measured as being ``closer'' that the neighbor ($SG$), the link to the later
+measured as being ``closer'' than the neighbor ($SG$), the link to the later
 is cut and a new connection is established. In the opposite situation where
-distance $SG$ is closer than $SH$ then if $SH$ is shorter that $GH$ then $H$ is
+distance $SG$ is closer than $SH$ then if $SH$ is shorter than $GH$, then $H$ is
 preserved as $S$'s new neighbor. In the case where $SH$ is larger than $SG$ and
@@
 them and calculating communication costs. Calculation is achieved through
-time-stamp checking, so clocks must be synchronized. The latency information
+timestamp checking, so clocks must be synchronized. The latency information
 gathered is later used to evolve the overlay network into a more efficient one,
 without reducing the search scope. Each node examines the latency information
-among his direct neighbors and those with larger latency are put in to a
-will-cut list where they remain for a certain period of time until when they are
-ultimately cut and recorded to the peer's cut list. Thus, low productive
+among his direct neighbors and those with larger latency are placed on a
+will-cut list where they remain for a certain period of time after which they are
+cut and placed on the peer's cut list. Thus, low productive
 connections are dropped and replaced by more efficient ones, reducing the
@@
 color, say red or white (separating all peers into two groups). Then the peer
-that plays the role of the boot-strap host provides the joining peer with a list
+that plays the role of the bootstrap host provides the joining peer with a list
 of active peers along with their color information. The later uses this list to
@@
 neighbors. The red peers, equipped with the information of all two-hop
-neighbors ($N^2$), creates minimum spanning tree of these neighbors and
+neighbors ($N^2$), create minimum spanning tree of these neighbors and
 assigns efficient forwarding paths. Having a minimum spanning tree with two
@@
 tries to minimize the overlay hop costs. THANCS is considered a \emph{local
-search method}, in the sense that it targets in finding a locally optimum
-solution, by exploiting knowledge within a 2-hop radius. The algorithm consists
+search method} because it aims to find a locally optimum
+solution by exploiting knowledge within a 2-hop radius. The algorithm consists
 of two main components: \emph{piggybacking neighbor distance on queries} and
-\emph{neighbor comparison and selection} which are further discussed bellow.
-The first component dictates peers to probe distances (using network delay
+\emph{neighbor comparison and selection} which are further discussed below.
+The first component requires peers to probe distances (using network delay
 distance measuring) with its immediate neighbors and stores information
@@
 neighbor of $n$, $p$ chooses one of the following two paths. First, if $q$ is a
-direct neighbor of $p$, then the later will check the cost of the involved
+direct neighbor of $p$, then the latter will check the cost of the involved
 links. If the most cost-intensive link is one of $pq$ or $pn$, then it is put
@@
 $q$ is within two-hop radius from $p$, the later probes the former (if it hasn't
-got the distance information yet) and again compares distances $pq$, $pn$ and
+got the distance information yet) and again compares distances $pq$, $pn$, and
 $nq$. If $pq$ is the most cost-intensive, $p$ will not establish the connection.
@@
 \subsubsection{Adaptive Peer Selection (APS)}
-Bernstein et al., contributed on the use of machine learning for building peer
+Bernstein et al., use machine learning techniques for building peer
 selection strategies from past experience \cite{BFLZ2003}. A decision tree is
@@
 collected. These can be load, bandwidth, and past uploading experience. Then a
-by Markov decision process is incorporated as a mechanism to shape the policy
-for switching among the peers. The problem with this approach is that the peer
+Markov decision process is incorporated as a mechanism to shape the policy
+for switching among the peers. Although novel, the downside of this approach is that the peer
 selection algorithm is slow due to the learning process and the complexity of
@@
 construction of unstructured P2P networks. The paper's proposition is twofold.
-Overlay construction and search method. For the overlay construction it employs
+Overlay construction and search method. For the overlay construction, it employs
 the notion of \emph{short} and \emph{long} connections. Assuming $N$ is the
 number of all network participants, $\alpha \leq 1 $ a system-wide magic number
-and $x$ an $\alpha$-related latency threshold, the boot-strapping peer randomly
-selects $\alpha \ast N$ close (latency bellow $x$) and
+and $x$ an $\alpha$-related latency threshold, the bootstrapping peer randomly
+selects $\alpha \ast N$ close (latency below $x$) and
 $\left( 1 - \alpha \right) \ast N$ distant (latency above $x$) nodes as its
@@
 flood to its distant neighborhood with $TTL = 1$. Then, peers that receive a
-query over a long link, start a local flood with $TTL = ttl$, were $ttl$ is a
+query over a long link start a local flood with $TTL = ttl$, where $ttl$ is a
 system defined parameter.
 
-The goal of the overlay construction is to create a network that exposes low
+The goal of the overlay construction is to create a network that exhibits low
 clustering, a beneficial characteristic of random graphs which can lead to a
 larger coverage of peers with the same number of messages and reduced
-duplication. This means more efficient information lookup that exposes low or no
+duplication. This means more efficient information lookup that has minimal or no
 negative impact whatsoever to the rest of the mechanisms employed in
@@
 peer, randomly connects to an already participating node through a bootstrap
-server. Being connected, it starts receiving information throughout the
+server. When connected, it starts receiving information throughout the
 link-state mechanism and thus, after some time, it has a complete picture of the
-overlay graph. Then it estimates the delay to all other nodes in order to
+overlay graph. Then it estimates the delay to all other nodes to
 determine its potential neighbors and ultimately connect to the overlay using
 some policy\footnote{For example, minimization of the average delay to all its
-neighbors.}. It is obvious that there is extensive resource usage for updating
-the wiring of the all nodes in the system. The authors claim that they can
+neighbors.}. EGOIST requires extensive resource usage to continuously update the
+wiring of all the nodes in the system. The authors claim that they can
 reduce the load imposed by the monitoring and announcing process from $O(n^2)$
@@
 peer at hand. Consequently, the remaining $35 - k$ peers are from the same ISP.
-These $k$ nodes are used in order to preserve a more extended, global view of a
-network, but in the same localize the load within the limits of a single ISP.
+These $k$ nodes are used to provide a more extended, global view of a
+network, but balance the load within the limits of a single ISP.
 This prevents clients from exchanging traffic through an expensive transit link,
-if there are alternative local connections that could offer the same, faster and
+if there are alternative local connections that could offer the same service faster and
 at no additional cost for the ISP. Implementation can either be done through
-modifications on the tracker side, to identify ISP locality, or through P2P
-traffic shaping devices installed on ISP's edge routers. The paper also proposes
-combination of biased neighbor selection along with bandwidth throttling and
-some caching approach for near optimal results. Unfortunately, implementation
-need, in some extend, contribution (exposure of ISP Autonomous System mappings)
+modifications on the tracker side to identify ISP locality, or through P2P
+traffic shaping devices installed on the ISP's edge routers. The paper also proposes
+a combination of biased neighbor selection along with bandwidth throttling and
+some caching techniques for near optimal results. Unfortunately, implementation
+and deployment
+need knowledge of ISP Autonomous System mappings
 or infrastructure changes (installation of P2P traffic shaping devices) on the
-ISP level itself, making the adoption difficult in the general case.
+ISP, making widespread adoption.
 
@@
 Choffnes and Bustamante propose Ono \cite{CB2008}, a protocol for managing
-BitTorrent traffic so that it significantly reduces cross-ISP traffic and in the
+BitTorrent traffic so that it significantly reduces cross-ISP traffic and at the
 same time enhances downloading rates by favoring connections within the borders
@@
 \cite{BCCMSBZ2006}, this work can lead to improved performance without requiring
-any cooperation between ISPs subscribers, or requiring any additional
+any cooperation between ISP subscribers, or requiring any additional
 infrastructure and network topology information. The selection approach is
@@
 (network measurement systems) information for their replica selection, so the
-authors claim that peers which exhibit similar redirection behavior, are very
-likely they are close to the replica servers and as a consequence to each other.
+authors claim that peers which exhibit similar redirection behavior, are 
+likely close to the replica servers and as a consequence close to each other.
 The redirection behavior is modeled in terms of \emph{ratio map} in the
 article's parlance, a vector of <replica-server,ratio> tuples, where ratio is
-the percentage of times CDN redirects the peer to the specific replica-server.
+the percentage of times the CDN redirects the peer to the specific replica-server.
 The bootstrapping phase consists of the peer performing DNS lookup to CDN names
-in order to build its redirection information. The interval for polling DNSs
+to build its redirection information. The interval for polling DNSs
 starts from 30 seconds and increases by one minute every time redirection
 information to the CDN is found to remain unchanged after a lookup. On the other
-hand, the interval is halved whenever is found to have been changed. In order
-not to avoid the bootstrapping phase if the ratio map is sufficiently fresh, the
+hand, the interval is halved whenever is found to have been changed.  
+To avoid the bootstrapping phase when the ratio map is sufficiently fresh, the
 protocol persists the information after the end of a BitTorrent session.
@@
 piece is calculated as the mean value of the distances of the peers that posses
-it. The paper conducts several measurements on the efficiency of the above
-scheme to reach the conclusion that the major design decision for all locality
+it. The paper conducts several measurements on the efficiency of this scheme 
+and reaches 
+the conclusion that the major design decision for all locality
 aware P2P systems is the trade-off of optimizing inter- Autonomous System
-traffic, which their solution achieves, and of the fairness among peers, a
-section that their locality-based approach does not do well in comparison to the
+traffic versus achieving fairness amongst peers.  While their
+locality-based approach achieves traffic optimization, it does not do well
+when compared with the 
 random approach of the standard BitTorrent protocol.
@@
 form a comprehensive metric to select peers with high download rate, low
-reciprocal upload demand and low routing hops. The TopBT metric is applied on
+reciprocal upload demand and low routing hops. The TopBT metric is applied to
 several parts of the original BitTorrent protocol in order to inject topology
-awareness for better peer selection; more specifically on the peer list the
-tracker returns when someone first tries to connect to the swarm, on the initial
-connection establishment on the connection replacement and of course the
+awareness for better peer selection. Specifically, the parts of the protocol
+that are modified include:  the peer list the
+tracker returns when someone first tries to connect to the swarm, the initial
+connection establishment, and the
 unchoking mechanism.
@@
 UTAPS \cite{LCY2008} is another peer selection strategy for the BitTorrent
-protocol. Its first step is to collect the needed information in order to
+protocol. Its first step is to collect the needed information to
 construct a picture of the underlying topology and the second is to make the
@@
 algorithm uses network tomography, a technique which suggests probing only
-from a large network's end points in order to infer its internal characteristics
+from a large network's endpoints to infer its internal characteristics
 \cite{chny_tomography_2002}. Upon a new peer arrival, the tracker trace-routes
@@
 e.t.c). The more the peers in the swarm the better picture a tracker has for the
-underlying infrastructure, which can then return to the new coming peer in the
-form of the bootstrapping peer list. The peers further enhance this picture by
+underlying infrastructure.  The tracker provides this information to the newcomer
+in the form
+of the bootstrapping peer list. The peers further enhance this picture by
 trace-routing the returned peers themselves and reporting back to the tracker.
@@
 reduce the network delay among their connections. This ``collaboration'', is
-what differentiates this approach from others by letting two peers to optimize
+what differentiates this approach from others by allowing two peers to optimize
 their neighborhood environment, than simply letting each node to ``selfishly''
@@
 % Exchanging neighbors within a radius of 2 (TTL value) reduces stretch
-% significantly, while keeping the additional overhead at very reasonable levels
+% significantly, while keeping the additional overhead at reasonable levels
 low &
@@
 domain names. The result is a flat overlay topology but with some changes it
-can be utilized in super-peer architectures as well. According to the algorithm
+can be utilized in super-peer architectures as well. According to the algorithm,
 half of the possible connections of a node is used to connect to local peers
@@
 the later secure the connectivity of the structure and prevent partitioning.
-Upon arrival the newcomer asks a \emph{host-cache} in order to establish its
-random connections. For establishing its sibling connections, the new comer
-multi-casts a special walker message in order to help it locate the right
+Upon arrival the newcomer asks a \emph{host-cache} to establish its
+random connections. For establishing its sibling connections, the newcomer
+multicasts a special walker message to help it locate the right
 candidate. If it is found then the candidate as well as its own siblings will
@@
 % overhead of split-hash and dnMatch is of low cost (TODO: double check this).
-% Walker message is multi-casted only when sibling connections must be
+% Walker message is multicasted only when sibling connections must be
 % established.
@@
 revision} in the paper's context, the AM metric is used to store hosts'
-addresses. read from \emph{X-Try} headers during handshake or from
+addresses read from \emph{X-Try} headers during handshake or from
 \emph{QueryHit} messages. Additionally, when a node reaches the max neighbor
-connections, node disconnects the neighbors with the lowest AM.
+connections, the node disconnects the neighbors with the lowest AM.
 
@@
 physical locations need to be installed in strategic positions accross the the
-Internet. Authors estimate that a number of 12 such servers can prove sufficient
+Internet. The authors estimate that a number of 12 such servers can prove sufficient
 for this. Every newly arriving node measures its distance from these landmarks
@@
 applied with no significant changes to both structured and unstructured P2P
-systems. \cite{RHMKS2002} has the detailed description of such algorithm for
+systems. \cite{RHMKS2002} has the detailed description of the algorithm for
 both architectures. Landmark binning is also considered as a good candidate for
-content distribution networks in particular. The major disadvantage of landmark
-binning is to install and maintain landmark servers on different autonomous
+content distribution networks. The major disadvantage of landmark
+binning is that landmark servers must be installed and maintained
+on different autonomous
 systems, all over the world. A typical P2P network usually has a couple of
@@
 \emph{mOverlay} \cite{ZZZSZ2004} tries to addresses the scalability issues that
-might be imposed to networks (i.e. load unbalance) when using static landmark
+might be imposed on networks (i.e. load unbalance) when using static landmark
 servers by introducing dynamic landmarks. Zhang et al. introduce the notion of a
 \emph{group} which is a set of peers considered close to each other with respect
-to any position in the underlying network, proximity being defined by some
+to any position in the underlying network, where proximity is defined by some
 user-defined cost metric like RTT, latency or something else. mOverlay is
@@
 grouped based on their distance to the groups already in the network, rendering
-the later (dynamic) landmarks for the process. The authors formally prove
+the latter (dynamic) landmarks for the process. The authors formally prove
 that any new node can reach its group by exchanging at most $O(logN)$ messages
@@
 
-In this section we introduced the topology mismatching problem from the
-perspective of decentralized unstructured architectures and we visited a series
+In this section, we introduced the topology mismatching problem from the
+perspective of unstructured P2P systems and we visited a series
 of algorithms that have been proposed in the literature to tackle its negative
-influence to networks' performance. Even though, at first glance, the algorithms
-seem to do things their own way actually we can distinguish four main
-methodologies that are being utilized in order for every algorithm to achieve
-its intended result. These methodologies are:
+influence on performance. In surveying the above algorithms, we identify
+four main methologolies that are used either selectively or in combination
+by each algorithm.  
+These methodologies are:
 \begin{enumerate}[\itshape i\upshape)]
@@
 \end{enumerate}
-
-Table~\ref{unstructured:table} offers a summarized overview of the algorithms
-we saw earlier in the section and tracks down the aforementioned methodologies
-each of them incorporates along with their highlights and a rough estimation of
-positive and negative aspects of their implementation.
 
@@
 The topology adaptation based protocols modify the topology of the P2P network
-using various techniques. The two commonly used approaches that can
-back such feature and were investigated in this section are creating
+using various techniques. The two most commonly used approaches to 
+topology adaptation are creating
 \emph{spanning tree}s using connection graphs, and creating \emph{clusters} of
@@
 this graph that can efficiently route messages among peers. Even though spanning
-trees provide efficient query performance, their maintenance costs can be
-proven a real blocker. That is why the original Narada algorithm did not have
-any real topology adaptation mechanism meaning that it was designed for small
-groups (therefore not scalable). Algorithms like AOTO, LTM or SBO try to
+trees provide efficient query performance, their maintenance costs have 
+proven to be a significant limitation. This is why the original Narada 
+algorithm 
+was intended for small
+groups and is thus, not scalable.  Algorithms like AOTO, LTM or SBO try to
 overcome these obstacles with clever tricks like forming minimum spanning trees
 for the two-hop neighbors ($N^2$) of each node, partitioning the graph into two
-random groups where each group is responsible of different tasks, or performing
+random groups where each group is responsible for different tasks, or performing
 local optimizations dynamically on the overlay graph. The advantage of building
 minimum spanning trees is that they maintain the connectivity on the network in
-an efficient manner, while still preserving the overall search scope. However,
+an efficient manner while still preserving the overall search scope. However,
 their construction and the update costs, especially in dynamic, high-churn
-environments, cause large additional traffic overhead to the underlying
+environments, cause large additional traffic overhead on the underlying
 network \cite{CRZ2000,CRSZ2001,CRSZ2002}.
 
-The cluster based approaches, on the other hand, select to link physically
-closer nodes with each other. T2MC, for example, uses trace-route logs and DDNO
-uses domain names to cluster close by nodes. Unfortunately, commonly used
+The cluster-based approaches, on the other hand, link physically
+closer nodes to each other. T2MC, for example, uses trace-route logs and DDNO
+uses domain names to cluster nearby nodes. Unfortunately, commonly used
 methods for proximity detection across Internet do not always return
-reliable results, therefore mapping accuracy is not guaranteed. Moreover,
-trace-route, is a heavy-weight utility to be frequently used on the
-network, let alone the fact that many network equipment vendors do not allow
+reliable results and therefore mapping accuracy is not guaranteed. Moreover,
+trace-route 
+is a heavy-weight utility and many network equipment vendors do not allow
 trace-route calls at all. The most problematic aspect of clustering, though, is
@@
 significantly shrink the search scope, negatively affecting the query response
-time the P2P end-user experiences. Of course, techniques that try to round up
-such edges and balance the efficiency of the clustering approach with improved
-connectivity, have also been proposed, like DDNO, which attempts to address this
+time that the P2P end-user experiences. Techniques, like DDNO, that try to 
+balance the efficiency of the clustering approach with improved
+connectivity have also been proposed.  Specifically, DDNO attempt to address this
 limitation by forcing half of each node's connections to be with other, randomly
@@
 
-Concluding, topology adaptation ultimately targets on the reduction of the
-average cost of the paths from one node to another. This way someone can deduce
-that the overlay network is better fit on top of the physical, underlying one.
-This deduction is fair in the general case but not always. For example an
+In brief, topology adaptation ultimately aims to reduce the 
+average cost of the paths from one node to another so that 
+the overlay network is better fit on top of the underlying IP network.
+This is fair in the general case but not always. For example an
 adaptive algorithm can exchange a slow edge in the overlay with a faster one
 thus reducing the average latency of the network communication. The pitfall here
-is that this new virtual link can traverse a fast Autonomous System -to-
-Autonomous System link meaning that even though is reduced message
-round-trip-time, it has more cost in terms of communication network economics
+is that this new virtual link can traverse a fast AS-to-AS
+link meaning that even though message
+round-trip-time is reduced, it has more cost in terms of communication network economics
 and management.
@@
 
-Forwarding optimization can reduce the emitted messages flooding the overlay and
-as a consequence the overall burden, network resources have to handle. In the
-first realization of the Gnutella protocol, each query request received by a
-peer, was then forwarded to all of its neighbors. This was the source of
-unnecessary, sometimes redundant traffic. Forwarding-optimization-based
+%Forwarding optimization can reduce the emitted messages flooding the overlay and
+%as a consequence the overall burden, network resources have to handle. In the
+%first realization of the Gnutella protocol, each query request received by a
+%peer, was then forwarded to all of its neighbors. This was the source of
+%unnecessary, sometimes redundant traffic.
+
+Forwarding-optimization-based
 approaches, propose intelligent forwarding link selection for a message's next
@@
 algorithm, but they mainly use one or more statistical heuristics. These can
-span from some obvious ones, like the latency of the link between the nodes or
+range from simple obvious ones, like the latency of the link between the nodes or
 the exposure of specific features, such as high processing bandwidth, high
@@
 the far right is not reached) thus limiting the scalability of the whole
-network. Moreover they don't actually address the problem of topology mismatch
-since they don't give any guarantees that overlay and underlying topologies are
+network. Moreover they do not actually address the problem of topology mismatch
+since they do not give any guarantees that overlay and underlying topologies are
 aligned with each other let alone quantify the mismatch and try to alleviate.
 Like all other approaches, forwarding based methods are commonly applied in
-conjunction with other to improve the quality of the P2P systems.
+conjunction with other methods to improve the quality of the P2P systems.
 
@@
 source of the query, the TTL value set for the messages, the current
-interconnection of peers and the high volatility of the environment. So, in
-order to develop a successful caching system for a P2P architecture, these
+interconnection of peers and the high volatility of the environment. Thus,
+to develop a successful caching system for a P2P architecture, these
 parameters also have to be considered. P2P caching/replication can be applied
-on two different levels, namely caching indices or pointers to data (see
+at two different levels, namely caching indices or pointers to data (see
 Sub-figure~\ref{figure:replication:index}) or caching the data itself (see
@@
 compares its list to the list of the potential neighbor. Peers with similar
-measured distances to the these landmark points is very likely to be close to
+measured distances to the these landmark points are likely to be close to
 each other as well.
 
-Landmark based protocols has four important drawbacks. Firstly the network
+Landmark based protocols have four important drawbacks. First, the network
 delay is not a reliable distance estimation method. For example, based on the
-load on the network the delay to certain nodes or networks can change from time
+load on the network, the delay to certain nodes or networks can change from time
 to time, which will eventually affect the distance measurements and wrong
 measurements will lead to wrong estimated positions for the nodes, or incorrect
-and non optimal clusterings of the nodes. Secondly, relying on predefined nodes
+and non optimal clusterings of the nodes. Second, relying on predefined nodes
 make the whole paradigm not fully distributed and the landmark system prone to
-become a single point of failure. Moreover, the third drawback of using
-landmark servers is the cost of installing and maintaining this kind of
+become a single point of failure. Third, using
+landmark servers requires costly installation and maintenance of landmark
 infrastructure across the whole Internet and for all the different autonomous
 system domains. As popular P2P file sharing applications usually have millions
-of peers connected at any time, it is not false to assume that the network costs
-of maintaining these landmark servers will be quite high. A possible solution to
+of peers connected at any time, the network costs
+of maintaining these landmark servers will likely be quite high. A possible solution to
 the scalability problem of the static landmark servers is to use ordinary nodes
 as dynamic landmarks once they estimate their own positions. Even though this
-approach scales much better than static landmark servers, still the measurement
+approach scales much better than static landmark servers, the measurement
 accuracy problem affects the overall performance of the system. Moreover,it can
@@
 
+Table~\ref{unstructured:table} offers a summary overview of the algorithms
+we saw earlier in this section and for each algorithm, shows which of the
+aforementioned methodologies it incorporates, along with its 
+highlights and a rough estimation of the
+positive and negative aspects of its implementation.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-01-27 conclusion.tex
2013.01.21 (Alex) Title changed. Abstract, introduction and conclusion reworked. Fixed massive bibtex problems.

--- 2014-01-21 conclusion.tex
@@
 
-Peer-to-peer architecture has been in the center of research attention in the
-last decade. Especially the decentralized unstructured genre exploits the
-advantages of loose coupling and self organization of computing nodes to form
-application-layer networks on top of the physical, best-effort infrastructure of
-the Internet that exhibit interesting properties. Scalability problems arouse
-quickly, though, because of the inefficient construction of this overlay that
-was built with no concern for the underlying physical network that causes a
-great deal of redundant traffic. The problem was identified by the research
-community as the topology mismatch problem between the overlay and the
-corresponding underlying physical network and a great deal of effort has been
-set towards alleviating it. Some fruits of this effort have been gathered and
-presented in this survey. Measurement of link cost through latency or RTT and
-deletion of inefficient established connections are the key concepts of almost
-all approaches. Others, manage to address the problem through hierarchical peer
-clustering (e.g. best IP matching). Additionally, some protocols focus on
-specific problems that furtherly arise, such as overlay partitioning, search
-scope reduction or convergence speed, to name just a few. Another desire is to
-form a protocol that could be applied to both decentralized unstructured and
-decentralized structured networks. Unfortunately no approach has equally
-addressed these problems in order to form a robust solution. So the field seems
-to be, still, fertile for any, new, clever idea.
+Research and development for computer systems
+built using the \p\ paradigm have flourished  in the last decade.
+By and large, such systems exploit the loose-coupling and self-organization 
+of participating nodes to shape substrates upon which diverse applications 
+can run; the latter offer services including but not limited to 
+distributed naming, 
+Internet telephony and video conferencing, 
+storage and indexing, 
+content and file sharing, 
+web crawling and caching, 
+event notification and subscribe/publish.
+These services are provided atop the 
+best-effort infrastructure of today's Internet 
+and they have to effectively deal with an array of key issues
+emanating from the fact that \p\ nodes 
+operate at the edge of the Internet; such critical issues include
+node availability, robustness, scalability, load-sharing,
+quality-of-service and user anonymity.
+The majority of such issues emerges as 
+application-level \p--networks are formed without paying 
+adequate attention to the design of the physical network beneath.
+The latter constitutes what is known as the
+topology mismatch problem between the \p-overlays and the physical networks.
+In addition, the development of \p--systems and applications
+typically involved juggling a set of tradeoffs (e.g.,
+choosing higher accuracy at the cost of increased system overhead.)
+
+% Computer systems design is typically about juggling a set of tradeoffs (e.g.,
+% choosing higher accuracy at the cost of increased system overhead.)  
+In surveying more than
+a decade's worth of research efforts aimed at solving the topology mismatch
+problem in both unstructured and structured \p\ networks, we find this
+tussle-of-tradeoffs property to hold.  
+With regards to the three criteria we used --efficiency,
+overhead, and scalability-- we find that none of the proposed solutions 
+is superior to the others on all three fronts.  This is not surprising.
+Nonetheless, by
+1)~presenting an analysis of each of the solutions, 
+including what distinguishes it from the
+others and its advantages and disadvantages and 
+2)~offering a pictorial
+comparison of how each solution fares compared with the others with regards
+to efficiency, overhead, and scalability, we hope to have provided \p\
+researchers and developers with enough perspective so that depending 
+on the needs of their particular application, they are able to select
+amongst these surveyed solutions or propose a new solution of their own.

### 2014-01-27 intro.tex
2013.01.21 (Alex) Title changed. Abstract, introduction and conclusion reworked. Fixed massive bibtex problems.

@@
 
-\emph{Peer-to-peer (P2P)} networks are self-organizing distributed systems where
+\emph{Peer-to-peer (\p)} networks are self-organizing distributed systems where
 participating nodes, called \emph{peers}, act as both resource providers and
-resource consumers, in contrast to the traditional \emph{client-server} model
-where nodes take on the role of pure servers or pure clients. Peer-to-peer
-networks have enjoyed immense attention from Internet users, companies, and
-researchers and have been widely deployed for over a decade, primarily because
-of the great number of features they offer distributed applications that are
-built on top of them. These features include:  high availability and robustness,
+resource consumers in contrast to the traditional \emph{client-server} model
+where nodes undertake specific roles.
+For over a decade, \p\ networks have been widely deployed and have
+enjoyed immense popularity from Internet communities, primarily because
+of the great number of features they offer to distributed applications 
+built atop them. 
+Such diverse features include:  high availability and robustness,
 load-balancing, quality of service, scalability, decentralized administration,
@@
 
-Two ``killer'' applications of the peer-to-peer paradigm are 
-file-sharing and telephony. P2P file-sharing was originally introduced by
-Napster (see Section~\ref{section:background}), which by 2001 was widely
-acknowledged as the ``fastest growing Internet application ever" with 26 million
-users sharing 80 million songs at the height of its popularity. Other file-sharing 
-applications quickly followed suit including Gnutella and its descendant,
-Limewire, enjoying 3 million concurrently connected peers by the end of 2006
-and BitTorrent, connecting over 150 million users. P2P telephony, on the
-other hand, has seen
-explosive growth with the advent of Skype. Since its introduction in 2003,
-Skype has become extremely popular with more than 650 million users in 2011
-\cite{skypetotalusers} and 35 million skypers concurrently online on the 5th of
-March 2012 \cite{skypesymusers}.
+The peer-to-peer paradigm gave impetus to two ``killer'' applications:
+file-sharing and Internet telephony.
+The {\sl Napster} file-sharing was widely acknowledged as the 
+``fastest growing Internet application ever" in $2001$ when 
+$26$ million users shared $80$ million songs.
+% File-sharing, originally introduced by
+% {\sl Napster} (Section~\ref{section:background}), was widely
+% acknowledged by $2001$ as the ``fastest growing Internet application ever" 
+% with $26$ million users sharing $80$ million songs.
+% % at the height of its popularity. 
+Other file-sharing applications followed suit 
+including {\sl Gnutella} and {\sl Limewire} enjoying $3$ 
+million concurrently-connected peers as well as 
+{\sl BitTorrent} connecting over $150$ million users by the end of $2006$.
+\p\ telephone saw explosive growth with the advent of {\sl Skype}. 
+Since its introduction in $2003$,
+{\sl Skype} has become extremely popular with more than $650$ million users 
+in $2011$~\cite{skypetotalusers}; on March $5$th, $2012$,  
+the astonishing figure of $35$ million concurrently-online users  
+was reported~\cite{skypesymusers}.
+%%
+In addition, a startling number of diverse and successful
+applications have been built on top of \p\ systems including:
+distributed search engines~\cite{yaci}, 
+distributed data-storage systems~\cite{kbc_oceanstore_2000,bdet_fsdfs_2000,dkkms_cfs_2001,dr_pastutility_2001,abc_farsite_2002,mmfc_ivy_2002,arla,agebh_dks_2003},
+Web caches, archives and publishing systems~\cite{ird_squirrel_2002,bags_youserv_2002,wrc_publius_2000,wm_tangler_2001},
+messaging applications~\cite{threedegrees}, 
+event-notification infrastructures~\cite{rkcd_scribe_2001,cdkr_scribe_2002,agebh_dks_2003}, 
+naming services~\cite{cmm_chorddns_2002}, 
+censor-resistant stores~\cite{cswh_freenet_2001} and
+lately, even cloud-based platforms~\cite{mgpj_cloudsnap_2011}.
 
-In addition to file-sharing and telephony, a large number of diverse
-applications have been built on top of peer-to-peer systems including:
-distributed search engines \cite{yaci}, distributed data storage systems
-\cite{kbc_oceanstore_2000,bdet_fsdfs_2000,dkkms_cfs_2001,dr_pastutility_2001,abc_farsite_2002,mmfc_ivy_2002,arla,agebh_dks_2003},
-Web caches, archives and publishing systems
-\cite{ird_squirrel_2002,bags_youserv_2002,wrc_publius_2000,wm_tangler_2001},
-messaging applications \cite{threedegrees}, event notification infrastructures
-\cite{rkcd_scribe_2001,cdkr_scribe_2002,agebh_dks_2003}, naming services
-\cite{cmm_chorddns_2002}, censor-resistant stores \cite{cswh_freenet_2001} and
-lately even cloud-based platforms \cite{mgpj_cloudsnap_2011}.
-
-Peer-to-peer systems are implemented using overlay networks. An \emph{overlay
-network} is a virtual system of nodes and logical links, that is built on top of
-an existing network intending to provide added value to it (e.g., offering QoS
-content distribution over the best effort Internet network). Nodes in the
-overlay communicate via virtual connections each of which may correspond to a
-path of possibly many physical links in the underlying network.
+\p\ systems are implemented using \emph{overlay networks}. 
+An overlay network is a virtual system of nodes and logical links
+created atop an existing network; overlays provide 
+added value to underlying network by for instance offering QoS
+content-distribution over the best effort Internet. 
+Overlay nodes communicate through \emph{virtual connections} each 
+of which may correspond to a path of possibly many physical links 
+in the underlying network.
 Figure~\ref{figure:overlay} illustrates a simple four-node overlay constructed
 over a wide-area network.
-
-\begin{figure}
+%%
+\begin{figure}[ht]
 \centering
-  \includegraphics[scale=0.7]{img/pdf/under-over-lay.pdf}
+  \includegraphics[scale=0.6]{img/pdf/under-over-lay.pdf}
 \caption{An example overlay network.}
@@
 
-One of the fundamental issues that defines the efficiency of an overlay network
+The single key issue that determines the efficiency of an overlay network
 is how well the overlay maps to the underlying network topology on which it
-rests. Consider two nodes\footnote{In the case of peer-to-peer networks,
-typically the participating nodes are user PCs sitting at the edge of the
-Internet.} that share a link in an overlay network which maps to multiple hops
-in the underlying IP topology. If the traffic patterns of the application
-are such that traffic flows heavily over the overlay link between the two nodes,
-then it is beneficial to construct the overlay topology such that the number of
+``rests''. 
+Consider two nodes\footnote{In \p\ networks,
+the participating nodes are typically user-PCs operating at the edge of the
+Internet.} that share a overlay network link which ultimately 
+maps to multiple hops in the underlying IP topology. 
+If the application running on the nodes generates traffic flows 
+the heavily use the link in question, it would be beneficial to 
+construct the overlay topology in a way that the number of 
 underlying IP links between these two nodes is minimized. 
+% If the traffic patterns of the application
+% are such that traffic flows heavily over the overlay link 
+% between the two nodes,
+% then it is beneficial to construct the overlay 
+% topology such that the number of
+% underlying IP links between these two nodes is minimized.
+%%
+Should the overlay network be constructed  so that 
+it does not match the underlying topology well, 
+the inherent \emph{topology mismatch} gives two
+significant problems: first, the performance of the application per se, can be
+adversely affected since traffic must flow over a larger,
+redundant, number of physical hops resulting in poor user experience 
+entailing noticeable latencies or jittering. 
+%%
+Second, other applications running
+on the underlying network infrastructure are also adversely affected.
+Studies have shown that highly popular \p\ applications contribute 
+the largest portion of overall 
+Internet traffic~\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}, with some reporting that more than $60$\% of this traffic 
+to be \p-related~\cite{cachelogic,ipoque2007,ipoque2009};
+it was also projected that this traffic would reach
+$8\;$\emph{Petabytes}-per-month by the end of $2012$~\cite{multinteligence}. 
+Evidently, this constitutes a major burden for 
+Internet Service Providers (ISPs) who must route 
+all of this traffic to destinations at the edge of the Internet. 
+If the \p\ overlay topology is poorly designed, 
+the demand on the Internet's backbone infrastructure may 
+substantially increase as traffic might have to flow
+``back and forth'' several times between two neighboring ISPs 
+while trying to travel from the source to its destination node
+in the overlay.
+Hence, it is very critical that \p\ networks be laid out 
+in ways that their topology matches as closely as possible the 
+underlying IP topology.
 
-If the overlay network is constructed such that it does not match the underlying
-topology well, the resulting \emph{topology mismatch} 
-can result in two
-significant problems. First, the performance of the application per se, can be
-adversely affected since application traffic must flow over a larger,
-redundunt, number of physical hops resulting in poor user experience like
-extensive search latency or jitter. Second, other applications running
-on the underlying network infrastructure are also adversely affected.
-Studies have shown that highly popular P2P applications contribute the largest
-portion of overall Internet traffic
-\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}, with
-some studies reporting that more than 60\% of Internet traffic is
-P2P-related~\cite{cachelogic,ipoque2007,ipoque2009} and others predicting that it
-will reach almost
-8 petabytes per month by the end of
-2012~\cite{multinteligence}. This is a significant burden for Internet Service
-Providers (ISPs) who must route all of this traffic to destination nodes sitting
-at the edge of the Internet. If the overlay topology of a peer-to-peer system is
-not efficiently constructed, then the burden on the Internet's backbone
-infrastructure could be even greater because traffic might have to flow
-back and forth several times between two neighboring ISPs as it travels from its
-source node in the overlay to its destination node. Thus, it is fundamentally
-important that P2P overlay networks are constructed and maintained so that
-their topology matches as closely as possible the underlying IP topology.
-
-For over a decade, several researchers have published profusely on the topology
-mismatch problem.  In this paper, we survey this research area and provide a
+For over a decade, researchers have extensively investigated  
+various aspects of this topology mismatch problem.  
+The main objective of this paper is to offer a comprehensive 
+survey of the work done in this are and provide a
 taxonomy of the proposed solutions. We point out synergies as well as
-similarities and differences in the published approaches. Ultimately, our goal
-is to help readers sift through the vast amount of literature, to help them
-understand the advantages and disadvantages of each work, and to provide
-them with enough perspective so that when the need arises, they are able to
+similarities and differences in the published approaches. 
+Ultimately, our goal is to help readers sift through 
+the voluminous literature, to help them
+understand the advantages and disadvantages of each work, and 
+to provide them with enough perspective so that 
+when the need arises, they are able to
 select amongst the different approaches for their particular application.
-
-The rest of the paper is organized as follows: In
-Section~\ref{section:background}, we provide background information on
-peer-to-peer overlay architectures including centralized, decentralized
-unstructured and decentralized structured peer-to-peer systems. We also formally
-define the problem of topology mismatch between an overlay and its underlying
-physical network and present the rationale behind our 
-evaluation of the techniques presented throughout this survey. In
-Sections~\ref{section:unstructured} and \ref{section:structured}, we present the
-various research efforts for decentralized unstructured and decentralized
-structured peer-to-peer systems, respectively. Finally, in 
-Section~\ref{section:conclusion}, we conclude the paper.
+%%
+The rest of the paper is organized as follows: 
+Section~\ref{section:background} provides background on
+overlay architectures including centralized, decentralized
+unstructured and decentralized-structured \p\ systems. 
+We also formally define the problem of topology mismatch 
+% between an overlay and its underlying
+% physical network 
+and offer the rationale behind our 
+evaluation of the techniques discussed in this paper. 
+Sections~\ref{section:unstructured} and~\ref{section:structured}
+respectively outline the surveyed research efforts 
+for decentralized-unstructured and decentralized-structured \p\ systems.
+Conclusions can be found in Section~\ref{section:conclusion}.

### 2014-01-27 mard-survey.tex
2013.01.21 (Alex) Title changed. Abstract, introduction and conclusion reworked. Fixed massive bibtex problems.

@@
 \usetikzlibrary{patterns}
+\usepackage{url}
 
@@
 \newdef{remark}[theorem]{Remark}
+
+\newcommand{\p}{\emph{P2P}}
 
@@
 
-\title{Alleviating Topology Mismatch in Distributed Overlay Networks: A Survey}
-% TODO: We added here "P2P". I'm not sure, because for example Narada is not P2P
-% TODO: I deleted "P2P"
+\title{Alleviating the Topology Mismatch Problem in Distributed Overlay Networks}
 
@@
 DELIS$^1$\\
-$^1$\emph{National and Kapodistrian University of Athens}, \emph{Athens},
-\emph{Greece}\\
+$^1$\emph{University of Athens}, \emph{Athens}, \emph{Greece}\\
 $^2$\emph{Izmir University of Economics}, \emph{Izmir}, \emph{Turkey}
@@
 \begin{abstract}
-Peer-to-peer (P2P) systems have enjoyed immense attention and have been
-widely deployed on the Internet for well over a decade.  P2P systems 
-are implemented via
-an overlay network abstraction on top of the best-effort IP infrastructure
-of the Internet. These systems offer a plethora
-of desirable features to distributed applications that are built on top of them
-including anonymity, high availability and robustness, load balancing,
-quality of service, scalability, and others.  Unfortunately, weaknesses of
-early deployments of P2P systems prevented applications from leveraging 
-these features.  One major weakness, identified early on, is the topology
+Peer-to-peer (\p) systems have enjoyed immense attention and have been
+widely deployed on the Internet for well over a decade.  
+They are implemented via an overlay network abstraction on top 
+of the best-effort IP infrastructure of the Internet. 
+\p\ systems offer a plethora of desirable features to 
+distributed applications that are built on top of them including
+anonymity, high availability and robustness, 
+load balancing, quality of service, scalability, and others.  
+Unfortunately, weaknesses of early deployments of \p\ systems prevented 
+applications from leveraging these features.  
+One major weakness, identified early on, is the topology
 mismatch problem between the overlay network and the underlying IP topology.
@@
 network resources and can adversely affect the scalability of applications
-built on top of P2P systems.  
+built atop \p\ systems.  
 In this paper, we survey over a decade's worth of research efforts
 aimed at alleviating the topology mismatch problem, in both unstructured
-and structured P2P systems.   
+and structured \p\ systems.   
 We provide a fine-grained categorization of the
 proposed solutions and discuss the advantages, disadvantages, and novelties
-introduced by each solution.  Finally, we provide an analysis and pictorial 
-comparisons 
-of the proposed solutions in an attempt to provide a quick one-stop-shop
-reference for anyone interested in the topic, from researchers and students,
-to system architects and developers.
+introduced by each solution.  
+Finally, we provide an analysis as well as pictorial comparisons 
+of the proposed solutions since we aim to offer a comprehensive 
+reference for system architects, developers, and researchers in the field.
+% in an attempt to provide a comprehensive quick one-stop-shop
+% reference for anyone interested in the topic, from researchers and students,
+% to system architects and developers.
 
@@
 \begin{bottomstuff}
-Authors' addresses: Department of Informatics and Telecommunications, National
-and Kapodistrian University of Athens, Athens, Greece. Department of Software
+Authors' addresses: Deprt. of Informatics and Telecommunications,
+University of Athens, Athens, Greece. Deprt. of Software
 Engineering, Izmir University of Economics, Izmir, Turkey.\\

### 2014-01-28 conclusion.tex
2013.01.23 (Vassilis) Text refinements in introduction and conclusion sections.

@@
 
-Research and development for computer systems
-built using the \p\ paradigm have flourished  in the last decade.
+Research and development around application systems that follow the 
+\p\ architectural paradigm, have flourished in the last ten years or so.
 By and large, such systems exploit the loose-coupling and self-organization 
-of participating nodes to shape substrates upon which diverse applications 
-can run; the latter offer services including but not limited to 
+of participating nodes, to shape substrates upon which diverse applications 
+can run; the latter offer services including, but not limited to, 
 distributed naming, 
@@
 web crawling and caching, 
-event notification and subscribe/publish.
+event notification and subscribing/publishing.
 These services are provided atop the 
 best-effort infrastructure of today's Internet 
-and they have to effectively deal with an array of key issues
-emanating from the fact that \p\ nodes 
-operate at the edge of the Internet; such critical issues include
-node availability, robustness, scalability, load-sharing,
-quality-of-service and user anonymity.
-The majority of such issues emerges as 
-application-level \p--networks are formed without paying 
-adequate attention to the design of the physical network beneath.
-The latter constitutes what is known as the
-topology mismatch problem between the \p-overlays and the physical networks.
-In addition, the development of \p--systems and applications
-typically involved juggling a set of tradeoffs (e.g.,
-choosing higher accuracy at the cost of increased system overhead.)
+and they have to effectively deal with an array of key quality attributes that
+they may need or have to offer, such as node availability, robustness,
+scalability, load-sharing, quality-of-service and user anonymity.
+The difficulty to handle such features effectively, emanates from the fact that
+peer nodes operate at the edge of the Internet and no adequate attention to the
+structure of the physical network is paid during application-level network
+formation. The deviation of the structure of the overlay network from the
+optimal, according to the underlying physical structure, constitutes what is
+termed topology mismatch and the problem is known as the topology mismatch
+problem.
 
+Development of \p--systems and applications
+typically involve juggling a set of tradeoffs (e.g.,
+choosing higher accuracy at the cost of increased system overhead.).
 % Computer systems design is typically about juggling a set of tradeoffs (e.g.,
@@
 Nonetheless, by
-1)~presenting an analysis of each of the solutions, 
+\begin{inparaenum}[\itshape i\upshape)]
+  \item presenting an analysis of each of the solutions,
 including what distinguishes it from the
 others and its advantages and disadvantages and
-2)~offering a pictorial
+  \item offering a pictorial
 comparison of how each solution fares compared with the others with regards
-to efficiency, overhead, and scalability, we hope to have provided \p\
-researchers and developers with enough perspective so that depending 
-on the needs of their particular application, they are able to select
-amongst these surveyed solutions or propose a new solution of their own.
+to efficiency, overhead, and scalability,
+\end{inparaenum}
+we hope to have provided \p\
+developers and researchers with enough perspective so that depending
+on the particular problem in hand, they are able to make a
+correct choice amongst these surveyed here or, by firmly standing on
+the shoulders of giants, propose a new solution of their own.

### 2014-01-28 intro.tex
2013.01.23 (Vassilis) Text refinements in introduction and conclusion sections.

@@
 
-\emph{Peer-to-peer (\p)} networks are self-organizing distributed systems where
+\emph{Peer-to-peer} (\p) networks are self-organizing, distributed systems where
 participating nodes, called \emph{peers}, act as both resource providers and
-resource consumers in contrast to the traditional \emph{client-server} model
+resource consumers, in contrast to the traditional \emph{client-server} model
 where nodes undertake specific roles.
@@
 The {\sl Napster} file-sharing was widely acknowledged as the 
-``fastest growing Internet application ever" in $2001$ when 
+``fastest growing Internet application ever'' in $2001$ when
 $26$ million users shared $80$ million songs.
@@
 % % at the height of its popularity. 
-Other file-sharing applications followed suit 
+Other file-sharing applications followed suit, 
 including {\sl Gnutella} and {\sl Limewire} enjoying $3$ 
@@
 {\sl BitTorrent} connecting over $150$ million users by the end of $2006$.
-\p\ telephone saw explosive growth with the advent of {\sl Skype}. 
+\p\ telephony saw explosive growth with the advent of {\sl Skype}.
 Since its introduction in $2003$,
@@
 In addition, a startling number of diverse and successful
-applications have been built on top of \p\ systems including:
+applications have been built based on a \p\ architecture. Some of them include:
 distributed search engines~\cite{yaci}, 
@@
 \p\ systems are implemented using \emph{overlay networks}. 
-An overlay network is a virtual system of nodes and logical links
-created atop an existing network; overlays provide 
-added value to underlying network by for instance offering QoS
+An overlay network, is a virtual system of nodes and logical links
+created above an existing network; overlays provide
+added value to the underlying layer by, for instance, offering QoS
 content-distribution over the best effort Internet. 
@@
 
-The single key issue that determines the efficiency of an overlay network
+The single key issue that determines the efficiency of an overlay network,
 is how well the overlay maps to the underlying network topology on which it
@@
 the participating nodes are typically user-PCs operating at the edge of the
-Internet.} that share a overlay network link which ultimately 
-maps to multiple hops in the underlying IP topology. 
-If the application running on the nodes generates traffic flows 
-the heavily use the link in question, it would be beneficial to 
+Internet.} that are connected with each other via a path of overlay links.
+If the application running on the nodes, generates heavy traffic along
+the overlay path, it would be beneficial to
 construct the overlay topology in a way that the number of 
@@
 Studies have shown that highly popular \p\ applications contribute 
-the largest portion of overall 
+the largest portion of the overall 
 Internet traffic~\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}, with some reporting that more than $60$\% of this traffic 
@@
 Hence, it is very critical that \p\ networks be laid out 
-in ways that their topology matches as closely as possible the 
+in ways that their topology matches, as closely as possible, the 
 underlying IP topology.
@@
 For over a decade, researchers have extensively investigated  
-various aspects of this topology mismatch problem.  
+various aspects of the topology mismatch problem.
 The main objective of this paper is to offer a comprehensive 
-survey of the work done in this are and provide a
-taxonomy of the proposed solutions. We point out synergies as well as
+survey of the work done in this and provide a
+taxonomy of the proposed solutions. We point out synergies, as well as
 similarities and differences in the published approaches. 
@@
 when the need arises, they are able to
-select amongst the different approaches for their particular application.
+select, amongst the different approaches, the one that is most suitable for
+their particular application.
 %%

### 2014-01-28 mard-survey.tex
2013.01.23 (Vassilis) Text refinements in introduction and conclusion sections.

@@
 \begin{abstract}
-Peer-to-peer (\p) systems have enjoyed immense attention and have been
-widely deployed on the Internet for well over a decade.  
-They are implemented via an overlay network abstraction on top 
+\emph{Peer-to-peer} (\p) systems have enjoyed immense attention and have been
+widely deployed on the Internet for, well, over a decade.  
+They are implemented via an overlay network abstraction, on top 
 of the best-effort IP infrastructure of the Internet. 
@@
 anonymity, high availability and robustness, 
-load balancing, quality of service, scalability, and others.  
-Unfortunately, weaknesses of early deployments of \p\ systems prevented 
+load balancing, quality of service, scalability, to name just a few.
+Unfortunately, weaknesses of early deployments of \p\ systems, prevented 
 applications from leveraging these features.
-One major weakness, identified early on, is the topology
-mismatch problem between the overlay network and the underlying IP topology.
+One major weakness, identified early on, is the \emph{topology mismatch problem}
+between the overlay network and the underlying IP topology.
 This mismatch can impose an extraordinary amount of unnecessary stress on
-network resources and can adversely affect the scalability of applications
-built atop \p\ systems.  
+network resources and can adversely affect the scalability and efficiency of
+applications built atop.
 In this paper, we survey over a decade's worth of research efforts
@@
 proposed solutions and discuss the advantages, disadvantages, and novelties
-introduced by each solution.  
+introduced by each of them.
 Finally, we provide an analysis as well as pictorial comparisons 
-of the proposed solutions since we aim to offer a comprehensive 
-reference for system architects, developers, and researchers in the field.
+of the reviewed approaches, since we aim to offer a comprehensive
+reference for developers, system architects and researchers in the field.
 % in an attempt to provide a comprehensive quick one-stop-shop
@@
 Engineering, Izmir University of Economics, Izmir, Turkey.\\
-emails: \{b.moystakas,mema,ad\}@di.uoa.gr, huseyin.akcan@ieu.edu.tr
+emails: \{vsmoustakas,mema,ad\}@di.uoa.gr, huseyin.akcan@ieu.edu.tr
 \end{bottomstuff}

### 2014-01-28 background.tex
2013.01.23 (Alex) Introduction, background and conclusion sections reworked.

@@
 In this section, we provide some background information on the types of
-peer-to-peer architectures that have been implemented and deployed. We also
-describe and motivate more formally the topology mismatch problem tackled by
-many P2P researchers over the past several years.
+\p\ architectures that have been implemented and deployed. We also
+describe and motivate more formally the topology mismatch problem 
+whose various aspects have been tackled over the past several years.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Peer-to-Peer (P2P) Overlay Architectures}
+\subsection{Peer-to-Peer (\p) Overlay Architectures}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Circa 2000, the first P2P file-sharing systems introduced the \emph{servent}
-concept \cite{gnutella}, a \emph{portmanteau} that blends the notions of server
-and client to denote the twofold role of a node participating in the P2P
-network. Such server-less systems have proven to be able to achieve outstanding
+Circa 2000, the first file-sharing \p-system introduced the \emph{servent}
+concept~\cite{gnutella}, a \emph{portmanteau} that blends the notions of server
+and client to denote the dual role of a node participating in \p-network. 
+Such server-less systems have proven to be able to achieve outstanding
 aggregate resource capacities as more and more participants join the system
-without requiring additional expenditure for infrastructure. While certain
-undesirable features such as 
+without requiring additional expenditure for infrastructure. 
+While certain undesirable features such as 
 free-riding~\cite{saroiu_measurefileshare_2002,adar_gnutellafreeriders_2000,hughes_gnutellafreeride_2005},
-the distribution of illegal content and other \emph{socio-technical}
-issues \cite{hughes_socp2p_2008}, have arisen, peer-to-peer systems have
-continued to gain popularity throughout the years primarily because they have successfully supported
-popular applications (i.e. file-sharing) amongst vast numbers of participating
-users.
-
-For over a decade, excitement over the immense potential of P2P systems has
+the distribution of illegal content and other~\emph{socio-technical}
+issues \cite{hughes_socp2p_2008}, have arisen, \p\ systems have
+continued to gain popularity throughout the years primarily 
+because they have successfully supported popular applications 
+(i.e. file-sharing) amongst \emph{vast} numbers of participating users.
+%%
+For over a decade, excitement over the immense potential of \p\ systems has
 generated a flurry of research and development, resulting in a wide range of
-popular protocols, networks, and applications. As P2P systems evolved, three
+popular protocols, networks, and applications. As \p\ systems evolved, three
 main architectures have emerged, namely
@@
 
-Designers of \emph{centralized} architectures (see
-Figure~\ref{figure:p2p-archs:centralized}) were the first to recognize that
-requests for resources (e.g. CPU cycles or popular content) need not be sent to
-a dedicated server but instead could be handled by the many hosts that already
-posess the resources. The downside of this approach is that it requires a
-centralized searching facility which becomes a single point of failure, a
-scalability bottleneck, and a target of malicious attacks (e.g.,
-\emph{Denial-of-Service (DoS) attacks}).
+Designers of \emph{centralized} architectures 
+(Figure~\ref{figure:p2p-archs:centralized}) were the first 
+to recognize that requests for resources 
+(e.g., CPU cycles or popular content) need not be sent to
+a dedicated server. Instead, requests could be handled by 
+many hosts that already possessed the resources in question. 
+The downside of this approach was that it required 
+centralized search that became a single point of failure, a
+scalability bottleneck, and a target of malicious including \emph{DoS} attacks. 
+% \emph{Denial-of-Service (DoS) attacks}).
 
-The most successful incarnation of the centralized approach is the file-sharing
-application \emph{Napster}\cite{napster}. Napster maintained a \emph{central
-index server} based on file lists provided by participating peers. The central
-index server was queried by users and it returned pointers to the actual
-content. Thus, by centralizing search while distributing downloads,  Napster
-achieved a highly functional design that, at its height, attracted 26 million
-users sharing approximately 80 million songs\cite{jmm_naptopusage_2001}.
-Napter's  centralized nature was what ultimately brought it down after Recording
-Industry Association of America (RIAA) filed a lawsuit for copyright infringemnt
-of the intellectual property of artists whose music was distributed by the
-Napster application. RIAA was able to point to the centralized indexing
-component of Napster as the entity resposible for illegal song-trading.
+{\sl Napster} was the most successful incarnation of the
+centralized approach in file-sharing~\cite{napster}.
+It maintained a \emph{central index server} based on file-lists 
+that were peer-provided.
+Users would query the index server that would ultimately 
+yield pointers to the actual content.
+Thus, by centralizing search while distributing downloads,  {\sl Napster}
+achieved a highly functional design that, at its height, attracted $26$ million
+users sharing approximately $80$ million songs\cite{jmm_naptopusage_2001}.
+%%
+The central indexing played a vital role in the demise of 
+{\sl Napter} as the \emph{Recording
+Industry Association of America} successfully argued in court 
+copyright infringement of the artists songs traded in the system. 
+%
+% Napter's  centralized nature was what ultimately brought it down after Recording
+% Industry Association of America (RIAA) filed a lawsuit for copyright infringement
+% of the intellectual property of artists whose music was distributed by the
+% Napster application. RIAA was able to point to the centralized indexing
+% component of Napster as the entity responsible for illegal song-trading.
 
-The centralized approach was quickly replaced by architectures that distribute
-both search and resource provision capabilities (see
-Figure~\ref{figure:p2p-archs:unstructured-full}). In these \emph{decentralized}
-architectures, resource placement within the overlay topology is
-random~\cite{YG-M2002}. For this reason they are referred to as
-\emph{unstructured}. The most important properties of such systems are that they
-support inherent heterogeneity of peers, are highly resilient to peer failures,
-and incur low maintenance overhead at handling the dynamics of peer
-participation \cite{stutzbach_churn_2006}. In the literature, they are also
-known as \emph{broadcast-based} systems, because they use message flooding among
-peers to propagate search queries. A well-known example of the fully
-decentralized
-unstructured approach is Gnutella, in its first realization (v.0.4)
-\cite{gnutellav04}.
+Architectures that distributed both search and resource provision 
+capabilities (Figure~\ref{figure:p2p-archs:unstructred-full}) soon followed.
+In these \emph{decentralized} approaches, 
+resource placement within the overlay topology is random~\cite{YG-M2002};
+for this reason, such architectures are referred to as \emph{unstructured}. 
+Key properties  of such \p\ systems are that they
+support inherent heterogeneity of peers, 
+are highly resilient to peer failures,
+and incur low maintenance overhead at handling 
+the dynamics of peer participation \cite{stutzbach_churn_2006}. 
+Unstructured systems are also known as 
+\emph{broadcast-based} systems as they use message flooding among
+peers to propagate search queries. 
+In its first realization,
+{\sl Gnutella}(\emph{v.0.4}) is a well-known example of the fully
+decentralized unstructured approach.
 
@@
 the network since queries travel within the network randomly, visiting nodes
-that do not have the resource of interest and thus wasting bandwidth.  To save
-on bandwidth, unstructured networks typically limit how long a query travels
-within the overlay (via a time-to-live (TTL) parameter). When a query is
-received by a node, it decreases the TTL by one unit. If the TTL has not reached
-zero, the node forwards the query to its neighbors.  Otherwise, it drops the
-query. While this approach prevents queries from visiting the entire network, it
-does not guarantee that the resource of interest, (e.g., a rare recording 
-of a musical group from the 1930s) will be found.
-
-To improve search scope and scalability and to lighten network load, 
-unstructured systems evolved into ``hierarchical'' systems that dynamically
-assign indexing functions to special peers, called \emph{ultra-peers}  or
-\emph{super-peers} (see
-Figure~\ref{figure:p2p-archs:unstructured-hierarchical}). Peers with reliable
-network connections and/or high compute power became ultra/super-peers
-facilitating less powerful peers (called \emph{leaf-nodes}) to find resources of
-interest. Well-known examples of these hierarchical unstructured systems include
-Gnutella v.0.6/Limewire~\cite{gnutella}, KaZaA~\cite{kazaa} and
-Skype~\cite{skype}. While the hierarchical approach relieves network pressure as
-search queries are flooded over a much smaller subset of peers (the
-ultra/super-peers), the disadvantages of reduced search scope and inability to
-locate rare resources remain.
-
-To fully address the disadvantages of decentralized unstructured approaches,
-\emph{decentralized structured} schemes have been proposed (see
-Figure~\ref{figure:p2p-archs:structured}). The key goal of these systems is to
-provide a self-organizing infrastructure for large-scale P2P applications
-\cite{ratnasamy_can_2001,stoica_chord_2001,antony_pastry_2001,zhao_tapestry_2001,maymounkov_kademlia_2002,rgrk_bamboo_2004}.
-They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes
-through a deterministic mechanism. The DHT provides a guaranteed bound on the
-number of overlay routing hops that have to be crossed to locate a resource, even
-in the case when only a single copy exists in the system.  This requires
-$O \left( log n \right)$ hops, compared to unstructured networks that require
-$O \left( n \right)$ to reliably locate any object. Unfortunately this
-efficiency does not come without cost. Structured overlays can only support
-exact-match queries (i.e., queries that identify resources by name) as opposed
-to content-based retrieval provided by unstructured overlays. This means
-that unstructured overlays can support more versatile resource location queries
-such as keyword and partial matching searches. Several popular file-sharing
-peer-to-peer applications have been based on DHTs including
-Overnet/eDonkey~\cite{overnet}, Kademlia/Kad
-Network~\cite{maymounkov_kademlia_2002}, some flavors of
-BitTorrent~\cite{c_bittorrent_2003}, and many more.  Moreover, DHTs have sparked
-a flurry of research activity and development for a large range of applications:
-from distributed search engines to event notification infrastructures to
-cloud-based platforms. We cite several examples in Section~\ref{section:intro}.
-
+that do not have the sought resource and thus wasting bandwidth.  
+To save on bandwidth, unstructured networks typically limit how long 
+a query travels within the overlay via the \emph{time-to-live} parameter
+(\emph{TTL}). 
+When a query is received by a node, it decreases its \emph{TTL} by one unit. 
+If the \emph{TTL} has not reached zero, the node forwards the query 
+to its neighbors.  Otherwise, it drops the query. 
+While this approach prevents queries from visiting the entire network, it
+does not guarantee that the resource of interest, (e.g., a rare recording) 
+% of a musical group from the 1930s) 
+will be ultimately found.
+%%%
 \begin{figure}[ht]
 \centering
-\subfigure[Centralized (SETI@Home or Napster). Peers query the centralized
+\subfigure[Centralized ({\sl SETI@Home} or {\sl Napster}). Peers query the centralized
 entity (dotted lines) and then establish the connection to the data source.] {
-  \includegraphics[scale=0.25]{img/pdf/centralized.pdf}
+  \includegraphics[scale=0.20]{img/pdf/centralized.pdf}
   \label{figure:p2p-archs:centralized}
 }\qquad\qquad
-\subfigure[Fully unstructured (Gnutella v.0.4). Data is distributed randomly,
+\subfigure[Fully unstructured ({\sl Gnutella} \emph{v.0.4}). Data is distributed randomly,
 the topology is also random and all participants are equal.] {
-  \includegraphics[scale=0.25]{img/pdf/unstructured-full.pdf}
-  \label{figure:p2p-archs:unstructured-full}
+  \includegraphics[scale=0.20]{img/pdf/unstructured-full.pdf}
+  \label{figure:p2p-archs:unstructred-full}
 }\qquad\qquad
-\subfigure[Hierarchical unstructured (Gnutella v.0.6 or FastTrack). Certain
+\subfigure[Hierarchical unstructured ({\sl Gnutella} \emph{v.0.6} or {\sl FastTrack}). Certain
 peers (bigger circles) shoulder specific responsibilities (e.g., data
 indexing).] {
-  \includegraphics[scale=0.25]{img/pdf/unstructured-hierarchical.pdf}
+  \includegraphics[scale=0.20]{img/pdf/unstructured-hierarchical.pdf}
   \label{figure:p2p-archs:unstructured-hierarchical}
 }\qquad\qquad
-\subfigure[Structured (Chord and Kademlia). Peers and data are placed in the
-overlay in a deterministic manner.] {
-  \includegraphics[scale=0.25]{img/pdf/structured.pdf}
+\subfigure[Structured ({\sl Chord} and {\sl Kademlia}). 
+Peers and data are placed in the overlay in a deterministic manner.] {
+  \includegraphics[scale=0.20]{img/pdf/structured.pdf}
   \label{figure:p2p-archs:structured}
 }
-\caption{The various peer-to-peer architectures.}
+\caption{The various \p\ architectures.}
 \label{figure:p2p-archs}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+To shorten search times, improve scalability, and lighten network load, 
+unstructured configurations evolved into \emph{hierarchical} systems 
+that dynamically assigned indexing functions to special peers, 
+called \emph{ultra-}  or \emph{super-peers} 
+(Figure~\ref{figure:p2p-archs:unstructured-hierarchical}). 
+Peers with reliable network connections and/or high compute power 
+became \emph{ultra/super-peers}
+facilitating less-powerful \emph{leaf-nodes} to find resources of
+interest. 
+Examples of such hierarchical unstructured systems include
+{\sl Gnutella}(\emph{v.0.6})/{\sl Limewire}~\cite{gnutella}, 
+{\sl KaZaA}~\cite{kazaa} and {\sl Skype}~\cite{skype}. 
+Although the hierarchical approach relieves network pressure as
+search queries were flooded over the much smaller subset of 
+ultra/super-peers, the disadvantages of reduced search scope and 
+inability to locate rare resources remained.
+
+The \emph{decentralized structured schemes}
+(Figure~\ref{figure:p2p-archs:structured}) sought to address the 
+disadvantages of the above decentralized unstructured \p--approaches.
+Here, the main objective was to offer 
+a self-organizing infrastructure for large-scale \p\ applications
+\cite{ratnasamy_can_2001,stoica_chord_2001,antony_pastry_2001,zhao_tapestry_2001,maymounkov_kademlia_2002,rgrk_bamboo_2004}.
+\emph{Decentralized structured schemes}
+implement a \emph{Distributed Hash Table (DHT)} that maps objects 
+to nodes through a deterministic mechanism. 
+\emph{DHT}s provide a guaranteed bound on the
+number of overlay routing hops that have to be ``traveled''
+% crossed 
+to locate a resource, even in the case when only a \emph{single} copy exists.  
+This requires
+$O$($log(n)$) hops, compared to unstructured networks that require
+$O(n)$ to reliably locate any object. 
+Unfortunately this efficiency does not come without cost. 
+Structured overlays can only support exact-match queries 
+(i.e., queries that identify resources by name) as opposed
+to content-based retrieval provided by unstructured overlays. 
+This means that unstructured overlays can support more versatile 
+resource location queries such as keyword and partial matching searches. 
+Several popular file-sharing
+\p\ applications have been based on \emph{DHT}s including
+{\sl Overnet}/{\sl eDonkey}~\cite{overnet}, 
+{\sl Kademlia}/{\sl Kad Network}~\cite{maymounkov_kademlia_2002},  and 
+some flavors of {\sl BitTorrent}~\cite{c_bittorrent_2003}.  
+Moreover, \emph{DHT}s sparked
+a flurry of research for many diverse of applications ranging
+from distributed search engines to event-notification and cloud-based
+platforms~\cite{kbc_oceanstore_2000,rkcd_scribe_2001,mgpj_cloudsnap_2011}.
+
 %
@@
 %imposes very small demands on individual nodes,
-%and more speci?cally it allows nodes to join or leave
+%and more specifically?cally it allows nodes to join or leave
 %the network without signi?cantly a?ecting the sys-
@@
 
-One of the major issues that defines the efficiency of an overlay network is the
-mapping of the overlay structure to the underlying infrastructure. Recall,, in
-Figure~\ref{figure:overlay} nodes $A$ and $B$ are on the same local network,
-while $C$ and $D$ are in different networks. The top layer represents the
-overlay interconnection formed by these nodes at a higher level. Link
-connections there, can change as needed by the running environment, without any
-particular constraint by the underlying physical topology. Consider the following
-example. 
-Assume nodes $A$, $B$, $C$ and $D$ are
-connected via the physical network shown in Figure~\ref{figure:phys}, where
-the network costs are given in milliseconds, and peer $A$ sends a message to
-peer $D$.
-
-\begin{figure}
+One of the major issues that defines the efficiency of 
+an overlay network is its mapping to 
+the underlying physical infrastructure. 
+Recall, in Figure~\ref{figure:overlay} nodes $A$ and $B$ are 
+on the same local network, while $C$ and $D$ are in different networks. 
+The top layer represents the overlay interconnection formed by 
+these four nodes at a higher level. 
+Link connections there, can change as needed by 
+the running environment, without any
+particular constraint by the underlying physical topology. 
+%%
+For instance, assume nodes $A$, $B$, $C$ and $D$ being 
+connected via the physical network shown in Figure~\ref{figure:phys}
+in which costs are depicted in milliseconds. 
+%%
+\begin{figure}[ht]
 \centering
-  \includegraphics[scale=0.7]{img/pdf/example-physical.pdf}
+  \includegraphics[scale=0.4]{img/pdf/example-physical.pdf}
 \caption{Interconnection of nodes in the physical level.}
@@
 \end{figure}
-
-Depending on the configuration of the overlay,
-peers will experience different performance.
+%%
+When peer $A$ sends a message its $D$ counterpart
+and depending on the overlay configuration, 
+nodes will experience different performance.
 For example, in the overlay depicted in
@@
 Figure~\ref{figure:overlay-2}, the path will be: $3 \rightarrow 8 \rightarrow
-4$. The total cost of the first sequence is $45 ms$, while the second yields a
-mere $15 ms$. Therefore, we can conclude that the second overlay is more
-congruent with the underlying physical network than the first one and, thus, is
-more efficient.
+4$. 
+The respective costs are $45 ms$ and $15 ms$.
+% The total cost of the first sequence is $45 ms$, while the second yields a
+% mere $15 ms$. 
+Evidently, the second overlay is more
+congruent with the underlying physical network in comparison 
+to the first and, thus, is more efficient.
 
@@
 \centering
-\subfigure[Nodes B and C have direct overlay connection.] {
-  \includegraphics[scale=0.5]{img/pdf/example-overlay-inefficient.pdf}
+\subfigure[Nodes $B$ and $C$ have direct overlay connection.] {
+  \includegraphics[scale=0.4]{img/pdf/example-overlay-inefficient.pdf}
   \label{figure:overlay-1}
 }\qquad\qquad
-\subfigure[Nodes A and D have direct overlay connection.] {
-  \includegraphics[scale=0.5]{img/pdf/example-overlay-efficient.pdf}
+\subfigure[Nodes $A$ and $D$ have direct overlay connection.] {
+  \includegraphics[scale=0.4]{img/pdf/example-overlay-efficient.pdf}
   \label{figure:overlay-2}
@@
 
-Ideally, an overlay network should be constructed to achieve an optimal 
+Ideally, an overlay should be constructed to achieve an \emph{optimal} 
 mapping between the overlay paths and the underlying network links to avoid
-inefficient states where redundant physical resources are utilised for the
-operation of the same link in the overlay. The problem of constructing an optimal
+inefficient states where redundant physical resources are utilized for the
+operation of the same link in the overlay. 
+The problem of constructing an optimal
 overlay is referred to as the \emph{topology mismatch problem}, and is formally
@@
 \begin{definition}
-Let $V = \{v_1, ..., v_n\}$ be a set of points denoting the network nodes,
+Let $V$=$\{v_1$, $\dots$, $v_n\}$ be a set of points denoting the network nodes,
 $\{v_i, v_j\} \in E$ be the set of unicast distances between nodes $v_i$ and
@@
 Early incarnations of the overlay protocols did not make use of the optimal
-mapping with the underlying network topology. Gnutella, for example, is
-considered far from scalable \cite{ritter_gnucantscale_2001} because every peer
+mapping with the underlying network topology. {\sl Gnutella}, for example, is
+considered far from scalable~\cite{ritter_gnucantscale_2001} since every peer
 chooses its neighbors in the overlay without any knowledge of the underlying
-network, resulting in a mismatch between the two topologies. Queries may be
-flooded over multiple paths merging in one peer while one of the paths would
-suffice.  Moreover, peers may forward the same message to each other
+network, resulting in mismatch. 
+Queries may be flooded over multiple paths merging in one peer 
+while one of the paths would suffice.  
+Moreover, peers may forward the same message to each other
 before they receive the query messages from each other.
-\cite{matei_mapgnutella_2002} showed that, even when 95\% of any two nodes are
-less than 7 hops away from each other, a \emph{flooding-based} algorithm, like
-the one used by Gnutella, can generate 330TB/month in a 50.000 node network. A
-topology mismatch can thus have a serious effect on Internet router performance.
+\cite{matei_mapgnutella_2002} showed that, even when $95$\% of any two nodes 
+are less than $7$ hops away from each other, 
+a \emph{flooding-based} {\sl Gnutella}-like algorithm 
+can generate $330$TB/month in a $50,000$ node network. 
+A topology mismatch can thus have a serious effect 
+on Internet router performance.
 
@@
 This means that even though the target node can be reached with a logarithmic
-number of overlay hops, the distance traveled in the underlying network, during the
-overlay routing process, can be far from optimal.  The ratio of the actual IP
-network distance a query travels to reach an object of interest (via overlay
-routing) to the minimal IP network distance to that object (i.e. through IP) is
-known as \emph{stretch} or \emph{Relative Delay Penalty (RDP)} \cite{CRZ2000}.
+number of overlay hops, the distance traveled in the underlying network, 
+during the overlay routing process, can be far from optimal.  
+%%AD the above text is disconnected with what follows just below... 
+%%	- does it give a metric for the "optimality" we seek???
+The ratio of the actual \emph{IP} network distance a query travels to 
+reach an object of interest (via overlay routing) to 
+the minimal \emph{IP} network distance to that object 
+(i.e. through \emph{IP}) is known as \emph{stretch} 
+or \emph{Relative Delay Penalty (RDP)}~\cite{CRZ2000}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-A topology unaware overlay network is able to control the sequence of peers a
+A topology--unaware overlay network is able to control the sequence of peers a
 message traverses before reaching its destination, but it is completely unaware
 of how the actual packets are switched at the underlying infrastructure along
-the overlay path. For example, a single logical point-to-point link on the
+the overlay path. 
+In particular, a single logical point-to-point link on the
 overlay typically corresponds to multiple physical links in the
-underlying layer. Additionally, a link in the underlying network often lies
-on several overlay paths causing increase of the traffic on the
-link, known as the link's \emph{stress}~\cite{CRSZ2002}. Even in the case
-of an approach that provides mechanisms to a new-coming node to detect close-by
-peers at join time, the stochastic behavior with which peers join and leave
-the overlay network, called the network's churn, further undermines the effort
-for maintaining the best position in the network throughout the life-time of a
-peer.
+underlying layer. 
+Moreover, a link in the physical network often lies
+on several overlay paths inadvertently increasing the traffic on the
+link; this is known as the link's \emph{stress}~\cite{CRSZ2002}. 
+There is also the stochastic phenomenon during which peers 
+depart and others arrive in the network, known 
+as \emph{churn}~\cite{stutzbach_churn_2006}.
+Even when an approach offers the best possible location
+when a node joins in, the churn is set to  
+continuously undermine the effort to maintain the best possible position
+of the  peer in the network.
+%%AD the paragraph below is ... LONG - rephrased as above...
+% Even in the case of an approach that provides mechanisms 
+% to a new-coming node to detect close-by peers at join time, 
+% the stochastic behavior with which peers join and leave
+% the overlay network, called the network's churn, 
+% further undermines the effort for maintaining 
+% the best position in the network throughout the life-time of a peer.
 
 The topology mismatch problem is exacerbated by the extremely large amounts of
-traffic generated by highly popular P2P applications (refer to the discussion
-made in Section~\ref{section:intro}). Thus, solving the topology mismatch
-problem is a fundamental challenge in contributing to the overall efficiency of
-the Internet. Unfortunately, topology mismatch is known to be an NP-Hard problem
-\cite{NPBOOK,C2000}. Adding to the complexity of the problem
-is the fact that end-to-end latencies demonstrate \emph{triangle inequality
+traffic generated by highly popular \p\ applications~\cite{byl-book}.
+Addressing this problem remains a fundamental challenge 
+in contributing to the overall efficiency of the Internet. 
+Unfortunately, topology mismatch is known to be 
+an \emph{NP-Hard} problem \cite{C2000,NPBOOK}. 
+There is also additional complexity due to the fact that 
+end-to-end latencies demonstrate \emph{triangle inequality
 violations (TIVs)} which are a consequence of the Internet's structure and
-routing policies and thus will remain a property of the Internet for the
-foreseeable future \cite{zheng_irprtt_2005}. TIVs affect both network coordinate
+routing policies; as such, they will remain a property of the Internet 
+for the foreseeable future \cite{zheng_irprtt_2005}. 
+\emph{TIV}s affect both network coordinate
 \cite{cox_vivaldi_2004,wong_meridian_2005} and positioning \cite{ng_gnp_2001}
-systems and makes the construction of IP-topology-aware overlays difficult.
+systems and makes the construction of \emph{IP}-topology-aware 
+overlays difficult.
 
-Over the past several years, researchers have proposed and studied an extensive
-array of heuristic solutions to the topology mismatch problem. The purpose of
-this survey is to gather and organize the extensive knowledge that has been
-produced. We hope to assist the reader to understand the different approaches
-that have been proposed to tackle the topology mismatch problem. In order to do
-so, we dedicate a discussion subsection after the presentation of the algorithms
-for both unstructured and structured architectures. This section offers:
-\begin{inparaenum}[\itshape i\upshape)]
-  \item further discussion that can help the reader identify, through
-a taxonomy, the pluses and minuses  of the presented algorithms,
-  \item a comprehensive tabular organization of the key-point information, and
-  \item an attempt to give a pictorial comparison of the relative performance of
-the algorithms.
-\end{inparaenum}
+Over the past several years, researchers have investigated an extensive
+array of heuristic solutions to the topology mismatch problem. 
+The purpose of this survey is to gather and organize 
+this extensive body of produced knowledge. 
+We aspire to assist the reader in better understanding the 
+different approaches suggested and to accomplish this we 
+present methods for both structured and unstructured distributed 
+overlay networks.
+We outline the respective algorithmic aspects and we categorize
+the surveyed approaches so that pros and cons of each method 
+can become apparent in their context.
 
-We hope to offer the reader a deeper insight for the observations that
-pushed research toward some direction. Thus, we try to categorize these
-approaches so that we can organize directions making them comprehensible
-and helping the reader build critical thinking on the advantages and
-disadvantages of each approach.
+We employ both tabular and pictorial means to 
+articulate features of the proposed approaches and outline 
+both strengths and weaknesses.
+To ascertain the effectiveness of each proposal, we predominantly use 
+three criteria that capture the behavior of the algorithms 
+discussed:
+\emph{efficiency}, \emph{overhead} and \emph{scalability}.
+For each of these criteria, we use a coarse-grained,
+three-level, gradation to denote how well the algorithms do,
+namely \emph{low}, \emph{medium} and \emph{high}.
+Each criterion has its own
+interpretation of the above literal values. 
+\emph{Efficiency} means how well the
+algorithm performs its search and topology adaptation functions, without
+affecting other characteristics of the networks 
+(e.g. BitTorrent's optimal download performance). 
+%%AD I am not sure what the above line says.. or connects. Something is 
+%%	missing...
+Efficiency is negatively effected every time an algorithm 
+reduces its search scope.
+For the \emph{overhead} criterion, we track the cost 
+to compute and communicate peer join, message forwarding, 
+and topology maintenance. 
+For example, when exchanging route tables only amongst neighbors, 
+we consider the traffic generated to be low.
+%%AD Again I am not sure how the example above ties with the definition...
+For the \emph{scalability} criterion, we consider system behavior 
+with respect to the number of participating peers. 
+We consider systems with low, medium, and high scalability 
+to mean they can support hundreds, tenths of thousands, 
+and millions of nodes, respectively.
 
-Tables provide a one-stop-shop overview of the algorithms and serve as a 
-tool both as a companion to the reader while reading the survey and a reference for
-future use.
+% we dedicate a discussion subsection after the presentation of the algorithms
+% for both unstructured and structured architectures. This section offers:
+% \begin{inparaenum}[\itshape i\upshape)]
+  % \item further discussion that can help the reader identify, through
+% a taxonomy, the pluses and minuses  of the presented algorithms,
+  % \item a comprehensive tabular organization of the key-point information, and
+  % \item an attempt to give a pictorial comparison of the relative performance of
+% the algorithms.
+% \end{inparaenum}
 
-Pictorial comparison consists of bar charts that give a relative feeling of the
-performance of the various algorithms, discussed in the corresponding section,
-in
-terms of \emph{efficiency}, \emph{overhead} and \emph{scalability}; three
-important characteristics for every distributed algorithm. For each of these
-criteria, we use a
-coarse-grained, three-level, gradation to denote how well the algorithms do,
-namely \emph{low}, \emph{medium} and \emph{high}. Each criterion has its own
-interpretation of the above literal values. Efficiency means how well the
-algorithm performs its search and topology adaptation functions, without
-affecting other characteristics of the networks (e.g. BitTorrent's optimal
-download performance). We assume that efficiency is
-negatively effected when an algorithm reduces the search scope. 
-%Moreover, the
+% We hope to offer the reader a deeper insight for the observations that
+% pushed research toward some direction. Thus, we try to categorize these
+% approaches so that we can organize directions making them comprehensible
+% and helping the reader build critical thinking on the advantages and
+% disadvantages of each approach.
+
+% Tables provide a one-stop-shop overview of the algorithms and serve as a 
+% tool both as a companion to the reader while reading the survey and a reference for
+% future use.
+% 
+% Pictorial comparison consists of bar charts that give a relative feeling of the
+% performance of the various algorithms, discussed in the corresponding section,
+% in
+% 
+% terms of \emph{efficiency}, \emph{overhead} and \emph{scalability}; three
+% important characteristics for every distributed algorithm. 
+% For each of these criteria, we use a coarse-grained, 
+% three-level, gradation to denote how well the algorithms do,
+%%  namely \emph{low}, \emph{medium} and \emph{high}. 
+% Each criterion has its own
+% interpretation of the above literal values. Efficiency means how well the
+% algorithm performs its search and topology adaptation functions, without
+% affecting other characteristics of the networks (e.g. BitTorrent's optimal
+% download performance). We assume that efficiency is
+% negatively effected when an algorithm reduces the search scope. 
+% %Moreover, the
 
@@
 %efficiency.
-For the overhead criterion, we track the cost to compute and communicate peer
-join, message forwarding, and topology maintenance. For example, when exchanging
-route tables only amongst neighbors, we consider the traffic generated to
-be low.
-%Also duplicates minimization is a plus and lazy or heart-beat-based
-%overlay control mechanisms impose different amount of stress to the network.
-%Nevertheless, caching/replication helps reducing the overall overhead by
-%constraining the query from spreading too much. Low convergence speed of
-%minimum-spanning-tree-based approaches, we mentioned above for efficiency,
-%results also in additional computational and communicational overhead.
-% TODO: We check
-% complexity in the asymptotic sense, using \emph{Big O} notation % in order to
-% check how the algorithm responds t. We take into account topology % creation and
-% topology maintenance...
-% TODO: HELP NEEDED!
-For the scalability criterion, we consider system behavior with respect to the number of
-participating peers. We consider systems with low, medium, and high scalability 
-to mean they can support 
-hundreds, tenths of thousands, and millions of nodes,
-respectively. 
-%Fully distributed approaches is a plus (no system-centric entities
-
-%involved, no need for global knowledge) and topology adaptation strengthens
-
-%scalability in highly dynamic environments. Reduced search scope, on the other
-
-%hand, penalizes the scalability since the more peers coming in the network the
-
-%more potentially unsatisfied queries will there be. Furthermore, clock
-
-%synchronization is a limiting factor to scalability of a distributed system.
-
-%
-% TODO: SOME DISCUSSION FROM \cite{XTZ2003) CONTRADICTS TO THE
-%TRIANGLE INEQUALITY ARGUMENT ABOVE
-%
-%[snip]
-%Studies [14] have shown that triangle inequality may not hold in Internet
-%topology. In fact, study from Pastry has shown that the proximity approximation
-% is much worse when using the Mercator topology that is based on the real
-% measurements of the Internet [3].
-%

### 2014-01-28 conclusion.tex
2013.01.23 (Alex) Introduction, background and conclusion sections reworked.

--- 2014-01-27 conclusion.tex
@@
 
-Research and development for computer systems
-built using the \p\ paradigm have flourished  in the last decade.
+Research and development on application systems that follow the 
+\p\ architectural paradigm, have flourished over the last decade.
 By and large, such systems exploit the loose-coupling and self-organization 
 of participating nodes to shape substrates upon which diverse applications 
-can run; the latter offer services including but not limited to 
+can run; the latter offer services including, but not limited to, 
 distributed naming, 
@@
 web crawling and caching, 
-event notification and subscribe/publish.
+event notification and subscribing/publishing.
 These services are provided atop the 
-best-effort infrastructure of today's Internet 
-and they have to effectively deal with an array of key issues
-emanating from the fact that \p\ nodes 
-operate at the edge of the Internet; such critical issues include
-node availability, robustness, scalability, load-sharing,
-quality-of-service and user anonymity.
-The majority of such issues emerges as 
-application-level \p--networks are formed without paying 
-adequate attention to the design of the physical network beneath.
-The latter constitutes what is known as the
-topology mismatch problem between the \p-overlays and the physical networks.
-In addition, the development of \p--systems and applications
-typically involved juggling a set of tradeoffs (e.g.,
-choosing higher accuracy at the cost of increased system overhead.)
+best-effort infrastructure of today's Internet;
+services have to comply with an array of key requirements and 
+offer quality attributes that include node availability, robustness,
+scalability, load-sharing, quality-of-service and user anonymity. 
+%%AD rephrased below as above - too long sentence it was!
+% and they have to effectively deal with an array of key quality attributes that
+% they may need or have to offer node availability, robustness,
+% scalability, load-sharing, quality-of-service and user anonymity.
+%%
+The difficulty to handle such characteristics effectively, 
+emanates from the fact that
+peers  operate at the edge of the Internet and no 
+adequate attention is paid to the structure of the physical network 
+during application-level network design. 
+The deviation of the logical distributed overlay formation from the 
+physical network yields suboptimal use of the underlying 
+infrastructure and is known as the topology mismatch
+problem.
+%%AD rephrased what is below as above..
+% The deviation of the structure of the overlay network from the
+% optimal, according to the underlying physical structure, constitutes what is
+% termed topology mismatch and the problem is known as the topology mismatch
+% problem.
 
+Development of \p--systems and applications
+typically involve juggling a set of tradeoffs (e.g.,
+choosing higher accuracy at the cost of increased system overhead).
 % Computer systems design is typically about juggling a set of tradeoffs (e.g.,
@@
 Nonetheless, by
-1)~presenting an analysis of each of the solutions, 
+\begin{inparaenum}[\itshape i\upshape)]
+  \item presenting an analysis of each of the solutions,
 including what distinguishes it from the
-others and its advantages and disadvantages and 
-2)~offering a pictorial
-comparison of how each solution fares compared with the others with regards
-to efficiency, overhead, and scalability, we hope to have provided \p\
-researchers and developers with enough perspective so that depending 
-on the needs of their particular application, they are able to select
-amongst these surveyed solutions or propose a new solution of their own.
+others and its pros and cons and
+  \item offering a pictorial
+comparison of how each solution fares in regards to others as far as
+efficiency, overhead, and scalability are concerned,
+\end{inparaenum}
+we hope to have provided \p\
+developers and researchers with enough insight and perspective;
+as they face specific problems, they will be readily able to 
+draw the best possible design decisions for their own application
+environments.
+%%AD sentence was too long - this "giant" thing to much ass-kissing... no.
+% when facing a specific problem, they would 
+% when they face a specific problem, th 
+% so that depending
+% on the particular problem in hand, they are able to make a
+% correct choice amongst these surveyed here or, by firmly standing on
+% the shoulders of giants, propose a new solution of their own.

### 2014-01-28 intro.tex
2013.01.23 (Alex) Introduction, background and conclusion sections reworked.

@@
 In addition, a startling number of diverse and successful
-applications have been built based on a \p\ architecture. Some of them include:
+applications have been built based on a \p\ architecture. 
+Some of them include:
 distributed search engines~\cite{yaci}, 
@@
 Web caches, archives and publishing systems~\cite{ird_squirrel_2002,bags_youserv_2002,wrc_publius_2000,wm_tangler_2001},
-messaging applications~\cite{threedegrees}, 
+messaging and dissemination applications~\cite{threedegrees,icpp08-pd}, 
 event-notification infrastructures~\cite{rkcd_scribe_2001,cdkr_scribe_2002,agebh_dks_2003}, 

### 2014-01-28 mard-survey.tex
2013.01.23 (Alex) Introduction, background and conclusion sections reworked.

@@
 \emph{Peer-to-peer} (\p) systems have enjoyed immense attention and have been
-widely deployed on the Internet for, well, over a decade.  
+widely deployed on the Internet for well over a decade.  %% no commas here
 They are implemented via an overlay network abstraction, on top 
@@
 load balancing, quality of service, scalability, to name just a few.
-Unfortunately, weaknesses of early deployments of \p\ systems, prevented 
+Unfortunately, inherent weaknesses of \p\ systems, prevented 
 applications from leveraging these features.
@@
 We provide a fine-grained categorization of the
-proposed solutions and discuss the advantages, disadvantages, and novelties
-introduced by each of them.
+suggested solutions and discuss the novelty as well as 
+the  advantages and disadvantages of each such proposal.
 Finally, we provide an analysis as well as pictorial comparisons 
@@
 reference for developers, system architects and researchers in the field.
-% in an attempt to provide a comprehensive quick one-stop-shop
-% reference for anyone interested in the topic, from researchers and students,
-% to system architects and developers.
 

### 2014-01-28 background.tex
2013.01.24 (Vassilis) Addressing Alex's comments in previous iteration for background and conclusion sections.

@@
 \p\ architectures that have been implemented and deployed. We also
-describe and motivate more formally the topology mismatch problem 
+describe and motivate more formally the topology mismatch problem, 
 whose various aspects have been tackled over the past several years.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Circa 2000, the first file-sharing \p-system introduced the \emph{servent}
-concept~\cite{gnutella}, a \emph{portmanteau} that blends the notions of server
-and client to denote the dual role of a node participating in \p-network. 
+%% VM: Don't know if we should start with this Gnutell specific material maybe
+% it needs to go to the respective paragraph of dece.unstructured bellow.
+Circa 2000, pioneering file-sharing \p-system,
+Gnutella~\cite{gnutellav04,gnutella} introduced the \emph{servent }concept, a
+\emph{portmanteau} that blends the notion of server
+and client to denote the dual role of a node participating in its network.
 Such server-less systems have proven to be able to achieve outstanding
-aggregate resource capacities as more and more participants join the system
-without requiring additional expenditure for infrastructure. 
+aggregate resource capacities, as more and more participants join the system,
+but without the additional expenditure for infrastructure.
 While certain undesirable features such as 
@@
 issues \cite{hughes_socp2p_2008}, have arisen, \p\ systems have
-continued to gain popularity throughout the years primarily 
+continued to gain popularity throughout the years, primarily 
 because they have successfully supported popular applications 
@@
 to recognize that requests for resources 
-(e.g., CPU cycles or popular content) need not be sent to
-a dedicated server. Instead, requests could be handled by 
+(e.g., CPU cycles or popular content) need not be fulfilled by
+a dedicated server. Instead, requests could be satisfied by
 many hosts that already possessed the resources in question. 
-The downside of this approach was that it required 
-centralized search that became a single point of failure, a
-scalability bottleneck, and a target of malicious including \emph{DoS} attacks. 
+The downside of this approach was that it required a
+centralized search service, that was instantly rendered a single point of
+failure,  a scalability bottleneck, and a target of malicious acts including
+\emph{DoS attacks}.
 % \emph{Denial-of-Service (DoS) attacks}).
@@
 {\sl Napster} was the most successful incarnation of the
-centralized approach in file-sharing~\cite{napster}.
+centralized approach, in file-sharing~\cite{napster}\footnote{Ofcourse, the
+SETI@home project is considered the pioneer, but on a different application
+domain \cite{seti}.}.
 It maintained a \emph{central index server} based on file-lists 
@@
 yield pointers to the actual content.
-Thus, by centralizing search while distributing downloads,  {\sl Napster}
+Thus, by centralizing search, while distributing downloads, {\sl Napster}
 achieved a highly functional design that, at its height, attracted $26$ million
-users sharing approximately $80$ million songs\cite{jmm_naptopusage_2001}.
+users, sharing approximately $80$ million music
+files\cite{jmm_naptopusage_2001}.
 %%
-The central indexing played a vital role in the demise of 
-{\sl Napter} as the \emph{Recording
+The central indexing, though, played a vital role in the demise of 
+{\sl Napter}. After \emph{Recording
 Industry Association of America} successfully argued in court
-copyright infringement of the artists songs traded in the system. 
+copyright infringement of the artists' work, it was then easy to point to the
+indexing server and ultimately bring it down.
 %
@@
 To save on bandwidth, unstructured networks typically limit how long 
-a query travels within the overlay via the \emph{time-to-live} parameter
-(\emph{TTL}). 
-When a query is received by a node, it decreases its \emph{TTL} by one unit. 
-If the \emph{TTL} has not reached zero, the node forwards the query 
-to its neighbors.  Otherwise, it drops the query. 
+a query travels within the overlay via the \emph{time-to-live} mechanism. 
+When a query is received by a node, it decreases its \emph{TTL} parameter by one
+unit. If the \emph{TTL} value has not reached zero, the node forwards the
+message to its neighbors; otherwise, it drops it.
 While this approach prevents queries from visiting the entire network, it
@@
 % of a musical group from the 1930s) 
-will be ultimately found.
+will, ultimately, be found.
 %%%
@@
 \centering
-\subfigure[Centralized ({\sl SETI@Home} or {\sl Napster}). Peers query the centralized
+\subfigure[Centralized ({\sl SETI@Home} or {\sl Napster}). Peers
+query the centralized 
 entity (dotted lines) and then establish the connection to the data source.] {
@@
 (Figure~\ref{figure:p2p-archs:structured}) sought to address the 
-disadvantages of the above decentralized unstructured \p--approaches.
-Here, the main objective was to offer 
+disadvantages of the, above mentioned, decentralized unstructured
+\p--approaches. Here, the main objective was to offer 
 a self-organizing infrastructure for large-scale \p\ applications
@@
 Moreover, \emph{DHT}s sparked
-a flurry of research for many diverse of applications ranging
+a flurry of research for many diverse of applications, ranging
 from distributed search engines to event-notification and cloud-based
@@
 %%
-When peer $A$ sends a message its $D$ counterpart
+When peer $A$ sends a message to its $D$ counterpart
 and depending on the overlay configuration, 
@@
 
+%% VM The next part was previously at the end of this paragraph and didn't seem
+%% to fit right there. I moved it here since it is a theoretical optimality
+%% definition used by some authors we review and I think it can help the reader
+%% better understand what problem. Also I have rewritten it.
+%%
+The ratio of the physical distance (e.g., in hops) a query travels to
+reach its target via overlay routing, to the actual minimal network distance via
+IP routing, is termed in the literature as \emph{stretch} or \emph{Relative
+Delay Penalty (RDP)}~\cite{CRZ2000}.
+%%
 Ideally, an overlay should be constructed to achieve an \emph{optimal} 
@@
 during the overlay routing process, can be far from optimal.
-%%AD the above text is disconnected with what follows just below... 
-%%	- does it give a metric for the "optimality" we seek???
-The ratio of the actual \emph{IP} network distance a query travels to 
-reach an object of interest (via overlay routing) to 
-the minimal \emph{IP} network distance to that object 
-(i.e. through \emph{IP}) is known as \emph{stretch} 
-or \emph{Relative Delay Penalty (RDP)}~\cite{CRZ2000}.
 
@@
 In particular, a single logical point-to-point link on the
-overlay typically corresponds to multiple physical links in the
+overlay, typically corresponds to multiple physical links in the
 underlying layer. 
@@
 of the  peer in the network.
-%%AD the paragraph below is ... LONG - rephrased as above...
-% Even in the case of an approach that provides mechanisms 
-% to a new-coming node to detect close-by peers at join time, 
-% the stochastic behavior with which peers join and leave
-% the overlay network, called the network's churn, 
-% further undermines the effort for maintaining 
-% the best position in the network throughout the life-time of a peer.
 
+%% VM: This section has a lot of information that I belieave are worth
+%% mentioning on the paper's conclusion section. We say for example why no
+%% approach we reviewed is best of all (cause it is an NP-Hard problem); or why
+%% virtual coordinate approaches will never be a reliable tool in today's
+%% Internet. Also it doesn't add much to the purpose of this section that is to
+%% say what motivated us to write the paper. If you also agree we can move it.
 The topology mismatch problem is exacerbated by the extremely large amounts of
@@
 overlays difficult.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 array of heuristic solutions to the topology mismatch problem.
+%% VM: If we move the above though we have to mention why all solutions are
+%% heuristic.
 The purpose of this survey is to gather and organize 
@@
 We aspire to assist the reader in better understanding the 
-different approaches suggested and to accomplish this we 
+different approaches suggested and to accomplish this, we 
 present methods for both structured and unstructured distributed 
@@
 We outline the respective algorithmic aspects and we categorize
-the surveyed approaches so that pros and cons of each method 
+the surveyed material so that pros and cons of each method
 can become apparent in their context.
@@
 \emph{Efficiency} means how well the
-algorithm performs its search and topology adaptation functions, without
-affecting other characteristics of the networks 
-(e.g. BitTorrent's optimal download performance). 
-%%AD I am not sure what the above line says.. or connects. Something is 
-%%	missing...
-Efficiency is negatively effected every time an algorithm 
-reduces its search scope.
-For the \emph{overhead} criterion, we track the cost 
+algorithm performs its search and topology adaptation functions (if any). There
+can be many criteria that effect this metric, either positivelly or negatively.
+For example, efficiency, is negatively effected every time an algorithm
+reduces the system's average query search scope, or when the method employed
+uses an unreliable metric to estimate physical proximity.
+For the \emph{overhead} criterion, we roughly assess the cost
 to compute and communicate peer join, message forwarding, 
-and topology maintenance. 
-For example, when exchanging route tables only amongst neighbors, 
-we consider the traffic generated to be low.
-%%AD Again I am not sure how the example above ties with the definition...
-For the \emph{scalability} criterion, we consider system behavior 
-with respect to the number of participating peers. 
-We consider systems with low, medium, and high scalability 
-to mean they can support hundreds, tenths of thousands, 
-and millions of nodes, respectively.
+and topology maintenance. As an example, algorithms that dictate only local
+exchange of routing information during topology adaptation, can be safelly
+considered to incur far less overhead, to the system, than those which
+require global knowledge.
+With \emph{scalability} we consider system behavior
+with respect to the number of participating peers. We consider systems with low,
+medium, and high scalability to mean they can drive hundreds, tenths of
+thousands, and millions of nodes, respectively.
+The heuristic nature of the solutions, inherently involve balancing trade-offs;
+so we pay extra attention how enhancing the above criteria may, in response,
+effect other aspects of the application. If, for example, tuning
+topology-awareness of the BitTorrent environment results in the degradation of
+the unrivaled optimal download performance of the vanilla-protocol, then this
+cannot be considered a successful answer to the problem.
 

### 2014-01-29 background.tex
2013.03.29 (Alex) Section on unstructured algorithms, massively reworked (from formatting to phrasing). Minor updates to background and conclusion sections.

--- 2014-01-28 background.tex
@@
 affecting other characteristics of the networks 
-(e.g. BitTorrent's optimal download performance). 
+(e.g., BitTorrent's optimal download performance). 
 %%AD I am not sure what the above line says.. or connects. Something is 

### 2014-01-29 conclusion.tex
2013.01.24 (Vassilis) Addressing Alex's comments in previous iteration for background and conclusion sections.
2013.03.29 (Alex) Section on unstructured algorithms, massively reworked (from formatting to phrasing). Minor updates to background and conclusion sections.

@@
 
-Research and development on application systems that follow the 
-\p\ architectural paradigm, have flourished over the last decade.
+Research and development in application systems that follow the 
+\p\ architectural paradigm have flourished for more than a decade now.
 By and large, such systems exploit the loose-coupling and self-organization 
@@
 event notification and subscribing/publishing.
+%%
 These services are provided atop the 
-best-effort infrastructure of today's Internet;
-services have to comply with an array of key requirements and 
-offer quality attributes that include node availability, robustness,
+best-effort infrastructure of today's Internet 
+and they have to effectively deal with an array of key quality attributes that
+they may have to offer, such as node availability, robustness,
 scalability, load-sharing, quality-of-service and user anonymity.
-%%AD rephrased below as above - too long sentence it was!
-% and they have to effectively deal with an array of key quality attributes that
-% they may need or have to offer node availability, robustness,
-% scalability, load-sharing, quality-of-service and user anonymity.
 %%
-The difficulty to handle such characteristics effectively, 
-emanates from the fact that
-peers  operate at the edge of the Internet and no 
-adequate attention is paid to the structure of the physical network 
-during application-level network design. 
-The deviation of the logical distributed overlay formation from the 
+The difficulty to handle such features effectively, emanates from the fact that
+peer nodes operate at the edge of the Internet and no adequate attention 
+is paid to the structure of the physical network during the 
+network formation at the application-level.
+The deviation of the distributed logically-formed overlay from the 
 physical network yields suboptimal use of the underlying 
@@
 problem.
-%%AD rephrased what is below as above..
-% The deviation of the structure of the overlay network from the
-% optimal, according to the underlying physical structure, constitutes what is
-% termed topology mismatch and the problem is known as the topology mismatch
-% problem.
 
@@
   \item presenting an analysis of each of the solutions,
-including what distinguishes it from the
-others and its pros and cons and
+including what distinguishes it from 
+others as well as its advantages and disadvantages and 
   \item offering a pictorial
@@
 environments.
-%%AD sentence was too long - this "giant" thing to much ass-kissing... no.
-% when facing a specific problem, they would 
-% when they face a specific problem, th 
-% so that depending
-% on the particular problem in hand, they are able to make a
-% correct choice amongst these surveyed here or, by firmly standing on
-% the shoulders of giants, propose a new solution of their own.
+%%AD let's avoids "giants" etc...

### 2014-01-29 mard-survey.tex
2013.03.29 (Alex) Section on unstructured algorithms, massively reworked (from formatting to phrasing). Minor updates to background and conclusion sections.

@@
 load balancing, quality of service, scalability, to name just a few.
-Unfortunately, inherent weaknesses of \p\ systems, prevented 
+Unfortunately, inherent weaknesses of early deployments of \p\ systems, prevented 
 applications from leveraging these features.
@@
 and structured \p\ systems.   
-We provide a fine-grained categorization of the
-suggested solutions and discuss the novelty as well as 
-the  advantages and disadvantages of each such proposal.
-Finally, we provide an analysis as well as pictorial comparisons 
+We provide a fine-grained categorization of the suggested solutions
+by discussing their novelty, advantages and weaknesses.
+Finally, we offer an analysis as well as pictorial comparisons 
 of the reviewed approaches, since we aim to offer a comprehensive

### 2014-01-29 unstructured.tex
2013.03.29 (Alex) Section on unstructured algorithms, massively reworked (from formatting to phrasing). Minor updates to background and conclusion sections.

@@
-\section{Unstructured P2P Networks}
+\section{Unstructured \p\ Networks}
 \label{section:unstructured}
 
-This section is dedicated to unstructured P2P networks. Algorithms
-for tackling the topology mismatch problem in such
-networks are discussed, analyzed, and categorized based on the use of the
-overlay structure, the message forwarding approach for peer communication, and
-techniques proposed for detecting proximity in order to optimize overlay
-topology.
+In this section, we present algorithms that tackle the topology mismatch
+problem in unstructured \p\ networks and classify them based on their 
+use of the overlay structure, their message forwarding scheme 
+for peer communication as well as techniques used for detecting proximity 
+in order to optimize their overlay topology.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Improving search in peer-to-peer networks}
+% \subsubsection{Improving search in peer-to-peer networks}
 
@@
 
-\cite{YG-M2002} proposes an easy to implement and practical solution to the
-inefficiency of Gnutella's ``blind flooding'' approach. The paper introduces
-three different forwarding methods, namely \emph{iterative deepening},
+To improve over {\sl Gnutella}'s ``blind flooding'' approach,
+\cite{YG-M2002} proposed a practical and easy to implement solution 
+weaved around three different message forwarding methods:
+\emph{iterative deepening},
 \emph{directed BFS}, and \emph{local indices}.
-
-In iterative deepening, %(see Figure~\ref{figure:dbfs:iterdeep})
+%%AD local indices is not "a message forwarding method" really...  but ok.
+
+In \emph{iterative deepening}, 
+%(see Figure~\ref{figure:dbfs:iterdeep})
 the search is
-performed on a BFS tree with multiple preset depths. The depth limit is
-iteratively increased by the source node for each query, based on the quality of
-the results. The source node may issue a new request with increased depth limit
-that will trigger the nodes at the last depth level to resume the search. The
-iterative approach avoids restarting the whole search process from scratch at
-each iteration and reduces the load of the nodes on the upper levels of the
-tree. Its major drawback is the delay between successive iterations, as the
-source node needs to examine the results at each iteration before deciding to
-quit or resume the query.
-
-The directed BFS
-tries to avoid this delay by forwarding the query messages only to a select
-set of neighbors, in which the selection criteria varies from the number of
+performed on a BFS tree with multiple preset depths. 
+The depth limit is iteratively increased by the source node for each query, 
+based on the quality of the results. 
+The source node may issue a new request with increased depth limit
+that will trigger the nodes at the last depth level to resume the search. 
+The iterative approach avoids restarting the entire search process 
+from scratch at each iteration and reduces the load on the nodes of 
+the upper levels of the tree. 
+Its major drawback is the delay between successive iterations, as the
+source-node has to examine the results at each iteration before deciding to
+either quit or resume the query.
+
+The \emph{directed BFS}
+tries to avoid the aforementioned delay by forwarding 
+the query messages only to a select set of neighbors, in which the selection criteria varies from the number of
 results received previously, distance in terms of hops, bandwidth, or the query
 load of the neighbor.
-
-In local indices
-each node
-indexes data within a radius of $r$ hops and uses this local index to answer
-queries on behalf of them without generating additional traffic. Local indices
-greatly reduce the aggregate bandwidth usage of the network and improves query
-efficiency, however, updating the indices in cases with frequent node joins and
-leaves introduces a serious overhead into the system if the radius is kept broad.
-
+%
+In \emph{local indices},
+each node indexes data within a radius of $r$ hops and uses 
+this local index to answer queries % on behalf of them 
+without generating additional traffic. Local indices
+greatly reduce the aggregate bandwidth usage of the network and 
+improve query efficiency. However, updating such indices in 
+the presence of frequent node joins/departures 
+may introduces significant overhead should the radius is kept broad.
+Below, we outline how the techniques in~\cite{YG-M2002}
+fare against the $3$ major performance criteria 
+of Section~\ref{section:background}.
 % \begin{center}
@@
 \begin{center}
+{\footnotesize
 \begin{tabular}{rccc}
@@
 % processing power needs to be consumed since more nodes are involved (death
-% spirral).
+% spiral).
 medium &
@@
 \end{tabular}
+}
 \end{center}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{ \bf Delay Aware P2P System}
-A new \emph{Delay Aware P2P System (DAPS)} is introduced in \cite{ZL2005}. Its
-main goal is to reduce the time of a look-up request by dividing the routing
-tables of peers into several sectors of increasing delay. The source node
-emitting the query message defines a delay boundary or the \emph{pruning
-factor} in the paper's context. Request messages are then forwarded only to
-nodes whose delay is less than or equal that boundary. 
+% \paragraph*{ \bf Delay Aware P2P System}
+% A new \emph{Delay Aware P2P System (DAPS)} is introduced in \cite{ZL2005}. 
+A \emph{delay--aware} \p\ system termed {\sl DAPS} 
+was introduced in~\cite{ZL2005}.
+{\sl DAPS}  seeks to attain reduced look-up times by dividing 
+peer routing tables into several sectors of increasing delay. 
+The source node that issues the query designates the delay 
+boundary it may tolerate providing so a \emph{pruning factor}.
+In this context, user requests are forwarded only to 
+nodes whose expected delay is less than or equal to the indicated boundary. 
+{\sl DAPS} primarily focuses on user ``experience'' and deploys a
+end-to-end delay monitoring mechanism that  may enable
+the clustering of routing tables. 
+In an dynamic environment, the formation 
+of very accurate such routing tables might be however an elusive goal.
 %With the clustered
 %routing tables and the loose organization of DAPS' overlay network it is
-%considered by is between structured and unstructured.
-
+%considered by is be tween structured and unstructured.
+%%
 % TODO: MAYBE REVIEW THE EVALUATION
+In terms of the $3$ performance criteria, {\sl DAPS} stands as follows:
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Gia}
-\emph{Gia} \cite{CRBLS2003} is an effort to solve the scalability
-problem of the unstructured P2P file sharing systems, in particular Gnutella.
-The main novelty in the design of Gia is the replacement of the blind flooding
-approach of Gnutella with random walks \cite{lv_randomwalks_2002}. Although
-random walks is a step in the right direction, issuing only a single copy of the
-query within the network reduces the search scope, thus affecting negatively the
-success rate of the query.  In order to overcome this limitation, Gia introduces
-a token-based flow control mechanism, which is essentially an intelligent flow
-control algorithm that gradually redirects the queries to nodes which are more
-likely to answer. In order to prevent overloading of nodes with query requests,
-Gia uses a token-based flow control algorithm in which each node announces the
-number of query requests it can handle in terms of tokens to its neighbors, so
-that neighbors only forward query requests to nodes that they previously
-received tokens from. Gia also acknowledges the heterogeneity in peer bandwidth,
-processing power, disk speed, etc, of the nodes in P2P networks and uses this
-information when connecting nodes to each other and by using a topology
-adaptation algorithm, Gia ensures that high capacity nodes have high degrees and
-low capacity nodes are within short proximity to high capacity ones.
+% \subsubsection{Gia}
+
+The key objective of the {\sl Gia} system~\cite{CRBLS2003} is 
+to help alleviate the scalability omnipresent
+in unstructured \p\ file-sharing systems.
+At first, {\sl Gia} replaced {\sl Gnutella}'s blind flooding 
+with random walks.
+Although this adoption was 
+a step in the right direction~\cite{lv_randomwalks_2002},
+issuing a single copy of the query within the network
+effectively reduces the search scope and may negatively affect 
+the success rate of the query in question.
+To overcome this limitation, {\sl Gia} introduces
+a token-based flow control mechanism
+% , which is essentially an intelligent flow control algorithm 
+%%AD what is "intelligent"?? - I think you can live without this above line... 
+that gradually redirects queries to nodes which are more
+likely to answer. 
+%%
+This flow control mechanism also helps prevent node overloading
+as each peer ``announces'' the number of query requests it 
+can handle in terms of tokens to its neighbors; to this end, 
+peers only forward query requests to nodes that they previously
+received tokens from.
+%%AD changed below as above... I have the impression that this token 
+%%	based mechanism is one and the same (the way the writing appeared 
+%%	was that they were two mechanisms!  check!!!
+%%%%
+% In order to prevent overloading of nodes with query requests,
+% Gia uses a token-based flow control algorithm in which each node announces the
+% number of query requests it can handle in terms of tokens to its neighbors, so
+% that neighbors only forward query requests to nodes that they previously
+% received tokens from. 
+%%%%%%%%%%%%%%%%%
+{\sl Gia} also acknowledges the heterogeneity in peer bandwidth,
+processing power, disk speed, etc., of \p\ nodes and uses this
+information when connecting nodes to each other.
+By using a topology adaptation algorithm, 
+{\sl Gia} places low capacity nodes within 
+short proximity to peers with high performance features.
+This topology adaptation algorithm is based on the metric 
+each node maintains about its satisfaction --ranging between $0$ and $1$--
+for the neighbors it finds itself associated with. 
+Through message exchange, a peer can establish new connections 
+or drop superseded one in order to improve its satisfaction.
+%%AD the 2 lines below make no sense.. I rephrased as above... 
+% \emph{Gia} ensures that high capacity nodes have high degrees and
+% low capacity nodes are within short proximity to high capacity ones.
+%%%%%%%%AD this (above) phrase is almost verbatim out of the paper! NO good idea ;-)
+Despite the fact that {\sl Gia}'s topology adaptation algorithm 
+that primarily focuses on the satisfaction of peers improves the network's 
+scalability, it falls short in addressing the mismatch problem
+as considerations for the underlying network are not handled in explicit
+terms.
+%%AD the comment TODO is somewhat correct. Changed the wording here..... 
 % TODO: THIS FOLLOWING REMARK SOUNDS LIKE THE ALGORITHM SHOULDN'T BE IN THIS
 %       SURVEY AFTER ALL!!
-Although the topology adaptation algorithm Gia uses improves the scalability of
-the network, it does not help much in solving the topology mismatch problem,
-since it does not consider the underlying physical topology.
-
+% Although the topology adaptation algorithm Gia uses improves the scalability of
+% the network, it does not help much in solving the topology mismatch problem,
+% since it does not consider the underlying physical topology.
 %
@@
 %
-
+In terms of the $3$ stated criteria, {\sl Gia}'s approach is as follows:
+%%
+%%AD from the discussion I think that Efficiency might me medium-how??? 
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Distributed Cycle Minimization Protocol (DCMP)}
-We have already discussed that blind flooding often causes messages to be routed in a direction
-that is far from the path that can
-ultimately fulfill the query request. Another problem with topology unaware
-systems is the duplication of messages due to cycles, even along the correct
-forwarding path. The authors of \cite{ZKB2008} focused on that exact deficiency
-of overlay networks and introduced Distributed Cycle Minimization Protocol
-(DCMP), a dynamic, fully distributed method to remove cycles without sacrificing
-overlay connectivity, resilience and other important properties of unstructured
-architectures by avoiding a hierarchical organization of peers. In DCMP, once a
-cycle is detected, the most powerful node in that cycle is elected as the
-\emph{Gate Peer} and the cycle is then broken at a strategic place so that it
-will result in the minimization of the distance between the Gate Peer and all
-other nodes that participated in that cycle. The process is managed by using two
-special message types namely \emph{Information Collection (IC)} and \emph{Cut
-Message (CM)}. One disadvantage of DCMP is that since the distance a forwarded
-message can travel is limited by the TTL value, which is practically $7$ in most
-cases, DCMP cannot detect cycles formed by more than 7 peers. Even though cycle
-elimination improves the network performance, it does not actually solve the
-topology mismatch problem.
-
+% \subsubsection{Distributed Cycle Minimization Protocol (DCMP)}
+% We have already discussed that blind flooding often causes messages 
+% to be routed in a direction
+% that is far from the path that can
+% ultimately fulfill the query request. 
+%%AD the above is just repetition...
+
+Another problem with topology unaware
+systems is the duplication of messages due to cycles that appears 
+ even along the correct forwarding path. 
+\cite{ZKB2008}~focuses on this exact deficiency
+of overlay networks and introduces the \emph{Distributed Cycle Minimization Protocol
+(DCMP)}, a dynamic, fully distributed method that removes cycles;
+this is accomplished without sacrificing
+overlay connectivity, resilience and other key properties of unstructured
+\p\ architectures and by avoiding a hierarchical organization of peers. 
+Once a cycle is detected in {\it DCMP}, the most powerful node in that cycle 
+is elected as the \emph{Gate Peer} and the cycle is then broken 
+%% at a strategic place  %%AD I am not sure what this adds?? strategic?? on what??
+so that it results in the minimization of the distance between the 
+\emph{Gate Peer} and all other nodes that are currently part of the cycle. 
+This process is managed by using two specialized message types 
+namely, \emph{Information Collection (IC)} and \emph{Cut Message (CM)}. 
+\emph{DCMP} bases its operation on messages whose travel is limited by 
+imposed $TTL$ whose maximum value is set at  $7$ for most cases. 
+This inherently limits the protocol as \emph{DCMP} is unlikely that 
+can efficiently detect cycles that span for more than $7$ nodes.
+%%AD rephrased as above
+% One disadvantage of DCMP is that since the distance a forwarded
+% message can travel is limited by the TTL value, which is practically $7$ in most
+% cases, DCMP cannot detect cycles formed by more than 7 peers. 
+Even though cycle elimination does improve network performance, 
+it cannot 
+% actually solve 
+directly contribute in the solution of 
+the topology mismatch problem.
+%%
 %
@@
 %\paragraph{}
-
 %
@@
 %network.
-%
-
+The expected \emph{DCMP} behavior as far as our $3$ criteria is:
+%
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Replication Strategies in Unstructured Peer-to-Peer Networks}
-\cite{CS2002} aim to improve the inefficient blind search algorithm by
-replicating the data in a P2P network. The main intuition behind the
-idea of replication, or using cached copies, is that as the number of copies for
-each item increases in the network, it would be easier for a search algorithm,
-even a random one, to locate these items. To analyze the feasibility of
-such an approach, the authors investigate three different replication
-strategies, namely uniform, proportional, and square root allocation. In the
-uniform model, the copies of items are uniformly replicated in the network,
-while in the proportional model, the items are replicated based on their query
-rate, so that frequently queried items are replicated more. The square root
-allocation is a strategy proposed by Cohen and Shenker, which is a model between
-the uniform and proportional allocation.  To evaluate the outcome of
-the replication models, the overall costs of successful and unsuccessful
-searches in the network are compared. Results argue that the uniform allocation
-model minimizes the maximum search size, therefore reduces the time spent on an
-unsuccessful search. The proportional model, on the other hand, promotes the
-more frequently queried items by replicating them more and therefore decreases the
-search time for popular items, but suffers when needing to locate the rare
-items. The authors also claim that the expected successful search size is the
-same for uniform and proportional models, and any approach between them would
-always behave much better. Finally, they propose the square root allocation
-approach, which is a replication strategy that minimizes the expected search
-size of successful queries in P2P networks.
-
+% \subsubsection{Replication Strategies in Unstructured Peer-to-Peer Networks}
+
+In~\cite{CS2002}, replication is used as a way 
+to improve inefficient blind search.
+As the number of replicas and/or cached copies increases in the network,
+it would relatively easy to locate items even if the search is random.
+% \cite{CS2002} aim to improve the inefficient blind search algorithm by
+% replicating the data in a P2P network. 
+% The main intuition behind the
+% idea of replication, or using cached copies, is that as the number of copies for
+% each item increases in the network, it would be easier for a search algorithm,
+% even a random one, to locate these items. 
+To analyze the feasibility of
+such an approach, \cite{CS2002} investigates $3$ replication
+strategies: uniform, proportional, and square-root allocation. In the
+uniform model, the copies of items are uniformly spread over the network,
+while in the proportional, items are replicated based on their query
+rate, so that frequently queried items can be found more often. 
+The square root is a compromise between uniform 
+and proportional allocation models.
+% allocation is a strategy proposed by Cohen and Shenker, which is a model between
+% the uniform and proportional allocation.  
+%%AD why u need to say the name here????
+The overall costs for both successful and unsuccessful
+searches are compared to ascertain the effectiveness of the $3$ suggested 
+approaches.
+% To evaluate the outcome of the replication models, 
+% the overall costs of successful and unsuccessful
+% searches in the network are compared. 
+Experimental outcomes indicate that the uniform allocation 
+minimizes the maximum search length and so it can 
+reduce the time spent on unsuccessful searches.
+% Results argue that the uniform allocation %%AD results argue???? --
+					    %% ALSO, "search size??? See above
+% model minimizes the maximum search size, therefore reduces the time spent on an
+% unsuccessful search. 
+The proportional strategy effectively decreases the search time 
+for popular items, but suffers when needing to locate the rare entities.
+% The proportional model, on the other hand, promotes the
+% more frequently queried items by replicating them more and therefore decreases the
+% search time for popular items, but suffers when needing to locate the rare
+% items. 
+When it comes to expected successful search size, it is the same for 
+for both uniform and proportional models and any approach in between 
+to always behave much better.
+% The authors also claim that the expected successful search size is the
+% same for uniform and proportional models, and any approach between them would
+% always behave much better. 
+The square root allocation  aims at minimizing the 
+expected search size for successful queries in \p\ networks.
+% Finally, they propose the square root allocation
+% approach, which is a replication strategy that minimizes the expected search
+% size of successful queries in P2P networks.
+%%
+The table below outlines how the $3$ strategies fare:
+%%
 \begin{center}
+{\footnotesize
 \begin{tabular}{rccc}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Narada}
+% \subsubsection{Narada}
 
@@
 
-\emph{Narada} \cite{CRZ2000,CRSZ2001,CRSZ2002} is a generic
-protocol for creating self-adapting overlay networks, that can achieve
-application-layer multicast communication without requiring IP multicast
-infrastructure at the network layer\footnote{
-  IP multicast is the term referring to the method of sending IP datagrams to a
-  group of receivers in a single transmission. Even though the method is
-  available some years now, IP multicast has not taken off as anticipated since
-  many believe that it violates the stateless design of the current Internet.
-  For this reason, it is widely deployed in contained environments like
-  enterprises (i.e. IPTV applications - distance learning, video conferencing),
-  commercial stock exchanges, and multimedia content delivery networks.
-}. Although it was not originally designed as a P2P system, Narada is
-a pioneer in that 
-%(and the accompanying \emph{Scattercast} \cite{C2000})
-it was the first to consider the feasibility of overlay-based, application
-layer, services over the Internet that can take into account bandwidth and
-latency properties of the physical underlying infrastructure. Additionally,
-authors realized the inefficiency caused by the topology mismatch problem and
-tried to address it by building a richer connected graph, called a
-\emph{mesh}, and building per source minimum spanning trees\footnote{
-  ENarada \cite{LYL2008} used Gossip protocol for the
-construction
-}.
+{\sl Narada}~\cite{CRZ2000,CRSZ2001,CRSZ2002} is a generic
+protocol for creating self-adapting overlay networks capable of 
+application-layer multicast communications without requiring IP multicast
+infrastructure at the network layer.
+Although IP--multicast would present a choice in implementing
+{\sl Narada}, it is in general considered that it violates the 
+stateless design of the current Internet.
+Despite the fact the {\sl Narada} was not designed as \p\ system per se,
+it was the first to consider the feasibility of overlay-based and application
+layer services over the Internet while taking into account 
+bandwidth and latency properties of the underlying physical infrastructure.
+%%AD you cannot have such a long footnote for God sake!
+% \footnote{
+  % IP multicast is the term referring to the method of sending IP datagrams to a
+  % group of receivers in a single transmission. Even though the method is
+  % available some years now, IP multicast has not taken off as anticipated since
+  % many believe that it violates the stateless design of the current Internet.
+  % For this reason, it is widely deployed in contained environments like
+  % enterprises (i.e. IPTV applications - distance learning, video conferencing),
+  % commercial stock exchanges, and multimedia content delivery networks.
+% }. 
+% Although it was not originally designed as a P2P system, Narada is
+% a pioneer in that 
+% %(and the accompanying \emph{Scattercast} \cite{C2000})
+% it was the first to consider the feasibility of overlay-based, application
+% layer, services over the Internet that can take into account bandwidth and
+% latency properties of the physical underlying infrastructure. 
+In the context of this work, the authors came across the 
+inefficiency caused by the topology mismatch and attempted to address it 
+by building an enhanced connected graph termed \emph{mesh}  %%AD why this graph
+							    %% is rich/enhanced???
+with each source featuring its own minimum spanning tree.
+A gossip-protocol was deployed for the creation of these 
+spanning trees~\cite{LYL2008}.
+%%
+%%AD in general avoid footnotes...  The sentence below does not make sense!
+%% "richer" to what???  ENarada is what???
+%% 	--- I rephrased as above - not sure correct.. check -----
+%% 
+% Additionally,
+% authors realized the inefficiency caused by the topology mismatch problem and
+% tried to address it by building a richer connected graph, called a
+% \emph{mesh}, and building per source minimum spanning trees\footnote{
+  % ENarada \cite{LYL2008} used Gossip protocol for the
+% construction
+% }.
 %as shown in Figure~\ref{figure:narada}.
-They also keep the graph and the trees
-dynamically updated while nodes continue to join and leave the network.
+Both graph and trees are dynamically updated as nodes keep joining 
+or departing the network.
+% They also keep the graph and the trees
+% dynamically updated while nodes continue to join and leave the network.
 The protocol aims to ease the physical link stress, the overall
-resource usage and the relative delay among end systems. Unfortunately, the most
-important limitation of Narada is that although it works reasonably well for
-small groups, it does not scale well for larger networks, therefore it is not
-suitable for, potentially very large, P2P file sharing application networks.
-
+resource usage as well as the relative delay among end-systems. 
+Unfortunately, the main limitation for {\sl Narada} is that 
+although it works reasonably well for small groups, 
+it does not scale well for larger networks. 
+Hence, it is not a suitable choice for potentially large \p\ networks:
+%%
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Adaptive Overlay Topology Optimization (AOTO)}
-\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{LZXN2003} is one
-of the first attempts, along with Narada, to address the topology mismatch
-problem. AOTO is a distributed algorithm that works in two steps, namely
-\emph{Selective Flooding} and \emph{Active Topology} to optimize the usage of
-the underlying physical resources. In \emph{Selective Flooding}, a minimum
-spanning tree is built among each peer and its immediate neighbors so that
-queries are not flooding the whole network and at the same time without
-shrinking the search scope. During \emph{Active Topology}, the overlay
-links are revised so that they depict the physical network topology as close as
-possible. This is done independently, by each peer by replacing non-flooding
-neighbors, with closer nodes. Picking a replacement out of these non-flooding
-neighbors follows a random policy (called \emph{Randomized AT} algorithm in the
-paper's context). For these steps a peer needs to keep track of communication
-costs to all its neighbors (e.g. the network delay) as well as between any
-pair of neighbors (meaning also that additional, cost probing, message types
-need to be added to the Gnutella protocol). The randomized AT algorithm is
-applied by the source peer every time the neighbour list is updated or an updated
-neighbour cost table is received.
-
+% \subsubsection{Adaptive Overlay Topology Optimization (AOTO)}
+
+Along with {\sl Narada}, the \emph{Adaptive Overlay Topology 
+Optimization (AOTO)}~\cite{LZXN2003} is one
+of the first attempts to address the topology mismatch problem. 
+\emph{AOTO} is a distributed algorithm that seeks to optimize the usage of
+the underlying physical resources and operates in $2$ phases:
+\emph{Selective Flooding} and \emph{Active Topology}.
+In \emph{Selective Flooding}, a minimum
+spanning tree is built 
+% among  %%AD among???  put "for"...
+for each peer and its immediate neighbors so that
+queries do not flood the entire network 
+% and at the same time 
+without simultaneously shrinking their search scope. 
+During \emph{Active Topology}, the overlay
+links are revised so that they can more closely reflect 
+the physical network topology. 
+%%
+%%AD what is a non-flooding neighbor?
+Each peer replaces independently non-flooding neighbors, with closer nodes.
+% This is done independently, by each peer by replacing non-flooding
+% neighbors, with closer nodes. 
+Picking a replacement out of these non-flooding
+neighbors follows a random policy called \emph{Randomized AT} algorithm.
+To accomplish the above actions, a peer has to keep track of 
+its communication costs with all its neighbors (e.g., network delays)
+as well as the costs between any pair of neighbors.
+%%AD I am not sure what the parenthesis below says.. Rephrased as above
+% as well as between any
+% pair of neighbors (meaning also that additional, cost probing, message types
+% need to be added to the Gnutella protocol). 
+The \emph{randomized AT} algorithm is applied by a source peer 
+every time its neighbor list is updated or an updated neighbor cost 
+table is received.
+%%AD do you also obtain neighbor cost tables???
+%%
 %\paragraph*{Selective Flooding (SF)}
@@
 %steps taken.
-
+%%
+\emph{AOTO}'s performance regarding the $3$ criteria is as follows:
+%%
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Adaptive Connection Establishment (ACE)}
-\emph{Adaptive Connection Establishment (ACE)} \cite{LZXN2004} uses the network
-delay as a metric to calculate the costs between nodes in a spanning tree.
-Each peer calculates the costs to its logical one-hop
-neighbors and forms a \emph{neighbor cost table (NCT)} using a special
-routing message type. Two neighboring peers exchange their NCTs in order for
-every peer to obtain the cost between any pair of its local neighbors forming a
-small overlay topology. Moreover, based on obtained NCTs a minimum spanning tree
-among each peer and its one-hop neighbors is built. Ultimately, physically
-far away neighbors are replaced by physically close neighbors. Specifically, a peer
-$S$ probes the distance between one of its non-flooding neighbor's neighbors,
-say $G$ and $H$ respectively. Assuming that the neighbor's neighbor ($SH$) is
-measured as being ``closer'' than the neighbor ($SG$), the link to the later
-is cut and a new connection is established. In the opposite situation where
-distance $SG$ is closer than $SH$ then if $SH$ is shorter than $GH$, then $H$ is
-preserved as $S$'s new neighbor. In the case where $SH$ is larger than $SG$ and
+% \subsubsection{Adaptive Connection Establishment (ACE)}
+The \emph{Adaptive Connection Establishment (ACE)} 
+approach~\cite{LZXN2004} uses the network delay as a metric 
+to estimate the costs between nodes in a spanning tree.
+%%AD when u say spanning tree u mean the ENTIRE network???
+Each peer computes the costs to its logical one-hop
+neighbors and forms a \emph{neighbor cost table (NCT)} 
+using a special routing message type. 
+Any pair of neighboring peers exchange their \emph{NCT}s 
+and so a minimal overlay topology can be formed.
+%%AD the wording of the following is a bit messed up! Simplify! 
+%%	reworked as above... check
+% Two neighboring peers exchange their \emph{NCT}s in order for
+% every peer to obtain the cost between any pair 
+% of its local neighbors forming a
+% small overlay topology. 
+Based on the obtained \emph{NCT}s a minimum spanning tree
+for each peer and its one-hop neighbors is created.
+%%AD rephrased as above.. This expression "... among each peer" makes 
+%%	absolutely NO sense... 
+% Moreover, based on obtained NCTs a minimum spanning tree
+% among each peer and its one-hop neighbors is built. 
+Finally, neighbors located physically far away 
+are replaced by physically-closer counterparts. 
+In particular, a peer $S$ iteratively probes the distance between 
+itself and every of its non-flooding neighbor nodes $G$  as well as 
+the distance between $S$ and $G$'s neighbors designated as $H$.
+%%AD I do not understand what on earth the following sentence SAYS!!!!
+%%	when you say "between" you need to have 2 parties - where is this???
+%%	I rephrased as above.. Check correctness...
+% Specifically, a peer
+% $S$ probes the distance between one of its non-flooding neighbor's neighbors,
+% say $G$ and $H$ respectively. 
+%%
+If the distance between a peer and a neighbor \emph{SG} is measured to 
+be larger than that with a neighbor's neighbor \emph{SH}, the link 
+between $S$ and $G$ is replaced with a new connection between $S$ and $H$.
+%%AD the following is complicated to diff to parse... Tried as above... 
+% Assuming that the neighbor's neighbor ($SH$) is
+% measured as being ``closer'' than the neighbor ($SG$), the link to the later
+% is cut and a new connection is established. 
+%%%
+%%	%%AD the opposite situation does not appear to be consistent!!!
+%	%% with what you say above... 
+% In the opposite situation where
+% distance $SG$ is closer than $SH$ then if $SH$ is shorter than $GH$, then $H$ is
+% preserved as $S$'s new neighbor. 
+%%%
+%%AD is the distance $GH$ in discussion here as well???  I am confused.
+In the case where $SH$ is larger than $SG$ and
 $GH$, no connection will be established and $S$ will continue probing another
-neighbor's neighbor. The above optimization is conducted within $1$-neighbor
-closure (among its source peer and all its direct neighbors) but the scope can
-be extended. The larger the scope, the better topology matching improvement but
-also the greater the computational overhead.
-
+neighbor's neighbor. 
+%%%
+The above optimization is conducted within $1$-neighbor
+closure using as base a peer and checking all its direct neighbors.
+Evidently the scope of such an operation could be extended.
+Should  a larger scope be used, a better topology matching can be obtained 
+at a greater computational overhead.
+%%
+%%AD This is NOT English.. No verb??? I am not sure...
+% The larger the scope, the better topology matching improvement but
+% also the greater the computational overhead.
 %
@@
 %the algorithm still suffers from the discrepancies caused by mis-located nodes.
-
+%%
+%%
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Location-aware Topology Matching (LTM)}
-\emph{Location-aware Topology Matching (LTM)} \cite{LLXNZ2004} proposes a
-method to optimize the overlay structure of the P2P network based on the
-physical topology. To achieve this, peers issue special messages called
-\textit{TTL-detector}s with TTL values of 0 and 1 so that the receiving peers
-discover one- and two-hop neighbor sets ($N$ and $N^2$ respectively) around
-them and calculating communication costs. Calculation is achieved through
-timestamp checking, so clocks must be synchronized. The latency information
-gathered is later used to evolve the overlay network into a more efficient one,
-without reducing the search scope. Each node examines the latency information
-among his direct neighbors and those with larger latency are placed on a
-will-cut list where they remain for a certain period of time after which they are
-cut and placed on the peer's cut list. Thus, low productive
-connections are dropped and replaced by more efficient ones, reducing the
-latency on the overall network. Although LTM improves the overall efficiency of
-the P2P network, it does not actually use real physical topology information,
-just latency metrics, therefore it does not offer a guaranteed safety net for
-the topology mismatch problem.
-
+% \subsubsection{Location-aware Topology Matching (LTM)}
+
+\emph{Location-aware Topology Matching (LTM)}~\cite{LLXNZ2004} 
+seeks to optimize an overlay \p\ structure based on the physical topology.
+To achieve this, peers issue special messages called
+\textit{TTL-detector}s whose \emph{TTL} values are $0$ or $1$;
+in this regard, peers 
+discover $1$- and $2$-hop neighbor sets ($N$ and $N^2$ respectively) %%AD not 
+						%% sure what the N and N2 are..
+and proceed to compute communication costs.
+% around them and calculating communication costs. 
+Time-stamps are used to derive network latency measurements that are then used 
+to improve the overlay network without sacrificing the search scope.
+% Calculation is achieved through
+% timestamp checking, so clocks must be synchronized. 
+% The latency information
+% gathered is later used to evolve the overlay network into a more efficient one,
+% without reducing the search scope. 
+Each node compares the latency information
+received from its direct neighbors;
+peers with longer latencies are placed on a
+\emph{will-cut} list where they remain for a certain period of time 
+after which they are finally eliminated 
+and are placed on the peer's \emph{cut-list}. 
+Thus, low-productivity connections are dropped and replaced by 
+more efficient ones, reducing the
+latency on the overall network. 
+Although \emph{LTM} improves the overall efficiency of
+the \p\ network, it is unable to offer any warranty for 
+effectively addressing the mismatch problem. %%AD shortened and consolidated
+%% 
+% use real physical topology information,
+% just latency metrics, therefore it does not offer a guaranteed safety net for
+% the topology mismatch problem.
 %
@@
 %
-
+\emph{LTM} fares as follows regarding our $3$  criteria:
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Scalable Bipartite Overlay (SBO)}
-\emph{Scalable Bipartite Overlay (SBO)}
-\cite{LXN2004,LXN2007} reduces the overhead of creating
-and maintaining a minimum spanning tree cost by randomly dividing the nodes into
-two groups, namely red and white and assigning different tasks to different
-groups. When a peer joins the network, it is randomly assigned with an initial
-color, say red or white (separating all peers into two groups). Then the peer
-that plays the role of the bootstrap host provides the joining peer with a list
-of active peers along with their color information. The later uses this list to
-establish connections to differently colored peers. This way, all peers form a
-bipartite overlay. The white peers measure distances to neighbors (red only) by
-using the network delay as a metric and report their corresponding red
-neighbors. The red peers, equipped with the information of all two-hop
-neighbors ($N^2$), create minimum spanning tree of these neighbors and
-assigns efficient forwarding paths. Having a minimum spanning tree with two
-hops, a red peer is able to send its queries within that range. Some white
-peers, though, have become non-forwarding neighbors. In this phase, such a
-(white) neighbor will try to find another red peer being two hops away from its
-current red neighbor to replace the later as its new neighbor. The white peers
-can further optimize their positions within the overlay if this is required.
-
+% \subsubsection{Scalable Bipartite Overlay (SBO)}
+
+The \emph{Scalable Bipartite Overlay (SBO)}~\cite{LXN2004,LXN2007} 
+reduces the overhead of creating
+and maintaining a minimum spanning tree 
+by randomly dividing the nodes into
+two groups --\emph{red}s and \emph{white}s-- 
+and assigning different tasks to the two groups.
+When a peer joins the network, it is randomly assigned with an initial
+color (red or white).
+%  (separating all peers into two groups). 
+Then the network bootstrap host node furnishes the joining peer with a list
+of active peers along with their color information. 
+%%AD how many nodes are provided here??? 
+The joining peer uses this list to 
+establish connections to differently colored peers. 
+In this regard, all peers form a a bipartite overlay. 
+Using the network delay as a metric, white-peers 
+peers measure distances from red counterparts
+and report the encountered red neighbors.
+Having information on all their $2$-hop neighbors 
+($N^2$) %%AD I am not sure what this N^2 means here!
+red-peers create minimum spanning trees for the neighbors in question and
+% assigns %%AD this should be plural
+assign efficient forwarding paths. 
+Having a minimal $2$-hop spanning tree, 
+a red-peer dispatches its queries only within this range. 
+Some white-peers, though, may have become non-forwarding neighbors. 
+%%AD I am not sure how you reach the above sentence... 
+%%
+%%AD The following sentence is very convoluted. I cannot parse it!
+In this phase, such a (white) neighbor will try to find another 
+red peer being two hops away from its
+current red neighbor to replace the later as its new neighbor. 
+%%%%
+The white-peers can further optimize their positions within 
+the overlay if this is required. 
+%%AD What are the implication of the above statement?? 
+%%   What is the end result of the approach in terms of the mismatch problem??? 
+%%   Something is missing... 
 %
@@
 % on traffic cost and about 60 percent reduction on query response time.
-
+In regards to the $3$ stated criteria, \emph{SBO} behaves as follows:
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Two-Hop-Away Neighbor Comparison and Selection (THANCS)}
-Work in \cite{LNXE2005,L2008} proposes a distributed heuristic
-called \emph{Two-Hop-Away Neighbor Comparison and Selection (THANCS)} that
-tries to minimize the overlay hop costs. THANCS is considered a \emph{local
-search method} because it aims to find a locally optimum
-solution by exploiting knowledge within a 2-hop radius. The algorithm consists
+% \subsubsection{Two-Hop-Away Neighbor Comparison and Selection (THANCS)}
+
+The distributed heuristic termed 
+\emph{Two-Hop-Away Neighbor Comparison and Selection (THANCS)}~\cite{LNXE2005,L2008} attempts to minimize overlay hop costs.
+\emph{THANCS} is essentially a \emph{local search method} as it aims 
+to find a locally optimum solution by exploiting knowledge within 
+a $2$-hop radius. 
+The algorithm consists
 of two main components: \emph{piggybacking neighbor distance on queries} and
-\emph{neighbor comparison and selection} which are further discussed below.
-The first component requires peers to probe distances (using network delay
-distance measuring) with its immediate neighbors and stores information
-locally. This is done by introducing a special query message type, the
-\emph{Piggy Message (PM)} which includes information about the neighbor
-identification and measured distance. A peer $p$ constructs a PM for its
+\emph{neighbor comparison and selection}.
+
+The \emph{piggybacking component} requires peers 
+to probe immediate neighbors using delay distance measurements 
+and store information locally. 
+This is done by introducing a special query message type, the
+\emph{Piggy Message (PM)} which includes information about 
+the neighbor identification and measured distance. 
+A peer $p$ constructs a \emph{PM} for its
 neighbor $n$, which contains $n$'s IP address and the distance between them.
-When $p$ receives a query from $n$, this PM will be piggybacked to the query
-message that will be then forwarded to all other neighbors of peer $p$. 
-Each neighbor will detach the PM information (it will not be further forwarded),
-record the $pq$ distance and process the query as always after receiving a query
-message. Selection of which incoming queries should piggyback a PM is proposed
+%%
+If $p$ receives a query from $n$, the respective \emph{PM}-message will 
+be piggybacked to the query and $p$ will forward this enlarged message to
+of its neighbors.
+% When $p$ receives a query from $n$, this PM will be piggybacked to the query
+% message that will be then forwarded to all other neighbors of peer $p$. 
+In turn, each neighbor $q$ will detach the \emph{PM}-information,
+% (it will not be further forwarded) 
+record the distance $pq$ and further process the query. %%AD added "further" - correct??
+%%%
+%%AD rephrased as above... When you use "this" "that" etc BE careful...
+% Each neighbor will detach the PM information 
+% (it will not be further forwarded),
+% record the $pq$ distance and process the query 
+% as always after receiving a query message.  %% THIS "as... " threw me off!!!!
+%%%%%%%%%%%%%%%%%%%%
+Selection of which incoming queries should piggyback a \emph{PM} is proposed
 by either \emph{pure probability-based (PPB)} or \emph{new neighbor triggered
-(NNT)} policies. The second component, namely neighbor comparison and selection,
-dictates peer $p$ to probe the distance to all his known, not yet probed,
-two-hop neighbors ($ N^2(p)$). The distance of $pn$ is known to $p$. Upon
-receiving a PM from node $n$ with the distance of $nq$, where $q$ is a direct
-neighbor of $n$, $p$ chooses one of the following two paths. First, if $q$ is a
+(NNT)} policies.
+
+The \emph{neighbor comparison and selection} component
+helps peer $p$ probe all his neighbors placed upto 
+$2$-hops away ($ N^2(p)$) %%AD what is this?? complexity???
+not probed so far.
+The distance of $pn$ is known to $p$. %%AD I am not sure what you say!
+Upon receiving a \emph{PM} from node $n$ with the distance of $nq$, 
+where $q$ is a direct
+neighbor of $n$, $p$ chooses one of the following two paths. 
+\begin{itemize} %%AD put the two directives here...
+\item[--]
+	if $q$ is a $1$-hope neighbor of $p$, then 
+	$p$ 
+\item[--]
+\end{itemize}
+%%AD I cannot follow what on earth it is said here... this needs 
+%%	rewriting and better structuring..  Use the above list for 
+%%	the expected operation...
+First, if $q$ is a
 direct neighbor of $p$, then the latter will check the cost of the involved
@@
 into a \emph{will-cut} list. If it is $nq$ then nothing is done (since either
-$n$ or $q$ will have the chance to handle it themselves).  On the other hand, if
+$n$ or $q$ will have the chance to handle it themselves).  
+%%AD this is what I think should go into the second clause above... 
+On the other hand, if
 $q$ is within two-hop radius from $p$, the later probes the former (if it hasn't
@@
 If it is $pn$, $p$ will establish connection $pq$ and put $pn$ in the will-cut
-list. If it is $nq$, $p$ will stay connected with both $n$ and $q$,
+list. 
+If it is $nq$, $p$ will stay connected with both $n$ and $q$,
 such that eventually in later stages either $n$ or $q$ will disconnect link $nq$.  
-
+%%AD The above paragraphs need clearing up.. The text cannot be read.. 
+%%
+%%AD Also, make the connection with the mismatch problem... and justify 
+%%	somewhat convincingly why the rating below comes true...
 %
@@
 %systems.
-
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Hops Adaptive Neighbor Discovery (HAND)}
-\cite{CLZHC2006} proposes the \emph{HAND} algorithm which uses a
-fully distributed triple hop adjustment strategy to address the topology
-mismatch problem. The approach's ultimate goal is to shape the current overlay
-graph $G$ into the \emph{Logical Communication Network (LCN)} $G^{*}$ which is
-how the optimal overlay is referred in the paper's context. A formal
-definition of achieving the above match is only if all peer hop sequences $(v_1,
-v_2, \ldots, v_k)$ in $G$ exist in $G^{*}$ and in the same order\footnote{In
-practice triple sequences $(v_1, v_2, v_3)$ are used.}. The mismatching
-detection is done in the following way. Suppose we want to verify peer a
-sequence, say $v_2-v_1-v_3$. A pair of probing messages are sent from $v_1$ to
-$v_2$ and $v_3$. Suppose delays of $(v_1,v_2)$ and $(v_1,v_3)$ are denoted as
-$x$ and $z$, respectively. When the probing message arrives to $v_2$ it forwards
-it directly to $v_3$. Similarly, when the probing message arrives to $v_3$, it
-forwards it directly to $v_2$. These last steps are performed in order to obtain
-delays of $(v_2,v_3)$ and $(v_3,v_2)$ physical paths, respectively, denoted by
-$y$. If $y=z-x\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched and should
-be adjusted to $v_1-v_2-v_3$ by deleting edge $(v_1,v_3)$ and adding a new
-$(v_2,v_3)$. If $y=x-z\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched and
+The \emph{Hops Adaptive Neighbor Discovery (HAND)} algorithm~\cite{CLZHC2006}
+uses a fully-distributed triple--hop adjustment strategy
+to address the topology mismatch problem.
+\emph{HAND}'s goal is to shape the current overlay
+graph $G$ into an optimal \emph{Logical Communication Network} 
+%(LCN)} %%AD why u need this acronym? u never use it!
+$G^{*}$.
+% \subsubsection{Hops Adaptive Neighbor Discovery (HAND)}
+% \cite{CLZHC2006} proposes the \emph{HAND} algorithm which uses a
+% fully distributed triple hop adjustment strategy to address the topology
+% mismatch problem. The approach's ultimate goal is to shape the current overlay
+% graph $G$ into the \emph{Logical Communication Network (LCN)} $G^{*}$ which is
+% how the optimal overlay is referred in the paper's context. 
+This optimality is attained if all peer hop 
+sequences $(v_1, v_2, \ldots, v_k)$ of $G$ exist in $G^{*}$ 
+and are in the same order.
+The latter indicates that in 
+practice triple sequences $(v_1, v_2, v_3)$ are used.
+%%AD this used to be a footnote - avoid footnotes in general!
+%%
+A mismatch is detected as follows:
+suppose we want to verify peer a
+sequence, say $v_2-v_1-v_3$. 
+Two probing messages are dispatched from $v_1$ to
+$v_2$ and $v_3$ and yield delays of $x$ and $z$ respectively.
+%
+When the probing message arrives at $v_2$, it gets forwarded
+directly to $v_3$. 
+Similarly, the message reaching $v_3$, it is directly forwarded to $v_2$.
+The above two forwarding actions seek to quantify 
+the corresponding delays of $(v_2,v_3)$ and $(v_3,v_2)$ physical paths
+denoted by $y$. 
+%%
+If $y=z-x\pm\varepsilon$, the sequence $v_2-v_1-v_3$ is mismatched and should
+be adjusted to $v_1-v_2-v_3$ by deleting edge $(v_1,v_3)$ and 
+then adding a new $(v_2,v_3)$. 
+If $y=x-z\pm\varepsilon$, sequence $v_2-v_1-v_3$ is mismatched and
 should be adjusted to $v_1-v_3-v_2$ by deleting edge $(v_1,v_2)$ and adding a
-new $(v_3,v_2)$, where $\varepsilon$ is a small positive real number denoting
-additional delays caused by possible forwarding and jitter delays. The
-advantages of the HAND algorithm compared to other approaches are that:
+new $(v_3,v_2)$. 
+The  $\varepsilon$ is a small positive real number denoting
+additional delays caused by possible forwarding and jitter delays. 
+%%
+The advantages of \emph{HAND} are that it
 \begin{inparaenum}[\itshape i\upshape)]
-  \item it does not need any clock synchronization,
-  \item it is a fully distributed algorithm making it suitable for
-        decentralized systems,
-  \item the traffic overhead is low,
-  \item it can be used in dynamic peer-to-peer environments, and
-  \item maintains lower query response times.
+  \item does not need any clock synchronization,
+  \item is a fully distributed algorithm.
+  \item involves low traffic overhead,
+  \item can be used in dynamic \p\ environments, and 
+  \item furnishes low query response times.
 \end{inparaenum}
-
 %
@@
 %LTM by an average of 55 percent.
-
+In terms of the $3$ stated criteria, \emph{HAND} fares as follows:
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 % TODO: DOUBLE CHECK IF THIS ALGORITHM GOES IN THIS SECTION/SUBSECTION
-\subsubsection{Adaptive Peer Selection (APS)}
-Bernstein et al., use machine learning techniques for building peer
-selection strategies from past experience \cite{BFLZ2003}. A decision tree is
-used to rate peers based on information about connection characteristics that is
-collected. These can be load, bandwidth, and past uploading experience. Then a
-Markov decision process is incorporated as a mechanism to shape the policy
-for switching among the peers. Although novel, the downside of this approach is that the peer
-selection algorithm is slow due to the learning process and the complexity of
-the method.
-
+% \subsubsection{Adaptive Peer Selection (APS)}
+The \emph{Adaptive Peer Selection (APS)} approach uses 
+machine learning techniques to form peer
+selection strategies based on past experience~\cite{BFLZ2003}. 
+A decision tree is used to rate peers based on information 
+collected for the characteristics of established connections.
+Such features include load, bandwidth, and past uploading experience. 
+%%AD load on what?? The dispatching end? the receiving end??
+Subsequently, a Markov decision process is introduced 
+as a mechanism to shape the policy
+for switching among the peers. 
+The success of approach depends on how fast the introduced 
+peer selection functions. Admittedly, \emph{APS} is slow 
+due to the learning process and its inherent complexity.
+As far as the $3$ criteria is concerned, \emph{APS} has as follows:
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Innocuous Topology Aware (ITA) Overlay Construction}
-\cite{PRFM2009} introduces an algorithm for Innocuous Topology Aware
-construction of unstructured P2P networks. The paper's proposition is twofold.
-Overlay construction and search method. For the overlay construction, it employs
-the notion of \emph{short} and \emph{long} connections. Assuming $N$ is the
-number of all network participants, $\alpha \leq 1 $ a system-wide magic number
-and $x$ an $\alpha$-related latency threshold, the bootstrapping peer randomly
-selects $\alpha \ast N$ close (latency below $x$) and
-$\left( 1 - \alpha \right) \ast N$ distant (latency above $x$) nodes as its
-neighbors. The search method has two steps. First, the querying node initiates a
-flood to its distant neighborhood with $TTL = 1$. Then, peers that receive a
-query over a long link start a local flood with $TTL = ttl$, where $ttl$ is a
-system defined parameter.
-
-The goal of the overlay construction is to create a network that exhibits low
-clustering, a beneficial characteristic of random graphs which can lead to a
-larger coverage of peers with the same number of messages and reduced
-duplication. This means more efficient information lookup that has minimal or no
-negative impact whatsoever to the rest of the mechanisms employed in
-unstructured P2P systems\footnote{For example the 1-hop replication and the
-dynamic querying mechanisms of later versions of Gnutella.}. The paper reports
-50\% reduction in communication latency among peers by cutting off some 20\% of
-the IP-level message traffic generation.
-
+The \emph{Innocuous Topology Aware Overlay Construction (ITA)} approach~\cite{PRFM2009} seeks to offer both an overlay formation approach and furnish 
+an effective way to carry out searching.
+% \subsubsection{Innocuous Topology Aware (ITA) Overlay Construction}
+% \cite{PRFM2009} introduces an algorithm for Innocuous Topology Aware
+% construction of unstructured P2P networks. The paper's proposition is twofold.
+% Overlay construction and search method. 
+When constructing the overlay, \emph{ITA} exploits the notion of 
+of \emph{short} and \emph{long} connections.
+% For the overlay construction, it employs
+% the notion of \emph{short} and \emph{long} connections. 
+Should $N$ be the number of network peers, 
+$\alpha \leq 1 $ a system-wide magic number
+and $x$ an $\alpha$-related latency threshold,
+the bootstrapping peer randomly selects $\alpha \ast N$ 
+``close'' (latency below $x$) and 
+$\left( 1 - \alpha \right) \ast N$ distant (latency above $x$) nodes
+as its neighbors. 
+%%AD how do they select all these magic constants???
+%%
+Searching occurs in two phases: in the initial phase, the querying node
+floods its distant neighborhood  with $TTL = 1$.
+Subsequently, peers receiving a query over a ``long link'' commence a 
+local flood with $TTL\;=\;ttl$ where \emph{ttl} is system-defined.
+%%AD how are you supposed to define these params?? in a fixed manner??
+
+The main objective of the overlay construction phase
+is to yield a network that exhibits low clustering. 
+In turn, this is 
+a beneficial characteristic for random graphs as it can lead to a
+larger coverage of peers with the same number of messages and 
+at the same time helps reduce duplication. 
+%%AD same number of messages in comparison to what?? 
+%%	same: duplication of what?? messages right???
+This combination offers efficient lookups that pose minimal 
+or no negative impact to other mechanisms possibly employed by 
+the \p\ systems.
+% This means more efficient information lookup that has minimal or no
+% negative impact whatsoever to the rest of the mechanisms employed in
+% unstructured P2P systems\footnote{For example the 1-hop replication and the
+% dynamic querying mechanisms of later versions of Gnutella.}. 
+A 50\% reduction in communication latency among peers by cutting off 
+some 20\% of the \emph{IP}-level traffic generated 
+in reported in~\cite{PRFM2009}.
+%%
+\emph{ITA} places as follows in the context of the $3$ criteria:
+%
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{EGOIST}
-EGOIST \cite{SLLBBR2008} is a set of algorithms to construct and manage overlay
-networks. It utilizes a selfish approach in the sense that every participating
+% \subsubsection{EGOIST}
+\emph{EGOIST}~\cite{SLLBBR2008} is a set of algorithms 
+to construct and manage overlay networks. 
+\emph{EGOIST} utilizes a selfish approach in the sense that every participating
 peer continuously updates its neighbors so as to minimize the sum of distances
-to all destinations under shortest-path routing. In EGOIST, a newly arriving
-peer, randomly connects to an already participating node through a bootstrap
-server. When connected, it starts receiving information throughout the
-link-state mechanism and thus, after some time, it has a complete picture of the
-overlay graph. Then it estimates the delay to all other nodes to
-determine its potential neighbors and ultimately connect to the overlay using
-some policy\footnote{For example, minimization of the average delay to all its
-neighbors.}. EGOIST requires extensive resource usage to continuously update the
-wiring of all the nodes in the system. The authors claim that they can
-reduce the load imposed by the monitoring and announcing process from $O(n^2)$
-to only $O(kn)$, where $n$ is the number of all nodes in the network and $k$ is
-the number of links a node chooses to establish.
-
+to all destinations under shortest-path routing. 
+A newly arriving peer, randomly connects to an already participating 
+node through a bootstrap server. 
+Once connected, the peer starts receiving information throughout the
+link-state mechanism and thus, after some time, 
+it has a complete picture of the overlay graph. 
+Then the peer estimates the delay to all other nodes to
+determine its potential neighbors and ultimately connects 
+to the overlay using some policy; such a policy might be 
+for example the minimization of the average delay to all its
+neighbors. 
+\emph{EGOIST} requires extensive resource usage to continuously update the
+wiring of all the nodes in the system. 
+%%AD I am not sure I get the "wiring of all"??? clarify.. 
+The approach can reduce the load imposed by the monitoring 
+and announcing process from $O(n^2)$ to only $O(kn)$, 
+where $n$ is the number of all network nodes 
+and $k$ is the number of links a node chooses to establish.
+%%AD is this an average number?? What kind of number???
+Regarding the $3$ criteria, \emph{EGOIST}'s behavior is as follows:
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Biased Neighbor Selection (BNS)}
-Bindal et al. \cite{BCCMSBZ2006} claim to strengthen BitTorrent
-protocol\cite{c_bittorrent_2003} locality by selecting most of a peer's
-neighbors to be out of the same ISP, while preserving the near optimal download
-performance of the protocol itself. BitTorrent systems employ most of the time
-mechanisms that can be proven very aggressive to an ISP's networking and
-accounting, basically, due to their lack of knowledge of Autonomous System
-boarder limits. BitTorrent specification, by default allows for each peer, 35
-connections. Biased neighbor selection experiments on some number $k$ which
-denotes a number of neighboring peers which are not on the same ISP than the
-peer at hand. Consequently, the remaining $35 - k$ peers are from the same ISP.
-These $k$ nodes are used to provide a more extended, global view of a
-network, but balance the load within the limits of a single ISP.
-This prevents clients from exchanging traffic through an expensive transit link,
-if there are alternative local connections that could offer the same service faster and
-at no additional cost for the ISP. Implementation can either be done through
-modifications on the tracker side to identify ISP locality, or through P2P
-traffic shaping devices installed on the ISP's edge routers. The paper also proposes
-a combination of biased neighbor selection along with bandwidth throttling and
-some caching techniques for near optimal results. Unfortunately, implementation
-and deployment
-need knowledge of ISP Autonomous System mappings
-or infrastructure changes (installation of P2P traffic shaping devices) on the
-ISP, making widespread adoption.
-
+
+The main objective of the \emph{Biased Neighbor Selection (BNS)} 
+approach~\cite{BCCMSBZ2006} is to 
+strengthen {\sl BitTorrent}'s~\cite{c_bittorrent_2003} locality 
+by carefully choosing most of the neighbors of a peer to come from 
+the same \emph{ISP} while at the same time adhering to the near-optimal
+download performance of the protocol.
+%%
+By and large, {\sl BitTorrent}s deploy mechanisms that have proved 
+aggressive for \emph{ISP}s' networking and accounting as they function 
+in a ``without-borders'' approach.
+% BitTorrent systems employ most of the time
+% mechanisms that can be proven very aggressive to an ISP's networking and
+% accounting, basically, due to their lack of knowledge of Autonomous System
+% boarder limits. 
+The default {\sl BitTorrent} specification designates $35$ connections 
+for each peer regardless if such connections are placed within or outside 
+the borders of the \emph{ISP} that hosts the node.
+%%
+\emph{BNS} proposes that only $k$ out of these $35$ connections are 
+established with nodes beyond the {\it ISP}-borders;
+the $k$ connections offer an extended and ``global'' view of 
+the network at large and seek to strike a balance between 
+the load imposed inside and outside the \emph{ISP}.
+The latter refrains peers from exchanging traffic through
+expensive network links should there exist alternative local connections
+offering the sought services.
+%%AD rewrote as above...
+% Biased neighbor selection experiments on some number $k$ which
+% denotes a number of neighboring peers which are not on the same ISP than the
+% peer at hand. 
+% Consequently, the remaining $35 - k$ peers are from the same ISP.
+% These $k$ nodes are used to provide a more extended, global view of a
+% network, but balance the load within the limits of a single ISP.
+% This prevents clients from exchanging traffic through an expensive transit link,
+% if there are alternative local connections that could offer the same service faster and
+% at no additional cost for the ISP. 
+\emph{BNS} is realized either through 
+the modification on the tracker so that \emph{ISP}-local resources are rapidly 
+identified,
+or through \p\ traffic shaping devices installed on \emph{ISP} edge routers. 
+%%
+In~\cite{BCCMSBZ2006}, a combination of 
+biased neighbor selection along with bandwidth throttling
+and caching techniques for near-optimal lookup results is proposed.
+% The paper also proposes
+% a combination of biased neighbor selection along with bandwidth throttling and
+% some caching techniques for near optimal results. 
+%%%%%%%%%%%%%%%%%%%%%
+% Unfortunately, implementation
+% and deployment
+% need knowledge of ISP Autonomous System mappings %%AD What are these Aut. Sys?
+% or infrastructure changes (installation of P2P traffic shaping devices) on the
+% ISP, making widespread adoption.
+%%AD I am not sure what the above sentence says! I rephrased as follows:
+Unfortunately, \emph{BNS} deployment requires knowledge of 
+\emph{ISP} network mappings and awareness of possible changes 
+that occur in the infrastructure limiting its widespread adoption;
+the following table qualitatively depicts the expected behavior of \emph{BNS}:
+%%
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Ono}
-Choffnes and Bustamante propose Ono \cite{CB2008}, a protocol for managing
-BitTorrent traffic so that it significantly reduces cross-ISP traffic and at the
-same time enhances downloading rates by favoring connections within the borders
-of a single autonomous system. Contrary to biased peer selection proposed in
-\cite{BCCMSBZ2006}, this work can lead to improved performance without requiring
-any cooperation between ISP subscribers, or requiring any additional
-infrastructure and network topology information. The selection approach is
-landmark-based and leverages existing CDN infrastructure for peer distance
-estimation. CDNs already use both static (i.e. geographical) and dynamic
+\emph{Ono}~\cite{CB2008} is a protocol that helps contain
+{\sl BitTorrent} traffic within individual \emph{ISP}s
+and enhances the downloading rates by favoring connections
+within the borders of a single autonomous system.
+% \subsubsection{Ono}
+% Choffnes and Bustamante propose Ono \cite{CB2008}, a protocol for managing
+% BitTorrent traffic so that it significantly reduces cross-ISP traffic and at the
+% same time enhances downloading rates by favoring connections within the borders
+% of a single autonomous system. 
+Contrary to \emph{BSN}~\cite{BCCMSBZ2006},
+\emph{Oro} leads to improved performance 
+without requiring any explicit cooperation between 
+\emph{ISP} subscribers or any additional
+infrastructure and network topology information.
+% Contrary to biased peer selection proposed in
+% \cite{BCCMSBZ2006}, this work can lead to 
+% improved performance without requiring
+% any cooperation between ISP subscribers, or requiring any additional
+% infrastructure and network topology information. 
+\emph{Oro}'s selection approach is 
+landmark-based and leverages existing {\sl CDN}-infrastructure 
+for peer distance estimation. 
+%%AD I do not understand the following sentence... I am not even sure what 
+%%   who is close to each other??? Confusing sentence. Rephrase...
+{\sl CDN}s already use both static (i.e., geographical) and dynamic
 (network measurement systems) information for their replica selection, so the
@@
 likely close to the replica servers and as a consequence close to each other.
-The redirection behavior is modeled in terms of \emph{ratio map} in the
-article's parlance, a vector of <replica-server,ratio> tuples, where ratio is
-the percentage of times the CDN redirects the peer to the specific replica-server.
-The bootstrapping phase consists of the peer performing DNS lookup to CDN names
-to build its redirection information. The interval for polling DNSs
-starts from 30 seconds and increases by one minute every time redirection
-information to the CDN is found to remain unchanged after a lookup. On the other
-hand, the interval is halved whenever is found to have been changed.  
-To avoid the bootstrapping phase when the ratio map is sufficiently fresh, the
-protocol persists the information after the end of a BitTorrent session.
-
+%%
+The redirection behavior is modeled in terms of \emph{ratio map}, 
+a vector of \emph{$<$replica-server,ratio$>$} tuples, where \emph{ratio} is
+the percentage of times the {\sl CDN} redirects the peer 
+to the specific \emph{replica-server}.
+%%
+The bootstrapping phase %%AD of what?? the protocol's bootstrapping???
+consists of the peer performing {\sl DNS} lookup to {\sl CDN}-names
+%%AD when you say "CDN names" - what are they????
+to build its redirection information.
+%%
+Initially, the {\sl DNS} polling interval is set to $30$ seconds 
+and increases by $1$ minute every time redirection information 
+to the {\sl CDN} remains unchanged after the lookup.
+The {\sl DNS} polling interval is halved if pertinent 
+redirection information has changed.
+To avoid the bootstrapping phase when the \emph{ratio map} 
+is sufficiently fresh, 
+\emph{Oro} persists the information 
+after the end of a BitTorrent session.
+%%AD what information is this?? DNS look ups?? 
+%%
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Locality-Awareness in BitTorrent-like P2P Applications}
-In \cite{LCLX2009} the authors study and compare three different
-approaches in injecting locality awareness in BitTorrent-like applications. The
-first acts at a \emph{macroscopic-level} and targets neighbor selection. When a
-peer asks the tracker to join, the later sorts the swarm peers according to
-their distance to the requesting peer in Autonomous System hop count and send it
-the first
-(e.g. 50) peers in the list. The second manipulates the chocking/unchocking
-BitTorrent mechanisms at an \emph{intermediate-level}. A peer unchockes its 4
-closest in terms of Autonomous System hop count interested neighbors. The same
-applies also
-when the peer turns to the seeding state\footnote{Thus this approach favors
-least distance, contrary to the original BitTorrent implementation which favors
-uploading speed.}. On a \emph{microscopic-level}, the rare-first chunk picking
-policy is substituted by the locality-first policy. The distance value of a
-piece is calculated as the mean value of the distances of the peers that posses
-it. The paper conducts several measurements on the efficiency of this scheme 
-and reaches 
-the conclusion that the major design decision for all locality
-aware P2P systems is the trade-off of optimizing inter- Autonomous System
-traffic versus achieving fairness amongst peers.  While their
-locality-based approach achieves traffic optimization, it does not do well
-when compared with the 
-random approach of the standard BitTorrent protocol.
-
+Three approaches that ``inject'' locality awareness 
+in {\sl BitTorrent}-like applications are discussed in~\cite{LCLX2009}.
+% \subsubsection{Locality-Awareness in BitTorrent-like P2P Applications}
+% In \cite{LCLX2009} the authors study and compare three different
+% approaches in injecting locality awareness in BitTorrent-like applications. 
+The first \emph{macroscopic-level} approach 
+focuses on improving the neighbor selection process. 
+When a peer asks a tracker to join, 
+the later sorts its swarm peers according to
+their hop-count distance from the requesting peer 
+send out the top-$k$ list of nodes (e.g., $k$=$50$).
+%% Autonomous System  hope distance... 
+%%AD Got this out - do u need it?? What does it mean "Autonomic System Hop Dist"???
+%%
+The second approach manipulates the {\sl BitTorrent} chocking/unchocking.
+% at an \emph{intermediate-level}. 
+A peer unchockes its $4$ closest in terms of 
+% Autonomous System 
+hop-count interested neighbors. 
+The same applies also when the peer turns to 
+the seeding state;  this  \emph{intermediate-level} approach favors
+least distance and is in contrast to the original {\sl BitTorrent} 
+implementation which favors uploading speed.
+%%
+Finally, the third approach that operates at 
+a \emph{microscopic-level}, substitutes the rare-first chunk picking
+policy by a locality-first policy. 
+%%AD is it "rarest"?? What does it mean "locality-first"???
+The distance value of a piece is computed as the mean value 
+of the distances of the peers that posses it. 
+%%AD What is this "piece"??? I am not sure how the average gets computed? 
+%%   It is very cryptic description...
+Through experimentation, %%AD simulation? real?
+the locality-based approach achieves traffic optimization.
+Nevertheless, it does not do that well when compared with the 
+random approach of the standard {\sl BitTorrent} protocol.
+%%AD rephrased and shortened the following as above...
+% The paper conducts several measurements on the efficiency of this scheme 
+% and reaches 
+% the conclusion that the major design decision for all locality
+% aware P2P systems is the trade-off of optimizing inter- Autonomous System
+% traffic versus achieving fairness amongst peers.  While their
+% locality-based approach achieves traffic optimization, it does not do well
+% when compared with the 
+% random approach of the standard BitTorrent protocol.
+%%%%%%
+%%AD so in general how does this approach help the mismatch problem?? 
+%%	need a statement here (1-2 sentences explaining this).
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{TopBT}
-\cite{RTLCGZ2010} proposed an approach for enhancing proximity awareness
-in the BitTorrent protocol without any need of any infrastructure installed. It
-suggests that a good peer selection metric should take into account both the
-downloading speed and the network topology. Thus, it proposes the use of the
-ratio of download rate to upload rate, $\frac{d}{u}$, divided by link-level or
-Autonomous System -level routing hops, $l$ or $a$ respectively, in an attempt to
-form a comprehensive metric to select peers with high download rate, low
-reciprocal upload demand and low routing hops. The TopBT metric is applied to
-several parts of the original BitTorrent protocol in order to inject topology
-awareness for better peer selection. Specifically, the parts of the protocol
-that are modified include:  the peer list the
-tracker returns when someone first tries to connect to the swarm, the initial
-connection establishment, and the
-unchoking mechanism.
-
+\emph{TopBT}~\cite{RTLCGZ2010} seeks to enhance 
+proximity awareness in the {\sl BitTorrent}
+without the need of any additional infrastructure.
+% \subsubsection{TopBT}
+% \cite{RTLCGZ2010} proposed an approach for enhancing proximity awareness
+% in the BitTorrent protocol without any need of any infrastructure installed. 
+The key conjecture upon which \emph{TopBT} is built is that 
+a good peer selection metric should take into account both the
+downloading speed and the network topology. 
+In this context, the approach suggests a metric that designates
+peers demonstrating high download rate, low reciprocal upload demand 
+and low routing hops.
+The metric is defined as the ratio of download to upload rate ${d}/{u}$
+divided by the either the link-level $l$ or the number of routing hops required $a$.
+%%AD The following is not a readable sentence... rephrase as above...
+% Thus, it proposes the use of the
+% ratio of download rate to upload rate, $\frac{d}{u}$, divided by link-level or
+% Autonomous System -level routing hops, $l$ or $a$ respectively, in an attempt to
+% form a comprehensive metric to select peers with high download rate, low
+% reciprocal upload demand and low routing hops.  %%AD What is "Autonomous System -level routing hops?????"
+%%%%%%%%%%%%%%%%%%%%%%%%
+The above \emph{TopBT}-metric is applied in various
+aspects of the original {\sl BitTorrent} protocol to ``inject'' topology
+awareness for improved peer selection. 
+%%
+In a effort to essentially improve on the inherent topology mismatch problem,
+\emph{TopBT} modifies the peer list the
+tracker returns when someone first tries to connect to the swarm,
+the initial connection establishment, and the
+unchoking mechanism in the {\sl BitTorrent}.
+%%AD how do they specifically are modified??? What is the difference 
+%%	with the previous paper?
+%%
+%
+%
 %A peer that runs the TopBT protocol evaluates its neighbors by periodically
@@
 %\emph{AS-hops}, using BGP dumps.
-
-Through experimentation on hundreds of PlanetLab as well as residential hosts,
-the authors claim more than 25\% traffic reduction and 15\% faster downloads,
-both with a lightweight cost when compared to popular BitTorrent
-implementations.
-
+%%
+Through experimentation on hundreds of {\it PlanetLab} and residential hosts,
+\emph{TopBT} offers more than 25\% traffic reduction and 15\% faster downloads
+if compared with popular {\sl BitTorrent} implementations.
+%%AD rephrased what is below as above ...
+% the authors claim more than 25\% traffic reduction and 15\% faster downloads,
+% both with a lightweight cost when compared to popular BitTorrent
+% implementations.
+%
+In reference to the stated $3$ criteria, \emph{TopBT} fares as follows:
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Underlying Topology-Aware Peer Selection (UTAPS)}
-UTAPS \cite{LCY2008} is another peer selection strategy for the BitTorrent
-protocol. Its first step is to collect the needed information to
-construct a picture of the underlying topology and the second is to make the
-peer selection, per se, based on this knowledge. For the first part, the
-algorithm uses network tomography, a technique which suggests probing only
-from a large network's endpoints to infer its internal characteristics
-\cite{chny_tomography_2002}. Upon a new peer arrival, the tracker trace-routes
-it to get some knowledge (IP address, routers involved, RTT and hops conducted
-e.t.c). The more the peers in the swarm the better picture a tracker has for the
-underlying infrastructure.  The tracker provides this information to the newcomer
-in the form
-of the bootstrapping peer list. The peers further enhance this picture by
+The \emph{Underlying Topology-Aware Peer Selection (UTAPS)}~\cite{LCY2008}
+is an effort to furnish an enhanced peer selection strategy.
+% \subsubsection{Underlying Topology-Aware Peer Selection (UTAPS)}
+% UTAPS \cite{LCY2008} is another peer selection strategy for the BitTorrent
+% protocol. 
+\emph{UTAPS} essentially works in two stages: in the first,
+it collects information to develop a picture of the underlying topology 
+and in the second, it proceeds to make the peer selection based on 
+this knowledge.
+In the first stage, \emph{UTAPS} employs
+network tomography, a technique that probes from a large network's 
+endpoints to infer its internal characteristics~\cite{chny_tomography_2002}. 
+%%
+Upon arrival of a new peer, the tracker trace-routes 
+it to obtain some basic knowledge 
+including IP address, routers involved, \emph{RTT} and number of 
+hops traversed.
+The more the peers in the swarm the better 
+picture a tracker has for the underlying infrastructure.  
+The tracker provides this information to the newcomer in the form
+of the bootstrapping peer list. 
+%%
+%%AD I am lost among the peers in the following sentence... Who traces whom???
+The peers further enhance this picture by
 trace-routing the returned peers themselves and reporting back to the tracker.
-For the selection part, a family of heuristics are proposed, which in general
-target those peers that expose low RTT and are within certain hop-count away of
-measured routers. The author's claim that peers running UTAPS instead of random
-peer selection can achieve better downloading rates with reduced burden on the
-underlying ISP infrastructure though their technique is coarse-grained and their
-evaluation configuration was far too small for extracting further safe
-conclusions.
-
+%%
+In the second stage,  \emph{UTAPS} employs a set of heuristics
+which target those peers that expose low \emph{RTT}
+and are found within certain hop-counts away from 
+observed routers. %%AD u had "measured routers" - how do you select these routers?
+% For the selection part, a family of heuristics are proposed, which in general
+% target those peers that expose low RTT and are within certain hop-count away of
+% measured routers. 
+%%
+%%AD if you leave this "the authors claim..." and an such an author gets 
+%%	to spend 30 secs with the paper, you are dead! dead dead!!! 
+%%
+% The author's claim that peers running UTAPS instead of random
+% peer selection can achieve better downloading rates with reduced burden on the
+% underlying ISP infrastructure though their technique is coarse-grained and their
+% evaluation configuration was far too small for extracting further safe
+% conclusions.
+It is reported in~\cite{LCY2008} that peers running \emph{UTAPS}
+instead of the random peer selection, achieve better downloading rates
+and at the same time offer reduced burden on the underlying \emph{ISP}s. 
+\emph{UTAPS} stands as follows in regard to the $3$ criteria:
+%%
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{An Effective Network-Aware Peer Selection Algorithm in
-BitTorrent}
-Qin et al., in their work in \cite{QLZG2009}, propose a
-clustering approach that differentiates peers in a swarm into local, intra-ISP
-and inter-ISP neighbors. The classification is done through the routers they
-use in order to communicate with the tracker. For this, a newly arriving
-BitTorrent peer trace-routes the tracker and sends the results in a form of a
-vector of triples that contain the IP and the hop counter for the router on the
-trace-route path as well as the link latency of the hop that was made in order
-to arrive to this router. The tracker uses this information in order to classify
-the router through a $k$-means classification algorithm, with $k = 3$. For peer
-selection, the authors propose a biased approach where the tracker returns $c$
-close peers and $d = N - c$ distant peers, where $N$ is the length of the
-returned list. In choosing $c$, the tracker employs an iterative search process
-first in the peer's local neighborhood. If not enough peers are available
-there, it goes to intra-ISP and subsequently, if needed, to the inter-ISP
-cluster. The evaluation of the algorithm is done in a similar artificial,
-contained laboratory environment as in \cite{LCY2008} and it shows up to
-5\% faster downloading times, and up to 22\% reduction of the total cross-ISP
-traffic.
-
+A clustering approach that differentiates peers encountered in a swarm
+in those that are local, intra-\emph{ISP} and inter-\emph{ISP} is 
+presented in~\cite{QLZG2009}.
+The classification is carried with the help of the routers used 
+while peers communicate with the tracker.
+% \subsubsection{An Effective Network-Aware Peer Selection Algorithm in
+% BitTorrent}
+% Qin et al., in their work in \cite{QLZG2009}, propose a
+% clustering approach that differentiates peers in a swarm into local, intra-ISP
+% and inter-ISP neighbors. 
+% The classification is done through the routers they
+% use in order to communicate with the tracker. 
+%%
+%%AD In the following sentence it is not clear who sends to WHOM and WHY?
+For this, a newly arriving
+{\sl BitTorrent} peer trace-routes the tracker and 
+sends the results in a form of a
+vector of triples that contain the \emph{IP} 
+and the hop counter for the router on the
+trace-route path as well as the link latency 
+of the hop that was made in order to arrive to this router. 
+%%
+The tracker uses this information in order to classify
+the router through a $k$-means classification algorithm, with $k = 3$. 
+For peer selection, \cite{QLZG2009} proposes a biased approach 
+where the tracker returns $c$ close peers 
+and $d\;=\;N-c$ distant peers, where $N$ is the length of the returned list. 
+In choosing $c$, the tracker employs an iterative search process
+first in the peer's local neighborhood. If not sufficient number of 
+peers are available there, the tracker goes to intra-\emph{ISP} 
+and subsequently, if needed, to the inter-\emph{ISP} cluster. 
+%%
+Experimentation in a controlled environment indicates up to
+5\% faster downloading times as well as up to 22\% reduction of 
+the total cross-\emph{ISP} traffic.
+% The evaluation of the algorithm is done in a similar artificial,
+% contained laboratory environment as in \cite{LCY2008} and it shows up to
+% 5\% faster downloading times, and up to 22\% reduction of the total cross-ISP
+% traffic.
+This clustering approach fares as follows in regard 
+to the $3$ criteria:
+%%
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Peer-exchange Routing Optimization Protocols (PROP)}
-\cite{QCYCZ2007} introduces two protocols, one generic and one optimized,
-collectively referred to as \emph{Peer-exchange Routing Optimization Protocols
-(PROP)} to adjust the neighborhood graph of an overlay network in order to
-reduce its overall link latency. PROP algorithms are based on the exchange of
-neighbors among peers, driven by the mutual benefit of the participants to
-reduce the network delay among their connections. This ``collaboration'', is
-what differentiates this approach from others by allowing two peers to optimize
+
+In~\cite{QCYCZ2007}, two protocols termed 
+\emph{Peer-exchange Routing Optimization Protocols (PROP)} 
+are introduced to help adjust the neighborhood graph of an overlay network
+while aiming at reducing the network's overall link latency.
+%%
+% \subsubsection{Peer-exchange Routing Optimization Protocols (PROP)}
+% \cite{QCYCZ2007} introduces two protocols, one generic and one optimized,
+% collectively referred to as \emph{Peer-exchange Routing Optimization Protocols
+% (PROP)} to adjust the neighborhood graph of an overlay network in order to
+% reduce its overall link latency. 
+\emph{PROP} algorithms are weaved around the concept of 
+``exchange'' of neighbors among peers; 
+this is carried out so that participant node can collectively 
+benefit from the attained reduction in network delays. 
+%%
+This ``collaboration'', is what differentiates this approach 
+from others by allowing two peers to optimize
 their neighborhood environment, than simply letting each node to ``selfishly''
-choose its own strategy. In PROP-G (Generic), a peer can exchange all of its
-neighbors with another peer, while in PROP-O (Optimized) two peers exchange a
-selected subset of their neighbors. PROP-G is a generic protocol that guarantees
-the connectivity of the overlay graph during exchanges, therefore can be applied
+choose its own strategy. 
+In the generic form of the \emph{PROP-G} protocol,
+a peer can exchange all of its
+neighbors with those of another peer. 
+The optimized version \emph{PROP-O} permits peers to 
+exchange to exchange selective sets of neighbors.
+%%AD give a sentence or two how this gets done..
+The generic \emph{PROP-G} protocol guarantees
+the connectivity of the overlay graph during exchanges, 
+therefore, it can be applied
 to both unstructured and structured overlay networks.
-
+The following table depicts the behavior of the \emph{PROP} 
+algorithms regarding the stated criteria:
+%%
 \begin{center}
+{\footnotesize
 \begin{tabular}{rccc}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Distributed Domain Name Order (DDNO)}
-\emph{Distributed Domain Name Order (DDNO)} \cite{Z-YK2005} uses the domain
-names to detect topologically close nodes based on the assumption that nodes
-within the same domain are topologically close to each other. DDNO is a
-heuristic approach to solve the topology mismatch problem and uses
-\emph{Split-Hash} and \emph{dnMatch} algorithms to detect locality by using
-domain names. The result is a flat overlay topology but with some changes it
-can be utilized in super-peer architectures as well. According to the algorithm,
-half of the possible connections of a node is used to connect to local peers
-(called \emph{sibling} connections), and the other half is used to randomly
-connect to the peers anywhere on the network (called \emph{random} connections).
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+% \subsubsection{Distributed Domain Name Order (DDNO)}
+The \emph{Distributed Domain Name Order (DDNO)} approach~\cite{Z-YK2005} 
+uses domain names to detect topologically-close nodes.
+The fundamental assumption of the approach is the nodes found in 
+the same domain are also topologically close.
+While seeking to address the mismatch problem, \emph{DDNO} 
+is a heuristic approach that uses
+\emph{Split-Hash} and \emph{dnMatch} algorithms 
+to detect locality based on domain names. 
+%%
+\emph{DDNO} outcome is 
+a flat overlay topology which with some adjustments 
+can be utilized in super-peer architectures as well. 
+According to the algorithm,
+half of the possible connections of a node are used to connect to local peers
+(called \emph{sibling} connections) and rest are used to randomly
+connect to the peers anywhere on the network 
+(called \emph{random} connections).
 The former ensure the reduction of long distance traveling for messages, while
 the later secure the connectivity of the structure and prevent partitioning.
+%%AD partitioning of what? network partitioning??
 Upon arrival the newcomer asks a \emph{host-cache} to establish its
-random connections. For establishing its sibling connections, the newcomer
-multicasts a special walker message to help it locate the right
-candidate. If it is found then the candidate as well as its own siblings will
+random connections. 
+For establishing its sibling connections, the newcomer
+multicasts a specialized \emph{walker message} to 
+locate the right candidate(s). 
+%%AD I am not sure what the following sentence says... It is too convoluted..
+If it is found then the candidate as well as its own siblings will
 respond to the newcomer about their willingness to establish connections with
 it.
-
+%%AD I am not sure what this paragraph offers.. I am somewhat puzzled: is 
+%%	this about what exactly?? Some refitting of the short description 
+%%	with the mismatch problem I think is needed...
+%%	In general, the description ends somewhat abruptly...
+%%
+%%
+In terms of the $3$ criteria, \emph{DDNO} fares are follows:
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Critical Topology-Aware Grouping (CTAG)}
-\emph{Critical Topology-Aware Grouping (CTAG)} presented in \cite{ZL2006}, is a
-grouping algorithm. The grouping strategy is based on the \emph{Internet
-Assigned Numbers Authority (IANA)} and the respective \emph{Regional Internet
-Registry (RIR)}'s IP assignment strategies, according to which nodes within the
-same organization are always addressed from the same block. The paper proposes
-the \emph{Adjacency Measurement (AM)}, a technique which uses the longest
-matching IP segment criterion to calculate node proximity. CTAG focuses on both
-the construction of the overlay as well as its constant and dynamic adaptation.
-In the first, called \emph{bootstrapping grouping}, the \emph{Gnutella Web
-Caching} mechanism is modified in order for a newcomer to choose the closest
-cache to retrieve the bootstrapping list. In the second, called \emph{dynamic
-revision} in the paper's context, the AM metric is used to store hosts'
+% \subsubsection{Critical Topology-Aware Grouping (CTAG)}
+The \emph{Critical Topology-Aware Grouping (CTAG)} approach~\cite{ZL2006}
+advocates a grouping strategy of peers based on the 
+\emph{Internet Assigned Numbers Authority (IANA)} and 
+the respective \emph{Regional Internet Registry (RIR)}'s IP assignments.
+Based on these assignments, nodes that belong in the same 
+organization are always addressed from the same block.
+%%AD above, I simply do not understand the ..."from the same block"
+%%	What does it mean??? 
+%%
+% \emph{Critical Topology-Aware Grouping (CTAG)} presented in \cite{ZL2006}, is a
+% grouping algorithm. The grouping strategy is based on the \emph{Internet
+% Assigned Numbers Authority (IANA)} and the respective \emph{Regional Internet
+% Registry (RIR)}'s IP assignment strategies, according to which nodes within the
+% same organization are always addressed from the same block. 
+\emph{CTAG} proposes the \emph{Adjacency Measurement (AM)}, 
+a technique which uses the longest
+matching \emph{IP} segment criterion to compute node proximity. 
+\emph{CTAG} focuses on both
+the construction of the overlay as well as its 
+constant and dynamic adaptation.
+In the first phase, called \emph{bootstrapping grouping}, 
+the {\sl Gnutella} web caching mechanism is modified 
+so that a newcomer may choose the closest %%AD closest here means what???
+cache to retrieve the bootstrapping list. 
+%%AD for sentence below - what is AM matric??  X-Try?? - 
+%%		can you rework the sentence?
+During the second phase called \emph{dynamic revision}, 
+the AM metric is used to store hosts' 
 addresses read from \emph{X-Try} headers during handshake or from
-\emph{QueryHit} messages. Additionally, when a node reaches the max neighbor
-connections, the node disconnects the neighbors with the lowest AM.
-
+\emph{QueryHit} messages. 
+When a node reaches its maximum number of neighbor
+connections, it disconnects those neighbors with the lowest AM.
+\emph{CTAG} behaves as follows regarding the $3$ criteria:
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Landmark Binning (LM)}\label{sec:landmark_binning}
+% \subsubsection{Landmark Binning (LM)}\label{sec:landmark_binning}
 % \begin{figure*}
@@
 
-Landmark Binning \cite{RHMKS2002} is an approach to partition close
-by nodes into bins based on their distance to well known anchor nodes across the
+\emph{Landmark Binning}~\cite{RHMKS2002} partitions close-by nodes 
+into bins based on their distance from well known anchor nodes across the
 Internet.%, as shown in Figure~\ref{fig:landmark_binning}.
-In order to detect locality, nodes use network latency (i.e. RTT) as a
-measurement technique. The network latency, even though not always accurate, is
-selected in this work due to its non-intrusive, transparent and easy to apply
-behavior. In order for the binning to work, a few anchor servers with known
-physical locations need to be installed in strategic positions accross the the
-Internet. The authors estimate that a number of 12 such servers can prove sufficient
-for this. Every newly arriving node measures its distance from these landmarks
-and unilaterally decides to join a specific bin based on these results.
+To detect locality, peers mainly use network latency (i.e. \emph{RTT}).
+%  as a measurement technique. 
+Despite the fact that such delays are not always accurate,
+they are used in~\cite{RHMKS2002} for they are 
+non-intrusive, transparent and easy to apply.
+%%
+For the binning to work, a few anchor servers with known
+physical locations need to be installed in strategic positions 
+across the the Internet. 
+It is conjectured~\cite{RHMKS2002} that $12$ such servers can prove
+sufficient for the task.
+An arriving node measures its distance from these landmarks
+and unilaterally decides to join a specific bin based on its measurements.
 Specifically, the node measures its round-trip time to each of the landmarks and
-orders the values in a decreasing order. The ordering represents a \emph{bin},
+orders the resulting \emph{RTT}-values in a decreasing order. 
+%%AD I do not understand what the following sentence says..
+%%	for each peer there is a different bin right? 
+The ordering represents a \emph{bin},
 in the sense of close-by nodes having the same landmark ordering and hence
-belong to the same bin. This means that a landmark system consisting of $m$ such
-nodes results in $m!$ potential different bins.The operation of the method is
-independent of the model incorporated by the overlay network and thus it can be
-applied with no significant changes to both structured and unstructured P2P
-systems. \cite{RHMKS2002} has the detailed description of the algorithm for
-both architectures. Landmark binning is also considered as a good candidate for
-content distribution networks. The major disadvantage of landmark
-binning is that landmark servers must be installed and maintained
-on different autonomous
-systems, all over the world. A typical P2P network usually has a couple of
-million nodes connected at any given time, thus rendering landmark servers an
-important parameter when designing for large scales.  The authors claim that
-latency estimation does not drain much from the network resources but in case
-scalability problems arise, they claim the solution is replacing single landmark
-servers with clusters of servers within the same physical area. However, this
-approach does not reduce the possibility of excessive network traffic flow
-through these landmark servers. One other possible problem is the incorrect
-binning caused by the use of inaccurate methods for delay measurement
-(latency is not proven to be a safe way to estimate distance).
-
+belong to the same bin. 
+%%
+%%AD I think the above sentence is the main information related in this 
+%%   piece of description so it requires a bit better description...
+Should there $m$ landmarks be adopted, potentially $m!$  different bins exist.
+% This means that a landmark system consisting of $m$ such
+% nodes results in $m!$ potential different bins.
+%%
+The operation of the method is independent of the model 
+incorporated by the overlay network and thus it can be
+applied with no significant changes to 
+both structured and unstructured \p\ systems. 
+% \cite{RHMKS2002} has the detailed description of the algorithm for
+% both architectures. %%AD obviously!!!!
+%%
+%%AD why is it important to say this about CDNs???
+% Landmark binning is also considered as a good candidate for
+% content distribution networks. 
+The main disadvantage of the approach is that 
+landmark servers must be installed and maintained throughout the world.
+Provided that a \p-system often may have a few million nodes 
+connected at any given point in time, landmark servers do 
+inherently play a pivotal role in the operation of the network.
+%%AD rephrased what is below as above...
+% A typical P2P network usually has a couple of
+% million nodes connected at any given time, thus rendering landmark servers an
+% important parameter when designing for large scales.  
+%%%%%%%%%%%%%%%%%%
+%%AD rephrased what is below as above...
+Although latency estimation does not drain network resources,
+the scale of the approach does call for an alternative design.
+In this respect, \cite{RHMKS2002} advocates the replacement of 
+each landmark server by a cluster of servers.
+However, this approach may not entirely avoid the creation of 
+excessive network traffic flow through its landmark server-clusters.
+%%AD rephrased what is below as above...
+% The authors claim that
+% latency estimation does not drain much from the network resources but in case
+% scalability problems arise, they claim the solution is replacing single landmark
+% servers with clusters of servers within the same physical area. However, this
+% approach does not reduce the possibility of excessive network traffic flow
+% through these landmark servers. One other possible problem is the incorrect
+% binning caused by the use of inaccurate methods for delay measurement
+% (latency is not proven to be a safe way to estimate distance).
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%
+%%
 %The algorithm can be considered scalable, as nodes need only compute distances
@@
 %This is implemented with an inter-bin to an intra-bin latency ratio.
-
 %
@@
 %topologies\cite{merugu_str2unstr_2003}.
-
 %
@@
 %
-
+%%
+The landmark binning behaves as follows:
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{mOverlay}
-\emph{mOverlay} \cite{ZZZSZ2004} tries to addresses the scalability issues that
-might be imposed on networks (i.e. load unbalance) when using static landmark
-servers by introducing dynamic landmarks. Zhang et al. introduce the notion of a
-\emph{group} which is a set of peers considered close to each other with respect
-to any position in the underlying network, where proximity is defined by some
-user-defined cost metric like RTT, latency or something else. mOverlay is
-considered a clustering approach since it tries to recreate small-world-like
-properties to the overlay network by creating a two-level hierarchical
-structure, where on the top level we have connections between groups while on
-the bottom we have connections between peers inside groups. Finding the correct
-closest group is the most important part of the overlay construction. Nodes are
-grouped based on their distance to the groups already in the network, rendering
-the latter (dynamic) landmarks for the process. The authors formally prove
-that any new node can reach its group by exchanging at most $O(logN)$ messages
-within the network. Finally mOverlay also considers a second important function
-in its protocol which is the constant maintenance of the overlay.
-
+% \subsubsection{mOverlay}
+The \emph{mOverlay} approach addresses scalability issues
+that might be arise when static landmark servers are in use. 
+To this end, the use of dynamic landmarks is proposed.
+% \emph{mOverlay} \cite{ZZZSZ2004} tries to 
+% addresses the scalability issues that
+% might be imposed on networks (i.e. load unbalance) when using static landmark
+% servers by introducing dynamic landmarks. 
+\emph{mOverlay}'s founding notion is that of a \emph{group} that 
+designates a set of peers found in close proximity.
+This proximity is user-defined in the protocol and 
+may involve metrics including \emph{RTT}s and network latencies.
+%%AD rephrased what is below as above... 
+% Zhang et al. introduce the notion of a
+% \emph{group} which is a set of peers considered close to each other with respect
+% to any position in the underlying network, where proximity is defined by some
+% user-defined cost metric like RTT, latency or something else. 
+%%%%%%%%%%%%%%%%%%%%%%%
+By and large a clustering approach, \emph{mOverlay} seeks to 
+recreate small-world-like properties by producing a 
+two-level hierarchical structure:
+at the top level, there are only connections among groups 
+while at the bottom, only \emph{intra-group} connections occur among peers.
+%%AD rephrased what is below as above... 
+% mOverlay is considered a clustering approach since 
+% it tries to recreate small-world-like
+% properties to the overlay network by creating a two-level hierarchical
+% structure, where on the top level we have connections between groups while on
+% the bottom we have connections between peers inside groups. 
+%%%%%%%%%%%%%%%%%%%%%%%
+Clearly, identifying groups and accurately finding the closest group 
+to a peer is a fundamental concern in the creation of the overlay.
+% Finding the correct closest group is the most important part 
+% of the overlay construction. 
+Nodes are grouped based on their distance to 
+the groups already in the network, rendering
+the latter be the \emph{dynamic landmarks} in the process. 
+%%AD how do you identify the initial groups????
+%%
+It is shown that in the proposed scheme, a new peer can 
+reach its group by expending at most $O(logN)$ messages
+within the network. 
+Finally, \emph{mOverlay} maintains stability and constant 
+overheads when a host either fails or departs the network;
+this is achieved through periodic cache updates and group 
+leader selections, should one either leaves or dies.
+%%AD rephrased what is below as above...
+% The authors formally prove
+% that any new node can reach its group by exchanging at most $O(logN)$ messages
+% within the network. Finally mOverlay also considers a second important function
+% in its protocol which is the constant maintenance of the overlay.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
 %
@@
 %of proximity, in the overlay, in order to efficiently detect the neighboring
-%groups of a newly formated group from the set of adjacent groups of its closest
+%groups of a newly formed group from the set of adjacent groups of its closest
 %neighbor. Additionally, during \emph{group joining} the corresponding protocol
@@
 %
-
+%%
+In terms of the stated criteria, \emph{mOverlay} fares as follows:
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 influence on performance. In surveying the above algorithms, we identify
-four main methologolies that are used either selectively or in combination
+four main methodologies that are used either selectively or in combination
 by each algorithm.  

### 2014-01-29 unstructured.tex
2013.04.03 (Alex) More on the unstructured algorithms text.

@@
 
-In this section, we introduced the topology mismatching problem from the
-perspective of unstructured P2P systems and we visited a series
-of algorithms that have been proposed in the literature to tackle its negative
-influence on performance. In surveying the above algorithms, we identify
-four main methodologies that are used either selectively or in combination
-by each algorithm.  
-These methodologies are:
+While surveying efforts to overcome the mismatch problem
+in the area of unstructured \p\ systems,
+we came to identify four key methodologies 
+utilized by the discussed approaches; these methodologies are:
+%%AD shorten below as above...
+% In this section, we introduced the topology mismatching problem from the
+% perspective of unstructured P2P systems and we visited a series
+% of algorithms that have been proposed in the literature to tackle its negative
+% influence on performance. In surveying the above algorithms, we identify
+% four main methodologies that are used either selectively or in combination
+% by each algorithm.  
+% These methodologies are:
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{enumerate}[\itshape i\upshape)]
@@
 \end{enumerate}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Topology adaptation}
+Each surveyed approach uses either one or more of the
+aforementioned four methodologies. 
+Below, we discuss how the proposed protocols use 
+elements of the four methodologies and we offer 
+a summary qualitative comparison for all surveyed approaches 
+applicable for \emph{unstructured} \p \emph{systems}.
+%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Topology Adaptation Methodology}
+
+Protocols based on topology adaptation 
+modify the topology of the \p\ network
+using various techniques. 
+The two most commonly used approaches to 
+topology adaptation are creating
+\emph{spanning tree}s using connection graphs 
+and creating \emph{clusters} of physically close nodes.
+
+\emph{Narada} and subsequent algorithms including 
+\emph{AOTO, LTM, SBO}, try to solve the problem
+by building a richer connected graph and forming 
+minimum spanning trees over
+this graph that can efficiently route messages among peers. 
+Even though spanning trees provide efficient query performance, 
+their maintenance costs have proven to be a significant limitation. 
+This is why the original \emph{Narada} proposal
+focused on small groups and thus, is not scalable.  
+%%
+The \emph{AOTO, LTM} and \emph{SBO} try to 
+% Algorithms like AOTO, LTM or SBO try to %%AD this "like" is colloquial 
+					  %%   try to avoid it if you can!
+overcome these obstacles with 
+ingeneous schemes
+% clever tricks %%AD this is a bit colloquial
+that entail forming minimum spanning trees
+for the $2$-hop neighbors ($N^2$) for each node, 
+%%AD u imply each has N neighbors above right???
+partitioning the graph into two
+random groups where each group is responsible for different tasks, 
+or performing local optimizations dynamically on the overlay graph. 
+%%AD what are these "optimizations" (in 2-3 words  - in sentence)
+The advantage of building minimum spanning trees is that 
+they maintain the connectivity on the network in
+an efficient manner while still preserving the overall search scope. 
+However, their construction and the update costs, 
+especially in dynamic, high-churn environments, 
+causes large additional traffic overhead on the underlying
+network~\cite{CRZ2000,CRSZ2001,CRSZ2002}.
+
+%%
+%%AD this figure HAS to be referenced in the text!!! 
+%%   ALSO - is this ANY different from Fig 4 of the paper??? 
+%%   WHY this Fig is being repeated here??? Grand Mystery!!!!
+%%	Perhaps we should put one fig similar but more sexy??? ;-)
 \begin{figure}[ht]
@@
 \subfigure[Inefficient overlay topology.] {
-  \includegraphics[scale=0.4]{img/pdf/topology-adaptation-before.pdf}
+  \includegraphics[scale=0.3]{img/pdf/topology-adaptation-before.pdf}
   \label{figure:topology-adaptation:before}
@@
 \subfigure[Efficient overlay topology after adaptation.] {
-  \includegraphics[scale=0.4]{img/pdf/topology-adaptation-after.pdf}
+  \includegraphics[scale=0.3]{img/pdf/topology-adaptation-after.pdf}
   \label{figure:topology-adaptation:after}
 }
-\caption{A simple example of how an overlay topology can be adapted.}
+\caption{A simplified example on how overlay topology adaptation may improve matters.}
+% A simple example of how an overlay topology can be adapted.}
 \label{figure:topology-adaptation}
 \end{figure}
-
-The topology adaptation based protocols modify the topology of the P2P network
-using various techniques. The two most commonly used approaches to 
-topology adaptation are creating
-\emph{spanning tree}s using connection graphs, and creating \emph{clusters} of
-physically close nodes.
-
-Narada and subsequent algorithms like AOTO, LTM, SBO, try to solve the problem
-by building a richer connected graph and forming minimum spanning trees over
-this graph that can efficiently route messages among peers. Even though spanning
-trees provide efficient query performance, their maintenance costs have 
-proven to be a significant limitation. This is why the original Narada 
-algorithm 
-was intended for small
-groups and is thus, not scalable.  Algorithms like AOTO, LTM or SBO try to
-overcome these obstacles with clever tricks like forming minimum spanning trees
-for the two-hop neighbors ($N^2$) of each node, partitioning the graph into two
-random groups where each group is responsible for different tasks, or performing
-local optimizations dynamically on the overlay graph. The advantage of building
-minimum spanning trees is that they maintain the connectivity on the network in
-an efficient manner while still preserving the overall search scope. However,
-their construction and the update costs, especially in dynamic, high-churn
-environments, cause large additional traffic overhead on the underlying
-network \cite{CRZ2000,CRSZ2001,CRSZ2002}.
-
-The cluster-based approaches, on the other hand, link physically
-closer nodes to each other. T2MC, for example, uses trace-route logs and DDNO
-uses domain names to cluster nearby nodes. Unfortunately, commonly used
+%%
+
+The cluster-based approaches, on the other hand, 
+link physically-close nodes to each other. 
+\emph{T2MC}, for example, uses trace-route logs and 
+\emph{DDNO} exploits domain names to cluster nodes in proximity. 
+Unfortunately, commonly used
 methods for proximity detection across Internet do not always return
-reliable results and therefore mapping accuracy is not guaranteed. Moreover,
-trace-route 
-is a heavy-weight utility and many network equipment vendors do not allow
-trace-route calls at all. The most problematic aspect of clustering, though, is
-its nature per se. Limited connectivity among the various local domains can
+reliable results and therefore, mapping accuracy is not guaranteed. 
+Moreover, trace-route is a computing resource ``heavy-weigth'' utility 
+and so  a number of network equipment vendors do not allow
+trace-route calls at all. %%AD do you have a refeerence or two in here???
+%%
+The most problematic aspect of clustering, though, is
+its nature per se. 
+Limited connectivity among the various local domains can
 significantly shrink the search scope, negatively affecting the query response
-time that the P2P end-user experiences. Techniques, like DDNO, that try to 
-balance the efficiency of the clustering approach with improved
-connectivity have also been proposed.  Specifically, DDNO attempt to address this
-limitation by forcing half of each node's connections to be with other, randomly
-selected, nodes over the network.
-
-In brief, topology adaptation ultimately aims to reduce the 
-average cost of the paths from one node to another so that 
-the overlay network is better fit on top of the underlying IP network.
-This is fair in the general case but not always. For example an
-adaptive algorithm can exchange a slow edge in the overlay with a faster one
-thus reducing the average latency of the network communication. The pitfall here
-is that this new virtual link can traverse a fast AS-to-AS
+time that the \p\ user experiences. 
+% Techniques, like DDNO, that try to 
+Among others, \emph{DDNO} tries to 
+balance the efficiency of the clustering approach 
+with enhanced node connectivity. 
+In particular, \emph{DDNO} attemps to address this limitation
+by forcing half of each node's connections to be with other, randomly
+selected, nodes in the network.
+%%AD you mean the "logical network"? the p2p network???
+
+As a methodology, 
+topology adaptation ultimately aims to reduce the 
+average path traversing cost from one node to another.
+In doing so, topology adaptation  re-arranges the 
+overlay network so that it becomes a better fit 
+for the underlying \emph{IP} network
+as Figure~\ref{figure:topology-adaptation} depicts.
+Despite the fact that the above is a plausible proposition,
+it is not always advantageous.
+% This is fair in the general case but not always. 
+%%
+For example, an topology adaptive algorithm can exchange 
+a slow edge in the overlay with a faster one
+thus reducing the average latency of the network communication. 
+The pitfall here is that this new virtual link 
+can traverse a fast \emph{AS}--to--\emph{AS} 	
+			%%AD you mean Autonomous Systems here?? 
+		  	%%    What is this AS???
 link meaning that even though message
-round-trip-time is reduced, it has more cost in terms of communication network economics
-and management.
+round-trip-time is reduced, it has more cost 
+in terms of communication network economics and management.
+%%AD just check the above sentence for correctness..
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Forwarding optimization}
-
-\begin{figure}[ht]
-\centering
-\subfigure[Typical blind flooding.] {
-\includegraphics[scale=0.25]{img/pdf/forwarding-optimization-before.pdf}
-  \label{figure:forwarding-optimisation:before}
-}\qquad\qquad
-\subfigure[A node can decide where to forward the messages.] {
-  \includegraphics[scale=0.25]{img/pdf/forwarding-optimization-after.pdf}
-  \label{figure:forwarding-optimisation:after}
-}
-\caption{Forwarding optimization in constraining message flooding.}
-\label{figure:forwarding-optimisation}
-\end{figure}
+\subsubsection{Forwarding Optimization Methodology}
 
@@
 % as well as content distribution networks. Application layer multi-cast,
-however,
+% however,
 % is not as efficient as the one enabled by IP and still struggles to ease the
@@
 
-Forwarding-optimization-based
-approaches, propose intelligent forwarding link selection for a message's next
-hop across the routing path. The selection criteria vary depending on the
-algorithm, but they mainly use one or more statistical heuristics. These can
-range from simple obvious ones, like the latency of the link between the nodes or
-the exposure of specific features, such as high processing bandwidth, high
-capacity, or low latency, to more sophisticated alternatives that may even take
-into account application level requirements such as the reliability or
-friendship. Figure~\ref{figure:forwarding-optimisation} shows a simple example
-of how forwarding optimization can work.
-Sub-figure~\ref{figure:forwarding-optimisation:before} shows an algorithm that
-floods the entire network for search of an object the existence of which is
-denoted by black nodes. Alternatively, a node can decide to forward its
-messages not to every output link but to a specific or subset of its output
-links. In Sub-figure~\ref{figure:forwarding-optimisation:after}, the node in the
-middle, does not flood its neighbors. Instead it picks one of them. The dashed
-line on the right, shows that, for this routing process, the algorithm has
+%%AD I think you will need to sprinkle some references of those surveyed 
+%%	in this subsubsection.. As opposed to what done above, 
+%%	this subsubsection has no reference in the text!!!!
+Approaches based on forwarding-optimization 
+propose intelligent forwarding link selection for a message's next
+hop across a routing path. 
+The selection criteria vary depending on the
+algorithm, but they mainly use one or more statistical heuristics. 
+These can range from simple and rather obvious ones, 
+including the latency of the link between the nodes or
+the exposure of specific features such as high processing bandwidth, high
+capacity, or low latency, 
+to more sophisticated alternatives that may even take
+into account application level requirements such as 
+the reliability or friendship.
+
+Figure~\ref{figure:forwarding-optimisation} shows a simple example
+of how forwarding optimization may work.
+%%
+\begin{figure}[ht]
+\centering
+\subfigure[Typical blind flooding.] {
+\includegraphics[scale=0.20]{img/pdf/forwarding-optimization-before.pdf}
+  \label{figure:forwarding-optimisation:before}
+}\qquad\qquad
+\subfigure[A node can decide where to forward the messages.] {
+  \includegraphics[scale=0.20]{img/pdf/forwarding-optimization-after.pdf}
+  \label{figure:forwarding-optimisation:after}
+}
+\caption{Forwarding optimization in constraining message flooding.}
+\label{figure:forwarding-optimisation}
+\end{figure}
+%%
+Fig.~\ref{figure:forwarding-optimisation:before} 
+% Sub-figure~\ref{figure:forwarding-optimisation:before} 
+shows a protocol that simply floods the entire network 
+in search of object found in the nodes colored in black.
+% for search of an object the existence of which is
+% denoted by black nodes. 
+Alternatively, a node can decide to forward its
+messages not to every output link but to a specific 
+or subset of its outgoing links. 
+In Sub-figure~\ref{figure:forwarding-optimisation:after}, the node in the
+In Fig.~\ref{figure:forwarding-optimisation:after}, 
+the node in the middle, does not flood its neighbors as instead 
+in only picks one of them.
+The dashed line on the right, shows that, 
+for this routing process, the protocol has
 rendered two potential forwarding paths as inefficient (e.g., the target nodes
-show overloading signs). These approaches have the advantage of enhancing the
-search responsiveness and reducing the aggregate resource usage of the physical
-network. On the other hand, they suffer from drastic reduction of the search
-scope (in Sub-figure~\ref{figure:forwarding-optimisation:after} the object on
+show overloading signs). 
+On one hand, forward
+optimisation approaches have 
+the advantage of enhancing the
+search responsiveness and reducing the aggregate resource 
+usage of the physical network. On the other, 
+they suffer from drastic reduction of the search
+scope (in Fig.~\ref{figure:forwarding-optimisation:after} the object on
 the far right is not reached) thus limiting the scalability of the whole
-network. Moreover they do not actually address the problem of topology mismatch
-since they do not give any guarantees that overlay and underlying topologies are
-aligned with each other let alone quantify the mismatch and try to alleviate.
-Like all other approaches, forwarding based methods are commonly applied in
-conjunction with other methods to improve the quality of the P2P systems.
+network. 
+Moreover, they adress the problem of the mismatch in a limited manner
+% do not actually address the problem of topology mismatch
+since they do not provide any guarantees 
+that overlay and underlying topologies are
+aligned with each other let alone quantify the mismatch 
+and try to alleviate.
+% Like all other methodologies, 
+Forwarding based optimisations are commonly applied in
+conjunction with other methodologies to improve 
+the quality of the \p\ systems.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Caching and replication}
-
+\subsubsection{Caching and Replication Methodology}
+
+%%AD in this portion there is NOT much about replication... 
+%%	ALSO, references to surveyed works as missing.. (other than
+%%	KAzaa!) -  Sprinkle some around the text...
+Caching is widely used to exploit locality and minimize redundant
+transfer of data. Caching has with much success been successfully 
+adopted by web and file server application environments. 
+Since peers in a \p\ system also operate as servers,
+it is intuitively expected that \p\ file sharing systems can also benefit from
+caching in improving performance and reducing overall resource usage. 
+However, the design of caches in this context 
+is non-trivial compared to the web-based caching. 
+Due to the fact nodes play the double role of server and client,
+two important issues have to be considered at design time. 
+First, the lifetime of a query is short, as the nodes join 
+and leave frequently. Second, the result
+of a single query string is not always the same, as this depends on the
+source of the query, the \emph{TTL} value set for the messages, 
+the current interconnection of peers and the 
+high volatility of the environment. 
+Thus, to develop a successful caching system for a \p\ architecture, these
+parameters also have to be carefully considered. 
+\p\ caching/replication can be applied
+at two different levels, namely caching indices or pointers to data 
+(Fig.~\ref{figure:replication:index}) or caching the data itself 
+(Fig.~\ref{figure:replication:data}). 
+%%%
 \begin{figure}[ht]
@@
 \end{figure}
-
-Caching is a widely used technique to exploit locality and minimize redundant
-transfer of data, previously successfully adopted by web and file server
-application environments. Since peers in a P2P system also operate as servers,
-it is intuitively expected that P2P file sharing systems can also benefit from
-caching in improving performance and reducing overall resource usage. However,
-the design of caches in this context is non-trivial compared to the web-based
-caching. Due to the fact that each node can act as both a server and a client,
-two important issues have to be considered at design time. First, the lifetime
-of a query is short, as the nodes join and leave frequently. Second, the result
-of a single query string is not always the same, as it is dependent on the
-source of the query, the TTL value set for the messages, the current
-interconnection of peers and the high volatility of the environment. Thus,
-to develop a successful caching system for a P2P architecture, these
-parameters also have to be considered. P2P caching/replication can be applied
-at two different levels, namely caching indices or pointers to data (see
-Sub-figure~\ref{figure:replication:index}) or caching the data itself (see
-Sub-figure~\ref{figure:replication:data}). Successful implementations have
-already been developed in some commercial P2P systems, like KazaA, which uses
-caching of indices on its super-peer level. Unfortunately, even though the
-state of the art in P2P algorithms using caching methods reduce the resource
-usage of the network, they also fail to consistently address typical problems
-of a topologically mismatched environment such as duplication of messages.
+%%
+Successful implementations have
+already been developed in some commercial \p\ systems, like 
+\emph{KaZaA}, which uses
+caching of indices on its super-peer level. 
+%%%
+Even though the state of the art in \p\ protocols using caching methods
+helps reduce the burden of network resources, 
+their contribution in addressing the mismatch problem 
+remains limited.
+%%AD I am not sure what the following piece says....
+%%	tried to rephrase as above...
+% Unfortunately, even though the state of the art 
+% in \p\ protocols using caching methods reduce the resource
+% usage of the network, 
+% they also fail to consistently address typical problems
+% of a topologically mismatched environment such as duplication of messages.
+%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+In landmark-based algorithms, nodes use network delay (e.g., \emph{RTT}) as a
+distance measurement method to position themselves with respect to ``a priori''
+known servers on the Internet. 
+These landmark servers are used
+by nodes to estimate their positions based on the intuitive assumption that
+nodes
+with similar distances to a set of landmarks, are physically close to each
+other, as well, over the network. 
+In Figure~\ref{figure:landmarking} the newly
+arriving peer, measures its distance to an array of landmark point denoted as
+$Ln$ and and makes a sorted list of the peers, say in increasing order. Then in
+order to choose the already participating peers with which to connect to,
+compares its list to the list of the potential neighbor. Peers with similar
+measured distances to the these landmark points are likely to be close to
+each other as well.
+%%
 \begin{figure}[ht]
@@
 
-In landmark-based algorithms, nodes use network delay (e.g. RTT) as a
-distance measurement method to position themselves with respect to ``a priori''
-known servers (the landmark servers) on the Internet. Landmark servers are used
-by nodes to estimate their positions based on the intuitive assumption that
-nodes
-with similar distances to a set of landmarks, are physically close to each
-other, as well, over the network. In Figure~\ref{figure:landmarking} the newly
-arriving peer, measures its distance to an array of landmark point denoted as
-$Ln$ and and makes a sorted list of the peers, say in increasing order. Then in
-order to choose the already participating peers with which to connect to,
-compares its list to the list of the potential neighbor. Peers with similar
-measured distances to the these landmark points are likely to be close to
-each other as well.
-
-Landmark based protocols have four important drawbacks. First, the network
+Landmark based protocols have four important drawbacks: First, the network
 delay is not a reliable distance estimation method. For example, based on the
@@
 infrastructure across the whole Internet and for all the different autonomous
-system domains. As popular P2P file sharing applications usually have millions
+system domains. 
+As popular \p\ file sharing applications usually have millions
 of peers connected at any time, the network costs
-of maintaining these landmark servers will likely be quite high. A possible solution to
+of maintaining these landmark servers will likely be quite high. 
+A possible solution to
 the scalability problem of the static landmark servers is to use ordinary nodes
@@
 close distance to each other.
+%%AD ok how about ay advantages?? Any referencess??? 
+%%	Also, a bit of a discussion is missing on hybrid approaches surveyed
+%%	for example which approaches use a combo of these methodologies???
 

### 2014-01-30 background.tex
2013.05.20 (Vassilis) Minor changes to abstract, introduction, background and unstructured (tex, plot and tables) sections. Useless comment clean-up. Various rephrasing on some reviewed content. Added references to the tables. Some image revisited. Discussion text added explaining the fact that many of the papers combine multiple methodologies.

--- 2014-01-28 background.tex
@@
 
-%The abstraction of the overlay network has been proposed as a way to implement
-%efficient, fully distributed, application-layer services such as  
-%routing messages to destinations that are not known in advance or building
-%one-to-many addressing schemes 
-% TODO find citation for Narada & Adobe's Real Time Media Flow Protocol (RTMFP)
-%\cite{} that provide quality-of-service (QoS) guarantees. \emph{IP
-%multicast}, as well as other proposals such as \emph{IntServ} and
-%\emph{DiffServ} \cite{cisco_diffserv_2005}, have still not seen wide deployment
-%largely due to the fact they require support from the IP
-%infrastructure. On the other hand, an overlay network can be deployed on
-%end-systems running the overlay protocol software, without the cooperation of ISPs.
-%Academic research includes
-%\cite{chu_esm_2000,jannotti_overcast_2000,kwon_tag_2002} among others.
-
-In this section, we provide some background information on the types of
+In this section, we provide some basic background information on the types of
 \p\ architectures that have been implemented and deployed. We also
@@
 concept~\cite{gnutella}, a \emph{portmanteau} that blends the notions of server
-and client to denote the dual role of a node participating in \p-network. 
+and client to denote the dual role of a node participating in the \p-network. 
 Such server-less systems have proven to be able to achieve outstanding
@@
 (Figure~\ref{figure:p2p-archs:centralized}) were the first 
-to recognize that requests for resources 
+to observe that requests for resources
 (e.g., CPU cycles or popular content) need not be sent to
-a dedicated server. Instead, requests could be handled by 
+a dedicated server. Instead, such requests could be handled by 
 many hosts that already possessed the resources in question. 
-The downside of this approach was that it required 
-centralized search that became a single point of failure, a
-scalability bottleneck, and a target of malicious including \emph{DoS} attacks. 
-% \emph{Denial-of-Service (DoS) attacks}).
+The downside of this approach was that it required a 
+centralized search directory that invevitably becames a single point of failure,
+a scalability bottleneck or a target of malicious behavior such as \emph{DoS}
+attacks.
 
@@
 %%
-The central indexing played a vital role in the demise of 
-{\sl Napter} as the \emph{Recording
-Industry Association of America} successfully argued in court 
-copyright infringement of the artists songs traded in the system. 
-%
-% Napter's  centralized nature was what ultimately brought it down after Recording
-% Industry Association of America (RIAA) filed a lawsuit for copyright infringement
-% of the intellectual property of artists whose music was distributed by the
-% Napster application. RIAA was able to point to the centralized indexing
-% component of Napster as the entity responsible for illegal song-trading.
+The central indexing component played a vital role in the demise of 
+{\sl Napter}, as the \emph{Recording
+Industry Association of America} successfully argued in court that this software
+entity was enabling copyright infringement of the artists' songs exchanged
+through the system.
 
@@
 for this reason, such architectures are referred to as \emph{unstructured}. 
-Key properties  of such \p\ systems are that they
-support inherent heterogeneity of peers, 
-are highly resilient to peer failures,
-and incur low maintenance overhead at handling 
-the dynamics of peer participation \cite{stutzbach_churn_2006}. 
-Unstructured systems are also known as 
+Key properties of such \p\ systems are that they support inherent heterogeneity
+of peers, are highly resilient to peer failures, and incur low maintenance
+overhead at handling the dynamics of peer participation
+\cite{stutzbach_churn_2006}. Unstructured systems are also known as 
 \emph{broadcast-based} systems as they use message flooding among
@@
 does not guarantee that the resource of interest, (e.g., a rare recording)
-% of a musical group from the 1930s) 
-will be ultimately found.
+will, ultimately, be found.
 %%%
@@
 search queries were flooded over the much smaller subset of 
-ultra/super-peers, the disadvantages of reduced search scope and 
-inability to locate rare resources remained.
+ultra/super-peers, it fails to address the reduced search scope and
+inability to locate rare resources.
 
-The \emph{decentralized structured schemes}
-(Figure~\ref{figure:p2p-archs:structured}) sought to address the 
-disadvantages of the above decentralized unstructured \p--approaches.
+\emph{Decentralized structured schemes} were introduced as an answer to the above
+problems of decentralized unstructured \p--approaches
+(Figure~\ref{figure:p2p-archs:structured}).
 Here, the main objective was to offer 
@@
 \cite{ratnasamy_can_2001,stoica_chord_2001,antony_pastry_2001,zhao_tapestry_2001,maymounkov_kademlia_2002,rgrk_bamboo_2004}.
-\emph{Decentralized structured schemes}
-implement a \emph{Distributed Hash Table (DHT)} that maps objects 
-to nodes through a deterministic mechanism. 
-\emph{DHT}s provide a guaranteed bound on the
-number of overlay routing hops that have to be ``traveled''
-% crossed 
-to locate a resource, even in the case when only a \emph{single} copy exists.  
-This requires
-$O$($log(n)$) hops, compared to unstructured networks that require
-$O(n)$ to reliably locate any object. 
-Unfortunately this efficiency does not come without cost. 
-Structured overlays can only support exact-match queries 
+To achieve that, they implement a \emph{Distributed Hash Table (DHT)} that maps
+objects to nodes through a deterministic mechanism. \emph{DHT}s provide a
+guaranteed bound on the number of overlay routing hops that have to be
+``traveled'' to locate a resource, even in the case when only a \emph{single}
+copy exists whithin the system. This requires
+$O$($log(n)$) hops, compared to the unstructured networks that require
+$O(n)$ to reliably locate any object. Unfortunately this efficiency does not
+come without cost. Structured overlays can only support exact-match queries 
 (i.e., queries that identify resources by name) as opposed
@@
 resource location queries such as keyword and partial matching searches. 
-Several popular file-sharing
-\p\ applications have been based on \emph{DHT}s including
-{\sl Overnet}/{\sl eDonkey}~\cite{overnet}, 
+Several popular file-sharing \p\ applications have been based on \emph{DHT}s
+including {\sl Overnet}/{\sl eDonkey}~\cite{overnet}, 
 {\sl Kademlia}/{\sl Kad Network}~\cite{maymounkov_kademlia_2002},  and 
@@
 Moreover, \emph{DHT}s sparked
-a flurry of research for many diverse of applications ranging
+a flurry of research for many, diverse, applications ranging
 from distributed search engines to event-notification and cloud-based
@@
 The respective costs are $45 ms$ and $15 ms$.
-% The total cost of the first sequence is $45 ms$, while the second yields a
-% mere $15 ms$. 
 Evidently, the second overlay is more
@@
 Ideally, an overlay should be constructed to achieve an \emph{optimal} 
-mapping between the overlay paths and the underlying network links to avoid
+mapping with its underlying network links, to avoid
 inefficient states where redundant physical resources are utilized for the
-operation of the same link in the overlay. 
-The problem of constructing an optimal
-overlay is referred to as the \emph{topology mismatch problem}, and is formally
-defined as follows:
+operation of the same virtual link in the overlay. The ratio of the actual
+\emph{IP} network distance a message travels to reach an object of interest
+(via overlay routing) to the minimal \emph{IP} network distance to that object
+(i.e. through \emph{IP}) is known as \emph{stretch} or
+\emph{Relative Delay Penalty (RDP)}~\cite{CRZ2000}.
+
+The problem of constructing an optimal overlay is referred to as the
+\emph{topology mismatch problem}, and is formally defined as follows:
+
+%%VM: Are we 100\% sure about the definition?
 \begin{definition}
@@
 
-Early incarnations of the overlay protocols did not make use of the optimal
-mapping with the underlying network topology. {\sl Gnutella}, for example, is
+Early overlay protocol implementations did not pay to much attention to map
+optimally to the underlying network topology. {\sl Gnutella}, for example, is
 considered far from scalable~\cite{ritter_gnucantscale_2001} since every peer
-chooses its neighbors in the overlay without any knowledge of the underlying
+chooses its neighbors without any knowledge of the underlying
 network, resulting in mismatch. 
@@
 A topology mismatch can thus have a serious effect 
-on Internet router performance.
+on Internet's router performance.
 
@@
 A topology--unaware overlay network is able to control the sequence of peers a
-message traverses before reaching its destination, but it is completely unaware
-of how the actual packets are switched at the underlying infrastructure along
+message traverses before reaching its destination, but it completely ignores
+how the actual packets are switched at the underlying infrastructure along
 the overlay path. 
@@
 Moreover, a link in the physical network often lies
-on several overlay paths inadvertently increasing the traffic on the
+on several overlay paths inadvertently increasing the traffic on that
 link; this is known as the link's \emph{stress}~\cite{CRSZ2002}. 
@@
 of the  peer in the network.
-%%AD the paragraph below is ... LONG - rephrased as above...
-% Even in the case of an approach that provides mechanisms 
-% to a new-coming node to detect close-by peers at join time, 
-% the stochastic behavior with which peers join and leave
-% the overlay network, called the network's churn, 
-% further undermines the effort for maintaining 
-% the best position in the network throughout the life-time of a peer.
 
@@
 We employ both tabular and pictorial means to 
-articulate features of the proposed approaches and outline 
+articulate features of the proposed methodologies and outline
 both strengths and weaknesses.
@@
 \emph{Efficiency} means how well the
-algorithm performs its search and topology adaptation functions, without
-affecting other characteristics of the networks 
-(e.g. BitTorrent's optimal download performance). 
-%%AD I am not sure what the above line says.. or connects. Something is 
+algorithm performs its search and topology adaptation functions. We also take
+take into account how this might effect other characteristics of the networks
+under consideration\footnote{For example we don't actually want a topologically
+optimized BitTorrent swarm if this results in a degraded download performance,
+a characteristic that is the protocol's corner stone.}.
+%%VM Does the above sentence with the footnote now connect?
+%%
+%% PREVIOUS COMMENT AD I am not sure what the above line says.. or connects. Something is
 %%	missing...
-Efficiency is negatively effected every time an algorithm 
+
+%% SOME ADDITIONAL EXAMPLES FOR EFFICIENCY IF WE WANT TO ADD THEM
+%minimum-spanning-tree-based techniques have slow convergence speed in topology
+%adaptation or the high clustering impacts load balancing and constrains search
+%scope (or at least increases search satisfaction time), all reducing overall
+%efficiency.
+
+Moreover, efficiency is negatively effected every time an algorithm
 reduces its search scope.
-For the \emph{overhead} criterion, we track the cost 
+For the \emph{overhead} criterion, we evaluate the cost
 to compute and communicate peer join, message forwarding, 
-and topology maintenance. 
-For example, when exchanging route tables only amongst neighbors, 
-we consider the traffic generated to be low.
-%%AD Again I am not sure how the example above ties with the definition...
+and topology maintenance. For example, when a protocol relies on peers
+exchanging routing tables with other peers within a 2-hop radius with each other
+, then we can be safelly deduce that it incurs lower overhead when compared to a
+protocol that requires global knowledge of the overlay network.
+%%VM Does the above sentence with the footnote now connect?
+%%
+%%PREVIOUS COMMENT AD Again I am not sure how the example above ties with the definition...
 For the \emph{scalability} criterion, we consider system behavior 
@@
 and millions of nodes, respectively.
-
-% we dedicate a discussion subsection after the presentation of the algorithms
-% for both unstructured and structured architectures. This section offers:
-% \begin{inparaenum}[\itshape i\upshape)]
-  % \item further discussion that can help the reader identify, through
-% a taxonomy, the pluses and minuses  of the presented algorithms,
-  % \item a comprehensive tabular organization of the key-point information, and
-  % \item an attempt to give a pictorial comparison of the relative performance of
-% the algorithms.
-% \end{inparaenum}
-
-% We hope to offer the reader a deeper insight for the observations that
-% pushed research toward some direction. Thus, we try to categorize these
-% approaches so that we can organize directions making them comprehensible
-% and helping the reader build critical thinking on the advantages and
-% disadvantages of each approach.
-
-% Tables provide a one-stop-shop overview of the algorithms and serve as a 
-% tool both as a companion to the reader while reading the survey and a reference for
-% future use.
-% 
-% Pictorial comparison consists of bar charts that give a relative feeling of the
-% performance of the various algorithms, discussed in the corresponding section,
-% in
-% 
-% terms of \emph{efficiency}, \emph{overhead} and \emph{scalability}; three
-% important characteristics for every distributed algorithm. 
-% For each of these criteria, we use a coarse-grained, 
-% three-level, gradation to denote how well the algorithms do,
-%%  namely \emph{low}, \emph{medium} and \emph{high}. 
-% Each criterion has its own
-% interpretation of the above literal values. Efficiency means how well the
-% algorithm performs its search and topology adaptation functions, without
-% affecting other characteristics of the networks (e.g. BitTorrent's optimal
-% download performance). We assume that efficiency is
-% negatively effected when an algorithm reduces the search scope. 
-% %Moreover, the
-
-%minimum-spanning-tree-based techniques have slow convergence speed in topology
-
-%adaptation or the high clustering impacts load balancing and constrains search
-
-%scope (or at least increases search satisfaction time), all reducing overall
-
-%efficiency.

### 2014-01-30 intro.tex
2013.05.20 (Vassilis) Minor changes to abstract, introduction, background and unstructured (tex, plot and tables) sections. Useless comment clean-up. Various rephrasing on some reviewed content. Added references to the tables. Some image revisited. Discussion text added explaining the fact that many of the papers combine multiple methodologies.

@@
 file-sharing and Internet telephony.
-The {\sl Napster} file-sharing was widely acknowledged as the 
-``fastest growing Internet application ever'' in $2001$ when
-$26$ million users shared $80$ million songs.
-% File-sharing, originally introduced by
-% {\sl Napster} (Section~\ref{section:background}), was widely
-% acknowledged by $2001$ as the ``fastest growing Internet application ever" 
-% with $26$ million users sharing $80$ million songs.
-% % at the height of its popularity. 
+The {\sl Napster} file-sharing system was widely acknowledged as the 
+``fastest growing Internet application ever'' in $2001$ when it topped
+$26$ million users sharing over $80$ million songs.
 Other file-sharing applications followed suit, 
 including {\sl Gnutella} and {\sl Limewire} enjoying $3$ 
-million concurrently-connected peers as well as 
-{\sl BitTorrent} connecting over $150$ million users by the end of $2006$.
+million concurrently-connected peers, as well as 
+{\sl BitTorrent} connecting over $150$ million monthly users by January $2012$~\cite{bittorrentusers}.
 \p\ telephony saw explosive growth with the advent of {\sl Skype}.
@@
 {\sl Skype} has become extremely popular with more than $650$ million users 
-in $2011$~\cite{skypetotalusers}; on March $5$th, $2012$,  
-the astonishing figure of $35$ million concurrently-online users  
+in $2011$~\cite{skypetotalusers} while on January $21$st, $2013$,
+the astonishing figure of $50$ million concurrently-online users
 was reported~\cite{skypesymusers}.
 %%
-In addition, a startling number of diverse and successful
-applications have been built based on a \p\ architecture. 
-Some of them include:
+Apart from the above, though, a startling number of diverse and successful
+applications have been built based on \p\ architectures. Some of them include:
 distributed search engines~\cite{yaci}, 
@@
 
+% TODO SOME EXTRA REFERENCES IF NEEDED FOR THE USE OF OVERLAYS IN GENERAL
+%
+% IF ULTIMATELY DELETED CHECK TO REMOVE THE REFERENCES INCLUDED SINCE MOST
+% PROBABLY ARE NOT USED ELSEWHERE.
+%
+%find citation for Narada & Adobe's Real Time Media Flow Protocol (RTMFP)
+%\cite{} that provide quality-of-service (QoS) guarantees. \emph{IP
+%multicast}, as well as other proposals such as \emph{IntServ} and
+%\emph{DiffServ} \cite{cisco_diffserv_2005}, have still not seen wide deployment
+%largely due to the fact they require support from the IP
+%infrastructure. On the other hand, an overlay network can be deployed on
+%end-systems running the overlay protocol software, without the cooperation of ISPs.
+%Academic research includes
+%\cite{chu_esm_2000,jannotti_overcast_2000,kwon_tag_2002} among others.
+
 \p\ systems are implemented using \emph{overlay networks}. 
 An overlay network, is a virtual system of nodes and logical links
-created above an existing network; overlays provide
-added value to the underlying layer by, for instance, offering QoS
-content-distribution over the best effort Internet. 
-Overlay nodes communicate through \emph{virtual connections} each 
-of which may correspond to a path of possibly many physical links 
+created above an existing network; overlays provide an abstraction that
+enables the implementaion of efficient, fully distributed, application-layer
+services such as routing messages to destinations that are not known in advance
+or offering QoS guarantees (i.e., in content-distribution) over best effort
+infrastructures. Overlay nodes communicate through \emph{virtual connections}
+each of which may correspond to a path of possibly many physical links 
 in the underlying network.
@@
 underlying IP links between these two nodes is minimized.
-% If the traffic patterns of the application
-% are such that traffic flows heavily over the overlay link 
-% between the two nodes,
-% then it is beneficial to construct the overlay 
-% topology such that the number of
-% underlying IP links between these two nodes is minimized.
-%%
 Should the overlay network be constructed so that
@@
 the inherent \emph{topology mismatch} gives two
-significant problems: first, the performance of the application per se, can be
+significant problems. First, the performance of the application per se, can be
 adversely affected since traffic must flow over a larger,
@@
 Section~\ref{section:background} provides background on
-overlay architectures including centralized, decentralized
-unstructured and decentralized-structured \p\ systems. 
+overlay architectures including centralized, decentralized-unstructured and
+decentralized-structured \p\ systems.
 We also formally define the problem of topology mismatch 
-% between an overlay and its underlying
-% physical network 
 and offer the rationale behind our 

### 2014-01-30 mard-survey.tex
2013.05.20 (Vassilis) Minor changes to abstract, introduction, background and unstructured (tex, plot and tables) sections. Useless comment clean-up. Various rephrasing on some reviewed content. Added references to the tables. Some image revisited. Discussion text added explaining the fact that many of the papers combine multiple methodologies.

@@
 \emph{Peer-to-peer} (\p) systems have enjoyed immense attention and have been
-widely deployed on the Internet for well over a decade.  %% no commas here
+widely deployed on the Internet for well over a decade.
 They are implemented via an overlay network abstraction, on top 
-of the best-effort IP infrastructure of the Internet. 
-\p\ systems offer a plethora of desirable features to 
+of the Internet's best-effort IP infrastructure. 
+\p\ systems support a plethora of desirable features to
 distributed applications that are built on top of them including
@@
 Unfortunately, inherent weaknesses of early deployments of \p\ systems, prevented 
-applications from leveraging these features.
+applications from leveraging the full potential of the paradigm.
 One major weakness, identified early on, is the \emph{topology mismatch problem}
@@
 
-\keywords{Distributed systems, peer-to-peer architectures, overlay network,
+\keywords{Distributed systems, peer-to-peer architectures, p2p, overlay network,
 topology awareness}
@@
 
-% \item Proximity Neighbor Selection\\
-%Finally, the third approach, constructs the routing tables using proximity
-%knowledge. Tapestry and Pastry's mechanisms of routing table
-%maintenance try to minimize the distance to nodes appearing in a peer's routing
-%table. Since routing is based on longest node ID prefix match, messages 
-%are gradually forwarded to nearby nodes at each routing step.
-%\cite{castro_proximityp2p_2002} argues on how Pastry exploits proximity
-%neighbor selection in order to create a scheme that is (more) location-aware
-%compared to the other well-known DHTs (CAN, Chord).
-%
-%\end{itemize}
-%
-%{\sethlcolor{yellow}\hl{
-%HA: Possible criteria: based on which protocol (eg Gnutella, CAN etc), peer
-%selection (topology selection, cluster, cache etc), supports dynamic update,
-%runtimes}}
-%%\textbf{Algorithm} & \textbf{Overlay structure} & \textbf{Forwarding} &
-%\textbf{Cache} & \textbf{Overlay optimization} & \textbf{Proximity information}
-%& \textbf{Base protocol} & \textbf{Dynamic update} & \textbf{Runtime} \\
-%
-%
-%{\sethlcolor{yellow}\hl{
-%HA: Possible criteria2: Overlay optimization structure, base protocol (eg
-%Gnutella, CAN etc), dynamic update, runtimes, scalability}}
-%
-%{\sethlcolor{yellow}\hl{
-%HA: maybe add also the year of publication, and see if there is a pattern in
-%terms of the method and the year??}}
-
 \input{unstructured}
@@
 % TODO: SOME INFORMATION ON ISP-ASSISTED TOPOLOGY AWARENESS
+%
+% IN CASE REVIEWERS POINT OUT THAT WE SHOULD ADD THOSE AS WELL.
+%
 %       ATTENTION: NEEDS REWORDING

### 2014-01-30 structured.tex
2013.05.20 (Vassilis) Minor changes to abstract, introduction, background and unstructured (tex, plot and tables) sections. Useless comment clean-up. Various rephrasing on some reviewed content. Added references to the tables. Some image revisited. Discussion text added explaining the fact that many of the papers combine multiple methodologies.

@@
 
-Maintaining several host states at different layers, makes any content migration
-costly. Additionally, the method does not make any continuing effort to remap
-the overlay structure after a node successfully joins in order to adapt its
-state to any condition change. Although this approach greatly reduces the
+Maintainece of several host states at different layers, renders any content
+migration a costly process. Additionally, the method does not make any
+continuing effort to remap
+the overlay structure after nodes successfully join resulting into poor
+adaptation to changing conditions. Although this approach greatly reduces the
 routing latency to far nodes, it is unable to dynamically identify nodes that
@@
 \emph{Mithos} \cite{WR2003} is a P2P protocol which incorporates directed
-incremental probing to find near optimal node placement, and is classified by
-its authors as an integration of geographic layout and proximity routing overlay
-optimization methods.
-
-The bootstrap phase of Mithos starts with a subset of existing members as the
-first set of candidates, and while the iteratively closest nodes are detected by
+incremental probing to find near optimal node placement, and is an integration
+of geographic layout and proximity routing overlay optimization methods.
+
+The bootstrap phase of \emph{Mithos} starts with a subset of existing members as
+the first set of candidates, and while the iteratively closest nodes are detected by
 probing the neighborhood, the candidate neighbor list is updated. To
@@
 selection is an effective proximity based method.
+
+% \item Proximity Neighbor Selection\\
+%Finally, the third approach, constructs the routing tables using proximity
+%knowledge. Tapestry and Pastry's mechanisms of routing table
+%maintenance try to minimize the distance to nodes appearing in a peer's routing
+%table. Since routing is based on longest node ID prefix match, messages
+%are gradually forwarded to nearby nodes at each routing step.
+%\cite{castro_proximityp2p_2002} argues on how Pastry exploits proximity
+%neighbor selection in order to create a scheme that is (more) location-aware
+%compared to the other well-known DHTs (CAN, Chord).
+%
+%\end{itemize}
+%
+%{\sethlcolor{yellow}\hl{
+%HA: Possible criteria: based on which protocol (eg Gnutella, CAN etc), peer
+%selection (topology selection, cluster, cache etc), supports dynamic update,
+%runtimes}}
+%%\textbf{Algorithm} & \textbf{Overlay structure} & \textbf{Forwarding} &
+%\textbf{Cache} & \textbf{Overlay optimization} & \textbf{Proximity information}
+%& \textbf{Base protocol} & \textbf{Dynamic update} & \textbf{Runtime} \\
+%
+%
+%{\sethlcolor{yellow}\hl{
+%HA: Possible criteria2: Overlay optimization structure, base protocol (eg
+%Gnutella, CAN etc), dynamic update, runtimes, scalability}}
+%
+%{\sethlcolor{yellow}\hl{
+%HA: maybe add also the year of publication, and see if there is a pattern in
+%terms of the method and the year??}}

### 2014-01-30 unstructured.tex
2013.05.20 (Vassilis) Minor changes to abstract, introduction, background and unstructured (tex, plot and tables) sections. Useless comment clean-up. Various rephrasing on some reviewed content. Added references to the tables. Some image revisited. Discussion text added explaining the fact that many of the papers combine multiple methodologies.

@@
 In this section, we present algorithms that tackle the topology mismatch
-problem in unstructured \p\ networks and classify them based on their 
+problem in unstructured \p\ networks. We classify them based on their
 use of the overlay structure, their message forwarding scheme 
-for peer communication as well as techniques used for detecting proximity 
-in order to optimize their overlay topology.
-
-% TODO: READ A Near-Optimal Algorithm Attacking the Topology Mismatch Problem in
-%Unstructured Peer-to-Peer Networks (This is an Approximation alg for Top-mis
-%problem)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% TODO: HOW CAN UNSTRUCTURED SCHEMES BE REFINED
-%
-%For this reasons, efforts have been placed for optimizing the efficiency of
-%decentralized unstructured peer-to-peer networks. Research mainly focuses on
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item reducing unnecessary, redundant communication traffic, and
-%  \item exploiting physical locality to reduce communication response.
-%\end{inparaenum}
-%The goal can be achieved at, both, the application-level network as well as the
-%underlying physical one. In the first case by refining the message relay
-%techniques, while in the second one, by adaptively reconstructing the
-%application network to map as well as possible to the the physical network.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+for peer communication and the techniques they use for detecting
+proximity, to be used to optimize their overlay topology.
 
 \subsection{Algorithms for Unstructured Architectures}
+
+% TODO: READ A Near-Optimal Algorithm Attacking the Topology Mismatch Problem in Unstructured Peer-to-Peer Networks)
 
@@
 \emph{directed BFS}, and \emph{local indices}.
-%%AD local indices is not "a message forwarding method" really...  but ok.
 
@@
 Its major drawback is the delay between successive iterations, as the
-source-node has to examine the results at each iteration before deciding to
-either quit or resume the query.
+source-node has to examine the results at each attempt before deciding to
+either quit or ``unfreeze'' the query.
 
@@
 tries to avoid the aforementioned delay by forwarding 
-the query messages only to a select set of neighbors, in which the selection criteria varies from the number of
-results received previously, distance in terms of hops, bandwidth, or the query
-load of the neighbor.
+query messages only to a selected subset of available neighbors. The selection
+criteria varies, from the number of results received previously or the distance
+in terms of hops traveled to locate these results, to the bandwidth, or the
+query load of the neighbor. That way, fewer nodes are visited but quality of
+query responses is maintained, to a large degree, on the premise that the
+selected heuristics can direct the search to the right path.
 %
@@
 the presence of frequent node joins/departures 
-may introduces significant overhead should the radius is kept broad.
+may introduce significant overhead should the radius is kept broad.
+%%
 Below, we outline how the techniques in~\cite{YG-M2002}
@@
 % \paragraph*{ \bf Delay Aware P2P System}
-% A new \emph{Delay Aware P2P System (DAPS)} is introduced in \cite{ZL2005}. 
 A \emph{delay--aware} \p\ system termed {\sl DAPS} 
@@
 The source node that issues the query designates the delay 
-boundary it may tolerate providing so a \emph{pruning factor}.
+boundary it may tolerate, providing so a \emph{pruning factor}.
 In this context, user requests are forwarded only to 
 nodes whose expected delay is less than or equal to the indicated boundary. 
-{\sl DAPS} primarily focuses on user ``experience'' and deploys a
+{\sl DAPS} primarily focuses on user ``experience'' and deploys an
 end-to-end delay monitoring mechanism that  may enable
@@
 %%
-% TODO: MAYBE REVIEW THE EVALUATION
 In terms of the $3$ performance criteria, {\sl DAPS} stands as follows:
@@
 \hline
-%
+% Does not check for close by nodes, it just sorts the contents of a node's
+% routing table.
+% Also in a high churn environment a lot of time could be spend probing for
+% distances, resorting routing tables and creating sectors. Moreover it just
+% prunes distant neighbout
 low &
-%
+% Sorting routing table entries is done locally so it is not affecting.
 low &
@@
 a token-based flow control mechanism 
-% , which is essentially an intelligent flow control algorithm 
-%%AD what is "intelligent"?? - I think you can live without this above line... 
 that gradually redirects queries to nodes which are more
-likely to answer. 
-%%
-This flow control mechanism also helps prevent node overloading
-as each peer ``announces'' the number of query requests it 
-can handle in terms of tokens to its neighbors; to this end, 
+likely to answer. This flow control mechanism also helps prevent node
+overloading as each peer ``announces'' the number of query requests it 
+can handle, in terms of tokens, to its neighbors; to this end, 
 peers only forward query requests to nodes that they previously
-received tokens from.
-%%AD changed below as above... I have the impression that this token 
-%%	based mechanism is one and the same (the way the writing appeared 
-%%	was that they were two mechanisms!  check!!!
-%%%%
-% In order to prevent overloading of nodes with query requests,
-% Gia uses a token-based flow control algorithm in which each node announces the
-% number of query requests it can handle in terms of tokens to its neighbors, so
-% that neighbors only forward query requests to nodes that they previously
-% received tokens from. 
-%%%%%%%%%%%%%%%%%
+received tokens from. Further refinement to the search mechanism is the
+support for \emph{one--hop replication} of pointers to content.
 {\sl Gia} also acknowledges the heterogeneity in peer bandwidth,
@@
 Through message exchange, a peer can establish new connections 
-or drop superseded one in order to improve its satisfaction.
-%%AD the 2 lines below make no sense.. I rephrased as above... 
-% \emph{Gia} ensures that high capacity nodes have high degrees and
-% low capacity nodes are within short proximity to high capacity ones.
-%%%%%%%%AD this (above) phrase is almost verbatim out of the paper! NO good idea ;-)
+or drop superseded ones in order to improve its satisfaction.
 Despite the fact that {\sl Gia}'s topology adaptation algorithm 
@@
 terms.
-%%AD the comment TODO is somewhat correct. Changed the wording here..... 
-% TODO: THIS FOLLOWING REMARK SOUNDS LIKE THE ALGORITHM SHOULDN'T BE IN THIS
-%       SURVEY AFTER ALL!!
-% Although the topology adaptation algorithm Gia uses improves the scalability of
-% the network, it does not help much in solving the topology mismatch problem,
-% since it does not consider the underlying physical topology.
-%
-%More specifically, there are four key components in the design of
-%\emph{Gia} which are summarized bellow:
-%\begin{enumerate}
-%  \item A \emph{dynamic topology adaptation} protocol that puts
-%participating nodes within short reach of high capacity nodes so that these
-%\emph{high-degree} nodes, which due to their high connectivity receive most of
-%the queries, actually have the capacity to handle them.
-%  \item An \emph{active flow control} scheme is used to avoid overloaded
-% hot-spots. Heterogeneity is detected and flow control tokens are given to
-%nodes based on the available capacity.
-%  \item Every node maintains pointers to to the content that is offered by
-% their immediate neighbors, creating a \emph{one-hop replication} of pointers,
-%scheme.
-%  \item A \emph{search protocol} based on random walks, that is biased towards
-% directing queries to high-capacity nodes that are typically best able to
-%answer these queries.
-%\end{enumerate}
 %
 In terms of the $3$ stated criteria, {\sl Gia}'s approach is as follows:
-%%
-%%AD from the discussion I think that Efficiency might me medium-how??? 
-\begin{center}
-{\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
-% Drastically reduces the search scope.
-low &
+\begin{center}
+{\footnotesize
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% Reduces the search scope.
+medium &
 % Only one copy of the query is issued to the network.
@@
 
-% \begin{center}
-% \begin{tikzpicture}
-% \begin{axis}[
-%   ybar,
-%    symbolic x coords={
-%      Efficiency,
-%      Overhead,
-%      Scalability
-%    },
-%    symbolic y coords={lo, med, hi},
-%    x tick label style={rotate=45,anchor=east},
-%    xtick=data, ytick=data
-% ]
-
-% \addplot coordinates {
-%   (Efficiency,lo)
-%   (Overhead,lo)
-%   (Scalability,lo)
-% };
-
-% \end{axis}
-% \end{tikzpicture}
-% \end{center}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % \subsubsection{Distributed Cycle Minimization Protocol (DCMP)}
-% We have already discussed that blind flooding often causes messages 
-% to be routed in a direction
-% that is far from the path that can
-% ultimately fulfill the query request. 
-%%AD the above is just repetition...
-
 Another problem with topology unaware
-systems is the duplication of messages due to cycles that appears 
- even along the correct forwarding path. 
-\cite{ZKB2008}~focuses on this exact deficiency
-of overlay networks and introduces the \emph{Distributed Cycle Minimization Protocol
+systems is the duplication of messages due to cycles in the network graph.
+Interestingly, such cycles appear even along the correct forwarding path.
+\cite{ZKB2008}~focuses on this exact deficiency of overlay networks and
+introduces the \emph{Distributed Cycle Minimization Protocol
 (DCMP)}, a dynamic, fully distributed method that removes cycles;
@@
 Once a cycle is detected in {\it DCMP}, the most powerful node in that cycle 
-is elected as the \emph{Gate Peer} and the cycle is then broken 
-%% at a strategic place  %%AD I am not sure what this adds?? strategic?? on what??
-so that it results in the minimization of the distance between the 
+is elected as the \emph{Gate Peer} and the cycle is then broken in that place
+that it will result in the minimization of the distance between the
 \emph{Gate Peer} and all other nodes that are currently part of the cycle. 
 This process is managed by using two specialized message types 
-namely, \emph{Information Collection (IC)} and \emph{Cut Message (CM)}. 
-\emph{DCMP} bases its operation on messages whose travel is limited by 
-imposed $TTL$ whose maximum value is set at  $7$ for most cases. 
-This inherently limits the protocol as \emph{DCMP} is unlikely that 
-can efficiently detect cycles that span for more than $7$ nodes.
-%%AD rephrased as above
-% One disadvantage of DCMP is that since the distance a forwarded
-% message can travel is limited by the TTL value, which is practically $7$ in most
-% cases, DCMP cannot detect cycles formed by more than 7 peers. 
+namely, \emph{Information Collection Message (ICM)} and \emph{Cut Message (CM)}. 
+\emph{DCMP} bases its operation on messages whose travel is limited by an
+imposed $TTL$ value (max set to $7$ for most cases). This inherently limits
+the protocol in detecting cycles that span for more than $TTL$ nodes.
 Even though cycle elimination does improve network performance, 
-it cannot 
-% actually solve 
-directly contribute in the solution of 
-the topology mismatch problem.
-%%
-%
-%\subsection{Distributed Cycle Minimization Protocol}
-%
-%\paragraph{}
-%The first step after detecting a duplicate message by some peer is to gather
-% information from all peers in the cycle using a new type of control message
-%called \emph{Information Collecting Message} or \emph{ICM}. ICM contains:
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item a \emph{GUID}\footnote{Globally Unique Identifier assigned to every
-% query message generated by any node.} field same as the one of the duplicate
-%message,
-%  \item \emph{DetectionID} field which represents the direction of the
-% connection where the duplicate was identified\footnote{This ensures the
-%uniqueness of the ICM messages because as it travels through many cyclic paths,
-%multiple peers will detect the duplicates and initiate an ICM message.}, and
-%  \item \emph{Node Information Vector (NIV)} which contains information
-% (bandwidth, CPU power, etc) about peers that propagated the ICM.
-%\end{inparaenum}
-%
-%\paragraph{}
-%Suppose $A$ detected the duplicate as depicted in Figure~\ref{figure:dcmp}. It
-% then emits an ICM to $B$ and $F$, that initially contains information only
-%about itself. Each peer that receives the ICM, appends its information and
-%propagates it along the reverse path of the original message. Since two copies
-%of ICM are sent, at some point, a peer, say peer $D$, will receive a duplicate
-%ICM. Using the information in the NIVs of the ICMs, $D$, decides to cut (for
-%example) the EF connection. To inform the other peers about its decision, it
-%emits a \emph{Cut Message (CM)} which contains the GUID and DetectionID of the
-%corresponding ICM and an additional field that identifies the connection to be
-%cut. $D$, then, forwards the CM in the reverse directions from where the ICM
-%arrived. Similarly CMs received by any peer are propagated toward the reverse
-%path of the corresponding ICM. Eventually, either peer $E$ or peer $F$ will
-%receive the CM and cut the connection, thus eliminating the cycle.
-%
-%Receiving a duplicate ICM denotes the existence of a cycle. The opposite is not
-% true though. For example, if the cycle contains $2 \times TTL$ edges, it will
-%not be detected, because the ICM messages will be discarded before they locate
-%it. There is a trade-off between preserving the connectivity of the network and
-%minimizing the duplicates that makes such a possibility to be safely ignored.
+it cannot  directly contribute in the solution of the topology mismatch problem.
 %
@@
 %
-%\paragraph{}
-%
 % TODO: SOME DISCUSSION
-%
 %Actually, the duplicate packet problem seems to hurt more the active nodes;
@@
 \hline
-% It return more result than LTM according to experiments in
+% It returns more result than LTM according to experiments in
 % \cite{ZKB2008}
@@
 
-In~\cite{CS2002}, replication is used as a way 
-to improve inefficient blind search.
-As the number of replicas and/or cached copies increases in the network,
-it would relatively easy to locate items even if the search is random.
-% \cite{CS2002} aim to improve the inefficient blind search algorithm by
-% replicating the data in a P2P network. 
-% The main intuition behind the
-% idea of replication, or using cached copies, is that as the number of copies for
-% each item increases in the network, it would be easier for a search algorithm,
-% even a random one, to locate these items. 
-To analyze the feasibility of
-such an approach, \cite{CS2002} investigates $3$ replication
-strategies: uniform, proportional, and square-root allocation. In the
-uniform model, the copies of items are uniformly spread over the network,
-while in the proportional, items are replicated based on their query
-rate, so that frequently queried items can be found more often. 
-The square root is a compromise between uniform 
-and proportional allocation models.
-% allocation is a strategy proposed by Cohen and Shenker, which is a model between
-% the uniform and proportional allocation.  
-%%AD why u need to say the name here????
-The overall costs for both successful and unsuccessful
+In~\cite{CS2002}, replication is used as a way to improve inefficient blind
+search. As the number of replicas and/or cached copies increases in the network,
+it would relatively easy to locate items even if search is random.
+To analyze the feasibility of such an approach, \cite{CS2002} investigates
+$3$ replication strategies: \emph{uniform}, \emph{proportional}, and
+\emph{square root allocation}. In the uniform model, the copies of items are
+uniformly spread over the network, while in the proportional, items are
+replicated based on their query rate, so that frequently queried items can be
+found more often.
+The square root is a compromise between uniform and proportional allocation
+models. The overall costs for both successful and unsuccessful
 searches are compared to ascertain the effectiveness of the $3$ suggested 
-approaches.
-% To evaluate the outcome of the replication models, 
-% the overall costs of successful and unsuccessful
-% searches in the network are compared. 
-Experimental outcomes indicate that the uniform allocation 
-minimizes the maximum search length and so it can 
-reduce the time spent on unsuccessful searches.
-% Results argue that the uniform allocation %%AD results argue???? --
-					    %% ALSO, "search size??? See above
-% model minimizes the maximum search size, therefore reduces the time spent on an
-% unsuccessful search. 
+approaches. Experimental outcomes indicate that the uniform allocation 
+minimizes the maximum search length and so it can reduce the time spent on
+unsuccessful searches. 
 The proportional strategy effectively decreases the search time 
 for popular items, but suffers when needing to locate the rare entities.
-% The proportional model, on the other hand, promotes the
-% more frequently queried items by replicating them more and therefore decreases the
-% search time for popular items, but suffers when needing to locate the rare
-% items. 
 When it comes to expected successful search size, it is the same for 
 for both uniform and proportional models and any approach in between 
-to always behave much better.
-% The authors also claim that the expected successful search size is the
-% same for uniform and proportional models, and any approach between them would
-% always behave much better. 
-The square root allocation  aims at minimizing the 
+to always behave much better. The square root allocation  aims at minimizing the 
 expected search size for successful queries in \p\ networks.
-% Finally, they propose the square root allocation
-% approach, which is a replication strategy that minimizes the expected search
-% size of successful queries in P2P networks.
-%%
 The table below outlines how the $3$ strategies fare:
@@
 Despite the fact the {\sl Narada} was not designed as \p\ system per se,
-it was the first to consider the feasibility of overlay-based and application
-layer services over the Internet while taking into account 
-bandwidth and latency properties of the underlying physical infrastructure.
-%%AD you cannot have such a long footnote for God sake!
-% \footnote{
-  % IP multicast is the term referring to the method of sending IP datagrams to a
-  % group of receivers in a single transmission. Even though the method is
-  % available some years now, IP multicast has not taken off as anticipated since
-  % many believe that it violates the stateless design of the current Internet.
-  % For this reason, it is widely deployed in contained environments like
-  % enterprises (i.e. IPTV applications - distance learning, video conferencing),
-  % commercial stock exchanges, and multimedia content delivery networks.
-% }. 
-% Although it was not originally designed as a P2P system, Narada is
-% a pioneer in that 
-% %(and the accompanying \emph{Scattercast} \cite{C2000})
-% it was the first to consider the feasibility of overlay-based, application
-% layer, services over the Internet that can take into account bandwidth and
-% latency properties of the physical underlying infrastructure. 
-In the context of this work, the authors came across the 
-inefficiency caused by the topology mismatch and attempted to address it 
-by building an enhanced connected graph termed \emph{mesh}  %%AD why this graph
-							    %% is rich/enhanced???
-with each source featuring its own minimum spanning tree.
-A gossip-protocol was deployed for the creation of these 
-spanning trees~\cite{LYL2008}.
-%%
-%%AD in general avoid footnotes...  The sentence below does not make sense!
-%% "richer" to what???  ENarada is what???
-%% 	--- I rephrased as above - not sure correct.. check -----
-%% 
-% Additionally,
-% authors realized the inefficiency caused by the topology mismatch problem and
-% tried to address it by building a richer connected graph, called a
-% \emph{mesh}, and building per source minimum spanning trees\footnote{
-  % ENarada \cite{LYL2008} used Gossip protocol for the
-% construction
-% }.
+it was the first (along with \emph{Scattercast} \cite{C2000}) to consider the
+feasibility of overlay-based, application-layer services over the Internet while
+taking into account bandwidth and latency properties of the underlying physical
+infrastructure. In the context of this work, the inefficiency caused by the
+topology mismatch and attempted to be tackled by building an highly connected
+graph termed \emph{mesh}
+with each source featuring its own minimum spanning tree. A gossip-protocol was
+deployed for the creation of these spanning trees~\cite{LYL2008}.
 %as shown in Figure~\ref{figure:narada}.
 Both graph and trees are dynamically updated as nodes keep joining 
-or departing the network.
-% They also keep the graph and the trees
-% dynamically updated while nodes continue to join and leave the network.
-The protocol aims to ease the physical link stress, the overall
-resource usage as well as the relative delay among end-systems. 
+or departing the network. The protocol aims to ease the physical link stress,
+the overall resource usage as well as the relative delay among end-systems. 
 Unfortunately, the main limitation for {\sl Narada} is that 
@@
 %
-low &
+medium &
 % The overhead of Narada is proportional to the multicast group size
 high &
-%
+% Maintaining the mesh for large multicast groups is expensive.
 low
@@
 Along with {\sl Narada}, the \emph{Adaptive Overlay Topology 
-Optimization (AOTO)}~\cite{LZXN2003} is one
-of the first attempts to address the topology mismatch problem. 
-\emph{AOTO} is a distributed algorithm that seeks to optimize the usage of
-the underlying physical resources and operates in $2$ phases:
+Optimization (AOTO)}~\cite{LZXN2003} is one of the first attempts to address the
+topology mismatch problem. \emph{AOTO} is a distributed algorithm that seeks to
+optimize the usage of the underlying physical resources and operates in
+$2$ phases:
 \emph{Selective Flooding} and \emph{Active Topology}.
-In \emph{Selective Flooding}, a minimum
-spanning tree is built 
-% among  %%AD among???  put "for"...
-for each peer and its immediate neighbors so that
-queries do not flood the entire network 
-% and at the same time 
-without simultaneously shrinking their search scope. 
-During \emph{Active Topology}, the overlay
-links are revised so that they can more closely reflect 
-the physical network topology. 
-%%
-%%AD what is a non-flooding neighbor?
-Each peer replaces independently non-flooding neighbors, with closer nodes.
-% This is done independently, by each peer by replacing non-flooding
-% neighbors, with closer nodes. 
-Picking a replacement out of these non-flooding
+In \emph{Selective Flooding}, a minimum spanning tree is built for each peer and
+its immediate neighbors so that queries do not flood the entire network and at
+the same time preventing their search scope from shrinking. This way some
+neighbors become non--flooding. During \emph{Active Topology}, each peer
+replaces independently such non--flooding neighbors, with closer nodes as an attempt
+to revise the overlay links so that they can reflect more closely the physical
+network topology. Picking a replacement out of these non-flooding
 neighbors follows a random policy called \emph{Randomized AT} algorithm.
@@
 its communication costs with all its neighbors (e.g., network delays)
-as well as the costs between any pair of neighbors.
-%%AD I am not sure what the parenthesis below says.. Rephrased as above
-% as well as between any
-% pair of neighbors (meaning also that additional, cost probing, message types
-% need to be added to the Gnutella protocol). 
-The \emph{randomized AT} algorithm is applied by a source peer 
+as well as the costs between any pair of neighbors. The \emph{randomized AT}
+algorithm is applied by a source peer
 every time its neighbor list is updated or an updated neighbor cost 
 table is received.
-%%AD do you also obtain neighbor cost tables???
 %%
@@
 approach~\cite{LZXN2004} uses the network delay as a metric 
-to estimate the costs between nodes in a spanning tree.
-%%AD when u say spanning tree u mean the ENTIRE network???
-Each peer computes the costs to its logical one-hop
-neighbors and forms a \emph{neighbor cost table (NCT)} 
+to estimate the costs between nodes. Each peer computes the costs to its
+logical one-hop neighbors and forms a \emph{neighbor cost table (NCT)} 
 using a special routing message type. 
 Any pair of neighboring peers exchange their \emph{NCT}s 
-and so a minimal overlay topology can be formed.
-%%AD the wording of the following is a bit messed up! Simplify! 
-%%	reworked as above... check
-% Two neighboring peers exchange their \emph{NCT}s in order for
-% every peer to obtain the cost between any pair 
-% of its local neighbors forming a
-% small overlay topology. 
-Based on the obtained \emph{NCT}s a minimum spanning tree
-for each peer and its one-hop neighbors is created.
-%%AD rephrased as above.. This expression "... among each peer" makes 
-%%	absolutely NO sense... 
-% Moreover, based on obtained NCTs a minimum spanning tree
-% among each peer and its one-hop neighbors is built. 
-Finally, neighbors located physically far away 
+and so a minimal overlay topology can be formed. Based on the obtained
+\emph{NCT}s a minimum spanning tree for each peer and its one-hop neighbors is
+created. Finally, neighbors located physically far away 
 are replaced by physically-closer counterparts. 
-In particular, a peer $S$ iteratively probes the distance between 
+In particular, a peer $S$ iteratively probes the distance $d$ between
 itself and every of its non-flooding neighbor nodes $G$ as well as
 the distance between $S$ and $G$'s neighbors designated as $H$.
-%%AD I do not understand what on earth the following sentence SAYS!!!!
-%%	when you say "between" you need to have 2 parties - where is this???
-%%	I rephrased as above.. Check correctness...
-% Specifically, a peer
-% $S$ probes the distance between one of its non-flooding neighbor's neighbors,
-% say $G$ and $H$ respectively. 
-%%
-If the distance between a peer and a neighbor \emph{SG} is measured to 
-be larger than that with a neighbor's neighbor \emph{SH}, the link 
-between $S$ and $G$ is replaced with a new connection between $S$ and $H$.
-%%AD the following is complicated to diff to parse... Tried as above... 
-% Assuming that the neighbor's neighbor ($SH$) is
-% measured as being ``closer'' than the neighbor ($SG$), the link to the later
-% is cut and a new connection is established. 
-%%%
-%%	%%AD the opposite situation does not appear to be consistent!!!
-%	%% with what you say above... 
-% In the opposite situation where
-% distance $SG$ is closer than $SH$ then if $SH$ is shorter than $GH$, then $H$ is
-% preserved as $S$'s new neighbor. 
-%%%
-%%AD is the distance $GH$ in discussion here as well???  I am confused.
-In the case where $SH$ is larger than $SG$ and
-$GH$, no connection will be established and $S$ will continue probing another
-neighbor's neighbor. 
+If $d_{SG} > d_{SH}$, then link $SG$ is replaced by link $SH$. If, on the other
+hand, $d_{SG} < d_{SH}$ then also $d_{GH}$ and then we have the following
+options. Either $d_{SH} < d_{GH}$ in which case link $SH$ is kept or
+$d_{SH} > d_{GH}$ in which case $S$ will keep probing other $G$'s direct
+neighbors.
 %%%
 The above optimization is conducted within $1$-neighbor
-closure using as base a peer and checking all its direct neighbors.
+closure using as base, a peer and checking all its direct neighbors.
 Evidently the scope of such an operation could be extended.
@@
 at a greater computational overhead.
-%%
-%%AD This is NOT English.. No verb??? I am not sure...
-% The larger the scope, the better topology matching improvement but
-% also the greater the computational overhead.
 %
@@
 in this regard, peers 
-discover $1$- and $2$-hop neighbor sets ($N$ and $N^2$ respectively) %%AD not 
-						%% sure what the N and N2 are..
-and proceed to compute communication costs.
-% around them and calculating communication costs. 
+discover $1$- and $2$-hop neighbor sets, designated as $N$ and $N^2$
+respectively, and proceed to compute communication costs.
 Time-stamps are used to derive network latency measurements that are then used 
 to improve the overlay network without sacrificing the search scope.
-% Calculation is achieved through
-% timestamp checking, so clocks must be synchronized. 
-% The latency information
-% gathered is later used to evolve the overlay network into a more efficient one,
-% without reducing the search scope. 
 Each node compares the latency information
@@
 the \p\ network, it is unable to offer any warranty for 
-effectively addressing the mismatch problem. %%AD shortened and consolidated
-%% 
-% use real physical topology information,
-% just latency metrics, therefore it does not offer a guaranteed safety net for
-% the topology mismatch problem.
+effectively addressing the mismatch problem.
 %
@@
 color (red or white).
-%  (separating all peers into two groups). 
 Then the network bootstrap host node furnishes the joining peer with a list
 of active peers along with their color information. 
-%%AD how many nodes are provided here??? 
 The joining peer uses this list to 
@@
 Having information on all their $2$-hop neighbors 
-($N^2$) %%AD I am not sure what this N^2 means here!
+($N^2$)
 red-peers create minimum spanning trees for the neighbors in question and
-% assigns %%AD this should be plural
-assign efficient forwarding paths. 
-Having a minimal $2$-hop spanning tree, 
-a red-peer dispatches its queries only within this range. 
-Some white-peers, though, may have become non-forwarding neighbors. 
-%%AD I am not sure how you reach the above sentence... 
-%%
-%%AD The following sentence is very convoluted. I cannot parse it!
-In this phase, such a (white) neighbor will try to find another 
-red peer being two hops away from its
-current red neighbor to replace the later as its new neighbor. 
-%%%%
-The white-peers can further optimize their positions within 
-the overlay if this is required. 
-%%AD What are the implication of the above statement?? 
-%%   What is the end result of the approach in terms of the mismatch problem??? 
-%%   Something is missing... 
+assign efficient $2$-hop forwarding paths. This process can render a white
+peer $E$ non--forwarding neighbor of the red peer $P$. Direct neighbor
+replacement is a process conducted by $E$ in order to replace $P$ with a
+$P' \in N^2(P)$ as its new neighbor. This adaptation tackles the
+topology mismatch problem by reducing message duplication and shorten response
+times caused by the problem identified as \emph{Revisit Not known (RN)}. RN is
+the situation when a node is visited some times as a relay stop before it is
+visited as an overlay peer, thus creating duplicate unnecessary messages to the
+network and delaying query responses.
 %
@@
 The distributed heuristic termed 
-\emph{Two-Hop-Away Neighbor Comparison and Selection (THANCS)}~\cite{LNXE2005,L2008} attempts to minimize overlay hop costs.
+\emph{Two-Hop-Away Neighbor Comparison and Selection (THANCS)}~\cite{LNXE2005,L2008}
+attempts to minimize overlay hop costs.
 \emph{THANCS} is essentially a \emph{local search method} as it aims 
 to find a locally optimum solution by exploiting knowledge within 
-a $2$-hop radius. 
-The algorithm consists
-of two main components: \emph{piggybacking neighbor distance on queries} and
+a $2$-hop radius. The algorithm consists of two main components:
+\emph{piggybacking neighbor distance on queries} and
 \emph{neighbor comparison and selection}.
@@
 to probe immediate neighbors using delay distance measurements 
-and store information locally. 
-This is done by introducing a special query message type, the
-\emph{Piggy Message (PM)} which includes information about 
-the neighbor identification and measured distance. 
-A peer $p$ constructs a \emph{PM} for its
-neighbor $n$, which contains $n$'s IP address and the distance between them.
-%%
-If $p$ receives a query from $n$, the respective \emph{PM}-message will 
-be piggybacked to the query and $p$ will forward this enlarged message to
-of its neighbors.
-% When $p$ receives a query from $n$, this PM will be piggybacked to the query
-% message that will be then forwarded to all other neighbors of peer $p$. 
-In turn, each neighbor $q$ will detach the \emph{PM}-information,
-% (it will not be further forwarded) 
-record the distance $pq$ and further process the query. %%AD added "further" - correct??
-%%%
-%%AD rephrased as above... When you use "this" "that" etc BE careful...
-% Each neighbor will detach the PM information 
-% (it will not be further forwarded),
-% record the $pq$ distance and process the query 
-% as always after receiving a query message.  %% THIS "as... " threw me off!!!!
-%%%%%%%%%%%%%%%%%%%%
-Selection of which incoming queries should piggyback a \emph{PM} is proposed
-by either \emph{pure probability-based (PPB)} or \emph{new neighbor triggered
-(NNT)} policies.
-
-The \emph{neighbor comparison and selection} component
-helps peer $p$ probe all his neighbors placed upto 
-$2$-hops away ($ N^2(p)$) %%AD what is this?? complexity???
-not probed so far.
-The distance of $pn$ is known to $p$. %%AD I am not sure what you say!
-Upon receiving a \emph{PM} from node $n$ with the distance of $nq$, 
-where $q$ is a direct
-neighbor of $n$, $p$ chooses one of the following two paths. 
-\begin{itemize} %%AD put the two directives here...
-\item[--]
-	if $q$ is a $1$-hope neighbor of $p$, then 
-	$p$ 
-\item[--]
+and store information locally. The algorithm introduces a special query message
+type, the \emph{Piggy Message (PM)} which includes information about 
+the neighbor identification (IP) and measured distance ($d$). PMs are
+piggybacked to normal
+Gnutella messages. When node $P$ receives a query from $Q$ it constructs a
+$PM_{PQ}$, piggybacks it to the query and forwards it to $P$'s neighbors. When
+the neighbors recieve it they detach the $PM_{PQ}$, record the distance
+$d_{PQ}$ and process the query normaly. Selection of which incoming queries
+should be piggybacked with a \emph{PM} is proposed by either
+\emph{pure probability-based (PPB)} or \emph{new neighbor triggered (NNT)}
+policies.
+
+The \emph{neighbor comparison and selection} component defines that a peer $S$
+probe all his $2$-hops away neighbors (a set denoted as $ N^2(S)$) not probed
+thus far. Let $P$ be a $1$-hop distance neighbor of $S$, and $Q$ be a probed
+peer by $S$. When $S$ receives a message piggybacked with a $PM_{PQ}$ we
+identify the following cases.
+\begin{itemize}
+  \item If $Q$ is already a direct neighbor of $S$ then we check the distances
+  $d_{SQ}$, $d_{SP}$ and $d_{PQ}$. If either $d_{SQ}$ or $d_{SP}$ is the longest, then the longest
+  link will be put in a \emph{will-cut} list\footnote{A peer will not send or
+  forward queries to connections in its will-cut list but will preserve them
+  for some time in order not to effect ongoing response messages travelling the
+  inverse path.}. If $d_{PQ}$ is the longest then nothing is done by $S$; being
+  fully distributed, neighbor comparison and selection process, will have the
+  oportunity to handle $PQ$ link when initiated by peers $P$ or $Q$.
+
+  \item If $Q$ is strictly a $2$-hop neighbor of $S$ and have never probed $Q$
+  in the past, stores distance $d_{SQ}$ and checks distances $d_{SQ}$, $d_{SP}$ and $d_{PQ}$.
+  If $d_{SQ}$ is the longest, $S$ will not create link $SQ$. If $d_{SP}$ is the
+  longest, $SQ$ will be created and $SP$ will be put into the will-cut list.
+  If $PQ$ is the longest, links $SP$ and $SQ$ will be preserved expecting that
+  $P$ or $Q$ will disconnect $PQ$ later.
 \end{itemize}
-%%AD I cannot follow what on earth it is said here... this needs 
-%%	rewriting and better structuring..  Use the above list for 
-%%	the expected operation...
-First, if $q$ is a
-direct neighbor of $p$, then the latter will check the cost of the involved
-links. If the most cost-intensive link is one of $pq$ or $pn$, then it is put
-into a \emph{will-cut} list. If it is $nq$ then nothing is done (since either
-$n$ or $q$ will have the chance to handle it themselves).  
-%%AD this is what I think should go into the second clause above... 
-On the other hand, if
-$q$ is within two-hop radius from $p$, the later probes the former (if it hasn't
-got the distance information yet) and again compares distances $pq$, $pn$, and
-$nq$. If $pq$ is the most cost-intensive, $p$ will not establish the connection.
-If it is $pn$, $p$ will establish connection $pq$ and put $pn$ in the will-cut
-list. 
-If it is $nq$, $p$ will stay connected with both $n$ and $q$,
-such that eventually in later stages either $n$ or $q$ will disconnect link $nq$.  
-%%AD The above paragraphs need clearing up.. The text cannot be read.. 
-%%
-%%AD Also, make the connection with the mismatch problem... and justify 
-%%	somewhat convincingly why the rating below comes true...
+
 %
@@
 %systems.
+
+As seen above, the method is fully distributed and needs only minimum knowledge
+of, at most, a $2$-hop distant peers. This makes it a good candidate for big
+overlay networks. Also topology adaptation helps construct a well fit overlay
+network with restect to the underlying one (at least with respect to network
+distances). Morever will-cut list and distance cache (which stores already
+probed peers) minimize the unnecessary messages for the network maintenance.
+These characteristics are depicted to the $3$ evaluation criteria we use in this
+survey as follows:
+
 \begin{center}
@@
 The \emph{Hops Adaptive Neighbor Discovery (HAND)} algorithm~\cite{CLZHC2006}
-uses a fully-distributed triple--hop adjustment strategy
-to address the topology mismatch problem.
-\emph{HAND}'s goal is to shape the current overlay
-graph $G$ into an optimal \emph{Logical Communication Network} 
-%(LCN)} %%AD why u need this acronym? u never use it!
-$G^{*}$.
-% \subsubsection{Hops Adaptive Neighbor Discovery (HAND)}
-% \cite{CLZHC2006} proposes the \emph{HAND} algorithm which uses a
-% fully distributed triple hop adjustment strategy to address the topology
-% mismatch problem. The approach's ultimate goal is to shape the current overlay
-% graph $G$ into the \emph{Logical Communication Network (LCN)} $G^{*}$ which is
-% how the optimal overlay is referred in the paper's context. 
-This optimality is attained if all peer hop 
-sequences $(v_1, v_2, \ldots, v_k)$ of $G$ exist in $G^{*}$ 
-and are in the same order.
-The latter indicates that in 
-practice triple sequences $(v_1, v_2, v_3)$ are used.
-%%AD this used to be a footnote - avoid footnotes in general!
-%%
-A mismatch is detected as follows:
-suppose we want to verify peer a
-sequence, say $v_2-v_1-v_3$. 
-Two probing messages are dispatched from $v_1$ to
+uses a fully-distributed triple--hop adjustment strategy, applied to a network
+graph $G$ in order to create an optimal graph $G^{*}$, free of the implications
+of topology mismatch. This optimality is attained if all peer hop 
+sequences $(v_1, v_2, \ldots, v_k)$ of $G$ exist in $G^{*}$ and are in the same
+order. The latter indicates that in practice triple sequences $(v_1, v_2, v_3)$
+are used. A mismatch is detected as follows: supposing that we want to verify
+a sequence, say $v_2-v_1-v_3$, two probing messages are dispatched from $v_1$ to
 $v_2$ and $v_3$ and yield delays of $x$ and $z$ respectively.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% TODO: DOUBLE CHECK IF THIS ALGORITHM GOES IN THIS SECTION/SUBSECTION
 % \subsubsection{Adaptive Peer Selection (APS)}
 The \emph{Adaptive Peer Selection (APS)} approach uses 
-machine learning techniques to form peer
-selection strategies based on past experience~\cite{BFLZ2003}. 
-A decision tree is used to rate peers based on information 
-collected for the characteristics of established connections.
-Such features include load, bandwidth, and past uploading experience. 
-%%AD load on what?? The dispatching end? the receiving end??
-Subsequently, a Markov decision process is introduced 
-as a mechanism to shape the policy
-for switching among the peers. 
-The success of approach depends on how fast the introduced 
-peer selection functions. Admittedly, \emph{APS} is slow 
-due to the learning process and its inherent complexity.
-As far as the $3$ criteria is concerned, \emph{APS} has as follows:
+machine learning techniques to form peer selection strategies based on past
+experience~\cite{BFLZ2003}. A decision tree is used to rate peers based on
+information collected for the characteristics of established connections. Such
+features include connection link load, bandwidth, and past uploading experience.
+Subsequently, a Markov decision process is introduced as a mechanism to shape
+the policy for switching among the peers. The success of approach depends on
+how fast the introduced peer selection functions. Admittedly, \emph{APS} is slow 
+due to the learning process and its inherent complexity. As far as the $3$
+criteria is concerned, \emph{APS} has as follows:
 \begin{center}
@@
 high &
-%
+% Complexity of computation, and slow convergence speed prevents the approach
+% to scale, especially in high churn environments.
 low
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The \emph{Innocuous Topology Aware Overlay Construction (ITA)} approach~\cite{PRFM2009} seeks to offer both an overlay formation approach and furnish 
-an effective way to carry out searching.
-% \subsubsection{Innocuous Topology Aware (ITA) Overlay Construction}
-% \cite{PRFM2009} introduces an algorithm for Innocuous Topology Aware
-% construction of unstructured P2P networks. The paper's proposition is twofold.
-% Overlay construction and search method. 
-When constructing the overlay, \emph{ITA} exploits the notion of 
-of \emph{short} and \emph{long} connections.
-% For the overlay construction, it employs
-% the notion of \emph{short} and \emph{long} connections. 
-Should $N$ be the number of network peers, 
-$\alpha \leq 1 $ a system-wide magic number
-and $x$ an $\alpha$-related latency threshold,
-the bootstrapping peer randomly selects $\alpha \ast N$ 
-``close'' (latency below $x$) and 
+The \emph{Innocuous Topology Aware Overlay Construction (ITA)}
+approach~\cite{PRFM2009} seeks to offer both an overlay formation approach and
+an effective way to carry out searching. When constructing the overlay,
+\emph{ITA} exploits the notion of \emph{short} and \emph{long} connections.
+Should $N$ be the number of network peers,  $\alpha \leq 1 $ a system-wide magic
+number and $x$ an $\alpha$-related latency threshold, the bootstrapping peer
+randomly selects $\alpha \ast N$ ``close'' (latency below $x$) and 
 $\left( 1 - \alpha \right) \ast N$ distant (latency above $x$) nodes
@@
 %%AD how do they select all these magic constants???
-%%
+%%VM This α constant seems to be the percentage of close neighbours selected by
+% a peer out of its total degree. Can't find the criterion with which it is
+% chosen though. Maybe we can ask... Mema :)
 Searching occurs in two phases: in the initial phase, the querying node
@@
 Subsequently, peers receiving a query over a ``long link'' commence a 
-local flood with $TTL\;=\;ttl$ where \emph{ttl} is system-defined.
+local flood with $TTL=ttl$ where \emph{ttl} is system-defined.
 %%AD how are you supposed to define these params?? in a fixed manner??
-
-The main objective of the overlay construction phase
-is to yield a network that exhibits low clustering. 
-In turn, this is 
-a beneficial characteristic for random graphs as it can lead to a
-larger coverage of peers with the same number of messages and 
-at the same time helps reduce duplication. 
-%%AD same number of messages in comparison to what?? 
-%%	same: duplication of what?? messages right???
-This combination offers efficient lookups that pose minimal 
-or no negative impact to other mechanisms possibly employed by 
-the \p\ systems.
-% This means more efficient information lookup that has minimal or no
-% negative impact whatsoever to the rest of the mechanisms employed in
-% unstructured P2P systems\footnote{For example the 1-hop replication and the
-% dynamic querying mechanisms of later versions of Gnutella.}. 
+%%
+%% VM Don't understand... the paper says that for increased awareness we need a
+%% small α (page 7). But in the begining they give that short links (close-by)
+%% are given by α*Ν. I guess this gets smaller for smaller α.... Also on
+%% page 7 they say they choose α to be 0.1, 0.05 and 0.033 and that
+%% this equals to ``10, 20, 30 long and and the same number of short links''.
+%% I don't understand how they deduce these numbers. Since short links is given
+%% by α*Ν and long by (1-α)*Ν at least the numbers should be different from
+%% short and long links...
+The main objective of the overlay construction phase is to yield a network that
+exhibits low clustering. In turn, this is a beneficial characteristic for
+random graphs as it can lead to a larger coverage of peers and at the same time
+helps reduce duplication of messages travelling around the network. This
+combination offers efficient lookups that pose minimal or no negative impact to
+other mechanisms possibly employed by the \p\ systems (e.g., 1-hop replication
+or dynamic querying mechanisms).
 A 50\% reduction in communication latency among peers by cutting off 
@@
 A newly arriving peer, randomly connects to an already participating 
-node through a bootstrap server. 
-Once connected, the peer starts receiving information throughout the
-link-state mechanism and thus, after some time, 
+node through a bootstrap server. Once connected, the peer starts receiving
+information throughout the link-state mechanism and thus, after some time, 
 it has a complete picture of the overlay graph. 
@@
 for example the minimization of the average delay to all its
-neighbors. 
-\emph{EGOIST} requires extensive resource usage to continuously update the
-wiring of all the nodes in the system. 
-%%AD I am not sure I get the "wiring of all"??? clarify.. 
-The approach can reduce the load imposed by the monitoring 
-and announcing process from $O(n^2)$ to only $O(kn)$, 
-where $n$ is the number of all network nodes 
-and $k$ is the number of links a node chooses to establish.
-%%AD is this an average number?? What kind of number???
+neighbors. \emph{EGOIST} requires extensive resource usage to continuously
+update the overlay connections of all the nodes in the system. This needs
+$O(n^2)$ measurements. Fortunatelly, each node does not need to do these
+measurements for maintenance (monitoring and updating) with all other
+participating nodes, but just with a number of $k<n$ nodes that it chooses to
+establish links with. This gives a reduced $O(kn)$ time complexity for
+maintenance while $O(n^2)$ is needed less frequently, when nodes reevaluate
+their connections.
+
 Regarding the $3$ criteria, \emph{EGOIST}'s behavior is as follows:
@@
 
-% TODO: CHECK IF THE FOLLOWING BITTORRENT ALGORITMS BELONG IN THIS SUBSECTION
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 The main objective of the \emph{Biased Neighbor Selection (BNS)} 
@@
 by carefully choosing most of the neighbors of a peer to come from 
-the same \emph{ISP} while at the same time adhering to the near-optimal
+the same \emph{ISP}, while at the same time adhering to the near-optimal
 download performance of the protocol.
 %%
-By and large, {\sl BitTorrent}s deploy mechanisms that have proved 
+By and large, {\sl BitTorrent} deploy mechanisms that have proved
 aggressive for \emph{ISP}s' networking and accounting as they function 
-in a ``without-borders'' approach.
-% BitTorrent systems employ most of the time
-% mechanisms that can be proven very aggressive to an ISP's networking and
-% accounting, basically, due to their lack of knowledge of Autonomous System
-% boarder limits. 
-The default {\sl BitTorrent} specification designates $35$ connections 
-for each peer regardless if such connections are placed within or outside 
-the borders of the \emph{ISP} that hosts the node.
+in a ``without-borders'' approach. The default {\sl BitTorrent} specification
+designates $35$ connections for each peer regardless if such connections are
+placed within or outside the borders of the \emph{ISP} that hosts the node.
 %%
@@
 offering the sought services. 
-%%AD rewrote as above...
-% Biased neighbor selection experiments on some number $k$ which
-% denotes a number of neighboring peers which are not on the same ISP than the
-% peer at hand. 
-% Consequently, the remaining $35 - k$ peers are from the same ISP.
-% These $k$ nodes are used to provide a more extended, global view of a
-% network, but balance the load within the limits of a single ISP.
-% This prevents clients from exchanging traffic through an expensive transit link,
-% if there are alternative local connections that could offer the same service faster and
-% at no additional cost for the ISP. 
 \emph{BNS} is realized either through 
@@
 and caching techniques for near-optimal lookup results is proposed.
-% The paper also proposes
-% a combination of biased neighbor selection along with bandwidth throttling and
-% some caching techniques for near optimal results. 
-%%%%%%%%%%%%%%%%%%%%%
-% Unfortunately, implementation
-% and deployment
-% need knowledge of ISP Autonomous System mappings %%AD What are these Aut. Sys?
-% or infrastructure changes (installation of P2P traffic shaping devices) on the
-% ISP, making widespread adoption.
-%%AD I am not sure what the above sentence says! I rephrased as follows:
 Unfortunately, \emph{BNS} deployment requires knowledge of 
@@
 and enhances the downloading rates by favoring connections
-within the borders of a single autonomous system.
-% \subsubsection{Ono}
-% Choffnes and Bustamante propose Ono \cite{CB2008}, a protocol for managing
-% BitTorrent traffic so that it significantly reduces cross-ISP traffic and at the
-% same time enhances downloading rates by favoring connections within the borders
-% of a single autonomous system. 
-Contrary to \emph{BSN}~\cite{BCCMSBZ2006},
+within the borders of a single autonomous system. Contrary to
+\emph{BSN}~\cite{BCCMSBZ2006},
 \emph{Oro} leads to improved performance 
@@
 infrastructure and network topology information.
-% Contrary to biased peer selection proposed in
-% \cite{BCCMSBZ2006}, this work can lead to 
-% improved performance without requiring
-% any cooperation between ISP subscribers, or requiring any additional
-% infrastructure and network topology information. 
 \emph{Oro}'s selection approach is 
 landmark-based and leverages existing {\sl CDN}-infrastructure 
-for peer distance estimation. 
-%%AD I do not understand the following sentence... I am not even sure what 
-%%   who is close to each other??? Confusing sentence. Rephrase...
-{\sl CDN}s already use both static (i.e., geographical) and dynamic
-(network measurement systems) information for their replica selection, so the
-authors claim that peers which exhibit similar redirection behavior, are 
-likely close to the replica servers and as a consequence close to each other.
+for peer distance estimation. {\sl CDN}s already use both static (i.e.,
+geographical, IP-clustering) and dynamic (i.e., network delay measurement)
+information for their replica selection. Thus, the algorithm leverages the
+observation that peers that are redirected to a specific CDN replica does not
+only mean that these are close by this replica but to each other as well.
 %%
@@
 %%
-The bootstrapping phase %%AD of what?? the protocol's bootstrapping???
-consists of the peer performing {\sl DNS} lookup to {\sl CDN}-names
-%%AD when you say "CDN names" - what are they????
-to build its redirection information.
+The protocol's bootstrapping phase consists of the newcomming peer performing
+{\sl DNS} lookup to {\sl CDN}-names to build its redirection information.
 %%
@@
 is sufficiently fresh, 
-\emph{Oro} persists the information 
-after the end of a BitTorrent session.
-%%AD what information is this?? DNS look ups?? 
+\emph{Oro} persists the map after the end of a BitTorrent session.
 %%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \subsubsection{Locality-Awareness in BitTorrent-like P2P Applications}
 Three approaches that ``inject'' locality awareness 
-in {\sl BitTorrent}-like applications are discussed in~\cite{LCLX2009}.
-% \subsubsection{Locality-Awareness in BitTorrent-like P2P Applications}
-% In \cite{LCLX2009} the authors study and compare three different
-% approaches in injecting locality awareness in BitTorrent-like applications. 
-The first \emph{macroscopic-level} approach 
+in {\sl BitTorrent}-like applications are discussed in~\cite{LCLX2009}. The
+first \emph{macroscopic-level} approach
 focuses on improving the neighbor selection process. 
@@
 send out the top-$k$ list of nodes (e.g., $k$=$50$).
-%% Autonomous System  hope distance... 
-%%AD Got this out - do u need it?? What does it mean "Autonomic System Hop Dist"???
-%%
-The second approach manipulates the {\sl BitTorrent} chocking/unchocking.
-% at an \emph{intermediate-level}. 
-A peer unchockes its $4$ closest in terms of 
-% Autonomous System 
+%%
+The second approach, applied in an \emph{intermediate-level}, manipulates the
+{\sl BitTorrent} chocking/unchocking mechanism. A peer unchockes its $4$
+closest in terms of 
+% VM I put ``autonomous system'' back in again because in the context of this
+% paper we are dealing on the number of AS crossings (one-hop = one AS crossing)
+% untli we reach a neighbor.
+Autonomous System
 hop-count interested neighbors. 
@@
 Finally, the third approach that operates at 
-a \emph{microscopic-level}, substitutes the rare-first chunk picking
-policy by a locality-first policy. 
-%%AD is it "rarest"?? What does it mean "locality-first"???
-The distance value of a piece is computed as the mean value 
-of the distances of the peers that posses it. 
-%%AD What is this "piece"??? I am not sure how the average gets computed? 
-%%   It is very cryptic description...
-Through experimentation, %%AD simulation? real?
-the locality-based approach achieves traffic optimization.
+a \emph{microscopic-level}, substitutes the rare-first chunk picking of the
+original {\sl BitTorrent} protocol that picks the rarest of the available
+chunks to download next, with a locality-first policy which picks first the
+chunck that is closer. The distance value of a chunk is computed as the mean
+value of the distances of the peers that posses it.
+Through experimentation on a real-world Internet topology simulated on
+PlanetLab, the locality-based approach achieves traffic optimization.
 Nevertheless, it does not do that well when compared with the
 random approach of the standard {\sl BitTorrent} protocol.
-%%AD rephrased and shortened the following as above...
-% The paper conducts several measurements on the efficiency of this scheme 
-% and reaches 
-% the conclusion that the major design decision for all locality
-% aware P2P systems is the trade-off of optimizing inter- Autonomous System
-% traffic versus achieving fairness amongst peers.  While their
-% locality-based approach achieves traffic optimization, it does not do well
-% when compared with the 
-% random approach of the standard BitTorrent protocol.
-%%%%%%
-%%AD so in general how does this approach help the mismatch problem?? 
-%%	need a statement here (1-2 sentences explaining this).
+
+All methods target in exposing the locality of the basic functions of the
+BitTorrent protocol. Even though they do not restructure the network to map
+better to the underlying infrastructure they favour the intra-AS communication
+over inter-AS communication thus reducing the general cost of communication.
+
 \begin{center}
@@
 low &
-%
-high
-\end{tabular}
-}
-\end{center}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% does not result in a better map to the underlying topology, it just eases
+% the implications of complete locality unawareness.
+medium
+\end{tabular}
+}
+\end{center}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \subsubsection{TopBT}
 \emph{TopBT}~\cite{RTLCGZ2010} seeks to enhance 
 proximity awareness in the {\sl BitTorrent}
-without the need of any additional infrastructure.
-% \subsubsection{TopBT}
-% \cite{RTLCGZ2010} proposed an approach for enhancing proximity awareness
-% in the BitTorrent protocol without any need of any infrastructure installed. 
-The key conjecture upon which \emph{TopBT} is built is that 
+without the need of any additional infrastructure. The key conjecture upon
+which \emph{TopBT} is built is that
 a good peer selection metric should take into account both the
@@
 divided by the either the link-level $l$ or the number of routing hops required $a$.
-%%AD The following is not a readable sentence... rephrase as above...
-% Thus, it proposes the use of the
-% ratio of download rate to upload rate, $\frac{d}{u}$, divided by link-level or
-% Autonomous System -level routing hops, $l$ or $a$ respectively, in an attempt to
-% form a comprehensive metric to select peers with high download rate, low
-% reciprocal upload demand and low routing hops.  %%AD What is "Autonomous System -level routing hops?????"
-%%%%%%%%%%%%%%%%%%%%%%%%
+This metric should guarantee selection of peers with high download rate, low
+reciprocal upload demand and decreased routing hops.
 The above \emph{TopBT}-metric is applied in various
@@
 %%
-In a effort to essentially improve on the inherent topology mismatch problem,
-\emph{TopBT} modifies the peer list the
-tracker returns when someone first tries to connect to the swarm,
-the initial connection establishment, and the
-unchoking mechanism in the {\sl BitTorrent}.
-%%AD how do they specifically are modified??? What is the difference 
-%%	with the previous paper?
-%%
-%
+In an effort to essentially improve on the inherent topology mismatch problem,
+\emph{TopBT} modifies the peer list the tracker returns when someone first tries
+to connect to the swarm, the initial connection establishment, and the
+unchoking mechanism in the {\sl BitTorrent}. The important feature of TopBT is
+that it does not need wide deployment of its same types in order to see results.
+It also supports both link-hop and AS-hop metrics to identify proximity in
+different levels.
 %
@@
 if compared with popular {\sl BitTorrent} implementations.
-%%AD rephrased what is below as above ...
-% the authors claim more than 25\% traffic reduction and 15\% faster downloads,
-% both with a lightweight cost when compared to popular BitTorrent
-% implementations.
-%
 In reference to the stated $3$ criteria, \emph{TopBT} fares as follows:
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \subsubsection{Underlying Topology-Aware Peer Selection (UTAPS)}
 The \emph{Underlying Topology-Aware Peer Selection (UTAPS)}~\cite{LCY2008}
-is an effort to furnish an enhanced peer selection strategy.
-% \subsubsection{Underlying Topology-Aware Peer Selection (UTAPS)}
-% UTAPS \cite{LCY2008} is another peer selection strategy for the BitTorrent
-% protocol. 
-\emph{UTAPS} essentially works in two stages: in the first,
+is an effort to furnish an enhanced peer selection strategy. \emph{UTAPS}
+essentially works in two stages: in the first,
 it collects information to develop a picture of the underlying topology 
 and in the second, it proceeds to make the peer selection based on 
-this knowledge.
-In the first stage, \emph{UTAPS} employs
+this knowledge. In the first stage, \emph{UTAPS} employs
 network tomography, a technique that probes from a large network's 
@@
 The tracker provides this information to the newcomer in the form
-of the bootstrapping peer list. 
-%%
-%%AD I am lost among the peers in the following sentence... Who traces whom???
-The peers further enhance this picture by
-trace-routing the returned peers themselves and reporting back to the tracker.
+of the bootstrapping peer list. These newly joined peers traceroute the tracker
+peer list themselves and return back to the tracker in order to further enhance
+the tracker's view of the infrastructure.
 %%
 In the second stage, \emph{UTAPS} employs a set of heuristics
-which target those peers that expose low \emph{RTT}
-and are found within certain hop-counts away from 
-observed routers. %%AD u had "measured routers" - how do you select these routers?
-% For the selection part, a family of heuristics are proposed, which in general
-% target those peers that expose low RTT and are within certain hop-count away of
-% measured routers. 
-%%
-%%AD if you leave this "the authors claim..." and an such an author gets 
-%%	to spend 30 secs with the paper, you are dead! dead dead!!! 
-%%
-% The author's claim that peers running UTAPS instead of random
-% peer selection can achieve better downloading rates with reduced burden on the
-% underlying ISP infrastructure though their technique is coarse-grained and their
-% evaluation configuration was far too small for extracting further safe
-% conclusions.
+which target those peers that expose low \emph{RTT} and are found within
+certain hop-counts away from
+the routers observed during the initial traceroute.
+
 It is reported in~\cite{LCY2008} that peers running \emph{UTAPS}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \subsubsection{An Effective Network-Aware Peer Selection Algorithm in
+% BitTorrent}
 A clustering approach that differentiates peers encountered in a swarm
@@
 presented in~\cite{QLZG2009}.
-The classification is carried with the help of the routers used 
-while peers communicate with the tracker.
-% \subsubsection{An Effective Network-Aware Peer Selection Algorithm in
-% BitTorrent}
-% Qin et al., in their work in \cite{QLZG2009}, propose a
-% clustering approach that differentiates peers in a swarm into local, intra-ISP
-% and inter-ISP neighbors. 
-% The classification is done through the routers they
-% use in order to communicate with the tracker. 
-%%
-%%AD In the following sentence it is not clear who sends to WHOM and WHY?
+The classification is carried with the help of the core routers used 
+while peers communicate with the tracker. It is assumed that even though the
+network may be highly volatile, these core routers are are usually more stable.
 For this, a newly arriving
-{\sl BitTorrent} peer trace-routes the tracker and 
-sends the results in a form of a
-vector of triples that contain the \emph{IP} 
-and the hop counter for the router on the
-trace-route path as well as the link latency 
-of the hop that was made in order to arrive to this router. 
-%%
-The tracker uses this information in order to classify
-the router through a $k$-means classification algorithm, with $k = 3$. 
+{\sl BitTorrent} peer, trace-routes the tracker and stores vectors containing
+info like the IP address and the hop number of a traced router, as well as the
+link delay of the traced router and the previous hop along the trace-route path.
+These vectors are reported to the tracker and the latter uses this information
+to classify the router through a $k$-means classification algorithm, with $k = 3$. 
 For peer selection, \cite{QLZG2009} proposes a biased approach 
@@
 the total cross-\emph{ISP} traffic.
-% The evaluation of the algorithm is done in a similar artificial,
-% contained laboratory environment as in \cite{LCY2008} and it shows up to
-% 5\% faster downloading times, and up to 22\% reduction of the total cross-ISP
-% traffic.
 This clustering approach fares as follows in regard 
@@
 % TODO: how intensive with regards to computing power is an on-line k-means
-(with
-% k=3) classification???
+% (with k=3) classification???
+% VM: do we have an idea?
 medium
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+% \subsubsection{Peer-exchange Routing Optimization Protocols (PROP)}
 In~\cite{QCYCZ2007}, two protocols termed 
@@
 while aiming at reducing the network's overall link latency.
-%%
-% \subsubsection{Peer-exchange Routing Optimization Protocols (PROP)}
-% \cite{QCYCZ2007} introduces two protocols, one generic and one optimized,
-% collectively referred to as \emph{Peer-exchange Routing Optimization Protocols
-% (PROP)} to adjust the neighborhood graph of an overlay network in order to
-% reduce its overall link latency. 
 \emph{PROP} algorithms are weaved around the concept of
@@
 choose its own strategy.
-In the generic form of the \emph{PROP-G} protocol,
-a peer can exchange all of its
-neighbors with those of another peer. 
-The optimized version \emph{PROP-O} permits peers to 
-exchange to exchange selective sets of neighbors.
-%%AD give a sentence or two how this gets done..
-The generic \emph{PROP-G} protocol guarantees
-the connectivity of the overlay graph during exchanges, 
-therefore, it can be applied
-to both unstructured and structured overlay networks.
+
+%%
+%% VM ADDING SOME OPERATIONAL DETAILS BELOW.
+The PROP algorithms operate, basically, into two phases. During the \emph{warm-up}
+phase a node $u$ probes its neighbors to collect distance information. Then it
+contacts, at a fixed time rate $timer$, a random node $v$, $nhops$ away from it. New distance
+information is calculated independently by nodes $u$ and $v$, now for the
+hypothetical state when the potential exchange would occur. If the benefit
+gained, in terms of reducing the average distance among the peers involved, is
+above some predefined threshold then the exchange will actually occur,
+otherwise no further action will be taken. The \emph{maintenance} phase, on the
+other hand, differs from the warm-up one, in that the probability
+of peers to be probed and the $timer$ interval that this happens, now depends on
+past peer exchange results.
+
+In the generic form of the \emph{PROP-G} protocol, a peer can exchange all of
+its neighbors with those of another peer. This essentially results in the two
+nodes exchanging positions so the topology and connectivity of the overlay is
+not affected by the operation of the algorithm. The optimized version
+\emph{PROP-O} permits peers to exchange selected sets of their neighbors. It is
+not allowed to exchange neighbors allong the path of the nodes that perform the
+exchange in order to ensure that in the end of the process the peers will remain
+connected. Additionally, the algorithm needs to preserve the characteristics of
+the network (i.e., the graph remains connected, preserve the degree of the
+nodes), so the exchange always involves equal number of neighbors.
+
 The following table depicts the behavior of the \emph{PROP} 
@@
 the same domain are also topologically close.
-While seeking to address the mismatch problem, \emph{DDNO} 
-is a heuristic approach that uses
-\emph{Split-Hash} and \emph{dnMatch} algorithms 
-to detect locality based on domain names. 
-%%
-\emph{DDNO} outcome is 
-a flat overlay topology which with some adjustments 
-can be utilized in super-peer architectures as well. 
-According to the algorithm,
+%%
+\emph{DDNO}'s outcome is a flat overlay topology which, with some adjustments, 
+can be utilized in super-peer architectures as well. According to the algorithm,
 half of the possible connections of a node are used to connect to local peers
 (called \emph{sibling} connections) and rest are used to randomly
-connect to the peers anywhere on the network 
-(called \emph{random} connections).
+connect to the peers anywhere on the network (called \emph{random} connections).
 The former ensure the reduction of long distance traveling for messages, while
-the later secure the connectivity of the structure and prevent partitioning.
-%%AD partitioning of what? network partitioning??
-Upon arrival the newcomer asks a \emph{host-cache} to establish its
-random connections. 
-For establishing its sibling connections, the newcomer
-multicasts a specialized \emph{walker message} to 
-locate the right candidate(s). 
-%%AD I am not sure what the following sentence says... It is too convoluted..
-If it is found then the candidate as well as its own siblings will
-respond to the newcomer about their willingness to establish connections with
-it.
-%%AD I am not sure what this paragraph offers.. I am somewhat puzzled: is 
-%%	this about what exactly?? Some refitting of the short description 
-%%	with the mismatch problem I think is needed...
-%%	In general, the description ends somewhat abruptly...
-%%
-%%
-In terms of the $3$ criteria, \emph{DDNO} fares are follows:
+the later secure the connectivity of the structure and prevent network
+partitioning.
+%% VM NEW ATTEMPT TO EXPLAIN THE ALGORITHM
+Upon arrival the newcomer peer $n$ asks a \emph{host-cache} to establish its
+random connections. For establishing its sibling connections, the newcomer
+multicasts a specialized \emph{message} $l$, to locate the right candidate(s).
+Initially, the behaviour of this message, is modeled as a random walker,
+until $l$ reaches some peer $d$ capable to guide it through its next step.
+This capability is enabled through a \emph{ZoneCache}, a data structure that
+contains information for nodes accessible in an $r$-hop radius from $d$.
+Ultimatelly $l$ will reach a node $m$, candidate to become $n$'s sibling. $m$
+will issue a broadcast message to its own siblings and then all togeather will
+inform the initiating peer $n$ of who is willing to accept new connections.
+
+The population and maintenance of ZoneCaches across the network is an important
+task for the efficiency of the constructed overlay. While seeking to address the
+mismatch problem, \emph{DDNO} is a heuristic approach that uses
+\emph{Split-Hash} and \emph{dnMatch} algorithms to effectively encode domain
+names and determine domain locality respectively. In terms of the $3$ criteria,
+\emph{DDNO} fares are follows:
 \begin{center}
@@
 the respective \emph{Regional Internet Registry (RIR)}'s IP assignments.
+%% VM CHANGED DESCRIPTION
 Based on these assignments, nodes that belong in the same 
-organization are always addressed from the same block.
-%%AD above, I simply do not understand the ..."from the same block"
-%%	What does it mean??? 
-%%
-% \emph{Critical Topology-Aware Grouping (CTAG)} presented in \cite{ZL2006}, is a
-% grouping algorithm. The grouping strategy is based on the \emph{Internet
-% Assigned Numbers Authority (IANA)} and the respective \emph{Regional Internet
-% Registry (RIR)}'s IP assignment strategies, according to which nodes within the
-% same organization are always addressed from the same block. 
-\emph{CTAG} proposes the \emph{Adjacency Measurement (AM)}, 
-a technique which uses the longest
-matching \emph{IP} segment criterion to compute node proximity. 
-\emph{CTAG} focuses on both
-the construction of the overlay as well as its 
-constant and dynamic adaptation.
-In the first phase, called \emph{bootstrapping grouping}, 
-the {\sl Gnutella} web caching mechanism is modified 
-so that a newcomer may choose the closest %%AD closest here means what???
-cache to retrieve the bootstrapping list. 
-%%AD for sentence below - what is AM matric??  X-Try?? - 
-%%		can you rework the sentence?
-During the second phase called \emph{dynamic revision}, 
-the AM metric is used to store hosts' 
-addresses read from \emph{X-Try} headers during handshake or from
-\emph{QueryHit} messages. 
-When a node reaches its maximum number of neighbor
+organization are always addressed from the same block of IP addresses.
+\emph{CTAG}, exploits this observation to construct topology-aware unstructured
+overlays. On top of this, \emph{Adjacency Measurement (AM)} is proposed, as
+a technique which uses the longest matching \emph{IP} segment criterion to
+compute node proximity. \emph{CTAG} focuses on both the construction of the
+overlay as well as its constant and dynamic adaptation.
+In the first phase,
+called \emph{bootstrapping grouping},
+the {\sl Gnutella} web caching mechanism
+is modified so that a newcomer may choose the closest, in terms of AM,
+cache to retrieve the bootstrapping list. Similarly, during the second phase,
+called \emph{dynamic revision}, the standard connection establishement mechanism
+is also altered so that nodes with the highest AM metric are chosen. When a node
+reaches its maximum number of neighbor
 connections, it disconnects those neighbors with the lowest AM.
@@
 %%	for each peer there is a different bin right?
-The ordering represents a \emph{bin},
-in the sense of close-by nodes having the same landmark ordering and hence
-belong to the same bin. 
-%%
-%%AD I think the above sentence is the main information related in this 
-%%   piece of description so it requires a bit better description...
-Should there $m$ landmarks be adopted, potentially $m!$  different bins exist.
-% This means that a landmark system consisting of $m$ such
-% nodes results in $m!$ potential different bins.
+%
+%The ordering represents a \emph{bin},
+%in the sense of close-by nodes having the same landmark ordering and hence
+%belong to the same bin. 
+%
+%% VM I CHANGED THE ABOVE AS BELOW TO REPHRASE IT.
+Each permutation of the set of landmarks represents a specific bin. Should there
+$m$ landmarks be adopted, potentially $m!$  different bins exist. Peers that
+end up with the same such ordering, belong to the same bin, in the sense that
+if they experience similar overall latencies from fixed network points then it
+is very likelly that they are close to each other.
 %%
@@
 both structured and unstructured \p\ systems.
-% \cite{RHMKS2002} has the detailed description of the algorithm for
-% both architectures. %%AD obviously!!!!
-%%
-%%AD why is it important to say this about CDNs???
-% Landmark binning is also considered as a good candidate for
-% content distribution networks. 
+%%
 The main disadvantage of the approach is that 
@@
 inherently play a pivotal role in the operation of the network.
-%%AD rephrased what is below as above...
-% A typical P2P network usually has a couple of
-% million nodes connected at any given time, thus rendering landmark servers an
-% important parameter when designing for large scales.  
-%%%%%%%%%%%%%%%%%%
-%%AD rephrased what is below as above...
 Although latency estimation does not drain network resources,
@@
 excessive network traffic flow through its landmark server-clusters.
-%%AD rephrased what is below as above...
-% The authors claim that
-% latency estimation does not drain much from the network resources but in case
-% scalability problems arise, they claim the solution is replacing single landmark
-% servers with clusters of servers within the same physical area. However, this
-% approach does not reduce the possibility of excessive network traffic flow
-% through these landmark servers. One other possible problem is the incorrect
-% binning caused by the use of inaccurate methods for delay measurement
-% (latency is not proven to be a safe way to estimate distance).
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-%
-%%
-%The algorithm can be considered scalable, as nodes need only compute distances
-% to a small number of predefined nodes and thus without exchanging any
-%information.
+%%
 %To answer the question as to whether the algorithm actually contributes
@@
 medium &
-% The overhead is confined to the communication of a node with, most probably,
-% 12 landmark servers at most.
+% The algorithm needs only nodes to compute distances to a small number of
+% predefined nodes without exchanging any additional information.
 low &
@@
 % \subsubsection{mOverlay}
-The \emph{mOverlay} approach addresses scalability issues
-that might be arise when static landmark servers are in use. 
+The \emph{mOverlay} \cite{ZZZSZ2004} approach addresses scalability issues
+that might arise when static landmark servers are in use.
 To this end, the use of dynamic landmarks is proposed.
-% \emph{mOverlay} \cite{ZZZSZ2004} tries to 
-% addresses the scalability issues that
-% might be imposed on networks (i.e. load unbalance) when using static landmark
-% servers by introducing dynamic landmarks. 
 \emph{mOverlay}'s founding notion is that of a \emph{group} that 
@@
 may involve metrics including \emph{RTT}s and network latencies.
-%%AD rephrased what is below as above... 
-% Zhang et al. introduce the notion of a
-% \emph{group} which is a set of peers considered close to each other with respect
-% to any position in the underlying network, where proximity is defined by some
-% user-defined cost metric like RTT, latency or something else. 
-%%%%%%%%%%%%%%%%%%%%%%%
 By and large a clustering approach, \emph{mOverlay} seeks to 
@@
 while at the bottom, only \emph{intra-group} connections occur among peers.
-%%AD rephrased what is below as above... 
-% mOverlay is considered a clustering approach since 
-% it tries to recreate small-world-like
-% properties to the overlay network by creating a two-level hierarchical
-% structure, where on the top level we have connections between groups while on
-% the bottom we have connections between peers inside groups. 
-%%%%%%%%%%%%%%%%%%%%%%%
+
+%%AD how do you identify the initial groups????
+%%
+%% VM REPHRASING IN ORDER TO ANSWER THE QUESTION ABOVE
 Clearly, identifying groups and accurately finding the closest group 
 to a peer is a fundamental concern in the creation of the overlay.
-% Finding the correct closest group is the most important part 
-% of the overlay construction. 
 Nodes are grouped based on their distance to 
 the groups already in the network, rendering
-the latter be the \emph{dynamic landmarks} in the process. 
-%%AD how do you identify the initial groups????
-%%
+the latter be the \emph{dynamic landmarks} in the process. For a newcomming
+peer $Q$ the \emph{grouping critirion} says that when the distance of $Q$ and
+some group $A$’s neighbor groups is the same as the distance
+between group $A$ and group $A$’s neighbor groups, then host $Q$
+should belong to group $A$. During network initialization or when grouping
+criterion is not met, new groups are created.
 It is shown that in the proposed scheme, a new peer can 
@@
 leader selections, should one either leaves or dies.
-%%AD rephrased what is below as above...
-% The authors formally prove
-% that any new node can reach its group by exchanging at most $O(logN)$ messages
-% within the network. Finally mOverlay also considers a second important function
-% in its protocol which is the constant maintenance of the overlay.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
 %
@@
 
-% The introduction of landmark servers renders the approach not fully
-% decentralized, thus preventing it from scaling smoothly. Communicating
-% and overloading landmark servers in high-churn systems is another scalability
-% concern.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% TODO: HOW CAN UNSTRUCTURED SCHEMES BE REFINED
+%
+%For this reasons, efforts have been placed for optimizing the efficiency of
+%decentralized unstructured peer-to-peer networks. Research mainly focuses on
+%\begin{inparaenum}[\itshape i\upshape)]
+%  \item reducing unnecessary, redundant communication traffic, and
+%  \item exploiting physical locality to reduce communication response.
+%\end{inparaenum}
+%The goal can be achieved at, both, the application-level network as well as the
+%underlying physical one. In the first case by refining the message relay
+%techniques, while in the second one, by adaptively reconstructing the
+%application network to map as well as possible to the the physical network.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 utilized by the discussed approaches; these methodologies are:
-%%AD shorten below as above...
-% In this section, we introduced the topology mismatching problem from the
-% perspective of unstructured P2P systems and we visited a series
-% of algorithms that have been proposed in the literature to tackle its negative
-% influence on performance. In surveying the above algorithms, we identify
-% four main methodologies that are used either selectively or in combination
-% by each algorithm.  
-% These methodologies are:
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{enumerate}[\itshape i\upshape)]
@@
 \end{enumerate}
-Each surveyed approach uses either one or more of the
-aforementioned four methodologies. 
-Below, we discuss how the proposed protocols use 
-elements of the four methodologies and we offer 
-a summary qualitative comparison for all surveyed approaches 
-applicable for \emph{unstructured} \p \emph{systems}.
-%%
+%%VM Added the following to mention the hybrid/combo methodologies
+An important aspect of many implementations, is that since the above
+methodologies use heuristics to solve topology mismatch, an
+intrinsically difficult problem, no single approach can offer a robust solution.
+Thus, many of the algorithms we surveyed are actually combinations of multiple
+such methodologies, in an attempt to refine the results of using just a single one.
+For example, \emph{mOverlay} combines topology adaptation with landmark binning or \emph{Gia}
+combines all three topology adaptation, forwarding optimization and caching
+methodologies. Some times combination results in win/loose trade-offs. For
+instance, the effectiveness of a caching/replication component can be undermined
+by a continuously adapting overlay that removes important links between peers.
+For this reason, the reader is advised to criticaly approach either the overall
+algorithms or the separate methodologies when evaluating them for different
+application domains.
+
+Below, we discuss how the proposed protocols use elements of the four
+methodologies and we offer a summary qualitative comparison for all surveyed
+approaches applicable for \emph{unstructured} \p \emph{systems}.
+%%VM I brought the following paragraph here from the end of this subsection.
+Moreover, Table~\ref{unstructured:table} offers a summary overview for each effort,
+showing which of the aforementioned methodologies it incorporates, along with its
+highlights and a rough estimation of the positive and negative aspects of its
+implementation.
 
@@
 modify the topology of the \p\ network
-using various techniques. 
-The two most commonly used approaches to 
+using various techniques. The two most commonly used approaches to
 topology adaptation are creating
@@
 minimum spanning trees over
-this graph that can efficiently route messages among peers. 
-Even though spanning trees provide efficient query performance, 
-their maintenance costs have proven to be a significant limitation. 
-This is why the original \emph{Narada} proposal
-focused on small groups and thus, is not scalable.  
-%%
-The \emph{AOTO, LTM} and \emph{SBO} try to 
-% Algorithms like AOTO, LTM or SBO try to %%AD this "like" is colloquial 
-					  %%   try to avoid it if you can!
-overcome these obstacles with 
-ingeneous schemes
-% clever tricks %%AD this is a bit colloquial
-that entail forming minimum spanning trees
-for the $2$-hop neighbors ($N^2$) for each node, 
-%%AD u imply each has N neighbors above right???
-partitioning the graph into two
-random groups where each group is responsible for different tasks, 
-or performing local optimizations dynamically on the overlay graph. 
-%%AD what are these "optimizations" (in 2-3 words  - in sentence)
-The advantage of building minimum spanning trees is that 
+this graph that can efficiently route messages among peers. The \emph{AOTO, LTM}
+and \emph{SBO} try to
+overcome this limitation with ingeneous schemes
+like forming minimum spanning trees
+for the $2$-hop away neighbors for each node, seperating participant nodes into
+groups, sometimes with different responsiblities and tasks at hand (e.g.,
+\emph{SBO}). The advantage of building minimum spanning trees is that
 they maintain the connectivity on the network in
@@
 
-%%
-%%AD this figure HAS to be referenced in the text!!! 
-%%   ALSO - is this ANY different from Fig 4 of the paper??? 
-%%   WHY this Fig is being repeated here??? Grand Mystery!!!!
-%%	Perhaps we should put one fig similar but more sexy??? ;-)
-\begin{figure}[ht]
-\centering
-\subfigure[Inefficient overlay topology.] {
-  \includegraphics[scale=0.3]{img/pdf/topology-adaptation-before.pdf}
-  \label{figure:topology-adaptation:before}
-}\qquad\qquad
-\subfigure[Efficient overlay topology after adaptation.] {
-  \includegraphics[scale=0.3]{img/pdf/topology-adaptation-after.pdf}
-  \label{figure:topology-adaptation:after}
-}
-\caption{A simplified example on how overlay topology adaptation may improve matters.}
-% A simple example of how an overlay topology can be adapted.}
-\label{figure:topology-adaptation}
-\end{figure}
-%%
-
 The cluster-based approaches, on the other hand, 
@@
 \emph{T2MC}, for example, uses trace-route logs and 
-\emph{DDNO} exploits domain names to cluster nodes in proximity. 
-Unfortunately, commonly used
-methods for proximity detection across Internet do not always return
-reliable results and therefore, mapping accuracy is not guaranteed. 
-Moreover, trace-route is a computing resource ``heavy-weigth'' utility 
-and so  a number of network equipment vendors do not allow
-trace-route calls at all. %%AD do you have a refeerence or two in here???
-%%
-The most problematic aspect of clustering, though, is
+\emph{DDNO} exploits domain names to cluster nodes in proximity at peer join.
+Further enhancements may include dynamic local restructuring of the overlay
+graph through neighbour exchange like in \emph{PROP} or cycle-cut like in
+\emph{DCMP} to achieve continoous adaptation thoughout a peer's lifecycle.
+Unfortunately, commonly used methods for proximity detection across Internet do
+not always return
+reliable results and therefore, mapping accuracy is not guaranteed. Specifically
+for traceroute, its overhead is not negligible and routers or firewalls in a
+network may have already been configured to disable traceroute response from
+the start. The most problematic aspect of clustering, though, is
 its nature per se. 
@@
 significantly shrink the search scope, negatively affecting the query response
-time that the \p\ user experiences. 
-% Techniques, like DDNO, that try to 
-Among others, \emph{DDNO} tries to 
+time that the \p\ user experiences. Among others, \emph{DDNO} tries to 
 balance the efficiency of the clustering approach 
-with enhanced node connectivity. 
-In particular, \emph{DDNO} attemps to address this limitation
-by forcing half of each node's connections to be with other, randomly
-selected, nodes in the network.
-%%AD you mean the "logical network"? the p2p network???
-
-As a methodology, 
-topology adaptation ultimately aims to reduce the 
+with enhanced node connectivity by forcing half of each node's connections to be
+with other, randomly selected, nodes.
+
+As a methodology, topology adaptation, ultimately aims to reduce the 
 average path traversing cost from one node to another.
@@
 overlay network so that it becomes a better fit 
-for the underlying \emph{IP} network
-as Figure~\ref{figure:topology-adaptation} depicts.
+for the underlying \emph{IP} network.
+% as Figure~\ref{figure:topology-adaptation} depicts.
 Despite the fact that the above is a plausible proposition,
-it is not always advantageous.
-% This is fair in the general case but not always. 
-%%
-For example, an topology adaptive algorithm can exchange 
-a slow edge in the overlay with a faster one
+it is not always advantageous though. For example, a topology adaptive
+algorithm can exchange a slow edge in the overlay with a faster one
 thus reducing the average latency of the network communication.
 The pitfall here is that this new virtual link
-can traverse a fast \emph{AS}--to--\emph{AS} 	
-			%%AD you mean Autonomous Systems here?? 
-		  	%%    What is this AS???
-link meaning that even though message
-round-trip-time is reduced, it has more cost 
-in terms of communication network economics and management.
-%%AD just check the above sentence for correctness..
-
-% These approaches include creating spanning trees using
-%connection graphs, creating cluster of physically close nodes, or using latency
-%information to detect proximity. The brief description of each category is
-%presented below:
-%
-%  \begin{itemize}
-%    \item \emph{Spanning tree based}. These approaches construct of a rich
-%    graph  based on the network connections and build minimum spanning tree
-%    on the graphs, causing large traffic overhead to the system
-%    \cite{chu_esm_2000,chu_esm_2002}.
-%
-%    \item \emph{Cluster based}. These approaches select to link physically
-%    closer nodes with each other, therefore shrink the search scope
-%    significantly while mapping accuracy is not always guaranteed.
-%
-%    \item \emph{Minimum latency first}. Use of latency as a metric to calculate
-%    distance among peers. They require global latency information
-%    ``landmarks''\footnote{Measuring latency between peers and stable Internet
-%    servers.}.
-%
-%  \end{itemize}
+can traverse a fast \emph{AS}--to--\emph{AS} link meaning that even though
+message round-trip-time is reduced, it has actually additional cost in terms of
+inter-ISP communication accounting and management.
+
+\begin{figure}[ht]
+\centering
+\subfigure[Inefficient overlay topology averaging $\simeq 16$ delay units.] {
+  \includegraphics[scale=0.4]{img/pdf/topology-adaptation-before.pdf}
+  \label{figure:topology-adaptation:before}
+}\qquad\qquad
+\subfigure[Efficient overlay topology after adaptation averaging $\simeq 11$ delay units.] {
+  \includegraphics[scale=0.4]{img/pdf/topology-adaptation-after.pdf}
+  \label{figure:topology-adaptation:after}
+}
+\caption{A simplified example on how overlay topology adaptation may improve matters.}
+\label{figure:topology-adaptation}
+\end{figure}
+%%
 
@@
 \subsubsection{Forwarding Optimization Methodology}
-
-% Researchers initially started considering alternatives for services who
-% implement one-to-many communication schemes, like IP multi-cast, on the
-% application layer (i.e. Narada). Application level multi-cast protocols
-% are generic protocols that can be applied to both P2P file sharing systems,
-% as well as content distribution networks. Application layer multi-cast,
-% however,
-% is not as efficient as the one enabled by IP and still struggles to ease the
-% difficulties imposed by the topology mismatch problem.
-
-%Forwarding optimization can reduce the emitted messages flooding the overlay and
-%as a consequence the overall burden, network resources have to handle. In the
-%first realization of the Gnutella protocol, each query request received by a
-%peer, was then forwarded to all of its neighbors. This was the source of
-%unnecessary, sometimes redundant traffic.
-
-%%AD I think you will need to sprinkle some references of those surveyed 
-%%	in this subsubsection.. As opposed to what done above, 
-%%	this subsubsection has no reference in the text!!!!
 Approaches based on forwarding-optimization 
@@
 algorithm, but they mainly use one or more statistical heuristics. 
-These can range from simple and rather obvious ones, 
-including the latency of the link between the nodes or
-the exposure of specific features such as high processing bandwidth, high
-capacity, or low latency, 
-to more sophisticated alternatives that may even take
-into account application level requirements such as 
-the reliability or friendship.
+These can range from simple connection speed, to exploiting specific peer
+attributes such as high processing capacity e.t.c. An example is \emph{Gia} that
+exploits the so called heterogeneity of peers as well as employs a token-based
+flow control mechanism to guide the messages towards peers that are more likelly
+to answer
+a query. More sophisticated alternatives also exist,
+taking into account application level requirements or exploiting machine
+learning techniques to adjust to overlay network routing, like the \emph{APS}
+algorithm.
 
@@
 Fig.~\ref{figure:forwarding-optimisation:before} 
-% Sub-figure~\ref{figure:forwarding-optimisation:before} 
 shows a protocol that simply floods the entire network 
 in search of object found in the nodes colored in black.
-% for search of an object the existence of which is
-% denoted by black nodes. 
 Alternatively, a node can decide to forward its
@@
 or subset of its outgoing links.
-In Sub-figure~\ref{figure:forwarding-optimisation:after}, the node in the
 In Fig.~\ref{figure:forwarding-optimisation:after}, 
@@
 Moreover, they adress the problem of the mismatch in a limited manner
-% do not actually address the problem of topology mismatch
 since they do not provide any guarantees 
@@
 and try to alleviate.
-% Like all other methodologies, 
 Forwarding based optimisations are commonly applied in
@@
 
-%Expanding the search scope, on the other hand, is no easy task
-%because the overhead of forming multi-cast trees is proportional to the
-%multi-cast group size.
-%forwarding based schemes do not consider dynamic joining
-%and leaving of peers so they do not scale well on dynamic environments.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-%%AD in this portion there is NOT much about replication... 
-%%	ALSO, references to surveyed works as missing.. (other than
-%%	KAzaa!) -  Sprinkle some around the text...
 Caching is widely used to exploit locality and minimize redundant
@@
 already been developed in some commercial \p\ systems, like 
-\emph{KaZaA}, which uses
-caching of indices on its super-peer level. 
+\emph{KaZaA} or less well known like \emph{Gia} and \emph{BNS} as we saw during
+the survey.
 %%%
@@
 helps reduce the burden of network resources, 
-their contribution in addressing the mismatch problem 
-remains limited.
-%%AD I am not sure what the following piece says....
-%%	tried to rephrase as above...
-% Unfortunately, even though the state of the art 
-% in \p\ protocols using caching methods reduce the resource
-% usage of the network, 
-% they also fail to consistently address typical problems
-% of a topologically mismatched environment such as duplication of messages.
-%%%%%%%%%%%%%%%%%%%%%%%
-
-%\subsection{Cache Based}
-%
-%Caching based protocols are effectively used to reduce traffic costs and
-%response
-%times. The caching policy varies depending on the way protocol handles the
+their contribution in addressing the actual mismatch between the overlay and
+and the underlying networks remains limited.
+
+% TODO SOME DISCUSSION
+%The caching policy varies depending on the way protocol handles the
 %index and the
@@
 distance measurement method to position themselves with respect to ``a priori''
-known servers on the Internet. 
+known servers on the Internet, like \emph{Ono} which uses the CDN-infrastructure
+for this purpose.
 These landmark servers are used
@@
 arriving peer, measures its distance to an array of landmark point denoted as
-$Ln$ and and makes a sorted list of the peers, say in increasing order. Then in
+$L_n$ and and makes a sorted list of the peers, say in increasing order. Then in
 order to choose the already participating peers with which to connect to,
@@
 the scalability problem of the static landmark servers is to use ordinary nodes
-as dynamic landmarks once they estimate their own positions. Even though this
+as dynamic landmarks, like in \emph{mOverlay} once they estimate their own positions. Even though this
 approach scales much better than static landmark servers, the measurement
@@
 close distance to each other.
-%%AD ok how about ay advantages?? Any referencess??? 
-%%	Also, a bit of a discussion is missing on hybrid approaches surveyed
-%%	for example which approaches use a combo of these methodologies???
-
-Table~\ref{unstructured:table} offers a summary overview of the algorithms
-we saw earlier in this section and for each algorithm, shows which of the
-aforementioned methodologies it incorporates, along with its 
-highlights and a rough estimation of the
-positive and negative aspects of its implementation.
 

### 2014-01-30 structured.tex
2013.05.26 (Vassilis) Structured algorithms section reworked.

@@
-\section{Structured P2P Networks}
+\section{Structured \p\ Networks}
 \label{section:structured}
@@
 This section presents a description and analysis of algorithms proposed
-to address the topology mismatch problem in structured P2P networks.
-In structured P2P networks, the construction of the routing tables
-held by participating nodes determines the efficiency of message forwarding.
-Routing tables that closely match the underlying IP topology
-achieve increased performance. As a result, the proposed solutions for
-the topology mismatch problem for structured P2P networks 
-use different levels of proximity
-information to optimize the routing tables used by peers.  
-The categorization of the algorithms we present here is based on the
-categorization suggested by previous 
-work~\cite{CDHR2002,CDCR2002,RSS2002}.
+to address the topology mismatch problem in structured \p\ networks. The
+categorization of the approaches we present here is based on how different
+levels of proximity information is gathered and the way peer routing tables are
+maintained and used in an effort make overlay hops as efficient as possible. The
+categorization is based on previous work found in the
+literature~\cite{CDHR2002,CDCR2002,RSS2002}.
+
+%% VM Rephrased as above to keep it analogous to the opening paragraph of
+%% the unstructured part.
+%%
+% This section presents a description and analysis of algorithms proposed
+% to address the topology mismatch problem in structured \p\ networks.
+% In structured \p\ networks, the construction of the routing tables
+% held by participating nodes determines the efficiency of message forwarding.
+% Routing tables that closely match the underlying IP topology
+% achieve increased performance. As a result, the proposed solutions for
+% the topology mismatch problem for structured \p\ networks
+% use different levels of proximity
+% information to optimize the routing tables used by peers.
+% The categorization of the algorithms we present here is based on the
+% categorization suggested by previous
+% work~\cite{CDHR2002,CDCR2002,RSS2002}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Global Soft-State}
+% \subsubsection{Global Soft-State}
 
@@
 
-\textit{Global Soft-State} \cite{XTZ2003} builds a global map to help
+\emph{Global Soft-State} \cite{XTZ2003} builds a global map to help
 choose shorter routing paths, combining the landmark binning method and small
 scale distance probes to reveal the proximity properties of the underlying
-network to the overlay. %Figure \ref{fig:global_softstate} presents the landmark
+network. %Figure \ref{fig:global_softstate} presents the landmark
 %space and the mapping of the position of the nodes.
 This global view of the
-state is made available to all nodes to help them find the best way to
-route their messages. \textit{Global Soft-State} operates in two phases: 
+state, is made available to all nodes to help them find the best way to
+route their messages. \emph{Global Soft-State} operates in two phases:
 generation and use of the proximity information. For generating the proximity
-information a hybrid approach is proposed, which uses landmark clustering as a
+information, a hybrid approach is proposed, which uses landmark clustering as a
 preprocessing step to select a number of potential nearest neighbor
@@
 round-trip-time-based scheme to ultimately choose the closest node. For using
-the proximity information, the algorithm chooses a different path from the
-classic gossiping approaches for constructing and maintaining the overlay. It is
-based on landmark clustering for strategic placement of proximity information on
-the overlay enabling any node to access such information using a landmark number
-that reflects its physical position in the network. For various logical
-regions\footnote{This might be a high-order zone in the eCAN\cite{xu_ecan_2002}
-context or a set of nodes sharing a particular prefix in overlays such as
-Pastry\cite{antony_pastry_2001}.} maps of physical information are built and
-published where each node may appear in a maximum of $log\left( N \right)$ such
-maps. To dynamically adapt to changing network conditions, a node subscribes to
-relevant \emph{soft states} that utilize a notification system in order to
-initialize any necessary neighbor re-selection.
-
-Maintainece of several host states at different layers, renders any content
-migration a costly process. Additionally, the method does not make any
-continuing effort to remap
-the overlay structure after nodes successfully join resulting into poor
-adaptation to changing conditions. Although this approach greatly reduces the
-routing latency to far nodes, it is unable to dynamically identify nodes that
-are close to routers and gateways in order to construct the secondary overlay.
-Nevertheless, static recognition of such nodes is currently done based on BGP
-reports and pre-chosen landmarks, sacrificing the self-organizing attribute of
-traditional DHTs.
+the proximity information, the algorithm follows a different path from that of the
+classic gossiping approaches for the construction and maintenance of the
+overlay. It is based on landmark clustering for generation and strategic
+placement of proximity information on the overlay. This information is called a
+``map'' and concerns a region of the overlay (part of the Cartesian space in
+structures like eCAN\cite{xu_ecan_2002} or a set of nodes sharing a particular
+prefix in overlays such as Pastry\cite{antony_pastry_2001}). Thus, each node not
+only can access such information from these landmarks but can also subscribe to
+relevant \emph{soft states} and get notified on dynamic network changes. A
+node may appear in a maximum of $log\left( N \right)$ such
+maps.
+%%
+Although this approach greatly reduces the routing latency to far nodes, it
+comes expensive. It is possible to end up with a very large number of regions
+each of which has to maintain a map. The fact that a peer may appear in multiple
+such maps we can conclude that additional traffic for probing and notifications
+renders the approach does not scalable. Moreover, to refine measurements
+additional inner-bin measurement s are required\cite{WZS2004}. Also improvement
+is done within a region because of little knowledge about neighboring zones.
+
+%% VM Removed most of the paragraph below which contains info that we cannot back
+%% Rescanned papers and wrote the paragraph below.
+%% DO NOT PERMANENTLY DELETE!
+%%
+% Unfortunately, maintained of several host states at different layers,
+% renders any content migration a costly process. Additionally, the method does
+% not make any continuing effort to remap
+% the overlay structure after nodes successfully join resulting into poor
+% adaptation to changing conditions. Although this approach greatly reduces the
+% routing latency to far nodes, it is unable to dynamically identify nodes that
+% are close to routers and gateways in order to construct the secondary overlay.
+% Nevertheless, static recognition of such nodes is currently done based on BGP
+% reports and pre-chosen landmarks, sacrificing the self-organizing attribute of
+% traditional DHTs.
 
@@
 % Statically implemented sacrificing the decentralized nature of DHTs and
-% thus poorly scaling.
+% thus poorly scaling. Also we can end up with a very large number of regions.
+% Because of the fact that a peer may appear in multiple
+% such maps we can conclude that additional traffic for probing and notifications
+% renders the approach does not scalable.
 low
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Mithos}
-\emph{Mithos} \cite{WR2003} is a P2P protocol which incorporates directed
+% \subsubsection{Mithos}
+\emph{Mithos} \cite{WR2003} is a \p\ protocol which incorporates directed
 incremental probing to find near optimal node placement, and is an integration
@@
 The bootstrap phase of \emph{Mithos} starts with a subset of existing members as
-the first set of candidates, and while the iteratively closest nodes are detected by
+the first set of candidates, and while the closest nodes are detected by iteratively
 probing the neighborhood, the candidate neighbor list is updated. To
-avoid a local minima, Mithos probes all the neighbors within a two-hop distance
+avoid local minima, \emph{Mithos} probes all the neighbors within a two-hop distance
 from the current minimum before concluding the process.
 
-After finding its first neighbor, the newcoming peer is assigned an ID using
-information gathered during the iterative neighbor selection phase. Virtual
-coordinates are assigned to the newcoming peer by using the distances of the two
-closest nodes and their neighbors, so that Euclidean distances between the node
-and all known hosts predict the network latency between them
-\cite{cox_vivaldi_2004}. The benefit of these synthetic coordinates are that
+After finding its first neighbor, the new-coming peer is assigned an ID using
+information gathered during the iterative neighbor selection phase. Based on the
+two closest nodes found, along with these nodes neighbors and the corresponding
+distances a system of virtual springs is set so that the tension of each spring
+is inversely proportional to the distance. The system is then left to settle,
+and the state in which it settles is used for its ID. The benefit of these
+synthetic coordinates are that
 they are explicitly used as the node's ID, and distance computations among nodes
@@
 
-The last step of the algorithm is the interconnection among neighbors. Mithos
+The last step of the algorithm is the interconnection among neighbors. \emph{Mithos}
 uses a \emph{quadrant-based} mechanism in which each node establishes
@@
 
-One major limitation of \textit{Mithos} is that the protocol cannot handle
-dynamic arrivals and departures from the network.
-
-%
-% TODO: review this part
-%
-%In order to avoid local minima during neighbor detection, extensive probing
+One major limitation of \emph{Mithos} is that the protocol cannot efficiently handle
+dynamic arrivals and departures from the network. Moreover, parts of this spring
+relaxation technique requires a centralized
+implementation~\cite{cox_vivaldi_2004}.
+
+%% VM I have read this somewhere but I didn't keep the reference ... :(
+%%
+% In order to avoid local minima during neighbor detection, extensive probing
 % must be undertaken. In simulation, unfortunately, only very small-sized
@@
 % conclusions can be made as for the behavior of an extensively large,
-% real-world p2p deployment of the scheme.
+% real-world \p\ deployment of the scheme.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{LAPTOP}
+% \subsubsection{LAPTOP}
 \emph{LAPTOP} \cite{WLH2007}, organizes the overlay into a tree-based hierarchy
-with the goal of reducing hops during message routing as well as minimizing
+with the goals of reducing hops during message routing as well as minimizing
 maintenance overhead. Additionally, a caching scheme is also incorporated so as
-to further reduce routing table update costs. The authors theoretically show
-that in LAPTOP, routing path length is bounded by $O(log_d N)$ and node joining
+to further reduce routing table update costs. It is theoretically shown
+that in \emph{LAPTOP}, routing path length is bounded by $O(log_d N)$ and node joining
 and leaving in the overlay network is bounded by $O\left( d log_d N \right)$
 hops in a balanced overlay tree, where $N$ is the number of nodes, and $d$ is
-the maximum degree of each node. LAPTOP implements the geographical layout
+the maximum degree of each node. \emph{LAPTOP} implements the geographical layout
 approach  and constructs its layout in a self-organizing and efficient fashion.
@@
 the destination. To improve scalability, the number of child nodes and the
-size of the routing cache are limited. In terms of overlay maintenance, LAPTOP
+size of the routing cache are limited. In terms of overlay maintenance, \emph{LAPTOP}
 incorporates a simple \emph{heartbeat}-based technique where each parent node is
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Proximity in Kademlia}
+% \subsubsection{Proximity in Kademlia}
 % Before discussing {\em Proximity in Kademlia}, it would be helpful to briefly
@@
 % \cite{maymounkov_kademlia_2002} is a distributed hash table (DHT) based
-% protocol for P2P networks, which uses an iteration based lookup algorithm. The
+% protocol for \p\ networks, which uses an iteration based lookup algorithm. The
 % protocol uses the standard $160$ bit ID system for nodes and locates the nodes
@@
 
-\emph{Proximity in Kademlia} \cite{KLKP2008} introduces proximity
+\cite{KLKP2008} introduces proximity
 controls over the vanilla Kademlia protocol's random ID assignment to
-optimize the underlying network usage of the protocol. The work in this paper
-can also be considered as a more generic study of incorporating proximity into
-iterative lookup algorithms. 
-The overlay tier improves its routing
-performance using a cost metric that is provided by the underlying tier
-and quantifies the suitability of establishing a communication link between peers as
-a cost function, based on the used proximity criteria (RTT, ISP locality, etc.).
-%An underlay metric provides information about the underlay network. The paper
-%  distinguishes between three kinds of possibilities to quantify it:
-%\begin{itemize}
-% \item Using measurements gained by previous lookups
-% \item Using measurements acquired by other peers or jointly calculated with
-%  others
-% \item Using a local database to look up information
-%\end{itemize}
-
-The paper proposes two approaches for overlay optimization: 
-{\em proximity routing} and {\em proximity neighbor selection}.
-In proximity routing, the goal
-is to choose the best next hop during routing a message. Due
-to the fact that the routing in Kademlia is iterative, it is the initiator of
-the lookup that chooses each next hop from a set of candidate nodes. Standard
-Kademlia always chooses the closest node with respect to the XOR metric but with
-proximity routing selection, Kademlia chooses the node with the smallest underlay
-metric cost. Thus, as in all such approaches, there is a trade-off between the
-overlay and underlay distances.
-
-Towards the enhancement of traffic locality, the authors also use information
-from the MaxMind GeoIP database\footnote{MaxMind Geolocation Technology.
-http://www.maxmind.com} to retrieve proximity information of a given peer based
-on its region, country, or ISP location, and use this information to form
-clusters of nearby peers. An alternative method could use Vivaldi
-coordinates \cite{cox_vivaldi_2004}. However, the authors report that their clustering
-approach performs better than the Vivaldi protocol based on their experiments.
-Authors also report that {\em Proximity routing} protocol successfully worked
-with the Kademlia protocol and improved the locality of the connections between
-peers.
-
-In \emph{proximity neighbor selection}, the goal is to
-keep peers with the least contact cost in the routing table. As Kademlia
-continuously learns about new peers via incoming requests and an iterative
-lookup process, no 
-special algorithm for searching more cost-effective peers is necessary and
-it simply chooses the best peers seen so far.
+optimize the underlying network usage.
+% The work in this paper
+% can also be considered as a more generic study of incorporating proximity into
+% iterative lookup algorithms.
+The overlay tier improves its routing performance using a cost metric that is
+provided by the underlying tier. Such information can be acquired by
+\begin{inparaenum}[\itshape i\upshape)]
+  \item measuring previous lookups,
+  \item exchanging measurements of other peers or jointly calculating them with
+  others, or
+  \item using a local database to look up information.
+\end{inparaenum}
+
+Regarding routing optimization both techniques of \emph{proximity routing
+selection (PRS)} and \emph{proximity neighbor selection (PNS)} are studied.
+%%
+In \emph{PRS}, the goal is to choose the best next hop during routing a message.
+% Due to the fact that the routing in Kademlia is iterative, it is the initiator
+% of the lookup that chooses each next hop from a set of candidate nodes.
+Plain old Kademlia (\emph{POK}) always chooses the closest node with respect to
+the XOR metric but with proximity routing, Kademlia should also choose the node
+that exhibits the smallest underlay metric cost, balancing a trade-off between
+the overlay and underlay progress distances.
+%%
+In \emph{PNS}, on the other hand, the goal is to keep peers with the least
+contact cost in the routing table. As Kademlia continuously learns about new
+peers via incoming requests and an iterative lookup process, no special
+algorithm for searching more cost-effective peers is necessary and it simply
+chooses the best peers seen so far.
+
+Proximity injection in Kademlia improves locality of connections between peers.
+Towards the enhancement of traffic locality, it is suggested to use information
+from the MaxMind GeoIP database\footnote{http://www.maxmind.com} and use it to
+form clusters of nearby peers. An other alternative method could use Vivaldi
+coordinates \cite{cox_vivaldi_2004}, however, it is suggested that the
+clustering approach performs better. Specifically when combined with both
+\emph{PRS} and \emph{PNS} preserves more than $40\%$ of the messages within an
+ISP and allows only a small $6\%$ to go through intercontinental backbones.
 
@@
 % approach in dynamic systems
-medium &
-% 
-medium &
-%
+% PNS reduces lookup latencies by 66\% compared to POK
+medium &
+% 
+medium &
+% MaxMind is commercial and cannot be used by FOSS P2P systems. Trying to
+% expose information from P2P client can result in security issues.
 medium
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{CHOord considering Proximity on IPv6 (CHOP6)}
+% \subsubsection{CHOord considering Proximity on IPv6 (CHOP6)}
 \emph{CHOP6} \cite{MT2007} roughly estimates the proximity among
-nodes by exploiting the IPv6 address format and RTT information if available.
+nodes by exploiting the IPv6 address format and RTT information, if available.
 The proximity estimation is achieved by introducing a 64-bit ID scheme in which
@@
 low &
-%
+% Its static nature imposes low overhead to the network
 low &
-%
+% IPv6 is still not widely applied and supported.
 medium
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{T2MC}
-
-T2MC \cite{SLCGZ2008} uses traceroute logs to detect the AS boundaries and
+% \subsubsection{T2MC}
+\emph{T2MC} \cite{SLCGZ2008} uses traceroute logs to detect the AS boundaries and
 cluster nearby peers together to reduce the redundant multiple message
-passes between the AS boundaries. T2MC uses a customized k-mean classification
-algorithm with k = 2 to perform the classification, and exploits the stable
-structure of the internet routers to guide clustering. Even though traceroute
-provides detailed information about the network structure, use of traceroute
+passes between the AS boundaries. \emph{T2MC} uses a customized k-mean classification
+algorithm with $k = 2$ to perform the classification, and exploits the stable
+structure of the Internet routers to guide clustering. Even though trace-route
+provides detailed information about the network structure, use of trace-route
 creates overhead on the overall network structure. For this same reason, it is
 not unusual for network administrators to disable this support on their network
-routers, which may affect the performance of the T2MC algorithm.
+routers, which may affect the performance of the \emph{T2MC} algorithm.
+
+The results of T2MC are encouraging but it is also considered coarse-grained
+since it classifies peers into only ``remote'' and ``close''
+groups \cite{QLZG2009}.
 
@@
 %router, and registers it along with all near clusters into the DHT of the
-%P2P overlay. As new peers join the network, those that share the same edge
+%\p\ overlay. As new peers join the network, those that share the same edge
 %router or any of the members of the near router clusters, would form a close
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{PChord}
-\emph{PChord}\cite{HLYW2005} is based on the Chord DHT and adds
-proximity routing into its routing mechanism. The main modification over the
-vanilla Chord is the inclusion of a \emph{proximity list} into its routing table
+% \subsubsection{PChord}
+\emph{PChord} \cite{HLYW2005} is based on the Chord DHT and adds
+proximity into its routing mechanism. The main modification over the
+vanilla Chord, is the inclusion of a \emph{proximity list} into its routing table
 so that the next hop is decided based on both the best progress
-towards the key and the physical proximity of the candidate nodes as well.
-The list is, at join time, empty but as the PChord node starts to interact with
+towards the key and the physical proximity of the candidate nodes.
+The list is, at join time, empty but as the \emph{PChord} node starts to interact with
 other nodes it applies a heuristic mechanism to fill the list. Entries are
 dynamically added or removed as the network state changes. The overlay
-maintenance cost can be considered similar to Chord's. Besides node join/leave
-and cost of heartbeat messages for connection verification, PChord adds
-heartbeat of nodes inside the proximity list but since this list always contains
-nodes with the lowest RTT the added cost is kept to a minimum.
+maintenance cost is similar to Chord's. Apart from the costs incurred by node
+join/leave and by heart-beat messages for connection verification, \emph{PChord} adds
+heart-beat among nodes inside the proximity list. Fortunately, since the list
+contains nodes with the lowest RTT, the added cost is kept to a minimum.
 
 Due to the fact that, with each step, there is always progress towards the
-target node in the ID space, PChord will result in a reduced hop number than
-Chord, since each hop in PChord is larger or at least equal in the key space.
-Passing through proximity links in the underlying network means reduced routing
+target node in the ID space, \emph{PChord} will result in a reduced hop number than
+Chord, since each hop in \emph{PChord} is larger or at least equal in the key space.
+Passing through proximity links, in the underlying network, means reduced routing
 cost. Additionally, if the number of entries in the proximity list is the same
-as the number of network partitions, PChord prevents hops from jumping in and
+as the number of network partitions, \emph{PChord} prevents hops from jumping in and
 out of the network partition the current node belongs to.
 
-% TODO: SOME DISCUSSION
-% TODO: NEEDS REWORDING
-% PChord’s main routing optimizations are of less overlay hops, passing
-% proximity links of the underly network and passing the same network partition
-% only once in the routing path, which directly results in the lower RDP of
-% routing efficiency. Meanwhile, PChord keeps light-weight maintenance cost as
-% Chord does.
+\emph{PChord}’s advantages include the fact that the main routing optimizations are of
+less overlay hops, following proximity links and crossing the same network
+partition only once along a routing path while in the same time keeping Chord's
+light-weight maintenance cost. On the other hand, it exhibits slow convergence
+speed and inefficiency in a high churn environment\cite{DK2006}.
 
@@
 \hline
-% Small decrease in needed overlay hops per object publishing
+% Small decrease in needed overlay hops per object publishing but rigid in
+% high churn environments.
 medium &
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{AChord}
-\emph{AChord} \cite{DK2006} uses IPv6's anycast mechanism to
+% \subsubsection{AChord}
+\emph{AChord} \cite{DK2006} uses IPv6's any-cast mechanism to
 relieve the protocol from complex joining procedures, and achieve high accuracy
-of network proximity and thus, high routing efficiency. Anycast delivers a message
-coming from the outside of an anycast group to the physically closest node in
-that group. AChord organizes all nodes participating in the overlay network into
-an anycast group. Any joining node comes outside of that group and, thus, is
+of network proximity and thus, high routing efficiency. Any-cast delivers a message
+coming from the outside of an any-cast group to the physically closest node in
+that group \cite{M2002}. \emph{AChord} organizes all nodes participating in the
+overlay network into an any-cast group. Any joining node comes outside of that group
+and, thus, is
 automatically forwarded to the physically nearest node in order to bootstrap,
 avoiding the need to, explicitly, maintain such nodes, and the effort of finding
-a way of locating the physically nearest among them. The ID of the newcoming
+a way of locating the physically nearest among them. The ID of the new-coming
 node is computed based on the ID of the bootstrap node and the bootstrap's
@@
 called \emph{adjacency table} which stores information about the closest known
-node. The routing decision is made by using both the adjacency and the finger
-table. The authors claim this to be a lightweight approach that requires few
-changes to Chord while keeping its advantageous characteristics.
-
-% TODO: MAYBE REVIEW THE EVALUATION
+node. The routing decision is made by using both the adjacency and finger
+tables.
+
 \begin{center}
@@
 \hline
-% good routing efficiency
+% good routing efficiency but no adaptation to high churn environments.
 medium &
@@
 medium &
-% assumes the operation of ideal underlying anycast mechanism (meaning that
+% assumes the operation of ideal underlying any-cast mechanism (meaning that
 % outside messages are always routed to the nearest node which in the current
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Chord6}
+% \subsubsection{Chord6}
 \emph{Chord6} \cite{XZHL2005} is another Chord variant that tries to
@@
 reduces inter-domain traffic between service providers. The difference between
-Chord6 and Chord relies on the identifier definition. Therefore, the approach
-can be easily portable to other DHTs such as CAN, Pastry and Tapestry. In Chord6
+\emph{Chord6} and Chord relies on the identifier definition. Therefore, the approach
+can be easily portable to other DHTs such as CAN, Pastry and Tapestry. In \emph{Chord6}
 the identifier contains two parts: the higher bits are obtained by hashing the
@@
 \hline
-% Can reduce intra-domain average path length but even though nodes in the same
+% Can reduce inter-domain average path length but even though nodes in the same
 % domain would have close identifiers, nodes in two close domains may have very
 % different ones reducing the overall efficiency of the approach in general when
-% considering inter-domain average path length.
+% considering inter-domain average path length. No active adaptation mechanism
+% for high churn environments.
 low &
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{DHT-PNS}
+% \subsubsection{DHT-PNS}
 The work in \cite{DLTZZ2006} propose a proximity neighbor selection
@@
 regions and towards an upper layer identifier space until one is found or the
-first layer reached.
-
-% TODO: MAYBE REVIEW THE EVALUATION
+first layer reached. Simulations suggest that the scheme has good convergence
+property of the delay distribution and good optimizing performance.
+Unfortunately, even though scaling seems to be also smooth, the simulations
+where made on rather small test beds (of at most 2048 nodes), so no safe
+conclusions can be deduced. Moreover it assumes uniform node distribution which
+is not always the case, especially in a synthetic coordinate system.\\
+
 \begin{center}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Quasi-Chord}
-The approach that is proposed by Sun and Zhang in \cite{SZ2008}, tries
-to address the topology mismatch problem, at the construction phase of the
-Chord identifier cycle, on the basis of underlying topology information.
-The proposed model constructs the \emph{Quasi-Chord} overlay network in three
-steps. First, each host acquires its coordinates in the geometric space
-utilizing the \emph{global network position (GNP)} protocol \cite{ng_gnp_2001}.
+% \subsubsection{Quasi-Chord}
+In \cite{SZ2008} proposes \emph{Quasi-Chord}, a Chord-based network that
+constructs its overlay network with underlying topology in mind. The network
+construction follows three steps. First, each host acquires its coordinates in
+the geometric space utilizing the \emph{global network position (GNP)} protocol
+\cite{ng_gnp_2001} and a set of landmarks.
 Second, using the \emph{Cantor space filling curve}, the $2$-dimensional space
 is converted to a $1$-dimensional one, used in the last, third, step to build
-the Quasi-Chord circle according to the host's Cantor value. In Quasi-Chord
+the \emph{Quasi-Chord} circle according to the host's Cantor value. Physical
+proximity is thus denoted by Cantor value proximity. The Chord cycle is
+constructed by sorting nodes in ascending order. In \emph{Quasi-Chord}
 each node stores 2 finger tables; one clockwise and one counter-clockwise.
@@
 %First of all the host must be positioned in the geometric space. The algorithm
-%  models the P2P network with a well defined distance function in such a way,
+%  models the \p\ network with a well defined distance function in such a way,
 % it can predict, with high accuracy, the distance between any two points in the
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{IP-based Clustering (IPBC)}
+% \subsubsection{IP-based Clustering (IPBC)}
 Proximity neighbor selection algorithms use probing and other techniques to
@@
 overload the network. \emph{IP-based clustering} \cite{KM2007} is a proximity
-neighbor selection based algorithm in which the authors propose to use the IP
-address prefixes (16 bit for IPv4) in order to detect proximity.
-\cite{freedman_iploc_2005} states that $97\%$ of prefixes larger than $24$ bits
-belong to a single geographical location. However, using a smaller number of bits
-creates less precise results and a larger number of bits increases the burden on the
-network and reduces the possible number of neighbors. Therefore, a careful
-selection is required in terms of performance/accuracy trade-off. IP-based
-clustering generates a key and its prefix and stores the prefix in the DHT
-itself, so that any newly joining nodes with the same IP prefix can query it and
+neighbor selection based algorithm which uses the IP address prefixes (16 bit
+for IPv4) in order to detect proximity. \emph{IPBC} clustering generates a key
+and its prefix and stores the prefix in the DHT itself, so that any newly
+joining nodes with the same IP prefix can query it and
 identify all the neighbors with the same prefix easily.  Nodes
 periodically update their entries in the DHT by removing those who leave, or
-those who have been timed out with no further activity detected from the node.
+those who have been timed-out.
 
@@
 % published by many peers when choosing to hash a short one. Their verdict is
-for
+% for
 % the use of a 16-bit wide prefix for real world systems deployed in an Internet
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Cone}
+% \subsubsection{Cone}
 \emph{Cone} \cite{HY2007} extends Chord using a proximity neighbor
@@
 generated using landmarks and round-trip-time-based distances to landmarks.
-Cone uses a two-layered identifier space. The first, named Chord-layer
+\emph{Cone} uses a two-layered identifier space. The first, named Chord-layer
 identifier, denoted as $Id_{Chord}$, is the same as in Chord. The second is the
@@
 third table called \emph{group} maintains information about other on-line peers
-within the current node's group in a way that entries are now close in the ID
+within the current node's group in a way that entries are now close in the id
 space.
@@
 % to exploit proximity information thus incurring low overall additional
-% operational overhead but imposing extra stress to the landmarks themselves 
+% operational overhead but imposing extra stress to the landmarks themselves.
+% Additional information must be maintained, three finger tables (front, back and
+% group)
 medium &
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{DynaMo}
+% \subsubsection{DynaMO}
 \emph{DynaMO} \cite{WZS2004} tries to build an overlay that not only 
 considers physical proximity amongst peers but their mobility attributes as well.
-DynaMO is based on the Pastry overlay
-because of Pastry's built-in locality
-heuristics which are thoroughly analyzed in the literature \cite{CDCR2002a}.  
-To adapt to mobile, ad-hoc networks, though, special care has been put in 
-the maintenance of an even overlay ID distribution so that hotspots are
-avoided. DynaMO requires a newcoming node to gather information concerning its
-physical neighborhood and uses it to assign itself an appropriate overlay ID.
-DynaMO tries to capitalize on the fact that in Pastry each node's table consists
-of rows equal to the number of digits in the overlay ID and columns equal to
-the ID's base. As we go down the table rows, the matching prefix between the
-current node's ID and the row's entries increases by one. Thus, the leaf set
-contains the closest nodes in the ID space. Additionally, as the prefix match
-increases by one, the result is exponentially fewer candidates that can fill the
-tables entries as the row increases. This leads to the observation
-\cite{antony_pastry_2001,CDCR2002a} that from overlay routing hop to overlay
-routing hop, the physical distance between nodes is likely to increase leading to
-a dominating last routing step in terms of the overall physical routing path
-distance traveled during a key lookup. Since the last routing step is usually
-taken from the leaf set, DynaMO focuses on making this last step as physically
-close as possible. In this context, two approaches are considered namely
-\emph{Random Landmarking (RLM)} and \emph{Closest Neighbor Prefix Assignment
+To adapt to mobile, ad-hoc networks, special care is put in by \emph{DynaMO}
+to maintain an evenly distributed overlay ID space so that hot-spots are
+avoided. The paper's implementation is based on the Pastry overlay due to its
+built-in locality heuristics that have been extensively studied
+\cite{CDCR2002a} but \emph{DynaMO}'s mechanisms are DHT-independent so it can be
+applied to other DHT's.
+
+\emph{DynaMO}'s approach can be viewed bottom-up. A new-coming node is required
+to compute physical information about its neighborhood and use it to
+unilaterally assign itself the an ID. \emph{DynaMO} is building on the
+observation that with every new overlay routing hop the physical distance
+between nodes is likely to increase, resulting in a dominating last routing
+step \cite{antony_pastry_2001,CDCR2002a} and for this two methods are discussed,
+\emph{random landmarking (RLM)} and \emph{closest neighbor prefix assignment
 (CNPA)}.
-
-The RLM lookup mechanism is used to locate nodes responsible for a fixed set of carefully
-chosen (in a way that the ID space is divided into equal portions) landmark
-keys. This means that when a node is assigned a landmark key, it plays the
-role of a temporary landmark for the network. Any joining node will measure its
-distance to this landmark node and assign to itself an ID consisting
-of a prefix taken from the closest landmark and a remainder that could be
-assigned randomly or generated by a mechanism that takes into account physical
-neighborhood. The length of the prefix can be determined as
-$prefix_length=|log_b k|$, where $b$ is the ID base and $k$ the number of
-landmarks. The above scheme results in physically close nodes, forming regions
-of common ID prefix that are likely to be close to each other in the ID space
-and thus bringing a node's leaf set closer to itself. Additionally, the
-advantage of dynamic landmarking is that in the case of failed landmark servers
-they can instantly be replaced by new substitutes, that share similar physical
-attributes with the failed and thus preserve the balance of the network.
-
-RLM may incur more traffic especially on landmark nodes, load that in some cases
-is not acceptable. CPNA on the other hand takes advantage of Pastry's
-specification that a newcoming node is bootstrapped by a physically close
-node.
-The newcomer then assumes the ID prefix of that neighbor while the rest of the ID is 
-generated similarly to RLM. Unfortunately, less overhead comes at the expense
-of being more coarse-grained.
+%%
+In \emph{RLM} a set of peers are chosen to be responsible for a set of landmark
+keys. These keys have to be chosen in a way that they equally divide the overlay
+space. The new-coming peer then computes distances to those
+landmarks, sorts them and assigns itself an ID based on the landmark ordering.
+In more detail, the peer adopts an ID prefix of length
+$prefix_length=|log_b k|$ (where $b$ is the ID base and $k$ the number of
+landmarks) from the landmark closest it and the rest can be assigned either
+randomly or computed with further proximity optimization in mind.
+What is achieved is that
+\begin{inparaenum}[\itshape i\upshape)]
+  \item landmark assignment is dynamic, and
+  \item a node’s leaf set is likely to reference physically close nodes.
+\end{inparaenum}
+Especially, the latter observation reduces the average routing cost because the
+leaf set is utilized for the last routing step (since it contains the nodes
+closest in the overlay space).
+%%
+\emph{CNPA} is proposed for network setups that want the bootstrap phase as lightweight
+as possible. It delegates to Pastry, the burden to identify the physically
+close node with which the new-coming peer will bootstrap. The new-comer assumes
+the bootstrapping nodes's ID prefix and fills the rest like \emph{RLM} does.
+
+\emph{RLM} achieves good proximity evaluation in a load balanced and adaptable to
+high churn environments fashion. Unfortunately it comes with more bootstrap
+overhead that can especially effect the set of temporary landmarks. On the other
+hand, \emph{CNPA}'s less incurred overhead comes at the expense of being more
+coarse-grained.
+
+% \emph{DynaMO} tries to capitalize on the fact that in Pastry each node's table consists
+% of rows equal to the number of digits in the overlay ID and columns equal to
+% the ID's base. As we go down the table rows, the matching prefix between the
+% current node's ID and the row's entries increases by one. Thus, the leaf set
+% contains the closest nodes in the ID space. Additionally, as the prefix match
+% increases by one, the result is exponentially fewer candidates that can fill the
+% tables entries as the row increases. This leads to the observation
+%  In this context, two approaches are considered namely
+% \emph{Random Landmarking (RLM)} and \emph{Closest Neighbor Prefix Assignment
+% (CNPA)}.
 
@@
 \begin{center}
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
-%
-medium &
-% The paper suggests that this approach generates significantly lower overhead
-% than the vanilla Pastry algorithm
-low &
+{\footnotesize
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{RLM} &
+%
+medium &
+%
+medium &
 % Dynamic landmarking approach helps the scalability of the algorithm. The
@@
 % degression rates.
-high
-\end{tabular}
-\end{center}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{SAT-Match: Self-Adaptive Topology Matching}
-
+high \\
+\emph{CNPA} &
+%
+medium &
+% % The paper suggests that this approach generates significantly lower overhead
+% than the vanilla Pastry algorithm
+low &
+% The queries are directed through fast nodes so it is important such nodes
+% be part of the network
+medium \\
+\end{tabular}
+}
+\end{center}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \subsubsection{SAT-Match: Self-Adaptive Topology Matching}
 % \begin{figure}[ht]
@@
 
-\emph{SAT-Match} \cite{RGJZ2004}, similar to the other approaches in proximity
-neighbor selection, uses probing to detect nearby peers and obtain proximity
-information. However, as an additional feature, SAT-Match peers can do selective
-jumps to adjust their positioning in the DHT, if it reduces the \emph{stretch}
-of its one-hop neighborhood. The paper defines stretch 
+\emph{SAT-Match} \cite{RGJZ2004} defines an iterative process that attempts to
+adaptively change the overlay network to ease the topology mismatch problem. It
+consists of a \emph{probing phase} when a peer probes nearby nodes for distances
+and a \emph{jumping phase} when the peer decides to change neighborhood to
+reduce its average neighbor distances. This process is performed unilaterally
+until no additional optimization is needed but it achieves optimization within a
+sufficiently large scope.
+%% VM This is used in other papers. Maybe we can put somewhere in the
+% definition of Topology Mismatch Problem in the intro/background.
+The paper defines \emph{stretch}
 $S = \frac{\bar{L_l}}{\bar{L_p}}$, where $\bar{L_l}$ is the average logical link
 latency while the $\bar{L_p}$ is the average physical link latency, and uses it
-a way to quantify the \emph{topology match degree} of the constructed overlay. For
-the probing phase SAT-Match uses a small TTL value for the probing messages in
-order to reduce redundancy \cite{jiang_lightflood_2008}. This process begins as
-soon as the node joins the network using a DHT mechanism. Each probing message
-contains information about the source and a small TTL value. The recipient of
-such a message returns information about itself to the source and forwards the
-probing message to its neighbors if the TTL is non-zero. The discovered nodes
-are referred to as $TTL-k$ neighborhood of the source node based on the $TTL$
-distance to the source node. Blindly selecting the peer with the smallest RTT
-as neighbor is, generally, not the optimal decision to make to achieve
-global stretch reduction. In a structured scheme, when a node jumps to connect
-to a physically close node, it may need to connect to other distant nodes to
-maintain the structure's integrity, thus creating an overall increase in the
-overlay's stretch. The two nodes with the smallest RTT are then used to
-select one zone to jump in this phase. The algorithm is as follows: The source
-node $S$ calculates the stretch change of its $TTL-1$ neighborhood and that of
-the $TTL-1$ neighborhood of the first of the previously selected peers. These
-calculations are made as if the jump has been made. If the stretch reduction is
-over a predefined threshold the jump is performed, otherwise the second selected
-candidate is picked and the same computations are performed. If again, the
-threshold is not met, then no jump is ultimately done. In case of a jump, this
-is performed as a combination of \emph{leave} and \emph{join} operations, in the
-CAN context. Even though no theoretical proof is given, valid results are
-obtained and demonstrated through experimentations.
+a way to quantify the \emph{topology match degree} of the constructed overlay.
+
+The \emph{probing phase} starts when the new-comer joins the DHT using the
+standard mechanism of the algorithm at hand. The starts probing the neighborhood
+the messages of small TTL value $k$ \cite{jiang_lightflood_2008}. Recipients return
+information to the source and forward the message if not expired. After
+discovering its TTL-$k$ neighborhood, the source measures RTT times, sorts them
+and picks the two nodes with the smallest RTTs.
+
+In the \emph{jumping phase}, the source node calculates the stretch change of
+its $TTL-1$ neighborhood and that of the $TTL-1$ neighborhood of the first of
+the peers picked during the previous phase. These calculations are made as if
+the jump has been made. If the stretch reduction is over a predefined threshold
+the jump is performed, otherwise the second selected candidate is picked and the
+same computations are performed. If again, the threshold is not met, then no
+jump is ultimately done. In case of a jump, this is performed as a combination
+of \emph{leave} and \emph{join} operations, in the underlying context
+(i.e. CAN).
+
+Even though no theoretical proof is given, valid results are
+obtained and demonstrated through experimentations. It is argued that since it
+is observed to achieve a $27\%$ stretch reduction when compared to a landmark
+enhanced CAN-system, it can be deduced that it performs better than landmark
+binning in terms of the matching degree.
+
 % Figure~\ref{figure:sat_match} demonstrates the selective jump on an
@@
 % secures the fast adaptation to frequent network changes. It also considered
-% lightweight and can easily be embedded into current p2p systems, as well
+% lightweight and can easily be embedded into current \p\ systems, as well
 % as effectively combined with other techniques, such as landmark binning.
@@
 % The self-adaptation nature of the algorithm and the selective jump mechanisms
-% scale much better than, for example, Mithos.
+% scale much better than, for example, Mithos. The test are on a relatively
+% small amount of peers, thus, no safe conclusion can be made with respect to the
+% algorithm's scalling performance.
 medium
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Structured P2P network algorithms use a global distributed hash table or a
-prefix tree structure to uniquely lookup peers or their data in the overlay
-network. As all the data is kept within the overlay, each node behaves as a
-client and a server, therefore nodes join and leave according to rules
-determined by the integrity of the global data structure. The main advantage of
-the structured P2P topology is that with the help of the global data structure,
-peers or their data can be found within the network even if there is only a
-single copy of that item present. However, each node join and leave creates
-maintenance overhead for the network due to updates required by the global data
-structure, and for networks with frequent node arrivals and departures, the
-overlay uses valuable network resources just to update the global structure.
-Nodes join the network by using a key value, which determines the location and
-the neighborhood of the new node within the network. However, assigning
-random key values to the newly inserted nodes creates non-optimal matching with
-the underlying physical network topology, therefore, increasing the overhead of
-the network even more. One solution for handling the topology mismatch problem
-is to consider the proximity of the peers when generating the key and joining
-the node to the network, so that nodes within the same network domains are
-selected as peers, or neighbors, during the overlay topology construction.
+
+%% VM I commented out the following paragraph since it feels out of place
+% (maybe better in the intro or background) or duplicated (from the intro or
+% background).
+%% 
+% Structured \p\ network algorithms use a global distributed hash table or a
+% prefix tree structure to uniquely lookup peers or their data in the overlay
+% network. As all the data is kept within the overlay, each node behaves as a
+% client and a server, therefore nodes join and leave according to rules
+% determined by the integrity of the global data structure. The main advantage of
+% the structured \p\ topology is that with the help of the global data structure,
+% peers or their data can be found within the network even if there is only a
+% single copy of that item present. However, each node join and leave creates
+% maintenance overhead for the network due to updates required by the global data
+% structure, and for networks with frequent node arrivals and departures, the
+% overlay uses valuable network resources just to update the global structure.
+% Nodes join the network by using a key value, which determines the location and
+% the neighborhood of the new node within the network. However, assigning
+% random key values to the newly inserted nodes creates non-optimal matching with
+% the underlying physical network topology, therefore, increasing the overhead of
+% the network even more. One solution for handling the topology mismatch problem
+% is to consider the proximity of the peers when generating the key and joining
+% the node to the network, so that nodes within the same network domains are
+% selected as peers, or neighbors, during the overlay topology construction.
 
@@
 \cite{CDHR2002,CDCR2002,RSS2002}:
-\begin{inparaenum}[\itshape i\upshape)]
+\begin{enumerate}[\itshape i\upshape)]
   \item geographic-layout-based,
@@
   \item proximity neighbor selection based,
-\end{inparaenum}
+\end{enumerate}
+
+%%VM Added the following to mention the hybrid/combo methodologies
+Similarly to unstructured approaches, most of the reviewed structured ones,
+incorporate multiple of the above methodologies. For example, \cite{KLKP2008}
+uses both proximity routing and proximity neighbor selection and achieves
+topology adaptation, or \emph{LAPTOP} combines geographic layout with
+replication strategies to refine the end results. The heuristic nature
+of the proposed solutions makes it a game of trade-offs, that in most cases
+is highly affected by the application domain and the implementation itself.
+Thus, the reader is advised to filter the results of each approach and try to
+think about how the points presented will work in her own particular problem.
 
@@
 key properties of each algorithm that differentiate it from the others and
-highlight 
-each algorithm's advantages and disadvantages.
-
-\input{structured-table}
-
-\input{structured-plot}
+highlight each algorithm's advantages and disadvantages.
 
@@
 
-In the geographic layout category fall methods which practically
+In the \emph{geographic layout (GL)} category fall methods which practically
 position physically nearby nodes together in the application space as well
@@
 proximity information that describe the overall geographic positioning of the
-peer nodes. Landmark servers and RTT measurements are two popular methods, which
+peer nodes. Some of the algorithms we saw were \emph{Global Soft-State},
+\emph{Mithos} and \emph{LAPTOP} which all try clustering close by nodes in
+order to reduce the network's average delay.
+Landmark servers and RTT measurements are two popular methods, which
 can also be used in conjunction, to discover physically nearby peers over the
-network. However, as discussed in Section~\ref{sec:landmark_binning}, these
+network. However, as already discussed, these
 methods do not always give reliable estimates for the node positions over the
 Internet. The landmark servers are not self-organizing and have maintenance
-overheads. To serve a P2P network with millions of peers, multiple landmark
+overheads. To serve a \p\ network with millions of peers, multiple landmark
 servers distributed over the whole world is required, which is hard to manage.
@@
 
-Even though managing the overlay structure based on the geographic layout of the
+Even though managing the overlay structure based on the \emph{GL} of the
 nodes improves the query efficiency of the system, on the other hand, it tends
-to create hotspots, and the needed failure resilience is undermined by the fact
-that neaerby nodes are more likely to suffer collective failures.
+to create hot-spots, and the needed failure resilience is undermined by the fact
+that nearby nodes are more likely to suffer collective failures.
 
@@
 
-Proximity routing does not require routing tables to be built using any
+\emph{Proximity routing (PR)} does not require routing tables to be built using any
 knowledge about network proximity. On the other hand it exploits such knowledge
 to choose the best next hop when routing a message as can be seen in
-Figure~\ref{figure:proximity-routing}. This approach tries to balance between
-choosing the node that will further progress the routing towards the destination
-and choosing the closest entry in the routing table, in terms of network
-proximity. As this is very difficult to achieve, the system might end up
+Figure~\ref{figure:proximity-routing}. We visited an array of methodologies that
+implemented this method, either on its own like in \emph{Mithos} and
+\emph{PChord} or in conjunction with proximity neighbor selection like in
+\emph{DynaMO}. \emph{PR} tries to balance between choosing the node that will
+further progress the routing towards the destination and choosing the closest
+entry in the routing table, in terms of network proximity. As this is very
+difficult to achieve, the system might end up
 using a greedy approach that selects low latency paths on the overlay, which
@@
 routing, however, joint uses of these two protocols are also possible. Moreover,
-they are considered relatively less effective than geographical layout when
-applied to CAN(-like) implementations. Moreover, the technique has been
-incorporated into a version of Chord causing an increase in the overhead of node
-joins and the size as well as maintenance cost of finger tables.
+they are less effective than geographical layout when applied to CAN(-like)
+implementations.
+
+% Moreover, the technique has been
+% incorporated into a version of Chord causing an increase in the overhead of node
+% joins and the size as well as maintenance cost of finger tables.
 
@@
 
-\emph{Proximity Neighbor Selection} constructs the routing tables using
+\emph{Proximity Neighbor Selection (PNS)} constructs the routing tables using
 proximity knowledge (see Figure~\ref{figure:proximity-neighbour-selection}). The
@@
 landmark-based systems described in Section~\ref{section:geographic_layout}.
-Here time-to-live values between nodes, or directly node ID prefixes are used to
-detect proximity. Tapestry, Pastry, and CAN successfully incorporated 
+Here time-to-live values between nodes (i.e. \emph{SAT-Match}) or IP address
+prefixes (i.e. \emph{DynaMO} or \emph{Cone}) are used to detect
+proximity. \cite{freedman_iploc_2005} states that $97\%$ of prefixes larger
+than $24$ bits belong to a single geographical location. However, using a
+smaller number of bits creates less precise results and a larger number of bits
+increases the burden on the network and reduces the possible number of
+neighbors. Moreover, in \cite{HLYW2005} it is argued that such approaches can
+destroy the uniform distribution in the key space and do not work in
+one-dimensional key space where the mapping is constrained on the overlay layer.
+Therefore, a careful selection is required in terms of
+performance/accuracy trade-off. Tapestry, Pastry, and CAN successfully incorporated
 proximity neighbor selection into their algorithms. The routing protocol in
@@
 %terms of the method and the year??}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\input{structured-table}
+
+\input{structured-plot}

### 2014-02-02 background.tex
2013.07.28 (Vassilis) Further investigation conducted to find research work that references the papers/algorithms we survey and cite. Text enriched.

@@
 %
-%\cite{matei_mapgnutella_2002, lv_randomwalks_2002, merugu_str2unstr_2003}
+%\cite{matei_mapgnutella_2002, LCCLS2002, merugu_str2unstr_2003}
 %showed that the ad-hoc network topology of unstructured overlay networks that
@@
 \emph{TIV}s affect both network coordinate
-\cite{cox_vivaldi_2004,wong_meridian_2005} and positioning \cite{ng_gnp_2001}
+\cite{cox_vivaldi_2004,WSS2005} and positioning \cite{ng_gnp_2001}
 systems and makes the construction of \emph{IP}-topology-aware 

### 2014-02-02 structured.tex
2013.07.28 (Vassilis) Further investigation conducted to find research work that references the papers/algorithms we survey and cite. Text enriched.

@@
 such maps we can conclude that additional traffic for probing and notifications
-renders the approach does not scalable. Moreover, to refine measurements
+renders the approach not scalable \cite{RGJZ2004}. Moreover, to refine measurements
 additional inner-bin measurement s are required\cite{WZS2004}. Also improvement
@@
 
-One major limitation of \emph{Mithos} is that the protocol cannot efficiently handle
-dynamic arrivals and departures from the network. Moreover, parts of this spring
-relaxation technique requires a centralized
-implementation~\cite{cox_vivaldi_2004}.
-
-%% VM I have read this somewhere but I didn't keep the reference ... :(
-%%
-% In order to avoid local minima during neighbor detection, extensive probing
-% must be undertaken. In simulation, unfortunately, only very small-sized
-% overlay topologies (of 200 to 1000 nodes) have been used and thus no safe
-% conclusions can be made as for the behavior of an extensively large,
-% real-world \p\ deployment of the scheme.
+One major limitation of \emph{Mithos} is that the protocol cannot effectively
+handle dynamic arrivals and departures from the network. In \cite{CCRK2004} it
+is claimed to need extensive probing to determine distances and at the same time
+its results are evaluated as less accurate than, for example,
+GNP \cite{ng_gnp_2001}. Maybe this has also to do with its iterative neighbor
+selection process which is reported to be prone to prematurely terminate at
+local minima \cite{WSS2005}. Another point that has negative effect to the
+scalability of the algorithm is that parts of its spring relaxation technique
+require a centralized implementation~\cite{cox_vivaldi_2004}.
 
@@
 to further reduce routing table update costs. It is theoretically shown
-that in \emph{LAPTOP}, routing path length is bounded by $O(log_d N)$ and node joining
-and leaving in the overlay network is bounded by $O\left( d log_d N \right)$
-hops in a balanced overlay tree, where $N$ is the number of nodes, and $d$ is
-the maximum degree of each node. \emph{LAPTOP} implements the geographical layout
-approach  and constructs its layout in a self-organizing and efficient fashion.
-The physical distances are estimated with round trip times to a few existing
-nodes in the overlay network.
+that in \emph{LAPTOP}, routing path length is bounded by $O(log_d N)$ and node
+joining and leaving in the overlay network is bounded by
+$O\left( d log_d N \right)$ hops in a balanced overlay tree, where $N$ is the
+number of nodes, and $d$ is the maximum degree of each node. \emph{LAPTOP}
+implements the geographical layout approach  and constructs its layout in a
+self-organizing and efficient fashion. The physical distances are estimated
+with round trip times to a few existing nodes in the overlay network.
 
@@
 % Trace-route is categorized as too heavy-weighted and intrusive for use in a
-% larger scale system \cite{RHMKS2002}. Additionally, disabling
+% larger scale system \cite{RKHS2002}. Additionally, disabling
 % ICMP is a common administrative policy for edge sites to enforce security,
@@
 light-weight maintenance cost. On the other hand, it exhibits slow convergence
-speed and inefficiency in a high churn environment\cite{DK2006}.
+speed and inefficiency in a high churn environment \cite{DK2006}.
 
@@
 % not increase.
-high
+medium
 \end{tabular}
@@
 
+This approach is able to reduce the average path length of the system to be
+logarithmically proportional to the number of available domains instead of the
+number of participating peers. However due to the large number of Internet
+domains and the potentially small number of system peers within each domain,
+this approach doesn't seem scalable and efficient \cite{DK2006}.
+
 \begin{center}
@@
 % Is designed on the basis of IPv6 protocol which is not widely operational yet.
-medium
+low
 \end{tabular}
@@
 
+The algorithm's overhead is very low because detection of proximity is only
+a comparison of prefixes. On the other hand, the efficiency of the proximity
+is caught into a vicious cycle with the storage space needed for DHT object
+publishing (how many bits will be used for either). Moreover, its static
+nature (no online measurements used) makes it difficult to handle
+\cite{EWF2011}.
+
 % Moreover, observing the
@@
 % is caught in a vicious cycle with the storage space needed for DHT object
-% publishing (how many bits will be used for either)
+% publishing (how many bits will be used for either).
 low &
@@
 \emph{RLM} &
-%
+%%
+%% TAKEN FROM
+%% Efficient service discovery mechanism for wireless sensor networks
+%%
+%%
+% common prefix of the overlay ID with the closest landmark. The
+% RLM utilizes the overlay lookup capabilities by reducing traffic
+% overhead within an overlay cluster. However, it generates much
+% control traffic for periodically measuring distances to all landmarks.
+% Moreover, the hop stretch overhead still increases when
+% performing overlay routing among overlay clusters.
 medium &
@@
 % work cooperatively with landmark binning itself.
+% Security is compromised with arbitrary jumps!
 high &
@@
 nodes improves the query efficiency of the system, on the other hand, it tends
-to create hot-spots, and the needed failure resilience is undermined by the fact
-that nearby nodes are more likely to suffer collective failures.
+to undermine the uniform distribution of peer identifiers which consequently
+tend to create hot-spots, failing resilience is undermined by the fact
+that nearby nodes are more likely to suffer collective failures \cite{HY2007}.
+
+% must
+% cope with the inherent trade os associated with selecting
+% identiers based on location: randomly selected identiers
+% are essential for the scalability and security of the system
+% and help to balance load while location-aware identiers pre-
+% vent messages from taking extensive scenic routes (on the
+% physical level).
 

### 2014-02-02 unstructured.tex
2013.07.28 (Vassilis) Further investigation conducted to find research work that references the papers/algorithms we survey and cite. Text enriched.

@@
 % to start from the last level of nodes and results in a restart of the whole
-% process
+% process.
 low &
-% the process in not started from the beginning at each iteration step
+% the process in not started from the beginning at each iteration step. For
+% popular files this can substantially reduce overhead but for less popular
+% ones, its performance might be uncertain due to the repeated restart of
+% flooding
 low &
@@
 Although this adoption was 
-a step in the right direction~\cite{lv_randomwalks_2002},
+a step in the right direction \cite{LCCLS2002},
 issuing a single copy of the query within the network
@@
 that primarily focuses on the satisfaction of peers improves the network's 
-scalability, it falls short in addressing the mismatch problem
+scalability it falls short in addressing the mismatch problem
 as considerations for the underlying network are not handled in explicit
-terms.
+terms \cite{LXLNZ2005}. Moreover it reduces the search scope, exhibits poor
+performance in the worst case when matching data is not found quickly
+\cite{PR2004,HJ2004} and can potentially build  disconnected topoloagies
+\cite{MBL2006}.
 %
@@
 Interestingly, such cycles appear even along the correct forwarding path.
-\cite{ZKB2008}~focuses on this exact deficiency of overlay networks and
+\cite{ZKB2008} focuses on this exact deficiency of overlay networks and
 introduces the \emph{Distributed Cycle Minimization Protocol
@@
 Even though cycle elimination does improve network performance, 
-it cannot  directly contribute in the solution of the topology mismatch problem.
+it cannot  directly contribute to the solution of the topology mismatch
+problem, while its perfomance in a high traffic and churn environments and its
+ability to retain important characteristics of a Gnutella-like topology (i.e.,
+like degree distribution, average peer distance, diameter e.t.c.) has been
+doubted \cite{SCG2009,CSG2010}. On the other hand, even though some of the overhead
+saved by the dublicate message reduction is spent to control the cycle cutting
+mechanism it is reported that still stays for example one to two orders of
+magnitude less than LTM due to its ``lazy'' broadcasting of control messages
+(as opposed to LTM's periodic approach) \cite{ZKB2008}.
 %
@@
 % topology mismatch and thus DCMP minimizes this as possible.
-% The control overhead is also one to two orders of magnitude less than LTM due
-% to its ``lazy'' broadcasting of control messages (as opposed to LTM's periodic
-% approach) - \cite{ZKB2008}
 low &
@@
 
-In~\cite{CS2002}, replication is used as a way to improve inefficient blind
-search. As the number of replicas and/or cached copies increases in the network,
-it would relatively easy to locate items even if search is random.
-To analyze the feasibility of such an approach, \cite{CS2002} investigates
-$3$ replication strategies: \emph{uniform}, \emph{proportional}, and
-\emph{square root allocation}. In the uniform model, the copies of items are
-uniformly spread over the network, while in the proportional, items are
-replicated based on their query rate, so that frequently queried items can be
-found more often.
-The square root is a compromise between uniform and proportional allocation
-models. The overall costs for both successful and unsuccessful
-searches are compared to ascertain the effectiveness of the $3$ suggested 
-approaches. Experimental outcomes indicate that the uniform allocation 
-minimizes the maximum search length and so it can reduce the time spent on
-unsuccessful searches. 
-The proportional strategy effectively decreases the search time 
-for popular items, but suffers when needing to locate the rare entities.
-When it comes to expected successful search size, it is the same for 
-for both uniform and proportional models and any approach in between 
-to always behave much better. The square root allocation  aims at minimizing the 
-expected search size for successful queries in \p\ networks.
-The table below outlines how the $3$ strategies fare:
+In \cite{CS2002} and \cite{LCCLS2002}, replication is used as a way to improve
+inefficient blind search. The observation is that as the number of replicas
+increases in the network, it would be relatively easy to locate items even if
+search is random.
+
+The work discusses $3$ approaches for replication allocation namely \emph{uniform},
+\emph{proportional}, and \emph{square-root replication}. In the uniform model,
+all objects are replicated without taking into account the query distribution,
+while in the proportional, objects are replicated analogously to their query
+rate, so that frequently queried items can be found more often. Experimental
+outcomes indicate that the uniform allocation minimizes the maximum search
+length and so it can reduce the time spent on unsuccessful searches. The
+proportional strategy effectively decreases the search time for popular items,
+but suffers when needing to locate the rare entities. When it comes to expected
+successful search size, it is the same for for both uniform and proportional
+models and any approach in between to always behave much better. Finally, the
+square-root model is an attempt to find the golden ratio between uniform and
+proportional allocation models. It suggests setting the number of object
+replicas to the square root of the searching rate for an object\footnote{An
+interesting application of the square-root principle has also been explored
+in the context of topology reconstruction, by ensuring node degree to be
+proportional to the square-root of its content popularity \cite{C2005}.}.
+Theory in \cite{CS2002} suggest that square-root minimizes the expected search size on
+successful queries, claim that is supported by simulations \cite{LCCLS2002}.
+
+In \cite{LCCLS2002}, a \emph{k-walker} query algorithm is presented and shown
+that using it along side square-root allocation results in perfomance similar to
+Gnutella's query flooding method but incurring up to two
+orders of magnitude less network traffic. Also checks $3$ strategies for
+replica placement, namely \emph{owner replication} where an object is replicated
+at the requesting node, \emph{path replication} that caches the object along the
+path from the requesting to the providing node and \emph{random replication}
+where the replication is taking place randomly on visited nodes. Observations in
+this work are not directly applicable to structured DHTs, because it is assumed
+that the lookup time for an object depends only on the number of replicas and
+not the placement strategy \cite{RS2004}.
+
+%%
+%% SOME EXTRA DISCUSSION TAKEN FROM
+%% IRM: Integrated File Replication and Consistency Maintenance in P2P Systems
+%% 
+%% Lv et al. [21] and Cohen and
+%% Shenker [22] showed that replicating objects proportionally
+%% to their popularity achieves optimal load balance but has
+%% varied search latency, while uniform replication has the
+%% same average search latency for all files but causes load
+%% imbalance. Square-Root replication method replicating files
+%% proportionally to the square root of their popularity is such
+%% that both average search size and capacity utilization rate
+%% vary per object, but the variance in utilization is consider-
+%% ably smaller than with uniform replication, and the variance
+%% in average search size is considerably smaller than with
+%% proportional replication.
+%%
+
+The table below outlines how the $3$ allocation approaches fare:
 %%
@@
 Along with {\sl Narada}, the \emph{Adaptive Overlay Topology 
-Optimization (AOTO)}~\cite{LZXN2003} is one of the first attempts to address the
+Optimization (AOTO)} \cite{LZXN2003} is one of the first attempts to address the
 topology mismatch problem. \emph{AOTO} is a distributed algorithm that seeks to
@@
 neighbors.
-%%%
+%
 The above optimization is conducted within $1$-neighbor
@@
 \hline
-% compared to LTM it has slow convergence speed.
+% compared to LTM it has slow convergence speed because of the information
+% exchange among each peer and all its neighbours.
 medium &
@@
 
-\emph{Location-aware Topology Matching (LTM)}~\cite{LLXNZ2004} 
+\emph{Location-aware Topology Matching (LTM)} \cite{LLXNZ2004}
 seeks to optimize an overlay \p\ structure based on the physical topology.
@@
 network and delaying query responses.
+Even though this attempt is based on a simple and elegant solution it is
+reported to not guarantee performance in terms of average communication delay
+or search scope \cite{HLH2009}.
 %
@@
 %percent.
+%
+% THE LINES BELOW NEED REWORDING
+% ==============================
+% \cite{LXN2004} Based on the above discussions, we can see the major
+% advantages of using a bipartite overlay. First, the operation
+% overhead is reduced. In ACE, every peer is required to
+% probe the neighbor distance and compute the MST, whereas
+% in SBO, white peers do the probing and red peers do the
+% computing. Second, SBO extends ACE into two-hop
+% neighbors, without additional information exchange
 %
@@
 probed peers) minimize the unnecessary messages for the network maintenance.
+There are works in the literature though that critisize, through simulation
+results, the algorithm's average broadcast delay and overlay improvement
+guarantee \cite{HLY2010}.
 These characteristics are depicted to the $3$ evaluation criteria we use in this
@@
 how fast the introduced peer selection functions. Admittedly, \emph{APS} is slow 
-due to the learning process and its inherent complexity. As far as the $3$
-criteria is concerned, \emph{APS} has as follows:
+due its on-line learning process and inherent complexity \cite{ZZLZ2009}. As far
+as the $3$ criteria is concerned, \emph{APS} has as follows:
 \begin{center}
@@
 random graphs as it can lead to a larger coverage of peers and at the same time
-helps reduce duplication of messages travelling around the network. This
+help reduce duplication of messages travelling around the network. This
 combination offers efficient lookups that pose minimal or no negative impact to
 other mechanisms possibly employed by the \p\ systems (e.g., 1-hop replication
-or dynamic querying mechanisms).
-A 50\% reduction in communication latency among peers by cutting off 
-some 20\% of the \emph{IP}-level traffic generated 
-in reported in~\cite{PRFM2009}.
+or dynamic querying mechanisms). Experimental results in the same paper suggest
+a 50\% reduction in communication latency among peers by cutting off some 20\%
+of the \emph{IP}-level traffic generated.
 %%
@@
 biased neighbor selection along with bandwidth throttling
-and caching techniques for near-optimal lookup results is proposed.
-Unfortunately, \emph{BNS} deployment requires knowledge of 
-\emph{ISP} network mappings and awareness of possible changes 
-that occur in the infrastructure limiting its widespread adoption;
-the following table qualitatively depicts the expected behavior of \emph{BNS}:
+and caching techniques for near-optimal lookup results is proposed. On the other
+hand, the effectiveness of \emph{BNS} is arguable by other studies, whose
+experimental results suggest that most peers in a swarm reside in different ASs,
+thus a client in most of the cases does not have enough peers to leverage
+intra-AS connections \cite{RTLCGZ2010}. Moreover, \emph{BNS} deployment
+requires knowledge of \emph{ISP} network mappings and awareness of possible
+changes that occur in the infrastructure limiting its widespread adoption; the
+following table qualitatively depicts the expected behavior of \emph{BNS}:
 %%
@@
 within the borders of a single autonomous system. Contrary to
-\emph{BSN}~\cite{BCCMSBZ2006},
+\emph{BNS}~\cite{BCCMSBZ2006},
 \emph{Oro} leads to improved performance 
@@
 %%
+Even though the results in the paper claim minimization of inter-ISP traffic
+cost while enhancing the, already, near optimal performance of BitTorrent
+download, similar to \cite{BCCMSBZ2006}, this ISP-friendly design of \emph{Ono}
+is questioned, especially on a Internet-scale-basis \cite{PMJKA2009,LCLX2009,CLYSR2011}.
+%%
 \begin{center}
@@
 % already, near optimal performance of BitTorrent download.
-high &
+medium &
 % periodic DNS lookups and CDN redirection imposes medium control overhead to
@@
 over inter-AS communication thus reducing the general cost of communication.
+Again, it is pointed out that there is a trade-off between reducing inter-domain
+traffic and fairness among peers in terms of the data the peers upload
 
@@
 a good peer selection metric should take into account both the
-downloading speed and the network topology. 
-In this context, the approach suggests a metric that designates
-peers demonstrating high download rate, low reciprocal upload demand 
-and low routing hops.
-The metric is defined as the ratio of download to upload rate ${d}/{u}$
-divided by the either the link-level $l$ or the number of routing hops required $a$.
+downloading speed and the network topology. The metric proposed is defined as
+the ratio of download to upload rate ${d}/{u}$
+divided by either the link-level $l$ or the number of routing hops required $a$.
 This metric should guarantee selection of peers with high download rate, low
@@
 Through experimentation on hundreds of {\it PlanetLab} and residential hosts,
-\emph{TopBT} offers more than 25\% traffic reduction and 15\% faster downloads
+\emph{TopBT} offers more than $25\%$ traffic reduction and $15\%$ faster downloads
 if compared with popular {\sl BitTorrent} implementations.
@@
 
-% In \cite{hsiao_redblue_2009}, Hsiao et al, claim to construct topology-aware
+% In \cite{HLH2009}, Hsiao et al, claim to construct topology-aware
 % unstructured overlays that \emph{guarantee} performance qualities in terms of
@@
 The former ensure the reduction of long distance traveling for messages, while
-the later secure the connectivity of the structure and prevent network
+the latter secure the connectivity of the structure and prevent network
 partitioning.
@@
 random connections. For establishing its sibling connections, the newcomer
-multicasts a specialized \emph{message} $l$, to locate the right candidate(s).
+multicasts a specialized message $l$, to locate the right candidate(s).
 Initially, the behaviour of this message, is modeled as a random walker,
@@
 medium &
-% overhead of split-hash and dnMatch is of low cost (TODO: double check this).
+% overhead of split-hash and dnMatch is of low cost
 % Walker message is multicasted only when sibling connections must be
@@
 
-\emph{Landmark Binning}~\cite{RHMKS2002} partitions close-by nodes 
+\emph{Landmark Binning}~\cite{RHKS2002} partitions close-by nodes
 into bins based on their distance from well known anchor nodes across the
@@
 Despite the fact that such delays are not always accurate,
-they are used in~\cite{RHMKS2002} for they are 
+they are used in~\cite{RHKS2002} for they are
 non-intrusive, transparent and easy to apply.
@@
 across the the Internet. 
-It is conjectured~\cite{RHMKS2002} that $12$ such servers can prove
+It is conjectured~\cite{RHKS2002} that $12$ such servers can prove
 sufficient for the task.
@@
 connected at any given point in time, landmark servers do 
-inherently play a pivotal role in the operation of the network.
-Although latency estimation does not drain network resources,
-the scale of the approach does call for an alternative design.
-In this respect, \cite{RHMKS2002} advocates the replacement of 
-each landmark server by a cluster of servers.
-However, this approach may not entirely avoid the creation of 
-excessive network traffic flow through its landmark server-clusters.
+inherently play a pivotal role in the operation of the network. Moreover,
+studies like \cite{GSG2002}, pin point the fact that end-host shoulder the
+latency measurments as the scheme's scalability bottlneck while others argue
+about the uneven distribution of nodes potentially resulting in load imbalance
+\cite{CDCR2002,M2003,ZZZSZ2004}, its coarse-graned nature that fails to
+effectively distinguish among realatively close nodes \cite{RGJZ2004} and the
+potentially increased maintenance cost \cite{XTZ2003}.
+
+% Although latency estimation does not drain network resources,
+% the scale of the approach does call for an alternative design.
+% In this respect, \cite{RHKS2002} advocates the replacement of
+% each landmark server by a cluster of servers.
+% However, this approach may not entirely avoid the creation of 
+% excessive network traffic flow through its landmark server-clusters.
 
@@
 \subsubsection{Forwarding Optimization Methodology}
-Approaches based on forwarding-optimization 
-propose intelligent forwarding link selection for a message's next
-hop across a routing path. 
-The selection criteria vary depending on the
-algorithm, but they mainly use one or more statistical heuristics. 
-These can range from simple connection speed, to exploiting specific peer
-attributes such as high processing capacity e.t.c. An example is \emph{Gia} that
-exploits the so called heterogeneity of peers as well as employs a token-based
-flow control mechanism to guide the messages towards peers that are more likelly
-to answer
-a query. More sophisticated alternatives also exist,
-taking into account application level requirements or exploiting machine
-learning techniques to adjust to overlay network routing, like the \emph{APS}
-algorithm.
+In unstructured systems there is no way to know where data are located. That is
+why early works like Gnutella \cite{gnutellav04} used blind flooding, a Breadth
+First Search approach with depth $d$, a system-maximum TTL value (in terms of
+hops) for messages. This is inefficient since it generates a large number of
+messages, a lot of them duplicates. Approaches based on
+\emph{forwarding-optimization} propose intelligent search mechanisms to try to
+make searching more efficient.
 
@@
 show overloading signs).
-On one hand, forward
-optimisation approaches have 
+
+Many alternative schemes have been proposed for this category, some of which we
+have surveyed; iterative deepening, modified BFS, local indices \cite{YG-M2002},
+$k$-walker random walks \cite{LCCLS2002} or heterogeneity exploitation and
+token-based flow control mechanisms \cite{CRBLS2003}. More sophisticated
+alternatives also exist, taking into account application level requirements
+or even exploiting machine learning techniques to adjust overlay network
+routing \cite{BFLZ2003}.
+
+Forwarding-optimization schemes in unstructured \p\ systems can be classified
+as DFS- or BFS-based. Routing indices \cite{CG-M2002} for example is a
+DFS-based technique while all the aforementioned are BFS. Also they can
+be classified as deterministic or probabilistic (probabilistic, random or
+ranking-based query forwarding). Iterative deepening and local indices
+\cite{YG-M2002} are examples of deterministic approaches. Moreover, in the
+literature algorithms are also taxonomised as blind or informed depending on
+whether nodes keep some metadata to facilitate the search. For example,
+$k$-walker random walks or iterative deepening are considered blind searches
+while local indices informed.
+
+On one hand, forward-optimization approaches have 
 the advantage of enhancing the
@@
 network. 
-Moreover, they adress the problem of the mismatch in a limited manner
+Moreover, they address the problem of the mismatch in a limited manner
 since they do not provide any guarantees 

### 2014-02-02 intro.tex
2013.10.20 (Alex) Text reworked and reviewed for introduction and unstructured sections. BibTeX file reformed and debugged (once more :-)

@@
 participating nodes, called \emph{peers}, act as both resource providers and
-resource consumers, in contrast to the traditional \emph{client-server} model
+resource consumers in contrast to the conventional \emph{client-server} model
 where nodes undertake specific roles.
@@
 %%
-Apart from the above, though, a startling number of diverse and successful
+Apart from the above, a startling number of diverse and successful
 applications have been built based on \p\ architectures. Some of them include:
@@
 created above an existing network; overlays provide an abstraction that
-enables the implementaion of efficient, fully distributed, application-layer
+enables the implementation of efficient, fully distributed, application-layer
 services such as routing messages to destinations that are not known in advance
@@
 in the overlay.
-Hence, it is very critical that \p\ networks be laid out 
+Hence, it is critical that \p\ networks be laid out 
 in ways that their topology matches, as closely as possible, the 

### 2014-02-02 unstructured.tex
2013.10.20 (Alex) Text reworked and reviewed for introduction and unstructured sections. BibTeX file reformed and debugged (once more :-)

@@
 performance in the worst case when matching data is not found quickly
-\cite{PR2004,HJ2004} and can potentially build  disconnected topoloagies
+\cite{PR2004,HJ2004} and can potentially build  disconnected topologies
 \cite{MBL2006}.
@@
 \emph{DCMP} bases its operation on messages whose travel is limited by an
-imposed $TTL$ value (max set to $7$ for most cases). This inherently limits
-the protocol in detecting cycles that span for more than $TTL$ nodes.
+imposed \emph{TTL} value (max set to $7$ for most cases). This inherently limits
+the protocol in detecting cycles that span for more than \emph{TTL} nodes.
 Even though cycle elimination does improve network performance, 
 it cannot  directly contribute to the solution of the topology mismatch
-problem, while its perfomance in a high traffic and churn environments and its
-ability to retain important characteristics of a Gnutella-like topology (i.e.,
+problem, while its performance in a high traffic and churn environments and its
+ability to retain important characteristics of a {\sl Gnutella}-like topology (i.e.,
 like degree distribution, average peer distance, diameter e.t.c.) has been
-doubted \cite{SCG2009,CSG2010}. On the other hand, even though some of the overhead
-saved by the dublicate message reduction is spent to control the cycle cutting
+doubted~\cite{CSG2010}. 
+On the other hand, even though some of the overhead
+saved by the duplicate message reduction is spent to control the cycle cutting
 mechanism it is reported that still stays for example one to two orders of
@@
 In \cite{LCCLS2002}, a \emph{k-walker} query algorithm is presented and shown
-that using it along side square-root allocation results in perfomance similar to
-Gnutella's query flooding method but incurring up to two
+that using it along side square-root allocation results in performance similar to
+{\sl Gnutella}'s query flooding method but incurring up to two
 orders of magnitude less network traffic. Also checks $3$ strategies for
@@
 % compared to LTM it has slow convergence speed because of the information
-% exchange among each peer and all its neighbours.
+% exchange among each peer and all its neighbors.
 medium &
@@
 to probe immediate neighbors using delay distance measurements 
-and store information locally. The algorithm introduces a special query message
-type, the \emph{Piggy Message (PM)} which includes information about 
-the neighbor identification (IP) and measured distance ($d$). PMs are
+and store this information locally. 
+The algorithm introduces a special query message type, 
+the \emph{Piggy Message (PM)} which includes information about 
+the neighbor identification (\emph{IP}) and measured distance ($d$). PMs are
 piggybacked to normal
-Gnutella messages. When node $P$ receives a query from $Q$ it constructs a
-$PM_{PQ}$, piggybacks it to the query and forwards it to $P$'s neighbors. When
-the neighbors recieve it they detach the $PM_{PQ}$, record the distance
-$d_{PQ}$ and process the query normaly. Selection of which incoming queries
-should be piggybacked with a \emph{PM} is proposed by either
-\emph{pure probability-based (PPB)} or \emph{new neighbor triggered (NNT)}
-policies.
+{\sl Gnutella} messages. 
+When a node $P$ receives a query from $Q$, it constructs a
+$PM_{PQ}$, piggybacks it to the query and forwards it to $P$'s neighbors. 
+As soon as a neighbor detaches a received $PM_{PQ}$, records the distance 
+$d_{PQ}$ and processes the query.
+% When the neighbors receive it they detach the $PM_{PQ}$, 
+% record the distance $d_{PQ}$ and process the query normaly. 
+Selection of which incoming queries
+should be piggybacked with a \emph{PM} is determined using either a
+% proposed by either
+\emph{pure probability-based (PPB)} or a \emph{new neighbor triggered (NNT)}
+policy.
 
@@
 thus far. Let $P$ be a $1$-hop distance neighbor of $S$, and $Q$ be a probed
-peer by $S$. When $S$ receives a message piggybacked with a $PM_{PQ}$ we
-identify the following cases.
-\begin{itemize}
-  \item If $Q$ is already a direct neighbor of $S$ then we check the distances
-  $d_{SQ}$, $d_{SP}$ and $d_{PQ}$. If either $d_{SQ}$ or $d_{SP}$ is the longest, then the longest
-  link will be put in a \emph{will-cut} list\footnote{A peer will not send or
+peer by $S$. When $S$ receives a message piggybacked with message $PM_{PQ}$, 
+% we identify the following cases.  %%AD put it passive as below.. 
+the following $2$ cases are identified:
+
+\begin{enumerate}
+  \item If $Q$ is already a direct neighbor of $S$ then we check 
+	the distances $d_{SQ}$, $d_{SP}$ and $d_{PQ}$. 
+	If either $d_{SQ}$ or $d_{SP}$ is the longest, then the longest
+  	link will be placed in a \emph{will-cut} list\footnote{A peer will not send or
   	forward queries to connections in its will-cut list but will preserve them
   	for some time in order not to effect ongoing response messages travelling the
-  inverse path.}. If $d_{PQ}$ is the longest then nothing is done by $S$; being
+  	inverse path.}. 
+	If $d_{PQ}$ is the longest, then nothing is done by $S$; being
   	fully distributed, neighbor comparison and selection process, will have the
-  oportunity to handle $PQ$ link when initiated by peers $P$ or $Q$.
+  	opportunity to handle $PQ$ link when initiated by peers $P$ or $Q$.
 
   \item If $Q$ is strictly a $2$-hop neighbor of $S$ and have never probed $Q$
-  in the past, stores distance $d_{SQ}$ and checks distances $d_{SQ}$, $d_{SP}$ and $d_{PQ}$.
+  	in the past, stores distance $d_{SQ}$ and checks 
+	distances $d_{SQ}$, $d_{SP}$ and $d_{PQ}$.
   	If $d_{SQ}$ is the longest, $S$ will not create link $SQ$. If $d_{SP}$ is the
@@
   	$P$ or $Q$ will disconnect $PQ$ later.
-\end{itemize}
+\end{enumerate}
 
@@
 
-As seen above, the method is fully distributed and needs only minimum knowledge
-of, at most, a $2$-hop distant peers. This makes it a good candidate for big
-overlay networks. Also topology adaptation helps construct a well fit overlay
-network with restect to the underlying one (at least with respect to network
-distances). Morever will-cut list and distance cache (which stores already
+As discussed above, \emph{THANCS} is fully distributed and needs only minimum knowledge
+of, at most, a $2$-hop distant peers. This features renders 
+the method a good candidate for large overlay networks. 
+Also its topology adaptation helps construct a well fit overlay
+with respect to the underlying network--at least with respect to network
+distances. In addition, will-cut list and distance caches (which store already
 probed peers) minimize the unnecessary messages for the network maintenance.
-There are works in the literature though that critisize, through simulation
-results, the algorithm's average broadcast delay and overlay improvement
-guarantee \cite{HLY2010}.
-These characteristics are depicted to the $3$ evaluation criteria we use in this
-survey as follows:
-
+%%
+In \cite{HLY2010}, the simulation-derived results of \emph{THANCS} as far
+average broadcast delay and overlay improvement guarantees are discussed;
+overall, \emph{THANCS} fares as follows regarding the $3$ stated criteria:
+%%AD rephrased more diplomatically the following as above.. 
+% There are works in the literature though that criticize, through simulation
+% results, the algorithm's average broadcast delay and overlay improvement
+% guarantee \cite{HLY2010}.
+% These characteristics are depicted to the $3$ evaluation criteria we use in this
+% survey as follows:
+%%%%%%%%%%%%%%%%%%%%%%
 \begin{center}
@@
 order. The latter indicates that in practice triple sequences $(v_1, v_2, v_3)$
-are used. A mismatch is detected as follows: supposing that we want to verify
+are used. A mismatch is detected as follows: should we want to verify
 a sequence, say $v_2-v_1-v_3$, two probing messages are dispatched from $v_1$ to
@@
 directly to $v_3$. 
-Similarly, the message reaching $v_3$, it is directly forwarded to $v_2$.
+Similarly, the message reaching $v_3$ is directly forwarded to $v_2$.
 The above two forwarding actions seek to quantify 
-the corresponding delays of $(v_2,v_3)$ and $(v_3,v_2)$ physical paths
-denoted by $y$. 
+the corresponding delays of $(v_2,v_3)$ and $(v_3,v_2)$ 
+% physical paths denoted by $y$. 
+for the physical path $y$.
+%%AD check above... it is one path (not paths right?)
 %%
@@
 features include connection link load, bandwidth, and past uploading experience.
-Subsequently, a Markov decision process is introduced as a mechanism to shape
-the policy for switching among the peers. The success of approach depends on
+Subsequently, a Markov decision process is used
+% introduced as a mechanism 
+to shape the policy for switching among the peers. The success of approach depends on
 how fast the introduced peer selection functions. Admittedly, \emph{APS} is slow 
-due its on-line learning process and inherent complexity \cite{ZZLZ2009}. As far
-as the $3$ criteria is concerned, \emph{APS} has as follows:
+due its on-line learning process and inherent complexity \cite{ZZLZ2009}. 
+\emph{APS}' behavior regarding the $3$ criteria has as follows:
+% As far as the $3$ criteria is concerned, \emph{APS} has as follows:
 \begin{center}
@@
 as its neighbors.
-%%AD how do they select all these magic constants???
-%%VM This α constant seems to be the percentage of close neighbours selected by
+%%QUE how do they select all these magic constants???
+%%ANS This α constant seems to be the percentage of close neighbors selected by
 % a peer out of its total degree. Can't find the criterion with which it is
 % chosen though. Maybe we can ask... Mema :)
-Searching occurs in two phases: in the initial phase, the querying node
-floods its distant neighborhood  with $TTL = 1$.
+Searching occurs in $2$ phases: in the initial, the querying node
+floods its distant neighborhood  with \emph{TTL=1}.
 Subsequently, peers receiving a query over a ``long link'' commence a 
-local flood with $TTL=ttl$ where \emph{ttl} is system-defined.
-%%AD how are you supposed to define these params?? in a fixed manner??
-%%
-%% VM Don't understand... the paper says that for increased awareness we need a
-%% small α (page 7). But in the begining they give that short links (close-by)
+local flood with \emph{TTL=ttl} where \emph{ttl} is system-defined.
+%%QUE how are you supposed to define these parameters?? in a fixed manner??
+%%
+%%ANS Don't understand... the paper says that for increased awareness we need a
+%% small α (page 7). But in the beginning they give that short links (close-by)
 %% are given by α*Ν. I guess this gets smaller for smaller α.... Also on
@@
 combination offers efficient lookups that pose minimal or no negative impact to
-other mechanisms possibly employed by the \p\ systems (e.g., 1-hop replication
+other mechanisms possibly employed by the \p\ systems (e.g., $1$-hop replication
 or dynamic querying mechanisms). Experimental results in the same paper suggest
-a 50\% reduction in communication latency among peers by cutting off some 20\%
+a $50$\% reduction in communication latency among peers by cutting off some $20$\%
 of the \emph{IP}-level traffic generated.
 %%
-\emph{ITA} places as follows in the context of the $3$ criteria:
+Regarding the $3$ criteria, \emph{ITA} is placed as follows:
 %
@@
 % \subsubsection{EGOIST}
-\emph{EGOIST}~\cite{SLLBBR2008} is a set of algorithms 
+\emph{EGOIST}~\cite{SLLBBR2008} is a set of algorithms used 
 to construct and manage overlay networks. 
@@
 it has a complete picture of the overlay graph. 
-Then the peer estimates the delay to all other nodes to
+Then, the peer estimates the delay to all other nodes to
 determine its potential neighbors and ultimately connects 
@@
 for example the minimization of the average delay to all its
-neighbors. \emph{EGOIST} requires extensive resource usage to continuously
-update the overlay connections of all the nodes in the system. This needs
-$O(n^2)$ measurements. Fortunatelly, each node does not need to do these
+neighbors. 
+\emph{EGOIST} requires extensive resource usage to continuously
+update the overlay connections for all nodes in the system. This needs
+$O(n^2)$ measurements. 
+Fortunately, each node does not need to do these
 measurements for maintenance (monitoring and updating) with all other
 participating nodes, but just with a number of $k<n$ nodes that it chooses to
-establish links with. This gives a reduced $O(kn)$ time complexity for
-maintenance while $O(n^2)$ is needed less frequently, when nodes reevaluate
-their connections.
-
-Regarding the $3$ criteria, \emph{EGOIST}'s behavior is as follows:
+establish links with. 
+The latter yields a reduced $O(kn)$ time complexity for
+maintenance while complexity $O(n^2)$ is required less frequently 
+and only when nodes reevaluate their connections.
+%
+\emph{EGOIST} fares as follows:
+% Regarding the $3$ criteria, \emph{EGOIST}'s behavior is as follows:
 \begin{center}
@@
 by carefully choosing most of the neighbors of a peer to come from 
-the same \emph{ISP}, while at the same time adhering to the near-optimal
+the same \emph{ISP}, while simultaneously adhering to the near-optimal
 download performance of the protocol.
 %%
-By and large, {\sl BitTorrent} deploy mechanisms that have proved
-aggressive for \emph{ISP}s' networking and accounting as they function 
-in a ``without-borders'' approach. The default {\sl BitTorrent} specification
+By and large, {\sl BitTorrent} deploys mechanisms that have proved
+aggressive to \emph{ISP} networking and accounting as they function 
+in a ``without-borders'' approach. 
+The default {\sl BitTorrent} specification
 designates $35$ connections for each peer regardless if such connections are
@@
 %%
-\emph{BNS} proposes that only $k$ out of these $35$ connections are 
+\emph{BNS} proposes that only $k$ out of these $35$ connections be
 established with nodes beyond the {\it ISP}-borders;
@@
 expensive network links should there exist alternative local connections
-offering the sought services. 
-\emph{BNS} is realized either through 
-the modification on the tracker so that \emph{ISP}-local resources are rapidly 
-identified,
-or through \p\ traffic shaping devices installed on \emph{ISP} edge routers. 
+offering the required services. 
+%%
+\emph{BNS} is realized through 
+\begin{inparaenum}[1)]
+\item
+the modification of the tracker so that 
+\emph{ISP}-local resources are rapidly identified
+\item
+the use of \p--traffic--shaping devices on \emph{ISP} edge--routers. 
+\end{inparaenum}
 %%
@@
 biased neighbor selection along with bandwidth throttling
-and caching techniques for near-optimal lookup results is proposed. On the other
-hand, the effectiveness of \emph{BNS} is arguable by other studies, whose
-experimental results suggest that most peers in a swarm reside in different ASs,
-thus a client in most of the cases does not have enough peers to leverage
-intra-AS connections \cite{RTLCGZ2010}. Moreover, \emph{BNS} deployment
+and caching techniques for near-optimal lookup results is proposed. 
+The effectiveness of \emph{BNS} is debated 
+in~\cite{RTLCGZ2010} where experimental results suggest that 
+most peers in a swarm reside in different authonomous systems (\emph{AS}s).
+%%AD AS = autonomous systems??? - put it so. First time discussed!
+Consequently, clients frequently do not have enough peers to leverage
+intra-\emph{AS} connections \cite{RTLCGZ2010}.
+%%AD rephrased as above - what is AS??? Acronym... 
+% On the other hand, the effectiveness of \emph{BNS}
+% is arguable by other studies, whose
+% experimental results suggest that most peers in a swarm reside in different ASs,
+% thus a client in most of the cases does not have enough peers to leverage
+% intra-AS connections \cite{RTLCGZ2010}. 
+Moreover, \emph{BNS} deployment
 requires knowledge of \emph{ISP} network mappings and awareness of possible
@@
 for peer distance estimation. {\sl CDN}s already use both static (i.e.,
-geographical, IP-clustering) and dynamic (i.e., network delay measurement)
+geographical, \emph{IP}-clustering) and dynamic (i.e., network delay measurement)
 information for their replica selection. Thus, the algorithm leverages the
-observation that peers that are redirected to a specific CDN replica does not
+observation that peers that are redirected to a specific {\sl CDN} replica does not
 only mean that these are close by this replica but to each other as well.
@@
 %%
-The protocol's bootstrapping phase consists of the newcomming peer performing
-{\sl DNS} lookup to {\sl CDN}-names to build its redirection information.
+The protocol's bootstrapping phase consists of the incoming peer performing
+{\sl DNS} lookups to {\sl CDN}-names to build its redirection information.
 %%
@@
 is sufficiently fresh, 
-\emph{Oro} persists the map after the end of a BitTorrent session.
-%%
-Even though the results in the paper claim minimization of inter-ISP traffic
-cost while enhancing the, already, near optimal performance of BitTorrent
-download, similar to \cite{BCCMSBZ2006}, this ISP-friendly design of \emph{Ono}
-is questioned, especially on a Internet-scale-basis \cite{PMJKA2009,LCLX2009,CLYSR2011}.
-%%
+\emph{Oro} persists the map after the end of a {\sl BitTorrent} session.
+%%
+Experimentation with \emph{Oro} yields minimization of
+inter-\emph{ISP} traffic costs and achieves near optimal performance 
+of {\sl BitTorrent} download~\cite{BCCMSBZ2006}.
+The \emph{ISP}-friendly design of \emph{Ono} however has been questioned when 
+it comes to Internet-scale~\cite{PMJKA2009,LCLX2009,CLYSR2011};
+its qualitative rating stands as follows:
+%%AD rephrased as above...
+% Even though the results in the paper claim minimization of inter-ISP traffic
+% cost while enhancing the, already, near optimal performance of BitTorrent
+% download, similar to \cite{BCCMSBZ2006}, this ISP-friendly design of \emph{Ono}
+% is questioned, especially on a Internet-scale-basis \cite{PMJKA2009,LCLX2009,CLYSR2011}.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{center}
@@
 % \subsubsection{Locality-Awareness in BitTorrent-like P2P Applications}
-Three approaches that ``inject'' locality awareness 
-in {\sl BitTorrent}-like applications are discussed in~\cite{LCLX2009}. The
-first \emph{macroscopic-level} approach
+In~\cite{LCLX2009}, three techniques that 
+``inject'' locality awareness in {\sl BitTorrent}-like applications are discussed.
+The first \emph{macroscopic-level} technique
 focuses on improving the neighbor selection process. 
@@
 %%
-The second approach, applied in an \emph{intermediate-level}, manipulates the
+The second technique, applied in an \emph{intermediate-level}, manipulates the
 {\sl BitTorrent} chocking/unchocking mechanism. A peer unchockes its $4$
 closest in terms of 
-% VM I put ``autonomous system'' back in again because in the context of this
+%%COMM-VM I put ``autonomous system'' back in again because in the context of this
 % paper we are dealing on the number of AS crossings (one-hop = one AS crossing)
-% untli we reach a neighbor.
-Autonomous System
-hop-count interested neighbors. 
+% until we reach a neighbor.
+autonomous system hop-count interested neighbors. 
 The same applies also when the peer turns to 
@@
 %%
-Finally, the third approach that operates at 
-a \emph{microscopic-level}, substitutes the rare-first chunk picking of the
+The last technique operates at 
+a \emph{microscopic-level} and substitutes the rare-first chunk picking of the
 original {\sl BitTorrent} protocol that picks the rarest of the available
-chunks to download next, with a locality-first policy which picks first the
-chunck that is closer. The distance value of a chunk is computed as the mean
+chunks to download next, with a locality-first policy; this picks first the
+chunck is located closer. 
+%%AD check slight rewording above
+The distance value of a chunk is computed as the mean
 value of the distances of the peers that posses it.
-Through experimentation on a real-world Internet topology simulated on
-PlanetLab, the locality-based approach achieves traffic optimization.
+Through experimentation on a real-world Internet topology simulated in
+{\sf PlanetLab}, the locality-based approach achieves traffic optimization.
 Nevertheless, it does not do that well when compared with the
@@
 
-All methods target in exposing the locality of the basic functions of the
-BitTorrent protocol. Even though they do not restructure the network to map
-better to the underlying infrastructure they favour the intra-AS communication
-over inter-AS communication thus reducing the general cost of communication.
-Again, it is pointed out that there is a trade-off between reducing inter-domain
-traffic and fairness among peers in terms of the data the peers upload
-
+%%AD changed the wording - check.. 
+All $3$ techniques in~\cite{LCLX2009} seek to harness 
+% in exposing 
+the locality of the basic functions of the
+{\sl BitTorrent} protocol. 
+Even though they do not restructure the network to map
+better to the underlying infrastructure, they favor the intra-\emph{AS} 
+over inter-\emph{AS} communications thus, reducing the overall communication cost.
+Again, there is a trade-off between reducing inter-domain
+traffic and fairness among peers in terms of the data the peers upload~\cite{LCLX2009};
+as far as the $3$ criteria, the overall approach fares as follows:
+%%
 \begin{center}
@@
 % \subsubsection{TopBT}
-\emph{TopBT}~\cite{RTLCGZ2010} seeks to enhance 
-proximity awareness in the {\sl BitTorrent}
-without the need of any additional infrastructure. The key conjecture upon
-which \emph{TopBT} is built is that
+The key objective of \emph{TopBT}~\cite{RTLCGZ2010} is to enhance
+proximity awareness in {\sl BitTorrent}
+without the need of additional infrastructure. 
+which \emph{TopBT} is built on the conjecture that
 a good peer selection metric should take into account both the
@@
 In an effort to essentially improve on the inherent topology mismatch problem,
-\emph{TopBT} modifies the peer list the tracker returns when someone first tries
-to connect to the swarm, the initial connection establishment, and the
-unchoking mechanism in the {\sl BitTorrent}. The important feature of TopBT is
-that it does not need wide deployment of its same types in order to see results.
-It also supports both link-hop and AS-hop metrics to identify proximity in
+\emph{TopBT} modifies: 
+\begin{inparaenum}[1)]
+\item the peer list the tracker returns when someone first tries
+to connect to the swarm, 
+\item the initial connection establishment, and 
+\item the unchoking mechanism in the {\sl BitTorrent}.
+\end{inparaenum}
+\emph{TopBT}'s key feature is
+that it does not need wide deployment.
+% of its same types in order to see results. %%AD I do not get the second part of the sentence
+% that it does not need wide deployment of its same types in order to see results.
+It also supports both link-hop and \emph{AS}-hop metrics to identify proximity in
 different levels.
@@
 The \emph{Underlying Topology-Aware Peer Selection (UTAPS)}~\cite{LCY2008}
-is an effort to furnish an enhanced peer selection strategy. \emph{UTAPS}
-essentially works in two stages: in the first,
+presents an enhanced peer selection strategy. 
+\emph{UTAPS} operates in two stages: in the first,
 it collects information to develop a picture of the underlying topology 
@@
 it to obtain some basic knowledge 
-including IP address, routers involved, \emph{RTT} and number of 
+including \emph{IP} address, routers involved, \emph{RTT} and number of 
 hops traversed.
@@
 The tracker provides this information to the newcomer in the form
-of the bootstrapping peer list. These newly joined peers traceroute the tracker
+of the bootstrapping peer list. 
+These newly joined peers traceroute the tracker
 peer list themselves and return back to the tracker in order to further enhance
@@
 the routers observed during the initial traceroute.
-
-It is reported in~\cite{LCY2008} that peers running \emph{UTAPS}
+Peers running \emph{UTAPS}
+% It is reported in~\cite{LCY2008} that peers running \emph{UTAPS}
 instead of the random peer selection, achieve better downloading rates
-and at the same time offer reduced burden on the underlying \emph{ISP}s. 
+and at the same time offer reduced burden on the underlying \emph{ISP}s~\cite{LCY2008}. 
 \emph{UTAPS} stands as follows in regard to the $3$ criteria:
@@
 {\sl BitTorrent} peer, trace-routes the tracker and stores vectors containing
-info like the IP address and the hop number of a traced router, as well as the
+information like the \emph{IP} address and the hop number of a traced router, as well as the
 link delay of the traced router and the previous hop along the trace-route path.
-These vectors are reported to the tracker and the latter uses this information
+These vectors are reported to the tracker which uses
+this information
 to classify the router through a $k$-means classification algorithm, with $k = 3$. 
@@
 % \subsubsection{Peer-exchange Routing Optimization Protocols (PROP)}
-In~\cite{QCYCZ2007}, two protocols termed 
-\emph{Peer-exchange Routing Optimization Protocols (PROP)} 
+In~\cite{QCYCZ2007}, two 
+\emph{Peer-exchange Routing Optimization Protocols (PROPs)} 
 are introduced to help adjust the neighborhood graph of an overlay network
 while aiming at reducing the network's overall link latency.
-\emph{PROP} algorithms are weaved around the concept of
+\emph{PROPs} are weaved around the concept of
 ``exchange'' of neighbors among peers; 
@@
 choose its own strategy. 
-
-%%
-%% VM ADDING SOME OPERATIONAL DETAILS BELOW.
-The PROP algorithms operate, basically, into two phases. During the \emph{warm-up}
-phase a node $u$ probes its neighbors to collect distance information. Then it
-contacts, at a fixed time rate $timer$, a random node $v$, $nhops$ away from it. New distance
-information is calculated independently by nodes $u$ and $v$, now for the
-hypothetical state when the potential exchange would occur. If the benefit
-gained, in terms of reducing the average distance among the peers involved, is
-above some predefined threshold then the exchange will actually occur,
-otherwise no further action will be taken. The \emph{maintenance} phase, on the
-other hand, differs from the warm-up one, in that the probability
-of peers to be probed and the $timer$ interval that this happens, now depends on
-past peer exchange results.
+%%
+%%COMM--VM ADDING SOME OPERATIONAL DETAILS BELOW.
+The \emph{PROPs} follow two phases: during the \emph{warm-up},
+a node $u$ probes its neighbors to collect distance information. 
+Subsequently, $u$ contacts, at a fixed time rate \emph{timer}, 
+a random node $v$, $nhops$ away. % from it. 
+New distance information is calculated independently 
+by nodes $u$ and $v$, now for the
+hypothetical state when the potential exchange would occur. 
+If the benefit gained, in terms of reducing the average distance among the peers involved, is
+above some predefined threshold then the exchange does occur.
+Otherwise, no further action is taken. 
+The second \emph{maintenance} phase
+differs from the warm-up in that the probability
+of peers to be probed and the \emph{timer} interval that this happens
+now depends on past peer exchange results.
 
@@
 \emph{PROP-O} permits peers to exchange selected sets of their neighbors. It is
-not allowed to exchange neighbors allong the path of the nodes that perform the
+not allowed to exchange neighbors along the path of the nodes that perform the
 exchange in order to ensure that in the end of the process the peers will remain
@@
 nodes), so the exchange always involves equal number of neighbors.
-
-The following table depicts the behavior of the \emph{PROP} 
-algorithms regarding the stated criteria:
+%%
+The following table depicts the \emph{PROP}s behavior 
+regarding the stated criteria:
 %%
@@
 partitioning.
-%% VM NEW ATTEMPT TO EXPLAIN THE ALGORITHM
-Upon arrival the newcomer peer $n$ asks a \emph{host-cache} to establish its
-random connections. For establishing its sibling connections, the newcomer
-multicasts a specialized message $l$, to locate the right candidate(s).
-Initially, the behaviour of this message, is modeled as a random walker,
+
+%%AD slight rewording throughout...
+Upon arrival, a peer $n$ asks a \emph{host-cache} to establish its
+random connections. 
+To identify its siblings, $n$ multicasts a specialized message $l$ 
+to locate the right candidate(s).
+Initially, the behavior of this message, is modeled as a random walker,
 until $l$ reaches some peer $d$ capable to guide it through its next step.
@@
 contains information for nodes accessible in an $r$-hop radius from $d$.
-Ultimatelly $l$ will reach a node $m$, candidate to become $n$'s sibling. $m$
-will issue a broadcast message to its own siblings and then all togeather will
+Ultimately $l$ will reach a node $m$ that is candidate to become $n$'s sibling. 
+Then, $m$ issues a broadcast message to its own siblings and then all in the group will
 inform the initiating peer $n$ of who is willing to accept new connections.
-
-The population and maintenance of ZoneCaches across the network is an important
-task for the efficiency of the constructed overlay. While seeking to address the
-mismatch problem, \emph{DDNO} is a heuristic approach that uses
+%%
+The population as well as the maintenance of \emph{ZoneCaches} across the network
+are key concerns when it comes to the efficient construction of the overlay.
+While seeking to address the mismatch problem, 
+\emph{DDNO} is a heuristic approach that uses
 \emph{Split-Hash} and \emph{dnMatch} algorithms to effectively encode domain
-names and determine domain locality respectively. In terms of the $3$ criteria,
-\emph{DDNO} fares are follows:
+names and determine domain locality respectively. 
+In terms of our $3$ criteria, \emph{DDNO} fares are follows:
 \begin{center}
@@
 \emph{Internet Assigned Numbers Authority (IANA)} and 
-the respective \emph{Regional Internet Registry (RIR)}'s IP assignments.
-%% VM CHANGED DESCRIPTION
+the respective \emph{Regional Internet Registry (RIR)}'s \emph{IP} assignments.
+%%
 Based on these assignments, nodes that belong in the same 
-organization are always addressed from the same block of IP addresses.
+organization are always addressed from the same block of \emph{IP} addresses.
 \emph{CTAG}, exploits this observation to construct topology-aware unstructured
@@
 overlay as well as its constant and dynamic adaptation.
-In the first phase,
-called \emph{bootstrapping grouping},
+In its first phase, called \emph{bootstrapping grouping},
 the {\sl Gnutella} web caching mechanism
-is modified so that a newcomer may choose the closest, in terms of AM,
-cache to retrieve the bootstrapping list. Similarly, during the second phase,
-called \emph{dynamic revision}, the standard connection establishement mechanism
-is also altered so that nodes with the highest AM metric are chosen. When a node
-reaches its maximum number of neighbor
-connections, it disconnects those neighbors with the lowest AM.
+is modified so that a newcomer may choose the closest, in terms of \emph{AM},
+cache to retrieve its bootstrapping list. 
+Similarly, during its second stage,
+called \emph{dynamic revision}, the standard connection establishment mechanism
+is also altered so that nodes with the highest \emph{AM} metric are chosen. 
+When a node reaches its maximum number of neighbor
+connections, it disconnects those neighbors with the lowest \emph{AM}.
 \emph{CTAG} behaves as follows regarding the $3$ criteria:
@@
 Internet.%, as shown in Figure~\ref{fig:landmark_binning}.
-To detect locality, peers mainly use network latency (i.e. \emph{RTT}).
+To detect locality, peers mainly use network latency (i.e., \emph{RTT}).
 %  as a measurement technique. 
@@
 physical locations need to be installed in strategic positions 
-across the the Internet. 
+across the Internet. 
 It is conjectured~\cite{RHKS2002} that $12$ such servers can prove
@@
 %% VM I CHANGED THE ABOVE AS BELOW TO REPHRASE IT.
-Each permutation of the set of landmarks represents a specific bin. Should there
-$m$ landmarks be adopted, potentially $m!$  different bins exist. Peers that
-end up with the same such ordering, belong to the same bin, in the sense that
-if they experience similar overall latencies from fixed network points then it
-is very likelly that they are close to each other.
-%%
-The operation of the method is independent of the model 
-incorporated by the overlay network and thus it can be
-applied with no significant changes to 
-both structured and unstructured \p\ systems.
-%%
-The main disadvantage of the approach is that 
-landmark servers must be installed and maintained throughout the world.
+Each permutation of the set of landmarks represents a specific bin. 
+Should there $m$ landmarks be adopted, $m!$  different bins potentially exist. 
+Peers that end up with the same ordering, belong to the same bin, in the sense that
+if they experience similar overall latencies from fixed network points; it
+is also  very likely that the peers in question are close to each other.
+%%
+
+As the operation of the method is independent of the model 
+incorporated by the overlay network, the approach can be
+applied with no significant changes to either
+structured or unstructured \p\ systems.
+%%
+The rather obvious disadvantage of the approach lies in the
+landmark servers that must be installed and maintained throughout the Internet.
 Provided that a \p-system often may have a few million nodes 
 connected at any given point in time, landmark servers do 
-inherently play a pivotal role in the operation of the network. Moreover,
-studies like \cite{GSG2002}, pin point the fact that end-host shoulder the
-latency measurments as the scheme's scalability bottlneck while others argue
-about the uneven distribution of nodes potentially resulting in load imbalance
-\cite{CDCR2002,M2003,ZZZSZ2004}, its coarse-graned nature that fails to
-effectively distinguish among realatively close nodes \cite{RGJZ2004} and the
-potentially increased maintenance cost \cite{XTZ2003}.
-
+inherently play a pivotal role in the operation of the network. 
+The approach has received a fair amount of criticism.
+For instance, \cite{GSG2002} points out that end-hosts may bear the
+cost of latency measurements and so, they may become bottlenecks.
+\cite{CDCR2002,M2003,ZZZSZ2004}~argue that uneven distribution of nodes 
+ultimately lead to load imbalances and \cite{RGJZ2004} indicates 
+that its coarse-grained nature fails to
+effectively distinguish among relatively close nodes.
+Finally, \cite{XTZ2003} underscores the potentially increased maintenance cost(s).
+%%AD rephrased what is below as above... The ONE sentence below cannot be read!!!!
+%%%
+% Moreover, studies like \cite{GSG2002}, pin point the fact that end-host shoulder the
+% latency measurements as the scheme's scalability bottleneck while others argue
+% about the uneven distribution of nodes potentially resulting in load imbalance
+% \cite{CDCR2002,M2003,ZZZSZ2004}, its coarse-grained nature that fails to
+% effectively distinguish among relatively close nodes \cite{RGJZ2004} and the
+% potentially increased maintenance cost \cite{XTZ2003}.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%%
+%
 % Although latency estimation does not drain network resources,
@@
 % excessive network traffic flow through its landmark server-clusters.
-
+%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % \subsubsection{mOverlay}
-The \emph{mOverlay} \cite{ZZZSZ2004} approach addresses scalability issues
+The \emph{mOverlay}~\cite{ZZZSZ2004} approach addresses scalability issues
 that might arise when static landmark servers are in use.
@@
 while at the bottom, only \emph{intra-group} connections occur among peers.
-
-%%AD how do you identify the initial groups????
-%%
-%% VM REPHRASING IN ORDER TO ANSWER THE QUESTION ABOVE
+%%
+
 Clearly, identifying groups and accurately finding the closest group 
@@
 the groups already in the network, rendering
-the latter be the \emph{dynamic landmarks} in the process. For a newcomming
-peer $Q$ the \emph{grouping critirion} says that when the distance of $Q$ and
+the latter be the \emph{dynamic landmarks} in the process. 
+%%
+For a incoming peer $Q$, 
+the \emph{grouping criterion} designates that when the distance of $Q$ and
 some group $A$’s neighbor groups is the same as the distance
 between group $A$ and group $A$’s neighbor groups, then host $Q$
-should belong to group $A$. During network initialization or when grouping
+should belong to group $A$. 
+During network initialization or when the grouping
 criterion is not met, new groups are created.
-It is shown that in the proposed scheme, a new peer can 
-reach its group by expending at most $O(logN)$ messages
-within the network. 
+A peer may reach its group by expending at most $O(logN)$ messages.
+%%AD simplified sentence below as above...
+% It is shown that in the proposed scheme, a new peer can 
+% reach its group by expending at most $O(logN)$ messages
+% within the network. 
 Finally, \emph{mOverlay} maintains stability and constant 
@@
 this is achieved through periodic cache updates and group 
-leader selections, should one either leaves or dies.
+leader selections, should a node either leaves or dies.
 %
@@
 %%
-In terms of the stated criteria, \emph{mOverlay} fares as follows:
+In terms of the stated $3$ criteria, \emph{mOverlay} fares as follows:
 \begin{center}
@@
 we came to identify four key methodologies 
-utilized by the discussed approaches; these methodologies are:
+utilized by the discussed approaches; they are:
 \begin{enumerate}[\itshape i\upshape)]
@@
 \end{enumerate}
-%%VM Added the following to mention the hybrid/combo methodologies
-An important aspect of many implementations, is that since the above
-methodologies use heuristics to solve topology mismatch, an
-intrinsically difficult problem, no single approach can offer a robust solution.
-Thus, many of the algorithms we surveyed are actually combinations of multiple
-such methodologies, in an attempt to refine the results of using just a single one.
-For example, \emph{mOverlay} combines topology adaptation with landmark binning or \emph{Gia}
-combines all three topology adaptation, forwarding optimization and caching
-methodologies. Some times combination results in win/loose trade-offs. For
-instance, the effectiveness of a caching/replication component can be undermined
+%
+A key characteristic shared by many of the presented techniques 
+is that they do not strictly adhere to a single methodology.
+As discussed approaches attempt to address the inherently challenging problem 
+of topology mismatch, they resort to heuristics and so, 
+they do not often furnish ``pure'' solutions (i.e., solutions that exclusively belong to 
+one of the $4$ aforementioned methodologies). 
+%%AD rephrased as above..
+% An important aspect of many implementations, is that since the above
+% methodologies use heuristics to solve topology mismatch, an
+% intrinsically difficult problem, no single approach can offer a robust solution.
+% Thus, many of the algorithms we surveyed are actually combinations of multiple
+% such methodologies, in an attempt to refine the results of using just a single one.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+For instance, \emph{mOverlay} combines topology adaptation with landmark binning 
+while \emph{Gia} weaves together $3$ methodologies:  topology adaptation, 
+forwarding optimization and caching. 
+This combination inevitably leads to feature/performance trade-offs: for example, 
+the effectiveness of a caching/replication component can be undermined
 by a continuously adapting overlay that removes important links between peers.
-For this reason, the reader is advised to criticaly approach either the overall
-algorithms or the separate methodologies when evaluating them for different
-application domains.
-
-Below, we discuss how the proposed protocols use elements of the four
+If a technique is to be evaluated for a specific application domain,
+its overall approach as well as its underlying used methodologies have to be considered.
+%%AD rephrased as above... did not like much this "advised" - is not wrong BUT sounds funny
+% For this reason, the reader is advised to critically approach either the overall
+% algorithms or the separate methodologies when evaluating them for different
+% application domains.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Below, we outline how the proposed protocols use elements of the four
 methodologies and we offer a summary qualitative comparison for all surveyed
@@
 %%VM I brought the following paragraph here from the end of this subsection.
-Moreover, Table~\ref{unstructured:table} offers a summary overview for each effort,
-showing which of the aforementioned methodologies it incorporates, along with its
-highlights and a rough estimation of the positive and negative aspects of its
-implementation.
+Moreover, Table~\ref{unstructured:table} offers a summary overview for each effort;
+in particular, the Table shows which of the aforementioned methodologies 
+every effort incorporates, the highlights of the effort as well as 
+a rough estimation of the pros and cons of in its implementation.
 
@@
 modify the topology of the \p\ network
-using various techniques. The two most commonly used approaches to
-topology adaptation are creating
-\emph{spanning tree}s using connection graphs
-and creating \emph{clusters} of physically close nodes.
+using various schemes. The two most commonly used
+create \emph{spanning tree}s using connection graphs
+as well as  \emph{clusters} of physically close nodes.
 
 \emph{Narada} and subsequent algorithms including 
-\emph{AOTO, LTM, SBO}, try to solve the problem
-by building a richer connected graph and forming 
+\emph{AOTO, LTM, SBO} try to solve the problem
+by building ``a richer connected graph'' and by forming 
 minimum spanning trees over
-this graph that can efficiently route messages among peers. The \emph{AOTO, LTM}
-and \emph{SBO} try to
-overcome this limitation with ingeneous schemes
+this graph that can efficiently route messages among peers. 
+\emph{AOTO, LTM} and \emph{SBO} try to
+overcome this limitation with ingenious schemes
 like forming minimum spanning trees
-for the $2$-hop away neighbors for each node, seperating participant nodes into
-groups, sometimes with different responsiblities and tasks at hand (e.g.,
-\emph{SBO}). The advantage of building minimum spanning trees is that
+for the $2$-hop away neighbors for each node, separating participant nodes into
+groups, sometimes with different responsibilities and tasks at hand (e.g.,
+\emph{SBO}). 
+The advantage of building minimum spanning trees is that
 they maintain the connectivity on the network in
 an efficient manner while still preserving the overall search scope.
-However, their construction and the update costs,
-especially in dynamic, high-churn environments,
-causes large additional traffic overhead on the underlying
+However, their construction and their update costs,
+especially in dynamic and high-churn environments,
+may cause large traffic overheads on the underlying
 network~\cite{CRZ2000,CRSZ2001,CRSZ2002}.
@@
 Further enhancements may include dynamic local restructuring of the overlay
-graph through neighbour exchange like in \emph{PROP} or cycle-cut like in
-\emph{DCMP} to achieve continoous adaptation thoughout a peer's lifecycle.
+graph through neighbor exchange like in \emph{PROP} or cycle-cut like in
+\emph{DCMP} to achieve continuous adaptation throughout a peer's life-cycle.
 Unfortunately, commonly used methods for proximity detection across Internet do
@@
 average path traversing cost from one node to another.
-In doing so, topology adaptation re-arranges the
+In doing so, the methodology re-arranges the
 overlay network so that it becomes a better fit 
-for the underlying \emph{IP} network.
+for the underlying \emph{IP} network
+as Figure~\ref{figure:topology-adaptation} depicts.
+%%AD u always need a reference of a Figure in the text!!!
 % as Figure~\ref{figure:topology-adaptation} depicts.
@@
 inter-ISP communication accounting and management.
-
+%%
 \begin{figure}[ht]
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%AD slight wording changes below - check
 \subsubsection{Forwarding Optimization Methodology}
-In unstructured systems there is no way to know where data are located. That is
-why early works like Gnutella \cite{gnutellav04} used blind flooding, a Breadth
-First Search approach with depth $d$, a system-maximum TTL value (in terms of
-hops) for messages. This is inefficient since it generates a large number of
-messages, a lot of them duplicates. Approaches based on
-\emph{forwarding-optimization} propose intelligent search mechanisms to try to
+In unstructured systems, there is no way to know where data are actually located. 
+This is why early works like Gnutella \cite{gnutellav04} used blind flooding, 
+a \emph{BFS}--approach with depth $d$ and a system-maximum \emph{TTL} value 
+(in terms of hops) for messages. 
+This is inefficient since it generates a large number of
+messages, a lot of which are duplicates. 
+Approaches based on \emph{forwarding-optimization} propose 
+intelligent search mechanisms in an attempt to
 make searching more efficient.
@@
 Many alternative schemes have been proposed for this category, some of which we
-have surveyed; iterative deepening, modified BFS, local indices \cite{YG-M2002},
+have surveyed; iterative deepening, modified \emph{BFS}, local indices \cite{YG-M2002},
 $k$-walker random walks \cite{LCCLS2002} or heterogeneity exploitation and
 token-based flow control mechanisms \cite{CRBLS2003}. More sophisticated
-alternatives also exist, taking into account application level requirements
+alternatives also exist, taking into account application--level requirements
 or even exploiting machine learning techniques to adjust overlay network
@@
 Forwarding-optimization schemes in unstructured \p\ systems can be classified
-as DFS- or BFS-based. Routing indices \cite{CG-M2002} for example is a
-DFS-based technique while all the aforementioned are BFS. Also they can
+as \emph{DFS-} or \emph{BFS}-based. Routing indices \cite{CG-M2002} for example is a
+\emph{DFS}-based technique while all the aforementioned are \emph{BFS}. Also,
+the schemes in question can
 be classified as deterministic or probabilistic (probabilistic, random or
@@
 scope (in Fig.~\ref{figure:forwarding-optimisation:after} the object on
-the far right is not reached) thus limiting the scalability of the whole
+the far right is not reached) thus, limiting the scalability of the whole
 network. 
-Moreover, they address the problem of the mismatch in a limited manner
-since they do not provide any guarantees 
+Forward-optimization suggestions address the problem of the mismatch 
+in a limited manner since they do not provide any guarantees 
 that overlay and underlying topologies are
-aligned with each other let alone quantify the mismatch 
-and try to alleviate.
-Forwarding based optimisations are commonly applied in
+aligned with each other.
+%% let alone quantify the mismatch and try to alleviate.
+%%AD I got the above out as superfluous
+To this end, forward-optimization techniques are commonly applied in
 conjunction with other methodologies to improve 
-the quality of the \p\ systems.
+the overall quality of \p\ systems.
 
@@
 transfer of data. Caching has with much success been successfully 
-adopted by web and file server application environments. 
+adopted by web-- and file--server application environments. 
 Since peers in a \p\ system also operate as servers,
-it is intuitively expected that \p\ file sharing systems can also benefit from
+it is intuitively expected that \p\ file--sharing systems can also benefit from
 caching in improving performance and reducing overall resource usage. 
@@
 already been developed in some commercial \p\ systems, like 
-\emph{KaZaA} or less well known like \emph{Gia} and \emph{BNS} as we saw during
-the survey.
+\emph{KaZaA} or less popular ones including \emph{Gia} and \emph{BNS}.
 %%%
@@
 distance measurement method to position themselves with respect to ``a priori''
-known servers on the Internet, like \emph{Ono} which uses the CDN-infrastructure
+known servers on the Internet, like \emph{Ono} which uses the \emph{CDN}-infrastructure
 for this purpose.
-These landmark servers are used
-by nodes to estimate their positions based on the intuitive assumption that
-nodes
+These landmark--servers are used by nodes to estimate 
+their positions based on the intuitive assumption that nodes
 with similar distances to a set of landmarks, are physically close to each
 other, as well, over the network. 
-In Figure~\ref{figure:landmarking} the newly
+In Figure~\ref{figure:landmarking}, the newly
 arriving peer, measures its distance to an array of landmark point denoted as
@@
 order to choose the already participating peers with which to connect to,
-compares its list to the list of the potential neighbor. Peers with similar
+the peer compares its list to the list of the potential neighbor. Peers with similar
 measured distances to the these landmark points are likely to be close to
@@
 
-Landmark based protocols have four important drawbacks: First, the network
+% Landmark--based protocols have four important drawbacks: First, the network
+%%AD the correct here is "first" NOT "First"!
+Landmark--based protocols have four important drawbacks: first, the network
 delay is not a reliable distance estimation method. For example, based on the
-load on the network, the delay to certain nodes or networks can change from time
-to time, which will eventually affect the distance measurements and wrong
-measurements will lead to wrong estimated positions for the nodes, or incorrect
-and non optimal clusterings of the nodes. Second, relying on predefined nodes
-make the whole paradigm not fully distributed and the landmark system prone to
-become a single point of failure. Third, using
-landmark servers requires costly installation and maintenance of landmark
-infrastructure across the whole Internet and for all the different autonomous
-system domains. 
-As popular \p\ file sharing applications usually have millions
-of peers connected at any time, the network costs
-of maintaining these landmark servers will likely be quite high. 
+load on the network, the delay to certain nodes or networks can significantly 
+change over time; this eventually affects the distance measurements and yields wrong
+measurements that ultimately lead to wrong estimated positions for the nodes 
+or incorrect and non--optimal clusterings of nodes. 
+Second, relying on predefined nodes
+makes the whole paradigm not fully distributed and the landmark system prone to
+single point of failure. Third, 
+landmark infrastructure require costly installation and maintenance 
+across the Internet and the different autonomous system domains. 
+As popular \p\ file--sharing applications usually have millions
+of peers connected to at any time, the network costs
+of maintaining these landmark servers can be quite high. 
 A possible solution to
 the scalability problem of the static landmark servers is to use ordinary nodes
-as dynamic landmarks, like in \emph{mOverlay} once they estimate their own positions. Even though this
-approach scales much better than static landmark servers, the measurement
-accuracy problem affects the overall performance of the system. Moreover,it can
+as dynamic landmarks in an approach reminiscent to that of \emph{mOverlay}.
+% once they estimate their own positions. 
+Even though this scales better, the accuracy of measurements
+may affect the overall performance of the system. Moreover,it can
 be characterized as a rather coarse-grained approximation, therefore not
-particularly well suited for detecting the correct positions of nodes within
+particularly well suited for identifying the correct positions of nodes within
 close distance to each other.
+%%AD check the above made various small changes...
 

### 2014-02-02 structured.tex
2013.10.29 (Alex) Structured section underwent several reworks/rephrasings. Text reviewed. Minor changes to unstructured section.

@@
 
-This section presents a description and analysis of algorithms proposed
-to address the topology mismatch problem in structured \p\ networks. The
-categorization of the approaches we present here is based on how different
-levels of proximity information is gathered and the way peer routing tables are
-maintained and used in an effort make overlay hops as efficient as possible. The
-categorization is based on previous work found in the
+This section offers description and analysis of algorithms proposed
+to address the topology mismatch problem in structured \p\ networks. 
+The categorization of the approaches is based on how different
+levels of proximity information is gathered as well as  
+the way peer routing tables are maintained and used in an effort make hops in
+overlays as efficient as possible. 
+%%AD what does it really mean this below???
+The categorization is based on previous work found in the
 literature~\cite{CDHR2002,CDCR2002,RSS2002}.
 
-%% VM Rephrased as above to keep it analogous to the opening paragraph of
+%%COMM Rephrased as above to keep it analogous to the opening paragraph of
 %% the unstructured part.
@@
 \emph{Global Soft-State} \cite{XTZ2003} builds a global map to help
-choose shorter routing paths, combining the landmark binning method and small
-scale distance probes to reveal the proximity properties of the underlying
-network. %Figure \ref{fig:global_softstate} presents the landmark
+choose shorter routing paths. This is accomplished by 
+combining the landmark binning method with small
+scale distance probes to reveal the proximity properties 
+of the underlying network. 
+%Figure \ref{fig:global_softstate} presents the landmark
 %space and the mapping of the position of the nodes.
-This global view of the
-state, is made available to all nodes to help them find the best way to
-route their messages. \emph{Global Soft-State} operates in two phases:
-generation and use of the proximity information. For generating the proximity
-information, a hybrid approach is proposed, which uses landmark clustering as a
+This global view of the state is made available to 
+\emph{all} nodes to help them find the best way to route their messages. 
+%%AD when you say "all" above means what?? all in the structured network?? 
+%%	how is the structured network organized?  (a line or 2 for this?)
+\emph{Global Soft-State} operates in $2$ phases:
+\emph{generation} and \emph{use} of the proximity information. 
+For \emph{generating} the proximity information, 
+a hybrid approach is proposed: it uses landmark clustering as a
 preprocessing step to select a number of potential nearest neighbor
-candidates and then refine the selection by incorporating a
-round-trip-time-based scheme to ultimately choose the closest node. For using
-the proximity information, the algorithm follows a different path from that of the
-classic gossiping approaches for the construction and maintenance of the
-overlay. It is based on landmark clustering for generation and strategic
-placement of proximity information on the overlay. This information is called a
-``map'' and concerns a region of the overlay (part of the Cartesian space in
-structures like eCAN\cite{xu_ecan_2002} or a set of nodes sharing a particular
-prefix in overlays such as Pastry\cite{antony_pastry_2001}). Thus, each node not
-only can access such information from these landmarks but can also subscribe to
-relevant \emph{soft states} and get notified on dynamic network changes. A
-node may appear in a maximum of $log\left( N \right)$ such
-maps.
-%%
-Although this approach greatly reduces the routing latency to far nodes, it
-comes expensive. It is possible to end up with a very large number of regions
-each of which has to maintain a map. The fact that a peer may appear in multiple
-such maps we can conclude that additional traffic for probing and notifications
-renders the approach not scalable \cite{RGJZ2004}. Moreover, to refine measurements
-additional inner-bin measurement s are required\cite{WZS2004}. Also improvement
-is done within a region because of little knowledge about neighboring zones.
-
-%% VM Removed most of the paragraph below which contains info that we cannot back
+candidates and then, refines the selection by incorporating a
+round-trip-time-based scheme to ultimately choose the closest node. 
+%% 
+The algorithm follows a different path from that of the
+classic gossiping approaches for the construction and 
+maintenance of the overlay when 
+it comes to \emph{use} the proximity information.
+%%AD isn't the following sentence somewhat of repetition??? (of what 
+%%	is stated above?? 
+The approach is based on landmark clustering 
+for generation and strategic
+placement of proximity information on the overlay. 
+This information is called a ``map'' and concerns a 
+region of the overlay; 
+the latter may be part of the Cartesian space in
+structures like \emph{eCAN}~\cite{xu_ecan_2002} or 
+a set of nodes sharing a particular
+prefix in overlays such as \emph{Pastry}\cite{antony_pastry_2001}. 
+%%
+Thus, each node not only can access such information from 
+these landmarks but can also subscribe to
+relevant \emph{soft states} and get notified on dynamic network changes. 
+A node may appear in a maximum of $log(N)$ such maps.
+%%
+Although this approach does reduce the routing latency to far nodes, it
+can become expensive as it is 
+possible to end up with a very large number of regions
+each of which has to maintain a map. 
+The fact that a peer may appear in multiple
+such maps points into additional traffic for probing and notifications
+and so renders the approach not scalable \cite{RGJZ2004}. 
+Moreover, to refine measurements
+additional inner-bin measurements are needed~\cite{WZS2004}.
+Also, improvements within a region can be minimal as there exists
+limited knowledge about neighboring zones.
+%%AD rephrased what is below -2 lines- as above..
+% Also improvement
+% is done within a region because of little knowledge about neighboring zones.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%% COMM Removed most of the paragraph below which 
+%%		contains info that we cannot back
+%%
 %% Rescanned papers and wrote the paragraph below.
@@
 % traditional DHTs.
-
+%
 %
@@
 %
-
 %
@@
 %
-
+In terms of our $3$ criteria, the approach fares as follows:
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 \emph{Mithos} \cite{WR2003} is a \p\ protocol which incorporates directed
-incremental probing to find near optimal node placement, and is an integration
-of geographic layout and proximity routing overlay optimization methods.
-
-The bootstrap phase of \emph{Mithos} starts with a subset of existing members as
-the first set of candidates, and while the closest nodes are detected by iteratively
-probing the neighborhood, the candidate neighbor list is updated. To
-avoid local minima, \emph{Mithos} probes all the neighbors within a two-hop distance
-from the current minimum before concluding the process.
-
-After finding its first neighbor, the new-coming peer is assigned an ID using
-information gathered during the iterative neighbor selection phase. Based on the
-two closest nodes found, along with these nodes neighbors and the corresponding
+incremental probing to find near optimal node placement and integrates
+geographic layout and proximity routing overlay optimization methods.
+%%AD in general some mention of the mismatch problem must be made right???
+%%		so that a reasonable connection can be established...
+
+The bootstrap phase of \emph{Mithos} starts with a subset 
+of existing members as the initial set of candidates and 
+while the closest nodes are detected by iterative
+probing of the neighborhood, the candidate neighbor--list gets updated. 
+To avoid local minima, \emph{Mithos} probes all the neighbors 
+within a two-hop distance from the current minimum 
+before concluding the process.
+
+After finding its first neighbor, an 
+% new-coming   %%AD overall the term new-coming is somewhat odd.. incoming? 
+incoming peer is assigned 
+an \emph{ID} using information gathered during 
+the iterative neighbor selection phase. 
+%%AD use SIMPLE sentences that are well structured to say what u want
+Based on the two closest nodes found, along with these nodes neighbors and the corresponding
 distances a system of virtual springs is set so that the tension of each spring
-is inversely proportional to the distance. The system is then left to settle,
-and the state in which it settles is used for its ID. The benefit of these
-synthetic coordinates are that
-they are explicitly used as the node's ID, and distance computations among nodes
-can be done in ID space without requiring physical measurements.
-
-The last step of the algorithm is the interconnection among neighbors. \emph{Mithos}
-uses a \emph{quadrant-based} mechanism in which each node establishes
-a link to the closest neighbor in each quadrant. During forwarding, the next
-hop is performed towards a neighbor in the same quadrant as the final
-destination. The new coming node may not know of other neighbors in all
+is inversely proportional to the distance. 
+%%AD the above is impossible to read.. "based on,... along with... " NO!!!
+%%AD I am not sure what you are talking about here - I mean below... rework
+%%	and simplify.. 
+The system is then left to settle,
+and the state in which it settles is used for its \emph{ID}. 
+The benefit of these synthetic coordinates are that
+they are explicitly used as the node's \emph{ID}, 
+and distance computations among nodes
+can be done in \emph{ID} space without requiring physical measurements.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% upto here %%%%%%%%%%%%%%%%
+
+The last step of the algorithm is the 
+interconnection among neighbors. 
+\emph{Mithos} uses a \emph{quadrant-based} mechanism in which 
+each node establishes a link to the closest neighbor in each quadrant. 
+%%AD what is the quadrant??? Hugh??? 
+During forwarding, the next hop is performed towards a neighbor 
+in the same quadrant of the final destination. 
+An incoming node may not know of other neighbors in all
 quadrants, therefore, the node first identifies neighbors in all quadrants
-using a mechanism based on ideas similar to a perimeter walk\footnote{Used in
-the Greedy Perimeter Stateless Routing (GPSR) protocol.} and then improves the
-results using parallel path processing by taking into account further geometric
-properties of node relationships.
+using a mechanism based on ideas similar to a perimeter walk~\footnote{Used in
+the Greedy Perimeter Stateless Routing (\emph{GPSR}) protocol.}  %%AD ref??
+and then improves the results using parallel path processing 
+by taking into account further geometric properties of node relationships.
+%%AD the above 5-6 lines sounds impressive but there is little help 
+%%	for the average ignorant reader.. what is perimeter walk in 
+%%	general? how quadrants are formed? What is their relationship?
 
 One major limitation of \emph{Mithos} is that the protocol cannot effectively
-handle dynamic arrivals and departures from the network. In \cite{CCRK2004} it
-is claimed to need extensive probing to determine distances and at the same time
-its results are evaluated as less accurate than, for example,
-GNP \cite{ng_gnp_2001}. Maybe this has also to do with its iterative neighbor
-selection process which is reported to be prone to prematurely terminate at
-local minima \cite{WSS2005}. Another point that has negative effect to the
-scalability of the algorithm is that parts of its spring relaxation technique
-require a centralized implementation~\cite{cox_vivaldi_2004}.
-
+handle dynamic arrivals and departures from the network. 
+\cite{CCRK2004}~points out the extensive probing to determine 
+distances and evaluates \emph{Mithos} results as less accurate
+than those derived with \emph{GNP}~\cite{ng_gnp_2001}. 
+%%AD what is GNP??? How did it "land" here???
+%%
+The latter has been also attributed by~\cite{WSS2005} to 
+the iterative neighbor selection process which that is 
+reportedly prone to premature termination at local minima.
+%%
+\cite{cox_vivaldi_2004} argues that elements of the protocol's
+spring relaxation technique require a centralized implementation 
+affecting \emph{Mithos} scalability.
+%%AD rephrased what is below as above...
+% Another point that has negative effect to the
+% scalability of the algorithm is that parts of its spring relaxation technique
+% require a centralized implementation~\cite{cox_vivaldi_2004}.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In regard to the $3$ criteria, \emph{Mithos} stands as follows:
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 % \subsubsection{LAPTOP}
-\emph{LAPTOP} \cite{WLH2007}, organizes the overlay into a tree-based hierarchy
-with the goals of reducing hops during message routing as well as minimizing
-maintenance overhead. Additionally, a caching scheme is also incorporated so as
-to further reduce routing table update costs. It is theoretically shown
-that in \emph{LAPTOP}, routing path length is bounded by $O(log_d N)$ and node
+\emph{LAPTOP} \cite{WLH2007} organizes the overlay 
+%%AD what is the nature of the "structure" overlay here???
+into a tree-based hierarchy so that 
+hops required during message routing be reduced and 
+maintenance overhead be minimized.
+A caching scheme is also used so as
+to further lower routing table update costs. 
+It is theoretically shown that 
+\emph{LAPTOP} routing path length is bounded by $O(log_d(N))$ and node
 joining and leaving in the overlay network is bounded by
-$O\left( d log_d N \right)$ hops in a balanced overlay tree, where $N$ is the
-number of nodes, and $d$ is the maximum degree of each node. \emph{LAPTOP}
-implements the geographical layout approach  and constructs its layout in a
-self-organizing and efficient fashion. The physical distances are estimated
+$O(d*log_d(N))$ hops in a balanced overlay tree with $N$ representing the
+number of nodes and $d$ the maximum degree of each node. 
+\emph{LAPTOP} 
+%% implements 
+adopts the geographical layout approach  
+and constructs its layout in a self-organizing and 
+efficient fashion. 
+The physical distances are estimated
 with round trip times to a few existing nodes in the overlay network.
-
-Each node is assigned a dot-formated address (e.g 1.3.4). Each octet ranges from
-$1$ to $d$, where $d$ is the maximum degree of the nodes. The assignment process
-is done by appending a unique octet to the address of each node's parent, while
+%%
+Each node is assigned a dot-formatted address (e.g., 1.3.4). 
+Each octet ranges from
+$1$ to $d$.
+%  where $d$ is the maximum degree of the nodes. 
+%%AD this definition was provided just above.. 
+The assignment process is done by appending a unique octet 
+to the address of the parent of each node, while
 the root node is assigned address $1$.
@@
 
-The routing scheme is similar to the longest-prefix IP matching scheme. At each
-forwarding hop, any message travels up the tree until the first common ancestor
-of both source and destination nodes is reached and then starts descending to
-arrive to its target. During tree traversals, special entries in the routing
-tables, called \emph{routing cache}, are maintained to increase routing
-efficiency and achieve finer load balance. Caching enables a node to forward a
-message to a better longest-prefix match than that of its direct ancestor making
-a larger, quicker and more cost effective step through the overlay and toward
-the destination. To improve scalability, the number of child nodes and the
-size of the routing cache are limited. In terms of overlay maintenance, \emph{LAPTOP}
-incorporates a simple \emph{heartbeat}-based technique where each parent node is
-responsible for monitoring its children.
-
+\emph{LAPTOP}'s routing scheme is similar to the 
+longest-prefix \emph{IP} matching scheme. 
+At each forwarding hop, a message travels up the tree 
+until the first common ancestor of both source and destination nodes 
+is reached and then, the message starts descending towards its target. 
+During tree traversals, special entries in the routing tables, 
+called \emph{routing cache}, 
+%%AD clarify - routing caches are the special entries OR the routing tables???
+are maintained to increase routing efficiency and achieve finer--load balance. 
+Caching enables a node to forward a message to a better 
+longest-prefix match than that of its direct ancestor yielding
+a larger, quicker and more cost--effective step through the overlay towards
+the destination. 
+To improve scalability, the number of child nodes and the
+size of the routing cache are limited. 
+To contain maintenance costs, \emph{LAPTOP} employs 
+a simple \emph{heartbeat}-based regime in which 
+a parent is responsible for the monitor of its own children.
+%%AD rephrased what is below as above.. 
+% In terms of overlay maintenance, \emph{LAPTOP}
+% incorporates a simple \emph{heartbeat}-based technique 
+% where each parent node is
+% responsible for monitoring its children.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 %
-
+%
 % At join process the new coming node is assigned its level label as well as its
@@
 %the change in the hierarchy.
-
+%%%
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
-
+\emph{LAPTOP}'s behavior for the $3$ criteria stands as follows:
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 
+In order to optimize the underlying network use,
 \cite{KLKP2008} introduces proximity
-controls over the vanilla Kademlia protocol's random ID assignment to
-optimize the underlying network usage.
+controls over the \emph{vanilla Kademlia} 
+protocol's random \emph{ID} assignment.
 % The work in this paper
@@
 % iterative lookup algorithms.
-The overlay tier improves its routing performance using a cost metric that is
-provided by the underlying tier. Such information can be acquired by
+In its quest to improve routing performance, the overlay uses 
+a cost ``yardstick'' (i.e., metric) provided by the underlying network.
+% The overlay tier improves its routing performance using a cost--metric that is
+% provided by the underlying tier. 
+Such cost information can be acquired by
 \begin{inparaenum}[\itshape i\upshape)]
   \item measuring previous lookups,
-  \item exchanging measurements of other peers or jointly calculating them with
+  \item exchanging or jointly-computing measurements with 
 	others, or 
-  \item using a local database to look up information.
+  \item using a local database to look--up information.
 \end{inparaenum}
 
-Regarding routing optimization both techniques of \emph{proximity routing
-selection (PRS)} and \emph{proximity neighbor selection (PNS)} are studied.
-%%
-In \emph{PRS}, the goal is to choose the best next hop during routing a message.
-% Due to the fact that the routing in Kademlia is iterative, it is the initiator
-% of the lookup that chooses each next hop from a set of candidate nodes.
-Plain old Kademlia (\emph{POK}) always chooses the closest node with respect to
-the XOR metric but with proximity routing, Kademlia should also choose the node
-that exhibits the smallest underlay metric cost, balancing a trade-off between
-the overlay and underlay progress distances.
-%%
-In \emph{PNS}, on the other hand, the goal is to keep peers with the least
-contact cost in the routing table. As Kademlia continuously learns about new
-peers via incoming requests and an iterative lookup process, no special
-algorithm for searching more cost-effective peers is necessary and it simply
-chooses the best peers seen so far.
-
-Proximity injection in Kademlia improves locality of connections between peers.
-Towards the enhancement of traffic locality, it is suggested to use information
-from the MaxMind GeoIP database\footnote{http://www.maxmind.com} and use it to
-form clusters of nearby peers. An other alternative method could use Vivaldi
-coordinates \cite{cox_vivaldi_2004}, however, it is suggested that the
-clustering approach performs better. Specifically when combined with both
-\emph{PRS} and \emph{PNS} preserves more than $40\%$ of the messages within an
-ISP and allows only a small $6\%$ to go through intercontinental backbones.
-
+\cite{KLKP2008} exploits $2$ routing techniques: 
+\emph{proximity routing selection (PRS)} and 
+\emph{proximity neighbor selection (PNS)}.
+%%
+In \emph{PRS}, the goal is to choose the best next hop while routing a message.
+\emph{Plain old Kademlia} (\emph{POK}) always selects the closest--node 
+with respect to the \emph{XOR} metric.
+With \emph{PRS} in place however, \emph{Kademlia} may also choose 
+the node that exhibits the smallest underlay metric cost; in this way, 
+it balances a trade-off between the logical overlay and 
+actual underlying distances.
+%%AD rephrased as above.. the "underlay" I am not sure what it is...
+%  but with proximity routing, Kademlia should also choose the node
+% that exhibits the smallest underlay metric cost, balancing a trade-off between
+% the overlay and underlay progress distances.
+%%
+In \emph{PNS}, on the other hand, the goal is to keep peers 
+with the least contact cost in the routing table. 
+%%AD WHAT is this "least contact cost"?? Sounds funny! CHECK
+As \emph{Kademlia} continually learns about new
+peers via incoming requests and an iterative lookup process, 
+%%AD what is this iterative lookup process???
+no specialized algorithm is required for searching for 
+more cost-effective peers 
+and the system simply goes for the best peers ``seen'' thus far.
+
+\emph{Kademlia} proximity injection improves 
+locality of connections among peers.
+To enhance traffic locality, the use of 
+{\sl MaxMind GeoIP} database~\footnote{\url{http://www.maxmind.com}} 
+is suggested to form clusters of nearby peers.
+To this end, Vivaldi coordinates~\cite{cox_vivaldi_2004} could be 
+also used as an alternative for morphing locality. 
+\cite{KLKP2008} however suggests that the clustering approach 
+when combined with \emph{PRS}/\emph{PNS} helps contain 
+$40\%$ of messages within an \emph{ISP} and allows only a $6\%$ 
+to go through the backbone.
+The behavior of the proposal with respect to the stated criteria is deemed as follows:
+%%
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 \emph{CHOP6} \cite{MT2007} roughly estimates the proximity among
-nodes by exploiting the IPv6 address format and RTT information, if available.
-The proximity estimation is achieved by introducing a 64-bit ID scheme in which
-the least significant bit part is the IPv6 global routing prefix and thus
-enabling a longest prefix match scheme. The protocol is designed based on the
-observation that it is possible to estimate a node's geographical location by
-simply examining the upper 32-bits of its IPv6 address. Moreover, similar to
-Chord\cite{stoica_chord_2001}, CHOP6 uses a finger table, whose entries hold
+nodes by exploiting the {\sl IPv6} address format and if available, 
+\emph{RTT} information.
+The proximity estimation is achieved by introducing a $64$-bit 
+\emph{ID} scheme in which
+the least significant bits are the {\sl IPv6} global routing prefix.
+Thus, a longest--prefix match scheme is established.
+% enabling a longest prefix match scheme. 
+%%AD check is the word "longest" is correct..
+The protocol is designed based on the
+observation that it is feasible to estimate a node's geographical location by
+examining the upper $32$-bits of its {\sl IPv6} address. 
+Moreover, similar to \emph{Chord}~\cite{stoica_chord_2001}, 
+\emph{CHOP6} uses a finger table, whose entries hold
 more than one candidate node.
-
 %
@@
 %
-
+With respect to the $3$ qualitative criteria, \emph{CHOP6} stands as follows:
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 \end{tabular}
+}
 \end{center}
+%%AD how does CHOP6 helps with the mismatch problem??
 
@@
 % \subsubsection{T2MC}
-\emph{T2MC} \cite{SLCGZ2008} uses traceroute logs to detect the AS boundaries and
-cluster nearby peers together to reduce the redundant multiple message
-passes between the AS boundaries. \emph{T2MC} uses a customized k-mean classification
-algorithm with $k = 2$ to perform the classification, and exploits the stable
-structure of the Internet routers to guide clustering. Even though trace-route
-provides detailed information about the network structure, use of trace-route
-creates overhead on the overall network structure. For this same reason, it is
-not unusual for network administrators to disable this support on their network
-routers, which may affect the performance of the \emph{T2MC} algorithm.
-
-The results of T2MC are encouraging but it is also considered coarse-grained
-since it classifies peers into only ``remote'' and ``close''
+\emph{T2MC} \cite{SLCGZ2008} uses traceroute logs to detect 
+the boundaries of an autonomous system (\emph{AS}) and
+cluster nearby peers; the objective here is to 
+diminish the redundant multiple messages that may 
+unnecessarily cross \emph{AS}-boundaries.
+\emph{T2MC} uses a customized $k$-mean classification
+algorithm with $k = 2$ to perform the classification 
+and exploits the stable structure of the Internet routers 
+to help guide the clustering in question. 
+Even though {\sl traceroute}
+provides detailed information about the network structure, 
+it extensive use may create overhead on the overall network structure. 
+In this respect, it is often the case that administrators 
+disable {\sl traceroute} support on their routers.
+The latter may affect \emph{T2MC}'s performance.
+%%
+Although overall experimental results have been encouraging, 
+\emph{T2MC} is considered coarse-grained
+as it classifies peers into only ``remote'' and ``close''
 groups \cite{QLZG2009}.
-
+%%AD a better connection with the mismatch problem has to be made... 
+%%		.. in the form of 1-2 lines (or short statements). 
+%%
+%
 %T2MC \cite{SLCGZ2008} uses trace-route logs to detect different autonomous
@@
 %and edge gateways.
-
+%
 %
@@
 %
-
+\emph{T2MC} fares as follows:
+%%
 \begin{center}
@@
 % \subsubsection{PChord}
-\emph{PChord} \cite{HLYW2005} is based on the Chord DHT and adds
-proximity into its routing mechanism. The main modification over the
-vanilla Chord, is the inclusion of a \emph{proximity list} into its routing table
-so that the next hop is decided based on both the best progress
+\emph{PChord} \cite{HLYW2005} is based on the \emph{Chord DHT} and adds
+proximity into its routing mechanism. 
+The main modification over its predecessor is the inclusion 
+of a \emph{proximity list} into its routing table;
+in this way, the next hop can be decided based on both the best progress
 towards the key and the physical proximity of the candidate nodes.
-The list is, at join time, empty but as the \emph{PChord} node starts to interact with
-other nodes it applies a heuristic mechanism to fill the list. Entries are
-dynamically added or removed as the network state changes. The overlay
-maintenance cost is similar to Chord's. Apart from the costs incurred by node
-join/leave and by heart-beat messages for connection verification, \emph{PChord} adds
-heart-beat among nodes inside the proximity list. Fortunately, since the list
-contains nodes with the lowest RTT, the added cost is kept to a minimum.
-
-Due to the fact that, with each step, there is always progress towards the
-target node in the ID space, \emph{PChord} will result in a reduced hop number than
-Chord, since each hop in \emph{PChord} is larger or at least equal in the key space.
-Passing through proximity links, in the underlying network, means reduced routing
-cost. Additionally, if the number of entries in the proximity list is the same
-as the number of network partitions, \emph{PChord} prevents hops from jumping in and
-out of the network partition the current node belongs to.
-
-\emph{PChord}’s advantages include the fact that the main routing optimizations are of
-less overlay hops, following proximity links and crossing the same network
-partition only once along a routing path while in the same time keeping Chord's
-light-weight maintenance cost. On the other hand, it exhibits slow convergence
-speed and inefficiency in a high churn environment \cite{DK2006}.
-
+%%AD I am not sure about the "best progress" term - what does it really mean???
+%%
+At join time, this list in empty. 
+As the \emph{PChord} node starts to interact with
+others, it applies a heuristic mechanism to fill the list. 
+Entries are dynamically added or removed as the network state changes. 
+The overlay maintenance cost is similar to that of \emph{Chord}. 
+%%
+Apart from the costs incurred when a  node joins in or departs and 
+by heart-beat messages used for connection verification, \emph{PChord} adds
+heart-beat among nodes it the proximity list. 
+Fortunately the latter is kept to a minimum provided that 
+the list entails nodes presenting very low \emph{RTT}s.
+%%AD removed "lowest" to use lowest u must have a set or something...
+
+As there is always progress towards the target in the \emph{ID}-space,
+\emph{PChord} yields a reduced number of hopes that \emph{Chord}.
+%%AD rephrased as above - the 2 last lines of the text below are 
+%%		are difficult to decipher... 
+% Due to the fact that, with each step, there is always 
+% progress towards the target node in the ID space, 
+% \emph{PChord} will result in a reduced hop number than
+% Chord, since each hop in \emph{PChord} is larger or 
+% at least equal in the key space.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+Passing through proximity links, in the underlying network, 
+means reduced routing cost. 
+Moreover, if the number of entries in the proximity list is the same
+as the number of network partitions, 
+\emph{PChord} prevents hops from jumping in and
+out of network partitions the current node belongs to.
+%%
+On one hand, 
+\emph{PChord}’s advantageous  points are that 
+its routing attains fewer overlay hops, 
+it crosses network partition(s) only once by following proximity links
+and it maintains the light-weight maintenance cost of \emph{Chord}.
+On the other, \emph{PChord} exhibits slow convergence speed 
+and inefficiency in high-churn environments~\cite{DK2006}.
+%%AD rephrased what is below as above... 
+% \emph{PChord}’s advantages include the fact 
+% that the main routing optimizations are of
+% less overlay hops, following proximity links and crossing the same network
+% partition only once along a routing path while in the same time keeping Chord's
+% light-weight maintenance cost. 
+% On the other hand, it exhibits slow convergence
+	% speed and inefficiency in a high churn environment \cite{DK2006}.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\emph{PChord} qualitatively behaves as follows:
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 % \subsubsection{AChord}
-\emph{AChord} \cite{DK2006} uses IPv6's any-cast mechanism to
-relieve the protocol from complex joining procedures, and achieve high accuracy
-of network proximity and thus, high routing efficiency. Any-cast delivers a message
-coming from the outside of an any-cast group to the physically closest node in
-that group \cite{M2002}. \emph{AChord} organizes all nodes participating in the
-overlay network into an any-cast group. Any joining node comes outside of that group
-and, thus, is
-automatically forwarded to the physically nearest node in order to bootstrap,
-avoiding the need to, explicitly, maintain such nodes, and the effort of finding
-a way of locating the physically nearest among them. The ID of the new-coming
-node is computed based on the ID of the bootstrap node and the bootstrap's
-predecessor in a way that the constructed ID will be positioned between the
-aforementioned two. After joining, the finger table is built the same way as in
-the Chord protocol. Moreover, additional nodes are maintained in a structure
-called \emph{adjacency table} which stores information about the closest known
-node. The routing decision is made by using both the adjacency and finger
-tables.
-
+\emph{AChord} \cite{DK2006} uses the {\sl IPv6} any-cast mechanism to
+relieve the protocol from complex joining procedures,
+harness network proximity and achieve high routing efficiency. 
+Any-cast delivers a message coming from the outside of 
+an any-cast group to the physically closest node in
+that group~\cite{M2002}. 
+\emph{AChord} organizes all nodes participating in the
+overlay network into an any-cast group. 
+Any joining node comes outside of that group and, thus, is
+automatically forwarded to the physically nearest node for bootstrapping;
+this avoids the need to explicitly maintain such nodes, %%AD such nodes???
+and the effort of locating the physically nearest. 
+%%
+The \emph{ID} of an incoming 
+% the new-coming %%AD changed...
+node is computed based on the \emph{ID} of the bootstrap node 
+and the bootstrap's predecessor in a way that the constructed 
+\emph{ID} is positioned between the aforementioned two. 
+After joining, a \emph{finger} table is created ala \emph{Chord}. 
+Moreover, an \emph{adjacency table} maintains 
+information about the closest--known peers. 
+The routing decision is made by using both adjacency and finger tables.
+%%AD how are these 2 combined???
+\emph{AChord} behavior has as follows:
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 % \subsubsection{Chord6}
-\emph{Chord6} \cite{XZHL2005} is another Chord variant that tries to
-exploit the hierarchical features of IPv6 to create a substrate that
-reduces inter-domain traffic between service providers. The difference between
-\emph{Chord6} and Chord relies on the identifier definition. Therefore, the approach
-can be easily portable to other DHTs such as CAN, Pastry and Tapestry. In \emph{Chord6}
-the identifier contains two parts: the higher bits are obtained by hashing the
-node's IPv6 address prefix of specific length, while the remaining lower bits
-are the hash value of the rest of the IPv6 address. As a result of this
-assignment, nodes in a domain will be mapped onto a continuous key space on the
-overlay network, which avoids unnecessary message forwarding across different
-service providers, thus minimizing overall routing cost.
-
-This approach is able to reduce the average path length of the system to be
-logarithmically proportional to the number of available domains instead of the
-number of participating peers. However due to the large number of Internet
-domains and the potentially small number of system peers within each domain,
-this approach doesn't seem scalable and efficient \cite{DK2006}.
-
+\emph{Chord6}~\cite{XZHL2005} is a \emph{Chord}-variant that
+exploits {\sl IPv6}'s  hierarchical features in order to create 
+a substrate that reduces inter-domain traffic among service providers. 
+The key difference between
+\emph{Chord6} and \emph{Chord} lies on the identifier definition. 
+Therefore, the approach can be easily portable to other 
+\emph{DHT}s including \emph{CAN}, \emph{Pastry} and \emph{Tapestry}. 
+\emph{Chord6}'s identifier contains two parts: 
+the higher bits are obtained by hashing the
+node's {\sl IPv6} address prefix of specific length, 
+while the remaining lower bits are the hash-value 
+of the rest of the {\sl IPv6} address. 
+As a result of this assignment, nodes in a domain are mapped 
+onto a continuous key space on the overlay network;
+the latter avoids unnecessary message--forwarding across different
+service providers and helps minimize overall routing cost.
+
+\emph{Chord6} does reduce the average system path length to be
+logarithmically proportional to the number of available domains 
+instead of the number of participating peers. 
+However, due to the large number of Internet
+domains and the possibility of occasionally having only 
+small number of peers within each domain,
+\cite{DK2006} challenges the approach's scalability and efficiency.
+%%
+Overall, the approach fares as follows:
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 % \subsubsection{DHT-PNS}
-The work in \cite{DLTZZ2006} propose a proximity neighbor selection
-scheme on top of the Chord DHT. The main goal is to use proximity information
-to group physically close by nodes as neighbors in the DHT table. In order to
-detect proximity, virtual network coordinates of peers are used, by using the
-Vivaldi protocol \cite{cox_vivaldi_2004}. The virtual coordinates are then used
-by the nodes to map to identifier space in the DHT. The space is partitioned
-using a \emph{concentric circle clustering scheme} where successive cycles of
-radiuses $\rho$, $2\rho$, $3\rho$ and so on, are constructed. Then the formed
-annuluses are divided into $2\chi-1$ \emph{sectors}, where $\chi$ denotes the
-sequence number of the annulus starting from $\chi = 1$ for the center cycle. It
-is proved in the paper, that this way each sector occupies the same area as does
-the center cycle. Assuming uniform node distribution, this characteristic,
-favors a more load balanced clustering operation. Every sector in this $2d$
-coordinate space is mapped to a unique \emph{region} in the DHT space forming a
-multi-layer node identifier space.  Thus, any nodes that belong to the same
+A proximity neighbor selection scheme  functioning atop 
+the \emph{Chord DHT} is introduced in~\cite{DLTZZ2006}.
+Its key objective here has been to group physically-close nodes as 
+neighbors in the \emph{DHT} table while exploiting proximity information.
+To detect proximity, virtual network coordinates for 
+peers are used with the help of the Vivaldi protocol~\cite{cox_vivaldi_2004}. 
+The virtual coordinates are then used to help map nodes to 
+the identifier--space of the \emph{DHT}.
+%%
+The space in question is partitioned using a 
+\emph{concentric circle clustering scheme} where successive cycles of
+radiuses $\rho$, $2\rho$, $3\rho$ and so on, are constructed. 
+The formed annuluses are then divided into 
+$2\chi-1$ \emph{sectors}, where $\chi$ denotes the
+sequence number of the annulus starting from $\chi = 1$ 
+for the center cycle. 
+In this way, 
+\cite{DLTZZ2006} prove that each sector
+occupies the same area as does the center cycle. 
+Assuming uniform node distribution, this characteristic,
+favors a more load--balanced clustering operation. 
+Every sector in this $2d$ space is mapped to 
+a unique \emph{region} in the \emph{DHT} space forming a
+multi-layer node identifier space.  
+Consequently, nodes belonging to the same
 sector, are mapped to the same region as well, preserving their proximity
-relationship unveiled by the use of the Vivaldi protocol. The individual pieces
-are mapped to the identifier space uniquely, allowing logarithmic lookup
-operations with high probability on Chord. The system described in the previous
-section allows any node to obtain its DHT key and its region key, in a fully
-distributed manner, just by applying a consistent hash function. Using an RPC
-call in Chord, the node can obtain the region's master node, called
-\emph{Cluster Node (CN)} which is responsible for clustering the nodes belonging
-in the same sector or region with that of the node at hand. An additional RPC
-call, registers it to its corresponding region and publishes its information,
-through the special peer CN. The node can, additionally, ask CN for other
+relationship.
+%% unveiled by the use of the Vivaldi protocol. 
+%%AD I am not sure u need the above line at the end of the previous sentence.
+Individual elements
+are uniquely mapped to the identifier--space, 
+allowing logarithmic lookup operations with high 
+probability on \emph{Chord}. 
+%%
+
+%%AD what is the section u say here??? -- also the first sentence 
+%%	here has to do a better connection with what was discussed before!
+The system described in the previous section 
+allows any node to obtain its \emph{DHT} key and its region key, in a fully
+distributed manner, just by applying a consistent hash function. 
+%%AD where these RPCs come from? kind of disconnected???
+Using an {\sl RPC} call in \emph{Chord}, 
+the node can obtain the region's master node, 
+called \emph{Cluster Node (CN)} which is responsible for clustering 
+the nodes belonging in the same sector or region with that of the node at hand. 
+An additional {\sl RPC} call, registers it to its corresponding region 
+and publishes its information, through the special peer \emph{CN}. 
+The node can, additionally, ask an \emph{CN} for other
 nodes that have previously joined the region to add them into its
-neighbor set for future routing table optimization. Even in the case when no
+neighbor set for future routing table optimization. Q
+%%AD the one below is not a parse-able sentence!!! What does it say???
+%%%		restrict the length of your sentences to say at most 2.5 lines!
+Even in the case when no
 neighbor is detected in the current region, the search is expanded to adjacent
@@
 property of the delay distribution and good optimizing performance.
-Unfortunately, even though scaling seems to be also smooth, the simulations
-where made on rather small test beds (of at most 2048 nodes), so no safe
-conclusions can be deduced. Moreover it assumes uniform node distribution which
-is not always the case, especially in a synthetic coordinate system.\\
-
+%%AD up to here the text in this second paragraph for Chord DHT remains 
+%%%%%		... rather cryptic - does need work and better connection 
+%%%%%		with the 1st paragraph...
+%%
+Unfortunately, even though scaling appeared promising, only limited 
+simulations carried out with configuration involving up to 2,048 nodes 
+were carried out.
+% where made on rather small test beds (of at most 2048 nodes), so no safe
+% conclusions can be deduced. 
+Moreover the approach assumes uniform node distribution which
+is not always the case, especially in a synthetic coordinate system.
+%%
+\emph{Chord DHT} fares as follows in reference to the $3$ criteria:
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 % \subsubsection{Quasi-Chord}
-In \cite{SZ2008} proposes \emph{Quasi-Chord}, a Chord-based network that
-constructs its overlay network with underlying topology in mind. The network
-construction follows three steps. First, each host acquires its coordinates in
-the geometric space utilizing the \emph{global network position (GNP)} protocol
+\cite{SZ2008} proposes \emph{Quasi-Chord}, a \emph{Chord}-based network that
+constructs its overlay network while attempting to match the underlying 
+network topology.
+The overlay creation follows $3$ stages:
+first, each host acquires its coordinates in
+a geometric space using the \emph{Global Network Position (GNP)} protocol
 \cite{ng_gnp_2001} and a set of landmarks.
-Second, using the \emph{Cantor space filling curve}, the $2$-dimensional space
-is converted to a $1$-dimensional one, used in the last, third, step to build
-the \emph{Quasi-Chord} circle according to the host's Cantor value. Physical
-proximity is thus denoted by Cantor value proximity. The Chord cycle is
-constructed by sorting nodes in ascending order. In \emph{Quasi-Chord}
-each node stores 2 finger tables; one clockwise and one counter-clockwise.
-
+Second, using the \emph{Cantor} space--filling curve, 
+the $2$-dimensional is converted to $1$-dimensional space. 
+The latter is employed in the third and final stage to build 
+the \emph{Quasi-Chord} circle according to the host's \emph{Cantor} value. 
+Physical proximity is thus denoted by \emph{Cantor} value proximity. 
+The \emph{Chord} cycle is constructed by sorting nodes in 
+ascending order. In \emph{Quasi-Chord},
+each node stores $2$ finger tables; one clockwise and one counter-clockwise.
+%%AD what is the purpose of this? efficiency to go back and forth???
+%%
+%
 %In the following paragraphs, these steps are further discussed.
@@
 %
-
+%
 %
@@
 %
-
+%%
+\emph{Quasi-Chord} qualitative behavior has as follows:
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 % \subsubsection{IP-based Clustering (IPBC)}
-Proximity neighbor selection algorithms use probing and other techniques to
-detect proximity. However, such methods are either not precise enough, or
-overload the network. \emph{IP-based clustering} \cite{KM2007} is a proximity
-neighbor selection based algorithm which uses the IP address prefixes (16 bit
-for IPv4) in order to detect proximity. \emph{IPBC} clustering generates a key
-and its prefix and stores the prefix in the DHT itself, so that any newly
-joining nodes with the same IP prefix can query it and
-identify all the neighbors with the same prefix easily.  Nodes
-periodically update their entries in the DHT by removing those who leave, or
-those who have been timed-out.
-
-The algorithm's overhead is very low because detection of proximity is only
+Proximity neighbor selection algorithms often use probing 
+% and other techniques
+%%AD say what are these "other" than simply say so... 
+to detect proximity. 
+However, such methods suffer by imprecisions often due to network overload.
+{\sl IP}-based clustering~\cite{KM2007} 
+% is a proximity neighbor selection algorithm which  %%AD u do not need this!
+uses {\sl IP} address prefixes ($16$ bit for {\sl IPv4}) 
+to detect proximity. 
+%%AD where this IPBC came from?
+\emph{IPBC} clustering generates a key and its prefix %%AD this ... "and" is not 						      %%    clear!!!
+and stores the prefix in the \emph{DHT} itself;
+in this way, newly joining nodes with the same {\sl IP} prefix 
+can query the \emph{DHT} and readily
+identify all neighbors with the same prefix.  
+Nodes periodically update their entries in the \emph{DHT} 
+by removing those who leave or have timed-out.
+%%
+
+The overhead of the algorithm is very low as proximity detection requires only
 a comparison of prefixes. On the other hand, the efficiency of the proximity
-is caught into a vicious cycle with the storage space needed for DHT object
-publishing (how many bits will be used for either). Moreover, its static
-nature (no online measurements used) makes it difficult to handle
-\cite{EWF2011}.
-
+is caught into a vicious cycle with the storage space needed for \emph{DHT} 
+object publishing (how many bits will be used for either). 
+%%AD did not get the parenthesis above - what is "either"???
+%%	I am not sure what the sentence below says and 
+%%	how is connect to the above
+Moreover, its static nature (no online measurements used) 
+makes it difficult to handle~\cite{EWF2011}.
+%%
 % Moreover, observing the
@@
 % Registries (RIRs), nodes inside will be physically close to each other.
-
+%
 %
@@
 %
-
+The suggested approach fares as follows:
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 \end{tabular}
+}
 \end{center}
 
+%%AD how the 2-level thing helps the mismatch  problem?? 
+%%	ALSO, why the symbolism of the layers is required?? Ld_Cord Ld_Cone
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % \subsubsection{Cone}
-\emph{Cone} \cite{HY2007} extends Chord using a proximity neighbor
+\emph{Cone} \cite{HY2007} extends \emph{Chord} using a proximity neighbor
 selection topology optimization algorithm. The proximity information is
-generated using landmarks and round-trip-time-based distances to landmarks.
-\emph{Cone} uses a two-layered identifier space. The first, named Chord-layer
-identifier, denoted as $Id_{Chord}$, is the same as in Chord. The second is the
-Cone-layer identifier, $Id_{Cone}$ which is constructed using two component
-identifiers. The first, known as \emph{group identifier (gid)} denotes a
-relevant group to which the node belongs while the second, namely \emph{local
-identifier (lid)} indicates the local identifier within the group. The group
-concept is a way of dividing nodes according to a
-common $Id_{Chord}$ prefix.  The structure of a Cone overlay, retains the
-Chord's circular topology. The difference lies in the fact that, now, two rings
-are created. A big ring, where nodes with the same \emph{gid} are arranged at
-each position. Each of these positions are a smaller ring for the particular
-group's \emph{lid}s. Routing is achieved in both clockwise and
+generated using landmarks and round-trip-time-based distances to 
+those landmarks. %%AD added those.. 
+%%AD how is this better/worse than the one just above based on IP clustering???
+\emph{Cone} uses a two-layered identifier space. The first layer is 
+identical to that of \emph{Chord} and is denoted as $Id_{Chord}$.
+% \emph{Chord}-layer identifier and denoted as $Id_{Chord}$, 
+% is the same as in Chord. 
+The second layer identifier known as \emph{Cone}-layer and denoted as 
+$Id_{Cone}$ is generated using two IDs. 
+The first, known as \emph{group identifier (gid)} denotes a
+relevant group to which the node belongs to while the second, 
+namely \emph{local identifier (lid)} indicates the % local 
+identifier within the group. 
+The group is used here as means to divide nodes according to  a
+% concept is a way of dividing nodes according to a
+common $Id_{Chord}$ prefix.  
+The \emph{Cone} overlay retains the \emph{Chord}'s circular topology. 
+The difference lies in the fact that, now, $2$ rings are created. 
+A big one in which nodes with the same \emph{gid} are arranged 
+% at each position.  %%AD I do not get thus at EACH position!!!
+in the same position.
+Each of these positions constitute a smaller ring for the particular
+group's \emph{lid}s. 
+%%%
+Routing is achieved in both clockwise and
 counter-clockwise directions in the big-ring, for which two routing tables are
-maintained, namely \emph{front} and \emph{back} finger tables. Entries in these
-tables, display physical network proximity with the current node. Moreover, a
-third table called \emph{group} maintains information about other on-line peers
-within the current node's group in a way that entries are now close in the id
-space.
-
+maintained: \emph{front} and \emph{back} finger tables. 
+Entries in these tables, display physical network proximity 
+with the current node. 
+Moreover, a third table called \emph{group} maintains 
+information about other on-line peers
+within the current node's group in a way that 
+entries are now close in the \emph{id} space. %%AD check rephrasing...
 %
@@
 %
-
+\emph{Cone} fares as follows: 
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 % \subsubsection{DynaMO}
-\emph{DynaMO} \cite{WZS2004} tries to build an overlay that not only 
-considers physical proximity amongst peers but their mobility attributes as well.
-To adapt to mobile, ad-hoc networks, special care is put in by \emph{DynaMO}
-to maintain an evenly distributed overlay ID space so that hot-spots are
-avoided. The paper's implementation is based on the Pastry overlay due to its
-built-in locality heuristics that have been extensively studied
-\cite{CDCR2002a} but \emph{DynaMO}'s mechanisms are DHT-independent so it can be
-applied to other DHT's.
-
-\emph{DynaMO}'s approach can be viewed bottom-up. A new-coming node is required
-to compute physical information about its neighborhood and use it to
-unilaterally assign itself the an ID. \emph{DynaMO} is building on the
-observation that with every new overlay routing hop the physical distance
-between nodes is likely to increase, resulting in a dominating last routing
-step \cite{antony_pastry_2001,CDCR2002a} and for this two methods are discussed,
-\emph{random landmarking (RLM)} and \emph{closest neighbor prefix assignment
-(CNPA)}.
-%%
-In \emph{RLM} a set of peers are chosen to be responsible for a set of landmark
-keys. These keys have to be chosen in a way that they equally divide the overlay
-space. The new-coming peer then computes distances to those
-landmarks, sorts them and assigns itself an ID based on the landmark ordering.
-In more detail, the peer adopts an ID prefix of length
-$prefix_length=|log_b k|$ (where $b$ is the ID base and $k$ the number of
-landmarks) from the landmark closest it and the rest can be assigned either
-randomly or computed with further proximity optimization in mind.
-What is achieved is that
+\emph{DynaMO}~\cite{WZS2004} attempts to create an overlay that 
+does not only consider physical proximity amongst peers but also
+takes into account mobility attributes of nodes.
+To adapt to mobile ad-hoc networks, \emph{DynaMO} strives
+to maintain an evenly--distributed overlay \emph{ID}-space 
+so that hot-spots are avoided. 
+%%
+Although based on the \emph{Pastry} overlay and 
+it well understood built-in locality heuristics ~\cite{CDCR2002a},
+\emph{DynaMO}'s own mechanisms are \emph{DHT}-independent and 
+so, they can be applied to other \emph{DHT}s.
+%%AD rephrased/shortened above - check
+%%
+\emph{DynaMO}'s approach may be viewed as bottom-up. 
+An incoming node computes physical information about its neighborhood
+and uses the outcome to unilaterally assign itself an \emph{ID}.
+The proposal is based the fact that with every new overlay  routing-hop
+the physical distance between nodes is likely to increase
+resulting in a dominating 
+last routing step~\cite{antony_pastry_2001,CDCR2002a}.
+To identify this step, two techniques are used: 
+\emph{Random LandMarking (RLM)} and 
+\emph{Closest Neighbor Prefix Assignment (CNPA)}.
+%%AD rephrased as above - check!
+% A new-coming node is required
+% to compute physical information about its neighborhood 
+% and use it to unilaterally assign itself the an ID. 
+% \emph{DynaMO} is building on the
+% observation that with every new overlay routing hop the physical distance
+% between nodes is likely to increase, resulting in a dominating last routing
+% step \cite{antony_pastry_2001,CDCR2002a} and for this two methods are discussed,
+% \emph{random landmarking (RLM)} and \emph{closest neighbor prefix assignment
+% (CNPA)}.
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In \emph{RLM}, a set of peers are chosen to be responsible 
+for a set of landmark keys. 
+These keys have to be chosen in a way that they equally divide the overlay
+space. 
+%%AD how do you accomplish the above?? 
+An incoming peer then computes distances from those landmarks,
+sorts them, and assigns itself an \emph{ID} on this 
+landmark ordering as follows:
+% The new-coming peer then computes distances to those
+% landmarks, sorts them and assigns itself an ID based on the landmark ordering.
+the peer adopts as \emph{ID} having a prefix from its closest
+landmark; the length is 
+$l_{prefix}=|log_b k|$, where $b$ is the \emph{ID} base 
+and $k$ the number of landmarks. 
+The remaining of bits of the peer-\emph{ID} are assigned either 
+randomly or computed with additional proximity optimizations.
+In this way, two desirable features are attained:
+% What is achieved is that
 \begin{inparaenum}[\itshape i\upshape)]
   \item landmark assignment is dynamic, and
-  \item a node’s leaf set is likely to reference physically close nodes.
+  \item the set of a node leaf may reference peers that are 
+	physically close by.
+	%%AD I am not sure why is that?? But it is ok...
 \end{inparaenum}
-Especially, the latter observation reduces the average routing cost because the
-leaf set is utilized for the last routing step (since it contains the nodes
-closest in the overlay space).
-%%
-\emph{CNPA} is proposed for network setups that want the bootstrap phase as lightweight
-as possible. It delegates to Pastry, the burden to identify the physically
-close node with which the new-coming peer will bootstrap. The new-comer assumes
-the bootstrapping nodes's ID prefix and fills the rest like \emph{RLM} does.
-
-\emph{RLM} achieves good proximity evaluation in a load balanced and adaptable to
-high churn environments fashion. Unfortunately it comes with more bootstrap
-overhead that can especially effect the set of temporary landmarks. On the other
-hand, \emph{CNPA}'s less incurred overhead comes at the expense of being more
-coarse-grained.
-
+The latter in particular helps reduce average routing cost(s)
+for the leaf-set of a node is used to efficiently determine the 
+last routing step.
+% (since it contains the nodes closest in the overlay space).
+%%AD I do not think you need the above.. %%AD Also rewriting..
+%%
+%%
+\emph{CNPA} is proposed for network setups for which a lightweight
+bootstrap phase is required.
+It delegates to \emph{Pastry} the task to identify the physically
+close--node with which an incoming peer gets bootstrapped. 
+The new-comer assumes the bootstrapping nodes's \emph{ID} prefix 
+and fills the remaining of its key using the techniques of \emph{RLM}.
+
+\emph{RLM} gets evaluated as presenting good proximity in a load-balanced 
+environments. Moreover, it is adaptable to high-churn settings.
+Unfortunately, \emph{RLM} bootstrap overhead is substantial and can 
+effect  the set of temporary landmarks.
+\emph{CNPA} on the other hand, incurs less overhead but this come 
+at the expense of being more coarse-grained.
+%%AD  Rephrased as above.  -- In the above I am not clear 
+%%	a) what are the "temporary" landmarks? b) what is coarse-grained 
+%%	and why??? 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \emph{RLM} achieves good proximity evaluation in a load balanced and adaptable to
+% high churn environments fashion. Unfortunately it comes with more bootstrap
+% overhead that can especially effect the set of temporary landmarks. On the other
+% hand, \emph{CNPA}'s less incurred overhead comes at the expense of being more
+% coarse-grained.
+%%%
+%%
 % \emph{DynaMO} tries to capitalize on the fact that in Pastry each node's table consists
@@
 % (CNPA)}.
-
 %
@@
 %
-
+\emph{DynoMO}'s elements fare as follows regarding our $3$ criteria:
 \begin{center}
@@
 % prevent them to maintain a good ratio in the presence of high landmark
-% degression rates.
+% digression rates.
 high \\
@@
 
-\emph{SAT-Match} \cite{RGJZ2004} defines an iterative process that attempts to
-adaptively change the overlay network to ease the topology mismatch problem. It
-consists of a \emph{probing phase} when a peer probes nearby nodes for distances
-and a \emph{jumping phase} when the peer decides to change neighborhood to
-reduce its average neighbor distances. This process is performed unilaterally
+\emph{SAT-Match}~\cite{RGJZ2004} defines an iterative process that 
+adaptively changes the overlay network to ease the topology mismatch problem. 
+It consists of a \emph{probing phase} in which a peer queries nearby nodes 
+for distances and a \emph{jumping phase} in which the peer decides 
+to change its neighborhood in order to reduce average neighbor distances. 
+This process is performed unilaterally
 until no additional optimization is needed but it achieves optimization within a
@@
 % definition of Topology Mismatch Problem in the intro/background.
-The paper defines \emph{stretch}
-$S = \frac{\bar{L_l}}{\bar{L_p}}$, where $\bar{L_l}$ is the average logical link
-latency while the $\bar{L_p}$ is the average physical link latency, and uses it
-a way to quantify the \emph{topology match degree} of the constructed overlay.
-
-The \emph{probing phase} starts when the new-comer joins the DHT using the
-standard mechanism of the algorithm at hand. The starts probing the neighborhood
-the messages of small TTL value $k$ \cite{jiang_lightflood_2008}. Recipients return
-information to the source and forward the message if not expired. After
-discovering its TTL-$k$ neighborhood, the source measures RTT times, sorts them
-and picks the two nodes with the smallest RTTs.
-
-In the \emph{jumping phase}, the source node calculates the stretch change of
-its $TTL-1$ neighborhood and that of the $TTL-1$ neighborhood of the first of
-the peers picked during the previous phase. These calculations are made as if
-the jump has been made. If the stretch reduction is over a predefined threshold
-the jump is performed, otherwise the second selected candidate is picked and the
-same computations are performed. If again, the threshold is not met, then no
-jump is ultimately done. In case of a jump, this is performed as a combination
+The notion of \emph{stretch} is used as th means for quantifying 
+the \emph{topology match degree} of the constructed overlay;
+stretch is defined as: 
+$S = \bar{L_l}/\bar{L_p}$ where $\bar{L_l}$ is the average logical link
+latency and $\bar{L_p}$ is the average physical link latency.
+%%AD what are the indexes: l and p -- p for peer and l for new peer???
+%%AD rephrased what is below as above... 
+% 
+% The paper defines \emph{stretch}
+% $S = \bar{L_l}/\bar{L_p}$, where $\bar{L_l}$ is the average logical link
+% latency while the $\bar{L_p}$ is the average physical link latency, and uses it
+% a way to quantify the \emph{topology match degree} of the constructed overlay.
+
+The \emph{probing phase} starts when the new-comer joins the \emph{DHT} 
+using the standard mechanism of the algorithm at hand. 
+The peer starts probing the neighborhood with 
+messages of small \emph{TTL=k}~\cite{jiang_lightflood_2008}. 
+Recipients return information to the source and forward the message 
+if not expired. 
+After discovering its \emph{TTL}-$k$ neighborhood, 
+the source measures \emph{RTT}-times, sorts them
+and picks the two nodes with the smallest \emph{RTT}s.
+
+During the \emph{jumping phase}, the source node calculates the 
+stretch change of its \emph{TTL--1} neighbors %%AD each of its neighbors???
+and that of the \emph{TTL-1} neighbors of the first of
+the peers picked during the previous phase. 
+These calculations are made as if the jump had been made. 
+Should the stretch reduction is over a predefined threshold
+the jump is actually performed. Otherwise,
+the second selected candidate is selected and the
+same computation is performed. 
+%%AD is this done only 2 times?? Or is it done recursively??
+If again, the threshold is not met, then no jump is ultimately carried out. 
+In case of a jump, this is performed as a combination
 of \emph{leave} and \emph{join} operations, in the underlying context
-(i.e. CAN).
-
-Even though no theoretical proof is given, valid results are
-obtained and demonstrated through experimentations. It is argued that since it
-is observed to achieve a $27\%$ stretch reduction when compared to a landmark
-enhanced CAN-system, it can be deduced that it performs better than landmark
-binning in terms of the matching degree.
-
+(i.e., \emph{CAN}).
+%%AD I am not sure of the last sentence + CAN... What does it say?? context???
+
+%%AD The following sentence says what exactly????? 
+% Even though no theoretical proof is given, valid results are
+% obtained and demonstrated through experimentations. 
+%%%%%%%
+% It is argued that since it
+% is observed to achieve a $27\%$ stretch reduction when compared to a landmark
+% enhanced CAN-system, it can be deduced that it performs better than landmark
+% binning in terms of the matching degree.
+%%AD rephrased above as below.. .
+Experimentation shows that \emph{SAT-Match} 
+attains a $27\%$ stretch reduction over 
+a landmark-enhanced \emph{CAN}-alternative.
+Based on this result, it is conjured that \emph{SAT-Match} 
+may perform better than landmark binning in terms of the matching degree. 
+\emph{SAT-Match} behaves as follows with respect to the stated $3$ criteria:
+%%AD matching degree is unqualified here... 
+%%
 % Figure~\ref{figure:sat_match} demonstrates the selective jump on an
 % example topology.
-
 %
@@
 % as effectively combined with other techniques, such as landmark binning.
-
+%%
 %
@@
 % selective jumps.
-
+%%
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 
-This section gathered together some of the most distinctive research work that
-can be found in the literature for addressing the inefficient network resource
-utilization in structured architectures imposed by the topology mismatch
-problem. A categorization of the methods used for achieving their goals is the
-following based on previous work from
-\cite{CDHR2002,CDCR2002,RSS2002}:
+This section outlined
+% gathered together  some of the most distinctive 
+representative research that attempts 
+to address the issue of inefficient network resource utilization 
+as this is created by the network topology mismatch
+in structured \p-systems.
+Depending on the salient features the surveyed techniques seek to explore,
+\cite{CDHR2002,CDCR2002,RSS2002} categorize the solutions based on:
+%%AD the categorization is due to 3 above citations??? Us?? ;-) 
+% A categorization of the methods used for achieving their goals is the
+% following based on previous work from
+% \cite{CDHR2002,CDCR2002,RSS2002}:
 \begin{enumerate}[\itshape i\upshape)]
-  \item geographic-layout-based,
-  \item proximity-routing-based,
-  \item proximity neighbor selection based,
+  \item geographic-layout,
+  \item proximity routing, and 
+  \item proximity neighbor selection,
 \end{enumerate}
 
-%%VM Added the following to mention the hybrid/combo methodologies
-Similarly to unstructured approaches, most of the reviewed structured ones,
-incorporate multiple of the above methodologies. For example, \cite{KLKP2008}
-uses both proximity routing and proximity neighbor selection and achieves
-topology adaptation, or \emph{LAPTOP} combines geographic layout with
-replication strategies to refine the end results. The heuristic nature
-of the proposed solutions makes it a game of trade-offs, that in most cases
-is highly affected by the application domain and the implementation itself.
-Thus, the reader is advised to filter the results of each approach and try to
-think about how the points presented will work in her own particular problem.
-
-A quick overview can be obtained in Table~\ref{structured:table}. Topology
-adaptation, landmarking and caching/replication have been visited during the
-discussion on unstructured systems in Section~\ref{section:unstructured} so
-we will not elaborate more here. As in the previous section, we provide the
-key properties of each algorithm that differentiate it from the others and
-highlight each algorithm's advantages and disadvantages.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Geographic layout} \label{section:geographic_layout}
-
+%%COM-VM Added the following to mention the hybrid/combo methodologies
+Similarly to unstructured approaches, most of the approaches 
+reviewed in this section entail features from multiple of the above categories. 
+For example, \cite{KLKP2008}~uses both proximity routing 
+and proximity neighbor selection to achieve topology adaptation 
+or \emph{LAPTOP} combines geographic--layout with
+replication strategies to refine its results. 
+%%%
+The heuristic nature of the suggested solutions opens them up 
+to trade-offs which in most cases are 
+highly affected by the application domain and the implementation itself.
+%%AD this is like instructions for manual - I think the point is made and 
+%%	so we can avoid this.. 
+% Thus, the reader is advised to filter the results of each approach and try to
+% think about how the points presented will work in her own particular problem.
+
+Table~\ref{structured:table} offers a quick overview of how the surveyed 
+approaches behave.  
+The topics of topology adaptation, landmarking and caching/replication
+are not further elaborated as they have been discussed 
+in Section~\ref{section:unstructured}.
+In a similar fashion to Table~\ref{unstructured:table}, we offer 
+key properties of every approach 
+that differentiate it from competitors and
+highlight each algorithm's pros and cons.
+%%AD rephrased what is below as above..
+% A quick overview can be obtained in Table~\ref{structured:table}. 
+% Topology adaptation, landmarking and caching/replication have been visited during the
+% discussion on unstructured systems in Section~\ref{section:unstructured} so
+% we will not elaborate more here. 
+% As in the previous section, we provide the
+% key properties of each algorithm that differentiate it from the others and
+% highlight each algorithm's advantages and disadvantages.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Geographic-layout} \label{section:geographic_layout}
+
+% In the \emph{geographic layout (GL)} category fall methods which practically
+This category consists of methods that 
+position physically nearby nodes together in the application space as 
+Figure~\ref{figure:geographic-layout} depicts. 
+This is accomplished by adjusting and maintaining the routing tables 
+of all involved peers by exploiting proximity information 
+that describe the overall geographic positioning of peers. 
+%%AD I am not sure I follow the 1-line above... 
+%%
 \begin{figure}[ht]
@@
 \caption{Successive nodes in the overlay circle can be chosen to be from the
-same or near autonomous systems.}
+same or near autonomous systems (\emph{AS}).}
 \label{figure:geographic-layout}
 \end{figure}
-
-In the \emph{geographic layout (GL)} category fall methods which practically
-position physically nearby nodes together in the application space as well
-(check Figure~\ref{figure:geographic-layout}). This is done by adjusting and
-maintaining the routing tables of all involved peers by exploiting
-proximity information that describe the overall geographic positioning of the
-peer nodes. Some of the algorithms we saw were \emph{Global Soft-State},
-\emph{Mithos} and \emph{LAPTOP} which all try clustering close by nodes in
+%%
+A number of the reviewed algorithms including 
+\emph{Global Soft-State},
+\emph{Mithos} and \emph{LAPTOP} resort to clustering close-by nodes in
 order to reduce the network's average delay.
-Landmark servers and RTT measurements are two popular methods, which
-can also be used in conjunction, to discover physically nearby peers over the
-network. However, as already discussed, these
-methods do not always give reliable estimates for the node positions over the
-Internet. The landmark servers are not self-organizing and have maintenance
-overheads. To serve a \p\ network with millions of peers, multiple landmark
-servers distributed over the whole world is required, which is hard to manage.
-The RTT measurements also can measure the delay between peers, but it is a
-greedy method, which can result in non-optimal overlay topologies especially if
-close by nodes have low bandwidth connections among themselves.
-
-Even though managing the overlay structure based on the \emph{GL} of the
-nodes improves the query efficiency of the system, on the other hand, it tends
-to undermine the uniform distribution of peer identifiers which consequently
-tend to create hot-spots, failing resilience is undermined by the fact
-that nearby nodes are more likely to suffer collective failures \cite{HY2007}.
+Landmark servers and \emph{RTT} measurements are two popular methods, which
+can also be used in conjunction, to discover peers in physical proximity.
+However, as already discussed, these
+methods do not always yield reliable estimates for node positions over the
+Internet. 
+Landmark servers are not self-organizing and present maintenance
+overheads. 
+To serve a \p--network with millions of peers, multiple landmark
+servers distributed over the world is required, a difficult proposition.
+\emph{RTT}s can measure the delay between peers, but their use constitutes a
+greedy method that can result in non-optimal overlay topologies;
+the latter is especially true if nearby 
+nodes happen to be connected through low bandwidth connections.
+
+On one hand, managing the overlay structure based on \emph{geographic layout} 
+of the nodes improves the query efficiency of the system in general. 
+On the other, it tends to undermine the uniform distribution 
+of peer identifiers which in turn terns to create 
+hot-spots and limited node resilience to failures.
+Such resilience appears to suffer as
+nearby nodes are more likely to suffer collective failures~\cite{HY2007}.
+%%AD changed the wording above - check...
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Proximity Routing}
-
+\subsubsection{Proximity Routing (PR)}
+
+This category of approaches % ({\emph{PR}) 
+does not require routing tables to be built using any
+knowledge about network proximity. 
+%%AD so if no knowledge is required what happens? How you go about 
+%%%
+%%AD I am not sure what the following sentence says... ;-) 
+On the other hand it exploits such knowledge
+to choose the best next hop when routing a message as can be seen in
+Figure~\ref{figure:proximity-routing}. 
+%%AD The above opening 2 sentences require re-calibration.. 
+%%
 \begin{figure}[ht]
@@
 \end{figure}
-
-\emph{Proximity routing (PR)} does not require routing tables to be built using any
-knowledge about network proximity. On the other hand it exploits such knowledge
-to choose the best next hop when routing a message as can be seen in
-Figure~\ref{figure:proximity-routing}. We visited an array of methodologies that
-implemented this method, either on its own like in \emph{Mithos} and
-\emph{PChord} or in conjunction with proximity neighbor selection like in
-\emph{DynaMO}. \emph{PR} tries to balance between choosing the node that will
-further progress the routing towards the destination and choosing the closest
-entry in the routing table, in terms of network proximity. As this is very
-difficult to achieve, the system might end up
-using a greedy approach that selects low latency paths on the overlay, which
-maps to suboptimal longer paths on the physical topology. Thus, in general,
-proximity neighbor selection is considered as a superior method to proximity
-routing, however, joint uses of these two protocols are also possible. Moreover,
-they are less effective than geographical layout when applied to CAN(-like)
+%%
+We reviewed an array of methodologies that strictly adhere to 
+the tenets of this category such as \emph{Mithos} and
+\emph{PChord} as well as others that function along with 
+proximity neighbor selection such as \emph{DynaMO}.
+%%
+\emph{PR} tries to balance between choosing the node that will
+further progress the routing towards the destination 
+and picking the closest entry in the routing table
+in terms of network proximity. 
+%%AD just above in this section you have not clarified the way 
+%%	routing tables get populated - u have to clarify this point 
+%%	in order for the contrast here to be valid.. 
+As this is very difficult to achieve,  %%AD Why so? Why is it difficult??? 
+the system might end up
+using a greedy approach that selects low latency paths on the overlay
+which ultimately maps to (suboptimal) longer paths on the physical topology. 
+%%AD in the following you make a point that should be made in the 
+%%	follow up segment - no???
+Thus, in general, proximity neighbor selection is considered 
+as a superior method to proximity routing, however, joint uses of these two protocols are also possible. 
+Moreover, they are  %%AD who is they? Confused!
+less effective than geographical layout when applied to CAN(-like)
 implementations.
+%%AD re-work carefully this segment...
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Proximity neighbor selection}
-
+\subsubsection{Proximity Neighbor Selection (PNS)}
+
+% \emph{Proximity Neighbor Selection (PNS)} 
+Algorithms in this category populate their routing tables using
+proximity knowledge as Figure~\ref{figure:proximity-neighbour-selection} 
+depicts. 
+%%
 \begin{figure}[ht]
@@
 \end{figure}
-
-\emph{Proximity Neighbor Selection (PNS)} constructs the routing tables using
-proximity knowledge (see Figure~\ref{figure:proximity-neighbour-selection}). The
-proximity information used in this method is different from the the
-landmark-based systems described in Section~\ref{section:geographic_layout}.
-Here time-to-live values between nodes (i.e. \emph{SAT-Match}) or IP address
-prefixes (i.e. \emph{DynaMO} or \emph{Cone}) are used to detect
-proximity. \cite{freedman_iploc_2005} states that $97\%$ of prefixes larger
-than $24$ bits belong to a single geographical location. However, using a
-smaller number of bits creates less precise results and a larger number of bits
-increases the burden on the network and reduces the possible number of
-neighbors. Moreover, in \cite{HLYW2005} it is argued that such approaches can
+%%
+The proximity information used here is different from that of 
+landmark-based systems described in Section~\ref{section:geographic_layout};
+here, \emph{TTL} values between nodes (i.e., \emph{SAT-Match}) or 
+\emph{IP} address prefixes (i.e., \emph{DynaMO} or \emph{Cone}) 
+are predominantly used to detect proximity. 
+\cite{freedman_iploc_2005} states that $97\%$ of prefixes larger
+than $24$ bits belong to a single geographical location. 
+However, using a smaller number of bits creates less precise results 
+and a larger number of bits may increase the burden in the network
+and simultaneously reduce the possible number of neighbors. 
+%%
+Moreover, \cite{HLYW2005} argue that such approaches can
 destroy the uniform distribution in the key space and do not work in
-one-dimensional key space where the mapping is constrained on the overlay layer.
+one-dimensional key space where the mapping is restricted to 
+operate at the overlay.
 Therefore, a careful selection is required in terms of
-performance/accuracy trade-off. Tapestry, Pastry, and CAN successfully incorporated
-proximity neighbor selection into their algorithms. The routing protocol in
-Pastry is based on longest node ID prefix matching, while CAN uses RTT values to
+performance/accuracy trade-off. 
+\emph{Tapestry}, \emph{Pastry}, and \emph{CAN} successfully incorporated
+proximity neighbor selection into their algorithms. 
+The routing protocol in \emph{Pastry} is based on longest node 
+\emph{ID} prefix matching, while \emph{CAN} uses \emph{RTT} values to
 detect nearby nodes. \cite{CDCR2002a} reports that proximity neighbor
 selection is an effective proximity based method.
+%%AD perhaps the ending of the previous segment should be used here...
 

### 2014-02-02 unstructured.tex
2013.10.29 (Alex) Structured section underwent several reworks/rephrasings. Text reviewed. Minor changes to unstructured section.

@@
   \item caching and replication, and
-  \item landmarking
+  \item landmarking.
 \end{enumerate}
@@
 is that they do not strictly adhere to a single methodology.
-As discussed approaches attempt to address the inherently challenging problem 
+As approaches attempt to address the challenging problem 
 of topology mismatch, they resort to heuristics and so, 
-they do not often furnish ``pure'' solutions (i.e., solutions that exclusively belong to 
+they do not often furnish ``pure'' approaches 
+(i.e., solutions that exclusively belong to 
 one of the $4$ aforementioned methodologies). 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-For instance, \emph{mOverlay} combines topology adaptation with landmark binning 
+For instance, \emph{mOverlay} combines topology adaptation 
+with landmark binning 
 while \emph{Gia} weaves together $3$ methodologies:  topology adaptation, 
@@
 network may have already been configured to disable traceroute response from
-the start. The most problematic aspect of clustering, though, is
+the start. 
+The most problematic aspect of clustering, though, is
 its nature per se. 
@@
 significantly shrink the search scope, negatively affecting the query response
-time that the \p\ user experiences. Among others, \emph{DDNO} tries to 
+time that the \p\ user experiences. 
+%%
+Among others, \emph{DDNO} tries to 
 balance the efficiency of the clustering approach 
@@
 message round-trip-time is reduced, it has actually additional cost in terms of
-inter-ISP communication accounting and management.
+inter-\emph{ISP} communication accounting and management.
 %%

### 2014-10-05 background.tex
2014.10.05 (Vassilis) Clean-up, rewording and reviewing the unstructured section. Efficiency, overhead and scalability definitions reworked in background section.

@@
 \emph{Efficiency} means how well the
-algorithm performs its search and topology adaptation functions. We also take
-take into account how this might effect other characteristics of the networks
-under consideration\footnote{For example we don't actually want a topologically
-optimized BitTorrent swarm if this results in a degraded download performance,
-a characteristic that is the protocol's corner stone.}.
-%%VM Does the above sentence with the footnote now connect?
+algorithm performs its search and topology adaptation functions. Our assessment
+also considers any negative effects that might come along with the application
+of the algorithm on intrinsic characteristics of the system (i.e., search scope
+coverage, download performance e.t.c).
+%% VM Reworked above definition for "efficiency".
+
+For the \emph{overhead} criterion, we evaluate the cost
+of computing and communicating peer join, message forwarding, 
+and overlay topology maintenance. For instance, the overhead that an algorithm,
+which needs global knowledge of the overlay topology, imposes to the system is
+greater compared to a protocol that only relies on peers exchanging routing
+tables within a a 2-hop radius from each other.
+%%VM Reworked definition of overhead and explanatory sentence.
 %%
-%% PREVIOUS COMMENT AD I am not sure what the above line says.. or connects. Something is
-%%	missing...
 
-%% SOME ADDITIONAL EXAMPLES FOR EFFICIENCY IF WE WANT TO ADD THEM
-%minimum-spanning-tree-based techniques have slow convergence speed in topology
-%adaptation or the high clustering impacts load balancing and constrains search
-%scope (or at least increases search satisfaction time), all reducing overall
-%efficiency.
-
-Moreover, efficiency is negatively effected every time an algorithm
-reduces its search scope.
-For the \emph{overhead} criterion, we evaluate the cost
-to compute and communicate peer join, message forwarding, 
-and topology maintenance. For example, when a protocol relies on peers
-exchanging routing tables with other peers within a 2-hop radius with each other
-, then we can be safelly deduce that it incurs lower overhead when compared to a
-protocol that requires global knowledge of the overlay network.
-%%VM Does the above sentence with the footnote now connect?
+For the \emph{scalability} criterion, we consider system behavior with respect
+to the number of participating peers it can reliably support. We characterize
+protocols as being lowly, mediumly or highly scalable if the are proven to
+effectivelly support hundreds, tenths of thousands or millions of nodes,
+respectively.
+%%%VM Reworked definition.
 %%
-%%PREVIOUS COMMENT AD Again I am not sure how the example above ties with the definition...
-For the \emph{scalability} criterion, we consider system behavior 
-with respect to the number of participating peers. 
-We consider systems with low, medium, and high scalability 
-to mean they can support hundreds, tenths of thousands, 
-and millions of nodes, respectively.

### 2014-10-05 unstructured.tex
2014.10.05 (Vassilis) Clean-up, rewording and reviewing the unstructured section. Efficiency, overhead and scalability definitions reworked in background section.

@@
 for peer communication and the techniques they use for detecting
-proximity, to be used to optimize their overlay topology.
+proximity.
 
@@
 \cite{YG-M2002} proposed a practical and easy to implement solution 
-weaved around three different message forwarding methods:
-\emph{iterative deepening},
-\emph{directed BFS}, and \emph{local indices}.
+weaved around three different message forwarding methods, namely
+\emph{iterative deepening}, \emph{directed BFS} and \emph{local indices}.
 
@@
 The source node may issue a new request with increased depth limit
-that will trigger the nodes at the last depth level to resume the search. 
-The iterative approach avoids restarting the entire search process 
-from scratch at each iteration and reduces the load on the nodes of 
-the upper levels of the tree. 
+that will trigger the nodes at the last visited depth level to resume the
+search thus avoiding restart of the entire search process and consequently 
+reducing the load on the nodes of the upper levels of the tree. 
 Its major drawback is the delay between successive iterations, as the
@@
 query messages only to a selected subset of available neighbors. The selection
-criteria varies, from the number of results received previously or the distance
-in terms of hops traveled to locate these results, to the bandwidth, or the
+criteria varies, from the number of results received or the distance (hops)
+traveled to locate these results, to the bandwidth, or the
 query load of the neighbor. That way, fewer nodes are visited but quality of
@@
 In \emph{local indices},
-each node indexes data within a radius of $r$ hops and uses 
-this local index to answer queries % on behalf of them 
-without generating additional traffic. Local indices
+each node indexes data hosted by neighboring nodes within a radius of $r$ hops
+and uses this local index to answer queries on behalf of other nodes.
+That way, local indices,
 greatly reduce the aggregate bandwidth usage of the network and 
@@
 of Section~\ref{section:background}.
-% \begin{center}
-% \begin{tabular}{ccc}
-% \textbf{Efficiency} & \textbf{Overhead} & \textbf{Scalability} \\
-% \hline
-% Iterative Deepening & ? & ? & ?
-% Directed BFS & ? & ? & ?
-% Local Indices & ? & ? & ?
-% \end{tabular}
-% \end{center}
+
 \begin{center}
@@
 % \subsubsection{Gia}
-
 The key objective of the {\sl Gia} system~\cite{CRBLS2003} is 
@@
 The \emph{neighbor comparison and selection} component defines that a peer $S$
-probe all his $2$-hops away neighbors (a set denoted as $ N^2(S)$) not probed
+probes all his $2$-hops away neighbors (a set denoted as $ N^2(S)$) not probed
 thus far. Let $P$ be a $1$-hop distance neighbor of $S$, and $Q$ be a probed
 peer by $S$. When $S$ receives a message piggybacked with message $PM_{PQ}$, 
-% we identify the following cases.  %%AD put it passive as below.. 
 the following $2$ cases are identified:
@@
 overall, \emph{THANCS} fares as follows regarding the $3$ stated criteria:
-%%AD rephrased more diplomatically the following as above.. 
-% There are works in the literature though that criticize, through simulation
-% results, the algorithm's average broadcast delay and overlay improvement
-% guarantee \cite{HLY2010}.
-% These characteristics are depicted to the $3$ evaluation criteria we use in this
-% survey as follows:
+
 %%%%%%%%%%%%%%%%%%%%%%
@@
 the corresponding delays of $(v_2,v_3)$ and $(v_3,v_2)$
-% physical paths denoted by $y$. 
 for the physical path $y$.
-%%AD check above... it is one path (not paths right?)
 %%
@@
 most peers in a swarm reside in different authonomous systems (\emph{AS}s).
-%%AD AS = autonomous systems??? - put it so. First time discussed!
 Consequently, clients frequently do not have enough peers to leverage
 intra-\emph{AS} connections \cite{RTLCGZ2010}.
-%%AD rephrased as above - what is AS??? Acronym... 
-% On the other hand, the effectiveness of \emph{BNS}
-% is arguable by other studies, whose
-% experimental results suggest that most peers in a swarm reside in different ASs,
-% thus a client in most of the cases does not have enough peers to leverage
-% intra-AS connections \cite{RTLCGZ2010}. 
 Moreover, \emph{BNS} deployment
@@
 %%
-Experimentation with \emph{Oro} yields minimization of
+Experimentation with \emph{Ono} yields minimization of
 inter-\emph{ISP} traffic costs and achieves near optimal performance 
@@
 its qualitative rating stands as follows:
-%%AD rephrased as above...
-% Even though the results in the paper claim minimization of inter-ISP traffic
-% cost while enhancing the, already, near optimal performance of BitTorrent
-% download, similar to \cite{BCCMSBZ2006}, this ISP-friendly design of \emph{Ono}
-% is questioned, especially on a Internet-scale-basis \cite{PMJKA2009,LCLX2009,CLYSR2011}.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%
-The second technique, applied in an \emph{intermediate-level}, manipulates the
-{\sl BitTorrent} chocking/unchocking mechanism. A peer unchockes its $4$
-closest in terms of 
-%%COMM-VM I put ``autonomous system'' back in again because in the context of this
-% paper we are dealing on the number of AS crossings (one-hop = one AS crossing)
-% until we reach a neighbor.
-autonomous system hop-count interested neighbors. 
+The second technique, applied in an \emph{intermediate-level}, manipulates
+{\sl BitTorrent}'s chocking/unchocking mechanism. A peer unchockes its $4$
+closest, in terms of 
+autonomous system hop-count, interested neighbors. 
 The same applies also when the peer turns to 
@@
 %%
-The last technique operates at 
-a \emph{microscopic-level} and substitutes the rare-first chunk picking of the
-original {\sl BitTorrent} protocol that picks the rarest of the available
-chunks to download next, with a locality-first policy; this picks first the
-chunck is located closer. 
-%%AD check slight rewording above
+The last technique operates at a \emph{microscopic-level} by chosing to pick
+the next chunk to download based on a locality-first policy; this picks first
+the chunk that is closer, as opposed to {\sl BitTorrent}'s rare-first chunk
+picking policy.
+%%VM I once more reworded the above, seemed to complex to me
 The distance value of a chunk is computed as the mean
@@
 
-%%AD changed the wording - check.. 
 All $3$ techniques in~\cite{LCLX2009} seek to harness 
-% in exposing 
 the locality of the basic functions of the
@@
 without the need of additional infrastructure. 
-which \emph{TopBT} is built on the conjecture that
+\emph{TopBT} is built on the conjecture that
 a good peer selection metric should take into account both the
@@
 aspects of the original {\sl BitTorrent} protocol to ``inject'' topology
-awareness for improved peer selection. 
-%%
-In an effort to essentially improve on the inherent topology mismatch problem,
-\emph{TopBT} modifies: 
+awareness for improved peer selection. More concretely, \emph{TopBT} modifies: 
 \begin{inparaenum}[1)]
@@
 \item the initial connection establishment, and 
-\item the unchoking mechanism in the {\sl BitTorrent}.
+\item the unchoking mechanism of {\sl BitTorrent}.
 \end{inparaenum}
-\emph{TopBT}'s key feature is
-that it does not need wide deployment.
-% of its same types in order to see results. %%AD I do not get the second part of the sentence
-% that it does not need wide deployment of its same types in order to see results.
+\emph{TopBT}'s key features are that it does not rely on any Internet infrastructures
+(ISPs or CDNs) whatsoever and it cooperates with already deployed BitTorrent clients.
 It also supports both link-hop and \emph{AS}-hop metrics to identify proximity in
-different levels.
+different levels in order to reduce overall traffic while preserving downloading
+performance.
 %
@@
 % the algorithm does not involve extensive calculations.
+% TopBT only generates a peak number of ping messages at the start during its
+% unchoking period (<16 in 5 minutes), and that number quickly drops to a few
+% (2 in 5 minutes). For traceroute, the number is higher, as each traceroute
+% message triggers a sequence of ICMP packets. It is about 20 times more than
+% the number shown in the figure. But overall, TopBT is light-weight in terms
+% of extra probing messages generated.
 low &
@@
 
-%%AD slight rewording throughout...
 Upon arrival, a peer $n$ asks a \emph{host-cache} to establish its
@@
 and unilaterally decides to join a specific bin based on its measurements.
+%
+%%VM Rewrote the following. Please double check that it makes sense. 
+%%
 Specifically, the node measures its round-trip time to each of the landmarks and
-orders the resulting \emph{RTT}-values in a decreasing order. 
-%%AD I do not understand what the following sentence says..
-%%	for each peer there is a different bin right?
-%
-%The ordering represents a \emph{bin},
-%in the sense of close-by nodes having the same landmark ordering and hence
-%belong to the same bin. 
-%
-%% VM I CHANGED THE ABOVE AS BELOW TO REPHRASE IT.
+orders the landmarks in decreasing order of those \emph{RTT}-values.
 Each permutation of the set of landmarks represents a specific bin. 
@@
 Finally, \cite{XTZ2003} underscores the potentially increased maintenance cost(s).
-%%AD rephrased what is below as above... The ONE sentence below cannot be read!!!!
-%%%
-% Moreover, studies like \cite{GSG2002}, pin point the fact that end-host shoulder the
-% latency measurements as the scheme's scalability bottleneck while others argue
-% about the uneven distribution of nodes potentially resulting in load imbalance
-% \cite{CDCR2002,M2003,ZZZSZ2004}, its coarse-grained nature that fails to
-% effectively distinguish among relatively close nodes \cite{RGJZ2004} and the
-% potentially increased maintenance cost \cite{XTZ2003}.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 A peer may reach its group by expending at most $O(logN)$ messages.
-%%AD simplified sentence below as above...
-% It is shown that in the proposed scheme, a new peer can 
-% reach its group by expending at most $O(logN)$ messages
-% within the network. 
 Finally, \emph{mOverlay} maintains stability and constant 
@@
 of topology mismatch, they resort to heuristics and so, 
-they do not often furnish ``pure'' approaches 
-(i.e., solutions that exclusively belong to 
-one of the $4$ aforementioned methodologies). 
-%%AD rephrased as above..
-% An important aspect of many implementations, is that since the above
-% methodologies use heuristics to solve topology mismatch, an
-% intrinsically difficult problem, no single approach can offer a robust solution.
-% Thus, many of the algorithms we surveyed are actually combinations of multiple
-% such methodologies, in an attempt to refine the results of using just a single one.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+they do not often furnish ``pure'' approaches that exclusively
+belong to one of the $4$ aforementioned methodologies.
 For instance, \emph{mOverlay} combines topology adaptation 
@@
 its overall approach as well as its underlying used methodologies have to be considered.
-%%AD rephrased as above... did not like much this "advised" - is not wrong BUT sounds funny
-% For this reason, the reader is advised to critically approach either the overall
-% algorithms or the separate methodologies when evaluating them for different
-% application domains.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 approaches applicable for \emph{unstructured} \p \emph{systems}.
-%%VM I brought the following paragraph here from the end of this subsection.
 Moreover, Table~\ref{unstructured:table} offers a summary overview for each effort;
@@
 as Figure~\ref{figure:topology-adaptation} depicts.
-%%AD u always need a reference of a Figure in the text!!!
-% as Figure~\ref{figure:topology-adaptation} depicts.
 Despite the fact that the above is a plausible proposition,
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%AD slight wording changes below - check
 \subsubsection{Forwarding Optimization Methodology}
@@
 aligned with each other.
-%% let alone quantify the mismatch and try to alleviate.
-%%AD I got the above out as superfluous
 To this end, forward-optimization techniques are commonly applied in
@@
 
-% Landmark--based protocols have four important drawbacks: First, the network
-%%AD the correct here is "first" NOT "First"!
-Landmark--based protocols have four important drawbacks: first, the network
-delay is not a reliable distance estimation method. For example, based on the
-load on the network, the delay to certain nodes or networks can significantly 
-change over time; this eventually affects the distance measurements and yields wrong
-measurements that ultimately lead to wrong estimated positions for the nodes 
-or incorrect and non--optimal clusterings of nodes. 
-Second, relying on predefined nodes
-makes the whole paradigm not fully distributed and the landmark system prone to
-single point of failure. Third, 
-landmark infrastructure require costly installation and maintenance 
-across the Internet and the different autonomous system domains. 
-As popular \p\ file--sharing applications usually have millions
-of peers connected to at any time, the network costs
-of maintaining these landmark servers can be quite high. 
-A possible solution to
-the scalability problem of the static landmark servers is to use ordinary nodes
-as dynamic landmarks in an approach reminiscent to that of \emph{mOverlay}.
-% once they estimate their own positions. 
-Even though this scales better, the accuracy of measurements
-may affect the overall performance of the system. Moreover,it can
-be characterized as a rather coarse-grained approximation, therefore not
-particularly well suited for identifying the correct positions of nodes within
-close distance to each other.
-%%AD check the above made various small changes...
+Landmark--based protocols have four important drawbacks:
+%%%%
+first, the network delay is not a reliable distance estimation method. For
+example, based on the load on the network, the delay to certain nodes or
+networks can significantly change over time; this eventually affects the
+distance measurements and yields wrong measurements that ultimately lead to
+wrong estimated positions for the nodes or incorrect and non--optimal
+clusterings of nodes.
+%%%%
+Second, it can be characterized as a rather coarse-grained approximation,
+therefore not particularly well suited for identifying the correct positions
+of nodes within close distance to each other.
+%%%%
+Third, relying on predefined nodes makes the whole paradigm not fully
+distributed and the landmark system prone to single point of failure.
+%%%%
+Lastly, landmark infrastructure requires costly installation and maintenance 
+across the Internet and the different autonomous system domains. As popular
+\p\ file--sharing applications usually have millions of peers connected to at
+any time, the network costs of maintaining these landmark servers can be quite
+high. A possible solution to the scalability problem of the static landmark
+servers is to use ordinary nodes as dynamic landmarks in an approach
+reminiscent to that of \emph{mOverlay}. Even though this scales better, the
+accuracy of measurements may affect the overall performance of the system,
+especially in an environment with high churn.
+%%VM Some rewording and restructuring. Please have a look.
+%
 

### 2014-11-30 structured.tex
2014.11.30 (Vassilis) Structured section massively reworked. Clean-up, rewording and reviewing comments from Alex. Removed Cone algorithm as its text was badly written and the paper was equaly badly written. New reference added (.bib file updated and KK2000.pdf file added).

@@
 %%AD what does it really mean this below???
+%%VM it means that this split into i) geographic layout, ii) proximity routing
+%%   and iii) proximity neighbour selection types of algorithms are terms
+%%   coined in those papers
 The categorization is based on previous work found in the
 literature~\cite{CDHR2002,CDCR2002,RSS2002}.
-
-%%COMM Rephrased as above to keep it analogous to the opening paragraph of
-%% the unstructured part.
-%%
-% This section presents a description and analysis of algorithms proposed
-% to address the topology mismatch problem in structured \p\ networks.
-% In structured \p\ networks, the construction of the routing tables
-% held by participating nodes determines the efficiency of message forwarding.
-% Routing tables that closely match the underlying IP topology
-% achieve increased performance. As a result, the proposed solutions for
-% the topology mismatch problem for structured \p\ networks
-% use different levels of proximity
-% information to optimize the routing tables used by peers.
-% The categorization of the algorithms we present here is based on the
-% categorization suggested by previous
-% work~\cite{CDHR2002,CDCR2002,RSS2002}.
 
@@
 
-\emph{Global Soft-State} \cite{XTZ2003} builds a global map to help
-choose shorter routing paths. This is accomplished by 
-combining the landmark binning method with small
-scale distance probes to reveal the proximity properties 
+%% VM Most of this paragraph on global soft-state rewritten/restructured.
+%% Have a look...
+
+\emph{Global Soft-State} \cite{XTZ2003} builds a global view to help nodes
+choose shorter routing paths. To generate the proximity information, it 
+combines the landmark binning method with small
+scale distance probes to reveal the properties 
 of the underlying network. 
@@
 %space and the mapping of the position of the nodes.
-This global view of the state is made available to 
-\emph{all} nodes to help them find the best way to route their messages. 
-%%AD when you say "all" above means what?? all in the structured network?? 
-%%	how is the structured network organized?  (a line or 2 for this?)
-\emph{Global Soft-State} operates in $2$ phases:
-\emph{generation} and \emph{use} of the proximity information. 
-For \emph{generating} the proximity information, 
-a hybrid approach is proposed: it uses landmark clustering as a
-preprocessing step to select a number of potential nearest neighbor
-candidates and then, refines the selection by incorporating a
-round-trip-time-based scheme to ultimately choose the closest node. 
-%% 
-The algorithm follows a different path from that of the
-classic gossiping approaches for the construction and 
-maintenance of the overlay when 
-it comes to \emph{use} the proximity information.
-%%AD isn't the following sentence somewhat of repetition??? (of what 
-%%	is stated above?? 
-The approach is based on landmark clustering 
-for generation and strategic
-placement of proximity information on the overlay. 
-This information is called a ``map'' and concerns a 
-region of the overlay; 
-the latter may be part of the Cartesian space in
-structures like \emph{eCAN}~\cite{xu_ecan_2002} or 
-a set of nodes sharing a particular
-prefix in overlays such as \emph{Pastry}\cite{antony_pastry_2001}. 
-%%
-Thus, each node not only can access such information from 
-these landmarks but can also subscribe to
-relevant \emph{soft states} and get notified on dynamic network changes. 
-A node may appear in a maximum of $log(N)$ such maps.
+
+This global view of the network state is used by organizing it into ``maps'' for
+regions of the overlay, that are strategically placed
+across the network and made available to any node. The notion of a
+region is a well defined property for structured \p\ networks and can be part of
+the Cartesian space in overlays like \emph{eCAN}~\cite{xu_ecan_2002} or a set of
+nodes sharing a particular prefix in overlays such as
+\emph{Pastry}~\cite{antony_pastry_2001}. Moreover, a node may appear in a maximum
+of $log(N)$ such maps.
+
+Direct access to this information, also coined \emph{soft states}, helps nodes
+find the best way to route their messages, while a publish/subscribe mechanism
+allows them to get notified on dynamic network changes.
+
 %%
@@
 limited knowledge about neighboring zones.
-%%AD rephrased what is below -2 lines- as above..
-% Also improvement
-% is done within a region because of little knowledge about neighboring zones.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % \subsubsection{Mithos}
-\emph{Mithos} \cite{WR2003} is a \p\ protocol which incorporates directed
-incremental probing to find near optimal node placement and integrates
-geographic layout and proximity routing overlay optimization methods.
+
+%% VM Whole paragraph rewritten
+
+%% VM: PREVIOUS COMMENTS BELOW REGARDING PREVIOUS VERSION
 %%AD in general some mention of the mismatch problem must be made right???
 %%		so that a reasonable connection can be established...
-
-The bootstrap phase of \emph{Mithos} starts with a subset 
-of existing members as the initial set of candidates and 
-while the closest nodes are detected by iterative
-probing of the neighborhood, the candidate neighbor--list gets updated. 
-To avoid local minima, \emph{Mithos} probes all the neighbors 
-within a two-hop distance from the current minimum 
-before concluding the process.
-
-After finding its first neighbor, an 
-% new-coming   %%AD overall the term new-coming is somewhat odd.. incoming? 
-incoming peer is assigned 
-an \emph{ID} using information gathered during 
-the iterative neighbor selection phase. 
-%%AD use SIMPLE sentences that are well structured to say what u want
-Based on the two closest nodes found, along with these nodes neighbors and the corresponding
-distances a system of virtual springs is set so that the tension of each spring
-is inversely proportional to the distance. 
-%%AD the above is impossible to read.. "based on,... along with... " NO!!!
-%%AD I am not sure what you are talking about here - I mean below... rework
-%%	and simplify.. 
-The system is then left to settle,
-and the state in which it settles is used for its \emph{ID}. 
-The benefit of these synthetic coordinates are that
-they are explicitly used as the node's \emph{ID}, 
-and distance computations among nodes
-can be done in \emph{ID} space without requiring physical measurements.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% upto here %%%%%%%%%%%%%%%%
-
-The last step of the algorithm is the 
-interconnection among neighbors. 
-\emph{Mithos} uses a \emph{quadrant-based} mechanism in which 
-each node establishes a link to the closest neighbor in each quadrant. 
 %%AD what is the quadrant??? Hugh???
-During forwarding, the next hop is performed towards a neighbor 
-in the same quadrant of the final destination. 
-An incoming node may not know of other neighbors in all
-quadrants, therefore, the node first identifies neighbors in all quadrants
-using a mechanism based on ideas similar to a perimeter walk~\footnote{Used in
-the Greedy Perimeter Stateless Routing (\emph{GPSR}) protocol.}  %%AD ref??
-and then improves the results using parallel path processing 
-by taking into account further geometric properties of node relationships.
+
+%%AD ref?? (σ.σ. GPSR protocol)
+
 %%AD the above 5-6 lines sounds impressive but there is little help 
@@
 
+%%AD what is GNP??? How did it "land" here???
+
+\emph{Mithos}~\cite{WR2003} ensures that neighbouring nodes in the overlay are
+close also in terms of the underlying layer by integrating proximity layout
+and proximity routing overlay optimizations.
+
+During bootstrap, an incoming peer, say J, starts with a set of
+candidates to establish links with and iteratively
+probes for minimizing distances, based on a given metric (i.e., network delay)
+and concluding when no further refinement can be achieved. To tackle the local
+minima problem that such a gradient descent method suffers from, Mithos probes
+all neighbours within a two-hop radius, from the currrent node that minimize
+the distance, before concluding the process.
+
+Next, the incomming peer is assigned a synthetic \emph{ID} that
+represents the state of minimum distance from its close--by neighbourhood as
+this was identified during the iterative neighbor selection phase described
+above. The benefit of these synthetic coordinates is that distance computations
+among nodes can be done in the \emph{ID} space without requiring physical
+measurements. The actual link establishment with the peers, that will be J's
+neighbours, is happening next. With peer J dividing the plane into four regions,
+the \emph{quadrant-based} mechanism Mithos employs, suggests that the node
+establishes a link to the closest neighbor in each quadrant of the plane. The
+previous steps
+do not provide any knowledge as to which nodes these would be, thus J has to
+identify them. It does so, by starting from a known node in the neighbourhood
+and scanning towards all quadrant borders ~\cite{KK2000,R2002}.
+
 One major limitation of \emph{Mithos} is that the protocol cannot effectively
 handle dynamic arrivals and departures from the network. 
-\cite{CCRK2004}~points out the extensive probing to determine 
-distances and evaluates \emph{Mithos} results as less accurate
-than those derived with \emph{GNP}~\cite{ng_gnp_2001}. 
-%%AD what is GNP??? How did it "land" here???
-%%
-The latter has been also attributed by~\cite{WSS2005} to 
-the iterative neighbor selection process which that is 
+\cite{CCRK2004}~argues about extensive probing to determine 
+distances and less accurate proximity results compared to other
+virtual coordinate systems. The latter has been also attributed
+by~\cite{WSS2005} to the iterative neighbor selection process which that is 
 reportedly prone to premature termination at local minima.
@@
 \cite{cox_vivaldi_2004} argues that elements of the protocol's
-spring relaxation technique require a centralized implementation 
-affecting \emph{Mithos} scalability.
-%%AD rephrased what is below as above...
-% Another point that has negative effect to the
-% scalability of the algorithm is that parts of its spring relaxation technique
-% require a centralized implementation~\cite{cox_vivaldi_2004}.
+synthetic ID generation require a centralized implementation 
+affecting \emph{Mithos}' scalability.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % \subsubsection{LAPTOP}
+%
+%%AD what is the nature of the "structure" overlay here???
+%% VM the structure is that it is organized as a tree hierarchy.
 \emph{LAPTOP} \cite{WLH2007} organizes the overlay 
-%%AD what is the nature of the "structure" overlay here???
 into a tree-based hierarchy so that 
@@
 number of nodes and $d$ the maximum degree of each node. 
-\emph{LAPTOP} 
-%% implements 
-adopts the geographical layout approach  
+\emph{LAPTOP} adopts the geographical layout approach  
 and constructs its layout in a self-organizing and 
@@
 $1$ to $d$.
-%  where $d$ is the maximum degree of the nodes. 
-%%AD this definition was provided just above.. 
 The assignment process is done by appending a unique octet 
@@
 During tree traversals, special entries in the routing tables,
-called \emph{routing cache}, 
-%%AD clarify - routing caches are the special entries OR the routing tables???
 are maintained to increase routing efficiency and achieve finer--load balance.
-Caching enables a node to forward a message to a better 
+These special entries constitute the \emph{routing cache} which enables a node
+to forward a message to a better 
 longest-prefix match than that of its direct ancestor yielding
@@
 a parent is responsible for the monitor of its own children.
-%%AD rephrased what is below as above.. 
-% In terms of overlay maintenance, \emph{LAPTOP}
-% incorporates a simple \emph{heartbeat}-based technique 
-% where each parent node is
-% responsible for monitoring its children.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 With \emph{PRS} in place however, \emph{Kademlia} may also choose 
-the node that exhibits the smallest underlay metric cost; in this way, 
+the node that exhibits the smallest routing cost; in this way, 
 it balances a trade-off between the logical overlay and 
@@
 %%
+%% VM rephrased to not contain the word "underlay". Clean up comment block if OK.
+%%
 In \emph{PNS}, on the other hand, the goal is to keep peers 
-with the least contact cost in the routing table. 
+with the least routing cost in the routing table. 
 %%AD WHAT is this "least contact cost"?? Sounds funny! CHECK
-As \emph{Kademlia} continually learns about new
-peers via incoming requests and an iterative lookup process, 
-%%AD what is this iterative lookup process???
-no specialized algorithm is required for searching for 
-more cost-effective peers 
-and the system simply goes for the best peers ``seen'' thus far.
+%% VM rephrased to not contain the word "contact". Clean up comment block if OK.
+
+%%
+%% VM Commenting out since this info is not that important I believe
+%%
+%As \emph{Kademlia} continually learns about new
+%peers via incoming requests and an iterative lookup process, 
+%%%AD what is this iterative lookup process???
+%no specialized algorithm is required for searching for 
+%more cost-effective peers 
+%and the system simply goes for the best peers ``seen'' thus far.
 
@@
 % \subsubsection{CHOord considering Proximity on IPv6 (CHOP6)}
-\emph{CHOP6} \cite{MT2007} roughly estimates the proximity among
-nodes by exploiting the {\sl IPv6} address format and if available, 
-\emph{RTT} information.
-The proximity estimation is achieved by introducing a $64$-bit 
-\emph{ID} scheme in which
-the least significant bits are the {\sl IPv6} global routing prefix.
-Thus, a longest--prefix match scheme is established.
-% enabling a longest prefix match scheme. 
-%%AD check is the word "longest" is correct..
-The protocol is designed based on the
-observation that it is feasible to estimate a node's geographical location by
-examining the upper $32$-bits of its {\sl IPv6} address. 
-Moreover, similar to \emph{Chord}~\cite{stoica_chord_2001}, 
-\emph{CHOP6} uses a finger table, whose entries hold
-more than one candidate node.
+
+%%AD how does CHOP6 helps with the mismatch problem??
+%% VM rewrote the whole part to answer this question
+%% By and large, rtt measurements and id assignment help choose the next best
+%% hop during routing that is i) toward the key and ii) has small cost
+%% either in terms of explicit RTT or implicit RTT (taking regional/national
+%% hops as a proxy of proximity - international RTT > regional RTT > national RTT)
+
+\emph{CHOP6} \cite{MT2007} is another Chord-variant DHT; it exploits
+the {\sl IPv6} address format, and if available \emph{RTT} information, to 
+enable proximity estimation among peers during message routing.
+
+Similar to \emph{Chord}~\cite{stoica_chord_2001}, 
+\emph{CHOP6} uses a finger table but each entry now holds hold k>1 candidate
+nodes. The next hop for a message is selected to be the one closest to the
+destination among those k candidate nodes. This proximity is determined either
+by exploiting RTT measurements or, if there is no such yet (i.e. a new peer), by
+estimating distance on overlay's ID space. When joining the network, each node
+receives a $64$-bit \emph{ID}. It comprises of a random part and a IPv6 prefix
+part. The random part is obtained by hashing the node's IP address and port
+number and is used to preserve the uniformity of
+node assignment. The IPv6 prefix is based on IPv6's global routing prefix.
+Because IANA  assigns address blocks to regions (RIRs) which are further
+distributed on a national level (NIRs), examining the upper $32$-bits
+of an IPv6 address someone can estimate the country or the region to which the
+node is connected. Thus, CHOP6's IPv6 prefix ID is used for proximity estimation
+if no RTT information is available during routing.
+
 %
@@
 \end{center}
-%%AD how does CHOP6 helps with the mismatch problem??
 
@@
 % \subsubsection{T2MC}
-\emph{T2MC} \cite{SLCGZ2008} uses traceroute logs to detect 
-the boundaries of an autonomous system (\emph{AS}) and
-cluster nearby peers; the objective here is to 
-diminish the redundant multiple messages that may 
-unnecessarily cross \emph{AS}-boundaries.
-\emph{T2MC} uses a customized $k$-mean classification
-algorithm with $k = 2$ to perform the classification 
-and exploits the stable structure of the Internet routers 
-to help guide the clustering in question. 
+
+%%AD a better connection with the mismatch problem has to be made... 
+%%		.. in the form of 1-2 lines (or short statements). 
+%%
+%% VM Added some extra lines and restructured the text
+\emph{T2MC}'s~\cite{SLCGZ2008} objective is to diminish the redundant multiple
+messages that may unnecessarily cross \emph{AS}-boundaries as these are the
+links that cost more to ISPs thus minimizing the effects of the topology
+mismatch problem. To achieve that, each node exploits the stable structure of
+the Internet routers to help solve a clustering problem. Uses traceroute
+logs to detect inter-\emph{AS} boundaries and clusters other peers to near-by and
+far-away peers by applying a customized $k$-mean classification
+algorithm with $k = 2$ to perform the classification.
+Specifically when joining the network, the peer randomly picks an IP and uses
+traceroute. Skipping the details of the algorithm, it picks the routers with
+minimum and maximum latencies in the above traceroute path as initial centroids
+for the ``near'' and ``remote'' sets respectively and after some iterations it
+ends up with two sets having the minimum intra-set latency variance. The
+router with the minimum hops in the ``remote'' set is set up as the hop
+threshold and anything above that is classified as the ``remote'' router cluster
+for that peer. Similarly, the router with the maximum hops in the ``near'' set
+is important because it denotes an edge gateway. Peers that share the same
+edge gateway or other members of the ``near'' router cluster are considered
+close-by.
+
 Even though {\sl traceroute}
@@
 Although overall experimental results have been encouraging, 
-\emph{T2MC} is considered coarse-grained
-as it classifies peers into only ``remote'' and ``close''
-groups \cite{QLZG2009}.
-%%AD a better connection with the mismatch problem has to be made... 
-%%		.. in the form of 1-2 lines (or short statements). 
-%%
+\emph{T2MC} is considered rather coarse-grained as it only
+as it classifies neighbours into two groups \cite{QLZG2009}.
 %
@@
 \emph{PChord} \cite{HLYW2005} is based on the \emph{Chord DHT} and adds
-proximity into its routing mechanism. 
-The main modification over its predecessor is the inclusion 
-of a \emph{proximity list} into its routing table;
-in this way, the next hop can be decided based on both the best progress
-towards the key and the physical proximity of the candidate nodes.
-%%AD I am not sure about the "best progress" term - what does it really mean???
-%%
-At join time, this list in empty. 
-As the \emph{PChord} node starts to interact with
-others, it applies a heuristic mechanism to fill the list. 
-Entries are dynamically added or removed as the network state changes. 
-The overlay maintenance cost is similar to that of \emph{Chord}. 
-%%
-Apart from the costs incurred when a  node joins in or departs and 
-by heart-beat messages used for connection verification, \emph{PChord} adds
-heart-beat among nodes it the proximity list. 
-Fortunately the latter is kept to a minimum provided that 
-the list entails nodes presenting very low \emph{RTT}s.
-%%AD removed "lowest" to use lowest u must have a set or something...
-
-As there is always progress towards the target in the \emph{ID}-space,
-\emph{PChord} yields a reduced number of hopes that \emph{Chord}.
+proximity into its routing mechanism. To achieve that, the standard finger table
+is augmented with a \emph{proximity list} which is a structure that contains
+nodes in the network that are close-by, in terms of RTT, with the node at hand.
+%in this way, the next hop can be decided based on both progress
+%towards the key and the physical proximity of the candidate nodes.
+%%AD
+%% I am not sure about the "best progress" term - what does it really mean???
+%% VM changed the above slightly to address the above comment. Generally
+%% in DHTs when a node receives a search query for a key that it does not have
+%% the data for, it routes the request to that node in its routing table that
+%% is closest to the key. That is what I mean with best progress to the keys
+%% In topology aware systems and topology aware routing we have the requirement
+%% that the next hop should be the closest one which in many cases can be
+%% a competing requirement with the best progress to the key requirement I
+%% just described. 
+In this way, the next hop during routing can be decided based on both the
+direction towards the search key and minimizing the cost of that hop.
+
+At join time, this list in empty, but during the node's lifetime it is created
+and modified in a heuristic fashion. During routing, the next hop is chosen
+as the closest node to the target key from both the finger and the proximity
+lists. This means that the next hop defined in the finger list can only be
+substituted by a better hop (closest to the key and in less physical distance)
+with PChord resulting to equal or even more efficient routing than the pure
+vanilla Chord.
+%% VM
+%% Removed the text below that was a rephrasing of one of my own a little
+%% bit more below. Hope it is way more clear now.
+%%
+% As there is always progress towards the target in the \emph{ID}-space,
+% \emph{PChord} yields a reduced number of hopes that \emph{Chord}.
 %%AD rephrased as above - the 2 last lines of the text below are 
@@
 % at least equal in the key space.
+
+Unfortunatelly the overlay maintenance cost is a little bit more than
+\emph{Chord} additional heart-beat messages are required for
+connection verification of the proximity list. Fortunately, these nodes
+are already close-by so the additional cost is kept to a minimum but
+it has been reported to exhibit slow convergence speed 
+and inefficiency in high-churn environments~\cite{DK2006}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%
-Passing through proximity links, in the underlying network, 
-means reduced routing cost. 
-Moreover, if the number of entries in the proximity list is the same
-as the number of network partitions, 
-\emph{PChord} prevents hops from jumping in and
-out of network partitions the current node belongs to.
-%%
-On one hand, 
-\emph{PChord}’s advantageous  points are that 
-its routing attains fewer overlay hops, 
-it crosses network partition(s) only once by following proximity links
-and it maintains the light-weight maintenance cost of \emph{Chord}.
-On the other, \emph{PChord} exhibits slow convergence speed 
-and inefficiency in high-churn environments~\cite{DK2006}.
-%%AD rephrased what is below as above... 
-% \emph{PChord}’s advantages include the fact 
-% that the main routing optimizations are of
-% less overlay hops, following proximity links and crossing the same network
-% partition only once along a routing path while in the same time keeping Chord's
-% light-weight maintenance cost. 
-% On the other hand, it exhibits slow convergence
-	% speed and inefficiency in a high churn environment \cite{DK2006}.
+%% VM:
+%% Removed this whole chuck since most of it sounded like
+%% repetition of the rest of the text below after all.
+%% Please review and remove if you aggree.
+%%
+% Passing through proximity links, in the underlying network, 
+% means reduced routing cost. 
+% Moreover, if the number of entries in the proximity list is the same
+% as the number of network partitions, 
+% \emph{PChord} prevents hops from jumping in and
+% out of network partitions the current node belongs to.
+% On one hand, 
+% \emph{PChord}’s advantageous  points are that 
+% its routing attains fewer overlay hops, 
+% it crosses network partition(s) only once by following proximity links
+% and it maintains the light-weight maintenance cost of \emph{Chord}.
+% On the other, \emph{PChord} exhibits slow convergence speed 
+% and inefficiency in high-churn environments~\cite{DK2006}.
+% %%AD rephrased what is below as above... 
+% % \emph{PChord}’s advantages include the fact 
+% % that the main routing optimizations are of
+% % less overlay hops, following proximity links and crossing the same network
+% % partition only once along a routing path while in the same time keeping Chord's
+% % light-weight maintenance cost. 
+% % On the other hand, it exhibits slow convergence
+% 	% speed and inefficiency in a high churn environment \cite{DK2006}.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 an any-cast group to the physically closest node in
-that group~\cite{M2002}. 
-\emph{AChord} organizes all nodes participating in the
-overlay network into an any-cast group. 
-Any joining node comes outside of that group and, thus, is
-automatically forwarded to the physically nearest node for bootstrapping;
-this avoids the need to explicitly maintain such nodes, %%AD such nodes???
+that group~\cite{M2002}. Thus, \emph{AChord} organizes all nodes participating
+in the overlay network into an any-cast group. Any joining node comes (from
+outside) and is automatically forwarded to the physically nearest node for
+bootstrapping; this avoids the need to explicitly maintain participants
 and the effort of locating the physically nearest. 
 %%
+
 The \emph{ID} of an incoming 
-% the new-coming %%AD changed...
 node is computed based on the \emph{ID} of the bootstrap node 
@@
 information about the closest--known peers.
-The routing decision is made by using both adjacency and finger tables.
+
+The routing decision is made by using both the adjacency and finger tables
+based on the closest node to the target key in the key space (same as
+Chord).
 %%AD how are these 2 combined???
+% VM:
+% Check above.
+% This AChord algo looks pretty similar to PChord.
+% If you agree I would suggest we merge the two.
+
 \emph{AChord} behavior has as follows:
@@
 % \subsubsection{Chord6}
-\emph{Chord6}~\cite{XZHL2005} is a \emph{Chord}-variant that
+\emph{Chord6}~\cite{XZHL2005} is another \emph{Chord}-variant that
 exploits {\sl IPv6}'s  hierarchical features in order to create 
@@
 sector, are mapped to the same region as well, preserving their proximity
-relationship.
-%% unveiled by the use of the Vivaldi protocol. 
-%%AD I am not sure u need the above line at the end of the previous sentence.
-Individual elements
-are uniquely mapped to the identifier--space, 
+relationship. Individual elements are uniquely mapped to the identifier--space,
 allowing logarithmic lookup operations with high 
-probability on \emph{Chord}. 
-%%
-
-%%AD what is the section u say here??? -- also the first sentence 
-%%	here has to do a better connection with what was discussed before!
-The system described in the previous section 
-allows any node to obtain its \emph{DHT} key and its region key, in a fully
-distributed manner, just by applying a consistent hash function. 
-%%AD where these RPCs come from? kind of disconnected???
-Using an {\sl RPC} call in \emph{Chord}, 
-the node can obtain the region's master node, 
-called \emph{Cluster Node (CN)} which is responsible for clustering 
+probability on \emph{Chord}. Also that way a node can obtain its \emph{DHT}
+key and its region key, in a fully distributed manner, just by applying a
+consistent hash function. The node can then communicate with the region's
+master node called \emph{Cluster Node (CN)} which is responsible for clustering 
 the nodes belonging in the same sector or region with that of the node at hand. 
-An additional {\sl RPC} call, registers it to its corresponding region 
-and publishes its information, through the special peer \emph{CN}. 
-The node can, additionally, ask an \emph{CN} for other
+It registers to its corresponding region and publishes its information, through
+the special peer \emph{CN} and can also ask the latter for other
 nodes that have previously joined the region to add them into its
-neighbor set for future routing table optimization. Q
+neighbor set for future routing table optimization.
+%
+If no other nodes are on that region, the search is expanded to neighbouring
+regions.
+%
+%% VM rephrase the one bellow as above in a super compact form
 %%AD the one below is not a parse-able sentence!!! What does it say???
 %%%		restrict the length of your sentences to say at most 2.5 lines!
-Even in the case when no
-neighbor is detected in the current region, the search is expanded to adjacent
-regions and towards an upper layer identifier space until one is found or the
-first layer reached. Simulations suggest that the scheme has good convergence
-property of the delay distribution and good optimizing performance.
+%Even in the case when no
+%neighbor is detected in the current region, the search is expanded to adjacent
+%regions and towards an upper layer identifier space until one is found or the
+%first layer reached. Simulations suggest that the scheme has good convergence
+%property of the delay distribution and good optimizing performance.
+%
+%
 %%AD up to here the text in this second paragraph for Chord DHT remains 
@@
 Unfortunately, even though scaling appeared promising, only limited 
-simulations carried out with configuration involving up to 2,048 nodes 
-were carried out.
-% where made on rather small test beds (of at most 2048 nodes), so no safe
-% conclusions can be deduced. 
+simulations were carried out with configurations involving up to 2,048 nodes.
 Moreover the approach assumes uniform node distribution which
@@
 The \emph{Chord} cycle is constructed by sorting nodes in 
-ascending order. In \emph{Quasi-Chord},
-each node stores $2$ finger tables; one clockwise and one counter-clockwise.
+ascending order.
+%In \emph{Quasi-Chord},
+%each node stores $2$ finger tables; one clockwise and one counter-clockwise.
 %%AD what is the purpose of this? efficiency to go back and forth???
+% VM not relevant to the topology awareness so I removed the sentence above
 %%
@@
 % \subsubsection{IP-based Clustering (IPBC)}
-Proximity neighbor selection algorithms often use probing 
-% and other techniques
-%%AD say what are these "other" than simply say so... 
-to detect proximity. 
+Proximity neighbor selection algorithms often use probing to detect proximity. 
 However, such methods suffer by imprecisions often due to network overload.
-{\sl IP}-based clustering~\cite{KM2007} 
-% is a proximity neighbor selection algorithm which  %%AD u do not need this!
-uses {\sl IP} address prefixes ($16$ bit for {\sl IPv4}) 
-to detect proximity. 
-%%AD where this IPBC came from?
-\emph{IPBC} clustering generates a key and its prefix %%AD this ... "and" is not 						      %%    clear!!!
-and stores the prefix in the \emph{DHT} itself;
-in this way, newly joining nodes with the same {\sl IP} prefix 
-can query the \emph{DHT} and readily
-identify all neighbors with the same prefix.  
-Nodes periodically update their entries in the \emph{DHT} 
+{\sl IP}-based clustering (IPBC)~\cite{KM2007} uses {\sl IP} address prefixes
+($16$ bit for {\sl IPv4}) to detect proximity.
+Such proximity information (IP addresses of network participants) can be
+easily stored in the DHT itself.
+
+Each peer in the overlay has a unique ID and and an IP address to
+characterize it. To publish itself it hashes a fixed-length prefix
+its IP to obtain a key and stores its ID and IP in the DHT itself. 
+Newly joining nodes with the same {\sl IP} prefix 
+can, thus, query the \emph{DHT} and readily identify all neighbors with the
+same prefix. Nodes periodically update their entries in the \emph{DHT} 
 by removing those who leave or have timed-out.
-%%
-
 The overhead of the algorithm is very low as proximity detection requires only
@@
 is caught into a vicious cycle with the storage space needed for \emph{DHT} 
-object publishing (how many bits will be used for either). 
-%%AD did not get the parenthesis above - what is "either"???
-%%	I am not sure what the sentence below says and 
-%%	how is connect to the above
-Moreover, its static nature (no online measurements used) 
-makes it difficult to handle~\cite{EWF2011}.
+object publishing.
+Moreover, its static nature assumes high correlation of common IP prefix
+length of participating peers and latency (with no other measurements),
+which might not always be the case, especially in cases of high-churn
+environments.~\cite{EWF2011}.
 %%
@@
 
-%%AD how the 2-level thing helps the mismatch  problem?? 
-%%	ALSO, why the symbolism of the layers is required?? Ld_Cord Ld_Cone
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \subsubsection{Cone}
-\emph{Cone} \cite{HY2007} extends \emph{Chord} using a proximity neighbor
-selection topology optimization algorithm. The proximity information is
-generated using landmarks and round-trip-time-based distances to 
-those landmarks. %%AD added those.. 
-%%AD how is this better/worse than the one just above based on IP clustering???
-\emph{Cone} uses a two-layered identifier space. The first layer is 
-identical to that of \emph{Chord} and is denoted as $Id_{Chord}$.
-% \emph{Chord}-layer identifier and denoted as $Id_{Chord}$, 
-% is the same as in Chord. 
-The second layer identifier known as \emph{Cone}-layer and denoted as 
-$Id_{Cone}$ is generated using two IDs. 
-The first, known as \emph{group identifier (gid)} denotes a
-relevant group to which the node belongs to while the second, 
-namely \emph{local identifier (lid)} indicates the % local 
-identifier within the group. 
-The group is used here as means to divide nodes according to  a
-% concept is a way of dividing nodes according to a
-common $Id_{Chord}$ prefix.  
-The \emph{Cone} overlay retains the \emph{Chord}'s circular topology. 
-The difference lies in the fact that, now, $2$ rings are created. 
-A big one in which nodes with the same \emph{gid} are arranged 
-% at each position.  %%AD I do not get thus at EACH position!!!
-in the same position.
-Each of these positions constitute a smaller ring for the particular
-group's \emph{lid}s. 
-%%%
-Routing is achieved in both clockwise and
-counter-clockwise directions in the big-ring, for which two routing tables are
-maintained: \emph{front} and \emph{back} finger tables. 
-Entries in these tables, display physical network proximity 
-with the current node. 
-Moreover, a third table called \emph{group} maintains 
-information about other on-line peers
-within the current node's group in a way that 
-entries are now close in the \emph{id} space. %%AD check rephrasing...
-%
-%Routing in Cone, comprises of the inter-group algorithm and the intra-group
+%% VM Removing this algorithm in general
+%  The paper is so badly written that probably shows something about its
+%  quality...
+% 
+% %%AD how the 2-level thing helps the mismatch  problem?? 
+% %%	ALSO, why the symbolism of the layers is required?? Ld_Cord Ld_Cone
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% % \subsubsection{Cone}
+% \emph{Cone} \cite{HY2007} extends \emph{Chord} using a proximity neighbor
+% selection topology optimization algorithm. The proximity information is
+% generated using landmarks and round-trip-time-based distances to 
+% those landmarks.
+% 
+% \emph{Cone} uses a two-layered identifier space. The first layer is 
+% identical to that of Chord (from now on refered to as \emph{Chord}-ID)
+% and the second known as \emph{Cone}-ID which comprise a
+% \emph{group identifier (gid)} part and a \emph{local identifier (lid)} part
+% both being the prefix and the suffix of the Chord-ID itself, respectively.
+% In escence, the \emph{gid} denotes the relevant (based on a Chord-ID prefix) group to
+% which the node belongs to while the \emph{lid} indicates the local identifier
+% within that group. The Cone topology can be visualised as another ring on top
+% of Chord whose nodes are the groups and each group is each such Cone-node is a
+% ring with the nodes of that group.
+% 
+% Proximity information is used during initialization. After a new node computes
+% its Chord- and Cone-IDs
+% 
+% Routing in Cone, comprises of the inter-group algorithm and the intra-group
 % algorithm. First the group of nodes to which the desired key lies is
@@
 % 
-\emph{Cone} fares as follows: 
-\begin{center}
-{\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
-% The result of experiments show that the Cone, compared with Chord, has obvious
-% improved on the delay of route and the hops of overlay networks…
-medium &
-% The approach incorporates a static landmark + RTT to these landmarks in order
-% to exploit proximity information thus incurring low overall additional
-% operational overhead but imposing extra stress to the landmarks themselves.
-% Additional information must be maintained, three finger tables (front, back and
-% group)
-medium &
-% landmark based approaches are not full distributed thus not smoothly scalable.
-medium
-\end{tabular}
-}
-\end{center}
+% 
+% \emph{Cone} overlay retains the \emph{Chord}'s circular topology. 
+% The difference lies in the fact that, now, $2$ rings are created. 
+% A big one in which nodes with the same \emph{gid} are arranged 
+% % at each position.  %%AD I do not get thus at EACH position!!!
+% in the same position.
+% Each of these positions constitute a smaller ring for the particular
+% group's \emph{lid}s. 
+% %%%
+% Routing is achieved in both clockwise and
+% counter-clockwise directions in the big-ring, for which two routing tables are
+% maintained: \emph{front} and \emph{back} finger tables. 
+% Entries in these tables, display physical network proximity 
+% with the current node. 
+% Moreover, a third table called \emph{group} maintains 
+% information about other on-line peers
+% within the current node's group in a way that 
+% entries are now close in the \emph{id} space. %%AD check rephrasing...
+% %
+% %Routing in Cone, comprises of the inter-group algorithm and the intra-group
+% %  algorithm. First the group of nodes to which the desired key lies is
+% % detected, exploiting physical proximity information (front and back finger
+% % tables). Second during the next and last hop, the message is forwarded to the
+% % exact node that contains the desired key.
+% %
+% 
+% Compared to IP-based clustering we saw before measuring RTTs incurs more
+% overhead to the network but its efficiency, especially in a very dynamic
+% environment is better.
+% 
+% \emph{Cone} fares as follows: 
+% \begin{center}
+% {\footnotesize
+% \begin{tabular}{ccc}
+% \emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+% \hline
+% % The result of experiments show that the Cone, compared with Chord, has obvious
+% % improved on the delay of route and the hops of overlay networks…
+% medium &
+% % The approach incorporates a static landmark + RTT to these landmarks in order
+% % to exploit proximity information thus incurring low overall additional
+% % operational overhead but imposing extra stress to the landmarks themselves.
+% % Additional information must be maintained, three finger tables (front, back and
+% % group)
+% medium &
+% % landmark based approaches are not full distributed thus not smoothly scalable.
+% medium
+% \end{tabular}
+% }
+% \end{center}
 
@@
 Although based on the \emph{Pastry} overlay and 
-it well understood built-in locality heuristics ~\cite{CDCR2002a},
+its well understood built-in locality heuristics~\cite{CDCR2002a},
 \emph{DynaMO}'s own mechanisms are \emph{DHT}-independent and 
 so, they can be applied to other \emph{DHT}s.
-%%AD rephrased/shortened above - check
-%%
 \emph{DynaMO}'s approach may be viewed as bottom-up. 
@@
 \emph{Closest Neighbor Prefix Assignment (CNPA)}.
-%%AD rephrased as above - check!
-% A new-coming node is required
-% to compute physical information about its neighborhood 
-% and use it to unilaterally assign itself the an ID. 
-% \emph{DynaMO} is building on the
-% observation that with every new overlay routing hop the physical distance
-% between nodes is likely to increase, resulting in a dominating last routing
-% step \cite{antony_pastry_2001,CDCR2002a} and for this two methods are discussed,
-% \emph{random landmarking (RLM)} and \emph{closest neighbor prefix assignment
-% (CNPA)}.
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 In \emph{RLM}, a set of peers are chosen to be responsible 
-for a set of landmark keys. 
-These keys have to be chosen in a way that they equally divide the overlay
-space. 
-%%AD how do you accomplish the above?? 
-An incoming peer then computes distances from those landmarks,
+for a set of landmark keys. These keys have to be chosen in a way that they
+equally divide the overlay space (i.e. in a 16 base an appropriate split
+would be 000..0, 100..0, 200..00, ... , F00..00).
+An new-coming peer then computes distances from those landmarks,
 sorts them, and assigns itself an \emph{ID} on this 
 landmark ordering as follows:
-% The new-coming peer then computes distances to those
-% landmarks, sorts them and assigns itself an ID based on the landmark ordering.
 the peer adopts as \emph{ID} having a prefix from its closest
@@
 In this way, two desirable features are attained:
-% What is achieved is that
 \begin{inparaenum}[\itshape i\upshape)]
@@
 	physically close by.
-	%%AD I am not sure why is that?? But it is ok...
 \end{inparaenum}
@@
 last routing step.
-% (since it contains the nodes closest in the overlay space).
-%%AD I do not think you need the above.. %%AD Also rewriting..
-%%
-%%
+
 \emph{CNPA} is proposed for network setups for which a lightweight
@@
 Unfortunately, \emph{RLM} bootstrap overhead is substantial and can 
-effect  the set of temporary landmarks.
+effect the set of landmarks.
 \emph{CNPA} on the other hand, incurs less overhead but this come 
-at the expense of being more coarse-grained.
-%%AD  Rephrased as above.  -- In the above I am not clear 
-%%	a) what are the "temporary" landmarks? b) what is coarse-grained 
-%%	and why??? 
+at the expense of being more coarse-grained as it is based on prefixes
+to determine proximity during.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % definition of Topology Mismatch Problem in the intro/background.
-The notion of \emph{stretch} is used as th means for quantifying 
+The notion of \emph{stretch} is used as the means for quantifying 
 the \emph{topology match degree} of the constructed overlay;
@@
 %%AD what are the indexes: l and p -- p for peer and l for new peer???
+%% VM: no l for logical and p for physical
 %%AD rephrased what is below as above... 
@@
 During the \emph{jumping phase}, the source node calculates the 
-stretch change of its \emph{TTL--1} neighbors %%AD each of its neighbors???
-and that of the \emph{TTL-1} neighbors of the first of
-the peers picked during the previous phase. 
-These calculations are made as if the jump had been made. 
-Should the stretch reduction is over a predefined threshold
-the jump is actually performed. Otherwise,
-the second selected candidate is selected and the
-same computation is performed. 
+stretch change to all its own direct neighbours and to the direct neighbours
+of the first peer that was selected in the probing phase, as if
+the jump has been performed. Using heuristics the node decides if the stretch
+reduction is enough to perform the actual jump, otherwise it selects the
+second peer from the probing phase and repeats the computation. If again, the
+threshold is not met, then no jump is ultimately carried out.
 %%AD is this done only 2 times?? Or is it done recursively??
-If again, the threshold is not met, then no jump is ultimately carried out. 
-In case of a jump, this is performed as a combination
-of \emph{leave} and \emph{join} operations, in the underlying context
-(i.e., \emph{CAN}).
-%%AD I am not sure of the last sentence + CAN... What does it say?? context???
-
-%%AD The following sentence says what exactly????? 
-% Even though no theoretical proof is given, valid results are
-% obtained and demonstrated through experimentations. 
-%%%%%%%
-% It is argued that since it
-% is observed to achieve a $27\%$ stretch reduction when compared to a landmark
-% enhanced CAN-system, it can be deduced that it performs better than landmark
-% binning in terms of the matching degree.
-%%AD rephrased above as below.. .
+%%VM no it stops
+%% Quoting from the paper
+% Our experiments
+% show that targeting the two zones associated with the two
+% closest nodes is most effective. This is because adaptive
+% and iterative local optimizations will quickly or eventually
+% group the physically close nodes in the same overlay zone.
+% If a node outside a zone finds a close node in the zone, the
+% rest of the nodes will be likely to be or eventually to be close
+% to it.
 Experimentation shows that \emph{SAT-Match} 
@@
 Based on this result, it is conjured that \emph{SAT-Match} 
-may perform better than landmark binning in terms of the matching degree. 
+may perform better than landmark binning in terms of the matching degree
+of the logical to the physical networks. 
 \emph{SAT-Match} behaves as follows with respect to the stated $3$ criteria:
-%%AD matching degree is unqualified here... 
 %%
@@
 %%AD the categorization is due to 3 above citations??? Us?? ;-)
+%% VM: As mentioned in the intro this categorization is due to the 3 above
+% citations
+%
 % A categorization of the methods used for achieving their goals is the
@@
 highly affected by the application domain and the implementation itself.
-%%AD this is like instructions for manual - I think the point is made and 
-%%	so we can avoid this.. 
-% Thus, the reader is advised to filter the results of each approach and try to
-% think about how the points presented will work in her own particular problem.
 
@@
 highlight each algorithm's pros and cons.
-%%AD rephrased what is below as above..
-% A quick overview can be obtained in Table~\ref{structured:table}. 
-% Topology adaptation, landmarking and caching/replication have been visited during the
-% discussion on unstructured systems in Section~\ref{section:unstructured} so
-% we will not elaborate more here. 
-% As in the previous section, we provide the
-% key properties of each algorithm that differentiate it from the others and
-% highlight each algorithm's advantages and disadvantages.
 
@@
 
-% In the \emph{geographic layout (GL)} category fall methods which practically
-This category consists of methods that 
-position physically nearby nodes together in the application space as 
-Figure~\ref{figure:geographic-layout} depicts. 
-This is accomplished by adjusting and maintaining the routing tables 
-of all involved peers by exploiting proximity information 
-that describe the overall geographic positioning of peers. 
+%% In the \emph{geographic layout (GL)} category fall methods which practically
+%This category consists of methods that 
+%position physically nearby nodes together in the application space as 
+%Figure~\ref{figure:geographic-layout} depicts. 
+%This is accomplished by adjusting and maintaining the routing tables 
+%of all involved peers by exploiting proximity information 
+%that describe the overall geographic positioning of peers. 
 %%AD I am not sure I follow the 1-line above...
-%%
+
+%% VM: I reworded the above as below.
+This category consists of methods that try to map the overlay id space on to
+the physical one, as depicted in Figure~\ref{figure:geographic-layout}.
+Routing tables of involved peers, are adjusted and maintained by exploiting
+proximity information that describe the overall (geographic) positioning of
+the participants, in contrast to proximity routing and proximity neighbour
+selection where decisions are taken based on and effecting a peer's nearby
+environment.
+
 \begin{figure}[ht]
@@
 Landmark servers are not self-organizing and present maintenance
-overheads. 
-To serve a \p--network with millions of peers, multiple landmark
-servers distributed over the world is required, a difficult proposition.
+overheads. Moreover, to serve a \p--network with millions of peers, multiple
+such servers are required to be distributed over the world, which is a difficult
+proposition per se.
 \emph{RTT}s can measure the delay between peers, but their use constitutes a
@@
 
-On one hand, managing the overlay structure based on \emph{geographic layout} 
-of the nodes improves the query efficiency of the system in general. 
-On the other, it tends to undermine the uniform distribution 
-of peer identifiers which in turn terns to create 
-hot-spots and limited node resilience to failures.
-Such resilience appears to suffer as
-nearby nodes are more likely to suffer collective failures~\cite{HY2007}.
-%%AD changed the wording above - check...
-
-% must
-% cope with the inherent trade os associated with selecting
-% identiers based on location: randomly selected identiers
-% are essential for the scalability and security of the system
-% and help to balance load while location-aware identiers pre-
-% vent messages from taking extensive scenic routes (on the
-% physical level).
+On one hand, approaches based on \emph{geographic layout} improve the query
+efficiency of the system in general and are especially effective when applied
+to CAN(-like) implementations. Unfortunatelly, in the same time, undermine the
+uniform distribution of peer identifiers which in turn tend to create 
+hot-spots and limited node resilience to failures. Such resilience appears to
+suffer as nearby nodes are more likely to suffer collective
+failures~\cite{HY2007}.
+%% VM reworded the above paragraph, please check it out.
 
@@
 
-This category of approaches % ({\emph{PR}) 
-does not require routing tables to be built using any
-knowledge about network proximity. 
-%%AD so if no knowledge is required what happens? How you go about 
+%This category of approaches % ({\emph{PR}) 
+%does not require routing tables to be built using any
+%knowledge about network proximity. 
+%%%AD so if no knowledge is required what happens? How you go about 
+%%%%
+%%%AD I am not sure what the following sentence says... ;-) 
+%On the other hand it exploits such knowledge
+%to choose the best next hop when routing a message as can be seen in
+%Figure~\ref{figure:proximity-routing}. 
+%%%AD The above opening 2 sentences require re-calibration.. 
 %%%
-%%AD I am not sure what the following sentence says... ;-) 
-On the other hand it exploits such knowledge
-to choose the best next hop when routing a message as can be seen in
-Figure~\ref{figure:proximity-routing}. 
-%%AD The above opening 2 sentences require re-calibration.. 
-%%
+
+%% VM Rewrote above as below
+This category of approaches use physical proximity knowledge to decide the best
+next hop during routing. As shown in Figure~\ref{figure:proximity-routing}
+maintenance of routing tables themselves is a separate process based only on the
+system's DHT construction mechanism and overlay id distribution.
+
 \begin{figure}[ht]
@@
 %%
-\emph{PR} tries to balance between choosing the node that will
-further progress the routing towards the destination 
-and picking the closest entry in the routing table
+When \emph{PR} based algorithms route messages, they try to balance between
+choosing the node that will further progress the routing towards the
+destination and picking the closest entry in the routing table
 in terms of network proximity. 
@@
 %%	in order for the contrast here to be valid..
-As this is very difficult to achieve,  %%AD Why so? Why is it difficult??? 
-the system might end up
-using a greedy approach that selects low latency paths on the overlay
-which ultimately maps to (suboptimal) longer paths on the physical topology. 
-%%AD in the following you make a point that should be made in the 
-%%	follow up segment - no???
-Thus, in general, proximity neighbor selection is considered 
-as a superior method to proximity routing, however, joint uses of these two protocols are also possible. 
-Moreover, they are  %%AD who is they? Confused!
-less effective than geographical layout when applied to CAN(-like)
-implementations.
+%%
+%% VM Is the above now better with the rewording of the section's opening paragraph?
+In the degenerated case that the system ends up using a greedy approach that
+selects only the lowest latency paths on the overlay, can, ultimately, result in
+more hops to reach the destination thus yielding suboptimal, longer paths on
+the physical topology as well. PR is also less less effective than geographical
+layout when applied to CAN(-like) implementations.
 %%AD re-work carefully this segment...
+%% VM how is it now?
 
@@
 % \emph{Proximity Neighbor Selection (PNS)} 
-Algorithms in this category populate their routing tables using
-proximity knowledge as Figure~\ref{figure:proximity-neighbour-selection} 
-depicts. 
+Algorithms in this category use proximity data during routing table construction
+and maintenance as shown in Figure~\ref{figure:proximity-neighbour-selection}.
 %%
@@
 here, \emph{TTL} values between nodes (i.e., \emph{SAT-Match}) or 
-\emph{IP} address prefixes (i.e., \emph{DynaMO} or \emph{Cone}) 
+\emph{IP} address prefixes (i.e., \emph{DynaMO}) % or \emph{Cone}) %%VM removing Cone from paper
 are predominantly used to detect proximity. 
@@
 %%AD perhaps the ending of the previous segment should be used here...
+%% VM: added below
+In general, proximity neighbor selection is considered 
+as a superior method to proximity routing, however, joint uses of the 
+approaches are possible.
 

### 2015-04-19 intro.tex
2014.10.05 (Vassilis) Clean-up, rewording and reviewing the unstructured section. Efficiency, overhead and scalability definitions reworked in background section.
2015.04.19 (Vassilis) Reformatting using the IEEE LaTeX template. Various integration changes.

@@
 
-\emph{Peer-to-peer} (\p) networks are self-organizing, distributed systems where
+\IEEEPARstart{P}{eer-to-peer (\p)} networks are self-organizing, distributed systems where
 participating nodes, called \emph{peers}, act as both resource providers and
@@
 
-% TODO SOME EXTRA REFERENCES IF NEEDED FOR THE USE OF OVERLAYS IN GENERAL
-%
-% IF ULTIMATELY DELETED CHECK TO REMOVE THE REFERENCES INCLUDED SINCE MOST
-% PROBABLY ARE NOT USED ELSEWHERE.
-%
-%find citation for Narada & Adobe's Real Time Media Flow Protocol (RTMFP)
-%\cite{} that provide quality-of-service (QoS) guarantees. \emph{IP
-%multicast}, as well as other proposals such as \emph{IntServ} and
-%\emph{DiffServ} \cite{cisco_diffserv_2005}, have still not seen wide deployment
-%largely due to the fact they require support from the IP
-%infrastructure. On the other hand, an overlay network can be deployed on
-%end-systems running the overlay protocol software, without the cooperation of ISPs.
-%Academic research includes
-%\cite{chu_esm_2000,jannotti_overcast_2000,kwon_tag_2002} among others.
-
 \p\ systems are implemented using \emph{overlay networks}. 
@@
 \centering
-  \includegraphics[scale=0.6]{img/pdf/under-over-lay.pdf}
+  \includegraphics[scale=0.53]{img/pdf/under-over-lay.pdf}
 \caption{An example overlay network.}

### 2015-04-19 mard-survey.tex
2013.07.28 (Vassilis) Further investigation conducted to find research work that references the papers/algorithms we survey and cite. Text enriched.
2015.04.19 (Vassilis) Reformatting using the IEEE LaTeX template. Various integration changes.

@@
-%\documentclass[a4paper,10pt]{article}
-\documentclass[acmcsur,acmnow]{acmtrans2m}
-%\documentclass[acmtocl,acmnow]{acmtrans2m}
+%% Based on:
+%% bare_jrnl.tex
+%% V1.4a
+%% 2014/09/17
+%% by Michael Shell
+%% see http://www.michaelshell.org/
+%% for current contact information.
+%%
+%% This is a skeleton file demonstrating the use of IEEEtran.cls
+%% (requires IEEEtran.cls version 1.8a or later) with an IEEE
+%% journal paper.
+%%
+%% Support sites:
+%% http://www.michaelshell.org/tex/ieeetran/
+%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
+%% and
+%% http://www.ieee.org/
 
-%\acmVolume{V}
-%\acmNumber{N}
-%\acmYear{YY}
-%\acmMonth{Month}
+%%*************************************************************************
+%% Legal Notice:
+%% This code is offered as-is without any warranty either expressed or
+%% implied; without even the implied warranty of MERCHANTABILITY or
+%% FITNESS FOR A PARTICULAR PURPOSE! 
+%% User assumes all risk.
+%% In no event shall IEEE or any contributor to this code be liable for
+%% any damages or losses, including, but not limited to, incidental,
+%% consequential, or any other damages, resulting from the use or misuse
+%% of any information contained here.
+%%
+%% All comments are the opinions of their respective authors and are not
+%% necessarily endorsed by the IEEE.
+%%
+%% This work is distributed under the LaTeX Project Public License (LPPL)
+%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
+%% distributed and modified. A copy of the LPPL, version 1.3, is included
+%% in the base LaTeX documentation of all distributions of LaTeX released
+%% 2003/12/01 or later.
+%% Retain all contribution notices and credits.
+%% ** Modified files should be clearly indicated as such, including  **
+%% ** renaming them and changing author support contact information. **
+%%
+%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
+%%                    bare_conf.tex, bare_jrnl.tex, bare_conf_compsoc.tex,
+%%                    bare_jrnl_compsoc.tex, bare_jrnl_transmag.tex
+%%*************************************************************************
+
+% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
+% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
+% *** with production work. IEEE's font choices and paper sizes can       ***
+% *** trigger bugs that do not appear when using other class files.       ***                          ***
+% The testflow support page is at:
+% http://www.michaelshell.org/tex/testflow/
+
+\documentclass[journal]{IEEEtran}
+%
+% If IEEEtran.cls has not been installed into the LaTeX system files,
+% manually specify the path to it like:
+% \documentclass[journal]{../sty/IEEEtran}
+
+% Some very useful LaTeX packages include:
+% (uncomment the ones you want to load)
+
+% *** MISC UTILITY PACKAGES ***
+%
+%\usepackage{ifpdf}
+% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
+% compilation based on whether the output is pdf or dvi.
+% usage:
+% \ifpdf
+%   % pdf code
+% \else
+%   % dvi code
+% \fi
+% The latest version of ifpdf.sty can be obtained from:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
+% Also, note that IEEEtran.cls V1.7 and later provides a builtin
+% \ifCLASSINFOpdf conditional that works the same way.
+% When switching from latex to pdflatex and vice-versa, the compiler may
+% have to be run twice to clear warning/error messages.
+
+% definitions
+%\newtheorem{theorem}{Theorem}[section]
+%\newtheorem{conjecture}[theorem]{Conjecture}
+%\newtheorem{corollary}[theorem]{Corollary}
+%\newtheorem{proposition}[theorem]{Proposition}
+%\newtheorem{lemma}[theorem]{Lemma}
+%\newdef{definition}[theorem]{Definition}
+%\newdef{remark}[theorem]{Remark}
+\newtheorem{definition}{Definition}
+
+\newcommand{\p}{\emph{P2P}}
+
+% *** CITATION PACKAGES ***
+%
+%\usepackage{cite}
+% cite.sty was written by Donald Arseneau
+% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
+% \cite{} output to follow that of IEEE. Loading the cite package will
+% result in citation numbers being automatically sorted and properly
+% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
+% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
+% \cite will automatically add leading space, if needed. Use cite.sty's
+% noadjust option (cite.sty V3.8 and later) if you want to turn this off
+% such as if a citation ever needs to be enclosed in parenthesis.
+% cite.sty is already installed on most LaTeX systems. Be sure and use
+% version 5.0 (2009-03-20) and later if using hyperref.sty.
+% The latest version can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
+% The documentation is contained in the cite.sty file itself.
+
+% *** GRAPHICS RELATED PACKAGES ***
+%
+\ifCLASSINFOpdf
+  \usepackage[pdftex]{graphicx}
+  % declare the path(s) where your graphic files are
+  \graphicspath{{./img/pdf/}}
+  % and their extensions so you won't have to specify these with
+  % every instance of \includegraphics
+  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
+\else
+  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
+  % will default to the driver specified in the system graphics.cfg if no
+  % driver is specified.
+  \usepackage[dvips]{graphicx}
+  % declare the path(s) where your graphic files are
+  \graphicspath{{./img/eps/}}
+  % and their extensions so you won't have to specify these with
+  % every instance of \includegraphics
+  % \DeclareGraphicsExtensions{.eps}
+\fi
+% graphicx was written by David Carlisle and Sebastian Rahtz. It is
+% required if you want graphics, photos, etc. graphicx.sty is already
+% installed on most LaTeX systems. The latest version and documentation
+% can be obtained at: 
+% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
+% Another good source of documentation is "Using Imported Graphics in
+% LaTeX2e" by Keith Reckdahl which can be found at:
+% http://www.ctan.org/tex-archive/info/epslatex/
+%
+% latex, and pdflatex in dvi mode, support graphics in encapsulated
+% postscript (.eps) format. pdflatex in pdf mode supports graphics
+% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
+% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
+% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
+% which can result in "jaggedy"/blurry rendering of lines and letters as
+% well as large increases in file sizes.
+%
+% You can find documentation about the pdfTeX application at:
+% http://www.tug.org/applications/pdftex
+
+% *** MATH PACKAGES ***
+%
+%\usepackage[cmex10]{amsmath}
+% A popular package from the American Mathematical Society that provides
+% many useful and powerful commands for dealing with mathematics. If using
+% it, be sure to load this package with the cmex10 option to ensure that
+% only type 1 fonts will utilized at all point sizes. Without this option,
+% it is possible that some math symbols, particularly those within
+% footnotes, will be rendered in bitmap form which will result in a
+% document that can not be IEEE Xplore compliant!
+%
+% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
+% thus preventing page breaks from occurring within multiline equations. Use:
+%\interdisplaylinepenalty=2500
+% after loading amsmath to restore such page breaks as IEEEtran.cls normally
+% does. amsmath.sty is already installed on most LaTeX systems. The latest
+% version and documentation can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
+
+% *** SPECIALIZED LIST PACKAGES ***
+%
+%\usepackage{algorithmic}
+% algorithmic.sty was written by Peter Williams and Rogerio Brito.
+% This package provides an algorithmic environment fo describing algorithms.
+% You can use the algorithmic environment in-text or within a figure
+% environment to provide for a floating algorithm. Do NOT use the algorithm
+% floating environment provided by algorithm.sty (by the same authors) or
+% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
+% algorithm float types and packages that provide these will not provide
+% correct IEEE style captions. The latest version and documentation of
+% algorithmic.sty can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
+% There is also a support site at:
+% http://algorithms.berlios.de/index.html
+% Also of interest may be the (relatively newer and more customizable)
+% algorithmicx.sty package by Szasz Janos:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
+
+% *** ALIGNMENT PACKAGES ***
+%
+%\usepackage{array}
+% Frank Mittelbach's and David Carlisle's array.sty patches and improves
+% the standard LaTeX2e array and tabular environments to provide better
+% appearance and additional user controls. As the default LaTeX2e table
+% generation code is lacking to the point of almost being broken with
+% respect to the quality of the end results, all users are strongly
+% advised to use an enhanced (at the very least that provided by array.sty)
+% set of table tools. array.sty is already installed on most systems. The
+% latest version and documentation can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/required/tools/
+
+% IEEEtran contains the IEEEeqnarray family of commands that can be used to
+% generate multiline equations as well as matrices, tables, etc., of high
+% quality.
+\usepackage{longtable}
+
+% *** SUBFIGURE PACKAGES ***
+%\ifCLASSOPTIONcompsoc
+%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
+%\else
+%  \usepackage[caption=false,font=footnotesize]{subfig}
+%\fi
+% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
+% for subfigure.sty, the latter of which is no longer maintained and is
+% incompatible with some LaTeX packages including fixltx2e. However,
+% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
+% which will override IEEEtran.cls' handling of captions and this will result
+% in non-IEEE style figure/table captions. To prevent this problem, be sure
+% and invoke subfig.sty's "caption=false" package option (available since
+% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
+% handling of captions.
+% Note that the Computer Society format requires a larger sans serif font
+% than the serif footnote size font used in traditional IEEE formatting
+% and thus the need to invoke different subfig.sty package options depending
+% on whether compsoc mode has been enabled.
+%
+% The latest version and documentation of subfig.sty can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
+\usepackage{subfigure}
+
+% *** FLOAT PACKAGES ***
+%
+%\usepackage{fixltx2e}
+% fixltx2e, the successor to the earlier fix2col.sty, was written by
+% Frank Mittelbach and David Carlisle. This package corrects a few problems
+% in the LaTeX2e kernel, the most notable of which is that in current
+% LaTeX2e releases, the ordering of single and double column floats is not
+% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
+% single column figure to be placed prior to an earlier double column
+% figure. The latest version and documentation can be found at:
+% http://www.ctan.org/tex-archive/macros/latex/base/
+
+%\usepackage{stfloats}
+% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
+% the ability to do double column floats at the bottom of the page as well
+% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
+% LaTeX2e). It also provides a command:
+%\fnbelowfloat
+% to enable the placement of footnotes below bottom floats (the standard
+% LaTeX2e kernel puts them above bottom floats). This is an invasive package
+% which rewrites many portions of the LaTeX2e float routines. It may not work
+% with other packages that modify the LaTeX2e float routines. The latest
+% version and documentation can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
+% Do not use the stfloats baselinefloat ability as IEEE does not allow
+% \baselineskip to stretch. Authors submitting work to the IEEE should note
+% that IEEE rarely uses double column equations and that authors should try
+% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
+% packages (also by Sigitas Tolusis) as IEEE does not format its papers in
+% such ways.
+% Do not attempt to use stfloats with fixltx2e as they are incompatible.
+% Instead, use Morten Hogholm'a dblfloatfix which combines the features
+% of both fixltx2e and stfloats:
+%
+% \usepackage{dblfloatfix}
+% The latest version can be found at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/dblfloatfix/
+
+%\ifCLASSOPTIONcaptionsoff
+%  \usepackage[nomarkers]{endfloat}
+% \let\MYoriglatexcaption\caption
+% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
+%\fi
+% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
+% Axel Sommerfeldt. This package may be useful when used in conjunction with 
+% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
+% submissions have lists of figures/tables at the end of the paper and that
+% figures/tables without any captions are placed on a page by themselves at
+% the end of the document. If needed, the draftcls IEEEtran class option or
+% \CLASSINPUTbaselinestretch interface can be used to increase the line
+% spacing as well. Be sure and use the nomarkers option of endfloat to
+% prevent endfloat from "marking" where the figures would have been placed
+% in the text. The two hack lines of code above are a slight modification of
+% that suggested by in the endfloat docs (section 8.4.1) to ensure that
+% the full captions always appear in the list of figures/tables - even if
+% the user used the short optional argument of \caption[]{}.
+% IEEE papers do not typically make use of \caption[]'s optional argument,
+% so this should not be an issue. A similar trick can be used to disable
+% captions of packages such as subfig.sty that lack options to turn off
+% the subcaptions:
+% For subfig.sty:
+% \let\MYorigsubfloat\subfloat
+% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
+% However, the above trick will not work if both optional arguments of
+% the \subfloat command are used. Furthermore, there needs to be a
+% description of each subfigure *somewhere* and endfloat does not add
+% subfigure captions to its list of figures. Thus, the best approach is to
+% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
+% and instead reference/explain all the subfigures within the main caption.
+% The latest version of endfloat.sty and its documentation can obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
+%
+% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
+% later in the document, say, to conditionally put the References on a 
+% page by themselves.
+
+% *** PDF, URL AND HYPERLINK PACKAGES ***
+%
+\usepackage{url}
+% url.sty was written by Donald Arseneau. It provides better support for
+% handling and breaking URLs. url.sty is already installed on most LaTeX
+% systems. The latest version and documentation can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/url/
+% Basically, \url{my_url_here}.
+
+% *** Do not adjust lengths that control margins, column widths, etc. ***
+% *** Do not use packages that alter fonts (such as pslatex).         ***
+% There should be no need to do such things with IEEEtran.cls V1.6 and later.
+% (Unless specifically asked to do so by the journal or conference you plan
+% to submit to, of course. )
 
@@
 \usepackage{paralist}
-\usepackage{subfigure}
-\usepackage{graphicx}
 \usepackage{colortbl}
 \usepackage{color,soul}
-\usepackage{longtable}
 \usepackage{wasysym}
@@
 \usetikzlibrary{patterns}
-\usepackage{url}
+\usepackage{supertabular}
 
-% definitions
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{conjecture}[theorem]{Conjecture}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{lemma}[theorem]{Lemma}
-\newdef{definition}[theorem]{Definition}
-\newdef{remark}[theorem]{Remark}
+% correct bad hyphenation here
+\hyphenation{op-tical net-works semi-conduc-tor}
 
-\newcommand{\p}{\emph{P2P}}
+\begin{document}
+%
+% paper title
+% Titles are generally capitalized except for words such as a, an, and, as,
+% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
+% not capitalized unless they are the first or last word of the title.
+% Linebreaks \\ can be used within to get better formatting as desired.
+% Do not put math or special symbols in the title.
+\title{Alleviating the Topology Mismatch Problem in Distributed Overlay Networks}
+%
+%
+% author names and IEEE memberships
+% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
+% a structure at a ~ so this keeps an author's name from being broken across
+% two lines.
+% use \thanks{} to gain access to the first footnote area
+% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
+% was not built to handle multiple paragraphs
+%
 
-%\newtheorem{theorem}{Theorem}[section]
-%\newtheorem{definition}[theorem]{Definition}
+\author{Vassilis Moustakas,
+        H\"useyin Akcan,
+        Mema Roussopoulos,
+        and~Alex Delis,~\IEEEmembership{Member,~IEEE}% <-this % stops a space
+\thanks{Vassilis Moustakas, Mema Roussopoulos and Alex Delis are with the Department
+of Informatics and Telecommunications, National and Kapodistrian University of Athens,
+GR15784, Athens, Greece (e-mail: vsmoustakas@gmail.com, mema@di.uoa.gr, ad@di.uoa.gr.}% <-this % stops a space
+\thanks{H\"useyin Akcan is with the Department of Software Engineering,
+Izmir University of Economics, Izmir, Turkey (email: huseyin.akcan@ieu.edu.tr).}% <-this % stops a space
+\thanks{Manuscript received June 30, 2015; revised July 31, 2015.}}
 
-%\textwidth      6.5in
-%\textheight     9.2in
-%\oddsidemargin  -0.5cm
-%\evensidemargin -0.5cm
-%\parindent      0pt              
-%\topmargin      -0.3in
-%\headsep        20pt
+% note the % following the last \IEEEmembership and also \thanks - 
+% these prevent an unwanted space from occurring between the last author name
+% and the end of the author line. i.e., if you had this:
 % 
-%\parskip        5pt
-%\parindent      10pt
+% \author{....lastname \thanks{...} \thanks{...} }
+%                     ^------------^------------^----Do not want these spaces!
+%
+% a space would be appended to the last name and could cause every name on that
+% line to be shifted left slightly. This is one of those "LaTeX things". For
+% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
+% "AB" then you have to do: "\textbf{A}\textbf{B}"
+% \thanks is no different in this regard, so shield the last } of each \thanks
+% that ends a line with a % and do not let a space in before the next \thanks.
+% Spaces after \IEEEmembership other than the last one are OK (and needed) as
+% you are supposed to have spaces between the names. For what it is worth,
+% this is a minor point as most people would not even notice if the said evil
+% space somehow managed to creep in.
 
-\markboth{}{}{}
+% The paper headers
+\markboth{IEEE COMMUNICATION SURVEYS \& TUTORIALS, VOL. XX, NO. YY, Z QUARTER YYYY}%
+{Moustakas \MakeLowercase{\textit{et al.}}: Alleviating the Topology Mismatch Problem in Distributed Overlay Networks}
+% The only time the second header will appear is for the odd numbered pages
+% after the title page when using the twoside option.
+% 
+% *** Note that you probably will NOT want to include the author's ***
+% *** name in the headers of peer review papers.                   ***
+% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
+% you desire.
 
-\title{Alleviating the Topology Mismatch Problem in Distributed Overlay Networks}
+% If you want to put a publisher's ID mark on the page you can do it like
+% this:
+%\IEEEpubid{0000--0000/00\$00.00~\copyright~2014 IEEE}
+% Remember, if you use this you must call \IEEEpubidadjcol in the second
+% column for its text to clear the IEEEpubid mark.
 
-\author{
-VASSILIS S. MOUSTAKAS$^1$, H\"USEYIN AKCAN$^2$, MEMA ROUSSOPOULOS$^1$ and ALEX
-DELIS$^1$\\
-$^1$\emph{University of Athens}, \emph{Athens}, \emph{Greece}\\
-$^2$\emph{Izmir University of Economics}, \emph{Izmir}, \emph{Turkey}
-}
+% use for special paper notices
+%\IEEEspecialpapernotice{(Invited Paper)}
 
+% make the title area
+\maketitle
+
+% As a general rule, do not put math, special symbols or citations
+% in the abstract or keywords.
 \begin{abstract}
 \emph{Peer-to-peer} (\p) systems have enjoyed immense attention and have been
-widely deployed on the Internet for well over a decade.
+widely deployed on the Internet for well ov, er a decade.
 They are implemented via an overlay network abstraction, on top 
@@
 reference for developers, system architects and researchers in the field.
-
 \end{abstract}
 
-\category{A.1}{General Literature}{Introductory and Survey}
-\category{C.2.4}{Distributed Systems}{Distributed Applications}
+% Note that keywords are not normally used for peerreview papers.
+\begin{IEEEkeywords}
+Distributed systems, peer-to-peer architectures, p2p, overlay network,
+topology awareness
+\end{IEEEkeywords}
 
-\terms{Topology mismatch}
-
-\keywords{Distributed systems, peer-to-peer architectures, p2p, overlay network,
-topology awareness}
-
-\begin{document}
-
-\begin{bottomstuff}
-Authors' addresses: Deprt. of Informatics and Telecommunications,
-University of Athens, Athens, Greece. Deprt. of Software
-Engineering, Izmir University of Economics, Izmir, Turkey.\\
-emails: \{vsmoustakas,mema,ad\}@di.uoa.gr, huseyin.akcan@ieu.edu.tr
-\end{bottomstuff}
-
-\maketitle
+% For peer review papers, you can put extra information on the cover
+% page as needed:
+% \ifCLASSOPTIONpeerreview
+% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
+% \fi
+%
+% For peerreview papers, this IEEEtran command inserts a page break and
+% creates the second title. It will be ignored for other modes.
+\IEEEpeerreviewmaketitle
 
@@
 \input{structured}
+
+%\section{Introduction}
+% The very first letter is a 2 line initial drop letter followed
+% by the rest of the first word in caps.
+% 
+% form to use if the first word consists of a single letter:
+% \IEEEPARstart{A}{demo} file is ....
+% 
+% form to use if you need the single drop letter followed by
+% normal text (unknown if ever used by IEEE):
+% \IEEEPARstart{A}{}demo file is ....
+% 
+% Some journals put the first two words in caps:
+% \IEEEPARstart{T}{his demo} file is ....
+% 
+% Here we have the typical use of a "T" for an initial drop letter
+% and "HIS" in caps to complete the first word.
+%\IEEEPARstart{T}{his} demo file is intended to serve as a ``starter file''
+%for IEEE journal papers produced under \LaTeX\ using
+%IEEEtran.cls version 1.8a and later.
+%% You must have at least 2 lines in the paragraph with the drop letter
+%% (should never be an issue)
+%I wish you the best of success.
+%
+%\hfill mds
+% 
+%\hfill September 17, 2014
+%
+%\subsection{Subsection Heading Here}
+%Subsection text here.
+%
+%% needed in second column of first page if using \IEEEpubid
+%%\IEEEpubidadjcol
+%
+%\subsubsection{Subsubsection Heading Here}
+%Subsubsection text here.
+
+% An example of a floating figure using the graphicx package.
+% Note that \label must occur AFTER (or within) \caption.
+% For figures, \caption should occur after the \includegraphics.
+% Note that IEEEtran v1.7 and later has special internal code that
+% is designed to preserve the operation of \label within \caption
+% even when the captionsoff option is in effect. However, because
+% of issues like this, it may be the safest practice to put all your
+% \label just after \caption rather than within \caption{}.
+%
+% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
+% option should be used if it is desired that the figures are to be
+% displayed while in draft mode.
+%
+%\begin{figure}[!t]
+%\centering
+%\includegraphics[width=2.5in]{myfigure}
+% where an .eps filename suffix will be assumed under latex, 
+% and a .pdf suffix will be assumed for pdflatex; or what has been declared
+% via \DeclareGraphicsExtensions.
+%\caption{Simulation results for the network.}
+%\label{fig_sim}
+%\end{figure}
+
+% Note that IEEE typically puts floats only at the top, even when this
+% results in a large percentage of a column being occupied by floats.
+
+% An example of a double column floating figure using two subfigures.
+% (The subfig.sty package must be loaded for this to work.)
+% The subfigure \label commands are set within each subfloat command,
+% and the \label for the overall figure must come after \caption.
+% \hfil is used as a separator to get equal spacing.
+% Watch out that the combined width of all the subfigures on a 
+% line do not exceed the text width or a line break will occur.
+%
+%\begin{figure*}[!t]
+%\centering
+%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
+%\label{fig_first_case}}
+%\hfil
+%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
+%\label{fig_second_case}}
+%\caption{Simulation results for the network.}
+%\label{fig_sim}
+%\end{figure*}
+%
+% Note that often IEEE papers with subfigures do not employ subfigure
+% captions (using the optional argument to \subfloat[]), but instead will
+% reference/describe all of them (a), (b), etc., within the main caption.
+% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
+% labels, the optional argument to \subfloat must be present. If a
+% subcaption is not desired, just leave its contents blank,
+% e.g., \subfloat[].
+
+% An example of a floating table. Note that, for IEEE style tables, the
+% \caption command should come BEFORE the table and, given that table
+% captions serve much like titles, are usually capitalized except for words
+% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
+% and up, which are usually not capitalized unless they are the first or
+% last word of the caption. Table text will default to \footnotesize as
+% IEEE normally uses this smaller font for tables.
+% The \label must come after \caption as always.
+%
+%\begin{table}[!t]
+%% increase table row spacing, adjust to taste
+%\renewcommand{\arraystretch}{1.3}
+% if using array.sty, it might be a good idea to tweak the value of
+% \extrarowheight as needed to properly center the text within the cells
+%\caption{An Example of a Table}
+%\label{table_example}
+%\centering
+%% Some packages, such as MDW tools, offer better commands for making tables
+%% than the plain LaTeX2e tabular which is used here.
+%\begin{tabular}{|c||c|}
+%\hline
+%One & Two\\
+%\hline
+%Three & Four\\
+%\hline
+%\end{tabular}
+%\end{table}
+
+% Note that the IEEE does not put floats in the very first column
+% - or typically anywhere on the first page for that matter. Also,
+% in-text middle ("here") positioning is typically not used, but it
+% is allowed and encouraged for Computer Society conferences (but
+% not Computer Society journals). Most IEEE journals/conferences use
+% top floats exclusively. 
+% Note that, LaTeX2e, unlike IEEE journals/conferences, places
+% footnotes above bottom floats. This can be corrected via the
+% \fnbelowfloat command of the stfloats package.
+
+%\section{Conclusion}
+%The conclusion goes here.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliographystyle{acmtrans}
+% if have a single appendix:
+%\appendix[Proof of the Zonklar Equations]
+% or
+%\appendix  % for no appendix heading
+% do not use \section anymore after \appendix, only \section*
+% is possibly needed
+
+% use appendices with more than one appendix
+% then use \section to start each appendix
+% you must declare a \section before using any
+% \subsection or using \label (\appendices by itself
+% starts a section numbered zero.)
+%
+
+%\appendices
+%\section{Proof of the First Zonklar Equation}
+%Appendix one text goes here.
+%
+%% you can choose not to have a title for an appendix
+%% if you want by leaving the argument blank
+%\section{}
+%Appendix two text goes here.
+
+% use section* for acknowledgment
+%\section*{Acknowledgment}
+
+%The authors would like to thank...
+
+% Can use something like this to put references on a page
+% by themselves when using endfloat and the captionsoff option.
+\ifCLASSOPTIONcaptionsoff
+  \newpage
+\fi
+
+% trigger a \newpage just before the given reference
+% number - used to balance the columns on the last page
+% adjust value as needed - may need to be readjusted if
+% the document is modified later
+%\IEEEtriggeratref{8}
+% The "triggered" command can be changed if desired:
+%\IEEEtriggercmd{\enlargethispage{-5in}}
+
+% references section
+
+% can use a bibliography generated by BibTeX as a .bbl file
+% BibTeX documentation can be easily obtained at:
+% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
+% The IEEEtran BibTeX style support page is at:
+% http://www.michaelshell.org/tex/ieeetran/bibtex/
+\bibliographystyle{IEEEtran}
+% argument is your BibTeX string definitions and bibliography database(s)
 \bibliography{mard-survey}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% <OR> manually copy in the resultant .bbl file
+% set second argument of \begin to the number of references
+% (used to reserve space for the reference number labels box)
+%\begin{thebibliography}{1}
+%
+%\bibitem{IEEEhowto:kopka}
+%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
+%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
+%
+%\end{thebibliography}
 
-\begin{received}
-Received Month Year;
-revised Month Year; accepted Month Year
-\end{received}
+% biography section
+% 
+% If you have an EPS/PDF photo (graphicx package needed) extra braces are
+% needed around the contents of the optional argument to biography to prevent
+% the LaTeX parser from getting confused when it sees the complicated
+% \includegraphics command within an optional argument. (You could create
+% your own custom macro containing the \includegraphics command to make things
+% simpler here.)
+%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
+% or if you just want to reserve a space for a photo:
 
+\begin{IEEEbiography}{Vassilis Moustakas}
+Biography text here.
+\end{IEEEbiography}
+
+% insert where needed to balance the two columns on the last page with
+% biographies
+%\newpage
+
+\begin{IEEEbiography}{H\"useyin Akcan}
+Biography text here.
+\end{IEEEbiography}
+
+\begin{IEEEbiography}{Mema Roussopoulos}
+Biography text here.
+\end{IEEEbiography}
+
+\begin{IEEEbiography}{Alex Delis}
+Biography text here.
+\end{IEEEbiography}
+
+% if you will not have a photo at all:
+%\begin{IEEEbiographynophoto}{John Doe}
+%Biography text here.
+%\end{IEEEbiographynophoto}
+
+% You can push biographies down or up by placing
+% a \vfill before or after them. The appropriate
+% use of \vfill depends on what kind of text is
+% on the last page and whether or not the columns
+% are being equalized.
+
+%\vfill
+
+% Can be used to pull up biographies so that the bottom of the last one
+% is flush with the other column.
+%\enlargethispage{-5in}
+
+% that's all folks
 \end{document}

### 2015-04-19 intro.tex
2015.04.19 (Vassilis) Adding a recent reference (from 2014).

@@
 {\sl Skype} has become extremely popular with more than $650$ million users 
-in $2011$~\cite{skypetotalusers} while on January $21$st, $2013$,
-the astonishing figure of $50$ million concurrently-online users
-was reported~\cite{skypesymusers}.
+in $2011$~\cite{skypetotalusers}, an astonishing $50$+ million concurrently-online
+users since 2013~\cite{skypesymusers} and an ever-growing Skype-to-Skype traffic
+recently estimated to 214 billion minutes~\cite{skypetraffic}.
+
 %%

### 2015-05-11 intro.tex
2015.04.24 (Vassilis) Added one paper from 2014. Refreshed some of the references (i.e., total skype users and p2p traffic future projection). Updated a conference paper to a more recent journal paper. Fine-tuned the typesetting, especially for tabular data.

--- 2015-04-19 intro.tex
@@
 {\sl Skype} has become extremely popular with more than $650$ million users 
-in $2011$~\cite{skypetotalusers} while on January $21$st, $2013$,
-the astonishing figure of $50$ million concurrently-online users
-was reported~\cite{skypesymusers}.
+in $2011$~\cite{skypetotalusers}, an astonishing $50$+ million concurrently-online
+users~\cite{skypesymusers} and slowly but steadily replacing traditional
+carrier traffic for services like telephony services, which was recently estimated to
+have reached 214 billion minutes for year 2013~\cite{skypetraffic}.
+
 %%
@@
 Internet traffic~\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}, with some reporting that more than $60$\% of this traffic 
-to be \p-related~\cite{cachelogic,ipoque2007,ipoque2009};
+to be \p-related~\cite{cachelogic,ipoque2009};
 it was also projected that this traffic would reach
-$8\;$\emph{Petabytes}-per-month by the end of $2012$~\cite{multinteligence}. 
+$7$~\emph{Exabytes}-per-month by $2018$~\cite{CVNI2014}! 
 Evidently, this constitutes a major burden for 

### 2015-05-11 mard-survey.tex
2015.04.24 (Vassilis) Added one paper from 2014. Refreshed some of the references (i.e., total skype users and p2p traffic future projection). Updated a conference paper to a more recent journal paper. Fine-tuned the typesetting, especially for tabular data.

@@
 
+\usepackage{adjustbox}
+
 % *** SUBFIGURE PACKAGES ***
@@
 \begin{document}
+
+\newcolumntype{R}[2]{%
+    >{\adjustbox{angle=#1,lap=\width-(#2)}\bgroup}%
+    l%
+    <{\egroup}%
+}
+\newcommand*\rot{\multicolumn{1}{R{45}{1em}}}% no optional argument here, please!
+
 %
@@
 of Informatics and Telecommunications, National and Kapodistrian University of Athens,
-GR15784, Athens, Greece (e-mail: vsmoustakas@gmail.com, mema@di.uoa.gr, ad@di.uoa.gr.}% <-this % stops a space
+GR15784, Athens, Greece (e-mails: {vsmoustakas,mema,ad}@di.uoa.gr.}% <-this % stops a space
 \thanks{H\"useyin Akcan is with the Department of Software Engineering,
@@
 \begin{IEEEkeywords}
-Distributed systems, peer-to-peer architectures, p2p, overlay network,
+Distributed systems, peer-to-peer, overlay network, topology mismatch,
 topology awareness

### 2015-05-11 structured.tex
2015.04.24 (Vassilis) Added one paper from 2014. Refreshed some of the references (i.e., total skype users and p2p traffic future projection). Updated a conference paper to a more recent journal paper. Fine-tuned the typesetting, especially for tabular data.

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Another work that targets mobile networks is \emph{3D Overlay}~\cite{AOS2014}. As the
+name suggests, the logical space of the network is a 3-dimensional space and
+the (transient) logical IDs (LIDs) are of the form of tuples of $M$-bit identifiers
+for each of the $x$, $y$ and $z$ axes. 
+When a peer wants to join the network it broadcasts a join request using
+expanded ring search. This ensures to locate peers that are physically adjacent
+to itself in the physical space. Peers in the network reply to the request
+sending their LID together with the list of their neighbours and the
+corresponding distances to them. Upon receiving the replies, the joining peer
+stores the information to its routing table thus forming an undirected connected
+graph of itself and both its $1$- and $2$-hop neighbours. Using hop distance
+obtained by the underlying (MANET) routing algorithm, builds a
+minimum spanning tree (MST) with itself as the root node and selects as direct
+neighbours the peers that are adjacent in the MST. Moreover, removes all the
+peers that were selected as directs during routing table construction but now
+found to be not adjacent in the MST. This ensures that neighboring peers on the
+overlay are also adjacent in the physical layer. Then the joining peer generates
+its LID which puts it in the 3 dimentional space. The LID computation is repeated
+as part of the update process that is required as a mechanism to maintain
+conectivity and adapt to new network states. Similarly to peer LIDs,
+keys of the files stored in the network are also generated within afformentioned
+3D space. The peer closest to the key of the file is responsible for it. Thus,
+for a peer to search for an object, it needs to generate the key for it and then
+forward the query. The neighbour to be forwarded the query, is
+selected from the routing table so that its LID and the key of the queried
+resource, brings closest the query in every part of the tuple.
+The interesting proporty of this algorithm is the construction of a multidimentional
+space that results in better load balance and finer grained information to ensure
+that each hop on the overlay during object location, is also a step closer physically
+toward the node that stores it. On the other hand, the use of TTL for its query messages
+renders the network not a pure DHT as it looses the ability that traditional
+DHTs have for zero false negative queries. Moreover, as the authors also admit, it
+does not adapt quickly to network changes and the evaluation was based on the assumption
+of low mobility scenarios. For the above reasons, 3DO exhibits the following performance
+with respect to the comparison criteria:
+
+\begin{center}
+{\footnotesize
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% 
+low &
+% 
+low &
+% 
+medium
+\end{tabular}
+}
+\end{center}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-05-11 unstructured.tex
2015.04.24 (Vassilis) Added one paper from 2014. Refreshed some of the references (i.e., total skype users and p2p traffic future projection). Updated a conference paper to a more recent journal paper. Fine-tuned the typesetting, especially for tabular data.

@@
 
-The \emph{Scalable Bipartite Overlay (SBO)}~\cite{LXN2004,LXN2007} 
+The \emph{Scalable Bipartite Overlay (SBO)}~\cite{LXN2007} 
 reduces the overhead of creating
@@
 % ==============================
-% \cite{LXN2004} Based on the above discussions, we can see the major
+% \cite{LXN2007} Based on the above discussions, we can see the major
 % advantages of using a bipartite overlay. First, the operation
@@
 The \emph{Innocuous Topology Aware Overlay Construction (ITA)}
-approach~\cite{PRFM2009} seeks to offer both an overlay formation approach and
+approach~\cite{PRFM2013} seeks to offer both an overlay formation approach and
 an effective way to carry out searching. When constructing the overlay,
@@
 $\left( 1 - \alpha \right) \ast N$ distant (latency above $x$) nodes
-as its neighbors.
+as its neighbors. Of course, the bootstrapping peer is not able check all the
+peers in the network to find the global best for the above sets. Authors
+prove that latency measurements against $30/ \alpha$ randomly selected peers
+result to a good estimation. 
 %%QUE how do they select all these magic constants???
@@
 or dynamic querying mechanisms). Experimental results in the same paper suggest
-a $50$\% reduction in communication latency among peers by cutting off some $20$\%
-of the \emph{IP}-level traffic generated.
+a $50$\% reduction in communication latency among peers by cutting off up to 
+$25$\% of the \emph{IP}-level traffic generated.
 %%
@@
 methodologies and we offer a summary qualitative comparison for all surveyed
-approaches applicable for \emph{unstructured} \p \emph{systems}.
+approaches applicable for \emph{unstructured}~\p~\emph{systems}.
 Moreover, Table~\ref{unstructured:table} offers a summary overview for each effort;

### 2015-05-17 abstract.tex
2015.05.16 (Alex & Vassilis) Extensive re-wording throughout the paper. Abstract on its one file. Better figure scaling and alignment. Re-ordering of algorithms in the histogram figures to match paper order of appearence and the table ordering.

--- 2014-01-30 mard-survey.tex
@@
-%\documentclass[a4paper,10pt]{article}
-\documentclass[acmcsur,acmnow]{acmtrans2m}
-%\documentclass[acmtocl,acmnow]{acmtrans2m}
-
-%\acmVolume{V}
-%\acmNumber{N}
-%\acmYear{YY}
-%\acmMonth{Month}
-
-% packages
-\usepackage[utf-8]{inputenc}
-\usepackage{paralist}
-\usepackage{subfigure}
-\usepackage{graphicx}
-\usepackage{colortbl}
-\usepackage{color,soul}
-\usepackage{longtable}
-\usepackage{wasysym}
-\usepackage{pdflscape}
-%\usepackage{tikz}
-\usepackage{pgfplots}
-\usetikzlibrary{patterns}
-\usepackage{url}
-
-% definitions
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{conjecture}[theorem]{Conjecture}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{lemma}[theorem]{Lemma}
-\newdef{definition}[theorem]{Definition}
-\newdef{remark}[theorem]{Remark}
-
-\newcommand{\p}{\emph{P2P}}
-
-%\newtheorem{theorem}{Theorem}[section]
-%\newtheorem{definition}[theorem]{Definition}
-
-%\textwidth      6.5in
-%\textheight     9.2in
-%\oddsidemargin  -0.5cm
-%\evensidemargin -0.5cm
-%\parindent      0pt              
-%\topmargin      -0.3in
-%\headsep        20pt
-%
-%\parskip        5pt
-%\parindent      10pt
-
-\markboth{}{}{}
-
-\title{Alleviating the Topology Mismatch Problem in Distributed Overlay Networks}
-
-\author{
-VASSILIS S. MOUSTAKAS$^1$, H\"USEYIN AKCAN$^2$, MEMA ROUSSOPOULOS$^1$ and ALEX
-DELIS$^1$\\
-$^1$\emph{University of Athens}, \emph{Athens}, \emph{Greece}\\
-$^2$\emph{Izmir University of Economics}, \emph{Izmir}, \emph{Turkey}
-}
+% As a general rule, do not put math, special symbols or citations
+% in the abstract or keywords.
 
@@
 widely deployed on the Internet for well over a decade.
-They are implemented via an overlay network abstraction, on top 
-of the Internet's best-effort IP infrastructure. 
+They are often implemented via an overlay network abstraction
+atop the Internet's best-effort IP infrastructure. 
 \p\ systems support a plethora of desirable features to
-distributed applications that are built on top of them including
-anonymity, high availability and robustness, 
-load balancing, quality of service, scalability, to name just a few.
-Unfortunately, inherent weaknesses of early deployments of \p\ systems, prevented 
+distributed applications including
+anonymity, high availability, robustness, 
+load balancing, quality of service and scalability to name just a few.
+Unfortunately, inherent weaknesses 
+of early deployments of \p\ systems, prevented 
 applications from leveraging the full potential of the paradigm.
-One major weakness, identified early on, is the \emph{topology mismatch problem}
+One major weakness, identified early on, 
+is the \emph{topology mismatch problem}
 between the overlay network and the underlying IP topology.
 This mismatch can impose an extraordinary amount of unnecessary stress on
-network resources and can adversely affect the scalability and efficiency of
-applications built atop.
+network resources and can adversely affect 
+both the scalability and efficiency of the operating applications.
 In this paper, we survey over a decade's worth of research efforts
-aimed at alleviating the topology mismatch problem, in both unstructured
-and structured \p\ systems.   
+aimed at alleviating the topology mismatch problem in both 
+\emph{structured} and \emph{unstructured} \p\ systems.   
 We provide a fine-grained categorization of the suggested solutions
@@
 Finally, we offer an analysis as well as pictorial comparisons 
-of the reviewed approaches, since we aim to offer a comprehensive
+of the reviewed approaches since we aim to offer a comprehensive
 reference for developers, system architects and researchers in the field.
-
 \end{abstract}
 
-\category{A.1}{General Literature}{Introductory and Survey}
-\category{C.2.4}{Distributed Systems}{Distributed Applications}
-
-\terms{Topology mismatch}
-
-\keywords{Distributed systems, peer-to-peer architectures, p2p, overlay network,
-topology awareness}
-
-\begin{document}
-
-\begin{bottomstuff}
-Authors' addresses: Deprt. of Informatics and Telecommunications,
-University of Athens, Athens, Greece. Deprt. of Software
-Engineering, Izmir University of Economics, Izmir, Turkey.\\
-emails: \{vsmoustakas,mema,ad\}@di.uoa.gr, huseyin.akcan@ieu.edu.tr
-\end{bottomstuff}
-
-\maketitle
-
-\input{intro}
-\input{background}
-
-\input{unstructured}
-
-\input{structured}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% TODO: SOME INFORMATION ON ISP-ASSISTED TOPOLOGY AWARENESS
-%
-% IN CASE REVIEWERS POINT OUT THAT WE SHOULD ADD THOSE AS WELL.
-%
-%       ATTENTION: NEEDS REWORDING
-%
-% P4P [16],instead, aimed to provide a general framework for P2P
-%applications by developing ISP-application interfaces so that
-%application software can use the network layer information,
-%such as congestion status and network topology, for better
-%performance and resource utilization.
-
-%P4P [37] project attempts to address
-%the problem through custom trackers, both for ISPs and P2P
-%systems, using an interface based on a primal-dual decomposition
-%of an optimization problem. This interface design simplifies
-%the realization of traffic-engineering objectives from each parties’
-%perspective and ensures the extensibility of the approach. Through
-%simulation-based studies and limited experimental deployments,
-%these collaborative approaches have been shown to effectively
-%reduce network costs while minimally impacting application performance.
-%A clear advantage of these proposals is that they
-%allow ISPs to incorporate aggregated traffic policies in their tracker
-%recommendations (e.g., a particular traffic balance ratio between
-%peering providers). However, all of them require deployments
-%of oracles for each participating ISP and their effectiveness is
-%ultimately predicated on their adoption by P2P applications and a
-%trust relationship between P2P users and their ISPs.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\input{conclusion}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliographystyle{acmtrans}
-\bibliography{mard-survey}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{received}
-Received Month Year;
-revised Month Year; accepted Month Year
-\end{received}
-
-\end{document}
+% Note that keywords are not normally used for peerreview papers.
+\begin{IEEEkeywords}
+Distributed systems, peer-to-peer, overlay network, topology mismatch,
+topology awareness
+\end{IEEEkeywords}

### 2015-05-17 background.tex
2015.05.16 (Alex & Vassilis) Extensive re-wording throughout the paper. Abstract on its one file. Better figure scaling and alignment. Re-ordering of algorithms in the histogram figures to match paper order of appearence and the table ordering.

@@
 free-riding~\cite{saroiu_measurefileshare_2002,adar_gnutellafreeriders_2000,hughes_gnutellafreeride_2005},
-the distribution of illegal content and other~\emph{socio-technical}
-issues \cite{hughes_socp2p_2008}, have arisen, \p\ systems have
+distribution of illegal content and other~\emph{socio-technical}
+issues \cite{hughes_socp2p_2008} have arisen, \p\ systems have
 continued to gain popularity throughout the years primarily 
-because they have successfully supported popular applications 
-(i.e. file-sharing) amongst \emph{vast} numbers of participating users.
+due to their successful file-sharing applications
+amongst \emph{vast} numbers of participating users.
 %%
@@
 popular protocols, networks, and applications. As \p\ systems evolved, three
-main architectures have emerged, namely
+main architectures have emerged, namely:
 \begin{inparaenum}[\itshape i\upshape)]
@@
 The downside of this approach was that it required a 
-centralized search directory that invevitably becames a single point of failure,
-a scalability bottleneck or a target of malicious behavior such as \emph{DoS}
-attacks.
+centralized search directory that inevitably became 
+a single point of failure, a scalability bottleneck or 
+a target of malicious behavior such as \emph{DoS} attacks.
 
@@
 overhead at handling the dynamics of peer participation
-\cite{stutzbach_churn_2006}. Unstructured systems are also known as 
-\emph{broadcast-based} systems as they use message flooding among
+\cite{stutzbach_churn_2006}. 
+Unstructured systems are also known as 
+\emph{broadcast-based} systems for they use message flooding among %% it is FOR!
 peers to propagate search queries. 
 In its first realization,
-{\sl Gnutella}(\emph{v.0.4}) is a well-known example of the fully
+{\sl Gnutella}(\emph{v.0.4}) became a well-known example of the fully
 decentralized unstructured approach.
@@
 entity (dotted lines) and then establish the connection to the data source.] {
-  \includegraphics[scale=0.20]{img/pdf/centralized.pdf}
+  \includegraphics[scale=0.17]{img/pdf/centralized.pdf}
   \label{figure:p2p-archs:centralized}
@@
 the topology is also random and all participants are equal.] {
-  \includegraphics[scale=0.20]{img/pdf/unstructured-full.pdf}
+  \includegraphics[scale=0.17]{img/pdf/unstructured-full.pdf}
   \label{figure:p2p-archs:unstructred-full}
@@
 indexing).] {
-  \includegraphics[scale=0.20]{img/pdf/unstructured-hierarchical.pdf}
+  \includegraphics[scale=0.16]{img/pdf/unstructured-hierarchical.pdf}
   \label{figure:p2p-archs:unstructured-hierarchical}
-}\qquad\qquad
-\subfigure[Structured ({\sl Chord} and {\sl Kademlia}). 
+}\qquad%\qquad
+\subfigure[Structured ({\sl Chord}~\cite{stoica_chord_2001} and {\sl Kademlia}~\cite{maymounkov_kademlia_2002}). 
 Peers and data are placed in the overlay in a deterministic manner.] {
-  \includegraphics[scale=0.20]{img/pdf/structured.pdf}
+  \includegraphics[scale=0.17]{img/pdf/structured.pdf}
   \label{figure:p2p-archs:structured}
@@
 \centering
-  \includegraphics[scale=0.4]{img/pdf/example-physical.pdf}
+  \includegraphics[scale=0.36]{img/pdf/example-physical.pdf}
 \caption{Interconnection of nodes in the physical level.}
@@
 \subfigure[Nodes $B$ and $C$ have direct overlay connection.] {
-  \includegraphics[scale=0.4]{img/pdf/example-overlay-inefficient.pdf}
+  \includegraphics[scale=0.28]{img/pdf/example-overlay-inefficient.pdf}
   \label{figure:overlay-1}
@@
 \subfigure[Nodes $A$ and $D$ have direct overlay connection.] {
-  \includegraphics[scale=0.4]{img/pdf/example-overlay-efficient.pdf}
+  \includegraphics[scale=0.28]{img/pdf/example-overlay-efficient.pdf}
   \label{figure:overlay-2}
@@
 
-Early overlay protocol implementations did not pay to much attention to map
+Early overlay protocol implementations did not pay much attention to map
 optimally to the underlying network topology. {\sl Gnutella}, for example, is
-considered far from scalable~\cite{ritter_gnucantscale_2001} since every peer
+considered far from scalable~\cite{ritter_gnucantscale_2001} for every peer
 chooses its neighbors without any knowledge of the underlying
-network, resulting in mismatch. 
-Queries may be flooded over multiple paths merging in one peer 
-while one of the paths would suffice.  
-Moreover, peers may forward the same message to each other
-before they receive the query messages from each other.
+network, resulting in mismatches. 
+Queries are flooded over multiple paths in an effort 
+to reach the node(s) having the sought content. 
+In addition, 
+pairs of nodes may simultaneously forward the same query to each other
+in a superflous mode of operation in 
+this opportunistic and without coordination environment.
+%%
 \cite{matei_mapgnutella_2002} showed that, even when $95$\% of any two nodes 
@@
 a \emph{flooding-based} {\sl Gnutella}-like algorithm 
-can generate $330$TB/month in a $50,000$ node network. 
+can generate $330$TB/month traffic in a $50,000$ node network. 
 A topology mismatch can thus have a serious effect 
-on Internet's router performance.
+on Internet's routing performance.
 
@@
 \cite{cox_vivaldi_2004,WSS2005} and positioning \cite{ng_gnp_2001}
-systems and makes the construction of \emph{IP}-topology-aware 
+systems and make the construction of \emph{IP}-topology-aware 
 overlays difficult.
@@
 three-level, gradation to denote how well the algorithms do,
-namely \emph{low}, \emph{medium} and \emph{high}.
+namely, \emph{low}, \emph{medium} and \emph{high}.
 Each criterion has its own
@@
 of the algorithm on intrinsic characteristics of the system (i.e., search scope
-coverage, download performance e.t.c).
-%% VM Reworked above definition for "efficiency".
+coverage, download performance etc.).
 

### 2015-05-17 conclusion.tex
2015.05.16 (Alex & Vassilis) Extensive re-wording throughout the paper. Abstract on its one file. Better figure scaling and alignment. Re-ordering of algorithms in the histogram figures to match paper order of appearence and the table ordering.

@@
 
-Research and development in application systems that follow the 
+Research and development in applications and systems that follow the 
 \p\ architectural paradigm have flourished for more than a decade now.
-By and large, such systems exploit the loose-coupling and self-organization 
+By and large, such artifacts exploit the loose-coupling and self-organization 
 of participating nodes to shape substrates upon which diverse applications 
@@
 environments.
-%%AD let's avoids "giants" etc...

### 2015-05-17 intro.tex
2015.05.16 (Alex & Vassilis) Extensive re-wording throughout the paper. Abstract on its one file. Better figure scaling and alignment. Re-ordering of algorithms in the histogram figures to match paper order of appearence and the table ordering.

@@
 {\sl Skype} has become extremely popular with more than $650$ million users 
-in $2011$~\cite{skypetotalusers}, an astonishing $50$+ million concurrently-online
-users~\cite{skypesymusers} and slowly but steadily replacing traditional
-carrier traffic for services like telephony services, which was recently estimated to
-have reached 214 billion minutes for year 2013~\cite{skypetraffic}.
+in $2011$~\cite{skypetotalusers} and an astonishing 
+$50$+ million concurrently-online users~\cite{skypesymusers}.
+Moreover, it has consistently eroded  the traffic handled by 
+traditional telephony carriers by slicing away a staggering 
+$214$ billion minutes of pertinent traffic in $2013$ alone~\cite{skypetraffic}.
 
@@
 \p\ systems are implemented using \emph{overlay networks}. 
-An overlay network, is a virtual system of nodes and logical links
+An overlay network, is a virtual system of nodes featuring logical interconnects (or links)
 created above an existing network; overlays provide an abstraction that
@@
 \centering
-  \includegraphics[scale=0.53]{img/pdf/under-over-lay.pdf}
+  \includegraphics[scale=0.45]{img/pdf/under-over-lay.pdf}
 \caption{An example overlay network.}
@@
 the inherent \emph{topology mismatch} gives two
-significant problems. First, the performance of the application per se, can be
+major problems. 
+First, the performance of the application per se, can be
 adversely affected since traffic must flow over a larger,
@@
 Second, other applications running
-on the underlying network infrastructure are also adversely affected.
+on the underlying network infrastructure may be adversely affected as well.
 Studies have shown that highly popular \p\ applications contribute 
@@
 The main objective of this paper is to offer a comprehensive 
-survey of the work done in this and provide a
+survey of the work carried out in the area
+and provide a
 taxonomy of the proposed solutions. We point out synergies, as well as

### 2015-05-17 mard-survey.tex
2015.05.16 (Alex & Vassilis) Extensive re-wording throughout the paper. Abstract on its one file. Better figure scaling and alignment. Re-ordering of algorithms in the histogram figures to match paper order of appearence and the table ordering.

@@
 
+\usepackage{comment}
+
 % correct bad hyphenation here
@@
         and~Alex Delis,~\IEEEmembership{Member,~IEEE}% <-this % stops a space
-\thanks{Vassilis Moustakas, Mema Roussopoulos and Alex Delis are with the Department
-of Informatics and Telecommunications, National and Kapodistrian University of Athens,
-GR15784, Athens, Greece (e-mails: {vsmoustakas,mema,ad}@di.uoa.gr.}% <-this % stops a space
+\thanks{Vassilis Moustakas, Mema Roussopoulos and Alex Delis are 
+	with the Department of Informatics and Telecommunications, 
+	\emph{University of Athens},
+GR15784, Athens, Greece, e-mails: \emph{\{vsmoustakas,mema,ad\}@di.uoa.gr}}% <-this % stops a space
 \thanks{H\"useyin Akcan is with the Department of Software Engineering,
-Izmir University of Economics, Izmir, Turkey (email: huseyin.akcan@ieu.edu.tr).}% <-this % stops a space
-\thanks{Manuscript received June 30, 2015; revised July 31, 2015.}}
+\emph{Izmir University of Economics}, 35330 Balcova, Izmir, Turkey, email: \emph{huseyin.akcan@ieu.edu.tr}}% <-this % stops a space
+\thanks{Manuscript received April 30, 2015; revised April 30, 2015.}}
 
@@
 
-% As a general rule, do not put math, special symbols or citations
-% in the abstract or keywords.
-\begin{abstract}
-\emph{Peer-to-peer} (\p) systems have enjoyed immense attention and have been
-widely deployed on the Internet for well ov, er a decade.
-They are implemented via an overlay network abstraction, on top 
-of the Internet's best-effort IP infrastructure. 
-\p\ systems support a plethora of desirable features to
-distributed applications that are built on top of them including
-anonymity, high availability and robustness, 
-load balancing, quality of service, scalability, to name just a few.
-Unfortunately, inherent weaknesses of early deployments of \p\ systems, prevented 
-applications from leveraging the full potential of the paradigm.
-One major weakness, identified early on, is the \emph{topology mismatch problem}
-between the overlay network and the underlying IP topology.
-This mismatch can impose an extraordinary amount of unnecessary stress on
-network resources and can adversely affect the scalability and efficiency of
-applications built atop.
-In this paper, we survey over a decade's worth of research efforts
-aimed at alleviating the topology mismatch problem, in both unstructured
-and structured \p\ systems.   
-We provide a fine-grained categorization of the suggested solutions
-by discussing their novelty, advantages and weaknesses.
-Finally, we offer an analysis as well as pictorial comparisons 
-of the reviewed approaches, since we aim to offer a comprehensive
-reference for developers, system architects and researchers in the field.
-\end{abstract}
-
-% Note that keywords are not normally used for peerreview papers.
-\begin{IEEEkeywords}
-Distributed systems, peer-to-peer, overlay network, topology mismatch,
-topology awareness
-\end{IEEEkeywords}
+\input{abstract}
 

### 2015-05-17 structured.tex
2015.05.16 (Alex & Vassilis) Extensive re-wording throughout the paper. Abstract on its one file. Better figure scaling and alignment. Re-ordering of algorithms in the histogram figures to match paper order of appearence and the table ordering.

--- 2014-11-30 structured.tex
@@
 levels of proximity information is gathered as well as  
-the way peer routing tables are maintained and used in an effort make hops in
-overlays as efficient as possible. 
+the way peer routing tables are maintained and used in an effort to make 
+overlays as efficient as possible in terms of hops traversed.
 %%AD what does it really mean this below???
@@
 %%   coined in those papers
-The categorization is based on previous work found in the
-literature~\cite{CDHR2002,CDCR2002,RSS2002}.
+Our categirization is based on previous work 
+found in~\cite{CDHR2002,CDCR2002,RSS2002}.
 
@@
 limited knowledge about neighboring zones.
-
+%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
-In terms of our $3$ criteria, the approach fares as follows:
+In terms of our $3$ stated criteria, the approach fares as follows:
 \begin{center}
@@
 
-During bootstrap, an incoming peer, say J, starts with a set of
-candidates to establish links with and iteratively
-probes for minimizing distances, based on a given metric (i.e., network delay)
-and concluding when no further refinement can be achieved. To tackle the local
+During bootstrap, an incoming peer $J$ obtains from a bootstrap node a set of
+candidate peers to establish links with. 
+Subsequently, the node works iteratively to identify peers that minimize
+a given metric (i.e., network delay).
+The process stops when no additional peers complying with the above metric 
+can be found. 
+% 
+% and iteratively
+% probes for minimizing distances, based on a given metric (i.e., network delay)
+% and concluding when no further refinement can be achieved. 
+To tackle the local
 minima problem that such a gradient descent method suffers from, Mithos probes
-all neighbours within a two-hop radius, from the currrent node that minimize
-the distance, before concluding the process.
-
-Next, the incomming peer is assigned a synthetic \emph{ID} that
-represents the state of minimum distance from its close--by neighbourhood as
-this was identified during the iterative neighbor selection phase described
-above. The benefit of these synthetic coordinates is that distance computations
+nodes within a $2$-hop radius from the current node that helps minimize
+the distance before concluding the process.
+
+$J$ is then assigned a synthetic \emph{ID} that
+represents the state of minimum distance from its close-by $2$-$d$ neighbourhood as
+this was discovered during the aforementioned phase. % iterative selection phase.
+The benefit of adopting synthetic coordinates is that distance computations
 among nodes can be done in the \emph{ID} space without requiring physical
-measurements. The actual link establishment with the peers, that will be J's
-neighbours, is happening next. With peer J dividing the plane into four regions,
-the \emph{quadrant-based} mechanism Mithos employs, suggests that the node
-establishes a link to the closest neighbor in each quadrant of the plane. The
-previous steps
-do not provide any knowledge as to which nodes these would be, thus J has to
-identify them. It does so, by starting from a known node in the neighbourhood
+measurements. 
+Mithos proceeds then to designate $J$'s closest neighbours as 
+previous steps do not reveal with certainty which these nodes are.
+%%
+Using Mithos \emph{quadrant-based} mechanism,
+$J$  divides the $2$--$d$ space of neighbours in four regions and 
+subsequently, establishes a link 
+to the closest neighbor in each quadrant of the plane. 
+% With peer $J$ dividing the plane into four regions,
+% the \emph{quadrant-based} mechanism Mithos employs, suggests that the node
+% establishes a link to the closest neighbor in each quadrant of the plane. 
+%%
+It does so, by starting from a known node in the neighbourhood
 and scanning towards all quadrant borders ~\cite{KK2000,R2002}.
 
-One major limitation of \emph{Mithos} is that the protocol cannot effectively
-handle dynamic arrivals and departures from the network. 
-\cite{CCRK2004}~argues about extensive probing to determine 
+A noted Mithos limitation is that the protocol cannot effectively
+handle dynamic arrivals and departures from the network~\cite{RGJZ2004}. 
+\cite{CCRK2004}~points out extensive probing for Mithos to determine 
 distances and less accurate proximity results compared to other
-virtual coordinate systems. The latter has been also attributed
-by~\cite{WSS2005} to the iterative neighbor selection process which that is 
+virtual coordinate systems. The last issue  has been also attributed
+by~\cite{WSS2005} to the iterative neighbor selection process that is 
 reportedly prone to premature termination at local minima.
@@
 synthetic ID generation require a centralized implementation 
-affecting \emph{Mithos}' scalability.
+affecting Mithos' scalability.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 $O(d*log_d(N))$ hops in a balanced overlay tree with $N$ representing the
-number of nodes and $d$ the maximum degree of each node. 
+number of nodes and $d$ the maximum degree of each node (i.e., number of links). 
 \emph{LAPTOP} adopts the geographical layout approach  
@@
 %%
-Each node is assigned a dot-formatted address (e.g., 1.3.4). 
-Each octet ranges from
-$1$ to $d$.
+Each node is assigned a dot-formatted address (e.g., 1.3.4) with
+each octet ranging from $1$ to $d$.
 The assignment process is done by appending a unique octet 
@@
 is reached and then, the message starts descending towards its target. 
-During tree traversals, special entries in the routing tables,
+During tree traversals, special entries in the routing tables
 are maintained to increase routing efficiency and achieve finer--load balance.
@@
 to forward a message to a better 
-longest-prefix match than that of its direct ancestor yielding
-a larger, quicker and more cost--effective step through the overlay towards
+longest-prefix match than that of its direct ancestor;
+this yields a quicker and more cost--effective step through the overlay towards
 the destination. 
@@
 \cite{KLKP2008} introduces proximity
-controls over the \emph{vanilla Kademlia} 
+controls over the \emph{vanilla Kademlia}~\cite{maymounkov_kademlia_2002} 
 protocol's random \emph{ID} assignment.
@@
 % provided by the underlying tier. 
-Such cost information can be acquired by
+Such cost information can be acquired by:
 \begin{inparaenum}[\itshape i\upshape)]
   \item measuring previous lookups,
-  \item exchanging or jointly-computing measurements with 
-	others, or 
+  \item exchanging or jointly-computing measurements with others, or 
   \item using a local database to look--up information.
@@
 
-\cite{KLKP2008} exploits $2$ routing techniques: 
+\cite{KLKP2008} exploits two routing techniques: 
 \emph{proximity routing selection (PRS)} and 
@@
 In \emph{PRS}, the goal is to choose the best next hop while routing a message.
-\emph{Plain old Kademlia} (\emph{POK}) always selects the closest--node 
-with respect to the \emph{XOR} metric.
+\emph{Kademlia} always selects the closest--node 
+with respect to the \emph{XOR} metric~\cite{maymounkov_kademlia_2002}.
 With \emph{PRS} in place however, \emph{Kademlia} may also choose 
@@
 To enhance traffic locality, the use of 
-{\sl MaxMind GeoIP} database~\footnote{\url{http://www.maxmind.com}} 
+{\sl MaxMind GeoIP} database\footnote{\url{http://www.maxmind.com}} 
 is suggested to form clusters of nearby peers.
@@
 to go through the backbone.
-The behavior of the proposal with respect to the stated criteria is deemed as follows:
+The behavior of \cite{KLKP2008} with respect to the stated criteria is deemed as follows:
 %%
@@
 
-\emph{CHOP6} \cite{MT2007} is another Chord-variant DHT; it exploits
-the {\sl IPv6} address format, and if available \emph{RTT} information, to 
+\emph{CHOP6} \cite{MT2007} is a Chord-variant DHT that exploits
+{\sl IPv6} address format as well as available \emph{RTT} information to 
 enable proximity estimation among peers during message routing.
-
+%%
 Similar to \emph{Chord}~\cite{stoica_chord_2001}, 
-\emph{CHOP6} uses a finger table but each entry now holds hold k>1 candidate
-nodes. The next hop for a message is selected to be the one closest to the
-destination among those k candidate nodes. This proximity is determined either
-by exploiting RTT measurements or, if there is no such yet (i.e. a new peer), by
-estimating distance on overlay's ID space. When joining the network, each node
-receives a $64$-bit \emph{ID}. It comprises of a random part and a IPv6 prefix
-part. The random part is obtained by hashing the node's IP address and port
-number and is used to preserve the uniformity of
-node assignment. The IPv6 prefix is based on IPv6's global routing prefix.
-Because IANA  assigns address blocks to regions (RIRs) which are further
-distributed on a national level (NIRs), examining the upper $32$-bits
-of an IPv6 address someone can estimate the country or the region to which the
-node is connected. Thus, CHOP6's IPv6 prefix ID is used for proximity estimation
+\emph{CHOP6} uses a finger table with each entry now holding $k$$>$$1$ candidate
+nodes. 
+The next hop for a message is selected to be the one closest to the
+destination among those $k$ candidate peers. 
+This proximity is determined either
+by exploiting RTT measurements or, if there is no such information yet 
+(i.e., a new peer appears), by
+estimating distance on the overlay's ID space.
+
+When joining the network, each node
+receives a $64$-bit \emph{ID}. 
+It comprises of a random part and the IPv6 prefix which 
+is obtained by hashing the node's IP address and port
+number; this combination is deployed in an effort  
+to preserve the uniformity of node assignment. 
+The prefix in question is based on IPv6's global routing prefix.
+As IANA\footnote{\url{http://www.iana.org/}}  
+assigns address blocks to regions (RIRs) which are further
+distributed at national level (NIRs), the first $32$-bits of the address
+can estimate the region or the country the node is connected in. 
+Thus, CHOP6's IPv6 prefix ID is used for proximity estimation
 if no RTT information is available during routing. 
-
+%%
 %
@@
 \emph{T2MC}'s~\cite{SLCGZ2008} objective is to diminish the redundant multiple
-messages that may unnecessarily cross \emph{AS}-boundaries as these are the
-links that cost more to ISPs thus minimizing the effects of the topology
-mismatch problem. To achieve that, each node exploits the stable structure of
-the Internet routers to help solve a clustering problem. Uses traceroute
-logs to detect inter-\emph{AS} boundaries and clusters other peers to near-by and
-far-away peers by applying a customized $k$-mean classification
-algorithm with $k = 2$ to perform the classification.
-Specifically when joining the network, the peer randomly picks an IP and uses
-traceroute. Skipping the details of the algorithm, it picks the routers with
-minimum and maximum latencies in the above traceroute path as initial centroids
-for the ``near'' and ``remote'' sets respectively and after some iterations it
-ends up with two sets having the minimum intra-set latency variance. The
-router with the minimum hops in the ``remote'' set is set up as the hop
-threshold and anything above that is classified as the ``remote'' router cluster
-for that peer. Similarly, the router with the maximum hops in the ``near'' set
-is important because it denotes an edge gateway. Peers that share the same
-edge gateway or other members of the ``near'' router cluster are considered
-close-by.
+messages that may unnecessarily cross autonomous system (\emph{AS}) boundaries.
+The latter represent links that are the costliest to ISPs and hence, 
+mininizing topology mismatches  greatly affects pertinent operational expenditure.
+%%
+To achieve that, each node exploits the stable structure of
+the Internet routers to help solve a clustering problem. 
+The algorithm uses {\sl traceroute}-logs to detect inter-\emph{AS} boundaries 
+and clusters peers to either nearby or 
+far-away groups by applying a customized $k$-mean classification ($k=2$).
+%%
+When joining the network, the peer randomly picks an IP and uses
+traceroute and selects the routers with
+minimum and maximum latencies in the above {\sl traceroute} path 
+as initial centroids for the ``nearby'' and ``far-away'' sets respectively.
+Following an iterative procedure, the node comes up with the two sets 
+demostrating minimum intra-set latency variance.
+At this stage, routers in the ``far-away'' set are not considered critical.
+%%
+%%AD skipping the details???? JUST CHECK
+% Skipping the details of the algorithm, it picks the routers with
+% minimum and maximum latencies in the above traceroute path as initial centroids
+% for the ``near'' and ``remote'' sets respectively and after some iterations it
+% ends up with two sets having the minimum intra-set latency variance. 
+%%
+% The router from the ``far-away'' set that presents the closest distance 
+% from the peer in terms of hops, is designated as beacon.
+% Any node portaying dinstance longer that the above is classified 
+% as 
+% 
+% to the peer, in terms of hops, is designated as a beacon.
+% Any node whose distance is measured to be longer that the distance 
+% between 
+% 
+% The router with the minimum hops in the ``remote'' set is set up as the hop
+% threshold and anything above that is classified as the ``remote'' router cluster
+% for that peer. 
+% Similarly, 
+In contrast, the router from the ``near'' set
+having the maximum number of hops 
+from the just-joining peer is key as it denotes an edge--gateway. 
+Peers that share the same edge--gateway or other members in the ``nearby'' 
+router cluster are considered intra-\emph{AS} neighbors.
+%%AD the final phrase for this is missing! So what??? What is achieved? 1 sentence.
 
@@
 provides detailed information about the network structure, 
-it extensive use may create overhead on the overall network structure. 
-In this respect, it is often the case that administrators 
-disable {\sl traceroute} support on their routers.
-The latter may affect \emph{T2MC}'s performance.
+its extensive use may create overhead on the overall network structure. 
+In this respect, administrators frequently
+disable {\sl traceroute} support on their routers this 
+undoubtly affects \emph{T2MC}'s operation.
 %%
@@
 \emph{T2MC} is considered rather coarse-grained as it only
-as it classifies neighbours into two groups \cite{QLZG2009}.
+classifies neighbours into two groups \cite{QLZG2009}.
 %
@@
 \begin{center}
+{\footnotesize
 \begin{tabular}{ccc}
@@
 \emph{PChord} \cite{HLYW2005} is based on the \emph{Chord DHT} and adds
-proximity into its routing mechanism. To achieve that, the standard finger table
-is augmented with a \emph{proximity list} which is a structure that contains
-nodes in the network that are close-by, in terms of RTT, with the node at hand.
+proximity into its routing mechanism. 
+To achieve that, the standard finger table
+is augmented with a \emph{proximity list} that contains
+nodes in the network found close-by in terms of RTT.
 %in this way, the next hop can be decided based on both progress
@@
 %% just described. 
-In this way, the next hop during routing can be decided based on both the
+In this respect, the next hop during routing can be decided based on both the
 direction towards the search key and minimizing the cost of that hop.
 
-At join time, this list in empty, but during the node's lifetime it is created
-and modified in a heuristic fashion. During routing, the next hop is chosen
-as the closest node to the target key from both the finger and the proximity
-lists. This means that the next hop defined in the finger list can only be
-substituted by a better hop (closest to the key and in less physical distance)
-with PChord resulting to equal or even more efficient routing than the pure
-vanilla Chord.
+When a peer joins the network, its proximity list is empty.
+This list is maintained in a heuristic fashion during the node's lifetime.
+While a node attempts routing, the next hop is selected
+as the closest node to the target key 
+from \emph{both} its finger table and proximity list. 
+An initial choice for the next hop made out of the finger list 
+can be substituted by a shorter one from the proximity list if  
+it is both closer to the key and in less physical distance.
+This yields an equal or even more efficient routing than traditional Chord.
 %% VM
@@
 
-Unfortunatelly the overlay maintenance cost is a little bit more than
-\emph{Chord} additional heart-beat messages are required for
-connection verification of the proximity list. Fortunately, these nodes
-are already close-by so the additional cost is kept to a minimum but
-it has been reported to exhibit slow convergence speed 
+Unfortunatelly the overlay maintenance cost is somewhat higher than
+\emph{Chord} as additional heart-beat messages are required for
+connection verification for the entries of the proximity list. 
+Fortunately, such entries represent nodes 
+already close-by so the additional cost is kept to a minimum.
+However, \emph{PChord} has been reported to exhibit slow convergence speed 
 and inefficiency in high-churn environments~\cite{DK2006}.
-
+%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % \subsubsection{AChord}
-\emph{AChord} \cite{DK2006} uses the {\sl IPv6} any-cast mechanism to
+\emph{AChord} \cite{DK2006} exploits IPv6's \emph{anycast} mechanism to
 relieve the protocol from complex joining procedures,
 harness network proximity and achieve high routing efficiency. 
-Any-cast delivers a message coming from the outside of 
-an any-cast group to the physically closest node in
-that group~\cite{M2002}. Thus, \emph{AChord} organizes all nodes participating
-in the overlay network into an any-cast group. Any joining node comes (from
-outside) and is automatically forwarded to the physically nearest node for
-bootstrapping; this avoids the need to explicitly maintain participants
+When sending a message to an IPv6 anycast group, 
+the originator accomplishes this delivery by interacting with
+the physically closest node in that group~\cite{M2002}.
+Thus, \emph{AChord} organizes all nodes participating
+in the overlay network into a single anycast group. 
+An arriving peer is automatically forwarded to the physically 
+nearest node for
+bootstrapping; this curtails the need to explicitly maintain participants
 and the effort of locating the physically nearest. 
@@
 
-The \emph{ID} of an incoming 
-node is computed based on the \emph{ID} of the bootstrap node 
-and the bootstrap's predecessor in a way that the constructed 
-\emph{ID} is positioned between the aforementioned two. 
+The \emph{ID} of an incoming node is computed based on 
+both the identifier of the bootstrap node 
+and the bootstrap's predecessor;
+this is  done in a way that the \emph{ID} of the 
+incoming peer is positioned \emph{between} the two identifiers mentioned above. 
 After joining, a \emph{finger} table is created ala \emph{Chord}. 
 Moreover, an \emph{adjacency table} maintains 
-information about the closest--known peers.
-
-The routing decision is made by using both the adjacency and finger tables
-based on the closest node to the target key in the key space (same as
-Chord).
-%%AD how are these 2 combined???
-% VM:
-% Check above.
-% This AChord algo looks pretty similar to PChord.
-% If you agree I would suggest we merge the two.
-
+information about the $k$ closest--known peers. 
+%%
+The routing decision takes place in a way reminiscent to \emph{Chord}
+and uses both adjacency and finger tables.
+%%
 \emph{AChord} behavior has as follows:
@@
 \emph{Chord6}~\cite{XZHL2005} is another \emph{Chord}-variant that
-exploits {\sl IPv6}'s  hierarchical features in order to create 
+uses IPv6's  hierarchical features to create 
 a substrate that reduces inter-domain traffic among service providers. 
@@
 \emph{Chord6} and \emph{Chord} lies on the identifier definition. 
-Therefore, the approach can be easily portable to other 
+The adopted approach is deemed as easily portable to other 
 \emph{DHT}s including \emph{CAN}, \emph{Pastry} and \emph{Tapestry}. 
+%%
 \emph{Chord6}'s identifier contains two parts: 
 the higher bits are obtained by hashing the
-node's {\sl IPv6} address prefix of specific length, 
-while the remaining lower bits are the hash-value 
-of the rest of the {\sl IPv6} address. 
-As a result of this assignment, nodes in a domain are mapped 
+node's IPv6 address prefix (of specific length), 
+while the remaining lower bits of the identifier 
+is the hash-value of the remaining of the IPv6 address. 
+%%
+As a result, nodes in a domain are mapped 
 onto a continuous key space on the overlay network;
@@
 
-\emph{Chord6} does reduce the average system path length to be
-logarithmically proportional to the number of available domains 
-instead of the number of participating peers. 
-However, due to the large number of Internet
+\emph{Chord6} does reduce the average 
+length of hops traversed
+% system path length 
+to be logarithmically proportional to the number 
+of available domains instead of the number of participating peers. 
+Due however to the large number of Internet
 domains and the possibility of occasionally having only 
-small number of peers within each domain,
+a small number of peers within each domain,
 \cite{DK2006} challenges the approach's scalability and efficiency.
@@
 the \emph{Chord DHT} is introduced in~\cite{DLTZZ2006}.
-Its key objective here has been to group physically-close nodes as 
+Its key objective was to group physically-close nodes as 
 neighbors in the \emph{DHT} table while exploiting proximity information.
@@
 \cite{DLTZZ2006} prove that each sector
-occupies the same area as does the center cycle. 
-Assuming uniform node distribution, this characteristic,
-favors a more load--balanced clustering operation. 
+This characteristic favors a more load--balanced clustering operation
+should we assume uniform node distribution.
 Every sector in this $2d$ space is mapped to 
@@
 multi-layer node identifier space.  
-Consequently, nodes belonging to the same
-sector, are mapped to the same region as well, preserving their proximity
-relationship. Individual elements are uniquely mapped to the identifier--space,
+Consequently, nodes from the same
+sector are mapped to the same region as well, 
+preserving their proximity relationship.
+
+Individual elements are uniquely mapped to the identifier--space,
 allowing logarithmic lookup operations with high 
-probability on \emph{Chord}. Also that way a node can obtain its \emph{DHT}
-key and its region key, in a fully distributed manner, just by applying a
-consistent hash function. The node can then communicate with the region's
-master node called \emph{Cluster Node (CN)} which is responsible for clustering 
-the nodes belonging in the same sector or region with that of the node at hand. 
-It registers to its corresponding region and publishes its information, through
-the special peer \emph{CN} and can also ask the latter for other
-nodes that have previously joined the region to add them into its
-neighbor set for future routing table optimization.
-%
-If no other nodes are on that region, the search is expanded to neighbouring
-regions.
+probability on \emph{Chord}. 
+To this end, 
+a node can obtain its \emph{DHT} key and its region key, 
+in a fully distributed manner, just by applying a consistent hash function. 
+The node can then communicate with the region's
+master node called \emph{Cluster Node (CN)} which 
+is responsible for clustering 
+the nodes belonging in the same sector or 
+region with that of the node at hand. 
+%%
+The node registers to its corresponding region  and publishes
+its information, through the special peer \emph{CN};
+the peer can also ask \emph{CN} for other
+nodes that have previously joined the region to enhance 
+the peer's neighbor set for future routing table optimization.
+%
+If no other nodes are found within the region, 
+the search is expanded to neighbouring regions.
 %
@@
 %%
-Unfortunately, even though scaling appeared promising, only limited 
+Even though scaling appeared promising, only limited 
 simulations were carried out with configurations involving up to 2,048 nodes.
@@
 %%
-\emph{Chord DHT} fares as follows in reference to the $3$ criteria:
+\emph{Chord DHT} fares as follows in reference to the stated $3$ criteria:
 \begin{center}
@@
 % \subsubsection{IP-based Clustering (IPBC)}
-Proximity neighbor selection algorithms often use probing to detect proximity. 
+\emph{Proximity neighbor selection} algorithms often 
+use probing to detect proximity. 
 However, such methods suffer by imprecisions often due to network overload.
-{\sl IP}-based clustering (IPBC)~\cite{KM2007} uses {\sl IP} address prefixes
-($16$ bit for {\sl IPv4}) to detect proximity.
-Such proximity information (IP addresses of network participants) can be
-easily stored in the DHT itself.
-
-Each peer in the overlay has a unique ID and and an IP address to
-characterize it. To publish itself it hashes a fixed-length prefix
-its IP to obtain a key and stores its ID and IP in the DHT itself. 
+IP-based clustering or \emph{IPBC}~\cite{KM2007} uses IP address prefixes
+($16$~bits for IPv4) to detect proximity.
+Such proximity information 
+% (IP addresses of network participants) 
+can be easily stored in the DHT itself.
+
+Each peer in the overlay has a unique \emph{ID} and and an IP address to
+characterize it. 
+To publish itself, the peer hashes a fixed-length prefix of 
+its IP to obtain a key and stores the resulting \emph{ID} and IP 
+in the DHT. 
 Newly joining nodes with the same {\sl IP} prefix 
-can, thus, query the \emph{DHT} and readily identify all neighbors with the
-same prefix. Nodes periodically update their entries in the \emph{DHT} 
+can, thus, query the \emph{DHT} and readily identify all 
+neighbors with the same prefix. 
+Nodes periodically update their entries in the \emph{DHT} 
 by removing those who leave or have timed-out.
+%%
+
 The overhead of the algorithm is very low as proximity detection requires only
@@
 %
-The suggested approach fares as follows:
+The suggested \emph{IPBC} approach fares as follows:
 \begin{center}
@@
 %%
-Although based on the \emph{Pastry} overlay and 
+Although based on \emph{Pastry} and 
 its well understood built-in locality heuristics~\cite{CDCR2002a},
-\emph{DynaMO}'s own mechanisms are \emph{DHT}-independent and 
-so, they can be applied to other \emph{DHT}s.
+\emph{DynaMO}'s own mechanisms are {DHT}-independent and 
+so, they can be applied to other overlays.
 \emph{DynaMO}'s approach may be viewed as bottom-up.
+
 An incoming node computes physical information about its neighborhood
 and uses the outcome to unilaterally assign itself an \emph{ID}.
-The proposal is based the fact that with every new overlay routing-hop
+The approach suggests that for every new overlay routing-hop 
 the physical distance between nodes is likely to increase
@@
 last routing step~\cite{antony_pastry_2001,CDCR2002a}.
-To identify this step, two techniques are used: 
+To identify this dominating step, two techniques are used: 
 \emph{Random LandMarking (RLM)} and 
@@
 for a set of landmark keys. These keys have to be chosen in a way that they
-equally divide the overlay space (i.e. in a 16 base an appropriate split
-would be 000..0, 100..0, 200..00, ... , F00..00).
-An new-coming peer then computes distances from those landmarks,
+equally divide the overlay space (i.e., in a $16$--base an appropriate split
+would be $000..0$, $100..0$, $200..00$, ... , $F00..00$).
+%
+
+A new-coming peer then computes distances from those landmarks,
 sorts them, and assigns itself an \emph{ID} on this 
 landmark ordering as follows:
-the peer adopts as \emph{ID} having a prefix from its closest
-landmark; the length is 
-$l_{prefix}=|log_b k|$, where $b$ is the \emph{ID} base 
+the peer adopts sets the prefix of its \emph{ID} to be equal to that 
+of its closest landmark.
+The length of this prefix is 
+$l_{prefix}=|log_b k|$, where $b$ is the \emph{ID}--base 
 and $k$ the number of landmarks. 
+%%
 The remaining of bits of the peer-\emph{ID} are assigned either 
@@
 bootstrap phase is required.
-It delegates to \emph{Pastry} the task to identify the physically
-close--node with which an incoming peer gets bootstrapped. 
-The new-comer assumes the bootstrapping nodes's \emph{ID} prefix 
-and fills the remaining of its key using the techniques of \emph{RLM}.
-
+It delegates the task of identifying the physically
+close--node to \emph{Pastry}; this node is used by the 
+incoming peer to get bootstrapped.
+The new-comer takes the bootstrapping nodes's \emph{ID} prefix 
+and fills the remaining bits of its key using the \emph{RLM} technique.
+%%
 \emph{RLM} gets evaluated as presenting good proximity in a load-balanced 
@@
 Unfortunately, \emph{RLM} bootstrap overhead is substantial and can 
-effect the set of landmarks.
+effect the set of landmarks~\cite{WZS2004}.
 \emph{CNPA} on the other hand, incurs less overhead but this come 
@@
 medium \\
+\end{tabular}
+}
+\end{center}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\emph{3D Overlay (3DO)}~\cite{AOS2014} targets mobile networks.
+As the name suggests, the logical space of the network is 
+a $3$--$d$ space and the used ``transient'' (continually changing) 
+logical \emph{ID}s (\emph{LID}s) consist of $3$ axes, $x$, $y$ and $z$ 
+in the form of a tuple;
+each such axis value is represented by an $M$-bit identifier.
+%%
+A newlly-arriving peer broadcasts a join request using
+an TTL-based expanding ring search. 
+This helps locate peers that are physically adjacent.
+Responding peers send in 
+their \emph{LID}s along with a list of their neighbors 
+and their corresponding distances from those neightbors. 
+%%
+Once the joining peer receives the above responses,
+it stores requisite information in its own routing table 
+and so, it generates an undirected connected
+graph of itself with both its $1$- and $2$-hop neighbours. 
+%%
+Starting off this graph and with the help of hop 
+distances obtained from the underlying MANET,
+the incoming peer builds a minimum spanning tree (MST)
+with itself as the root node; 
+in this tree, the nodes adjacent to the root become the
+direct neighbors of the peer.
+%%
+In addition, the entries of the initial routing table 
+that do not correspond to above direct neighbors are
+eliminated.
+
+The procedure above ascertains that close peers on the
+overlay are also adjacent at the physical layer. 
+%%
+At this stage, a peer can generate its \emph{LID}
+that automatically places it in the $3$-$d$ network space.
+Updates in the peer's connectivity infromation 
+reflecting continual changes in the mobile network,
+trigger updates on the composition of \emph{LID} itself.
+
+%%AD work on what the following says.. 
+%%VM I wrote the following paragraph as 
+\emph{3DO} stores objects in the same $3$-$d$ space. For each
+file $f$, a key $k$ of the form of an $x$, $y$ and $z$ tuple
+is hashed out of the file's unique ID~(e.g., filename).
+The node whose \emph{LID} is closer to $k$ stores the
+file itself.
+%%
+To retrieve the file, the requesting node calculates $f$'s $k$
+and issues a query message. The message is a associated with a TTL
+value. Each peer forwards that query to the node,
+in its routing table, that is closest to $k$ up until the peer
+that stores the file is reached or the query's lifetime expires.
+
+\emph{3DO} focuses on
+the creation of a $3$-$d$ space that yields improved load balancing 
+and uses fine--grained information to ensure
+that each hop on the overlay during object location, 
+is also closer physically to the sought object.
+%%
+In contrast, the use of TTL in query messages
+renders the network not--a--pure DHT as it abolishes 
+the zero false--negative queries that traditional DHTs maintain.
+\cite{AOS2014} also concedes that the approach
+does not rapidly adapt to network changes and so
+its evaluation was solely based on low mobility scenarios. 
+For the above reasons, \emph{3DO} exhibits the following behavior:
+%%
+\begin{center}
+{\footnotesize
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% 
+low &
+% 
+low &
+% 
+medium
 \end{tabular}
@@
 to change its neighborhood in order to reduce average neighbor distances. 
-This process is performed unilaterally
-until no additional optimization is needed but it achieves optimization within a
+This process is performed iteratively
+until no additional optimization is required; 
+yet, the methods achieves optimization within a
 sufficiently large scope.
@@
 
-The \emph{probing phase} starts when the new-comer joins the \emph{DHT} 
-using the standard mechanism of the algorithm at hand. 
-The peer starts probing the neighborhood with 
-messages of small \emph{TTL=k}~\cite{jiang_lightflood_2008}. 
+% The \emph{probing phase} starts when the new-comer joins the \emph{DHT} 
+% using the standard mechanism of the algorithm at hand. 
+When a peer joins the \emph{DHT}, it starts 
+its \emph{probing phase} by contacting the neighborhood with 
+messages having small \emph{TTL=k}~\cite{jiang_lightflood_2008}. 
 Recipients return information to the source and forward the message 
@@
 During the \emph{jumping phase}, the source node calculates the 
-stretch change to all its own direct neighbours and to the direct neighbours
+stretch change to all its direct neighbours and to the direct neighbours
 of the first peer that was selected in the probing phase, as if
-the jump has been performed. Using heuristics the node decides if the stretch
-reduction is enough to perform the actual jump, otherwise it selects the
-second peer from the probing phase and repeats the computation. If again, the
+the jump has been performed. 
+Using heuristics, the node determines if the stretch
+reduction is sufficient to perform the actual jump; 
+otherwise, it selects the second peer from the probing phase 
+and repeats the computation. If again, the
 threshold is not met, then no jump is ultimately carried out.
-%%AD is this done only 2 times?? Or is it done recursively??
-%%VM no it stops
-%% Quoting from the paper
-% Our experiments
-% show that targeting the two zones associated with the two
-% closest nodes is most effective. This is because adaptive
-% and iterative local optimizations will quickly or eventually
-% group the physically close nodes in the same overlay zone.
-% If a node outside a zone finds a close node in the zone, the
-% rest of the nodes will be likely to be or eventually to be close
-% to it.
+
 Experimentation shows that \emph{SAT-Match} 
@@
 a landmark-enhanced \emph{CAN}-alternative.
-Based on this result, it is conjured that \emph{SAT-Match} 
+This result shows that \emph{SAT-Match} 
 may perform better than landmark binning in terms of the matching degree
@@
 
-This section outlined
-% gathered together  some of the most distinctive 
-representative research that attempts 
-to address the issue of inefficient network resource utilization 
-as this is created by the network topology mismatch
-in structured \p-systems.
-Depending on the salient features the surveyed techniques seek to explore,
-\cite{CDHR2002,CDCR2002,RSS2002} categorize the solutions based on:
-%%AD the categorization is due to 3 above citations??? Us?? ;-)
-%% VM: As mentioned in the intro this categorization is due to the 3 above
-% citations
-%
-% A categorization of the methods used for achieving their goals is the
-% following based on previous work from
-% \cite{CDHR2002,CDCR2002,RSS2002}:
+So far in this section, we have discussed research efforts 
+trying to address the issue of inefficient network resource utilization 
+manifested by network--topology mismatches in structured \p-systems.
+Based on the salient feature of outlined techniques
+\cite{CDHR2002,CDCR2002,RSS2002}, we classify furnished solutions 
+using the following aspects:
 \begin{enumerate}[\itshape i\upshape)]
@@
 
-%%COM-VM Added the following to mention the hybrid/combo methodologies
-Similarly to unstructured approaches, most of the approaches 
+%%VM Added the following to mention the hybrid/combo methodologies
+Similarly to unstructured approaches, most of the methods
 reviewed in this section entail features from multiple of the above categories. 
 For example, \cite{KLKP2008}~uses both proximity routing 
-and proximity neighbor selection to achieve topology adaptation 
-or \emph{LAPTOP} combines geographic--layout with
+and proximity neighbor selection to achieve topology adaptation,
+while \emph{LAPTOP} combines geographic--layout with
 replication strategies to refine its results.
@@
 The heuristic nature of the suggested solutions opens them up 
-to trade-offs which in most cases are 
+to trade-offs which, in most cases, are 
 highly affected by the application domain and the implementation itself.
@@
 in Section~\ref{section:unstructured}.
-In a similar fashion to Table~\ref{unstructured:table}, we offer 
-key properties of every approach 
-that differentiate it from competitors and
-highlight each algorithm's pros and cons.
+In a similar fashion to Table~\ref{unstructured:table}, 
+we outline key properties of each surveyed approach 
+that helps diffferentiate techniques from competitors.
+We also highlight pros and cons for all proposals discussed.
 
@@
 Routing tables of involved peers, are adjusted and maintained by exploiting
-proximity information that describe the overall (geographic) positioning of
-the participants, in contrast to proximity routing and proximity neighbour
-selection where decisions are taken based on and effecting a peer's nearby
-environment.
+proximity information that describes the overall (geographic) positioning of
+the participants. This comes in contrast to proximity routing and proximity neighbour
+selection (discussed later on) where decisions are taken based on and effecting
+a peer's nearby environment.
 
@@
 \centering
-  \includegraphics[scale=0.4]{img/pdf/geographic-layout.pdf}
+  \includegraphics[scale=0.36]{img/pdf/geographic-layout.pdf}
 \caption{Successive nodes in the overlay circle can be chosen to be from the
@@
 Landmark servers are not self-organizing and present maintenance
-overheads. Moreover, to serve a \p--network with millions of peers, multiple
+overhead. Moreover, to serve a \p--network with potentially millions of peers, multiple
 such servers are required to be distributed over the world, which is a difficult
@@
 \emph{RTT}s can measure the delay between peers, but their use constitutes a
-greedy method that can result in non-optimal overlay topologies;
+greedy method that can result in sub-optimal overlay topologies;
 the latter is especially true if nearby 
@@
 efficiency of the system in general and are especially effective when applied
-to CAN(-like) implementations. Unfortunatelly, in the same time, undermine the
-uniform distribution of peer identifiers which in turn tend to create 
-hot-spots and limited node resilience to failures. Such resilience appears to
-suffer as nearby nodes are more likely to suffer collective
-failures~\cite{HY2007}.
+to CAN-like implementations. On the other hand, they inherently undermine the
+uniform distribution of peer identifiers. This is an important downside as it
+creates hot-spots that hurt system scalability. Moreover, as nearby nodes
+are more likely to suffer collective failures~\cite{HY2007}, it adversely affects
+system availability.
 %% VM reworded the above paragraph, please check it out.
@@
 This category of approaches use physical proximity knowledge to decide the best
-next hop during routing. As shown in Figure~\ref{figure:proximity-routing}
+next hop during routing. As shown in Figure~\ref{figure:proximity-routing},
 maintenance of routing tables themselves is a separate process based only on the
-system's DHT construction mechanism and overlay id distribution.
+system's DHT construction mechanism and overlay ID distribution.
 
@@
 %%
-When \emph{PR} based algorithms route messages, they try to balance between
+When \emph{PR}--based algorithms route messages, they try to balance between
 choosing the node that will further progress the routing towards the
@@
 %% VM Is the above now better with the rewording of the section's opening paragraph?
-In the degenerated case that the system ends up using a greedy approach that
-selects only the lowest latency paths on the overlay, can, ultimately, result in
-more hops to reach the destination thus yielding suboptimal, longer paths on
-the physical topology as well. PR is also less less effective than geographical
-layout when applied to CAN(-like) implementations.
+In the degenerate case, a system could use a greedy approach to select
+only the lowest latency candidate peers during forwarding. That would to
+longer, sub--optimal, paths to the destinations; thus imposing
+extra stress to network resources. PR is also less
+effective than geographical layout when applied to CAN--like implementations.
 %%AD re-work carefully this segment...
@@
 Algorithms in this category use proximity data during routing table construction
-and maintenance as shown in Figure~\ref{figure:proximity-neighbour-selection}.
+and maintenance along the lines of Figure~\ref{figure:proximity-neighbour-selection}.
 %%
@@
 %%
-The proximity information used here is different from that of 
+The proximity information used is different from that of 
 landmark-based systems described in Section~\ref{section:geographic_layout};
-here, \emph{TTL} values between nodes (i.e., \emph{SAT-Match}) or 
+here, \emph{RTT} values between nodes (i.e., \emph{SAT-Match}) or 
 \emph{IP} address prefixes (i.e., \emph{DynaMO}) % or \emph{Cone}) %%VM removing Cone from paper
@@
 %%
-Moreover, \cite{HLYW2005} argue that such approaches can
+Moreover, \cite{HLYW2005} argue that prefix--based approaches can
 destroy the uniform distribution in the key space and do not work in
@@
 operate at the overlay.
+
 Therefore, a careful selection is required in terms of
@@
 \emph{Tapestry}, \emph{Pastry}, and \emph{CAN} successfully incorporated
-proximity neighbor selection into their algorithms. 
+proximity neighbor selection into their methods. 
 The routing protocol in \emph{Pastry} is based on longest node 
@@
 In general, proximity neighbor selection is considered 
-as a superior method to proximity routing, however, joint uses of the 
-approaches are possible.
+superior to proximity routing, however, simultaneous use of both
+are also possible.
 

### 2015-05-17 unstructured.tex
2015.05.16 (Alex & Vassilis) Extensive re-wording throughout the paper. Abstract on its one file. Better figure scaling and alignment. Re-ordering of algorithms in the histogram figures to match paper order of appearence and the table ordering.

@@
 \subfigure[Inefficient overlay topology averaging $\simeq 16$ delay units.] {
-  \includegraphics[scale=0.4]{img/pdf/topology-adaptation-before.pdf}
+  \includegraphics[scale=0.36]{img/pdf/topology-adaptation-before.pdf}
   \label{figure:topology-adaptation:before}
@@
 \subfigure[Efficient overlay topology after adaptation averaging $\simeq 11$ delay units.] {
-  \includegraphics[scale=0.4]{img/pdf/topology-adaptation-after.pdf}
+  \includegraphics[scale=0.36]{img/pdf/topology-adaptation-after.pdf}
   \label{figure:topology-adaptation:after}
@@
 \subfigure[Typical blind flooding.] {
-\includegraphics[scale=0.20]{img/pdf/forwarding-optimization-before.pdf}
+\includegraphics[scale=0.17]{img/pdf/forwarding-optimization-before.pdf}
   \label{figure:forwarding-optimisation:before}
@@
 \subfigure[A node can decide where to forward the messages.] {
-  \includegraphics[scale=0.20]{img/pdf/forwarding-optimization-after.pdf}
+  \includegraphics[scale=0.19]{img/pdf/forwarding-optimization-after.pdf}
   \label{figure:forwarding-optimisation:after}
@@
 \subfigure[Indexing can reduce the cost of the last hop.] {
-\includegraphics[scale=0.3]{img/pdf/replication-index.pdf}
+\includegraphics[scale=0.25]{img/pdf/replication-index.pdf}
   \label{figure:replication:index}
@@
 \subfigure[Data replication along the path of a successful query.] {
-  \includegraphics[scale=0.3]{img/pdf/replication-data.pdf}
+  \includegraphics[scale=0.25]{img/pdf/replication-data.pdf}
   \label{figure:replication:data}
@@
 \centering
-  \includegraphics[scale=0.4]{img/pdf/landmarking.pdf}
+  \includegraphics[scale=0.35]{img/pdf/landmarking.pdf}
 \caption{Landmark binning during node bootstrap.}
@@
 \end{figure}
+
+\input{unstructured-table}
+\input{unstructured-plot}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input{unstructured-table}
-
-\input{unstructured-plot}

### 2015-05-24 structured.tex
2015.05.17 (Alex) Minor re-wording for the text on the 3DO algorithm. Added connection sentence for discussion on structured approaches between introductory paragraph and main body of the section. Some additional re-wording and figure re-scaling in the (structured) discussion text.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\emph{3D Overlay (3DO)}~\cite{AOS2014} targets mobile networks.
-As the name suggests, the logical space of the network is 
+\emph{3D Overlay (3DO)}~\cite{AOS2014} seeks to address the 
+network mismatch problem  for peers operating in mobile networks.
+As its name suggests, the network logical space is 
 a $3$--$d$ space and the used ``transient'' (continually changing) 
@@
 A newlly-arriving peer broadcasts a join request using
-an TTL-based expanding ring search. 
-This helps locate peers that are physically adjacent.
+an \emph{TTL}-based expanding ring search helping locate 
+peers that are physically adjacent.
 Responding peers send in 
@@
 %%
-Once the joining peer receives the above responses,
+Once a joining peer receives the above responses,
 it stores requisite information in its own routing table 
@@
 with itself as the root node; 
-in this tree, the nodes adjacent to the root become the
+in this MST, the nodes adjacent to the root become the
 direct neighbors of the peer.
 %%
-In addition, the entries of the initial routing table 
+Moreover, the entries of the initial routing table 
 that do not correspond to above direct neighbors are
@@
 
-The procedure above ascertains that close peers on the
+The above procedure ascertains that close peers on the
 overlay are also adjacent at the physical layer. 
@@
 
-%%AD work on what the following says.. 
-%%VM I wrote the following paragraph as 
 \emph{3DO} stores objects in the same $3$-$d$ space. For each
 file $f$, a key $k$ of the form of an $x$, $y$ and $z$ tuple
-is hashed out of the file's unique ID~(e.g., filename).
+is hashed out of the file's unique \emph{ID}~(e.g., filename).
 The node whose \emph{LID} is closer to $k$ stores the
@@
 To retrieve the file, the requesting node calculates $f$'s $k$
-and issues a query message. The message is a associated with a TTL
-value. Each peer forwards that query to the node,
-in its routing table, that is closest to $k$ up until the peer
-that stores the file is reached or the query's lifetime expires.
+and issues a query message. 
+The message is a associated with a \emph{TTL} value. 
+Each peer forwards that query to the node,
+in its routing table, that is closest to $k$ until either the peer
+that stores the file is located or the query's lifetime expires.
 
 \emph{3DO} focuses on
-the creation of a $3$-$d$ space that yields improved load balancing 
-and uses fine--grained information to ensure
+the creation of a $3$-$d$ space whose stated objective is to
+yield improved load balancing 
+and exploit fine--grained information to ensure
 that each hop on the overlay during object location, 
@@
 %%
-In contrast, the use of TTL in query messages
+However, the use of \emph{TTL} in query messages
 renders the network not--a--pure DHT as it abolishes 
@@
 its evaluation was solely based on low mobility scenarios. 
-For the above reasons, \emph{3DO} exhibits the following behavior:
+For the above reasons, \emph{3DO} exhibits the following behavior 
+in conjunction with our three stated criteria:
 %%
@@
 latency and $\bar{L_p}$ is the average physical link latency.
-%%AD what are the indexes: l and p -- p for peer and l for new peer???
-%% VM: no l for logical and p for physical
-%%AD rephrased what is below as above... 
-% 
-% The paper defines \emph{stretch}
-% $S = \bar{L_l}/\bar{L_p}$, where $\bar{L_l}$ is the average logical link
-% latency while the $\bar{L_p}$ is the average physical link latency, and uses it
-% a way to quantify the \emph{topology match degree} of the constructed overlay.
-
+% 
 % The \emph{probing phase} starts when the new-comer joins the \emph{DHT} 
@@
 
-So far in this section, we have discussed research efforts 
+In this section, we have discussed research efforts 
 trying to address the issue of inefficient network resource utilization 
-manifested by network--topology mismatches in structured \p-systems.
+manifested by network--topology mismatches in \emph{structured} \p-systems.
 Based on the salient feature of outlined techniques
 \cite{CDHR2002,CDCR2002,RSS2002}, we classify furnished solutions 
-using the following aspects:
+based on the following aspects:
 \begin{enumerate}[\itshape i\upshape)]
@@
 \end{enumerate}
-
-%%VM Added the following to mention the hybrid/combo methodologies
+%%
+%%COM Added the following to mention the hybrid/combo methodologies
 Similarly to unstructured approaches, most of the methods
@@
 
-Table~\ref{structured:table} offers a quick overview of how the surveyed 
-approaches behave.
+Table~\ref{structured:table} offers a comparative view 
+on how our surveyed approaches behave.
 The topics of topology adaptation, landmarking and caching/replication
@@
 We also highlight pros and cons for all proposals discussed.
+
+%%AD Connection???  to the groups???
+In what follows, we place the surveyed structured \p2p-methods in 
+the three categories mentioned above and which are formulated based 
+on utilized proximity and geographic aspects.
 
@@
 \centering
-  \includegraphics[scale=0.36]{img/pdf/geographic-layout.pdf}
+  \includegraphics[scale=0.44]{img/pdf/geographic-layout.pdf}
 \caption{Successive nodes in the overlay circle can be chosen to be from the
@@
 Landmark servers are not self-organizing and present maintenance
-overhead. Moreover, to serve a \p--network with potentially millions of peers, multiple
-such servers are required to be distributed over the world, which is a difficult
-proposition per se.
+overhead. Moreover, to serve a structured \p--network with potentially 
+tens of thousands or more 
+% millions 
+peers, multiple such servers are required to be distributed over the world, 
+which is a difficult proposition per se.
 \emph{RTT}s can measure the delay between peers, but their use constitutes a
@@
 efficiency of the system in general and are especially effective when applied
-to CAN-like implementations. On the other hand, they inherently undermine the
+to \emph{CAN}-like implementations. On the other hand, they inherently undermine the
 uniform distribution of peer identifiers. This is an important downside as it
-creates hot-spots that hurt system scalability. Moreover, as nearby nodes
-are more likely to suffer collective failures~\cite{HY2007}, it adversely affects
-system availability.
-%% VM reworded the above paragraph, please check it out.
+creates hot-spots that hurt system scalability. 
+Moreover, it adversely affects system availability as nearby nodes
+are more likely to suffer collective failures~\cite{HY2007}.
 
@@
 \centering
-  \includegraphics[scale=0.4]{img/pdf/proximity-routing.pdf}
+  \includegraphics[scale=0.48]{img/pdf/proximity-routing.pdf}
 \caption{Proximity routing takes proximity information into account during the
@@
 %%
-\begin{figure}[ht]
+\begin{figure}[hbft]
 \centering
-  \includegraphics[scale=0.4]{img/pdf/proximity-neighbor-selection.pdf}
+  \includegraphics[scale=0.48]{img/pdf/proximity-neighbor-selection.pdf}
 \caption{Proximity neighbor selection dictates that proximity is taken into

### 2015-05-24 structured.tex
2015.05.17 (Vassilis) Enabled ancoring for the motivation subsection of the background section. Enabled ancoring for histogram figures (efficiency, overhead, scalability). Elaborated on connection sentences for both (unstructured and structured) discussion sections. Some minor re-wording and image re-scaling for better typesetting.

@@
 A newlly-arriving peer broadcasts a join request using
-an \emph{TTL}-based expanding ring search helping locate 
+a \emph{TTL}-based expanding ring search helping locate 
 peers that are physically adjacent.
@@
 
-Table~\ref{structured:table} offers a comparative view 
-on how our surveyed approaches behave.
-The topics of topology adaptation, landmarking and caching/replication
-are not further elaborated as they have been discussed 
-in Section~\ref{section:unstructured}.
-In a similar fashion to Table~\ref{unstructured:table}, 
-we outline key properties of each surveyed approach 
-that helps diffferentiate techniques from competitors.
-We also highlight pros and cons for all proposals discussed.
-
 %%AD Connection???  to the groups???
-In what follows, we place the surveyed structured \p2p-methods in 
+%%VM I like it. I put one similar to the unstructured part
+In what follows, we place the surveyed structured \p-methods in 
 the three categories mentioned above and which are formulated based 
 on utilized proximity and geographic aspects.
+With Table~\ref{structured:table}, we offer a comparative view
+of the algorithms in a similar fashion to Table~\ref{unstructured:table}, 
+to outline the key properties of each surveyed approach 
+that helps differentiate techniques from
+competitors~\footnote{The topics of topology adaptation, landmarking and
+caching/replication are not further elaborated as they have been discussed 
+in Section~\ref{section:unstructured}.}.
+Finally, Figures~\ref{structured:plot:efficiency}, \ref{structured:plot:overhead}
+and \ref{structured:plot:scalability} help with the pictorial comparison
+of the algorithms' bahavior with respect to the three effectiveness
+criteria we defined in Subsection~\ref{background:motivation}.
 

### 2015-05-24 unstructured.tex
2015.05.17 (Vassilis) Enabled ancoring for the motivation subsection of the background section. Enabled ancoring for histogram figures (efficiency, overhead, scalability). Elaborated on connection sentences for both (unstructured and structured) discussion sections. Some minor re-wording and image re-scaling for better typesetting.

@@
 
-Below, we outline how the proposed protocols use elements of the four
-methodologies and we offer a summary qualitative comparison for all surveyed
+Below, we outline how the proposed protocols use elements of the above four
+methodologies. Then we offer a summary qualitative comparison for all surveyed
 approaches applicable for \emph{unstructured}~\p~\emph{systems}.
-Moreover, Table~\ref{unstructured:table} offers a summary overview for each effort;
-in particular, the Table shows which of the aforementioned methodologies 
-every effort incorporates, the highlights of the effort as well as 
-a rough estimation of the pros and cons of in its implementation.
+Specifically, Table~\ref{unstructured:table} summarises, for each
+effort, the methodologies it incorporates, its special highlights as well as 
+a rough estimation of the pros and cons in its implementation. Last but
+not least, with Figures~\ref{unstructured:plot:efficiency}, \ref{unstructured:plot:overhead}
+and \ref{unstructured:plot:scalability} we pictorially compare the
+algorithms in terms of \emph{efficiency}, \emph{overhead} and \emph{scalability}
+respectively; criteria we defined in Subsection~\ref{background:motivation}
+and which we used to evaluate the effectiveness of the each surveyed algorithm.
 
@@
 for this purpose.
-These landmark--servers are used by nodes to estimate 
-their positions based on the intuitive assumption that nodes
+For example, in Figure~\ref{figure:landmarking}, the newly
+arriving peer, measures its distance to an array of landmark points denoted as
+$L_n$. Then it creates an ordered list of the landmarks based on its measured
+distances uses that information to select its neighbours. Specifically,
+the landmark--servers are used by nodes to estimate 
+their positions based on the intuitive assumption that peers
 with similar distances to a set of landmarks, are physically close to each
 other, as well, over the network.  
-In Figure~\ref{figure:landmarking}, the newly
-arriving peer, measures its distance to an array of landmark point denoted as
-$L_n$ and and makes a sorted list of the peers, say in increasing order. Then in
-order to choose the already participating peers with which to connect to,
-the peer compares its list to the list of the potential neighbor. Peers with similar
-measured distances to the these landmark points are likely to be close to
-each other as well.
 %%
@@
 \centering
-  \includegraphics[scale=0.35]{img/pdf/landmarking.pdf}
+  \includegraphics[scale=0.34]{img/pdf/landmarking.pdf}
 \caption{Landmark binning during node bootstrap.}

### 2015-05-24 background.tex
2015.05.17 (Vassilis) Enabled ancoring for the motivation subsection of the background section. Enabled ancoring for histogram figures (efficiency, overhead, scalability). Elaborated on connection sentences for both (unstructured and structured) discussion sections. Some minor re-wording and image re-scaling for better typesetting.
2015.05.18 (Alex) Gone through spelling correction. Added minimal biography text for authors. Some more re-wording on the connection sentences in the discussion section. Moving table and plots to improve overall typesetting in the structured section.

@@
 ``traveled'' to locate a resource, even in the case when only a \emph{single}
-copy exists whithin the system. This requires
+copy exists within the system. This requires
 $O$($log(n)$) hops, compared to the unstructured networks that require
@@
 pairs of nodes may simultaneously forward the same query to each other
-in a superflous mode of operation in 
+in a superfluous mode of operation in 
 this opportunistic and without coordination environment.
@@
 \subsection{Motivation and Goal for this Survey}
+\label{background:motivation}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 protocols as being lowly, mediumly or highly scalable if the are proven to
-effectivelly support hundreds, tenths of thousands or millions of nodes,
+effectively support hundreds, tenths of thousands or millions of nodes,
 respectively.

### 2015-05-24 mard-survey.tex
2015.05.18 (Alex) Gone through spelling correction. Added minimal biography text for authors. Some more re-wording on the connection sentences in the discussion section. Moving table and plots to improve overall typesetting in the structured section.

@@
 \begin{IEEEbiography}{Vassilis Moustakas}
-Biography text here.
+% Biography text here.
+is a researcher with the University of Athens in Athens, Greece.
 \end{IEEEbiography}
@@
 \begin{IEEEbiography}{H\"useyin Akcan}
-Biography text here.
+% Biography text here.
+is an Associate Professor of Computer Science at Izmir University of Economics in Izmir, Turkey.
 \end{IEEEbiography}
@@
 \begin{IEEEbiography}{Mema Roussopoulos}
-Biography text here.
+% Biography text here.
+is an Associate Professor of Computer Science at the Univ. of Athens in Athens, Greece.
 \end{IEEEbiography}
@@
 \begin{IEEEbiography}{Alex Delis}
-Biography text here.
+% Biography text here.
+is a Professor of Computer Science at the Univ. of Athens in Athens, Greece.
 \end{IEEEbiography}

### 2015-05-24 structured.tex
2015.05.18 (Alex) Gone through spelling correction. Added minimal biography text for authors. Some more re-wording on the connection sentences in the discussion section. Moving table and plots to improve overall typesetting in the structured section.

--- 2015-05-24 structured.tex
@@
 %%VM it means that this split into i) geographic layout, ii) proximity routing
-%%   and iii) proximity neighbour selection types of algorithms are terms
+%%   and iii) proximity neighbor selection types of algorithms are terms
 %%   coined in those papers
-Our categirization is based on previous work 
+Our categorization is based on previous work 
 found in~\cite{CDHR2002,CDCR2002,RSS2002}.
@@
 
-\emph{Mithos}~\cite{WR2003} ensures that neighbouring nodes in the overlay are
+\emph{Mithos}~\cite{WR2003} ensures that neighboring nodes in the overlay are
 close also in terms of the underlying layer by integrating proximity layout
@@
 $J$ is then assigned a synthetic \emph{ID} that
-represents the state of minimum distance from its close-by $2$-$d$ neighbourhood as
+represents the state of minimum distance from its close-by $2$-$d$ neighborhood as
 this was discovered during the aforementioned phase. % iterative selection phase.
@@
 measurements. 
-Mithos proceeds then to designate $J$'s closest neighbours as 
+Mithos proceeds then to designate $J$'s closest neighbors as 
 previous steps do not reveal with certainty which these nodes are.
@@
 Using Mithos \emph{quadrant-based} mechanism,
-$J$  divides the $2$--$d$ space of neighbours in four regions and 
+$J$  divides the $2$--$d$ space of neighbors in four regions and 
 subsequently, establishes a link 
@@
 %%
-It does so, by starting from a known node in the neighbourhood
+It does so, by starting from a known node in the neighborhood
 and scanning towards all quadrant borders ~\cite{KK2000,R2002}.
@@
 % case when its parent has reached its maximum degree.
-% \item Each node is assigned a dot-formated address (e.g 1.3.4). Each octet
+% \item Each node is assigned a dot-formatted address (e.g 1.3.4). Each octet
 % ranges from
@@
 The latter represent links that are the costliest to ISPs and hence, 
-mininizing topology mismatches  greatly affects pertinent operational expenditure.
+minimizing topology mismatches  greatly affects pertinent operational expenditure.
 %%
@@
 Following an iterative procedure, the node comes up with the two sets 
-demostrating minimum intra-set latency variance.
+demonstrating minimum intra-set latency variance.
 At this stage, routers in the ``far-away'' set are not considered critical.
@@
 % from the peer in terms of hops, is designated as beacon.
-% Any node portaying dinstance longer that the above is classified 
+% Any node portraying distance longer that the above is classified 
 % as 
@@
 \emph{T2MC} is considered rather coarse-grained as it only
-classifies neighbours into two groups \cite{QLZG2009}.
+classifies neighbors into two groups \cite{QLZG2009}.
 %
@@
 
-Unfortunatelly the overlay maintenance cost is somewhat higher than
+Unfortunately the overlay maintenance cost is somewhat higher than
 \emph{Chord} as additional heart-beat messages are required for
@@
 %% repetition of the rest of the text below after all.
-%% Please review and remove if you aggree.
+%% Please review and remove if you agree.
 %%
@@
 If no other nodes are found within the region, 
-the search is expanded to neighbouring regions.
+the search is expanded to neighboring regions.
 %
@@
 % which the node belongs to while the \emph{lid} indicates the local identifier
-% within that group. The Cone topology can be visualised as another ring on top
+% within that group. The Cone topology can be visualized as another ring on top
 % of Chord whose nodes are the groups and each group is each such Cone-node is a
@@
 %%
-A newlly-arriving peer broadcasts a join request using
-an \emph{TTL}-based expanding ring search helping locate 
+A newly-arriving peer broadcasts a join request using
+a \emph{TTL}-based expanding ring search helping locate 
 peers that are physically adjacent.
@@
 their \emph{LID}s along with a list of their neighbors 
-and their corresponding distances from those neightbors. 
+and their corresponding distances from those neighbors. 
 %%
@@
 and so, it generates an undirected connected
-graph of itself with both its $1$- and $2$-hop neighbours. 
+graph of itself with both its $1$- and $2$-hop neighbors. 
 %%
@@
 that automatically places it in the $3$-$d$ network space.
-Updates in the peer's connectivity infromation 
+Updates in the peer's connectivity information 
 reflecting continual changes in the mobile network,
@@
 During the \emph{jumping phase}, the source node calculates the 
-stretch change to all its direct neighbours and to the direct neighbours
+stretch change to all its direct neighbors and to the direct neighbors
 of the first peer that was selected in the probing phase, as if
@@
 
+In what follows, we place the surveyed \emph{structured} \p-methods in 
+the three above categories formulated based on 
+utilized proximity and geographic aspects.
 Table~\ref{structured:table} offers a comparative view
-on how our surveyed approaches behave.
-The topics of topology adaptation, landmarking and caching/replication
-are not further elaborated as they have been discussed 
-in Section~\ref{section:unstructured}.
-In a similar fashion to Table~\ref{unstructured:table}, 
-we outline key properties of each surveyed approach 
-that helps diffferentiate techniques from competitors.
-We also highlight pros and cons for all proposals discussed.
-
-%%AD Connection???  to the groups???
-In what follows, we place the surveyed structured \p2p-methods in 
-the three categories mentioned above and which are formulated based 
-on utilized proximity and geographic aspects.
+of the algorithms in a way reminiscent to that used in Table~\ref{unstructured:table}.
+Table~\ref{structured:table} outlines key properties for all surveyed approach 
+that helps differentiate and better position competing techniques.
+Clearly, we do not further elaborate topics that we have included and discussed 
+at length in Section~\ref{section:unstructured}; these aspects entail 
+topology adaptation, landmarking, and caching/replication.
+Finally, Figures~\ref{structured:plot:efficiency}, \ref{structured:plot:overhead}
+and \ref{structured:plot:scalability} help with the pictorial comparison
+of the algorithms' bahavior with respect to the three effectiveness
+criteria we defined in subsection~\ref{background:motivation}.
 
@@
 proximity information that describes the overall (geographic) positioning of
-the participants. This comes in contrast to proximity routing and proximity neighbour
+the participants. This comes in contrast to proximity routing and proximity neighbor
 selection (discussed later on) where decisions are taken based on and effecting
@@
 order to reduce the network's average delay.
+%%
+\input{structured-table}
+%%
+\input{structured-plot}
+%%
 Landmark servers and \emph{RTT} measurements are two popular methods, which
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input{structured-table}
-
-\input{structured-plot}

### 2015-05-24 unstructured.tex
2015.05.18 (Alex) Gone through spelling correction. Added minimal biography text for authors. Some more re-wording on the connection sentences in the discussion section. Moving table and plots to improve overall typesetting in the structured section.

@@
 % distances, resorting routing tables and creating sectors. Moreover it just
-% prunes distant neighbout
+% prunes distant neighbor
 low &
@@
 % TODO: I DON'T REMEMBER WHY THE FOLLOWING IS MENTIONING LTM WHICH IS ANOTHER
-%       ALGORITM. MAYBE THIS CAN BE USED AS A PART OF DISCUSSION AND/OR
+%       ALGORITHM. MAYBE THIS CAN BE USED AS A PART OF DISCUSSION AND/OR
 %       COMPARISON OF AOTO AND LTM
@@
 in~\cite{RTLCGZ2010} where experimental results suggest that 
-most peers in a swarm reside in different authonomous systems (\emph{AS}s).
+most peers in a swarm reside in different autonomous systems (\emph{AS}s).
 Consequently, clients frequently do not have enough peers to leverage
@@
 %%
-The last technique operates at a \emph{microscopic-level} by chosing to pick
+The last technique operates at a \emph{microscopic-level} by choosing to pick
 the next chunk to download based on a locality-first policy; this picks first
@@
 
-Below, we outline how the proposed protocols use elements of the above four
-methodologies. Then we offer a summary qualitative comparison for all surveyed
+In what follows, we outline how the surveyed protocols 
+use elements of the above four methodologies. 
+We then  offer a summary qualitative comparison for all surveyed
 approaches applicable for \emph{unstructured}~\p~\emph{systems}.
-Specifically, Table~\ref{unstructured:table} summarises, for each
+Specifically, Table~\ref{unstructured:table} summarizes, for each
 effort, the methodologies it incorporates, its special highlights as well as 
-a rough estimation of the pros and cons in its implementation. Last but
-not least, with Figures~\ref{unstructured:plot:efficiency}, \ref{unstructured:plot:overhead}
-and \ref{unstructured:plot:scalability} we pictorially compare the
-algorithms in terms of \emph{efficiency}, \emph{overhead} and \emph{scalability}
-respectively; criteria we defined in Subsection~\ref{background:motivation}
-and which we used to evaluate the effectiveness of the each surveyed algorithm.
+a rough estimation of the pros and cons in its implementation. 
+Last but not least, Figures~\ref{unstructured:plot:efficiency}, \ref{unstructured:plot:overhead}
+and \ref{unstructured:plot:scalability} pictorially compare the
+surveyed techniques in terms of \emph{efficiency}, \emph{overhead} and \emph{scalability};
+we defined these three criteria in Subsection~\ref{background:motivation}
+and we collectively used them as a yardstick to qualitatively 
+evaluate the presented approaches.
 
@@
 $L_n$. Then it creates an ordered list of the landmarks based on its measured
-distances uses that information to select its neighbours. Specifically,
+distances uses that information to select its neighbors. Specifically,
 the landmark--servers are used by nodes to estimate 

### 2015-05-30 background.tex
2015.05.30 (Vassilis) Changes added based on Hüseyin's comments. Mainly minor re-wording and fixing typos.

@@
 different approaches suggested and to accomplish this, we
-present methods for both structured and unstructured distributed 
+present methods for unstructured and structured
 overlay networks.
@@
 greater compared to a protocol that only relies on peers exchanging routing
-tables within a a 2-hop radius from each other.
-%%VM Reworked definition of overhead and explanatory sentence.
-%%
+tables within a 2-hop radius from each other.
 
@@
 to the number of participating peers it can reliably support. We characterize
-protocols as being lowly, mediumly or highly scalable if the are proven to
+protocols as being lowly, mediumly or highly scalable if they are proven to
 effectively support hundreds, tenths of thousands or millions of nodes,
 respectively.
-%%%VM Reworked definition.
-%%

### 2015-05-30 intro.tex
2015.05.30 (Vassilis) Changes added based on Hüseyin's comments. Mainly minor re-wording and fixing typos.

@@
 various aspects of the topology mismatch problem.
+This research area started to become active after  The mismatch problem is Interest around All interesting work around the topology We omitted centralized Centralized systems are not
+
 The main objective of this paper is to offer a comprehensive 
@@
 evaluation of the techniques discussed in this paper. 
-Sections~\ref{section:unstructured} and~\ref{section:structured}
-respectively outline the surveyed research efforts 
-for decentralized-unstructured and decentralized-structured \p\ systems.
+Work in the area was inspired mostly due to the volatile and fully
+distributed nature of the decentrilized overlays so we skip
+centralized systems and in Sections~\ref{section:unstructured} and~\ref{section:structured}
+we outline the surveyed research efforts
+for unstructured and structured \p\ overlays, respectively.
 Conclusions can be found in Section~\ref{section:conclusion}.

### 2015-05-30 structured.tex
2015.05.30 (Vassilis) Changes added based on Hüseyin's comments. Mainly minor re-wording and fixing typos.

@@
 %
-In terms of our $3$ stated criteria, the approach fares as follows:
+In terms of our three stated criteria, the approach fares as follows:
 \begin{center}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In regard to the $3$ criteria, \emph{Mithos} stands as follows:
+In regard to the three criteria, \emph{Mithos} stands as follows:
 \begin{center}
@@
 %
-\emph{LAPTOP}'s behavior for the $3$ criteria stands as follows:
+\emph{LAPTOP}'s behavior for the three criteria stands as follows:
 \begin{center}
@@
 %
-With respect to the $3$ qualitative criteria, \emph{CHOP6} stands as follows:
+With respect to the three qualitative criteria, \emph{CHOP6} stands as follows:
 \begin{center}
@@
 messages that may unnecessarily cross autonomous system (\emph{AS}) boundaries.
-The latter represent links that are the costliest to ISPs and hence, 
+Inter-\emph{AS} links are the costliest to ISPs and hence, 
 minimizing topology mismatches greatly affects pertinent operational expenditure.
@@
 %%
-\emph{AChord} behavior has as follows:
+\emph{AChord} behaves as follows:
 \begin{center}
@@
 In this way, 
-\cite{DLTZZ2006} prove that each sector
-This characteristic favors a more load--balanced clustering operation
+\cite{DLTZZ2006} prove that for each sector
+this characteristic favors a more load--balanced clustering operation
 should we assume uniform node distribution.
@@
 %%
-\emph{Chord DHT} fares as follows in reference to the stated $3$ criteria:
+\emph{Chord DHT} fares as follows in reference to the stated three criteria:
 \begin{center}
@@
 Second, using the \emph{Cantor} space--filling curve, 
-the $2$-dimensional is converted to $1$-dimensional space. 
+the $2$-dimensional space is converted to $1$-dimensional space. 
 The latter is employed in the third and final stage to build 
@@
 %%
-\emph{Quasi-Chord} qualitative behavior has as follows:
+\emph{Quasi-Chord}'s qualitative behavior is as follows:
 \begin{center}
@@
 
-Each peer in the overlay has a unique \emph{ID} and and an IP address to
+Each peer in the overlay has a unique \emph{ID} and an IP address to
 characterize it. 
@@
 landmark ordering as follows:
-the peer adopts sets the prefix of its \emph{ID} to be equal to that 
+the peer sets the prefix of its \emph{ID} to be equal to that 
 of its closest landmark.
@@
 effect the set of landmarks~\cite{WZS2004}.
-\emph{CNPA} on the other hand, incurs less overhead but this come 
+\emph{CNPA} on the other hand, incurs less overhead but this comes 
 at the expense of being more coarse-grained as it is based on prefixes
-to determine proximity during.
+to determine proximity during bootstrap.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
-\emph{DynoMO}'s elements fare as follows regarding our $3$ criteria:
+\emph{DynoMO}'s elements fare as follows regarding our three criteria:
 \begin{center}
@@
 and issues a query message. 
-The message is a associated with a \emph{TTL} value. 
+The message is associated with a \emph{TTL} value. 
 Each peer forwards that query to the node,
@@
 of the logical to the physical networks.
-\emph{SAT-Match} behaves as follows with respect to the stated $3$ criteria:
+\emph{SAT-Match} behaves as follows with respect to the stated three criteria:
 %%
@@
 and \ref{structured:plot:scalability} help with the pictorial comparison
-of the algorithms' bahavior with respect to the three effectiveness
+of the algorithms' behavior with respect to the three effectiveness
 criteria we defined in subsection~\ref{background:motivation}.
@@
 the physical one, as depicted in Figure~\ref{figure:geographic-layout}.
-Routing tables of involved peers, are adjusted and maintained by exploiting
-proximity information that describes the overall (geographic) positioning of
-the participants. This comes in contrast to proximity routing and proximity neighbor
-selection (discussed later on) where decisions are taken based on and effecting
-a peer's nearby environment.
-
 \begin{figure}[ht]
@@
 \end{figure}
-%%
+Routing tables of involved peers, are adjusted and maintained by exploiting
+proximity information that describes the overall (geographic) positioning of
+the participants. This comes in contrast to proximity routing and proximity neighbor
+selection (discussed later on) where decisions are taken based on and effecting
+a peer's nearby environment.
+
 A number of the reviewed algorithms including 
@@
 order to reduce the network's average delay.
-%%
-\input{structured-table}
-%%
-\input{structured-plot}
-%%
 Landmark servers and \emph{RTT} measurements are two popular methods, which
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%
+\input{structured-table}
+%%
+\input{structured-plot}
+%%

### 2015-05-30 unstructured.tex
2015.05.30 (Vassilis) Changes added based on Hüseyin's comments. Mainly minor re-wording and fixing typos.

@@
 Below, we outline how the techniques in~\cite{YG-M2002}
-fare against the $3$ major performance criteria 
+fare against the three major performance criteria 
 of Section~\ref{section:background}.
@@
 the clustering of routing tables. 
-In an dynamic environment, the formation 
+In a dynamic environment, the formation 
 of very accurate such routing tables might be however an elusive goal.
@@
 %%
-In terms of the $3$ performance criteria, {\sl DAPS} stands as follows:
+In terms of the three performance criteria, {\sl DAPS} stands as follows:
 \begin{center}
@@
 that primarily focuses on the satisfaction of peers improves the network's
-scalability it falls short in addressing the mismatch problem
+scalability, it falls short in addressing the mismatch problem
 as considerations for the underlying network are not handled in explicit
@@
 %
-In terms of the $3$ stated criteria, {\sl Gia}'s approach is as follows:
+In terms of the three stated criteria, {\sl Gia}'s approach is as follows:
 \begin{center}
@@
 % search scope.
+medium
+\end{tabular}
+}
+\end{center}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \subsubsection{Location-aware Topology Matching (LTM)}
+
+\emph{Location-aware Topology Matching (LTM)} \cite{LLXNZ2004}
+seeks to optimize an overlay \p\ structure based on the physical topology.
+To achieve this, peers issue special messages called
+\textit{TTL-detector}s whose \emph{TTL} values are $0$ or $1$;
+in this regard, peers 
+discover $1$- and $2$-hop neighbor sets, designated as $N$ and $N^2$
+respectively, and proceed to compute communication costs.
+Time-stamps are used to derive network latency measurements that are then used 
+to improve the overlay network without sacrificing the search scope.
+Each node compares the latency information
+received from its direct neighbors;
+peers with longer latencies are placed on a
+\emph{will-cut} list where they remain for a certain period of time 
+after which they are finally eliminated 
+and are placed on the peer's \emph{cut-list}. 
+Thus, low-productivity connections are dropped and replaced by 
+more efficient ones, reducing the
+latency on the overall network. 
+Although \emph{LTM} improves the overall efficiency of
+the \p\ network, it is unable to offer any warranty for 
+effectively addressing the mismatch problem.
+%
+% TODO: SOME DISCUSSION
+%
+%\paragraph*{Low productive connection cutting}
+%There are three cases for any peer $P$ who receives $d(i, S, v)$ multiple
+% times:
+%\begin{inparaenum}[\itshape i\upshape)]
+%  \item $P$ receives both $d(i, S, 1)$ and $d(i, S, 0)$
+%  \item $P$ receives multiple $d(i, S, 0)$s from different paths, and P
+% randomly chooses to process one
+%  \item $P$ receives one $d(i, S, 1)$ and multiple $d(i, S, 0)$s, and $P$
+% processes $d(i, S, 1)$ and one randomly selected $d(i, S, 0)$
+%\end{inparaenum}
+%If the link with the largest cost is found and is a direct neighbor then the
+% connection is put in a will-cut list and stays there for a certain period of
+%time. If it is not, then it is handled by other peers. After that period,
+%connections are cut and recorded to $P$'s cut-list.
+%
+%\paragraph*{Source probing}
+%For a peer $P\in(N^2(S) - N(S))$ who receives only one $d(i, S, 0)$, the cost
+% of $PS$ is obtained (with list look-up or probing). Then $P$ compares it with
+%the cost from each hop and if $PS$ has the largest cost, $P$ will not keep this
+%connection, while otherwise the connection will be created.
+%
+%\paragraph{}
+%Supposing $n$ is the number of peers, $c_n$ is the average number of neighbors
+% and $c_e$ is the average cost of logical links, then in the flooding-based
+%search the traffic incurred by one query from an arbitrary peer in a
+%peer-to-peer network is $O(n)$. As observed in the Gnutella network
+%\cite{sripanidkulchai_gnutella_2001}, each peer issues $0.3$ queries per minute
+%in average, thus the per minute traffic incurred by the network with $n$ peers
+%is $O(n^2)$. Because each $d(i, S, v)$ has a TTL of $2$ in each source peer,
+%the traffic for one time LTM optimization in all peers is at most $2nc_n^2c_e$.
+%If each peer uses LTM $k$ times per minute, the total traffic incurred is
+%$2knc_n^2c_e$. Simulation shows the best value for $k$ is $2$ or $3$ So, the
+%traffic overhead caused by LTM to the network is $O(n)$.
+%
+%TTLj-detectors, with $j > 2$, would detect and break cycles with more than 4
+% links. LTM though, does not use such detectors because detector-flood traffic
+%would increase significantly, and cut links between two end-peers, could cause
+%queries initiated by them to traverse a path much more expensive than the cost
+%on the the cut link.
+%
+%\paragraph{}
+%LTM disadvantages are
+%\begin{inparaenum}[\itshape i\upshape)]
+%  \item disagreement of measured delay due to unsynchronized clocks causes
+% problems when deciding the cut positions, which can influence the network
+%connectivity, and
+%  \item the network delay metric mainly focuses on disabling the connections
+% between peers physically far away without considering the shortcuts created by
+%powerful peers.
+%\end{inparaenum}
+%
+\emph{LTM} fares as follows regarding our three  criteria:
+\begin{center}
+{\footnotesize
+\begin{tabular}{ccc}
+\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
+\hline
+% Authors claim 75% reduction on traffic cost and about 65% reduction on query
+% response time.
+medium &
+%
+medium &
+%
 medium
@@
 saved by the duplicate message reduction is spent to control the cycle cutting
-mechanism it is reported that still stays for example one to two orders of
-magnitude less than LTM due to its ``lazy'' broadcasting of control messages
-(as opposed to LTM's periodic approach) \cite{ZKB2008}.
+mechanism, it is reported that it does a lot better in that area
+due to its ``lazy'' broadcasting of control messages, as opposed to, for example,
+LTM's periodic approach \cite{ZKB2008}.
 %
@@
 %network.
-The expected \emph{DCMP} behavior as far as our $3$ criteria is:
+The expected \emph{DCMP} behavior as far as our three criteria is:
 %
@@
 % topology mismatch and thus DCMP minimizes this as possible.
+% It is reported though that still stays for example one to two orders of
+% magnitude less than LTM due to its ``lazy'' broadcasting of control messages
+% (as opposed to LTM's periodic approach) \cite{ZKB2008}.
 low &
@@
 increases in the network, it would be relatively easy to locate items even if
-search is random.
-
-The work discusses $3$ approaches for replication allocation namely \emph{uniform},
-\emph{proportional}, and \emph{square-root replication}. In the uniform model,
-all objects are replicated without taking into account the query distribution,
+search is random. Concretely, three three approaches for replication allocation
+are discussed: \emph{uniform}, \emph{proportional}, and
+\emph{square-root replication}.
+
+In the uniform model, all objects are replicated without taking into account
+the query distribution,
 while in the proportional, objects are replicated analogously to their query
@@
 {\sl Gnutella}'s query flooding method but incurring up to two
-orders of magnitude less network traffic. Also checks $3$ strategies for
+orders of magnitude less network traffic. Also checks three strategies for
 replica placement, namely \emph{owner replication} where an object is replicated
@@
 
-The table below outlines how the $3$ allocation approaches fare:
+The table below outlines how the three allocation approaches fare:
 %%
@@
 infrastructure. In the context of this work, the inefficiency caused by the
-topology mismatch and attempted to be tackled by building an highly connected
-graph termed \emph{mesh}
+topology mismatch is attempted to be tackled by building a highly connected
+graph, termed \emph{mesh},
 with each source featuring its own minimum spanning tree. A gossip-protocol was
@@
 %as shown in Figure~\ref{figure:narada}.
-Both graph and trees are dynamically updated as nodes keep joining 
+The graph and trees are dynamically updated as nodes keep joining 
 or departing the network. The protocol aims to ease the physical link stress,
@@
 %%
-\emph{AOTO}'s performance regarding the $3$ criteria is as follows:
+\emph{AOTO}'s performance regarding the three criteria is as follows:
 %%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \subsubsection{Location-aware Topology Matching (LTM)}
-
-\emph{Location-aware Topology Matching (LTM)} \cite{LLXNZ2004}
-seeks to optimize an overlay \p\ structure based on the physical topology.
-To achieve this, peers issue special messages called
-\textit{TTL-detector}s whose \emph{TTL} values are $0$ or $1$;
-in this regard, peers 
-discover $1$- and $2$-hop neighbor sets, designated as $N$ and $N^2$
-respectively, and proceed to compute communication costs.
-Time-stamps are used to derive network latency measurements that are then used 
-to improve the overlay network without sacrificing the search scope.
-Each node compares the latency information
-received from its direct neighbors;
-peers with longer latencies are placed on a
-\emph{will-cut} list where they remain for a certain period of time 
-after which they are finally eliminated 
-and are placed on the peer's \emph{cut-list}. 
-Thus, low-productivity connections are dropped and replaced by 
-more efficient ones, reducing the
-latency on the overall network. 
-Although \emph{LTM} improves the overall efficiency of
-the \p\ network, it is unable to offer any warranty for 
-effectively addressing the mismatch problem.
-%
-% TODO: SOME DISCUSSION
-%
-%\paragraph*{Low productive connection cutting}
-%There are three cases for any peer $P$ who receives $d(i, S, v)$ multiple
-% times:
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item $P$ receives both $d(i, S, 1)$ and $d(i, S, 0)$
-%  \item $P$ receives multiple $d(i, S, 0)$s from different paths, and P
-% randomly chooses to process one
-%  \item $P$ receives one $d(i, S, 1)$ and multiple $d(i, S, 0)$s, and $P$
-% processes $d(i, S, 1)$ and one randomly selected $d(i, S, 0)$
-%\end{inparaenum}
-%If the link with the largest cost is found and is a direct neighbor then the
-% connection is put in a will-cut list and stays there for a certain period of
-%time. If it is not, then it is handled by other peers. After that period,
-%connections are cut and recorded to $P$'s cut-list.
-%
-%\paragraph*{Source probing}
-%For a peer $P\in(N^2(S) - N(S))$ who receives only one $d(i, S, 0)$, the cost
-% of $PS$ is obtained (with list look-up or probing). Then $P$ compares it with
-%the cost from each hop and if $PS$ has the largest cost, $P$ will not keep this
-%connection, while otherwise the connection will be created.
-%
-%\paragraph{}
-%Supposing $n$ is the number of peers, $c_n$ is the average number of neighbors
-% and $c_e$ is the average cost of logical links, then in the flooding-based
-%search the traffic incurred by one query from an arbitrary peer in a
-%peer-to-peer network is $O(n)$. As observed in the Gnutella network
-%\cite{sripanidkulchai_gnutella_2001}, each peer issues $0.3$ queries per minute
-%in average, thus the per minute traffic incurred by the network with $n$ peers
-%is $O(n^2)$. Because each $d(i, S, v)$ has a TTL of $2$ in each source peer,
-%the traffic for one time LTM optimization in all peers is at most $2nc_n^2c_e$.
-%If each peer uses LTM $k$ times per minute, the total traffic incurred is
-%$2knc_n^2c_e$. Simulation shows the best value for $k$ is $2$ or $3$. So, the
-%traffic overhead caused by LTM to the network is $O(n)$.
-%
-%TTLj-detectors, with $j > 2$, would detect and break cycles with more than 4
-% links. LTM though, does not use such detectors because detector-flood traffic
-%would increase significantly, and cut links between two end-peers, could cause
-%queries initiated by them to traverse a path much more expensive than the cost
-%on the the cut link.
-%
-%\paragraph{}
-%LTM disadvantages are
-%\begin{inparaenum}[\itshape i\upshape)]
-%  \item disagreement of measured delay due to unsynchronized clocks causes
-% problems when deciding the cut positions, which can influence the network
-%connectivity, and
-%  \item the network delay metric mainly focuses on disabling the connections
-% between peers physically far away without considering the shortcuts created by
-%powerful peers.
-%\end{inparaenum}
-%
-\emph{LTM} fares as follows regarding our $3$  criteria:
-\begin{center}
-{\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
-% Authors claim 75% reduction on traffic cost and about 65% reduction on query
-% response time.
-medium &
-%
-medium &
-%
-medium
-\end{tabular}
-}
-\end{center}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % \subsubsection{Scalable Bipartite Overlay (SBO)}
@@
 establish connections to differently colored peers. 
-In this regard, all peers form a a bipartite overlay. 
-Using the network delay as a metric, white-peers 
-peers measure distances from red counterparts
+In this regard, all peers form a bipartite overlay. 
+Using the network delay as metric, white-peers 
+measure distances from red counterparts
 and report the encountered red neighbors.
@@
 red-peers create minimum spanning trees for the neighbors in question and
-assign efficient $2$-hop forwarding paths. This process can render a white
-peer $E$ non--forwarding neighbor of the red peer $P$. Direct neighbor
+assign efficient $2$-hop forwarding paths. This process can render a white-peer
+$E$ non--forwarding neighbor of the red-peer $P$. Direct neighbor
 replacement is a process conducted by $E$ in order to replace $P$ with a
@@
 % on traffic cost and about 60 percent reduction on query response time.
-In regards to the $3$ stated criteria, \emph{SBO} behaves as follows:
+In regards to the three stated criteria, \emph{SBO} behaves as follows:
 \begin{center}
@@
 %%
-In \cite{HLY2010}, the simulation-derived results of \emph{THANCS} as far
-average broadcast delay and overlay improvement guarantees are discussed;
-overall, \emph{THANCS} fares as follows regarding the $3$ stated criteria:
+\emph{THANCS} reportedly improves brodcast performance the extent of which, though,
+depends on the underlying network since \emph{THANCS} does not provide the
+performance guarantee for the improvement of any given overlay topology~\cite{HLY2010}.
 
@@
 %LTM by an average of 55 percent.
-In terms of the $3$ stated criteria, \emph{HAND} fares as follows:
+In terms of the three stated criteria, \emph{HAND} fares as follows:
 \begin{center}
@@
 due its on-line learning process and inherent complexity \cite{ZZLZ2009}. 
-\emph{APS}' behavior regarding the $3$ criteria has as follows:
-% As far as the $3$ criteria is concerned, \emph{APS} has as follows:
+\emph{APS}'s behavior regarding the three criteria is as follows:
+% As far as the three criteria is concerned, \emph{APS} has as follows:
 \begin{center}
@@
 $\left( 1 - \alpha \right) \ast N$ distant (latency above $x$) nodes
-as its neighbors. Of course, the bootstrapping peer is not able check all the
+as its neighbors. Of course, the bootstrapping peer is not able to check all the
 peers in the network to find the global best for the above sets. Authors
@@
 %%
-Regarding the $3$ criteria, \emph{ITA} is placed as follows:
+Regarding the three criteria, \emph{ITA} is placed as follows:
 %
@@
 maintenance while complexity $O(n^2)$ is required less frequently 
-and only when nodes reevaluate their connections.
+and only when nodes re-evaluate their connections.
 %
 \emph{EGOIST} fares as follows:
-% Regarding the $3$ criteria, \emph{EGOIST}'s behavior is as follows:
+% Regarding the three criteria, \emph{EGOIST}'s behavior is as follows:
 \begin{center}
@@
 \emph{BNS}~\cite{BCCMSBZ2006},
-\emph{Oro} leads to improved performance 
+\emph{Ono} leads to improved performance 
 without requiring any explicit cooperation between 
@@
 infrastructure and network topology information.
-\emph{Oro}'s selection approach is 
+\emph{Ono}'s selection approach is 
 landmark-based and leverages existing {\sl CDN}-infrastructure 
@@
 is sufficiently fresh, 
-\emph{Oro} persists the map after the end of a {\sl BitTorrent} session.
+\emph{Ono} persists the map after the end of a {\sl BitTorrent} session.
 %%
@@
 the later sorts its swarm peers according to
-their hop-count distance from the requesting peer 
-send out the top-$k$ list of nodes (e.g., $k$=$50$).
+their hop-count distance from the requesting peer and 
+sends out the top-$k$ list of nodes (e.g., $k$=$50$).
 %%
@@
 
-All $3$ techniques in~\cite{LCLX2009} seek to harness 
+All three techniques in~\cite{LCLX2009} seek to harness 
 the locality of the basic functions of the
@@
 traffic and fairness among peers in terms of the data the peers upload~\cite{LCLX2009};
-as far as the $3$ criteria, the overall approach fares as follows:
+as far as the three criteria, the overall approach fares as follows:
 %%
@@
 if compared with popular {\sl BitTorrent} implementations.
-In reference to the stated $3$ criteria, \emph{TopBT} fares as follows:
+In reference to the stated three criteria, \emph{TopBT} fares as follows:
 \begin{center}
@@
 and at the same time offer reduced burden on the underlying \emph{ISP}s~\cite{LCY2008}. 
-\emph{UTAPS} stands as follows in regard to the $3$ criteria:
+\emph{UTAPS} stands as follows in regard to the three criteria:
 %%
@@
 while peers communicate with the tracker. It is assumed that even though the
-network may be highly volatile, these core routers are are usually more stable.
+network may be highly volatile, these core routers are usually more stable.
 For this, a newly arriving
@@
 This clustering approach fares as follows in regard 
-to the $3$ criteria:
+to the three criteria:
 %%
@@
 Subsequently, $u$ contacts, at a fixed time rate \emph{timer}, 
-a random node $v$, $nhops$ away. % from it. 
+a random node $v$, $n$-hops away. % from it. 
 New distance information is calculated independently 
@@
 names and determine domain locality respectively. 
-In terms of our $3$ criteria, \emph{DDNO} fares are follows:
+In terms of our three criteria, \emph{DDNO} fares are follows:
 \begin{center}
@@
 connections, it disconnects those neighbors with the lowest \emph{AM}.
-\emph{CTAG} behaves as follows regarding the $3$ criteria:
+\emph{CTAG} behaves as follows regarding the three criteria:
 \begin{center}
@@
 %%
-For a incoming peer $Q$, 
+For an incoming peer $Q$, 
 the \emph{grouping criterion} designates that when the distance of $Q$ and
-some group $A$’s neighbor groups is the same as the distance
-between group $A$ and group $A$’s neighbor groups, then host $Q$
+some group $A$'s neighbor groups is the same as the distance
+between group $A$ and group $A$'s neighbor groups, then host $Q$
 should belong to group $A$. 
@@
 %%
-In terms of the stated $3$ criteria, \emph{mOverlay} fares as follows:
+In terms of the stated three criteria, \emph{mOverlay} fares as follows:
 \begin{center}
@@
 modify the topology of the \p\ network
-using various schemes. The two most commonly used
-create \emph{spanning tree}s using connection graphs
-as well as  \emph{clusters} of physically close nodes.
+using various schemes, the two most common of which
+create either \emph{spanning tree}s using connection graphs
+or \emph{clusters} of physically close nodes.
 
@@
 the node in the middle, does not flood its neighbors as instead 
-in only picks one of them.
+it only picks one of them.
 The dashed line on the right, shows that, 
@@
 helps reduce the burden of network resources, 
-their contribution in addressing the actual mismatch between the overlay and
+their contribution in addressing the actual mismatch between the overlay
 and the underlying networks remains limited.
@@
 for this purpose.
+
+\begin{figure}[ht]
+\centering
+  \includegraphics[scale=0.34]{img/pdf/landmarking.pdf}
+\caption{Landmark binning during node bootstrap.}
+\label{figure:landmarking}
+\end{figure}
+
+\input{unstructured-table}
+\input{unstructured-plot}
+%%
 For example, in Figure~\ref{figure:landmarking}, the newly
@@
 other, as well, over the network.
-%%
-\begin{figure}[ht]
-\centering
-  \includegraphics[scale=0.34]{img/pdf/landmarking.pdf}
-\caption{Landmark binning during node bootstrap.}
-\label{figure:landmarking}
-\end{figure}
-
-\input{unstructured-table}
-\input{unstructured-plot}
 
@@
 especially in an environment with high churn.
-%%VM Some rewording and restructuring. Please have a look.
-%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-06-03 background.tex
2015.06.03 (Mema & Vassilis) Mema's latest pass. Also added paper for citation LCCLS2002.

@@
 that were peer-provided.
-Users would query the index server that would ultimately 
-yield pointers to the actual content.
+Users would query the index server that would provide 
+pointers to the actual content.
 Thus, by centralizing search while distributing downloads,  {\sl Napster}
-achieved a highly functional design that, at its height, attracted $26$ million
+achieved a highly efficient design that, at its height, attracted $26$ million
 users sharing approximately $80$ million songs\cite{jmm_naptopusage_2001}.
@@
 for this reason, such architectures are referred to as \emph{unstructured}. 
-Key properties of such \p\ systems are that they support inherent heterogeneity
+Key properties of such \p\ systems are that they support the inherent heterogeneity
 of peers, are highly resilient to peer failures, and incur low maintenance
-overhead at handling the dynamics of peer participation
+overhead while handling the dynamics of peer participation
 \cite{stutzbach_churn_2006}. 
@@
 (Figure~\ref{figure:p2p-archs:structured}).
-Here, the main objective was to offer 
+Here, the main objective is to offer 
 a self-organizing infrastructure for large-scale \p\ applications
 \cite{ratnasamy_can_2001,stoica_chord_2001,antony_pastry_2001,zhao_tapestry_2001,maymounkov_kademlia_2002,rgrk_bamboo_2004}.
-To achieve that, they implement a \emph{Distributed Hash Table (DHT)} that maps
+To achieve this, they implement a \emph{Distributed Hash Table (DHT)} that maps
 objects to nodes through a deterministic mechanism. \emph{DHT}s provide a
@@
 Moreover, \emph{DHT}s sparked
-a flurry of research for many, diverse, applications ranging
+a flurry of research for many diverse applications, ranging
 from distributed search engines to event-notification and cloud-based
@@
 %%
-For instance, assume nodes $A$, $B$, $C$ and $D$ being 
+For instance, assume nodes $A$, $B$, $C$ and $D$ are 
 connected via the physical network shown in Figure~\ref{figure:phys}
@@
 %%
-When peer $A$ sends a message its $D$ counterpart
-and depending on the overlay configuration, 
+When peer $A$ sends a message to peer $D$, 
+depending on the overlay configuration, 
 nodes will experience different performance.
@@
 
-Early overlay protocol implementations did not pay much attention to map
-optimally to the underlying network topology. {\sl Gnutella}, for example, is
+Early overlay protocol implementations gave litte attention to the 
+topology mismatch problem.
+{\sl Gnutella}, for example, is
 considered far from scalable~\cite{ritter_gnucantscale_2001} for every peer
@@
 Queries are flooded over multiple paths in an effort 
-to reach the node(s) having the sought content. 
+to reach nodes that have the sought content. 
 In addition, 
@@
 Even when an approach offers the best possible location
-when a node joins in, the churn is set to  
-continuously undermine the effort to maintain the best possible position
+when a node joins, churn   
+continuously undermines the effort to maintain the best possible position
 of the  peer in the network.
@@
 and overlay topology maintenance. For instance, the overhead that an algorithm,
-which needs global knowledge of the overlay topology, imposes to the system is
+which needs global knowledge of the overlay topology, imposes on the system is
 greater compared to a protocol that only relies on peers exchanging routing
@@
 protocols as being lowly, mediumly or highly scalable if they are proven to
-effectively support hundreds, tenths of thousands or millions of nodes,
+effectively support hundreds, tenths of thousands, or millions of nodes,
 respectively.

### 2015-06-03 conclusion.tex
2015.06.03 (Mema & Vassilis) Mema's latest pass. Also added paper for citation LCCLS2002.

@@
 
-Research and development in applications and systems that follow the 
-\p\ architectural paradigm have flourished for more than a decade now.
-By and large, such artifacts exploit the loose-coupling and self-organization 
-of participating nodes to shape substrates upon which diverse applications 
-can run; the latter offer services including, but not limited to, 
-distributed naming, 
-Internet telephony and video conferencing, 
-storage and indexing, 
-content and file sharing, 
-web crawling and caching, 
-event notification and subscribing/publishing.
-%%
-These services are provided atop the 
-best-effort infrastructure of today's Internet 
-and they have to effectively deal with an array of key quality attributes that
-they may have to offer, such as node availability, robustness,
-scalability, load-sharing, quality-of-service and user anonymity.
-%%
-The difficulty to handle such features effectively, emanates from the fact that
-peer nodes operate at the edge of the Internet and no adequate attention 
-is paid to the structure of the physical network during the 
-network formation at the application-level.
-The deviation of the distributed logically-formed overlay from the 
-physical network yields suboptimal use of the underlying 
-infrastructure and is known as the topology mismatch
-problem.
-
+%% Removing after Mema's suggestion
+%Research and development in applications and systems that follow the 
+%\p\ architectural paradigm have flourished for more than a decade now.
+%By and large, such artifacts exploit the loose-coupling and self-organization 
+%of participating nodes to shape substrates upon which diverse applications 
+%can run; the latter offer services including, but not limited to, 
+%distributed naming, 
+%Internet telephony and video conferencing, 
+%storage and indexing, 
+%content and file sharing, 
+%web crawling and caching, 
+%event notification and subscribing/publishing.
+%%%
+%These services are provided atop the 
+%best-effort infrastructure of today's Internet 
+%and they have to effectively deal with an array of key quality attributes that
+%they may have to offer, such as node availability, robustness,
+%scalability, load-sharing, quality-of-service and user anonymity.
+%%%
+%The difficulty to handle such features effectively, emanates from the fact that
+%peer nodes operate at the edge of the Internet and no adequate attention 
+%is paid to the structure of the physical network during the 
+%network formation at the application-level.
+%The deviation of the distributed logically-formed overlay from the 
+%physical network yields suboptimal use of the underlying 
+%infrastructure and is known as the topology mismatch
+%problem.
+%
 Development of \p--systems and applications
@@
   \item offering a pictorial
-comparison of how each solution fares in regards to others as far as
+comparison of how each solution fares with regards to others 
+as far as
 efficiency, overhead, and scalability are concerned,
 \end{inparaenum}
-we hope to have provided \p\
+we hope to provide \p\
 developers and researchers with enough insight and perspective;
 as they face specific problems, they will be readily able to 
-draw the best possible design decisions for their own application
-environments.
+draw the best possible design decisions for their applications.

### 2015-06-03 intro.tex
2015.06.03 (Mema & Vassilis) Mema's latest pass. Also added paper for citation LCCLS2002.

@@
 services such as routing messages to destinations that are not known in advance
-or offering QoS guarantees (i.e., in content-distribution) over best effort
+or offering QoS guarantees (i.e., in content-distribution) over best-effort
 infrastructures. Overlay nodes communicate through \emph{virtual connections}
@@
 it does not match the underlying topology well, 
-the inherent \emph{topology mismatch} gives two
+the inherent \emph{topology mismatch} creates two
 major problems. 
@@
 redundant, number of physical hops resulting in poor user experience 
-entailing noticeable latencies or jittering. 
+entailing noticeable latencies or jitter. 
 %%
@@
 $7$~\emph{Exabytes}-per-month by $2018$~\cite{CVNI2014}! 
-Evidently, this constitutes a major burden for 
+This constitutes a major burden for 
 Internet Service Providers (ISPs) who must route 
@@
 Hence, it is critical that \p\ networks be laid out 
-in ways that their topology matches, as closely as possible, the 
-underlying IP topology.
+in ways that their topology matches the underlying IP topology as 
+closely as possible.
 
@@
 The rest of the paper is organized as follows: 
-Section~\ref{section:background} provides background on
+In Section~\ref{section:background}, we provide background on
 overlay architectures including centralized, decentralized-unstructured and
@@
 evaluation of the techniques discussed in this paper. 
-Work in the area was inspired mostly due to the volatile and fully
+Work in the area was inspired mostly by the volatile and fully
 distributed nature of the decentrilized overlays so we skip
@@
 for unstructured and structured \p\ overlays, respectively.
-Conclusions can be found in Section~\ref{section:conclusion}.
+We conclude in Section~\ref{section:conclusion}.

### 2015-06-03 structured.tex
2015.06.03 (Mema & Vassilis) Mema's latest pass. Also added paper for citation LCCLS2002.

@@
 across the network and made available to any node. The notion of a
-region is a well defined property for structured \p\ networks and can be part of
+region is a well-defined property for structured \p\ networks and can be part of
 the Cartesian space in overlays like \emph{eCAN}~\cite{xu_ecan_2002} or a set of
@@
 find the best way to route their messages, while a publish/subscribe mechanism
-allows them to get notified on dynamic network changes.
+allows them to get notified about dynamic network changes.
 
@@
 The fact that a peer may appear in multiple
-such maps points into additional traffic for probing and notifications
-and so renders the approach not scalable \cite{RGJZ2004}. 
+such maps incurs additional traffic for probing and notifications
+and so renders the approach unscalable \cite{RGJZ2004}. 
 Moreover, to refine measurements
@@
 During bootstrap, an incoming peer $J$ obtains from a bootstrap node a set of
-candidate peers to establish links with. 
+candidate peers with which to establish links. 
 Subsequently, the node works iteratively to identify peers that minimize
@@
 $J$ is then assigned a synthetic \emph{ID} that
-represents the state of minimum distance from its close-by $2$-$d$ neighborhood as
-this was discovered during the aforementioned phase. % iterative selection phase.
+represents the state of minimum distance from its nearby $2$-$d$ neighborhood 
+discovered during the aforementioned phase. % iterative selection phase.
 The benefit of adopting synthetic coordinates is that distance computations
@@
 measurements. 
-Mithos proceeds then to designate $J$'s closest neighbors as 
-previous steps do not reveal with certainty which these nodes are.
+Mithos then proceeds to designate $J$'s closest neighbors as 
+previous steps do not reveal with certainty which nodes are its closest
+neighbors.
 %%
@@
 into a tree-based hierarchy so that 
-hops required during message routing be reduced and 
-maintenance overhead be minimized.
-A caching scheme is also used so as
-to further lower routing table update costs. 
+hops required during message routing are reduced and 
+maintenance overhead is minimized.
+A caching scheme is also used 
+to further reduce routing table update costs. 
 It is theoretically shown that 
@@
 The physical distances are estimated
-with round trip times to a few existing nodes in the overlay network.
+with round-trip times to a few existing nodes in the overlay network.
 %%
@@
 
-In order to optimize the underlying network use,
+To optimize the underlying network use,
 \cite{KLKP2008} introduces proximity
@@
 when combined with \emph{PRS}/\emph{PNS} helps contain 
-$40\%$ of messages within an \emph{ISP} and allows only a $6\%$ 
+$40\%$ of messages within an \emph{ISP} and allows only $6\%$ 
 to go through the backbone.
@@
 (i.e., a new peer appears), by
-estimating distance on the overlay's ID space.
+estimating the distance in the overlay's ID space.
 
@@
 distributed at national level (NIRs), the first $32$-bits of the address
-can estimate the region or the country the node is connected in. 
+can estimate the region or the country in which the node is connected. 
 Thus, CHOP6's IPv6 prefix ID is used for proximity estimation
@@
 %%
-To achieve that, each node exploits the stable structure of
+To achieve this, each node exploits the stable structure of
 the Internet routers to help solve a clustering problem. 
@@
 its extensive use may create overhead on the overall network structure. 
-In this respect, administrators frequently
-disable {\sl traceroute} support on their routers this 
+Administrators thus frequently
+disable {\sl traceroute} support on their routers and this 
 undoubtly affects \emph{T2MC}'s operation.
@@
 % \subsubsection{PChord}
-\emph{PChord} \cite{HLYW2005} is based on the \emph{Chord DHT} and adds
-proximity into its routing mechanism. 
-To achieve that, the standard finger table
+\emph{PChord} \cite{HLYW2005} is based on the \emph{Chord DHT} and incorporates
+proximity in its routing mechanism. 
+To achieve this, the standard finger table
 is augmented with a \emph{proximity list} that contains
-nodes in the network found close-by in terms of RTT.
+nodes in the network found nearby in terms of RTT.
 %in this way, the next hop can be decided based on both progress
@@
 This list is maintained in a heuristic fashion during the node's lifetime.
-While a node attempts routing, the next hop is selected
-as the closest node to the target key 
-from \emph{both} its finger table and proximity list. 
+When a node needs to route a message, it chooses as the next hop 
+the node closest to the target key 
+in \emph{both} its finger table and proximity list. 
 An initial choice for the next hop made out of the finger list 
 can be substituted by a shorter one from the proximity list if  
-it is both closer to the key and in less physical distance.
+it is both closer to the key (ID-wise) and closer in physical distance.
 This yields an equal or even more efficient routing than traditional Chord.
@@
 incoming peer is positioned \emph{between} the two identifiers mentioned above. 
-After joining, a \emph{finger} table is created ala \emph{Chord}. 
+After joining, a \emph{finger} table is created as in \emph{Chord}. 
 Moreover, an \emph{adjacency table} maintains 
@@
 
-\emph{Chord6} does reduce the average 
+\emph{Chord6} reduces the average 
 length of hops traversed
@@
 this characteristic favors a more load--balanced clustering operation
-should we assume uniform node distribution.
+assuming uniform node distribution.
 Every sector in this $2d$ space is mapped to 
@@
 % \subsubsection{DynaMO}
-\emph{DynaMO}~\cite{WZS2004} attempts to create an overlay that 
-does not only consider physical proximity amongst peers but also
+\emph{DynaMO}~\cite{WZS2004} creates an overlay that 
+not only considers physical proximity amongst peers but also
 takes into account mobility attributes of nodes.
@@
 
-A new-coming peer then computes distances from those landmarks,
+An incoming node then computes distances from those landmarks,
 sorts them, and assigns itself an \emph{ID} on this 
@@
 The latter in particular helps reduce average routing cost(s)
-for the leaf-set of a node is used to efficiently determine the 
+as the leaf-set of a node is used to efficiently determine the 
 last routing step.
@@
 It delegates the task of identifying the physically
-close--node to \emph{Pastry}; this node is used by the 
+close--node in a \emph{Pastry}-based network; this node is used by the 
 incoming peer to get bootstrapped.
@@
 %%
-\emph{RLM} gets evaluated as presenting good proximity in a load-balanced 
+\emph{RLM} is evaluated and exhibits good proximity in load-balanced 
 environments. Moreover, it is adaptable to high-churn settings.
@@
 For the above reasons, \emph{3DO} exhibits the following behavior 
-in conjunction with our three stated criteria:
+with respect to our three stated criteria:
 %%
@@
 for distances and a \emph{jumping phase} in which the peer decides 
-to change its neighborhood in order to reduce average neighbor distances. 
+to change its neighborhood to reduce average neighbor distances. 
 This process is performed iteratively
@@
 Table~\ref{structured:table} offers a comparative view
-of the algorithms in a way reminiscent to that used in Table~\ref{unstructured:table}.
+of the algorithms and
 Table~\ref{structured:table} outlines key properties for all surveyed approach 
@@
 and \ref{structured:plot:scalability} help with the pictorial comparison
-of the algorithms' behavior with respect to the three effectiveness
+of the algorithms' behavior with respect to the three 
 criteria we defined in subsection~\ref{background:motivation}.
@@
 the participants. This comes in contrast to proximity routing and proximity neighbor
-selection (discussed later on) where decisions are taken based on and effecting
+selection (discussed later on) where decisions are taken based on and affecting
 a peer's nearby environment.
@@
 \emph{Global Soft-State},
-\emph{Mithos} and \emph{LAPTOP} resort to clustering close-by nodes in
-order to reduce the network's average delay.
+\emph{Mithos} and \emph{LAPTOP} resort to clustering nearby nodes 
+to reduce the network's average delay.
 Landmark servers and \emph{RTT} measurements are two popular methods, which
@@
 the latter is especially true if nearby 
-nodes happen to be connected through low bandwidth connections.
+nodes happen to be connected through low-bandwidth connections.
 
@@
 In the degenerate case, a system could use a greedy approach to select
-only the lowest latency candidate peers during forwarding. That would to
-longer, sub--optimal, paths to the destinations; thus imposing
-extra stress to network resources. PR is also less
+only the lowest latency candidate peers during forwarding. This leads to
+longer, sub-optimal, paths to the destinations and thus imposes
+extra stress on the network. PR is also less
 effective than geographical layout when applied to CAN--like implementations.
@@
 detect nearby nodes. \cite{CDCR2002a} reports that proximity neighbor
-selection is an effective proximity based method.
+selection is an effective proximity-based method.
 %%AD perhaps the ending of the previous segment should be used here...

### 2015-06-03 unstructured.tex
2015.06.03 (Mema & Vassilis) Mema's latest pass. Also added paper for citation LCCLS2002.

@@
 Its major drawback is the delay between successive iterations, as the
-source-node has to examine the results at each attempt before deciding to
+source-node has to examine the results with each attempt before deciding to
 either quit or ``unfreeze'' the query.
@@
 The source node that issues the query designates the delay 
-boundary it may tolerate, providing so a \emph{pruning factor}.
+boundary it may tolerate, by providing a \emph{pruning factor}.
 In this context, user requests are forwarded only to 
@@
 the clustering of routing tables. 
-In a dynamic environment, the formation 
-of very accurate such routing tables might be however an elusive goal.
+In a dynamic environment, the formation of such 
+accurate routing tables may be an elusive goal.
 %With the clustered
@@
 each node maintains about its satisfaction --ranging between $0$ and $1$--
-for the neighbors it finds itself associated with. 
-Through message exchange, a peer can establish new connections 
-or drop superseded ones in order to improve its satisfaction.
+for the neighbors with which it finds itself associated. 
+Through message exchanges, a peer can establish new connections 
+or drop superseded ones to improve its satisfaction.
 Despite the fact that {\sl Gia}'s topology adaptation algorithm 
-that primarily focuses on the satisfaction of peers improves the network's
+improves the network's
 scalability, it falls short in addressing the mismatch problem
@@
 it cannot  directly contribute to the solution of the topology mismatch
-problem, while its performance in a high traffic and churn environments and its
+problem, while its performance in high-traffic and churn environments and its
 ability to retain important characteristics of a {\sl Gnutella}-like topology (i.e.,
-like degree distribution, average peer distance, diameter e.t.c.) has been
+degree distribution, average peer distance, diameter e.t.c.) has been
 doubted~\cite{CSG2010}. 
-On the other hand, even though some of the overhead
-saved by the duplicate message reduction is spent to control the cycle cutting
-mechanism, it is reported that it does a lot better in that area
+On the other hand, overhead is greatly reduced 
 due to its ``lazy'' broadcasting of control messages, as opposed to, for example,
@@
 search is random. Concretely, three three approaches for replication allocation
-are discussed: \emph{uniform}, \emph{proportional}, and
+are proposed: \emph{uniform}, \emph{proportional}, and
 \emph{square-root replication}.
@@
 proportional strategy effectively decreases the search time for popular items,
-but suffers when needing to locate the rare entities. When it comes to expected
-successful search size, it is the same for for both uniform and proportional
-models and any approach in between to always behave much better. Finally, the
-square-root model is an attempt to find the golden ratio between uniform and
-proportional allocation models. It suggests setting the number of object
+but suffers when needing to locate rare items. For allocation ratio ``between''
+1 and the ratio of the query rates, the authors
+prove that the average performance on successful queries is always better
+than in uniform and proportional models. Finally, with the
+square-root model they attempt to find the golden ratio in ``between'', by
+suggesting to set the number of object
 replicas to the square root of the searching rate for an object\footnote{An
@@
 Theory in \cite{CS2002} suggest that square-root minimizes the expected search size on
-successful queries, claim that is supported by simulations \cite{LCCLS2002}.
-
-In \cite{LCCLS2002}, a \emph{k-walker} query algorithm is presented and shown
+successful queries, a claim that is supported by simulations \cite{LCCLS2002}.
+
+Lv et al.~\cite{LCCLS2002}, present a \emph{k-walker} query algorithm and show
 that using it along side square-root allocation results in performance similar to
-{\sl Gnutella}'s query flooding method but incurring up to two
+{\sl Gnutella}'s query flooding method but incurs up to two
 orders of magnitude less network traffic. Also checks three strategies for
@@
 taking into account bandwidth and latency properties of the underlying physical
-infrastructure. In the context of this work, the inefficiency caused by the
-topology mismatch is attempted to be tackled by building a highly connected
+infrastructure. To overcome the inefficiency caused by 
+topology mismatch, Narada builds a highly connected
 graph, termed \emph{mesh},
@@
 In \emph{Selective Flooding}, a minimum spanning tree is built for each peer and
-its immediate neighbors so that queries do not flood the entire network and at
-the same time preventing their search scope from shrinking. This way some
+its immediate neighbors so that queries do not flood the entire network 
+while preventing their search scope from shrinking. This way some
 neighbors become non--flooding. During \emph{Active Topology}, each peer
@@
 In particular, a peer $S$ iteratively probes the distance $d$ between
-itself and every of its non-flooding neighbor nodes $G$ as well as
+itself and each of its non-flooding neighbor nodes $G$ as well as
 the distance between $S$ and $G$'s neighbors designated as $H$.
 If $d_{SG} > d_{SH}$, then link $SG$ is replaced by link $SH$. If, on the other
-hand, $d_{SG} < d_{SH}$ then also $d_{GH}$ and then we have the following
+hand, $d_{SG} < d_{SH}$ then we have the following
 options. Either $d_{SH} < d_{GH}$ in which case link $SH$ is kept or
@@
 $E$ non--forwarding neighbor of the red-peer $P$. Direct neighbor
-replacement is a process conducted by $E$ in order to replace $P$ with a
+replacement is a process conducted by $E$ to replace $P$ with a
 $P' \in N^2(P)$ as its new neighbor. This adaptation tackles the
@@
   	forward queries to connections in its will-cut list but will preserve them
-  	for some time in order not to effect ongoing response messages travelling the
+  	for some time in order not to affect ongoing response messages traveling the
   	inverse path.}. 
@@
 uses a fully-distributed triple--hop adjustment strategy, applied to a network
-graph $G$ in order to create an optimal graph $G^{*}$, free of the implications
+graph $G$ to create an optimal graph $G^{*}$, free of the implications
 of topology mismatch. This optimality is attained if all peer hop 
@@
 as its neighbors. Of course, the bootstrapping peer is not able to check all the
-peers in the network to find the global best for the above sets. Authors
+peers in the network to find the global best for the above sets. The authors
 prove that latency measurements against $30/ \alpha$ randomly selected peers
@@
 % chosen though. Maybe we can ask... Mema :)
-Searching occurs in $2$ phases: in the initial, the querying node
+Searching occurs in two phases: first, the querying node
 floods its distant neighborhood  with \emph{TTL=1}.
@@
 random graphs as it can lead to a larger coverage of peers and at the same time
-help reduce duplication of messages travelling around the network. This
+help reduce duplication of messages traveling around the network. This
 combination offers efficient lookups that pose minimal or no negative impact to
@@
 node through a bootstrap server. Once connected, the peer starts receiving
-information throughout the link-state mechanism and thus, after some time, 
+information via a link-state mechanism and thus, after some time, 
 it has a complete picture of the overlay graph. 
@@
 The second technique, applied in an \emph{intermediate-level}, manipulates
-{\sl BitTorrent}'s chocking/unchocking mechanism. A peer unchockes its $4$
+{\sl BitTorrent}'s chocking/unchocking mechanism. A peer unchokes its $4$
 closest, in terms of 
@@
 It also supports both link-hop and \emph{AS}-hop metrics to identify proximity in
-different levels in order to reduce overall traffic while preserving downloading
+different levels to reduce overall traffic while preserving downloading
 performance.
@@
 \emph{TopBT} offers more than $25\%$ traffic reduction and $15\%$ faster downloads
-if compared with popular {\sl BitTorrent} implementations.
+compared to popular {\sl BitTorrent} implementations.
 In reference to the stated three criteria, \emph{TopBT} fares as follows:
@@
 it collects information to develop a picture of the underlying topology 
-and in the second, it proceeds to make the peer selection based on 
+and in the second, it makes its peer selection based on 
 this knowledge. In the first stage, \emph{UTAPS} employs
@@
 hops traversed.
-The more the peers in the swarm the better 
+The more the peers in the swarm, the better 
 picture a tracker has for the underlying infrastructure.  
@@
 These newly joined peers traceroute the tracker
-peer list themselves and return back to the tracker in order to further enhance
+peer list themselves and return back to the tracker to further enhance
 the tracker's view of the infrastructure.
@@
 In choosing $c$, the tracker employs an iterative search process
-first in the peer's local neighborhood. If not sufficient number of 
-peers are available there, the tracker goes to intra-\emph{ISP} 
+first in the peer's local neighborhood. If there is not a sufficient number of 
+peers available, the tracker goes to intra-\emph{ISP} 
 and subsequently, if needed, to the inter-\emph{ISP} cluster. 
@@
 the total cross-\emph{ISP} traffic.
-This clustering approach fares as follows in regard 
+This clustering approach fares as follows with regards 
 to the three criteria:
@@
 are introduced to help adjust the neighborhood graph of an overlay network
-while aiming at reducing the network's overall link latency.
+while reducing the network's overall link latency.
 \emph{PROPs} are weaved around the concept of
@@
 from others by allowing two peers to optimize
-their neighborhood environment, than simply letting each node to ``selfishly''
+their neighborhood environment, in contrast with simply allowing each node to ``selfishly''
 choose its own strategy. 
@@
 not allowed to exchange neighbors along the path of the nodes that perform the
-exchange in order to ensure that in the end of the process the peers will remain
+exchange to ensure that in the end of the process the peers will remain
 connected. Additionally, the algorithm needs to preserve the characteristics of
-the network (i.e., the graph remains connected, preserve the degree of the
-nodes), so the exchange always involves equal number of neighbors.
+the network (i.e., the graph remains connected and the degree of the
+nodes is unchanged), so the exchange always involves equal number of neighbors.
 %%
@@
 effort, the methodologies it incorporates, its special highlights as well as 
-a rough estimation of the pros and cons in its implementation. 
+a brief presentation of the pros and cons in its implementation. 
 Last but not least, Figures~\ref{unstructured:plot:efficiency}, \ref{unstructured:plot:overhead}
@@
 Despite the fact that the above is a plausible proposition,
-it is not always advantageous though. For example, a topology adaptive
+it is not always advantageous. For example, a topology adaptive
 algorithm can exchange a slow edge in the overlay with a faster one
@@
 In Fig.~\ref{figure:forwarding-optimisation:after}, 
-the node in the middle, does not flood its neighbors as instead 
-it only picks one of them.
+the node in the middle, does not flood its neighbors; instead 
+it picks only one of them.
 The dashed line on the right, shows that, 

### 2015-07-18 mard-survey.tex
2015.06.25 (Vassilis) Minor changes to "thanks" and "bio" sections.

@@
         and~Alex Delis,~\IEEEmembership{Member,~IEEE}% <-this % stops a space
-\thanks{Vassilis Moustakas, Mema Roussopoulos and Alex Delis are 
+\thanks{Vassilis Moustakas is
+with \emph{Amazon Development Center Germany GmbH},
+DE80807, Munich Germany, email: \emph{vsmoustakas@gmail.com}
+(this work was done while a graduate student at the Department of Informatics and Telecommunications, \emph{University of Athens})}% <-this % stops a space
+\thanks{H\"useyin Akcan is with the Department of Software Engineering,
+\emph{Izmir University of Economics}, 35330 Balcova, Izmir, Turkey, email: \emph{huseyin.akcan@ieu.edu.tr}}% <-this % stops a space
+\thanks{Mema Roussopoulos and Alex Delis are 
 	with the Department of Informatics and Telecommunications, 
 	\emph{University of Athens},
-GR15784, Athens, Greece, e-mails: \emph{\{vsmoustakas,mema,ad\}@di.uoa.gr}}% <-this % stops a space
-\thanks{H\"useyin Akcan is with the Department of Software Engineering,
-\emph{Izmir University of Economics}, 35330 Balcova, Izmir, Turkey, email: \emph{huseyin.akcan@ieu.edu.tr}}% <-this % stops a space
-\thanks{Manuscript received April 30, 2015; revised April 30, 2015.}}
+GR15784, Athens, Greece, e-mails: \emph{\{mema,ad\}@di.uoa.gr}}% <-this % stops a space
+\thanks{Manuscript received June 28, 2015; revised June 28, 2015.}}
 
@@
 % Biography text here.
-is a researcher with the University of Athens in Athens, Greece.
+is a software development engineer at Amazon.com and a researcher with the University of Athens in Athens, Greece.
 \end{IEEEbiography}

### 2015-07-19 intro.tex
2015.07.19 (Vassilis) Reformatting for Elsevier publication targeting the Journal of Systems and Software (JSS).

@@
 
-\IEEEPARstart{P}{eer-to-peer (\p)} networks are self-organizing, distributed systems where
+Peer-to-peer (\p) networks are self-organizing, distributed systems where
 participating nodes, called \emph{peers}, act as both resource providers and

### 2015-07-19 mard-survey.tex
2015.07.19 (Vassilis) Reformatting for Elsevier publication targeting the Journal of Systems and Software (JSS).

--- 2015-05-17 mard-survey.tex
@@
-%% Based on:
-%% bare_jrnl.tex
-%% V1.4a
-%% 2014/09/17
-%% by Michael Shell
-%% see http://www.michaelshell.org/
-%% for current contact information.
 %% 
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.8a or later) with an IEEE
-%% journal paper.
+%% Copyright 2007, 2008, 2009 Elsevier Ltd
 %% 
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
-%% and
-%% http://www.ieee.org/
+%% This file is part of the 'Elsarticle Bundle'.
+%% ---------------------------------------------
+%% 
+%% It may be distributed under the conditions of the LaTeX Project Public
+%% License, either version 1.2 of this license or (at your option) any
+%% later version.  The latest version of this license is in
+%%    http://www.latex-project.org/lppl.txt
+%% and version 1.2 or later is part of all distributions of LaTeX
+%% version 1999/12/01 or later.
+%% 
+%% The list of all files belonging to the 'Elsarticle Bundle' is
+%% given in the file `manifest.txt'.
+%% 
+%% Template article for Elsevier's document class `elsarticle'
+%% with harvard style bibliographic references
+%% SP 2008/03/01
 
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE! 
-%% User assumes all risk.
-%% In no event shall IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%
-%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
-%%                    bare_conf.tex, bare_jrnl.tex, bare_conf_compsoc.tex,
-%%                    bare_jrnl_compsoc.tex, bare_jrnl_transmag.tex
-%%*************************************************************************
+%%\documentclass[preprint,12pt,authoryear]{elsarticle}
 
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. IEEE's font choices and paper sizes can       ***
-% *** trigger bugs that do not appear when using other class files.       ***                          ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
+%% Use the option review to obtain double line spacing
+%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
 
-\documentclass[journal]{IEEEtran}
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[journal]{../sty/IEEEtran}
+%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
+%% for a journal layout:
+%% \documentclass[final,1p,times,authoryear]{elsarticle}
+%% \documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
+%% \documentclass[final,3p,times,authoryear]{elsarticle}
+%% \documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
+%% \documentclass[final,5p,times,authoryear]{elsarticle}
+\documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}
 
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
+%% packages
 
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
+%% For including figures, graphicx.sty has been loaded in
+%% elsarticle.cls. If you prefer to use the old commands
+%% please give \usepackage{epsfig}
 
-% definitions
-%\newtheorem{theorem}{Theorem}[section]
-%\newtheorem{conjecture}[theorem]{Conjecture}
-%\newtheorem{corollary}[theorem]{Corollary}
-%\newtheorem{proposition}[theorem]{Proposition}
-%\newtheorem{lemma}[theorem]{Lemma}
-%\newdef{definition}[theorem]{Definition}
-%\newdef{remark}[theorem]{Remark}
-\newtheorem{definition}{Definition}
+%% The amssymb package provides various useful mathematical symbols
+\usepackage{amssymb}
+%% The amsthm package provides extended theorem environments
+%% \usepackage{amsthm}
 
-\newcommand{\p}{\emph{P2P}}
+\usepackage{subfigure}
+\usepackage{longtable}
+\usepackage{adjustbox}
+\usepackage{url}
 
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off
-% such as if a citation ever needs to be enclosed in parenthesis.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 5.0 (2009-03-20) and later if using hyperref.sty.
-% The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
-% The documentation is contained in the cite.sty file itself.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-  \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  \graphicspath{{./img/pdf/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  \graphicspath{{./img/eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
-\fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation
-% can be obtained at: 
-% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found at:
-% http://www.ctan.org/tex-archive/info/epslatex/
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-%\usepackage[cmex10]{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics. If using
-% it, be sure to load this package with the cmex10 option to ensure that
-% only type 1 fonts will utilized at all point sizes. Without this option,
-% it is possible that some math symbols, particularly those within
-% footnotes, will be rendered in bitmap form which will result in a
-% document that can not be IEEE Xplore compliant!
-%
-% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
-% There is also a support site at:
-% http://algorithms.berlios.de/index.html
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/tools/
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-\usepackage{longtable}
-
-\usepackage{adjustbox}
-
-% *** SUBFIGURE PACKAGES ***
-%\ifCLASSOPTIONcompsoc
-%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false,font=footnotesize]{subfig}
-%\fi
-% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
-% for subfigure.sty, the latter of which is no longer maintained and is
-% incompatible with some LaTeX packages including fixltx2e. However,
-% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
-% which will override IEEEtran.cls' handling of captions and this will result
-% in non-IEEE style figure/table captions. To prevent this problem, be sure
-% and invoke subfig.sty's "caption=false" package option (available since
-% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
-% handling of captions.
-% Note that the Computer Society format requires a larger sans serif font
-% than the serif footnote size font used in traditional IEEE formatting
-% and thus the need to invoke different subfig.sty package options depending
-% on whether compsoc mode has been enabled.
-%
-% The latest version and documentation of subfig.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
-\usepackage{subfigure}
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure. The latest version and documentation can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/base/
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
-% Do not use the stfloats baselinefloat ability as IEEE does not allow
-% \baselineskip to stretch. Authors submitting work to the IEEE should note
-% that IEEE rarely uses double column equations and that authors should try
-% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
-% packages (also by Sigitas Tolusis) as IEEE does not format its papers in
-% such ways.
-% Do not attempt to use stfloats with fixltx2e as they are incompatible.
-% Instead, use Morten Hogholm'a dblfloatfix which combines the features
-% of both fixltx2e and stfloats:
-%
-% \usepackage{dblfloatfix}
-% The latest version can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/dblfloatfix/
-
-%\ifCLASSOPTIONcaptionsoff
-%  \usepackage[nomarkers]{endfloat}
-% \let\MYoriglatexcaption\caption
-% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
-%\fi
-% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
-% Axel Sommerfeldt. This package may be useful when used in conjunction with 
-% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
-% submissions have lists of figures/tables at the end of the paper and that
-% figures/tables without any captions are placed on a page by themselves at
-% the end of the document. If needed, the draftcls IEEEtran class option or
-% \CLASSINPUTbaselinestretch interface can be used to increase the line
-% spacing as well. Be sure and use the nomarkers option of endfloat to
-% prevent endfloat from "marking" where the figures would have been placed
-% in the text. The two hack lines of code above are a slight modification of
-% that suggested by in the endfloat docs (section 8.4.1) to ensure that
-% the full captions always appear in the list of figures/tables - even if
-% the user used the short optional argument of \caption[]{}.
-% IEEE papers do not typically make use of \caption[]'s optional argument,
-% so this should not be an issue. A similar trick can be used to disable
-% captions of packages such as subfig.sty that lack options to turn off
-% the subcaptions:
-% For subfig.sty:
-% \let\MYorigsubfloat\subfloat
-% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
-% However, the above trick will not work if both optional arguments of
-% the \subfloat command are used. Furthermore, there needs to be a
-% description of each subfigure *somewhere* and endfloat does not add
-% subfigure captions to its list of figures. Thus, the best approach is to
-% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
-% and instead reference/explain all the subfigures within the main caption.
-% The latest version of endfloat.sty and its documentation can obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
-%
-% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
-% later in the document, say, to conditionally put the References on a 
-% page by themselves.
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/url/
-% Basically, \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
-% packages
 \usepackage[utf-8]{inputenc}
@@
 
+%% The lineno packages adds line numbers. Start line numbering with
+%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
+%% for the whole article with \linenumbers.
+%% \usepackage{lineno}
+
+\journal{Systems and Software}
+
+\newtheorem{definition}{Definition}
+\newcommand{\p}{\emph{P2P}}
+
 \begin{document}
@@
 
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
+\begin{frontmatter}
+
+%% Title, authors and addresses
+
+%% use the tnoteref command within \title for footnotes;
+%% use the tnotetext command for theassociated footnote;
+%% use the fnref command within \author or \address for footnotes;
+%% use the fntext command for theassociated footnote;
+%% use the corref command within \author for corresponding author footnotes;
+%% use the cortext command for theassociated footnote;
+%% use the ead command for the email address,
+%% and the form \ead[url] for the home page:
+%% \title{Title\tnoteref{label1}}
+%% \tnotetext[label1]{}
+%% \author{Name\corref{cor1}\fnref{label2}}
+%% \ead{email address}
+%% \ead[url]{home page}
+%% \fntext[label2]{}
+%% \cortext[cor1]{}
+%% \address{Address\fnref{label3}}
+%% \fntext[label3]{}
+
 \title{Alleviating the Topology Mismatch Problem in Distributed Overlay Networks}
-%
-%
-% author names and IEEE memberships
-% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
-% a structure at a ~ so this keeps an author's name from being broken across
-% two lines.
-% use \thanks{} to gain access to the first footnote area
-% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
-% was not built to handle multiple paragraphs
-%
+
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1,label2]{}
+%% \address[label1]{}
+%% \address[label2]{}
 
@@
         Mema Roussopoulos,
-        and~Alex Delis,~\IEEEmembership{Member,~IEEE}% <-this % stops a space
-\thanks{Vassilis Moustakas, Mema Roussopoulos and Alex Delis are 
-	with the Department of Informatics and Telecommunications, 
-	\emph{University of Athens},
-GR15784, Athens, Greece, e-mails: \emph{\{vsmoustakas,mema,ad\}@di.uoa.gr}}% <-this % stops a space
-\thanks{H\"useyin Akcan is with the Department of Software Engineering,
-\emph{Izmir University of Economics}, 35330 Balcova, Izmir, Turkey, email: \emph{huseyin.akcan@ieu.edu.tr}}% <-this % stops a space
-\thanks{Manuscript received April 30, 2015; revised April 30, 2015.}}
+        and~Alex Delis}
 
-% note the % following the last \IEEEmembership and also \thanks - 
-% these prevent an unwanted space from occurring between the last author name
-% and the end of the author line. i.e., if you had this:
-% 
-% \author{....lastname \thanks{...} \thanks{...} }
-%                     ^------------^------------^----Do not want these spaces!
-%
-% a space would be appended to the last name and could cause every name on that
-% line to be shifted left slightly. This is one of those "LaTeX things". For
-% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
-% "AB" then you have to do: "\textbf{A}\textbf{B}"
-% \thanks is no different in this regard, so shield the last } of each \thanks
-% that ends a line with a % and do not let a space in before the next \thanks.
-% Spaces after \IEEEmembership other than the last one are OK (and needed) as
-% you are supposed to have spaces between the names. For what it is worth,
-% this is a minor point as most people would not even notice if the said evil
-% space somehow managed to creep in.
+\address{
+\emph{vsmoustakas@gmail.com}
+\emph{huseyin.akcan@ieu.edu.tr}
+\emph{\{mema,ad\}@di.uoa.gr}
+}
 
-% The paper headers
-\markboth{IEEE COMMUNICATION SURVEYS \& TUTORIALS, VOL. XX, NO. YY, Z QUARTER YYYY}%
-{Moustakas \MakeLowercase{\textit{et al.}}: Alleviating the Topology Mismatch Problem in Distributed Overlay Networks}
-% The only time the second header will appear is for the odd numbered pages
-% after the title page when using the twoside option.
-% 
-% *** Note that you probably will NOT want to include the author's ***
-% *** name in the headers of peer review papers.                   ***
-% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
-% you desire.
+\begin{abstract}
+\input{abstract}
+\end{abstract}
 
-% If you want to put a publisher's ID mark on the page you can do it like
-% this:
-%\IEEEpubid{0000--0000/00\$00.00~\copyright~2014 IEEE}
-% Remember, if you use this you must call \IEEEpubidadjcol in the second
-% column for its text to clear the IEEEpubid mark.
+\begin{keyword}
+Distributed systems, peer-to-peer, overlay network, topology mismatch,topology awareness
 
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
+%% keywords here, in the form: keyword \sep keyword
 
-% make the title area
-\maketitle
+%% PACS codes here, in the form: \PACS code \sep code
 
-\input{abstract}
+%% MSC codes here, in the form: \MSC code \sep code
+%% or \MSC[2008] code \sep code (2000 is the default)
 
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle
+\end{keyword}
+
+\end{frontmatter}
+
+%% \linenumbers
+
+%% main text
 
@@
 
-%\section{Introduction}
-% The very first letter is a 2 line initial drop letter followed
-% by the rest of the first word in caps.
-% 
-% form to use if the first word consists of a single letter:
-% \IEEEPARstart{A}{demo} file is ....
-% 
-% form to use if you need the single drop letter followed by
-% normal text (unknown if ever used by IEEE):
-% \IEEEPARstart{A}{}demo file is ....
-% 
-% Some journals put the first two words in caps:
-% \IEEEPARstart{T}{his demo} file is ....
-% 
-% Here we have the typical use of a "T" for an initial drop letter
-% and "HIS" in caps to complete the first word.
-%\IEEEPARstart{T}{his} demo file is intended to serve as a ``starter file''
-%for IEEE journal papers produced under \LaTeX\ using
-%IEEEtran.cls version 1.8a and later.
-%% You must have at least 2 lines in the paragraph with the drop letter
-%% (should never be an issue)
-%I wish you the best of success.
-%
-%\hfill mds
-% 
-%\hfill September 17, 2014
-%
-%\subsection{Subsection Heading Here}
-%Subsection text here.
-%
-%% needed in second column of first page if using \IEEEpubid
-%%\IEEEpubidadjcol
-%
-%\subsubsection{Subsubsection Heading Here}
-%Subsubsection text here.
-
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure}
-
-% Note that IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command,
-% and the \label for the overall figure must come after \caption.
-% \hfil is used as a separator to get equal spacing.
-% Watch out that the combined width of all the subfigures on a 
-% line do not exceed the text width or a line break will occur.
-%
-%\begin{figure*}[!t]
-%\centering
-%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
-%\label{fig_second_case}}
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat[]), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
-% labels, the optional argument to \subfloat must be present. If a
-% subcaption is not desired, just leave its contents blank,
-% e.g., \subfloat[].
-
-% An example of a floating table. Note that, for IEEE style tables, the
-% \caption command should come BEFORE the table and, given that table
-% captions serve much like titles, are usually capitalized except for words
-% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
-% and up, which are usually not capitalized unless they are the first or
-% last word of the caption. Table text will default to \footnotesize as
-% IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that the IEEE does not put floats in the very first column
-% - or typically anywhere on the first page for that matter. Also,
-% in-text middle ("here") positioning is typically not used, but it
-% is allowed and encouraged for Computer Society conferences (but
-% not Computer Society journals). Most IEEE journals/conferences use
-% top floats exclusively. 
-% Note that, LaTeX2e, unlike IEEE journals/conferences, places
-% footnotes above bottom floats. This can be corrected via the
-% \fnbelowfloat command of the stfloats package.
-
-%\section{Conclusion}
-%The conclusion goes here.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% TODO: SOME INFORMATION ON ISP-ASSISTED TOPOLOGY AWARENESS
-%
-% IN CASE REVIEWERS POINT OUT THAT WE SHOULD ADD THOSE AS WELL.
-%
-%       ATTENTION: NEEDS REWORDING
-%
-% P4P [16],instead, aimed to provide a general framework for P2P
-%applications by developing ISP-application interfaces so that
-%application software can use the network layer information,
-%such as congestion status and network topology, for better
-%performance and resource utilization.
-
-%P4P [37] project attempts to address
-%the problem through custom trackers, both for ISPs and P2P
-%systems, using an interface based on a primal-dual decomposition
-%of an optimization problem. This interface design simplifies
-%the realization of traffic-engineering objectives from each parties’
-%perspective and ensures the extensibility of the approach. Through
-%simulation-based studies and limited experimental deployments,
-%these collaborative approaches have been shown to effectively
-%reduce network costs while minimally impacting application performance.
-%A clear advantage of these proposals is that they
-%allow ISPs to incorporate aggregated traffic policies in their tracker
-%recommendations (e.g., a particular traffic balance ratio between
-%peering providers). However, all of them require deployments
-%of oracles for each participating ISP and their effectiveness is
-%ultimately predicated on their adoption by P2P applications and a
-%trust relationship between P2P users and their ISPs.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \input{conclusion}
 
-% if have a single appendix:
-%\appendix[Proof of the Zonklar Equations]
-% or
-%\appendix  % for no appendix heading
-% do not use \section anymore after \appendix, only \section*
-% is possibly needed
+%% The Appendices part is started with the command \appendix;
+%% appendix sections are then done as normal sections
+%% \appendix
 
-% use appendices with more than one appendix
-% then use \section to start each appendix
-% you must declare a \section before using any
-% \subsection or using \label (\appendices by itself
-% starts a section numbered zero.)
-%
+%% \section{}
+%% \label{}
 
-%\appendices
-%\section{Proof of the First Zonklar Equation}
-%Appendix one text goes here.
-%
-%% you can choose not to have a title for an appendix
-%% if you want by leaving the argument blank
-%\section{}
-%Appendix two text goes here.
+%% If you have bibdatabase file and want bibtex to generate the
+%% bibitems, please use
+%%
+\bibliographystyle{elsarticle-harv} 
+\bibliography{mard-survey}
 
-% use section* for acknowledgment
-%\section*{Acknowledgment}
+%% else use the following coding to input the bibitems directly in the
+%% TeX file.
 
-%The authors would like to thank...
+%%\begin{thebibliography}{00}
 
-% Can use something like this to put references on a page
-% by themselves when using endfloat and the captionsoff option.
-\ifCLASSOPTIONcaptionsoff
-  \newpage
-\fi
+%% \bibitem[Author(year)]{label}
+%% Text of bibliographic item
 
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
+%%\bibitem[ ()]{}
 
-% references section
+%%\end{thebibliography}
+\end{document}
 
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-\bibliography{mard-survey}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-%\begin{thebibliography}{1}
-%
-%\bibitem{IEEEhowto:kopka}
-%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-%
-%\end{thebibliography}
-
-% biography section
-% 
-% If you have an EPS/PDF photo (graphicx package needed) extra braces are
-% needed around the contents of the optional argument to biography to prevent
-% the LaTeX parser from getting confused when it sees the complicated
-% \includegraphics command within an optional argument. (You could create
-% your own custom macro containing the \includegraphics command to make things
-% simpler here.)
-%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
-% or if you just want to reserve a space for a photo:
-
-\begin{IEEEbiography}{Vassilis Moustakas}
-Biography text here.
-\end{IEEEbiography}
-
-% insert where needed to balance the two columns on the last page with
-% biographies
-%\newpage
-
-\begin{IEEEbiography}{H\"useyin Akcan}
-Biography text here.
-\end{IEEEbiography}
-
-\begin{IEEEbiography}{Mema Roussopoulos}
-Biography text here.
-\end{IEEEbiography}
-
-\begin{IEEEbiography}{Alex Delis}
-Biography text here.
-\end{IEEEbiography}
-
-% if you will not have a photo at all:
-%\begin{IEEEbiographynophoto}{John Doe}
-%Biography text here.
-%\end{IEEEbiographynophoto}
-
-% You can push biographies down or up by placing
-% a \vfill before or after them. The appropriate
-% use of \vfill depends on what kind of text is
-% on the last page and whether or not the columns
-% are being equalized.
-
-%\vfill
-
-% Can be used to pull up biographies so that the bottom of the last one
-% is flush with the other column.
-%\enlargethispage{-5in}
-
-% that's all folks
-\end{document}
+\endinput
+%%
+%% End of file `elsarticle-template-harv.tex'.

### 2015-07-26 cover-letter/comst/mard-coverletter-comst-ieee.tex
2015.07.23 (Alex) Renaming the paper to add the word "survey" in order to better fit JSS. Some minor fixes to author names and affiliations. Adding letter to JSS Editor-In-Chief and submission receipt + other resources.

%% start of file `template.tex'.
%% Copyright 2006-2013 Xavier Danaux (xdanaux@gmail.com).
%
% This work may be distributed and/or modified under the
% conditions of the LaTeX Project Public License version 1.3c,
% available at http://www.latex-project.org/lppl/.

\documentclass[12pt,a4paper,sans]{moderncv}        % possible options include font size ('10pt', '11pt' and '12pt'), paper size ('a4paper', 'letterpaper', 'a5paper', 'legalpaper', 'executivepaper' and 'landscape') and font family ('sans' and 'roman')

% moderncv themes
\moderncvstyle{classic}                            % style options are 'casual' (default), 'classic', 'oldstyle' and 'banking'
\moderncvcolor{blue}                              % color options 'blue' (default), 'orange', 'green', 'red', 'purple', 'grey' and 'black'
%\renewcommand{\familydefault}{\sfdefault}         % to set the default font; use '\sfdefault' for the default sans serif font, '\rmdefault' for the default roman one, or any tex font name
%\nopagenumbers{}                                  % uncomment to suppress automatic page numbering for CVs longer than one page

% character encoding
\usepackage[utf8]{inputenc}                       % if you are not using xelatex ou lualatex, replace by the encoding you are using
%\usepackage{CJKutf8}                              % if you need to use CJK to typeset your resume in Chinese, Japanese or Korean

% adjust the page margins
\usepackage[scale=0.75]{geometry}
%\setlength{\hintscolumnwidth}{3cm}                % if you want to change the width of the column with the dates
%\setlength{\makecvtitlenamewidth}{10cm}           % for the 'classic' style, if you want to force the width allocated to your name and avoid line breaks. be careful though, the length is normally calculated to avoid any overlap with your personal info; use this at your own typographical risks...

% personal data
\name{Vassilis}{Moustakas}
\title{Software Development Engineer, M.Sc.}                               % optional, remove / comment the line if not wanted
\address{SDE at Amazon.com}{Marcel-Breuer-Str. 12}{80807 Munich, Germany}% optional, remove / comment the line if not wanted; the "postcode city" and and "country" arguments can be omitted or provided empty
\phone[mobile]{+49 (0) 176 55194610}                   % optional, remove / comment the line if not wanted
\phone[fixed]{+49 (0) 89 20097833}                    % optional, remove / comment the line if not wanted
%\phone[fax]{+3~(456)~789~012}                      % optional, remove / comment the line if not wanted
\email{vsmoustakas@gmail.com}                      % optional, remove / comment the line if not wanted
\homepage{de.linkedin.com/in/vsmoustakas}          % optional, remove / comment the line if not wanted
%\extrainfo{additional information}                 % optional, remove / comment the line if not wanted
%\photo[64pt][0.4pt]{picture}                       % optional, remove / comment the line if not wanted; '64pt' is the height the picture must be resized to, 0.4pt is the thickness of the frame around it (put it to 0pt for no frame) and 'picture' is the name of the picture file
%\quote{Some quote}                                 % optional, remove / comment the line if not wanted

% to show numerical labels in the bibliography (default is to show no labels); only useful if you make citations in your resume
%\makeatletter
%\renewcommand*{\bibliographyitemlabel}{\@biblabel{\arabic{enumiv}}}
%\makeatother
%\renewcommand*{\bibliographyitemlabel}{[\arabic{enumiv}]}% CONSIDER REPLACING THE ABOVE BY THIS

% bibliography with mutiple entries
%\usepackage{multibib}
%\newcites{book,misc}{{Books},{Others}}
%----------------------------------------------------------------------------------
%            content
%----------------------------------------------------------------------------------
\begin{document}
%-----       letter       ---------------------------------------------------------
% recipient data
\recipient{Ekram Hossain}{Professor/Editor-In-Chief IEEE COMST\\Department of Electrical and Computer Engineering\\University of Manitoba\\75A Chancellor's Circle\\Winnipeg, MB R3T 5V6 CANADA}
\date{June 28, 2015}
\opening{Dear Professor Hossain,}
\closing{Sincerely,}
%\enclosure[Attached]{curriculum vit\ae{}}          % use an optional argument to use a string other than "Enclosure", or redefine \enclname
\makelettertitle

We are submitting a survey paper entitled ``Alleviating the Topology Mismatch Problem in Distributed Overlay Networks'' for publication in the Journal IEEE Communications Surveys \& Tutorials (IEEE COMST). The paper is co-authored with Assoc. Prof. H\"useyin Akcan of Izmir University of Economics in Izmir, Turkey, Prof. Mema Roussopoulos of the University of Athens in Athens, Greece and Prof. Alex Delis of the University of Athens in Athens, Greece.

Thank you very much for your kind consideration and your effort to have our work refereed in a timely fashion. We look forward to hearing from you.

\makeletterclosing

\end{document}

%% end of file `template.tex'.

### 2015-07-26 mard-survey.tex
2015.07.23 (Alex) Renaming the paper to add the word "survey" in order to better fit JSS. Some minor fixes to author names and affiliations. Adding letter to JSS Editor-In-Chief and submission receipt + other resources.

@@
 
-\title{Alleviating the Topology Mismatch Problem in Distributed Overlay Networks}
+\title{Alleviating the Topology Mismatch Problem in Distributed Overlay Networks: A Survey}
 
@@
 
-\author{Vassilis Moustakas,
-        H\"useyin Akcan,
-        Mema Roussopoulos,
-        and~Alex Delis}
+% \author{Vassilis Moustakas,
+        % H\"useyin Akcan,
+        % Mema Roussopoulos,
+        % and~Alex Delis}
+% 
+% \address{
+% Amazon Services\emph{vsmoustakas@gmail.com,}
+% Izmir Economics Univ.  \emph{huseyin.akcan@ieu.edu.tr,}
+% Univ. of Athens \emph{\{mema,ad\}@di.uoa.gr}
+% }
 
-\address{
-\emph{vsmoustakas@gmail.com}
-\emph{huseyin.akcan@ieu.edu.tr}
-\emph{\{mema,ad\}@di.uoa.gr}
-}
+\author[rvt0]{Vassilis Moustakas} % \fnref{fn1}}
+\ead{vsmoustakas@gmail.com}
+\author[focal]{H\"useyin Akcan} % \fnref{fn2}}
+\ead{huseyin.akcan@ieu.edu.tr}
+\author[rvt]{Mema Roussopoulos} % \fnref{fn3}}
+\ead{mema@di.uoa.gr}
+\author[rvt]{Alex Delis\corref{cor1}} % \fnref{fn3}}
+\ead{ad@di.uoa.gr}
+\cortext[cor1]{Corresponding author}
+% \cortext[cor2]{Principal corresponding author}
+% \fntext[fn1]{Department of Informatics and Telecommunications, University of Athens, Athens, 15703, Greece.}
+% \fntext[fn2]{Department of Software Engineering, Izmir University of Economics, 35330 Balcova, Izmir, Turkey}
+% \fntext[fn3]{Yet another author footnote. Indeed, you can have
+% any number of author footnotes.}
+\address[rvt0]{Amazon Development Center Germany GmbH, Marcel-Breuer Str. 12, 80807 Munich, Germany}
+\address[focal]{Department of Software Engineering, Izmir University of Economics, 35330 Balcova, Izmir, Turkey}
+\address[rvt]{Department of Informatics and Telecommunications, University of Athens, Athens, 15703, Greece}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2015-09-13 acknowledge.tex
2015.09.20 (Vassilis) Added acknowledgements section. Updated metric definitions. Introduced timeline figures to capture termoral and trending dimensions of surveyed material. Minor fixes and typos.

\section{Acknowledgements}
\label{section:acknowledge}

We would also like to express our gratitude to all the anonymous reviewers who helped us improve the presentation of our work with their insightful comments. This research has been partially supported by \emph{EC FP7 Workprogramme (2007-2013)} under grant agreement n. 283610 \emph{``European Scalable EarthScience Service Environment (EarthServer)''} and \emph{EC H2020} under grant agreement n. 654367 \emph{``Agile Analytics on Big Data Cubes (EarthServer 2)''}.

This  work  has  been  partially  supported  by \emph{SocWeb}, \emph{iMarine}, and \emph{Sucre} \emph{FP7 EU projects}.

### 2015-09-13 background.tex
2015.09.20 (Vassilis) Added acknowledgements section. Updated metric definitions. Introduced timeline figures to capture termoral and trending dimensions of surveyed material. Minor fixes and typos.

@@
 
-{\sl Napster} was the most successful incarnation of the
+\emph{Napster} was the most successful incarnation of the
 centralized approach in file-sharing~\cite{napster}.
@@
 pointers to the actual content.
-Thus, by centralizing search while distributing downloads,  {\sl Napster}
+Thus, by centralizing search while distributing downloads,  \emph{Napster}
 achieved a highly efficient design that, at its height, attracted $26$ million
-users sharing approximately $80$ million songs\cite{jmm_naptopusage_2001}.
+users sharing approximately $80$ million songs~\cite{jmm_naptopusage_2001}.
 %%
 The central indexing component played a vital role in the demise of 
-{\sl Napter}, as the \emph{Recording
+\emph{Napter}, as the \emph{Recording
 Industry Association of America} successfully argued in court that this software
@@
 In its first realization,
-{\sl Gnutella}(\emph{v.0.4}) became a well-known example of the fully
+\emph{Gnutella}(\emph{v.0.4}) became a well-known example of the fully
 decentralized unstructured approach.
@@
 \centering
-\subfigure[Centralized ({\sl SETI@Home} or {\sl Napster}). Peers query the centralized
+\subfigure[Centralized (\emph{SETI@Home} or \emph{Napster}). Peers query the centralized
 entity (dotted lines) and then establish the connection to the data source.] {
@@
 }\qquad\qquad
-\subfigure[Fully unstructured ({\sl Gnutella} \emph{v.0.4}). Data is distributed randomly,
+\subfigure[Fully unstructured (\emph{Gnutella} \emph{v.0.4}). Data is distributed randomly,
 the topology is also random and all participants are equal.] {
@@
 }\qquad\qquad
-\subfigure[Hierarchical unstructured ({\sl Gnutella} \emph{v.0.6} or {\sl FastTrack}). Certain
+\subfigure[Hierarchical unstructured (\emph{Gnutella} \emph{v.0.6} or \emph{FastTrack}). Certain
 peers (bigger circles) shoulder specific responsibilities (e.g., data
@@
 }\qquad%\qquad
-\subfigure[Structured ({\sl Chord}~\cite{stoica_chord_2001} and {\sl Kademlia}~\cite{maymounkov_kademlia_2002}). 
+\subfigure[Structured (\emph{Chord}~\cite{stoica_chord_2001} and \emph{Kademlia}~\cite{maymounkov_kademlia_2002}). 
 Peers and data are placed in the overlay in a deterministic manner.] {
@@
 Examples of such hierarchical unstructured systems include
-{\sl Gnutella}(\emph{v.0.6})/{\sl Limewire}~\cite{gnutella}, 
-{\sl KaZaA}~\cite{kazaa} and {\sl Skype}~\cite{skype}. 
+\emph{Gnutella}(\emph{v.0.6})/\emph{Limewire}~\cite{gnutella}, 
+\emph{KaZaA}~\cite{kazaa} and \emph{Skype}~\cite{skype}. 
 Although the hierarchical approach relieves network pressure as
@@
 Several popular file-sharing \p\ applications have been based on \emph{DHT}s
-including {\sl Overnet}/{\sl eDonkey}~\cite{overnet}, 
-{\sl Kademlia}/{\sl Kad Network}~\cite{maymounkov_kademlia_2002},  and 
-some flavors of {\sl BitTorrent}~\cite{c_bittorrent_2003}.  
+including \emph{Overnet}/\emph{eDonkey}~\cite{overnet}, 
+\emph{Kademlia}/\emph{Kad Network}~\cite{maymounkov_kademlia_2002},  and 
+some flavors of \emph{BitTorrent}~\cite{c_bittorrent_2003}.  
 Moreover, \emph{DHT}s sparked
@@
 topology mismatch problem.
-{\sl Gnutella}, for example, is
+\emph{Gnutella}, for example, is
 considered far from scalable~\cite{ritter_gnucantscale_2001} for every peer
@@
 are less than $7$ hops away from each other, 
-a \emph{flooding-based} {\sl Gnutella}-like algorithm 
+a \emph{flooding-based} \emph{Gnutella}-like algorithm 
 can generate $330$TB/month traffic in a $50,000$ node network. 
@@
 interpretation of the above literal values.
-\emph{Efficiency} means how well the
-algorithm performs its search and topology adaptation functions. Our assessment
-also considers any negative effects that might come along with the application
-of the algorithm on intrinsic characteristics of the system (i.e., search scope
-coverage, download performance etc.).
 
-For the \emph{overhead} criterion, we evaluate the cost
-of computing and communicating peer join, message forwarding, 
-and overlay topology maintenance. For instance, the overhead that an algorithm,
-which needs global knowledge of the overlay topology, imposes on the system is
-greater compared to a protocol that only relies on peers exchanging routing
-tables within a 2-hop radius from each other.
+We use the term \emph{efficiency} to gauge the competence of a given
+algorithm to match the overlay to the underlying network. Minimization
+of duplicate messages traveling the network, quick object location and
+preservation of search scope as well as ability and
+effectiveness to adapt in high churn environments are some of the
+characteristics we consider when grading an approach on this criterion.
+
+With \emph{overhead}, we, coarsely, quantify the additional cost a given approach
+imposes to the underlying resources when applied to
+\begin{inparaenum}[\itshape i\upshape)]
+  \item create,
+  \item adapt, and
+  \item maintain
+\end{inparaenum}
+the overlay network while preserving its topology awareness. Computing and
+communicating peer join and departure, message forwarding, updating routing tables
+are some of things that we take into account to perform our assessement. Note
+that this is the one criterion of the three we use that its literal values
+``low'' and ``high'' must be interpreted as ``better'' and ``worse'' respectivelly.
 
@@
 effectively support hundreds, tenths of thousands, or millions of nodes,
-respectively.
+respectively. One important consideration here is the extent of self-organization
+of a system's peers and thus the overall distributed nature of the overlay network.
+The least a network relies on global knowledge to perform it functions the smoother
+it scales.

### 2015-09-13 intro.tex
2015.09.20 (Vassilis) Added acknowledgements section. Updated metric definitions. Introduced timeline figures to capture termoral and trending dimensions of surveyed material. Minor fixes and typos.

@@
 file-sharing and Internet telephony.
-The {\sl Napster} file-sharing system was widely acknowledged as the 
+The \emph{Napster} file-sharing system was widely acknowledged as the 
 ``fastest growing Internet application ever'' in $2001$ when it topped
@@
 Other file-sharing applications followed suit, 
-including {\sl Gnutella} and {\sl Limewire} enjoying $3$ 
+including \emph{Gnutella} and \emph{Limewire} enjoying $3$ 
 million concurrently-connected peers, as well as 
-{\sl BitTorrent} connecting over $150$ million monthly users by January $2012$~\cite{bittorrentusers}.
-\p\ telephony saw explosive growth with the advent of {\sl Skype}.
+\emph{BitTorrent} connecting over $150$ million monthly users by January $2012$~\cite{bittorrentusers}.
+\p\ telephony saw explosive growth with the advent of \emph{Skype}.
 Since its introduction in $2003$,
-{\sl Skype} has become extremely popular with more than $650$ million users 
+\emph{Skype} has become extremely popular with more than $650$ million users 
 in $2011$~\cite{skypetotalusers} and an astonishing 
@@
 various aspects of the topology mismatch problem.
-This research area started to become active after  The mismatch problem is Interest around All interesting work around the topology We omitted centralized Centralized systems are not
-
 The main objective of this paper is to offer a comprehensive 

### 2015-09-13 structured.tex
2015.09.20 (Vassilis) Added acknowledgements section. Updated metric definitions. Introduced timeline figures to capture termoral and trending dimensions of surveyed material. Minor fixes and typos.

@@
 
-\emph{Global Soft-State} \cite{XTZ2003} builds a global view to help nodes
+\textbf{\emph{Global Soft-State}} \cite{XTZ2003} builds a global view to help nodes
 choose shorter routing paths. To generate the proximity information, it 
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{Global Soft-State} &
 % Unable to identify nodes that are close to routers but it can reduce routing
@@
 
-\emph{Mithos}~\cite{WR2003} ensures that neighboring nodes in the overlay are
+\textbf{\emph{Mithos}}~\cite{WR2003} ensures that neighboring nodes in the overlay are
 close also in terms of the underlying layer by integrating proximity layout
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{Mithos} &
 % It has very good results in reducing communication latency among peers but its
@@
 %% VM the structure is that it is organized as a tree hierarchy.
-\emph{LAPTOP} \cite{WLH2007} organizes the overlay 
+\textbf{\emph{LAPTOP}}~\cite{WLH2007} organizes the overlay 
 into a tree-based hierarchy so that 
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{LAPTOP} &
 % It minimizes routing length, but may suffer load imbalance since nodes higher
@@
 To optimize the underlying network use,
-\cite{KLKP2008} introduces proximity
+\textbf{\cite{KLKP2008}} introduces proximity
 controls over the \emph{vanilla Kademlia}~\cite{maymounkov_kademlia_2002} 
@@
 To enhance traffic locality, the use of 
-{\sl MaxMind GeoIP} database\footnote{\url{http://www.maxmind.com}} 
+\emph{MaxMind GeoIP} database\footnote{\url{http://www.maxmind.com}} 
 is suggested to form clusters of nearby peers.
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{\cite{KLKP2008}} &
 % Uses static information to cluster nodes which is not the most efficient
@@
 
-\emph{CHOP6} \cite{MT2007} is a Chord-variant DHT that exploits
-{\sl IPv6} address format as well as available \emph{RTT} information to 
+\textbf{\emph{CHOP6}} \cite{MT2007} is a Chord-variant DHT that exploits
+\emph{IPv6} address format as well as available \emph{RTT} information to 
 enable proximity estimation among peers during message routing.
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{CHOP6} &
 % Simple approach that exploits proximity but it is coarse grained and thus not
@@
 %% VM Added some extra lines and restructured the text
-\emph{T2MC}'s~\cite{SLCGZ2008} objective is to diminish the redundant multiple
+\textbf{\emph{T2MC}}'s~\cite{SLCGZ2008} objective is to diminish the redundant multiple
 messages that may unnecessarily cross autonomous system (\emph{AS}) boundaries.
@@
 the Internet routers to help solve a clustering problem. 
-The algorithm uses {\sl traceroute}-logs to detect inter-\emph{AS} boundaries 
+The algorithm uses \emph{traceroute}-logs to detect inter-\emph{AS} boundaries 
 and clusters peers to either nearby or 
@@
 traceroute and selects the routers with
-minimum and maximum latencies in the above {\sl traceroute} path 
+minimum and maximum latencies in the above \emph{traceroute} path 
 as initial centroids for the ``nearby'' and ``far-away'' sets respectively.
@@
 
-Even though {\sl traceroute}
+Even though \emph{traceroute}
 provides detailed information about the network structure, 
@@
 Administrators thus frequently
-disable {\sl traceroute} support on their routers and this 
+disable \emph{traceroute} support on their routers and this 
 undoubtly affects \emph{T2MC}'s operation.
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{T2MC} &
 % The efficiency of choosing close by peers is very good and reported almost
@@
 % \subsubsection{PChord}
-\emph{PChord} \cite{HLYW2005} is based on the \emph{Chord DHT} and incorporates
+\textbf{\emph{PChord}} \cite{HLYW2005} is based on the \emph{Chord DHT} and incorporates
 proximity in its routing mechanism. 
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{PChord} &
 % Small decrease in needed overlay hops per object publishing but rigid in
@@
 % \subsubsection{AChord}
-\emph{AChord} \cite{DK2006} exploits IPv6's \emph{anycast} mechanism to
+\textbf{\emph{AChord}} \cite{DK2006} exploits IPv6's \emph{anycast} mechanism to
 relieve the protocol from complex joining procedures,
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{AChord} &
 % good routing efficiency but no adaptation to high churn environments.
@@
 % \subsubsection{Chord6}
-\emph{Chord6}~\cite{XZHL2005} is another \emph{Chord}-variant that
+\textbf{\emph{Chord6}}~\cite{XZHL2005} is another \emph{Chord}-variant that
 uses IPv6's  hierarchical features to create 
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{Chord6} &
 % Can reduce inter-domain average path length but even though nodes in the same
@@
 A proximity neighbor selection scheme functioning atop 
-the \emph{Chord DHT} is introduced in~\cite{DLTZZ2006}.
+the \emph{Chord DHT} is introduced by \cite{DLTZZ2006} (\textbf{\emph{DHT-PNS}}).
 Its key objective was to group physically-close nodes as 
@@
 %%
-\emph{Chord DHT} fares as follows in reference to the stated three criteria:
+\emph{DHT-PNS} fares as follows in reference to the stated three criteria:
 \begin{center}
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{DHT-PNS} &
 %
@@
 % \subsubsection{Quasi-Chord}
-\cite{SZ2008} proposes \emph{Quasi-Chord}, a \emph{Chord}-based network that
+\cite{SZ2008} propose \textbf{\emph{Quasi-Chord}}, a \emph{Chord}-based network that
 constructs its overlay network while attempting to match the underlying 
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{Quasi-Chord} &
 % The algorithm shows to reduce end-to-end delay as well as not produce
@@
 However, such methods suffer by imprecisions often due to network overload.
-IP-based clustering or \emph{IPBC}~\cite{KM2007} uses IP address prefixes
+\textbf{\emph{IP-based clustering}} or \textbf{\emph{IPBC}}~\cite{KM2007} uses IP address prefixes
 ($16$~bits for IPv4) to detect proximity.
@@
 in the DHT. 
-Newly joining nodes with the same {\sl IP} prefix 
+Newly joining nodes with the same \emph{IP} prefix 
 can, thus, query the \emph{DHT} and readily identify all 
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{IPBC} &
 % The efficiency of proximity detection through the prefix matching technique
@@
 % \subsubsection{DynaMO}
-\emph{DynaMO}~\cite{WZS2004} creates an overlay that 
+\textbf{\emph{DynaMO}}~\cite{WZS2004} creates an overlay that 
 not only considers physical proximity amongst peers but also
@@
 To identify this dominating step, two techniques are used: 
-\emph{Random LandMarking (RLM)} and 
-\emph{Closest Neighbor Prefix Assignment (CNPA)}.
+\textbf{\emph{Random LandMarking (RLM)}} and 
+\textbf{\emph{Closest Neighbor Prefix Assignment (CNPA)}}.
 In \emph{RLM}, a set of peers are chosen to be responsible 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\emph{3D Overlay (3DO)}~\cite{AOS2014} seeks to address the 
+\textbf{\emph{3D Overlay (3DO)}}~\cite{AOS2014} seeks to address the 
 network mismatch problem  for peers operating in mobile networks.
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{3DO} &
 % 
@@
 
-\emph{SAT-Match}~\cite{RGJZ2004} defines an iterative process that 
+\textbf{\emph{SAT-Match}}~\cite{RGJZ2004} defines an iterative process that 
 adaptively changes the overlay network to ease the topology mismatch problem. 
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{SAT-Match} &
 % SAT-Match seems to outperform landmark binning in the sense of providing a
@@
 %%
-%%COM Added the following to mention the hybrid/combo methodologies
+Trends for structured systems depicted in Figure~\ref{structured:timeline}
+show the same lack of paradigm shift we identified for the unstructured
+approaches (check Figure~\ref{unstructured:timeline}). Leveraging proximity
+information for selecting the next hop or to construct and maintain
+routing tables as a whole, seem to be the elements more used in the
+literature throught the years that this area is researched.
+%%%
 Similarly to unstructured approaches, most of the methods
@@
 while \emph{LAPTOP} combines geographic--layout with
-replication strategies to refine its results.
-%%%
-The heuristic nature of the suggested solutions opens them up 
-to trade-offs which, in most cases, are 
+replication strategies to refine its results. What looks a bit different
+compared the trends in unstructred systems is the fact that, here, researchers
+seem to juggle with more methods at the same time. This, and the
+fact that these suggested solutions are also heuristic in nature, opens them up 
+to even harder trade-offs which, in most cases, are 
 highly affected by the application domain and the implementation itself.
@@
 utilized proximity and geographic aspects.
-Table~\ref{structured:table} offers a comparative view
-of the algorithms and
 Table~\ref{structured:table} outlines key properties for all surveyed approach 
@@
 topology adaptation, landmarking, and caching/replication.
-Finally, Figures~\ref{structured:plot:efficiency}, \ref{structured:plot:overhead}
+Figures~\ref{structured:plot:efficiency}, \ref{structured:plot:overhead}
 and \ref{structured:plot:scalability} help with the pictorial comparison
 of the algorithms' behavior with respect to the three 
-criteria we defined in subsection~\ref{background:motivation}.
+criteria we defined in subsection~\ref{background:motivation}. Finally,
+we embed algorithms and characteristics on the temporal axis in
+Figure~\ref{structured:timeline} to help depict emerging trends during
+all these years of research.
 
@@
 %%
+\input{structured-timeline}
+%%

### 2015-09-13 unstructured.tex
2015.07.19 (Vassilis) Reformatting for Elsevier publication targeting the Journal of Systems and Software (JSS).
2015.09.20 (Vassilis) Added acknowledgements section. Updated metric definitions. Introduced timeline figures to capture termoral and trending dimensions of surveyed material. Minor fixes and typos.

@@
 
-To improve over {\sl Gnutella}'s ``blind flooding'' approach,
+To improve over \emph{Gnutella}'s ``blind flooding'' approach,
 \cite{YG-M2002} proposed a practical and easy to implement solution 
@@
 
-In \emph{iterative deepening}, 
+In \textbf{\emph{iterative deepening}},
 %(see Figure~\ref{figure:dbfs:iterdeep})
@@
 
-The \emph{directed BFS}
+The \textbf{\emph{directed BFS}}
 tries to avoid the aforementioned delay by forwarding 
@@
 selected heuristics can direct the search to the right path.
-%
-In \emph{local indices},
+
+In \textbf{\emph{local indices}},
 each node indexes data hosted by neighboring nodes within a radius of $r$ hops
@@
 may introduce significant overhead should the radius is kept broad.
-%%
+
 Below, we outline how the techniques in~\cite{YG-M2002}
@@
 % \paragraph*{ \bf Delay Aware P2P System}
-A \emph{delay--aware} \p\ system termed {\sl DAPS} 
+A \emph{delay--aware} \p\ system termed \textbf{\emph{DAPS}} 
 was introduced in~\cite{ZL2005}.
-{\sl DAPS}  seeks to attain reduced look-up times by dividing 
+\emph{DAPS}  seeks to attain reduced look-up times by dividing 
 peer routing tables into several sectors of increasing delay. 
@@
 nodes whose expected delay is less than or equal to the indicated boundary. 
-{\sl DAPS} primarily focuses on user ``experience'' and deploys an
+\emph{DAPS} primarily focuses on user ``experience'' and deploys an
 end-to-end delay monitoring mechanism that may enable
@@
 %%
-In terms of the three performance criteria, {\sl DAPS} stands as follows:
-\begin{center}
-{\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+In terms of the three performance criteria, \emph{DAPS} stands as follows:
+\begin{center}
+{\footnotesize
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{DAPS} &
 % Does not check for close by nodes, it just sorts the contents of a node's
@@
 % \subsubsection{Gia}
-The key objective of the {\sl Gia} system~\cite{CRBLS2003} is 
+The key objective of the \textbf{\emph{Gia}} system~\cite{CRBLS2003} is 
 to help alleviate the scalability omnipresent
 in unstructured \p\ file-sharing systems.
-At first, {\sl Gia} replaced {\sl Gnutella}'s blind flooding 
+At first, \emph{Gia} replaced \emph{Gnutella}'s blind flooding 
 with random walks.
@@
 the success rate of the query in question.
-To overcome this limitation, {\sl Gia} introduces
+To overcome this limitation, \emph{Gia} introduces
 a token-based flow control mechanism 
@@
 support for \emph{one--hop replication} of pointers to content.
-{\sl Gia} also acknowledges the heterogeneity in peer bandwidth,
+\emph{Gia} also acknowledges the heterogeneity in peer bandwidth,
 processing power, disk speed, etc., of \p\ nodes and uses this
@@
 By using a topology adaptation algorithm, 
-{\sl Gia} places low capacity nodes within 
+\emph{Gia} places low capacity nodes within 
 short proximity to peers with high performance features.
@@
 or drop superseded ones to improve its satisfaction.
-Despite the fact that {\sl Gia}'s topology adaptation algorithm 
+Despite the fact that \emph{Gia}'s topology adaptation algorithm 
 improves the network's
@@
 \cite{MBL2006}.
-%
-In terms of the three stated criteria, {\sl Gia}'s approach is as follows:
-\begin{center}
-{\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+
+In terms of the three stated criteria, \emph{Gia}'s approach is as follows:
+\begin{center}
+{\footnotesize
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{Gia} &
 % Reduces the search scope.
@@
 
-\emph{Location-aware Topology Matching (LTM)} \cite{LLXNZ2004}
+\textbf{\emph{Location-aware Topology Matching (LTM)}} \cite{LLXNZ2004}
 seeks to optimize an overlay \p\ structure based on the physical topology.
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{LTM} &
 % Authors claim 75% reduction on traffic cost and about 65% reduction on query
@@
 \cite{ZKB2008} focuses on this exact deficiency of overlay networks and
-introduces the \emph{Distributed Cycle Minimization Protocol
-(DCMP)}, a dynamic, fully distributed method that removes cycles;
+introduces the \textbf{\emph{Distributed Cycle Minimization Protocol(DCMP)}},
+a dynamic, fully distributed method that removes cycles;
 this is accomplished without sacrificing
@@
 problem, while its performance in high-traffic and churn environments and its
-ability to retain important characteristics of a {\sl Gnutella}-like topology (i.e.,
+ability to retain important characteristics of a \emph{Gnutella}-like topology (i.e.,
 degree distribution, average peer distance, diameter e.t.c.) has been
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{DCMP} &
 % It returns more result than LTM according to experiments in
@@
 search is random. Concretely, three three approaches for replication allocation
-are proposed: \emph{uniform}, \emph{proportional}, and
-\emph{square-root replication}.
-
-In the uniform model, all objects are replicated without taking into account
+are proposed: \textbf{\emph{uniform}}, \textbf{\emph{proportional}}, and
+\textbf{\emph{square-root replication}}.
+
+In the \emph{uniform model}, all objects are replicated without taking into account
 the query distribution,
-while in the proportional, objects are replicated analogously to their query
+while in the \emph{proportional}, objects are replicated analogously to their query
 rate, so that frequently queried items can be found more often. Experimental
@@
 than in uniform and proportional models. Finally, with the
-square-root model they attempt to find the golden ratio in ``between'', by
+\emph{square-root model} they attempt to find the golden ratio in ``between'', by
 suggesting to set the number of object
@@
 that using it along side square-root allocation results in performance similar to
-{\sl Gnutella}'s query flooding method but incurs up to two
+\emph{Gnutella}'s query flooding method but incurs up to two
 orders of magnitude less network traffic. Also checks three strategies for
@@
 low \\
-\emph{Square Root Replication Allocation} &
+\emph{Square-Root Replication} &
 % All (strictly) in-between strategies are (strictly) better than Uniform and
@@
 
-{\sl Narada}~\cite{CRZ2000,CRSZ2001,CRSZ2002} is a generic
+\textbf{\emph{Narada}}~\cite{CRZ2000,CRSZ2001,CRSZ2002} is a generic
 protocol for creating self-adapting overlay networks capable of 
@@
 Although IP--multicast would present a choice in implementing
-{\sl Narada}, it is in general considered that it violates the 
+\emph{Narada}, it is in general considered that it violates the 
 stateless design of the current Internet.
-Despite the fact the {\sl Narada} was not designed as \p\ system per se,
+Despite the fact the \emph{Narada} was not designed as \p\ system per se,
 it was the first (along with \emph{Scattercast} \cite{C2000}) to consider the
@@
 the overall resource usage as well as the relative delay among end-systems. 
-Unfortunately, the main limitation for {\sl Narada} is that 
+Unfortunately, the main limitation for \emph{Narada} is that 
 although it works reasonably well for small groups, 
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{Narada} &
 %
@@
 
-Along with {\sl Narada}, the \emph{Adaptive Overlay Topology 
-Optimization (AOTO)} \cite{LZXN2003} is one of the first attempts to address the
+Along with \emph{Narada}, the \textbf{\emph{Adaptive Overlay Topology 
+Optimization (AOTO)}} \cite{LZXN2003} is one of the first attempts to address the
 topology mismatch problem. \emph{AOTO} is a distributed algorithm that seeks to
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{AOTO} &
 %
@@
 % \subsubsection{Adaptive Connection Establishment (ACE)}
-The \emph{Adaptive Connection Establishment (ACE)} 
+The \textbf{\emph{Adaptive Connection Establishment (ACE)}} 
 approach~\cite{LZXN2004} uses the network delay as a metric 
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{ACE} &
 % compared to LTM it has slow convergence speed because of the information
@@
 
-The \emph{Scalable Bipartite Overlay (SBO)}~\cite{LXN2007} 
+The \textbf{\emph{Scalable Bipartite Overlay (SBO)}}~\cite{LXN2007} 
 reduces the overhead of creating
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{SBO} &
 % SBO reduces traffic cost between 85 and 90 percent
@@
 The distributed heuristic termed 
-\emph{Two-Hop-Away Neighbor Comparison and Selection (THANCS)}~\cite{LNXE2005,L2008}
+\textbf{\emph{Two-Hop-Away Neighbor Comparison and Selection (THANCS)}}~\cite{LNXE2005,L2008}
 attempts to minimize overlay hop costs.
@@
 piggybacked to normal
-{\sl Gnutella} messages. 
+\emph{Gnutella} messages. 
 When a node $P$ receives a query from $Q$, it constructs a
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{THANCS} &
 % 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The \emph{Hops Adaptive Neighbor Discovery (HAND)} algorithm~\cite{CLZHC2006}
+The \textbf{\emph{Hops Adaptive Neighbor Discovery (HAND)}} algorithm~\cite{CLZHC2006}
 uses a fully-distributed triple--hop adjustment strategy, applied to a network
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{HAND} &
 % similar traffic reduction rate savings to LTM
@@
 % \subsubsection{Adaptive Peer Selection (APS)}
-The \emph{Adaptive Peer Selection (APS)} approach uses 
+The \textbf{\emph{Adaptive Peer Selection (APS)}} approach uses 
 machine learning techniques to form peer selection strategies based on past
@@
 % introduced as a mechanism 
-to shape the policy for switching among the peers. The success of approach depends on
+to shape the policy for switching among the peers. The success of this approach depends on
 how fast the introduced peer selection functions. Admittedly, \emph{APS} is slow 
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{APS} &
 % machine learning the best way to adapt the overlay 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The \emph{Innocuous Topology Aware Overlay Construction (ITA)}
+The \textbf{\emph{Innocuous Topology Aware Overlay Construction (ITA)}}
 approach~\cite{PRFM2013} seeks to offer both an overlay formation approach and
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{ITA} &
 % low clustering with larger coverage of peers with the same number of messages
@@
 % \subsubsection{EGOIST}
-\emph{EGOIST}~\cite{SLLBBR2008} is a set of algorithms used 
+\textbf{\emph{EGOIST}}~\cite{SLLBBR2008} is a set of algorithms used 
 to construct and manage overlay networks. 
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{EGOIST} &
 % peer neighborhood selected to minimize, for example, average delay
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The main objective of the \emph{Biased Neighbor Selection (BNS)} 
+The main objective of the \textbf{\emph{Biased Neighbor Selection (BNS)}} 
 approach~\cite{BCCMSBZ2006} is to 
-strengthen {\sl BitTorrent}'s~\cite{c_bittorrent_2003} locality 
+strengthen \emph{BitTorrent}'s~\cite{c_bittorrent_2003} locality 
 by carefully choosing most of the neighbors of a peer to come from 
@@
 %%
-By and large, {\sl BitTorrent} deploys mechanisms that have proved
+By and large, \emph{BitTorrent} deploys mechanisms that have proved
 aggressive to \emph{ISP} networking and accounting as they function 
 in a ``without-borders'' approach. 
-The default {\sl BitTorrent} specification
+The default \emph{BitTorrent} specification
 designates $35$ connections for each peer regardless if such connections are
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{BNS} &
 % inter-ISP cost minimized while retaining the near optimal performance of
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\emph{Ono}~\cite{CB2008} is a protocol that helps contain
-{\sl BitTorrent} traffic within individual \emph{ISP}s
+\textbf{\emph{Ono}}~\cite{CB2008} is a protocol that helps contain
+\emph{BitTorrent} traffic within individual \emph{ISP}s
 and enhances the downloading rates by favoring connections
@@
 \emph{Ono}'s selection approach is 
-landmark-based and leverages existing {\sl CDN}-infrastructure 
-for peer distance estimation. {\sl CDN}s already use both static (i.e.,
+landmark-based and leverages existing \emph{CDN}-infrastructure 
+for peer distance estimation. \emph{CDN}s already use both static (i.e.,
 geographical, \emph{IP}-clustering) and dynamic (i.e., network delay measurement)
 information for their replica selection. Thus, the algorithm leverages the
-observation that peers that are redirected to a specific {\sl CDN} replica does not
+observation that peers that are redirected to a specific \emph{CDN} replica does not
 only mean that these are close by this replica but to each other as well.
@@
 a vector of \emph{$<$replica-server,ratio$>$} tuples, where \emph{ratio} is
-the percentage of times the {\sl CDN} redirects the peer 
+the percentage of times the \emph{CDN} redirects the peer 
 to the specific \emph{replica-server}.
@@
 The protocol's bootstrapping phase consists of the incoming peer performing
-{\sl DNS} lookups to {\sl CDN}-names to build its redirection information.
-%%
-Initially, the {\sl DNS} polling interval is set to $30$ seconds 
+\emph{DNS} lookups to \emph{CDN}-names to build its redirection information.
+%%
+Initially, the \emph{DNS} polling interval is set to $30$ seconds 
 and increases by $1$ minute every time redirection information 
-to the {\sl CDN} remains unchanged after the lookup.
-The {\sl DNS} polling interval is halved if pertinent 
+to the \emph{CDN} remains unchanged after the lookup.
+The \emph{DNS} polling interval is halved if pertinent 
 redirection information has changed.
@@
 is sufficiently fresh, 
-\emph{Ono} persists the map after the end of a {\sl BitTorrent} session.
+\emph{Ono} persists the map after the end of a \emph{BitTorrent} session.
 %%
@@
 inter-\emph{ISP} traffic costs and achieves near optimal performance 
-of {\sl BitTorrent} download~\cite{BCCMSBZ2006}.
+of \emph{BitTorrent} download~\cite{BCCMSBZ2006}.
 The \emph{ISP}-friendly design of \emph{Ono} however has been questioned when 
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{Ono} &
 % Authors claim minimization of inter-ISP traffic cost while enhancing the,
@@
 % \subsubsection{Locality-Awareness in BitTorrent-like P2P Applications}
-In~\cite{LCLX2009}, three techniques that 
-``inject'' locality awareness in {\sl BitTorrent}-like applications are discussed.
+In~\textbf{\cite{LCLX2009}}, three techniques that 
+``inject'' locality awareness in \emph{BitTorrent}-like applications are discussed.
 The first \emph{macroscopic-level} technique
@@
 The second technique, applied in an \emph{intermediate-level}, manipulates
-{\sl BitTorrent}'s chocking/unchocking mechanism. A peer unchokes its $4$
+\emph{BitTorrent}'s chocking/unchocking mechanism. A peer unchokes its $4$
 closest, in terms of 
@@
 the seeding state;  this  \emph{intermediate-level} approach favors
-least distance and is in contrast to the original {\sl BitTorrent} 
+least distance and is in contrast to the original \emph{BitTorrent} 
 implementation which favors uploading speed.
@@
 the next chunk to download based on a locality-first policy; this picks first
-the chunk that is closer, as opposed to {\sl BitTorrent}'s rare-first chunk
+the chunk that is closer, as opposed to \emph{BitTorrent}'s rare-first chunk
 picking policy.
@@
 Nevertheless, it does not do that well when compared with the
-random approach of the standard {\sl BitTorrent} protocol.
+random approach of the standard \emph{BitTorrent} protocol.
 
@@
 the locality of the basic functions of the
-{\sl BitTorrent} protocol. 
+\emph{BitTorrent} protocol. 
 Even though they do not restructure the network to map
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{\cite{LCLX2009}} &
 % minimization of inter-AS traffic but unfair among peers.
@@
 % \subsubsection{TopBT}
-The key objective of \emph{TopBT}~\cite{RTLCGZ2010} is to enhance
-proximity awareness in {\sl BitTorrent}
+The key objective of \textbf{\emph{TopBT}}~\cite{RTLCGZ2010} is to enhance
+proximity awareness in \emph{BitTorrent}
 without the need of additional infrastructure. 
@@
 The above \emph{TopBT}-metric is applied in various
-aspects of the original {\sl BitTorrent} protocol to ``inject'' topology
+aspects of the original \emph{BitTorrent} protocol to ``inject'' topology
 awareness for improved peer selection. More concretely, \emph{TopBT} modifies: 
@@
 \item the initial connection establishment, and 
-\item the unchoking mechanism of {\sl BitTorrent}.
+\item the unchoking mechanism of \emph{BitTorrent}.
 \end{inparaenum}
@@
 \emph{TopBT} offers more than $25\%$ traffic reduction and $15\%$ faster downloads
-compared to popular {\sl BitTorrent} implementations.
+compared to popular \emph{BitTorrent} implementations.
 In reference to the stated three criteria, \emph{TopBT} fares as follows:
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{TopBT} &
 % it takes into account both dl speed and network topology. The later though is
@@
 % \subsubsection{Underlying Topology-Aware Peer Selection (UTAPS)}
-The \emph{Underlying Topology-Aware Peer Selection (UTAPS)}~\cite{LCY2008}
+The \textbf{\emph{Underlying Topology-Aware Peer Selection (UTAPS)}}~\cite{LCY2008}
 presents an enhanced peer selection strategy. 
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{UTAPS} &
 % coarse grained approach of the network tomography doesn't provide a detailed
@@
 in those that are local, intra-\emph{ISP} and inter-\emph{ISP} is 
-presented in~\cite{QLZG2009}.
+presented by \textbf{\cite{QLZG2009}}.
 The classification is carried with the help of the core routers used 
@@
 For this, a newly arriving
-{\sl BitTorrent} peer, trace-routes the tracker and stores vectors containing
+\emph{BitTorrent} peer, trace-routes the tracker and stores vectors containing
 information like the \emph{IP} address and the hop number of a traced router, as well as the
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{\cite{QLZG2009}} &
 % It incorporates clustering of peers to inter- and intra ISP which is a
@@
 In~\cite{QCYCZ2007}, two 
-\emph{Peer-exchange Routing Optimization Protocols (PROPs)} 
+\textbf{\emph{Peer-exchange Routing Optimization Protocols (PROPs)}} 
 are introduced to help adjust the neighborhood graph of an overlay network
@@
 % \subsubsection{Distributed Domain Name Order (DDNO)}
-The \emph{Distributed Domain Name Order (DDNO)} approach~\cite{Z-YK2005} 
+The \textbf{\emph{Distributed Domain Name Order (DDNO)}} approach~\cite{Z-YK2005}
 uses domain names to detect topologically-close nodes.
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{DDNO} &
 % Clustering approaches reduce search scope (or at least increase search
@@
 % \subsubsection{Critical Topology-Aware Grouping (CTAG)}
-The \emph{Critical Topology-Aware Grouping (CTAG)} approach~\cite{ZL2006}
+The \textbf{\emph{Critical Topology-Aware Grouping (CTAG)}} approach~\cite{ZL2006}
 advocates a grouping strategy of peers based on the 
@@
 In its first phase, called \emph{bootstrapping grouping},
-the {\sl Gnutella} web caching mechanism
+the \emph{Gnutella} web caching mechanism
 is modified so that a newcomer may choose the closest, in terms of \emph{AM},
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{CTAG} &
 % The technique partitions the system reducing the search efficiency
@@
 
-\emph{Landmark Binning}~\cite{RHKS2002} partitions close-by nodes
+\textbf{\emph{Landmark Binning}}~\cite{RHKS2002} partitions close-by nodes
 into bins based on their distance from well known anchor nodes across the
@@
 Each permutation of the set of landmarks represents a specific bin. 
-Should there $m$ landmarks be adopted, $m!$  different bins potentially exist. 
+Should there be $m$ landmarks adopted, $m!$  different bins potentially exist. 
 Peers that end up with the same ordering, belong to the same bin, in the sense that
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{Landmark Binning} &
 % The technique is coarse grained thus doesn't achieve optimal results
@@
 % \subsubsection{mOverlay}
-The \emph{mOverlay}~\cite{ZZZSZ2004} approach addresses scalability issues
+The \textbf{\emph{mOverlay}}~\cite{ZZZSZ2004} approach addresses scalability issues
 that might arise when static landmark servers are in use.
@@
 {\footnotesize
-\begin{tabular}{ccc}
-\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
-\hline
+\begin{tabular}{rccc}
+\multicolumn{1}{r}{} &
+\multicolumn{1}{c}{\emph{Efficiency}} &
+\multicolumn{1}{c}{\emph{Overhead}} &
+\multicolumn{1}{c}{\emph{Scalability}}
+\\
+\cline{2-4}
+\emph{mOverlay} &
 % The technique is coarse grained thus doesn't achieve optimal results
@@
 forwarding optimization and caching.
-This combination inevitably leads to feature/performance trade-offs: for example, 
+%%
+The timeline in Figure~\ref{unstructured:timeline} further emphasises this trend
+in the broader picture. All characteristics we discuss,
+seem to have been identified and put into practice, early on in the literature.
+In that sense, no breakthrough
+or any revolutionary ideas came out lately but mostly people fine-tuned the
+heuristic parameters in pursue of the goldset that performs best. That doesn't
+render all that work less exciting or important since it led to better theoretical
+understanding of the problem and its difficulty. Moreover, the combination of
+multiple methodologies inevitably leads to interesting feature/performance trade-offs
+that researchers need to tackle in real world applications: for example, 
 the effectiveness of a caching/replication component can be undermined
 by a continuously adapting overlay that removes important links between peers.
+Topology adaptation and forwarding optimization seem to be constant areas of
+concentration for the researchers. This is of no surprise as all seem to
+understand the importance of the volatile nature of unstructured \p systems. 
+%%
 If a technique is to be evaluated for a specific application domain,
 its overall approach as well as its underlying used methodologies have to be considered.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 a brief presentation of the pros and cons in its implementation. 
-Last but not least, Figures~\ref{unstructured:plot:efficiency}, \ref{unstructured:plot:overhead}
+Figures~\ref{unstructured:plot:efficiency}, \ref{unstructured:plot:overhead}
 and \ref{unstructured:plot:scalability} pictorially compare the
@@
 and we collectively used them as a yardstick to qualitatively 
-evaluate the presented approaches.
+evaluate the presented approaches. Last but not least,
+Figure~\ref{unstructured:timeline} creates a timeline of the algorithms
+and their elements where we capture the temporal and trending dimensions
+of the surveyed material.
 
@@
 
-\input{unstructured-table}
-\input{unstructured-plot}
 %%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%
+\input{unstructured-table}
+%%
+\input{unstructured-plot}
+%%
+\input{unstructured-timeline}
+%%

### 2015-09-26 journals/jss/rebuttal/rebuttal.tex
2015.09.28 (Alex, Mema & Vassilis) Rephrasing and fixing typos. Adding rebutal letter and new letter to area editor. Including revised (r1) submission. Minor file reorganization and Makefile cleanup.

\documentclass[10pt,letterpaper]{article}

\usepackage{color}
\usepackage[margin=3cm, a4paper]{geometry}
\usepackage[hyphens]{url}
\usepackage{paralist}

\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt}

\begin{document}

\title{
	\textbf{\textsf{
	Revised Manuscript JSS-D-15-00624\\
	Authors' Responses to the Reviewers' Comments
	}}
}
\author{Vassilis Moustakas, H\"useyin Akcan, Mema Roussopoulos, and Alex Delis}
\maketitle

% Introduction
\section*{}
We would like to thank the reviewers for their comments and recommendations 
that helped us a lot while revising our manuscript.

In this new version of our survey, 
we have attempted to address all feasible comments 
and sufficiently answer all pertinent questions.
As a consequence, the 
manuscript has undergone a number 
of important changes and updates, which are:

\begin{enumerate}

\item 
Section 2.3 has been revised to offer more accurate definitions 
of the evaluation criteria ``efficiency'' and ``overhead'' that 
better distinguish the two and more accurately articulate
their distinct purpose in our evaluation.
\item 
We have added timeline figures for unstructured and structured families 
of approaches to depict
the temporal and trending characteristics of applied methodologies.
\item 
Sections 3.2 and 4.2 have been updated with discussion and comments 
that address aspects of the timeline figures.
\item 
Mini-tables have been updated to include 
algorithm names throughout the text. 
\item 
Figures of pages 19-21 and 32-34 have been updated to indicate 
the year next to the algorithm name.

\end{enumerate}

In the following two segments, we provide detailed replies to 
the comments and questions raised by the reviewers. 
Each individual comment has been copied \emph{as-is} from 
the original decision letter that was sent to us by 
the Editor--in-Chief followed by our corresponding response.

% Responses to the First Reviewer
\clearpage
\subsection*{\textbf{\textsf{Responses to the First Reviewer}}}

%%R1,C1
\subsubsection*{\textbf{\textsf{Reviewer's Comment\#1}}}
\noindent
\emph{I have a minor comment to make to improve the quality of the paper. Most of
the evaluation is carried out in a ``descriptive'' way, which is fine by me.
However it would be interesting to see some experimental evaluation (for at least
one of the metrics, e.g. efficient) using various data sets. I understand this
could be a lot of work, however if you could focus on three systems and test
efficiency/scalability, the results would be more convincing. These also will
support the claims you have made.}\\[6pt]

\noindent
\textbf{\textsf{Response}}:
%%
We would like to thank the reviewer for the thoughtful suggestion. 
It would have been indeed 
both an exciting and tremendous exercise 
to put all surveyed approaches under a \emph{common experimental setting} 
(to the extent possible) 
and test them against diverse workloads.

As the reviewer has kindly pointed out, this would have 
translated to an excessive amount of effort with rather uncertain results. 
More importantly, such an
effort would have yielded not a survey-type manuscript ---our 
stated objective--- but rather something much more resembling 
to an experiment-based monograph. 
Frankly, we never considered going into this level of evaluation 
work for the numerous details and diverse points of the algorithms 
would have posed insurmountable levels of challenge.

We believe that the suggestion of picking $3$ (three)
out of the $N$ approaches presented and evaluating
them quantitatively has important practical difficulties and 
does not guarantee added value to the surveyed material.

For example, what would be the methodology used to pick those
three systems? Does one choose three at random, 
the three most popular/widely deployed or
the ``best'' three w.r.t. the metric of interest, 
based on our prediction of which we think will
fare best? 
Does one choose the three that are simplest to implement?

There are reasons for and against following each of the 
above selection approaches and this apparently non-systematic approach makes
it unclear whether presenting results for just three of 
the surveyed approaches would render much added value
to our survey manuscript.

Enabling comparison through developing functional prototypes and subsequent 
quantitative evaluation would have been truly an undertaking of 
enormous dimensions. 
Not only do we lack pertinent resources, but also 
performing a fair and complete comparison would be arguably 
beyond the scope of this survey.

Although we have clearly given the reviewer's comment
the due consideration, for all the above reasons, it would have been 
unfair for the manuscript and its carefully crafted balance in 
presenting a trove of approaches to venture in such an experimental  direction. 
We do appreciate the comment however and we are thankful for the positive 
overall evaluation of our survey.

% Responses to the Third Reviewer
\clearpage
\subsection*{\textbf{\textsf{Responses to the Third Reviewer}}}

%% R3.C1
\subsubsection*{\textbf{\textsf{Reviewer's Comment\#1}}}
\noindent
\emph{I do recommend changing terminologies to highlight the differences between
``efficiency'' and ``overhead''. Maybe instead you can use
``Topology adaptation cost'' instead of efficiency, and break overhead into both
``Topology maintenance cost'' and ``Communication overhead''. Again this is just a
recommendation, mainly to clarify the underlying differences between what is meant
by efficiency and overhead.}\\[6pt]

\noindent
\textbf{\textsf{Response}}:
Thank you for this suggestion.

Although a potential change of the 
term ``Efficiency'' to ``Topology Adaptation Cost''
followed by a break up of the term 
``Overhead'' to ``Topology Maintenance Cost'' and ``Communication Overhead''
would have been feasible, we believe that such modifications would ultimately 
render the way the overall text has been written somewhat problematic.

In particular, we use the term ``Efficiency'' to identify how \emph{closely matched} are the
overlay and underlying networks after applying a specific approach. 
We experience good efficiency when few(er) packets travel across the network
while at the same time, an approach can preserve 
discoverability of resources (not reducing the search scope). 
In this context, ``Topology Adaptation Cost'' is intended to describe something 
different (as stated below).

We use the ``Overhead'' term to quantify the (mainly communication) cost
that a given approach carries when applied, to (i) create, (ii) adapt and (iii) maintain
the topology awareness of the P2P network.

The proposed distinction of
``Topology Maintenance Cost'' and ``Communication Overhead'' is, for this reason, not
accurate as it is not complete in terms of the intented abstraction of the measure we
want to use in this context.  
``Communication Cost'' can be thought of as an all encompassing cost 
that includes ``Topology Maintenance Cost'' (added communication to maintain topology
awareness in the presence of network churn).

The above choice also may clarify our intention to have
``Topology Adaptation Cost'' fall into the ``Overhead'' family of measures and
not in ``Efficiency'' as suggested by the reviewer above.

Having (hopefully successfully) argued for the use of terms as they appeared in the 
first version of the manuscript, we also understood the point raised and so, 
we have attempted to provide an improved definition of pertinent terms and metrics.
To this end, we updated
subsection ``2.3 Motivation and Goal of this Survey'' with revised/re-written 
discussion on the evaluation criteria. We also provided additional examples of
the features we account for in each of the metrics to further clarify their
nature/objective(s).

We believe the revised text to have clarified the issue raised and 
to present a more straightforward reading at this time.

%% R3.C2
\subsubsection*{\textbf{\textsf{Reviewer's Comment\#2}}}
\noindent
\emph{I also recommend looking for potential trends/patterns in P2P overlay network
approaches over time of introducing these approaches. Maybe listing approaches by time
while studying the evaluated metrics.}\\[6pt]

\noindent
\textbf{\textsf{Response}}:
That is a very good point! 
We have updated our manuscript by
creating timeline plots for depicting families of all surveyed approaches 
and elaborated accordingly
in the respective discussion sections.

Specifically, the timeline figures depict algorithms and methodologies. In one hand, we embed
algorithms/papers per year and on the other hand we show which of the characteristic methodologies
are used per surveyed year. We also emphasize when each methodology first appears in the
literature. That way we are capturing the temporal and trending dimensions of the
surveyed material. Moreover, we have updated paragraphs 3.2 and 4.2 (for the unstructured
and structured sections, respectively) by adding discussion points on each of the timeline figures.

\baselineskip=0.95\baselineskip

%% R3.C3 
\subsubsection*{\textbf{\textsf{Reviewer's Comment\#3}}}
\noindent
\emph{On this note, you should add the year next to each examined approach in figures 19-21, 31-33.}\\[6pt]

\noindent
\textbf{\textsf{Response}}:
We have updated the respective figures mentioned by the reviewer to contain a year indication
next to the examined approach.

%% R3.C4 
\subsubsection*{\textbf{\textsf{Reviewer's Comment\#4}}}
\noindent
\emph{I recommend to add the name of each approach to the left of the
``efficiency overhead scalability'' mini-tables, and not only with the tables that have
more than one approach. I also recommend highlighting the name of each approach (in bold)
at the start of each respective paragraph.}\\[6pt]

\noindent
\textbf{\textsf{Response}}:
Mini tables updated and highlighting approach name added as proposed by the reviewer.

%% R3.C5
\subsubsection*{\textbf{\textsf{Reviewer's Comment\#5}}}
\noindent
\emph{Page 2, bottom of left column, ``This research area started to become active after The mismatch problem is Interest ...''}\\[6pt]

\noindent
\textbf{\textsf{Response}}:
Removed the line as it was not supposed to be there. Thank you for pointing this out.

%% R3.C6
\subsubsection*{\textbf{\textsf{Reviewer's Comment\#6}}}
\noindent
\emph{Page 3, first paragraph ``... approximately 80 million songsJupiterMedi...'' insert space after ``songs''}\\[6pt]

\noindent
\textbf{\textsf{Response}}:
Added a space character in the position correctly suggested by the reviewer.

%% R3.C7
\subsubsection*{\textbf{\textsf{Reviewer's Comment\#7}}}
\noindent
\emph{Page 8 in the APS paragraph, 8th line ``The success of THIS approach ...'' (add ``this'')}\\[6pt]

\noindent
\textbf{\textsf{Response}}:
Added the word ``this'' to the respective position.

%% R3.C7
\subsubsection*{\textbf{\textsf{Reviewer's Comment\#7}}}
\noindent
\emph{Page 11, last line, ``Should there BE m landmarks adopted...'' (move ``be'')}\\[6pt]

\noindent
\textbf{\textsf{Response}}:
Moved the verb ``be'' to its correct place in the sentence as pointed out by the reviewer.

\end{document}

### 2015-09-26 acknowledge.tex
2015.09.28 (Alex, Mema & Vassilis) Rephrasing and fixing typos. Adding rebutal letter and new letter to area editor. Including revised (r1) submission. Minor file reorganization and Makefile cleanup.

@@
 
-We would also like to express our gratitude to all the anonymous reviewers who helped us improve the presentation of our work with their insightful comments.
-This research has been partially supported by \emph{EC FP7 Workprogramme (2007-2013)} under grant agreement n.
-283610 \emph{``European Scalable EarthScience Service Environment (EarthServer)''} and \emph{EC H2020} under grant agreement n.
-654367 \emph{``Agile Analytics on Big Data Cubes (EarthServer 2)''}.
-
-This  work  has  been  partially  supported  by \emph{SocWeb}, \emph{iMarine}, and \emph{Sucre} \emph{FP7 EU projects}.
+We would like to thank the anonymous reviewers who helped us 
+improve the presentation of our work with their insightful comments. 
+This  work  has  been  partially  supported  by
+\emph{iMarine} \emph{FP7 EU} project and \emph{ERC Starting Grant n. 279237}.

### 2015-09-26 background.tex
2015.09.28 (Alex, Mema & Vassilis) Rephrasing and fixing typos. Adding rebutal letter and new letter to area editor. Including revised (r1) submission. Minor file reorganization and Makefile cleanup.

@@
 
-Early overlay protocol implementations gave litte attention to the 
+Early overlay protocol implementations gave little attention to the 
 topology mismatch problem.
@@
 that this is the one criterion of the three we use that its literal values
-``low'' and ``high'' must be interpreted as ``better'' and ``worse'' respectivelly.
+``low'' and ``high'' must be interpreted as ``better'' and ``worse'' respectively.
 

### 2015-09-26 conclusion.tex
2015.09.20 (Vassilis) Added acknowledgements section. Updated metric definitions. Introduced timeline figures to capture termoral and trending dimensions of surveyed material. Minor fixes and typos.
2015.09.28 (Alex, Mema & Vassilis) Rephrasing and fixing typos. Adding rebutal letter and new letter to area editor. Including revised (r1) submission. Minor file reorganization and Makefile cleanup.

@@
-% TODO: Should be reviewed
 \section{Conclusion}
@@
 Development of \p--systems and applications
-typically involve juggling a set of tradeoffs (e.g.,
+typically involve juggling a set of trade-offs (e.g.,
 choosing higher accuracy at the cost of increased system overhead).
@@
 problem in both unstructured and structured \p\ networks, we find this
-tussle-of-tradeoffs property to hold.  
+tussle-of-trade-offs property to hold.  
 With regards to the three criteria we used --efficiency,

### 2015-09-26 intro.tex
2015.09.28 (Alex, Mema & Vassilis) Rephrasing and fixing typos. Adding rebutal letter and new letter to area editor. Including revised (r1) submission. Minor file reorganization and Makefile cleanup.

@@
 Work in the area was inspired mostly by the volatile and fully
-distributed nature of the decentrilized overlays so we skip
+distributed nature of the decentralized overlays so we skip
 centralized systems and in Sections~\ref{section:unstructured} and~\ref{section:structured}

### 2015-09-26 journals/jss/bios/bios.tex
2015.09.28 (Alex, Mema & Vassilis) Rephrasing and fixing typos. Adding rebutal letter and new letter to area editor. Including revised (r1) submission. Minor file reorganization and Makefile cleanup.

\documentclass[10pt,letterpaper]{article}

\usepackage{color}
\usepackage[margin=3cm, a4paper]{geometry}
\usepackage[hyphens]{url}
\usepackage{paralist}

\setlength{\parindent}{0pt}
\setlength{\parskip}{10pt}

\begin{document}

\vspace*{0.5cm}
\centerline{\large \sf Biographical Sketches}
\vspace*{0.5cm}
\vspace*{0.5cm}

\emph{Vassilis Moustakas} received his MS Computer Science from the Univ. of Athens and his BS in Informatics from the Technological Educational Institute of Athens.  He is currently a member of the technical staff at the Amazon Development Center Germany in Munich as part of the A9.com Product Search group implementing scalable search and discovery technologies for Amazon.com and other websites. Prior to that, he worked as a software engineer for a number of companies as well as a researcher for various FP7 EU projects. His interests lie in Information Retrieval, Data Mining and Distributed Computing.

\emph{H\"useyin Akcan} is an Associate Professor in the Department of Software Engineering at Izmir University of Economics. His research interest are in computer systems, sensor networks, computational geometry and computer gaming. He has received both an PhD and an MS in Computer Science from the NYU Polytechnic School of Engineering and his BS from Instabul Technical University.

\emph{Mema Roussopoulos}  is an Associate Professor of Computer Science at the Univ. of Athens. Her interests  are  in  the  areas  of  distributed systems, networking, and digital preservation. She is a recipient of the NSF CAREER Award, the ERC Starting Grant Award, and Best Paper Award at ACM SOSP 2003.  She received her PhD in Computer Science from Stanford University and her BS from the Univ. of Maryland at College Park.

\emph{Alex Delis} is a Professor of Computer Science at the University of Athens. His research interests are in distributed computing and data management.  His work has been supported by agencies and organizations in Australia, US and the European Union. He holds both a PhD and an MS in Computer Science from the Univ. of Maryland at College Park and received his Eng. Diploma from the Univ. of Patras.

\end{document}

### 2015-09-26 journals/jss/highlights/highlights.tex
2015.09.28 (Alex, Mema & Vassilis) Rephrasing and fixing typos. Adding rebutal letter and new letter to area editor. Including revised (r1) submission. Minor file reorganization and Makefile cleanup.

\documentclass[10pt,letterpaper]{article}

\usepackage{color}
\usepackage[margin=3cm, a4paper]{geometry}
\usepackage[hyphens]{url}
\usepackage{paralist}

\setlength{\parindent}{0pt}
\setlength{\parskip}{10pt}

\begin{document}

\vspace*{0.5cm}
\centerline{\large\sf Highlights for JSS-D-15-00624}
\vspace*{0.5cm}
\vspace*{0.5cm}

\begin{itemize}
\item
The topology mismatch problem of P2P architectures is systematically defined.
\item
We aggregate research tackling the topology mismatch problem in P2P systems.
\item
Both unstructured and structured overlay paradigms are surveyed.
\item
We provide tabular data of salient features of all surveyed approaches.
% \item
% Abstracted metrics that help provide a pictorial comparison of methods.
\item
Trending and temporal characteristics of algorithms are presented.
% \item
% No definite solution found, rather a heuristic tussle-of-trade-offs property holds.
% \item
% A solid reference for system architects and researchers in the field is furnished.
\end{itemize}

\end{document}

### 2015-09-26 journals/jss/letter-to-editor/letter-to-editor-r1.tex
2015.09.28 (Alex, Mema & Vassilis) Rephrasing and fixing typos. Adding rebutal letter and new letter to area editor. Including revised (r1) submission. Minor file reorganization and Makefile cleanup.

\documentclass[11pt]{article}

%% inclusion of packages
\usepackage{diletter-wsig}
% \usepackage{diletter-w-sig}
\usepackage{times}
\usepackage{url}

%% paragraph setting
\setlength{\parindent}{0pt}
\setlength{\parskip}{15pt}

%% style-specific parameters
\todayis{September 28th, 2015}
\myname{Alex Delis \\ Professor}
\myemail{\url{ad@di.uoa.gr}}
\myweb{\url{www.di.uoa.gr/~ad}}
\myphone{+30-210-727.5212}
\myfax{+30-210-727.5214}
\mycountry{Greece}
\mytitle{}

\begin{document}
\recipient{%
Professor Eleni Karatza \\
Area Editor - \emph{Journal of Systems and Software} \\
Department of Informatics \\
Aristotle University of Thessaloniki \\
54124 Thessaloniki, Greece
}

\begin{letter}{Dear Professor Karatza, }

\baselineskip=1.13\baselineskip

We are submitting the revision of our paper entitled 
``Alleviating the Topology Mismatch Problem in Distributed Overlay Networks
: A Survey''. We have initially submitted this paper to 
{\it Journal of Systems and Software} on July 21st, 2015.

The paper is co-authored with Vassilis Moustakas 
of Amazon Development Center in Munich, Germany, 
Huseyin Akcan of Izmir University of Economics in 
Izmir, Turkey,  and 
Mema Roussopoulos of the Univ. of Athens in Athens, Greece.

We also provide specific replies to comments made by the 
two reviewers.

Thank you very much for kindly taking the time to handle our 
revised submission and we look forward to hearing from you.

\myclosing{Sincerely yours,}

\end{letter}

\end{document}

### 2015-09-26 journals/jss/letter-to-editor/letter-to-editor.tex
2015.09.28 (Alex, Mema & Vassilis) Rephrasing and fixing typos. Adding rebutal letter and new letter to area editor. Including revised (r1) submission. Minor file reorganization and Makefile cleanup.

--- 2015-09-26 journals/jss/letter-to-editor/letter-to-editor-r1.tex
@@
 %% style-specific parameters
-\todayis{September 28th, 2015}
+\todayis{July 21st, 2015}
 \myname{Alex Delis \\ Professor}
@@
 \recipient{%
-Professor Eleni Karatza \\
-Area Editor - \emph{Journal of Systems and Software} \\
-Department of Informatics \\
-Aristotle University of Thessaloniki \\
-54124 Thessaloniki, Greece
+Professor Hans van Vliet \\
+Editor-in-Chief, \emph{Journal of Systems and Software}\\
+Dept. of Computer Science, VU University, \\
+De Boelelaan 1081, 1081 HV, \\
+Amsterdam, The Netherlands
 }
 
-\begin{letter}{Dear Professor Karatza, }
+\begin{letter}{Dear Professor van Vliet,}
 
@@
 
-We are submitting the revision of our paper entitled 
+We are submitting a paper entitled 
 ``Alleviating the Topology Mismatch Problem in Distributed Overlay Networks
-: A Survey''. We have initially submitted this paper to 
-{\it Journal of Systems and Software} on July 21st, 2015.
+: A Survey''
+for consideration 
+to the {\it Journal of Systems and Software}.
 
@@
 
-We also provide specific replies to comments made by the 
-two reviewers.
-
-Thank you very much for kindly taking the time to handle our 
-revised submission and we look forward to hearing from you.
+Thank you very much for kindly taking the time to handle our submission and
+we look forward to hearing from you soon.
 

### 2015-09-26 structured.tex
2015.09.28 (Alex, Mema & Vassilis) Rephrasing and fixing typos. Adding rebutal letter and new letter to area editor. Including revised (r1) submission. Minor file reorganization and Makefile cleanup.

@@
 disable \emph{traceroute} support on their routers and this 
-undoubtly affects \emph{T2MC}'s operation.
+undoubtedly affects \emph{T2MC}'s operation.
 %%
@@
 incoming peer to get bootstrapped.
-The new-comer takes the bootstrapping nodes's \emph{ID} prefix 
+The new-comer takes the bootstrapping node's \emph{ID} prefix 
 and fills the remaining bits of its key using the \emph{RLM} technique.
@@
 information for selecting the next hop or to construct and maintain
-routing tables as a whole, seem to be the elements more used in the
-literature throught the years that this area is researched.
+routing tables as a whole, are the two predominant elements used 
+throughout the years in 
+the structured systems literature.
 %%%
@@
 while \emph{LAPTOP} combines geographic--layout with
-replication strategies to refine its results. What looks a bit different
-compared the trends in unstructred systems is the fact that, here, researchers
+replication strategies to refine its results. What looks different
+compared with the trends in unstructured systems is the fact that, here, researchers
 seem to juggle with more methods at the same time. This, and the
 fact that these suggested solutions are also heuristic in nature, opens them up 
-to even harder trade-offs which, in most cases, are 
-highly affected by the application domain and the implementation itself.
+to trade-offs that are more difficult to manage and are highly 
+affected by both the application domain and the implementation itself.
 
@@
 utilized proximity and geographic aspects.
-Table~\ref{structured:table} outlines key properties for all surveyed approach 
+Table~\ref{structured:table} outlines key properties for all surveyed approaches 
 that helps differentiate and better position competing techniques.

### 2015-09-26 unstructured.tex
2015.09.28 (Alex, Mema & Vassilis) Rephrasing and fixing typos. Adding rebutal letter and new letter to area editor. Including revised (r1) submission. Minor file reorganization and Makefile cleanup.

@@
 %%
-\emph{THANCS} reportedly improves brodcast performance the extent of which, though,
+\emph{THANCS} reportedly improves broadcast performance the extent of which, though,
 depends on the underlying network since \emph{THANCS} does not provide the
@@
 they do not often furnish ``pure'' approaches that exclusively
-belong to one of the $4$ aforementioned methodologies.
+belong to one of the four aforementioned methodologies.
 For instance, \emph{mOverlay} combines topology adaptation 
 with landmark binning 
-while \emph{Gia} weaves together $3$ methodologies:  topology adaptation, 
+while \emph{Gia} weaves together three methodologies:  topology adaptation, 
 forwarding optimization and caching.
 %%
-The timeline in Figure~\ref{unstructured:timeline} further emphasises this trend
-in the broader picture. All characteristics we discuss,
-seem to have been identified and put into practice, early on in the literature.
-In that sense, no breakthrough
-or any revolutionary ideas came out lately but mostly people fine-tuned the
-heuristic parameters in pursue of the goldset that performs best. That doesn't
-render all that work less exciting or important since it led to better theoretical
+The timeline in Figure~\ref{unstructured:timeline} further emphasizes this trend
+in the broader picture. All methodologies we discuss,
+have been identified and put into practice, early on in the literature,
+while in subsequent years, the research community has mostly focused on fine-tuning
+the heuristic parameters of these methodologies in pursuit of the
+combination that performs best.
+This has led to better theoretical 
 understanding of the problem and its difficulty. Moreover, the combination of
-multiple methodologies inevitably leads to interesting feature/performance trade-offs
-that researchers need to tackle in real world applications: for example, 
+multiple methodologies has inevitably led to interesting feature/performance trade-offs
+that researchers need to tackle in real world applications. For example, 
 the effectiveness of a caching/replication component can be undermined
 by a continuously adapting overlay that removes important links between peers.
-Topology adaptation and forwarding optimization seem to be constant areas of
-concentration for the researchers. This is of no surprise as all seem to
-understand the importance of the volatile nature of unstructured \p systems. 
+In particular, topology adaptation and forwarding optimization have been constant areas of
+focus of the research community. This is of no surprise given 
+the volatile nature of unstructured \p systems. 
 %%
